### 2017-05-23 chronology.tex
add a chronology of things done

\documentclass[a4paper,11pt,oneside,onecolumn,openany,final,svgnames]{memoir}
\input{packages}
\input{abbreviations}
\title{PhD plans}
\title{Géraud Le Falher}
\begin{document}
\maketitle
\begin{description}
	\item[Oct. 14] State of the art for \pcc{} (see
		\href{run:support/cc.pdf}{A few words on Correlation Clustering})
	\item[Nov. 14 --- Jan. 15] Trying different ways of closing the
		triangles of a general graph without introducing bad cycle in
		order to transfer the \ccPivot{} algorithm (see
		\href{run:support/december.pdf}{Combinatorial
		Correlation Clustering on general unweighted graphs} and the Dating presentation
		\href{run:support/presentationeng.pdf}{Signed graphs: clustering and link prediction})
	\item[Jan. 15 --- Apr. 15]{Predicting link signs by building a low
		stretch spanning tree dubbed \gtx{}, which works by extracting
		stars, connecting them in a graph, and repeating on this reduced graph
		(see \href{run:support/nips2015.pdf}{some experimental results}).}
	\item[Mar. 15] Visit to Claudio. The setting was a graph where nodes have profiles and be
		similar along various ways, and the problems were either answering online
		similarity queries or recovering hidden profiles (see
		\href{run:support/contextual_clustering.pdf}{Contextual Clustering}).
	\item[May. 15] Implementing \shz{}~\autocite{Vitale2012} (actually it took a week, but I wasn't so
		much in the lab in May)
	\item[Jun. 15 --- Aug. 15] 
		I supervised the internship of Paul Dennetiere on the implementation of a parallel
		algorithm for \pcc{} described by \textcite{Pan2014} (see the
		\href{run:support/stage_cc.pdf}{internship description}, a
		\href{run:support/draft_paul.pdf}{draft of the technical part of Paul's report} and
		some ideas on \href{run:support/parallel_postprocessing.pdf}{parallel post
		processing} we didn't have time to explore).

		I worked on building trees that minimized the stretch of the distances between a
		set of revealed nodes and the other ones (see a
		\href{run:support/sujet2_lowstretch.pdf}{a MVA project topic} and
		\href{run:support/constrained_low_stretch.pdf}{some initial experiments}).

		Two others summaries of what I did until then can be found in my
		\href{run:support/postermlss.pdf}{MLSS poster} and my
		\href{run:support/progress.pdf}{first year progress report}.
	\item[Sep. 15]
		When Fabio came back from Finland, he told me about a problem on graph
		reconstruction. I can't fully remember the setting, but I think the general idea
		was to make some queries about the existence of certain edges. Trivially, by
		making $\Theta(|V|^2)$ queries, we know the full topology so the question of
		was can we do better. I didn't write much about except
		\href{run:support/permutation.pdf}{one thing about counting permutation on a
		line graph}, but this is very incomplete.
	\item[Oct. 15 --- Nov. 15] Following Michael's visit during the summer, we
		worked on a binary node classification problem where the bias is that a node
		is labeled $+1$ iff at least two of its neighbors are labeled $+1$. This
		lead to a propagation algorithm based on triangles chain (see
		\href{run:support/NodeClassification_triangle_mail_fabio_nov15.pdf}{some Fabio writeup on the topic}).
	\item[Nov. 15 --- May. 16] Edge classification in the troll trust model (see
		the \href{https://arxiv.org/pdf/1606.00182.pdf}{arxiv version of the AISTATS paper})
	\item[Jun.16 --- Dec. 16] Edge clustering
	\item[Dec. 16] Students of the MVA did a project on \pcc{} under perturbation,
		here is the \href{run:support/sujet1_ccpert.pdf}{topic description} and
		\href{run:support/Dubreuil_Ramzi_report.pdf}{their report}.
	\item[Jan. 17 --- Feb. 17] We worked with Fabio on a more robust version of
		the \shz{} algorithm that decreases its confidence to labeled nodes when
		they are implicated in further prediction mistake but
		\href{run:support/trees.pdf}{experimental results} are not very satisfying.
	\item[Mar. 17 --- May. 17] Edge clustering
\end{description}

During my thesis, I also published my Master's thesis in a
conference~\autocite{Thesis15} and kept working on Urban Informatics, which
resulted in poster~\autocite{WWWDemo16} and a journal
paper~\autocite{GeotopicTBD16}.

Furthermore I (loosely) maintained a list of references on
\href{run:support/dist.pdf}{distributed graph processing frameworks} and
\href{run:support/embedding.pdf}{neural network based graph embedding methods}.

\begingroup
% \setstretch{0.9}
\setlength\bibitemsep{2pt}
\printbibliography
\endgroup
\end{document}

### 2017-05-23 packages.tex
add a chronology of things done

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{setspace}
\usepackage[citestyle=alphabetic,bibstyle=ieee-alphabetic,isbn=false,maxnames=1,minnames=1,sorting=none,backend=biber,defernumbers=true]{biblatex}
\AtEveryBibitem{
   \clearfield{arxivId}
   % \clearfield{booktitle}
   % \clearfield{doi}
   \clearfield{eprint}
   \clearfield{eventdate}
   \clearfield{isbn}
   \clearfield{issn}
   % \clearfield{journaltitle}
   \clearfield{month}
   % \clearfield{number}
   % \clearfield{pages}
   \clearfield{series}
   \clearfield{url}
   \clearfield{urldate}
   \clearfield{venue}
   % \clearfield{volume}
   \clearlist{location} % alias to field 'address'
   \clearlist{publisher}
   \clearname{editor}
}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{%
    % draft,    % = no hyperlinking at all (useful in b/w printouts)
    colorlinks=true, linktocpage=true, pdfstartpage=3, pdfstartview=FitV,%
    % uncomment the following line if you want to have black links (e.g., for printing)
    %colorlinks=false, linktocpage=false, pdfborder={0 0 0}, pdfstartpage=3, pdfstartview=FitV,%
    breaklinks=true, pdfpagemode=UseNone, pageanchor=true, pdfpagemode=UseOutlines,%
    plainpages=false, bookmarksnumbered, bookmarksopen=true, bookmarksopenlevel=1,%
    hypertexnames=true, pdfhighlight=/O,%nesting=true,%frenchlinks,%
    urlcolor=Chocolate, linkcolor=DodgerBlue, citecolor=LimeGreen, %pagecolor=RoyalBlue,%
}
\addbibresource{/home/orphee/data/projects/biblio/library.bib}

### 2017-06-08 lowstretch_and_cc.tex
start looking for biological directed signed networks

\documentclass[a4paper,11pt,oneside,onecolumn,openany,notitlepage,final,svgnames]{article}
\input{packages}
\input{abbreviations}
\usepackage{charter}
\usepackage{fullpage}
\title{Edge sign prediction in general graphs and connection with \pcc{}}

\begin{document}
\maketitle

According to \href{https://web.stanford.edu/class/cs224w/slides/handout-bionets.pdf}%
{this description of various biological networks}, \emph{gene regulatory network} are directed and
contains activation and inhibition links, as \href{https://en.wikipedia.org/wiki/Gene_regulatory_network#Overview}%
{showed on Wikipedia}. There are some online databases such as
\href{http://regulondb.ccg.unam.mx/menu/download/datasets/index.jsp}%
{RegulonDB} or \href{http://www.pathwaycommons.org/pcviz/}{Pathway Commons} (this last one provide
network in the BioPAX format, which can be visualized by
\href{http://www.cytoscape.org/}{CytoScape}).
Another sources is the \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2708159/table/T1}{Table 1}
of \autocite{BioSigned09}.

\begingroup
\setstretch{0.9}
\setlength\bibitemsep{2pt}
\printbibliography
\endgroup

\end{document}

### 2017-06-09 lowstretch_and_cc.tex
start writing about the limitations of the troll method (+refs)

@@
-\documentclass[a4paper,11pt,oneside,onecolumn,openany,notitlepage,final,svgnames]{article}
+\documentclass[nobib,a4paper,11pt,oneside,onecolumn,openany,notitlepage,final,svgnames]{tufte-handout}
 \input{packages}
 \input{abbreviations}
-\usepackage{charter}
-\usepackage{fullpage}
-\title{Edge sign prediction in general graphs and connection with \pcc{}}
+% \usepackage{charter}
+% \usepackage{fullpage}
+
+\title{Edge sign prediction in general graphs and connection with Correlation Clustering}
 
@@
 \maketitle
+
+\section{Limitations of the troll method}
+\label{sec:limitations_of_the_troll_method}
+
+The method presented in the previous chapter hinges crucially upon our sign generative model. Yet
+one can imagine contexts where this model is not applicable, especially when nodes do not represent
+human beings. One way to alleviate this issue was introduced with the online algorithm, where this
+time, signs are generated by an arbitrary adversary. However, in that case, we are still facing two
+limitations\marginpar{limitations is not the right word, maybe \emph{strong constraints}…}
+\begin{enumerate}
+	\item Our bias remains that the labeling is regular (recall this means informally than the all
+		outgoing signs from a given node tend to be the same, and likewise for the incoming signs),
+		since irregularities are the cost payed by the adversary to make our algorithm mispredict. While
+		this bias is well suited to social networks, other applications may require other
+		bias\marginpar{namely the \pcc{} bias we'll introduce next}.
+	\item In the online setting, we evaluate our performance by the regret\marginpar{ref to equation},
+		whereas in general we are interested in more classical measure of predictive accuracy.
+\end{enumerate}
+
+Take for instance a biological network.
 
@@
 \href{http://www.cytoscape.org/}{CytoScape}).
-Another sources is the \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2708159/table/T1}{Table 1}
-of \autocite{BioSigned09}.
+Another source is the \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2708159/table/T1}{Table 1}
+of \cite{BioSigned09}.
+
+other domain? coref, images (is it directed?), entity resolution…\marginpar{Actually it's difficult
+to find other kind of graph because all those are constructed (like image, coreference,
+deduplication) are inherently symmetric}
+
+A much more serious limitation is that our sign model only apply to directed graph. Consider another
+model, where each node $i$ is endowed with an integer $c_i \in \{1, \ldots, k\}$ that define its
+cluster and let the sign of the undirected edge $i,j$ be $+1$ if $c_i = c_j$ and $-1$ otherwise.
+This corresponds naturally to situation modeled by the \pcc{} problem.
+
+For the 2 clusters case ($k=2$), characterization proven already in 1936 by \textcite{Konig36}, as noted in
+\cite{Zaslavsky2012} (commenting on \cite{harary1953}: \enquote{Although Thm. 3 was anticipated by
+\textcite[Theorem X.11]{Konig36}  without the terminology of signs, here is the
+first recognition of the crucial fact that labelling edges by elements of a
+group—specifically, the sign group—can lead to a general theory.})
+and \cite{Huffner2010} (\enquote{\textcite{Konig36} proved the following characterization of balanced graphs. For a
+	graph $G = (V , E)$, the following are equivalent:
+	\begin{enumerate}
+		\item V can be partitioned into two sets $V_1$ and $V_2$ called sides such that there is no negative edge $\{v, w\} \in E$ with both $v, w \in V_1$ or both $v, w \in V_2$ and no negative edge \{v, w\} with $v \in V_1$ and $w \in V_2$ .
+		\item $V$ can be colored with two colors such that for all $\{v, w\} \in E^-$, the vertices $v$ and $w$ have different colors, and for all $\{v, w\} \in E^+$, the vertices $v$ and $w$ have the same color.
+The color classes correspond to the sides.
+		\item $G$ does not contain cycles with an odd number of negative edges.
+	\end{enumerate}
+	Using the characterization by a coloring, it is easy to see that balance of a signed
+graph can be checked in linear time by depth-first search.})
 
@@
 
+% \bibliography{/home/orphee/data/projects/biblio/library.bib,more.bib}
+% \bibliographystyle{plainnat}
 \end{document}

### 2017-06-13 lowstretch_and_cc.tex
add outline for the lowstretch chapter

@@
-\documentclass[nobib,a4paper,11pt,oneside,onecolumn,openany,notitlepage,final,svgnames]{tufte-handout}
+\documentclass[nobib, a4paper, 11pt, oneside, onecolumn, openany, notitlepage, final,
+svgnames,  marginals=raggedouter,
+%symmetric, justified,
+]{tufte-handout}
+\setcounter{secnumdepth}{3}
+\setcounter{tocdepth}{3}
 \input{packages}
@@
 \maketitle
+% \tableofcontents
 
@@
 
+\subsection{Reliance on the sign generative model}
+\label{sub:reliance_on_the_sign_generative_model}
+
 The method presented in the previous chapter hinges crucially upon our sign generative model. Yet
 one can imagine contexts where this model is not applicable, especially when nodes do not represent
-human beings. One way to alleviate this issue was introduced with the online algorithm, where this
+human beings. One way to alleviate this issue was introduced with our online algorithm, where this
 time, signs are generated by an arbitrary adversary. However, in that case, we are still facing two
@@
 	\item Our bias remains that the labeling is regular (recall this means informally than the all
-		outgoing signs from a given node tend to be the same, and likewise for the incoming signs),
+		the outgoing signs from a given node tend to be the same, and likewise for the incoming signs),
 		since irregularities are the cost payed by the adversary to make our algorithm mispredict. While
@@
 		bias\marginpar{namely the \pcc{} bias we'll introduce next}.
-	\item In the online setting, we evaluate our performance by the regret\marginpar{ref to equation},
-		whereas in general we are interested in more classical measure of predictive accuracy.
+	\item In the online setting, we evaluate our performance by the regret\marginpar{ref to regret equation},
+		whereas in general we are interested in more classical measures of predictive accuracy.
 \end{enumerate}
+
+\paragraph{Experience on directed biological network}
+\label{par:experience_on_directed_biological_network}
 
@@
 of \cite{BioSigned09}.
+%DasGupta also describe (in
+%http://www.sciencedirect.com/science/article/pii/S0303264706001419#sec14) how
+%they built a directed signed network from this SBML file
+%https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1681468/bin/msb4100014-sd1.xml
 
 other domain? coref, images (is it directed?), entity resolution…\marginpar{Actually it's difficult
-to find other kind of graph because all those are constructed (like image, coreference,
+to find other kind of graph because all constructed ones (like image, coreference,
 deduplication) are inherently symmetric}
 
-A much more serious limitation is that our sign model only apply to directed graph. Consider another
-model, where each node $i$ is endowed with an integer $c_i \in \{1, \ldots, k\}$ that define its
-cluster and let the sign of the undirected edge $i,j$ be $+1$ if $c_i = c_j$ and $-1$ otherwise.
-This corresponds naturally to situation modeled by the \pcc{} problem.
+\subsection{Need for a directed graph}
+\label{sub:need_for_a_directed_graph}
 
+A more serious limitation is that our sign model only applies to directed graphs. Consider
+another model, where each node $i$ is endowed with an integer $c_i \in \{1, \ldots, k\}$ that
+specifies its cluster and let the sign of the undirected edge $i,j$ be $+$ if $c_i = c_j$ and $-$
+otherwise.  This corresponds naturally to the situation modeled by the \pcc{} problem.
+
+\section{\pcc{}}
+\label{sec:correlation_clustering}
+
+\subsection{Problem setting}
+\label{sub:problem_setting}
+
+\subsection{Relation with edge sign prediction}
+\label{sub:relation_with_edge_sign_prediction}
+
+\iffalse
 For the 2 clusters case ($k=2$), characterization proven already in 1936 by \textcite{Konig36}, as noted in
@@
 group—specifically, the sign group—can lead to a general theory.})
-and \cite{Huffner2010} (\enquote{\textcite{Konig36} proved the following characterization of balanced graphs. For a
-	graph $G = (V , E)$, the following are equivalent:
+and \cite{Huffner2010} (\enquote{\textcite{Konig36} proved the following characterization of
+	balanced graphs. For a graph $G = (V , E)$, the following are equivalent:\marginpar{There is a
+	proof in \autocite[p. 111]{BookKleinberg2010}, maybe I can rewrite it as well}
 	\begin{enumerate}
-		\item V can be partitioned into two sets $V_1$ and $V_2$ called sides such that there is no negative edge $\{v, w\} \in E$ with both $v, w \in V_1$ or both $v, w \in V_2$ and no negative edge \{v, w\} with $v \in V_1$ and $w \in V_2$ .
-		\item $V$ can be colored with two colors such that for all $\{v, w\} \in E^-$, the vertices $v$ and $w$ have different colors, and for all $\{v, w\} \in E^+$, the vertices $v$ and $w$ have the same color.
-The color classes correspond to the sides.
+		\item $V$ can be partitioned into two sets $V_1$ and $V_2$ called sides such that there is no
+			negative edge $\{v, w\} \in E$ with both $v, w \in V_1$ or both $v, w \in V_2$ and no negative
+			edge $\{v, w\}$ with $v \in V_1$ and $w \in V_2$ .
+		\item $V$ can be colored with two colors such that for all $\{v, w\} \in E^-$, the vertices $v$
+			and $w$ have different colors, and for all $\{v, w\} \in E^+$, the vertices $v$ and $w$ have
+			the same color. The color classes correspond to the sides.
 		\item $G$ does not contain cycles with an odd number of negative edges.
@@
 graph can be checked in linear time by depth-first search.})
+\fi
+
+\subsection{State of the art}
+\label{sub:state_of_the_art}
+
+\subsection{Variants and extensions}
+\label{sub:variants_and_extensions}
+
+\subsubsection{\pcc{} under stability assumption}
+\label{ssub:cc_under_stability_assumption}
+
+Haris Angelidakis, Konstantin Makarychev, and Yury Makarychev. 2017.
+Algorithms for Stable and Perturbation-Resilient Problems. STOC’17
+\href{http://ttic.uchicago.edu/~yury/papers/two-stable.pdf}{10.1145/3055399.3055487}
+improves over the one cited in the internship description
+
+\subsubsection{Parallel \pcc{}}
+\label{ssub:parallel_cc}
+
+\subsection{Empirical evaluation?}
+\label{sub:cc_empiracal_evaluation}
+
+\section{Low stretch trees and spanners}
+\label{sec:low_stretch_trees_and_spanners}
+
+\subsection{\gtx{}: a simple low stretch tree construction}
+\label{sub:gtx_a_simple_low_stretch_tree_construction}
+
+\subsection{State of the art}
+\label{sub:gtx_state_of_the_art}
+\url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
+for instance the Elkin paper \enquote{Our centralized randomized algorithm computes (with
+probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in
+$O(|E|)$ time, whenever $k = \Omega(\log n)$. Note that when $k = \omega(\log n)$, the number of
+edges is $n(1+o(1))$, i.e., in this range the algorithm computes an ultra-sparse spanner in $O(|E|)$
+time.} For instance, if $k=5\log n$, we get a $10\log n$-spanner with $n\left(1+O\left(\frac{\log\log
+n}{n}\right)\right)$ edges in $O(|E|)$ time.
+
+\subsection{Empirical evaluation}
+\label{sub:gtx_empiracal_evaluation}
+
+\section{Conclusion}
 

### 2017-06-13 packages.tex
start looking for biological directed signed networks
start writing about the limitations of the troll method (+refs)
add outline for the lowstretch chapter

@@
 \usepackage{setspace}
-\usepackage[citestyle=alphabetic,bibstyle=ieee-alphabetic,isbn=false,maxnames=1,minnames=1,sorting=none,backend=biber,defernumbers=true]{biblatex}
+\usepackage[citestyle=alphabetic, bibstyle=ieee-alphabetic, isbn=false, maxnames=1, minnames=1,
+sorting=nyvt, backref=true, backend=biber, defernumbers=true]{biblatex}
 \AtEveryBibitem{
@@
    \clearfield{series}
-   \clearfield{url}
+   % \clearfield{url}
    \clearfield{urldate}
@@
 }
+\addbibresource{/home/orphee/data/projects/biblio/library.bib}
+\addbibresource{more.bib}
 \usepackage{xcolor}
@@
 }
-\addbibresource{/home/orphee/data/projects/biblio/library.bib}

### 2017-06-14 gtx_experiments.tex
add a previous document on galaxy tree experiments, plus small layout change

In this \namecref{sub:gtx_empirical_evaluation}, we provide empirical evidences of the properties of
\gtx{} over several classes of graph, and compare it with a \bfs{} baseline.\marginpars{If time
allows, it would be interesting to implement some methods of \vref{sub:gtx_state_of_the_art} and add
them to the comparison}\todo*{implement more low stretch methods} Namely, we consider three kinds of
graph topology (with both synthetic and real world instances that carry signs on their edges) and
evaluate $(i)$ what average stretch is reached by various trees and $(ii)$ how accurate is the sign
prediction.

\subsubsection{Graph topology}

The three kinds of topology we consider are:
\begin{description}
	\item[\grid{}] which are 2D lattices, where each node has four neighbors except on the boundary.
		The synthetic ones are square, while the \enquote{real world} ones represents the four neighbors
		pixel connectivity of the pictures showed in \autoref{fig:gtx_xp_bwpics}.
	\item[\lpa{}] which are built synthetically according to the model of \textcite{Barabasi1999}.
		While this does not follow the more rigorous specification of \textcite{PAmodel04}, informally,
		we start with a line graph of $m$ nodes and add node one by one until the graph consists of $n$
		nodes. Each time a new node is added, it is connected to $m$ of the existing nodes with a
		probability proportional to their degree. Here we choose $m=3.13$, that is when adding a new
		node, we pick $3$ or $4$ existing neighbors such the initial expected number of neighbors for
		each new nodes is $3.13$. Such graphs are quite sparse and have short diameter, thus providing a
		crude but reasonable approximation of online social networks. Therefore, the real world
		instances of the \lpa{} model are \wik{}, \sla{} and \epi{}\marginpars{as used in the first
		chapter}\todo*{add a ref to first chapter} along with \gplus{}. The last one is constructed from
		ego networks of \gplus{}\footnote{Available at
		\url{http://snap.stanford.edu/data/egonets-Gplus.html}} by keeping the largest connected
		component of users whose gender is known. Basic statistics of those real \lpa{} graphs are
		presented in (\autoref{tab:gtx_xp_dataset}). 
	\item[\triangle{}] which consists of a Delaunay triangulation of random 2D points\footnote{As
		implemented by the \textsf{graph-tool} library (\url{https://graph-tool.skewed.de})}.
\end{description}

\begin{table}[htpb]
	\centering
	\caption{Dataset description }\label{tab:gtx_xp_dataset}
	\begin{tabular}{lrrcc}
		\toprule
             & $|V|$  & $|E|$    & fraction of $+$ edges & $\frac{2|E|}{|V|\cdot(|V|-1)}$ \\
		\midrule
		\wik{}   & \np{7065}   & \np{99936}    & 78.5\%                & $4.00\cdot 10^{3}$             \\
		\gplus{} & \np{74917}  & \np{10130461} & 67.6\%                & $3.61\cdot 10^{3}$             \\
		\sla{}   & \np{82052}  & \np{498527}   & 76.4\%                & $1.48\cdot 10^{4}$             \\
		\epi{}   & \np{119070} & \np{701569}   & 83.2\%                & $9.90\cdot 10^{5}$             \\
		\bottomrule
	\end{tabular}
\end{table}

\begin{figure}[t]
	\centering
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/zmonastery}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/zworld}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/nips_poster}
	\end{subfigure}

	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/zmonastery_bin}
		\caption{monastery}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/zworld_bin}
		\caption{world}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/nips_poster_bin}
		\caption{poster}
	\end{subfigure}

	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/nips_logo}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/space}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/waterfall}
	\end{subfigure}

	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/nips_logo_bin}
		\caption{logo}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/space_bin}
		\caption{space}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/waterfall_bin}
		\caption{waterfall}
	\end{subfigure}
	\caption{Real world pictures and their binarized version}\label{fig:gtx_xp_bwpics}
\end{figure}

\subsubsection{Stretch}

The first property of Galaxy trees we wish to evaluate is their stretch, which depends only of graph
topology. Namely, let $G$ be a graph over vertex set $V$ with $|V|=n$ and edge set
$E$.\Todo[MOVE]{Stretch definition is likely to happen somewhere earlier} Furthermore, let $T$ be a
spanning tree of $G$ and $\etest{}$ the edges of $G$ not in $T$. Then we define the \emph{average
test edge stretch} as $\frac{1}{|\etest{}|} \sum_{(u,v) \in \etest{}} |\mathrm{path}^T_{u,v}|$,
where $|\mathrm{path}^T_{u,v}|$ is the unique path between $u$ and $v$ in $T$.

As we consider unweighted graphs, we compare \gtx{} with a natural baseline, namely a spanning tree
rooted at the highest degree node and obtained through a breadth first visit of the graph. This
involves randomness in order in which nodes are visited. Likewise in \gtx{}, the choice of the edge
linking two stars is not always unique, meaning that we have to break ties at random.  Therefore,
for each graph, we repeat the tree construction 12 times and present the average result, noting that
the variance (showed as error bar in \autoref{fig:gtx_xp_st}) is small.

On \lpa{} and \triangle{}, we see that both trees exhibits logarithmic stretch, although with a
larger constant for \gtx{}. Note that this is also the case for others low stretch tree methods
\autocite[\S 5.3.1]{papplow}. On \grid{} however, \gtx{} preserves this logarithmic stretch growth
while this is visually no longer the case for \bfs{}.

\begin{figure}[tbh]
	\centering
	\begin{subfigure}[b]{0.9\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/gridst}
		\caption{\grid{} }\label{fig:gtx_xp_gridst}
	\end{subfigure}

	\begin{subfigure}[b]{0.9\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/past}
		\caption{\lpa{} }\label{fig:gtx_xp_past}
	\end{subfigure}

	\begin{subfigure}[b]{0.9\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/trst}
		\caption{\triangle{} }\label{fig:gtx_xp_trst}
	\end{subfigure}
	\caption{Stretch over graphs of increasing size}\label{fig:gtx_xp_st}
\end{figure}

\subsubsection{Sign prediction}

The second design goal of Galaxy trees is to accurately predict the sign of edges in $\etest{}$.
Except for the three real datasets that already include signs\footnote{We nonetheless perform some
preprocessing in order to make them undirected to remove the small proportion of conflicting edges
(e.g. positive from $u$ to $v$ but negative from $v$ to $u$).}, all the other are constructed,
meaning we have to set sign on their edges in the first place. This is done by partitioning the
nodes into two clusters. For \gplus{} we use node gender, for pictures we use node color (black or
white), and for all others, we propagate labels $0$ and $1$ from randomly selected high degree nodes.
Once each node belongs to one of the two clusters, we set the sign of an edge between two nodes to
be $+$ if they are in the same cluster and $-$ otherwise.  Predicting using path parity will thus
gives perfect result. To test performance in real or adversarial situation, we then add noise, that
is we select a fraction of edges uniformly at random and flip their sign.

We evaluate the performance of our prediction using the Matthews Correlation\Todo{merge this
definition of MCC with the one in first chapter} Coefficient (MCC)~\autocite{MCC00} \[
	\mathrm{MCC} = \frac{ TP \times TN - FP \times FN } {\sqrt{ (TP + FP) ( TP + FN ) (
			TN + FP ) ( TN + FN ) } } = \pm \sqrt{\frac{\chi^2}{n}}
\]
Since we do not have confidence score, we cannot use AUC. Yet we have to account for the large sign
unbalance and thus cannot rely on accuracy or $F_1$ measure.  Therefore we choose MCC, which
combines all the four numbers of the confusion matrix in a single metric. It ranges from $+1$
(perfect prediction) to $-1$ (inverse prediction) through $0$ (random prediction). As a
demonstration of MCC usefulness, predicting all edges but one to be positive on Slashdot gives
$.764$ accuracy, $.886$ $F_1$ score\marginpars{Actually the $F_1$ score is $.866$ for positive
edges, $0$ for negative ones and $.661$ if we can take an average weighted by class size.} but
$-0.0007$ MCC.

As showed in \autoref{fig:gtx_xp_mcc}, when the noise level is low, \gtx{} performs better than
\bfs{}. As the noise level gets higher, they have similar performance. Note also than in
\autoref{fig:gtx_xp_pasynthmcc}, \gtx{} is less sensible to the size of the graph.

\begin{figure}[tbh]
	\centering
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/grsynthmcc}
		\caption{Synthetic \grid{} }\label{fig:gtx_xp_grsynthmcc}
	\end{subfigure}~
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/grrwmcc}
		\caption{Pictures \grid{} }\label{fig:gtx_xp_grrwmcc}
	\end{subfigure}
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/pasynthmcc}
		\caption{Synthetic \lpa{} }\label{fig:gtx_xp_pasynthmcc}
	\end{subfigure}~
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/trmcc}
		\caption{\triangle{} }\label{fig:gtx_xp_trmcc}
	\end{subfigure}
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/parwmcc}
		\caption{Real world network }\label{fig:gtx_xp_parwmcc}
	\end{subfigure}
	\caption{MCC over various graphs}\label{fig:gtx_xp_mcc}
\end{figure}

To further assess the quality of our trees, we plug them in them into a successful heuristic method
to predict edge sign: \asym{}~\autocite{Kunegis2009}. \Todo{It might also be interesting to see if
that would be a good training set for our troll method, although it has to be checked it makes sense
from a running time point of view.} It computes the exponential of the adjacency matrix after it has
been reduce to $z$ dimension. This allows to count the sign of all paths between two pairs of nodes
with decreasing weight depending of their length. To simulate an active learning setting, we reveal
only a subset of edge in $A$. This subset can be: $i)$ the edges forming a \bfs{}, $ii)$ the edges
forming a \gtx{} $iii)$ $|V|-1$ edges chosen uniformly at random.

We set the parameter $z$ equal to $15$ because $i)$ it is one of the best in \cite[Fig.
11]{Kunegis2009}, $ii)$ it performs well on real dataset in \cite[Fig.3]{Cesa-Bianchi2012a}, and
$iii)$ it was good in our initial testing (\texttt{20150401\_wed\_spectral.ipynb}).

As the \asym{} has a $O(n^3)$ complexity and uses quite some memory at prediction time, the larger
graphs used previously are not all included. The conclusion of \autoref{fig:gtx_xp_asym} is that
except on social network, it is better to use spanning trees than random edges. Specifically, \gtx{}
on \grid{} and \bfs{} elsewhere.

\begin{figure}[tbh]
	\centering
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/grsynthasym}
		\caption{Synthetic \grid{} \label{fig:gtx_xp_grsynthasym}}
	\end{subfigure}~
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/grrwasym}
		\caption{\enquote{Real} \grid{} }\label{fig:gtx_xp_grrwasym}
	\end{subfigure}
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/pasynthasym}
		\caption{Synthetic \lpa{} }\label{fig:gtx_xp_pasynthasym}
	\end{subfigure}~
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/trasym}
		\caption{\triangle{} }\label{fig:gtx_xp_trasym}
	\end{subfigure}
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/parwasym}
		\caption{Real world network }\label{fig:gtx_xp_parwasym}
	\end{subfigure}
	\caption{\asym{} over various graphs}\label{fig:gtx_xp_asym}
\end{figure}

Finally\marginpars{Actually I never did it because \shz{} wasn't implemented at the time, so now is
a good occasion}\todo*{Run shazoo on galaxy tree} we also compare \gtx{} with \bfs{} and \rst{} on
the task of nodes prediction using \shz{} algorithm~\autocite{Vitale2012}.

### 2017-06-14 lowstretch_and_cc.tex
add a previous document on galaxy tree experiments, plus small layout change

@@
-\documentclass[nobib, a4paper, 11pt, oneside, onecolumn, openany, notitlepage, final,
-svgnames,  marginals=raggedouter,
+\documentclass[nobib, a4paper, 10pt, oneside, onecolumn, openany, notitlepage, final,
+svgnames, marginals=raggedouter, english,
 %symmetric, justified,
-]{tufte-handout}
+]{article}
+\usepackage{charter}
 \setcounter{secnumdepth}{3}
@@
 \input{abbreviations}
-% \usepackage{charter}
-% \usepackage{fullpage}
 
@@
 \begin{document}
+\newgeometry{vmargin=1.5cm,left=1.5cm,textwidth=130mm,marginparsep=5mm,marginparwidth=52mm}
 \maketitle
@@
 time, signs are generated by an arbitrary adversary. However, in that case, we are still facing two
-limitations\marginpar{limitations is not the right word, maybe \emph{strong constraints}…}
+limitations\marginpars{limitations is not the right word, maybe \emph{strong constraints}…}
 \begin{enumerate}
@@
 		this bias is well suited to social networks, other applications may require other
-		bias\marginpar{namely the \pcc{} bias we'll introduce next}.
-	\item In the online setting, we evaluate our performance by the regret\marginpar{ref to regret equation},
+		bias\marginpars{namely the \pcc{} bias we'll introduce next}.
+	\item In the online setting, we evaluate our performance by the regret\marginpars{ref to regret equation},
 		whereas in general we are interested in more classical measures of predictive accuracy.
@@
 
-\paragraph{Experience on directed biological network}
-\label{par:experience_on_directed_biological_network}
+\subsubsection{Experience on directed biological network}
+\label{ssub:experience_on_directed_biological_network}
 
@@
 
-other domain? coref, images (is it directed?), entity resolution…\marginpar{Actually it's difficult
+other domain? coref, images (is it directed?), entity resolution…\marginpars{Actually it's difficult
 to find other kind of graph because all constructed ones (like image, coreference,
@@
 and \cite{Huffner2010} (\enquote{\textcite{Konig36} proved the following characterization of
-	balanced graphs. For a graph $G = (V , E)$, the following are equivalent:\marginpar{There is a
+	balanced graphs. For a graph $G = (V , E)$, the following are equivalent:\marginpars{There is a
 	proof in \autocite[p. 111]{BookKleinberg2010}, maybe I can rewrite it as well}
@@
 
+\iffalse
 Haris Angelidakis, Konstantin Makarychev, and Yury Makarychev. 2017.
@@
 improves over the one cited in the internship description
+\fi
 
@@
 \label{sub:gtx_state_of_the_art}
+
+\iffalse
 \url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
@@
 
+While they have many applications [see first paragraph of \url{https://arxiv.org/pdf/1401.2454.pdf},
+which was later merged in a STOC'14 paper] (a major one being solving linear system), in some
+practical situations their advantages are less clear [from
+\url{https://link.springer.com/chapter/10.1007/978-3-319-20086-6_16}\enquote{for reasonable inputs
+the constant factors make the solver much slower than methods with higher asymptotic complexity.
+One other aspect predicted by theory is confirmed by our findings: Spanning trees with lower
+stretch indeed reduce the solver's running time. Yet, simple spanning tree algorithms perform
+better in practice than those with a guaranteed low stretch.} this is improved by
+\url{https://link.springer.com/chapter/10.1007%2F978-3-319-20086-6_17} although they seem to work
+	mostly with the Laplacian of the tree ]
+\fi
+
 \subsection{Empirical evaluation}
-\label{sub:gtx_empiracal_evaluation}
+\label{sub:gtx_empirical_evaluation}
+
+\input{gtx_experiments}
 
@@
 \begingroup
+\small
 \setstretch{0.9}
+\todos
+
 \setlength\bibitemsep{2pt}

### 2017-06-19 gtx_algorithm.tex
start describing galaxy tree algorithm

We first describe two primitives we use to implement \gtx{}, namely \extractStar{} and
\collapseStar{}.

\extractStar{} takes as input a graph $G_t=(V_t, E_t)$, and optionally a \emph{threshold function}
$t_f$ or a \emph{degree function} $d_f$. While the graph is not exhausted, it repeatedly samples a
star center $c_i$, creates a star $S_i$ with $c_i$ neighbors, removes all the nodes of $S_i$ from
$V_t$ and all the edges incident to $S_i$ from $E_t$, and finally decrements the degree of the 2-hop
neighbors of $c_i$.\Todo{add a figure} Upon completion, it returns a list of stars and a mapping of
each node of $V_t$ to the unique star it belongs to. We consider three heuristics to choose centers:

\begin{itemize}
  \item choose the node with the current highest degree, with ties broken arbitrarily
  \item if $n_i$ is the number of node remaining in $V_t$ before choosing the \ith{} center, choose
    a node \uar{} among those with a degree larger than $t_f(n_i)$. Setting the threshold function
    to be the identity therefore recovers the previous strategy, but the idea here is to choose
    among a small set of high degree nodes, for instance by letting $t_f(n) = \sqrt{n}$
  \item if $\degr(u)$ is the degree of node $u$, choose node proportionally to $d_f(\degr(u))$, where
    again the degree function favors high degree nodes, for instance $d_f(\degr(u)) = \degr(u)^2$.
\end{itemize}

Let us analyze the complexity of the highest degree heuristic when $|V_t|=n$ and $|E_t|=m$. We first
build a priority queue of all the nodes sorted by their degree, which takes $O(n)$ time.  Then, at
each iteration of the inner loop, we find the center of the next star by extracting the maximum of
the queue, and we decrease the priority (\ie the degree) of all nodes adjacent to the new star.
Since both operations require constant time when using a Strict Fibonacci
Heap~\autocite{FibonacciHeaps12} and there are at most $n$ iterations of that loop, a coarse
approximation of the runtime of \extractStar{} is $O(n^2)$. However, observe that there can be at
most $m$ decrease operations (since degree $0$ nodes form a singleton star in constant time),
reducing the complexity to $O(m)$.

\bigskip

The second routine, \collapseStar{} takes as input the result of \extractStar{}, along with $E_t$
and an optional $\emph{eccentricity}$ array we will describe soon. It builds a new graph $G_{t+1}$
where each star becomes a node and there is a link between two nodes $s_1$ and $s_2$ if the nodes
making up $s_1$ and $s_2$ are connected in $E_t$. For that, we first shuffle $E_t$ and go through
it. When we  find an edge w hose endpoints belong to two different stars not yet connected, we use
that edge to connect these two stars. This trivially takes $O(m)$ times.

A variant instead keeps track of all edges connecting each pair of stars to choose one that will
best contribute to our low stretch objective. Namely, when connecting two stars, we would prefer to
join their centers rather than two peripheral points. For that we maintain an eccentricity count for
all of the nodes of the original $G_0$, which is incremented by $1$ each time a node is chosen to be
on the periphery of a star.\Todo{Show an example of a small 3 levels nested stars}
For each pair of stars, we thus choose the edge across them with minimal sum of its endpoints'
eccentricity. This requires another pass over the edges, preserving the $O(m)$ runtime.

\bigskip

We are then ready to describe the complete \gtx{} algorithm.\footnote{Note that for clarity, we
removed some bookkeeping, mainly related to maintaining mapping between nodes at different level of
contraction. Yet the full python implementation is available at
\url{https://github.com/daureg/magnet/blob/master/veverica/new_galaxy.py}.} Intuitively, until each
connected component is reduced to a single node, we first extract stars from the current graph $G_t$
and then collapse them to form $G_{t+1}$.

### 2017-06-20 gtx_algorithm.tex
add a figure of a star

@@
 $t_f$ or a \emph{degree function} $d_f$. While the graph is not exhausted, it repeatedly samples a
-star center $c_i$, creates a star $S_i$ with $c_i$ neighbors, removes all the nodes of $S_i$ from
-$V_t$ and all the edges incident to $S_i$ from $E_t$, and finally decrements the degree of the 2-hop
-neighbors of $c_i$.\Todo{add a figure} Upon completion, it returns a list of stars and a mapping of
+star center $c_i$, creates a star $S_i^t$ with $c_i$ neighbors, removes all the nodes of $S_i^t$ from
+$V_t$ and all the edges incident to $S_i^t$ from $E_t$, and finally decrements the degree of the 2-hop
+neighbors of $c_i$.
+\begin{marginfigure}
+  \centering
+  \includegraphics[height=0.15\textheight]{assets/tikz/gtx_star_tikz.pdf}
+  \caption[A sample star]{A sample star created at the $t^\mathrm{th}$ level. The black node
+    % \tikz{\node[vertex,rare] {$c_i$};}
+    is the center $c_i$ of the star $S_i^t$, which is made of the four light gray peripheral nodes
+  % \tikz{\node[vertex,medium] {$p_1$};} to \tikz{\node[vertex,medium] {$p_4$};}
+  as well as the solid edges. The 2-hops neighbors of $c_i$ are the white nodes
+  % \tikz{\node[vertex] {$h_1$};} to \tikz{\node[vertex] {$h_3$};}
+  $h_1$ to $h_3$, whose degree will decrease once we $S_i^t$ is removed from $G_t$.}
+  \label{fig:gtx_star_simple}
+\end{marginfigure}
+Upon completion, it returns a list of stars and a mapping of
 each node of $V_t$ to the unique star it belongs to. We consider three heuristics to choose centers:

### 2017-06-21 assets/tikz/gtx_eccentricity_tikz.tex
start giving an example of an execution of galaxy tree

\documentclass[svgnames,11pt,tikz]{standalone}
\input{preambule}
\begin{document}
\tikzstyle{star}=[circle,DodgerBlue,very thick,fill opacity=.2]
\tikzstyle{star-label}=[opacity=1, , font=\Large]
\tikzstyle{removed}=[Gray,line width=1.15pt]
\begin{tikzpicture}[auto,vertex/.append style={minimum width=8mm}]
  \draw[star,Orange] (2.260, -1.063) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$S_1^2$};
  \draw[star,Orange] (6.504, -9.728) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$S_2^2$};
  \draw[star,Orange] (11.641, 1.189) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$S_3^2$};

  \draw[star] (0.000, -0.000) circle [radius=1.8]  node[star-label,xshift=-30, yshift=25] {$S_1^1$};
  \draw[star] (1.689, -3.448) circle [radius=1.8]  node[star-label,xshift=-30, yshift=5] {$S_2^1$};
  \draw[star] (3.734, 0.896) circle [radius=1.8]  node[star-label,xshift=-30, yshift=5] {$S_3^1$};
  \draw[star] (4.245, -8.664) circle [radius=1.8]  node[star-label,xshift=-22, yshift=-25] {$S_4^1$};
  \draw[star] (5.934, -12.113) circle [radius=1.8]  node[star-label,xshift=-15, yshift=35] {$S_5^1$};
  \draw[star] (7.978, -7.768) circle [radius=1.8]  node[star-label,xshift=30, yshift=25] {$S_6^1$};
  \draw[star] (9.381, 2.252) circle [radius=1.8]  node[star-label,xshift=-15, yshift=30] {$S_7^1$};
  \draw[star] (11.071, -1.196) circle [radius=1.8]  node[star-label,xshift=-30, yshift=25] {$S_8^1$};
  \draw[star] (13.115, 3.149) circle [radius=1.8]  node[star-label,xshift=-30, yshift=20	] {$S_9^1$};

  \node[vertex,fill=black!0,] (n1) at (0.000, -0.000) {1};
  \node[vertex,fill=black!25,] (n2) at (0.223, -1.179) {2};
  \node[vertex,fill=black!25,] (n3) at (-1.179, -0.223) {3};
  \node[vertex,fill=black!25,] (n4) at (-0.223, 1.179) {4};
  \node[vertex,fill=black!25,] (n5) at (1.179, 0.223) {5};
  \node[vertex,fill=black!25,] (n6) at (1.689, -3.448) {6};
  \node[vertex,fill=black!50,] (n7) at (2.408, -4.409) {7};
  \node[vertex,fill=black!50,] (n8) at (0.728, -4.167) {8};
  \node[vertex,fill=black!50,] (n9) at (0.971, -2.487) {9};
  \node[vertex,fill=black!50,] (n10) at (2.650, -2.730) {10};
  \node[vertex,fill=black!25,] (n11) at (3.734, 0.896) {11};
  \node[vertex,fill=black!50,] (n12) at (4.391, -0.108) {12};
  \node[vertex,fill=black!50,] (n13) at (2.730, 0.240) {13};
  \node[vertex,fill=black!50,] (n14) at (3.077, 1.901) {14};
  \node[vertex,fill=black!50,] (n15) at (4.738, 1.553) {15};
  \node[vertex,fill=black!25,] (n16) at (4.245, -8.664) {16};
  \node[vertex,fill=black!50,] (n17) at (4.450, -9.846) {17};
  \node[vertex,fill=black!50,] (n18) at (3.062, -8.870) {18};
  \node[vertex,fill=black!50,] (n19) at (4.039, -7.482) {19};
  \node[vertex,fill=black!50,] (n20) at (5.427, -8.459) {20};
  \node[vertex,fill=black!50,] (n21) at (5.934, -12.113) {21};
  \node[vertex,fill=black!75,text=white] (n22) at (5.614, -13.269) {22};
  \node[vertex,fill=black!75,text=white] (n23) at (4.777, -11.793) {23};
  \node[vertex,fill=black!75,text=white] (n24) at (6.254, -10.956) {24};
  \node[vertex,fill=black!75,text=white] (n25) at (7.090, -12.433) {25};
  \node[vertex,fill=black!50,] (n26) at (7.978, -7.768) {26};
  \node[vertex,fill=black!75,text=white] (n27) at (8.018, -8.967) {27};
  \node[vertex,fill=black!75,text=white] (n28) at (6.779, -7.808) {28};
  \node[vertex,fill=black!75,text=white] (n29) at (7.939, -6.568) {29};
  \node[vertex,fill=black!75,text=white] (n30) at (9.178, -7.728) {30};
  \node[vertex,fill=black!25,] (n31) at (9.381, 2.252) {31};
  \node[vertex,fill=black!50,] (n32) at (8.964, 1.127) {32};
  \node[vertex,fill=black!50,] (n33) at (8.256, 2.669) {33};
  \node[vertex,fill=black!50,] (n34) at (9.799, 3.377) {34};
  \node[vertex,fill=black!50,] (n35) at (10.507, 1.835) {35};
  \node[vertex,fill=black!50,] (n36) at (11.071, -1.196) {36};
  \node[vertex,fill=black!75,text=white] (n37) at (11.078, -2.396) {37};
  \node[vertex,fill=black!75,text=white] (n38) at (9.871, -1.203) {38};
  \node[vertex,fill=black!75,text=white] (n39) at (11.064, 0.004) {39};
  \node[vertex,fill=black!75,text=white] (n40) at (12.271, -1.189) {40};
  \node[vertex,fill=black!50,] (n41) at (13.115, 3.149) {41};
  \node[vertex,fill=black!75,text=white] (n42) at (13.437, 1.993) {42};
  \node[vertex,fill=black!75,text=white] (n43) at (11.959, 2.827) {43};
  \node[vertex,fill=black!75,text=white] (n44) at (12.794, 4.305) {44};
  \node[vertex,fill=black!75,text=white] (n45) at (14.271, 3.470) {45};

  \draw[edge] (n1) edge [bend left =0] node[] {} (n2);
  \draw[edge] (n1) edge [bend left =0] node[] {} (n3);
  \draw[edge] (n1) edge [bend left =0] node[] {} (n4);
  \draw[edge] (n1) edge [bend left =0] node[] {} (n5);
  \draw[edge] (n2) edge [bend left =25] node[] {} (n6);
  \draw[edge,removed] (n2) edge [bend left =0] node[] {} (n9);
  \draw[edge,removed] (n3) edge [bend left =-20] node[] {} (n12);
  \draw[edge] (n4) edge [bend left =20] node[] {} (n14);
  % \draw[edge] (n3) edge [bend left =-20] node[] {} (n18);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n7);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n8);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n9);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n10);
  % \draw[edge] (n7) edge [bend left =-15] node[] {} (n43);
  % \draw[edge] (n9) edge [bend left =-10] node[] {} (n32);
  \draw[edge] (n8) .. controls ++(-40:3) and ++(240: 2) .. (n37);
  \draw[edge] (n11) edge [bend left =0] node[] {} (n12);
  \draw[edge] (n11) edge [bend left =0] node[] {} (n13);
  \draw[edge] (n11) edge [bend left =0] node[] {} (n14);
  \draw[edge] (n11) edge [bend left =0] node[] {} (n15);
  \draw[edge,removed] (n13) edge [bend left =0] node[] {} (n29);
  % \draw[edge] (n12) edge [bend left =0] node[] {} (n18);
  \draw[edge] (n15) .. controls ++(-40:3) and ++(64: 3.5) .. (n26);
  \draw[edge] (n16) edge [bend left =0] node[] {} (n17);
  \draw[edge] (n16) edge [bend left =0] node[] {} (n18);
  \draw[edge] (n16) edge [bend left =0] node[] {} (n19);
  \draw[edge] (n16) edge [bend left =0] node[] {} (n20);
  \draw[edge] (n17) edge [bend left =0] node[] {} (n23);
  %\draw[edge] (n19) edge [bend left =0] node[] {} (n27);
  \draw[edge] (n19) edge [bend left =0] node[] {} (n29);
  \draw[edge,removed] (n20) edge [bend left =0] node[] {} (n24);
  \draw[edge] (n21) edge [bend left =0] node[] {} (n22);
  \draw[edge] (n21) edge [bend left =0] node[] {} (n23);
  \draw[edge] (n21) edge [bend left =0] node[] {} (n24);
  \draw[edge] (n21) edge [bend left =0] node[] {} (n25);
  \draw[edge,removed] (n25) edge [bend left =-20] node[] {} (n42);
  \draw[edge] (n26) edge [bend left =0] node[] {} (n27);
  \draw[edge] (n26) edge [bend left =0] node[] {} (n28);
  \draw[edge] (n26) edge [bend left =0] node[] {} (n29);
  \draw[edge] (n26) edge [bend left =0] node[] {} (n30);
  \draw[edge] (n31) edge [bend left =0] node[] {} (n32);
  \draw[edge] (n31) edge [bend left =0] node[] {} (n33);
  \draw[edge] (n31) edge [bend left =0] node[] {} (n34);
  \draw[edge] (n31) edge [bend left =0] node[] {} (n35);
  \draw[edge] (n32) edge [bend left =0] node[] {} (n39);
  \draw[edge] (n35) edge [bend left =0] node[] {} (n43);
  \draw[edge,removed] (n33) edge [bend left =-25] node[] {} (n38);
  \draw[edge,removed] (n34) edge [bend left =20] node[] {} (n43);
  \draw[edge] (n36) edge [bend left =0] node[] {} (n37);
  \draw[edge] (n36) edge [bend left =0] node[] {} (n38);
  \draw[edge] (n36) edge [bend left =0] node[] {} (n39);
  \draw[edge] (n36) edge [bend left =0] node[] {} (n40);
  \draw[edge] (n41) edge [bend left =0] node[] {} (n42);
  \draw[edge] (n41) edge [bend left =0] node[] {} (n43);
  \draw[edge] (n41) edge [bend left =0] node[] {} (n44);
  \draw[edge] (n41) edge [bend left =0] node[] {} (n45);

  \draw[Green, line width=2pt] (11.8, 6.5) 
  .. controls ++(190:1) and ++(55: 2) .. (-2, 2.5)
  .. controls  ++(55: -4) and ++(170:4) .. (6, -15) 
  .. controls  ++(-10:9) and ++(10:9) .. node[pos=.4,font=\Huge,xshift=-15pt] {$S_1^3$} cycle ;
\end{tikzpicture}
\end{document}

### 2017-06-21 assets/tikz/gtx_run_level1_tikz.tex
start giving an example of an execution of galaxy tree

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
\begin{tikzpicture}[auto,vertex/.append style={minimum width=8mm}]
  \node[vertex,fill=DodgerBlue,text=black] (s1) at (0.000, -0.000) {$S_1^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s2) at (0.845, -1.724) {$S_2^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s3) at (1.867, 0.448) {$S_3^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s4) at (2.122, -4.332) {$S_4^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s5) at (2.967, -6.056) {$S_5^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s6) at (3.989, -3.884) {$S_6^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s7) at (4.691, 1.126) {$S_7^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s8) at (5.535, -0.598) {$S_8^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s9) at (6.558, 1.574) {$S_9^1$};

  \foreach \source/\dest in {1/2, 1/3, 2/8,3/6,4/5,4/6,7/8,7/9}
  \draw[edge] (s\source) -- (s\dest);
  \draw[edge] (s5) edge [bend left =-17]  (s9);
\end{tikzpicture}
\end{document}

### 2017-06-21 gtx_algorithm.tex
start giving an example of an execution of galaxy tree

@@
 
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.98\linewidth]{tikz/gtx_eccentricity_tikz.pdf}
+  \caption[The hierarchical structure of stars created by \gtx{}]{%
+    The execution of the \gtx{} algorithm. The original graph is made of the solid edges
+    connecting the nodes labeled by their id. Edges forming the final spanning tree are in black
+    while the others are in gray. The four shades of gray, from white to dark gray
+    denote increasing node eccentricity (as computed at the end of the algorithm). The \ith{} star
+    created during the \jth{} iteration of the algorithm is denoted $S_i^j$. Refer to the main text
+    for a complete walk through.}
+  \label{fig:gtx_eccentricity}
+\end{figure}
+
+\begin{figure}[bthp]
+  \centering
+  \begin{subfigure}[b]{0.47\textwidth}
+    \centering
+    \includegraphics[height=7cm]{tikz/gtx_run_level1_tikz}
+    \caption{Resulting graph after the first iteration}\label{fig:gtx_run1}
+  \end{subfigure}~
+  \begin{subfigure}[b]{0.47\textwidth}
+    \centering
+    \includegraphics[height=3cm]{tikz/gtx_run_level2_tikz}
+    \caption{Resulting graph after the second iteration}\label{fig:gtx_run2}
+    \vspace{\baselineskip}
+    \includegraphics[height=3cm]{tikz/gtx_run_level3_tikz}
+    \caption{Resulting graph after the third iteration}\label{fig:gtx_run3}
+  \end{subfigure}~
+  \caption{The other iterations of \gtx{}}\label{fig:gtx_run}
+\end{figure}
+
 \bigskip
+
+Let us start with the initial graph $G_0$ depicted in \autoref{fig:gtx_eccentricity} and initialize
+the eccentricity of all nodes to $0$. When running \extractStar{}, we see that the maximum degree is
+$4$, achieved at nodes $\{1, 6, 11, 16, 21, 26, 31, 36, 41\}$. For the sake of simplicity, assume
+nodes are picked according to their id. First, node $1$ is gonna form the star
+$\textcolor{DodgerBlue}{S_1^1}$ with peripheral nodes $2$, $3$, $4$ and $5$. This will increment the
+eccentricity of those peripheral nodes by $1$. Then node $6$ forms its star
+$\textcolor{DodgerBlue}{S_2^1}$ with $7$,$8$,$9$ and $10$. The process continues until node $41$ is
+chosen to be the center of star $\textcolor{DodgerBlue}{S_9^1}$, at which point $V_0$ has been
+exhausted and \extractStar{} terminates.
+
+We then call \collapseStar{}, with the eccentricity reducing variant. This will create all possible
+links between pairs of star. For instance, the edge between nodes $19$ and $29$ leads to the edge
+between $\textcolor{DodgerBlue}{S_4^1}$ and $\textcolor{DodgerBlue}{S_6^1}$. This was actually the
+only possible edge between $\textcolor{DodgerBlue}{S_4^1}$ and $\textcolor{DodgerBlue}{S_6^1}$.
+Consider on the other hand the case of edges $(2, 6)$ and $(2, 9)$. They both connect
+$\textcolor{DodgerBlue}{S_1^1}$ and $\textcolor{DodgerBlue}{S_2^1}$. Yet at this point in time, the
+eccentricity of node $2$ is $1$, the eccentricity of node $6$ is $0$ and the eccentricity of node
+$9$ is $1$. The edge $(2, 6)$ has therefore the smallest total eccentricity and is chosen to connect
+$\textcolor{DodgerBlue}{S_1^1}$ and $\textcolor{DodgerBlue}{S_2^1}$. The full result of the
+\collapseStar{} procedure can be seen on \autoref{fig:gtx_run1}.
 

### 2017-06-22 notations.tex
add a notations table

\begin{table*}[htpb]
  \centering
  \caption{List of notations used in this thesis}\label{tab:notations}%
  \begin{tabulary}{187mm}{LL}
    \toprule
    Symbol & Meaning \\
    \midrule
    $G$ & A graph, either directed or undirected \\
    $V$, $V_t$ & The set of all the nodes of a graph. Unless noted otherwise, nodes are indexed from $1$ to $|V|$ \\
    $E$, $E_t$ & The set of all the edges of a graph \\
    $\degr(u)$ & The total degree of node $u$ (that is the number of edges incident to $u$
    regardless of their direction) \\
    $(u,v)$, $(i,j)$ & An undirected edges between nodes $u$ and $v$ (or $i$ and $j$) \\
    $\mathcal{N}(u)$ & The set of all neighbors of $u$, regardless of direction. We thus have
    $|\mathcal{N}(u)| = \degr(u)$ \\
    \bottomrule
  \end{tabulary}
\end{table*}

### 2017-06-22 packages.tex
add a previous document on galaxy tree experiments, plus small layout change
start describing galaxy tree algorithm
add support files for tikz drawing
add a notations table

@@
 \usepackage[T1]{fontenc}
+\usepackage{csquotes}
+\usepackage{babel}
+\usepackage[heightrounded]{geometry}
 \usepackage{setspace}
+\usepackage{microtype}
 \usepackage[citestyle=alphabetic, bibstyle=ieee-alphabetic, isbn=false, maxnames=1, minnames=1,
@@
 \usepackage{xcolor}
+\usepackage{amsmath}
+\usepackage{sidenotes}
+\renewcommand{\footnote}[1]{\sidenote{\footnotesize #1}}
+\usepackage{enumitem}
+\usepackage{tabulary}
+\usepackage[marginpar]{todo}
+\usepackage[np,autolanguage]{numprint}
+\input{drawing}
+\usepackage{varioref}
 \usepackage{hyperref}
@@
     hypertexnames=true, pdfhighlight=/O,%nesting=true,%frenchlinks,%
-    urlcolor=Chocolate, linkcolor=DodgerBlue, citecolor=LimeGreen, %pagecolor=RoyalBlue,%
+    urlcolor=Chocolate, linkcolor=RoyalBlue, citecolor=LimeGreen, %pagecolor=RoyalBlue,%
 }
+\newcommand{\marginpars}[1]{\marginpar{\small#1}}
+\usepackage{caption}
+\usepackage[margin=0pt,font+=small,labelformat=parens,labelsep=space,
+skip=6pt,list=false,hypcap=false]{subcaption}
+\captionsetup{compatibility=false}
+\usepackage{graphicx}
+\usepackage{booktabs}
+\graphicspath{{./assets/}}
+\usepackage[capitalize,noabbrev]{cleveref}

### 2017-06-22 gtx_algorithm.tex
add pseudo code for extract star and more implementation details

@@
-We first describe two primitives we use to implement \gtx{}, namely \extractStar{} and
-\collapseStar{}.
+The \gtx{} algorithm takes as input a graph topology $G_0=(V_0, E_0)$ and produces a sequence of
+graphs $\{G_t\}_{t=1}^T$ of decreasing size until each connected component of $G_0$ is reduced to a
+single node. As we will prove later, after reaching this point, the algorithm has selected $|V_0| -
+1$ edges that form a spanning tree of $G_0$. $G_{t+1}$ is obtained from $G_t$ by composing two
+primitives so that we can informally write $G_{t+1} = \left(\collapseStar{} \circ
+\extractStar{}\right)(G_t)$.
+
+\extractStar{} partitions the graph $G_t$ into a set of stars and \collapseStar{} build the graph
+made of those stars using the edges in $E_t$. We provide more details on those two operations in the
+following, as well as their complexity analysis. Then we state formally the \gtx{} algorithm and prove
+its termination and correctness.  Finally, we study its properties, such as the number of iterations
+needed to finish and the stretch of the resulting tree. For simplicity and without loss of
+generality, we assume that $G_0$ consist of a single connected component.
+
+\medskip
 
 \extractStar{} takes as input a graph $G_t=(V_t, E_t)$, and optionally a \emph{threshold function}
-$t_f$ or a \emph{degree function} $d_f$. While the graph is not exhausted, it repeatedly samples a
-star center $c_i$, creates a star $S_i^t$ with $c_i$ neighbors, removes all the nodes of $S_i^t$ from
-$V_t$ and all the edges incident to $S_i^t$ from $E_t$, and finally decrements the degree of the 2-hop
-neighbors of $c_i$.
+$t_f$ or a \emph{degree function} $d_f$. While the nodeset $V_t$ is not exhausted, it repeatedly samples a
+star center $c_i$, creates a star $S_i^t$ with the neighbors of $c_i$, removes all the nodes of $S_i^t$ from
+$V_t$ and all the edges incident to $S_i^t$ from $E_t$, and finally decrements accordingly the
+degree of the 2-hop neighbors of $c_i$ (see \autoref{fig:gtx_star_simple} for a visual
+representation of this notation).
 \begin{marginfigure}
@@
   \includegraphics[height=0.15\textheight]{assets/tikz/gtx_star_tikz.pdf}
-  \caption[A sample star]{A sample star created at the $t^\mathrm{th}$ level. The black node
+  \caption[A sample star]{A sample star created during the \tth{} extraction level. The black node
     % \tikz{\node[vertex,rare] {$c_i$};}
@@
 
-\begin{itemize}
+\begin{itemize}%[nosep]
   \item choose the node with the current highest degree, with ties broken arbitrarily
@@
     among a small set of high degree nodes, for instance by letting $t_f(n) = \sqrt{n}$
-  \item if $\degr(u)$ is the degree of node $u$, choose node proportionally to $d_f(\degr(u))$, where
-    again the degree function favors high degree nodes, for instance $d_f(\degr(u)) = \degr(u)^2$.
+  \item if $\degr(u)$ is the degree of node $u$, choose node proportionally to $d_f(\degr(u))$.
+    Again, the degree function is designed so that it favors the selection of high degree nodes. For
+    instance, one could use $d_f(\degr(u)) = \degr(u)^2$.
 \end{itemize}
 
-Let us analyze the complexity of the highest degree heuristic when $|V_t|=n$ and $|E_t|=m$. We first
-build a priority queue of all the nodes sorted by their degree, which takes $O(n)$ time.  Then, at
-each iteration of the inner loop, we find the center of the next star by extracting the maximum of
-the queue, and we decrease the priority (\ie the degree) of all nodes adjacent to the new star.
-Since both operations require constant time when using a Strict Fibonacci
-Heap~\autocite{FibonacciHeaps12} and there are at most $n$ iterations of that loop, a coarse
-approximation of the runtime of \extractStar{} is $O(n^2)$. However, observe that there can be at
-most $m$ decrease operations (since degree $0$ nodes form a singleton star in constant time),
-reducing the complexity to $O(m)$.
+We now give the pseudo code of \extractStar{} for the highest degree variant.\footnote{Note that for
+clarity, we removed some bookkeeping code in all listings, mainly the part related to maintaining
+mapping between nodes at different level of contraction. However, the full python implementation
+is available at \url{https://github.com/daureg/magnet/blob/master/veverica/new_galaxy.py\#L27}.}
+We assume that $G$ is the adjacency list of the graph, so that $G[u]$ is the set of neighbors of
+$u$, \ie{} $G[u] \equiv \mathcal{N}(u)$. The other piece of notation is $\textsc{Star}$, which
+simply create a star given a center and a list of peripheral nodes.  \vspace{-\baselineskip}
 
-\bigskip
+\begin{center}
+  \rule{\textwidth}{.3pt}
+  \begin{algorithmic}[1]
+    \Function{\extractStar{}}{$G_t=(V_t,E_t)$}
+      \State Let $Q$ be a max-priority queue. The key of element $x$ is $Q[x]$
+      \Let{$stars$}{[]}
+      \Let{$remaining$}{$\emptyset$}
+      \ForAll{node $u$ in $V_t$}
+        % \State $\textsc{Insert}\\left(Q,\,u\right)$ \Comment with the key $\degr(u)$
+        \State \Call{Insert}{$Q,\,u$} \Comment with the key $\degr(u)$
+        \Let{$remaining$}{$remaining \bigcup \left\{u\right\}$}
+      \EndFor
+      \While{$Q$ is not empty}
+        \Let{$c_i$}{\Call{Extract-Max}{$Q$}}
+        \If{$c_i$ not in $remaining$}
+          \State \textbf{continue} \Comment{$c_i$ is part of an existing star so there is
+          nothing to do}
+        \EndIf
+        \Let{$periphery$}{$G[c_i] \bigcap remaining$}
+        \Let{$stars$}{$stars \bigcap $\Call{Star}{$c_i,\, periphery$}}
+        \Let{$remaining$}{$remaining \setminus\left\{c_i\right\} \setminus periphery$}
+        \For{$p$ in $periphery$}
+          \For{$h$ in $G[p] \bigcap remaining$}
+            \State \Call{Decrease-Key}{$Q,\, h,\, Q[h]-1$}
+          \EndFor
+        \EndFor
+      \EndWhile
+      \State \textbf{return} $stars$
+    \EndFunction
+  \end{algorithmic}
+  \rule{\textwidth}{.3pt}
+\end{center}
+
+\extractStar{} terminates because at each iteration of the while loop line 8, we remove one node
+from $Q$ and never add any. Let us analyze the complexity when $|V_t|=n$ and $|E_t|=m$. We first
+build a priority queue of all the nodes sorted by their degree (line 5--7), which takes $O(n)$ time.
+Then, at each iteration of the inner loop, we find the center of the next star by extracting the
+maximum of the queue (line 9), we build the corresponding star (line 12--14) and we decrease the
+priority (\ie the degree) of all nodes adjacent to the new star (line 15--17).  Since both
+operations require constant time when using a Strict Fibonacci Heap~\autocite{FibonacciHeaps12} and
+there are $n$ iterations of that loop, a coarse approximation of the runtime of \extractStar{} is
+$O(n^2)$. However, observe that there can be at most $m$ decrease operations (since after that, all
+nodes still in the queue have an effective degree of $0$, meaning that $periphery$ will the empty
+set and lines 13--17 will run in constant time), reducing the complexity to $O(m+n)$.
+
+The other two variants are more time consuming because they require additional bookkeeping. Their
+randomization make them useful in an adversarial context but it also renders their analysis more
+challenging, not necessarily for the runtime of \extractStar{} but mostly for the tree construction.
+Therefore, we only briefly describe the implementations here.\footnote{Although they are available
+online at 
+\nolinkurl{https://github.com/daureg/magnet/blob/master/veverica/}%
+\{\href{https://github.com/daureg/magnet/blob/master/veverica/ThresholdSampler.py}%
+{ThresholdSampler.py}, \href{https://github.com/daureg/magnet/blob/master/veverica/NodeSampler.py}%
+{NodeSampler.py}\}.} For the threshold function, we
+maintain two queues, $high$ and $low$, containing nodes whose degree is respectively above and below
+the current threshold. We select a node \uar{} in $high$, remove the corresponding star from $G_t$,
+recompute the new threshold and if necessary, move nodes which fell under the threshold from $high$
+to $low$ and those who climb above the threshold from $low$ to $high$. For the degree function, we
+can draw any node as center proportionally to its weight (where the weight of node $u$ is defined as
+$d_f\left(\degr(u)\right)$), but we cannot use the standard method of computing the cumulative sum
+of weights since each iteration change some of them. Therefore, we construct a binary tree whose
+leaves are the nodes of $V_t$ and where each tree nodes maintain the sum of weights in its left and
+right subtrees. To sample, we draw a random number between $0$ and the total weight of the tree.
+When degrees are updated (or graph node removed), we update the weights along a path from the
+corresponding leaves to the root of the tree.
+
+\medskip
 
@@
 making up $s_1$ and $s_2$ are connected in $E_t$. For that, we first shuffle $E_t$ and go through
-it. When we  find an edge w hose endpoints belong to two different stars not yet connected, we use
+it. When we find an edge whose endpoints belong to two different stars not yet connected, we use
 that edge to connect these two stars. This trivially takes $O(m)$ times.
@@
 all of the nodes of the original $G_0$, which is incremented by $1$ each time a node is chosen to be
-on the periphery of a star.\Todo{Show an example of a small 3 levels nested stars}
+on the periphery of a star.\Todo{link that to the walk through example.}
 For each pair of stars, we thus choose the edge across them with minimal sum of its endpoints'
@@
   \centering
-  \includegraphics[width=0.98\linewidth]{tikz/gtx_eccentricity_tikz.pdf}
+  \includegraphics[width=0.78\linewidth]{tikz/gtx_eccentricity_tikz.pdf}
   \caption[The hierarchical structure of stars created by \gtx{}]{%
     The execution of the \gtx{} algorithm. The original graph is made of the solid edges
-    connecting the nodes labeled by their id. Edges forming the final spanning tree are in black
+    connecting the nodes labeled by  their index. Edges forming the final spanning tree are in black
     while the others are in gray. The four shades of gray, from white to dark gray
@@
     \centering
-    \includegraphics[height=7cm]{tikz/gtx_run_level1_tikz}
+    \includegraphics[height=5cm]{tikz/gtx_run_level1_tikz}
     \caption{Resulting graph after the first iteration}\label{fig:gtx_run1}
@@
     \centering
-    \includegraphics[height=3cm]{tikz/gtx_run_level2_tikz}
+    \includegraphics[height=2.2cm]{tikz/gtx_run_level2_tikz}
     \caption{Resulting graph after the second iteration}\label{fig:gtx_run2}
     \vspace{\baselineskip}
-    \includegraphics[height=3cm]{tikz/gtx_run_level3_tikz}
+    \includegraphics[height=2.2cm]{tikz/gtx_run_level3_tikz}
     \caption{Resulting graph after the third iteration}\label{fig:gtx_run3}
@@
 
-\bigskip
+\medskip
 
+We illustrate the operation of the \gtx{} algorithm on small (and somewhat contrived) example.
 Let us start with the initial graph $G_0$ depicted in \autoref{fig:gtx_eccentricity} and initialize
@@
 $4$, achieved at nodes $\{1, 6, 11, 16, 21, 26, 31, 36, 41\}$. For the sake of simplicity, assume
-nodes are picked according to their id. First, node $1$ is gonna form the star
-$\textcolor{DodgerBlue}{S_1^1}$ with peripheral nodes $2$, $3$, $4$ and $5$. This will increment the
+nodes are picked according to their index. First, node $1$ is forms the star
+$\textcolor{DodgerBlue}{S_1^1}$ with peripheral nodes $2$, $3$, $4$ and $5$. This increments the
 eccentricity of those peripheral nodes by $1$. Then node $6$ forms its star
-$\textcolor{DodgerBlue}{S_2^1}$ with $7$,$8$,$9$ and $10$. The process continues until node $41$ is
-chosen to be the center of star $\textcolor{DodgerBlue}{S_9^1}$, at which point $V_0$ has been
-exhausted and \extractStar{} terminates.
+$\textcolor{DodgerBlue}{S_2^1}$ with $7$, $8$, $9$ and $10$. The process continues until node $41$ is
+chosen to be the center of star $\textcolor{DodgerBlue}{S_9^1}$, at which point the max-priority
+queue has been exhausted and \extractStar{} finishes.
 
-We then call \collapseStar{}, with the eccentricity reducing variant. This will create all possible
-links between pairs of star. For instance, the edge between nodes $19$ and $29$ leads to the edge
-between $\textcolor{DodgerBlue}{S_4^1}$ and $\textcolor{DodgerBlue}{S_6^1}$. This was actually the
+We then call \collapseStar{}, with the eccentricity reducing variant. This will connect all possible
+pairs of star. For instance, the edge between nodes $19$ and $29$ leads to the edge
+between $\textcolor{DodgerBlue}{S_4^1}$ and $\textcolor{DodgerBlue}{S_6^1}$. This is actually the
 only possible edge between $\textcolor{DodgerBlue}{S_4^1}$ and $\textcolor{DodgerBlue}{S_6^1}$.
 Consider on the other hand the case of edges $(2, 6)$ and $(2, 9)$. They both connect
-$\textcolor{DodgerBlue}{S_1^1}$ and $\textcolor{DodgerBlue}{S_2^1}$. Yet at this point in time, the
-eccentricity of node $2$ is $1$, the eccentricity of node $6$ is $0$ and the eccentricity of node
+$\textcolor{DodgerBlue}{S_1^1}$ and $\textcolor{DodgerBlue}{S_2^1}$. Yet at this point of the algorithm,
+the eccentricity of node $2$ is $1$, the eccentricity of node $6$ is $0$ and the eccentricity of node
 $9$ is $1$. The edge $(2, 6)$ has therefore the smallest total eccentricity and is chosen to connect
@@
 \collapseStar{} procedure can be seen on \autoref{fig:gtx_run1}.
-
-We are then ready to describe the complete \gtx{} algorithm.\footnote{Note that for clarity, we
-removed some bookkeeping, mainly related to maintaining mapping between nodes at different level of
-contraction. Yet the full python implementation is available at
-\url{https://github.com/daureg/magnet/blob/master/veverica/new_galaxy.py}.} Intuitively, until each
-connected component is reduced to a single node, we first extract stars from the current graph $G_t$
-and then collapse them to form $G_{t+1}$.

### 2017-06-23 gtx_algorithm.tex
add missing pseudo code and finish example execution

@@
-The \gtx{} algorithm takes as input a graph topology $G_0=(V_0, E_0)$ and produces a sequence of
+The \gtx{} algorithm takes as input an unweighted graph topology $G_0=(V_0, E_0)$ and produces a sequence of
 graphs $\{G_t\}_{t=1}^T$ of decreasing size until each connected component of $G_0$ is reduced to a
@@
 
+\paragraph{\extractStar{}}\label{par:extractstar}%
 \extractStar{} takes as input a graph $G_t=(V_t, E_t)$, and optionally a \emph{threshold function}
@@
       \State Let $Q$ be a max-priority queue. The key of element $x$ is $Q[x]$
-      \Let{$stars$}{[]}
-      \Let{$remaining$}{$\emptyset$}
+      \Let{$stars$, $inner\_edges$, $remaining$}{$[]$, $\emptyset$, $\emptyset$}
       \ForAll{node $u$ in $V_t$}
-        % \State $\textsc{Insert}\\left(Q,\,u\right)$ \Comment with the key $\degr(u)$
         \State \Call{Insert}{$Q,\,u$} \Comment with the key $\degr(u)$
@@
         \Let{$stars$}{$stars \bigcap $\Call{Star}{$c_i,\, periphery$}}
+        \Let{$inner\_edges$}{$inner\_edges \bigcup \{(c_i, p): p \in periphery\}$}
         \Let{$remaining$}{$remaining \setminus\left\{c_i\right\} \setminus periphery$}
@@
       \EndWhile
-      \State \textbf{return} $stars$
+      \State \textbf{return} $stars$, $inner\_edges$, $membership$%
+      \Comment{$membership$ maps the nodes of $V_t$ to the index of the star they belong to}
     \EndFunction
@@
 challenging, not necessarily for the runtime of \extractStar{} but mostly for the tree construction.
+\marginpars{Which means this paragraph can be skipped at first.}
 Therefore, we only briefly describe the implementations here.\footnote{Although they are available
@@
 
+\paragraph{\collapseStar{}}\label{par:collapsestar}%
 The second routine, \collapseStar{} takes as input the result of \extractStar{}, along with $E_t$
-and an optional $\emph{eccentricity}$ array we will describe soon. It builds a new graph $G_{t+1}$
+and an optional $\emph{eccentricity}$ array we will describe soon\marginpars{Actually, we could make
+eccentricity the default behavior, which would simplify the pseudo code}. It builds a new graph $G_{t+1}$
 where each star becomes a node and there is a link between two nodes $s_1$ and $s_2$ if the nodes
-making up $s_1$ and $s_2$ are connected in $E_t$. For that, we first shuffle $E_t$ and go through
-it. When we find an edge whose endpoints belong to two different stars not yet connected, we use
-that edge to connect these two stars. This trivially takes $O(m)$ times.
+making up $s_1$ and $s_2$ are connected in $E_t$.
 
-A variant instead keeps track of all edges connecting each pair of stars to choose one that will
-best contribute to our low stretch objective. Namely, when connecting two stars, we would prefer to
-join their centers rather than two peripheral points. For that we maintain an eccentricity count for
-all of the nodes of the original $G_0$, which is incremented by $1$ each time a node is chosen to be
-on the periphery of a star.\Todo{link that to the walk through example.}
-For each pair of stars, we thus choose the edge across them with minimal sum of its endpoints'
-eccentricity. This requires another pass over the edges, preserving the $O(m)$ runtime.
+\begin{center}
+  \rule{\textwidth}{.3pt}
+  \begin{algorithmic}[1]
+    \Function{\collapseStar{}}{$G_t=(V_t,E_t),\,membership,\,eccentricity$}
+      \State Let $G_{t+1}$ be an empty graph
+      \State Let $cross\_edges$ be a HashMap mapping edges in $G_{t+1}$ to set of edges in $G_t$
+      \ForAll{edge $(u, v)$ in \Call{Shuffle}{$E_t$}}
+        \Let{$s_u,\,s_v$}{$membership[u]$, $membership[v]$}%
+        \Comment{swap if needed so that $s_u < s_v$}
+        \If{$s_u = s_v$}
+          \State \textbf{continue}
+        \EndIf
+        \If{$eccentricity$ is not \textbf{null}}
+          \Let{$cross\_edges[(s_u, s_v)]$}{$cross\_edges[(s_u, s_v)] \bigcup \{(u, v)\}$}
+          % \State $cross\_edges[(s_u, s_v)].$\Call{Add}{$(u, v)$}
+        \ElsIf{$(s_u, s_v)$ not in $cross\_edges$}
+          \Let{$cross\_edges[(s_u, s_v)]$}{$\{(u, v)\}$}
+          \State Add edge $(s_u, s_v)$ to $G_{t+1}$
+        \EndIf
+      \EndFor
+      \If{$eccentricity$ is not \textbf{null}}
+        \ForAll{$\Big(\underbrace{(u,v)}_{\text{key}},\,
+          \underbrace{candidate\_edges}_{\text{value}}\Big)$
+          in $cross\_edges.$\Call{Items}{$\,$}}
+          \Let{$(u_0, v_0)$}{$\argmin_{(u,v)\in candidate\_edges} eccentricity[u]+eccentricity[v]$}
+          \Let{$cross\_edges[(s_u, s_v)]$}{$\{(u_0, v_0)\}$}
+          \State Add edge $(s_u, s_v)$ to $G_{t+1}$
+        \EndFor
+      \EndIf
+      \State \textbf{return} $G_{t+1}$, $cross\_edges$
+    \EndFunction
+  \end{algorithmic}
+  \rule{\textwidth}{.3pt}
+\end{center}
+
+For that, we first shuffle $E_t$ and go through it (line 4). When we find an edge whose endpoints
+belong to two different stars not yet connected (line 10), we use that edge to connect these two
+stars (line 11--12) This trivially takes $O(m)$ times.  A variant instead keeps track of all edges
+connecting each pair of stars (line 9) to choose one that will best contribute to our low stretch
+objective. Namely, when connecting two stars, we would prefer to join their centers rather than two
+peripheral points. For that we maintain an eccentricity count for all of the nodes of the original
+$G_0$, which is incremented by $1$ each time a node is chosen to be on the periphery of a
+star.\footnote{We will discuss this construction in more details when describing the main \gtx{}
+function since it takes place there.} For each pair of stars, we thus choose the edge across them
+with minimal sum of its endpoints' eccentricity (line 15--17)\marginpars{Line 15 is actually a
+simplification since the $eccentricity$ array is only defined for nodes in $V_0$ so when
+\collapseStar{} is called on $G_3$ for instance, $u$ and $v$ are nodes of $V_3$ and we have to
+retrieve the actual corresponding edge in $E_0$. Maybe I could simply add a HashMap from $E_t$ to
+$E_0$ like in the python code…}. This only requires another pass over the edges, thus preserving the
+$O(m)$ runtime.
+
+\paragraph{Putting the pieces together}\label{par:full_gtx}%
+\extractStar{} and \collapseStar{} are truly the core the of \gtx{} algorithm but to obtain our
+final spanning tree, we need additional works, namely updating the eccentricity of the nodes of
+$V_0$ and keeping track of each edge within and between along every contraction level.
+
+\begin{algorithm}
+  \caption{\gtx{}($G_0=(V_0,E_0)$) \label{alg:gtx}}
+	\begin{algorithmic}[1]
+    \State Let $eccentricity$ be an array of size $|V_0|$ initially all set to $0$
+    \Let{$G_t$}{$G_0$}
+    \Let{$inner\_edges\_seq$, $outer\_edges\_seq$}{$[]$, $[]$}
+    \Repeat
+      \Let{$stars$, $inner\_edges$, $membership$}{\Call{Extract-Stars}{$G_t$}}
+      \State $inner\_edges\_seq$.\Call{Append}{$inner\_edges$}
+      \State \Call{Update-Eccentricity}{$stars$, $eccentricity$, $original\_node$}
+      \Let{$full\_membership$, $original\_node$}{\Call{Update-Nodes-Mapping}{$full\_membership$, $star\_membership$}}
+      \Let{$G_{t+1}$, $\!outer\_edges$}{\Call{Collapse-Stars}{$E_t \!\setminus\! inner\_edges$,
+      $\!membership$, $\!eccentricity$}}
+      \State $outer\_edges\_seq$.\Call{Append}{$outer\_edges$}
+      \Let{$G_{t}$}{$G_{t+1}$}
+    \Until{$|outer\_edges|>0$}
+    \State \textbf{return} \Call{Assemble-Spanning-Trees}{$inner\_edges\_seq$, $outer\_edges\_seq$}
+		\begin{center}
+      \vspace{-.5\baselineskip}
+			\rule{0.5\textwidth}{.2pt}
+		\end{center}
+    \includegraphics{assets/tmp-code/update_eccentricity.pdf}
+    \includegraphics{assets/tmp-code/update_nodes_mapping.pdf}
+    \includegraphics{assets/tmp-code/assemble_spanning_tree.pdf}
+	\end{algorithmic}
+\end{algorithm}
+
+Basically, at every collapse level, we first extract stars from the current graph (line 5), then
+update the eccentricity and nodes mapping (lines 7--8) and finally collapse the graph (line 9) We
+perform these operations until there are no edge connecting stars anymore. At this point, we revisit
+every outer edges to form the spanning tree (line 13). Because the main loop does not perform any
+work besides calling other function, assuming there are T collapse steps, the complexity of \gtx{}
+is $O(T(m+n)) = O(Tm)$. \marginpars{I didn't write pseudo code for the three helping functions but
+instead put their verbatim python version. They are mostly plumbing code, the only interesting
+aspect is their complexity.  \texttt{update\_centrality} and \texttt{update\_nodes\_mapping} takes
+$O(n_0)$ time since they go through every node of the original graph. As for
+\texttt{assemble\_spanning\_tree}, it goes again through every edges visited at every level of
+contraction, meaning it's $O(Tm)$. However, we already visited those edges at least once so it only
+adds a constant factor $2$ in the overall complexity of the \gtx{} algorithm.}
+
+\paragraph{Exemple of \gtx{}}
+\label{par:exemple_of_gtx}
+
+We illustrate the operation of the \gtx{} algorithm on small (and somewhat contrived) example.
+Let us start with the initial graph $G_0$ depicted in \autoref{fig:gtx_eccentricity}
+\vpageref{fig:gtx_eccentricity} and initialize
+the eccentricity of all nodes to $0$. When running \extractStar{}, we see that the maximum degree is
+$4$, achieved at nodes $\{1, 6, 11, 16, 21, 26, 31, 36, 41\}$. For the sake of simplicity, assume
+nodes are picked according to their index. First, node $1$ is forms the star
+$\textcolor{DodgerBlue}{S_1^1}$ with peripheral nodes $2$, $3$, $4$ and $5$. This increments the
+eccentricity of those peripheral nodes by $1$. Then node $6$ forms its star
+$\textcolor{DodgerBlue}{S_2^1}$ with $7$, $8$, $9$ and $10$. The process continues until node $41$ is
+chosen to be the center of star $\textcolor{DodgerBlue}{S_9^1}$, at which point the max-priority
+queue has been exhausted and \extractStar{} finishes.
 
@@
 
-\medskip
-
-We illustrate the operation of the \gtx{} algorithm on small (and somewhat contrived) example.
-Let us start with the initial graph $G_0$ depicted in \autoref{fig:gtx_eccentricity} and initialize
-the eccentricity of all nodes to $0$. When running \extractStar{}, we see that the maximum degree is
-$4$, achieved at nodes $\{1, 6, 11, 16, 21, 26, 31, 36, 41\}$. For the sake of simplicity, assume
-nodes are picked according to their index. First, node $1$ is forms the star
-$\textcolor{DodgerBlue}{S_1^1}$ with peripheral nodes $2$, $3$, $4$ and $5$. This increments the
-eccentricity of those peripheral nodes by $1$. Then node $6$ forms its star
-$\textcolor{DodgerBlue}{S_2^1}$ with $7$, $8$, $9$ and $10$. The process continues until node $41$ is
-chosen to be the center of star $\textcolor{DodgerBlue}{S_9^1}$, at which point the max-priority
-queue has been exhausted and \extractStar{} finishes.
-
 We then call \collapseStar{}, with the eccentricity reducing variant. This will connect all possible
@@
 $\textcolor{DodgerBlue}{S_1^1}$ and $\textcolor{DodgerBlue}{S_2^1}$. The full result of the
-\collapseStar{} procedure can be seen on \autoref{fig:gtx_run1}.
+\collapseStar{} procedure is $G_1$, which can be seen on \autoref{fig:gtx_run1}.
+
+We now run \extractStar{} on $G_1$. Because all nodes have degree $2$, they could all be chosen to
+be the center of a star yet we again consider them ordered by their index and therefore choose
+$\textcolor{DodgerBlue}{S_1^1}$ to be the center of the star $\textcolor{Orange}{S_1^2}$ with
+peripheral nodes $\textcolor{DodgerBlue}{S_2^1}$ and $\textcolor{DodgerBlue}{S_3^1}$. The original
+nodes belonging to those peripheral stars (nodes $4$ to $15$) have their eccentricity incremented by
+$1$. The next node with highest degree in $G_1$ is now $\textcolor{DodgerBlue}{S_4^1}$, which forms
+a star with $\textcolor{DodgerBlue}{S_5^1}$ and $\textcolor{DodgerBlue}{S_6^1}$. This choice means
+that nodes $21$ through $30$ have their eccentricity incremented by $1$. Finally,
+$\textcolor{DodgerBlue}{S_7^1}$ forms the last star with $\textcolor{DodgerBlue}{S_8^1}$ and
+$\textcolor{DodgerBlue}{S_9^1}$. Then \collapseStar{} connects the resulting three stars, and this
+time there is only a single choice between each pair of stars, leading to the graph $G_2$ showed in
+\autoref{fig:gtx_run2}
+
+The action of \extractStar{} on $G_2$ is quite simple, because there is only one star that can be
+created, so let say we choose $\textcolor{Orange}{S_1^2}$ as its center, with
+$\textcolor{Orange}{S_2^2}$ and $\textcolor{Orange}{S_3^2}$ as peripheral nodes.  This increases the
+eccentricity of their underlying $G_0$ nodes by $1$ (namely nodes $16$ to $45$).  Because there is
+only one star $\textcolor{Green}{S_1^3}$ left, \collapseStar{} returns $G_3$ showed in
+\autoref{fig:gtx_run3} and an empty list of outer edges, meaning that the inner loop of \gtx{} is
+finished and we can go through every edges we chose between stars at every level to recover the
+final spanning tree, showed with black edges in \autoref{fig:gtx_eccentricity}
+\vpageref{fig:gtx_eccentricity}. For completeness, we can also look at the edges which are not in
+the spanning tree and therefore contribute to the stretch of the tree. In that case the average
+stretch is $7$, as showed in \autoref{tab:gtx_example_stretch}.
+
+\begin{table}[htpb]
+  \centering
+  \caption{Stretch of the example tree}
+  \label{tab:gtx_example_stretch}
+  \begin{tabulary}{\linewidth}{lLr}
+    \toprule
+    test edge & path in the tree & length \\
+    \midrule
+    $2,9$   & $2$--$6$--$9$                       & $2$ \\
+    $3,12$  & $3$--$1$--$4$--$14$--$11$--$12$     & $6$ \\
+    $13,29$ & $13$--$11$--$15$--$26$--$29$        & $4$ \\
+    $20,24$ & $20$--$16$--$17$--$23$--$21$--$24$  & $5$ \\
+    $25,42$ & $25$--$21$--$23$--$17$--$16$--$19$--$29$--$26$--$15$--$11$--$14$--$4$--$1$--$2$--$6$--$8$--$37$--$36$--$39$-- $32$--$31$--$35$--$43$--$41$--$45$ & $24$ \\
+    $33,38$ & $33$--$31$--$32$--$39$--$36$--$38$  & $5$ \\
+    $34,43$ & $34$--$31$--$35$--$43$              & $3$ \\
+    \bottomrule
+  \end{tabulary}
+\end{table}
+
+\paragraph{Number of iterations needed}\label{par:number_of_iteration}%
+\textcolor{red}{\LARGE First give the example of line graph, say it's worst case because of barbell
+graph (use simple inkscape figures or paper), then tree and its longest path from root, then graph
+and spanning tree}

### 2017-06-25 gtx_algorithm.tex
termination and draft of number of iterations

@@
 
-Basically, at every collapse level, we first extract stars from the current graph (line 5), then
-update the eccentricity and nodes mapping (lines 7--8) and finally collapse the graph (line 9) We
-perform these operations until there are no edge connecting stars anymore. At this point, we revisit
-every outer edges to form the spanning tree (line 13). Because the main loop does not perform any
-work besides calling other function, assuming there are T collapse steps, the complexity of \gtx{}
-is $O(T(m+n)) = O(Tm)$. \marginpars{I didn't write pseudo code for the three helping functions but
+As described in \autoref{alg:gtx}, \marginpars{I didn't write pseudo code for the three helping functions but
 instead put their verbatim python version. They are mostly plumbing code, the only interesting
@@
 adds a constant factor $2$ in the overall complexity of the \gtx{} algorithm.} 
+at every collapse level, we first extract stars from the current graph (line 5), then
+update the eccentricity and nodes mapping (lines 7--8) and finally collapse the graph (line 9) We
+perform these operations until there are no edge connecting stars anymore. At this point, we revisit
+every outer edges to form the spanning tree (line 13). Because the main loop does not perform any
+work besides calling other function, assuming there are T collapse steps, the complexity of \gtx{}
+is $O(T(m+n)) = O(Tm)$.
+
+Later we hope to found an upper bound of $T$ in terms of $m$, and also to refine the analysis to
+leverage the fact that $\sum_{t=0}^T |E_t|$ is significantly smaller than $Tm$. For now, let us
+prove the termination of \gtx{} by observing that, assuming $|E_t|>0$, $|E_{t+1}| < |E_t|$ since
+at least two nodes in $G_t$ are connected and will therefore form a star, which will remove all the
+inner edges of that star. As for the fact that we get a tree, note that a star is a tree, therefore
+a star of star is a tree, and so on.\Todo{make the two arguments about \gtx{} correctness more
+formal, and also explain that since all nodes are part of the ultimate star, this tree is a spanning
+one. Motivate the name of \gtx{}.}
 
@@
 \paragraph{Number of iterations needed}\label{par:number_of_iteration}%
-\textcolor{red}{\LARGE First give the example of line graph, say it's worst case because of barbell
-graph (use simple inkscape figures or paper), then tree and its longest path from root, then graph
-and spanning tree}
+
+\textcolor{red}{\LARGE Draft}
+
+A crucial quantity of the \gtx{} algorithm, both in terms of complexity and resulting stretch, is
+the number of collapses $T$ needed before termination. While this is still elusive to express in the
+general case, let us first look at some simple cases. For instance, a very sparse example of graph
+is the line graph. While it is already a tree, let us look how the \gtx{} algorithm operates on it.
+Say we have $n$ nodes in that line. There a star is made at most of three consecutive nodes. In the
+worst case, centers will be chosen such that we have a succession of three and two nodes stars (as
+in \autoref{fig:gtx_line_graph}).%
+\begin{marginfigure}
+  \centering
+  \includegraphics[width=0.9\linewidth]{assets/tmp-code/line_graph.pdf}
+  \caption{A line graph with stars in red}
+  \label{fig:gtx_line_graph}
+\end{marginfigure}
+This will results in $\nicefrac{2n}{5}$ stars, which is less than half of $n$ and because in that
+case $n\approx m$, \gtx{} will finish after $O(\log m)$ iterations. Note that a barbell graph (two
+cliques connected by a line) would require a number of iterations proportional to the length of that
+central line, despite having many more edges. This suggests unsurprisingly that the diameter of the graph
+could be a good parameter to quantify the number of iteration needed. Indeed, take a tree a consider
+the length $p$ of its longest path from the root to a leaf. By a similar argument as the one used in
+the line case, it seems \gtx{} will terminate after $O(\log p)$ iterations.

### 2017-06-25 gtx_algorithm.tex
typos

@@
 single node. As we will prove later, after reaching this point, the algorithm has selected $|V_0| -
-1$ edges that form a spanning tree of $G_0$. $G_{t+1}$ is obtained from $G_t$ by composing two
+1$ edges that form a spanning tree of $G_0$. Intuitively, $G_{t+1}$ is obtained from $G_t$ by composing two
 primitives so that we can informally write $G_{t+1} = \left(\collapseStar{} \circ
@@
 needed to finish and the stretch of the resulting tree. For simplicity and without loss of
-generality, we assume that $G_0$ consist of a single connected component.
+generality, we assume that $G_0$ through $G_T$ consist of a single connected component.
 
@@
 $t_f$ or a \emph{degree function} $d_f$. While the nodeset $V_t$ is not exhausted, it repeatedly samples a
-star center $c_i$, creates a star $S_i^t$ with the neighbors of $c_i$, removes all the nodes of $S_i^t$ from
+node $c_i$, creates a star $S_i^t$ with $c_i$ at its center and the neighbors of $c_i$ on the
+periphery, removes all the nodes of $S_i^t$ from
 $V_t$ and all the edges incident to $S_i^t$ from $E_t$, and finally decrements accordingly the
@@
   \includegraphics[height=0.15\textheight]{assets/tikz/gtx_star_tikz.pdf}
-  \caption[A sample star]{A sample star created during the \tth{} extraction level. The black node
+  \caption[A sample star]{A sample star created during the \tth{} collapse level. The black node
     % \tikz{\node[vertex,rare] {$c_i$};}
-    is the center $c_i$ of the star $S_i^t$, which is made of the four light gray peripheral nodes
+    is the center $c_i$ of the star $S_i^t$, which is also made of the four light gray peripheral nodes
   % \tikz{\node[vertex,medium] {$p_1$};} to \tikz{\node[vertex,medium] {$p_4$};}
@@
   % \tikz{\node[vertex] {$h_1$};} to \tikz{\node[vertex] {$h_3$};}
-  $h_1$ to $h_3$, whose degree will decrease once we $S_i^t$ is removed from $G_t$.}
+  $h_1$ to $h_3$, whose degree will decrease once $S_i^t$ is removed from $G_t$.}
   \label{fig:gtx_star_simple}
-\end{marginfigure}
+\end{marginfigure}%
 Upon completion, it returns a list of stars and a mapping of
-each node of $V_t$ to the unique star it belongs to. We consider three heuristics to choose centers:
+each node of $V_t$ to the unique star it belongs to. We consider three heuristics to sample centers:
 
@@
     among a small set of high degree nodes, for instance by letting $t_f(n) = \sqrt{n}$
-  \item if $\degr(u)$ is the degree of node $u$, choose node proportionally to $d_f(\degr(u))$.
-    Again, the degree function is designed so that it favors the selection of high degree nodes. For
-    instance, one could use $d_f(\degr(u)) = \degr(u)^2$.
+  \item if $\degr_i(u)$ is the degree of node $u$ before iteration \ith{}, choose node
+    proportionally to $d_f(\degr_i(u))$.  Again, the degree function is designed so that it favors the
+    selection of high degree nodes. For instance, one could use $d_f(\degr_i(u)) = \degr_i(u)^2$.
 \end{itemize}
@@
         \Let{$periphery$}{$G[c_i] \bigcap remaining$}
-        \Let{$stars$}{$stars \bigcap $\Call{Star}{$c_i,\, periphery$}}
+        \Let{$stars$}{$stars \bigcup \{$\Call{Star}{$c_i,\, periphery$}\}}
         \Let{$inner\_edges$}{$inner\_edges \bigcup \{(c_i, p): p \in periphery\}$}
@@
 maximum of the queue (line 9), we build the corresponding star (line 12--14) and we decrease the
-priority (\ie the degree) of all nodes adjacent to the new star (line 15--17).  Since both
+priority (\ie{} the degree) of all nodes adjacent to the new star (line 15--17).  Since both
 operations require constant time when using a Strict Fibonacci Heap~\autocite{FibonacciHeaps12} and
@@
 The other two variants are more time consuming because they require additional bookkeeping. Their
-randomization make them useful in an adversarial context but it also renders their analysis more
-challenging, not necessarily for the runtime of \extractStar{} but mostly for the tree construction.
-\marginpars{Which means this paragraph can be skipped at first.}
+randomization make them useful in an adversarial context but it also renders the analysis  of the
+resulting spanning tree more challenging\marginpars{Which means this paragraph can be skipped at first.}
 Therefore, we only briefly describe the implementations here.\footnote{Although they are available
@@
 can draw any node as center proportionally to its weight (where the weight of node $u$ is defined as
-$d_f\left(\degr(u)\right)$), but we cannot use the standard method of computing the cumulative sum
-of weights since each iteration change some of them. Therefore, we construct a binary tree whose
+$d_f\left(\degr(u)\right)$). Yet we cannot use the standard method of computing the cumulative sum
+of weights since some of them change at each iteration. Therefore, we construct a binary tree whose
 leaves are the nodes of $V_t$ and where each tree nodes maintain the sum of weights in its left and
-right subtrees. To sample, we draw a random number between $0$ and the total weight of the tree.
+right subtrees. To sample, we draw a random number $r$ between $0$ and the total weight of the tree and
+go down from the root to the leaf spanning the weight interval containing $r$.
 When degrees are updated (or graph node removed), we update the weights along a path from the
@@
 eccentricity the default behavior, which would simplify the pseudo code}. It builds a new graph $G_{t+1}$
-where each star becomes a node and there is a link between two nodes $s_1$ and $s_2$ if the nodes
+where each star becomes a node and where there is a link between two nodes $s_1$ and $s_2$ if the nodes
 making up $s_1$ and $s_2$ are connected in $E_t$.
@@
   \begin{algorithmic}[1]
-    \Function{\collapseStar{}}{$G_t=(V_t,E_t),\,membership,\,eccentricity$}
+    \Function{\collapseStar{}}{$E_t,\,membership,\,eccentricity$}
       \State Let $G_{t+1}$ be an empty graph
-      \State Let $cross\_edges$ be a HashMap mapping edges in $G_{t+1}$ to set of edges in $G_t$
+      \State Let $cross\_edges$ be a mapping from edges in $G_{t+1}$ to set of edges in $G_t$
       \ForAll{edge $(u, v)$ in \Call{Shuffle}{$E_t$}}
@@
       \If{$eccentricity$ is not \textbf{null}}
-        \ForAll{$\Big(\underbrace{(u,v)}_{\text{key}},\,
+        \ForAll{$\Big(\underbrace{(s_u,s_v)}_{\text{key}},\,
           \underbrace{candidate\_edges}_{\text{value}}\Big)$
@@
 
-For that, we first shuffle $E_t$ and go through it (line 4). When we find an edge whose endpoints
+For that, we first shuffle $E_t$ and iterate on it (line 4). When we find an edge whose endpoints
 belong to two different stars not yet connected (line 10), we use that edge to connect these two
@@
 \extractStar{} and \collapseStar{} are truly the core the of \gtx{} algorithm but to obtain our
-final spanning tree, we need additional works, namely updating the eccentricity of the nodes of
-$V_0$ and keeping track of each edge within and between along every contraction level.
+final spanning tree, we need additional work, namely updating the eccentricity of the nodes of
+$V_0$ and keeping track of each edge within and between stars along every contraction level.
 
@@
       \State \Call{Update-Eccentricity}{$stars$, $eccentricity$, $original\_node$}
-      \Let{$full\_membership$, $original\_node$}{\Call{Update-Nodes-Mapping}{$full\_membership$, $star\_membership$}}
+      \Let{$full\_membership$, $original\_node$}{\Call{Update-Nodes-Mapping}{$full\_membership$, $membership$}}
       \Let{$G_{t+1}$, $\!outer\_edges$}{\Call{Collapse-Stars}{$E_t \!\setminus\! inner\_edges$,
@@
     created during the \jth{} iteration of the algorithm is denoted $S_i^j$. Refer to the main text
-    for a complete walk through.}
+    for the complete description of the execution.}
   \label{fig:gtx_eccentricity}
@@
 final spanning tree, showed with black edges in \autoref{fig:gtx_eccentricity}
-\vpageref{fig:gtx_eccentricity}. For completeness, we can also look at the edges which are not in
-the spanning tree and therefore contribute to the stretch of the tree. In that case the average
+\vpageref{fig:gtx_eccentricity}. For completeness, we can also look at the edges which are not part
+of the spanning tree and therefore contribute to the stretch of the tree. In that case the average
 stretch is $7$, as showed in \autoref{tab:gtx_example_stretch}.

### 2017-06-25 notations.tex
add missing pseudo code and finish example execution
typos

@@
   \caption{List of notations used in this thesis}\label{tab:notations}%
+  \vspace{-.5\baselineskip}
   \begin{tabulary}{187mm}{LL}
@@
     $(u,v)$, $(i,j)$ & An undirected edges between nodes $u$ and $v$ (or $i$ and $j$) \\
-    $\mathcal{N}(u)$ & The set of all neighbors of $u$, regardless of direction. We thus have
+    $\mathcal{N}(u)$ & The set of all neighbors of $u$, regardless of edge direction. We thus have
     $|\mathcal{N}(u)| = \degr(u)$ \\

### 2017-06-25 gtx_algorithm.tex
rename collapse to contract

@@
 1$ edges that form a spanning tree of $G_0$. Intuitively, $G_{t+1}$ is obtained from $G_t$ by composing two
-primitives so that we can informally write $G_{t+1} = \left(\collapseStar{} \circ
+primitives so that we can informally write $G_{t+1} = \left(\contractStar{} \circ
 \extractStar{}\right)(G_t)$.
 
-\extractStar{} partitions the graph $G_t$ into a set of stars and \collapseStar{} build the graph
+\extractStar{} partitions the graph $G_t$ into a set of stars and \contractStar{} build the graph
 made of those stars using the edges in $E_t$. We provide more details on those two operations in the
@@
   \includegraphics[height=0.15\textheight]{assets/tikz/gtx_star_tikz.pdf}
-  \caption[A sample star]{A sample star created during the \tth{} collapse level. The black node
+  \caption[A sample star]{A sample star created during the \tth{} contraction level. The black node
     % \tikz{\node[vertex,rare] {$c_i$};}
@@
 
-\paragraph{\collapseStar{}}\label{par:collapsestar}%
-The second routine, \collapseStar{} takes as input the result of \extractStar{}, along with $E_t$
+\paragraph{\contractStar{}}\label{par:contractstar}%
+The second routine, \contractStar{} takes as input the result of \extractStar{}, along with $E_t$
 and an optional $\emph{eccentricity}$ array we will describe soon\marginpars{Actually, we could make
@@
   \begin{algorithmic}[1]
-    \Function{\collapseStar{}}{$E_t,\,membership,\,eccentricity$}
+    \Function{\contractStar{}}{$E_t,\,membership,\,eccentricity$}
       \State Let $G_{t+1}$ be an empty graph
@@
 simplification since the $eccentricity$ array is only defined for nodes in $V_0$ so when
-\collapseStar{} is called on $G_3$ for instance, $u$ and $v$ are nodes of $V_3$ and we have to
+\contractStar{} is called on $G_3$ for instance, $u$ and $v$ are nodes of $V_3$ and we have to
 retrieve the actual corresponding edge in $E_0$. Maybe I could simply add a HashMap from $E_t$ to
@@
 \paragraph{Putting the pieces together}\label{par:full_gtx}%
-\extractStar{} and \collapseStar{} are truly the core the of \gtx{} algorithm but to obtain our
+\extractStar{} and \contractStar{} are truly the core the of \gtx{} algorithm but to obtain our
 final spanning tree, we need additional work, namely updating the eccentricity of the nodes of
@@
       \Let{$full\_membership$, $original\_node$}{\Call{Update-Nodes-Mapping}{$full\_membership$, $membership$}}
-      \Let{$G_{t+1}$, $\!outer\_edges$}{\Call{Collapse-Stars}{$E_t \!\setminus\! inner\_edges$,
+      \Let{$G_{t+1}$, $\!outer\_edges$}{\Call{Contract-Stars}{$E_t \!\setminus\! inner\_edges$,
       $\!membership$, $\!eccentricity$}}
@@
 adds a constant factor $2$ in the overall complexity of the \gtx{} algorithm.} 
-at every collapse level, we first extract stars from the current graph (line 5), then
-update the eccentricity and nodes mapping (lines 7--8) and finally collapse the graph (line 9) We
+at every contraction level, we first extract stars from the current graph (line 5), then
+update the eccentricity and nodes mapping (lines 7--8) and finally contract the graph (line 9) We
 perform these operations until there are no edge connecting stars anymore. At this point, we revisit
 every outer edges to form the spanning tree (line 13). Because the main loop does not perform any
-work besides calling other function, assuming there are T collapse steps, the complexity of \gtx{}
+work besides calling other function, assuming there are $T$ contract steps, the complexity of \gtx{}
 is $O(T(m+n)) = O(Tm)$.
@@
 
-We then call \collapseStar{}, with the eccentricity reducing variant. This will connect all possible
+We then call \contractStar{}, with the eccentricity reducing variant. This will connect all possible
 pairs of star. For instance, the edge between nodes $19$ and $29$ leads to the edge
@@
 $\textcolor{DodgerBlue}{S_1^1}$ and $\textcolor{DodgerBlue}{S_2^1}$. The full result of the
-\collapseStar{} procedure is $G_1$, which can be seen on \autoref{fig:gtx_run1}.
+\contractStar{} procedure is $G_1$, which can be seen on \autoref{fig:gtx_run1}.
 
@@
 $\textcolor{DodgerBlue}{S_7^1}$ forms the last star with $\textcolor{DodgerBlue}{S_8^1}$ and
-$\textcolor{DodgerBlue}{S_9^1}$. Then \collapseStar{} connects the resulting three stars, and this
+$\textcolor{DodgerBlue}{S_9^1}$. Then \contractStar{} connects the resulting three stars, and this
 time there is only a single choice between each pair of stars, leading to the graph $G_2$ showed in
@@
 eccentricity of their underlying $G_0$ nodes by $1$ (namely nodes $16$ to $45$).  Because there is
-only one star $\textcolor{Green}{S_1^3}$ left, \collapseStar{} returns $G_3$ showed in
+only one star $\textcolor{Green}{S_1^3}$ left, \contractStar{} returns $G_3$ showed in
 \autoref{fig:gtx_run3} and an empty list of outer edges, meaning that the inner loop of \gtx{} is
@@
 A crucial quantity of the \gtx{} algorithm, both in terms of complexity and resulting stretch, is
-the number of collapses $T$ needed before termination. While this is still elusive to express in the
+the number of contraction $T$ needed before termination. While this is still elusive to express in the
 general case, let us first look at some simple cases. For instance, a very sparse example of graph

### 2017-06-27 assets/tikz/gtx_eccentricity_tikz.tex
change the example graph to add color on its edges

@@
 \begin{document}
-\tikzstyle{star}=[circle,DodgerBlue,very thick,fill opacity=.2]
-\tikzstyle{star-label}=[opacity=1, , font=\Large]
-\tikzstyle{removed}=[Gray,line width=1.15pt]
-\begin{tikzpicture}[auto,vertex/.append style={minimum width=8mm}]
-  \draw[star,Orange] (2.260, -1.063) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$S_1^2$};
-  \draw[star,Orange] (6.504, -9.728) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$S_2^2$};
-  \draw[star,Orange] (11.641, 1.189) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$S_3^2$};
+\tikzstyle{star}=[DodgerBlue,line width=2pt]
+\tikzstyle{star-label}=[font=\Large]
+\tikzstyle{removed}=[dashed,line width=1.7pt]
+\begin{tikzpicture}[auto,vertex/.append style={minimum width=8mm},
+  edge/.append style={line width=2.2pt,color=DodgerBlue!90!Black}]
+  \draw[star,Orange] (2.260, -1.063) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$\mathbf{S_1^2}$};
+  \draw[star,Orange] (6.504, -9.728) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$\mathbf{S_2^2}$};
+  \draw[star,Orange] (11.641, 1.189) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$\mathbf{S_3^2}$};
 
-  \draw[star] (0.000, -0.000) circle [radius=1.8]  node[star-label,xshift=-30, yshift=25] {$S_1^1$};
-  \draw[star] (1.689, -3.448) circle [radius=1.8]  node[star-label,xshift=-30, yshift=5] {$S_2^1$};
-  \draw[star] (3.734, 0.896) circle [radius=1.8]  node[star-label,xshift=-30, yshift=5] {$S_3^1$};
-  \draw[star] (4.245, -8.664) circle [radius=1.8]  node[star-label,xshift=-22, yshift=-25] {$S_4^1$};
-  \draw[star] (5.934, -12.113) circle [radius=1.8]  node[star-label,xshift=-15, yshift=35] {$S_5^1$};
-  \draw[star] (7.978, -7.768) circle [radius=1.8]  node[star-label,xshift=30, yshift=25] {$S_6^1$};
-  \draw[star] (9.381, 2.252) circle [radius=1.8]  node[star-label,xshift=-15, yshift=30] {$S_7^1$};
-  \draw[star] (11.071, -1.196) circle [radius=1.8]  node[star-label,xshift=-30, yshift=25] {$S_8^1$};
-  \draw[star] (13.115, 3.149) circle [radius=1.8]  node[star-label,xshift=-30, yshift=20	] {$S_9^1$};
+  \draw[star] (0.000, -0.000) circle [radius=1.8]  node[star-label,xshift=-30, yshift=25] {$\mathbf{S_1^1}$};
+  \draw[star] (1.689, -3.448) circle [radius=1.8]  node[star-label,xshift=-30, yshift=5] {$\mathbf{S_2^1}$};
+  \draw[star] (3.734, 0.896) circle [radius=1.8]  node[star-label,xshift=-30, yshift=5] {$\mathbf{S_3^1}$};
+  \draw[star] (4.245, -8.664) circle [radius=1.8]  node[star-label,xshift=-22, yshift=-25] {$\mathbf{S_4^1}$};
+  \draw[star] (5.934, -12.113) circle [radius=1.8]  node[star-label,xshift=-15, yshift=35] {$\mathbf{S_5^1}$};
+  \draw[star] (7.978, -7.768) circle [radius=1.8]  node[star-label,xshift=30, yshift=25] {$\mathbf{S_6^1}$};
+  \draw[star] (9.381, 2.252) circle [radius=1.8]  node[star-label,xshift=-15, yshift=30] {$\mathbf{S_7^1}$};
+  \draw[star] (11.071, -1.196) circle [radius=1.8]  node[star-label,xshift=-30, yshift=25] {$\mathbf{S_8^1}$};
+  \draw[star] (13.115, 3.149) circle [radius=1.8]  node[star-label,xshift=-30, yshift=20	] {$\mathbf{S_9^1}$};
 
@@
   \draw[edge] (n1) edge [bend left =0] node[] {} (n5);
-  \draw[edge] (n2) edge [bend left =25] node[] {} (n6);
-  \draw[edge,removed] (n2) edge [bend left =0] node[] {} (n9);
-  \draw[edge,removed] (n3) edge [bend left =-20] node[] {} (n12);
-  \draw[edge] (n4) edge [bend left =20] node[] {} (n14);
+  \draw[edge,Orange!90!Black] (n2) edge [bend left =25] node[] {} (n6);
+  \draw[edge,removed,Orange!90!Black] (n2) edge [bend left =0] node[] {} (n9);
+  \draw[edge,removed,Orange!90!Black] (n3) edge [bend left =-20] node[] {} (n12);
+  \draw[edge,Orange!90!Black] (n4) edge [bend left =20] node[] {} (n14);
   % \draw[edge] (n3) edge [bend left =-20] node[] {} (n18);
@@
   % \draw[edge] (n9) edge [bend left =-10] node[] {} (n32);
-  \draw[edge] (n8) .. controls ++(-40:3) and ++(240: 2) .. (n37);
+  \draw[edge,Green!90!Black] (n8) .. controls ++(-40:3) and ++(240: 2) .. (n37);
   \draw[edge] (n11) edge [bend left =0] node[] {} (n12);
@@
   \draw[edge] (n11) edge [bend left =0] node[] {} (n15);
-  \draw[edge,removed] (n13) edge [bend left =0] node[] {} (n29);
+  \draw[edge,removed,Green!90!Black] (n13) edge [bend left =0] node[] {} (n29);
   % \draw[edge] (n12) edge [bend left =0] node[] {} (n18);
-  \draw[edge] (n15) .. controls ++(-40:3) and ++(64: 3.5) .. (n26);
+  \draw[edge,Green!90!Black] (n15) .. controls ++(-40:3) and ++(64: 3.5) .. (n26);
   \draw[edge] (n16) edge [bend left =0] node[] {} (n17);
@@
   \draw[edge] (n16) edge [bend left =0] node[] {} (n20);
-  \draw[edge] (n17) edge [bend left =0] node[] {} (n23);
+  \draw[edge,Orange!90!Black] (n17) edge [bend left =0] node[] {} (n23);
   %\draw[edge] (n19) edge [bend left =0] node[] {} (n27);
-  \draw[edge] (n19) edge [bend left =0] node[] {} (n29);
-  \draw[edge,removed] (n20) edge [bend left =0] node[] {} (n24);
+  \draw[edge,Orange!90!Black] (n19) edge [bend left =0] node[] {} (n29);
+  \draw[edge,removed,Orange!90!Black] (n20) edge [bend left =0] node[] {} (n24);
   \draw[edge] (n21) edge [bend left =0] node[] {} (n22);
@@
   \draw[edge] (n21) edge [bend left =0] node[] {} (n25);
-  \draw[edge,removed] (n25) edge [bend left =-20] node[] {} (n42);
+  \draw[edge,removed,Green!90!Black] (n25) edge [bend left =-20] node[] {} (n42);
   \draw[edge] (n26) edge [bend left =0] node[] {} (n27);
@@
   \draw[edge] (n31) edge [bend left =0] node[] {} (n35);
-  \draw[edge] (n32) edge [bend left =0] node[] {} (n39);
-  \draw[edge] (n35) edge [bend left =0] node[] {} (n43);
-  \draw[edge,removed] (n33) edge [bend left =-25] node[] {} (n38);
-  \draw[edge,removed] (n34) edge [bend left =20] node[] {} (n43);
+  \draw[edge,Orange!90!Black] (n32) edge [bend left =0] node[] {} (n39);
+  \draw[edge,Orange!90!Black] (n35) edge [bend left =0] node[] {} (n43);
+  \draw[edge,removed,Orange!90!Black] (n33) edge [bend left =-25] node[] {} (n38);
+  \draw[edge,removed,Orange!90!Black] (n34) edge [bend left =20] node[] {} (n43);
   \draw[edge] (n36) edge [bend left =0] node[] {} (n37);
@@
 
-  \draw[Green, line width=2pt] (11.8, 6.5) 
+  \draw[star, Green!90!Black] (11.8, 6.5) 
   .. controls ++(190:1) and ++(55: 2) .. (-2, 2.5)
   .. controls  ++(55: -4) and ++(170:4) .. (6, -15) 
-  .. controls  ++(-10:9) and ++(10:9) .. node[pos=.4,font=\Huge,xshift=-15pt] {$S_1^3$} cycle ;
+  .. controls  ++(-10:9) and ++(10:9) .. node[pos=.4,font=\Huge,xshift=-15pt] {$\mathbf{S_1^3}$} cycle ;
 \end{tikzpicture}

### 2017-06-27 gtx_algorithm.tex
change the example graph to add color on its edges

@@
   \caption[The hierarchical structure of stars created by \gtx{}]{%
-    The execution of the \gtx{} algorithm. The original graph is made of the solid edges
-    connecting the nodes labeled by  their index. Edges forming the final spanning tree are in black
-    while the others are in gray. The four shades of gray, from white to dark gray
+    The execution of the \gtx{} algorithm. The original graph is made of the solid and dashed edges
+    connecting the nodes labeled by their index. Edges forming the final spanning tree are solid
+    while the others are dashed. Their colors indicate at which iteration they were chosen to be inside a
+    star. The four shades of gray, from white to dark gray
     denote increasing node eccentricity (as computed at the end of the algorithm). The \ith{} star

### 2017-06-27 assets/tikz/gtx_noreduc_tikz.tex
add a note on singleton star

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
\begin{tikzpicture}[auto,vertex/.append style={minimum width=2mm}]

  \node[vertex,] (c1) at (0.000,0.000) {$c_1$};
  \node[vertex,medium] (p1) at (-3.000,1.500) {$$};
  \draw[edge,] (c1) edge [bend left =0] node[] {} (p1);
  \node[vertex,fill=Black!50] (h1) at (-2.559,2.107) {$$};
  \draw[edge,] (p1) edge [bend left =0] node[] {} (h1);
  \node[vertex,fill=Black!50] (h2) at (-2.844,2.234) {$$};
  \draw[edge,] (p1) edge [bend left =0] node[] {} (h2);
  \node[vertex,fill=Black!50] (h3) at (-3.156,2.234) {$$};
  \draw[edge,] (p1) edge [bend left =0] node[] {} (h3);
  \node[vertex,fill=Black!50] (h4) at (-3.441,2.107) {$$};
  \draw[edge,] (p1) edge [bend left =0] node[] {} (h4);
  \node[vertex,medium] (p2) at (-1.800,1.500) {$$};
  \draw[edge,] (c1) edge [bend left =0] node[] {} (p2);
  \node[vertex,fill=Black!50] (h5) at (-1.359,2.107) {$$};
  \draw[edge,] (p2) edge [bend left =0] node[] {} (h5);
  \node[vertex,fill=Black!50] (h6) at (-1.644,2.234) {$$};
  \draw[edge,] (p2) edge [bend left =0] node[] {} (h6);
  \node[vertex,fill=Black!50] (h7) at (-1.956,2.234) {$$};
  \draw[edge,] (p2) edge [bend left =0] node[] {} (h7);
  \node[vertex,fill=Black!50] (h8) at (-2.241,2.107) {$$};
  \draw[edge,] (p2) edge [bend left =0] node[] {} (h8);
  \node[vertex,medium] (p3) at (-0.600,1.500) {$$};
  \draw[edge,] (c1) edge [bend left =0] node[] {} (p3);
  \node[vertex,fill=Black!50] (h9) at (-0.159,2.107) {$$};
  \draw[edge,] (p3) edge [bend left =0] node[] {} (h9);
  \node[vertex,fill=Black!50] (h10) at (-0.444,2.234) {$$};
  \draw[edge,] (p3) edge [bend left =0] node[] {} (h10);
  \node[vertex,fill=Black!50] (h11) at (-0.756,2.234) {$$};
  \draw[edge,] (p3) edge [bend left =0] node[] {} (h11);
  \node[vertex,fill=Black!50] (h12) at (-1.041,2.107) {$$};
  \draw[edge,] (p3) edge [bend left =0] node[] {} (h12);
  \node[vertex,medium] (p4) at (0.600,1.500) {$$};
  \draw[edge,] (c1) edge [bend left =0] node[] {} (p4);
  \node[vertex,fill=Black!50] (h13) at (1.041,2.107) {$$};
  \draw[edge,] (p4) edge [bend left =0] node[] {} (h13);
  \node[vertex,fill=Black!50] (h14) at (0.756,2.234) {$$};
  \draw[edge,] (p4) edge [bend left =0] node[] {} (h14);
  \node[vertex,fill=Black!50] (h15) at (0.444,2.234) {$$};
  \draw[edge,] (p4) edge [bend left =0] node[] {} (h15);
  \node[vertex,fill=Black!50] (h16) at (0.159,2.107) {$$};
  \draw[edge,] (p4) edge [bend left =0] node[] {} (h16);
  \node[vertex,medium] (p5) at (1.800,1.500) {$$};
  \draw[edge,] (c1) edge [bend left =0] node[] {} (p5);
  \node[vertex,fill=Black!50] (h17) at (2.241,2.107) {$$};
  \draw[edge,] (p5) edge [bend left =0] node[] {} (h17);
  \node[vertex,fill=Black!50] (h18) at (1.956,2.234) {$$};
  \draw[edge,] (p5) edge [bend left =0] node[] {} (h18);
  \node[vertex,fill=Black!50] (h19) at (1.644,2.234) {$$};
  \draw[edge,] (p5) edge [bend left =0] node[] {} (h19);
  \node[vertex,fill=Black!50] (h20) at (1.359,2.107) {$$};
  \draw[edge,] (p5) edge [bend left =0] node[] {} (h20);
  \node[vertex,medium] (p6) at (3.000,1.500) {$$};
  \draw[edge,] (c1) edge [bend left =0] node[] {} (p6);
  \node[vertex,fill=Black!50] (h21) at (3.441,2.107) {$$};
  \draw[edge,] (p6) edge [bend left =0] node[] {} (h21);
  \node[vertex,fill=Black!50] (h22) at (3.156,2.234) {$$};
  \draw[edge,] (p6) edge [bend left =0] node[] {} (h22);
  \node[vertex,fill=Black!50] (h23) at (2.844,2.234) {$$};
  \draw[edge,] (p6) edge [bend left =0] node[] {} (h23);
  \node[vertex,fill=Black!50] (h24) at (2.559,2.107) {$$};
  \draw[edge,] (p6) edge [bend left =0] node[] {} (h24);
  \node[vertex,] (c2) at (0.000,4.500) {$c_2$};
  \node[vertex,medium] (pu1) at (-3.000,3.000) {$$};
  \draw[edge,] (c2) edge [bend left =0] node[] {} (pu1);
  \draw[edge,] (pu1) edge [bend left =0] node[] {} (h1);
  \draw[edge,] (pu1) edge [bend left =0] node[] {} (h2);
  \draw[edge,] (pu1) edge [bend left =0] node[] {} (h3);
  \draw[edge,] (pu1) edge [bend left =0] node[] {} (h4);
  \node[vertex,medium] (pu2) at (-1.800,3.000) {$$};
  \draw[edge,] (c2) edge [bend left =0] node[] {} (pu2);
  \draw[edge,] (pu2) edge [bend left =0] node[] {} (h5);
  \draw[edge,] (pu2) edge [bend left =0] node[] {} (h6);
  \draw[edge,] (pu2) edge [bend left =0] node[] {} (h7);
  \draw[edge,] (pu2) edge [bend left =0] node[] {} (h8);
  \node[vertex,medium] (pu3) at (-0.600,3.000) {$$};
  \draw[edge,] (c2) edge [bend left =0] node[] {} (pu3);
  \draw[edge,] (pu3) edge [bend left =0] node[] {} (h9);
  \draw[edge,] (pu3) edge [bend left =0] node[] {} (h10);
  \draw[edge,] (pu3) edge [bend left =0] node[] {} (h11);
  \draw[edge,] (pu3) edge [bend left =0] node[] {} (h12);
  \node[vertex,medium] (pu4) at (0.600,3.000) {$$};
  \draw[edge,] (c2) edge [bend left =0] node[] {} (pu4);
  \draw[edge,] (pu4) edge [bend left =0] node[] {} (h13);
  \draw[edge,] (pu4) edge [bend left =0] node[] {} (h14);
  \draw[edge,] (pu4) edge [bend left =0] node[] {} (h15);
  \draw[edge,] (pu4) edge [bend left =0] node[] {} (h16);
  \node[vertex,medium] (pu5) at (1.800,3.000) {$$};
  \draw[edge,] (c2) edge [bend left =0] node[] {} (pu5);
  \draw[edge,] (pu5) edge [bend left =0] node[] {} (h17);
  \draw[edge,] (pu5) edge [bend left =0] node[] {} (h18);
  \draw[edge,] (pu5) edge [bend left =0] node[] {} (h19);
  \draw[edge,] (pu5) edge [bend left =0] node[] {} (h20);
  \node[vertex,medium] (pu6) at (3.000,3.000) {$$};
  \draw[edge,] (c2) edge [bend left =0] node[] {} (pu6);
  \draw[edge,] (pu6) edge [bend left =0] node[] {} (h21);
  \draw[edge,] (pu6) edge [bend left =0] node[] {} (h22);
  \draw[edge,] (pu6) edge [bend left =0] node[] {} (h23);
  \draw[edge,] (pu6) edge [bend left =0] node[] {} (h24);

  \draw[black,line width=1pt,line cap=round, <-]  ($(c1) + (3:15pt)$) arc (0:180:15pt) 
  node[pos=-.2,yshift=-5pt,anchor=south west] {$p$ branches};
  \draw[black,line width=1pt,line cap=round, <-]  ($(p4) + (-5:10pt)$) arc (0:180:10pt) 
  node[pos=-.2,xshift=-34pt,rotate=35, yshift=-34pt,font=\small,anchor=south west] {$q$ branches};
\end{tikzpicture}
\end{document}

### 2017-06-27 assets/tikz/gtx_singleton_tikz.tex
add a note on singleton star

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
  \tikzstyle{star}=[DodgerBlue,line width=1pt]
  \tikzstyle{star-label}=[font=\Large]
  \begin{tikzpicture}[auto,vertex/.append style={minimum width=6.5mm}]
    \node[vertex,] (c1) at (-2.427,-0.000) {$c_1$};
    \node[vertex,] (c2) at (2.427,0.000) {$c_2$};
    \node[vertex,] (c3) at (0.000,0.000) {$c_3$};

    \draw[star] (0.000, -0.000) circle [radius=.4]  node[star-label,xshift=0, yshift=25] {$S_3^1$};
    \draw[star] (c1) circle [radius=2]  node[star-label,xshift=-30, yshift=30] {$S_1^1$};
    \draw[star] ( c2) circle [radius=2]  node[star-label,xshift=-30, yshift=30] {$S_2^1$};

    \node[vertex,medium] (p1) at (-1.283,-1.144) {$p_1$};
    \draw[edge,] (c1) edge [bend left =0] node[] {} (p1);
    \node[vertex,medium] (p2) at (-2.008,1.563) {$p_2$};
    \draw[edge,] (c1) edge [bend left =0] node[] {} (p2);
    \node[vertex,medium] (p3) at (-3.990,-0.419) {$p_3$};
    \draw[edge,] (c1) edge [bend left =0] node[] {} (p3);
    \node[vertex,medium] (p4) at (1.283,-1.144) {$p_4$};
    \draw[edge,] (c2) edge [bend left =0] node[] {} (p4);
    \node[vertex,medium] (p5) at (3.990,-0.419) {$p_5$};
    \draw[edge,] (c2) edge [bend left =0] node[] {} (p5);
    \node[vertex,medium] (p6) at (2.008,1.563) {$p_6$};
    \draw[edge,] (c2) edge [bend left =0] node[] {} (p6);
    \draw[edge,] (c3) edge [bend left =0] node[] {} (p4);
    \draw[edge,] (c3) edge [bend left =0] node[] {} (p1);

    \draw[->] (0,-2) -- (0,-2.9);

    \node[vertex,fill=DodgerBlue,text=black] (s1) at (-2,-3.7) {$S_1^1$};
    \node[vertex,fill=DodgerBlue,text=black] (s2) at (2,-3.7) {$S_2^1$};
    \node[vertex,fill=DodgerBlue,text=black] (s3) at (0,-3.7) {$S_3^1$};
    \foreach \source/\dest in {3/2, 1/3}
    \draw[edge] (s\source) -- (s\dest);
  \end{tikzpicture}
\end{document}

### 2017-06-27 gtx_algorithm.tex
add a note on singleton star

@@
 \extractStar{}\right)(G_t)$.
+% [from a CMU NIPS submission http://www.stat.cmu.edu/~arinaldo/papers/mutualfriends.pdf] We
+% consider Mutual-Friends to be *an algorithmic primitive*, by which we mean a kind of subroutine for
+% a more complicated function that iterates Mutual-Friends, similarly to [5].
 
@@
 generality, we assume that $G_0$ through $G_T$ consist of a single connected component.
+\marginpars{Actually, we only need to assume that $G_0$ is connected because then we can prove that
+the other subsequent $G_t$ will be as well.}
 
@@
 the line case, it seems \gtx{} will terminate after $O(\log p)$ iterations.
+
+While we assume that $G_t$ is connected, it might happen during the execution of \extractStar{} that%
+\begin{marginfigure}
+  \centering
+  \includegraphics[width=0.95\linewidth]{assets/tikz/gtx_singleton_tikz.pdf}
+  \caption{The formation of a singleton star}
+  \label{fig:gtx_singleton}
+\end{marginfigure}
+the degree of a node $u$ drops to zero because because all of its neighbors were claimed by the
+periphery of previous stars. Such a node then forms a \emph{singleton star}, as illustrated in
+\autoref{fig:gtx_singleton}, where after the creation of \starone{1} and \starone{2}, $c_3$ is the
+single node of the third star \starone{3}.%
+\begin{marginfigure}
+  \centering
+  \includegraphics[width=0.95\linewidth]{assets/tikz/gtx_noreduc_tikz.pdf}
+  \caption{A case were too many singletons \enquote{waste} one iteration of \gtx{}}
+  \label{fig:gtx_noreduc}
+\end{marginfigure}
+In the case no singleton are formed during one execution of \extractStar{}, the number of nodes is
+reduced by at least $2$ (because each star is made of at least two nodes). On the other hand, we can
+construct a graph with singletons where the factor of reduction in number of nodes and edges can be
+arbitrarily close to $1$. Consider \autoref{fig:gtx_noreduc} and assume that first extract two stars
+centered in $c_1$ and $c_2$.%

### 2017-06-28 notations.tex
clarify and expand notations

@@
-\begin{table*}[htpb]
+\begin{table*}[thpb]
   \centering
@@
     \midrule
-    $G$ & A graph, either directed or undirected \\
-    $V$, $V_t$ & The set of all the nodes of a graph. Unless noted otherwise, nodes are indexed from $1$ to $|V|$ \\
-    $E$, $E_t$ & The set of all the edges of a graph \\
-    $\degr(u)$ & The total degree of node $u$ (that is the number of edges incident to $u$
+    $G$ & An unweighted graph, that can be either directed or undirected \\
+    $V$ & The set of all the nodes of a graph, with $|V|=n$. Unless noted otherwise, nodes are indexed from $1$ to $n$ \\
+    $E$ & The set of all the edges of a graph, with $|E|=m$ \\
+    $u$ & A generic node of $G$. When referring to several nodes, we naturally use $u$, $v$, $w$ and so on. \\
+    $(u,v)$ & An undirected edge between nodes $u$ and $v$ \\
+    % \euv{} & A directed edge from node $u$ to node $v$ \\
+    $\yuv{}$ & The sign of the edge $(u,v)$, which can be either $+1$ or $-1$ \\
+    $Y(E)$ & The labeling of $E$, that is the set of all signs of $E$: $Y(E) = \{\yuv , (u,v)\in E\}$ \\
+    $\etrain$ & A subset of $E$, given or chosen, of which we observe the signs \\
+    $\degr(u)$ & The total degree of node $u$ (that is, the number of edges incident to $u$,
     regardless of their direction) \\
-    $(u,v)$, $(i,j)$ & An undirected edges between nodes $u$ and $v$ (or $i$ and $j$) \\
-    $\mathcal{N}(u)$ & The set of all neighbors of $u$, regardless of edge direction. We thus have
-    $|\mathcal{N}(u)| = \degr(u)$ \\
+    $\mathcal{N}(u)$ & The set of all neighbors of $u$, regardless of edge direction. It thus holds
+    that $|\mathcal{N}(u)| = \degr(u)$ \\
+    $T$ & An unweighted and undirected tree \\
+    $\pathtuv$ & The unique path between $u$ and $v$ in the tree $T$, represented by an ordered list of edges \\
+    $|\pathtuv|$ & The length of the path between $u$ and $v$ in the tree $T$, that is its number of edges \\
     \bottomrule

### 2017-06-28 lowstretch_and_cc.tex
start describing galaxy tree algorithm
add support files for tikz drawing
add a notations table
typos
add a short intro to explain why we build our tree

@@
 \title{Edge sign prediction in general graphs and connection with Correlation Clustering}
-
+% \usetikzlibrary{external}
+% \tikzexternalize[mode=list and make, failed ref warnings for={\ref,\cite,\pageref},]
 \begin{document}
-\newgeometry{vmargin=1.5cm,left=1.5cm,textwidth=130mm,marginparsep=5mm,marginparwidth=52mm}
+\newgeometry{vmargin=1.3cm,left=1.5cm,textwidth=130mm,marginparsep=5mm,marginparwidth=52mm}
 \maketitle
 % \tableofcontents
+
+% \input{notations}
 
@@
 
+% ICML paper this year that touch something very related multicut and give recent applications in
+% vision https://arxiv.org/abs/1503.03791
+
 \subsection{Variants and extensions}
@@
 
+\newpage
+
 \section{Low stretch trees and spanners}
@@
 
+\input{notations}
+
+Given a graph $G$, and following the social balance theory roughly summarized as \enquote{the enemy
+of my enemy is my friend}, assume that the labelling of $E$ is consistent with a two-clustering of
+$V$. Namely, assume that $V$ can be partitioned in two clusters such that edges within each cluster
+are positive and edges across clusters are negative. In that case, the following
+\emph{multiplicative rule} holds: for any nodes $u$, $v$ in $V$, and any path $p$ between $u$ and
+$v$ in $G$, the sign $\yuv$ is equal to the product of the signs along $p$ (we call this product the
+parity of $p$ and denote it $\pi(p)$). While it is a simple and convenient model, this is too strong
+of a requirement to be satisfied in practice. Therefore, we relax it by assuming that we start with
+a consistent labeling $Y$ but can only observe a randomly perturbed version $Y'$ of $Y$.
+Specifically, given a constant $q\in [0, \nicefrac{1}{2})$, each sign of $Y$ is flipped with a
+probability smaller than $q$.
+
+In this section, we are interested in active learning algorithms, that first query a subset
+$\etrain$ of the edges, observe the signs in $\etest$ and use them to predict the remaining signs.
+More precisely, we focus on an algorithm that query a spanning tree $T$ of $G$ and predict the sign
+of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. While this
+can\Todo{state it formally then ;)} be stated more formally~\autocite[Section
+4.1]{Cesa-Bianchi2012b}, it makes sense intuitively that since each sign has been potentially
+flipped, the longer the path in $T$, the more likely its parity will be not be equal to the truth
+$\yuv{}$. Therefore we would like each such path to be as short as possible and in the rest of this
+section, we describe a way to build spanning tree tailored for this situation.
+
 \subsection{\gtx{}: a simple low stretch tree construction}
 \label{sub:gtx_a_simple_low_stretch_tree_construction}
+
+\input{gtx_algorithm}
 
@@
 
-\input{gtx_experiments}
+% \input{gtx_experiments}
 

### 2017-06-28 gtx_algorithm.tex
improve description of extract star
- move the heuristics variant in a separate paragraph
- give more visibility to the proof of complexity
- streamline the pseudo code and give extra textual explanations

@@
 
-\medskip
-
 \paragraph{\extractStar{}}\label{par:extractstar}%
-\extractStar{} takes as input a graph $G_t=(V_t, E_t)$, and optionally a \emph{threshold function}
-$t_f$ or a \emph{degree function} $d_f$. While the nodeset $V_t$ is not exhausted, it repeatedly samples a
+\extractStar{} takes as input a graph $G_t=(V_t, E_t)$.
+While the nodeset $V_t$ is not exhausted, it repeatedly samples a
 node $c_i$, creates a star $S_i^t$ with $c_i$ at its center and the neighbors of $c_i$ on the
@@
 \end{marginfigure}%
-Upon completion, it returns a list of stars and a mapping of
-each node of $V_t$ to the unique star it belongs to. We consider three heuristics to sample centers:
+Upon completion, it returns a list of stars,
+a list of each star edges,
+and a map (or associative array) of each node of $V_t$ to the unique star it belongs to.
+According to the definition of \textcite{HashTableBook08}, an associative array is an abstract data
+type composed of a collection of (key, value) pairs, such that each possible key appears at most
+once in the collection. It efficiently supports the addition, removal and modification of a pair, as
+well as the lookup of a value associated with a particular key.
 
-\begin{itemize}%[nosep]
-  \item choose the node with the current highest degree, with ties broken arbitrarily
-  \item if $n_i$ is the number of node remaining in $V_t$ before choosing the \ith{} center, choose
-    a node \uar{} among those with a degree larger than $t_f(n_i)$. Setting the threshold function
-    to be the identity therefore recovers the previous strategy, but the idea here is to choose
-    among a small set of high degree nodes, for instance by letting $t_f(n) = \sqrt{n}$
-  \item if $\degr_i(u)$ is the degree of node $u$ before iteration \ith{}, choose node
-    proportionally to $d_f(\degr_i(u))$.  Again, the degree function is designed so that it favors the
-    selection of high degree nodes. For instance, one could use $d_f(\degr_i(u)) = \degr_i(u)^2$.
-\end{itemize}
-
-We now give the pseudo code of \extractStar{} for the highest degree variant.\footnote{Note that for
+As showed in the following pseudo code\footnote{Note that for
 clarity, we removed some bookkeeping code in all listings, mainly the part related to maintaining
-mapping between nodes at different level of contraction. However, the full python implementation
-is available at \url{https://github.com/daureg/magnet/blob/master/veverica/new_galaxy.py\#L27}.}
-We assume that $G$ is the adjacency list of the graph, so that $G[u]$ is the set of neighbors of
-$u$, \ie{} $G[u] \equiv \mathcal{N}(u)$. The other piece of notation is $\textsc{Star}$, which
-simply create a star given a center and a list of peripheral nodes.  \vspace{-\baselineskip}
+mapping between nodes at different collapse level. However, the full python implementation
+is available at \url{https://github.com/daureg/magnet/blob/master/veverica/new_galaxy.py\#L27}.}, we
+sample centers by choosing the node with the current highest degree, with ties broken
+arbitrarily.\footnote{We also consider more involved heuristics but in the interest of simplicity,
+they are presented later on~\vpageref{ssec:gtx_center_choice}.} This is achieved efficiently by
+maintaining a max-priority queue $Q$, initially containing all the nodes of $V_t$. The priority of a
+node is its current degree and we equipped $Q$ with two standard operations described
+by~\textcite[section 6.5]{CormenAlgo09}: \textsc{Extract-Max}$(Q)$ removes and returns the
+node of $Q$ with the largest degree and \textsc{Decrease-Key}$(Q$, $u$, $\Delta)$ decrements
+the degree of the node $u$ by an amount $\Delta$.
+We also assume that $G$ is the adjacency list of the graph, so that $G[u]$ is the set of neighbors of
+$u$, \ie{} $G[u] \equiv \mathcal{N}(u)$. Finally $membership$ is a map, initially empty, and we
+define the \textsc{Star} function, which creates a star given a center $c$, a list $periphery$ of
+peripheral nodes, and a star index $i$. After creating the \ith{} star, for every node $u$ belonging
+to that star, the \textsc{Star} function sets $membership[u] = i$.
+\vspace{-\baselineskip}
 
@@
     \Function{\extractStar{}}{$G_t=(V_t,E_t)$}
-      \State Let $Q$ be a max-priority queue. The key of element $x$ is $Q[x]$
-      \Let{$stars$, $inner\_edges$, $remaining$}{$[]$, $\emptyset$, $\emptyset$}
-      \ForAll{node $u$ in $V_t$}
-        \State \Call{Insert}{$Q,\,u$} \Comment with the key $\degr(u)$
-        \Let{$remaining$}{$remaining \bigcup \left\{u\right\}$}
-      \EndFor
+      \State Let $Q$ be the max-priority queue described above
+      \State Let $remaining$ be a set of nodes initially containing all the nodes in $V_t$
+      \State Let $membership$ be an empty map
+      \Let{$stars$}{$[]$}
+      \Let{$inner\_edges$}{$\emptyset$}
       \While{$Q$ is not empty}
-        \Let{$c_i$}{\Call{Extract-Max}{$Q$}}
-        \If{$c_i$ not in $remaining$}
-          \State \textbf{continue} \Comment{$c_i$ is part of an existing star so there is
+        \Let{$c$}{\Call{Extract-Max}{$Q$}}
+        \If{$c$ not in $remaining$}
+          \State \textbf{continue} \Comment{$c$ is part of an existing star so there is
           nothing to do}
         \EndIf
-        \Let{$periphery$}{$G[c_i] \bigcap remaining$}
-        \Let{$stars$}{$stars \bigcup \{$\Call{Star}{$c_i,\, periphery$}\}}
-        \Let{$inner\_edges$}{$inner\_edges \bigcup \{(c_i, p): p \in periphery\}$}
-        \Let{$remaining$}{$remaining \setminus\left\{c_i\right\} \setminus periphery$}
+        \Let{$periphery$}{$G[c] \bigcap remaining$}
+        \Let{$stars$}{$stars \bigcup \{$\Call{Star}{$c$, $periphery$, $|stars|$}\}}
+        \Let{$inner\_edges$}{$inner\_edges \bigcup \{(c, p): p \in periphery\}$}
+        \Let{$remaining$}{$remaining \setminus \left\{ \{c_i\} \cup \setminus periphery\right\}$}
         \For{$p$ in $periphery$}
           \For{$h$ in $G[p] \bigcap remaining$}
-            \State \Call{Decrease-Key}{$Q,\, h,\, Q[h]-1$}
+            \State \Call{Decrease-Key}{$Q$, $h$, $1$}
           \EndFor
@@
       \EndWhile
-      \State \textbf{return} $stars$, $inner\_edges$, $membership$%
-      \Comment{$membership$ maps the nodes of $V_t$ to the index of the star they belong to}
+      \State \textbf{return} $stars$, $inner\_edges$, $membership$
     \EndFunction
@@
 
-\extractStar{} terminates because at each iteration of the while loop line 8, we remove one node
-from $Q$ and never add any. Let us analyze the complexity when $|V_t|=n$ and $|E_t|=m$. We first
-build a priority queue of all the nodes sorted by their degree (line 5--7), which takes $O(n)$ time.
-Then, at each iteration of the inner loop, we find the center of the next star by extracting the
-maximum of the queue (line 9), we build the corresponding star (line 12--14) and we decrease the
+\begin{prop}
+  For any graph $G=(V,E)$, \extractStar$(G)$ terminates in $O(|E|)$ time.
+\end{prop}
+\begin{proof}
+\extractStar{} terminates because at each iteration of the while loop line 7, we remove one node
+from $Q$ and never add any. Let us now analyze the complexity. We first
+build a priority queue of all the nodes according to their degree (line 2), which takes $O(|V|)$ time.
+Then, at each iteration of the while loop, we find the center of the next star by extracting the
+maximum of the queue (line 8), we build the corresponding star (line 11--13) and we decrease the
 priority (\ie{} the degree) of all nodes adjacent to the new star (line 15--17).  Since both
-operations require constant time when using a Strict Fibonacci Heap~\autocite{FibonacciHeaps12} and
-there are $n$ iterations of that loop, a coarse approximation of the runtime of \extractStar{} is
-$O(n^2)$. However, observe that there can be at most $m$ decrease operations (since after that, all
-nodes still in the queue have an effective degree of $0$, meaning that $periphery$ will the empty
-set and lines 13--17 will run in constant time), reducing the complexity to $O(m+n)$.
-
-The other two variants are more time consuming because they require additional bookkeeping. Their
-randomization make them useful in an adversarial context but it also renders the analysis  of the
-resulting spanning tree more challenging\marginpars{Which means this paragraph can be skipped at first.}
-Therefore, we only briefly describe the implementations here.\footnote{Although they are available
-online at 
-\nolinkurl{https://github.com/daureg/magnet/blob/master/veverica/}%
-\{\href{https://github.com/daureg/magnet/blob/master/veverica/ThresholdSampler.py}%
-{ThresholdSampler.py}, \href{https://github.com/daureg/magnet/blob/master/veverica/NodeSampler.py}%
-{NodeSampler.py}\}.} For the threshold function, we
-maintain two queues, $high$ and $low$, containing nodes whose degree is respectively above and below
-the current threshold. We select a node \uar{} in $high$, remove the corresponding star from $G_t$,
-recompute the new threshold and if necessary, move nodes which fell under the threshold from $high$
-to $low$ and those who climb above the threshold from $low$ to $high$. For the degree function, we
-can draw any node as center proportionally to its weight (where the weight of node $u$ is defined as
-$d_f\left(\degr(u)\right)$). Yet we cannot use the standard method of computing the cumulative sum
-of weights since some of them change at each iteration. Therefore, we construct a binary tree whose
-leaves are the nodes of $V_t$ and where each tree nodes maintain the sum of weights in its left and
-right subtrees. To sample, we draw a random number $r$ between $0$ and the total weight of the tree and
-go down from the root to the leaf spanning the weight interval containing $r$.
-When degrees are updated (or graph node removed), we update the weights along a path from the
-corresponding leaves to the root of the tree.
-
-\medskip
+queue operations require constant time when using a Strict Fibonacci Heap~\autocite{FibonacciHeaps12} and
+there are $|V|$ iterations of that loop, a coarse approximation of the runtime of \extractStar{} is
+$O(|V|^2)$. However, observe that there can be at most $|E|$ decrease operations (since after that, all
+nodes still in the queue have an effective degree of $0$, meaning that $periphery$ will be the empty
+set and lines 12--17 will run in constant time), reducing the complexity to $O(|E|+|V|)$.
+\end{proof}
 
@@
 centered in $c_1$ and $c_2$.%
+
+\paragraph{Variants of \extractStar{}}
+\label{ssec:gtx_center_choice}
+
+The execution of \extractStar{} is mainly deterministic, except for the fact the ties between nodes
+with the same highest degree are broken arbitrarily. While this allows for an efficient
+implementation, and simplify the analysis of the resulting sequence of stars and therefore the
+induced spanning tree, in can be detrimental in an adversarial context, where we could end up with a
+tree forcing a lot of mistakes. We add an element of randomization to \extractStar{} by letting it
+use of two optional arguments, a \emph{threshold function} $\tau$ or a \emph{degree function}
+$\widetilde{d}$. Such functions modify the center sampling process in the following way:
+\begin{itemize}%[nosep]
+  \item if $n_{t,i}$ is the number of node remaining in $V_t$ before choosing the \ith{} center, choose
+    a node \uar{} among those with a degree larger than $\tau(n_{t,i})$. The idea is to choose
+    among a small set of high degree nodes, for instance by letting $\tau(n) = \sqrt{n}$. Note
+    however we cannot guarantee there will always be nodes with degree above the threshold, in which
+    case we default on the highest degree node
+  \item if $\degr_i(u)$ is the degree of node $u$ before choosing the \ith{} center, choose node
+    proportionally to $\widetilde{d}(\degr_i(u))$. Again, the degree function is designed so that it
+    favors the selection of high degree nodes. For instance, one could use
+    $\widetilde{d}(\degr_i(u)) = \degr_i(u)^2$.
+\end{itemize}
+
+These two variants are more time consuming because they require additional bookkeeping.
+Therefore, we don't provide a full complexity analysis and only briefly sketch their implementations
+here.\footnote{Although they are available online at 
+\nolinkurl{https://github.com/daureg/magnet/blob/master/veverica/}%
+\{\href{https://github.com/daureg/magnet/blob/master/veverica/ThresholdSampler.py}%
+{ThresholdSampler.py}, \href{https://github.com/daureg/magnet/blob/master/veverica/NodeSampler.py}%
+{NodeSampler.py}\}.} For the threshold function, we
+maintain two queues, $high$ and $low$, containing nodes whose degree is respectively above and below
+the current threshold. We select a node \uar{} in $high$, remove the corresponding star from $G_t$,
+recompute the new threshold and if necessary, move nodes which fell under the threshold from $high$
+to $low$ and those who climb above the threshold from $low$ to $high$. For the degree function, we
+can draw any node as center proportionally to its weight (where the weight of node $u$ is defined as
+$d_f\left(\degr(u)\right)$). Yet we cannot use the standard method of computing the cumulative sum
+of weights since some of them change at each iteration. Therefore, we construct a binary tree whose
+leaves are the nodes of $V_t$ and where each tree nodes maintain the sum of weights in its left and
+right subtrees. To sample, we draw a random number $r$ between $0$ and the total weight of the tree and
+go down from the root to the leaf spanning the weight interval containing $r$.
+When degrees are updated (or graph node removed), we update the weights along a path from the
+corresponding leaves to the root of the tree.

### 2017-06-28 notations.tex
improve description of extract star
- move the heuristics variant in a separate paragraph
- give more visibility to the proof of complexity
- streamline the pseudo code and give extra textual explanations

@@
     $\yuv{}$ & The sign of the edge $(u,v)$, which can be either $+1$ or $-1$ \\
-    $Y(E)$ & The labeling of $E$, that is the set of all signs of $E$: $Y(E) = \{\yuv , (u,v)\in E\}$ \\
+    $Y(E)$ & The labeling of $E$, that is the set of all signs of $E$: $Y(E) = \{\yuv : (u,v)\in E\}$ \\
     $\etrain$ & A subset of $E$, given or chosen, of which we observe the signs \\

### 2017-06-29 gtx_algorithm.tex
rewrite intro to galaxy tree

@@
-The \gtx{} algorithm takes as input an unweighted graph topology $G_0=(V_0, E_0)$ and produces a sequence of
-graphs $\{G_t\}_{t=1}^T$ of decreasing size until each connected component of $G_0$ is reduced to a
-single node. As we will prove later, after reaching this point, the algorithm has selected $|V_0| -
-1$ edges that form a spanning tree of $G_0$. Intuitively, $G_{t+1}$ is obtained from $G_t$ by composing two
-primitives so that we can informally write $G_{t+1} = \left(\collapseStar{} \circ
-\extractStar{}\right)(G_t)$.
+To achieve our objective of building a spanning tree that minimizes the distances between all
+connected pair of nodes in the original graph, we rely on a particular subgraph structure, namely
+the star.  We therefore introduce two algorithmic primitives: \extractStar{}, which partition a
+graph $G$ into a set of disjoint stars; and \collapseStar{}, which selects edges from $E$ to
+assemble these stars into a new, smaller graph. Given a graph topology $G_0=(V_0, E_0)$ and assuming
+for simplicity that $G_0$ consists of a single connected component,\footnote{For we can otherwise
+run our algorithm in parallel on each connected components of $G_0$.} the \gtx{} algorithm
+repeatedly applies these two primitives to produce a sequence of graphs $\{G_t\}_{t=0}^K$ of
+decreasing size, until $G_K$ is made of a single node. All the edges selected while reaching
+this stage then form the spanning tree we were looking for.
 % [from a CMU NIPS submission http://www.stat.cmu.edu/~arinaldo/papers/mutualfriends.pdf] We
@@
 
-\extractStar{} partitions the graph $G_t$ into a set of stars and \collapseStar{} build the graph
-made of those stars using the edges in $E_t$. We provide more details on those two operations in the
-following, as well as their complexity analysis. Then we state formally the \gtx{} algorithm and prove
-its termination and correctness.  Finally, we study its properties, such as the number of iterations
-needed to finish and the stretch of the resulting tree. For simplicity and without loss of
-generality, we assume that $G_0$ through $G_T$ consist of a single connected component.
-\marginpars{Actually, we only need to assume that $G_0$ is connected because then we can prove that
-the other subsequent $G_t$ will be as well.}
+In the following, we provide a more precise description of our two primitives and analyze their
+complexity. Then we state formally the complete \gtx{} algorithm, prove its termination and
+correctness, and show a detailed example of its execution.  Finally, we study its properties, such
+as the number of iterations needed to finish and the stretch of the resulting tree.
 

### 2017-06-29 lowstretch_and_cc.tex
rewrite intro to galaxy tree

@@
 In this section, we are interested in active learning algorithms, that first query a subset
-$\etrain$ of the edges, observe the signs in $\etest$ and use them to predict the remaining signs.
-More precisely, we focus on an algorithm that query a spanning tree $T$ of $G$ and predict the sign
+$\etrain$ of the edges, observe the signs in $\etrain$ and use them to predict the remaining signs.
+More precisely, we focus on an algorithm that queries spanning tree $T$ of $G$ and predicts the sign
 of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. While this
@@
 4.1]{Cesa-Bianchi2012b}, it makes sense intuitively that since each sign has been potentially
-flipped, the longer the path in $T$, the more likely its parity will be not be equal to the truth
-$\yuv{}$. Therefore we would like each such path to be as short as possible and in the rest of this
-section, we describe a way to build spanning tree tailored for this situation.
+flipped, the longer the path in $T$, the more likely its parity will be not be equal to the true
+sign $\yuv{}$. Therefore we would like each such path to be as short as possible. In the following,
+we describe a way to build spanning trees tailored for this situation.
 
-\subsection{\gtx{}: a simple low stretch tree construction}
+\subsection{\gtx{}: a spanning tree designed for sign prediction}
 \label{sub:gtx_a_simple_low_stretch_tree_construction}

### 2017-06-29 gtx_algorithm.tex
clarify extract star and more detailed complexity proof

@@
 representation of this notation).
+Upon completion, it returns a list of stars,
+the set of all the edges within a star,
+and a map (or associative array) that associates each node of $V_t$ to the index of the
+unique star it belongs to.
+According to the definition of \textcite{HashTableBook08}, an associative array is an abstract data
+type composed of a collection of (key, value) pairs, such that each possible key appears at most
+once in the collection. It efficiently supports the addition, removal and modification of a pair, as
+well as the lookup of a value associated with a particular key.
 \begin{marginfigure}
@@
   \label{fig:gtx_star_simple}
-\end{marginfigure}%
-Upon completion, it returns a list of stars,
-a list of each star edges,
-and a map (or associative array) of each node of $V_t$ to the unique star it belongs to.
-According to the definition of \textcite{HashTableBook08}, an associative array is an abstract data
-type composed of a collection of (key, value) pairs, such that each possible key appears at most
-once in the collection. It efficiently supports the addition, removal and modification of a pair, as
-well as the lookup of a value associated with a particular key.
+\end{marginfigure}
 
@@
 sample centers by choosing the node with the current highest degree, with ties broken
-arbitrarily.\footnote{We also consider more involved heuristics but in the interest of simplicity,
-they are presented later on~\vpageref{ssec:gtx_center_choice}.} This is achieved efficiently by
+arbitrarily.\footnote{We also consider more involved heuristics but, in the interest of simplicity,
+they are presented later~\vpageref{ssec:gtx_center_choice}.} This is achieved efficiently by
 maintaining a max-priority queue $Q$, initially containing all the nodes of $V_t$. The priority of a
-node is its current degree and we equipped $Q$ with two standard operations described
+node is its current degree and we equip $Q$ with two standard operations described
 by~\textcite[section 6.5]{CormenAlgo09}: \textsc{Extract-Max}$(Q)$ removes and returns the
@@
 We also assume that $G$ is the adjacency list of the graph, so that $G[u]$ is the set of neighbors of
-$u$, \ie{} $G[u] \equiv \mathcal{N}(u)$. Finally $membership$ is a map, initially empty, and we
+$u$, \ie{} $G[u] \equiv \mathcal{N}(u)$. Finally $membership$ is a map and we
 define the \textsc{Star} function, which creates a star given a center $c$, a list $periphery$ of
@@
       \State Let $Q$ be the max-priority queue described above
-      \State Let $remaining$ be a set of nodes initially containing all the nodes in $V_t$
+      \State Let $remaining$ be a set of nodes, initially containing all the nodes in $V_t$
       \State Let $membership$ be an empty map
@@
         \EndIf
-        \Let{$periphery$}{$G[c] \bigcap remaining$}
+        \Let{$periphery$}{$G_t[c] \bigcap remaining$}
         \Let{$stars$}{$stars \bigcup \{$\Call{Star}{$c$, $periphery$, $|stars|$}\}}
         \Let{$inner\_edges$}{$inner\_edges \bigcup \{(c, p): p \in periphery\}$}
-        \Let{$remaining$}{$remaining \setminus \left\{ \{c_i\} \cup \setminus periphery\right\}$}
+        \Let{$remaining$}{$remaining \setminus \left\{ \{c_i\} \cup periphery\right\}$}
         \For{$p$ in $periphery$}
-          \For{$h$ in $G[p] \bigcap remaining$}
+          \For{$h$ in $G_t[p] \bigcap remaining$}
             \State \Call{Decrease-Key}{$Q$, $h$, $1$}
@@
 \begin{prop}
-  For any graph $G=(V,E)$, \extractStar$(G)$ terminates in $O(|E|)$ time.
+  For any connected graph $G=(V,E)$, \extractStar$(G)$ terminates in $O(|E|)$ time.
 \end{prop}
@@
 \extractStar{} terminates because at each iteration of the while loop line 7, we remove one node
-from $Q$ and never add any. Let us now analyze the complexity. We first
-build a priority queue of all the nodes according to their degree (line 2), which takes $O(|V|)$ time.
-Then, at each iteration of the while loop, we find the center of the next star by extracting the
-maximum of the queue (line 8), we build the corresponding star (line 11--13) and we decrease the
-priority (\ie{} the degree) of all nodes adjacent to the new star (line 15--17).  Since both
-queue operations require constant time when using a Strict Fibonacci Heap~\autocite{FibonacciHeaps12} and
-there are $|V|$ iterations of that loop, a coarse approximation of the runtime of \extractStar{} is
-$O(|V|^2)$. However, observe that there can be at most $|E|$ decrease operations (since after that, all
-nodes still in the queue have an effective degree of $0$, meaning that $periphery$ will be the empty
-set and lines 12--17 will run in constant time), reducing the complexity to $O(|E|+|V|)$.
+from $Q$ and never add any. Let us now analyze its complexity. We first build a priority queue of
+all the nodes according to their degree (line 2), which requires $O(|V|)$ insertions into $Q$. Then
+we execute $|V|$ iterations of the while loop. However, the main idea here is that we process each
+node and each edge exactly once. We first find the center $c$ of the next star by extracting the
+maximum of the queue (line 8) and testing if $c$ is still part of the graph, which happens $|V|$
+time. Then we build the corresponding star (line 11--14). In total, we test the membership of $|V|$
+nodes in line 11, $\textsc{Star}$ updates the $membership$ map $|V|$ times in line 12,
+$inner\_edges$ consists of $|E|$ edges at most in line 13 and $remaining$ is only updated $|V|$
+times in line 14. Finally, we decrease the priority (\ie{} the degree) of all nodes adjacent to the
+new star (line 15--17). Each decrement is supported by a single edge, thus \textsc{Decrease-Keys} is
+called at most $|E|$ times. Since all queue operations require constant time when using a Strict
+Fibonacci Heap~\autocite{FibonacciHeaps12}, the complexity is $O(|E|+|V|)$, which is also $O(|E|)$
+since $G$ is connected.
 \end{proof}

### 2017-06-29 gtx_algorithm.tex
typos

@@
 \paragraph{\collapseStar{}}\label{par:collapsestar}%
-The second routine, \collapseStar{} takes as input the result of \extractStar{}, along with $E_t$
+The second primitive, \collapseStar{} takes as input the result of \extractStar{}, along with $E_t$
 and an optional $\emph{eccentricity}$ array we will describe soon\marginpars{Actually, we could make
@@
 
-We illustrate the operation of the \gtx{} algorithm on small (and somewhat contrived) example.
+We illustrate the operation of the \gtx{} algorithm on a small (and somewhat contrived) example.
 Let us start with the initial graph $G_0$ depicted in \autoref{fig:gtx_eccentricity}
@@
 \end{marginfigure}
-the degree of a node $u$ drops to zero because because all of its neighbors were claimed by the
+the degree of a node $u$ drops to zero because all of its neighbors were claimed by the
 periphery of previous stars. Such a node then forms a \emph{singleton star}, as illustrated in
@@
 construct a graph with singletons where the factor of reduction in number of nodes and edges can be
-arbitrarily close to $1$. Consider \autoref{fig:gtx_noreduc} and assume that first extract two stars
+arbitrarily close to $1$. Consider \autoref{fig:gtx_noreduc} and assume that we first extract two stars
 centered in $c_1$ and $c_2$.%

### 2017-06-29 gtx_algorithm.tex
use the star macros in gtx example

@@
 $4$, achieved at nodes $\{1, 6, 11, 16, 21, 26, 31, 36, 41\}$. For the sake of simplicity, assume
-nodes are picked according to their index. First, node $1$ is forms the star
-$\textcolor{DodgerBlue}{S_1^1}$ with peripheral nodes $2$, $3$, $4$ and $5$. This increments the
-eccentricity of those peripheral nodes by $1$. Then node $6$ forms its star
-$\textcolor{DodgerBlue}{S_2^1}$ with $7$, $8$, $9$ and $10$. The process continues until node $41$ is
-chosen to be the center of star $\textcolor{DodgerBlue}{S_9^1}$, at which point the max-priority
-queue has been exhausted and \extractStar{} finishes.
+nodes are picked according to their index. First, node $1$ forms the star $\starone{1}$ with
+peripheral nodes $2$, $3$, $4$ and $5$. This increments the eccentricity of those peripheral nodes
+by $1$. Then node $6$ forms its star $\starone{2}$ with $7$, $8$, $9$ and $10$. The process
+continues until node $41$ is chosen to be the center of star $\starone{9}$, at which point the
+max-priority queue has been exhausted and \extractStar{} finishes.
 
@@
 
-We then call \collapseStar{}, with the eccentricity reducing variant. This will connect all possible
-pairs of star. For instance, the edge between nodes $19$ and $29$ leads to the edge
-between $\textcolor{DodgerBlue}{S_4^1}$ and $\textcolor{DodgerBlue}{S_6^1}$. This is actually the
-only possible edge between $\textcolor{DodgerBlue}{S_4^1}$ and $\textcolor{DodgerBlue}{S_6^1}$.
-Consider on the other hand the case of edges $(2, 6)$ and $(2, 9)$. They both connect
-$\textcolor{DodgerBlue}{S_1^1}$ and $\textcolor{DodgerBlue}{S_2^1}$. Yet at this point of the algorithm,
-the eccentricity of node $2$ is $1$, the eccentricity of node $6$ is $0$ and the eccentricity of node
-$9$ is $1$. The edge $(2, 6)$ has therefore the smallest total eccentricity and is chosen to connect
-$\textcolor{DodgerBlue}{S_1^1}$ and $\textcolor{DodgerBlue}{S_2^1}$. The full result of the
+We then call \collapseStar{}. This will connect all possible pairs of star. For instance, the edge
+between nodes $19$ and $29$ leads to the edge between $\starone{4}$ and $\starone{6}$. This is
+actually the only possible edge between $\starone{4}$ and $\starone{6}$.  Consider on the other hand
+the case of edges $(2, 6)$ and $(2, 9)$. They both connect $\starone{1}$ and $\starone{2}$. Yet at
+this point of the algorithm, the eccentricity of node $2$ is $1$, the eccentricity of node $6$ is
+$0$ and the eccentricity of node $9$ is $1$. The edge $(2, 6)$ has therefore the smallest total
+eccentricity and is chosen to connect $\starone{1}$ and $\starone{2}$. The full result of the
 \collapseStar{} procedure is $G_1$, which can be seen on \autoref{fig:gtx_run1}.
@@
 We now run \extractStar{} on $G_1$. Because all nodes have degree $2$, they could all be chosen to
-be the center of a star yet we again consider them ordered by their index and therefore choose
-$\textcolor{DodgerBlue}{S_1^1}$ to be the center of the star $\textcolor{Orange}{S_1^2}$ with
-peripheral nodes $\textcolor{DodgerBlue}{S_2^1}$ and $\textcolor{DodgerBlue}{S_3^1}$. The original
-nodes belonging to those peripheral stars (nodes $4$ to $15$) have their eccentricity incremented by
-$1$. The next node with highest degree in $G_1$ is now $\textcolor{DodgerBlue}{S_4^1}$, which forms
-a star with $\textcolor{DodgerBlue}{S_5^1}$ and $\textcolor{DodgerBlue}{S_6^1}$. This choice means
-that nodes $21$ through $30$ have their eccentricity incremented by $1$. Finally,
-$\textcolor{DodgerBlue}{S_7^1}$ forms the last star with $\textcolor{DodgerBlue}{S_8^1}$ and
-$\textcolor{DodgerBlue}{S_9^1}$. Then \collapseStar{} connects the resulting three stars, and this
-time there is only a single choice between each pair of stars, leading to the graph $G_2$ showed in
-\autoref{fig:gtx_run2}
+be the center of a star yet we again they are picked according to their index and therefore we
+choose $\starone{1}$ to be the center of the star $\startwo{1}$ with peripheral nodes $\starone{2}$
+and $\starone{3}$. The original nodes belonging to those peripheral stars (nodes $4$ to $15$) have
+their eccentricity incremented by $1$. The next node with highest degree in $G_1$ is now
+$\starone{4}$, which forms a star with $\starone{5}$ and $\starone{6}$. This choice means that nodes
+$21$ through $30$ have their eccentricity incremented by $1$. Finally, $\starone{7}$ forms the last
+star with $\starone{8}$ and $\starone{9}$. Then \collapseStar{} connects the resulting three stars,
+and this time there is only a single choice between each pair of stars, leading to the graph $G_2$
+showed in \autoref{fig:gtx_run2}
 
 The action of \extractStar{} on $G_2$ is quite simple, because there is only one star that can be
-created, so let say we choose $\textcolor{Orange}{S_1^2}$ as its center, with
-$\textcolor{Orange}{S_2^2}$ and $\textcolor{Orange}{S_3^2}$ as peripheral nodes.  This increases the
-eccentricity of their underlying $G_0$ nodes by $1$ (namely nodes $16$ to $45$).  Because there is
-only one star $\textcolor{Green}{S_1^3}$ left, \collapseStar{} returns $G_3$ showed in
-\autoref{fig:gtx_run3} and an empty list of outer edges, meaning that the inner loop of \gtx{} is
-finished and we can go through every edges we chose between stars at every level to recover the
-final spanning tree, showed with black edges in \autoref{fig:gtx_eccentricity}
+created, so let say we choose $\startwo{1}$ as its center, with $\startwo{2}$ and $\startwo{3}$ as
+peripheral nodes. This increases the eccentricity of their underlying $G_0$ nodes by $1$ (namely
+nodes $16$ to $45$). Because there is only one star $\starthree{1}$ left, \collapseStar{} returns
+$G_3$ showed in \autoref{fig:gtx_run3} and an empty list of outer edges, meaning that the inner loop
+of \gtx{} is finished and we can go through every edges we chose between stars at every level to
+recover the final spanning tree, showed with solid edges in \autoref{fig:gtx_eccentricity}
 \vpageref{fig:gtx_eccentricity}. For completeness, we can also look at the edges which are not part

### 2017-06-29 assets/tikz/gtx_cross_edges_tikz.tex
clarify contractstar and its proof

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
\tikzstyle{star}=[DodgerBlue,line width=1.3pt,fill=White]
\tikzstyle{star-label}=[font=\Large]
\begin{tikzpicture}[auto,vertex/.append style={minimum width=7mm},
  edge/.append style={line width=1.5pt,color=DodgerBlue}]

  \draw[draw=none, fill opacity=.7, fill=Red] (-.30,-1.5) -- (.77,-3.34) -- (2.34,-2.42) -- (1.37,-.77) -- cycle;
  \draw[star] (0,0) circle [radius=1.7]  node[star-label,xshift=-30, yshift=25] {${S_1^1}$};
  \draw[star] (2.189,-3.948) circle [radius=1.7]  node[star-label,xshift=30, yshift=0] {${S_2^1}$};

  \node[vertex,fill=black!0,] (n1) at (0.000, -0.000) {1};
  \node[vertex,fill=black!25,] (n2) at (0.223, -1.179) {2};
  \node[vertex,fill=black!25,] (n3) at (-1.179, -0.223) {3};
  \node[vertex,fill=black!25,] (n4) at (-0.223, 1.179) {4};
  \node[vertex,fill=black!25,] (n5) at (1.179, 0.223) {5};
  \node[vertex,fill=black!0, ]  (n6) at (2.189,-3.948) {6};
  \node[vertex,fill=black!25,]  (n7) at (2.908,-4.909) {7};
  \node[vertex,fill=black!25,]  (n8) at (1.228,-4.667) {8};
  \node[vertex,fill=black!25,]  (n9) at (1.471,-2.987) {9};
  \node[vertex,fill=black!25,] (n10) at (3.150,-3.230) {10};

  \draw[edge] (n1) edge [bend left =0] node[] {} (n2);
  \draw[edge] (n1) edge [bend left =0] node[] {} (n3);
  \draw[edge] (n1) edge [bend left =0] node[] {} (n4);
  \draw[edge] (n1) edge [bend left =0] node[] {} (n5);

  \draw[edge,Orange] (n2) edge [bend left =29] node[pos=0.35,below=0pt] {\small $1$} (n6);
  \draw[edge,Orange] (n2) edge [bend left =0] node[above=0pt] {\small $2$} (n9);
  \draw[edge,Orange] (n3) edge [bend left =16] node[above=1pt] {\small $2$} (n8);
  \draw[edge,Orange] (n5) edge [bend left =-25] node[pos=0.3,below=2pt] {\small $2$} (n10);

  \draw[edge] (n6) edge [bend left =0] node[] {} (n7);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n8);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n9);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n10);
  \node[text=Red!80!White,anchor=west,font=\small] (ce) at (1.5, -1) {$cross\_edges[(s_1, s_2)]$};
\end{tikzpicture}
\end{document}

### 2017-06-29 gtx_algorithm.tex
clarify contractstar and its proof

@@
 \paragraph{\collapseStar{}}\label{par:collapsestar}%
-The second primitive, \collapseStar{} takes as input the result of \extractStar{}, along with $E_t$
-and an optional $\emph{eccentricity}$ array we will describe soon\marginpars{Actually, we could make
-eccentricity the default behavior, which would simplify the pseudo code}. It builds a new graph $G_{t+1}$
-where each star becomes a node and where there is a link between two nodes $s_1$ and $s_2$ if the nodes
-making up $s_1$ and $s_2$ are connected in $E_t$.
+
+The second primitive, \collapseStar{} takes as input the edges $E_t$ of the current graph, along
+with the $membership$ result of \extractStar{} and an $eccentricity$ array. It builds a new graph
+$G_{t+1}$ where each star from \extractStar{} becomes a node and where there is a link between two
+nodes $s_u$ and $s_v$ if the nodes in $V_t$ making up $s_u$ and $s_v$ are connected in $E_t$. The
+$eccentricity$ array is needed because when connecting two stars, we would prefer to join their
+center rather than two of their peripheral points. Therefore, we maintain an eccentricity count for
+all of the nodes of the original $G_0$, which is incremented by $1$ each time a node is chosen to be
+on the periphery of a star. In other words, the eccentricity of an original node quantifies to which
+extent it has been pushed to the border of the galaxy.
+
+\begin{marginfigure}
+  \centering
+  \includegraphics[width=0.92\linewidth]{assets/tikz/gtx_cross_edges_tikz.pdf}
+  \caption[Cross edge representation]{In that graphs, there are four possible edges between the two
+  stars $s_1$ and $s_2$, and all are part of $cross\_edges[(s_1, s_2)]$ here represented in light
+red. Those edges are labeled with the sum of eccentricity of their underlying endpoints, and the
+minimal one is linking a peripheral node of $s_1$ directly to the center of $s_2$.}
+  \label{fig:gtx_cross_edge}
+\end{marginfigure}%
+\collapseStar{} not only returns the new graph $G_{t+1}$ but also a map $cross\_edges$. This map is
+used to keep track of which edge in $E_t$ connects any pair of node in $V_{t+1}$ (as illustrated in
+\autoref{fig:gtx_cross_edge}). It associates to any edge in $E_{t+1}$ a set of edges from $E_t$.
+This set can have an arbitrary size during the execution of \collapseStar{}, yet it is guaranteed to
+contain a single edge upon termination. Although this does not appear in the following pseudo code,
+in practice we ensure that $(s_u, s_v)$ and $(s_v, s_u)$ refer to the same edge in $cross\_edges$.
 
@@
       \State Let $G_{t+1}$ be an empty graph
-      \State Let $cross\_edges$ be a mapping from edges in $G_{t+1}$ to set of edges in $G_t$
-      \ForAll{edge $(u, v)$ in \Call{Shuffle}{$E_t$}}
-        \Let{$s_u,\,s_v$}{$membership[u]$, $membership[v]$}%
-        \Comment{swap if needed so that $s_u < s_v$}
-        \If{$s_u = s_v$}
-          \State \textbf{continue}
-        \EndIf
-        \If{$eccentricity$ is not \textbf{null}}
+      \State Let $cross\_edges$ be the map described above
+      \ForAll{edge $(u, v)$ in $E_t$}
+      \Let{$s_u$}{$membership[u]$}
+      \Let{$s_v$}{$membership[v]$}
+        \If{$u$ and $v$ are not in the same star (\ie{} $s_u \neq s_v$)}
           \Let{$cross\_edges[(s_u, s_v)]$}{$cross\_edges[(s_u, s_v)] \bigcup \{(u, v)\}$}
-          % \State $cross\_edges[(s_u, s_v)].$\Call{Add}{$(u, v)$}
-        \ElsIf{$(s_u, s_v)$ not in $cross\_edges$}
-          \Let{$cross\_edges[(s_u, s_v)]$}{$\{(u, v)\}$}
-          \State Add edge $(s_u, s_v)$ to $G_{t+1}$
         \EndIf
       \EndFor
-      \If{$eccentricity$ is not \textbf{null}}
-        \ForAll{$\Big(\underbrace{(s_u,s_v)}_{\text{key}},\,
-          \underbrace{candidate\_edges}_{\text{value}}\Big)$
-          in $cross\_edges.$\Call{Items}{$\,$}}
-          \Let{$(u_0, v_0)$}{$\argmin_{(u,v)\in candidate\_edges} eccentricity[u]+eccentricity[v]$}
+      \ForAll{pair of will-be connected stars $(s_u,s_v)$ in $cross\_edges$}
+        \Let{$possible\_underlying\_edges$}{$cross\_edges[(s_u, s_v)]$}
+        % \Let{$(u_0, v_0)$}{$\argmin_{(u,v)\in candidate\_edges} eccentricity[u]+eccentricity[v]$}
+        \State $$(u_0, v_0) \gets \argmin_{(u,v)\in possible\_underlying\_edges} eccentricity[u]+eccentricity[v]$$
         \Let{$cross\_edges[(s_u, s_v)]$}{$\{(u_0, v_0)\}$}
-          \State Add edge $(s_u, s_v)$ to $G_{t+1}$
+        \Let{$E_{t+1}$}{$E_{t+1} \bigcup \{(s_u, s_v)\}$}
       \EndFor
-      \EndIf
       \State \textbf{return} $G_{t+1}$, $cross\_edges$
@@
 
-For that, we first shuffle $E_t$ and iterate on it (line 4). When we find an edge whose endpoints
-belong to two different stars not yet connected (line 10), we use that edge to connect these two
-stars (line 11--12) This trivially takes $O(m)$ times.  A variant instead keeps track of all edges
-connecting each pair of stars (line 9) to choose one that will best contribute to our low stretch
-objective. Namely, when connecting two stars, we would prefer to join their centers rather than two
-peripheral points. For that we maintain an eccentricity count for all of the nodes of the original
-$G_0$, which is incremented by $1$ each time a node is chosen to be on the periphery of a
-star.\footnote{We will discuss this construction in more details when describing the main \gtx{}
-function since it takes place there.} For each pair of stars, we thus choose the edge across them
-with minimal sum of its endpoints' eccentricity (line 15--17)\marginpars{Line 15 is actually a
-simplification since the $eccentricity$ array is only defined for nodes in $V_0$ so when
-\collapseStar{} is called on $G_3$ for instance, $u$ and $v$ are nodes of $V_3$ and we have to
-retrieve the actual corresponding edge in $E_0$. Maybe I could simply add a HashMap from $E_t$ to
-$E_0$ like in the python code…}. This only requires another pass over the edges, thus preserving the
-$O(m)$ runtime.
+\begin{prop}
+  \label{prop:gtx_time_collapsestar}
+  For any graph $G=(V,E)$, \collapseStar$(E$, $membership$, $eccentricity)$ terminates in $O(|E|)$
+  time.
+\end{prop}
+\begin{proof}
+  \marginpars{\small Line 11 is actually a simplification since the $eccentricity$ array is only
+  defined for nodes in $V_0$ so when \collapseStar{} is called on $G_3$ for instance, $u$ and $v$
+  are nodes of $V_3$ and we have to retrieve the actual corresponding edge in $E_0$. Maybe I could
+  simply add a map from $E_t$ to $E_0$ like in the real python code…}
+  The analysis of \collapseStar is rather straightforward because the function only executes two
+  loops over its edges argument. During the for loop of line 4, we only perform constant time
+  operations on maps. Likewise, in the for loop of line 9, the most expensive is finding the minimum
+  in line 11. Computing the eccentricity of an edge is done in constant time and only once for each
+  edge of $E$.  As a consequence, the total time of \collapseStar{} is indeed $O(|E|)$.
+\end{proof}
 

### 2017-06-30 assets/tikz/gtx_cross_edges_tikz.tex
typo: change name of stars in cross_edges figure

@@
   \draw[draw=none, fill opacity=.7, fill=Red] (-.30,-1.5) -- (.77,-3.34) -- (2.34,-2.42) -- (1.37,-.77) -- cycle;
-  \draw[star] (0,0) circle [radius=1.7]  node[star-label,xshift=-30, yshift=25] {${S_1^1}$};
-  \draw[star] (2.189,-3.948) circle [radius=1.7]  node[star-label,xshift=30, yshift=0] {${S_2^1}$};
+  \draw[star] (0,0) circle [radius=1.7]  node[star-label,xshift=-30, yshift=25] {${s_1}$};
+  \draw[star] (2.189,-3.948) circle [radius=1.7]  node[star-label,xshift=30, yshift=0] {${s_2}$};
 

### 2017-06-30 gtx_algorithm.tex
clarify gtx algorithm and start proving it

@@
         \Let{$possible\_underlying\_edges$}{$cross\_edges[(s_u, s_v)]$}
-        % \Let{$(u_0, v_0)$}{$\argmin_{(u,v)\in candidate\_edges} eccentricity[u]+eccentricity[v]$}
+        \marginpars{\small Line 11 is actually a simplification since the $eccentricity$ array is only
+        defined for nodes in $V_0$ so when \collapseStar{} is called on $G_3$ for instance, $u$ and $v$
+        are nodes of $V_3$ and we have to retrieve the actual corresponding edge in $E_0$. Maybe I could
+        simply add a map from $E_t$ to $E_0$ like in the real python code…}
         \State $$(u_0, v_0) \gets \argmin_{(u,v)\in possible\_underlying\_edges} eccentricity[u]+eccentricity[v]$$
@@
 \begin{proof}
-  \marginpars{\small Line 11 is actually a simplification since the $eccentricity$ array is only
-  defined for nodes in $V_0$ so when \collapseStar{} is called on $G_3$ for instance, $u$ and $v$
-  are nodes of $V_3$ and we have to retrieve the actual corresponding edge in $E_0$. Maybe I could
-  simply add a map from $E_t$ to $E_0$ like in the real python code…}
   The analysis of \collapseStar is rather straightforward because the function only executes two
@@
 final spanning tree, we need additional work, namely updating the eccentricity of the nodes of
-$V_0$ and keeping track of each edge within and between stars along every contraction level.
+$V_0$ and keeping track of each edge within and between stars along every collapse steps. Despite
+our earlier promise, this entails showing some bookkeeping code, because it has an influence on the
+runtime of \gtx{}. \Todo{Those maps could probably use better names}
+Namely in the listing of \autoref{alg:gtx} \vpageref{alg:gtx}, we use the three following maps:
 
-\begin{algorithm}
+\vspace{0.5\baselineskip}
+\noindent\begin{tabulary}{187mm}{lllL}
+  \toprule
+  name               & keys set at $t$ & values set at $t$   & comment \\
+  \midrule
+  $star\_membership$ & nodes in $V_t$  & nodes in $V_{t+1}$  & this is the map returned by
+  \extractStar{}$(G_t)$ \\
+  $full\_membership$ & nodes in $V_0$  & nodes in $V_{t+1}$  & at $t=0$, this is equals to
+  $star\_membership$ and then it gets updated at every iteration to maintain its original keys set.  \\
+  $original\_node$   & nodes in $V_t$  & sets of $V_0$ nodes & This can be seen as the reverse of
+  $full\_membership$ and it is the one needed to update the eccentricity of the original nodes. \\
+  \bottomrule
+\end{tabulary}
+\vspace{0.5\baselineskip}
+
+\begin{algorithm}[tbh]
   \caption{\gtx{}($G_0=(V_0,E_0)$) \label{alg:gtx}}
@@
       \Let{$stars$, $inner\_edges$, $membership$}{\Call{Extract-Stars}{$G_t$}}
-      \State $inner\_edges\_seq$.\Call{Append}{$inner\_edges$}
+      \State Add $inner\_edges$ to the list $inner\_edges\_seq$
       \State \Call{Update-Eccentricity}{$stars$, $eccentricity$, $original\_node$}
@@
       $\!membership$, $\!eccentricity$}}
-      \State $outer\_edges\_seq$.\Call{Append}{$outer\_edges$}
+      \State Add $outer\_edges$ to the list $outer\_edges\_seq$
       \Let{$G_{t}$}{$G_{t+1}$}
     \Until{$|outer\_edges|>0$}
-    \State \textbf{return} \Call{Assemble-Spanning-Trees}{$inner\_edges\_seq$, $outer\_edges\_seq$}
+    \State \textbf{return} \Call{Assemble-Spanning-Tree}{$inner\_edges\_seq$, $outer\_edges\_seq$}
 		\begin{center}
@@
 
-As described in \autoref{alg:gtx}, \marginpars{I didn't write pseudo code for the three helping functions but
-instead put their verbatim python version. They are mostly plumbing code, the only interesting
-aspect is their complexity.  \texttt{update\_centrality} and \texttt{update\_nodes\_mapping} takes
-$O(n_0)$ time since they go through every node of the original graph. As for
-\texttt{assemble\_spanning\_tree}, it goes again through every edges visited at every level of
-contraction, meaning it's $O(Tm)$. However, we already visited those edges at least once so it only
-adds a constant factor $2$ in the overall complexity of the \gtx{} algorithm.}
-at every collapse level, we first extract stars from the current graph (line 5), then
-update the eccentricity and nodes mapping (lines 7--8) and finally collapse the graph (line 9) We
-perform these operations until there are no edge connecting stars anymore. At this point, we revisit
-every outer edges to form the spanning tree (line 13). Because the main loop does not perform any
-work besides calling other function, assuming there are T collapse steps, the complexity of \gtx{}
-is $O(T(m+n)) = O(Tm)$.
+As described in \autoref{alg:gtx}, at every collapse level, we first extract stars from the current
+graph (line 5), then update the eccentricity and node mappings (lines 7--8) and finally collapse the
+graph (line 9). We perform these operations until there are no edge connecting stars anymore. At
+this point, we revisit every outer edges to build the spanning tree (line 13).
+
+\begin{prop}
+  \label{prop:gtx_correct}
+  For any connected graph $G_0$, \gtx$(G_0)$ terminates after $K\leq |V_0|$ iterations. Furthermore,
+  it runs in $O(K|E_0|)$ time and returns a spanning tree of $G_0$.
+\end{prop}
+We will need the following lemma.
+\begin{lemma}
+  \label{lem:gtx_stay_connected}
+  If $G_0$ is connected, then any subsequent graph $G_t,\, t \leq K$ is also connected.
+\end{lemma}
+\begin{proof}
+ Suppose not and assume, to the contrary, that there exists at least one (or more) disconnected
+ graph in $\{G_t\}_{t=1}^K$ and let $t_0$ be the smallest index such that $G_{t_0}$ is disconnected.
+ Then there exist two nodes $s_u$ and $s_v$ in $V_{t_0}$ with no path between them in $E_{t_0}$.
+ Letting $\mathcal{U}, \mathcal{V} \subset V_{t_0-1}$ be respectively the nodes forming stars $u$
+ and $v$, this implies there is no path between nodes in $\mathcal{U}$ and nodes in $\mathcal{V}$.
+ However, $G_{t_0-1}$ is connected by hypothesis, which leads to a contradiction.
+\end{proof}
+
+\begin{proof}[Proof of \autoref{prop:gtx_correct}]
+  Let us first show that \gtx{} terminates in less than $|V_0|$ iterations. This follows from the
+  fact that every time we collapse the graph $G_t$, we strictly reduce the number of nodes. Indeed,
+  according to \autoref{lem:gtx_stay_connected}, $G_t$ is connected so at least two nodes of $V_t$
+  are joined by an edge and will form a star, \ie{} a single node. We can thus claim that $|V_{t+1}| <
+  |V_t|$. Note also that for all $t$, $|V_t| > 0$ and that when $|V_t|=1$, \extractStar{} creates a
+  singleton and \collapseStar{} does not return any outer edges so \gtx{} finishes, proving that the
+  number of iterations $K$ satisfies $K \leq |V_0|$.
+
+ Then we analyze the time complexity. We already now that during the \tth{} iteration,
+ \extractStar{} and \collapseStar{} take time $O(|E_t|)$. We provide only the actual python code
+ instead of pseudo code for the remaining functions since they do not present any special
+ interesting algorithmic aspect. However, we can say that \textsc{Update-Eccentricity} and
+ \textsc{Update-Nodes-Mapping} take $O(|V_0|)$ time since they go through every node of the original
+ graph. Because $G_0$ is connected, $|V_0| \leq |E_0|$, and since for all $t$, $|E_t| \leq |E_0|$,
+ we have that the \gtx{} inner loop takes $O(K|E_0|)$ time. Finally, since
+ \textsc{assemble-spanning-tree} goes again through every edge visited at every collapse step, it
+ also takes $O(K|E_0|)$ time, which is thus the overall complexity of the \gtx{} algorithm.
+
+  Finally we prove that \gtx{} indeed returns a spanning tree of $G_0$.
+\end{proof}
 

### 2017-06-30 gtx_algorithm.tex
typos

@@
 \begin{prop}
+  \label{prop:gtx_time_extractstar}
   For any connected graph $G=(V,E)$, \extractStar$(G)$ terminates in $O(|E|)$ time.
@@
 from $Q$ and never add any. Let us now analyze its complexity. We first build a priority queue of
-all the nodes according to their degree (line 2), which requires $O(|V|)$ insertions into $Q$. Then
+all the nodes according to their degree (line 2), which requires $|V|$ insertions into $Q$. Then
 we execute $|V|$ iterations of the while loop. However, the main idea here is that we process each
@@
   \includegraphics[width=0.92\linewidth]{assets/tikz/gtx_cross_edges_tikz.pdf}
-  \caption[Cross edge representation]{In that graphs, there are four possible edges between the two
-  stars $s_1$ and $s_2$, and all are part of $cross\_edges[(s_1, s_2)]$ here represented in light
+  \caption[Cross edge representation]{In this graph, there are four possible edges between the two
+  stars $s_1$ and $s_2$, and all are part of $cross\_edges[(s_1, s_2)]$, here represented in light
 red. Those edges are labeled with the sum of eccentricity of their underlying endpoints, and the
@@
 \begin{proof}
-  The analysis of \collapseStar is rather straightforward because the function only executes two
-  loops over its edges argument. During the for loop of line 4, we only perform constant time
-  operations on maps. Likewise, in the for loop of line 9, the most expensive is finding the minimum
+  The analysis of \collapseStar{} is rather straightforward because the function only executes two
+  loops over $E$. During the for loop of line 4, it only performs constant time
+  operations on maps. Likewise, in the for loop of line 9, the most expensive operation is finding the minimum
   in line 11. Computing the eccentricity of an edge is done in constant time and only once for each
@@
 
-\paragraph{Exemple of \gtx{}}
+\paragraph{Example of \gtx{}}
 \label{par:exemple_of_gtx}
@@
 is the line graph. While it is already a tree, let us look how the \gtx{} algorithm operates on it.
-Say we have $n$ nodes in that line. There a star is made at most of three consecutive nodes. In the
-worst case, centers will be chosen such that we have a succession of three and two nodes stars (as
-in \autoref{fig:gtx_line_graph}).%
+Say we have $n$ nodes and $m$ edges in that line. There a star is made at most of three
+consecutive nodes. In the worst case, centers will be chosen such that we have a succession of three
+and two nodes stars, as in \autoref{fig:gtx_line_graph}.%
 \begin{marginfigure}
@@
 This will results in $\nicefrac{2n}{5}$ stars, which is less than half of $n$ and because in that
-case $n\approx m$, \gtx{} will finish after $O(\log m)$ iterations. Note that a barbell graph (two
-cliques connected by a line) would require a number of iterations proportional to the length of that
+case $n=m+1$, \gtx{} will finish after $O(\log m)$ iterations. Note that a barbell graph (two
+cliques connected by a line) would requires a number of iterations proportional to the length of that
 central line, despite having many more edges. This suggests unsurprisingly that the diameter of the graph
-could be a good parameter to quantify the number of iteration needed. Indeed, take a tree a consider
+could be a good parameter to quantify the number of iterations needed. Indeed, take a tree and consider
 the length $p$ of its longest path from the root to a leaf. By a similar argument as the one used in
@@
 
-While we assume that $G_t$ is connected, it might happen during the execution of \extractStar{} that%
+While all $G_t$ are connected, it might happen during the execution of \extractStar{} that%
 \begin{marginfigure}

### 2017-06-30 gtx_algorithm.tex
more text on the singleton problem

@@
 In the case no singleton are formed during one execution of \extractStar{}, the number of nodes is
-reduced by at least $2$ (because each star is made of at least two nodes). On the other hand, we can
+reduced by at least a factor $2$ (because each star is made of at least two nodes). On the other hand, we can
 construct a graph with singletons where the factor of reduction in number of nodes and edges can be
-arbitrarily close to $1$. Consider \autoref{fig:gtx_noreduc} and assume that we first extract two stars
-centered in $c_1$ and $c_2$.%
+arbitrarily close to $1$. Consider \autoref{fig:gtx_noreduc}, where $c_1$ and $c_2$ have a degree
+$p$ and their neighbors (in light gray) all have a degree $q$, which we can vary from $1$ to $p-1$.
+At this step $t$, we thus have $|V_t| = 2(1+p)+pq$ and $|E_t| = 2(p+qp)$. Next we run \extractStar{}
+and assume that we first extract the two stars centered in $c_1$ and $c_2$. This reduces the
+effective degree of all the middle nodes (in dark gray) to $0$ and they thus form singletons, that
+are all connected to the first two stars by \collapseStar{}. As a result, at step $t+1$ we have
+$|V_{t+1}| = 2 + pq$ and $|E_{t+1}| = 2pq$, so that
+\begin{equation*}
+\frac{|V_{t+1}|}{|V_t|} = \frac{pq+2}{p(q+2) + 2} \qquad \text{and} \qquad
+\frac{|E_{t+1}|}{|E_t|} = \frac{2pq}{2p(q+1)} = \frac{q}{q+1}
+\end{equation*}
+When we let $p$ goes to infinity, setting $q=1$ makes $\frac{|V_{t+1}|}{|V_t|}$ tend to
+$\nicefrac{1}{3}$ and $\frac{|E_{t+1}|}{|E_t|}$ to $\nicefrac{1}{2}$ while setting $q=p-1$ makes
+both $\frac{|V_{t+1}|}{|V_t|}$ and $\frac{|E_{t+1}|}{|E_t|}$ tend to $1$. Note however that in both
+cases, this is not really a problem because at the next iteration, there will be only two stars (one
+of them will be a singleton). Is it always like that? In other words, can we have a situation where
+a large number of singletons is carried over several consecutive iterations?
 

### 2017-06-30 gtx_algorithm.tex
mention a spanner variant of galaxy tree

@@
 corresponding leaves to the root of the tree.
+
+A variant of the \gtx{} algorithm as a whole we did not explore so much in practice is its ability
+to produce spanner by stopping early. Basically if we stop at iteration $t$, we output the graph
+$G_t$ (which in general is not a tree) with its edges unfolded to lie in $E_0$. This corresponds to
+a trade off between having more edges than $|V_0|-1$ but potentially making shorter connection and
+thus having a lower stretch.

### 2017-07-03 gtx_algorithm.tex
show that galaxy tree builds a spanning tree

@@
  Then we analyze the time complexity. We already now that during the \tth{} iteration,
- \extractStar{} and \collapseStar{} take time $O(|E_t|)$. We provide only the actual python code
- instead of pseudo code for the remaining functions since they do not present any special
- interesting algorithmic aspect. However, we can say that \textsc{Update-Eccentricity} and
+ \extractStar{} and \collapseStar{} take time $O(|E_t|)$. We only provide the actual python code
+ instead of pseudo code for the remaining functions since they do not present any relevant
+ algorithmic aspect. However, we can say that \textsc{Update-Eccentricity} and\marginpars{We can get
+ rid of this $O(|V_0|)$ at each iteration by ignoring eccentricity and taking a random edge between
+ star, although it's difficult to tell how it would affect the quality of the resulting tree.}
  \textsc{Update-Nodes-Mapping} take $O(|V_0|)$ time since they go through every node of the original
@@
  we have that the \gtx{} inner loop takes $O(K|E_0|)$ time. Finally, since
- \textsc{assemble-spanning-tree} goes again through every edge visited at every collapse step, it
+ \textsc{Assemble-Spanning-Tree} goes again through every edge visited at every collapse step, it
  also takes $O(K|E_0|)$ time, which is thus the overall complexity of the \gtx{} algorithm.
 
-  Finally we prove that \gtx{} indeed returns a spanning tree of $G_0$.
+  Finally we prove that \gtx{} indeed returns a spanning tree of $G_0$. Specifically, we show that
+  by starting from $G_K$ and going in reverse order, we build a spanning tree $T_t$ of every $G_t$,
+  including eventually $G_0$. This process is illustrated on \autoref{fig:gtx_spanning_tree} and
+  works as follow: when moving from $G_{t+1}$ to $G_t$, we expand the nodes of $G_{t+1}$ into star
+  of nodes of $V_t$ and update all the outer edges we met so far to have their endpoints in $V_t$.
+  These outer edges are exactly $T_{t+1}$ and by translating them into $E_t$ and adding the inner
+  star edges of $G_t$, we build $T_t$. $G_K$ consists of a single node with no edges so we have
+  trivially that $T_K$ = $G_K$. Then, assume we have a spanning tree $T_{t+1}$ of $G_{t+1}$, so that
+  all nodes in $V_{t+1}$ are connected without cycle. Expanding each node of $V_{t+1}$ into a star
+  of $V_t$ nodes ensures that the spanning property is preserved since each node of $V_t$ is covered
+  by construction. Those stars are also without cycle by construction. The second step, translating
+  the existing edges of $T_{t+1}$, only connects nodes belonging to different stars and because
+  those edges were cycle free in $G_{t+1}$, this remains the case in $G_t$. Therefore we have build
+  a spanning tree $T_t$ of $G_t$. By repeating this procedure, the \gtx{} algorithm eventually
+  builds a spanning tree of $G_0$.
+
 \end{proof}
+\begin{figure*}[h]
+  \centering
+	\begin{subfigure}[b]{37mm}
+    \centering
+		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g3_tikz.pdf}
+		\caption{Spanning tree of $G_3$}\label{fig:gtx_spanning_tree_3}
+	\end{subfigure}~
+	\begin{subfigure}[b]{45mm}
+    \centering
+		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g2_tikz.pdf}
+		\caption{Spanning tree of $G_2$}\label{fig:gtx_spanning_tree_2}
+	\end{subfigure}~
+	\begin{subfigure}[b]{49mm}
+    \centering
+		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g1_tikz.pdf}
+		\caption{Spanning tree of $G_1$}\label{fig:gtx_spanning_tree_1}
+	\end{subfigure}~
+	\begin{subfigure}[b]{53mm}
+    \centering
+		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g0_tikz.pdf}
+		\caption{Spanning tree of $G_0$}\label{fig:gtx_spanning_tree_0}
+	\end{subfigure}~
+  \caption{Unfolding stars to recover spanning trees}
+  \label{fig:gtx_spanning_tree}
+\end{figure*}
 
-Later we hope to found an upper bound of $T$ in terms of $m$, and also to refine the analysis to
-leverage the fact that $\sum_{t=0}^T |E_t|$ is significantly smaller than $Tm$. For now, let us
-prove the termination of \gtx{} by observing that, assuming $|E_t|>0$, $|E_{t+1}| < |E_t|$ since
-at least two nodes in $G_t$ are connected and will therefore form a star, which will remove all the
-inner edges of that star. As for the fact that we get a tree, note that a star is a tree, therefore
-a star of star is a tree, and so on.\Todo{make the two arguments about \gtx{} correctness more
-formal, and also explain that since all nodes are part of the ultimate star, this tree is a spanning
-one. Motivate the name of \gtx{}.}
+The bound of \autoref{prop:gtx_correct} amounts to $O(|V_0||E_0|)$, which we believe is overly
+pessimistic. First because we expect the number of nodes to decrease by more than one at each step,
+maybe leading to only $O(\log |V_0|)$ iterations. More importantly, instead of bounding
+$\sum_{t=0}^K |E_t|$ by $K|E_0|$, we hope to show that $E_{t+1} = rE_t$ with $r<1$ so that this sum
+is $O(|E_0|)$.
 

### 2017-07-03 assets/tikz/gtx_singleton_tikz.tex
minor wording and layout adjustments

@@
 
-    \draw[->] (0,-2) -- (0,-2.9);
+    \draw[edge,->] (0,-1) -- (0,-2.4);
 
-    \node[vertex,fill=DodgerBlue,text=black] (s1) at (-2,-3.7) {$S_1^1$};
-    \node[vertex,fill=DodgerBlue,text=black] (s2) at (2,-3.7) {$S_2^1$};
-    \node[vertex,fill=DodgerBlue,text=black] (s3) at (0,-3.7) {$S_3^1$};
+    \node[vertex,fill=DodgerBlue,text=black] (s1) at (-2,-3) {$S_1^1$};
+    \node[vertex,fill=DodgerBlue,text=black] (s2) at (2,-3) {$S_2^1$};
+    \node[vertex,fill=DodgerBlue,text=black] (s3) at (0,-3) {$S_3^1$};
     \foreach \source/\dest in {3/2, 1/3}

### 2017-07-03 gtx_algorithm.tex
minor wording and layout adjustments

@@
   builds a spanning tree of $G_0$.
-
 \end{proof}
+\vspace{-2\baselineskip}
 \begin{figure*}[h]
@@
 of \gtx{} is finished and we can go through every edges we chose between stars at every level to
-recover the final spanning tree, showed with solid edges in \autoref{fig:gtx_eccentricity}
-\vpageref{fig:gtx_eccentricity}. For completeness, we can also look at the edges which are not part
+recover the final spanning tree, a process illustrated in \autoref{fig:gtx_spanning_tree}
+\vpageref{fig:gtx_spanning_tree}. For completeness, we can also look at the edges which are not part
 of the spanning tree and therefore contribute to the stretch of the tree. In that case the average
@@
 A crucial quantity of the \gtx{} algorithm, both in terms of complexity and resulting stretch, is
-the number of collapses $T$ needed before termination. While this is still elusive to express in the
+the number of collapses $K$ needed before termination. While this is still elusive to express in the
 general case, let us first look at some simple cases. For instance, a very sparse example of graph
 is the line graph. While it is already a tree, let us look how the \gtx{} algorithm operates on it.
-Say we have $n$ nodes and $m$ edges in that line. There a star is made at most of three
+Say we have $n$ nodes and $m$ edges in that line. Here, a star is made at most of three
 consecutive nodes. In the worst case, centers will be chosen such that we have a succession of three
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{assets/tmp-code/line_graph.pdf}
-  \caption{A line graph with stars in red}
+  \includegraphics[width=0.9\linewidth]{assets/tikz/gtx_line_tikz.pdf}
+  \caption{A line graph with stars in blue}
   \label{fig:gtx_line_graph}
@@
 case $n=m+1$, \gtx{} will finish after $O(\log m)$ iterations. Note that a barbell graph (two
-cliques connected by a line) would requires a number of iterations proportional to the length of that
+cliques connected by a line) would requires a number of iterations logarithmic in the length of that
 central line, despite having many more edges. This suggests unsurprisingly that the diameter of the graph
@@
 
-While all $G_t$ are connected, it might happen during the execution of \extractStar{} that%
+While all $G_t$ are connected according to \autoref{lem:gtx_stay_connected}, it might happen during
+the execution of \extractStar{} that%
 \begin{marginfigure}
@@
 \autoref{fig:gtx_singleton}, where after the creation of \starone{1} and \starone{2}, $c_3$ is the
-single node of the third star \starone{3}.%
-\begin{marginfigure}
-  \centering
-  \includegraphics[width=0.95\linewidth]{assets/tikz/gtx_noreduc_tikz.pdf}
-  \caption{A case were too many singletons \enquote{waste} one iteration of \gtx{}}
-  \label{fig:gtx_noreduc}
-\end{marginfigure}
+single node of the third star \starone{3}.
 In the case no singleton are formed during one execution of \extractStar{}, the number of nodes is
@@
 effective degree of all the middle nodes (in dark gray) to $0$ and they thus form singletons, that
-are all connected to the first two stars by \collapseStar{}. As a result, at step $t+1$ we have
+are all connected to the first two stars by \collapseStar{}. As a result, at step $t+1$ we have%
+\begin{marginfigure}
+  \centering
+  \includegraphics[width=0.95\linewidth]{assets/tikz/gtx_noreduc_tikz.pdf}
+  \caption{A case were too many singletons \enquote{waste} one iteration of \gtx{}}
+  \label{fig:gtx_noreduc}
+\end{marginfigure}
 $|V_{t+1}| = 2 + pq$ and $|E_{t+1}| = 2pq$, so that

### 2017-07-04 gtx_related.tex
start writing low stretch related work

Looking for subgraphs $H$ of $G$ that best preserve the distance in $G$ while being sparse is an old
problem, especially driven by network design in fields such as transportation and electrical
circuit. For instance, \textcite{Johnson1978} define the following problem\footnote{We adapt their
notations to match ours}:  
\begin{problem}[Network Design Problem]
  \label{prob:gtx_ndp}
  Given an undirected integer-weighted graph $G=(V, E, w)$, a budget $B\in\Nbb$ and a criterion
  threshold $C\in \Nbb$, does there exist a spanning subgraph $G'=(V, E')$ of $G$ with weight
  $w(E') \leq B$ and criterion value $F(G') \leq C$, where the criterion function $F(G')$ denotes
  the sum of the weights of the shortest paths in $G'$ between all vertex pairs?
\end{problem}
Note that their definition of stretch includes all possible edges in $G$ and not only those in $E$.
They prove that this problem is \NPc{} by exhibiting a reduction from the \textsc{Knapsack}
problem. Finally they prove that the simpler problem of finding a spanning tree on an unweighted
graph, that is
\begin{problem}[Simple Network Design Problem]
  \autoref{prob:gtx_ndp} with $w$ being the equal to $1$ for all edges in $E$ and $B=|V|-1$.
\end{problem}
is also \NPc{} by reduction from \textsc{Exact 3-Cover}.

Earlier but slightly less related, \textcite{OptimalNetwork69} studies the following problem:
\begin{problem}[Optimal Network Problem]
  \label{prob:gtx_scott}
  Given a set $V$ of $n$ vertices, find a set of spanning edges $E\subset V^2$ that minimizes
  the sum of the length of the shortest paths  between all vertex pairs while the
  total length of the resulting network does not exceed some upper bound $B\in\Nbb$.
\end{problem}
This can be seen as a special case of \autoref{prob:gtx_ndp} with $G$ being the
unweighted\marginpars{according to the experiments, it's not clear whether there are weights or
not…} $n$-complete graph. He proposes a backtracking solution and two local search approximate
algorithms.

% Quandt R. E. (1960). Models of transportation and optimal network construction.  J. reg. Sri. 2,27-45.
% KNIGHT U. G. W. (1960a). The logical design of electrical networks using linear programming methods.  Proc. Instn elect.  Engrs A107, 306-319.
% KNIGHT U. G. W. (1960b).  Logical design of electrical networks. J. Instn elect. Engrs 6, 228-230.
% Then look back at Alon for Pelleg solution in the late 80's

### 2017-07-05 gtx_algorithm.tex
add references to Alon '95 paper
first version of the proof of logarithmic iteration of gtx

@@
 this stage then form the spanning tree we were looking for.
+This can be seen as a simple instantiation of a general procedure described in \autocite[Section
+5.2]{LowerBound95}.
 % [from a CMU NIPS submission http://www.stat.cmu.edu/~arinaldo/papers/mutualfriends.pdf] We
@@
 single node of the third star \starone{3}.
+
+Such singleton stars are problematic because when too many are created, we can end up having
+$|V_{t+1}|$ almost as large as $|V_t|$. Consider for instance the graph $G_t$ at the top of
+\autoref{fig:gtx_manysingles} (which is actually a tree). It has of order $k^3$ nodes and because
+all nodes at level 2 will be selected as center of their star, the graph $G_{t+1}$ (showed at the
+bottom of \autoref{fig:gtx_manysingles}) also have around $k^3$ nodes, meaning that
+$\frac{|V_{t+1}|}{|V_t|}$ tends to $1$ as $k$ goes to infinity. This is because
+$k\left(k(k+1)\right)$ singleton stars were created. Let us denote the fraction of singleton stars
+over the total number of nodes by $\alpha$. In that case,
+$$\alpha = \frac{k\left(k(k+1)\right)}{1+k+k(k+1)+k^2(k+1)}
+= 1 - \frac{1+2k+k^2}{1+2k+2k^2+k^3}
+= 1 - \frac{1+k}{1+k+k^2}$$
+We will now see that when so many singleton stars are created at one iteration, the next iteration
+greatly reduces the number of nodes so that over the course of two iterations, we have the following
+result:
+
+\begin{lemma}
+  \label{lemma:gtx_lessnodes}
+  During the execution of the \gtx{} algorithm, for all $t \leq K-2$, $|V_{t+2}| \leq
+  \frac{2}{3}|V_t|$.
+\end{lemma}
+\begin{marginfigure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/raw/lots_of_singletons_bw.png}
+  \caption{A case were too many singletons \enquote{waste} one iteration of \gtx{}}
+  \label{fig:gtx_manysingles}
+\end{marginfigure}
+\begin{proof}
+  Say that at iteration $t$ of \gtx{}, we have $n_t$ nodes and assume a fraction $\alpha$ of them
+  are gonna form singleton stars. This means there will be at most $\frac{1-\alpha}{2}n_t$
+  \emph{proper} stars, because such stars are made of at least two nodes. Now fix a constant
+  $\lambda \in (0,1)$, for instance $\lambda = \nicefrac{9}{10}$. If $\alpha \leq \lambda$, $n_{t+1}
+  \leq (\frac{1-\alpha}{2} + \alpha)n_t = \frac{1+\alpha}{2}n_t = \frac{19}{20}n_t$ and we get a
+  constant rate reduction of the number of nodes. On the other hand, if $\alpha > \lambda$, we could
+  be in the situation of the previous example and have $\alpha$ tends to $1$. Then we only have a
+  few proper stars and by construction, each singleton star is connected to least one proper star.
+  In the worst case, those connections are uniform so that each of these proper star has a degree
+  $\frac{\alpha n_t}{\frac{1-\alpha}{2}n_t} = \frac{2\alpha}{1-\alpha}$ which goes to infinity when
+  $\alpha$ goes to $1$ and is greater than $18$ when $\alpha$ is greater than $\nicefrac{9}{10}$.
+  Even ignoring the fact that those proper stars are connected, it follows that in the next
+  iteration, the number of node will be reduced by the degree of those proper stars. Setting in the
+  worst case $n_{t+1}=\frac{1+\alpha}{2}n_t$, we thus have $$n_{t+2} \leq
+  \frac{n_{t+1}}{\frac{2\alpha}{1-\alpha}} = \frac{1-\alpha^2}{4\alpha}n_t$$
+  \begin{marginfigure}
+    \centering
+    \includegraphics[width=\linewidth]{assets/tmp-code/gtx_reduc.pdf}
+    \caption{The fraction of $n_t$ nodes remaining in $n_{t+2}$ for the three cases described in the
+    main text.}
+    % \label{fig:gtx_manysingles}
+  \end{marginfigure}
+  This defines three ways to go from $n_t$ to $n_{t+2}$: if the first iteration has $\alpha>\lambda$,
+  then we just seen that in the worst case we have at least $n_{t+2}\leq
+  \frac{1-\lambda^2}{4\lambda}n_t$; if the first iteration has $\alpha < \lambda$, then either the
+  second also has  $\alpha < \lambda$ (in which case $n_{t+2}\leq
+  \left(\frac{1+\lambda}{2}\right)^2n_t$) or the second has $\alpha>\lambda$ (in which case
+  $n_{t+2}\leq \frac{1+\lambda}{2}n_t$). We can optimize over $\lambda$ by finding its value such
+  that $\frac{1-\lambda^2}{4\lambda}=\frac{1+\lambda}{2}$, yielding $\lambda = \frac{1}{3}$. Then,
+  that in any event, $n_{t+2}\leq \frac{2}{3}n_t$ as claimed.
+\end{proof}
+
+From that we conclude that the number of iterations of the \gtx{} algorithm satisfies $K=O(\log
+|V_0|)$. We would like to prove a similar result on the edges so that the overall complexity of the
+\gtx{} algorithm would be $O(|V_0|\log |V_0|+|E_0|)$. However, this does not follow directly from the
+reduction of number of nodes. Indeed we can have a situation where most of the edges of $E_t$ are
+between the stars of $G_t$ and thus will remain in the next iteration. Still, we hope that
+similarly to the previous proof, this high density of outer edges will lead to strong collapse in
+the next iteration.
+
+\iffalse
 In the case no singleton are formed during one execution of \extractStar{}, the number of nodes is
@@
 a large number of singletons is carried over several consecutive iterations?
+\fi
 

### 2017-07-06 gtx_algorithm.tex
typos

@@
   Say that at iteration $t$ of \gtx{}, we have $n_t$ nodes and assume a fraction $\alpha$ of them
-  are gonna form singleton stars. This means there will be at most $\frac{1-\alpha}{2}n_t$
+  will form singleton stars during the \tth{} iteration. This means there will be at most $\frac{1-\alpha}{2}n_t$
   \emph{proper} stars, because such stars are made of at least two nodes. Now fix a constant
   $\lambda \in (0,1)$, for instance $\lambda = \nicefrac{9}{10}$. If $\alpha \leq \lambda$, $n_{t+1}
-  \leq (\frac{1-\alpha}{2} + \alpha)n_t = \frac{1+\alpha}{2}n_t = \frac{19}{20}n_t$ and we get a
+  \leq (\frac{1-\alpha}{2} + \alpha)n_t = \frac{1+\alpha}{2}n_t$ and $\frac{1+\alpha}{2}n_t \leq
+  \frac{1+\lambda}{2}n_t = \frac{19}{20}n_t$ so we get a
   constant rate reduction of the number of nodes. On the other hand, if $\alpha > \lambda$, we could
-  be in the situation of the previous example and have $\alpha$ tends to $1$. Then we only have a
-  few proper stars and by construction, each singleton star is connected to least one proper star.
-  In the worst case, those connections are uniform so that each of these proper star has a degree
+  be in the situation of the previous example and have $\alpha$ be arbitrarily close to $1$. Then we
+  only have a few proper stars and by construction, each singleton star is connected to least one proper star.
+  In the worst case, those connections are uniform so that each of these proper stars has degree at least
   $\frac{\alpha n_t}{\frac{1-\alpha}{2}n_t} = \frac{2\alpha}{1-\alpha}$ which goes to infinity when
@@
   that $\frac{1-\lambda^2}{4\lambda}=\frac{1+\lambda}{2}$, yielding $\lambda = \frac{1}{3}$. Then,
-  that in any event, $n_{t+2}\leq \frac{2}{3}n_t$ as claimed.
+  in any event, $n_{t+2}\leq \frac{2}{3}n_t$ as claimed.
 \end{proof}
@@
 \gtx{} algorithm would be $O(|V_0|\log |V_0|+|E_0|)$. However, this does not follow directly from the
-reduction of number of nodes. Indeed we can have a situation where most of the edges of $E_t$ are
+reduction of number of nodes. Indeed, we can have a situation where most of the edges of $E_t$ are
 between the stars of $G_t$ and thus will remain in the next iteration. Still, we hope that

### 2017-07-06 gtx_related.tex
few more related works

@@
-Looking for subgraphs $H$ of $G$ that best preserve the distance in $G$ while being sparse is an old
-problem, especially driven by network design in fields such as transportation and electrical
-circuit. For instance, \textcite{Johnson1978} define the following problem\footnote{We adapt their
-notations to match ours}:  
+Looking for a subgraph $H$ of $G$ that best preserve the distance in $G$ while being sparse is an old
+problem, especially driven by network design in fields such as transportation~\autocite{RoadNetworks60}
+and electrical circuits~\autocite{electricalNetworks60}. For instance, \textcite{Johnson1978} define
+the following problem\footnote{We adapt their notations to match ours}:  
 \begin{problem}[Network Design Problem]
@@
 graph, that is
+\vspace{-.5\baselineskip}
 \begin{problem}[Simple Network Design Problem]
   \autoref{prob:gtx_ndp} with $w$ being the equal to $1$ for all edges in $E$ and $B=|V|-1$.
+\end{problem}%
+\vspace{-.5\baselineskip}
+\noindent is also \NPc{} by reduction from \textsc{Exact 3-Cover}.
+
+\begin{problem}[Optimal Communication Spanning Tree~\autocite{Requirements74}]
+Given a set of nodes $V=\{v_1, \ldots, v_n\}$, a set of distances $d_{ij}$ and a set of requirements
+$r_{ij}$ between $v_i$ and $v_j$, find a spanning tree connecting these $n$ nodes such that the
+total cost of communication of the spanning tree is a minimum among all spanning trees. The cost of
+communication for a pair of nodes is $r_{i,j}$ multiplied by the sum of the distances of arcs which
+form the unique path connecting $v_i$ and $v_j$ in the spanning tree. The cost of a spanning tree is
+the sum of costs over all pairs of nodes.
 \end{problem}
-is also \NPc{} by reduction from \textsc{Exact 3-Cover}.
+
+Given a weighted graph $G=(V,E,w)$, if we let $d_{ij}$ be $w((i,j))$ and $r_{ij}$ be $1$ if
+$(i,j)\in E$ and $0$ otherwise, it seems to me this correspond to the weighted low stretch problem.
+However, \textcite[Section 4]{lognMetricBoundConf03} claim a $O(\log n)$ approximation so maybe I'm
+wrong.
+
+This defines two kind of structures, spanning trees and spanners (which are still sparse subgraphs
+yet containing more than $|V|-1$ edges).
+
+\paragraph{Trees}
+\label{par:trees}
 
@@
 \end{problem}
-This can be seen as a special case of \autoref{prob:gtx_ndp} with $G$ being the
-unweighted\marginpars{according to the experiments, it's not clear whether there are weights or
+This can be seen as a special case of \autoref{prob:gtx_ndp} with $G$ being the unweighted
+\marginpars{according to the experiments, it's not clear whether there are weights or
 not…} $n$-complete graph. He proposes a backtracking solution and two local search approximate
-algorithms.
+algorithms. Some early branch and bound heuristic solutions to \autoref{prob:gtx_scott} are surveyed
+in~\autoref[Section 2.3.2]{networkDesignSurvey89} although they do not come with asymptotic
+guarantee on the stretch. Furthermore, \textcite{optimApproxNP80} proves that for any $\epsilon \in
+(0,1)$, finding a $|V|^{1-\epsilon}$ approximation is \NPc{}.
 
-% Quandt R. E. (1960). Models of transportation and optimal network construction.  J. reg. Sri. 2,27-45.
-% KNIGHT U. G. W. (1960a). The logical design of electrical networks using linear programming methods.  Proc. Instn elect.  Engrs A107, 306-319.
-% KNIGHT U. G. W. (1960b).  Logical design of electrical networks. J. Instn elect. Engrs 6, 228-230.
-% Then look back at Alon for Pelleg solution in the late 80's
+\paragraph{Spanners}
+\label{par:spanners}
+
+% check some surveys of the 80's
+% http://pubsonline.informs.org/doi/abs/10.1287/trsc.18.1.1
+% http://onlinelibrary.wiley.com/doi/10.1002/net.3230190305/full
+% early solutions
+% http://onlinelibrary.wiley.com/doi/10.1002/net.3230090104/full
+% http://onlinelibrary.wiley.com/doi/10.1002/net.3230130309/full
+% later solution?
+% http://ieeexplore.ieee.org/document/81738

### 2017-07-07 gtx_related.tex
typos

@@
 $(i,j)\in E$ and $0$ otherwise, it seems to me this correspond to the weighted low stretch problem.
-However, \textcite[Section 4]{lognMetricBoundConf03} claim a $O(\log n)$ approximation so maybe I'm
+However, \textcite[Section 3, page 453]{lognMetricBoundConf03} claim a $O(\log n)$ approximation so maybe I'm
 wrong.
@@
 algorithms. Some early branch and bound heuristic solutions to \autoref{prob:gtx_scott} are surveyed
-in~\autoref[Section 2.3.2]{networkDesignSurvey89} although they do not come with asymptotic
+in~\autocite[Section 2.3.2]{networkDesignSurvey89} although they do not come with asymptotic
 guarantee on the stretch. Furthermore, \textcite{optimApproxNP80} proves that for any $\epsilon \in

### 2017-07-07 lowstretch_and_cc.tex
start writing low stretch related work
typos

@@
 \begin{enumerate}
-	\item Our bias remains that the labeling is regular (recall this means informally than the all
+	\item Our bias remains that the labeling is regular (recall this means informally that the all
 		the outgoing signs from a given node tend to be the same, and likewise for the incoming signs),
@@
 
-\subsection{Problem setting}
+\subsection{Problem setting and applications}
 \label{sub:problem_setting}
@@
 
-\subsection{State of the art}
-\label{sub:gtx_state_of_the_art}
+\subsection{Related works}
+\label{sub:gtx_related_works}
+
+\input{gtx_related}
 

### 2017-07-08 gtx_algorithm.tex
typos
found a mistake in the node reduction proof :(

@@
     \Repeat
-      \Let{$stars$, $inner\_edges$, $membership$}{\Call{Extract-Stars}{$G_t$}}
+      \Let{$stars$, $inner\_edges$, $star\_membership$}{\Call{Extract-Stars}{$G_t$}}
       \State Add $inner\_edges$ to the list $inner\_edges\_seq$
       \State \Call{Update-Eccentricity}{$stars$, $eccentricity$, $original\_node$}
-      \Let{$full\_membership$, $original\_node$}{\Call{Update-Nodes-Mapping}{$full\_membership$, $membership$}}
+      \Let{$full\_membership$, $original\_node$}{\Call{Update-Nodes-Mapping}{$full\_membership$,
+      $star\_membership$}}
       \Let{$G_{t+1}$, $\!outer\_edges$}{\Call{Collapse-Stars}{$E_t \!\setminus\! inner\_edges$,
-      $\!membership$, $\!eccentricity$}}
+      $\!star\_membership$, $\!eccentricity$}}
       \State Add $outer\_edges$ to the list $outer\_edges\_seq$
@@
 result:
+{\color{blue}{\large the following is actually incorrect but I leave it for now for illustration
+purpose}
 
@@
   \label{lemma:gtx_lessnodes}
-  During the execution of the \gtx{} algorithm, for all $t \leq K-2$, $|V_{t+2}| \leq
+  During the execution of the \gtx{} algorithm, for all $0\leq t \leq K-2$, $|V_{t+2}| \leq
   \frac{2}{3}|V_t|$.
@@
   $\alpha$ goes to $1$ and is greater than $18$ when $\alpha$ is greater than $\nicefrac{9}{10}$.
-  Even ignoring the fact that those proper stars are connected, it follows that in the next
+  \textbf{Even ignoring the fact that those proper stars are connected}, it follows that in the next
   iteration, the number of node will be reduced by the degree of those proper stars. Setting in the
@@
 From that we conclude that the number of iterations of the \gtx{} algorithm satisfies $K=O(\log
-|V_0|)$. We would like to prove a similar result on the edges so that the overall complexity of the
+|V_0|)$.}
+The problem is that we cannot ignore the fact that proper stars are connected, since one of them can
+be connected to all others and have the highest degree. That proper star is then chosen as the first
+center, all other proper stars become its periphery and all their attached singletons star remain
+singleton stars at the next iteration. This actually what happens on \autoref{fig:gtx_manysingles},
+an example where $|V_{t+2}| \sim |V_t| \sim k^3$. This could even be worse if proper stars are
+connected in such a hierarchical way that singleton stars can survive for many iterations.
+
+We would like to prove a similar result on the edges so that the overall complexity of the
 \gtx{} algorithm would be $O(|V_0|\log |V_0|+|E_0|)$. However, this does not follow directly from the

### 2017-07-10 gtx_related.tex
add more related works on low stretch tree

@@
 \noindent is also \NPc{} by reduction from \textsc{Exact 3-Cover}.
+In this setting of a spanning tree minimizing the stretch over all pairs of nodes, \enquote{
+\textcite{AllPairStrech08} obtained a $(c log n)^{O(\log \Delta)}$ factor, where $c$ is the optimal
+maximum stretch and $\Delta$ is the diameter.  They also showed $O(1)$ approximation for the case
+where the graph is unweighted. The constant was later reduced to $6$ by \textcite{AllPairStrech10}}
 
@@
 $(i,j)\in E$ and $0$ otherwise, it seems to me this correspond to the weighted low stretch problem.
-However, \textcite[Section 3, page 453]{lognMetricBoundConf03} claim a $O(\log n)$ approximation so maybe I'm
-wrong.
+% However, \textcite[Section 3, page 453]{lognMetricBoundConf03} claim a $O(\log n)$ approximation so maybe I'm
+% wrong. Turns out, they refer to a distribution over trees and this $O(\log n)$ is the expected
+% stretch of a tree sampled from this distribution
 
@@
 
+Next is the seminal paper of \textcite{LowerBound95}. It touches on many topics, and frame the
+problem in a game theoretic way but here we only focus on two of their results: a lower bound of
+$\Omega(\log n)$ for the average stretch of any tree and their construction of a tree with $\exp
+O(\sqrt{\log n\log\log n})$ average stretch in time $O(m^2)$. The lower bound follows from an
+existing result in extremal graph theory~\autocite[pages 107--109]{ExtremalGraph04}: there is a
+positive constant $a$ such that for all $n\in \Nbb$, one can construct a graph $G$ with $n$ vertices
+and $2n$ edges such that every cycle $G$ has a length of at least $a\log n$. Now consider any
+spanning tree $T$ of $G$.  While all the $n-1$ edges of $T$ have a stretch of $1$, the $n+1$
+remaining ones form a cycle in $T$ hence in $G$ as well and thus incur a stretch of at least $a\log
+n$. This shows that the average stretch is at least $\frac{1}{2}a\log n$.
+
+They construct a low stretch spanning tree in a bottom up manner like the \gtx{} algorithm. First,
+they extend the definition of stretch to multigraph~\autocite[Section 4]{LowerBound95} and then
+describe a procedure to transform in linear time any multigraph $G$ with $n$ nodes to a multigraph
+$G'$ on the same nodeset with at most $n(n+1)$ edges such the average stretch of $G'$ is at most
+twice that of $G$~\autocite[Lemma 5.2]{LowerBound95}. The next ingredient is an algorithm to build a
+low diameter decomposition of a multigraph $G$, parametrized by a number $x(n)$ depending of $n$. It
+works by repeatedly selecting an arbitrary node and growing a ball around it until the number of
+edges leaving the ball is at most a fraction $\nicefrac{1}{x(n)}$ of the number of edges with both
+endpoints in the ball. The key property of this decomposition is that it yields a partition of $G$
+in clusters such that the radius of each cluster is small (namely at most $O(x(n)\log n)$) and there
+is most a fraction $\nicefrac{1}{x(n)}$ of edges between clusters. Finally, the iterative procedure
+is a follows: once a partition has been built, we compute a shortest path spanning tree in each
+cluster that are then collapsed into super nodes to form the next graph $G'$ and the process repeats.
+Another difference from \gtx{}, besides the partition procedure, is that $G'$ is a multigraph,
+taking into account the number of edges joining cluster, while \collapseStar{} picks only the most
+direct one.
+
+Another interesting idea from this paper is to consider a distribution over trees instead of a
+single instance, especially when one is concern about the maximum stretch instead of the average
+one. For instance, on a cycle with $n$ nodes, a tree is obtained by removing one edge, and that edge
+incurs a stretch of $n-1$. The uniform distribution over such trees has a maximum stretch of
+$2\left(1 - \frac{1}{n}\right)$~\autocite{circle2k89}.
+
+\begin{table}[htbp]
+  \centering
+  \caption{Reproduction of Table 1 from~\autocite{Abraham2012} showing the evolution of the best
+  asymptotic average stretch over time.}\label{tab:gtx_related}
+  \begin{tabular}{lll}
+    \toprule
+    work                      & average stretch                          & time                    \\
+    \midrule
+    \autocite{LowerBound95}   & $\exp(O(\sqrt{\log n\log\log n}))$       & $O(m^2)$                \\
+    \autocite{LowerStretch05} & $O((\log n)^2 \log \log n)$              & $O(m \log^2 n)$         \\
+    \autocite{nearlyTight08}  & $O(\log n(\log \log n)^3)$              & $O(m \log^2 n)$         \\
+    % \autocite{nearlyTight08}  & $O(\log n \log\log n(\log\log\log n)^3)$ & $O(m^2)$                \\
+    \autocite{TighterSDD11}   & $O(\log n(\log \log n)^3 )$              & $O(m \log n\log\log n)$ \\
+    \autocite{Abraham2012}    & $O(\log n \log \log n)$                  & $O(m \log n\log\log n)$ \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\autocite{LowerStretch05} is the star decomposition described in section 3 of \autocite{papplow},
+which was modified by \autocite{nearlyTight08} to improve the stretch. Then \textcite{TighterSDD11}
+improved the runtime by rounding the edge weights to the closest power of $2$ and using a modified
+implementation of the Dijkstra's algorithm in the case of at most $k$ distinct edge
+weights~\autocite{FastPathFewWeights10}. Finally, \textcite{Abraham2012} describe an even more
+complex but tighter petal decomposition.
+\begin{marginfigure}
+  \centering
+  \includegraphics[width=.95\textwidth]{assets/raw/star_decomp.pdf}
+  \caption{Star decomposition (reproduced from Figure 1 of~\autocite{LowerStretch05})}
+  \label{fig:gtx_star_decomp}
+\end{marginfigure}
+
+If the tree is non spanning (\ie{} it can include edges not from the original graph as long as the
+distances in tree remain larger the distances in the graph), the average stretch can be reduced to
+$\Theta(\log n)$~\autocite{lognMetricBoundConf03}. Furthermore, when the objective if to minimize
+the average stretch over all possible pairs of node, then \textcite{constantDistortion07} achieve a
+universal constant bound.
+
+Special case of graph
+series parallel
+\enquote{In a subsequent paper, \textcite{seriesParallel06} proved that every series-parallel
+unweighted graph admits a spanning tree of average stretch $O(log n)$. This bound is tight as it
+matches the lower bound established in~\autocite{cutsTrees99}.}
+
 \paragraph{Spanners}

### 2017-07-11 gtx_related.tex
related works on spanners

@@
 
+more special cases are in~\autocite{specialCase14}, although it's for the minimum max stretch
+$t^\star$.
+\enquote{Note also that a number of particular graph classes (like interval
+graphs, permutation graphs, asteroidal-triple–free graphs, strongly chordal graphs,
+dually chordal graphs, and others) admit tree $t$-spanners for small values of $t$}
+
+A tangential problem is \enquote{Light Approximate Shortest-path Trees (LASTs).  Given a set of
+“sources” and a sink, a LAST is a tree of weight close to the minimum spanning tree (MST) on the
+sources and the sink, such that the tree distance from any source to the sink is close to the
+shortest-path distance. Khuller, Raghavachari, and Young (KRY95) defined and studied LASTs in the
+offline setting and showed that one can get constant stretch with cost constant times the
+MST.}\footnote{FOCS'17 \href{https://arxiv.org/abs/1611.00052}{arxiv.org/abs/1611.00052}}
+
 \paragraph{Spanners}
 \label{par:spanners}
+
+the subgraph $H$ is said to be an $(\alpha, \beta)$-spanner of $G$ if, for a pair of parameters
+$\alpha \geq 1$, $\beta \geq 0$, and for every pair $u, v \in V$ of vertices, it holds that $d_H(u,
+v) \leq \alpha\cdot d_G(u, v)+\beta$.
+
+They have applications in computing approximately shortest
+paths [9, 22, 28, 37], routing [48], distance oracles and
+labeling schemes [49, 56, 36] and synchronization [7].
+
+From 6:They also appear in biology in the process of reconstructing phylogenetic trees from
+matrices, whose entries represent genetic distances among contemporary living species (H. J.
+Bandelt, A. W. M. Dress, Reconstructing the Shape of a Tree from Observed Dissimilarity Data, Adv.
+in Appl. Math. 7 (1986)). Robotics researchers have studied spanners under the constraints of
+Euclidean geometry, where vertices of the graph are points in space, and edges are line segments
+joining pairs of points (Chew), (Dobkin), $[DJ], [K], [KG], [LL]$.
+
+studied in 1; 4; 6; 9; 15; 19; 22; 24; 26; 28; 30; 31; 37; 43; 51; 52; 57; 58
+
+\begin{tabulary}{\textwidth}{LLLLL}
+  \toprule
+  work  & average stretch & edge size                              & weighted & time                                             \\
+  \midrule
+  46    & $4k + 1$        & $O(n^{1+\nicefrac{1}{k}})$             & no       & polynomial                                       \\
+  6     & $2k +1$         & $O(n\cdot \ceil{n^{\nicefrac{1}{k}}})$ & yes      & $O\left(m(n^{1+\nicefrac{1}{k}}+n\log n)\right)$ \\
+  40    & $2k-1$          & $O(n^{1+\nicefrac{1}{k}}+n)$           & no       & $O(m)$                                           \\
+  Elkin & $2k-1$          & $n \cdot (1 + O(\frac{\log k}{n}))$    & no       & $O(m)$                                           \\
+  \bottomrule
+\end{tabulary}
+
+% 22: E. Cohen, "Fast algorithms for constructing t-spanners and paths with stretch t," Proceedings
+% of 1993 IEEE 34th Annual Foundations of Computer Science, Palo Alto, CA, 1993, pp. 648-658.  doi:
+% 10.1109/SFCS.1993.366822
+We construct t-spanners of size (number of edges) Õ(n 1+(2+\epsilon)/t )
+(for any \epsilon > 0 and t such that t/(2+\epsilon) is integral). These spanners can be constructed
+by a randomized algorithm that runs in Õ(mn (2+\epsilon)/t ) time.
+
+% Halperin and Zwick [40].  Their deterministic algorithm, for an integer parame- ter k ≥ 1,
+% computes a (2k − 1)-spanner with n 1+1/k + n edges in O(|E|) time. (Their result improved previous
+% pioneering work by [46, 22].)
+
+Describe the greedy algorithm of 6. Using (55 On Dynamic Shortest Paths Problems Liam Roditty, Uri
+Zwick, 2004), it runs in $O(\alpha n^{2+\nicefrac{1}{\alpha}})$
+
+peleg 2007 hardness results
+
+46: Peleg, D. and Schäffer, A. A. (1989), Graph spanners. J. Graph Theory, 13: 99–116.
+doi:10.1002/jgt.3190130114
+
+study the problem on unweighted graph. Application to routing scheme (48: D. Peleg and E. Upfal, A
+tradeoff between space and efficiency for routing tables. 20th ACM Symposium on the Theory of
+Computing, Chicago (1988))
+Special case for the complete graph weighted by the distance in a 2D plan. Existing $\sqrt{10}$
+spanner for the $\ell_1$ metric (L. P. Chew, There is a planar graph almost as good as the complete
+graph.  Proceedings of the 2nd ACM Symposium on Computational Geometry, (1986)) (improved to
+$\sqrt{4+2\sqrt{2}}$ by N. Bonichon, C. Gavoille, N. Hanusse, L. Perkovic The stretch factor of
+$\ell_1$ and $\ell_{\infty}$ Delaunay triangulations European Symposium on Algorithms (ESA) (2012))
+and $\phi \pi$ for $\ell_2$ (D. P. Dobkin, S. J. Friedman, and K. J. Supowit, Delaunay graphs are
+almost as good as complete graphs. 28th IEEE Symposium on the Foundations of Computer Science,
+(1987)) (improved to $1.998$ by Ge Xia. 2011. Improved upper bound on the stretch factor of delaunay
+triangulations. In Proceedings of the twenty-seventh annual symposium on Computational geometry
+(SoCG '11)). See (Prosenjit Bose, Michiel Smid, On plane geometric spanners: A survey and open
+problems, Computational Geometry, Volume 46, Issue 7, 2013) for more on the 2D case.
+In undirected graph, finding a spanner with less than $k$ edges is \NPc{} (theorem 2.2)
+For $k<n$, one can construct in polynomial time a $(4\log_k n +1)$ spanner with less than $kn$ edges
+(theorem 2.4) giving for instance ($k=2$) $O(\log n)$ spanner with $O(n)$ edges and
+($k=n^{\nicefrac{1}{r}}$, $r\geq 1$) a $(4r+1)$ spanner with $O(n^{1+\nicefrac{1}{r}})$ edges
+(matching lower bound within constant factor). For every $d \geq 0$, the $d$-dimensional cube has a
+$3$-spanner with fewer than $7\time 2^d$ edges (Lemma 2.10 from 47: D. Peleg and J. D. Ullman. An
+optimal synchronizer for the hypercube. SIAM J. on Comput., 18:740–747, 1989)
+For chordal graphs: for every $n$-vertex chordal graph there exists a $2$-spanner with
+$O(n\sqrt{n})$ edges (matching lower bound), a $3$-spanner with $O(n \log n)$ edges and a $5$-spanner
+with $O(n)$ edges.
+Much more difficult for directed graph, according to Theorem 4.2: For every $t \geq 1$ there are
+infinitely many $n$-vertex directed graphs for which every $t$-spanner requires
+$\Omega(\nicefrac{n^2}{t^2})$ edges.
 

### 2017-07-21 cc_part.tex
move CC section to its own file

--- 2017-07-07 lowstretch_and_cc.tex
@@
-\documentclass[nobib, a4paper, 10pt, oneside, onecolumn, openany, notitlepage, final,
-svgnames, marginals=raggedouter, english,
-%symmetric, justified,
-]{article}
-\usepackage{charter}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\input{packages}
-\input{abbreviations}
-
-\title{Edge sign prediction in general graphs and connection with Correlation Clustering}
-% \usetikzlibrary{external}
-% \tikzexternalize[mode=list and make, failed ref warnings for={\ref,\cite,\pageref},]
-\begin{document}
-\newgeometry{vmargin=1.3cm,left=1.5cm,textwidth=130mm,marginparsep=5mm,marginparwidth=52mm}
-\maketitle
-% \tableofcontents
-
-% \input{notations}
-
-\section{Limitations of the troll method}
-\label{sec:limitations_of_the_troll_method}
-
-\subsection{Reliance on the sign generative model}
-\label{sub:reliance_on_the_sign_generative_model}
-
-The method presented in the previous chapter hinges crucially upon our sign generative model. Yet
-one can imagine contexts where this model is not applicable, especially when nodes do not represent
-human beings. One way to alleviate this issue was introduced with our online algorithm, where this
-time, signs are generated by an arbitrary adversary. However, in that case, we are still facing two
-limitations\marginpars{limitations is not the right word, maybe \emph{strong constraints}…}
-\begin{enumerate}
-	\item Our bias remains that the labeling is regular (recall this means informally that the all
-		the outgoing signs from a given node tend to be the same, and likewise for the incoming signs),
-		since irregularities are the cost payed by the adversary to make our algorithm mispredict. While
-		this bias is well suited to social networks, other applications may require other
-		bias\marginpars{namely the \pcc{} bias we'll introduce next}.
-	\item In the online setting, we evaluate our performance by the regret\marginpars{ref to regret equation},
-		whereas in general we are interested in more classical measures of predictive accuracy.
-\end{enumerate}
-
-\subsubsection{Experience on directed biological network}
-\label{ssub:experience_on_directed_biological_network}
-
-Take for instance a biological network.
-
-According to \href{https://web.stanford.edu/class/cs224w/slides/handout-bionets.pdf}%
-{this description of various biological networks}, \emph{gene regulatory network} are directed and
-contains activation and inhibition links, as \href{https://en.wikipedia.org/wiki/Gene_regulatory_network#Overview}%
-{showed on Wikipedia}. There are some online databases such as
-\href{http://regulondb.ccg.unam.mx/menu/download/datasets/index.jsp}%
-{RegulonDB} or \href{http://www.pathwaycommons.org/pcviz/}{Pathway Commons} (this last one provide
-network in the BioPAX format, which can be visualized by
-\href{http://www.cytoscape.org/}{CytoScape}).
-Another source is the \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2708159/table/T1}{Table 1}
-of \cite{BioSigned09}.
-%DasGupta also describe (in
-%http://www.sciencedirect.com/science/article/pii/S0303264706001419#sec14) how
-%they built a directed signed network from this SBML file
-%https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1681468/bin/msb4100014-sd1.xml
-
-other domain? coref, images (is it directed?), entity resolution…\marginpars{Actually it's difficult
-to find other kind of graph because all constructed ones (like image, coreference,
-deduplication) are inherently symmetric}
-
-\subsection{Need for a directed graph}
-\label{sub:need_for_a_directed_graph}
-
-A more serious limitation is that our sign model only applies to directed graphs. Consider
-another model, where each node $i$ is endowed with an integer $c_i \in \{1, \ldots, k\}$ that
-specifies its cluster and let the sign of the undirected edge $i,j$ be $+$ if $c_i = c_j$ and $-$
-otherwise.  This corresponds naturally to the situation modeled by the \pcc{} problem.
-
-\section{\pcc{}}
-\label{sec:correlation_clustering}
-
 \subsection{Problem setting and applications}
@@
 \label{sub:cc_empiracal_evaluation}
-
-\newpage
-
-\section{Low stretch trees and spanners}
-\label{sec:low_stretch_trees_and_spanners}
-
-\input{notations}
-
-Given a graph $G$, and following the social balance theory roughly summarized as \enquote{the enemy
-of my enemy is my friend}, assume that the labelling of $E$ is consistent with a two-clustering of
-$V$. Namely, assume that $V$ can be partitioned in two clusters such that edges within each cluster
-are positive and edges across clusters are negative. In that case, the following
-\emph{multiplicative rule} holds: for any nodes $u$, $v$ in $V$, and any path $p$ between $u$ and
-$v$ in $G$, the sign $\yuv$ is equal to the product of the signs along $p$ (we call this product the
-parity of $p$ and denote it $\pi(p)$). While it is a simple and convenient model, this is too strong
-of a requirement to be satisfied in practice. Therefore, we relax it by assuming that we start with
-a consistent labeling $Y$ but can only observe a randomly perturbed version $Y'$ of $Y$.
-Specifically, given a constant $q\in [0, \nicefrac{1}{2})$, each sign of $Y$ is flipped with a
-probability smaller than $q$.
-
-In this section, we are interested in active learning algorithms, that first query a subset
-$\etrain$ of the edges, observe the signs in $\etrain$ and use them to predict the remaining signs.
-More precisely, we focus on an algorithm that queries spanning tree $T$ of $G$ and predicts the sign
-of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. While this
-can\Todo{state it formally then ;)} be stated more formally~\autocite[Section
-4.1]{Cesa-Bianchi2012b}, it makes sense intuitively that since each sign has been potentially
-flipped, the longer the path in $T$, the more likely its parity will be not be equal to the true
-sign $\yuv{}$. Therefore we would like each such path to be as short as possible. In the following,
-we describe a way to build spanning trees tailored for this situation.
-
-\subsection{\gtx{}: a spanning tree designed for sign prediction}
-\label{sub:gtx_a_simple_low_stretch_tree_construction}
-
-\input{gtx_algorithm}
-
-\subsection{Related works}
-\label{sub:gtx_related_works}
-
-\input{gtx_related}
-
-\iffalse
-\url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
-for instance the Elkin paper \enquote{Our centralized randomized algorithm computes (with
-probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in
-$O(|E|)$ time, whenever $k = \Omega(\log n)$. Note that when $k = \omega(\log n)$, the number of
-edges is $n(1+o(1))$, i.e., in this range the algorithm computes an ultra-sparse spanner in $O(|E|)$
-time.} For instance, if $k=5\log n$, we get a $10\log n$-spanner with $n\left(1+O\left(\frac{\log\log
-n}{n}\right)\right)$ edges in $O(|E|)$ time.
-
-While they have many applications [see first paragraph of \url{https://arxiv.org/pdf/1401.2454.pdf},
-which was later merged in a STOC'14 paper] (a major one being solving linear system), in some
-practical situations their advantages are less clear [from
-\url{https://link.springer.com/chapter/10.1007/978-3-319-20086-6_16}\enquote{for reasonable inputs
-the constant factors make the solver much slower than methods with higher asymptotic complexity.
-One other aspect predicted by theory is confirmed by our findings: Spanning trees with lower
-stretch indeed reduce the solver's running time. Yet, simple spanning tree algorithms perform
-better in practice than those with a guaranteed low stretch.} this is improved by
-\url{https://link.springer.com/chapter/10.1007%2F978-3-319-20086-6_17} although they seem to work
-	mostly with the Laplacian of the tree ]
-\fi
-
-\subsection{Empirical evaluation}
-\label{sub:gtx_empirical_evaluation}
-
-% \input{gtx_experiments}
-
-\section{Conclusion}
-
-\begingroup
-\small
-\setstretch{0.9}
-\todos
-
-\setlength\bibitemsep{2pt}
-\printbibliography
-\endgroup
-
-% \bibliography{/home/orphee/data/projects/biblio/library.bib,more.bib}
-% \bibliographystyle{plainnat}
-\end{document}

### 2017-07-21 lowstretch_and_cc.tex
move CC section to its own file

@@
 deduplication) are inherently symmetric}
+\fi
 
@@
 
+\iffalse
 A more serious limitation is that our sign model only applies to directed graphs. Consider
@@
 otherwise.  This corresponds naturally to the situation modeled by the \pcc{} problem.
+\fi
 
@@
 
-\subsection{Problem setting and applications}
-\label{sub:problem_setting}
-
-\subsection{Relation with edge sign prediction}
-\label{sub:relation_with_edge_sign_prediction}
-
-\iffalse
-For the 2 clusters case ($k=2$), characterization proven already in 1936 by \textcite{Konig36}, as noted in
-\cite{Zaslavsky2012} (commenting on \cite{harary1953}: \enquote{Although Thm. 3 was anticipated by
-\textcite[Theorem X.11]{Konig36}  without the terminology of signs, here is the
-first recognition of the crucial fact that labelling edges by elements of a
-group—specifically, the sign group—can lead to a general theory.})
-and \cite{Huffner2010} (\enquote{\textcite{Konig36} proved the following characterization of
-	balanced graphs. For a graph $G = (V , E)$, the following are equivalent:\marginpars{There is a
-	proof in \autocite[p. 111]{BookKleinberg2010}, maybe I can rewrite it as well}
-	\begin{enumerate}
-		\item $V$ can be partitioned into two sets $V_1$ and $V_2$ called sides such that there is no
-			negative edge $\{v, w\} \in E$ with both $v, w \in V_1$ or both $v, w \in V_2$ and no negative
-			edge $\{v, w\}$ with $v \in V_1$ and $w \in V_2$ .
-		\item $V$ can be colored with two colors such that for all $\{v, w\} \in E^-$, the vertices $v$
-			and $w$ have different colors, and for all $\{v, w\} \in E^+$, the vertices $v$ and $w$ have
-			the same color. The color classes correspond to the sides.
-		\item $G$ does not contain cycles with an odd number of negative edges.
-	\end{enumerate}
-	Using the characterization by a coloring, it is easy to see that balance of a signed
-graph can be checked in linear time by depth-first search.})
-\fi
-
-\subsection{State of the art}
-\label{sub:state_of_the_art}
-
-% ICML paper this year that touch something very related multicut and give recent applications in
-% vision https://arxiv.org/abs/1503.03791
-
-\subsection{Variants and extensions}
-\label{sub:variants_and_extensions}
-
-\subsubsection{\pcc{} under stability assumption}
-\label{ssub:cc_under_stability_assumption}
-
-\iffalse
-Haris Angelidakis, Konstantin Makarychev, and Yury Makarychev. 2017.
-Algorithms for Stable and Perturbation-Resilient Problems. STOC’17
-\href{http://ttic.uchicago.edu/~yury/papers/two-stable.pdf}{10.1145/3055399.3055487}
-improves over the one cited in the internship description
-\fi
-
-\subsubsection{Parallel \pcc{}}
-\label{ssub:parallel_cc}
-
-\subsection{Empirical evaluation?}
-\label{sub:cc_empiracal_evaluation}
+\input{cc_part}
 

### 2017-07-21 lowstretch_and_cc.tex
formatting

@@
 
+\iffalse
 The method presented in the previous chapter hinges crucially upon our sign generative model. Yet
@@
 \end{enumerate}
+\fi
 
-\subsubsection{Experience on directed biological network}
-\label{ssub:experience_on_directed_biological_network}
+\subsubsection{Experience on directed biological networks}
+\label{ssub:experience_on_directed_biological_networks}
 
+\iffalse
 Take for instance a biological network.
@@
 
+\iffalse
 \section{Low stretch trees and spanners}
@@
 
-\input{gtx_algorithm}
+% \input{gtx_algorithm}
 
@@
 
-\input{gtx_related}
+% \input{gtx_related}
 
-\iffalse
 \url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
@@
 \begingroup
-\small
+\newgeometry{hmargin=2cm,vmargin=1.3cm}
 \setstretch{0.9}
@@
 \printbibliography
+\restoregeometry
 \endgroup

### 2017-07-21 cc_part.tex
describe the CC problem

@@
 \label{sub:problem_setting}
+
+Like other clustering frameworks, in \pcc{}, we are given a set of objects and we want to gather
+them into groups (called clusters) so that objects belonging to one cluster are similar to each
+other while being dissimilar to objects from all the other clusters.
+
+In \pcc{}, we formalize this problem by considering objects as the nodes of a graph $G$, whose edges
+weight encode similarity. Namely, in the most general case, for nodes $u$ and $v$, the edge between
+$u$ and $v$ is associated with two positive numbers:
+$w_{u,v}^+$ denotes the strength of the similarity between $u$ and $w$;
+$w_{u,v}^-$ denotes the strength of the dissimilarity between $u$ and $w$.
+Note however that in many applications, only one of these two numbers is non zero in which case we
+more conveniently set $w_{u,v} = \begin{cases}
+	 w_{u,v}^+ & \quad \text{if } w_{u,v}^+ > 0 \text{ and } w_{u,v}^-=0 \\
+	-w_{u,v}^- & \quad \text{if } w_{u,v}^+ = 0 \text{ and } w_{u,v}^->0 \\
+\end{cases}$
+
+Now consider a clustering \cluster{} of $V$, that is a function from $V$ to $\Nbb{}^{|V|}_{>0}$
+that assigns to each node a cluster index. For instance, $\cluster(u) = 3$ means that $u$ belongs
+to the third cluster. We can evaluate how \cluster{} fits our clustering paradigm in two ways,
+either by the number of \emph{agreements}, that is the weighted number of positive edges inside
+clusters plus the weighted number of negative edges across clusters; or by the number of
+\emph{disagreements}, that is the weighted number of negative edges inside clusters plus the
+weighted number of positive edges across clusters. Given a cost function $c$, which is usually the
+identity, \pcc{} can then be seen as graph optimization problem, either of maximizing agreements
+(\maxa{}):
+\begin{equation}
+	\max_{\cluster{}} \sum_{\cluster(u) = \cluster(v)} c(w_{uv}^+) +
+	\sum_{\cluster(u) \neq \cluster(v)} c(w_{uv}^-)
+	\label{eq:maxa}
+\end{equation}
+or minimizing disagreements (\mind{}):
+\begin{equation}
+	\min_{\cluster{}} \sum_{\cluster(u) = \cluster(v)} c(w_{uv}^-) +
+	\sum_{\cluster(u) \neq \cluster(v)} c(w_{uv}^+)
+	\label{eq:mind}
+\end{equation}
+
+Although an optimal clustering $\cluster^\star$ achieves the same value on both \eqref{eq:maxa}
+and \eqref{eq:mind}, we will see in \autoref{sub:state_of_the_art} that the latter objective is in
+some sense \enquote{easier}.
+
+\begin{figure}[hbt]
+	\centering
+	\includegraphics[width=0.9\linewidth]{assets/tikz/cc_objectives_tikz.pdf}
+	\caption[]{A clustering of a small with two disagreements due to two cycles with one negative edge.}
+	\label{fig:cc_objectives}
+\end{figure}
 

### 2017-07-31 cc_part.tex
add some applications of CC

@@
 and \eqref{eq:mind}, we will see in \autoref{sub:state_of_the_art} that the latter objective is in
-some sense \enquote{easier}.
+some sense \enquote{easier}. Another interesting feature of the \pcc{} problem is that contrary to
+other clustering formulations, it does not require us to set the number of clusters $k$ beforehand.
+Instead, $k$ emerges as a natural property of the solution. Since clustering is an unsupervised
+problem, this is generally handy. However, in some situations, we may have prior knowledge on how
+many clusters are the data, or external constraints. This can be handled with parametrized version
+of \pcc{} \autocites{Giotis2006}{Fomin2014}\marginpars{According to many articles, \pcc{} and
+cluster editing are equivalent, but even on complete graphs or on $(V,E^+)$ I don't see why.}.
+% The cluster edit distance of a graph is the smallest number of edges to change for it to admit a
+% perfect clustering (i.e., a union of cliques).  Equivalently, it is the cost of the optimal
+% correlation clustering. [From KDDTuto14, slides 204]
 
+In \autoref{fig:cc_objectives}, we show a simple instance of \pcc{} and one of its optimal solution.
 \begin{figure}[hbt]
 	\centering
-	\includegraphics[width=0.9\linewidth]{assets/tikz/cc_objectives_tikz.pdf}
-	\caption[]{A clustering of a small with two disagreements due to two cycles with one negative edge.}
+	\includegraphics[width=0.8\linewidth]{assets/tikz/cc_objectives_tikz.pdf}
+	\caption[Small example of \pcc{}]{A small graph with eight nodes and ten edges. Solid edges
+	represent positive edges and dashed edges represent negative edges. A clustering \cluster{}
+	is showed with 3 clusters: $\{1, 2, 4\}$, $\{3, 5, 6\}$ and $\{7, 8\}$. \cluster{} incurs two
+	disagreements: the negative edge between nodes $1$ and $2$ within the blue cluster, and the
+	positive edge $6,7$ between the orange and green clusters. Those disagreements are created
+	by two cycles with one negative edge and thus cannot be avoided, meaning that \cluster{} is
+	optimal. However it is not the unique solution, for instance merging the orange and green
+	clusters would also yields two disagreements.}
 	\label{fig:cc_objectives}
 \end{figure}
+
+\subsubsection{Applications}
+\label{ssub:cc_applications}
+
+According to \textcite[Section 5]{Demaine2006}, \pcc{} is well suited to several situations:
+\begin{itemize}
+   \item when the items to be clustered do not belong to a natural metric space (preventing
+      approaches such as $k$-means) but we still know for some pairs whether they are similar or
+      not.
+   \item when we do not know the number of clusters beforehand but we have a similarity measure. In
+      that case, we can select a problem-specific similarity threshold and set all edges with a
+      similarity larger than the threshold to be positive while the others are set to negative.
+   \item when we have a classic clustering problem (that is a set of objects, a distance between
+      them and an objective function to minimize) with additional pairwise constraints of the form
+      \emph{must-link/cannot-link}. Instead of restraining a clustering algorithm to the space of
+      feasible solution, we convert the distances between objects and the constraints into signed
+      edges and solve the resulting \pcc{} problem. 
+\end{itemize}
+
+In addition to these general considerations, \pcc{} has also been used in several domains:
+\begin{description}
+   \item[Computer Vision] For instance image segmentation
+      \autocites{Bagon2011}{Kim2011}{VolumeSegmentation12}.
+
+\Textcite{Shape3D17} develop a method to extract a network of descriptive curves from 3D shapes.
+After an initial stage of generating many such \emph{flowlines}, they describe in Section 6 a \pcc{}
+formulation to extract \emph{reliable} representative flowlines, using geometric constraints to give
+
+\textsc{Learning to Divide and Conquer for Online Multi-Target Tracking}
+In order to track several targets across sequential video frames, \textcite{multiTracking15} propose
+a multistage framework. One step revolves around a cost matrix $A$ and zones
+
+\item[Natural Language Processing] 
+   coreference resolution is solved by \textcite[Section 2.3]{graphicalCoreference04} using a
+undirected graphical model on which performing inference is equivalent to \pcc{}.
+	     \Textcite{Elsner2009} compared various heuristics with a bound of the optimal solution
+	     obtained through SDP relaxation and show that best performing ones are within few
+	     percents of it, provided they are followed by a local search step.
+Another application in NLP is clustering words based on distributional embedding while adding
+antonym constraints~\autocite{SignedWordRatings}.
+\item[Biology]
+
+		\Textcite{Mason2009} analyze a signed co-expression networks of genes
+		involved in embryonic stem cells to find which genes are related to
+		pluripotency or self-renewal.  Also something about haplotype assembly
+		\autocite{Das2015}.
+
+\Textcite{Ben-Dor99} assume there exists a perfect clustering, and receive as
+input a complete similarity matrix, which is corrupted by measurement error.
+They give an $O(n^2(\log n)^c)$ algorithm that recover the planted disjoint
+cliques with high probability.
+
+	   \enquote{They examined dynamical systems, where a gene is modeled as a vertex and an
+	      activating connection is modeled as a positive edge and an inhibiting connection is
+	      modeled as a negative edge. The claim is that biological dynamical systems are close
+	      to being balanced, and that finding a minimum set of edges to delete to make the graph
+	      balanced can be used to decompose the graph into “monotone subsystems”, which exhibit
+	   stable behavior and thus allow a better understanding of the dynamics of a
+	system.}\autocite{monotoneBiology07}
+
+     \item[Network science]
+study political vote \autocites{BrazilCC17}{Jiang2015}{Mendonca2015}. Also in politics
+community structure \autocites{Traag2009}{CommunityUN12} (more in the variants part)
+
+Visualization of signed social graph \autocite{Luca10}
+
+\item[Others]
+   \begin{itemize}
+	\item duplicate detection, also called entity resolution \autocite{DeDup09}
+
+statistical physics (Barahona 1982), Barahona F (1982) On the computational complexity of Ising
+	   spin glass models. J Phys A: Math Gen 15(10):3241–3253 \autocite{Ising82}
+
+portfolio risk analysis
+	   \enquote{In this context, each security is represented by a vertex in the signed graph
+	      while the correlation between securities is represented by the set of signed edges. A
+	      balanced signed graph with only positive edges represents a speculative portfolio
+	      since all its securities tend to move in the same direction, either on the upside or
+	      on the downside. On the other hand, a balanced signed graph with at least one negative
+	      edge is associated with a limited risk portfolio. Such a portfolio is defined by two
+	      sets of securities, each set with a tendency to move in tandem, while some pair of
+	      assets (connected by negative edges) tend to move in opposite directions providing the
+	      investors with a hedging guarantee. According to Harary et al. (2003), an unbalanced
+	   signed graph represents an unpredictable portfolio.}\autocite{portfolio02}
+
+Given an electrical circuit layout, \textcite{circuitDesign07} extract a graph of its components
+(called shifter) that must be assigned one of two possible phases. Because two shifters next to some
+specific shape must be in opposite phase and two shifters separated by less than a specified
+distance must be of the same phase, the authors look for a two-clustering of the nodes that will
+minimize the number of disagreements.
+% [21] N. Gülpinar, G. Gutin, G. Mitra, and A. Zverovitch. Extracting pure network submatrices in linear programs using signed graphs. Discrete Applied Mathematics, 137:359–372, 2004
+% + Figueiredo2011
+   \end{itemize}
+\end{description}
 

### 2017-08-01 cc_part.tex
finish writing on CC applications

--- 2017-07-21 cc_part.tex
@@
 and \eqref{eq:mind}, we will see in \autoref{sub:state_of_the_art} that the latter objective is in
-some sense \enquote{easier}.
+some sense \enquote{easier}. Another interesting feature of the \pcc{} problem is that contrary to
+other clustering formulations, it does not require us to set the number of clusters $k$ beforehand.
+Instead, $k$ emerges as a natural property of the solution. Since clustering is an unsupervised
+problem, this is generally handy. However, in some situations, we may have prior knowledge on how
+many clusters are the data, or external constraints. This can be handled with parametrized version
+of \pcc{} \autocites{Giotis2006}{Fomin2014}\marginpars{According to many articles, \pcc{} and
+cluster editing are equivalent, but even on complete graphs or on $(V,E^+)$ I don't see why.}.
+% The cluster edit distance of a graph is the smallest number of edges to change for it to admit a
+% perfect clustering (i.e., a union of cliques).  Equivalently, it is the cost of the optimal
+% correlation clustering. [From KDDTuto14, slides 204]
 
+In \autoref{fig:cc_objectives}, we show a simple instance of \pcc{} and one of its optimal solution.
 \begin{figure}[hbt]
 	\centering
-	\includegraphics[width=0.9\linewidth]{assets/tikz/cc_objectives_tikz.pdf}
-	\caption[]{A clustering of a small with two disagreements due to two cycles with one negative edge.}
+	\includegraphics[width=0.8\linewidth]{assets/tikz/cc_objectives_tikz.pdf}
+	\caption[Small example of \pcc{}]{A small graph with eight nodes and ten edges. Solid edges
+	represent positive edges and dashed edges represent negative edges. A clustering \cluster{}
+	is showed with 3 clusters: $\{1, 2, 4\}$, $\{3, 5, 6\}$ and $\{7, 8\}$. \cluster{} incurs two
+	disagreements: the negative edge between nodes $1$ and $2$ within the blue cluster, and the
+	positive edge $6,7$ between the orange and green clusters. Those disagreements are created
+	by two cycles with one negative edge and thus cannot be avoided, meaning that \cluster{} is
+	optimal. However it is not the unique solution, for instance merging the orange and green
+	clusters would also yields two disagreements.}
 	\label{fig:cc_objectives}
 \end{figure}
+
+\subsubsection{Applications}
+\label{ssub:cc_applications}
+
+According to \textcite[Section 5]{Demaine2006}, \pcc{} is well suited to several situations:
+\begin{itemize}
+   \item when the items to be clustered do not belong to a natural metric space (preventing
+      approaches such as $k$-means) but we still know for some pairs whether they are similar or
+      not.
+   \item when we do not know the number of clusters beforehand but we have a similarity measure. In
+      that case, we can select a problem-specific similarity threshold and set all edges with a
+      similarity larger than the threshold to be positive while the others are set to negative.
+   \item when we have a classic clustering problem (that is a set of objects, a distance between
+      them and an objective function to minimize) with additional pairwise constraints of the form
+      \emph{must-link/cannot-link}. Instead of restraining a clustering algorithm to the space of
+      feasible solution, we convert the distances between objects and the constraints into signed
+      edges and solve the resulting \pcc{} problem. 
+\end{itemize}
+
+In addition to these general considerations, \pcc{} has also been used in several domains:
+\begin{description}
+   \item[Computer Vision]
+      The ubiquitous task of segmenting an image into homogeneous regions is prerequisite for many
+      further processing, and can be tackled with \pcc{}. For instance, to segment cell in
+      microscopy imagery, \textcite{CellSeg14} first use generic image features to classify pixel in
+      belonging to region boundaries or not. Then, they extract small scale regions called
+      superpixels. After building the adjacency graph of these superpixels, they assigned edge
+      weight by averaging the boundary probability over the pixels separating the adjacent
+      superpixels. They also add strong negative constraints between distant superpixels, and lastly
+      cluster these superpixels according to the \pcc{} objective to obtain the final segmentation.
+      A similar approach was used earlier in~\textcite{Kim2011}, which stressed the importance of
+      considering such higher order constraints between distant superpixels in order to achieve good
+      performance. This was also extended to 3D segmentation~\autocite{VolumeSegmentation12}, where
+      additional tunings allowed to segment a volume image of a mouse cortex with up to billions
+      voxels. \Textcite{Beier2015} segment 2D and 3D images with a energy based formulation of
+      \pcc{} and iteratively improve their solution by merging it with another clustering given by a
+      proposal generator. By developing another scalable energy based optimization procedure, and
+      with the help of few user provided cues, \textcite{Bagon2011} were able to apply \pcc{}
+      directly at the pixel level.
+      % one more? https://link.springer.com/chapter/10.1007/978-3-642-33783-3_41
+
+      Beyond image segmentation, \textcite{Shape3D17} develop a method to extract a network of
+      descriptive curves from 3D shapes.  After an initial stage of generating many such
+      \emph{flowlines}, they describe in Section 6 a \pcc{} formulation to extract \emph{reliable}
+      representative flowlines, using geometric constraints to obtain positive or negative cues that
+      two flowlines are from the same reliable representative.
+
+      Finally, in order to track several targets across sequential video frames,
+      \textcite{multiTracking15} propose a multistage framework. One step revolves around a matrix
+      $A$ that defines the cost of assigning an object tracked in previous frames to an object
+      detected in the current frame. This matrix is turned into a symmetric affinity matrix
+      $\bar{A}_{sym}$ that can be seen as a signed graph. \pcc{} is then used to extract clusters
+      (called zones), in which local processing is performed, whose complexity can be adapted to the
+      difficulty of each zone.
+
+   \item[Natural Language Processing] 
+      Coreference resolution is the task of finding all expressions that refer to the same entity in
+      a text. Like image segmentation, it is a preprocessing step that can later be used in document
+      summarization, question answering, and information extraction. \Textcite[Section
+      2.3]{graphicalCoreference04} tackle coreference resolution by using a undirected graphical
+      model on which performing inference is equivalent to \pcc{}. On small scale instances,
+      \textcite{Elsner2009} first obtain an upper bound of the optimal solution by solving a SDP
+      relaxation of the problem. They then compared various heuristics and show that best performing
+      ones are within few percents of the optimum, provided they are followed by a local search
+      step, such as the Best One Element Move~\autocite{Gionis2007}. Another possible task would be
+      to cluster words based on distributional embedding vector while adding antonym
+      constraints~\autocite{SignedWordRatings}, although the authors of that paper choose to use
+      signed spectral clustering.
+
+   \item[Biology]
+      The input is a similarity matrix between gene expressions in various experimental conditions
+      and the goal is to cluster those genes into groups which react similarly. \Textcite[Section
+      4]{Ben-Dor99} gives three examples: 112 genes involved in the rat central nervous system, 1246
+      genes of the roundworm \emph{C. elegans} and 2000 human genes obtained from 40 tumor and 22
+      normal colon samples. \Textcite{Mason2009} analyze a signed co-expression networks of genes
+      involved in embryonic stem cells to find which genes are related to pluripotency (the ability
+      to differentiate into any type of cell in the body) or self-renewal (the ability to replicate
+      indefinitely). Another application is to study the variation of one individual
+      DNA~\autocite{Das2015}. In the human organism, chromosomes are organized in pair, and both
+      chromosomes of a pair have similar but not identical DNA sequences. This is mostly because of
+      single nucleotide polymorphisms (SNPs), where a single base differs between the two DNA
+      sequences, leading to different alleles of the corresponding gene. An haplotype is the list of
+      all alleles at a contiguous sites in a region of a single chromosome and this information is
+      used in several medical applications. The high-throughput sequencing of one individual genome
+      yields many short \emph{reads} that provides information about the order of nucleotides in a
+      fragment of one chromosome and that can be used to assemble haplotypes. Namely the authors
+      build a graph of reads and define a similarity function between reads to assign weight on the
+      edges. The clusters of that graph correspond to haplotypes, and are obtained by a SDP
+      relaxation of the \pcc{} objective. \Textcite{monotoneBiology07} also consider graphs whose
+      nodes are genes, but in a different context. In this case, positive edges represent an
+      activating connection, while negative edges represent inhibiting connection. They also define
+      a \emph{monotone system} as a balanced subgraph, that is a subgraph which does not contain a
+      cycle with an odd number of edges. Such monotone system are stable, in the sense that
+      modifying the concentration of one gene will have a predictable effect, even ignoring the
+      precise kinematics of the chemical reactions involved. Their goal is to find the minimum
+      number of edges to remove in order to decompose a dynamics system into a collection of
+      monotone system. This corresponds to the \pcc{} \eqref{eq:mind} objective and allows to study
+      the complete system more easily.
+      % \enquote{They examined dynamical systems, where a gene is modeled as a vertex and an
+      % activating connection is modeled as a positive edge and an inhibiting connection is modeled as
+      % a negative edge. The claim is that biological dynamical systems are close to being balanced,
+      % and that finding a minimum set of edges to delete to make the graph balanced can be used to
+      % decompose the graph into “monotone subsystems”, which exhibit stable behavior and thus allow a
+      % better understanding of the dynamics of a system.}\autocite{monotoneBiology07}
+
+   \item[Network science]
+      As mentioned in the first Chapter, one early use of signed graphs was to model social
+      interactions and given the large body of work on community
+      detection~\autocite{FortunatoSurvey10}, it is therefore not surprising that \pcc{} has been
+      used to study such networks. For instance, one can extract all the votes of the member of
+      political parliament and form a graph whose node are politicians and edge weight quantify how
+      much they agree or disagree on various issues they have been voting on. This can be used to
+      study various social science questions such as loyalty, leadership, coalitions, political
+      crisis and polarization. It has been applied to the European
+      parliament~\autocite{Mendonca2015}, Slovenian parliament~\autocite{Jiang2015} and the
+      Brazilian parliament~\autocites{BrazilCC17}. This can also be used at international level. For
+      instance, by considering a dataset of military alliances and disputes, \textcite{Traag2009}
+      cluster countries into blocks that resemble those identified by Huntington in his \emph{Clash
+      of Civilizations} book. Another source of data is the vote on resolutions during the United
+      Nations General Assembly~\autocite{CommunityUN12}. Finally, one can also study how to exploit
+      the information contained within the negative links to enhance the visualization of social
+      networks~\autocite{Luca10}. Note however that those three last papers not explicitly use a
+      \pcc{} formulation.
+
+   \item[Others]
+      \begin{itemize}[leftmargin=*]
+	 \item
+	    Deduplication, also called duplicate detection or entity resolution, is the process
+	    of identifying objects from a real-world, noisy database that refer to the same entity.
+	    On a high level, a solution to this problem is to build a graph of all the available
+	    objects, define a similarity between them and run a \pcc{} algorithm. The main challenge
+	    thus lies in devising an appropriate similarity measure, given that object can have very
+	    different features from one database to another. \Textcite{LargeScaleDeDup09} propose a
+	    declarative language, expressing both hard constraints (that have to be satisfied) and
+	    soft constraints (that can be seen as cues guiding the process). Because of these hard
+	    constraints that admissible clusterings have to respect, the authors have to modify in
+	    non trivial way an existing \pcc{} algorithm. \pcc{} was also evaluated among other
+	    solutions to that problem by \textcite{DeDup09}, who note that their non optimized
+	    implementation does not perform the best.
+	 \item 
+	    Given an electrical circuit layout, \textcite{circuitDesign07} extract a graph of its
+	    components (called shifter) that must be assigned one of two possible phases. Because
+	    two shifters next to some specific shape must be in opposite phase and two shifters
+	    separated by less than a specified distance must be of the same phase, the authors look
+	    for a two-clustering of the nodes that will minimize the number of disagreements.
+      \end{itemize}
+\end{description}
 

### 2017-08-04 cc_part.tex
add one more ref of dedup application
add a draft section on connection CC and edge sign

@@
 In addition to these general considerations, \pcc{} has also been used in several domains:
+%TODO turn those items into paragraphs?
 \begin{description}
@@
 	    On a high level, a solution to this problem is to build a graph of all the available
-	    objects, define a similarity between them and run a \pcc{} algorithm. The main challenge
+	    objects, define a similarity between them and run a \pcc{} algorithm. This was indeed
+	    this kind of problems at Whizbang! Labs that partly motivated one of the early \pcc{}
+	    paper~\autocite{Bansal2002}. The main challenge
 	    thus lies in devising an appropriate similarity measure, given that object can have very
@@
 	    constraints that admissible clusterings have to respect, the authors have to modify in
-	    non trivial way an existing \pcc{} algorithm. \pcc{} was also evaluated among other
+	    non trivial ways an existing \pcc{} algorithm. Another example is given by
+	    \textcite{Crosslingual07}, who cluster together news article in different language
+	    covering the same event. \pcc{} was also evaluated among other
 	    solutions to that problem by \textcite{DeDup09}, who note that their non optimized
@@
 
-\iffalse
-For the 2 clusters case ($k=2$), characterization proven already in 1936 by \textcite{Konig36}, as noted in
-\cite{Zaslavsky2012} (commenting on \cite{harary1953}: \enquote{Although Thm. 3 was anticipated by
-\textcite[Theorem X.11]{Konig36}  without the terminology of signs, here is the
-first recognition of the crucial fact that labelling edges by elements of a
-group—specifically, the sign group—can lead to a general theory.})
-and \cite{Huffner2010} (\enquote{\textcite{Konig36} proved the following characterization of
-	balanced graphs. For a graph $G = (V , E)$, the following are equivalent:\marginpars{There is a
-	proof in \autocite[p. 111]{BookKleinberg2010}, maybe I can rewrite it as well}
-	\begin{enumerate}
-		\item $V$ can be partitioned into two sets $V_1$ and $V_2$ called sides such that there is no
-			negative edge $\{v, w\} \in E$ with both $v, w \in V_1$ or both $v, w \in V_2$ and no negative
-			edge $\{v, w\}$ with $v \in V_1$ and $w \in V_2$ .
-		\item $V$ can be colored with two colors such that for all $\{v, w\} \in E^-$, the vertices $v$
-			and $w$ have different colors, and for all $\{v, w\} \in E^+$, the vertices $v$ and $w$ have
-			the same color. The color classes correspond to the sides.
-		\item $G$ does not contain cycles with an odd number of negative edges.
-	\end{enumerate}
-	Using the characterization by a coloring, it is easy to see that balance of a signed
-graph can be checked in linear time by depth-first search.})
-\fi
+\input{cc_sign_relation}
 

### 2017-08-04 cc_sign_relation.tex
add a draft section on connection CC and edge sign

Now that we are familiar with the \pcc{} problem and its numerous applications, let us return to the
original problem of this chapter, predicting the edge sign of a signed graph. In order to do so, we
will first focus on the case where a signed graph has no disagreements (that is the minimum of
\eqref{eq:mind} is $0$), and the connection with the balance theory.

In his seminal work on interpersonal relations,
\citeauthor*{Heider46}~\autocites{Heider46}{HeiderBook58} argues by
psychological and sociological arguments that in order to reduce their cognitive dissonance, three
people always interact in a way that preserve social balance. This social balance requirements can be
succinctly summarized by four statements~\autocite{HeiderBook58}:

\begin{enumerate}
	\item \textcolor{DodgerBlue}{my friend's} \textcolor{Orange}{friend} is
		\textcolor{Olive}{my friend}\hspace{1em}%
		\tikz[baseline=5]{
			\node[draw,circle,inner sep=2pt] (me) at (0,0) {};
			\node[draw,circle,inner sep=2pt] (mine) at (.5,.5) {};
			\node[draw,circle,inner sep=2pt] (alter) at (1,0) {};
			\draw[] (me) -- node[left,DodgerBlue,pos=.9] {$+$} (mine) ;
			\draw[] (mine) -- node[right,Orange,pos=.1] {$+$} (alter) ;
			\draw[white] (me) -- node[text=Olive] {$+$} (alter) ;
		}
	\item \textcolor{DodgerBlue}{my friend's} \textcolor{Orange}{enemy} is
		\textcolor{Olive}{my enemy}\hspace{1em}%
		\tikz[baseline=5]{
			\node[draw,circle,inner sep=2pt] (me) at (0,0) {};
			\node[draw,circle,inner sep=2pt] (mine) at (.5,.5) {};
			\node[draw,circle,inner sep=2pt] (alter) at (1,0) {};
			\draw[] (me) -- node[left,DodgerBlue,pos=.9] {$+$} (mine) ;
			\draw[] (mine) -- node[right,Orange,pos=.1] {$-$} (alter) ;
			\draw[white] (me) -- node[text=Olive] {$-$} (alter) ;
		}
	\item \textcolor{DodgerBlue}{my enemy's} \textcolor{Orange}{friend} is
		\textcolor{Olive}{my enemy}\hspace{1em}%
		\tikz[baseline=5]{
			\node[draw,circle,inner sep=2pt] (me) at (0,0) {};
			\node[draw,circle,inner sep=2pt] (mine) at (.5,.5) {};
			\node[draw,circle,inner sep=2pt] (alter) at (1,0) {};
			\draw[] (me) -- node[left,DodgerBlue,pos=.9] {$-$} (mine) ;
			\draw[] (mine) -- node[right,Orange,pos=.1] {$+$} (alter) ;
			\draw[white] (me) -- node[text=Olive] {$-$} (alter) ;
		}
	\item \textcolor{DodgerBlue}{my enemy's} \textcolor{Orange}{enemy} is
		\textcolor{Olive}{my friend}\hspace{1em}%
		\tikz[baseline=5]{
			\node[draw,circle,inner sep=2pt] (me) at (0,0) {};
			\node[draw,circle,inner sep=2pt] (mine) at (.5,.5) {};
			\node[draw,circle,inner sep=2pt] (alter) at (1,0) {};
			\draw[] (me) -- node[left,DodgerBlue,pos=.9] {$-$} (mine) ;
			\draw[] (mine) -- node[right,Orange,pos=.1] {$-$} (alter) ;
			\draw[white] (me) -- node[text=Olive] {$+$} (alter) ;
		}
\end{enumerate}

This can be readily translated into graph properties. Given a path of length two over three nodes,
note that the first part of each statement (in blue and orange) defines the four possible sign
assignment of such a path. The last part of the statement (in green) then prescribes which sign
should close the triangle to respect the social balance. As we can see, this means that triangles can
only have zero or two negative signs, or equivalently, that no triangle can have an odd number
number of negative signs.

From this local observation, we define a graph as strongly balanced if none of its cycle has an odd
number of negative signs
in that case we have \textcite[Theorem 3]{harary1953} stating:
statement
proof
According to \textcite{Huffner2010}, a similar theorem was proved earlier by \textcite[Theorem
X.11]{Konig36}, although \textcite{Zaslavsky2012} notes that it was stated \enquote{without the
	terminology of signs, while \autocite{harary1953}
has the first recognition of the crucial fact that labelling edges by elements of a
group---specifically, the sign group---can lead to a general theory.}

when graph is strongly balanced, the 2 clusters of previous claim gives a natural partition with no
disagreements in the sense of \pcc{}. However, the existence of 2 such clusters is only a sufficient
but not necessary condition for having no disagreement. As hinted in
\autoref{sub:need_for_a_directed_graph}, when the nodes are partitioned into $k$ clusters, letting
edges within clusters to be positive and edges across clusters to be negative will by definition
results in zero disagreement. \Textcite{davis1967clustering} show that this $k>2$ clusters partition
is equivalent to the absence of cycle with a single negative edge, namely
statement
proof
In terms of balance, this corresponds to allowing triangle to have three negative edges, and is called
weak balance, as illustrated on \autoref{fig:cc_balance}.

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.8\linewidth]{assets/raw/triangle_smaller.pdf}
	\caption{The four possible undirected triads, as classified by the two structural balance theories
	introduced in the main text} \label{fig:cc_balance}
\end{figure}

In real data though, we do not expect neither strong nor weak balance to hold, for they are fairly
demanding model. Indeed, three of the real networks considered in \autoref{chap:troll} have been
repeatedly shown to be unbalanced, although the extent of this unbalance depends of the importance
given to longer cycle compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
{measureUnbalance17} For the same reason, in virtually any real instance of the \pcc{} problem, the
optimal solution will incur some disagreements. \Textcite[Section 5.5]{BookKleinberg2010}
nonetheless prove that even when only a fraction of the triangle are strongly balanced in a complete
graph, the two clusters structure is still present, although it does not cover the whole graph
anymore. Formally statement
The proof makes essential use of the strong balance (which is ok since next section also deal with
2-CC) and the completeness of the graph (which is a far less realistic assumption) yet it would be
interesting to see if it can be extended.

Said that, how is \pcc{} useful in predicting edge sign? At first sight, the two problems seem quite
different for the former is an unsupervised/agnostic clustering problem while the latter is a
supervised classification problem. Yet, getting an optimal solution of \pcc{} on the training set
would provide a principled heuristics for \esp{}. Unfortunately, as we shall see in
\autoref{sub:state_of_the_art}, exact or fixed parameter solution are not scalable and one has to
rely approximation. Under the Unique Games Conjecture, the approximation factor is lower bounded by
$\Theta(\log n)$ in the worst case. However, this is still a viable option, as our bias is that we
will operate on instances that are to some extent balanced. Therefore, we are not in the worst case
and we can find some comfort in results about the noisy (\vpageref{sub:variants_and_extensions}) and
stable (\autoref{ssub:cc_under_stability_assumption}) settings.

Another connection we will leverage in \autoref{sec:low_stretch_trees_and_spanners} is that the
value of the \mind{} objective is a complexity measure of the \esp{} problem~\autocite[Section
4.1]{Cesa-Bianchi2012b}. Before that, we will also present some variants and extentsion of \pcc{},
some of them bearing similarities with the Chapter 3's problem.

there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration index)
which they claimed has been peer reviewed? They also mention their method to compute frustration
exactly \url{https://arxiv.org/abs/1611.09030} along with existing work (for $n=50$)
\url{https://doi.org/10.1016/j.socnet.2013.09.002} (leading to an 1996 heuristics for \pcc{} in
section 3 of \url{https://doi.org/10.1016/0378-8733(95)00259-6})

\iffalse
strong balance random graph model \url{https://www.lri.fr/~yannis/randsigned.pdf}, line index is basically
the objective value of CC and if I'm reading their bound correctly,  basically you need to flip half
the edges of a random graph to make it balanced (also most of their results hold when the
probability of an existing edge is 1/2, but it was strengthened in a follow up dealing with weak
balance \url{http://people.maths.ox.ac.uk/harutyunyan/weak-balance.pdf})

For the 2 clusters case ($k=2$), characterization proven already in 1936 by \textcite{Konig36}, as noted in
\cite{Zaslavsky2012} (commenting on \cite{harary1953}: \enquote{Although Theorem 3 was anticipated by
\textcite[Theorem X.11]{Konig36}  without the terminology of signs, here is the
first recognition of the crucial fact that labelling edges by elements of a
group---specifically, the sign group---can lead to a general theory.})
and \cite{Huffner2010} (\enquote{\textcite{Konig36} proved the following characterization of
	balanced graphs. For a graph $G = (V, E)$, the following are equivalent:\marginpars{There is a
	proof in \autocite[p. 111]{BookKleinberg2010}, maybe I can rewrite it as well}
	\begin{enumerate}
		\item $V$ can be partitioned into two sets $V_1$ and $V_2$ called sides such that there is no
			negative edge $\{v, w\} \in E$ with both $v, w \in V_1$ or both $v, w \in V_2$ and no positive
			edge $\{v, w\}$ with $v \in V_1$ and $w \in V_2$ .
		\item $V$ can be colored with two colors such that for all $\{v, w\} \in E^-$, the vertices $v$
			and $w$ have different colors, and for all $\{v, w\} \in E^+$, the vertices $v$ and $w$ have
			the same color. The color classes correspond to the sides.
		\item $G$ does not contain cycles with an odd number of negative edges.
	\end{enumerate}
	Using the characterization by a coloring, it is easy to see that balance of a signed
graph can be checked in linear time by depth-first search.})
\fi

### 2017-08-04 cc_approx.tex
add a section on CC approximation results

\subsection{Hardness and approximation}

Although the same problem was studied earlier~\autocites{Early96}{Ben-Dor99}, \textcite{Bansal2002}
coined the term \pcc{} and were the first to study this problem complexity. Namely, for complete,
unweighted signed graphs, they show that both \mind{} and \maxa{} are \NPc{}. Along the way, they
give a 17429-approximation of \mind{} and a PTAS that, for any $\epsilon \in [0,1]$, runs in
\marginpars{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
minimization (respectively maximization) problem $\mathcal{P}$  in NP if given any $\epsilon>0$ and
any instance $x$ of $\mathcal{P}$ of size $n$, $\mathcal{A}$ produces, in polynomial time in $n$, a
solution that is within a factor $1+\epsilon$ (respectively $1-\epsilon$) of being optimal with
respect to $x$. Note that the time is not necessarily polynomial in $\epsilon$, so that a running
time of $O(n^{\frac{1}{\epsilon}})$ would qualify~\autocite[Definition 3.10]{CpxBook99}.}
$O(n^2e^{O(\frac{1}{\epsilon^{10}}\log\frac{1}{\epsilon})})$ and returns with probability
$1-\frac{\epsilon}{3}$ a solution with at most $\epsilon n^2$ fewer agreements than the optimal of
\maxa{}.

\begin{table}[htpb]
   \centering
   \small
   \caption{Hardness results of \pcc{}} \label{tab:cc_cpx}
   \begin{tabulary}{187mm}{lCCCC}
      \toprule
               & \multicolumn{2}{c}{\mind{}}   & \multicolumn{2}{c}{\maxa{}}                                   \\
      \cmidrule(r){2-3}
      \cmidrule(r){4-5}
      graph    & weighted                      & unweighted                                                     & weighted                                                     & unweighted                   \\
      \midrule
      Complete & \APXh{}~\autocite{Charikar2003} & \NPc{}~\autocite{Bansal2002}, \APXh{}~\autocite{Charikar2003}  &                                                              & \NPc{}~\autocite{Bansal2002} \\
      General  & \APXh{}~\autocites{Charikar2003}{Demaine2003} &  \APXh{}~\autocites{Charikar2003}{Emanuel2003}   & \multicolumn{2}{c}{\APXh{}~\autocite[Thm. 9]{Charikar2003}} \\
      \bottomrule
   \end{tabulary}
\end{table}

The next year, several authors independently strengthened these results and extended them to
weighted and general graphs, as summarized in \autoref{tab:cc_cpx}. In the most complete paper,
\textcite{Charikar2003} show that on complete graphs, minimizing disagreements is \APXh{},
\enquote{that is, is NP-hard to approximate within some constant factor greater than one} (it would
be nice to provide some intuition why it's more difficult to minimize disagreements but according to
the authors themselves, their reduction is \enquote{somewhat intricate}). They prove the same result
on general graphs, for \mind{} by using a reduction from the multicut problem \autocite[Theorem
8]{Charikar2003} (see \autocite{Emanuel2003} for the other direction and \autocite{Demaine2006} for
unified proof?) which ask for the minimum weight set of edges whose removal in $G$ disconnect the
$k$ pairs $(s_i, t_i)$ and for \maxa{} by a reduction from MAX 3SAT~\autocite[Theorem
9]{Charikar2003}. The multicut reduction yields a $O(\log n)$ approximation bound, which is achieved
by rounding a Linear Program, that we now describe. Assign a binary variable $x_{uv}$ to each pair
of nodes (so that $x_{uv}=x_{vu}$). For a given clustering \cluster{}, let $x_{uv} = 0$ if $u$ and
$v$ are in the same cluster and $x_{uv}=1$ is $u$ and $v$ are in different clusters. Noting that
$1-x_{uv}$ is $1$ if the edge $(u,v)$ is within a cluster and $0$ otherwise, the weighted
number of disagreements is then $w(\cluster{}) = \sum_{(u,v)\in E^-} w_{uv}(1-x_{uv}) +
\sum_{(u,v)\in E^+} w_{uv}x_{uv}$. By construction, if edges $(u,v)$ and $(v,w)$ are within the same
cluster, then $(v, w)$ is also within that cluster. In terms of $x$ variable, we have that $x_{uv}=0
\wedge x_{vw}=0 \implies x_{uw} = 0$. For $x$ to be a valid cluster assignment, we thus that all
variable are either $0$ or $1$ and respect the triangle inequality. We can thus relax the problem into
\begin{align}
   \label{eq:mindLP}
   \text{minimize } & \sum_{(u,v)\in E^-} w_{uv}(1-x_{uv}) + \sum_{(u,v)\in E^+} w_{uv}x_{uv} \\
   \text{subject to}& \quad x_{uw} \leq x_{uv} + x_{vw} \nonumber\\
   \phantom{subject to}& \quad 0 \leq x_{uv} \leq 1  \nonumber \\
   \phantom{subject to}& \quad x_{uv} = x_{vu}  \nonumber
\end{align}
Once the LP \eqref{eq:mindLP} is solved, we interpret $x_{uv}$ as a distance: the larger it its and
the more we want $u$ and $v$ to be in different cluster. We can then use the \emph{region growing}
method~\autocite{RegionGrowing93}. Namely, we pick a node at random and grow a ball based on these
distance until a threshold is met, at which point we create a cluster from this ball, remove the
corresponding nodes from the graph and repeat until all nodes are clustered.

However, \textcite[Theorem 2]{Charikar2003} note that the LP formulation has a poor integrality gap
when it comes to \maxa{}, thus they turn to a Semi Definite Program. Say that each cluster is
associated with a basis vector, then for each node $u$ in a cluster, we set $a_u$ to be the
corresponding basis vector. If $u$ and $v$ are in the same cluster, we then have $a_u\cdot a_v = 1$
while if they belong to different clusters, $a_u\cdot a_v = 0$. The weighted number of agreements
can then be represented by
\begin{align}
   \label{eq:maxaSDP}
   \text{maximize } & \sum_{(u,v)\in E^+} w_{uv}(a_u\cdot a_v) + \sum_{(u,v)\in E^-} w_{uv}(1-a_u\cdot a_v) \\
   \text{subject to}& \quad a_u\cdot a_u=1 \nonumber\\
   \phantom{subject to}& \quad a_u\cdot a_v\geq 0  \nonumber
\end{align}
After solving the SDP, a clustering can be obtained by a general rounding technique $H_t$: pick $t$
random hyperplanes and divides the nodes in $2^t$ clusters. \Textcite[Theorem 3]{Charikar2003} prove
that taking the best results of $H_2$ and $H_3$ gives in a $0.7664$ approximation on general graph.
This was slightly improved to $0.7666$ by \textcite{Swamy2004} with a different rounding: pick $k$
random unit vectors (called \emph{spokes}) and assign each $a_u$ to the closest spoke.

Combining \mind{} and \maxa{}, \textcite[Section 4]{Charikar2004} give a $\Omega(\frac{1}{\log n})$
approximation of the \textsc{MaxCorr} problem, which is maximizing \eqref{eq:maxa} - \eqref{eq:mind}
and can be formulated as a quadratic programming problem solved in polynomial time.

In complete graphs, \textcite[Section 3]{Charikar2003} also give an improved $4$-approximation to
\mind{}, by rounding the same LP and using its solution in randomized algorithm. We will not
describe it in detail since a similar idea was used by \textcite{CCPivotConf05} with a better
approximation. To explain it, we first describe their randomized combinatorial algorithm \ccpivot{},
which gives a $3$-approximation of \mind{} on complete unweighted graphs (it has later been
derandomized while preserving its approximation guarantee~\autocite{derandomCCPivot08}). At each
iteration, we pick a node $u$ \uar{} and we create a cluster containing $u$ and all its neighbors
linked by a positive edges. On weighted complete graphs, they tweak this algorithm by using the
solution of the  LP~\eqref{eq:mindLP} to obtain different approximation factor depending on the
constraints imposed on the weight.\Todo{Comment on those constraints, especially the triangular ones
in the context of consensus clustering.} Recall that in the general formulation of the problem, each
edge carries two positive numbers: $w_{u,v}^+$ and $w_{u,v}^-$. If the weights respect the
probability constraints stating that for all edge $(u,v)$ in $E$, $w_{u,v}^+ + w_{u,v}^- = 1$, they
get a $2.5$-approximation. Note that unweighted graphs naturally fit into that case. If the weights
additionally respect the triangular inequality constraints stating that $w_{u,v}^- \leq w_{u,w}^- +
w_{w,v}^-$, this become a $2$-approximation. After solving the LP~\eqref{eq:mindLP} with additional
probability constraints, when picking a node $u$, each of its neighbors $v$ is added to the cluster
of $u$ with probability $x_{uv}$. \Textcite{Chawla2014} improve these two factors to respectively
$2.06$ and $1.5$ by exploiting the same idea but setting the probability to include each neighbor
$v$ of $u$ in the cluster of $u$ to be $1-f^+(x_{uv})$ if $(u,v)\in E^+$ and $1-f^-(x_{uv})$ if
$(u,v)\in E^-$, with a careful choice of $f^+$ and $f^-$. They also give a derandomized version of
their algorithm in the full version of the paper~\autocite[Theorem 23]{ChawlaArxiv14}.

\begin{table}
   \begin{tabulary}{187mm}{llcLL}
      \toprule
                                &            & $k$   & \mind{}                                                                & \maxa{}                                                     \\
      \midrule
      \multirow{2}{*}{Complete} & unweighted &       & $2.06$ \autocite{Chawla2014}                                           & PTAS from \textcite{Bansal2002}  \\
      \cmidrule(r){4-4}
                                & weighted   &       & $1.5$ (with triangular inequality) \autocite{Chawla2014}               &  and by setting $k=\Omega(1/\epsilon)$ \autocite{Giotis2006}  \\
      \midrule
   \multirow{3}{*}{General}     & unweighted &       & \multicolumn{2}{c}{No specific results for the unweighted case}                                                                       \\
      \cmidrule(r){4-5}
                                & weighted   & $k=2$ & $O(\sqrt{\log n})$ \autocite{Giotis2006}                               & $0.884$ \autocite{Mitra2009}                                  \\
      \cmidrule(r){4-5}
                                & weighted   &       & $O(\log n)$ \autocite{Charikar2003}, optimal under the UCG conjecture  & $0.7666$ \autocite{Swamy2004}                                 \\
      \bottomrule
   \end{tabulary}
   \caption{Best results on various problem.\label{tab:cc_approx}}
\end{table}

This conclude the presentation of the known approximation results on \pcc{}, that we summarize in
\autoref{tab:cc_approx}. As mentioned earlier, not having to set the number of clusters is an
attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.
The problem was studied by \textcite{Giotis2006} on general graph and we compiled their results in
\autoref{tab:cc_fixed}. On general graph and with $k$ being the number of clusters, 
they provide PTAS for \maxa{} running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and
for \mind{} running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}k^6\log d\right)}$.

\begin{table}[htpb]
   \centering
   \caption{Approximation results for \pcc{} on general graph with $k$ clusters} \label{tab:cc_fixed}
   \begin{tabulary}{187mm}{lLL}
      \toprule
      $k$	 & 2 & $\geq 3$ \\
      \midrule
      \maxa{} & 0.878 (improved to 0.884 by \autocite{Mitra2009}) & 0.7666 \autocite{Swamy2004} \\
      \mind{} & $O(\sqrt{\log n})$ as it reduces to Min 2CNF Deletion for which \textcite{min2CNF05}
      give such an approximation &
      this can be reduced from $k$-coloring, which for any $\epsilon > 0$ is \NPc{} to approximate
      within $n^{1-\epsilon}$ \autocite{InnaproxChroma07} \\
      \bottomrule
   \end{tabulary}
\end{table}

### 2017-08-05 cc_approx.tex
add an extra PTAS for maxagree

@@
 \autoref{tab:cc_approx}. As mentioned earlier, not having to set the number of clusters is an
-attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.
-The problem was studied by \textcite{Giotis2006} on general graph and we compiled their results in
-\autoref{tab:cc_fixed}. On general graph and with $k$ being the number of clusters, 
-they provide PTAS for \maxa{} running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and
-for \mind{} running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
-\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}k^6\log d\right)}$.
+attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.  The
+problem was studied by \textcite{Giotis2006} on general graph and we compiled their results in
+\autoref{tab:cc_fixed}. On complete unweighted graph and with $k$ being the number of clusters, they
+provide PTAS for \maxa{} running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and for
+\mind{} running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
+\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}k^6\log d\right)}$
+(also with weight). For complete weighted graph, \textcite{WeightedMaxAPTAS08} provide a PTAS for
+\maxa{} under the condition that the ratio between the largest and smallest weights is bounded by a
+constant.
 

### 2017-08-09 notations.tex
few more abbreviations and more names in the bibliography

@@
     regardless of their direction) \\
-    $\mathcal{N}(u)$ & The set of all neighbors of $u$, regardless of edge direction. It thus holds
-    that $|\mathcal{N}(u)| = \degr(u)$ \\
+    $\nei(u)$ & The set of all neighbors of $u$, regardless of edge direction. It thus holds
+    that $|\nei(u)| = \degr(u)$ \\
     $T$ & An unweighted and undirected tree \\

### 2017-08-09 packages.tex
add pseudo code for extract star and more implementation details
add missing pseudo code and finish example execution
typos
improve description of extract star
- move the heuristics variant in a separate paragraph
- give more visibility to the proof of complexity
- streamline the pseudo code and give extra textual explanations
clarify gtx algorithm and start proving it
start writing low stretch related work
related works on spanners
describe the CC problem
add a section on CC approximation results
add an extra PTAS for maxagree
few more abbreviations and more names in the bibliography

@@
 \usepackage[citestyle=alphabetic, bibstyle=ieee-alphabetic, isbn=false, maxnames=1, minnames=1,
-sorting=nyvt, backref=true, backend=biber, defernumbers=true]{biblatex}
+sorting=nyvt, backref=true, backend=biber, defernumbers=true,maxbibnames=60]{biblatex}
 \AtEveryBibitem{
@@
    \clearfield{series}
-   % \clearfield{url}
+   \clearfield{url}
    \clearfield{urldate}
@@
 \addbibresource{more.bib}
+\addbibresource{gtx.bib}
 \usepackage{xcolor}
 \usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{mathtools}
 \usepackage{sidenotes}
@@
 \usepackage{tabulary}
+\usepackage{bm}
+\usepackage{multirow}
 \usepackage[marginpar]{todo}
 \usepackage[np,autolanguage]{numprint}
+\usepackage{listings}
+\lstset{frame=none, language=Python, tabsize=4,}
+% \usepackage{inconsolata}
+\usepackage{nicefrac}
 \input{drawing}
+
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\newcommand*\Let[2]{\State #1 $\gets$ #2}
+\def\algorithmautorefname{Algorithm}
+
 \usepackage{varioref}
@@
 \graphicspath{{./assets/}}
-\usepackage[capitalize,noabbrev]{cleveref}
+% \usepackage[capitalize,noabbrev]{cleveref}
+\usepackage[]{extdash}

### 2017-08-09 cc_approx.tex
one more approximation algo and some tiny changes

--- 2017-08-04 cc_approx.tex
@@
 derandomized while preserving its approximation guarantee~\autocite{derandomCCPivot08}). At each
-iteration, we pick a node $u$ \uar{} and we create a cluster containing $u$ and all its neighbors
+iteration, we pick a node $u$ \uar{} (called the pivot) and we create a cluster containing $u$ and all its neighbors
 linked by a positive edges. On weighted complete graphs, they tweak this algorithm by using the
@@
 \autoref{tab:cc_approx}. As mentioned earlier, not having to set the number of clusters is an
-attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.
-The problem was studied by \textcite{Giotis2006} on general graph and we compiled their results in
-\autoref{tab:cc_fixed}. On general graph and with $k$ being the number of clusters, 
-they provide PTAS for \maxa{} running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and
-for \mind{} running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
-\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}k^6\log d\right)}$.
+attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.  The
+problem was studied by \textcite{Giotis2006} on general graph and we compiled their results in
+\autoref{tab:cc_fixed}. On complete unweighted graph and with $k$ being the number of clusters, they
+provide PTAS for \maxa{}$_k$ running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and for
+\mind{}$_k$ running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
+\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}k^6\log d\right)}$
+and that can handle weighted graphs. For $k=2$, \mind$_2$ admits a faster local search method with a
+factor $2$ approximation~\autocite{Coleman2008}.
+For complete weighted graph, \textcite{WeightedMaxAPTAS08} provide a PTAS for
+\maxa{}$_k$ under the condition that the ratio between the largest and smallest weights is bounded by a
+constant.
 

### 2017-08-09 cc_sign_relation.tex
one more approximation algo and some tiny changes

@@
 given to longer cycle compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
-{measureUnbalance17} For the same reason, in virtually any real instance of the \pcc{} problem, the
+{measureUnbalance17} For the same reason, in virtually any real instance of the \pcc{} problem, the%
+% there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration
+% index) which they claimed has been peer reviewed?
 optimal solution will incur some disagreements. \Textcite[Section 5.5]{BookKleinberg2010}
@@
 some of them bearing similarities with the Chapter 3's problem.
-
-there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration index)
-which they claimed has been peer reviewed? They also mention their method to compute frustration
-exactly \url{https://arxiv.org/abs/1611.09030} along with existing work (for $n=50$)
-\url{https://doi.org/10.1016/j.socnet.2013.09.002} (leading to an 1996 heuristics for \pcc{} in
-section 3 of \url{https://doi.org/10.1016/0378-8733(95)00259-6})
 

### 2017-08-10 cc_part.tex
split CC section into several files
one more approximation algo and some tiny changes
add few more ref

--- 2017-08-04 cc_part.tex
@@
 many clusters are the data, or external constraints. This can be handled with parametrized version
-of \pcc{} \autocites{Giotis2006}{Fomin2014}\marginpars{According to many articles, \pcc{} and
-cluster editing are equivalent, but even on complete graphs or on $(V,E^+)$ I don't see why.}.
-% The cluster edit distance of a graph is the smallest number of edges to change for it to admit a
-% perfect clustering (i.e., a union of cliques).  Equivalently, it is the cost of the optimal
-% correlation clustering. [From KDDTuto14, slides 204]
+of \pcc{}.
 
@@
       % one more? https://link.springer.com/chapter/10.1007/978-3-642-33783-3_41
+      % 3D mesh segement https://link.springer.com/article/10.1007/s41095-016-0071-3
 
@@
       ones are within few percents of the optimum, provided they are followed by a local search
-      step, such as the Best One Element Move~\autocite{Gionis2007}. Another possible task would be
+      step, such as the Best One Element Move~\autocite{Gionis2007}. Further NLP tasks amenable to
+      \pcc{} formulation are referenced in their paper. Another possible task would be 
       to cluster words based on distributional embedding vector while adding antonym
@@
       monotone system. This corresponds to the \pcc{} \eqref{eq:mind} objective and allows to study
-      the complete system more easily.
+      the complete system more easily. Many more applications of \mind{} on weighted complete graphs
+      are presented in~\autocite[Section 6]{clusterEditSurvey13}.
       % \enquote{They examined dynamical systems, where a gene is modeled as a vertex and an
@@
 	    constraints that admissible clusterings have to respect, the authors have to modify in
-	    non trivial ways an existing \pcc{} algorithm. Another example is given by
+	    non trivial ways an existing \pcc{} algorithm. This was extended to weighted and partial
+	    constraints  by \url{https://doi.org/10.1007/978-3-319-10933-6_23}. Another example is given by
 	    \textcite{Crosslingual07}, who cluster together news article in different language
@@
 
-% ICML paper this year that touch something very related multicut and give recent applications in
-% vision https://arxiv.org/abs/1503.03791
+\input{cc_approx}
+\input{cc_proposed}
 
@@
 \label{sub:variants_and_extensions}
+
+\input{cc_variants}
 
@@
 improves over the one cited in the internship description
-\fi
 
@@
 \label{sub:cc_empiracal_evaluation}
+\fi

### 2017-08-12 cc/multimincut_stable.tex
put all CC material into one folder

--- 2017-08-10 cc_part.tex
@@
-\subsection{Problem setting and applications}
-\label{sub:problem_setting}
-
-Like other clustering frameworks, in \pcc{}, we are given a set of objects and we want to gather
-them into groups (called clusters) so that objects belonging to one cluster are similar to each
-other while being dissimilar to objects from all the other clusters.
-
-In \pcc{}, we formalize this problem by considering objects as the nodes of a graph $G$, whose edges
-weight encode similarity. Namely, in the most general case, for nodes $u$ and $v$, the edge between
-$u$ and $v$ is associated with two positive numbers:
-$w_{u,v}^+$ denotes the strength of the similarity between $u$ and $w$;
-$w_{u,v}^-$ denotes the strength of the dissimilarity between $u$ and $w$.
-Note however that in many applications, only one of these two numbers is non zero in which case we
-more conveniently set $w_{u,v} = \begin{cases}
-	 w_{u,v}^+ & \quad \text{if } w_{u,v}^+ > 0 \text{ and } w_{u,v}^-=0 \\
-	-w_{u,v}^- & \quad \text{if } w_{u,v}^+ = 0 \text{ and } w_{u,v}^->0 \\
-\end{cases}$
-
-Now consider a clustering \cluster{} of $V$, that is a function from $V$ to $\Nbb{}^{|V|}_{>0}$
-that assigns to each node a cluster index. For instance, $\cluster(u) = 3$ means that $u$ belongs
-to the third cluster. We can evaluate how \cluster{} fits our clustering paradigm in two ways,
-either by the number of \emph{agreements}, that is the weighted number of positive edges inside
-clusters plus the weighted number of negative edges across clusters; or by the number of
-\emph{disagreements}, that is the weighted number of negative edges inside clusters plus the
-weighted number of positive edges across clusters. Given a cost function $c$, which is usually the
-identity, \pcc{} can then be seen as graph optimization problem, either of maximizing agreements
-(\maxa{}):
-\begin{equation}
-	\max_{\cluster{}} \sum_{\cluster(u) = \cluster(v)} c(w_{uv}^+) +
-	\sum_{\cluster(u) \neq \cluster(v)} c(w_{uv}^-)
-	\label{eq:maxa}
-\end{equation}
-or minimizing disagreements (\mind{}):
-\begin{equation}
-	\min_{\cluster{}} \sum_{\cluster(u) = \cluster(v)} c(w_{uv}^-) +
-	\sum_{\cluster(u) \neq \cluster(v)} c(w_{uv}^+)
-	\label{eq:mind}
-\end{equation}
-
-Although an optimal clustering $\cluster^\star$ achieves the same value on both \eqref{eq:maxa}
-and \eqref{eq:mind}, we will see in \autoref{sub:state_of_the_art} that the latter objective is in
-some sense \enquote{easier}. Another interesting feature of the \pcc{} problem is that contrary to
-other clustering formulations, it does not require us to set the number of clusters $k$ beforehand.
-Instead, $k$ emerges as a natural property of the solution. Since clustering is an unsupervised
-problem, this is generally handy. However, in some situations, we may have prior knowledge on how
-many clusters are the data, or external constraints. This can be handled with parametrized version
-of \pcc{}.
-
-In \autoref{fig:cc_objectives}, we show a simple instance of \pcc{} and one of its optimal solution.
-\begin{figure}[hbt]
-	\centering
-	\includegraphics[width=0.8\linewidth]{assets/tikz/cc_objectives_tikz.pdf}
-	\caption[Small example of \pcc{}]{A small graph with eight nodes and ten edges. Solid edges
-	represent positive edges and dashed edges represent negative edges. A clustering \cluster{}
-	is showed with 3 clusters: $\{1, 2, 4\}$, $\{3, 5, 6\}$ and $\{7, 8\}$. \cluster{} incurs two
-	disagreements: the negative edge between nodes $1$ and $2$ within the blue cluster, and the
-	positive edge $6,7$ between the orange and green clusters. Those disagreements are created
-	by two cycles with one negative edge and thus cannot be avoided, meaning that \cluster{} is
-	optimal. However it is not the unique solution, for instance merging the orange and green
-	clusters would also yields two disagreements.}
-	\label{fig:cc_objectives}
-\end{figure}
-
-\subsubsection{Applications}
-\label{ssub:cc_applications}
-
-According to \textcite[Section 5]{Demaine2006}, \pcc{} is well suited to several situations:
-\begin{itemize}
-   \item when the items to be clustered do not belong to a natural metric space (preventing
-      approaches such as $k$-means) but we still know for some pairs whether they are similar or
-      not.
-   \item when we do not know the number of clusters beforehand but we have a similarity measure. In
-      that case, we can select a problem-specific similarity threshold and set all edges with a
-      similarity larger than the threshold to be positive while the others are set to negative.
-   \item when we have a classic clustering problem (that is a set of objects, a distance between
-      them and an objective function to minimize) with additional pairwise constraints of the form
-      \emph{must-link/cannot-link}. Instead of restraining a clustering algorithm to the space of
-      feasible solution, we convert the distances between objects and the constraints into signed
-      edges and solve the resulting \pcc{} problem. 
-\end{itemize}
-
-In addition to these general considerations, \pcc{} has also been used in several domains:
-%TODO turn those items into paragraphs?
-\begin{description}
-   \item[Computer Vision]
-      The ubiquitous task of segmenting an image into homogeneous regions is prerequisite for many
-      further processing, and can be tackled with \pcc{}. For instance, to segment cell in
-      microscopy imagery, \textcite{CellSeg14} first use generic image features to classify pixel in
-      belonging to region boundaries or not. Then, they extract small scale regions called
-      superpixels. After building the adjacency graph of these superpixels, they assigned edge
-      weight by averaging the boundary probability over the pixels separating the adjacent
-      superpixels. They also add strong negative constraints between distant superpixels, and lastly
-      cluster these superpixels according to the \pcc{} objective to obtain the final segmentation.
-      A similar approach was used earlier in~\textcite{Kim2011}, which stressed the importance of
-      considering such higher order constraints between distant superpixels in order to achieve good
-      performance. This was also extended to 3D segmentation~\autocite{VolumeSegmentation12}, where
-      additional tunings allowed to segment a volume image of a mouse cortex with up to billions
-      voxels. \Textcite{Beier2015} segment 2D and 3D images with a energy based formulation of
-      \pcc{} and iteratively improve their solution by merging it with another clustering given by a
-      proposal generator. By developing another scalable energy based optimization procedure, and
-      with the help of few user provided cues, \textcite{Bagon2011} were able to apply \pcc{}
-      directly at the pixel level.
-      % one more? https://link.springer.com/chapter/10.1007/978-3-642-33783-3_41
-      % 3D mesh segement https://link.springer.com/article/10.1007/s41095-016-0071-3
-
-      Beyond image segmentation, \textcite{Shape3D17} develop a method to extract a network of
-      descriptive curves from 3D shapes.  After an initial stage of generating many such
-      \emph{flowlines}, they describe in Section 6 a \pcc{} formulation to extract \emph{reliable}
-      representative flowlines, using geometric constraints to obtain positive or negative cues that
-      two flowlines are from the same reliable representative.
-
-      Finally, in order to track several targets across sequential video frames,
-      \textcite{multiTracking15} propose a multistage framework. One step revolves around a matrix
-      $A$ that defines the cost of assigning an object tracked in previous frames to an object
-      detected in the current frame. This matrix is turned into a symmetric affinity matrix
-      $\bar{A}_{sym}$ that can be seen as a signed graph. \pcc{} is then used to extract clusters
-      (called zones), in which local processing is performed, whose complexity can be adapted to the
-      difficulty of each zone.
-
-   \item[Natural Language Processing] 
-      Coreference resolution is the task of finding all expressions that refer to the same entity in
-      a text. Like image segmentation, it is a preprocessing step that can later be used in document
-      summarization, question answering, and information extraction. \Textcite[Section
-      2.3]{graphicalCoreference04} tackle coreference resolution by using a undirected graphical
-      model on which performing inference is equivalent to \pcc{}. On small scale instances,
-      \textcite{Elsner2009} first obtain an upper bound of the optimal solution by solving a SDP
-      relaxation of the problem. They then compared various heuristics and show that best performing
-      ones are within few percents of the optimum, provided they are followed by a local search
-      step, such as the Best One Element Move~\autocite{Gionis2007}. Further NLP tasks amenable to
-      \pcc{} formulation are referenced in their paper. Another possible task would be 
-      to cluster words based on distributional embedding vector while adding antonym
-      constraints~\autocite{SignedWordRatings}, although the authors of that paper choose to use
-      signed spectral clustering.
-
-   \item[Biology]
-      The input is a similarity matrix between gene expressions in various experimental conditions
-      and the goal is to cluster those genes into groups which react similarly. \Textcite[Section
-      4]{Ben-Dor99} gives three examples: 112 genes involved in the rat central nervous system, 1246
-      genes of the roundworm \emph{C. elegans} and 2000 human genes obtained from 40 tumor and 22
-      normal colon samples. \Textcite{Mason2009} analyze a signed co-expression networks of genes
-      involved in embryonic stem cells to find which genes are related to pluripotency (the ability
-      to differentiate into any type of cell in the body) or self-renewal (the ability to replicate
-      indefinitely). Another application is to study the variation of one individual
-      DNA~\autocite{Das2015}. In the human organism, chromosomes are organized in pair, and both
-      chromosomes of a pair have similar but not identical DNA sequences. This is mostly because of
-      single nucleotide polymorphisms (SNPs), where a single base differs between the two DNA
-      sequences, leading to different alleles of the corresponding gene. An haplotype is the list of
-      all alleles at a contiguous sites in a region of a single chromosome and this information is
-      used in several medical applications. The high-throughput sequencing of one individual genome
-      yields many short \emph{reads} that provides information about the order of nucleotides in a
-      fragment of one chromosome and that can be used to assemble haplotypes. Namely the authors
-      build a graph of reads and define a similarity function between reads to assign weight on the
-      edges. The clusters of that graph correspond to haplotypes, and are obtained by a SDP
-      relaxation of the \pcc{} objective. \Textcite{monotoneBiology07} also consider graphs whose
-      nodes are genes, but in a different context. In this case, positive edges represent an
-      activating connection, while negative edges represent inhibiting connection. They also define
-      a \emph{monotone system} as a balanced subgraph, that is a subgraph which does not contain a
-      cycle with an odd number of edges. Such monotone system are stable, in the sense that
-      modifying the concentration of one gene will have a predictable effect, even ignoring the
-      precise kinematics of the chemical reactions involved. Their goal is to find the minimum
-      number of edges to remove in order to decompose a dynamics system into a collection of
-      monotone system. This corresponds to the \pcc{} \eqref{eq:mind} objective and allows to study
-      the complete system more easily. Many more applications of \mind{} on weighted complete graphs
-      are presented in~\autocite[Section 6]{clusterEditSurvey13}.
-      % \enquote{They examined dynamical systems, where a gene is modeled as a vertex and an
-      % activating connection is modeled as a positive edge and an inhibiting connection is modeled as
-      % a negative edge. The claim is that biological dynamical systems are close to being balanced,
-      % and that finding a minimum set of edges to delete to make the graph balanced can be used to
-      % decompose the graph into “monotone subsystems”, which exhibit stable behavior and thus allow a
-      % better understanding of the dynamics of a system.}\autocite{monotoneBiology07}
-
-   \item[Network science]
-      As mentioned in the first Chapter, one early use of signed graphs was to model social
-      interactions and given the large body of work on community
-      detection~\autocite{FortunatoSurvey10}, it is therefore not surprising that \pcc{} has been
-      used to study such networks. For instance, one can extract all the votes of the member of
-      political parliament and form a graph whose node are politicians and edge weight quantify how
-      much they agree or disagree on various issues they have been voting on. This can be used to
-      study various social science questions such as loyalty, leadership, coalitions, political
-      crisis and polarization. It has been applied to the European
-      parliament~\autocite{Mendonca2015}, Slovenian parliament~\autocite{Jiang2015} and the
-      Brazilian parliament~\autocites{BrazilCC17}. This can also be used at international level. For
-      instance, by considering a dataset of military alliances and disputes, \textcite{Traag2009}
-      cluster countries into blocks that resemble those identified by Huntington in his \emph{Clash
-      of Civilizations} book. Another source of data is the vote on resolutions during the United
-      Nations General Assembly~\autocite{CommunityUN12}. Finally, one can also study how to exploit
-      the information contained within the negative links to enhance the visualization of social
-      networks~\autocite{Luca10}. Note however that those three last papers not explicitly use a
-      \pcc{} formulation.
-
-   \item[Others]
-      \begin{itemize}[leftmargin=*]
-	 \item
-	    Deduplication, also called duplicate detection or entity resolution, is the process
-	    of identifying objects from a real-world, noisy database that refer to the same entity.
-	    On a high level, a solution to this problem is to build a graph of all the available
-	    objects, define a similarity between them and run a \pcc{} algorithm. This was indeed
-	    this kind of problems at Whizbang! Labs that partly motivated one of the early \pcc{}
-	    paper~\autocite{Bansal2002}. The main challenge
-	    thus lies in devising an appropriate similarity measure, given that object can have very
-	    different features from one database to another. \Textcite{LargeScaleDeDup09} propose a
-	    declarative language, expressing both hard constraints (that have to be satisfied) and
-	    soft constraints (that can be seen as cues guiding the process). Because of these hard
-	    constraints that admissible clusterings have to respect, the authors have to modify in
-	    non trivial ways an existing \pcc{} algorithm. This was extended to weighted and partial
-	    constraints  by \url{https://doi.org/10.1007/978-3-319-10933-6_23}. Another example is given by
-	    \textcite{Crosslingual07}, who cluster together news article in different language
-	    covering the same event. \pcc{} was also evaluated among other
-	    solutions to that problem by \textcite{DeDup09}, who note that their non optimized
-	    implementation does not perform the best.
-	 \item 
-	    Given an electrical circuit layout, \textcite{circuitDesign07} extract a graph of its
-	    components (called shifter) that must be assigned one of two possible phases. Because
-	    two shifters next to some specific shape must be in opposite phase and two shifters
-	    separated by less than a specified distance must be of the same phase, the authors look
-	    for a two-clustering of the nodes that will minimize the number of disagreements.
-      \end{itemize}
-\end{description}
-
-\subsection{Relation with edge sign prediction}
-\label{sub:relation_with_edge_sign_prediction}
-
-\input{cc_sign_relation}
-
-\subsection{State of the art}
-\label{sub:state_of_the_art}
-
-\input{cc_approx}
-\input{cc_proposed}
-
-\subsection{Variants and extensions}
-\label{sub:variants_and_extensions}
-
-\input{cc_variants}
-
-\subsubsection{\pcc{} under stability assumption}
-\label{ssub:cc_under_stability_assumption}
-
-\iffalse
 Haris Angelidakis, Konstantin Makarychev, and Yury Makarychev. 2017.
@@
 improves over the one cited in the internship description
-
-\subsubsection{Parallel \pcc{}}
-\label{ssub:parallel_cc}
-
-\subsection{Empirical evaluation?}
-\label{sub:cc_empiracal_evaluation}
-\fi

### 2017-08-12 cc/problem_applications.tex
put all CC material into one folder

--- 2017-08-10 cc_part.tex
@@
-\subsection{Problem setting and applications}
-\label{sub:problem_setting}
-
 Like other clustering frameworks, in \pcc{}, we are given a set of objects and we want to gather
@@
 \end{description}
-
-\subsection{Relation with edge sign prediction}
-\label{sub:relation_with_edge_sign_prediction}
-
-\input{cc_sign_relation}
-
-\subsection{State of the art}
-\label{sub:state_of_the_art}
-
-\input{cc_approx}
-\input{cc_proposed}
-
-\subsection{Variants and extensions}
-\label{sub:variants_and_extensions}
-
-\input{cc_variants}
-
-\subsubsection{\pcc{} under stability assumption}
-\label{ssub:cc_under_stability_assumption}
-
-\iffalse
-Haris Angelidakis, Konstantin Makarychev, and Yury Makarychev. 2017.
-Algorithms for Stable and Perturbation-Resilient Problems. STOC’17
-\href{http://ttic.uchicago.edu/~yury/papers/two-stable.pdf}{10.1145/3055399.3055487}
-improves over the one cited in the internship description
-
-\subsubsection{Parallel \pcc{}}
-\label{ssub:parallel_cc}
-
-\subsection{Empirical evaluation?}
-\label{sub:cc_empiracal_evaluation}
-\fi

### 2017-08-12 cc/proposed.tex
describe proposed solutions to CC

We already described methods based on LP and SDP, as they have proved to be the best way to obtain
tight approximation in the worst case. Here we survey other kind of approaches with different aims.

\paragraph{Exact methods} Because of the complexity of the \pcc{} we presented earlier, one has to
rely on approximations to solve large instances of the problem. However, we can imagine offline
signed social networks with only few nodes, in which case it is reasonable to expect finding the
optimal clustering. Furthermore, this can also be useful to evaluate in practice the quality of
heuristic methods, albeit in non asymptotic setting. The most straightforward idea is to use the
formulation of the LP \eqref{eq:mindLP} with the additional constraints that all variable are
binary. \Textcite{ExactMIP13} solve this binary integer program on random instances and show that
depending of the negative edge density, the FICO Xpress solver starts to be unable to finish within
one hour time limit when $n \geq 40$.
\Textcite{Aref2016} describe four linear and quadratic  binary integer problems
to model \pcc{} with two clusters along with some preprocessing optimizations. With the Gurobi
software, they solve instances with up to $3200$ edges in less than a second. They also sketch
extensions to weighted graph and more than two clusters.
In a different direction, \textcite{Berg2015}, encode the linear and quadratic integer formulations
into weighted MaxSAT instances and use the state of the art solver
MaxHS\footnote{\href{http://www.maxhs.org/}{http://maxhs.org}}~\autocite{SATSolver13} to get the
exact solution on instances with at most $1000$ nodes in less than a few hours.

After defining the matrix $A$ by $A_{uv} = w_{uv}^+ - w_{uv}^-$, \textcite{LowRank16} show that the
\maxa{} problem can be written up to constant factor by defining one vector $x_u$ per node as:
\begin{align}
   \label{eq:cc_lowrank}
   \text{maximize } & \sum_{u<v} A_{uv}x_u^Tx_v \\
   \text{subject to}& \quad x_{uw} \in \{e_1,\ldots, e_n\} \nonumber
\end{align}
where $e_i$ are the canonical vectors of $\Rbb{}^n$. They show that when the matrix $A$ is positive
semidefinite of rank $k$, the \pcc{} problem can be solved exactly in $O(n^{k^2})$ time. More
practically, they also give an algorithm that closely approximate the objective of
\eqref{eq:cc_lowrank} in $O(nk)$ time.

\paragraph{Physics inspired energy methods}

The Potts model~\autocite{PottsSurvey82} describes a general model of \emph{spins} organized in a
lattice and being in one of $k$ possible states. A spin $u$ interacts with each of its neighbors $v$
through a coupling $J_{uv}$. The energy of this system, called the Hamiltonian, is defined by
$\mathcal{H}(\bm{\sigma}) = -\sum_{u,v} J_{uv} \delta(\sigma_u, \sigma_v)$ where $\bm{\sigma}$ is
the spin configuration (that is, $\sigma_u \in \{1,\ldots,k\}\,\forall u$) and $\delta(\sigma_u,
\sigma_v)$ is equal to one if $u$ and $v$ are in the same state and zero otherwise. It is a general
principle of physics that isolated systems tend to minimize their energy, which in this case amounts
to finding a spin configuration minimizing the Hamiltonian. Viewing the spins as nodes of a graph,
the couplings as the graph weighted edges and the $k$ possible states as clusters, it is quite
natural to formulate the clustering problem as a Hamiltonian minimization
problem~\autocite{CommunityPhysics06}. Letting $A$ be the matrix such that $A_{uv} = w_{uv}^+ -
w_{uv}^-$ on a general directed graph, \textcite{Traag2009} reward positive and absent negative
links within cluster and penalize negative and absent positive links across clusters to come up with
the following Hamiltonian: $\mathcal{H}(\bm{\sigma}) = -\sum_{u,v} \left(A_{uv}-(\gamma^+p^+_{uv} -
\gamma^-p^-_{uv})\right) \delta(\sigma_u, \sigma_v)$ where $\gamma^+$ and $\gamma^-$ are user
parameters and $p^\pm_{uv}$ are null model probabilities, which are equal to $p^\pm_{uv} =
\frac{|E^\pm|}{|V|(|V|-1)}$, or $p^\pm_{uv}=\frac{\dout^\pm(u) \din^\pm(v)}{|E^\pm|}$ in the degree
corrected model. They note that setting $\gamma^+ = 0 = \gamma^-$ make minimizing the Hamiltonian
equivalent to the \mind{} problem, which they do by simulated
annealing~\autocite{SimulatedAnnealing83}. The idea is to start from a random partition, and jump to
another partition by moving a single node from one cluster to another. Such moves are made with a
probability proportional to how each move reduces the Hamiltonian and as the procedure goes on, large
jumps are made less and less probable by reducing a parameter called the temperature.
One advantage of this energy formulation is that it requires only one variable per node instead of
one variable per edge as in the case of the linear program.

In the case of $2$-\pcc{}, \textcite{Facchetti2011isingmodel} rewrite the Hamiltonian as
$\mathcal{H}(\bm{\sigma}) = -\frac{1}{2}\bm{\sigma}^T A \bm{\sigma} = -\frac{1}{2}\bm{1}^TT_\sigma A
T_\sigma\bm{1}$. There $T_\sigma = diag(\bm{\sigma})$ (where $\bm{\sigma} \in \{0,1\}^n$) is the
outcome of a local search algorithm such that $A_\sigma = T_\sigma A T_\sigma$ has the same number
of disagreements as $A$ but the least number of negative edges. This is called a gauge
transformation in the spin glass literature and the benefit of that heuristic is that it scales
gracefully to large graphs.
\Textcite{Bagon2011} also write the \maxa{} objective as a Potts model, and show that it can be
interpreted as the log posterior of a partition matrix under a simple generative model and as a
pair-wise conditional random field energy without unary terms. This allows them to adapt existing
discrete energy optimization algorithms in order to cope with the following three challenges of the
\pcc{} energy: \enquote{(i) the energy is non sub-modular, (ii) the number of clusters is not known
in advance, and (iii) there is no unary term}. Doing so, the are able to handle large problems with
more than 100K nodes.  Also adopting an energy minimization approach, \textcite{Kappes2016} assign a
probability to each cut of a signed graph proportional to the exponential of the number of
disagreements of that cut. They also develop efficient cut sampling methods.

% something about local minima when defining another energy function, and how it is achieved by a
% dynamic model but to be honest it has little to do with clustering {Marvel2009landscape}
% Kappes2016 \enquote{Furthermore, due to the lack of an external field (unary terms), any permutation of an
% optimal assignment results in another optimal labeling.}
% like an ICML'17 paper that touches something very related (multicut) and gives recent applications in
% vision https://arxiv.org/abs/1503.03791 represents a cut by the set of interclusters edges

\paragraph{Distributed setting}

To handle the massive size of some large real world dataset, parallelizing existing algorithms is a
natural approach. Due to its simplicity and approximation guarantee, \ccpivot{} has been adapted
three times for that purpose. First, \textcite{Chierichetti2014} describe how to uniformly sample
several pivots in the same round, remove pivots that are adjacent through positive edges and then
grow the corresponding clusters with potential conflicts solved according to the node order in a
global permutation drawn at the beginning of the algorithm. On general graphs, this requires $O(\log
n\diam(G))$ rounds\marginpars{maybe I could adapt the proof of lemma 1 to prove the number of rounds
of \gtx{}, although this is quite probabilist.}, which on complete graphs reduces to $O(\log n)$.
Furthermore, this almost preserves the approximation factor, which is
$3+\frac{14\epsilon}{1-7\epsilon}$, where $\epsilon$ is a parameter smaller than $1$. Finally, this
allows to experimentally cluster graphs with millions of nodes and billions of edges. Second,
\textcite{ParallelCCNIPS15} describe an equivalent version of \ccpivot{} where a permutation $\pi$
of the nodes is drawn at the start of the algorithm and the pivots are chosen sequentially in the
order of $\pi$. The exact same partition can be obtained when several pivots are chosen at the same
time by different threads as long as they respect two concurrency rules: (i) two nodes $u$ and $v$
can become pivot at the same time if they are not connected with a positive edge, otherwise only the
one with the smallest index in $\pi$ becomes pivot; (ii) if $w$ is a positive neighbor of two pivots
$u$ and $v$, it is affected to the cluster of the pivot with the smallest index in $\pi$. Enforcing
these rules preserves the factor $3$ approximation and the algorithms terminate after $O(\log
n\diam(G))$ rounds. The authors also present a version where each round is faster as it does not
enforce rule (i). However, this weaken the approximation guarantee to $(3 + \epsilon)OPT +
O(\epsilon n\log^2 n)$. In practice, the solution is very close the one of \ccpivot{}, although it
degrades as the number of threads increases. The number of rounds preserving the $3$-approximation
is lowered to $O(\log\log n)$ by \textcite{Ahn2015}. They also present results obtained in a single
pass, which corresponds to the streaming model: the algorithm receives the edges of $G$ one by one
and upon seeing the last one outputs its result. The additional constraint is that this algorithm
can only use $O(n\polylog n)$ space. In this setting, the authors show a polynomial time
$(1-\epsilon)$-approximation of \maxa{} if the weights are bounded (and $0.766(1-\epsilon)$ if the
weights are arbitrary); and an $O(\log |E^-|)$ approximation of \mind{} in polynomial time with arbitrary
weights. This is done by combining graph sketching and a method to solve convex programs in a space
efficient manner.

\Textcite{Bonchi2013} suggest a different paradigm to solve \pcc{}, that can be applied in a
distributed setting to obtain a scalable approach. Namely, given a node $u$, they want to output a
globally consistent cluster index $\cluster(u)$ while making at most $t$ queries to an sign oracle.
Here $t$ is a parameter that depends on the quality of clustering produced but not on the size of
the graph. And because this procedure is local to each node, it can be run in parallel. Finally, one
can get a full clustering by computing $\cluster(u)$ for all the nodes in the graph. Despite the
problem being apparently more challenging, they obtain approximation factors that are close to the
best known (which in this model would make $\Omega(n^2)$ total queries). More precisely, they use
two techniques.
The first one is inspired by \ccpivot{} and first find a good set of pivots by seeing the problem as
independent set on a sampled part of the positive graph, then for a given node, find the closest
such pivot or create a singleton. Given a quality parameter $\epsilon\in(0,1)$, it yields a $4\cdot
OPT + \epsilon n^2$ approximation of \mind{} requiring $O(\frac{n}{\epsilon})$ time and
queries~\autocite[Theorem 3.3]{Bonchi2013}
Roughly stated, the second one relies on an existing low-rank approximation of the adjacency matrix,
that partition the graph into similar sized classes such that edges between those classes behave as
in a random graph. This initial partition is \enquote{coarsened} into a good clustering by
considering all possible ways of assigning classes to clusters. It gives an $OPT + \epsilon n^2$
additive approximation  for \mind{} that runs in time $n \cdot poly(\nicefrac{1}{\epsilon}) +
2^{poly(\nicefrac{1}{\epsilon})}$ \autocite[Corollary 3.7]{Bonchi2013}.

\paragraph{Spectral Clustering}

A classic method for clustering graphs is to leverage their spectral properties. Namely, if $A$ is
the adjacency matrix of $G$ and $D$ its degree diagonal matrix (that is $D_{u,u} = \degr(u)$), the
\emph{Laplacian} of $G$, defined by $L_G = D - A$, is a symmetric positive semidefinite matrix. As
such, it has $n$ real non-negative eigenvalues, and its spectrum provides additional information on
the connectivity of $G$. For instance, $0$ is always the smallest eigenvalue and its multiplicity is
equal to number of connected components of $G$, while ---if $G$ is connected--- the second
eigenvalue is the algebraic connectivity of $G$, whose magnitude is an indication of how well
connected is the graph. This matrix is typically used for clustering by computing its first $k$
eigenvectors, which embeds the $n$ nodes of $G$ in $\Rbb^k$, where there are then clustered with the
$k$-means algorithm. This can be seen as a relaxation of the discrete \rcut{} objective, which asks
for the partition $\{C_1, \ldots, C_k\}$ minimizing $\frac{1}{2}\sum_{i=1}^k \frac{\mathrm{cut}(C_i,
\bar{C_i})}{|C_i|}$, where $\bar{C_i}$ is the complement of $C_i$ in $V$ and $\mathrm{cut}(B, C) =
\sum_{u\in B, v\in C} w_{uv}$ is the total weight of the edges between $B$ and
$C$~\autocite{tutoSpectralClustering07}. By considering the symmetric normalized Laplacian
$L_{sym}  = D^{\nicefrac{1}{2}}LD^{\nicefrac{1}{2}}$, it is also possible to approximate the
normalized cut objective (\ncut{}), where $|C_i|$ is replaced by $vol(C_i) = \sum_{u\in C_i}
\degr(u)$. We will now see how these kinds of approaches can be extended to signed graphs, noting
first that they require to fix the number of clusters beforehand and are looking for clusters
balanced in size, which makes the problem related but not equivalent to \pcc{}.

The first line of research consider only \mind{}$_2$. For instance, \textcite{NcutAnd2CC08} show
that both normalized cut and \mind{}$_2$ objectives can be written as a SDP (or equivalently as
eigenvalue problems) and thus combined, the intuition being that we look for \ncut{} solutions whose
number of disagreements is not too much more than the approximate optimal.

Letting \ncut{}$(C, \bar{C}) = \frac{\mathrm{cut}(C, \bar{C})}{\mathrm{bal}(C)}$ with
$\mathrm{bal}(C) = 2\frac{vol(C)vol(\bar{C})}{vol(V)}$, \textcite{mOneCC12} define a new objective:
\begin{equation*}
  \hat{F}_\gamma(C) = \frac{\mathrm{cut}(C, \bar{C}) + \gamma\left(\hat{M}(C)+\hat{N}(C)\right)}{\mathrm{bal}(C)}
\end{equation*}
where $\gamma \in \Rbb{}_+$ is a parameter, while $\hat{M}(C)$ and $\hat{N}(C)$ are respectively the
number of positive and negative disagreements of the $(C, \bar{C})$ clustering.
They show how to optimize a tight continuous relaxation of $\hat{F}_\gamma$ as the non negative
ratio of a difference of convex function and a convex function.

On the other hand, one can also adapt these two cut objectives to directly include negative edges.
\Textcite{Luca10} define the signed Laplacian as $\bar{L} = \bar{D} - A$, where $\bar{D}$ is the
signed degree matrix such that $\bar{D}_{uu} = \sum_{v\in \nei(u)} |A_{uv}|$, as well as a signed
variant of the symmetric normalized Laplacian $\bar{L}_{sym}  = \bar{D}^{\nicefrac{1}{2}} L
\bar{D}^{\nicefrac{1}{2}}$. They show that the signed Laplacian is positive semidefinite, and
even positive-definite as soon as the graph is unbalanced (that is contains a cycle with a odd
number of negative edges). From positive and negative cuts defined as $\mathrm{cut}^+(B, C) =
\sum_{u\in B, v\in C} w^+_{uv}$ and $\mathrm{cut}^-(B, C) = \sum_{u\in B, v\in C} w^-_{uv}$, a
natural signed cut is $\mathrm{scut}(B, C) = 2\mathrm{cut}^+(B, C) + \mathrm{cut}^-(B, B) +
\mathrm{cut}^-(C, C)$ which then be used to defined signed \rcut{} and \ncut{}. Arguing that those
definitions force negatively linked nodes to be symmetric around the origin, do not take into
account the balance of negative edges in each cluster and are difficult to extend to more than two
clusters, \textcite{SignedEmbedding15} instead propose two new normalized cuts:
\begin{align*}
  SNScut(C_1, \ldots, C_k) &= \sum_{i=1}^k
  \frac{\mathrm{cut}^+(C_i, \bar{C_i})-\mathrm{cut}^-(C_i, \bar{C_i})}{vol(C_i)}  \\
  BNScut(C_1, \ldots, C_k) &= \sum_{i=1}^k
  \frac{\mathrm{cut}^+(C_i, \bar{C_i})-\mathrm{cut}^-(C_i, \bar{C_i})+vol^-(C_i)}{vol(C_i)}
\end{align*}

Noting that if $x_i\in\Rbb^n$ is the vector indicator of cluster $C_i$ (that is the \uth{} entry of
$x_i$ is $1$ is $u$ belongs to $C_i$ and $0$ otherwise), $x_i^T\bar{L}x_i = 2\mathrm{cut}^-(C_i,
C_i) + \mathrm{cut}^-(C_i, \bar{C}_i) + \mathrm{cut}^+(C_i, \bar{C}_i)$,
\textcite{mSemanticWordCC17} introduce the following cut objective:
\begin{equation*}
  sNcut(C_1, \ldots, C_k) = \sum_{i=1}^k
  \frac{2\mathrm{cut}^-(C_i, C_i) + \mathrm{cut}^-(C_i, \bar{C}_i) + \mathrm{cut}^+(C_i, \bar{C}_i)}{vol(C_i)}
\end{equation*}

% Kai-Yang Chiang, Joyce Jiyoung Whang, and Inderjit S. Dhillon. 2012. Scalable clustering of signed
% networks using balance normalized cut. In CIKM '12. 615-624.
% DOI=http://dx.doi.org/10.1145/2396761.2396841

Finally, \textcite{mGeometricMean16} show that the Laplacians defined so far can be seen as
arithmetic means of the Laplacian $L^+$ of the positive subgraph $G^+=(V, E^+)$ and the signless
Laplacian $Q^-$ of the negative subgraph $G^-=(V, E^-)$, where $Q^- = D^- + A^-$. They suggest
instead to use a geometric mean defined for two positive matrices $A$ and $B$ as $A\#B =
A^{\shalf{}} \left(A^{-\shalf{}} BA^{-\shalf{}} \right)^{\shalf{}} A^{\shalf{}}$. This suggestion is
based on the fact that if $u$ is a common eigenvector of both $A$ and $B$ with eigenvalue $\lambda$
and $\mu$ respectively, then $u$ is an eigenvector of $A+B$ with eigenvalue $\lambda+\mu$ and an
eigenvector of $A\#B$ with eigenvalue $\sqrt{\lambda\mu}$. Therefore, the $k$ smallest eigenvalues
of the geometric mean Laplacian will be influenced by both smallest eigenvalues of $L^+$
(corresponding to assortative clusters in $G^+$) and of $Q^-$ (corresponding to disassortative
clusters in $G^-$), while this is note the case for the arithmetic mean of Laplacians.

\paragraph{Heuristic methods}

While all\marginpars{or most? not physic ones so move them afterwards} the methods we discuss so far
are either exact or come with some approximation guarantees, practitioners have also develop
approaches that are designed to efficiently reach a solution that is satisfying enough for the
application at hand. For instance, while studying small scale signed social networks,
\textcite{Early96} describe the following procedure. Start with a random initial $k$ clustering of
the graphs and for $T$ steps, sample randomly $P$ neighboring partitions, compute their number of
disagreements and move the one with the least disagreements. Neighboring partitions are obtained
either by moving one node from cluster to another or by exchanging a pair of nodes between two
clusters. The overall complexity is $O(nTP)$. This is quite similar to the \emph{Best One Element
Move} described in~\autocite{Gionis2007}, except that the exchange operation is replaced by moving
one node to its own singleton cluster. The \emph{Cluster Affinity Search Technique}
algorithm~\textcite{Ben-Dor99} instead grows clusters one by one by maintaining for the current
cluster the affinity of all nodes, which is the sum of the weight between that node and the nodes in
the cluster. Nodes above a certain threshold are added to the current cluster and nodes below the
threshold are removed, with the affinity to the current cluster being recomputed after each addition
or deletion. After defining the net weight of an edge to be $w^\pm_{uv} = w^+_{uv} - w^-_{uv}$,
\textcite{Elsner2009} describe three folklore heuristics that start with empty clusters and add node
one by one: \enquote{The \textsc{Best} algorithm adds each node $u$ to the cluster with the
strongest $w^\pm$ connecting to $u$, or to a new singleton if none of the $w^\pm$ are positive.
The \textsc{First} algorithm adds each node $u$ to the cluster containing the most recently
considered node $v$ with $w^\pm_{uv} > 0$. The \textsc{Vote} algorithm adds each node to the cluster
that minimizes the \pcc{} clustering objective, \ie{} to the cluster maximizing the total net weight
or to a singleton if no total is positive.} Empirically, \textsc{Vote} turns out to be the best.
Among other related heuristics, \textcite{mergingHeuristics14} describe the random maximum merging
algorithm, that starts with singleton clusters and keep merging two clusters chosen at random among
those whose merge would result in the maximum improvement of the score function.  This runs in
$O(n^2\log n)$, provides a $2$-approximation of \maxa{}, a $O(n\times \text{size of the largest
cluster})$-approximation of \mind and empirically results in less disagreements than \ccpivot{}.
Building upon their previous GRASP work~\autocite{GRASP13}, an iterated local search heuristic is
presented in~\autocites{Levorato2015}{Levorato2017}. Each iteration of this algorithm starts by
greedily build a clustering in fashion similar to \textsc{Vote}, albeit with more randomness in the
node ordering. This clustering is locally improved by moving blocks of $r\in\{1,2\}$ nodes from one
cluster to another as long as the number of disagreements decreases, a phase called neighborhood
descent. The algorithm then enters an inner loop where the current clustering is perturbed by $t$
random one-node-moves and updated if a subsequent neighborhood descent can improve it compared with
before the perturbation. The authors note that both the outer loop and the neighborhood descent can
be run in parallel, which allow them to process a \np{10000} nodes graph on 10 cores in around 700
seconds.
This local search followed by random perturbation is reminiscent of genetic algorithms, and such an
approach is sketched in~\autocite{GeneticCC08}.\marginpars{but that's a terrible paper!} Instead of
moving nodes, \textcite{restoreNeighborhood13} starts from the observation that in a balanced
graphs, $G^+$ is a disjoint union of cliques in which all node a given cluster share the shame
neighbors. Their algorithm is initialized with the set $E_s^+$ of edges where $w^+_{uv} > w^-_{uv}$,
and repeatedly sample an edge $(u,v)$ from $E_s^+$ before trying to make the neighborhoods of $u$
and $v$ coincide by adding or removing edges in $E_s^+$.  Yet another idea is modify the LP
\eqref{eq:mindLP} to replace the binary variable $x_{uv}$ by a $k$ clusters indicator matrix
$L\in\Rbb^{k\times n}$ where $L_{iu}$ is $1$ is $u\in C_i$ and $-1$
otherwise~\autocite{AltOptimLP13}. Indeed $x_{u\cdot} \equiv L_{\cluster(u)\cdot}$, where $\cluster$
is the cluster assignment. By relaxing the integer constraints on $L$, it is possible to do
alternate optimization on $L$ and $\cluster$.

### 2017-08-12 cc/variants.tex
describe some variants of CC

\paragraph{non binary edge labelling}

In the so called \textsc{Chromatic}-\pcc{} setting, \enquote{positive} edges are now associated with
one of $L$ possible color and the goal is to form clusters mostly made up of edges with the one same
color. Namely, a disagreement is now a negative edge between clusters or an within-cluster edge
whose color differs from the majority color of that cluster. This is motivated by edge-labeled graph
in social networks, biology and citation networks and will discuss such applications in
\autoref{chap:vector}. As a generalisation of \pcc{}, it is \NPc{} but \textcite{Bonchi2012a}
present a modification of the \ccpivot{} algorithm that pick edges instead of nodes as pivots, and
grow clusters by adding monochromatic triangle. This gives an approximation factor of six times the
maximum degree of the graph. They also present a method when the number $k$ of clusters is fixed
beforehand, starting with an initial partition and then alternating between finding the majority
color of clusters and finding better clusters. Unfortunately, the maximum degree of a graph can be
as large as $n$. However, \textcite{Anava2015} present constant factor approximations. Namely, they
show the problem can be reduce to classical \pcc{} by setting all edges incident to a node $u$ to
negative if they are not of the majority color of $u$. They then apply the regular \ccpivot{} and
show this gives a $11$-approximation to the original problem. Furthermore, they also write a linear
program and round it using the region growing technique of \textcite{Charikar2003} to obtain
approximation factor of $4$. \Textcite{multiChromatic15} extend their work to the case were a single
edge can carry a \emph{set} of labels and adapt their randomized algorithm so that the approximation
factor is multiplied by the size of the input label set.

\paragraph{Overlapping \pcc{}}

While in \pcc{}, each node is assigned to a single cluster, in other settings we may want to relax
this constraint. Given a complete weighted graph, \textcite{Bonchi2012} want to output a clustering
\cluster{} that minimizes the following cost: \[ \sum_{(u,v)\in E} \left| H(\cluster(u),
\cluster(v)) - w_{uv}\right|\] where $H$ is a similarity function between two sets of labels, chosen
here to be the Jaccard similarity or a $0/1$ indicator of non empty intersection. These problems are
showed \NPc{} and approximated by a local search algorithm, iteratively optimizing the assignment of
one node while all others are fixed. As one of the demonstration on their theoretical work,
\textcite{WeightedTheta15} show a faster solution based on a weighted extension of the Lovász's
theta function, the corresponding geometric embedding of graphs and a solver derived from one-class
SVM, while \textcite{GeneticOCC14} propose a genetic algorithm to solve this problem. Finally,
\Textcite{StochasticCC13} also deals with overlapping clustering by relaxing the problem to a
stochastic setting and using \enquote{the Baum-Eagon inequality, which provides an effective
iterative means for maximizing polynomial functions in probability domains}.

\paragraph{\msc{}}

In \msc{}, the goal is to output a clustering which best summarizes (or agrees with) the several
given input clusterings of the same set of objects.  Motivations includes robustness --by using an
ensemble of clusterings from diverse methods-- and privacy --if the clusterings were computed by
different parties each considering only a subset of the objects attributes. We can build the
complete graph of these objects, with weights set to the fraction of clusterings that place two
objects in different clusters, thus representing a kind of distance in the space of clusterings.  As
first show by \textcite{Gionis2007}, finding the optimal clustering is therefore an instance of
\pcc{} where the weights obey the triangular inequality. They give a deterministic $3$-approximation
using the region growing method of \textcite{Charikar2003}. Later \textcite{Bonizzoni2008} show that
the minimization version is \APXh{}, even when the input is made of three clusterings and give a
combinatorial $\frac{4}{5}$-approximation for the maximization problem. Experimental evaluations are
conducted by \textcite{Bertolacci07} and \textcite{Filkov08}. The former describe a scalable
approach that first samples a small portion of the data, runs a (potentially computationally
expensive) approximation algorithm and finally augment the resulting partition by adding to it the
unsampled nodes one by one.  Experiments confirm that the running time is greatly improved compared
with the linear program methods while the resulting objective value is essentially the same. Note
however that LP methods can be applied in practice thanks to some tricks~\autocite{ConsensusLP10}.
% On the parameterized complexity of consensus clustering, Theoretical Computer Science, 2014,
% \url{http://dx.doi.org/10.1016/j.tcs.2014.05.002}

### 2017-08-12 cc/variants.tex
add two variants: noisy and local

@@
 % \url{http://dx.doi.org/10.1016/j.tcs.2014.05.002}
+
+\paragraph{Local \pcc{}}
+
+% that one is not really local, except for the node penalty
+\Textcite{Puleo2014} adapt the linear program of~\autocite{Charikar2003} (and its rounding by the
+region growing method. Namely after solving the LP, repeatedly pick a ball center $u$ \uar{} with
+radius \shalf{}: if the average distance of the nodes in the ball to $u$ is less than
+\nicefrac{1}{4}, the ball forms a cluster, otherwise $\{u\}$ forms a singleton cluster.) to the case
+where all clusters have to contain less than $K$ nodes, by assigning to each node $u$ a penalty
+$\mu_u$. If $u$ is placed in a cluster $C_i$, the original \mind{} objective is penalized by an
+extra $\mu_u\left(|C_i| - (K+1)\right)$. By varying $\mu_v$ between $0$ and $1$ and because the
+positive weights are assumed to smaller than $1$, this cluster size constraint can be made hard or
+soft. They also handle more general weights, since they allow $w^-_{uv}$ to be as large as $\tau$
+for $\tau\in [1,\infty)$ while still guaranteeing a $5-\nicefrac{1}{\tau}$-approximation on complete
+graphs, and adapt \ccpivot{} to unweighted graphs with the hard cluster size constraint, obtaining a
+randomized $7$-approximation.
+Those soft constraints are for instance used in biological application 
+\emph{A new correlation clustering method for cancer mutation analysis}
+\url{https://arxiv.org/abs/1601.06476}
+\autocite{Hou2016}
+
+\Textcite{pmlr-v48-puleo16} also modify the \mind{} objective to make it more general. Based on the
+classic \pcc{} linear program, they define a \enquote{\emph{fractional clustering} of $G$ as a
+vector $x$ indexed by $V$ such that $x_{uv} \in [0, 1]$ for all $uv \in \binom{V}{2}$ and such that
+$x_{vz} \leq x_{vw} + x_{wz}$ for all distinct $v, w, z \in V$}. They also define \enquote{The
+error vector $err(x)$ of $x$, as a real vector indexed by $V$ whose coordinates are}
+\begin{equation*}
+  \mathrm{err}(x)_u = \sum_{v\in\nei^+(u)} x_{uv} + \sum_{v\in\nei^-(u)} (1-x_{uv})
+\end{equation*}
+
+Given a function $f: \Rbb^n_{\leq 0} \rightarrow \Rbb$ verifying two elementary conditions, the
+problem is then to find a fractional clustering $x$ minimizing $f(err(x))$. The classical \pcc{}
+corresponds to setting $f(x) = \lhalf{}\ell^1(x)$ whereas the authors here are interested in Minimax
+\pcc{} that arises by setting $f(x) = \ell^\infty(x)$. Minimizing the maximum number of
+disagreements incurred by a single node is motivated by the example of recommendation systems: if
+errors corresponds to unsatisfying recommendations, we do not want a single user to suffer a large
+number of them. Minimax \pcc{} is \NPc{} on both complete graphs and complete bipartite graphs but
+by modifying the region growing method of \textcite{Charikar2003}, they give respectively a $48$ and
+$10$ approximation, the latter for the one-sided error (that consider the nodes in only one of the
+two clusters). The idea is to chose pivot not randomly but by maximizing a given criteria and
+to grow ball with a ratio $\alpha$ computed numerically to optimize the approximation factor.
+Interestingly, and in contrast with the classic \pcc{} situation, minimax \maxa{} is not easier than
+minimax \mind{} and seems not to have a constant factor approximation, even on complete graphs.
+Furthermore these algorithms are deterministic, as opposed to many \pcc{} approximation, since
+bounds on expected disagreements on a edge does not translate easily on their maximum.
+\Textcite{Charikar2017} improve these two factors to $7$, using a simpler version of the algorithm
+of \textcite{pmlr-v48-puleo16}. Namely, find the ball of radius $\nicefrac{1}{7}$ with the largest
+number node and create a cluster from its center with a radius of $\nicefrac{3}{7}$. They also show
+that on general weighted graphs, the LP has a large integrality gap of $\nicefrac{n}{2}$ yet they
+combine it with a combinatorial approach to reach a $O(\sqrt{n})$ approximation. Finally they
+consider the complementary problem of maximizing the minimum number of agreements reach at a single
+node, and provide a $\frac{1}{2+\epsilon}$ approximation.
+
+\paragraph{recovery under noise}
+
+As we have seen, assuming the Unique Games Conjecture, the minimum Multicut problem, and therefore
+the \pcc{} problem on a general graph, cannot be approximated to a constant factor in the worst
+case.  But maybe we can do better in the average case, which motivates the study of semi-random
+model, where real graphs are seen as being obtained from the controlled perturbation of a perfectly
+clusterable graph. In the simplest case, each edge sign is independently flipped with probability
+$p \in [0, \shalf)$. This situation on complete graphs was considered in \autocite[Section
+6]{Bansal2002}, showing a simple algorithm that with high probability makes
+$\tilde{O}(n^\frac{3}{2})$ mistakes, and in \autocite[Theorem 2.6]{Ben-Dor99}, with an algorithm
+recovering with high probability the planted partition of an unweighted graph in $O(n^2(\log n)^c)$,
+where $c$ depends on the size of the smallest cluster and the noise probability.
+
+\Textcite{Joachims2005} analyze a more refined weighted model where weights are generated by a
+probability distribution whose mean on true positive edges is larger than $\mu^+>0$ and whose mean
+on true negative edges is smaller than $\mu^-<0$. They give a finite-sample bounds on the number of
+node misclustered w.r.t the planted partition as a function on the probability distribution
+parameters. Indeed, as pointed out by \textcite{plantedAilon09}, independent and uniform noise is
+not a good model of real situations, where the input of \pcc{}, \ie{} the similarity between nodes,
+is often the result of a preprocessing, which may present strong correlations. Therefore, instead of
+measuring the quality of a solution against in the input (\ie{} the similarity information), they
+argue it is more sensible to measure it against the (unknown) true optimal clustering that gave rise
+to the output, and show that \ccpivot{} allows that thanks to a new analysis. \Textcite{Mathieu2010}
+also consider an adversarial model, where all edges are flipped with probability $p$ but the
+adversary then decide whether to show us the true sign or the flipped sign. On complete unweighted
+graphs, they find a solution of \mind{} at most $1 + O(n^{-\nicefrac{1}{6}})$ times the optimal
+whenever $p \leq \shalf - n^{-\nicefrac{1}{3}}$ by rounding the usual SDP solution. If in addition
+$p\leq\nicefrac{1}{3}$, there is no adversary and each planted cluster has at least
+$\Theta(\sqrt{n})$ nodes, then the planted partition can be recovered exactly.
+\Textcite{Makarychev2014} study the same adversarial model on general weighted graphs, giving a PTAS
+for \mind{} when $p\leq \nicefrac{1}{4}$. Under additional assumptions on the density of edges, they
+present another algorithm that finds the ground truth clustering with an arbitrarily small
+classification error.
+% http://www.jmlr.org/papers/v15/chen14a.html

### 2017-08-13 cc/variants.tex
add active variant of CC

@@
 % http://www.jmlr.org/papers/v15/chen14a.html
+
+\paragraph{Online \& active setting}
+
+Except for the local algorithms of \textcite{Bonchi2013}, all works presented so far considered the
+batch case of \pcc{}, where the whole graph and all the signs are available at all time and with no
+cost. This might not always be the case, for instance if the graph is too large to fit in memory or
+if the edge labels are given by an expensive external procedure.
+
+% \emph{Still have to write this part. Is there a link between mistake bound and approximation
+% ratio?} say I have a binary classifier that perfectly predict sign wrt to the optimal clustering
+% C* those below are more concerned with sign prediction than clustering per se Beside batch
+% setting, one can also consider active \autocites{Cesa-Bianchi2012b}{Cesa-Bianchi2012a} and online
+% \autocite{Gentile2013} framework.
+% \enquote{the work of Cesa-Bianchi et al. [CBGV+12] that take a learning-theoretic perspective on the
+% problem of predicting signs. They consider three types of models: batch, online, and active
+% learning, and provide theoretical bounds for prediction mistakes for each setting. They use the
+% correlation clustering objective as their learning bias, and they show that the risk of the
+% empirical risk minimizer is controlled by the correlation clustering objective.}
+
+Case in point, in the context of entities resolution, \textcite{activeCoref07} consider the \pcc{}
+problem where by querying an oracle, one can either reduce the uncertainty about one edge weight or
+add an extra node with edges connecting it to existing nodes. However, this requires web queries are
+resource bounded and thus yields an active problem, where one has to choose the most informative
+queries given a budget. A formal definition is given in~\autocite{queryMoreEdgeCC07} and in
+practice, after finding an initially good partition, they select in each cluster a node to be its
+centroid, and query the edges connecting the centroid as ordered by an entropy-based criterion.
+\textcite[Section 5]{Ailon2014} present another active algorithm that solves \mind$[k]$ with a
+linear number of queries but an exponential running time in $n$.  In a noisy setting,
+\textcite{Mitzenmacher2016} assume there is a planted $k$-partition and that we have access to an
+oracle that for $u,v\in V^2$ returns $\cluster(u) - \cluster(v) \mod k$ with probability $1-p$ and a
+noisy $\cluster(u) - \cluster(v) \pm 1 \mod k$ otherwise. For $k=2$, this is exactly $2$-\pcc{}, for
+$k>2$, the oracle gives more information than simply the sign of the path from $u$ to $v$. Whenever
+$p < \shalf$, they show that $O(n^{1+\frac{1}{\log\log n}}\log n)$ queries are enough are to recover
+the planted partition in polynomial time and with high probability. Those queries are actually
+random (\ie{} non-adaptive), and the clusters are found by looking at almost edge-disjoint path
+between all pairs of nodes.
+
+In the online setting, \textcite{greedyOnline10} give a greedy algorithm that upon node arrival creates a
+singleton cluster and then merge all pairs of clusters for which it increases the total number of
+agreements. For \mind{}, this is $O(n)$-competitive algorithm and they show such ratio is optimal by
+exhibiting an instance\footnote{two positive cliques $A$ and $B$ joined by positive edges except
+between $a\in A$ and $\{b_1,\ldots, b_k\}\in B$. Those nodes are given first and thus form a cluster
+which yield at least one disagreement every time one the $n-(k+1)$ remaining nodes is added.} on
+which any strategy ends up with $n - k$ disagreements whereas optimal cost is $k$. On the \maxa{}
+side, this greedy strategy result in a $0.5$-competitive algorithm. If it is randomly mixed with a
+\textsc{Dense} variation, it increases to $0.5+\eta$, still far from the demonstrated $0.834$ upper
+bound.
+% when the graph is given node by node,
+% \url{https://link.springer.com/chapter/10.1007/978-3-319-18173-8_7} show to cluster it in static
+% clique, so it sounds related to cluster editing, except their objective is to maximize the
+% fraction of edge in cliques (ie avoid small cluster)

### 2017-08-13 cc/variants.tex
add cluster editing variant

@@
 % fraction of edge in cliques (ie avoid small cluster)
+
+\paragraph{\textsc{Cluster editing}}
+
+% \url{http://www.sciencedirect.com/science?_ob=ArticleListURL&_method=list&_ArticleListID=-1231039636&_sort=r&_st=4&md5=018b782f17918959c0cc389c33ba2cd7&searchtype=a}
+
+In the cluster editing problem, one is given a general unsigned input graph $H=(V,E)$ and wants to
+find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
+union of cliques. This is equivalent to \pcc{} on complete graphs. To see why, let $G=(V,(E^+,E^-))$
+be a complete signed graph with nodeset $V$, $E^+$ be the set $E$ of edges of $H$ and $E^-$ be all
+the edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
+disagreements within clusters are edges added to $H$ and the positive disagreements between clusters
+are deleted from $H$.  To the best of our knowledge, the problem was first introduced under this
+name by \textcite{Shamir02}.\marginpars{However, the same problem was studied before and we refer the
+reader to the comprehensive survey of \textcite{clusterEditSurvey13} for additional details,
+whereas we shall only give important and more recent pointers here.} They show the problem is
+\NPc{}, even if the number of clusters $k\geq2$ is set beforehand\footnote{The reduction is from
+3-exact 3 cover, see \autocite[Theorems 1, 2 and Corollary 1]{Shamir02}.} and provide a
+$0.878$-approximation in the weighted $k=2$ case using the standard Goemans-Williamson SDP
+relaxation.
+
+If we parameterize the problem by the number $d$ of edges that need to be added or deleted (that is
+the number of disagreements), then it can be solved in polynomial time in the size of the input
+graph (but not in $d$). The best known approximation so far is
+$O(1.62^d+m+n)$~\autocite{GoldenCE12}, which search for conflict triple (\ie{} unbalanced triangle
+in the sign language) and branch by either deleting or merging one positive edge. If we additionally
+look for exactly $k$ clusters (\ie \mind$[k]$), there is a fixed parameters algorithm running in
+$O(2^{O(\sqrt{kd})}+n+m)$~\autocite{Fomin2014}. If at most $a$ edges can be added and at most $b$
+edges can be deleted at each node, and if the minimum size of a cluster is at least $2(a+b)$, then
+the problem can be solved in polynomial time~\autocite{Abu-Khzam2015}. Finally, On planar graph,
+there is a PTAS running in $O(n2^{\epsilon^{-1}\log(\epsilon^{-1})})$ obtained by dividing the graph
+into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}

### 2017-08-14 cc/variants.tex
add communities detection variant of CC

@@
 into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}
+
+\paragraph{Communities detection}
+
+The clustering problem is often named communities detection in the context of social networks, and
+several methods developed by practitioners have been extended to signed graphs. While they do not
+necessarily considered the \pcc{} objective, and especially not its optimum, we still give a brief
+overview of them, as they tend to have been more tested experimentally. For instance, to find the
+cluster of node $u$, \textcite{Yang2007} use a random walk approach on the positive subgraph $G^+$
+to compute the probability of each node to reach $u$ in $k$ steps, sort the nodes accordingly and
+then find a threshold  based on number of disagreements.
+The one node moves local heuristic that we described in the Physics inspired paragraph
+\vpageref{par:cc_physics} can also be formulated as genetic algorithms that simultaneously try to
+minimize the number of disagreements and maximize a signed variant of the
+modularity~\autocites{Li2013}{Amelio2013}. \Textcite{Anchuri2012} also consider these two objectives
+by seeing them as eigenvalue problems and devise a iterative splitting procedure.
+The overlapping community detection variant is considered by \textcite{Chen14}, who used a signed
+probabilistic mixture model. Namely, an edge selects a pair of cluster $r,s$ with probability
+$\omega_{rs}$ (where $r=s$ if the edge is positive and $r\neq s$ otherwise) and choose its endpoint
+$u$ and $v$ with probability $\theta_{ru}$ and $\theta_{sv}$. $\theta_{ru}$ is therefore the soft
+membership of node $u$ to cluster $r$, and those parameters are estimated using the
+expectation-maximization algorithm. The same model is extended to directed graph by
+\textcite{Jiang2015}, who strangely enough names it stochastic blockmodel, although the focus is
+still on edge and not nodes.
+In a similar spirit to \maxa{}, \textcite{SignedGang} focus on finding $k$ subgraphs dense in
+positive edges and densely connected by negative edges to each others. They dub such graph
+\emph{Oppositive Cohesive Groups}, or more vivedely \emph{Gangs in War} and after formulating the
+problem as a constrained quadratic optimization, they propose a faster iterative local search
+heuristic.
+% relaxed structural balance \autocite{Doreian2009} but used the same approach as \textcite{Early96}

### 2017-08-14 cc/multimincut_stable.tex
start writing about multicut reduction

@@
-Haris Angelidakis, Konstantin Makarychev, and Yury Makarychev. 2017.
-Algorithms for Stable and Perturbation-Resilient Problems. STOC’17
-\href{http://ttic.uchicago.edu/~yury/papers/two-stable.pdf}{10.1145/3055399.3055487}
-improves over the one cited in the internship description
+We first give more details on the equivalence between \mmc{} and \pcc{},
+following~\autocite{Demaine2006} but omitting the full proofs to avoid redundancy. Recall that given
+an weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and sinks
+$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, in \mmc{} we want to find the lightest set of edges $T\in E_H$
+whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
+polynomial reduction from \mmc{} to \pcc{}. Namely given an instance $(H, S)$ of \mmc{}, we let
+$W_H=\sum_{e\in E_H} w_H(e)$ be the total weight of $H$ and $G_H$ be the same graph as $H$ with all
+its edges labeled positively. We then connect every pair $(s_i, t_i)$ by a negative edge of weight
+$W_H+1$. One can check (\autocite[Theorem 4.7]{Demaine2006}) that a multicut $T$ in $H$ with weight
+$W$ induces a clustering in $G_H$ of weighted disagreement $W$ by the connected components of
+$G^+\setminus T$. Likewise a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
+weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges
+are simulated in the following way: include every source and sink in a clique of $n$ nodes and
+connect that clique to the corresponding source or sink by $n$ negative edges.
+
+We next present the polynomial reduction from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
+$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v)$ of
+weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
+weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-sink pair added to $S_G$. This
+construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
+4.4]{Demaine2006})\Todo{give the intuition of the proof} that it is easy to construct a multicut of
+weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.
+
+\begin{figure}[htpb]
+   \centering
+   \includegraphics[width=0.8\linewidth]{assets/raw/cc_to_mmc.png}
+   \caption{The transformation from \pcc{} on $G$ to \mmc{} on $H$ (temporarily reproduced from
+   \autocite{Demaine2006}) \label{fig:cc_mmc}}
+\end{figure}
+
+Whereas clustering objective functions are \NPh{} to optimize, we expect that meaningful instances
+in which we are interested have additional structure which allows for guaranteed polynomial time
+algorithms, see for instance \autocite{clusteringFeasibility15} for a critical overview of some
+proposed such notions of structure. Informally, a general idea is that the clustering should not
+change (or at least very little) if the data are slightly perturbed. For instance, a weighted graph
+is \emph{$\alpha$-stable} (with $\alpha>1$) for some partition objective if its optimal partition
+remains the same whenever every weight $w_i$ is multiplied by a factor $c_i$ between $1$ and
+$\alpha$. Finally, we say that an algorithm is \emph{robust} if, given an instance $\mathcal{I}$ and
+in polynomial time, it: returns the optimal solution of $\mathcal{I}$ if $\mathcal{I}$ is
+$\alpha$-stable; and if $\mathcal{I}$ is not $\alpha$-stable, either returns the optimal solution of
+$\mathcal{I}$ or reports that $\mathcal{I}$ is not $\alpha$-stable. This is a handy property, as in
+general, we cannot check the stability of an instance in practice as we do not know the optimal
+solution hence we cannot tell whether it changes or not under perturbations.
+
+\Textcite{StableCC17} provide a robust algorithm for $2-\nicefrac{2}{k}$-stable instance of
+$k$-\textsc{Minimum Multiway Cut}, which therefore is not really useful in our case because I confused
+multiway cut and multicut. To be fair, it seems I'm not the only one, \href{http://pages.cs.wisc.edu/~shuchi/papers/multicut-hardness-full.pdf}{on the hardness of
+   approximating multicut}
+   \enquote{multicut is known to be APX-hard
+      (\href{https://pdfs.semanticscholar.org/1cf6/4c2bdd4f1c384a55910606a64c8d831a96ba.pdf}%
+   {Dahlhaus et al. 1994}).} by that 1994 paper define the multiway cut problem, see
+   \autoref{fig:cc_multiwhat}. \Textcite[Theorem 24]{Bansal2004} show that we can go from
+   \textsc{Minimum Multiway Cut} to \pcc{} but I doubt about the other direction, mostly because
+   \textsc{Minimum Multiway Cut} can actually be approximated with factor smaller than $1.3$.
+   see \url{http://www.nowozin.net/sebastian/papers/nowozin2009lpstability.pdf} about LP multicut
+   stability.
+\begin{figure}[htpb]
+   \centering
+   \includegraphics[width=0.9\linewidth]{assets/raw/multicut_vs_multiway.png}
+   \caption{Confusing statement} \label{fig:cc_multiwhat}
+\end{figure}
+
+% \enquote{We prove that if the Unique Games Conjecture of Khot (2002) is true, then for every constant L > 0
+% it is NP-hard to approximate Multicut within factor L. If a quantitatively stronger version of the
+% conjecture is true, then Multicut is NP-hard to approximate within a factor of $\Omega(\sqrt{\log
+% \log n})$.}
+
+% It's probably limited because LP with a lot of constraints do not scale that well (someone claimed
+% $O(n^{4.5})$)
+
+% SDP Gaps and UGC Hardness for Multiway Cut, 0-Extension and Metric Labelling,
+% On the Unique Games Conjecture: Subhash Khot
+% under the UGC, one cannot get better than $O(\log n)$ approximation using LP or SDP relaxation.

### 2017-08-14 cc/variants.tex
typos

@@
 membership of node $u$ to cluster $r$, and those parameters are estimated using the
-expectation-maximization algorithm. The same model is extended to directed graph by
+expectation-maximization algorithm. The same model is extended to directed graphs by
 \textcite{Jiang2015}, who strangely enough names it stochastic blockmodel, although the focus is
 still on edge and not nodes.
-In a similar spirit to \maxa{}, \textcite{SignedGang} focus on finding $k$ subgraphs dense in
-positive edges and densely connected by negative edges to each others. They dub such graph
-\emph{Oppositive Cohesive Groups}, or more vivedely \emph{Gangs in War} and after formulating the
+In a similar spirit to \maxa{}$[k]$, \textcite{SignedGang} focus on finding $k$ subgraphs dense in
+positive edges and densely connected by negative edges to each others. They dub such subgraph
+\emph{Oppositive Cohesive Groups}, or more vividly \emph{Gangs in War}, and after formulating the
 problem as a constrained quadratic optimization, they propose a faster iterative local search

### 2017-08-15 cc/sign_relation.tex
expand balance theory

Now that we are familiar with the \pcc{} problem and its numerous applications, let us return to the
original problem of this chapter, predicting the edge sign of a signed graph. In order to do so, we
will first focus on the case where a signed graph has no disagreements (that is the minimum of
\eqref{eq:mind} is $0$) and is unweighted, as it highlights more clearly the connection with the
balance theory.

In his seminal work on interpersonal relations,
\citeauthor*{Heider46}~\autocites{Heider46}{HeiderBook58} posits through
psychological and sociological arguments that in order to reduce their cognitive dissonance, three
people always interact in a way that preserve social balance. This social balance requirements can be
succinctly summarized by four statements~\autocite{HeiderBook58}:

\begin{enumerate}
	\item \textcolor{DodgerBlue}{my friend's} \textcolor{Orange}{friend} is
		\textcolor{Olive}{my friend}\hspace{1em}%
		\tikz[baseline=5]{
			\node[draw,circle,inner sep=2pt] (me) at (0,0) {};
			\node[draw,circle,inner sep=2pt] (mine) at (.5,.5) {};
			\node[draw,circle,inner sep=2pt] (alter) at (1,0) {};
			\draw[] (me) -- node[left,DodgerBlue,pos=.9] {$+$} (mine) ;
			\draw[] (mine) -- node[right,Orange,pos=.1] {$+$} (alter) ;
			\draw[white] (me) -- node[text=Olive] {$+$} (alter) ;
		}
	\item \textcolor{DodgerBlue}{my friend's} \textcolor{Orange}{enemy} is
		\textcolor{Olive}{my enemy}\hspace{1em}%
		\tikz[baseline=5]{
			\node[draw,circle,inner sep=2pt] (me) at (0,0) {};
			\node[draw,circle,inner sep=2pt] (mine) at (.5,.5) {};
			\node[draw,circle,inner sep=2pt] (alter) at (1,0) {};
			\draw[] (me) -- node[left,DodgerBlue,pos=.9] {$+$} (mine) ;
			\draw[] (mine) -- node[right,Orange,pos=.1] {$-$} (alter) ;
			\draw[white] (me) -- node[text=Olive] {$-$} (alter) ;
		}
	\item \textcolor{DodgerBlue}{my enemy's} \textcolor{Orange}{friend} is
		\textcolor{Olive}{my enemy}\hspace{1em}%
		\tikz[baseline=5]{
			\node[draw,circle,inner sep=2pt] (me) at (0,0) {};
			\node[draw,circle,inner sep=2pt] (mine) at (.5,.5) {};
			\node[draw,circle,inner sep=2pt] (alter) at (1,0) {};
			\draw[] (me) -- node[left,DodgerBlue,pos=.9] {$-$} (mine) ;
			\draw[] (mine) -- node[right,Orange,pos=.1] {$+$} (alter) ;
			\draw[white] (me) -- node[text=Olive] {$-$} (alter) ;
		}
	\item \textcolor{DodgerBlue}{my enemy's} \textcolor{Orange}{enemy} is
		\textcolor{Olive}{my friend}\hspace{1em}%
		\tikz[baseline=5]{
			\node[draw,circle,inner sep=2pt] (me) at (0,0) {};
			\node[draw,circle,inner sep=2pt] (mine) at (.5,.5) {};
			\node[draw,circle,inner sep=2pt] (alter) at (1,0) {};
			\draw[] (me) -- node[left,DodgerBlue,pos=.9] {$-$} (mine) ;
			\draw[] (mine) -- node[right,Orange,pos=.1] {$-$} (alter) ;
			\draw[white] (me) -- node[text=Olive] {$+$} (alter) ;
		}
\end{enumerate}

This can be readily translated into graph properties. Given a path of length two over three nodes,
note that the first part of each statement (in blue and orange) defines the four possible sign
assignments of such a path. The last part of the statement (in green) then prescribes which sign
should close the triangle to respect the social balance. Such triangles are called strongly balanced.
\begin{definition}[strongly balanced triangle]
	A triangle is strongly balanced if it has zero or two negative signs. Otherwise it is unbalanced.
\end{definition}

We will use triangles as the building block of our definition of strongly balanced graphs.
\begin{definition}[strongly balanced complete graphs]
	A complete signed graph is strongly balanced if all its triangles are strongly balanced.
\end{definition}

It is interesting to note that this local property of triangles gives raise to a global structure.
Namely a strongly balanced complete graph can be split in two (possibly empty) clusters $L$ and $R$
such that all edges within $L$ and $R$ are positive, while all edges between $L$ and $R$ are
negative. Indeed let us pick an arbitrary node $u$ and let $L=\neip(u)$ and $R=V\setminus L$. Now
let $v,w\in L\setminus \{u\}$ and $x,y\in R$. From \autoref{fig:cc_twostrongcc} and the fact that
every triangle is balanced, we can see that $(v,w)$ and $(x,y)$ are positive edges, while $(w,y)$ is
a negative edge.
\begin{marginfigure}
  \centering
  \includegraphics[width=28mm]{assets/raw/2cc.pdf}
  \caption{A two-clustering a complete strongly balanced graph}
  \label{fig:cc_twostrongcc}
\end{marginfigure}
As we will see later, these two facts are actually equivalent.

Those two clusters gives a natural partition with no disagreements in the sense of \pcc{}. However,
the existence of two such clusters is only a sufficient but not necessary condition for having no
disagreement. As hinted in \autoref{sub:need_for_a_directed_graph}, when the nodes are partitioned
into $k\geq 2$ clusters, letting edges within clusters to be positive and edges across clusters to
be negative will by construction results in zero disagreement. Furthermore, it was noted by
sociologists that among triangles with odd number of negative edges, one is more stable and common
than that the other one. Therefore, \textcite{davis1967clustering} relax the strong balance into the
weak balance by considering triangles with three negative edges to be balanced, as illustrated on
\autoref{fig:cc_balance}. Formally, we can modify our two previous definitions to arrive at weak
balance:
\begin{definition}[Weak balance]
	A triangle is weakly balanced if it has zero, two or three negative signs. Otherwise it is
	unbalanced. \\
	A complete signed graph is weakly balanced if all its triangles are weakly balanced.
\end{definition}

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.8\linewidth]{assets/raw/triangle_smaller.pdf}
	\caption{The four possible undirected triads, as classified by the two structural balance theories
	introduced in the main text} \label{fig:cc_balance}
\end{figure}

In that case as well, the local property of triangle has structural implication for the whole graph.
Namely, we can partition a complete weakly balanced graph into $k$ clusters of positively connected
nodes that are negatively with the rest of the graph. Consider the same construction as in
\autoref{fig:cc_twostrongcc}, putting the positive neighborhood of a node $u$ in one cluster $L$ and
the rest of the nodes in $R$. For $v,w\in L\setminus \{u\}$ and $x,y\in R$, we can still conclude
that $(v,w)$ is positive and $(v,y)$ is negative. On the other hand, this time weak balance allows
$(x,y)$ to be either positive or negative. We can split $R$ further into the positive neighborhood of
$x$ and the rest, and keep doing so until we form the $k$ clusters. We will make this argument more
formal in the following.

We gave the definitions of balance in the context of complete graphs in order to build intuition
about the partition consequence and because it easier to reason on triangles. Furthermore, if we are
given a set of objects and a similarity function, we can indeed build a complete graph. However,
they are situations were it is no realistic to make the completeness assumption, especially for
social networks that are typically very sparse. We therefore extend the concept of balance to
general graphs in the following way:
\begin{definition}[balanced general graphs]
	A general graph is strongly (respectively weakly) balanced if there is a sign assignment of all its
	missing edges such that the resulting complete signed graph is strongly (respectively weakly)
	balanced.
\end{definition}
To similarly extend the characterization of signed graphs by consistent clustering, we need to
consider longer cycles than triangles.
\begin{definition}[balanced cycles]
	A cycle is strongly (respectively weakly) unbalanced if it has an odd number of negative edges
	(respectively exactly one negative edge). Otherwise it is strongly (respectively weakly) balanced.
\end{definition}

Being strongly balanced is equivalent to having two consistent clusters, as proved in the seminal
theorem of \textcite[Theorem 3]{harary1953}:
\begin{theorem}[Structural Theorem]
  A graph $G=(V,E)$ is strongly balanced if and only if $V$ can be divided in two disjoint clusters
  such that all edges within clusters are positive and all edges between clusters are negative.
\end{theorem}

According to \textcite{Huffner2010}, a similar theorem was proved earlier by \textcite[Theorem
X.11]{Konig36}, although \textcite{Zaslavsky2012} notes that it was stated \enquote{without the
terminology of signs, while \autocite{harary1953} has the first recognition of the crucial fact that
labelling edges by elements of a group---specifically, the sign group---can lead to a general
theory.}

A useful characterization, whose proof can be found in~\autocite[page 122]{BookKleinberg2010}, is the
following:
\begin{theorem}
  A graph $G=(V,E)$ is strongly balanced if and only if all its cycles are strongly balanced.
\end{theorem}

The same results also hold for weak balance.
\begin{theorem}
  A graph $G=(V,E)$ is weakly balanced if and only if $V$ can be divided in $k$ disjoint clusters
  such that all edges within clusters are positive and all edges between clusters are negative.
\end{theorem}
\begin{proof}
  Let first assume that $G$ is weakly balanced. According to the definition, we can choose sign for
  the missing edges such that it becomes a complete graph with all its triangles weakly balanced. As
  mentioned earlier, we can then pick a node $u_1$ and let $\cluster_1 = \neip(u) \cup \{u\}$. All
  nodes in $\cluster_1$ are connected positively with each and negatively with $V\setminus
  \cluster_1$. Note also that $|V\setminus \cluster_1| < |V|$ and therefore we can repeat this
  procedure until $V$ is exhausted, at which point we have obtained our $k$ clusters verifying the
  theorem's statement conditions.

  Conversely, assume we have such $k$ clusters. We can complete the graph $G$ by letting the edge
  $(u,v)$ be positive if $u$ and $v$ are in the same cluster and negative otherwise. Let us the pick
  three arbitrary nodes $u,v,w\in V$. There are three cases: they are either all in the same
  cluster, all in different clusters or two of them are in a first cluster and the third node is in
  a second cluster. In every case, one can check they form a weakly balanced triangle and therefore
  $G$ is weakly balanced.
\end{proof}
\begin{theorem}[Theorem 1 of \autocite{davis1967clustering}]
  A graph $G=(V,E)$ is weakly balanced if and only if all its cycles are weakly balanced.
\end{theorem}

In real data though, we do not expect neither strong nor weak balance to hold, for they are fairly
demanding model. Indeed, three of the real networks considered in \autoref{chap:troll} have been
repeatedly shown to be unbalanced, although the extent of this unbalance depends of the importance
given to longer cycle compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
{measureUnbalance17} For the same reason, in virtually any real instance of the \pcc{} problem, the%
% there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration
% index) which they claimed has been peer reviewed?
\marginpars{The proof makes essential use of the strong balance (which is ok since next section also
deal with 2-CC) and the completeness of the graph (which is a far less realistic assumption) yet
it would be interesting to see if it can be extended.}
optimal solution will incur some disagreements. \Textcite[Section 5.5]{BookKleinberg2010}
nonetheless prove that even when only a fraction of the triangle are strongly balanced in a complete
graph, the two clusters structure is still present, although it does not cover the whole graph
anymore.

\begin{theorem}
  Let $\epsilon < \nicefrac{1}{8}$ and $\delta = \sqrt[3]\epsilon$. If at least $1-\epsilon$ of all
  triangles in a signed complete graph are strongly balanced, then either
  \begin{enumerate}[(a),nosep]
    \item there is a set consisting of at least $1-\delta$ of the nodes in which at least $1-\delta$
      of all edges are positive, or else
    \item the nodes can be divided into two groups $L$ and $R$ such that
      \begin{enumerate}[(i),nosep]
	\item at least $1-\delta$ of the edges in $L$ are positive,
	\item at least $1-\delta$ of the edges in $R$ are positive, and
	\item at least $1-\delta$ of the edges between $L$ and $R$ are negative.
      \end{enumerate}
  \end{enumerate}
\end{theorem}

We can conclude that balance theory helps characterize zero disagreement instances of \pcc{}.
However, how is \pcc{} useful in predicting edge sign? At first sight, the two problems seem quite
different for the former is an unsupervised/agnostic clustering problem while the latter is a
supervised classification problem. Yet, getting an optimal solution of \pcc{} on the training set
would provide a principled heuristics for \esp{}. Unfortunately, as we shall see in
\autoref{sub:state_of_the_art}, exact or fixed parameter solution are not scalable and one has to
rely approximation. Under the Unique Games Conjecture, the approximation factor is lower bounded by
$\Theta(\log n)$ in the worst case. However, this is still a viable option, as our bias is that we
will operate on instances that are to some extent balanced. Therefore, we are not in the worst case
and we can find some comfort in results about the noisy (\vpageref{sub:variants_and_extensions}) and
stable (\autoref{ssub:cc_under_stability_assumption}) settings.

Another connection we will leverage in \autoref{sec:low_stretch_trees_and_spanners} is that the
value of the \mind{} objective is a complexity measure of the \esp{} problem~\autocite[Section
4.1]{Cesa-Bianchi2012b}. Before that, we will also present some variants and extensions of \pcc{},
some of them bearing similarities with the Chapter 3's problem.

\iffalse
strong balance random graph model \url{https://www.lri.fr/~yannis/randsigned.pdf}, line index is basically
the objective value of CC and if I'm reading their bound correctly,  basically you need to flip half
the edges of a random graph to make it balanced (also most of their results hold when the
probability of an existing edge is 1/2, but it was strengthened in a follow up dealing with weak
balance \url{http://people.maths.ox.ac.uk/harutyunyan/weak-balance.pdf})

For the 2 clusters case ($k=2$), characterization proven already in 1936 by \textcite{Konig36}, as noted in
\cite{Zaslavsky2012} (commenting on \cite{harary1953}: \enquote{Although Theorem 3 was anticipated by
\textcite[Theorem X.11]{Konig36}  without the terminology of signs, here is the
first recognition of the crucial fact that labelling edges by elements of a
group---specifically, the sign group---can lead to a general theory.})
and \cite{Huffner2010} (\enquote{\textcite{Konig36} proved the following characterization of
	balanced graphs. For a graph $G = (V, E)$, the following are equivalent:\marginpars{There is a
	proof in \autocite[p. 111]{BookKleinberg2010}, maybe I can rewrite it as well}
	\begin{enumerate}
		\item $V$ can be partitioned into two sets $V_1$ and $V_2$ called sides such that there is no
			negative edge $\{v, w\} \in E$ with both $v, w \in V_1$ or both $v, w \in V_2$ and no positive
			edge $\{v, w\}$ with $v \in V_1$ and $w \in V_2$ .
		\item $V$ can be colored with two colors such that for all $\{v, w\} \in E^-$, the vertices $v$
			and $w$ have different colors, and for all $\{v, w\} \in E^+$, the vertices $v$ and $w$ have
			the same color. The color classes correspond to the sides.
		\item $G$ does not contain cycles with an odd number of negative edges.
	\end{enumerate}
	Using the characterization by a coloring, it is easy to see that balance of a signed
graph can be checked in linear time by depth-first search.})
\fi

### 2017-08-15 cc/problem_applications.tex
small typos at beginning of CC

@@
 
-In \pcc{}, we formalize this problem by considering objects as the nodes of a graph $G$, whose edges
-weight encode similarity. Namely, in the most general case, for nodes $u$ and $v$, the edge between
+In \pcc{}, we formalize this problem by considering objects as the nodes of a graph $G$, whose edge
+weights encode similarity. Namely, in the most general case, for nodes $u$ and $v$, the edge between
 $u$ and $v$ is associated with two positive numbers:
-$w_{u,v}^+$ denotes the strength of the similarity between $u$ and $w$;
-$w_{u,v}^-$ denotes the strength of the dissimilarity between $u$ and $w$.
-Note however that in many applications, only one of these two numbers is non zero in which case we
-more conveniently set $w_{u,v} = \begin{cases}
+$w_{u,v}^+$ denotes the strength of the similarity between $u$ and $v$;
+$w_{u,v}^-$ denotes the strength of the dissimilarity between $u$ and $v$.
+Note, however, that in many applications, only one of these two numbers is non zero, in which case
+we more conveniently set $w_{u,v} = \begin{cases}
 	 w_{u,v}^+ & \quad \text{if } w_{u,v}^+ > 0 \text{ and } w_{u,v}^-=0 \\
@@
 Now consider a clustering \cluster{} of $V$, that is a function from $V$ to $\Nbb{}^{|V|}_{>0}$
-that assigns to each node a cluster index. For instance, $\cluster(u) = 3$ means that $u$ belongs
+that assigns to each node a cluster index.\marginpars{We will also abuse the notation and let
+\cluster{} be a set of clusters $\cluster=\{C_1, C_2, \ldots, C_k\}$ that form a partition of $V$.
+That is, $\forall i\,C_i\subset V$, $\bigcup_{i=1}^k C_i = V$ and $\forall i\neq j,\, C_i\cup C_j =
+\emptyset$.} For instance, $\cluster(u) = 3$ means that $u$ belongs
 to the third cluster. We can evaluate how \cluster{} fits our clustering paradigm in two ways,
@@
 weighted number of positive edges across clusters. Given a cost function $c$, which is usually the
-identity, \pcc{} can then be seen as graph optimization problem, either of maximizing agreements
+identity, \pcc{} can then be seen as a graph optimization problem, either of maximizing agreements
 (\maxa{}):
@@
 many clusters are the data, or external constraints. This can be handled with parametrized version
-of \pcc{}.
+of \pcc{}, namely \maxa$[k]$ and \mind$[k]$ where the optimization is over clustering with exactly
+$k$ clusters.
 
@@
 	disagreements: the negative edge between nodes $1$ and $2$ within the blue cluster, and the
-	positive edge $6,7$ between the orange and green clusters. Those disagreements are created
+	positive edge $(6,7)$ between the orange and green clusters. Those disagreements are created
 	by two cycles with one negative edge and thus cannot be avoided, meaning that \cluster{} is
-	optimal. However it is not the unique solution, for instance merging the orange and green
+	optimal. However it is not the unique solution: for instance, merging the orange and green
 	clusters would also yields two disagreements.}
@@
    \item[Computer Vision]
-      The ubiquitous task of segmenting an image into homogeneous regions is prerequisite for many
+      The ubiquitous task of segmenting an image into homogeneous regions is a prerequisite for many
       further processing, and can be tackled with \pcc{}. For instance, to segment cell in
@@
       superpixels. After building the adjacency graph of these superpixels, they assigned edge
-      weight by averaging the boundary probability over the pixels separating the adjacent
+      weights by averaging the boundary probabilities of the pixels separating adjacent
       superpixels. They also add strong negative constraints between distant superpixels, and lastly
       cluster these superpixels according to the \pcc{} objective to obtain the final segmentation.
-      A similar approach was used earlier in~\textcite{Kim2011}, which stressed the importance of
+      A similar approach was used earlier in~\textcite{Kim2011}, who stress the importance of
       considering such higher order constraints between distant superpixels in order to achieve good
       performance. This was also extended to 3D segmentation~\autocite{VolumeSegmentation12}, where
-      additional tunings allowed to segment a volume image of a mouse cortex with up to billions
+      additional tuning allows to segment a volume image of a mouse cortex with up to billions
       voxels. \Textcite{Beier2015} segment 2D and 3D images with a energy based formulation of
@@
       proposal generator. By developing another scalable energy based optimization procedure, and
-      with the help of few user provided cues, \textcite{Bagon2011} were able to apply \pcc{}
+      with the help of few user provided cues, \textcite{Bagon2011} are able to apply \pcc{}
       directly at the pixel level.
       % one more? https://link.springer.com/chapter/10.1007/978-3-642-33783-3_41
-      % 3D mesh segement https://link.springer.com/article/10.1007/s41095-016-0071-3
+      % 3D mesh segment https://link.springer.com/article/10.1007/s41095-016-0071-3
 
@@
       detected in the current frame. This matrix is turned into a symmetric affinity matrix
-      $\bar{A}_{sym}$ that can be seen as a signed graph. \pcc{} is then used to extract clusters
-      (called zones), in which local processing is performed, whose complexity can be adapted to the
-      difficulty of each zone.
+      $\bar{A}_{sym}$ that can be seen as a signed graph adjacency matrix. \pcc{} is then used to extract clusters
+      (called zones), in which local processing is performed. This is beneficial since the
+      complexity of these local methods can be adapted to the difficulty of each zone.
 
@@
       \textcite{Elsner2009} first obtain an upper bound of the optimal solution by solving a SDP
-      relaxation of the problem. They then compared various heuristics and show that best performing
+      relaxation of the problem. They then compare various heuristics and show that best performing
       ones are within few percents of the optimum, provided they are followed by a local search
@@
    \item[Biology]
-      The input is a similarity matrix between gene expressions in various experimental conditions
+      Clustering tasks also abounds in Biology.
+      A typical input is a similarity matrix between genes expression level in various experimental conditions
       and the goal is to cluster those genes into groups which react similarly. \Textcite[Section
@@
       genes of the roundworm \emph{C. elegans} and 2000 human genes obtained from 40 tumor and 22
-      normal colon samples. \Textcite{Mason2009} analyze a signed co-expression networks of genes
+      normal colon samples. \Textcite{Mason2009} analyze a signed co-expression network of genes
       involved in embryonic stem cells to find which genes are related to pluripotency (the ability
@@
       sequences, leading to different alleles of the corresponding gene. An haplotype is the list of
-      all alleles at a contiguous sites in a region of a single chromosome and this information is
+      all alleles at a contiguous region of a single chromosome, and this information is
       used in several medical applications. The high-throughput sequencing of one individual genome
       yields many short \emph{reads} that provides information about the order of nucleotides in a
-      fragment of one chromosome and that can be used to assemble haplotypes. Namely the authors
-      build a graph of reads and define a similarity function between reads to assign weight on the
+      fragment of one chromosome and that can be used to assemble haplotypes. To do so, the authors
+      build a graph of reads and define a similarity function between reads to assign weights on the
       edges. The clusters of that graph correspond to haplotypes, and are obtained by a SDP
@@
       number of edges to remove in order to decompose a dynamics system into a collection of
-      monotone system. This corresponds to the \pcc{} \eqref{eq:mind} objective and allows to study
+      monotone system. This corresponds to the \mind{} problem and allows to study
       the complete system more easily. Many more applications of \mind{} on weighted complete graphs
@@
       detection~\autocite{FortunatoSurvey10}, it is therefore not surprising that \pcc{} has been
-      used to study such networks. For instance, one can extract all the votes of the member of
-      political parliament and form a graph whose node are politicians and edge weight quantify how
+      used to study such networks. For instance, one can extract all the votes of the members of a
+      political parliament and form a graph whose nodes are politicians and edge weights quantify how
       much they agree or disagree on various issues they have been voting on. This can be used to
@@
 	    non trivial ways an existing \pcc{} algorithm. This was extended to weighted and partial
-	    constraints  by \url{https://doi.org/10.1007/978-3-319-10933-6_23}. Another example is given by
-	    \textcite{Crosslingual07}, who cluster together news article in different language
+	    constraints  by \textcite{WeightedER14}. Another example is given by
+	    \textcite{Crosslingual07}, who cluster together news articles in different languages
 	    covering the same event. \pcc{} was also evaluated among other

### 2017-08-15 cc/sign_relation.tex
small typos at beginning of CC

@@
 original problem of this chapter, predicting the edge sign of a signed graph. In order to do so, we
-will first focus on the case where a signed graph has no disagreements (that is the minimum of
+will first focus on the case where a signed graph has no disagreement (that is the minimum of
 \eqref{eq:mind} is $0$) and is unweighted, as it highlights more clearly the connection with the
@@
 psychological and sociological arguments that in order to reduce their cognitive dissonance, three
-people always interact in a way that preserve social balance. This social balance requirements can be
+people always interact in a way that preserve social balance. These social balance requirements can be
 succinctly summarized by four statements~\autocite{HeiderBook58}:
@@
 such that all edges within $L$ and $R$ are positive, while all edges between $L$ and $R$ are
-negative. Indeed let us pick an arbitrary node $u$ and let $L=\neip(u)$ and $R=V\setminus L$. Now
+negative. Indeed let us pick an arbitrary node $u$ and let $L=\neip(u)\setminus \{u\}$ and $R=V\setminus L$. Now
 let $v,w\in L\setminus \{u\}$ and $x,y\in R$. From \autoref{fig:cc_twostrongcc} and the fact that
@@
 
-Those two clusters gives a natural partition with no disagreements in the sense of \pcc{}. However,
+Those two clusters represent a natural partition with no disagreements in the sense of \pcc{}. However,
 the existence of two such clusters is only a sufficient but not necessary condition for having no
@@
 sociologists that among triangles with odd number of negative edges, one is more stable and common
-than that the other one. Therefore, \textcite{davis1967clustering} relax the strong balance into the
+than that the other. Therefore, \textcite{davis1967clustering} relax the strong balance into the
 weak balance by considering triangles with three negative edges to be balanced, as illustrated on
@@
 
-In that case as well, the local property of triangle has structural implication for the whole graph.
+In that case as well, the local property of triangles has structural implication for the whole graph.
 Namely, we can partition a complete weakly balanced graph into $k$ clusters of positively connected
-nodes that are negatively with the rest of the graph. Consider the same construction as in
-\autoref{fig:cc_twostrongcc}, putting the positive neighborhood of a node $u$ in one cluster $L$ and
+nodes that are negatively tied with the rest of the graph. Consider the same construction as in
+\autoref{fig:cc_twostrongcc}, putting a node $u$ and its positive neighborhood in one cluster $L$ and
 the rest of the nodes in $R$. For $v,w\in L\setminus \{u\}$ and $x,y\in R$, we can still conclude
@@
 We gave the definitions of balance in the context of complete graphs in order to build intuition
-about the partition consequence and because it easier to reason on triangles. Furthermore, if we are
+about the partition consequences and because it easier to reason on triangles. Furthermore, if we are
 given a set of objects and a similarity function, we can indeed build a complete graph. However,
-they are situations were it is no realistic to make the completeness assumption, especially for
+they are situations were it is not realistic to make the completeness assumption, especially for
 social networks that are typically very sparse. We therefore extend the concept of balance to
@@
   the missing edges such that it becomes a complete graph with all its triangles weakly balanced. As
-  mentioned earlier, we can then pick a node $u_1$ and let $\cluster_1 = \neip(u) \cup \{u\}$. All
-  nodes in $\cluster_1$ are connected positively with each and negatively with $V\setminus
-  \cluster_1$. Note also that $|V\setminus \cluster_1| < |V|$ and therefore we can repeat this
+  mentioned earlier, we can then pick a node $u_1$ and let $C_1 = \neip(u) \cup \{u\}$. All
+  nodes in $C_1$ are connected positively with each other and negatively with nodes in $V\setminus
+  C_1$. Note also that $|V\setminus C_1| < |V|$ and therefore we can repeat this
   procedure until $V$ is exhausted, at which point we have obtained our $k$ clusters verifying the
@@
 repeatedly shown to be unbalanced, although the extent of this unbalance depends of the importance
-given to longer cycle compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
-{measureUnbalance17} For the same reason, in virtually any real instance of the \pcc{} problem, the%
+given to longer cycles compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}.%
+{measureUnbalance17} For the same reason, in virtually any real instances of the \pcc{} problem, the%
 % there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration
@@
 We can conclude that balance theory helps characterize zero disagreement instances of \pcc{}.
-However, how is \pcc{} useful in predicting edge sign? At first sight, the two problems seem quite
-different for the former is an unsupervised/agnostic clustering problem while the latter is a
+However, how is \pcc{} useful in predicting edge sign? At first sight, the two problems seem to be quite
+different, for the former is an unsupervised/agnostic clustering problem, while the latter is a
 supervised classification problem. Yet, getting an optimal solution of \pcc{} on the training set
 would provide a principled heuristics for \esp{}. Unfortunately, as we shall see in
-\autoref{sub:state_of_the_art}, exact or fixed parameter solution are not scalable and one has to
-rely approximation. Under the Unique Games Conjecture, the approximation factor is lower bounded by
+\autoref{sub:state_of_the_art}, exact or fixed parameter solutions are not scalable and one has to
+rely on approximations. Under the Unique Games Conjecture, the approximation factor is lower bounded by
 $\Theta(\log n)$ in the worst case. However, this is still a viable option, as our bias is that we
@@
 4.1]{Cesa-Bianchi2012b}. Before that, we will also present some variants and extensions of \pcc{},
-some of them bearing similarities with the Chapter 3's problem.
+in \autoref{sub:variants_and_extensions}, some of them bearing similarities with the Chapter 3's problem.
 

### 2017-08-16 cc/approx.tex
fix headers
some proofreading and small additions

--- 2017-08-09 cc_approx.tex
@@
-\subsection{Hardness and approximation}
-
 Although the same problem was studied earlier~\autocites{Early96}{Ben-Dor99}, \textcite{Bansal2002}
@@
 unweighted signed graphs, they show that both \mind{} and \maxa{} are \NPc{}. Along the way, they
-give a 17429-approximation of \mind{} and a PTAS that, for any $\epsilon \in [0,1]$, runs in
-\marginpars{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
+give a $17429$-approximation of \mind{} and a PTAS%
+\footnote{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
 minimization (respectively maximization) problem $\mathcal{P}$  in NP if given any $\epsilon>0$ and
-any instance $x$ of $\mathcal{P}$ of size $n$, $\mathcal{A}$ produces, in polynomial time in $n$, a
+any instance $x$ of $\mathcal{P}$ of size $n$, $\mathcal{A}$ produces, in time polynomial in $n$, a
 solution that is within a factor $1+\epsilon$ (respectively $1-\epsilon$) of being optimal with
@@
 time of $O(n^{\frac{1}{\epsilon}})$ would qualify~\autocite[Definition 3.10]{CpxBook99}.}
+that, for any $\epsilon \in [0,1]$, runs in
 $O(n^2e^{O(\frac{1}{\epsilon^{10}}\log\frac{1}{\epsilon})})$ and returns with probability
@@
 
-\begin{table}[htpb]
+\begin{table}[bt]
    \centering
@@
 on general graphs, for \mind{} by using a reduction from the multicut problem \autocite[Theorem
-8]{Charikar2003} (see \autocite{Emanuel2003} for the other direction and \autocite{Demaine2006} for
-unified proof?) which ask for the minimum weight set of edges whose removal in $G$ disconnect the
+8]{Charikar2003}\footnote{This reduction was actually turned into an equivalence, as we describe
+in more details in \autoref{ssub:cc_under_stability_assumption}.}
+which asks for the minimum weight set of edges whose removal in $G$ disconnect the
 $k$ pairs $(s_i, t_i)$ and for \maxa{} by a reduction from MAX 3SAT~\autocite[Theorem
@@
 cluster, then $(v, w)$ is also within that cluster. In terms of $x$ variable, we have that $x_{uv}=0
-\wedge x_{vw}=0 \implies x_{uw} = 0$. For $x$ to be a valid cluster assignment, we thus that all
-variable are either $0$ or $1$ and respect the triangle inequality. We can thus relax the problem into
+\wedge x_{vw}=0 \implies x_{uw} = 0$. For $x$ to be a valid cluster assignment, we thus require that all
+variables are either $0$ or $1$ and respect the triangle inequality. We can thus relax the problem into
 \begin{align}
@@
 \end{align}
-Once the LP \eqref{eq:mindLP} is solved, we interpret $x_{uv}$ as a distance: the larger it its and
-the more we want $u$ and $v$ to be in different cluster. We can then use the \emph{region growing}
-method~\autocite{RegionGrowing93}. Namely, we pick a node at random and grow a ball based on these
-distance until a threshold is met, at which point we create a cluster from this ball, remove the
-corresponding nodes from the graph and repeat until all nodes are clustered.
+Once the LP \eqref{eq:mindLP} is solved, we interpret $x_{uv}$ as a distance: the larger it is and
+the more we want $u$ and $v$ to be in different clusters. We can then use the \regionGrow{}
+method~\autocite{RegionGrowing93}. Namely, we pick a ball center $u$ \uar{} with radius \shalf{}: if
+the average distance of the nodes in the ball to $u$ is less than $\nicefrac{1}{4}$, the ball forms a
+cluster, otherwise $\{u\}$ forms a singleton cluster. We then remove the corresponding nodes from
+the graph and repeat until all nodes are clustered. As we shall see, this method has been re used
+with variations on the LP or on the two thresholds \shalf{} and $\nicefrac{1}{4}$.
 
@@
 edge carries two positive numbers: $w_{u,v}^+$ and $w_{u,v}^-$. If the weights respect the
-probability constraints stating that for all edge $(u,v)$ in $E$, $w_{u,v}^+ + w_{u,v}^- = 1$, they
-get a $2.5$-approximation. Note that unweighted graphs naturally fit into that case. If the weights
+probability constraints stating that for all edge $(u,v)$ in $E$, $w_{u,v}^+ + w_{u,v}^- = 1$, this
+tweaking provide a $2.5$-approximation. Note that unweighted graphs naturally fit into that case, as
+each edge is either labeled $+$ or $-$. If the weights
 additionally respect the triangular inequality constraints stating that $w_{u,v}^- \leq w_{u,w}^- +
@@
 
-This conclude the presentation of the known approximation results on \pcc{}, that we summarize in
+This concludes the presentation of the known approximation results on \pcc{}, that we summarize in
 \autoref{tab:cc_approx}. As mentioned earlier, not having to set the number of clusters is an
 attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.  The
-problem was studied by \textcite{Giotis2006} on general graph and we compiled their results in
-\autoref{tab:cc_fixed}. On complete unweighted graph and with $k$ being the number of clusters, they
-provide PTAS for \maxa{}$_k$ running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and for
-\mind{}$_k$ running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
+problem was studied by \textcite{Giotis2006} on general graphs and we compile their results in
+\autoref{tab:cc_fixed}. On complete unweighted graphs and with $k$ being the number of clusters, they
+provide PTASs for \maxa{}$[k]$ running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and for
+\mind{}$[k]$ running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
 \textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}k^6\log d\right)}$
-and that can handle weighted graphs. For $k=2$, \mind$_2$ admits a faster local search method with a
+and that can handle weighted graphs. For $k=2$, \mind$[2]$ admits a faster local search method with a
 factor $2$ approximation~\autocite{Coleman2008}.
 For complete weighted graph, \textcite{WeightedMaxAPTAS08} provide a PTAS for
-\maxa{}$_k$ under the condition that the ratio between the largest and smallest weights is bounded by a
+\maxa{}$[k]$ under the condition that the ratio between the largest and smallest weights is bounded by a
 constant.

### 2017-08-16 cc/multimincut_stable.tex
some proofreading and small additions

@@
 following~\autocite{Demaine2006} but omitting the full proofs to avoid redundancy. Recall that given
-an weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and sinks
+a weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and targets
 $S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, in \mmc{} we want to find the lightest set of edges $T\in E_H$
@@
 weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges
-are simulated in the following way: include every source and sink in a clique of $n$ nodes and
-connect that clique to the corresponding source or sink by $n$ negative edges.
+are simulated in the following way: include every source and target in a clique of $n$ nodes and
+connect that clique to the corresponding source or target by $n$ negative edges.
 
@@
 weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
-weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-sink pair added to $S_G$. This
+weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-target pair added to $S_G$. This
 construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
-4.4]{Demaine2006})\Todo{give the intuition of the proof} that it is easy to construct a multicut of
-weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.
+4.4]{Demaine2006})\Todo{give the intuition of the proof} that it takes a linear time to construct a
+multicut of weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.
 

### 2017-08-16 cc/proposed.tex
add communities detection variant of CC
some proofreading and small additions

@@
 We already described methods based on LP and SDP, as they have proved to be the best way to obtain
-tight approximation in the worst case. Here we survey other kind of approaches with different aims.
+tight approximations in the worst case. Here we survey other kind of approaches with different aims.
 
@@
 depending of the negative edge density, the FICO Xpress solver starts to be unable to finish within
-one hour time limit when $n \geq 40$.
+a one hour time limit when $n \geq 40$.
 \Textcite{Aref2016} describe four linear and quadratic  binary integer problems
@@
 software, they solve instances with up to $3200$ edges in less than a second. They also sketch
-extensions to weighted graph and more than two clusters.
+extensions to weighted graphs and more than two clusters.
 In a different direction, \textcite{Berg2015}, encode the linear and quadratic integer formulations
@@
 \eqref{eq:cc_lowrank} in $O(nk)$ time.
+On complete graphs, we may also turn on fixed-parameter tractable algorithms, where the parameter
+can be the number of disagreements, the number of clusters or the maximum number of disagreements
+per nodes: see \vpageref{par:cc_editing}.
 
 \paragraph{Physics inspired energy methods}
+\label{par:cc_physics}
 
+Here we describe a physics particle model that can readily be adapted to model \pcc{}.
 The Potts model~\autocite{PottsSurvey82} describes a general model of \emph{spins} organized in a
@@
 natural to formulate the clustering problem as a Hamiltonian minimization
-problem~\autocite{CommunityPhysics06}. Letting $A$ be the matrix such that $A_{uv} = w_{uv}^+ -
+problem~\autocite{CommunityPhysics06}.
+
+Letting $A$ be the matrix such that $A_{uv} = w_{uv}^+ -
 w_{uv}^-$ on a general directed graph, \textcite{Traag2009} reward positive and absent negative
@@
 distributed setting to obtain a scalable approach. Namely, given a node $u$, they want to output a
-globally consistent cluster index $\cluster(u)$ while making at most $t$ queries to an sign oracle.
+globally consistent cluster index $\cluster(u)$ while making at most $t$ queries to a sign oracle.
 Here $t$ is a parameter that depends on the quality of clustering produced but not on the size of
@@
 two techniques.
-The first one is inspired by \ccpivot{} and first find a good set of pivots by seeing the problem as
-independent set on a sampled part of the positive graph, then for a given node, find the closest
-such pivot or create a singleton. Given a quality parameter $\epsilon\in(0,1)$, it yields a $4\cdot
-OPT + \epsilon n^2$ approximation of \mind{} requiring $O(\frac{n}{\epsilon})$ time and
-queries~\autocite[Theorem 3.3]{Bonchi2013}
+The first one is inspired by \ccpivot{}. It starts by finding a good set of pivots, seeing the
+problem as finding a maximal independent set on a sampled part of the positive graph. Then, for a
+given node, it finds the closest such pivot or creates a singleton. Given a quality parameter
+$\epsilon\in(0,1)$, it yields a $4\cdot OPT + \epsilon n^2$ approximation of \mind{} requiring
+$O(\frac{n}{\epsilon})$ time and queries~\autocite[Theorem 3.3]{Bonchi2013}.
 Roughly stated, the second one relies on an existing low-rank approximation of the adjacency matrix,
@@
 connected is the graph. This matrix is typically used for clustering by computing its first $k$
-eigenvectors, which embeds the $n$ nodes of $G$ in $\Rbb^k$, where there are then clustered with the
+eigenvectors, which embed the $n$ nodes of $G$ in $\Rbb^k$, where there are then clustered with the
 $k$-means algorithm. This can be seen as a relaxation of the discrete \rcut{} objective, which asks
@@
 
-The first line of research consider only \mind{}$_2$. For instance, \textcite{NcutAnd2CC08} show
-that both normalized cut and \mind{}$_2$ objectives can be written as a SDP (or equivalently as
+The first line of research consider only \mind{}$[2]$. For instance, \textcite{NcutAnd2CC08} show
+that both normalized cut and \mind{}$[2]$ objectives can be written as a SDP (or equivalently as
 eigenvalue problems) and thus combined, the intuition being that we look for \ncut{} solutions whose
@@
 signed degree matrix such that $\bar{D}_{uu} = \sum_{v\in \nei(u)} |A_{uv}|$, as well as a signed
-variant of the symmetric normalized Laplacian $\bar{L}_{sym}  = \bar{D}^{\nicefrac{1}{2}} L
+variant of the symmetric normalized Laplacian $\bar{L}_{sym}  = \bar{D}^{\nicefrac{1}{2}} \bar{L}
 \bar{D}^{\nicefrac{1}{2}}$. They show that the signed Laplacian is positive semidefinite, and
-even positive-definite as soon as the graph is unbalanced (that is contains a cycle with a odd
+even positive-definite as soon as the graph is unbalanced (that is contains a cycle with an odd
 number of negative edges). From positive and negative cuts defined as $\mathrm{cut}^+(B, C) =
@@
 natural signed cut is $\mathrm{scut}(B, C) = 2\mathrm{cut}^+(B, C) + \mathrm{cut}^-(B, B) +
-\mathrm{cut}^-(C, C)$ which then be used to defined signed \rcut{} and \ncut{}. Arguing that those
+\mathrm{cut}^-(C, C)$ which can then be used to defined signed \rcut{} and \ncut{}. Arguing that those
 definitions force negatively linked nodes to be symmetric around the origin, do not take into
@@
 \end{equation*}
-
-% Kai-Yang Chiang, Joyce Jiyoung Whang, and Inderjit S. Dhillon. 2012. Scalable clustering of signed
-% networks using balance normalized cut. In CIKM '12. 615-624.
-% DOI=http://dx.doi.org/10.1145/2396761.2396841
+Additional cut formulations for $k$-clusters are also presented in~\autocite{moreSignedCut12}.
 
@@
 
-While all\marginpars{or most? not physic ones so move them afterwards} the methods we discuss so far
+While all the methods we discuss so far
 are either exact or come with some approximation guarantees, practitioners have also develop
@@
 clusters. The overall complexity is $O(nTP)$. This is quite similar to the \emph{Best One Element
-Move} described in~\autocite{Gionis2007}, except that the exchange operation is replaced by moving
+Move} described in~\autocite{Gionis2007}, except that there the exchange operation is replaced by moving
 one node to its own singleton cluster. The \emph{Cluster Affinity Search Technique}
-algorithm~\textcite{Ben-Dor99} instead grows clusters one by one by maintaining for the current
-cluster the affinity of all nodes, which is the sum of the weight between that node and the nodes in
+algorithm~\autocite{Ben-Dor99} instead grows clusters one by one by maintaining for the current
+cluster the affinity of all nodes, which is the sum of the weights between that node and the nodes in
 the cluster. Nodes above a certain threshold are added to the current cluster and nodes below the
 threshold are removed, with the affinity to the current cluster being recomputed after each addition
-or deletion. After defining the net weight of an edge to be $w^\pm_{uv} = w^+_{uv} - w^-_{uv}$,
+or deletion.
+
+After defining the net weight of an edge to be $w^\pm_{uv} = w^+_{uv} - w^-_{uv}$,
 \textcite{Elsner2009} describe three folklore heuristics that start with empty clusters and add node
@@
 those whose merge would result in the maximum improvement of the score function.  This runs in
-$O(n^2\log n)$, provides a $2$-approximation of \maxa{}, a $O(n\times \text{size of the largest
-cluster})$-approximation of \mind and empirically results in less disagreements than \ccpivot{}.
+$O(n^2\log n)$, provides a $2$-approximation of \maxa{}, a $O(n\cdot \text{size of the largest
+cluster})$-approximation of \mind{}, and empirically results in less disagreements than \ccpivot{}.
 Building upon their previous GRASP work~\autocite{GRASP13}, an iterated local search heuristic is
@@
 This local search followed by random perturbation is reminiscent of genetic algorithms, and such an
-approach is sketched in~\autocite{GeneticCC08}.\marginpars{but that's a terrible paper!} Instead of
+approach is sketched in~\autocite{GeneticCC08}.\footnote{but that's a terrible paper!} Instead of
 moving nodes, \textcite{restoreNeighborhood13} starts from the observation that in a balanced
-graphs, $G^+$ is a disjoint union of cliques in which all node a given cluster share the shame
+graph, $G^+$ is a disjoint union of cliques in which all the nodes of a given cluster share the shame
 neighbors. Their algorithm is initialized with the set $E_s^+$ of edges where $w^+_{uv} > w^-_{uv}$,
-and repeatedly sample an edge $(u,v)$ from $E_s^+$ before trying to make the neighborhoods of $u$
+and repeatedly samples an edge $(u,v)$ from $E_s^+$ before trying to make the neighborhoods of $u$
 and $v$ coincide by adding or removing edges in $E_s^+$.  Yet another idea is modify the LP

### 2017-08-16 cc/variants.tex
some proofreading and small additions

@@
-\paragraph{non binary edge labelling}
+% leftover:
+% - something about fuzzy edges, ie we're not sure they are in the graph or not (not super
+%   interesting in my opinion) Clustering with partial information, Theoretical Computer Science,
+%   Volume 411, Issue 7, 2010, Pages 1202-1211, ISSN 0304-3975,
+%   \url{http://dx.doi.org/10.1016/j.tcs.2009.12.016}
+% - \paragraph{Hypergraphs} \Textcite{Kim2011} show a LP relaxation on hypergraph and
+%   \textcite{Ricatte13} describe a class of hypergraphs that can be reduced to signed graph.  Jörg
+%   Hendrik Kappes, Markus Speth, Gerhard Reinelt, Christoph Schnörr, Higher-order segmentation via
+%   multicuts, Computer Vision and Image Understanding, Volume 143, 2016, Pages 104-119, ISSN
+%   1077-3142, \url{http://dx.doi.org/10.1016/j.cviu.2015.11.005}
+% - Interactive \pcc{}?~\autocite{interactiveCC16}, could it go into active?  Geerts, F. & Ndindi,
+%   R. Bounded correlation clustering. Int. J. Data Sci. Anal. 1, 17–35 (2016)
+%   \url{https://doi.org/10.1007/s41060-016-0005-2}
+
+So far we focused on \pcc{} in its rawest form, that is solving the \mind{} and \maxa{} objectives
+in the case where the general binary-labeled graph is known in advance. We will now see first some
+special cases, namely when the weights obey the triangle inequality (to solve \msc{}) or when the
+graph is bipartite and then move to variants. We will consider more general objectives, when the
+edges are labeled categorically instead of binary, when nodes can belong simultaneously to several
+cluster or when we optimize local objectives per nodes instead of global ones. We will also look at
+clustering in signed graphs in general, using spectral methods or heuristics from the community
+detection literature. Finally we will shift perspective and review methods beyond the batch case and
+the worst-case analysis.
+
+% \paragraph{\msc{}}                                | special weights
+% \paragraph{Bipartite \pcc{}}                      | special graphs
+% \paragraph{Categorical edge labelling}            | more general objective
+% \paragraph{Overlapping \pcc{}}                    | more general objective
+% \paragraph{Local \pcc{}}                          | more general objective or at least different
+% \paragraph{Spectral clustering}                   | related objective
+% \paragraph{Communities detection}                 | related objective
+% \paragraph{Online \& active setting}              | different setting
+% \paragraph{recovery under noise}                  | non worst case analysis
+% \subsubsection{\pcc{} under stability assumption} | non worst case analysis
+
+\paragraph{Bipartite \pcc{}}
+
+Bipartite graphs are an interesting special case for \pcc{}, as they often appear in the context of
+recommendation systems, where users rate products positively or negatively, although in this setting
+we cannot expect to have complete bipartite graphs in practice. The first results was given by
+\textcite{Amit04}, who obtains a $11$-approximation for \mind{} by adapting the \regionGrow{} method.
+The \ccpivot{} is adapted to the bipartite case by
+\textcite{Bipartite12}, who prove it results in a randomized $4$-approximation (and provide a matching
+deterministic approximation by rounding a LP). By using their idea of rounding the results of the LP
+differently for positive, negative and in that case same-side absent edges, \textcite{Chawla2014}
+bring down this approximation factor to $3$, even for $K\geq 2$-partite graphs. By formulating the
+\maxa{}$[k]$ problem as a bilinear maximization problem and computing a low-rank approximation of
+the graph biadjacency matrix, \textcite{Asteris2016} obtain a efficient PTAS, that is a $(1-\delta)$
+approximation running in time exponential in $k$ and $\delta^{-1}$ but linear in $n$. By an
+appropriate choice of $k$, it is possible to use this PTAS to solve the general \maxa{} problem.
+
+\paragraph{Categorical edge labelling}
 
 In the so called \textsc{Chromatic}-\pcc{} setting, \enquote{positive} edges are now associated with
-one of $L$ possible color and the goal is to form clusters mostly made up of edges with the one same
-color. Namely, a disagreement is now a negative edge between clusters or an within-cluster edge
+one of $L$ possible colors and the goal is to form clusters mostly made up of edges with the one same
+color. Namely, a disagreement is now a negative edge between clusters or a within-cluster edge
 whose color differs from the majority color of that cluster. This is motivated by edge-labeled graph
@@
 present a modification of the \ccpivot{} algorithm that pick edges instead of nodes as pivots, and
-grow clusters by adding monochromatic triangle. This gives an approximation factor of six times the
+grow clusters by adding monochromatic triangles. This gives an approximation factor of six times the
 maximum degree of the graph. They also present a method when the number $k$ of clusters is fixed
@@
 show this gives a $11$-approximation to the original problem. Furthermore, they also write a linear
-program and round it using the region growing technique of \textcite{Charikar2003} to obtain
-approximation factor of $4$. \Textcite{multiChromatic15} extend their work to the case were a single
-edge can carry a \emph{set} of labels and adapt their randomized algorithm so that the approximation
-factor is multiplied by the size of the input label set.
+program and round it using the \regionGrow{} method to obtain an
+approximation factor of $4$. \Textcite{multiChromatic15} extend this line of work to the case were a
+single edge can carry a \emph{set} of labels and adapt their randomized algorithm so that the
+approximation factor is multiplied by the size of the input label set.
 
@@
 In \msc{}, the goal is to output a clustering which best summarizes (or agrees with) the several
-given input clusterings of the same set of objects.  Motivations includes robustness --by using an
-ensemble of clusterings from diverse methods-- and privacy --if the clusterings were computed by
+given input clusterings of the same set of objects.  Motivations include robustness ---by using an
+ensemble of clusterings from diverse methods--- and privacy ---if the clusterings were computed by
 different parties each considering only a subset of the objects attributes. We can build the
@@
 \pcc{} where the weights obey the triangular inequality. They give a deterministic $3$-approximation
-using the region growing method of \textcite{Charikar2003}. Later \textcite{Bonizzoni2008} show that
+using the \regionGrow{} method. Later \textcite{Bonizzoni2008} show that
 the minimization version is \APXh{}, even when the input is made of three clusterings and give a
-combinatorial $\frac{4}{5}$-approximation for the maximization problem. Experimental evaluations are
+combinatorial $\nicefrac{4}{5}$-approximation for the maximization problem. Experimental evaluations are
 conducted by \textcite{Bertolacci07} and \textcite{Filkov08}. The former describe a scalable
@@
 unsampled nodes one by one.  Experiments confirm that the running time is greatly improved compared
-with the linear program methods while the resulting objective value is essentially the same. Note
-however that LP methods can be applied in practice thanks to some tricks~\autocite{ConsensusLP10}.
-% On the parameterized complexity of consensus clustering, Theoretical Computer Science, 2014,
-% \url{http://dx.doi.org/10.1016/j.tcs.2014.05.002}
+with the linear program methods while the resulting objective value is essentially the same. Note,
+however, that LP methods can be applied in practice thanks to some tricks~\autocite{ConsensusLP10}.
+If we have $k$ input clusterings $\cluster^1,\ldots,\cluster^k$ and we parameterized the problem by
+$t$, which is the sum over the input clusterings of the number of pairs of objects that are clustered
+differently by a solution $\cluster^\star$ and $\cluster^i$, then there is a polynomial algorithm
+running in $O(4.24^{\nicefrac{t}{k}}\cdot \nicefrac{t}{k}^3 +
+kn^2)$~\autocite{parameterizedConsensus14}.
 
@@
 
+In classical \pcc{}, all nodes have an identical role, in the sense that they contribute equally to
+the final objective in terms of (dis)agreements. Here we instead look at approaches where we either
+add a local penalty to each in order to better control their behavior or where we altogether modify
+the objective to focus on (dis)agreements at specific nodes.
+
 % that one is not really local, except for the node penalty
-\Textcite{Puleo2014} adapt the linear program of~\autocite{Charikar2003} (and its rounding by the
-region growing method. Namely after solving the LP, repeatedly pick a ball center $u$ \uar{} with
-radius \shalf{}: if the average distance of the nodes in the ball to $u$ is less than
-\nicefrac{1}{4}, the ball forms a cluster, otherwise $\{u\}$ forms a singleton cluster.) to the case
+\Textcite{Puleo2014} adapt the linear program of~\autocite{Charikar2003} and its \regionGrow{}
+method to the case
 where all clusters have to contain less than $K$ nodes, by assigning to each node $u$ a penalty
@@
 randomized $7$-approximation.
-Those soft constraints are for instance used in biological application 
-\emph{A new correlation clustering method for cancer mutation analysis}
-\url{https://arxiv.org/abs/1601.06476}
-\autocite{Hou2016}
+These soft constraints are for instance used in a biological application where nodes are genes and
+where singleton and giant clusters are uninformative~\autocite{Hou2016}
 
@@
 
-Given a function $f: \Rbb^n_{\leq 0} \rightarrow \Rbb$ verifying two elementary conditions, the
+Given a function $f: \Rbb^n_{\geq 0} \rightarrow \Rbb$ verifying two elementary conditions, the
 problem is then to find a fractional clustering $x$ minimizing $f(err(x))$. The classical \pcc{}
@@
 disagreements incurred by a single node is motivated by the example of recommendation systems: if
-errors corresponds to unsatisfying recommendations, we do not want a single user to suffer a large
+errors correspond to unsatisfying recommendations, we do not want a single user to suffer a large
 number of them. Minimax \pcc{} is \NPc{} on both complete graphs and complete bipartite graphs but
-by modifying the region growing method of \textcite{Charikar2003}, they give respectively a $48$ and
-$10$ approximation, the latter for the one-sided error (that consider the nodes in only one of the
-two clusters). The idea is to chose pivot not randomly but by maximizing a given criteria and
-to grow ball with a ratio $\alpha$ computed numerically to optimize the approximation factor.
+by modifying the \regionGrow{} method, the authors respectively a $48$ and
+$10$ approximation, the latter for the one-sided error (that only counts disagreements for the nodes
+in one of the two clusters).
+The idea is to chose pivot not randomly but by maximizing a given criteria and
+to grow balls with a radius $\alpha$ computed numerically to optimize the approximation factor.
 Interestingly, and in contrast with the classic \pcc{} situation, minimax \maxa{} is not easier than
 minimax \mind{} and seems not to have a constant factor approximation, even on complete graphs.
-Furthermore these algorithms are deterministic, as opposed to many \pcc{} approximation, since
-bounds on expected disagreements on a edge does not translate easily on their maximum.
+Furthermore, these algorithms are deterministic, as opposed to many \pcc{} approximation, since
+bounds on expected disagreements of an edge does not translate easily on their maximum.
 \Textcite{Charikar2017} improve these two factors to $7$, using a simpler version of the algorithm
@@
 
-\paragraph{recovery under noise}
+\paragraph{\pcc{} under noise}
 
-As we have seen, assuming the Unique Games Conjecture, the minimum Multicut problem, and therefore
-the \pcc{} problem on a general graph, cannot be approximated to a constant factor in the worst
+As we have seen, assuming the Unique Games Conjecture, the \mmc{} problem, and therefore
+the \pcc{} problem on a general graph, cannot be approximated to within a constant factor in the worst
 case.  But maybe we can do better in the average case, which motivates the study of semi-random
@@
 
-\Textcite{Joachims2005} analyze a more refined weighted model where weights are generated by a
+\Textcite{Joachims2005} analyze a more refined weighted model, where weights are generated by a
 probability distribution whose mean on true positive edges is larger than $\mu^+>0$ and whose mean
-on true negative edges is smaller than $\mu^-<0$. They give a finite-sample bounds on the number of
+on true negative edges is smaller than $\mu^-<0$. They give a finite-sample bound on the number of
 node misclustered w.r.t the planted partition as a function on the probability distribution
@@
 argue it is more sensible to measure it against the (unknown) true optimal clustering that gave rise
-to the output, and show that \ccpivot{} allows that thanks to a new analysis. \Textcite{Mathieu2010}
+to the input, and show that \ccpivot{} allows that thanks to a new analysis. \Textcite{Mathieu2010}
 also consider an adversarial model, where all edges are flipped with probability $p$ but the
@@
 $\Theta(\sqrt{n})$ nodes, then the planted partition can be recovered exactly.
-\Textcite{Makarychev2014} study the same adversarial model on general weighted graphs, giving a PTAS
+\Textcite{Makarychev2014} study the same adversarial model but on general weighted graphs, giving a PTAS
 for \mind{} when $p\leq \nicefrac{1}{4}$. Under additional assumptions on the density of edges, they
 present another algorithm that finds the ground truth clustering with an arbitrarily small
-classification error.
-% http://www.jmlr.org/papers/v15/chen14a.html
+classification error.\footnote{They also mention a potentially interesting related
+work~\autocite{unobservedCC14}.} % http://www.jmlr.org/papers/v15/chen14a.html
 
@@
 
-Case in point, in the context of entities resolution, \textcite{activeCoref07} consider the \pcc{}
+Case in point, in the context of entity resolution, \textcite{activeCoref07} consider the \pcc{}
 problem where by querying an oracle, one can either reduce the uncertainty about one edge weight or
-add an extra node with edges connecting it to existing nodes. However, this requires web queries are
-resource bounded and thus yields an active problem, where one has to choose the most informative
+add an extra node with edges connecting it to existing nodes. However, this requires web queries
+that are resource bounded and thus yields an active setting learning problem, where one has to
+choose the most informative
 queries given a budget. A formal definition is given in~\autocite{queryMoreEdgeCC07} and in
@@
 oracle that for $u,v\in V^2$ returns $\cluster(u) - \cluster(v) \mod k$ with probability $1-p$ and a
-noisy $\cluster(u) - \cluster(v) \pm 1 \mod k$ otherwise. For $k=2$, this is exactly $2$-\pcc{}, for
-$k>2$, the oracle gives more information than simply the sign of the path from $u$ to $v$. Whenever
+noisy $\cluster(u) - \cluster(v) \pm 1 \mod k$ otherwise. For $k=2$, this is exactly $2$-\pcc{},
+whereas for $k>2$, the oracle gives more information than simply the sign of the path from $u$ to
+$v$. Whenever
 $p < \shalf$, they show that $O(n^{1+\frac{1}{\log\log n}}\log n)$ queries are enough are to recover
@@
 agreements. For \mind{}, this is $O(n)$-competitive algorithm and they show such ratio is optimal by
-exhibiting an instance\footnote{two positive cliques $A$ and $B$ joined by positive edges except
-between $a\in A$ and $\{b_1,\ldots, b_k\}\in B$. Those nodes are given first and thus form a cluster
-which yield at least one disagreement every time one the $n-(k+1)$ remaining nodes is added.} on
+exhibiting an instance\footnote{Namely two positive cliques $A$ and $B$ joined by positive edges except
+between $a\in A$ and $\{b_1,\ldots, b_k\}\in B$. Those nodes are given first and thus form a cluster,
+which yields at least one disagreement every time one the $n-(k+1)$ remaining nodes is added.} on
 which any strategy ends up with $n - k$ disagreements whereas optimal cost is $k$. On the \maxa{}
-side, this greedy strategy result in a $0.5$-competitive algorithm. If it is randomly mixed with a
-\textsc{Dense} variation, it increases to $0.5+\eta$, still far from the demonstrated $0.834$ upper
+side, this greedy strategy result in a $\shalf$-competitive algorithm. If it is randomly mixed with a
+\textsc{Dense} variation, it increases to $\shalf+\eta$, still far from the demonstrated $0.834$ upper
 bound.
@@
 \paragraph{\textsc{Cluster editing}}
+\label{par:cc_editing}
 
@@
 
-If we parameterize the problem by the number $d$ of edges that need to be added or deleted (that is
+If we parameterize the problem by the number $d$ of edges that need to be added or deleted (that is,
 the number of disagreements), then it can be solved in polynomial time in the size of the input
 graph (but not in $d$). The best known approximation so far is
-$O(1.62^d+m+n)$~\autocite{GoldenCE12}, which search for conflict triple (\ie{} unbalanced triangle
+$O(1.62^d+m+n)$~\autocite{GoldenCE12}, which search for conflict triples (\ie unbalanced triangles
 in the sign language) and branch by either deleting or merging one positive edge. If we additionally
@@
 edges can be deleted at each node, and if the minimum size of a cluster is at least $2(a+b)$, then
-the problem can be solved in polynomial time~\autocite{Abu-Khzam2015}. Finally, On planar graph,
+the problem can be solved in polynomial time~\autocite{Abu-Khzam2015}. Finally, for planar graph,
 there is a PTAS running in $O(n2^{\epsilon^{-1}\log(\epsilon^{-1})})$ obtained by dividing the graph
-into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}
+into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}.
 
-\paragraph{Communities detection}
+\paragraph{Community detection}
 
-The clustering problem is often named communities detection in the context of social networks, and
+The clustering problem is often named community detection in the context of social networks, and
 several methods developed by practitioners have been extended to signed graphs. While they do not
@@
 cluster of node $u$, \textcite{Yang2007} use a random walk approach on the positive subgraph $G^+$
-to compute the probability of each node to reach $u$ in $k$ steps, sort the nodes accordingly and
-then find a threshold  based on number of disagreements.
-The one node moves local heuristic that we described in the Physics inspired paragraph
+to compute the probability of each node to reach $u$ in $T$ steps, sort the nodes accordingly and
+then find a threshold  based on the number of disagreements.
+The one node moves local heuristic that we described in the Physics-inspired paragraph
 \vpageref{par:cc_physics} can also be formulated as genetic algorithms that simultaneously try to
@@
 probabilistic mixture model. Namely, an edge selects a pair of cluster $r,s$ with probability
-$\omega_{rs}$ (where $r=s$ if the edge is positive and $r\neq s$ otherwise) and choose its endpoint
+$\omega_{rs}$ (where $r=s$ if the edge is positive and $r\neq s$ otherwise) and chooses its endpoints
 $u$ and $v$ with probability $\theta_{ru}$ and $\theta_{sv}$. $\theta_{ru}$ is therefore the soft

### 2017-08-16 cc/proposed.tex
move sections around

@@
 can be the number of disagreements, the number of clusters or the maximum number of disagreements
-per nodes: see \vpageref{par:cc_editing}.
+per nodes, as we will now see.
 
-\paragraph{Physics inspired energy methods}
-\label{par:cc_physics}
+\paragraph{\textsc{Cluster editing}}
+\label{par:cc_editing}
 
-Here we describe a physics particle model that can readily be adapted to model \pcc{}.
-The Potts model~\autocite{PottsSurvey82} describes a general model of \emph{spins} organized in a
-lattice and being in one of $k$ possible states. A spin $u$ interacts with each of its neighbors $v$
-through a coupling $J_{uv}$. The energy of this system, called the Hamiltonian, is defined by
-$\mathcal{H}(\bm{\sigma}) = -\sum_{u,v} J_{uv} \delta(\sigma_u, \sigma_v)$ where $\bm{\sigma}$ is
-the spin configuration (that is, $\sigma_u \in \{1,\ldots,k\}\,\forall u$) and $\delta(\sigma_u,
-\sigma_v)$ is equal to one if $u$ and $v$ are in the same state and zero otherwise. It is a general
-principle of physics that isolated systems tend to minimize their energy, which in this case amounts
-to finding a spin configuration minimizing the Hamiltonian. Viewing the spins as nodes of a graph,
-the couplings as the graph weighted edges and the $k$ possible states as clusters, it is quite
-natural to formulate the clustering problem as a Hamiltonian minimization
-problem~\autocite{CommunityPhysics06}.
+% \url{http://www.sciencedirect.com/science?_ob=ArticleListURL&_method=list&_ArticleListID=-1231039636&_sort=r&_st=4&md5=018b782f17918959c0cc389c33ba2cd7&searchtype=a}
 
-Letting $A$ be the matrix such that $A_{uv} = w_{uv}^+ -
-w_{uv}^-$ on a general directed graph, \textcite{Traag2009} reward positive and absent negative
-links within cluster and penalize negative and absent positive links across clusters to come up with
-the following Hamiltonian: $\mathcal{H}(\bm{\sigma}) = -\sum_{u,v} \left(A_{uv}-(\gamma^+p^+_{uv} -
-\gamma^-p^-_{uv})\right) \delta(\sigma_u, \sigma_v)$ where $\gamma^+$ and $\gamma^-$ are user
-parameters and $p^\pm_{uv}$ are null model probabilities, which are equal to $p^\pm_{uv} =
-\frac{|E^\pm|}{|V|(|V|-1)}$, or $p^\pm_{uv}=\frac{\dout^\pm(u) \din^\pm(v)}{|E^\pm|}$ in the degree
-corrected model. They note that setting $\gamma^+ = 0 = \gamma^-$ make minimizing the Hamiltonian
-equivalent to the \mind{} problem, which they do by simulated
-annealing~\autocite{SimulatedAnnealing83}. The idea is to start from a random partition, and jump to
-another partition by moving a single node from one cluster to another. Such moves are made with a
-probability proportional to how each move reduces the Hamiltonian and as the procedure goes on, large
-jumps are made less and less probable by reducing a parameter called the temperature.
-One advantage of this energy formulation is that it requires only one variable per node instead of
-one variable per edge as in the case of the linear program.
+In the cluster editing problem, one is given a general unsigned input graph $H=(V,E)$ and wants to
+find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
+union of cliques. This is equivalent to \pcc{} on complete graphs. To see why, let $G=(V,(E^+,E^-))$
+be a complete signed graph with nodeset $V$, $E^+$ be the set $E$ of edges of $H$ and $E^-$ be all
+the edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
+disagreements within clusters are edges added to $H$ and the positive disagreements between clusters
+are deleted from $H$.  To the best of our knowledge, the problem was first introduced under this
+name by \textcite{Shamir02}.\marginpars{However, the same problem was studied before and we refer the
+reader to the comprehensive survey of \textcite{clusterEditSurvey13} for additional details,
+whereas we shall only give important and more recent pointers here.} They show the problem is
+\NPc{}, even if the number of clusters $k\geq2$ is set beforehand\footnote{The reduction is from
+3-exact 3 cover, see \autocite[Theorems 1, 2 and Corollary 1]{Shamir02}.} and provide a
+$0.878$-approximation in the weighted $k=2$ case using the standard Goemans-Williamson SDP
+relaxation.
 
-In the case of $2$-\pcc{}, \textcite{Facchetti2011isingmodel} rewrite the Hamiltonian as
-$\mathcal{H}(\bm{\sigma}) = -\frac{1}{2}\bm{\sigma}^T A \bm{\sigma} = -\frac{1}{2}\bm{1}^TT_\sigma A
-T_\sigma\bm{1}$. There $T_\sigma = diag(\bm{\sigma})$ (where $\bm{\sigma} \in \{0,1\}^n$) is the
-outcome of a local search algorithm such that $A_\sigma = T_\sigma A T_\sigma$ has the same number
-of disagreements as $A$ but the least number of negative edges. This is called a gauge
-transformation in the spin glass literature and the benefit of that heuristic is that it scales
-gracefully to large graphs.
-\Textcite{Bagon2011} also write the \maxa{} objective as a Potts model, and show that it can be
-interpreted as the log posterior of a partition matrix under a simple generative model and as a
-pair-wise conditional random field energy without unary terms. This allows them to adapt existing
-discrete energy optimization algorithms in order to cope with the following three challenges of the
-\pcc{} energy: \enquote{(i) the energy is non sub-modular, (ii) the number of clusters is not known
-in advance, and (iii) there is no unary term}. Doing so, the are able to handle large problems with
-more than 100K nodes.  Also adopting an energy minimization approach, \textcite{Kappes2016} assign a
-probability to each cut of a signed graph proportional to the exponential of the number of
-disagreements of that cut. They also develop efficient cut sampling methods.
-
-% something about local minima when defining another energy function, and how it is achieved by a
-% dynamic model but to be honest it has little to do with clustering {Marvel2009landscape}
-% Kappes2016 \enquote{Furthermore, due to the lack of an external field (unary terms), any permutation of an
-% optimal assignment results in another optimal labeling.}
-% like an ICML'17 paper that touches something very related (multicut) and gives recent applications in
-% vision https://arxiv.org/abs/1503.03791 represents a cut by the set of interclusters edges
+If we parameterize the problem by the number $d$ of edges that need to be added or deleted (that is,
+the number of disagreements), then it can be solved in polynomial time in the size of the input
+graph (but not in $d$). The best known approximation so far is
+$O(1.62^d+m+n)$~\autocite{GoldenCE12}, which search for conflict triples (\ie unbalanced triangles
+in the sign language) and branch by either deleting or merging one positive edge. If we additionally
+look for exactly $k$ clusters (\ie \mind$[k]$), there is a fixed parameters algorithm running in
+$O(2^{O(\sqrt{kd})}+n+m)$~\autocite{Fomin2014}. If at most $a$ edges can be added and at most $b$
+edges can be deleted at each node, and if the minimum size of a cluster is at least $2(a+b)$, then
+the problem can be solved in polynomial time~\autocite{Abu-Khzam2015}. Finally, for planar graph,
+there is a PTAS running in $O(n2^{\epsilon^{-1}\log(\epsilon^{-1})})$ obtained by dividing the graph
+into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}.
 
@@
 
-\paragraph{Spectral Clustering}
-
-A classic method for clustering graphs is to leverage their spectral properties. Namely, if $A$ is
-the adjacency matrix of $G$ and $D$ its degree diagonal matrix (that is $D_{u,u} = \degr(u)$), the
-\emph{Laplacian} of $G$, defined by $L_G = D - A$, is a symmetric positive semidefinite matrix. As
-such, it has $n$ real non-negative eigenvalues, and its spectrum provides additional information on
-the connectivity of $G$. For instance, $0$ is always the smallest eigenvalue and its multiplicity is
-equal to number of connected components of $G$, while ---if $G$ is connected--- the second
-eigenvalue is the algebraic connectivity of $G$, whose magnitude is an indication of how well
-connected is the graph. This matrix is typically used for clustering by computing its first $k$
-eigenvectors, which embed the $n$ nodes of $G$ in $\Rbb^k$, where there are then clustered with the
-$k$-means algorithm. This can be seen as a relaxation of the discrete \rcut{} objective, which asks
-for the partition $\{C_1, \ldots, C_k\}$ minimizing $\frac{1}{2}\sum_{i=1}^k \frac{\mathrm{cut}(C_i,
-\bar{C_i})}{|C_i|}$, where $\bar{C_i}$ is the complement of $C_i$ in $V$ and $\mathrm{cut}(B, C) =
-\sum_{u\in B, v\in C} w_{uv}$ is the total weight of the edges between $B$ and
-$C$~\autocite{tutoSpectralClustering07}. By considering the symmetric normalized Laplacian
-$L_{sym}  = D^{\nicefrac{1}{2}}LD^{\nicefrac{1}{2}}$, it is also possible to approximate the
-normalized cut objective (\ncut{}), where $|C_i|$ is replaced by $vol(C_i) = \sum_{u\in C_i}
-\degr(u)$. We will now see how these kinds of approaches can be extended to signed graphs, noting
-first that they require to fix the number of clusters beforehand and are looking for clusters
-balanced in size, which makes the problem related but not equivalent to \pcc{}.
-
-The first line of research consider only \mind{}$[2]$. For instance, \textcite{NcutAnd2CC08} show
-that both normalized cut and \mind{}$[2]$ objectives can be written as a SDP (or equivalently as
-eigenvalue problems) and thus combined, the intuition being that we look for \ncut{} solutions whose
-number of disagreements is not too much more than the approximate optimal.
-
-Letting \ncut{}$(C, \bar{C}) = \frac{\mathrm{cut}(C, \bar{C})}{\mathrm{bal}(C)}$ with
-$\mathrm{bal}(C) = 2\frac{vol(C)vol(\bar{C})}{vol(V)}$, \textcite{mOneCC12} define a new objective:
-\begin{equation*}
-  \hat{F}_\gamma(C) = \frac{\mathrm{cut}(C, \bar{C}) + \gamma\left(\hat{M}(C)+\hat{N}(C)\right)}{\mathrm{bal}(C)}
-\end{equation*}
-where $\gamma \in \Rbb{}_+$ is a parameter, while $\hat{M}(C)$ and $\hat{N}(C)$ are respectively the
-number of positive and negative disagreements of the $(C, \bar{C})$ clustering.
-They show how to optimize a tight continuous relaxation of $\hat{F}_\gamma$ as the non negative
-ratio of a difference of convex function and a convex function.
-
-On the other hand, one can also adapt these two cut objectives to directly include negative edges.
-\Textcite{Luca10} define the signed Laplacian as $\bar{L} = \bar{D} - A$, where $\bar{D}$ is the
-signed degree matrix such that $\bar{D}_{uu} = \sum_{v\in \nei(u)} |A_{uv}|$, as well as a signed
-variant of the symmetric normalized Laplacian $\bar{L}_{sym}  = \bar{D}^{\nicefrac{1}{2}} \bar{L}
-\bar{D}^{\nicefrac{1}{2}}$. They show that the signed Laplacian is positive semidefinite, and
-even positive-definite as soon as the graph is unbalanced (that is contains a cycle with an odd
-number of negative edges). From positive and negative cuts defined as $\mathrm{cut}^+(B, C) =
-\sum_{u\in B, v\in C} w^+_{uv}$ and $\mathrm{cut}^-(B, C) = \sum_{u\in B, v\in C} w^-_{uv}$, a
-natural signed cut is $\mathrm{scut}(B, C) = 2\mathrm{cut}^+(B, C) + \mathrm{cut}^-(B, B) +
-\mathrm{cut}^-(C, C)$ which can then be used to defined signed \rcut{} and \ncut{}. Arguing that those
-definitions force negatively linked nodes to be symmetric around the origin, do not take into
-account the balance of negative edges in each cluster and are difficult to extend to more than two
-clusters, \textcite{SignedEmbedding15} instead propose two new normalized cuts:
-\begin{align*}
-  SNScut(C_1, \ldots, C_k) &= \sum_{i=1}^k
-  \frac{\mathrm{cut}^+(C_i, \bar{C_i})-\mathrm{cut}^-(C_i, \bar{C_i})}{vol(C_i)}  \\
-  BNScut(C_1, \ldots, C_k) &= \sum_{i=1}^k
-  \frac{\mathrm{cut}^+(C_i, \bar{C_i})-\mathrm{cut}^-(C_i, \bar{C_i})+vol^-(C_i)}{vol(C_i)}
-\end{align*}
-
-Noting that if $x_i\in\Rbb^n$ is the vector indicator of cluster $C_i$ (that is the \uth{} entry of
-$x_i$ is $1$ is $u$ belongs to $C_i$ and $0$ otherwise), $x_i^T\bar{L}x_i = 2\mathrm{cut}^-(C_i,
-C_i) + \mathrm{cut}^-(C_i, \bar{C}_i) + \mathrm{cut}^+(C_i, \bar{C}_i)$,
-\textcite{mSemanticWordCC17} introduce the following cut objective:
-\begin{equation*}
-  sNcut(C_1, \ldots, C_k) = \sum_{i=1}^k
-  \frac{2\mathrm{cut}^-(C_i, C_i) + \mathrm{cut}^-(C_i, \bar{C}_i) + \mathrm{cut}^+(C_i, \bar{C}_i)}{vol(C_i)}
-\end{equation*}
-Additional cut formulations for $k$-clusters are also presented in~\autocite{moreSignedCut12}.
-
-Finally, \textcite{mGeometricMean16} show that the Laplacians defined so far can be seen as
-arithmetic means of the Laplacian $L^+$ of the positive subgraph $G^+=(V, E^+)$ and the signless
-Laplacian $Q^-$ of the negative subgraph $G^-=(V, E^-)$, where $Q^- = D^- + A^-$. They suggest
-instead to use a geometric mean defined for two positive matrices $A$ and $B$ as $A\#B =
-A^{\shalf{}} \left(A^{-\shalf{}} BA^{-\shalf{}} \right)^{\shalf{}} A^{\shalf{}}$. This suggestion is
-based on the fact that if $u$ is a common eigenvector of both $A$ and $B$ with eigenvalue $\lambda$
-and $\mu$ respectively, then $u$ is an eigenvector of $A+B$ with eigenvalue $\lambda+\mu$ and an
-eigenvector of $A\#B$ with eigenvalue $\sqrt{\lambda\mu}$. Therefore, the $k$ smallest eigenvalues
-of the geometric mean Laplacian will be influenced by both smallest eigenvalues of $L^+$
-(corresponding to assortative clusters in $G^+$) and of $Q^-$ (corresponding to disassortative
-clusters in $G^-$), while this is note the case for the arithmetic mean of Laplacians.
-
 \paragraph{Heuristic methods}
@@
 alternate optimization on $L$ and $\cluster$.
+
+\paragraph{Physics inspired energy methods}
+\label{par:cc_physics}
+
+Here we describe a physics particle model that can readily be adapted to model \pcc{} and provide
+additional heuristic methods.
+The Potts model~\autocite{PottsSurvey82} describes a general model of \emph{spins} organized in a
+lattice and being in one of $k$ possible states. A spin $u$ interacts with each of its neighbors $v$
+through a coupling $J_{uv}$. The energy of this system, called the Hamiltonian, is defined by
+$\mathcal{H}(\bm{\sigma}) = -\sum_{u,v} J_{uv} \delta(\sigma_u, \sigma_v)$ where $\bm{\sigma}$ is
+the spin configuration (that is, $\sigma_u \in \{1,\ldots,k\}\,\forall u$) and $\delta(\sigma_u,
+\sigma_v)$ is equal to one if $u$ and $v$ are in the same state and zero otherwise. It is a general
+principle of physics that isolated systems tend to minimize their energy, which in this case amounts
+to finding a spin configuration minimizing the Hamiltonian. Viewing the spins as nodes of a graph,
+the couplings as the graph weighted edges and the $k$ possible states as clusters, it is quite
+natural to formulate the clustering problem as a Hamiltonian minimization
+problem~\autocite{CommunityPhysics06}.
+
+Letting $A$ be the matrix such that $A_{uv} = w_{uv}^+ -
+w_{uv}^-$ on a general directed graph, \textcite{Traag2009} reward positive and absent negative
+links within cluster and penalize negative and absent positive links across clusters to come up with
+the following Hamiltonian: $\mathcal{H}(\bm{\sigma}) = -\sum_{u,v} \left(A_{uv}-(\gamma^+p^+_{uv} -
+\gamma^-p^-_{uv})\right) \delta(\sigma_u, \sigma_v)$ where $\gamma^+$ and $\gamma^-$ are user
+parameters and $p^\pm_{uv}$ are null model probabilities, which are equal to $p^\pm_{uv} =
+\frac{|E^\pm|}{|V|(|V|-1)}$, or $p^\pm_{uv}=\frac{\dout^\pm(u) \din^\pm(v)}{|E^\pm|}$ in the degree
+corrected model. They note that setting $\gamma^+ = 0 = \gamma^-$ make minimizing the Hamiltonian
+equivalent to the \mind{} problem, which they do by simulated
+annealing~\autocite{SimulatedAnnealing83}. The idea is to start from a random partition, and jump to
+another partition by moving a single node from one cluster to another. Such moves are made with a
+probability proportional to how each move reduces the Hamiltonian and as the procedure goes on, large
+jumps are made less and less probable by reducing a parameter called the temperature.
+One advantage of this energy formulation is that it requires only one variable per node instead of
+one variable per edge as in the case of the linear program.
+
+In the case of $2$-\pcc{}, \textcite{Facchetti2011isingmodel} rewrite the Hamiltonian as
+$\mathcal{H}(\bm{\sigma}) = -\frac{1}{2}\bm{\sigma}^T A \bm{\sigma} = -\frac{1}{2}\bm{1}^TT_\sigma A
+T_\sigma\bm{1}$. There $T_\sigma = diag(\bm{\sigma})$ (where $\bm{\sigma} \in \{0,1\}^n$) is the
+outcome of a local search algorithm such that $A_\sigma = T_\sigma A T_\sigma$ has the same number
+of disagreements as $A$ but the least number of negative edges. This is called a gauge
+transformation in the spin glass literature and the benefit of that heuristic is that it scales
+gracefully to large graphs.
+\Textcite{Bagon2011} also write the \maxa{} objective as a Potts model, and show that it can be
+interpreted as the log posterior of a partition matrix under a simple generative model and as a
+pair-wise conditional random field energy without unary terms. This allows them to adapt existing
+discrete energy optimization algorithms in order to cope with the following three challenges of the
+\pcc{} energy: \enquote{(i) the energy is non sub-modular, (ii) the number of clusters is not known
+in advance, and (iii) there is no unary term}. Doing so, the are able to handle large problems with
+more than 100K nodes.  Also adopting an energy minimization approach, \textcite{Kappes2016} assign a
+probability to each cut of a signed graph proportional to the exponential of the number of
+disagreements of that cut. They also develop efficient cut sampling methods.
+
+% something about local minima when defining another energy function, and how it is achieved by a
+% dynamic model but to be honest it has little to do with clustering {Marvel2009landscape}
+% Kappes2016 \enquote{Furthermore, due to the lack of an external field (unary terms), any permutation of an
+% optimal assignment results in another optimal labeling.}
+% like an ICML'17 paper that touches something very related (multicut) and gives recent applications in
+% vision https://arxiv.org/abs/1503.03791 represents a cut by the set of interclusters edges

### 2017-08-16 cc/variants.tex
move sections around

@@
 % \subsubsection{\pcc{} under stability assumption} | non worst case analysis
+
+\paragraph{\msc{}}
+
+In \msc{}, the goal is to output a clustering which best summarizes (or agrees with) the several
+given input clusterings of the same set of objects.  Motivations include robustness ---by using an
+ensemble of clusterings from diverse methods--- and privacy ---if the clusterings were computed by
+different parties each considering only a subset of the objects attributes. We can build the
+complete graph of these objects, with weights set to the fraction of clusterings that place two
+objects in different clusters, thus representing a kind of distance in the space of clusterings.  As
+first show by \textcite{Gionis2007}, finding the optimal clustering is therefore an instance of
+\pcc{} where the weights obey the triangular inequality. They give a deterministic $3$-approximation
+using the \regionGrow{} method. Later \textcite{Bonizzoni2008} show that
+the minimization version is \APXh{}, even when the input is made of three clusterings and give a
+combinatorial $\nicefrac{4}{5}$-approximation for the maximization problem. Experimental evaluations are
+conducted by \textcite{Bertolacci07} and \textcite{Filkov08}. The former describe a scalable
+approach that first samples a small portion of the data, runs a (potentially computationally
+expensive) approximation algorithm and finally augment the resulting partition by adding to it the
+unsampled nodes one by one.  Experiments confirm that the running time is greatly improved compared
+with the linear program methods while the resulting objective value is essentially the same. Note,
+however, that LP methods can be applied in practice thanks to some tricks~\autocite{ConsensusLP10}.
+If we have $k$ input clusterings $\cluster^1,\ldots,\cluster^k$ and we parameterized the problem by
+$t$, which is the sum over the input clusterings of the number of pairs of objects that are clustered
+differently by a solution $\cluster^\star$ and $\cluster^i$, then there is a polynomial algorithm
+running in $O(4.24^{\nicefrac{t}{k}}\cdot \nicefrac{t}{k}^3 +
+kn^2)$~\autocite{parameterizedConsensus14}.
 
@@
 
-\paragraph{\msc{}}
-
-In \msc{}, the goal is to output a clustering which best summarizes (or agrees with) the several
-given input clusterings of the same set of objects.  Motivations include robustness ---by using an
-ensemble of clusterings from diverse methods--- and privacy ---if the clusterings were computed by
-different parties each considering only a subset of the objects attributes. We can build the
-complete graph of these objects, with weights set to the fraction of clusterings that place two
-objects in different clusters, thus representing a kind of distance in the space of clusterings.  As
-first show by \textcite{Gionis2007}, finding the optimal clustering is therefore an instance of
-\pcc{} where the weights obey the triangular inequality. They give a deterministic $3$-approximation
-using the \regionGrow{} method. Later \textcite{Bonizzoni2008} show that
-the minimization version is \APXh{}, even when the input is made of three clusterings and give a
-combinatorial $\nicefrac{4}{5}$-approximation for the maximization problem. Experimental evaluations are
-conducted by \textcite{Bertolacci07} and \textcite{Filkov08}. The former describe a scalable
-approach that first samples a small portion of the data, runs a (potentially computationally
-expensive) approximation algorithm and finally augment the resulting partition by adding to it the
-unsampled nodes one by one.  Experiments confirm that the running time is greatly improved compared
-with the linear program methods while the resulting objective value is essentially the same. Note,
-however, that LP methods can be applied in practice thanks to some tricks~\autocite{ConsensusLP10}.
-If we have $k$ input clusterings $\cluster^1,\ldots,\cluster^k$ and we parameterized the problem by
-$t$, which is the sum over the input clusterings of the number of pairs of objects that are clustered
-differently by a solution $\cluster^\star$ and $\cluster^i$, then there is a polynomial algorithm
-running in $O(4.24^{\nicefrac{t}{k}}\cdot \nicefrac{t}{k}^3 +
-kn^2)$~\autocite{parameterizedConsensus14}.
-
 \paragraph{Local \pcc{}}
@@
 
-\paragraph{\pcc{} under noise}
+\paragraph{Spectral Clustering}
 
-As we have seen, assuming the Unique Games Conjecture, the \mmc{} problem, and therefore
-the \pcc{} problem on a general graph, cannot be approximated to within a constant factor in the worst
-case.  But maybe we can do better in the average case, which motivates the study of semi-random
-model, where real graphs are seen as being obtained from the controlled perturbation of a perfectly
-clusterable graph. In the simplest case, each edge sign is independently flipped with probability
-$p \in [0, \shalf)$. This situation on complete graphs was considered in \autocite[Section
-6]{Bansal2002}, showing a simple algorithm that with high probability makes
-$\tilde{O}(n^\frac{3}{2})$ mistakes, and in \autocite[Theorem 2.6]{Ben-Dor99}, with an algorithm
-recovering with high probability the planted partition of an unweighted graph in $O(n^2(\log n)^c)$,
-where $c$ depends on the size of the smallest cluster and the noise probability.
+A classic method for clustering graphs is to leverage their spectral properties. Namely, if $A$ is
+the adjacency matrix of $G$ and $D$ its degree diagonal matrix (that is $D_{u,u} = \degr(u)$), the
+\emph{Laplacian} of $G$, defined by $L_G = D - A$, is a symmetric positive semidefinite matrix. As
+such, it has $n$ real non-negative eigenvalues, and its spectrum provides additional information on
+the connectivity of $G$. For instance, $0$ is always the smallest eigenvalue and its multiplicity is
+equal to number of connected components of $G$, while ---if $G$ is connected--- the second
+eigenvalue is the algebraic connectivity of $G$, whose magnitude is an indication of how well
+connected is the graph. This matrix is typically used for clustering by computing its first $k$
+eigenvectors, which embed the $n$ nodes of $G$ in $\Rbb^k$, where there are then clustered with the
+$k$-means algorithm. This can be seen as a relaxation of the discrete \rcut{} objective, which asks
+for the partition $\{C_1, \ldots, C_k\}$ minimizing $\frac{1}{2}\sum_{i=1}^k \frac{\mathrm{cut}(C_i,
+\bar{C_i})}{|C_i|}$, where $\bar{C_i}$ is the complement of $C_i$ in $V$ and $\mathrm{cut}(B, C) =
+\sum_{u\in B, v\in C} w_{uv}$ is the total weight of the edges between $B$ and
+$C$~\autocite{tutoSpectralClustering07}. By considering the symmetric normalized Laplacian
+$L_{sym}  = D^{\nicefrac{1}{2}}LD^{\nicefrac{1}{2}}$, it is also possible to approximate the
+normalized cut objective (\ncut{}), where $|C_i|$ is replaced by $vol(C_i) = \sum_{u\in C_i}
+\degr(u)$. We will now see how these kinds of approaches can be extended to signed graphs, noting
+first that they require to fix the number of clusters beforehand and are looking for clusters
+balanced in size, which makes the problem related but not equivalent to \pcc{}.
 
-\Textcite{Joachims2005} analyze a more refined weighted model, where weights are generated by a
-probability distribution whose mean on true positive edges is larger than $\mu^+>0$ and whose mean
-on true negative edges is smaller than $\mu^-<0$. They give a finite-sample bound on the number of
-node misclustered w.r.t the planted partition as a function on the probability distribution
-parameters. Indeed, as pointed out by \textcite{plantedAilon09}, independent and uniform noise is
-not a good model of real situations, where the input of \pcc{}, \ie{} the similarity between nodes,
-is often the result of a preprocessing, which may present strong correlations. Therefore, instead of
-measuring the quality of a solution against in the input (\ie{} the similarity information), they
-argue it is more sensible to measure it against the (unknown) true optimal clustering that gave rise
-to the input, and show that \ccpivot{} allows that thanks to a new analysis. \Textcite{Mathieu2010}
-also consider an adversarial model, where all edges are flipped with probability $p$ but the
-adversary then decide whether to show us the true sign or the flipped sign. On complete unweighted
-graphs, they find a solution of \mind{} at most $1 + O(n^{-\nicefrac{1}{6}})$ times the optimal
-whenever $p \leq \shalf - n^{-\nicefrac{1}{3}}$ by rounding the usual SDP solution. If in addition
-$p\leq\nicefrac{1}{3}$, there is no adversary and each planted cluster has at least
-$\Theta(\sqrt{n})$ nodes, then the planted partition can be recovered exactly.
-\Textcite{Makarychev2014} study the same adversarial model but on general weighted graphs, giving a PTAS
-for \mind{} when $p\leq \nicefrac{1}{4}$. Under additional assumptions on the density of edges, they
-present another algorithm that finds the ground truth clustering with an arbitrarily small
-classification error.\footnote{They also mention a potentially interesting related
-work~\autocite{unobservedCC14}.} % http://www.jmlr.org/papers/v15/chen14a.html
+The first line of research consider only \mind{}$[2]$. For instance, \textcite{NcutAnd2CC08} show
+that both normalized cut and \mind{}$[2]$ objectives can be written as a SDP (or equivalently as
+eigenvalue problems) and thus combined, the intuition being that we look for \ncut{} solutions whose
+number of disagreements is not too much more than the approximate optimal.
+
+Letting \ncut{}$(C, \bar{C}) = \frac{\mathrm{cut}(C, \bar{C})}{\mathrm{bal}(C)}$ with
+$\mathrm{bal}(C) = 2\frac{vol(C)vol(\bar{C})}{vol(V)}$, \textcite{mOneCC12} define a new objective:
+\begin{equation*}
+  \hat{F}_\gamma(C) = \frac{\mathrm{cut}(C, \bar{C}) + \gamma\left(\hat{M}(C)+\hat{N}(C)\right)}{\mathrm{bal}(C)}
+\end{equation*}
+where $\gamma \in \Rbb{}_+$ is a parameter, while $\hat{M}(C)$ and $\hat{N}(C)$ are respectively the
+number of positive and negative disagreements of the $(C, \bar{C})$ clustering.
+They show how to optimize a tight continuous relaxation of $\hat{F}_\gamma$ as the non negative
+ratio of a difference of convex function and a convex function.
+
+On the other hand, one can also adapt these two cut objectives to directly include negative edges.
+\Textcite{Luca10} define the signed Laplacian as $\bar{L} = \bar{D} - A$, where $\bar{D}$ is the
+signed degree matrix such that $\bar{D}_{uu} = \sum_{v\in \nei(u)} |A_{uv}|$, as well as a signed
+variant of the symmetric normalized Laplacian $\bar{L}_{sym}  = \bar{D}^{\nicefrac{1}{2}} \bar{L}
+\bar{D}^{\nicefrac{1}{2}}$. They show that the signed Laplacian is positive semidefinite, and
+even positive-definite as soon as the graph is unbalanced (that is contains a cycle with an odd
+number of negative edges). From positive and negative cuts defined as $\mathrm{cut}^+(B, C) =
+\sum_{u\in B, v\in C} w^+_{uv}$ and $\mathrm{cut}^-(B, C) = \sum_{u\in B, v\in C} w^-_{uv}$, a
+natural signed cut is $\mathrm{scut}(B, C) = 2\mathrm{cut}^+(B, C) + \mathrm{cut}^-(B, B) +
+\mathrm{cut}^-(C, C)$ which can then be used to defined signed \rcut{} and \ncut{}. Arguing that those
+definitions force negatively linked nodes to be symmetric around the origin, do not take into
+account the balance of negative edges in each cluster and are difficult to extend to more than two
+clusters, \textcite{SignedEmbedding15} instead propose two new normalized cuts:
+\begin{align*}
+  SNScut(C_1, \ldots, C_k) &= \sum_{i=1}^k
+  \frac{\mathrm{cut}^+(C_i, \bar{C_i})-\mathrm{cut}^-(C_i, \bar{C_i})}{vol(C_i)}  \\
+  BNScut(C_1, \ldots, C_k) &= \sum_{i=1}^k
+  \frac{\mathrm{cut}^+(C_i, \bar{C_i})-\mathrm{cut}^-(C_i, \bar{C_i})+vol^-(C_i)}{vol(C_i)}
+\end{align*}
+
+Noting that if $x_i\in\Rbb^n$ is the vector indicator of cluster $C_i$ (that is the \uth{} entry of
+$x_i$ is $1$ is $u$ belongs to $C_i$ and $0$ otherwise), $x_i^T\bar{L}x_i = 2\mathrm{cut}^-(C_i,
+C_i) + \mathrm{cut}^-(C_i, \bar{C}_i) + \mathrm{cut}^+(C_i, \bar{C}_i)$,
+\textcite{mSemanticWordCC17} introduce the following cut objective:
+\begin{equation*}
+  sNcut(C_1, \ldots, C_k) = \sum_{i=1}^k
+  \frac{2\mathrm{cut}^-(C_i, C_i) + \mathrm{cut}^-(C_i, \bar{C}_i) + \mathrm{cut}^+(C_i, \bar{C}_i)}{vol(C_i)}
+\end{equation*}
+Additional cut formulations for $k$-clusters are also presented in~\autocite{moreSignedCut12}.
+
+Finally, \textcite{mGeometricMean16} show that the Laplacians defined so far can be seen as
+arithmetic means of the Laplacian $L^+$ of the positive subgraph $G^+=(V, E^+)$ and the signless
+Laplacian $Q^-$ of the negative subgraph $G^-=(V, E^-)$, where $Q^- = D^- + A^-$. They suggest
+instead to use a geometric mean defined for two positive matrices $A$ and $B$ as $A\#B =
+A^{\shalf{}} \left(A^{-\shalf{}} BA^{-\shalf{}} \right)^{\shalf{}} A^{\shalf{}}$. This suggestion is
+based on the fact that if $u$ is a common eigenvector of both $A$ and $B$ with eigenvalue $\lambda$
+and $\mu$ respectively, then $u$ is an eigenvector of $A+B$ with eigenvalue $\lambda+\mu$ and an
+eigenvector of $A\#B$ with eigenvalue $\sqrt{\lambda\mu}$. Therefore, the $k$ smallest eigenvalues
+of the geometric mean Laplacian will be influenced by both smallest eigenvalues of $L^+$
+(corresponding to assortative clusters in $G^+$) and of $Q^-$ (corresponding to disassortative
+clusters in $G^-$), while this is note the case for the arithmetic mean of Laplacians.
+
+\paragraph{Community detection}
+
+The clustering problem is often named community detection in the context of social networks, and
+several methods developed by practitioners have been extended to signed graphs. While they do not
+necessarily considered the \pcc{} objective, and especially not its optimum, we still give a brief
+overview of them, as they tend to have been more tested experimentally. For instance, to find the
+cluster of node $u$, \textcite{Yang2007} use a random walk approach on the positive subgraph $G^+$
+to compute the probability of each node to reach $u$ in $T$ steps, sort the nodes accordingly and
+then find a threshold  based on the number of disagreements.
+The one node moves local heuristic that we described in the Physics-inspired paragraph
+\vpageref{par:cc_physics} can also be formulated as genetic algorithms that simultaneously try to
+minimize the number of disagreements and maximize a signed variant of the
+modularity~\autocites{Li2013}{Amelio2013}. \Textcite{Anchuri2012} also consider these two objectives
+by seeing them as eigenvalue problems and devise a iterative splitting procedure.
+The overlapping community detection variant is considered by \textcite{Chen14}, who used a signed
+probabilistic mixture model. Namely, an edge selects a pair of cluster $r,s$ with probability
+$\omega_{rs}$ (where $r=s$ if the edge is positive and $r\neq s$ otherwise) and chooses its endpoints
+$u$ and $v$ with probability $\theta_{ru}$ and $\theta_{sv}$. $\theta_{ru}$ is therefore the soft
+membership of node $u$ to cluster $r$, and those parameters are estimated using the
+expectation-maximization algorithm. The same model is extended to directed graphs by
+\textcite{Jiang2015}, who strangely enough names it stochastic blockmodel, although the focus is
+still on edge and not nodes.
+In a similar spirit to \maxa{}$[k]$, \textcite{SignedGang} focus on finding $k$ subgraphs dense in
+positive edges and densely connected by negative edges to each others. They dub such subgraph
+\emph{Oppositive Cohesive Groups}, or more vividly \emph{Gangs in War}, and after formulating the
+problem as a constrained quadratic optimization, they propose a faster iterative local search
+heuristic.
+% relaxed structural balance \autocite{Doreian2009} but used the same approach as \textcite{Early96}
 
@@
 
-\paragraph{\textsc{Cluster editing}}
-\label{par:cc_editing}
+\paragraph{\pcc{} under noise}
 
-% \url{http://www.sciencedirect.com/science?_ob=ArticleListURL&_method=list&_ArticleListID=-1231039636&_sort=r&_st=4&md5=018b782f17918959c0cc389c33ba2cd7&searchtype=a}
+As we have seen, assuming the Unique Games Conjecture, the \mmc{} problem, and therefore
+the \pcc{} problem on a general graph, cannot be approximated to within a constant factor in the worst
+case.  But maybe we can do better in the average case, which motivates the study of semi-random
+model, where real graphs are seen as being obtained from the controlled perturbation of a perfectly
+clusterable graph. In the simplest case, each edge sign is independently flipped with probability
+$p \in [0, \shalf)$. This situation on complete graphs was considered in \autocite[Section
+6]{Bansal2002}, showing a simple algorithm that with high probability makes
+$\tilde{O}(n^\frac{3}{2})$ mistakes, and in \autocite[Theorem 2.6]{Ben-Dor99}, with an algorithm
+recovering with high probability the planted partition of an unweighted graph in $O(n^2(\log n)^c)$,
+where $c$ depends on the size of the smallest cluster and the noise probability.
 
-In the cluster editing problem, one is given a general unsigned input graph $H=(V,E)$ and wants to
-find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
-union of cliques. This is equivalent to \pcc{} on complete graphs. To see why, let $G=(V,(E^+,E^-))$
-be a complete signed graph with nodeset $V$, $E^+$ be the set $E$ of edges of $H$ and $E^-$ be all
-the edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
-disagreements within clusters are edges added to $H$ and the positive disagreements between clusters
-are deleted from $H$.  To the best of our knowledge, the problem was first introduced under this
-name by \textcite{Shamir02}.\marginpars{However, the same problem was studied before and we refer the
-reader to the comprehensive survey of \textcite{clusterEditSurvey13} for additional details,
-whereas we shall only give important and more recent pointers here.} They show the problem is
-\NPc{}, even if the number of clusters $k\geq2$ is set beforehand\footnote{The reduction is from
-3-exact 3 cover, see \autocite[Theorems 1, 2 and Corollary 1]{Shamir02}.} and provide a
-$0.878$-approximation in the weighted $k=2$ case using the standard Goemans-Williamson SDP
-relaxation.
-
-If we parameterize the problem by the number $d$ of edges that need to be added or deleted (that is,
-the number of disagreements), then it can be solved in polynomial time in the size of the input
-graph (but not in $d$). The best known approximation so far is
-$O(1.62^d+m+n)$~\autocite{GoldenCE12}, which search for conflict triples (\ie unbalanced triangles
-in the sign language) and branch by either deleting or merging one positive edge. If we additionally
-look for exactly $k$ clusters (\ie \mind$[k]$), there is a fixed parameters algorithm running in
-$O(2^{O(\sqrt{kd})}+n+m)$~\autocite{Fomin2014}. If at most $a$ edges can be added and at most $b$
-edges can be deleted at each node, and if the minimum size of a cluster is at least $2(a+b)$, then
-the problem can be solved in polynomial time~\autocite{Abu-Khzam2015}. Finally, for planar graph,
-there is a PTAS running in $O(n2^{\epsilon^{-1}\log(\epsilon^{-1})})$ obtained by dividing the graph
-into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}.
-
-\paragraph{Community detection}
-
-The clustering problem is often named community detection in the context of social networks, and
-several methods developed by practitioners have been extended to signed graphs. While they do not
-necessarily considered the \pcc{} objective, and especially not its optimum, we still give a brief
-overview of them, as they tend to have been more tested experimentally. For instance, to find the
-cluster of node $u$, \textcite{Yang2007} use a random walk approach on the positive subgraph $G^+$
-to compute the probability of each node to reach $u$ in $T$ steps, sort the nodes accordingly and
-then find a threshold  based on the number of disagreements.
-The one node moves local heuristic that we described in the Physics-inspired paragraph
-\vpageref{par:cc_physics} can also be formulated as genetic algorithms that simultaneously try to
-minimize the number of disagreements and maximize a signed variant of the
-modularity~\autocites{Li2013}{Amelio2013}. \Textcite{Anchuri2012} also consider these two objectives
-by seeing them as eigenvalue problems and devise a iterative splitting procedure.
-The overlapping community detection variant is considered by \textcite{Chen14}, who used a signed
-probabilistic mixture model. Namely, an edge selects a pair of cluster $r,s$ with probability
-$\omega_{rs}$ (where $r=s$ if the edge is positive and $r\neq s$ otherwise) and chooses its endpoints
-$u$ and $v$ with probability $\theta_{ru}$ and $\theta_{sv}$. $\theta_{ru}$ is therefore the soft
-membership of node $u$ to cluster $r$, and those parameters are estimated using the
-expectation-maximization algorithm. The same model is extended to directed graphs by
-\textcite{Jiang2015}, who strangely enough names it stochastic blockmodel, although the focus is
-still on edge and not nodes.
-In a similar spirit to \maxa{}$[k]$, \textcite{SignedGang} focus on finding $k$ subgraphs dense in
-positive edges and densely connected by negative edges to each others. They dub such subgraph
-\emph{Oppositive Cohesive Groups}, or more vividly \emph{Gangs in War}, and after formulating the
-problem as a constrained quadratic optimization, they propose a faster iterative local search
-heuristic.
-% relaxed structural balance \autocite{Doreian2009} but used the same approach as \textcite{Early96}
+\Textcite{Joachims2005} analyze a more refined weighted model, where weights are generated by a
+probability distribution whose mean on true positive edges is larger than $\mu^+>0$ and whose mean
+on true negative edges is smaller than $\mu^-<0$. They give a finite-sample bound on the number of
+node misclustered w.r.t the planted partition as a function on the probability distribution
+parameters. Indeed, as pointed out by \textcite{plantedAilon09}, independent and uniform noise is
+not a good model of real situations, where the input of \pcc{}, \ie{} the similarity between nodes,
+is often the result of a preprocessing, which may present strong correlations. Therefore, instead of
+measuring the quality of a solution against in the input (\ie{} the similarity information), they
+argue it is more sensible to measure it against the (unknown) true optimal clustering that gave rise
+to the input, and show that \ccpivot{} allows that thanks to a new analysis. \Textcite{Mathieu2010}
+also consider an adversarial model, where all edges are flipped with probability $p$ but the
+adversary then decide whether to show us the true sign or the flipped sign. On complete unweighted
+graphs, they find a solution of \mind{} at most $1 + O(n^{-\nicefrac{1}{6}})$ times the optimal
+whenever $p \leq \shalf - n^{-\nicefrac{1}{3}}$ by rounding the usual SDP solution. If in addition
+$p\leq\nicefrac{1}{3}$, there is no adversary and each planted cluster has at least
+$\Theta(\sqrt{n})$ nodes, then the planted partition can be recovered exactly.
+\Textcite{Makarychev2014} study the same adversarial model but on general weighted graphs, giving a PTAS
+for \mind{} when $p\leq \nicefrac{1}{4}$. Under additional assumptions on the density of edges, they
+present another algorithm that finds the ground truth clustering with an arbitrarily small
+classification error.\footnote{They also mention a potentially interesting related
+work~\autocite{unobservedCC14}.} % http://www.jmlr.org/papers/v15/chen14a.html

### 2017-08-17 cc/approx.tex
add few more references

@@
 give a $17429$-approximation of \mind{} and a PTAS%
-\footnote{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
+\bottomfootnote{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
 minimization (respectively maximization) problem $\mathcal{P}$  in NP if given any $\epsilon>0$ and

### 2017-08-17 cc/proposed.tex
add few more references

@@
 cluster})$-approximation of \mind{}, and empirically results in less disagreements than \ccpivot{}.
-Building upon their previous GRASP work~\autocite{GRASP13}, an iterated local search heuristic is
+Building upon their previous GRASP work~\autocite{GRASP13}, an iterated local search (ILS) heuristic is
 presented in~\autocites{Levorato2015}{Levorato2017}. Each iteration of this algorithm starts by
@@
 be run in parallel, which allow them to process a \np{10000} nodes graph on 10 cores in around 700
-seconds.
+seconds. \Textcite{heuristicCE16} show how to use ILS to generate initial solutions of the Cluster
+Editing problem that are then fed to an integer program.
 This local search followed by random perturbation is reminiscent of genetic algorithms, and such an

### 2017-08-17 cc/sign_relation.tex
add few more references

@@
 \eqref{eq:mind} is $0$) and is unweighted, as it highlights more clearly the connection with the
-balance theory.
+balance theory. We focus here on balance theory, for it is both compelling and well established.
+Indeed, as illustrated by the account of \textcite{CSBalanceSurvey15}, it has been used extensively
+in the last half century to study signed networks\marginpars{It was even used to study fairy
+tales~\autocite{fairyTales80}.} and their dynamics.
+However, it has shortcomings, the main one being its inability to deal with directed graphs.
+Therefore alternative theories governing signs formation have been proposed, as surveyed
+in~\autocite{Yap2015}.
 
@@
 repeatedly shown to be unbalanced, although the extent of this unbalance depends of the importance
-given to longer cycles compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}.%
-{measureUnbalance17} For the same reason, in virtually any real instances of the \pcc{} problem, the%
+given to longer cycles compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
+{measureUnbalance17}. For the same reason, in virtually any real instances of the \pcc{} problem, the%
 % there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration

### 2017-08-17 cc/variants.tex
add few more references

@@
 \end{equation*}
-Additional cut formulations for $k$-clusters are also presented in~\autocite{moreSignedCut12}.
+Additional cut formulations for $k$-clusters are also presented in~\autocite{moreSignedCut12},
+although \textcite{Knyazev2017} argue that using the non signed Laplacian and considering negative
+eigenvalues might be just as effective, citing for instance numerical instability of signed
+Laplacian.
 

### 2017-08-17 cc/variants.tex
another paper

--- 2017-08-16 cc/variants.tex
@@
 \end{equation*}
-Additional cut formulations for $k$-clusters are also presented in~\autocite{moreSignedCut12}.
+Additional cut formulations for $k$-clusters are also presented in~\autocite{moreSignedCut12},
+although \textcite{Knyazev2017l} argue that using the non signed Laplacian and considering negative
+eigenvalues might be just as effective, citing for instance numerical instability of signed
+Laplacian.
 
@@
 centroid, and query the edges connecting the centroid as ordered by an entropy-based criterion.
+In the case the oracle answering the queries is consistent, then they exists an information
+theoretic bound on the number of queries needed to recover the clusters, and an algorithm matching
+that bound up to a $O(\log n)$ factor~\autocite{perfectActiveLong17}.
 \textcite[Section 5]{Ailon2014} present another active algorithm that solves \mind$[k]$ with a
@@
 work~\autocite{unobservedCC14}.} % http://www.jmlr.org/papers/v15/chen14a.html
+% \textcite{Noisy2CCLocality16} study the recovery of an arbitrary planted 2-partition under noisy
+% measurement in special kind of graphs where each node is connected with a local neighborhood and
+% show computationally efficient algorithms that only query an information theoretically optimal
+% number of pair. I think it was extended in \url{https://doi.org/10.1109/TIT.2016.2600566}

### 2017-08-21 cc/approx.tex
typos found by LanguageTool

@@
 $O(n^2e^{O(\frac{1}{\epsilon^{10}}\log\frac{1}{\epsilon})})$ and returns with probability
-$1-\frac{\epsilon}{3}$ a solution with at most $\epsilon n^2$ fewer agreements than the optimal of
-\maxa{}.
+$1-\frac{\epsilon}{3}$ a solution with at most $\epsilon n^2$ fewer agreements than the optimal
+solution of \maxa{}.
 

### 2017-08-21 cc/multimincut_stable.tex
typos found by LanguageTool

@@
 whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
-polynomial reduction from \mmc{} to \pcc{}. Namely given an instance $(H, S)$ of \mmc{}, we let
+polynomial reduction from \mmc{} to \pcc{}. Namely, given an instance $(H, S)$ of \mmc{}, we let
 $W_H=\sum_{e\in E_H} w_H(e)$ be the total weight of $H$ and $G_H$ be the same graph as $H$ with all
@@
 $W$ induces a clustering in $G_H$ of weighted disagreement $W$ by the connected components of
-$G^+\setminus T$. Likewise a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
+$G^+\setminus T$. Likewise, a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
 weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges

### 2017-08-21 cc/problem_applications.tex
add few more references
typos found by LanguageTool

@@
 $w_{u,v}^-$ denotes the strength of the dissimilarity between $u$ and $v$.
-Note, however, that in many applications, only one of these two numbers is non zero, in which case
+Note, however, that in many applications, only one of these two numbers is nonzero, in which case
 we more conveniently set $w_{u,v} = \begin{cases}
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{assets/tikz/cc_objectives_tikz.pdf}
+	\includegraphics[width=0.63\linewidth]{assets/tikz/cc_objectives_tikz.pdf}
 	\caption[Small example of \pcc{}]{A small graph with eight nodes and ten edges. Solid edges
@@
 	by two cycles with one negative edge and thus cannot be avoided, meaning that \cluster{} is
-	optimal. However it is not the unique solution: for instance, merging the orange and green
-	clusters would also yields two disagreements.}
+	optimal. However, it is not the unique solution: for instance, merging the orange and green
+	clusters would also yield two disagreements.}
 	\label{fig:cc_objectives}
@@
       additional tuning allows to segment a volume image of a mouse cortex with up to billions
-      voxels. \Textcite{Beier2015} segment 2D and 3D images with a energy based formulation of
+      voxels. \Textcite{Beier2015} segment 2D and 3D images with an energy based formulation of
       \pcc{} and iteratively improve their solution by merging it with another clustering given by a
       proposal generator. By developing another scalable energy based optimization procedure, and
-      with the help of few user provided cues, \textcite{Bagon2011} are able to apply \pcc{}
+      with the help of few user-provided cues, \textcite{Bagon2011} are able to apply \pcc{}
       directly at the pixel level.
@@
       summarization, question answering, and information extraction. \Textcite[Section
-      2.3]{graphicalCoreference04} tackle coreference resolution by using a undirected graphical
+      2.3]{graphicalCoreference04} tackle coreference resolution by using an undirected graphical
       model on which performing inference is equivalent to \pcc{}. On small scale instances,
@@
 	    constraints that admissible clusterings have to respect, the authors have to modify in
-	    non trivial ways an existing \pcc{} algorithm. This was extended to weighted and partial
+	    nontrivial ways an existing \pcc{} algorithm. This was extended to weighted and partial
 	    constraints  by \textcite{WeightedER14}. Another example is given by

### 2017-08-21 cc/proposed.tex
typos found by LanguageTool

@@
 binary. \Textcite{ExactMIP13} solve this binary integer program on random instances and show that
-depending of the negative edge density, the FICO Xpress solver starts to be unable to finish within
-a one hour time limit when $n \geq 40$.
+depending on the negative edge density, the FICO Xpress solver starts to be unable to finish within
+a one-hour time limit when $n \geq 40$.
 \Textcite{Aref2016} describe four linear and quadratic  binary integer problems
@@
 are either exact or come with some approximation guarantees, practitioners have also develop
-approaches that are designed to efficiently reach a solution that is satisfying enough for the
+approaches that are designed to efficiently reach a solution that is satisfactory enough for the
 application at hand. For instance, while studying small scale signed social networks,
@@
 considered node $v$ with $w^\pm_{uv} > 0$. The \textsc{Vote} algorithm adds each node to the cluster
-that minimizes the \pcc{} clustering objective, \ie{} to the cluster maximizing the total net weight
+that minimizes the \pcc{} objective, \ie{} to the cluster maximizing the total net weight
 or to a singleton if no total is positive.} Empirically, \textsc{Vote} turns out to be the best.
@@
 $O(n^2\log n)$, provides a $2$-approximation of \maxa{}, a $O(n\cdot \text{size of the largest
-cluster})$-approximation of \mind{}, and empirically results in less disagreements than \ccpivot{}.
+cluster})$-approximation of \mind{}, and empirically results in fewer disagreements than \ccpivot{}.
 Building upon their previous GRASP work~\autocite{GRASP13}, an iterated local search (ILS) heuristic is
@@
 and repeatedly samples an edge $(u,v)$ from $E_s^+$ before trying to make the neighborhoods of $u$
-and $v$ coincide by adding or removing edges in $E_s^+$.  Yet another idea is modify the LP
+and $v$ coincide by adding or removing edges in $E_s^+$.  Yet another idea is to modify the LP
 \eqref{eq:mindLP} to replace the binary variable $x_{uv}$ by a $k$ clusters indicator matrix
 $L\in\Rbb^{k\times n}$ where $L_{iu}$ is $1$ is $u\in C_i$ and $-1$
-otherwise~\autocite{AltOptimLP13}. Indeed $x_{u\cdot} \equiv L_{\cluster(u)\cdot}$, where $\cluster$
+otherwise~\autocite{AltOptimLP13}. Indeed, $x_{u\cdot} \equiv L_{\cluster(u)\cdot}$, where $\cluster$
 is the cluster assignment. By relaxing the integer constraints on $L$, it is possible to do
@@
 outcome of a local search algorithm such that $A_\sigma = T_\sigma A T_\sigma$ has the same number
-of disagreements as $A$ but the least number of negative edges. This is called a gauge
+of disagreements as $A$ but the smallest number of negative edges. This is called a gauge
 transformation in the spin glass literature and the benefit of that heuristic is that it scales
@@
 \pcc{} energy: \enquote{(i) the energy is non sub-modular, (ii) the number of clusters is not known
-in advance, and (iii) there is no unary term}. Doing so, the are able to handle large problems with
+in advance, and (iii) there is no unary term}. Doing so, they are able to handle large problems with
 more than 100K nodes.  Also adopting an energy minimization approach, \textcite{Kappes2016} assign a

### 2017-08-21 cc/sign_relation.tex
another paper
typos found by LanguageTool

@@
 \eqref{eq:mind} is $0$) and is unweighted, as it highlights more clearly the connection with the
-balance theory. We focus here on balance theory, for it is both compelling and well established.
+balance theory. We focus here on balance theory, for it is both compelling and well-established.
 Indeed, as illustrated by the account of \textcite{CSBalanceSurvey15}, it has been used extensively
@@
 However, it has shortcomings, the main one being its inability to deal with directed graphs.
-Therefore alternative theories governing signs formation have been proposed, as surveyed
+Therefore, alternative theories governing signs formation have been proposed, as surveyed
 in~\autocite{Yap2015}.
@@
 It is interesting to note that this local property of triangles gives raise to a global structure.
-Namely a strongly balanced complete graph can be split in two (possibly empty) clusters $L$ and $R$
+Namely, a strongly balanced complete graph can be split in two (possibly empty) clusters $L$ and $R$
 such that all edges within $L$ and $R$ are positive, while all edges between $L$ and $R$ are
-negative. Indeed let us pick an arbitrary node $u$ and let $L=\neip(u)\setminus \{u\}$ and $R=V\setminus L$. Now
+negative. Indeed, let us pick an arbitrary node $u$ and let $L=\neip(u)\setminus \{u\}$ and $R=V\setminus L$. Now
 let $v,w\in L\setminus \{u\}$ and $x,y\in R$. From \autoref{fig:cc_twostrongcc} and the fact that
@@
 $(x,y)$ to be either positive or negative. We can split $R$ further into the positive neighborhood of
-$x$ and the rest, and keep doing so until we form the $k$ clusters. We will make this argument more
-formal in the following.
+$x$ and the rest, and keep doing so until we form the $k$ clusters. In the following we will make
+this argument more formal.
 
@@
 
-The same results also hold for weak balance.
+Similar results also hold for weak balance.
 \begin{theorem}
@@
 
-In real data though, we do not expect neither strong nor weak balance to hold, for they are fairly
+In real data though, we do not expect either strong or weak balance to hold, for they are fairly
 demanding model. Indeed, three of the real networks considered in \autoref{chap:troll} have been
-repeatedly shown to be unbalanced, although the extent of this unbalance depends of the importance
+repeatedly shown to be unbalanced, although the extent of this unbalance depends on the importance
 given to longer cycles compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
@@
 graph, the two clusters structure is still present, although it does not cover the whole graph
-anymore.
+any more.
 
@@
 supervised classification problem. Yet, getting an optimal solution of \pcc{} on the training set
-would provide a principled heuristics for \esp{}. Unfortunately, as we shall see in
+would provide a principled heuristic for \esp{}. Unfortunately, as we shall see in
 \autoref{sub:state_of_the_art}, exact or fixed parameter solutions are not scalable and one has to

### 2017-08-21 cc/variants.tex
typos found by LanguageTool

@@
 edges are labeled categorically instead of binary, when nodes can belong simultaneously to several
-cluster or when we optimize local objectives per nodes instead of global ones. We will also look at
+clusters or when we optimize local objectives per nodes instead of global ones. We will also look at
 clustering in signed graphs in general, using spectral methods or heuristics from the community
-detection literature. Finally we will shift perspective and review methods beyond the batch case and
+detection literature. Finally, we will shift perspective and review methods beyond the batch case and
 the worst-case analysis.
@@
 we cannot expect to have complete bipartite graphs in practice. The first results was given by
-\textcite{Amit04}, who obtains a $11$-approximation for \mind{} by adapting the \regionGrow{} method.
+\textcite{Amit04}, who obtains an $11$-approximation for \mind{} by adapting the \regionGrow{} method.
 The \ccpivot{} is adapted to the bipartite case by
@@
 differently for positive, negative and in that case same-side absent edges, \textcite{Chawla2014}
-bring down this approximation factor to $3$, even for $K\geq 2$-partite graphs. By formulating the
+bring down this approximation factor to $3$, even for $K\geq 2$-partite graphs. Through formulating the
 \maxa{}$[k]$ problem as a bilinear maximization problem and computing a low-rank approximation of
@@
 as large as $n$. However, \textcite{Anava2015} present constant factor approximations. Namely, they
-show the problem can be reduce to classical \pcc{} by setting all edges incident to a node $u$ to
+show the problem can be reduced to classical \pcc{} by setting all edges incident to a node $u$ to
 negative if they are not of the majority color of $u$. They then apply the regular \ccpivot{} and
-show this gives a $11$-approximation to the original problem. Furthermore, they also write a linear
+show this gives an $11$-approximation to the original problem. Furthermore, they also write a linear
 program and round it using the \regionGrow{} method to obtain an
@@
 \cluster(v)) - w_{uv}\right|\] where $H$ is a similarity function between two sets of labels, chosen
-here to be the Jaccard similarity or a $0/1$ indicator of non empty intersection. These problems are
+here to be the Jaccard similarity or a $0/1$ indicator of nonempty intersection. These problems are
 showed \NPc{} and approximated by a local search algorithm, iteratively optimizing the assignment of
@@
 stochastic setting and using \enquote{the Baum-Eagon inequality, which provides an effective
-iterative means for maximizing polynomial functions in probability domains}.
+iterative scheme for maximizing polynomial functions in probability domains}.
 
@@
 disagreements incurred by a single node is motivated by the example of recommendation systems: if
-errors correspond to unsatisfying recommendations, we do not want a single user to suffer a large
-number of them. Minimax \pcc{} is \NPc{} on both complete graphs and complete bipartite graphs but
+errors correspond to unsatisfying recommendations, we do not want a single user to suffer many
+of them. Minimax \pcc{} is \NPc{} on both complete graphs and complete bipartite graphs but
 by modifying the \regionGrow{} method, the authors respectively a $48$ and
@@
 in one of the two clusters).
-The idea is to chose pivot not randomly but by maximizing a given criteria and
+The idea is to chose pivots not randomly but by maximizing a given criteria and
 to grow balls with a radius $\alpha$ computed numerically to optimize the approximation factor.
@@
 minimax \mind{} and seems not to have a constant factor approximation, even on complete graphs.
-Furthermore, these algorithms are deterministic, as opposed to many \pcc{} approximation, since
+Furthermore, these algorithms are deterministic, as opposed to many \pcc{} approximations, since
 bounds on expected disagreements of an edge does not translate easily on their maximum.
@@
 that on general weighted graphs, the LP has a large integrality gap of $\nicefrac{n}{2}$ yet they
-combine it with a combinatorial approach to reach a $O(\sqrt{n})$ approximation. Finally they
+combine it with a combinatorial approach to reach a $O(\sqrt{n})$ approximation. Finally, they
 consider the complementary problem of maximizing the minimum number of agreements reach at a single
@@
 number of positive and negative disagreements of the $(C, \bar{C})$ clustering.
-They show how to optimize a tight continuous relaxation of $\hat{F}_\gamma$ as the non negative
+They show how to optimize a tight continuous relaxation of $\hat{F}_\gamma$ as the non-negative
 ratio of a difference of convex function and a convex function.
@@
 In a similar spirit to \maxa{}$[k]$, \textcite{SignedGang} focus on finding $k$ subgraphs dense in
-positive edges and densely connected by negative edges to each others. They dub such subgraph
+positive edges and densely connected by negative edges to each other. Thee dub such subgraph
 \emph{Oppositive Cohesive Groups}, or more vividly \emph{Gangs in War}, and after formulating the
@@
 Case in point, in the context of entity resolution, \textcite{activeCoref07} consider the \pcc{}
-problem where by querying an oracle, one can either reduce the uncertainty about one edge weight or
+problem where, by querying an oracle, one can either reduce the uncertainty about one edge weight or
 add an extra node with edges connecting it to existing nodes. However, this requires web queries
@@
 centroid, and query the edges connecting the centroid as ordered by an entropy-based criterion.
-In the case the oracle answering the queries is consistent, then they exists an information
+In the case the oracle answering the queries is consistent, then there exists an information
 theoretic bound on the number of queries needed to recover the clusters, and an algorithm matching
@@
 agreements. For \mind{}, this is $O(n)$-competitive algorithm and they show such ratio is optimal by
-exhibiting an instance\footnote{Namely two positive cliques $A$ and $B$ joined by positive edges except
+exhibiting an instance\footnote{Namely, two positive cliques $A$ and $B$ joined by positive edges except
 between $a\in A$ and $\{b_1,\ldots, b_k\}\in B$. Those nodes are given first and thus form a cluster,

### 2017-08-21 cc/problem_applications.tex
add mention of CC as subspace clustering

@@
 \end{equation}
-or minimizing disagreements (\mind{}):
+or minimizing disagreements (\mind{})\footnote{Note that in the data mining literature, \pcc{} may
+refer to another problem. Namely, it is a special case of high-dimensional data clustering, where
+features are locally correlated in various ways across different clusters that reside in arbitrarily
+oriented subspaces~\autocite{otherCCproblem09}.}:
 \begin{equation}

### 2017-08-22 cc/variants.tex
a few things here and there

@@
 objects in different clusters, thus representing a kind of distance in the space of clusterings.  As
-first show by \textcite{Gionis2007}, finding the optimal clustering is therefore an instance of
-\pcc{} where the weights obey the triangular inequality. They give a deterministic $3$-approximation
-using the \regionGrow{} method. Later \textcite{Bonizzoni2008} show that
+first show by \textcite{EarlyConsensusClustering03}} , finding the optimal clustering is therefore
+an instance of \pcc{} where the weights obey the triangular inequality. \textcite{Gionis2007} give a
+deterministic $3$-approximation using the \regionGrow{} method. Later \textcite{Bonizzoni2008} show that
 the minimization version is \APXh{}, even when the input is made of three clusterings and give a

### 2017-08-23 cc/bias_directed.tex
reorganize files and write some intro

--- 2017-07-21 lowstretch_and_cc.tex
@@
-\documentclass[nobib, a4paper, 10pt, oneside, onecolumn, openany, notitlepage, final,
-svgnames, marginals=raggedouter, english,
-%symmetric, justified,
-]{article}
-\usepackage{charter}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\input{packages}
-\input{abbreviations}
-
-\title{Edge sign prediction in general graphs and connection with Correlation Clustering}
-% \usetikzlibrary{external}
-% \tikzexternalize[mode=list and make, failed ref warnings for={\ref,\cite,\pageref},]
-\begin{document}
-\newgeometry{vmargin=1.3cm,left=1.5cm,textwidth=130mm,marginparsep=5mm,marginparwidth=52mm}
-\maketitle
-% \tableofcontents
-
-% \input{notations}
-
-\section{Limitations of the troll method}
-\label{sec:limitations_of_the_troll_method}
-
-\subsection{Reliance on the sign generative model}
-\label{sub:reliance_on_the_sign_generative_model}
-
-The method presented in the previous chapter hinges crucially upon our sign generative model. Yet
-one can imagine contexts where this model is not applicable, especially when nodes do not represent
-human beings. One way to alleviate this issue was introduced with our online algorithm, where this
-time, signs are generated by an arbitrary adversary. However, in that case, we are still facing two
-limitations\marginpars{limitations is not the right word, maybe \emph{strong constraints}…}
-\begin{enumerate}
-	\item Our bias remains that the labeling is regular (recall this means informally that the all
-		the outgoing signs from a given node tend to be the same, and likewise for the incoming signs),
-		since irregularities are the cost payed by the adversary to make our algorithm mispredict. While
-		this bias is well suited to social networks, other applications may require other
-		bias\marginpars{namely the \pcc{} bias we'll introduce next}.
-	\item In the online setting, we evaluate our performance by the regret\marginpars{ref to regret equation},
-		whereas in general we are interested in more classical measures of predictive accuracy.
-\end{enumerate}
-
-\subsubsection{Experience on directed biological network}
-\label{ssub:experience_on_directed_biological_network}
-
-Take for instance a biological network.
-
-According to \href{https://web.stanford.edu/class/cs224w/slides/handout-bionets.pdf}%
-{this description of various biological networks}, \emph{gene regulatory network} are directed and
-contains activation and inhibition links, as \href{https://en.wikipedia.org/wiki/Gene_regulatory_network#Overview}%
-{showed on Wikipedia}. There are some online databases such as
-\href{http://regulondb.ccg.unam.mx/menu/download/datasets/index.jsp}%
-{RegulonDB} or \href{http://www.pathwaycommons.org/pcviz/}{Pathway Commons} (this last one provide
-network in the BioPAX format, which can be visualized by
-\href{http://www.cytoscape.org/}{CytoScape}).
-Another source is the \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2708159/table/T1}{Table 1}
-of \cite{BioSigned09}.
-%DasGupta also describe (in
-%http://www.sciencedirect.com/science/article/pii/S0303264706001419#sec14) how
-%they built a directed signed network from this SBML file
-%https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1681468/bin/msb4100014-sd1.xml
-
-other domain? coref, images (is it directed?), entity resolution…\marginpars{Actually it's difficult
-to find other kind of graph because all constructed ones (like image, coreference,
-deduplication) are inherently symmetric}
-\fi
-
-\subsection{Need for a directed graph}
-\label{sub:need_for_a_directed_graph}
-
-\iffalse
 A more serious limitation is that our sign model only applies to directed graphs. Consider
@@
 otherwise.  This corresponds naturally to the situation modeled by the \pcc{} problem.
-\fi
-
-\section{\pcc{}}
-\label{sec:correlation_clustering}
-
-\input{cc_part}
-
-\newpage
-
-\section{Low stretch trees and spanners}
-\label{sec:low_stretch_trees_and_spanners}
-
-\input{notations}
-
-Given a graph $G$, and following the social balance theory roughly summarized as \enquote{the enemy
-of my enemy is my friend}, assume that the labelling of $E$ is consistent with a two-clustering of
-$V$. Namely, assume that $V$ can be partitioned in two clusters such that edges within each cluster
-are positive and edges across clusters are negative. In that case, the following
-\emph{multiplicative rule} holds: for any nodes $u$, $v$ in $V$, and any path $p$ between $u$ and
-$v$ in $G$, the sign $\yuv$ is equal to the product of the signs along $p$ (we call this product the
-parity of $p$ and denote it $\pi(p)$). While it is a simple and convenient model, this is too strong
-of a requirement to be satisfied in practice. Therefore, we relax it by assuming that we start with
-a consistent labeling $Y$ but can only observe a randomly perturbed version $Y'$ of $Y$.
-Specifically, given a constant $q\in [0, \nicefrac{1}{2})$, each sign of $Y$ is flipped with a
-probability smaller than $q$.
-
-In this section, we are interested in active learning algorithms, that first query a subset
-$\etrain$ of the edges, observe the signs in $\etrain$ and use them to predict the remaining signs.
-More precisely, we focus on an algorithm that queries spanning tree $T$ of $G$ and predicts the sign
-of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. While this
-can\Todo{state it formally then ;)} be stated more formally~\autocite[Section
-4.1]{Cesa-Bianchi2012b}, it makes sense intuitively that since each sign has been potentially
-flipped, the longer the path in $T$, the more likely its parity will be not be equal to the true
-sign $\yuv{}$. Therefore we would like each such path to be as short as possible. In the following,
-we describe a way to build spanning trees tailored for this situation.
-
-\subsection{\gtx{}: a spanning tree designed for sign prediction}
-\label{sub:gtx_a_simple_low_stretch_tree_construction}
-
-\input{gtx_algorithm}
-
-\subsection{Related works}
-\label{sub:gtx_related_works}
-
-\input{gtx_related}
-
-\iffalse
-\url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
-for instance the Elkin paper \enquote{Our centralized randomized algorithm computes (with
-probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in
-$O(|E|)$ time, whenever $k = \Omega(\log n)$. Note that when $k = \omega(\log n)$, the number of
-edges is $n(1+o(1))$, i.e., in this range the algorithm computes an ultra-sparse spanner in $O(|E|)$
-time.} For instance, if $k=5\log n$, we get a $10\log n$-spanner with $n\left(1+O\left(\frac{\log\log
-n}{n}\right)\right)$ edges in $O(|E|)$ time.
-
-While they have many applications [see first paragraph of \url{https://arxiv.org/pdf/1401.2454.pdf},
-which was later merged in a STOC'14 paper] (a major one being solving linear system), in some
-practical situations their advantages are less clear [from
-\url{https://link.springer.com/chapter/10.1007/978-3-319-20086-6_16}\enquote{for reasonable inputs
-the constant factors make the solver much slower than methods with higher asymptotic complexity.
-One other aspect predicted by theory is confirmed by our findings: Spanning trees with lower
-stretch indeed reduce the solver's running time. Yet, simple spanning tree algorithms perform
-better in practice than those with a guaranteed low stretch.} this is improved by
-\url{https://link.springer.com/chapter/10.1007%2F978-3-319-20086-6_17} although they seem to work
-	mostly with the Laplacian of the tree ]
-\fi
-
-\subsection{Empirical evaluation}
-\label{sub:gtx_empirical_evaluation}
-
-% \input{gtx_experiments}
-
-\section{Conclusion}
-
-\begingroup
-\small
-\setstretch{0.9}
-\todos
-
-\setlength\bibitemsep{2pt}
-\printbibliography
-\endgroup
-
-% \bibliography{/home/orphee/data/projects/biblio/library.bib,more.bib}
-% \bibliographystyle{plainnat}
-\end{document}

### 2017-08-23 cc/bias_main.tex
reorganize files and write some intro

Let us recall first the sign model we used for \dssn{}. Each node $u$ is endowed with two parameters
$p$ and $q$ drawn from an arbitrary joint distribution $\mu$ over $[0,1]\times [0,1]$. $p$ can be
interpreted as the tendency of $u$ to send positive edges to other nodes (\ie{} the
\enquote{niceness} of $u$), while $q$ can be interpreted as the tendency of $u$ to receive positive
edges from other nodes (\ie{} the \enquote{popularity} of $u$). This suggests that nodes in the
graph have a form of agency, and this imposes that edges are directed. We show experimentally that
failing to meet these two assumptions harms the performance of our previous method. This is not
surprising, for our bias is no more justified in that case. Therefore, we suggest a different bias,
drawing heavily on social balance theories, although we shall see later this holds for many
nonsocial graphs as well.

\subsection{Reliance on the sign generative model}
\label{sub:reliance_on_the_sign_generative_model}

\input{bias_model}

\subsection{Need for a directed graph}
\label{sub:need_for_a_directed_graph}

\input{bias_directed}

### 2017-08-23 cc/bias_model.tex
reorganize files and write some intro

--- 2017-07-21 lowstretch_and_cc.tex
@@
-\documentclass[nobib, a4paper, 10pt, oneside, onecolumn, openany, notitlepage, final,
-svgnames, marginals=raggedouter, english,
-%symmetric, justified,
-]{article}
-\usepackage{charter}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\input{packages}
-\input{abbreviations}
-
-\title{Edge sign prediction in general graphs and connection with Correlation Clustering}
-% \usetikzlibrary{external}
-% \tikzexternalize[mode=list and make, failed ref warnings for={\ref,\cite,\pageref},]
-\begin{document}
-\newgeometry{vmargin=1.3cm,left=1.5cm,textwidth=130mm,marginparsep=5mm,marginparwidth=52mm}
-\maketitle
-% \tableofcontents
-
-% \input{notations}
-
-\section{Limitations of the troll method}
-\label{sec:limitations_of_the_troll_method}
-
-\subsection{Reliance on the sign generative model}
-\label{sub:reliance_on_the_sign_generative_model}
-
 The method presented in the previous chapter hinges crucially upon our sign generative model. Yet
@@
 
-\subsubsection{Experience on directed biological network}
-\label{ssub:experience_on_directed_biological_network}
+\subsubsection{Experience on directed biological networks}
+\label{ssub:experience_on_directed_biological_networks}
 
@@
 deduplication) are inherently symmetric}
-\fi
-
-\subsection{Need for a directed graph}
-\label{sub:need_for_a_directed_graph}
-
-\iffalse
-A more serious limitation is that our sign model only applies to directed graphs. Consider
-another model, where each node $i$ is endowed with an integer $c_i \in \{1, \ldots, k\}$ that
-specifies its cluster and let the sign of the undirected edge $i,j$ be $+$ if $c_i = c_j$ and $-$
-otherwise.  This corresponds naturally to the situation modeled by the \pcc{} problem.
-\fi
-
-\section{\pcc{}}
-\label{sec:correlation_clustering}
-
-\input{cc_part}
-
-\newpage
-
-\section{Low stretch trees and spanners}
-\label{sec:low_stretch_trees_and_spanners}
-
-\input{notations}
-
-Given a graph $G$, and following the social balance theory roughly summarized as \enquote{the enemy
-of my enemy is my friend}, assume that the labelling of $E$ is consistent with a two-clustering of
-$V$. Namely, assume that $V$ can be partitioned in two clusters such that edges within each cluster
-are positive and edges across clusters are negative. In that case, the following
-\emph{multiplicative rule} holds: for any nodes $u$, $v$ in $V$, and any path $p$ between $u$ and
-$v$ in $G$, the sign $\yuv$ is equal to the product of the signs along $p$ (we call this product the
-parity of $p$ and denote it $\pi(p)$). While it is a simple and convenient model, this is too strong
-of a requirement to be satisfied in practice. Therefore, we relax it by assuming that we start with
-a consistent labeling $Y$ but can only observe a randomly perturbed version $Y'$ of $Y$.
-Specifically, given a constant $q\in [0, \nicefrac{1}{2})$, each sign of $Y$ is flipped with a
-probability smaller than $q$.
-
-In this section, we are interested in active learning algorithms, that first query a subset
-$\etrain$ of the edges, observe the signs in $\etrain$ and use them to predict the remaining signs.
-More precisely, we focus on an algorithm that queries spanning tree $T$ of $G$ and predicts the sign
-of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. While this
-can\Todo{state it formally then ;)} be stated more formally~\autocite[Section
-4.1]{Cesa-Bianchi2012b}, it makes sense intuitively that since each sign has been potentially
-flipped, the longer the path in $T$, the more likely its parity will be not be equal to the true
-sign $\yuv{}$. Therefore we would like each such path to be as short as possible. In the following,
-we describe a way to build spanning trees tailored for this situation.
-
-\subsection{\gtx{}: a spanning tree designed for sign prediction}
-\label{sub:gtx_a_simple_low_stretch_tree_construction}
-
-\input{gtx_algorithm}
-
-\subsection{Related works}
-\label{sub:gtx_related_works}
-
-\input{gtx_related}
-
-\iffalse
-\url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
-for instance the Elkin paper \enquote{Our centralized randomized algorithm computes (with
-probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in
-$O(|E|)$ time, whenever $k = \Omega(\log n)$. Note that when $k = \omega(\log n)$, the number of
-edges is $n(1+o(1))$, i.e., in this range the algorithm computes an ultra-sparse spanner in $O(|E|)$
-time.} For instance, if $k=5\log n$, we get a $10\log n$-spanner with $n\left(1+O\left(\frac{\log\log
-n}{n}\right)\right)$ edges in $O(|E|)$ time.
-
-While they have many applications [see first paragraph of \url{https://arxiv.org/pdf/1401.2454.pdf},
-which was later merged in a STOC'14 paper] (a major one being solving linear system), in some
-practical situations their advantages are less clear [from
-\url{https://link.springer.com/chapter/10.1007/978-3-319-20086-6_16}\enquote{for reasonable inputs
-the constant factors make the solver much slower than methods with higher asymptotic complexity.
-One other aspect predicted by theory is confirmed by our findings: Spanning trees with lower
-stretch indeed reduce the solver's running time. Yet, simple spanning tree algorithms perform
-better in practice than those with a guaranteed low stretch.} this is improved by
-\url{https://link.springer.com/chapter/10.1007%2F978-3-319-20086-6_17} although they seem to work
-	mostly with the Laplacian of the tree ]
-\fi
-
-\subsection{Empirical evaluation}
-\label{sub:gtx_empirical_evaluation}
-
-% \input{gtx_experiments}
-
-\section{Conclusion}
-
-\begingroup
-\small
-\setstretch{0.9}
-\todos
-
-\setlength\bibitemsep{2pt}
-\printbibliography
-\endgroup
-
-% \bibliography{/home/orphee/data/projects/biblio/library.bib,more.bib}
-% \bibliographystyle{plainnat}
-\end{document}

### 2017-08-23 cc/cc_main.tex
reorganize files and write some intro

% *TODO*
% say more clearly there are 3 methods for approximation so far: LP & region growing; LP & distance
% pivot; combinatorial pivot
% say that in practice, general graphs are assumed to be complete with missing edges being negative
% - “The problem is agnostic in the sense that the clustering of the data is not taken into account
%   or even assumed to exist.” another way to look at it is that we're trying to get a binary
%   classifier of edges (+ or -) using an hypothesis class (nodes clustering) with limited
%   expression power (as it has to respect transitivity constraints, whereas signs can be inconsistent)
% - when the graph is weighted, there is no much point separating complete from
%   general since weights can be 0

In this section, we introduce the \pcc{} problem (\autoref{sub:problem_setting}) and motivate it
both by general applications (\autoref{ssub:cc_applications}) and by its close connection with the
\esp{} problem (\autoref{sub:relation_with_edge_sign_prediction}). After discussing its theoretical
complexity and various approximation methods (\autoref{ssub:cc_harness_approx}) along other kinds of
methods (\autoref{ssub:cc_methods}), we will review related problems stemming from the original
formulation (\autoref{sub:variants_and_extensions}). In doing so, we present a thorough and updated
counterpart to existing surveys the same
topic~\autocites{bonchi2014correlation}{surveyCC16}{CCWirth2017}.

% surveyCC16 is about mind on complete graph (ie cluster editing). They give a Russian article from
% 1971 that says it cluster editing on triangle-free graph can be solved in polynomial time, which
% include bipartite graph (so for CC, if G^+ is (complete? I think so) bipartite)
% Fridman, G.Š.: A graph approximation problem. Upravlyaemye Sistemy 8, 73–75 (1971). (in Russian)
% In [14] it was observed that the problem is polynomially solvable for instances of at most 2
% clusters??! Actually it means consensus clustering can be solved in polynomial time when there are
% only 2 clustering as input

\subsection{Problem setting and applications}
\label{sub:problem_setting}
\input{cc_problem_applications}

\subsection{Relation with edge sign prediction}
\label{sub:relation_with_edge_sign_prediction}
\input{cc_sign_relation}

\subsection{State of the art}
\label{sub:state_of_the_art}

\subsubsection{Hardness and approximation}
\label{ssub:cc_harness_approx}
\input{cc_approx}
\subsubsection{Proposed approaches}
\label{ssub:cc_methods}
\input{cc_proposed}

\subsection{Variants and extensions}
\label{sub:variants_and_extensions}
\input{cc_variants}

\subsubsection{\pcc{} under stability assumption}
\label{ssub:cc_under_stability_assumption}
\input{cc_multimincut_stable}

### 2017-08-23 cc/cc_variants.tex
reorganize files and write some intro

--- 2017-08-22 cc/variants.tex
@@
 objects in different clusters, thus representing a kind of distance in the space of clusterings.  As
-first show by \textcite{EarlyConsensusClustering03}} , finding the optimal clustering is therefore
+first show by \textcite{EarlyConsensusClustering03} , finding the optimal clustering is therefore
 an instance of \pcc{} where the weights obey the triangular inequality. \textcite{Gionis2007} give a

### 2017-08-23 cc/conclusions.tex
reorganize files and write some intro

\begin{itemize}
	\item improve stable \pcc{}
	\item give guarantees for \gtx{}
	\item handle weights in \gtx{}
\end{itemize}

### 2017-08-23 cc/gtx_main.tex
reorganize files and write some intro

--- 2017-07-21 lowstretch_and_cc.tex
@@
-\documentclass[nobib, a4paper, 10pt, oneside, onecolumn, openany, notitlepage, final,
-svgnames, marginals=raggedouter, english,
-%symmetric, justified,
-]{article}
-\usepackage{charter}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
-\input{packages}
-\input{abbreviations}
-
-\title{Edge sign prediction in general graphs and connection with Correlation Clustering}
-% \usetikzlibrary{external}
-% \tikzexternalize[mode=list and make, failed ref warnings for={\ref,\cite,\pageref},]
-\begin{document}
-\newgeometry{vmargin=1.3cm,left=1.5cm,textwidth=130mm,marginparsep=5mm,marginparwidth=52mm}
-\maketitle
-% \tableofcontents
-
-% \input{notations}
-
-\section{Limitations of the troll method}
-\label{sec:limitations_of_the_troll_method}
-
-\subsection{Reliance on the sign generative model}
-\label{sub:reliance_on_the_sign_generative_model}
-
-The method presented in the previous chapter hinges crucially upon our sign generative model. Yet
-one can imagine contexts where this model is not applicable, especially when nodes do not represent
-human beings. One way to alleviate this issue was introduced with our online algorithm, where this
-time, signs are generated by an arbitrary adversary. However, in that case, we are still facing two
-limitations\marginpars{limitations is not the right word, maybe \emph{strong constraints}…}
-\begin{enumerate}
-	\item Our bias remains that the labeling is regular (recall this means informally that the all
-		the outgoing signs from a given node tend to be the same, and likewise for the incoming signs),
-		since irregularities are the cost payed by the adversary to make our algorithm mispredict. While
-		this bias is well suited to social networks, other applications may require other
-		bias\marginpars{namely the \pcc{} bias we'll introduce next}.
-	\item In the online setting, we evaluate our performance by the regret\marginpars{ref to regret equation},
-		whereas in general we are interested in more classical measures of predictive accuracy.
-\end{enumerate}
-
-\subsubsection{Experience on directed biological network}
-\label{ssub:experience_on_directed_biological_network}
-
-Take for instance a biological network.
-
-According to \href{https://web.stanford.edu/class/cs224w/slides/handout-bionets.pdf}%
-{this description of various biological networks}, \emph{gene regulatory network} are directed and
-contains activation and inhibition links, as \href{https://en.wikipedia.org/wiki/Gene_regulatory_network#Overview}%
-{showed on Wikipedia}. There are some online databases such as
-\href{http://regulondb.ccg.unam.mx/menu/download/datasets/index.jsp}%
-{RegulonDB} or \href{http://www.pathwaycommons.org/pcviz/}{Pathway Commons} (this last one provide
-network in the BioPAX format, which can be visualized by
-\href{http://www.cytoscape.org/}{CytoScape}).
-Another source is the \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2708159/table/T1}{Table 1}
-of \cite{BioSigned09}.
-%DasGupta also describe (in
-%http://www.sciencedirect.com/science/article/pii/S0303264706001419#sec14) how
-%they built a directed signed network from this SBML file
-%https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1681468/bin/msb4100014-sd1.xml
-
-other domain? coref, images (is it directed?), entity resolution…\marginpars{Actually it's difficult
-to find other kind of graph because all constructed ones (like image, coreference,
-deduplication) are inherently symmetric}
-\fi
-
-\subsection{Need for a directed graph}
-\label{sub:need_for_a_directed_graph}
-
-\iffalse
-A more serious limitation is that our sign model only applies to directed graphs. Consider
-another model, where each node $i$ is endowed with an integer $c_i \in \{1, \ldots, k\}$ that
-specifies its cluster and let the sign of the undirected edge $i,j$ be $+$ if $c_i = c_j$ and $-$
-otherwise.  This corresponds naturally to the situation modeled by the \pcc{} problem.
-\fi
-
-\section{\pcc{}}
-\label{sec:correlation_clustering}
-
-\input{cc_part}
-
-\newpage
-
-\section{Low stretch trees and spanners}
-\label{sec:low_stretch_trees_and_spanners}
-
-\input{notations}
-
 Given a graph $G$, and following the social balance theory roughly summarized as \enquote{the enemy
@@
 
-\iffalse
-\url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
-for instance the Elkin paper \enquote{Our centralized randomized algorithm computes (with
-probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in
-$O(|E|)$ time, whenever $k = \Omega(\log n)$. Note that when $k = \omega(\log n)$, the number of
-edges is $n(1+o(1))$, i.e., in this range the algorithm computes an ultra-sparse spanner in $O(|E|)$
-time.} For instance, if $k=5\log n$, we get a $10\log n$-spanner with $n\left(1+O\left(\frac{\log\log
-n}{n}\right)\right)$ edges in $O(|E|)$ time.
-
-While they have many applications [see first paragraph of \url{https://arxiv.org/pdf/1401.2454.pdf},
-which was later merged in a STOC'14 paper] (a major one being solving linear system), in some
-practical situations their advantages are less clear [from
-\url{https://link.springer.com/chapter/10.1007/978-3-319-20086-6_16}\enquote{for reasonable inputs
-the constant factors make the solver much slower than methods with higher asymptotic complexity.
-One other aspect predicted by theory is confirmed by our findings: Spanning trees with lower
-stretch indeed reduce the solver's running time. Yet, simple spanning tree algorithms perform
-better in practice than those with a guaranteed low stretch.} this is improved by
-\url{https://link.springer.com/chapter/10.1007%2F978-3-319-20086-6_17} although they seem to work
-	mostly with the Laplacian of the tree ]
-\fi
-
 \subsection{Empirical evaluation}
@@
 
-% \input{gtx_experiments}
-
-\section{Conclusion}
-
-\begingroup
-\small
-\setstretch{0.9}
-\todos
-
-\setlength\bibitemsep{2pt}
-\printbibliography
-\endgroup
-
-% \bibliography{/home/orphee/data/projects/biblio/library.bib,more.bib}
-% \bibliographystyle{plainnat}
-\end{document}
+\input{gtx_experiments}

### 2017-08-23 cc/gtx_related.tex
reorganize files and write some intro

--- 2017-07-11 gtx_related.tex
@@
 
+\url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
+for instance the Elkin paper \enquote{Our centralized randomized algorithm computes (with
+probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in
+$O(|E|)$ time, whenever $k = \Omega(\log n)$. Note that when $k = \omega(\log n)$, the number of
+edges is $n(1+o(1))$, i.e., in this range the algorithm computes an ultra-sparse spanner in $O(|E|)$
+time.} For instance, if $k=5\log n$, we get a $10\log n$-spanner with $n\left(1+O\left(\frac{\log\log
+n}{n}\right)\right)$ edges in $O(|E|)$ time.
+
 the subgraph $H$ is said to be an $(\alpha, \beta)$-spanner of $G$ if, for a pair of parameters
@@
 % 10.1109/SFCS.1993.366822
-We construct t-spanners of size (number of edges) Õ(n 1+(2+\epsilon)/t )
-(for any \epsilon > 0 and t such that t/(2+\epsilon) is integral). These spanners can be constructed
-by a randomized algorithm that runs in Õ(mn (2+\epsilon)/t ) time.
+% We construct t-spanners of size (number of edges) Õ(n 1+(2+\epsilon)/t )
+% (for any \epsilon > 0 and t such that t/(2+\epsilon) is integral). These spanners can be constructed
+% by a randomized algorithm that runs in Õ(mn (2+\epsilon)/t ) time.
 
@@
 % http://ieeexplore.ieee.org/document/81738
+
+While they have many applications [see first paragraph of \url{https://arxiv.org/pdf/1401.2454.pdf},
+which was later merged in a STOC'14 paper] (a major one being solving linear systems), in some
+practical situations their advantages are less clear [from
+\url{https://link.springer.com/chapter/10.1007/978-3-319-20086-6_16}\enquote{for reasonable inputs
+the constant factors make the solver much slower than methods with higher asymptotic complexity.
+One other aspect predicted by theory is confirmed by our findings: Spanning trees with lower
+stretch indeed reduce the solver's running time. Yet, simple spanning tree algorithms perform
+better in practice than those with a guaranteed low stretch.} this is improved by
+\url{https://link.springer.com/chapter/10.1007%2F978-3-319-20086-6_17} although they seem to work
+	mostly with the Laplacian of the tree ]

### 2017-08-23 cc/main.tex
reorganize files and write some intro

In the previous chapter, we presented a learning bias for the \esp{} problem, namely our sign
generative model. By leveraging its rigorous theoretical guarantees, we have demonstrated
empirically its good performance on real \dssn{}. However, in this chapter, we consider other kinds
of signed networks and show they require another bias and different, more combinatorial algorithms.
More precisely, we start by pointing in \autoref{sec:new_bias} that our previous bias may not apply
to directed graphs from domains other than social science, or to undirected graphs. Motivated by
balance theories, we then introduce a new bias, assuming that ideally, nodes are grouped in $K$
clusters such that all edges within clusters are positive and all edges between cluster are
negative. Recovering such clusters from a given signed graph has been studied extensively in the
last decades under the \pcc{} name. In \autoref{sec:correlation_clustering}, we show how this
difficult combinatorial clustering problem is connected to our learning bias. We survey a wide range
of methods to solve it exactly, approximately or heuristically, paying special attention to settings
where such recovery is easier, such as noisy or stable instances.  Finally, focusing on the
important and convenient special case $K=2$, we develop in \autoref{sec:gtx} an algorithm that,
given an undirected general graph, produce a spanning tree designed to support fast and accurate
edge sign prediction.

\input{../notations}

\section{A bias for general graphs}
\label{sec:new_bias}

\input{bias_main}

\section{\pcc{}}
\label{sec:correlation_clustering}

\input{cc_main}

\section{Low stretch trees and spanners}
\label{sec:gtx}

\input{gtx_main}

\section{Conclusions}
\label{sec:cc_conclusions}

\input{conclusions}

### 2017-08-23 lowstretch_and_cc.tex
reorganize files and write some intro

@@
-\documentclass[nobib, a4paper, 10pt, oneside, onecolumn, openany, notitlepage, final,
+\documentclass[nobib, a4paper, 10pt, oneside, onecolumn, openany, notitlepage, draft,
 svgnames, marginals=raggedouter, english,
@@
 
-\title{Edge sign prediction in general graphs and connection with Correlation Clustering}
+\title{Edge sign prediction in general graphs and Correlation Clustering}
 % \usetikzlibrary{external}
@@
 \begin{document}
-\newgeometry{vmargin=1.3cm,left=1.5cm,textwidth=130mm,marginparsep=5mm,marginparwidth=52mm}
+\newgeometry{vmargin=1.3cm,left=1.5cm,textwidth=150mm,marginparsep=5mm,marginparwidth=32mm}
 \maketitle
@@
 
-% \input{notations}
-
-\section{Limitations of the troll method}
-\label{sec:limitations_of_the_troll_method}
-
-\subsection{Reliance on the sign generative model}
-\label{sub:reliance_on_the_sign_generative_model}
-
-\iffalse
-The method presented in the previous chapter hinges crucially upon our sign generative model. Yet
-one can imagine contexts where this model is not applicable, especially when nodes do not represent
-human beings. One way to alleviate this issue was introduced with our online algorithm, where this
-time, signs are generated by an arbitrary adversary. However, in that case, we are still facing two
-limitations\marginpars{limitations is not the right word, maybe \emph{strong constraints}…}
-\begin{enumerate}
-	\item Our bias remains that the labeling is regular (recall this means informally that the all
-		the outgoing signs from a given node tend to be the same, and likewise for the incoming signs),
-		since irregularities are the cost payed by the adversary to make our algorithm mispredict. While
-		this bias is well suited to social networks, other applications may require other
-		bias\marginpars{namely the \pcc{} bias we'll introduce next}.
-	\item In the online setting, we evaluate our performance by the regret\marginpars{ref to regret equation},
-		whereas in general we are interested in more classical measures of predictive accuracy.
-\end{enumerate}
-\fi
-
-\subsubsection{Experience on directed biological networks}
-\label{ssub:experience_on_directed_biological_networks}
-
-\iffalse
-Take for instance a biological network.
-
-According to \href{https://web.stanford.edu/class/cs224w/slides/handout-bionets.pdf}%
-{this description of various biological networks}, \emph{gene regulatory network} are directed and
-contains activation and inhibition links, as \href{https://en.wikipedia.org/wiki/Gene_regulatory_network#Overview}%
-{showed on Wikipedia}. There are some online databases such as
-\href{http://regulondb.ccg.unam.mx/menu/download/datasets/index.jsp}%
-{RegulonDB} or \href{http://www.pathwaycommons.org/pcviz/}{Pathway Commons} (this last one provide
-network in the BioPAX format, which can be visualized by
-\href{http://www.cytoscape.org/}{CytoScape}).
-Another source is the \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2708159/table/T1}{Table 1}
-of \cite{BioSigned09}.
-%DasGupta also describe (in
-%http://www.sciencedirect.com/science/article/pii/S0303264706001419#sec14) how
-%they built a directed signed network from this SBML file
-%https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1681468/bin/msb4100014-sd1.xml
-
-other domain? coref, images (is it directed?), entity resolution…\marginpars{Actually it's difficult
-to find other kind of graph because all constructed ones (like image, coreference,
-deduplication) are inherently symmetric}
-\fi
-
-\subsection{Need for a directed graph}
-\label{sub:need_for_a_directed_graph}
-
-\iffalse
-A more serious limitation is that our sign model only applies to directed graphs. Consider
-another model, where each node $i$ is endowed with an integer $c_i \in \{1, \ldots, k\}$ that
-specifies its cluster and let the sign of the undirected edge $i,j$ be $+$ if $c_i = c_j$ and $-$
-otherwise.  This corresponds naturally to the situation modeled by the \pcc{} problem.
-\fi
-
-\section{\pcc{}}
-\label{sec:correlation_clustering}
-
-\input{cc_part}
-
-\newpage
-
-\iffalse
-\section{Low stretch trees and spanners}
-\label{sec:low_stretch_trees_and_spanners}
-
-\input{notations}
-
-Given a graph $G$, and following the social balance theory roughly summarized as \enquote{the enemy
-of my enemy is my friend}, assume that the labelling of $E$ is consistent with a two-clustering of
-$V$. Namely, assume that $V$ can be partitioned in two clusters such that edges within each cluster
-are positive and edges across clusters are negative. In that case, the following
-\emph{multiplicative rule} holds: for any nodes $u$, $v$ in $V$, and any path $p$ between $u$ and
-$v$ in $G$, the sign $\yuv$ is equal to the product of the signs along $p$ (we call this product the
-parity of $p$ and denote it $\pi(p)$). While it is a simple and convenient model, this is too strong
-of a requirement to be satisfied in practice. Therefore, we relax it by assuming that we start with
-a consistent labeling $Y$ but can only observe a randomly perturbed version $Y'$ of $Y$.
-Specifically, given a constant $q\in [0, \nicefrac{1}{2})$, each sign of $Y$ is flipped with a
-probability smaller than $q$.
-
-In this section, we are interested in active learning algorithms, that first query a subset
-$\etrain$ of the edges, observe the signs in $\etrain$ and use them to predict the remaining signs.
-More precisely, we focus on an algorithm that queries spanning tree $T$ of $G$ and predicts the sign
-of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. While this
-can\Todo{state it formally then ;)} be stated more formally~\autocite[Section
-4.1]{Cesa-Bianchi2012b}, it makes sense intuitively that since each sign has been potentially
-flipped, the longer the path in $T$, the more likely its parity will be not be equal to the true
-sign $\yuv{}$. Therefore we would like each such path to be as short as possible. In the following,
-we describe a way to build spanning trees tailored for this situation.
-
-\subsection{\gtx{}: a spanning tree designed for sign prediction}
-\label{sub:gtx_a_simple_low_stretch_tree_construction}
-
-% \input{gtx_algorithm}
-
-\subsection{Related works}
-\label{sub:gtx_related_works}
-
-% \input{gtx_related}
-
-\url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
-for instance the Elkin paper \enquote{Our centralized randomized algorithm computes (with
-probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in
-$O(|E|)$ time, whenever $k = \Omega(\log n)$. Note that when $k = \omega(\log n)$, the number of
-edges is $n(1+o(1))$, i.e., in this range the algorithm computes an ultra-sparse spanner in $O(|E|)$
-time.} For instance, if $k=5\log n$, we get a $10\log n$-spanner with $n\left(1+O\left(\frac{\log\log
-n}{n}\right)\right)$ edges in $O(|E|)$ time.
-
-While they have many applications [see first paragraph of \url{https://arxiv.org/pdf/1401.2454.pdf},
-which was later merged in a STOC'14 paper] (a major one being solving linear system), in some
-practical situations their advantages are less clear [from
-\url{https://link.springer.com/chapter/10.1007/978-3-319-20086-6_16}\enquote{for reasonable inputs
-the constant factors make the solver much slower than methods with higher asymptotic complexity.
-One other aspect predicted by theory is confirmed by our findings: Spanning trees with lower
-stretch indeed reduce the solver's running time. Yet, simple spanning tree algorithms perform
-better in practice than those with a guaranteed low stretch.} this is improved by
-\url{https://link.springer.com/chapter/10.1007%2F978-3-319-20086-6_17} although they seem to work
-	mostly with the Laplacian of the tree ]
-\fi
-
-\subsection{Empirical evaluation}
-\label{sub:gtx_empirical_evaluation}
-
-% \input{gtx_experiments}
-
-\section{Conclusion}
+\subimport*{cc/}{main}
 
@@
 
-% \bibliography{/home/orphee/data/projects/biblio/library.bib,more.bib}
-% \bibliographystyle{plainnat}
 \end{document}

### 2017-08-23 cc/bias_directed.tex
rewrite the motivation for a new bias

@@
-A more serious limitation is that our sign model only applies to directed graphs. Consider
-another model, where each node $i$ is endowed with an integer $c_i \in \{1, \ldots, k\}$ that
-specifies its cluster and let the sign of the undirected edge $i,j$ be $+$ if $c_i = c_j$ and $-$
-otherwise.  This corresponds naturally to the situation modeled by the \pcc{} problem.
+Having two parameters per node, one for outgoing edges and another one for incoming edges clearly
+targets directed graphs. Many online interactions are inherently directed, for instance friendship,
+trust or communication. On the other hand, predicting edge signs in undirected graphs is an equally
+relevant objective. A prime example of such a situation is when we are given $n$ objects, some
+pairwise similarities them, and the similarity function itself, which unfortunately takes an
+exorbitant time to be evaluated. There is an underlying graph and being able to predict the sign of
+its undirected edges would save us expensive evaluations of the symmetric similarity function.
+
+A trivial way to turn an undirected graph $G$ into a directed graph $G'$ is to let $V'=V$ and, for
+every edge $(u,v)$ in $E$, to add both \euv{} and \evu{} to $E'$ with the same sign as $(u,v)$. In
+terms of our generative model, this corresponds to putting all the probability mass of $\mu$ on the
+$p=q$ diagonal. This is clearly not a very satisfactory solution, for it removes one degree of
+freedom from the model. To illustrate this point, we conduct the following experiment. We use our
+previous \dssn{} datasets and remove the edge direction. As mentioned earlier, for few pairs of
+nodes there are reciprocal edges of different signs, in which case we pick a sign arbitrarily. Given
+those undirected graphs, we orient them using the approach described above and compare our method
+with the \complowrank{} approach ran directly on the undirected graphs. This is a fair comparison,
+for \complowrank{} is designed to works solely on undirected graphs.\Todo{run the undirected
+experiments.}

### 2017-08-23 cc/bias_main.tex
rewrite the motivation for a new bias

@@
 Let us recall first the sign model we used for \dssn{}. Each node $u$ is endowed with two parameters
-$p$ and $q$ drawn from an arbitrary joint distribution $\mu$ over $[0,1]\times [0,1]$. $p$ can be
+$p_u$ and $q_u$ drawn from an arbitrary joint distribution $\mu$ over $[0,1]\times [0,1]$. $p_u$ can be
 interpreted as the tendency of $u$ to send positive edges to other nodes (\ie{} the
-\enquote{niceness} of $u$), while $q$ can be interpreted as the tendency of $u$ to receive positive
-edges from other nodes (\ie{} the \enquote{popularity} of $u$). This suggests that nodes in the
+\enquote{niceness} of $u$), while $q_u$ can be interpreted as the tendency of $u$ to receive positive
+edges from other nodes (\ie{} the \enquote{popularity} of $u$). The sign of an edge \euv{} is then
+chosen to be positive with probability $\frac{p_u+q_v}{2}$. This suggests that nodes in the
 graph have a form of agency, and this imposes that edges are directed. We show experimentally that
@@
 
-\subsection{Reliance on the sign generative model}
-\label{sub:reliance_on_the_sign_generative_model}
+\subsection{Sign generative model and behavior}
+\label{sub:bias_sign_behavior}
 
@@
 
-\subsection{Need for a directed graph}
+\subsection{Directed edges requirement}
 \label{sub:need_for_a_directed_graph}

### 2017-08-23 cc/bias_model.tex
rewrite the motivation for a new bias

@@
-The method presented in the previous chapter hinges crucially upon our sign generative model. Yet
-one can imagine contexts where this model is not applicable, especially when nodes do not represent
-human beings. One way to alleviate this issue was introduced with our online algorithm, where this
-time, signs are generated by an arbitrary adversary. However, in that case, we are still facing two
-limitations\marginpars{limitations is not the right word, maybe \emph{strong constraints}…}
-\begin{enumerate}
-	\item Our bias remains that the labeling is regular (recall this means informally that the all
-		the outgoing signs from a given node tend to be the same, and likewise for the incoming signs),
-		since irregularities are the cost payed by the adversary to make our algorithm mispredict. While
-		this bias is well suited to social networks, other applications may require other
-		bias\marginpars{namely the \pcc{} bias we'll introduce next}.
-	\item In the online setting, we evaluate our performance by the regret\marginpars{ref to regret equation},
-		whereas in general we are interested in more classical measures of predictive accuracy.
-\end{enumerate}
+We can interpret the description of our sign model as if, when establishing a new link, nodes were
+deciding the sign of this link based on their individual preferences. While this makes sense in
+social networks where nodes represent human beings, one can imagine contexts where this model is not
+applicable. Another model, \eg{} for proteins, could be that proteins belongs to functional groups
+and that two proteins interact positively if they belong to the same group, negatively otherwise.
+We presented a way to circumvent this notion of node behavior with our online algorithm, where this
+time, signs are generated by an arbitrary adversary. However, even in that case, our bias remains
+that the labeling is regular. Recall this means informally that all the outgoing signs from a given
+node tend to be the same, and likewise for the incoming signs. Indeed, irregularities are the cost
+payed by the adversary to make our algorithm mispredict. In other words, regularity hints at a
+consistent sign behavior from nodes, that we cannot assume carry out outside of social networks.
+Pursuing the proteins example, we can imagine that proteins interact half of the time with proteins
+of their own group, and half of time with proteins from other groups. This would correspond to
+maximum irregularity, yet it is a plausible situation. We now show experimentally, that in fact,
+biological networks do not lend themselves to our sign model bias.
 
-\subsubsection{Experience on directed biological networks}
-\label{ssub:experience_on_directed_biological_networks}
+% In the online setting, we evaluate our performance by the regret\marginpars{ref to regret equation},
+% whereas in general we are interested in more classical measures of predictive accuracy.
 
-Take for instance a biological network.
-
-According to \href{https://web.stanford.edu/class/cs224w/slides/handout-bionets.pdf}%
-{this description of various biological networks}, \emph{gene regulatory network} are directed and
-contains activation and inhibition links, as \href{https://en.wikipedia.org/wiki/Gene_regulatory_network#Overview}%
-{showed on Wikipedia}. There are some online databases such as
-\href{http://regulondb.ccg.unam.mx/menu/download/datasets/index.jsp}%
-{RegulonDB} or \href{http://www.pathwaycommons.org/pcviz/}{Pathway Commons} (this last one provide
-network in the BioPAX format, which can be visualized by
-\href{http://www.cytoscape.org/}{CytoScape}).
-Another source is the \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2708159/table/T1}{Table 1}
-of \cite{BioSigned09}.
-%DasGupta also describe (in
-%http://www.sciencedirect.com/science/article/pii/S0303264706001419#sec14) how
-%they built a directed signed network from this SBML file
-%https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1681468/bin/msb4100014-sd1.xml
-
-other domain? coref, images (is it directed?), entity resolution…\marginpars{Actually it's difficult
-to find other kind of graph because all constructed ones (like image, coreference,
-deduplication) are inherently symmetric}
+Namely, we consider \emph{gene regulatory networks}. The nodes of such directed graphs are various
+chemicals (such as genes, proteins or messenger RNAs) and a directed edge \euv{} indicates that one
+node $u$ influences the concentration of another node $v$ through a chemical reaction. This
+influence can be positive (that is, an increase of $u$ concentration results in an increase of $v$
+concentration) or negative (that is, an increase of $u$ concentration results in a decrease of $v$
+concentration). In this context, we could now interpret the \enquote{niceness} of $u$ as it ability
+to accelerate chemical reactions, and its \enquote{popularity} as its propensity to take part in
+reactions accelerated by other nodes. Intuitively, this is quite far-fetched. To demonstrate this,
+we borrow several gene regulatory networks from \autocite[Table 1]{BioSigned09} and perform the same
+experiments as in the previous chapter.\Todo{get the data and run the experiments.}

### 2017-08-23 cc/bias_balance.tex
move social balance to first section

--- 2017-08-21 cc/sign_relation.tex
@@
-Now that we are familiar with the \pcc{} problem and its numerous applications, let us return to the
-original problem of this chapter, predicting the edge sign of a signed graph. In order to do so, we
-will first focus on the case where a signed graph has no disagreement (that is the minimum of
-\eqref{eq:mind} is $0$) and is unweighted, as it highlights more clearly the connection with the
-balance theory. We focus here on balance theory, for it is both compelling and well-established.
+From a narrow practical point of view, balance theories specify which sign assignments of a triangle
+are possible. By forbidding triangles with a single negative edge, weak balance gives rise to a
+$K$-\emph{consistent} clusters structure on complete graphs. After making these notions more precise
+in the rest of this section, we conclude by precisely exposing our learning bias.
+
+\begin{aside}
+We focus here on balance theory, for it is both compelling and well-established.
 Indeed, as illustrated by the account of \textcite{CSBalanceSurvey15}, it has been used extensively
-in the last half century to study signed networks\marginpars{It was even used to study fairy
-tales~\autocite{fairyTales80}.} and their dynamics.
+in the last half century to study signed networks and their dynamics (an example of this far
+reaching influence is the analysis of the relation between characters in fairy
+tales~\autocite{fairyTales80}).
 However, it has shortcomings, the main one being its inability to deal with directed graphs.
@@
 in~\autocite{Yap2015}.
+\end{aside}
 
 In his seminal work on interpersonal relations,
-\citeauthor*{Heider46}~\autocites{Heider46}{HeiderBook58} posits through
+\textcite{Heider46} posits through
 psychological and sociological arguments that in order to reduce their cognitive dissonance, three
@@
 
-We will use triangles as the building block of our definition of strongly balanced graphs.
+We use triangles as the building block of our definition of strongly balanced graphs.
 \begin{definition}[strongly balanced complete graphs]
@@
   \includegraphics[width=28mm]{assets/raw/2cc.pdf}
-  \caption{A two-clustering a complete strongly balanced graph}
+  \caption{A two-clustering of a complete strongly balanced graph}
   \label{fig:cc_twostrongcc}
 \end{marginfigure}
-As we will see later, these two facts are actually equivalent.
+We say that $L$ and $R$ are $2$-consistent clusters and more generally, we define
+\begin{definition}[$K$-consistent clustering]
+  Given a graph $G=(V, E)$, a clustering $\cluster=\{C_1, \ldots, C_K\}$ of $V$ is \emph{consistent}
+  with the signs of $E$ if for every edge $(u,v)$ in $E$:
+  \begin{enumerate}[(i),nosep]
+    \item $(u,v)$ is positive if $\cluster(u)=\cluster(v)$, and
+    \item $(u,v)$ is negative if $\cluster(u)\neq\cluster(v)$.
+  \end{enumerate}
+\end{definition}
 
-Those two clusters represent a natural partition with no disagreements in the sense of \pcc{}. However,
-the existence of two such clusters is only a sufficient but not necessary condition for having no
-disagreement. As hinted in \autoref{sub:need_for_a_directed_graph}, when the nodes are partitioned
-into $k\geq 2$ clusters, letting edges within clusters to be positive and edges across clusters to
-be negative will by construction results in zero disagreement. Furthermore, it was noted by
+We just saw that if a complete graphs is strongly balanced, then it admits a $2$-consistent
+clustering (\autoref{thm:structural} proves the converse is true). It is natural to ask under which
+conditions a complete signed graph admits a $K>2$ consistent clustering. This is where the notion of
+weak balance comes handy. It was noted by
 sociologists that among triangles with odd number of negative edges, one is more stable and common
@@
 weak balance by considering triangles with three negative edges to be balanced, as illustrated on
-\autoref{fig:cc_balance}. Formally, we can modify our two previous definitions to arrive at weak
-balance:
+\autoref{fig:cc_balance}. Formally, we can modify our two previous definitions of balance to arrive
+at weak balance:
 \begin{definition}[Weak balance]
@@
 In that case as well, the local property of triangles has structural implication for the whole graph.
-Namely, we can partition a complete weakly balanced graph into $k$ clusters of positively connected
-nodes that are negatively tied with the rest of the graph. Consider the same construction as in
+Namely, we can partition a complete weakly balanced graph into $K$ consistent clusters. 
+Consider the same construction as in
 \autoref{fig:cc_twostrongcc}, putting a node $u$ and its positive neighborhood in one cluster $L$ and
@@
 \end{definition}
-To similarly extend the characterization of signed graphs by consistent clustering, we need to
+To recover the characterization of signed graphs by consistent clustering, we need to
 consider longer cycles than triangles.
@@
 
-Being strongly balanced is equivalent to having two consistent clusters, as proved in the seminal
+Being strongly balanced is equivalent to having a $2$-consistent clustering, as proved in the landmark
 theorem of \textcite[Theorem 3]{harary1953}:
-\begin{theorem}[Structural Theorem]
-  A graph $G=(V,E)$ is strongly balanced if and only if $V$ can be divided in two disjoint clusters
-  such that all edges within clusters are positive and all edges between clusters are negative.
+\begin{theorem}[Structural Theorem]\label{thm:structural}
+  A graph $G=(V,E)$ is strongly balanced if and only if $V$ admits a $2$-consistent clustering.
 \end{theorem}
 
+\begin{aside}
 According to \textcite{Huffner2010}, a similar theorem was proved earlier by \textcite[Theorem
@@
 theory.}
+\end{aside}
 
@@
 \begin{theorem}
-  A graph $G=(V,E)$ is weakly balanced if and only if $V$ can be divided in $k$ disjoint clusters
-  such that all edges within clusters are positive and all edges between clusters are negative.
+  A graph $G=(V,E)$ is weakly balanced if and only if $V$ admits a $K$-consistent clustering.
 \end{theorem}
@@
   C_1$. Note also that $|V\setminus C_1| < |V|$ and therefore we can repeat this
-  procedure until $V$ is exhausted, at which point we have obtained our $k$ clusters verifying the
-  theorem's statement conditions.
+  procedure until $V$ is exhausted, at which point we have obtained our $K$ consistent clusters.
 
-  Conversely, assume we have such $k$ clusters. We can complete the graph $G$ by letting the edge
+  Conversely, assume we have $K$ consistent clusters. We can complete the graph $G$ by letting the edge
   $(u,v)$ be positive if $u$ and $v$ are in the same cluster and negative otherwise. Let us the pick
@@
 given to longer cycles compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
-{measureUnbalance17}. For the same reason, in virtually any real instances of the \pcc{} problem, the%
+{measureUnbalance17}.
 % there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration
 % index) which they claimed has been peer reviewed?
+% For the same reason, in virtually any real instances of the \pcc{} problem, the optimal solution
+% will incur some disagreements. 
 \marginpars{The proof makes essential use of the strong balance (which is ok since next section also
@@
 it would be interesting to see if it can be extended.}
-optimal solution will incur some disagreements. \Textcite[Section 5.5]{BookKleinberg2010}
+\Textcite[Section 5.5]{BookKleinberg2010}
 nonetheless prove that even when only a fraction of the triangle are strongly balanced in a complete
@@
 
+Using these results on weak balance, we can now make our new bias explicit. Drawing parallel with
+the previous generative model, we assume that each node $u$ is endowed with an integer $\cluster(u)
+\in \{1, \ldots, K\}$ that specifies its cluster and we let the undirected edge $u,v$ be positive if
+$\cluster(u) = \cluster(v)$ and negative otherwise. In other words, this corresponds to picking an
+arbitrary partition of $V$ and setting the signs of $E$ in such a way that the partition becomes a
+consistent clustering. This can still be seen as a generative process because of the initial integer
+drawn \uar{},\Todo{Although I'm not sure how I would write this likelihood to be fair} but one that
+avoids the two pitfalls discussed before. Indeed, there is no decision to be made by nodes, since
+the previous probability has been replaced by a binary function (either both nodes are in the same
+cluster or not). Furthermore, since the clustering function is symmetric, the model is inherently
+suited to undirected graphs.
+
+\iffalse
 We can conclude that balance theory helps characterize zero disagreement instances of \pcc{}.
@@
 
-\iffalse
 strong balance random graph model \url{https://www.lri.fr/~yannis/randsigned.pdf}, line index is basically

### 2017-08-23 packages.tex
add few more ref
put all CC material into one folder
small typos at beginning of CC
some proofreading and small additions
add few more references
reorganize files and write some intro
rewrite the motivation for a new bias
move social balance to first section

@@
 \usepackage{setspace}
+\usepackage{xspace}
 \usepackage{microtype}
@@
 sorting=nyvt, backref=true, backend=biber, defernumbers=true,maxbibnames=60]{biblatex}
+\renewcommand*{\bibfont}{\normalfont\footnotesize}
 \AtEveryBibitem{
-   \clearfield{arxivId}
+   % \clearfield{arxivId}
    % \clearfield{booktitle}
    % \clearfield{doi}
-   \clearfield{eprint}
+   % \clearfield{eprint}
    \clearfield{eventdate}
@@
 \usepackage{amsthm}
+\usepackage{import}
 \usepackage{mathtools}
 \usepackage{sidenotes}
+\let\bottomfootnote\footnote
 \renewcommand{\footnote}[1]{\sidenote{\footnotesize #1}}
-\usepackage{enumitem}
+\usepackage[shortlabels]{enumitem}
 \usepackage{tabulary}
@@
 \def\algorithmautorefname{Algorithm}
+
+\usepackage{mdframed}
+\newenvironment{aside}%
+{\begin{mdframed}[hidealllines=true, leftline=true, innertopmargin=0.1ex, innerbottommargin=0.1ex,
+      innermargin=2.3em, leftmargin=2.3em, outermargin=0em, innerleftmargin=0.7em, innerrightmargin=0pt,
+      linewidth=2.0pt, skipabove=\baselineskip, skipbelow=\baselineskip, linecolor=Gray]\small}%
+{\end{mdframed}}
 

### 2017-08-24 cc/cc_main.tex
change intro of CC

@@
 
-In this section, we introduce the \pcc{} problem (\autoref{sub:problem_setting}) and motivate it
-both by general applications (\autoref{ssub:cc_applications}) and by its close connection with the
-\esp{} problem (\autoref{sub:relation_with_edge_sign_prediction}). After discussing its theoretical
-complexity and various approximation methods (\autoref{ssub:cc_harness_approx}) along other kinds of
-methods (\autoref{ssub:cc_methods}), we will review related problems stemming from the original
-formulation (\autoref{sub:variants_and_extensions}). In doing so, we present a thorough and updated
-counterpart to existing surveys the same
-topic~\autocites{bonchi2014correlation}{surveyCC16}{CCWirth2017}.
+As mentioned earlier, most real signed graphs are not fully balanced and thus do not have a
+perfectly consistent clustering. The problem of quantifying how much a given graph departs from this
+ideal situation is called \pcc{}. After defining it formally in \autoref{sub:problem_setting}, we
+show that \pcc{} is a learning problem on its own, and discuss how it relates to the binary
+classification problem of predicting edge signs, among many others applications. We then present
+numerous methods to solve \pcc{} in \autoref{sub:state_of_the_art}, ranging from exact methods (that
+implicitly leverage our bias) to heuristics (whose greedy and energy minimization frameworks are
+classic in unsupervised graph cut problems). We also devote a large section to approximation
+methods, in order to get a sense of why the problem is \NPc{}. This allows us to finish in
+\autoref{sub:cc_non_worst_case} by describing situations related to our bias where is possible to
+solve the problem efficiently and optimally. By doing so, we update existing surveys on the same
+topic~\autocites{bonchi2014correlation}{surveyCC16}{CCWirth2017}, highlighting the importance of
+\pcc{} to solve our binary learning problem.
 
@@
 
-\subsection{Problem setting and applications}
+\subsection{Problem setting and connection to \esp{}}
 \label{sub:problem_setting}
 \input{cc_problem_applications}
-
-\subsection{Relation with edge sign prediction}
-\label{sub:relation_with_edge_sign_prediction}
-\input{cc_sign_relation}
 

### 2017-08-24 cc/cc_problem_applications.tex
change intro of CC

--- 2017-08-21 cc/problem_applications.tex
@@
 
-Now consider a clustering \cluster{} of $V$, that is a function from $V$ to $\Nbb{}^{|V|}_{>0}$
-that assigns to each node a cluster index.\marginpars{We will also abuse the notation and let
-\cluster{} be a set of clusters $\cluster=\{C_1, C_2, \ldots, C_k\}$ that form a partition of $V$.
-That is, $\forall i\,C_i\subset V$, $\bigcup_{i=1}^k C_i = V$ and $\forall i\neq j,\, C_i\cup C_j =
-\emptyset$.} For instance, $\cluster(u) = 3$ means that $u$ belongs
-to the third cluster. We can evaluate how \cluster{} fits our clustering paradigm in two ways,
+%TODO replace N by [|K|] (and define that notation in the table)
+
+Now consider a clustering \cluster{} of $V$, that is a function from $V$ to $\Nbb{}^{|V|}_{>0}$ that
+assigns to each node a cluster index. For instance, $\cluster(u) = 3$ means that $u$ belongs to the
+third cluster. We will also abuse the notation and let \cluster{} be a set of clusters $\{C_1, C_2,
+\ldots, C_K\}$ that form a partition of $V$.\footnote{That is, $\forall i\,C_i\subset V$,
+$\bigcup_{i=1}^K C_i = V$ and $\forall i\neq j,\, C_i\cap C_j = \emptyset$.}
+We can evaluate how \cluster{} fits our clustering paradigm in two ways,
 either by the number of \emph{agreements}, that is the weighted number of positive edges inside
@@
 some sense \enquote{easier}. Another interesting feature of the \pcc{} problem is that contrary to
-other clustering formulations, it does not require us to set the number of clusters $k$ beforehand.
-Instead, $k$ emerges as a natural property of the solution. Since clustering is an unsupervised
+other clustering formulations, it does not require us to set the number of clusters $K$ beforehand.
+Instead, $K$ emerges as a natural property of the solution. Since clustering is an unsupervised
 problem, this is generally handy. However, in some situations, we may have prior knowledge on how
 many clusters are the data, or external constraints. This can be handled with parametrized version
-of \pcc{}, namely \maxa$[k]$ and \mind$[k]$ where the optimization is over clustering with exactly
-$k$ clusters.
+of \pcc{}, namely \maxa$[K]$ and \mind$[K]$ where the optimization is over clustering with exactly
+$K$ clusters.
 
@@
 
-\subsubsection{Applications}
+\subsubsection{Connection with \esp{} and other applications}
 \label{ssub:cc_applications}
@@
       are presented in~\autocite[Section 6]{clusterEditSurvey13}.
+      % Finding the 3D shapes of chromosome given pairwise contact frequencies of different regions
+      % \url{https://doi.org/10.1145/3107411.3108216}
       % \enquote{They examined dynamical systems, where a gene is modeled as a vertex and an

### 2017-08-24 cc/cc_problem_and_esp.tex
move applications of CC to intro

--- 2017-08-24 cc/cc_problem_applications.tex
@@
 In addition to these general considerations, \pcc{} has also been used in several domains:
-%TODO turn those items into paragraphs?
-\begin{description}
-   \item[Computer Vision]
-      The ubiquitous task of segmenting an image into homogeneous regions is a prerequisite for many
-      further processing, and can be tackled with \pcc{}. For instance, to segment cell in
-      microscopy imagery, \textcite{CellSeg14} first use generic image features to classify pixel in
-      belonging to region boundaries or not. Then, they extract small scale regions called
-      superpixels. After building the adjacency graph of these superpixels, they assigned edge
-      weights by averaging the boundary probabilities of the pixels separating adjacent
-      superpixels. They also add strong negative constraints between distant superpixels, and lastly
-      cluster these superpixels according to the \pcc{} objective to obtain the final segmentation.
-      A similar approach was used earlier in~\textcite{Kim2011}, who stress the importance of
-      considering such higher order constraints between distant superpixels in order to achieve good
-      performance. This was also extended to 3D segmentation~\autocite{VolumeSegmentation12}, where
-      additional tuning allows to segment a volume image of a mouse cortex with up to billions
-      voxels. \Textcite{Beier2015} segment 2D and 3D images with an energy based formulation of
-      \pcc{} and iteratively improve their solution by merging it with another clustering given by a
-      proposal generator. By developing another scalable energy based optimization procedure, and
-      with the help of few user-provided cues, \textcite{Bagon2011} are able to apply \pcc{}
-      directly at the pixel level.
-      % one more? https://link.springer.com/chapter/10.1007/978-3-642-33783-3_41
-      % 3D mesh segment https://link.springer.com/article/10.1007/s41095-016-0071-3
-
-      Beyond image segmentation, \textcite{Shape3D17} develop a method to extract a network of
-      descriptive curves from 3D shapes.  After an initial stage of generating many such
-      \emph{flowlines}, they describe in Section 6 a \pcc{} formulation to extract \emph{reliable}
-      representative flowlines, using geometric constraints to obtain positive or negative cues that
-      two flowlines are from the same reliable representative.
-
-      Finally, in order to track several targets across sequential video frames,
-      \textcite{multiTracking15} propose a multistage framework. One step revolves around a matrix
-      $A$ that defines the cost of assigning an object tracked in previous frames to an object
-      detected in the current frame. This matrix is turned into a symmetric affinity matrix
-      $\bar{A}_{sym}$ that can be seen as a signed graph adjacency matrix. \pcc{} is then used to extract clusters
-      (called zones), in which local processing is performed. This is beneficial since the
-      complexity of these local methods can be adapted to the difficulty of each zone.
-
-   \item[Natural Language Processing] 
-      Coreference resolution is the task of finding all expressions that refer to the same entity in
-      a text. Like image segmentation, it is a preprocessing step that can later be used in document
-      summarization, question answering, and information extraction. \Textcite[Section
-      2.3]{graphicalCoreference04} tackle coreference resolution by using an undirected graphical
-      model on which performing inference is equivalent to \pcc{}. On small scale instances,
-      \textcite{Elsner2009} first obtain an upper bound of the optimal solution by solving a SDP
-      relaxation of the problem. They then compare various heuristics and show that best performing
-      ones are within few percents of the optimum, provided they are followed by a local search
-      step, such as the Best One Element Move~\autocite{Gionis2007}. Further NLP tasks amenable to
-      \pcc{} formulation are referenced in their paper. Another possible task would be 
-      to cluster words based on distributional embedding vector while adding antonym
-      constraints~\autocite{SignedWordRatings}, although the authors of that paper choose to use
-      signed spectral clustering.
-
-   \item[Biology]
-      Clustering tasks also abounds in Biology.
-      A typical input is a similarity matrix between genes expression level in various experimental conditions
-      and the goal is to cluster those genes into groups which react similarly. \Textcite[Section
-      4]{Ben-Dor99} gives three examples: 112 genes involved in the rat central nervous system, 1246
-      genes of the roundworm \emph{C. elegans} and 2000 human genes obtained from 40 tumor and 22
-      normal colon samples. \Textcite{Mason2009} analyze a signed co-expression network of genes
-      involved in embryonic stem cells to find which genes are related to pluripotency (the ability
-      to differentiate into any type of cell in the body) or self-renewal (the ability to replicate
-      indefinitely). Another application is to study the variation of one individual
-      DNA~\autocite{Das2015}. In the human organism, chromosomes are organized in pair, and both
-      chromosomes of a pair have similar but not identical DNA sequences. This is mostly because of
-      single nucleotide polymorphisms (SNPs), where a single base differs between the two DNA
-      sequences, leading to different alleles of the corresponding gene. An haplotype is the list of
-      all alleles at a contiguous region of a single chromosome, and this information is
-      used in several medical applications. The high-throughput sequencing of one individual genome
-      yields many short \emph{reads} that provides information about the order of nucleotides in a
-      fragment of one chromosome and that can be used to assemble haplotypes. To do so, the authors
-      build a graph of reads and define a similarity function between reads to assign weights on the
-      edges. The clusters of that graph correspond to haplotypes, and are obtained by a SDP
-      relaxation of the \pcc{} objective. \Textcite{monotoneBiology07} also consider graphs whose
-      nodes are genes, but in a different context. In this case, positive edges represent an
-      activating connection, while negative edges represent inhibiting connection. They also define
-      a \emph{monotone system} as a balanced subgraph, that is a subgraph which does not contain a
-      cycle with an odd number of edges. Such monotone system are stable, in the sense that
-      modifying the concentration of one gene will have a predictable effect, even ignoring the
-      precise kinematics of the chemical reactions involved. Their goal is to find the minimum
-      number of edges to remove in order to decompose a dynamics system into a collection of
-      monotone system. This corresponds to the \mind{} problem and allows to study
-      the complete system more easily. Many more applications of \mind{} on weighted complete graphs
-      are presented in~\autocite[Section 6]{clusterEditSurvey13}.
-      % Finding the 3D shapes of chromosome given pairwise contact frequencies of different regions
-      % \url{https://doi.org/10.1145/3107411.3108216}
-      % \enquote{They examined dynamical systems, where a gene is modeled as a vertex and an
-      % activating connection is modeled as a positive edge and an inhibiting connection is modeled as
-      % a negative edge. The claim is that biological dynamical systems are close to being balanced,
-      % and that finding a minimum set of edges to delete to make the graph balanced can be used to
-      % decompose the graph into “monotone subsystems”, which exhibit stable behavior and thus allow a
-      % better understanding of the dynamics of a system.}\autocite{monotoneBiology07}
-
-   \item[Network science]
-      As mentioned in the first Chapter, one early use of signed graphs was to model social
-      interactions and given the large body of work on community
-      detection~\autocite{FortunatoSurvey10}, it is therefore not surprising that \pcc{} has been
-      used to study such networks. For instance, one can extract all the votes of the members of a
-      political parliament and form a graph whose nodes are politicians and edge weights quantify how
-      much they agree or disagree on various issues they have been voting on. This can be used to
-      study various social science questions such as loyalty, leadership, coalitions, political
-      crisis and polarization. It has been applied to the European
-      parliament~\autocite{Mendonca2015}, Slovenian parliament~\autocite{Jiang2015} and the
-      Brazilian parliament~\autocites{BrazilCC17}. This can also be used at international level. For
-      instance, by considering a dataset of military alliances and disputes, \textcite{Traag2009}
-      cluster countries into blocks that resemble those identified by Huntington in his \emph{Clash
-      of Civilizations} book. Another source of data is the vote on resolutions during the United
-      Nations General Assembly~\autocite{CommunityUN12}. Finally, one can also study how to exploit
-      the information contained within the negative links to enhance the visualization of social
-      networks~\autocite{Luca10}. Note however that those three last papers not explicitly use a
-      \pcc{} formulation.
-
-   \item[Others]
-      \begin{itemize}[leftmargin=*]
-	 \item
-	    Deduplication, also called duplicate detection or entity resolution, is the process
-	    of identifying objects from a real-world, noisy database that refer to the same entity.
-	    On a high level, a solution to this problem is to build a graph of all the available
-	    objects, define a similarity between them and run a \pcc{} algorithm. This was indeed
-	    this kind of problems at Whizbang! Labs that partly motivated one of the early \pcc{}
-	    paper~\autocite{Bansal2002}. The main challenge
-	    thus lies in devising an appropriate similarity measure, given that object can have very
-	    different features from one database to another. \Textcite{LargeScaleDeDup09} propose a
-	    declarative language, expressing both hard constraints (that have to be satisfied) and
-	    soft constraints (that can be seen as cues guiding the process). Because of these hard
-	    constraints that admissible clusterings have to respect, the authors have to modify in
-	    nontrivial ways an existing \pcc{} algorithm. This was extended to weighted and partial
-	    constraints  by \textcite{WeightedER14}. Another example is given by
-	    \textcite{Crosslingual07}, who cluster together news articles in different languages
-	    covering the same event. \pcc{} was also evaluated among other
-	    solutions to that problem by \textcite{DeDup09}, who note that their non optimized
-	    implementation does not perform the best.
-	 \item 
-	    Given an electrical circuit layout, \textcite{circuitDesign07} extract a graph of its
-	    components (called shifter) that must be assigned one of two possible phases. Because
-	    two shifters next to some specific shape must be in opposite phase and two shifters
-	    separated by less than a specified distance must be of the same phase, the authors look
-	    for a two-clustering of the nodes that will minimize the number of disagreements.
-      \end{itemize}
-\end{description}

### 2017-08-24 intro/use_of_sign_graphs.tex
move applications of CC to intro

--- 2017-08-21 cc/problem_applications.tex
@@
-Like other clustering frameworks, in \pcc{}, we are given a set of objects and we want to gather
-them into groups (called clusters) so that objects belonging to one cluster are similar to each
-other while being dissimilar to objects from all the other clusters.
-
-In \pcc{}, we formalize this problem by considering objects as the nodes of a graph $G$, whose edge
-weights encode similarity. Namely, in the most general case, for nodes $u$ and $v$, the edge between
-$u$ and $v$ is associated with two positive numbers:
-$w_{u,v}^+$ denotes the strength of the similarity between $u$ and $v$;
-$w_{u,v}^-$ denotes the strength of the dissimilarity between $u$ and $v$.
-Note, however, that in many applications, only one of these two numbers is nonzero, in which case
-we more conveniently set $w_{u,v} = \begin{cases}
-	 w_{u,v}^+ & \quad \text{if } w_{u,v}^+ > 0 \text{ and } w_{u,v}^-=0 \\
-	-w_{u,v}^- & \quad \text{if } w_{u,v}^+ = 0 \text{ and } w_{u,v}^->0 \\
-\end{cases}$
-
-Now consider a clustering \cluster{} of $V$, that is a function from $V$ to $\Nbb{}^{|V|}_{>0}$
-that assigns to each node a cluster index.\marginpars{We will also abuse the notation and let
-\cluster{} be a set of clusters $\cluster=\{C_1, C_2, \ldots, C_k\}$ that form a partition of $V$.
-That is, $\forall i\,C_i\subset V$, $\bigcup_{i=1}^k C_i = V$ and $\forall i\neq j,\, C_i\cup C_j =
-\emptyset$.} For instance, $\cluster(u) = 3$ means that $u$ belongs
-to the third cluster. We can evaluate how \cluster{} fits our clustering paradigm in two ways,
-either by the number of \emph{agreements}, that is the weighted number of positive edges inside
-clusters plus the weighted number of negative edges across clusters; or by the number of
-\emph{disagreements}, that is the weighted number of negative edges inside clusters plus the
-weighted number of positive edges across clusters. Given a cost function $c$, which is usually the
-identity, \pcc{} can then be seen as a graph optimization problem, either of maximizing agreements
-(\maxa{}):
-\begin{equation}
-	\max_{\cluster{}} \sum_{\cluster(u) = \cluster(v)} c(w_{uv}^+) +
-	\sum_{\cluster(u) \neq \cluster(v)} c(w_{uv}^-)
-	\label{eq:maxa}
-\end{equation}
-or minimizing disagreements (\mind{}):
-\begin{equation}
-	\min_{\cluster{}} \sum_{\cluster(u) = \cluster(v)} c(w_{uv}^-) +
-	\sum_{\cluster(u) \neq \cluster(v)} c(w_{uv}^+)
-	\label{eq:mind}
-\end{equation}
-
-Although an optimal clustering $\cluster^\star$ achieves the same value on both \eqref{eq:maxa}
-and \eqref{eq:mind}, we will see in \autoref{sub:state_of_the_art} that the latter objective is in
-some sense \enquote{easier}. Another interesting feature of the \pcc{} problem is that contrary to
-other clustering formulations, it does not require us to set the number of clusters $k$ beforehand.
-Instead, $k$ emerges as a natural property of the solution. Since clustering is an unsupervised
-problem, this is generally handy. However, in some situations, we may have prior knowledge on how
-many clusters are the data, or external constraints. This can be handled with parametrized version
-of \pcc{}, namely \maxa$[k]$ and \mind$[k]$ where the optimization is over clustering with exactly
-$k$ clusters.
-
-In \autoref{fig:cc_objectives}, we show a simple instance of \pcc{} and one of its optimal solution.
-\begin{figure}[hbt]
-	\centering
-	\includegraphics[width=0.63\linewidth]{assets/tikz/cc_objectives_tikz.pdf}
-	\caption[Small example of \pcc{}]{A small graph with eight nodes and ten edges. Solid edges
-	represent positive edges and dashed edges represent negative edges. A clustering \cluster{}
-	is showed with 3 clusters: $\{1, 2, 4\}$, $\{3, 5, 6\}$ and $\{7, 8\}$. \cluster{} incurs two
-	disagreements: the negative edge between nodes $1$ and $2$ within the blue cluster, and the
-	positive edge $(6,7)$ between the orange and green clusters. Those disagreements are created
-	by two cycles with one negative edge and thus cannot be avoided, meaning that \cluster{} is
-	optimal. However, it is not the unique solution: for instance, merging the orange and green
-	clusters would also yield two disagreements.}
-	\label{fig:cc_objectives}
-\end{figure}
-
-\subsubsection{Applications}
-\label{ssub:cc_applications}
-
-According to \textcite[Section 5]{Demaine2006}, \pcc{} is well suited to several situations:
-\begin{itemize}
-   \item when the items to be clustered do not belong to a natural metric space (preventing
-      approaches such as $k$-means) but we still know for some pairs whether they are similar or
-      not.
-   \item when we do not know the number of clusters beforehand but we have a similarity measure. In
-      that case, we can select a problem-specific similarity threshold and set all edges with a
-      similarity larger than the threshold to be positive while the others are set to negative.
-   \item when we have a classic clustering problem (that is a set of objects, a distance between
-      them and an objective function to minimize) with additional pairwise constraints of the form
-      \emph{must-link/cannot-link}. Instead of restraining a clustering algorithm to the space of
-      feasible solution, we convert the distances between objects and the constraints into signed
-      edges and solve the resulting \pcc{} problem. 
-\end{itemize}
-
-In addition to these general considerations, \pcc{} has also been used in several domains:
 %TODO turn those items into paragraphs?

### 2017-08-24 cc/cc_problem_and_esp.tex
elaborate on the connection between CC and ESP

@@
 
-\subsubsection{Connection with \esp{} and other applications}
+\paragraph{Connection with \esp{} and other applications}
 \label{ssub:cc_applications}
 
-According to \textcite[Section 5]{Demaine2006}, \pcc{} is well suited to several situations:
+\esp{} can clearly be casted as a supervised classification problem where, given a labelled training
+set of $m_0$ edges $\mathcal{X}=\{(e_1, y_1), \ldots, (e_{m_0}, y_{m_0})\}$, we can extract for
+every edge a feature vector in $\Rbb^d$, pick an hypothesis class like linear models, find among
+this class the hypothesis minimizing the empirical risk, and use it to predict the sign of edges in
+the testing set. In contrast, \pcc{} in an agnostic problem. More precisely, the hypothesis class is
+fixed, as it consists of all possible partitions of the nodeset. We are trying to find the partition
+that best approximates the observed signs, but each weakly unbalanced cycle is gonna cost us at
+least one unavoidable mistakes. Despite this fundamental difference in their nature, \pcc{} is
+connected with \esp{} at two levels, theoretical and practical. On the theoretical side, computing
+the minimum number of disagreements gives a measure of the \esp{} problem complexity.  Intuitively,
+given a labelled subgraph $H$ of $G$ (that is the training set), if there are few disagreements, the
+\enquote{quasi} connected components of $H^+$ should give a reasonable estimation of the underlying
+consistent clustering. Predicting the signs based on this clustering should results in few errors.
+More formally, one can show that by finding a clustering $\cluster^\star$ of $H$ that has less
+disagreements than $\kappa$ times the smallest number of disagreements $\Delta$, the number of
+errors on the testing set will be at most of order $\kappa\Delta + \sqrt{mn\log
+n}$~\autocite[Theorem 6]{Cesa-Bianchi2012b}. These theoretical considerations naturally lead to
+practical batch algorithm, which first finds a clustering minimizing as much as possible the number
+of disagreements on the training set and then predicts signs according to this clustering. Our
+assumption in doing so is that, according to our bias, our graph has few disagreements. Therefore,
+provided we can find a clustering close to the optimal, we will not make too many prediction
+mistakes.
+
+\bigskip
+
+Besides the special case of predicting edge sign, \textcite[Section 5]{Demaine2006} point out that
+\pcc{} is well suited to several more general situations:
 \begin{itemize}
@@
 
-In addition to these general considerations, \pcc{} has also been used in several domains:
+Furthermore, as mentioned in the introduction of this thesis, \pcc{} has also been used in several
+domains. For instance in computer vision, it is used to segment images in
+2D~\autocites{Kim2011}{Bagon2011}{CellSeg14} or 3D~\autocites{VolumeSegmentation12}{Beier2015}, to
+simplify 3D shapes~\autocite{Shape3D17} and to track target across sequential video
+frames~\autocite{multiTracking15}. When the nodes are words, \pcc{} is employed to identify
+coreference~\autocites{graphicalCoreference04}{Elsner2009} or cluster synonym
+words~\autocite{SignedWordRatings}. It plays a large role in biology, to clusters
+genes~\autocite{Ben-Dor99}, identify mutation regions in chromosomes~\autocite{Das2015} or finding
+stable subsystems~\autocite{monotoneBiology07}. Applied to social networks, it helps analyze
+political assemblies~\autocites{Mendonca2015}{Jiang2015}{BrazilCC17} and international
+relations~\autocites{Traag2009}{CommunityUN12}. Finally, given large databases with duplicated
+records, it naturally models entity
+resolution~\autocites{Crosslingual07}{DeDup09}{LargeScaleDeDup09}{WeightedER14}.

### 2017-08-24 cc/cc_multimincut_stable.tex
reorganize files and write some intro
move text around for a state of the art following more the general direction of chapter

--- 2017-08-14 cc/multimincut_stable.tex
@@
-We first give more details on the equivalence between \mmc{} and \pcc{},
-following~\autocite{Demaine2006} but omitting the full proofs to avoid redundancy. Recall that given
-an weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and sinks
-$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, in \mmc{} we want to find the lightest set of edges $T\in E_H$
-whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
-polynomial reduction from \mmc{} to \pcc{}. Namely given an instance $(H, S)$ of \mmc{}, we let
-$W_H=\sum_{e\in E_H} w_H(e)$ be the total weight of $H$ and $G_H$ be the same graph as $H$ with all
-its edges labeled positively. We then connect every pair $(s_i, t_i)$ by a negative edge of weight
-$W_H+1$. One can check (\autocite[Theorem 4.7]{Demaine2006}) that a multicut $T$ in $H$ with weight
-$W$ induces a clustering in $G_H$ of weighted disagreement $W$ by the connected components of
-$G^+\setminus T$. Likewise a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
-weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges
-are simulated in the following way: include every source and sink in a clique of $n$ nodes and
-connect that clique to the corresponding source or sink by $n$ negative edges.
-
-We next present the polynomial reduction from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
-$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v)$ of
-weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
-weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-sink pair added to $S_G$. This
-construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
-4.4]{Demaine2006})\Todo{give the intuition of the proof} that it is easy to construct a multicut of
-weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.
-
-\begin{figure}[htpb]
-   \centering
-   \includegraphics[width=0.8\linewidth]{assets/raw/cc_to_mmc.png}
-   \caption{The transformation from \pcc{} on $G$ to \mmc{} on $H$ (temporarily reproduced from
-   \autocite{Demaine2006}) \label{fig:cc_mmc}}
-\end{figure}
-
 Whereas clustering objective functions are \NPh{} to optimize, we expect that meaningful instances
@@
 
+\iffalse
 \Textcite{StableCC17} provide a robust algorithm for $2-\nicefrac{2}{k}$-stable instance of
@@
 \end{figure}
+\fi
 

### 2017-08-24 cc/cc_state_of_art.tex
move text around for a state of the art following more the general direction of chapter

--- 2017-08-21 cc/proposed.tex
@@
-We already described methods based on LP and SDP, as they have proved to be the best way to obtain
-tight approximations in the worst case. Here we survey other kind of approaches with different aims.
+As we saw, solving \pcc{} on a training set would give us a principled way of predicting edge signs.
+In this section, we thus look at existing methods for \pcc{}. We start with exact methods and fixed
+parameter algorithms on complete graphs, noting they implicitly assume our learning bias. However,
+in practice, these methods do not scale well, especially if their assumptions are not met. Indeed,
+we next present hardness results about \pcc{}, namely that it is \NPh{} to approximate with some
+constant factor greater than one, and then describe some proposed approximations. Whereas this
+sounds like bad news for us, we show that heuristics can solve large instances with satisfactory
+results. We conclude by further highlighting the learning aspect of \pcc{}, describing approaches in
+the active and online settings.
 
-\paragraph{Exact methods} Because of the complexity of the \pcc{} we presented earlier, one has to
-rely on approximations to solve large instances of the problem. However, we can imagine offline
-signed social networks with only few nodes, in which case it is reasonable to expect finding the
-optimal clustering. Furthermore, this can also be useful to evaluate in practice the quality of
-heuristic methods, albeit in non asymptotic setting. The most straightforward idea is to use the
-formulation of the LP \eqref{eq:mindLP} with the additional constraints that all variable are
-binary. \Textcite{ExactMIP13} solve this binary integer program on random instances and show that
-depending on the negative edge density, the FICO Xpress solver starts to be unable to finish within
-a one-hour time limit when $n \geq 40$.
-\Textcite{Aref2016} describe four linear and quadratic  binary integer problems
-to model \pcc{} with two clusters along with some preprocessing optimizations. With the Gurobi
-software, they solve instances with up to $3200$ edges in less than a second. They also sketch
-extensions to weighted graphs and more than two clusters.
-In a different direction, \textcite{Berg2015}, encode the linear and quadratic integer formulations
-into weighted MaxSAT instances and use the state of the art solver
-MaxHS\footnote{\href{http://www.maxhs.org/}{http://maxhs.org}}~\autocite{SATSolver13} to get the
-exact solution on instances with at most $1000$ nodes in less than a few hours.
+\subsubsection{Exact methods}
+\label{ssub:cc_exact_methods}
 
-After defining the matrix $A$ by $A_{uv} = w_{uv}^+ - w_{uv}^-$, \textcite{LowRank16} show that the
-\maxa{} problem can be written up to constant factor by defining one vector $x_u$ per node as:
-\begin{align}
-   \label{eq:cc_lowrank}
-   \text{maximize } & \sum_{u<v} A_{uv}x_u^Tx_v \\
-   \text{subject to}& \quad x_{uw} \in \{e_1,\ldots, e_n\} \nonumber
-\end{align}
-where $e_i$ are the canonical vectors of $\Rbb{}^n$. They show that when the matrix $A$ is positive
-semidefinite of rank $k$, the \pcc{} problem can be solved exactly in $O(n^{k^2})$ time. More
-practically, they also give an algorithm that closely approximate the objective of
-\eqref{eq:cc_lowrank} in $O(nk)$ time.
-On complete graphs, we may also turn on fixed-parameter tractable algorithms, where the parameter
-can be the number of disagreements, the number of clusters or the maximum number of disagreements
-per nodes, as we will now see.
-
-\paragraph{\textsc{Cluster editing}}
+\paragraph{Complete graphs}
 \label{par:cc_editing}
@@
 into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}.
+
+See \url{http://fpt.wikidot.com/fpt-races} and
+\url{http://fpt.wikidot.com/complexity-status-of-edge-modification-problems} for parameterized
+papers on \mmc{} and \textsc{Cluster Editing}.
+
+\paragraph{General graphs}
+
+Because of the complexity of the \pcc{} we presented earlier, one has to
+rely on approximations to solve large instances of the problem. However, we can imagine offline
+signed social networks with only few nodes, in which case it is reasonable to expect finding the
+optimal clustering. Furthermore, this can also be useful to evaluate in practice the quality of
+heuristic methods, albeit in non asymptotic setting. The most straightforward idea is to use the
+formulation of the LP \eqref{eq:mindLP} with the additional constraints that all variable are
+binary.
+
+Assign a binary variable $x_{uv}$ to each pair
+of nodes (so that $x_{uv}=x_{vu}$). For a given clustering \cluster{}, let $x_{uv} = 0$ if $u$ and
+$v$ are in the same cluster and $x_{uv}=1$ is $u$ and $v$ are in different clusters. Noting that
+$1-x_{uv}$ is $1$ if the edge $(u,v)$ is within a cluster and $0$ otherwise, the weighted
+number of disagreements is then $w(\cluster{}) = \sum_{(u,v)\in E^-} w_{uv}(1-x_{uv}) +
+\sum_{(u,v)\in E^+} w_{uv}x_{uv}$. By construction, if edges $(u,v)$ and $(v,w)$ are within the same
+cluster, then $(v, w)$ is also within that cluster. In terms of $x$ variable, we have that $x_{uv}=0
+\wedge x_{vw}=0 \implies x_{uw} = 0$. For $x$ to be a valid cluster assignment, we thus require that all
+variables are either $0$ or $1$ and respect the triangle inequality. We can thus relax the problem into
+\begin{align}
+   \label{eq:mindLP}
+   \text{minimize } & \sum_{(u,v)\in E^-} w_{uv}(1-x_{uv}) + \sum_{(u,v)\in E^+} w_{uv}x_{uv} \\
+   \text{subject to}& \quad x_{uw} \leq x_{uv} + x_{vw} \nonumber\\
+   \phantom{subject to}& \quad 0 \leq x_{uv} \leq 1  \nonumber \\
+   \phantom{subject to}& \quad x_{uv} = x_{vu}  \nonumber
+\end{align}
+
+\Textcite{ExactMIP13} solve this binary integer program on random instances and show that
+depending on the negative edge density, the FICO Xpress solver starts to be unable to finish within
+a one-hour time limit when $n \geq 40$.
+\Textcite{Aref2016} describe four linear and quadratic  binary integer problems
+to model \pcc{} with two clusters along with some preprocessing optimizations. With the Gurobi
+software, they solve instances with up to $3200$ edges in less than a second. They also sketch
+extensions to weighted graphs and more than two clusters.
+In a different direction, \textcite{Berg2015}, encode the linear and quadratic integer formulations
+into weighted MaxSAT instances and use the state of the art solver
+MaxHS\footnote{\href{http://www.maxhs.org/}{http://maxhs.org}}~\autocite{SATSolver13} to get the
+exact solution on instances with at most $1000$ nodes in less than a few hours.
+
+After defining the matrix $A$ by $A_{uv} = w_{uv}^+ - w_{uv}^-$, \textcite{LowRank16} show that the
+\maxa{} problem can be written up to constant factor by defining one vector $x_u$ per node as:
+\begin{align}
+   \label{eq:cc_lowrank}
+   \text{maximize } & \sum_{u<v} A_{uv}x_u^Tx_v \\
+   \text{subject to}& \quad x_{uw} \in \{e_1,\ldots, e_n\} \nonumber
+\end{align}
+where $e_i$ are the canonical vectors of $\Rbb{}^n$. They show that when the matrix $A$ is positive
+semidefinite of rank $k$, the \pcc{} problem can be solved exactly in $O(n^{k^2})$ time. More
+practically, they also give an algorithm that closely approximate the objective of
+\eqref{eq:cc_lowrank} in $O(nk)$ time.
+On complete graphs, we may also turn on fixed-parameter tractable algorithms, where the parameter
+can be the number of disagreements, the number of clusters or the maximum number of disagreements
+per nodes, as we will now see.
+
+exact describe the LP here
+fpa
+
+basically the problem is tractable when the number of disagreements is small, agreeing with our
+bias. Likewise, exact methods works well when small rank, smart polytope in LP
+
+\subsubsection{Hardness and approximations}
+\label{ssub:cc_harness_approx}
+
+Although the same problem was studied earlier~\autocites{Early96}{Ben-Dor99}, \textcite{Bansal2002}
+coined the term \pcc{} and were the first to study this problem complexity. Namely, for complete,
+unweighted signed graphs, they show that both \mind{} and \maxa{} are \NPc{}. Along the way, they
+give a $17429$-approximation of \mind{} and a PTAS%
+\bottomfootnote{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
+minimization (respectively maximization) problem $\mathcal{P}$  in NP if given any $\epsilon>0$ and
+any instance $x$ of $\mathcal{P}$ of size $n$, $\mathcal{A}$ produces, in time polynomial in $n$, a
+solution that is within a factor $1+\epsilon$ (respectively $1-\epsilon$) of being optimal with
+respect to $x$. Note that the time is not necessarily polynomial in $\epsilon$, so that a running
+time of $O(n^{\frac{1}{\epsilon}})$ would qualify~\autocite[Definition 3.10]{CpxBook99}.}
+that, for any $\epsilon \in [0,1]$, runs in
+$O(n^2e^{O(\frac{1}{\epsilon^{10}}\log\frac{1}{\epsilon})})$ and returns with probability
+$1-\frac{\epsilon}{3}$ a solution with at most $\epsilon n^2$ fewer agreements than the optimal
+solution of \maxa{}.
+
+\begin{table}[bt]
+   \centering
+   \small
+   \caption{Hardness results of \pcc{}} \label{tab:cc_cpx}
+   \begin{tabulary}{187mm}{lCCCC}
+      \toprule
+               & \multicolumn{2}{c}{\mind{}}   & \multicolumn{2}{c}{\maxa{}}                                   \\
+      \cmidrule(r){2-3}
+      \cmidrule(r){4-5}
+      graph    & weighted                      & unweighted                                                     & weighted                                                     & unweighted                   \\
+      \midrule
+      Complete & \APXh{}~\autocite{Charikar2003} & \NPc{}~\autocite{Bansal2002}, \APXh{}~\autocite{Charikar2003}  &                                                              & \NPc{}~\autocite{Bansal2002} \\
+      General  & \APXh{}~\autocites{Charikar2003}{Demaine2003} &  \APXh{}~\autocites{Charikar2003}{Emanuel2003}   & \multicolumn{2}{c}{\APXh{}~\autocite[Thm. 9]{Charikar2003}} \\
+      \bottomrule
+   \end{tabulary}
+\end{table}
+
+The next year, several authors independently strengthened these results and extended them to
+weighted and general graphs, as summarized in \autoref{tab:cc_cpx}. In the most complete paper,
+\textcite{Charikar2003} show that on complete graphs, minimizing disagreements is \APXh{},
+\enquote{that is, is NP-hard to approximate within some constant factor greater than one} (it would
+be nice to provide some intuition why it's more difficult to minimize disagreements but according to
+the authors themselves, their reduction is \enquote{somewhat intricate}). They prove the same result
+on general graphs, for \mind{} by using a reduction from the multicut problem \autocite[Theorem
+8]{Charikar2003}\footnote{This reduction was actually turned into an equivalence, as we describe
+in more details in \autoref{ssub:cc_under_stability_assumption}.}
+
+We first give more details on the equivalence between \mmc{} and \pcc{},
+following~\autocite{Demaine2006} but omitting the full proofs to avoid redundancy. Recall that given
+a weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and targets
+$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, in \mmc{} we want to find the lightest set of edges $T\in E_H$
+whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
+polynomial reduction from \mmc{} to \pcc{}. Namely, given an instance $(H, S)$ of \mmc{}, we let
+$W_H=\sum_{e\in E_H} w_H(e)$ be the total weight of $H$ and $G_H$ be the same graph as $H$ with all
+its edges labeled positively. We then connect every pair $(s_i, t_i)$ by a negative edge of weight
+$W_H+1$. One can check (\autocite[Theorem 4.7]{Demaine2006}) that a multicut $T$ in $H$ with weight
+$W$ induces a clustering in $G_H$ of weighted disagreement $W$ by the connected components of
+$G^+\setminus T$. Likewise, a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
+weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges
+are simulated in the following way: include every source and target in a clique of $n$ nodes and
+connect that clique to the corresponding source or target by $n$ negative edges.
+
+We next present the polynomial reduction from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
+$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v)$ of
+weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
+weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-target pair added to $S_G$. This
+construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
+4.4]{Demaine2006})\Todo{give the intuition of the proof} that it takes a linear time to construct a
+multicut of weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.
+
+\begin{figure}[htpb]
+   \centering
+   \includegraphics[width=0.8\linewidth]{assets/raw/cc_to_mmc.png}
+   \caption{The transformation from \pcc{} on $G$ to \mmc{} on $H$ (temporarily reproduced from
+   \autocite{Demaine2006}) \label{fig:cc_mmc}}
+\end{figure}
+
+which asks for the minimum weight set of edges whose removal in $G$ disconnect the
+$k$ pairs $(s_i, t_i)$ and for \maxa{} by a reduction from MAX 3SAT~\autocite[Theorem
+9]{Charikar2003}. The multicut reduction yields a $O(\log n)$ approximation bound, which is achieved
+by rounding the canonical Linear Program~\eqref{eq:mindLP}.
+Once the LP \eqref{eq:mindLP} is solved, we interpret $x_{uv}$ as a distance: the larger it is and
+the more we want $u$ and $v$ to be in different clusters. We can then use the \regionGrow{}
+method~\autocite{RegionGrowing93}. Namely, we pick a ball center $u$ \uar{} with radius \shalf{}: if
+the average distance of the nodes in the ball to $u$ is less than $\nicefrac{1}{4}$, the ball forms a
+cluster, otherwise $\{u\}$ forms a singleton cluster. We then remove the corresponding nodes from
+the graph and repeat until all nodes are clustered. As we shall see, this method has been re used
+with variations on the LP or on the two thresholds \shalf{} and $\nicefrac{1}{4}$.
+
+However, \textcite[Theorem 2]{Charikar2003} note that the LP formulation has a poor integrality gap
+when it comes to \maxa{}, thus they turn to a Semi Definite Program. Say that each cluster is
+associated with a basis vector, then for each node $u$ in a cluster, we set $a_u$ to be the
+corresponding basis vector. If $u$ and $v$ are in the same cluster, we then have $a_u\cdot a_v = 1$
+while if they belong to different clusters, $a_u\cdot a_v = 0$. The weighted number of agreements
+can then be represented by
+\begin{align}
+   \label{eq:maxaSDP}
+   \text{maximize } & \sum_{(u,v)\in E^+} w_{uv}(a_u\cdot a_v) + \sum_{(u,v)\in E^-} w_{uv}(1-a_u\cdot a_v) \\
+   \text{subject to}& \quad a_u\cdot a_u=1 \nonumber\\
+   \phantom{subject to}& \quad a_u\cdot a_v\geq 0  \nonumber
+\end{align}
+After solving the SDP, a clustering can be obtained by a general rounding technique $H_t$: pick $t$
+random hyperplanes and divides the nodes in $2^t$ clusters. \Textcite[Theorem 3]{Charikar2003} prove
+that taking the best results of $H_2$ and $H_3$ gives in a $0.7664$ approximation on general graph.
+This was slightly improved to $0.7666$ by \textcite{Swamy2004} with a different rounding: pick $k$
+random unit vectors (called \emph{spokes}) and assign each $a_u$ to the closest spoke.
+
+Combining \mind{} and \maxa{}, \textcite[Section 4]{Charikar2004} give a $\Omega(\frac{1}{\log n})$
+approximation of the \textsc{MaxCorr} problem, which is maximizing \eqref{eq:maxa} - \eqref{eq:mind}
+and can be formulated as a quadratic programming problem solved in polynomial time.
+
+In complete graphs, \textcite[Section 3]{Charikar2003} also give an improved $4$-approximation to
+\mind{}, by rounding the same LP and using its solution in randomized algorithm. We will not
+describe it in detail since a similar idea was used by \textcite{CCPivotConf05} with a better
+approximation. To explain it, we first describe their randomized combinatorial algorithm \ccpivot{},
+which gives a $3$-approximation of \mind{} on complete unweighted graphs (it has later been
+derandomized while preserving its approximation guarantee~\autocite{derandomCCPivot08}). At each
+iteration, we pick a node $u$ \uar{} (called the pivot) and we create a cluster containing $u$ and all its neighbors
+linked by a positive edges. On weighted complete graphs, they tweak this algorithm by using the
+solution of the  LP~\eqref{eq:mindLP} to obtain different approximation factor depending on the
+constraints imposed on the weight.\Todo{Comment on those constraints, especially the triangular ones
+in the context of consensus clustering.} Recall that in the general formulation of the problem, each
+edge carries two positive numbers: $w_{u,v}^+$ and $w_{u,v}^-$. If the weights respect the
+probability constraints stating that for all edge $(u,v)$ in $E$, $w_{u,v}^+ + w_{u,v}^- = 1$, this
+tweaking provide a $2.5$-approximation. Note that unweighted graphs naturally fit into that case, as
+each edge is either labeled $+$ or $-$. If the weights
+additionally respect the triangular inequality constraints stating that $w_{u,v}^- \leq w_{u,w}^- +
+w_{w,v}^-$, this become a $2$-approximation. After solving the LP~\eqref{eq:mindLP} with additional
+probability constraints, when picking a node $u$, each of its neighbors $v$ is added to the cluster
+of $u$ with probability $x_{uv}$. \Textcite{Chawla2014} improve these two factors to respectively
+$2.06$ and $1.5$ by exploiting the same idea but setting the probability to include each neighbor
+$v$ of $u$ in the cluster of $u$ to be $1-f^+(x_{uv})$ if $(u,v)\in E^+$ and $1-f^-(x_{uv})$ if
+$(u,v)\in E^-$, with a careful choice of $f^+$ and $f^-$. They also give a derandomized version of
+their algorithm in the full version of the paper~\autocite[Theorem 23]{ChawlaArxiv14}.
+
+\begin{table}
+   \begin{tabulary}{187mm}{llcLL}
+      \toprule
+                                &            & $k$   & \mind{}                                                                & \maxa{}                                                     \\
+      \midrule
+      \multirow{2}{*}{Complete} & unweighted &       & $2.06$ \autocite{Chawla2014}                                           & PTAS from \textcite{Bansal2002}  \\
+      \cmidrule(r){4-4}
+                                & weighted   &       & $1.5$ (with triangular inequality) \autocite{Chawla2014}               &  and by setting $k=\Omega(1/\epsilon)$ \autocite{Giotis2006}  \\
+      \midrule
+   \multirow{3}{*}{General}     & unweighted &       & \multicolumn{2}{c}{No specific results for the unweighted case}                                                                       \\
+      \cmidrule(r){4-5}
+                                & weighted   & $k=2$ & $O(\sqrt{\log n})$ \autocite{Giotis2006}                               & $0.884$ \autocite{Mitra2009}                                  \\
+      \cmidrule(r){4-5}
+                                & weighted   &       & $O(\log n)$ \autocite{Charikar2003}, optimal under the UCG conjecture  & $0.7666$ \autocite{Swamy2004}                                 \\
+      \bottomrule
+   \end{tabulary}
+   \caption{Best results on various problem.\label{tab:cc_approx}}
+\end{table}
+
+This concludes the presentation of the known approximation results on \pcc{}, that we summarize in
+\autoref{tab:cc_approx}.
 
@@
 
-\paragraph{Heuristic methods}
+As mentioned earlier, not having to set the number of clusters is an
+attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.  The
+problem was studied by \textcite{Giotis2006} on general graphs and we compile their results in
+\autoref{tab:cc_fixed}. On complete unweighted graphs and with $k$ being the number of clusters, they
+provide PTASs for \maxa{}$[k]$ running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and for
+\mind{}$[k]$ running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
+\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}k^6\log d\right)}$
+and that can handle weighted graphs. For $k=2$, \mind$[2]$ admits a faster local search method with a
+factor $2$ approximation~\autocite{Coleman2008}.
+For complete weighted graph, \textcite{WeightedMaxAPTAS08} provide a PTAS for
+\maxa{}$[k]$ under the condition that the ratio between the largest and smallest weights is bounded by a
+constant.
+
+\begin{table}[htpb]
+   \centering
+   \caption{Approximation results for \pcc{} on general graph with $k$ clusters} \label{tab:cc_fixed}
+   \begin{tabulary}{187mm}{lLL}
+      \toprule
+      $k$	 & 2 & $\geq 3$ \\
+      \midrule
+      \maxa{} & 0.878 (improved to 0.884 by \autocite{Mitra2009}) & 0.7666 \autocite{Swamy2004} \\
+      \mind{} & $O(\sqrt{\log n})$ as it reduces to Min 2CNF Deletion for which \textcite{min2CNF05}
+      give such an approximation &
+      this can be reduced from $k$-coloring, which for any $\epsilon > 0$ is \NPc{} to approximate
+      within $n^{1-\epsilon}$ \autocite{InnaproxChroma07} \\
+      \bottomrule
+   \end{tabulary}
+\end{table}
+
+summary APX hard by equivalence with multicut (what about complete)? 3 strategies for approximation
+(in rewrote) -> needs for heuristics
+To summarize, there are currently three main approaches to approximate \mind{}. On a general graph,
+one first solve the canonical linear program, and use the \regionGrow{} method to round the
+fractional solution, resulting in an O(log n) approximation. On complete graphs, the simplest
+solution is the \ccpivot{} algorithm, that grows clusters out of randomly selected pivots, in a way
+reminiscent of the proof of thm. The approximation factor of this general idea can be improved by
+interpreting the fractional solution of the linear program as probability to be included in the
+pivot's cluster
+
+\subsubsection{Heuristics}
+\label{ssub:cc_heuristics}
+
+\paragraph{Greedy methods}
+\label{par:cc_greedy}
 
@@
 % dynamic model but to be honest it has little to do with clustering {Marvel2009landscape}
+% does it have something to do with phase transition
 % Kappes2016 \enquote{Furthermore, due to the lack of an external field (unary terms), any permutation of an
@@
 % vision https://arxiv.org/abs/1503.03791 represents a cut by the set of interclusters edges
+
+\subsubsection{Active and online settings}
+
+Except for the local algorithms of \textcite{Bonchi2013}, all works presented so far considered the
+batch case of \pcc{}, where the whole graph and all the signs are available at all time and with no
+cost. This might not always be the case, for instance if the graph is too large to fit in memory or
+if the edge labels are given by an expensive external procedure.
+
+% \emph{Still have to write this part. Is there a link between mistake bound and approximation
+% ratio?} say I have a binary classifier that perfectly predict sign wrt to the optimal clustering
+% C* those below are more concerned with sign prediction than clustering per se Beside batch
+% setting, one can also consider active \autocites{Cesa-Bianchi2012b}{Cesa-Bianchi2012a} and online
+% \autocite{Gentile2013} framework.
+% \enquote{the work of Cesa-Bianchi et al. [CBGV+12] that take a learning-theoretic perspective on the
+% problem of predicting signs. They consider three types of models: batch, online, and active
+% learning, and provide theoretical bounds for prediction mistakes for each setting. They use the
+% correlation clustering objective as their learning bias, and they show that the risk of the
+% empirical risk minimizer is controlled by the correlation clustering objective.}
+
+Case in point, in the context of entity resolution, \textcite{activeCoref07} consider the \pcc{}
+problem where, by querying an oracle, one can either reduce the uncertainty about one edge weight or
+add an extra node with edges connecting it to existing nodes. However, this requires web queries
+that are resource bounded and thus yields an active setting learning problem, where one has to
+choose the most informative
+queries given a budget. A formal definition is given in~\autocite{queryMoreEdgeCC07} and in
+practice, after finding an initially good partition, they select in each cluster a node to be its
+centroid, and query the edges connecting the centroid as ordered by an entropy-based criterion.
+In the case the oracle answering the queries is consistent, then there exists an information
+theoretic bound on the number of queries needed to recover the clusters, and an algorithm matching
+that bound up to a $O(\log n)$ factor~\autocite{perfectActiveLong17}.
+\textcite[Section 5]{Ailon2014} present another active algorithm that solves \mind$[k]$ with a
+linear number of queries but an exponential running time in $n$.  In a noisy setting,
+\textcite{Mitzenmacher2016} assume there is a planted $k$-partition and that we have access to an
+oracle that for $u,v\in V^2$ returns $\cluster(u) - \cluster(v) \mod k$ with probability $1-p$ and a
+noisy $\cluster(u) - \cluster(v) \pm 1 \mod k$ otherwise. For $k=2$, this is exactly $2$-\pcc{},
+whereas for $k>2$, the oracle gives more information than simply the sign of the path from $u$ to
+$v$. Whenever
+$p < \shalf$, they show that $O(n^{1+\frac{1}{\log\log n}}\log n)$ queries are enough are to recover
+the planted partition in polynomial time and with high probability. Those queries are actually
+random (\ie{} non-adaptive), and the clusters are found by looking at almost edge-disjoint path
+between all pairs of nodes.
+
+In the online setting, \textcite{greedyOnline10} give a greedy algorithm that upon node arrival creates a
+singleton cluster and then merge all pairs of clusters for which it increases the total number of
+agreements. For \mind{}, this is $O(n)$-competitive algorithm and they show such ratio is optimal by
+exhibiting an instance\footnote{Namely, two positive cliques $A$ and $B$ joined by positive edges except
+between $a\in A$ and $\{b_1,\ldots, b_k\}\in B$. Those nodes are given first and thus form a cluster,
+which yields at least one disagreement every time one the $n-(k+1)$ remaining nodes is added.} on
+which any strategy ends up with $n - k$ disagreements whereas optimal cost is $k$. On the \maxa{}
+side, this greedy strategy result in a $\shalf$-competitive algorithm. If it is randomly mixed with a
+\textsc{Dense} variation, it increases to $\shalf+\eta$, still far from the demonstrated $0.834$ upper
+bound.
+% when the graph is given node by node,
+% \url{https://link.springer.com/chapter/10.1007/978-3-319-18173-8_7} show to cluster it in static
+% clique, so it sounds related to cluster editing, except their objective is to maximize the
+% fraction of edge in cliques (ie avoid small cluster)
+
+\bigskip
+
+full conc of the section: the problem is difficult. However, it can be solved with guarentees in
+theory (because LP are not practical) and in efficiently in practice (using pivot as initilizatin
+for instance). Furthermore, instances obeying our bias can be solved much better. We next present
+settings where we are close to that bias

### 2017-08-24 cc/cc_variants.tex
move text around for a state of the art following more the general direction of chapter

@@
 
-\paragraph{Online \& active setting}
-
-Except for the local algorithms of \textcite{Bonchi2013}, all works presented so far considered the
-batch case of \pcc{}, where the whole graph and all the signs are available at all time and with no
-cost. This might not always be the case, for instance if the graph is too large to fit in memory or
-if the edge labels are given by an expensive external procedure.
-
-% \emph{Still have to write this part. Is there a link between mistake bound and approximation
-% ratio?} say I have a binary classifier that perfectly predict sign wrt to the optimal clustering
-% C* those below are more concerned with sign prediction than clustering per se Beside batch
-% setting, one can also consider active \autocites{Cesa-Bianchi2012b}{Cesa-Bianchi2012a} and online
-% \autocite{Gentile2013} framework.
-% \enquote{the work of Cesa-Bianchi et al. [CBGV+12] that take a learning-theoretic perspective on the
-% problem of predicting signs. They consider three types of models: batch, online, and active
-% learning, and provide theoretical bounds for prediction mistakes for each setting. They use the
-% correlation clustering objective as their learning bias, and they show that the risk of the
-% empirical risk minimizer is controlled by the correlation clustering objective.}
-
-Case in point, in the context of entity resolution, \textcite{activeCoref07} consider the \pcc{}
-problem where, by querying an oracle, one can either reduce the uncertainty about one edge weight or
-add an extra node with edges connecting it to existing nodes. However, this requires web queries
-that are resource bounded and thus yields an active setting learning problem, where one has to
-choose the most informative
-queries given a budget. A formal definition is given in~\autocite{queryMoreEdgeCC07} and in
-practice, after finding an initially good partition, they select in each cluster a node to be its
-centroid, and query the edges connecting the centroid as ordered by an entropy-based criterion.
-In the case the oracle answering the queries is consistent, then there exists an information
-theoretic bound on the number of queries needed to recover the clusters, and an algorithm matching
-that bound up to a $O(\log n)$ factor~\autocite{perfectActiveLong17}.
-\textcite[Section 5]{Ailon2014} present another active algorithm that solves \mind$[k]$ with a
-linear number of queries but an exponential running time in $n$.  In a noisy setting,
-\textcite{Mitzenmacher2016} assume there is a planted $k$-partition and that we have access to an
-oracle that for $u,v\in V^2$ returns $\cluster(u) - \cluster(v) \mod k$ with probability $1-p$ and a
-noisy $\cluster(u) - \cluster(v) \pm 1 \mod k$ otherwise. For $k=2$, this is exactly $2$-\pcc{},
-whereas for $k>2$, the oracle gives more information than simply the sign of the path from $u$ to
-$v$. Whenever
-$p < \shalf$, they show that $O(n^{1+\frac{1}{\log\log n}}\log n)$ queries are enough are to recover
-the planted partition in polynomial time and with high probability. Those queries are actually
-random (\ie{} non-adaptive), and the clusters are found by looking at almost edge-disjoint path
-between all pairs of nodes.
-
-In the online setting, \textcite{greedyOnline10} give a greedy algorithm that upon node arrival creates a
-singleton cluster and then merge all pairs of clusters for which it increases the total number of
-agreements. For \mind{}, this is $O(n)$-competitive algorithm and they show such ratio is optimal by
-exhibiting an instance\footnote{Namely, two positive cliques $A$ and $B$ joined by positive edges except
-between $a\in A$ and $\{b_1,\ldots, b_k\}\in B$. Those nodes are given first and thus form a cluster,
-which yields at least one disagreement every time one the $n-(k+1)$ remaining nodes is added.} on
-which any strategy ends up with $n - k$ disagreements whereas optimal cost is $k$. On the \maxa{}
-side, this greedy strategy result in a $\shalf$-competitive algorithm. If it is randomly mixed with a
-\textsc{Dense} variation, it increases to $\shalf+\eta$, still far from the demonstrated $0.834$ upper
-bound.
-% when the graph is given node by node,
-% \url{https://link.springer.com/chapter/10.1007/978-3-319-18173-8_7} show to cluster it in static
-% clique, so it sounds related to cluster editing, except their objective is to maximize the
-% fraction of edge in cliques (ie avoid small cluster)
-
 \paragraph{\pcc{} under noise}

### 2017-08-25 cc/cc_main.tex
move applications of CC to intro
move text around for a state of the art following more the general direction of chapter
reorganize CC state of the art

@@
-% *TODO*
-% say more clearly there are 3 methods for approximation so far: LP & region growing; LP & distance
-% pivot; combinatorial pivot
-% say that in practice, general graphs are assumed to be complete with missing edges being negative
-% - “The problem is agnostic in the sense that the clustering of the data is not taken into account
-%   or even assumed to exist.” another way to look at it is that we're trying to get a binary
-%   classifier of edges (+ or -) using an hypothesis class (nodes clustering) with limited
-%   expression power (as it has to respect transitivity constraints, whereas signs can be inconsistent)
-% - when the graph is weighted, there is no much point separating complete from
-%   general since weights can be 0
-
 As mentioned earlier, most real signed graphs are not fully balanced and thus do not have a
@@
 classic in unsupervised graph cut problems). We also devote a large section to approximation
-methods, in order to get a sense of why the problem is \NPc{}. This allows us to finish in
+methods, in order to get a sense of why the problem is \NPc{} and even hard to approximate. This
+allows us to finish in
 \autoref{sub:cc_non_worst_case} by describing situations related to our bias where is possible to
-solve the problem efficiently and optimally. By doing so, we update existing surveys on the same
-topic~\autocites{bonchi2014correlation}{surveyCC16}{CCWirth2017}, highlighting the importance of
+solve the problem efficiently and optimally. In this section, we thus update existing surveys on the same
+topic~\autocites{bonchi2014correlation}{surveyCC16}{CCWirth2017} and highlight the importance of
 \pcc{} to solve our binary learning problem.
@@
 \label{sub:problem_setting}
-\input{cc_problem_applications}
+\input{cc_problem_and_esp}
 
@@
 
-\subsubsection{Hardness and approximation}
-\label{ssub:cc_harness_approx}
-\input{cc_approx}
-\subsubsection{Proposed approaches}
-\label{ssub:cc_methods}
-\input{cc_proposed}
+\input{cc_state_of_art}
 

### 2017-08-25 cc/cc_state_of_art.tex
reorganize CC state of the art

@@
 sounds like bad news for us, we show that heuristics can solve large instances with satisfactory
-results. We conclude by further highlighting the learning aspect of \pcc{}, describing approaches in
+results. We conclude by further highlighting the learning aspect of \pcc{} through approaches in
 the active and online settings.
@@
 \label{ssub:cc_exact_methods}
+
+Because of the complexity of the \pcc{}, one has to rely on approximations to solve large instances
+of the problem. However, we can imagine offline signed social networks with only few nodes, in which
+case it is reasonable to expect finding the optimal clustering. Furthermore, this can also be useful
+to evaluate in practice the quality of heuristic methods, albeit in non asymptotic setting. The
+methods used to solve \pcc{} exactly depends on the type of graph considered. More precisely, the
+fixed parameter algorithms we describe now only handle complete graphs, while the next set of
+methods also work on general graphs.
 
@@
 
-In the cluster editing problem, one is given a general unsigned input graph $H=(V,E)$ and wants to
+Let us first describe the cluster editing problem, where one is given a general unsigned input graph
+$H=(V,E)$ and wants to
 find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
-union of cliques. This is equivalent to \pcc{} on complete graphs. To see why, let $G=(V,(E^+,E^-))$
+union of cliques. It is possible to navigate between an instance $H$ of cluster editing and an
+instance $G$ of \pcc{} on complete graphs in polynomial time. To see why, let $G=(V,(E^+,E^-))$
 be a complete signed graph with nodeset $V$, $E^+$ be the set $E$ of edges of $H$ and $E^-$ be all
@@
 whereas we shall only give important and more recent pointers here.} They show the problem is
-\NPc{}, even if the number of clusters $k\geq2$ is set beforehand\footnote{The reduction is from
+\NPc{}, even if the number of clusters $K\geq2$ is set beforehand\footnote{The reduction is from
 3-exact 3 cover, see \autocite[Theorems 1, 2 and Corollary 1]{Shamir02}.} and provide a
-$0.878$-approximation in the weighted $k=2$ case using the standard Goemans-Williamson SDP
-relaxation.
+$0.878$-approximation in the weighted $K=2$ case using the celebrated Goemans-Williamson SDP
+relaxation~\autocite{MaxCutSDP95}.
 
@@
 in the sign language) and branch by either deleting or merging one positive edge. If we additionally
-look for exactly $k$ clusters (\ie \mind$[k]$), there is a fixed parameters algorithm running in
+look for exactly $K$ clusters (\ie \mind$[K]$), there is a fixed parameters algorithm running in
 $O(2^{O(\sqrt{kd})}+n+m)$~\autocite{Fomin2014}. If at most $a$ edges can be added and at most $b$
 edges can be deleted at each node, and if the minimum size of a cluster is at least $2(a+b)$, then
-the problem can be solved in polynomial time~\autocite{Abu-Khzam2015}. Finally, for planar graph,
+the problem can be solved in polynomial time~\autocite{Abu-Khzam2015} by using various reduction
+rules. Finally, for planar graph,
 there is a PTAS running in $O(n2^{\epsilon^{-1}\log(\epsilon^{-1})})$ obtained by dividing the graph
-into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}.
+into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}. In other words, if we
+are given a complete signed graph whose positive subgraph is planar, then we can minimize the number
+of disagreements arbitrarily close to the optimal, provided we have enough time to do so.
 
-See \url{http://fpt.wikidot.com/fpt-races} and
-\url{http://fpt.wikidot.com/complexity-status-of-edge-modification-problems} for parameterized
-papers on \mmc{} and \textsc{Cluster Editing}.
+% See \url{http://fpt.wikidot.com/fpt-races} and
+% \url{http://fpt.wikidot.com/complexity-status-of-edge-modification-problems} for parameterized
+% papers on \mmc{} and \textsc{Cluster Editing}.
 
@@
 
-Because of the complexity of the \pcc{} we presented earlier, one has to
-rely on approximations to solve large instances of the problem. However, we can imagine offline
-signed social networks with only few nodes, in which case it is reasonable to expect finding the
-optimal clustering. Furthermore, this can also be useful to evaluate in practice the quality of
-heuristic methods, albeit in non asymptotic setting. The most straightforward idea is to use the
-formulation of the LP \eqref{eq:mindLP} with the additional constraints that all variable are
-binary.
-
+On general graphs, a natural idea is to formulate the clustering problem as an integer problem and
+solve it optimally in polynomial time.
 Assign a binary variable $x_{uv}$ to each pair
@@
 \wedge x_{vw}=0 \implies x_{uw} = 0$. For $x$ to be a valid cluster assignment, we thus require that all
-variables are either $0$ or $1$ and respect the triangle inequality. We can thus relax the problem into
+variables are either $0$ or $1$ and respect the triangle inequality.
 \begin{align}
-   \label{eq:mindLP}
+   \label{eq:mindIP}
    \text{minimize } & \sum_{(u,v)\in E^-} w_{uv}(1-x_{uv}) + \sum_{(u,v)\in E^+} w_{uv}x_{uv} \\
    \text{subject to}& \quad x_{uw} \leq x_{uv} + x_{vw} \nonumber\\
-   \phantom{subject to}& \quad 0 \leq x_{uv} \leq 1  \nonumber \\
+   \phantom{subject to}& \quad \leq x_{uv} \in \{0, 1\} \label{eq:mindInteger} \\
    \phantom{subject to}& \quad x_{uv} = x_{vu}  \nonumber
 \end{align}
-
 \Textcite{ExactMIP13} solve this binary integer program on random instances and show that
@@
 
+There is also a fixed parameter algorithm to solve \mind{} for general unweighted graphs. Indeed, we
+shall see later than the problem is equivalent to \mmc{}, which in this case asks for the minimal
+number $d$ of edges to remove in order to separate every pair of nodes in $S=\{(s_i,t_i)\}_{i=1}^k
+\in V^2$. It turns out that \mmc{} can be solved exactly in time $2^{O(d^3)}\cdot
+n^{O(1)}$~\autocite{FPAMulticut14}.
+
 After defining the matrix $A$ by $A_{uv} = w_{uv}^+ - w_{uv}^-$, \textcite{LowRank16} show that the
@@
 \eqref{eq:cc_lowrank} in $O(nk)$ time.
-On complete graphs, we may also turn on fixed-parameter tractable algorithms, where the parameter
-can be the number of disagreements, the number of clusters or the maximum number of disagreements
-per nodes, as we will now see.
 
-exact describe the LP here
-fpa
+\bigskip
 
-basically the problem is tractable when the number of disagreements is small, agreeing with our
-bias. Likewise, exact methods works well when small rank, smart polytope in LP
+Reflecting on these methods, we can conclude that \pcc{} is solvable in (near) polynomial time
+when the problem instance follow our learning bias. Indeed, fixed parameter algorithms are efficient
+when the number of disagreements is low, which corresponds to a quasi consistent clustering.
+Likewise, the small rank assumption of the adjacency matrix is equivalent to assuming a consistent
+clustering~\autocite[Theorem 13]{LowRankCompletion14}. Even when solving integer programs, one can
+use the bias to add inequalities and thus tighten the polytope of admissible solutions. On the other
+hand, we will now see that in the worst case, for example if an instance does not obey our bias,
+then there is little hope to obtain an optimal solution.
 
@@
 
-Although the same problem was studied earlier~\autocites{Early96}{Ben-Dor99}, \textcite{Bansal2002}
-coined the term \pcc{} and were the first to study this problem complexity. Namely, for complete,
+Although the same problem was considered earlier~\autocites{Early96}{Ben-Dor99}, \textcite{Bansal2002}
+coin the term \pcc{} and are the first to study this problem complexity. Namely, for complete,
 unweighted signed graphs, they show that both \mind{} and \maxa{} are \NPc{}. Along the way, they
 give a $17429$-approximation of \mind{} and a PTAS%
-\bottomfootnote{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
+\marginpars{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
 minimization (respectively maximization) problem $\mathcal{P}$  in NP if given any $\epsilon>0$ and
@@
 that, for any $\epsilon \in [0,1]$, runs in
-$O(n^2e^{O(\frac{1}{\epsilon^{10}}\log\frac{1}{\epsilon})})$ and returns with probability
-$1-\frac{\epsilon}{3}$ a solution with at most $\epsilon n^2$ fewer agreements than the optimal
+$O(n^2e^{O(\epsilon^{-10}\log \epsilon^{-1})})$ and returns with probability
+$1-\nicefrac{\epsilon}{3}$ a solution with at most $\epsilon n^2$ fewer agreements than the optimal
 solution of \maxa{}.
 
-\begin{table}[bt]
+The next year, several authors independently strengthened these results and extended them to
+weighted and general graphs, as summarized in \autoref{tab:cc_cpx}. In the most complete paper,
+\textcite{Charikar2003} show that on complete graphs, in sharp contrast with \maxa{} that admits a
+PTAS, \mind{} is \APXh{}, \enquote{that is, is NP-hard to approximate within some constant factor
+greater than one}. Unfortunately, it is not easy to give intuition why, for this follows from a
+\enquote{somewhat intricate reduction from max 2-colorable subgraph problem on bounded degree
+3-uniform hypergraphs}. On general graphs though, both versions of the optimization problem are
+\APXh{}. It is showed for \maxa{} by a reduction from MAX 3SAT~\autocite[Theorem 9]{Charikar2003}
+and for \mind{} by using a reduction from the \mmc{} problem~\autocite[Theorem 8]{Charikar2003}. In
+the case of \mind{}, \textcite{Emanuel2003} show a reduction in the other direction. Because \mmc{}
+is an interesting graph cut problem, and its complexity has been well studied, we here give more
+details on its equivalence with \pcc{}.
+
+For that we follow~\autocite{Demaine2006} but omit the full proofs to avoid redundancy. Recall that
+given a weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and targets
+$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, in \mmc{} we want to find the lightest set of edges $T\in E_H$
+whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
+polynomial transformation from \mmc{} to \pcc{}. Namely, given an instance $(H, S)$ of \mmc{}, we let
+$W_H=\sum_{e\in E_H} w_H(e)$ be the total weight of $H$ and $G_H$ be the same graph as $H$ with all
+its edges labeled positively. We then connect every pair $(s_i, t_i)$ by a negative edge of weight
+$W_H+1$. One can check (\autocite[Theorem 4.7]{Demaine2006}) that a multicut $T$ in $H$ with weight
+$W$ induces a clustering in $G_H$ of weighted disagreement $W$ by the connected components of
+$G^+\setminus T$. Likewise, a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
+weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges
+are simulated in the following way: include every source and target in a clique of $n$ nodes and
+connect that clique to the corresponding source or target by $n$ negative edges.
+
+\begin{table}[b]
    \centering
@@
 
-The next year, several authors independently strengthened these results and extended them to
-weighted and general graphs, as summarized in \autoref{tab:cc_cpx}. In the most complete paper,
-\textcite{Charikar2003} show that on complete graphs, minimizing disagreements is \APXh{},
-\enquote{that is, is NP-hard to approximate within some constant factor greater than one} (it would
-be nice to provide some intuition why it's more difficult to minimize disagreements but according to
-the authors themselves, their reduction is \enquote{somewhat intricate}). They prove the same result
-on general graphs, for \mind{} by using a reduction from the multicut problem \autocite[Theorem
-8]{Charikar2003}\footnote{This reduction was actually turned into an equivalence, as we describe
-in more details in \autoref{ssub:cc_under_stability_assumption}.}
-
-We first give more details on the equivalence between \mmc{} and \pcc{},
-following~\autocite{Demaine2006} but omitting the full proofs to avoid redundancy. Recall that given
-a weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and targets
-$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, in \mmc{} we want to find the lightest set of edges $T\in E_H$
-whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
-polynomial reduction from \mmc{} to \pcc{}. Namely, given an instance $(H, S)$ of \mmc{}, we let
-$W_H=\sum_{e\in E_H} w_H(e)$ be the total weight of $H$ and $G_H$ be the same graph as $H$ with all
-its edges labeled positively. We then connect every pair $(s_i, t_i)$ by a negative edge of weight
-$W_H+1$. One can check (\autocite[Theorem 4.7]{Demaine2006}) that a multicut $T$ in $H$ with weight
-$W$ induces a clustering in $G_H$ of weighted disagreement $W$ by the connected components of
-$G^+\setminus T$. Likewise, a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
-weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges
-are simulated in the following way: include every source and target in a clique of $n$ nodes and
-connect that clique to the corresponding source or target by $n$ negative edges.
-
-We next present the polynomial reduction from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
+We next present the polynomial transformation from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
 $H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v)$ of
@@
 
-which asks for the minimum weight set of edges whose removal in $G$ disconnect the
-$k$ pairs $(s_i, t_i)$ and for \maxa{} by a reduction from MAX 3SAT~\autocite[Theorem
-9]{Charikar2003}. The multicut reduction yields a $O(\log n)$ approximation bound, which is achieved
-by rounding the canonical Linear Program~\eqref{eq:mindLP}.
-Once the LP \eqref{eq:mindLP} is solved, we interpret $x_{uv}$ as a distance: the larger it is and
-the more we want $u$ and $v$ to be in different clusters. We can then use the \regionGrow{}
-method~\autocite{RegionGrowing93}. Namely, we pick a ball center $u$ \uar{} with radius \shalf{}: if
-the average distance of the nodes in the ball to $u$ is less than $\nicefrac{1}{4}$, the ball forms a
-cluster, otherwise $\{u\}$ forms a singleton cluster. We then remove the corresponding nodes from
-the graph and repeat until all nodes are clustered. As we shall see, this method has been re used
-with variations on the LP or on the two thresholds \shalf{} and $\nicefrac{1}{4}$.
+If a certain conjecture in computational complexity is true\footnote{Namely, the Unique Game
+Conjecture of~\textcite{UGC02}.}, then for every constant $c>0$ it is \NPh{} to approximate \mmc{}
+within a factor $c$~\autocite{}. On general graphs, the best known approximation factor in $O(\log
+k)$~\autocite{RegionGrowing93}, which according to the presented reduction translates to
+$O(\log|E^-|)=O(\log n^2)=O(\log n)$. This approximation ratio is achieved by first solving the
+relaxation of the integer problem \eqref{eq:mindIP}, where the constraint \eqref{eq:mindInteger} is
+replaced by $x_{uv}\in [0,1]$. From now on, we refer to that relaxation as the \emph{canonical
+\mind{} LP}. Once this canonical \mind{} LP is solved, we interpret $x_{uv}$ as a distance: the
+larger it is and the more we want $u$ and $v$ to be in different clusters. We can then use the
+\regionGrow{} method~\autocite{RegionGrowing93}. The idea is to pick a random center $u$ and to add
+to $u$'s cluster all the nodes at distance less than $r$ from $u$ before removing that cluster and
+repeating the process. $r<\shalf$ is chosen adaptively such that the weight of positive edges
+leaving the cluster is less than $c\log(n+1)$ times the fractional weights of the positive edges
+inside the cluster.\Todo{The full proof can be found in pages 7--9 of
+\href{http://erikdemaine.org/papers/Clustering_TCS/paper.pdf}{this online pdf} but in my opinion, it
+doesn't have much to do with machine learning.}
 
@@
 In complete graphs, \textcite[Section 3]{Charikar2003} also give an improved $4$-approximation to
-\mind{}, by rounding the same LP and using its solution in randomized algorithm. We will not
-describe it in detail since a similar idea was used by \textcite{CCPivotConf05} with a better
-approximation. To explain it, we first describe their randomized combinatorial algorithm \ccpivot{},
-which gives a $3$-approximation of \mind{} on complete unweighted graphs (it has later been
-derandomized while preserving its approximation guarantee~\autocite{derandomCCPivot08}). At each
-iteration, we pick a node $u$ \uar{} (called the pivot) and we create a cluster containing $u$ and all its neighbors
+\mind{}, by rounding the same LP and using a simpler version of \regionGrow{} with fixed radius.
+Namely, we pick a ball center $u$ \uar{} with radius \shalf{}: if the average distance of the nodes
+in the ball to $u$ is less than $\nicefrac{1}{4}$, the ball forms a cluster, otherwise $\{u\}$ forms
+a singleton cluster. We then remove the corresponding nodes from the graph and repeat until all
+nodes are clustered.
+
+Not long after, \textcite{CCPivotConf05} came up with a better approximation. To explain it, we
+first describe their randomized combinatorial algorithm \ccpivot{}, which gives a $3$-approximation
+of \mind{} on complete unweighted graphs (and was later derandomized while preserving its
+approximation guarantee~\autocite{derandomCCPivot08}). At each iteration, we pick a node $u$ \uar{}
+(called the pivot) and we create a cluster containing $u$ and all its neighbors
 linked by a positive edges. On weighted complete graphs, they tweak this algorithm by using the
-solution of the  LP~\eqref{eq:mindLP} to obtain different approximation factor depending on the
+solution of the canonical LP to obtain different approximation factor depending on the
 constraints imposed on the weight.\Todo{Comment on those constraints, especially the triangular ones
@@
 additionally respect the triangular inequality constraints stating that $w_{u,v}^- \leq w_{u,w}^- +
-w_{w,v}^-$, this become a $2$-approximation. After solving the LP~\eqref{eq:mindLP} with additional
+w_{w,v}^-$, this become a $2$-approximation. After solving the canonical LP with additional
 probability constraints, when picking a node $u$, each of its neighbors $v$ is added to the cluster
@@
 \begin{table}
-   \begin{tabulary}{187mm}{llcLL}
+   \caption{Best current results on \pcc{} problems. The \enquote{easiest} setting is \maxa{} on
+      complete graphs, for it admits PTASs. All others cases are \APXh{}. However, we see that on
+      the diagonal (that is \mind{} on complete graphs and \maxa{} on general graphs), there exists
+      constant factor approximations. This is not the case for the most \enquote{difficult} problem,
+      \mind{} on general graphs. \label{tab:cc_approx}}
+   \begin{tabulary}{\textwidth}{lLL}
       \toprule
-                                &            & $k$   & \mind{}                                                                & \maxa{}                                                     \\
+      $G$      &  \mind{}                                                  & \maxa{}                                 \\
       \midrule
-      \multirow{2}{*}{Complete} & unweighted &       & $2.06$ \autocite{Chawla2014}                                           & PTAS from \textcite{Bansal2002}  \\
-      \cmidrule(r){4-4}
-                                & weighted   &       & $1.5$ (with triangular inequality) \autocite{Chawla2014}               &  and by setting $k=\Omega(1/\epsilon)$ \autocite{Giotis2006}  \\
-      \midrule
-   \multirow{3}{*}{General}     & unweighted &       & \multicolumn{2}{c}{No specific results for the unweighted case}                                                                       \\
-      \cmidrule(r){4-5}
-                                & weighted   & $k=2$ & $O(\sqrt{\log n})$ \autocite{Giotis2006}                               & $0.884$ \autocite{Mitra2009}                                  \\
-      \cmidrule(r){4-5}
-                                & weighted   &       & $O(\log n)$ \autocite{Charikar2003}, optimal under the UCG conjecture  & $0.7666$ \autocite{Swamy2004}                                 \\
+      Complete &  $2.06$ (and $1.5$ if $G$ is weighted and weights obey the triangular inequality) \autocite{Chawla2014}
+               & PTAS from \textcite{Bansal2002} and by setting $k=\Omega(1/\epsilon)$ in \autocite{Giotis2006}      \\
+      General  &  $O(\log n)$ \autocite{Charikar2003}                      & $0.7666$ \autocite{Swamy2004}           \\
       \bottomrule
    \end{tabulary}
-   \caption{Best results on various problem.\label{tab:cc_approx}}
 \end{table}
 
-This concludes the presentation of the known approximation results on \pcc{}, that we summarize in
+\medskip
+
+This concludes the presentation of the known approximation results on \pcc{}, that we gather in
 \autoref{tab:cc_approx}.
+To summarize, except for maximizing the number of agreements on a complete graphs, we cannot hope
+for exact solutions for the other problems in the worst case, since they are \APXh{}. Moreover, even
+constant factor approximation are ruled out for \mind{} on general graphs, as showed by the
+equivalence with \mmc{}.
+Currently there are three main approaches to approximate \mind{}. On a general graph,
+one first solve the canonical LP, and use the \regionGrow{} method to round the fractional solution,
+resulting in an $O(log n)$ approximation. On complete graphs, the simplest solution is the
+\ccpivot{} algorithm, that grows clusters out of randomly selected pivots, in a way reminiscent of
+the proof of \autoref{thm:weakbal-consistent}. The approximation factor of this general idea can be
+improved by interpreting the fractional solution of the linear program as probability to be included
+in the pivot's cluster. The cost of solving linear programs with so many variables and constraints
+being prohibitive, real world general graphs requires to move from methods with guarantees to more
+heuristic approaches we describe next. Before that, we make two remarks. The first one regards a
+line of work improving scalability of these approximations by presenting distributed solutions. The
+second one is concerned with the case where the number of clusters is set beforehand, and proves
+that the problem becomes easier.
 
-\paragraph{Distributed setting}
-
+\begin{aside}
+% \marginpars{maybe I could adapt the proof of lemma 1 to prove the number of rounds of \gtx{},
+% although this is quite probabilist.}
 To handle the massive size of some large real world dataset, parallelizing existing algorithms is a
@@
 global permutation drawn at the beginning of the algorithm. On general graphs, this requires $O(\log
-n\diam(G))$ rounds\marginpars{maybe I could adapt the proof of lemma 1 to prove the number of rounds
-of \gtx{}, although this is quite probabilist.}, which on complete graphs reduces to $O(\log n)$.
+n\diam(G))$ rounds, which on complete graphs reduces to $O(\log n)$.
 Furthermore, this almost preserves the approximation factor, which is
@@
 O(\epsilon n\log^2 n)$. In practice, the solution is very close the one of \ccpivot{}, although it
-degrades as the number of threads increases. The number of rounds preserving the $3$-approximation
+degrades as the number of threads increases.
+
+The number of rounds preserving the $3$-approximation
 is lowered to $O(\log\log n)$ by \textcite{Ahn2015}. They also present results obtained in a single
@@
 2^{poly(\nicefrac{1}{\epsilon})}$ \autocite[Corollary 3.7]{Bonchi2013}.
+\end{aside}
 
+\begin{aside}
 As mentioned earlier, not having to set the number of clusters is an
@@
 problem was studied by \textcite{Giotis2006} on general graphs and we compile their results in
-\autoref{tab:cc_fixed}. On complete unweighted graphs and with $k$ being the number of clusters, they
-provide PTASs for \maxa{}$[k]$ running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and for
-\mind{}$[k]$ running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
-\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}k^6\log d\right)}$
-and that can handle weighted graphs. For $k=2$, \mind$[2]$ admits a faster local search method with a
+\autoref{tab:cc_fixed}. On complete unweighted graphs and with $K$ being the number of clusters, they
+provide PTASs for \maxa{}$[K]$ running in $nk^{O(\epsilon^{-3}\log(\frac{K}{\epsilon}))}$ time and for
+\mind{}$[K]$ running in $n^{O\left(\epsilon^{-2} 9^K\right)}\log(n)$ time. The latter was improved by
+\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}K^6\log d\right)}$
+and that can handle weighted graphs. For $K=2$, \mind$[2]$ admits a faster local search method with a
 factor $2$ approximation~\autocite{Coleman2008}.
 For complete weighted graph, \textcite{WeightedMaxAPTAS08} provide a PTAS for
-\maxa{}$[k]$ under the condition that the ratio between the largest and smallest weights is bounded by a
+\maxa{}$[K]$ under the condition that the ratio between the largest and smallest weights is bounded by a
 constant.
-
-\begin{table}[htpb]
-   \centering
-   \caption{Approximation results for \pcc{} on general graph with $k$ clusters} \label{tab:cc_fixed}
-   \begin{tabulary}{187mm}{lLL}
+\begin{center}
+   \captionsetup{font=small}
+   \captionof{table}{\label{tab:cc_fixed} Approximation results for \pcc{} on general graphs with $K$
+   clusters}
+   % 187mm - 3em = https://tex.stackexchange.com/a/8263
+   \begin{tabulary}{177mm}{lLL}
       \toprule
-      $k$	 & 2 & $\geq 3$ \\
+      $K$	 & $2$ & $\geq 3$ \\
       \midrule
-      \maxa{} & 0.878 (improved to 0.884 by \autocite{Mitra2009}) & 0.7666 \autocite{Swamy2004} \\
+      \maxa{} & $0.878$ (improved to $0.884$ by \autocite{Mitra2009}) & $0.7666$
+      \autocite{Swamy2004} (and it can be proved that this creates at most $6$ clusters) \\
       \mind{} & $O(\sqrt{\log n})$ as it reduces to Min 2CNF Deletion for which \textcite{min2CNF05}
       give such an approximation &
-      this can be reduced from $k$-coloring, which for any $\epsilon > 0$ is \NPc{} to approximate
+      this can be reduced from $K$-coloring, which for any $\epsilon > 0$ is \NPc{} to approximate
       within $n^{1-\epsilon}$ \autocite{InnaproxChroma07} \\
@@
    \end{tabulary}
-\end{table}
-
-summary APX hard by equivalence with multicut (what about complete)? 3 strategies for approximation
-(in rewrote) -> needs for heuristics
-To summarize, there are currently three main approaches to approximate \mind{}. On a general graph,
-one first solve the canonical linear program, and use the \regionGrow{} method to round the
-fractional solution, resulting in an O(log n) approximation. On complete graphs, the simplest
-solution is the \ccpivot{} algorithm, that grows clusters out of randomly selected pivots, in a way
-reminiscent of the proof of thm. The approximation factor of this general idea can be improved by
-interpreting the fractional solution of the linear program as probability to be included in the
-pivot's cluster
+\end{center}
+\end{aside}
 
@@
 
-\paragraph{Greedy methods}
-\label{par:cc_greedy}
-
 While all the methods we discuss so far
@@
 approaches that are designed to efficiently reach a solution that is satisfactory enough for the
-application at hand. For instance, while studying small scale signed social networks,
+application at hand.
+
+\paragraph{Greedy methods}
+\label{par:cc_greedy}
+Several of such methods fall into the greedy framework.
+For instance, while studying small scale signed social networks,
 \textcite{Early96} describe the following procedure. Start with a random initial $k$ clustering of
@@
 those whose merge would result in the maximum improvement of the score function. This runs in
-$O(n^2\log n)$, provides a $2$-approximation of \maxa{}, a $O(n\cdot \text{size of the largest
-cluster})$-approximation of \mind{}, and empirically results in fewer disagreements than \ccpivot{}.
+$O(n^2\log n)$, and empirically results in fewer disagreements than \ccpivot{}.
 Building upon their previous GRASP work~\autocite{GRASP13}, an iterated local search (ILS) heuristic is
@@
 and repeatedly samples an edge $(u,v)$ from $E_s^+$ before trying to make the neighborhoods of $u$
-and $v$ coincide by adding or removing edges in $E_s^+$.  Yet another idea is to modify the LP
-\eqref{eq:mindLP} to replace the binary variable $x_{uv}$ by a $k$ clusters indicator matrix
+and $v$ coincide by adding or removing edges in $E_s^+$.  Yet another idea is to modify the
+canonical LP to replace the binary variable $x_{uv}$ by a $k$ clusters indicator matrix
 $L\in\Rbb^{k\times n}$ where $L_{iu}$ is $1$ is $u\in C_i$ and $-1$
@@
 
-full conc of the section: the problem is difficult. However, it can be solved with guarentees in
-theory (because LP are not practical) and in efficiently in practice (using pivot as initilizatin
-for instance). Furthermore, instances obeying our bias can be solved much better. We next present
-settings where we are close to that bias
+This state of the art depicts a nuanced landscape for the \pcc{} problem. On one hand, minimizing
+the disagreements is difficult, although it is the most natural way given our bias. On the other
+hand, it can be solved with approximation guarantees, at least in theory. Indeed, the best methods
+requires solving linear programs with $m$ variables and up to $O(n^3)$ constraints, which results in
+a high complexity of $O(m^{4.5})$~\autocite[Section 7.2]{LPBook07}. Fortunately we can obtain
+reasonable solutions in practice, thanks to efficient heuristics (using the output of \ccpivot{} as
+initialization for instance). Furthermore, instances obeying our bias can be solved more easily and
+almost exactly. We next present settings where instances are indeed close to our bias.
+Before that, we remark that whereas \ccpivot{} is often used on general graphs in practice by
+assuming the missing edges are negative, doing so does not preserve its approximation guarantee.
+\begin{marginfigure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/raw/pivotstar.pdf}
+  \caption{A positive star with few negative edges}
+  \label{fig:cc_star_pivot}
+\end{marginfigure}
+
+\begin{aside}
+   Consider the graph $G$ showed in \autoref{fig:cc_star_pivot}. Assume there are $n$ nodes
+   connected positively to the center of the star, and $k<n-1$ negative edges between the peripheral
+   nodes. We define $G'$ as the complete graph obtained from $G$ by setting all missing edges to be
+   negative. On $G'$, the optimal solution is to have a cluster with the center of the star and one
+   peripheral node, while others peripheral nodes are put in singleton clusters. This creates $n-1$
+   positive disagreements. When running \ccpivot{} on $G'$, if the first pivot is a peripheral node,
+   this is the solution we obtain. Otherwise, with probability $\nicefrac{1}{n}$, the center of the
+   star is chosen as pivot, creating a single big cluster that incurs $(\shalf)(n-1)(n-2)$ negative
+   disagreements. In expectation, \ccpivot{} thus achieves \[
+   \frac{1}{n}\left(\frac{(n-1)(n-2)}{2}+(n-1)^2\right)=\frac{3n-4}{2n}(n-1)\sim \frac{3}{2}n \]
+   disagreements, which is roughly a $\nicefrac{3}{2}$-approximation as $n$ grows larger.
+
+   Now let us get back to the original $G$. Here the optimal solution is to have a single big
+   cluster, which incurs $k$ negative disagreements. With the same reasoning as before, the expected
+   number of disagreements of \ccpivot{} is now $\frac{1}{n}\left(k+(n-1)^2\right)$. If for
+   instance $k=\floor{\sqrt{n}}$ (as in \autoref{fig:cc_star_pivot}), then the previous expression
+   is equivalent to $n$ as $n$ tends to infinity. \ccpivot{} is thus $\sqrt{n}$ worst in expectation
+   than the optimal when run on this general graph.
+\end{aside}

### 2017-08-25 notations.tex
reorganize CC state of the art

@@
+In this thesis we will sometimes write remarks in a smaller font and with a light blue edging.
+\begin{aside}
+  Such remarks provide additional information to the topic discussed above, but can be skipped
+  without harming the comprehension of the main material.
+\end{aside}
+
 \begin{table*}[thpb]
@@
     \midrule
-    $G$ & An unweighted graph, that can be either directed or undirected \\
+    \rangek{} & The natural integers from $1$ to $K$, \ie{} $\{1, 2, \ldots, K \}$ \\
+    $G$ & An arbitrary graph. It should be clear from the context whether it is weighted or not, and directed or not \\
     $V$ & The set of all the nodes of a graph, with $|V|=n$. Unless noted otherwise, nodes are indexed from $1$ to $n$ \\
+    $u$ & A generic node of $G$. When referring to several nodes, we naturally use $u$, $v$, $w$ and so on. \\
     $E$ & The set of all the edges of a graph, with $|E|=m$ \\
-    $u$ & A generic node of $G$. When referring to several nodes, we naturally use $u$, $v$, $w$ and so on. \\
     $(u,v)$ & An undirected edge between nodes $u$ and $v$ \\
-    % \euv{} & A directed edge from node $u$ to node $v$ \\
+    \euv{} & A directed edge from node $u$ to node $v$ \\
     $\yuv{}$ & The sign of the edge $(u,v)$, which can be either $+1$ or $-1$ \\

### 2017-08-25 packages.tex
reorganize CC state of the art

--- 2017-08-09 packages.tex
@@
 sorting=nyvt, backref=true, backend=biber, defernumbers=true,maxbibnames=60]{biblatex}
+\renewcommand*{\bibfont}{\normalfont\footnotesize}
 \AtEveryBibitem{
-   \clearfield{arxivId}
+   % \clearfield{arxivId}
    % \clearfield{booktitle}
    % \clearfield{doi}
-   \clearfield{eprint}
-   \clearfield{eventdate}
+   % \clearfield{eprint}
+   % \clearfield{eventdate}
    \clearfield{isbn}
@@
    % \clearfield{volume}
-   \clearlist{location} % alias to field 'address'
+   % \clearlist{location} % alias to field 'address'
    \clearlist{publisher}
@@
 \def\algorithmautorefname{Algorithm}
+
+\usepackage{mdframed}
+\newenvironment{aside}%
+{\begin{mdframed}[hidealllines=true, leftline=true, innertopmargin=0.5ex, innerbottommargin=0.5ex,
+      innermargin=2.3em, leftmargin=2.3em, outermargin=0em, innerleftmargin=0.7em, innerrightmargin=0pt,
+      linewidth=1.5pt, skipabove=\baselineskip, skipbelow=\baselineskip,
+   linecolor=DodgerBlue!30!White]\small}%
+{\end{mdframed}}
+\usepackage{stmaryrd}
 

### 2017-08-25 cc/cc_non_worst_case.tex
reorganize non worst case CC and add two more refs

--- 2017-08-24 cc/cc_variants.tex
@@
-% leftover:
-% - something about fuzzy edges, ie we're not sure they are in the graph or not (not super
-%   interesting in my opinion) Clustering with partial information, Theoretical Computer Science,
-%   Volume 411, Issue 7, 2010, Pages 1202-1211, ISSN 0304-3975,
-%   \url{http://dx.doi.org/10.1016/j.tcs.2009.12.016}
-% - \paragraph{Hypergraphs} \Textcite{Kim2011} show a LP relaxation on hypergraph and
-%   \textcite{Ricatte13} describe a class of hypergraphs that can be reduced to signed graph.  Jörg
-%   Hendrik Kappes, Markus Speth, Gerhard Reinelt, Christoph Schnörr, Higher-order segmentation via
-%   multicuts, Computer Vision and Image Understanding, Volume 143, 2016, Pages 104-119, ISSN
-%   1077-3142, \url{http://dx.doi.org/10.1016/j.cviu.2015.11.005}
-% - Interactive \pcc{}?~\autocite{interactiveCC16}, could it go into active?  Geerts, F. & Ndindi,
-%   R. Bounded correlation clustering. Int. J. Data Sci. Anal. 1, 17–35 (2016)
-%   \url{https://doi.org/10.1007/s41060-016-0005-2}
-
-So far we focused on \pcc{} in its rawest form, that is solving the \mind{} and \maxa{} objectives
-in the case where the general binary-labeled graph is known in advance. We will now see first some
-special cases, namely when the weights obey the triangle inequality (to solve \msc{}) or when the
-graph is bipartite and then move to variants. We will consider more general objectives, when the
-edges are labeled categorically instead of binary, when nodes can belong simultaneously to several
-clusters or when we optimize local objectives per nodes instead of global ones. We will also look at
-clustering in signed graphs in general, using spectral methods or heuristics from the community
-detection literature. Finally, we will shift perspective and review methods beyond the batch case and
-the worst-case analysis.
-
-% \paragraph{\msc{}}                                | special weights
-% \paragraph{Bipartite \pcc{}}                      | special graphs
-% \paragraph{Categorical edge labelling}            | more general objective
-% \paragraph{Overlapping \pcc{}}                    | more general objective
-% \paragraph{Local \pcc{}}                          | more general objective or at least different
-% \paragraph{Spectral clustering}                   | related objective
-% \paragraph{Communities detection}                 | related objective
-% \paragraph{Online \& active setting}              | different setting
-% \paragraph{recovery under noise}                  | non worst case analysis
-% \subsubsection{\pcc{} under stability assumption} | non worst case analysis
-
-\paragraph{\msc{}}
-
-In \msc{}, the goal is to output a clustering which best summarizes (or agrees with) the several
-given input clusterings of the same set of objects.  Motivations include robustness ---by using an
-ensemble of clusterings from diverse methods--- and privacy ---if the clusterings were computed by
-different parties each considering only a subset of the objects attributes. We can build the
-complete graph of these objects, with weights set to the fraction of clusterings that place two
-objects in different clusters, thus representing a kind of distance in the space of clusterings.  As
-first show by \textcite{EarlyConsensusClustering03} , finding the optimal clustering is therefore
-an instance of \pcc{} where the weights obey the triangular inequality. \textcite{Gionis2007} give a
-deterministic $3$-approximation using the \regionGrow{} method. Later \textcite{Bonizzoni2008} show that
-the minimization version is \APXh{}, even when the input is made of three clusterings and give a
-combinatorial $\nicefrac{4}{5}$-approximation for the maximization problem. Experimental evaluations are
-conducted by \textcite{Bertolacci07} and \textcite{Filkov08}. The former describe a scalable
-approach that first samples a small portion of the data, runs a (potentially computationally
-expensive) approximation algorithm and finally augment the resulting partition by adding to it the
-unsampled nodes one by one.  Experiments confirm that the running time is greatly improved compared
-with the linear program methods while the resulting objective value is essentially the same. Note,
-however, that LP methods can be applied in practice thanks to some tricks~\autocite{ConsensusLP10}.
-If we have $k$ input clusterings $\cluster^1,\ldots,\cluster^k$ and we parameterized the problem by
-$t$, which is the sum over the input clusterings of the number of pairs of objects that are clustered
-differently by a solution $\cluster^\star$ and $\cluster^i$, then there is a polynomial algorithm
-running in $O(4.24^{\nicefrac{t}{k}}\cdot \nicefrac{t}{k}^3 +
-kn^2)$~\autocite{parameterizedConsensus14}.
-
-\paragraph{Bipartite \pcc{}}
-
-Bipartite graphs are an interesting special case for \pcc{}, as they often appear in the context of
-recommendation systems, where users rate products positively or negatively, although in this setting
-we cannot expect to have complete bipartite graphs in practice. The first results was given by
-\textcite{Amit04}, who obtains an $11$-approximation for \mind{} by adapting the \regionGrow{} method.
-The \ccpivot{} is adapted to the bipartite case by
-\textcite{Bipartite12}, who prove it results in a randomized $4$-approximation (and provide a matching
-deterministic approximation by rounding a LP). By using their idea of rounding the results of the LP
-differently for positive, negative and in that case same-side absent edges, \textcite{Chawla2014}
-bring down this approximation factor to $3$, even for $K\geq 2$-partite graphs. Through formulating the
-\maxa{}$[k]$ problem as a bilinear maximization problem and computing a low-rank approximation of
-the graph biadjacency matrix, \textcite{Asteris2016} obtain a efficient PTAS, that is a $(1-\delta)$
-approximation running in time exponential in $k$ and $\delta^{-1}$ but linear in $n$. By an
-appropriate choice of $k$, it is possible to use this PTAS to solve the general \maxa{} problem.
-
-\paragraph{Categorical edge labelling}
-
-In the so called \textsc{Chromatic}-\pcc{} setting, \enquote{positive} edges are now associated with
-one of $L$ possible colors and the goal is to form clusters mostly made up of edges with the one same
-color. Namely, a disagreement is now a negative edge between clusters or a within-cluster edge
-whose color differs from the majority color of that cluster. This is motivated by edge-labeled graph
-in social networks, biology and citation networks and will discuss such applications in
-\autoref{chap:vector}. As a generalisation of \pcc{}, it is \NPc{} but \textcite{Bonchi2012a}
-present a modification of the \ccpivot{} algorithm that pick edges instead of nodes as pivots, and
-grow clusters by adding monochromatic triangles. This gives an approximation factor of six times the
-maximum degree of the graph. They also present a method when the number $k$ of clusters is fixed
-beforehand, starting with an initial partition and then alternating between finding the majority
-color of clusters and finding better clusters. Unfortunately, the maximum degree of a graph can be
-as large as $n$. However, \textcite{Anava2015} present constant factor approximations. Namely, they
-show the problem can be reduced to classical \pcc{} by setting all edges incident to a node $u$ to
-negative if they are not of the majority color of $u$. They then apply the regular \ccpivot{} and
-show this gives an $11$-approximation to the original problem. Furthermore, they also write a linear
-program and round it using the \regionGrow{} method to obtain an
-approximation factor of $4$. \Textcite{multiChromatic15} extend this line of work to the case were a
-single edge can carry a \emph{set} of labels and adapt their randomized algorithm so that the
-approximation factor is multiplied by the size of the input label set.
-
-\paragraph{Overlapping \pcc{}}
-
-While in \pcc{}, each node is assigned to a single cluster, in other settings we may want to relax
-this constraint. Given a complete weighted graph, \textcite{Bonchi2012} want to output a clustering
-\cluster{} that minimizes the following cost: \[ \sum_{(u,v)\in E} \left| H(\cluster(u),
-\cluster(v)) - w_{uv}\right|\] where $H$ is a similarity function between two sets of labels, chosen
-here to be the Jaccard similarity or a $0/1$ indicator of nonempty intersection. These problems are
-showed \NPc{} and approximated by a local search algorithm, iteratively optimizing the assignment of
-one node while all others are fixed. As one of the demonstration on their theoretical work,
-\textcite{WeightedTheta15} show a faster solution based on a weighted extension of the Lovász's
-theta function, the corresponding geometric embedding of graphs and a solver derived from one-class
-SVM, while \textcite{GeneticOCC14} propose a genetic algorithm to solve this problem. Finally,
-\Textcite{StochasticCC13} also deals with overlapping clustering by relaxing the problem to a
-stochastic setting and using \enquote{the Baum-Eagon inequality, which provides an effective
-iterative scheme for maximizing polynomial functions in probability domains}.
-
-\paragraph{Local \pcc{}}
-
-In classical \pcc{}, all nodes have an identical role, in the sense that they contribute equally to
-the final objective in terms of (dis)agreements. Here we instead look at approaches where we either
-add a local penalty to each in order to better control their behavior or where we altogether modify
-the objective to focus on (dis)agreements at specific nodes.
-
-% that one is not really local, except for the node penalty
-\Textcite{Puleo2014} adapt the linear program of~\autocite{Charikar2003} and its \regionGrow{}
-method to the case
-where all clusters have to contain less than $K$ nodes, by assigning to each node $u$ a penalty
-$\mu_u$. If $u$ is placed in a cluster $C_i$, the original \mind{} objective is penalized by an
-extra $\mu_u\left(|C_i| - (K+1)\right)$. By varying $\mu_v$ between $0$ and $1$ and because the
-positive weights are assumed to smaller than $1$, this cluster size constraint can be made hard or
-soft. They also handle more general weights, since they allow $w^-_{uv}$ to be as large as $\tau$
-for $\tau\in [1,\infty)$ while still guaranteeing a $5-\nicefrac{1}{\tau}$-approximation on complete
-graphs, and adapt \ccpivot{} to unweighted graphs with the hard cluster size constraint, obtaining a
-randomized $7$-approximation.
-These soft constraints are for instance used in a biological application where nodes are genes and
-where singleton and giant clusters are uninformative~\autocite{Hou2016}
-
-\Textcite{pmlr-v48-puleo16} also modify the \mind{} objective to make it more general. Based on the
-classic \pcc{} linear program, they define a \enquote{\emph{fractional clustering} of $G$ as a
-vector $x$ indexed by $V$ such that $x_{uv} \in [0, 1]$ for all $uv \in \binom{V}{2}$ and such that
-$x_{vz} \leq x_{vw} + x_{wz}$ for all distinct $v, w, z \in V$}. They also define \enquote{The
-error vector $err(x)$ of $x$, as a real vector indexed by $V$ whose coordinates are}
-\begin{equation*}
-  \mathrm{err}(x)_u = \sum_{v\in\nei^+(u)} x_{uv} + \sum_{v\in\nei^-(u)} (1-x_{uv})
-\end{equation*}
-
-Given a function $f: \Rbb^n_{\geq 0} \rightarrow \Rbb$ verifying two elementary conditions, the
-problem is then to find a fractional clustering $x$ minimizing $f(err(x))$. The classical \pcc{}
-corresponds to setting $f(x) = \lhalf{}\ell^1(x)$ whereas the authors here are interested in Minimax
-\pcc{} that arises by setting $f(x) = \ell^\infty(x)$. Minimizing the maximum number of
-disagreements incurred by a single node is motivated by the example of recommendation systems: if
-errors correspond to unsatisfying recommendations, we do not want a single user to suffer many
-of them. Minimax \pcc{} is \NPc{} on both complete graphs and complete bipartite graphs but
-by modifying the \regionGrow{} method, the authors respectively a $48$ and
-$10$ approximation, the latter for the one-sided error (that only counts disagreements for the nodes
-in one of the two clusters).
-The idea is to chose pivots not randomly but by maximizing a given criteria and
-to grow balls with a radius $\alpha$ computed numerically to optimize the approximation factor.
-Interestingly, and in contrast with the classic \pcc{} situation, minimax \maxa{} is not easier than
-minimax \mind{} and seems not to have a constant factor approximation, even on complete graphs.
-Furthermore, these algorithms are deterministic, as opposed to many \pcc{} approximations, since
-bounds on expected disagreements of an edge does not translate easily on their maximum.
-\Textcite{Charikar2017} improve these two factors to $7$, using a simpler version of the algorithm
-of \textcite{pmlr-v48-puleo16}. Namely, find the ball of radius $\nicefrac{1}{7}$ with the largest
-number node and create a cluster from its center with a radius of $\nicefrac{3}{7}$. They also show
-that on general weighted graphs, the LP has a large integrality gap of $\nicefrac{n}{2}$ yet they
-combine it with a combinatorial approach to reach a $O(\sqrt{n})$ approximation. Finally, they
-consider the complementary problem of maximizing the minimum number of agreements reach at a single
-node, and provide a $\frac{1}{2+\epsilon}$ approximation.
-
-\paragraph{Spectral Clustering}
-
-A classic method for clustering graphs is to leverage their spectral properties. Namely, if $A$ is
-the adjacency matrix of $G$ and $D$ its degree diagonal matrix (that is $D_{u,u} = \degr(u)$), the
-\emph{Laplacian} of $G$, defined by $L_G = D - A$, is a symmetric positive semidefinite matrix. As
-such, it has $n$ real non-negative eigenvalues, and its spectrum provides additional information on
-the connectivity of $G$. For instance, $0$ is always the smallest eigenvalue and its multiplicity is
-equal to number of connected components of $G$, while ---if $G$ is connected--- the second
-eigenvalue is the algebraic connectivity of $G$, whose magnitude is an indication of how well
-connected is the graph. This matrix is typically used for clustering by computing its first $k$
-eigenvectors, which embed the $n$ nodes of $G$ in $\Rbb^k$, where there are then clustered with the
-$k$-means algorithm. This can be seen as a relaxation of the discrete \rcut{} objective, which asks
-for the partition $\{C_1, \ldots, C_k\}$ minimizing $\frac{1}{2}\sum_{i=1}^k \frac{\mathrm{cut}(C_i,
-\bar{C_i})}{|C_i|}$, where $\bar{C_i}$ is the complement of $C_i$ in $V$ and $\mathrm{cut}(B, C) =
-\sum_{u\in B, v\in C} w_{uv}$ is the total weight of the edges between $B$ and
-$C$~\autocite{tutoSpectralClustering07}. By considering the symmetric normalized Laplacian
-$L_{sym}  = D^{\nicefrac{1}{2}}LD^{\nicefrac{1}{2}}$, it is also possible to approximate the
-normalized cut objective (\ncut{}), where $|C_i|$ is replaced by $vol(C_i) = \sum_{u\in C_i}
-\degr(u)$. We will now see how these kinds of approaches can be extended to signed graphs, noting
-first that they require to fix the number of clusters beforehand and are looking for clusters
-balanced in size, which makes the problem related but not equivalent to \pcc{}.
-
-The first line of research consider only \mind{}$[2]$. For instance, \textcite{NcutAnd2CC08} show
-that both normalized cut and \mind{}$[2]$ objectives can be written as a SDP (or equivalently as
-eigenvalue problems) and thus combined, the intuition being that we look for \ncut{} solutions whose
-number of disagreements is not too much more than the approximate optimal.
-
-Letting \ncut{}$(C, \bar{C}) = \frac{\mathrm{cut}(C, \bar{C})}{\mathrm{bal}(C)}$ with
-$\mathrm{bal}(C) = 2\frac{vol(C)vol(\bar{C})}{vol(V)}$, \textcite{mOneCC12} define a new objective:
-\begin{equation*}
-  \hat{F}_\gamma(C) = \frac{\mathrm{cut}(C, \bar{C}) + \gamma\left(\hat{M}(C)+\hat{N}(C)\right)}{\mathrm{bal}(C)}
-\end{equation*}
-where $\gamma \in \Rbb{}_+$ is a parameter, while $\hat{M}(C)$ and $\hat{N}(C)$ are respectively the
-number of positive and negative disagreements of the $(C, \bar{C})$ clustering.
-They show how to optimize a tight continuous relaxation of $\hat{F}_\gamma$ as the non-negative
-ratio of a difference of convex function and a convex function.
-
-On the other hand, one can also adapt these two cut objectives to directly include negative edges.
-\Textcite{Luca10} define the signed Laplacian as $\bar{L} = \bar{D} - A$, where $\bar{D}$ is the
-signed degree matrix such that $\bar{D}_{uu} = \sum_{v\in \nei(u)} |A_{uv}|$, as well as a signed
-variant of the symmetric normalized Laplacian $\bar{L}_{sym}  = \bar{D}^{\nicefrac{1}{2}} \bar{L}
-\bar{D}^{\nicefrac{1}{2}}$. They show that the signed Laplacian is positive semidefinite, and
-even positive-definite as soon as the graph is unbalanced (that is contains a cycle with an odd
-number of negative edges). From positive and negative cuts defined as $\mathrm{cut}^+(B, C) =
-\sum_{u\in B, v\in C} w^+_{uv}$ and $\mathrm{cut}^-(B, C) = \sum_{u\in B, v\in C} w^-_{uv}$, a
-natural signed cut is $\mathrm{scut}(B, C) = 2\mathrm{cut}^+(B, C) + \mathrm{cut}^-(B, B) +
-\mathrm{cut}^-(C, C)$ which can then be used to defined signed \rcut{} and \ncut{}. Arguing that those
-definitions force negatively linked nodes to be symmetric around the origin, do not take into
-account the balance of negative edges in each cluster and are difficult to extend to more than two
-clusters, \textcite{SignedEmbedding15} instead propose two new normalized cuts:
-\begin{align*}
-  SNScut(C_1, \ldots, C_k) &= \sum_{i=1}^k
-  \frac{\mathrm{cut}^+(C_i, \bar{C_i})-\mathrm{cut}^-(C_i, \bar{C_i})}{vol(C_i)}  \\
-  BNScut(C_1, \ldots, C_k) &= \sum_{i=1}^k
-  \frac{\mathrm{cut}^+(C_i, \bar{C_i})-\mathrm{cut}^-(C_i, \bar{C_i})+vol^-(C_i)}{vol(C_i)}
-\end{align*}
-
-Noting that if $x_i\in\Rbb^n$ is the vector indicator of cluster $C_i$ (that is the \uth{} entry of
-$x_i$ is $1$ is $u$ belongs to $C_i$ and $0$ otherwise), $x_i^T\bar{L}x_i = 2\mathrm{cut}^-(C_i,
-C_i) + \mathrm{cut}^-(C_i, \bar{C}_i) + \mathrm{cut}^+(C_i, \bar{C}_i)$,
-\textcite{mSemanticWordCC17} introduce the following cut objective:
-\begin{equation*}
-  sNcut(C_1, \ldots, C_k) = \sum_{i=1}^k
-  \frac{2\mathrm{cut}^-(C_i, C_i) + \mathrm{cut}^-(C_i, \bar{C}_i) + \mathrm{cut}^+(C_i, \bar{C}_i)}{vol(C_i)}
-\end{equation*}
-Additional cut formulations for $k$-clusters are also presented in~\autocite{moreSignedCut12},
-although \textcite{Knyazev2017l} argue that using the non signed Laplacian and considering negative
-eigenvalues might be just as effective, citing for instance numerical instability of signed
-Laplacian.
-
-Finally, \textcite{mGeometricMean16} show that the Laplacians defined so far can be seen as
-arithmetic means of the Laplacian $L^+$ of the positive subgraph $G^+=(V, E^+)$ and the signless
-Laplacian $Q^-$ of the negative subgraph $G^-=(V, E^-)$, where $Q^- = D^- + A^-$. They suggest
-instead to use a geometric mean defined for two positive matrices $A$ and $B$ as $A\#B =
-A^{\shalf{}} \left(A^{-\shalf{}} BA^{-\shalf{}} \right)^{\shalf{}} A^{\shalf{}}$. This suggestion is
-based on the fact that if $u$ is a common eigenvector of both $A$ and $B$ with eigenvalue $\lambda$
-and $\mu$ respectively, then $u$ is an eigenvector of $A+B$ with eigenvalue $\lambda+\mu$ and an
-eigenvector of $A\#B$ with eigenvalue $\sqrt{\lambda\mu}$. Therefore, the $k$ smallest eigenvalues
-of the geometric mean Laplacian will be influenced by both smallest eigenvalues of $L^+$
-(corresponding to assortative clusters in $G^+$) and of $Q^-$ (corresponding to disassortative
-clusters in $G^-$), while this is note the case for the arithmetic mean of Laplacians.
-
-\paragraph{Community detection}
-
-The clustering problem is often named community detection in the context of social networks, and
-several methods developed by practitioners have been extended to signed graphs. While they do not
-necessarily considered the \pcc{} objective, and especially not its optimum, we still give a brief
-overview of them, as they tend to have been more tested experimentally. For instance, to find the
-cluster of node $u$, \textcite{Yang2007} use a random walk approach on the positive subgraph $G^+$
-to compute the probability of each node to reach $u$ in $T$ steps, sort the nodes accordingly and
-then find a threshold  based on the number of disagreements.
-The one node moves local heuristic that we described in the Physics-inspired paragraph
-\vpageref{par:cc_physics} can also be formulated as genetic algorithms that simultaneously try to
-minimize the number of disagreements and maximize a signed variant of the
-modularity~\autocites{Li2013}{Amelio2013}. \Textcite{Anchuri2012} also consider these two objectives
-by seeing them as eigenvalue problems and devise a iterative splitting procedure.
-The overlapping community detection variant is considered by \textcite{Chen14}, who used a signed
-probabilistic mixture model. Namely, an edge selects a pair of cluster $r,s$ with probability
-$\omega_{rs}$ (where $r=s$ if the edge is positive and $r\neq s$ otherwise) and chooses its endpoints
-$u$ and $v$ with probability $\theta_{ru}$ and $\theta_{sv}$. $\theta_{ru}$ is therefore the soft
-membership of node $u$ to cluster $r$, and those parameters are estimated using the
-expectation-maximization algorithm. The same model is extended to directed graphs by
-\textcite{Jiang2015}, who strangely enough names it stochastic blockmodel, although the focus is
-still on edge and not nodes.
-In a similar spirit to \maxa{}$[k]$, \textcite{SignedGang} focus on finding $k$ subgraphs dense in
-positive edges and densely connected by negative edges to each other. Thee dub such subgraph
-\emph{Oppositive Cohesive Groups}, or more vividly \emph{Gangs in War}, and after formulating the
-problem as a constrained quadratic optimization, they propose a faster iterative local search
-heuristic.
-% relaxed structural balance \autocite{Doreian2009} but used the same approach as \textcite{Early96}
+We now look at two settings where our bias is likely to be present.  In the first one, we assume the
+input graph is randomly perturbed version of a consistent clustering. In the second one, we do not
+assume anything about such a consistent clustering but expect instead that the optimal solution is
+\enquote{clear} enough that it does not change when weights are multiplicatively perturbed.
 
@@
 % number of pair. I think it was extended in \url{https://doi.org/10.1109/TIT.2016.2600566}
+
+\paragraph{\pcc{} under stability assumption}
+
+Whereas clustering objective functions are \NPh{} to optimize, we expect that meaningful instances
+in which we are interested have additional structure which allows for guaranteed polynomial time
+algorithms, see for instance \autocite{clusteringFeasibility15} for a critical overview of some
+proposed such notions of structure. Informally, a general idea is that the clustering should not
+change (or at least very little) if the data are slightly perturbed. For instance, a weighted graph
+is \emph{$\alpha$-stable} (with $\alpha>1$) for some partition objective if its optimal partition
+remains the same whenever every weight $w_i$ is multiplied by a factor $c_i$ between $1$ and
+$\alpha$.
+Finally, we say that an algorithm is \emph{robust} if, given an instance $\mathcal{I}$ and
+in polynomial time, it: returns the optimal solution of $\mathcal{I}$ if $\mathcal{I}$ is
+$\alpha$-stable; and if $\mathcal{I}$ is not $\alpha$-stable, either returns the optimal solution of
+$\mathcal{I}$ or reports that $\mathcal{I}$ is not $\alpha$-stable. This is a handy property, as in
+general, we cannot check the stability of an instance in practice as we do not know the optimal
+solution hence we cannot tell whether it changes or not under perturbations.
+
+See \autocite{StableCC09} for application to \mind{} on complete graphs and \autocite{StableLP09}
+for more general study of the solution stability of the canonical \mind{} LP in any kind of graph.
+
+\iffalse
+\Textcite{StableCC17} provide a robust algorithm for $2-\nicefrac{2}{k}$-stable instance of
+$k$-\textsc{Minimum Multiway Cut}, which therefore is not really useful in our case because I confused
+multiway cut and multicut. To be fair, it seems I'm not the only one, \href{http://pages.cs.wisc.edu/~shuchi/papers/multicut-hardness-full.pdf}{on the hardness of
+   approximating multicut}
+   \enquote{multicut is known to be APX-hard
+      (\href{https://pdfs.semanticscholar.org/1cf6/4c2bdd4f1c384a55910606a64c8d831a96ba.pdf}%
+   {Dahlhaus et al. 1994}).} by that 1994 paper define the multiway cut problem, see
+   \autoref{fig:cc_multiwhat}. \Textcite[Theorem 24]{Bansal2004} show that we can go from
+   \textsc{Minimum Multiway Cut} to \pcc{} but I doubt about the other direction, mostly because
+   \textsc{Minimum Multiway Cut} can actually be approximated with factor smaller than $1.3$.
+   see \url{http://www.nowozin.net/sebastian/papers/nowozin2009lpstability.pdf} about LP multicut
+   stability.
+\begin{figure}[htpb]
+   \centering
+   \includegraphics[width=0.9\linewidth]{assets/raw/multicut_vs_multiway.png}
+   \caption{Confusing statement} \label{fig:cc_multiwhat}
+\end{figure}
+\fi
+
+% \enquote{We prove that if the Unique Games Conjecture of Khot (2002) is true, then for every constant L > 0
+% it is NP-hard to approximate Multicut within factor L. If a quantitatively stronger version of the
+% conjecture is true, then Multicut is NP-hard to approximate within a factor of $\Omega(\sqrt{\log
+% \log n})$.}
+
+% It's probably limited because LP with a lot of constraints do not scale that well (someone claimed
+% $O(n^{4.5})$)
+
+% SDP Gaps and UGC Hardness for Multiway Cut, 0-Extension and Metric Labelling,
+% On the Unique Games Conjecture: Subhash Khot
+% under the UGC, one cannot get better than $O(\log n)$ approximation using LP or SDP relaxation.

### 2017-08-25 cc/cc_variants.tex
reorganize non worst case CC and add two more refs

@@
 edges are labeled categorically instead of binary, when nodes can belong simultaneously to several
-clusters or when we optimize local objectives per nodes instead of global ones. We will also look at
+clusters or when we optimize local objectives per nodes instead of global ones. Finally we will also look at
 clustering in signed graphs in general, using spectral methods or heuristics from the community
-detection literature. Finally, we will shift perspective and review methods beyond the batch case and
-the worst-case analysis.
+detection literature.
 
@@
 % \paragraph{Communities detection}                 | related objective
+
 % \paragraph{Online \& active setting}              | different setting
@@
 % relaxed structural balance \autocite{Doreian2009} but used the same approach as \textcite{Early96}
-
-\paragraph{\pcc{} under noise}
-
-As we have seen, assuming the Unique Games Conjecture, the \mmc{} problem, and therefore
-the \pcc{} problem on a general graph, cannot be approximated to within a constant factor in the worst
-case.  But maybe we can do better in the average case, which motivates the study of semi-random
-model, where real graphs are seen as being obtained from the controlled perturbation of a perfectly
-clusterable graph. In the simplest case, each edge sign is independently flipped with probability
-$p \in [0, \shalf)$. This situation on complete graphs was considered in \autocite[Section
-6]{Bansal2002}, showing a simple algorithm that with high probability makes
-$\tilde{O}(n^\frac{3}{2})$ mistakes, and in \autocite[Theorem 2.6]{Ben-Dor99}, with an algorithm
-recovering with high probability the planted partition of an unweighted graph in $O(n^2(\log n)^c)$,
-where $c$ depends on the size of the smallest cluster and the noise probability.
-
-\Textcite{Joachims2005} analyze a more refined weighted model, where weights are generated by a
-probability distribution whose mean on true positive edges is larger than $\mu^+>0$ and whose mean
-on true negative edges is smaller than $\mu^-<0$. They give a finite-sample bound on the number of
-node misclustered w.r.t the planted partition as a function on the probability distribution
-parameters. Indeed, as pointed out by \textcite{plantedAilon09}, independent and uniform noise is
-not a good model of real situations, where the input of \pcc{}, \ie{} the similarity between nodes,
-is often the result of a preprocessing, which may present strong correlations. Therefore, instead of
-measuring the quality of a solution against in the input (\ie{} the similarity information), they
-argue it is more sensible to measure it against the (unknown) true optimal clustering that gave rise
-to the input, and show that \ccpivot{} allows that thanks to a new analysis. \Textcite{Mathieu2010}
-also consider an adversarial model, where all edges are flipped with probability $p$ but the
-adversary then decide whether to show us the true sign or the flipped sign. On complete unweighted
-graphs, they find a solution of \mind{} at most $1 + O(n^{-\nicefrac{1}{6}})$ times the optimal
-whenever $p \leq \shalf - n^{-\nicefrac{1}{3}}$ by rounding the usual SDP solution. If in addition
-$p\leq\nicefrac{1}{3}$, there is no adversary and each planted cluster has at least
-$\Theta(\sqrt{n})$ nodes, then the planted partition can be recovered exactly.
-\Textcite{Makarychev2014} study the same adversarial model but on general weighted graphs, giving a PTAS
-for \mind{} when $p\leq \nicefrac{1}{4}$. Under additional assumptions on the density of edges, they
-present another algorithm that finds the ground truth clustering with an arbitrarily small
-classification error.\footnote{They also mention a potentially interesting related
-work~\autocite{unobservedCC14}.} % http://www.jmlr.org/papers/v15/chen14a.html
-% \textcite{Noisy2CCLocality16} study the recovery of an arbitrary planted 2-partition under noisy
-% measurement in special kind of graphs where each node is connected with a local neighborhood and
-% show computationally efficient algorithms that only query an information theoretically optimal
-% number of pair. I think it was extended in \url{https://doi.org/10.1109/TIT.2016.2600566}

### 2017-08-25 cc/gtx_algorithm.tex
another failed proof of gtx number of iteration

--- 2017-07-08 gtx_algorithm.tex
@@
 
+Here is another proof that does not work, where we study the diameter of the graph. Indeed, it is
+initially bounded by the number of nodes $n$, and by the correctness of \gtx{}, it will be zero upon
+termination. If we show that the diameter shrinks by a constant factor at each iteration, then we
+can claim that \gtx{} terminates after a number of iterations logarithmic in $n$. For ease of
+exposition, we instead prove equivalently that when going backward in time, the diameter expands at
+least by a constant factor at each iteration. Let us denote by $\Delta_t$ the diameter of $G_t$ and
+take any path of length $\ell \leq \Delta_t$ in $G_t$. Let $[s_1, s_2, \ldots, s_{\ell+1}]$ be the
+nodes along this path, where the notation $s_i$ reflects the fact that these nodes were stars of
+nodes in $G_{t-1}$. If we choose any $u,v \in V_{t-1}^2$ such that $u$ is in the star $s_1$ and $v$
+is in the star $s_{\ell+1}$, then we can see on \autoref{fig:gtx_expand_path} that there exists a
+path of length $2\ell -1$ between $u$ and $v$ in $G_{t-1}$.
+{\color{blue}{\large The figure actually shows this is not always true.}
+More generally, for any path of length
+$\ell$ in $G_t$, there is a path of length $2\ell-1$ in $G_{t-1}$. In particular, for
+$\ell=\Delta_t$, there is a path of length $2\Delta_t-1$ in $G_{t-1}$, from which we conclude that
+$2\Delta_t-1\leq\Delta_{t-1}$, or equivalently $\Delta_t \leq
+\frac{1}{2}\left(\Delta_{t-1}+1\right)$. One can then check by recurrence that $\Delta_t \leq
+\frac{1}{2^t} \left(\Delta_0+2^t-1\right) \leq \frac{n}{2^t} + 1$.}
+
+\begin{figure*}[htpb]
+   \centering
+   \includegraphics[width=\linewidth]{raw/expand_path.pdf}
+   \caption[Star expansion of a path]{A path within $G_t$ in black, and the underlying graph
+     $G_{t-1}$ in blue. In the worst case, the endpoints of the path in $G_{t-1}$ are at the
+     periphery of the endpoint stars.  The intermediary nodes in $G_t$ could be proper or singleton
+     star, but there cannot be two adjacent singleton stars.}
+   \label{fig:gtx_expand_path}
+\end{figure*}
+
+\iffalse
 We would like to prove a similar result on the edges so that the overall complexity of the
@@
 
-\iffalse
 In the case no singleton are formed during one execution of \extractStar{}, the number of nodes is
@@
 a trade off between having more edges than $|V_0|-1$ but potentially making shorter connection and
-thus having a lower stretch.
+thus having a lower stretch. This would also be interesting in the sign prediction setting. Assuming
+the treewidth of $G_t$ is decreasing with $t$, maybe we could show there is \enquote{reasonable}
+number of paths between $u$ and $v$, compute their parity and take a majority vote on the results.

### 2017-08-25 cc/cc_state_of_art.tex
move definition of PTAS earlier

@@
 the problem can be solved in polynomial time~\autocite{Abu-Khzam2015} by using various reduction
-rules. Finally, for planar graph,
-there is a PTAS running in $O(n2^{\epsilon^{-1}\log(\epsilon^{-1})})$ obtained by dividing the graph
+rules. Finally, for planar graph, there is a PTAS%
+\bottomfootnote{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
+minimization (respectively maximization) problem $\mathcal{P}$  in NP if given any $\epsilon>0$ and
+any instance $x$ of $\mathcal{P}$ of size $n$, $\mathcal{A}$ produces, in time polynomial in $n$, a
+solution that is within a factor $1+\epsilon$ (respectively $1-\epsilon$) of being optimal with
+respect to $x$. Note that the time is not necessarily polynomial in $\epsilon$, so that a running
+time of $O(n^{\frac{1}{\epsilon}})$ would qualify~\autocite[Definition 3.10]{CpxBook99}.}
+running in $O(n2^{\epsilon^{-1}\log(\epsilon^{-1})})$ obtained by dividing the graph
 into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}. In other words, if we
@@
 give a $17429$-approximation of \mind{} and a PTAS%
-\marginpars{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
-minimization (respectively maximization) problem $\mathcal{P}$  in NP if given any $\epsilon>0$ and
-any instance $x$ of $\mathcal{P}$ of size $n$, $\mathcal{A}$ produces, in time polynomial in $n$, a
-solution that is within a factor $1+\epsilon$ (respectively $1-\epsilon$) of being optimal with
-respect to $x$. Note that the time is not necessarily polynomial in $\epsilon$, so that a running
-time of $O(n^{\frac{1}{\epsilon}})$ would qualify~\autocite[Definition 3.10]{CpxBook99}.}
 that, for any $\epsilon \in [0,1]$, runs in

### 2017-09-02 troll/algs_batch.tex
initial import of troll paper (with citation fix)

\section{Algorithms in the Batch Setting}\label{s:algbatch}
Given $G(Y) =(V,E(Y))$, we have at our disposal a training set $E_0$ of labeled edges from $E(Y)$, our goal being that of building a predictive model for the labels of the remaining edges.

Our first algorithm is an approximation to the Bayes optimal predictor $y^*(i,j)$. Let us denote by $\htr(i)$ and $\hun(i)$ the trollness and the untrustworthiness of node $i$ when both are computed on the subgraph induced by the training edges. We now design and analyze an edge classifier of the form
%
\begin{equation}
\label{eq:predictor}
	\sgn\Big(\big(1-\htr(i)\big) + \big(1-\hun(j)\big) - \tfrac{1}{2} -\tau\Big)~,
\end{equation}
%
where $\tau \ge 0$ is the only parameter to be trained. Despite its simplicity, this classifier works reasonably well in practice, as demonstrated by our experiments (see Section~\ref{s:exp}). Moreover, unlike previous edge sign prediction methods for directed graphs, our classifier comes with a rigorous theoretical motivation, since it approximates the Bayes optimal classifier $y^*(i,j)$ with respect to the generative model defined in Section~\ref{s:gen}. It is important to point out that when we use $1-\htr(i)$ and $1-\hun(j)$ to estimate $p_i$ and $q_j$, an additive bias shows up due to~(\ref{e:pout}). This motivates the need of a threshold parameter $\tau$ to cancel this bias. Yet, the presence of a prior distribution $\mu(p,q)$ ensures that this bias is the same for all edges $(i,j) \in E$.

%\subsection{Approximation to Bayes via Active Learning}\label{ss:active}
%
% We now introduce an active learning algorithm approximating the Bayes optimal predictions $y^*(i,j)$.

\iffalse
***********************************************************************
Given a positive integer parameter $Q < \tfrac{|E|}{2|V|+1}$ such that there exists a set\footnote
{
This set is needed to find an estimate $\tauhat$ of $\tau$ in~(\ref{eq:predictor}), see Step~3 of the algorithm. It can be any set of directed paths/cycles in $G$ that are pairwise vertex-disjoint.
}
$E_L \ss E$ of size $Q$ where each vertex $i \in V$ appearing as an endpoint of some edge in $E_L$ occurs at most once as origin ---i.e., $(i,j)$, and at most once as destination ---i.e., $(j,i)$. The algorithm performs the following steps:

\textbf{1.} For each $i \in V$ such that $\din(i) \ge Q$, $Q$ edges are drawn at random without replacement from $\Nin(i)$, and their are labels queried. Let
$
	\hdeltain(i) = \hdin^+(i)/Q
$,
where $\hdin^+(i)$ is the number of positive edges sampled from $\Nin(i)$;

\textbf{2.} For each $i \in V$ such that $\dout(i) \ge Q$, $Q$ edges are drawn at random without replacement from $\Nout(i)$ and their labels are queried. Let
$
	\hdeltaout(i) = \hdout^+(i)/Q
$,
where $\hdout^+(i)$ is the number of positive edges sampled from $\Nout(i)$;

\textbf{3.} Sample any edge in $E_L$ not yet sampled, and let $\tauhat$ be the fraction of positive edges in $E_L$;

\textbf{4.} Any remaining non-sampled edge $(i,j)$ is predicted as
$
	\yhat(i,j) = \sgn\big(\hdeltaout(i) + \hdeltain(j) - \tfrac{1}{2} - \tauhat \big)
$.
***********************************************************************
\fi

Our algorithm works under the assumption that for given parameters $Q$ (a positive integer) and $\alpha \in (0,1)$ there exists a set\footnote
{
$E_L$ is needed to find an estimate $\tauhat$ of $\tau$ in~(\ref{eq:predictor}) ---see Step~3 of the algorithm. Any undirected matching of $G$ of size $\mathcal{O}(\log|V|)$ can be used. In practice, however, we never computed $E_L$, and estimated $\tau$ on the entire training set $E_0$.
}
$E_L \ss E$ of size $\tfrac{2Q}{\alpha}$ where each vertex $i \in V$ appearing as an endpoint of some edge in $E_L$ occurs at most once as origin ---i.e., $(i,j)$--- and at most once as destination ---i.e., $(j,i)$. Moreover, we assume $E_0$ has been drawn from $E$ at random {\em without} replacement, with $m = |E_0| = \alpha\,|E|$. The algorithm performs the following steps:

%\textbf{1.} Draw a subset $E_0$ of $m = \alpha|E|$ edges from $E$ at random {\em without} replacement, and observe the corresponding labels.

\begin{enumerate}[leftmargin=*,nosep,label=\textbf{\arabic*.}]
	\item For each $j \in V$, let
		$
		\hun(j) = \hdin^-(j)/\hdin(j)
		$,
		i.e., the fraction of negative edges found in $\Nin(j) \cap E_0$.

	\item For each $i \in V$, let
		$
		\htr(i) = \hdout^-(i)/\hdout(i)
		$,
		i.e., the fraction of negative edges found in $\Nout(i) \cap E_0$.

	\item Let $\tauhat$ be the fraction of positive edges in $E_L\cap E_0$.

	\item Any remaining edge $(i,j) \in E\setminus E_0$ is predicted as
		$
		\yhat(i,j) = \sgn\Big(\big(1-\htr(i)\big) + \big(1-\hun(j)\big) - \tfrac{1}{2} -\tauhat\Big)
		$.
\end{enumerate}

The next result\footnote
{
All proofs are in the supplementary material.
} 
shows that if the graph is not too sparse, then the above algorithm can approximate the Bayes optimal predictor on nodes whose in-degree and out-degree is not too small.
%
% Let $n = |V|$. Given a function $f\,:\, \Nat \rightarrow \R$, such that $f(n) \rightarrow \infty$ as $n \rightarrow \infty$ (e.g., $f(n) = \ln n$) and a threshold $\theta = \frac{1}{n\,f(n)}$, let $\kintheta$ be the number of nodes $i \in V$ with $\din(i) \geq \theta$, and $\kouttheta$ be the number of nodes $i \in V$ with $\dout(i) \geq \theta$.
%

\begin{theorem}
\label{t:active}
Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the model in Section~\ref{s:gen}.
%, and let $E_0$ of $m = |E_0| = \alpha|E|$. 
If the algorithm is run with parameter $Q = \Omega(\ln|V|)$, and $\alpha \in (0,1)$ such that the above assumptions are satisfied, then $\yhat(i,j) = y^*(i,j)$ holds with high probability simultaneously for all test edges $(i,j) \in E$ such that $\dout(i),\din(j) = \Omega(\ln|V|)$, and $\eta(i,j) = \Pr(y_{i,j}=1)$ is bounded away from $\tfrac{1}{2}$. 
%The overall number of training edges is $\Theta\big(|V|\ln|V|\big)$.
\end{theorem}
%
% While learning the Bayes optimal predictor is alluring, in practice

%NCB The approach leading to Theorem~\ref{t:active} needs the graph to be sufficiently dense and the bias $\tau$ to be the same for all edges. In order to address these limitations, we now introduce a second algorithm.

%\textcolor{blue}{
% The approach leading to Theorem~\ref{t:active} needs the graph to be sufficiently dense and the bias $\tau$ to be the same for all edges. In order to address these limitations, we now introduce a second method, which is used to derive the \usrule\ algorithm assessed in our experiments of Section~\ref{s:exp}.%}
The approach leading to Theorem~\ref{t:active} lets us derive the \usrule{}
algorithm assessed in our experiments of Section~\ref{s:exp}, but it needs the graph to be
sufficiently dense and the bias $\tau$ to be the same for all edges. In order to
address these limitations, we now introduce a second method based on label
propagation.

% From this algorithm we derive the \textsc{blc} algorithm we use in our
% experiments of Section~\ref{s:exp}.
%we show that \textsc{blc} works well even in a passive learning setting.

%{\bf CG: also pointer to logistic regression from here?}

\subsection{Approximation to Maximum Likelihood via Label Propagation}\label{ss:passive}

For simplicity, assume the joint prior distribution $\mu(p,q)$ is uniform over $[0,1]^2$ with independent marginals, and suppose that we draw at random without replacement the training set $E_0 = \big((i_1,j_1),y_{i_1,j_1}), ((i_2,j_2),y_{i_2,j_2}), \ldots, ((i_m,j_m),y_{i_m,j_m}\big)$, with $m = |E_0|$. Then a reasonable approach to approximate $y^*(i,j)$ would be to resort to a maximum likelihood estimator of the parameters $\{p_i, q_i\}_{i=1}^{|V|}$ based on $E_0$.
% Working out the math (details in the supplementary material), one can easily see that
As showed in the supplementary material,
the gradient of the log-likelihood function w.r.t.\ $\{p_i, q_i\}_{i=1}^{|V|}$ satisfies
%
%
\begin{small}
\begin{align}
&\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial p_{\ell}}\label{e:mlp}\\
&\ \ = 
\sum_{k=1}^m
\frac{\Ind{i_k = \ell,y_{\ell,j_k}=+1}}{p_{\ell}+q_{j_k}}\, 
- \sum_{k=1}^m
\frac{\Ind{i_k = \ell,y_{\ell,j_k}=-1}}{2-p_{\ell}-q_{j_k}}\,,\notag
\end{align}
%
\begin{align}
&\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial q_{\ell}}\label{e:mlq}\\
&\ \ = 
\sum_{k=1}^m
\frac{\Ind{j_k = \ell,y_{i_k,\ell}=+1}}{p_{i_k}+q_{\ell}}\, 
- \sum_{k=1}^m
\frac{\Ind{j_k = \ell,y_{i_k,\ell}=-1}}{2-p_{i_k}-q_{\ell}}\,,\notag
\end{align}
\end{small}
%
where $\Ind{\cdot}$ is the indicator function of the event at argument.
Unfortunately, equating~(\ref{e:mlp}) and~(\ref{e:mlq}) to zero, and solving for parameters $\{p_i, q_i\}_{i=1}^{|V|}$ gives rise to a hard set of nonlinear equations.
Moreover, some such parameters may never occur in these equations, namely whenever $\Nout(i)$ or $\Nin(j)$ are not represented in $E_0$ for some $i,j\in V$. 
%Specifically, if $E_0 \cap \Nout(i) = \emptyset$ then $p_i$ does not occur, and if $E_0 \cap \Nin(j) = \emptyset$ then $q_j$ does not.
%
Our \emph{first approximation} is therefore to replace the nonlinear equations resulting from~(\ref{e:mlp}) and~(\ref{e:mlq}) by the following set of linear equations\footnote{Details are provided in the supplementary material.}, one for each $\ell \in V$:
%
\begin{align*}
\sum_{k=1}^m &\Ind{i_k = \ell,y_{\ell,j_k}=+1} \left(2-p_{\ell}-q_{j_k}\right)\\
=
&\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=-1}
(p_{\ell}+q_{j_k})%, \quad \ell = 1, \ldots, |V|
\end{align*}
%
%\ncb{Claudio, che significa la doppia somma qui sotto?}
% CG: un typo...
%
\begin{align*}
\sum_{k=1}^m &\Ind{j_k = \ell,y_{i_k,\ell}=+1} \left(2-p_{i_k}-q_{\ell}\right)\\
=
&\sum_{k=1}^m  \Ind{j_k = \ell,y_{i_k,\ell}=-1}
\left(p_{i_k}+q_{\ell}\right)~.%,\ \ \ \ell = 1, \ldots, |V|\,,
\end{align*}
%
%
%
\iffalse
************************************
Denoting for brevity ${\wh \dout(i)}$ and  ${\wh \din(i)}$ the out-degre and the in-degree 
of node $i$ over the training set, these can be rewritten as
% 
\begin{align*}
\sum_{k=1\,:\,i_k = \ell}^m \left(p_{\ell}+q_{j_k}\right) 
&= 
{\wh \dout(\ell)} + \sum_{k=1\,:\,i_k = \ell} y_{\ell,j_k}, \qquad \ell = 1, \ldots, |V|~,\\
\sum_{k=1\,:\,j_k = \ell}^m \left(p_{i_k}+q_{\ell}\right) 
&= 
{\wh \din(\ell)} + \sum_{k=1\,:\,j_k = \ell} y_{i_k,\ell}, \qquad \ell = 1, \ldots, |V|\,,
\end{align*}
%
************************************
\fi
The solution to these equations are precisely the points where the gradient w.r.t.\ $(\bp,\bq) =\{p_i, q_i\}_{i=1}^{|V|}$ of the quadratic function
\[
f_{E_0}(\bp,\bq) = \sum_{(i,j) \in E_0} \left(\frac{1+y_{i,j}}{2} - \frac{p_i+q_j}{2} \right)^2
\]
vanishes.
We follow a label propagation approach by adding to $f_{E_0}$ the corresponding test set function $f_{E\setminus E_0}$, and treat the sum of the two as the function to be minimized during training w.r.t.\ both $(\bp,\bq)$ and all $y_{i,j} \in [-1,+1]$ for $(i,j)\in E\setminus E_0$, i.e.,
%
\begin{equation}\label{e:quadratic}
\min_{(\bp,\bq), y_{i,j} \in [-1,+1],\,(i,j)\in E\setminus E_0} 
\left(f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\right)\,.
\end{equation}
%
Binary $\pm 1$ predictions on the test set $E\setminus E_0$ are then obtained by thresholding the obtained values $y_{i,j}$ at $0$.

\iffalse
********************************************************
and then solve for parameters $\{p_i, q_i\}_{i=1}^{|V|}$.
Yet, it may well be the case that some such parameters never occur in all these equations.\footnote
{
On the other hand, recall that no pairing $(i,j) \in E$ can occur more than once here, since $E_0$ is sampled without replacement.
}
This will happen precisely whenever $\Nout(i)$ or $\Nin(j)$ are not represented in $E_0$. Specifically, if $E_0 \cap \Nout(i) = \emptyset$ then $p_i$ does not occur, and if $E_0 \cap \Nin(j) = \emptyset$ then $q_j$ does not occur. Hence, for each unsampled edge $(\ell,j)\in E\setminus E_0$, we add to~(\ref{e:mlp2}) the equations
\[
p_{\ell}+q_j = 1+y_{\ell,j}~,
\]
motivated by the fact that $\E \left[\frac{1+y_{\ell,j}}{2}\,|\, (\ell,j)\right] =  \frac{p_{\ell}+q_j}{2}$. Similarly, we add to~(\ref{e:mlq2}) the equations
\[
p_i+q_{\ell} = 1+y_{i,\ell}~.
\]
This gives rise to the following set of equations
% 
\begin{align}
p_{i}+ \frac{1}{\dout(i)}\,\sum_{j \in \Nout(i)} q_{j}
&= 
\frac{1}{\dout(i)}\,\sum_{j=1}^{|V|} (1+y_{i,j}), \qquad i = 1, \ldots, |V|~,\label{e:mlpa}\\
q_{j} +  \frac{1}{\din(j)}\,\sum_{i \in \Nin(j)} p_{i}
&= 
\frac{1}{\din(j)}\,\sum_{i=1}^{|V|} (1+y_{i,j}), \qquad j = 1, \ldots, |V|\,,\label{e:mlqa}\\
1+y_{i,j} &= p_i+q_{j},\qquad\qquad\qquad\quad\,\, (i,j) \in E\setminus E_0~.\label{e:mlya}
\end{align}
%
*******************************************************
\fi

%\textcolor{red}{
We now proceed to solve \eqref{e:quadratic} via label propagation~\cite{LabelPropa03} on the graph $G''$ obtained through the $G \rightarrow G''$ reduction of Section~\ref{s:prel}.%} 
However, because of the presence of negative edge weights in $G''$, we first have to symmetrize\footnote{%
While we note here that such linear transformation of the variables does not change the problem, we provide more details in Section~1.3 of the supplementary material.} variables $p_i, q_i, y_{i,j}$ so as they all lie in the interval $[-1,+1]$.
After this step, one can 
%easily 
see that, once we get back to the original variables, label propagation computes the harmonic solution minimizing the function
%
%\hspace*{-0.01cm}\vbox{%
\begin{align*}
{\widehat f}&\big(\bp,\bq,{y_{i,j}}_{(i,j) \in E\setminus E_0}\big)
%&=
%\sum_{(i,j) \in E}
%\Biggl(
%\frac{1}{2}\left(\frac{1+y_{i,j}}{2}-p_i\right)^2 + \frac{1}{2}\left(\frac{1+y_{i,j}}{2}-q_j\right)^2 \\
%&\qquad\qquad+ \left(\frac{p_i+q_j-1}{2}\right)^2\Biggl)\\
= f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\\ 
& \qquad + \frac{1}{2}\sum_{i\in V} \left(\dout(i)\Bigl(p_i-\frac{1}{2}\Bigl)^2+\din(i)\Bigl(q_i-\frac{1}{2}\Bigl)^2\right)\,.
\end{align*}
%}
%
The function ${\widehat f}$ is thus a regularized version of the target function $f_{E_0} + f_{E\setminus E_0}$ in (\ref{e:quadratic}), where the regularization term 
%$\sum_{i\in V} \big(\dout(i)r(p_i)+\din(i)r(q_i)\big)$ 
tries to enforce the extra constraint that whenever a node $i$ has a high out-degree then the corresponding $p_i$ should be close to $1/2$. Thus, on any edge $(i,j)$ departing from $i$, the Bayes optimal predictor $y^*(i,j) = \sgn(p_i+q_j-1)$ will mainly depend on $q_j$ being larger or smaller than $\tfrac{1}{2}$ (assuming $j$ has small in-degree). Similarly, if $i$ has a high in-degree, then the corresponding $q_i$ should be close to $1/2$ implying that on any edge $(j,i)$ arriving at $i$ the Bayes optimal predictor $y^*(j,i)$ will mainly depend on $p_j$ (assuming $j$ has small out-degree). Put differently, a node having a huge out-neighborhood makes each outgoing edge ``count less" than a node having only a small number of outgoing edges, and similarly for in-neighborhoods.
%
%
\iffalse
trying to establish the relative contribution of $p_i$ and $q_j$ to $\Pr(y_{i,j}=1) = \frac{p_i+q_j}{2}$  as follows: If $(i,j)$ is such that $i$ has high outdegree, then $q_j$ is more relevant than $p_i$; symmetrically, if  $(i,j)$ is such that $j$ has high indegree, then $p_i$ is more relevant than $q_j$.
%when both are viewed as functions of $\bp$, $\bq$, and $y_{i,j} \in [-1,+1]$, for $(i,j)\in E\setminus E_0$. 
\fi
%
%
The label propagation algorithm operating on $G''$ does so (see again Figure \ref{f:etnr} (c)) by iteratively updating as follows:
%
\begin{align*}
	p_{i}  & \leftarrow \frac{-\sum_{j \in \NNout(i)} q_{j} + \sum_{j \in \NNout(i)} (1+y_{i,j})}{3\,\dout(i)}\,\quad\forall i\in V\\ 
	q_{j}  & \leftarrow \frac{-\sum_{i \in  \NNin(j)} p_{i} + \sum_{i \in \NNin(j) } (1+y_{i,j})}{3\,\din(j)}\qquad\forall j \in V\\
  y_{i,j} & \leftarrow \frac{p_i + q_j}{2}~ \quad \forall (i,j) \in E\setminus E_0~.
\end{align*}
%
The algorithm is guaranteed to converge~\cite{LabelPropa03} to the minimizer of ${\widehat f}$. Notice that the presence of negative weights on the edges of $G''$ does not prevent label propagation from converging. 
%Indeed, any node classification algorithm handling both positive and negative weights on the edges of $G''$ could be used instead of label propagation. One alternative would be the {\textsc wta} algorithm from \cite{}.
This is the algorithm we will be championing in our experiments of Section~\ref{s:exp}.

% \footnote
% {
% Notice that in the case of negative weights $w_{i,j}$ over the edges, the label propagation update should read as
% \(
% y_{i,j} = \frac{\sum_{j \in \Nout(i)} w_{i,j} y_{i,j}}{\sum_{j \in \Nout(i)} |w_{i,j}|}~.
% \)
% }
%

%A computation-friendly alternative is to follow a {\em method of moments}-like approach. We start off from the statistics $\{1-\htr(i),1-\hun(i)\}_{i=1}^{|V|}$, and observe that, even if we are sampling $E_0$ without replacement, $1-\htr(i)$ is an unbiased estimator of (\ref{e:pout}), and $1-\hun(i)$ is an unbiased estimator of (\ref{e:pin}).

\label{ssec:further_related}
{\bf Further related work.} The vast majority of existing edge sign prediction algorithms for directed graphs are based on the computation of local features of the graph. These features are evaluated on the subgraph induced by the training edges, and the resulting values are used to train a supervised classification algorithm (e.g., logistic regression). The most basic set of local features used to classify a given edge $(i,j)$ are defined by 
% the neighbourhoods\footnote
% {
% We abuse notation here, and define the quantities $\din^+(j),\din^-(j),\dout^+(i),\dout^-(i)$ on the subgraph induced by the training edges.
% }
$\din^+(j),\din^-(j),\dout^+(i),\dout^-(i)$ computed over the training set $E_0$, and by the embeddedness coefficient $\big|\Nout(i) \cap \Nin(j)\big|$. In turn, these can be used to define more complicated features, such as
\(
	\frac{\din^+(j) + |E^+|\uin(j)}{\din(j) + \uin(j)}
\quad\text{and}\quad
	\frac{\dout^+(i) + p^+\uout(i)}{\dout(i) + \uout(i)}
\)
introduced in~\cite{Bayesian15}, together with their negative counterparts, where $|E^+|$ is the overall fraction of positive edges, and $\uin(j),\uout(i)$ are, respectively, the number of test edges outgoing from $i$ and the number of test edges incoming to $j$.
Other types of features are derived from social status theory~(e.g., \cite{Leskovec2010}), and involve the so-called triads; namely, the triangles formed by $(i,j)$ together with $(i,w)$ and $(w,j)$ for any $w \in \NNout(i) \cap \NNin(j)$. A third group of features is based on node ranking scores. These scores are computed using a variery of methods, including Prestige~\cite{zolfaghar2010mining}, exponential ranking~\cite{traag2010exponential}, PageTrust~\cite{de2008pagetrust}, Bias and Deserve~\cite{mishra2011finding}, TrollTrust~\cite{wu2016troll}, and generalizations of PageRank and HITS to signed networks~\cite{shahriari2014ranking}. Examples of features using such scores are \textsl{reputation} and \textsl{optimism}~\cite{shahriari2014ranking}, defined for a node $i$ by
\(
	\frac{\sum_{j \in \NNin(i)} y_{j,i}\sigma(j)}{\sum_{j \in \NNin(i)} \sigma(j)}
\quad\text{and}\quad
	\frac{\sum_{j \in \NNout(i)} Y_{i,j}\sigma(j)}{\sum_{j \in \NNout(i)} \sigma(j)}\,,
\)
where $\sigma(j)$ is the ranking score assigned to node $j$. Some of these algorithms will be used as representative competitors in our experimental study of Section~\ref{s:exp}.

### 2017-09-02 troll/algs_online.tex
initial import of troll paper (with citation fix)

\section{Algorithms in the Online Setting}\label{s:algonline}

For the online scenario, we have the following result.
%
\begin{theorem}\label{t:online}
There exists a randomized online prediction algorithm $A$ whose expected number of mistakes satisfies
$
\E M_A(Y) = \Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V| \right)
$
on any edge-labeled graph $G(Y) = (V,E(Y))$.
\end{theorem}
%
The algorithm used in Theorem~\ref{t:online} is a combination of randomized Weighted Majority instances. Details are reported in the supplementary material. We complement the above result by providing a mistake lower bound. Like Theorem~\ref{t:online}, the following result holds for all graphs, and for all label irregularity levels $\Psi_G(Y)$.
%
\begin{theorem}
\label{t:mistake_bound}
Given any edge-labeled graph $G(Y) = (V,E(Y))$ and any integer $K \le \big\lfloor \tfrac{|E|}{2}\big\rfloor$, a randomized labeling $Y\in\spin^{|E|}$ exists such that $\Psi_G(Y) \leq K$, and the expected number of mistakes that any online algorithm $A$ can be forced to make satisfies
\(
\E M_A(Y) \ge \frac{K}{2}\,.
\)
Moreover, as $\frac{K}{|E|} \rightarrow 0$ then
\(
\E M_A(Y) = K
\).
\end{theorem}
%

%\footnote
%{
%In fact, we can refine the above statement by proving that, as $\frac{K}{|E|} \rightarrow 0$, the lower bound gets arbitrarily close to $K$ for any $G(Y)$, hence asymptotically matching the upper bound of Theorem~\ref{t:online} (\textit{proof in the supplementary material}).
%}

### 2017-09-02 troll/conclusions.tex
initial import of troll paper (with citation fix)

\section{Conclusions and Ongoing Research}
%
%FV
We have studied the edge sign prediction problem in directed graphs in both batch and online learning settings. In both cases, the underlying modeling assumption hinges on the trollness and (un)trustworthiness predictive features. We have introduced a simple generative model for the edge labels 
%FV that serves both as a theoretical justification for many successful heuristic methods to edge sign prediction and 
%FV as a way 
to craft this problem as a node sign prediction problem to be efficiently tackled by standard Label Propagation algorithms. 
%FV We then observed that the two features used
%FV notion of trollness and trustworthiness
%FV naturally lend themselves to define a notion of edge sign regularity that can also be analyzed in an (adversarial) online setting. In this setting, 
%FV 
Furthermore, we have studied the problem in an (adversarial) online setting providing 
upper and (almost matching) lower bounds on the expected number of prediction mistakes.

%FV Finally, we complemented the above with experiments on five real-world datasets in the small training set regime.

%FV 
Finally, we validated our theoretical results by experimentally assessing our methods on five real-world datasets in the small training set regime. 
Two interesting conclusions from our experiments are: i. Our generative model is robust, for it produces Bayes optimal predictors which tend to be empirically best also within the larger set of models that includes all logistic regressors based on trollness and trustworthiness alone; ii. 
%We showed that 
our methods are in practice either strictly better than their competitors in terms of prediction quality or, when they are not, they are faster.
We are currently engaged in 
%FV gathering more experimental evidence, as well as in 
extending our approach so as to incorporate further predictive features (e.g., side information, when available).

### 2017-09-02 troll/experiments.tex
initial import of troll paper (with citation fix)

\section{Experimental Analysis}\label{s:exp}
We now evaluate our edge sign classification methods on representative real-world datasets of varying density and label regularity,
% within the batch settings,
showing that our methods compete well against existing approaches in terms of both predictive and computational performance. We are especially interested in small training set regimes, and have restricted our comparison to the batch learning scenario since all competing methods we are aware of have been developed in that setting only.

{\bf Datasets.} We considered five real-world classification datasets. The first three are directed \ssn{} widely used as benchmarks for this task~(e.g.,\cite{Leskovec2010,shahriari2014ranking,wu2016troll}): In \wik{}, there is an edge from user $i$ to user $j$ if $j$ applies for an admin position and $i$ votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $i$ can tag other members $j$ as friends or foes. Finally, in \epi{}, an online shopping website, user $j$ reviews products and, based on these reviews, another user $i$ can display whether he considers $j$ to be reliable or not. In addition to these three datasets, we considered two other \ssn{} where the signs are inferred automatically.
%rather than given explicitly by the users.
In \kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $i$ to user $j$ indicates whether they edited the same article in a constructive manner or not.\footnote
{
This is the \href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of
the \enquote{Wikisigned} dataset}, from which we removed self-loops.
}
Finally, in the \aut{}~\cite{kumar2016structure} network, an author $i$ cites another author $j$ by either endorsing or criticizing $j$'s work. The edge sign is derived by classifying the citation sentiment with a simple, yet powerful, keyword-based technique using a list of positive and negative words. See \cite{kumar2016structure} for more details.\footnote{
We again removed self-loops and merged multi-edges which are all of the same sign.
}

\begin{table*}[bt]
  \centering
  \small
  \caption{Dataset properties. The 5th column gives the fraction of positive labels. The last two columns provide two different measures of label regularity ---see main text.\label{tab:dataset}}
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Dataset &       $|V|$ &       $|E|$ &$\frac{|E|}{|V|}$ & $\frac{|E^+|}{|E|}$ & $\frac{\Psi^2_{G''}(Y)}{|E|}$ & $\frac{\Psi_G(Y)}{|E|}$ \\ % & twin edges & twin disagreement \\
    \midrule
    \aut{} &   \np{4831} &  \np{39452} & 8.1  &              72.33\% &                    .076 &                       .191 \\ %&      5.07\% &            27.10\% \\
    \wik{} &   \np{7114} & \np{103108} & 14.5 &              78.79\% &                    .063 &                       .142 \\ %&      5.56\% &            10.04\% \\
    \sla{} &  \np{82140} & \np{549202} & 6.7  &              77.40\% &                    .059 &                       .143 \\ %&     17.74\% &             4.00\% \\
    \kiw{} & \np{138587} & \np{740106} & 5.3  &              87.89\% &                    .034 &                       .086 \\ %&      6.55\% &             2.09\% \\
    \epi{} & \np{131580} & \np{840799} & 6.4  &              85.29\% &                    .031 &                       .074 \\ %&     30.83\% &            14.64\% \\
    \bottomrule
  \end{tabular}
\end{table*}

\autoref{tab:dataset} summarizes statistics for these datasets. We note that most edge labels are positive. Hence, test set accuracy is not an appropriate measure of prediction performance. We instead evaluated our performance using the so-called Matthews Correlation Coefficient (MCC)~(e.g., \cite{MCC00}), defined as
%\footnote
%{
%For completeness, we also report accuracy results in the supplementary material.
%}  
\[
	\mathrm{MCC} = \frac{tp\times tn-fp\times fn} {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } }\,.
			% = \pm \sqrt{\frac{\chi^2}{n}}
\]
MCC combines all the four quantities found in a binary confusion matrix ($t$rue $p$ositive, $t$rue $n$egative, $f$alse $p$ositive and $f$alse $n$egative) into a single metric which ranges from $-1$ (when all predictions are incorrect) to $+1$ (when all predictions are correct).% through $0$ (when predictions are made uniformly at random).

Although the semantics of the edge signs is not the same across these networks, we can see from \autoref{tab:dataset} that our generative model essentially fits all of them. Specifically, the last two columns of the table report the rate of label (ir)regularity, as measured by $\Psi^2_{G''}(Y)/|E|$ (second-last column) and $\Psi_{G}(Y)/|E|$ (last column), where 
\[
\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \left(f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\right)\,,
\]
%$\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} f_{E}(\bp,\bq)$, 
%
$f_{E_0}$ and $f_{E\setminus E_0}$ being the quadratic criterions of Section~\ref{ss:passive}, viewed as functions of both $(\bp,\bq)$, and $y_{i,j}$,
% when all labels are known,
\iffalse
*****************
\[
\Psi_{G''}(Y) = \min_{\bp,\bq\in[0,1]^{|V|}}
\sum_{(i,j) \in E} \left(\frac{1+y_{i,j}}{2} - \frac{p_i+q_j}{2}\right)^2\,,
\]
i.e., the quadratic criterion $f_{E}(\bp,\bq)$ of Section~\ref{ss:passive} when all labels are known,
*****************
\fi
%essentially hinges upon, 
and $\Psi_{G}(Y)$ is the label regularity measure adopted in the online setting, as defined in Section~\ref{s:prel}. 
%The two measures have been crafted here so as to be on the same scale, hence they can be contrasted to each other on the same dataset. 
It is reasonable to expect that higher label irregularity corresponds to lower prediction performance. This trend is in fact confirmed by our experimental findings: whereas \epi{} tends to be easy, \aut{} tends to be hard, and this holds for all algorithms we tested, even if they do not explicitly comply with our inductive bias principles. Moreover, $\Psi^2_{G''}(Y)/|E|$ tends to be proportional to $\Psi_{G}(Y)/|E|$ across datasets, hence confirming the anticipated connection between the two regularity measures.

%
%Another notable feature of our datasets is the relative regularity of their
%labeling with respect to our two complexity measures defined in
%\autoref{s:analysis}
%(here  where the minimum
%is found using the L-BFGS-B algorithm)
%(especially in \epi{} case, which is confirmed by higher performance).
%

\iffalse
***********************************************************************
Finally, there is a low fraction of reciprocal (or \emph{twin}) edges (i.e.\
$\eij \in E$ and $\eji \in E$, which is a common mechanism of link formation in
directed networks~\cite{DirectedReciprocity04}), and in most cases they do not
disagree, meaning they have the same sign. In practice, we use that fact to
improve our accuracy at no additional computational cost: when predicting $\eij
\in E_{\mathrm{test}}$, if the reciprocal edge \eji{} is part of the training set, we
set $\sgn(\eij) = \sgn(\eji)$.

from http://www.nature.com/articles/srep02729

The study of link reciprocity in binary directed networks1,2, or the tendency
of vertex pairs to form mutual connections, has received an increasing
attention in recent years3,4,5,6,7,8,9,10,11,12,13,14. Among other things,
reciprocity has been shown to be crucial in order to classify3 and model4
directed networks, understand the effects of network structure on dynamical
processes (e.g. diffusion or percolation processes5,6,7), explain patterns of
growth in out-of-equilibrium networks (as in the case of the Wikipedia8 or the
World Trade Web9,10), and study the onset of higher-order structures such as
correlations11,12 and triadic motifs13,14,15,16
************************************************************************
\fi

{\bf Algorithms and parameter tuning.} 
We compared the following algorithms:
%
%\begin{itemize}%[nosep,leftmargin=2em]
%
%\item

{\bf 1.} The label propagation algorithm of Section~\ref{ss:passive} (referred to as \uslpropGsec{}). The actual binarizing threshold was set by cross-validation on the training set.
%
%\item

{\bf 2.} The algorithm analyzed at the beginning of Section~\ref{s:algbatch}, which we call \usrule{} (Bayes Learning Classifier based on {\em tr}ollness and {\em un}trustworthiness). After computing $\htr(i)$ and $\hun(i)$ on training set $E_0$ for all $i \in V$ (or setting those values to $\frac{1}{2}$ in case there is no outgoing or incoming edges for some node), we use Eq.~\eqref{eq:predictor} and estimate $\tau$ on $E_0$.
%
%\item

{\bf 3.} A logistic regression model where each edge $(i,j)$ is associated with the features $[1-\htr(i), 1-\hun(j)]$ computed again on $E_0$ (we call this method \uslogregp{}). Best binary thresholding is again computed on $E_0$. Experimenting with this logistic model serves to support the claim we made in the introduction that our generative model in Section~\ref{s:gen} is a good fit for the data.

%
%\item 
{\bf 4.} The solution obtained by directly solving the unregularized problem (\ref{e:quadratic}) through a fast constrained minimization algorithm (referred to as \qoptim{}). Again, the actual binarizing threshold was set by cross-validation on the training set.\footnote
{
We have also tried to minimize (\ref{e:quadratic}) by removing the $[-1,+1]$ constraints, but got similar MCC results
as the ones we report for \qoptim{}
}
%

{\bf 5.} The matrix completion method from~\cite{LowRankCompletion14} based on \complowrank{} matrix factorization. Since the authors showed their method to be robust to the choice of the rank parameter $k$, we picked $k=7$ in our experiments.
%
%\item

{\bf 6.} A logistic regression model built on \comptriads{} features derived from status theory~\cite{Leskovec2010}.
%
%\item

{\bf 7.} The PageRank-inspired algorithm from \cite{wu2016troll}, where a recursive notion of trollness is computed by solving a suitable set of nonlinear equations through an iterative method, and then used to assign ranking scores to nodes, from which (un)trustworthiness features are finally extracted for each edge. We call this method \compranknodes{}. As for hyperparameter tuning ($\beta$ and $\lambda_1$ in~\cite{wu2016troll}), we closely followed the authors' suggestion of doing cross validation.
%
%\item

{\bf 8.} The last competitor is the logistic regression model whose features have been build according to \cite{Bayesian15}. We call this method \compbayesian{}.
%\end{itemize}

The above methods can be roughly divided into {\em local} and {\em global} methods. A local method hinges on building local predictive features, based on neighborhoods: \usrule{}, \uslogregp{}, \comptriads{}, and \compbayesian{} essentially fall into this category. The remaining methods 
%(\uslpropGsec{}, \complowrank{}, and \compranknodes{}) 
are global in that their features are designed to depend on global properties of the graph topology.

\iffalse
***************************************************
%Given a training set \trainset{}, we exploit the revealed signs in two ways.
First, with label propagation (referred to as \uslprop{} in the following). We
initialize the labels vector $f_0$ to $|V'|$ random values drawn uniformly in
$[0, 1]$ and set the label of the training edges equal to their observed value.
% TODO: starting with zeros gives lower MCC and accuracy after a fixed number
% of iterations
Let the adjacency matrix of $G'$ be $A$, its diagonal degree matrix
be $D$, its diameter be $\diam{G'}$ and $P$ be the following sparse matrix $P=D^{-1}A$. This allow performing
one round of label propagation as $f_{t+1} = Pf_t$, followed by clamping the
training labels. Each round thus involves $2|E|$ multiplications and we do
$\diam{G'}$ of them\footnote{$\diam{G'}$ is a small constant, ranging from $16$
for \aut{} to $38$ for \epi{}}. With the computed labels, we can associate each
edge $\eij$ with $a_{i,j} = \nicefrac{1}{2}\left(f_{\diam{G'}}(\iout) +
f_{\diam{G'}}(\jin)\right)$.
% This is what I do in the code but now I realize this is just an extra
% propagation round only on the square node so maybe simplify the description
% (actually no, it's needed for ERM)
By sorting the $a$ values of the training edges, we can find the threshold $t$
that minimizes the number of mistakes on the training set while predicting
$\yij = \sgn\left(a_{i,j} - t\right)$ in $O(|\trainset{}|\log |\trainset{}|)$
time.

Then we again associate each
edge $\eij$ with $a_{i,j} = \left(1-\htr(i)\right) + \left(1-\hun(j)\right)$
and find the $\tau$ of \eqref{eq:predictor} that minimizes the empirical risk.
Another way to exploit those quantities is to
train a Logistic Regression model where each edge is associated with two
features: $[1-\htr(i), 1-\hun(j)]$ (we call this method \uslogregp{}).

We compare these methods with four competitors, each exploiting the sign
information and network topology in a different way:

Looking at undirected graphs with a global perspective,
\citet{LowRankCompletion14} consider the observed adjacency matrix $A$, made of
the edges in \trainset{}, as a noisy sampling of the adjacency matrix
$A^\star$ of an underlying complete graph satisfying the weakly balance
condition (that is with no triangle containing only one negative edge).
%TODO the previous sentence is too long
This condition implies the existence of a small number $k$ of node clusters
with positive links within clusters and negative links across clusters, which
in turn implies $\rank A^\star=k$.
By recovering a complete matrix $\tilde{A}$ which matches the non-zeros entries
of $A$, it is possible to predict the sign of $\eij \notin \trainset$
as $\yhat_{i,j} = \sgn(\tilde{A}_{i,j})$. Although the exact version of this
problem is NP-hard, authors assume that $k$ is an hyperparameter known
beforehand\footnote{The method is showed to be robust to the choice of its
value, thus we pick $k=7$.} and look for two matrices $W,H\in
\mathbb{R}^{k\times|V|}$ which minimise a sigmoidal reconstruction loss with
$A$, subject to a regularization term. We refer to this method as
\emph{\complowrank{}}.

The status theory heuristic posits that a positive link from $i$ to $j$ denotes
that user $i$ considers user $j$ as having a higher status or skill than
himself~\cite{Leskovec2010}. It implies that among the 16 possible triads of
three nodes with two directed signed edges among them, some must be more
represented than others. The \emph{\comptriads{}} method exploits this fact by
counting for each edge in the training set how frequently it is involved in
each of the 16 triad types. Based on these edge features and 7 degree features
of each edge endpoints, a Logistic Regression model is trained and used to
predict the sign of the test edges.

Inspired by the PageRank algorithm, \citet{wu2016troll} propose a recursive
definition of trollness based on the opinion of one node's neighbors weighted
by their own trollness, which allow to assign a trustworthiness $\pi(i)$ to
each node $i$ through a set of non linear equations solved by an iterative
method. This $\pi$ values are used to rank the nodes by computing their
Reputation and Optimism scores, thus providing four features for each edge,
which are in turn used to train a Logistic Regression model for the
classification task. We refer to this method as \emph{\compranknodes{}}.
Computing $\pi$ requires to choose two hyperparameters of the model, $\beta$
and $\lambda_1$ and the authors suggest to do it by holding out some training
data to perform cross validation but in practice we found little difference in
performance in our implementation while it consumes a lot of time.

The last competitors is also local. \citet{Bayesian15} note that  a node can
belong to one of the 16 types based of whether is number of its positive (resp.
negative) outgoing (resp. incoming) edges is zero or not.  The number of
unobserved incoming and outgoing edges of each node $i$ let us define its
16-dimensional vector $V_i$ containing the probability of transitioning to any
other state.  Then for each edge \eij we compute the outer product of $V_i$
with $V_j$ and also include additional features such as triads count and degree
information before training a Logistic Regression model. We refer to this
method as \emph{\compbayesian{}}.

***************************************************
\fi

\iffalse
****************************************************
\begin{table*}[t]
  \centering
\setlength{\tabcolsep}{3pt}
\scriptsize
% \begin{longtable}{lrccc|cccc}
\caption{MCC with increasing training set size, with one standard deviation over 12 random sampling of $\trainset$. The last four columns refer to the methods we took from the literature. For the sake of readability, we multiplied all MCC values by 100. The best number in each row is highlighted in \textbf{\textcolor{brown}{bold brown}} and the second one in \textit{\textcolor{red}{italic red}}. If the difference is statistically significant ($p$-value of a paired Student's $t$-test less than $0.005$), the best score is underlined. The ``time" rows contain the time taken to train on a $15\%$ training set.\label{tab:all_mcc}}
\begin{tabular}{lrccc|cccc}
\toprule
                                                  & $\frac{|\trainset{}|}{|E|}$ &                    \uslogregp{} &                       \usrule{} &                 \uslpropGsec{} &   \compranknodes{} &                 \compbayesian{} &     \complowrank{} &      \comptriads{} \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\aut{}}}
                                                  & $3\%$  &  $\vsecondSig{15.19} \pm 0.93$  &               $15.09 \pm 0.92$  &  $\vfirstSig{19.00} \pm 1.16$  &  $12.28 \pm 1.70$  &               $10.91 \pm 1.31$  &   $8.85 \pm 1.07$  &   $8.62 \pm 0.97$  \\
                                                  %& $5\%$  &  $\vsecondSig{19.54} \pm 0.48$  &               $19.53 \pm 0.43$  &  $\vfirstSig{24.11} \pm 0.68$  &  $16.65 \pm 0.59$  &               $14.53 \pm 0.85$  &  $12.07 \pm 0.88$  &  $11.47 \pm 1.23$  \\
                                                  %& $7\%$  &  $\vsecondSig{23.58} \pm 0.55$  &               $23.55 \pm 0.43$  &  $\vfirstSig{27.64} \pm 1.08$  &  $21.38 \pm 0.86$  &               $19.75 \pm 1.02$  &  $15.15 \pm 0.54$  &  $14.58 \pm 1.15$  \\
                                                  & $9\%$  &  $\vsecondSig{26.46} \pm 0.96$  &               $26.40 \pm 0.96$  &  $\vfirstSig{30.25} \pm 1.16$  &  $24.44 \pm 1.07$  &               $23.75 \pm 1.37$  &  $17.08 \pm 1.15$  &  $16.42 \pm 1.13$  \\
                                                  & $15\%$ &  $\vsecondSig{32.98} \pm 0.61$  &               $32.98 \pm 0.65$  &  $\vfirstSig{35.73} \pm 0.70$  &  $31.03 \pm 0.97$  &               $32.25 \pm 0.89$  &  $22.57 \pm 0.99$  &  $22.01 \pm 0.62$  \\
                                                  & $20\%$ &               $36.57 \pm 0.49$  &  $\vsecondSig{36.72} \pm 0.44$  &  $\vfirstSig{38.53} \pm 0.55$  &  $34.57 \pm 0.79$  &               $36.52 \pm 0.78$  &  $25.57 \pm 0.71$  &  $24.77 \pm 0.64$  \\
                                                  & $25\%$ &               $39.90 \pm 0.65$  &               $40.16 \pm 0.63$  &  $\vfirstSig{41.32} \pm 0.67$  &  $38.26 \pm 0.77$  &  $\vsecondSig{40.32} \pm 0.70$  &  $29.24 \pm 1.10$  &  $27.13 \pm 0.38$  \\
                                                  & time (ms)   &                           1.9 &                           0.4 &                          16.1 &              128 &                         5398 &               1894 &              4.6 \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\wik{}}} 
                                                  & $3\%$  &  $\vsecondSig{32.32} \pm 0.60$  &  $31.83 \pm 0.63$  &   $\vfirstSig{33.92} \pm 0.54$  &  $26.90 \pm 1.00$  &  $19.94 \pm 1.19$  &  $19.45 \pm 0.64$  &   $4.29 \pm 0.80$  \\
                                                  %& $5\%$  &  $\vsecondSig{38.62} \pm 0.63$  &  $37.82 \pm 0.66$  &   $\vfirstSig{39.44} \pm 0.44$  &  $33.47 \pm 1.12$  &  $26.68 \pm 0.72$  &  $24.18 \pm 1.04$  &  $11.36 \pm 1.32$  \\
                                                  %& $7\%$  &     $\vsecond{42.81} \pm 0.80$  &  $42.15 \pm 0.76$  &      $\vfirst{43.28} \pm 0.61$  &  $38.13 \pm 0.80$  &  $33.18 \pm 1.27$  &  $27.89 \pm 0.64$  &  $18.57 \pm 1.38$  \\
                                                  & $9\%$  &     $\vsecond{45.57} \pm 0.48$  &  $44.74 \pm 0.58$  &      $\vfirst{45.75} \pm 0.46$  &  $41.60 \pm 0.73$  &  $38.25 \pm 1.02$  &  $30.75 \pm 0.68$  &  $24.04 \pm 1.27$  \\
                                                  & $15\%$ &      $\vfirst{50.70} \pm 0.28$  &  $49.64 \pm 0.57$  &     $\vsecond{50.44} \pm 0.38$  &  $48.02 \pm 0.38$  &  $46.82 \pm 0.40$  &  $35.31 \pm 0.75$  &  $34.42 \pm 1.00$  \\
                                                  & $20\%$ &   $\vfirstSig{52.98} \pm 0.50$  &  $52.00 \pm 0.47$  &  $\vsecondSig{52.58} \pm 0.40$  &  $51.42 \pm 0.33$  &  $50.45 \pm 0.40$  &  $38.16 \pm 0.53$  &  $38.55 \pm 0.72$  \\
                                                  & $25\%$ &      $\vfirst{54.49} \pm 0.30$  &  $53.52 \pm 0.41$  &     $\vsecond{54.22} \pm 0.44$  &  $53.42 \pm 0.36$  &  $52.78 \pm 0.61$  &  $39.94 \pm 0.48$  &  $41.51 \pm 0.46$  \\
                                                  & time (ms)   &                           4.2 &              1.1 &                           35.3 &              210 &              14089 &               4858 &              10.6 \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\sla{}}} 
                                                  & $3\%$  &  $32.34 \pm 0.27$  &  $31.78 \pm 0.81$  &  $\vsecondSig{36.62} \pm 0.42$  &   $\vfirstSig{42.90} \pm 0.54$  &  $25.11 \pm 0.66$  &  $34.32 \pm 0.64$  &  $20.95 \pm 0.86$  \\
                                                  %& $5\%$  &  $36.93 \pm 0.32$  &  $36.25 \pm 0.63$  &  $\vsecondSig{40.72} \pm 0.24$  &   $\vfirstSig{45.47} \pm 0.37$  &  $29.50 \pm 0.17$  &  $36.95 \pm 0.44$  &  $28.26 \pm 1.06$  \\
                                                  %& $7\%$  &  $39.96 \pm 0.27$  &  $39.26 \pm 0.39$  &  $\vsecondSig{43.62} \pm 0.21$  &   $\vfirstSig{46.80} \pm 0.64$  &  $33.57 \pm 0.27$  &  $38.25 \pm 0.31$  &  $33.89 \pm 1.41$  \\
                                                  & $9\%$  &  $42.16 \pm 0.16$  &  $41.19 \pm 0.30$  &  $\vsecondSig{45.70} \pm 0.26$  &   $\vfirstSig{47.46} \pm 0.27$  &  $37.00 \pm 0.17$  &  $39.42 \pm 0.20$  &  $39.14 \pm 1.43$  \\
                                                  & $15\%$ &  $46.44 \pm 0.11$  &  $45.23 \pm 0.32$  &   $\vfirstSig{49.65} \pm 0.23$  &  $\vsecondSig{48.59} \pm 0.72$  &  $43.28 \pm 0.21$  &  $41.09 \pm 0.29$  &  $46.27 \pm 1.00$  \\
                                                  & $20\%$ &  $48.71 \pm 0.22$  &  $47.79 \pm 0.29$  &     $\vsecond{51.88} \pm 0.28$  &      $\vfirst{52.09} \pm 0.28$  &  $47.03 \pm 0.24$  &  $43.10 \pm 0.42$  &  $49.44 \pm 0.76$  \\
                                                  & $25\%$ &  $50.23 \pm 0.16$  &  $49.43 \pm 0.22$  &     $\vsecond{53.30} \pm 0.15$  &      $\vfirst{53.46} \pm 0.23$  &  $49.46 \pm 0.24$  &  $44.37 \pm 0.50$  &  $51.51 \pm 0.48$  \\
                                                  & time (ms)   &              21.2 &              5.7 &                           655 &                            1918 &              77042 &              56252 &              78 \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\epi{}}} 
                                                  & $3\%$  &  $43.51 \pm 0.35$  &  $41.39 \pm 1.11$  &     $\vsecond{51.47} \pm 0.28$  &     $\vfirst{52.04} \pm 0.52$  &  $31.00 \pm 0.66$  &  $36.84 \pm 0.43$  &  $34.42 \pm 3.59$  \\
                                                  %& $5\%$  &  $49.16 \pm 0.30$  &  $46.65 \pm 0.83$  &  $\vsecondSig{54.69} \pm 0.33$  &  $\vfirstSig{56.29} \pm 0.42$  &  $38.00 \pm 0.37$  &  $39.73 \pm 0.37$  &  $43.24 \pm 1.81$  \\
                                                  %& $7\%$  &  $52.35 \pm 0.21$  &  $50.27 \pm 0.55$  &  $\vsecondSig{56.77} \pm 0.33$  &  $\vfirstSig{58.40} \pm 0.46$  &  $43.91 \pm 0.43$  &  $41.93 \pm 0.28$  &  $46.05 \pm 1.17$  \\
                                                  & $9\%$  &  $54.85 \pm 0.21$  &  $53.23 \pm 0.62$  &  $\vsecondSig{58.43} \pm 0.34$  &  $\vfirstSig{60.21} \pm 0.48$  &  $48.24 \pm 0.38$  &  $43.95 \pm 0.54$  &  $49.94 \pm 1.50$  \\
                                                  & $15\%$ &  $59.29 \pm 0.15$  &  $57.76 \pm 0.42$  &  $\vsecondSig{61.41} \pm 0.19$  &  $\vfirstSig{62.69} \pm 0.34$  &  $56.88 \pm 0.56$  &  $48.61 \pm 0.88$  &  $54.56 \pm 1.27$  \\
                                                  & $20\%$ &  $61.45 \pm 0.13$  &  $60.06 \pm 0.31$  &  $\vsecondSig{63.14} \pm 0.18$  &  $\vfirstSig{64.13} \pm 0.19$  &  $61.49 \pm 0.29$  &  $51.43 \pm 0.75$  &  $56.96 \pm 1.49$  \\
                                                  & $25\%$ &  $62.95 \pm 0.18$  &  $61.93 \pm 0.29$  &  $\vsecondSig{64.47} \pm 0.29$  &  $\vfirstSig{65.22} \pm 0.61$  &  $64.45 \pm 0.39$  &  $54.51 \pm 0.99$  &  $58.73 \pm 1.65$  \\
                                                  & time (ms)   &              32 &              7.1 &                            1226 &                           2341 &             116838 &             121530 &              129 \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\kiw{}}} 
                                                  & $3\%$  &               $26.02 \pm 0.27$  &  $\vsecondSig{26.23} \pm 0.46$  &   $\vfirstSig{33.92} \pm 0.50$  &  $23.59 \pm 0.43$  &              $20.02 \pm 0.68$  &  $20.13 \pm 0.56$  &   $1.11 \pm 0.26$  \\
                                                  %& $5\%$  &               $30.76 \pm 0.18$  &  $\vsecondSig{30.95} \pm 0.29$  &   $\vfirstSig{36.46} \pm 0.53$  &  $26.60 \pm 0.53$  &              $26.33 \pm 0.77$  &  $23.09 \pm 0.47$  &   $3.35 \pm 0.58$  \\
                                                  %& $7\%$  &               $33.44 \pm 0.26$  &  $\vsecondSig{33.45} \pm 0.26$  &   $\vfirstSig{37.57} \pm 0.47$  &  $30.79 \pm 1.43$  &              $30.53 \pm 0.50$  &  $25.00 \pm 0.39$  &   $6.54 \pm 1.42$  \\
                                                  & $9\%$  &  $\vsecondSig{35.27} \pm 0.19$  &               $35.13 \pm 0.23$  &   $\vfirstSig{38.33} \pm 0.94$  &  $33.38 \pm 0.47$  &              $33.87 \pm 0.28$  &  $26.68 \pm 0.36$  &  $11.07 \pm 0.67$  \\
                                                  & $15\%$ &               $38.21 \pm 0.10$  &               $37.72 \pm 0.15$  &  $\vsecondSig{38.63} \pm 0.59$  &  $36.81 \pm 0.31$  &  $\vfirstSig{40.14} \pm 0.19$  &  $29.97 \pm 0.37$  &  $18.12 \pm 1.04$  \\
                                                  & $20\%$ &  $\vsecondSig{39.58} \pm 0.17$  &               $38.74 \pm 0.18$  &               $39.16 \pm 0.80$  &  $38.56 \pm 0.21$  &  $\vfirstSig{43.37} \pm 0.26$  &  $31.89 \pm 0.46$  &  $21.53 \pm 0.65$  \\
                                                  & $25\%$ &  $\vsecondSig{40.48} \pm 0.11$  &               $39.48 \pm 0.14$  &               $39.14 \pm 0.78$  &  $39.80 \pm 0.26$  &  $\vfirstSig{45.76} \pm 0.15$  &  $34.23 \pm 0.42$  &  $23.89 \pm 0.54$  \\
                                                  & time (ms)   &                           28 &                           6.6 &                           824 &               2938 &                         103264 &             130036 &              104 \\
\bottomrule
\end{tabular}
\end{table*}
***********************************************
\fi

\begin{table*}[t]
  \centering
\setlength{\tabcolsep}{3pt}
\scriptsize
% \begin{longtable}{lrccc|cccc}
\caption{MCC with increasing training set size, with one standard deviation over 12 random sampling of $\trainset$. The last four columns refer to the methods we took from the literature. For the sake of readability, we multiplied all MCC values by 100. The best number in each row is highlighted in \textbf{\textcolor{brown}{bold brown}} and the second one in \textit{\textcolor{red}{italic red}}. If the difference is statistically significant ($p$-value of a paired Student's $t$-test less than $0.005$), the best score is underlined. The ``time" rows contain the time taken to train on a $15\%$ training set.\label{tab:all_mcc}}
\begin{tabular}{lrcccc|cccc}
\toprule
                                                  & $\frac{|\trainset{}|}{|E|}$ &                 \uslpropGsec{} &                       \usrule{} &                    \uslogregp{} &         \qoptim{} &     \complowrank{} &      \comptriads{} &   \compranknodes{} &              \compbayesian{} \\
\midrule
% \multirow{7}{*}{\rotatebox[origin=c]{90}{\aut{}}} & $3\%$  &  $\vfirstSig{19.07} \pm 1.38$  &               $14.97 \pm 0.72$  &  $\vsecondSig{15.05} \pm 0.73$  &  $10.88 \pm 0.83$  &   $8.89 \pm 1.44$  &   $7.54 \pm 1.42$  &  $12.48 \pm 1.18$  &            $11.06 \pm 1.64$  \\
%                                                   & $9\%$  &  $\vfirstSig{29.64} \pm 0.74$  &  $\vsecondSig{25.96} \pm 0.61$  &               $25.90 \pm 0.67$  &  $23.51 \pm 0.72$  &  $17.05 \pm 0.70$  &  $16.13 \pm 0.64$  &  $23.84 \pm 0.91$  &            $23.44 \pm 0.81$  \\
\multirow{7}{*}{\rotatebox[origin=c]{90}{\aut{}}} & $5\%$                       & $\vfirstSig{24.54} \pm 0.69$  & $\vsecondSig{20.21} \pm 0.66$ & $20.19 \pm 0.71$              & $15.86 \pm 0.81$ & $12.76 \pm 0.65$ & $11.04 \pm 0.81$ & $17.18 \pm 1.11$             & $15.28 \pm 1.31$ \\
                                                  & $10\%$                      & $\vfirstSig{31.20} \pm 0.58$  & $\vsecondSig{27.54} \pm 0.56$ & $27.49 \pm 0.62$              & $25.36 \pm 0.78$ & $17.81 \pm 0.76$ & $16.99 \pm 0.63$ & $25.36 \pm 0.85$             & $24.74 \pm 0.59$ \\
                                                  & $15\%$ &  $\vfirstSig{35.66} \pm 0.68$  &  $\vsecondSig{32.87} \pm 0.58$  &               $32.79 \pm 0.60$  &  $31.39 \pm 0.75$  &  $22.58 \pm 0.53$  &  $21.55 \pm 0.91$  &  $30.60 \pm 0.87$  &            $31.71 \pm 0.99$  \\
                                                  & $20\%$ &  $\vfirstSig{38.67} \pm 0.48$  &  $\vsecondSig{36.94} \pm 0.51$  &               $36.86 \pm 0.48$  &  $35.47 \pm 0.41$  &  $25.80 \pm 0.94$  &  $24.27 \pm 0.56$  &  $35.01 \pm 0.83$  &            $36.13 \pm 0.75$  \\
                                                  & $25\%$ &     $\vfirst{41.05} \pm 0.73$  &               $39.83 \pm 0.58$  &               $39.76 \pm 0.59$  &  $38.48 \pm 0.55$  &  $29.67 \pm 0.78$  &  $26.85 \pm 0.87$  &  $38.06 \pm 0.86$  &  $\vsecond{40.34} \pm 0.94$  \\
                                                  & time   &                           19.6 &                             0.6 &                             2.6 &               2835 &               3279 &                6.2 &                155 &                         4813 \\
\midrule
% \multirow{7}{*}{\rotatebox[origin=c]{90}{\wik{}}} & $3\%$  &   $\vfirstSig{34.08} \pm 1.03$  &  $31.65 \pm 0.99$  &  $\vsecondSig{32.08} \pm 0.94$  &  $28.50 \pm 0.92$  &  $19.73 \pm 0.99$  &   $3.98 \pm 0.84$  &  $27.43 \pm 0.99$  &  $20.11 \pm 1.14$  \\
%                                                   & $9\%$  &      $\vfirst{45.80} \pm 0.43$  &  $44.92 \pm 0.48$  &     $\vsecond{45.72} \pm 0.42$  &  $43.02 \pm 0.52$  &  $30.24 \pm 0.54$  &  $23.50 \pm 1.66$  &  $41.36 \pm 0.66$  &  $38.05 \pm 0.93$  \\
\multirow{7}{*}{\rotatebox[origin=c]{90}{\wik{}}} & $5\%$                       & $\vfirstSig{39.46} \pm 0.79$  & $38.03 \pm 0.97$              & $\vsecondSig{38.50} \pm 0.87$ & $35.72 \pm 0.70$ & $24.58 \pm 1.18$ & $9.59 \pm 1.10$  & $33.60 \pm 0.64$             & $26.45 \pm 0.57$ \\
                                                  & $10\%$                      & $\vsecond{47.17} \pm 0.35$    & $46.03 \pm 0.49$              & $\vfirst{47.22} \pm 0.40$     & $44.53 \pm 0.48$ & $31.72 \pm 0.61$ & $26.36 \pm 0.83$ & $43.21 \pm 0.81$             & $40.28 \pm 0.69$ \\
                                                  & $15\%$ &     $\vsecond{50.49} \pm 0.33$  &  $49.89 \pm 0.40$  &      $\vfirst{50.87} \pm 0.36$  &  $49.08 \pm 0.33$  &  $35.77 \pm 0.58$  &  $33.64 \pm 0.83$  &  $48.50 \pm 0.47$  &  $47.07 \pm 0.38$  \\
                                                  & $20\%$ &  $\vsecondSig{52.74} \pm 0.31$  &  $52.24 \pm 0.49$  &   $\vfirstSig{53.13} \pm 0.27$  &  $51.79 \pm 0.35$  &  $37.90 \pm 0.27$  &  $38.41 \pm 0.53$  &  $51.49 \pm 0.43$  &  $50.54 \pm 0.39$  \\
                                                  & $25\%$ &     $\vsecond{54.00} \pm 0.63$  &  $53.42 \pm 0.59$  &      $\vfirst{54.26} \pm 0.37$  &  $53.31 \pm 0.37$  &  $40.16 \pm 0.57$  &  $41.34 \pm 1.07$  &  $53.30 \pm 0.37$  &  $52.92 \pm 0.48$  \\
                                                  & time   &                            41.9 &                1.6 &                             6.0 &              10629 &               8523 &               14.8 &                249 &              12507 \\
\midrule
% \multirow{7}{*}{\rotatebox[origin=c]{90}{\sla{}}} & $3\%$  &  $\vsecondSig{36.58} \pm 0.43$  &  $31.71 \pm 0.59$  &  $32.29 \pm 0.47$  &  $29.14 \pm 0.56$  &  $34.39 \pm 0.56$  &  $20.57 \pm 0.93$  &   $\vfirstSig{42.98} \pm 0.56$  &  $24.96 \pm 0.43$  \\
%                                                   & $9\%$  &  $\vsecondSig{45.71} \pm 0.28$  &  $41.06 \pm 0.38$  &  $42.19 \pm 0.18$  &  $39.60 \pm 0.25$  &  $39.39 \pm 0.34$  &  $37.92 \pm 1.57$  &   $\vfirstSig{47.40} \pm 0.42$  &  $36.83 \pm 0.36$  \\
\multirow{7}{*}{\rotatebox[origin=c]{90}{\sla{}}} & $5\%$                       & $\vsecondSig{40.77} \pm 0.20$ & $36.13 \pm 0.57$              & $37.00 \pm 0.29$              & $33.49 \pm 0.32$ & $36.83 \pm 0.47$ & $27.10 \pm 0.75$ & $\vfirstSig{45.16} \pm 0.59$ & $29.25 \pm 0.23$ \\
                                                  & $10\%$                      & $\vsecondSig{46.61} \pm 0.29$ & $41.89 \pm 0.39$              & $43.15 \pm 0.21$              & $40.92 \pm 0.23$ & $39.57 \pm 0.27$ & $40.38 \pm 1.47$ & $\vfirstSig{47.84} \pm 0.50$ & $38.25 \pm 0.21$ \\
                                                  & $15\%$ &   $\vfirstSig{49.62} \pm 0.22$  &  $45.42 \pm 0.36$  &  $46.42 \pm 0.16$  &  $45.56 \pm 0.19$  &  $41.21 \pm 0.19$  &  $45.88 \pm 1.01$  &  $\vsecondSig{48.75} \pm 0.71$  &  $43.47 \pm 0.16$  \\
                                                  & $20\%$ &     $\vsecond{51.88} \pm 0.24$  &  $47.78 \pm 0.25$  &  $48.66 \pm 0.10$  &  $48.10 \pm 0.30$  &  $42.74 \pm 0.44$  &  $48.79 \pm 0.57$  &      $\vfirst{52.10} \pm 0.33$  &  $46.89 \pm 0.27$  \\
                                                  & $25\%$ &     $\vsecond{53.12} \pm 0.20$  &  $49.39 \pm 0.24$  &  $50.22 \pm 0.12$  &  $50.11 \pm 0.20$  &  $44.24 \pm 0.44$  &  $50.62 \pm 0.53$  &      $\vfirst{53.29} \pm 0.22$  &  $49.42 \pm 0.22$  \\
                                                  & time   &                             677 &                8.3 &               32.8 &              78537 &              69988 &                131 &                            2441 &              68085 \\
\midrule
% \multirow{7}{*}{\rotatebox[origin=c]{90}{\epi{}}} & $3\%$  &  $\vsecondSig{51.43} \pm 0.29$  &  $41.42 \pm 0.84$  &  $43.41 \pm 0.34$  &  $36.56 \pm 0.49$  &  $37.01 \pm 0.57$  &  $31.71 \pm 6.20$  &  $\vfirstSig{52.33} \pm 0.42$  &            $31.25 \pm 0.80$  \\
%                                                   & $9\%$  &  $\vsecondSig{58.48} \pm 0.26$  &  $53.20 \pm 0.62$  &  $54.80 \pm 0.12$  &  $51.58 \pm 0.36$  &  $43.81 \pm 0.50$  &  $50.43 \pm 2.07$  &  $\vfirstSig{60.06} \pm 0.46$  &            $48.06 \pm 0.41$  \\
\multirow{7}{*}{\rotatebox[origin=c]{90}{\epi{}}} & $5\%$                       & $\vsecondSig{54.83} \pm 0.16$ & $46.94 \pm 0.80$              & $49.16 \pm 0.32$              & $42.79 \pm 0.34$ & $39.96 \pm 0.60$ & $42.94 \pm 2.06$ & $\vfirstSig{56.04} \pm 0.76$ & $37.99 \pm 0.49$ \\
                                                  & $10\%$                      & $\vsecondSig{58.94} \pm 0.27$ & $54.03 \pm 0.46$              & $55.90 \pm 0.13$              & $53.43 \pm 0.39$ & $44.50 \pm 0.52$ & $50.29 \pm 1.07$ & $\vfirstSig{60.60} \pm 0.32$ & $49.90 \pm 0.36$ \\
                                                  & $15\%$ &  $\vsecondSig{61.47} \pm 0.21$  &  $57.63 \pm 0.45$  &  $59.25 \pm 0.17$  &  $58.80 \pm 0.32$  &  $48.24 \pm 0.58$  &  $54.64 \pm 1.62$  &  $\vfirstSig{62.69} \pm 0.21$  &            $56.94 \pm 0.65$  \\
                                                  & $20\%$ &  $\vsecondSig{63.17} \pm 0.13$  &  $60.15 \pm 0.40$  &  $61.45 \pm 0.17$  &  $61.86 \pm 0.13$  &  $52.21 \pm 0.37$  &  $57.27 \pm 1.42$  &  $\vfirstSig{64.10} \pm 0.12$  &            $61.18 \pm 0.45$  \\
                                                  & $25\%$ &               $64.05 \pm 0.20$  &  $61.88 \pm 0.38$  &  $62.89 \pm 0.12$  &  $63.42 \pm 0.14$  &  $54.68 \pm 0.62$  &  $58.42 \pm 1.59$  &     $\vfirst{65.40} \pm 0.85$  &  $\vsecond{64.59} \pm 0.30$  \\
                                                  & time   &                            1329 &               10.1 &               54.0 &             143881 &             127654 &                209 &                           3174 &                       104305 \\
\midrule
% \multirow{7}{*}{\rotatebox[origin=c]{90}{\kiw{}}} & $3\%$  &   $\vfirstSig{33.88} \pm 0.54$  &  $\vsecondSig{26.02} \pm 0.80$  &               $26.02 \pm 0.42$  &  $15.13 \pm 0.40$  &  $19.93 \pm 0.56$  &   $0.86 \pm 0.11$  &  $23.50 \pm 0.54$  &              $20.10 \pm 0.60$  \\
%                                                   & $9\%$  &   $\vfirstSig{38.62} \pm 0.52$  &               $35.10 \pm 0.17$  &  $\vsecondSig{35.29} \pm 0.11$  &  $28.23 \pm 0.33$  &  $26.61 \pm 0.37$  &   $9.99 \pm 0.93$  &  $33.22 \pm 0.29$  &              $33.86 \pm 0.24$  \\
\multirow{7}{*}{\rotatebox[origin=c]{90}{\kiw{}}} & $5\%$                       & $\vfirstSig{36.36} \pm 0.53$  & $\vsecondSig{30.89} \pm 0.28$ & $30.81 \pm 0.20$              & $21.69 \pm 0.25$ & $23.15 \pm 0.26$ & $3.04 \pm 0.46$  & $26.63 \pm 0.44$             & $26.68 \pm 0.34$ \\
                                                  & $10\%$                      & $\vfirstSig{38.58} \pm 0.74$  & $35.68 \pm 0.22$              & $\vsecondSig{35.93} \pm 0.16$ & $29.75 \pm 0.21$ & $27.07 \pm 0.44$ & $12.34 \pm 0.79$ & $33.85 \pm 0.33$             & $35.00 \pm 0.34$ \\
                                                  & $15\%$ &  $\vsecondSig{39.08} \pm 0.55$  &               $37.77 \pm 0.22$  &               $38.27 \pm 0.19$  &  $33.61 \pm 0.11$  &  $30.05 \pm 0.29$  &  $17.95 \pm 0.92$  &  $36.88 \pm 0.32$  &  $\vfirstSig{40.00} \pm 0.26$  \\
                                                  & $20\%$ &               $39.04 \pm 0.69$  &               $38.88 \pm 0.36$  &  $\vsecondSig{39.55} \pm 0.11$  &  $35.04 \pm 0.17$  &  $32.17 \pm 0.31$  &  $21.44 \pm 0.67$  &  $38.60 \pm 0.31$  &  $\vfirstSig{43.32} \pm 0.22$  \\
                                                  & $25\%$ &               $38.90 \pm 0.45$  &               $39.41 \pm 0.16$  &  $\vsecondSig{40.44} \pm 0.14$  &  $36.18 \pm 0.20$  &  $33.94 \pm 0.74$  &  $23.41 \pm 0.41$  &  $39.75 \pm 0.32$  &  $\vfirstSig{45.76} \pm 0.29$  \\
                                                  & time   &                             927 &                             9.6 &                            46.8 &             219109 &             129460 &                177 &               3890 &                          92719 \\
\bottomrule
\end{tabular}
\end{table*}

{\bf Results.} Our main results are summarized in Table~\ref{tab:all_mcc}, reporting MCC test set performance after training on sets of varying size (from 5\% to 25\%). Results have been averaged over 12 repetitions. Because scalability is a major concern 
%when training 
on sizeable datasets, we also give an idea of relative training times (in milliseconds) by reporting the time it took to train a single run of each algorithm on a training set of size\footnote
{
Comparison of training time performances is fair since all algorithms have been carefully implemented using the same stack of Python libraries, and run on the same machine (16 Xeon cores and 192Gb Ram).
} 
15\% of $|E|$, and then predict on the test set. Though our experiments are not conclusive, some trends can be 
%readily 
spotted:
%
%\begin{itemize}%[nosep,leftmargin=2em]
%

%\item 
{\bf 1.}
Global methods tend to outperform local methods in terms of prediction performance, but are also significantly (or even much) slower (running times can differ by as much as three orders of magnitude). This is not surprising, and is in line with previous experimental findings (e.g., \cite{shahriari2014ranking,wu2016troll}). \compbayesian{} looks like an exception to this rule, but its running time is indeed in the same ballpark as global methods.
%

%\item
{\bf 2.} 
\uslpropGsec{} always ranks first or at least second in this comparison when MCC is considered. On top of it, \uslpropGsec{} is fastest among the global methods (one or even two orders of magnitude faster), thereby showing the benefit of our approach to edge sign prediction.
%\item

{\bf 3.} The regularized solution computed by \uslpropGsec{} is always better than the unregularized one computed by \qoptim{} in terms of both MCC and running time.

{\bf 4.} 
As claimed in the introduction, our Bayes approximator \usrule{} closely mirrors in performance the more involved \uslogregp{} model. In fact, supporting our generative model of Section~\ref{s:gen}, the logistic regression weights for features $1-\htr(i)$ and $1-\hun(j)$ are almost equal (see Table~2 in the supplementary material), thereby suggesting that predictor~\eqref{eq:predictor}, derived from the theoretical results at the beginning of Section~\ref{s:algbatch}, is {\em also} the best logistic model based on trollness and untrustworthiness.
%This also contributes to giving our generative model of Section\ref{s:gen} a desireable robustness.
%\end{itemize}
%

\iffalse
**********************************************
In our experiments, we sample uniformly at random $3$, $5$, $7$, $9$, $15$,
$20$, $30$, $40$, $50$, $70$, $80$ and $90$\% of the edges and train each
method on that sample before predicting the edges not sampled. This procedure
is repeated 10 times to account for the randomness of the sampling. As showed
in \autoref{tab:mcc_all}, in the low regime (that
is when $|\trainset{}|$ is less than $25\%$) our methods are always first or at
least second. Furthermore, we can compare their runtime performances since all
algorithms are implemented using the same stack of Python libraries and run
on the same machine (16 Xeon cores and 192Gb Ram). As showed in
\autoref{tab:time_per_edge}, our method are faster
(\uslprop{}), up to several orders of magnitude (\usrule{}).
**********************************************
\fi

\iffalse
**********************************************************
\begin{table}[b]
	\centering
	\caption{The time needed for each method to train on a $15\%$ training set and make its prediction.
		It is measured in number of test edges predicted per microsecond, to illustrate that our methods scale linearly with the number of edges.
		Out of fairness, we did not include the time needed to compute triads features needed by \compbayesian{} and \comptriads{} since our implementation is not optimal. 
		% However, this is a $O(|V|^3)$ operation which would necessarily comes at a cost.
	\label{tab:time_per_edge}}
	\input{exp_data/time_per_edge}
\end{table}

\begin{figure}[tb]
  \centering
  \begin{subfigure}[b]{\textwidth}
    \centering
    \caption{The time (in millisecond) needed for each method to train on a
      $15\%$ training set and make its prediction. Out of fairness, we did not
			include the time needed to compute triads features needed by \compbayesian{}
      and \comptriads{} since our implementation is not optimal. 
		% However, this is a $O(|V|^3)$ operation which would necessarily comes at a cost.
			\label{tab:time_total}}
      \input{exp_data/time_total}
  \end{subfigure}
  \begin{subfigure}[b]{\textwidth}
    \centering
    \caption{The same data scaled by the number of edges. More precisely, it
      measures the speed of each method by the number of test edges predicted
      per microsecond.  Note that our methods scale linearly with the number of
      edges. \label{tab:time_per_edge}}
      \input{exp_data/time_per_edge}
  \end{subfigure}
\end{figure}
**********************************************************
\fi

### 2017-09-02 troll/generative.tex
initial import of troll paper (with citation fix)

\section{Generative Model for Edge Labels}\label{s:gen}
We now define the stochastic generative model for edge labels we use in the batch learning setting. Given the graph $G = (V,E)$, let the label $y_{i,j} \in \spin$ of directed edge $(i,j) \in E$ be generated as follows. Each node $i \in V$ is endowed with two latent parameters $p_i, q_i \in [0,1]$, which we assume to be generated, for each node $i$, by an independent draw from a fixed but unknown joint prior distribution $\mu(p,q)$ over $[0,1]^2$. Each label $y_{i,j} \in \spin$ is then generated by an independent draw from the mixture of $p_i$ and $q_j$,
$
	\Pr\big( y_{i,j} = 1 \big) = \tfrac{p_i + q_j}{2}~.
$
The basic intuition is that the nature $y_{i,j}$ of a relationship $i\rightarrow j$ is stochastically determined by a mixture between how much node $i$ tends to like other people ($p_i$) and how much node $j$ tends to be liked by other people ($q_j$). In a certain sense, $1-tr(i)$ is the empirical counterpart to $p_i$, and $1-un(j)$ is the empirical counterpart to $q_j$.\footnote
{
One might view our model as reminiscent of standard models for link generation in social network analysis, like the classical $p_1$ model from \cite{hl81}. Yet, the similarity falls short, for all these models aim at representing the likelihood of the network topology, rather than the probability of edge signs, once the topology is {\em given}.
} 
Notice that the Bayes optimal prediction for $y_{i,j}$ is
$
	y^*(i,j) = \sgn\big(\eta(i,j) - \tfrac{1}{2}\big)~,
$
where $\eta(i,j) = \Pr\big( y_{i,j} = 1 \big)$. Moreover, the probability of drawing at random a $+1$-labeled edge from $\Nout(i)$ and the probability of drawing at random a $+1$-labeled edge from $\Nin(j)$ are respectively equal to
%

\begin{small}
\begin{equation}\label{e:pout}
\frac{1}{2}\,\Biggl(p_i + \frac{1}{\dout(i)}\!\!\sum_{j\in \NNout(i)} \!\!\! q_j \Biggl) \,\,\,\,\text{and}\,\,\,\,
%\end{equation}
%and similarly, the probability of drawing at random a $+1$-labeled edge from $\Nin(j)$ %equals
%
%\begin{equation}\label{e:pin}
\frac{1}{2}\,\Biggl(q_j + \frac{1}{\din(j)}\!\!\sum_{i\in \NNin(j)} \!\!\! p_i \Biggl)~.
\end{equation}
\end{small}
%

### 2017-09-02 troll/introduction.tex
initial import of troll paper (with citation fix)

\section{Introduction}
% say that we are facing edge sign prediction from many viewpoints.
% say about trolls and trustworthy people in a social network
% say that we are using reductions to node classification (novelty)
% generative model
% novel and cool
%
Connections in social networks are mostly driven by the {\em homophily} assumption: linked individuals tend to be similar, sharing personality traits, attitudes, or interests. However, homophily alone is clearly not sufficient to explain the variety of social links. In fact, sociologists have long studied networks, hereafter called \emph{signed} social networks, where also {\em negative} relationships ---like dissimilarity, disapproval or distrust--- are explicitly displayed. The presence of negative relationships is also a feature of many technology-mediated social networks. Known examples are \textsc{Ebay}, where users trust or distrust agents in the network based on their personal interactions, \textsc{Slashdot}, where each user can tag another user as friend or foe, and \textsc{Epinion}, where users can rate positively or negatively not only products, but also other users. Even in social networks where connections solely represent friendships, negative links can still emerge from the analysis of online debates among users.

%Consider, in particular, a user joining an online community within a social network. His/her behavior will often fit one of these two stereotypes: the new member could play well with other users, establishing positive relationships with those who have been helpful. Or, the new user could try and disrupt the community by engaging into anti-social behavior, and creating conflictual relationships with other members. This behavioral dichotomy is supported by decades of research in psychology, starting with the seminal work~\cite{Dissonance57} about cognitive dissonance (when someone is acting in contradiction with his/her personal beliefs, ideas, or values). This kind of attitude expressed publicly on social media leads to the definition of \enquote{trolls}~\cite{Hardaker10}: users whose real intentions are to cause disruption and/or to trigger or exacerbate conflict for the purposes of their own amusement. The paper~\cite{Shachaf10} elaborates on their motives, adding that boredom, attention seeking, and revenge motivate trolls; they find pleasure from causing damage to other people or to the community as a whole.

When the social network is signed, specific challenges arise in both network analysis and learning. On the one hand, novel methods are required to tackle standard tasks (e.g., user clustering, link prediction, targeted advertising/recommendation, 
%prediction of user interests, and 
analysis of the spreading of diseases in epidemiological models). On the other hand, new problems such as edge sign prediction, which we consider here, naturally emerge. Edge sign prediction is the problem of classifying the positive or negative nature of the links based on the network topology. Prior knowledge of the network topology is often a realistic assumption, for in several situations the discovery of the link sign can be more costly than acquiring the topological information of the network. For instance, when two users of an online social network communicate on a public web page, we immediately detect a link. Yet, the classification of the link sign as positive or negative may require complex techniques.

From the modeling and algorithmic viewpoints, because of the huge amount of available networked data, a major concern in developing learning methods for edge sign prediction is algorithmic scalability. Many successful, yet simple heuristics for edge sign prediction are based on the troll-trust features, i.e., on the fraction of outgoing negative links (trollness) and incoming positive links (trustworthiness) at each node. We study such heuristics by defining a probabilistic generative model for the signs on the directed links of a given network, and show that these heuristics can be understood and analyzed as approximators to the Bayes optimal classifier for our generative model. We also gather empirical evidence supporting our probabilistic model by observing that a logistic model trained on trollness and trustworthiness features alone is able to learn weights that, on all datasets considered in our experiments, consistently satisfy the properties predicted by our model.

We then introduce suitable graph transformations defining reductions from edge sign prediction to node sign prediction problems. This opens up the possibility of using the arsenal of known algorithmic techniques developed for node classification. In particular, we show that a Label Propagation algorithm, combined with our reduction, approximates the maximum likelihood estimator of our probabilistic generative model. Experiments on real-world data show the competitiveness of our approach in terms of both prediction performance (especially in the regime when training data are scarce) and scalability.

Finally, we point out that the notions of trollness and trustworthiness naturally define a measure of complexity, or learning bias, for the signed network that can also be used to design {\em online} (i.e., sequential) learning algorithms for the edge sign prediction problem. The learning bias encourages settings where the nodes in the network have polarized features (e.g., trollness/trustworthiness are either very high or very low). Our online analysis holds under adversarial conditions, namely, without any stochastic assumption on the assignment of signs to the network links.

%Hence, it is necessary to depart from well-established yet computationally expensive approaches by relying on novel algorithmic techniques.

%In this paper, we consider the problem of learning link classifiers, and study this problem from different viewpoints. Our contributions are summarized as follows.

\subsection{Related work}
%
Interest in signed networks can be traced back to the psychological theory of structural balance~\cite{Cartwright56,HeiderBook58} with its weak version~\cite{davis1967clustering}. The advent of online signed social networks has enabled a more thorough and quantitative understanding of that phenomenon. Among the several approaches related to our work, some extend the spectral properties of a graph to the signed case in order to find good embeddings for classification~\cite{Kunegis2009,SignedEmbedding15}. However, the use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which makes those methods hardly scalable to large graphs. Another approach is based on mining ego networks with SVM. Although this method seems to deliver good results~\cite{Papaoikonomou2014}, the running time makes it often impractical for large real-world datasets. An alternative approach, based on local features only and proposed in~\cite{Leskovec2010}, relies on the so-called status theory for directed graphs~\cite{guha2004propagation}. Some works in active learning, using a more sophisticated bias based on the correlation clustering (CC) index~\cite{Cesa-Bianchi2012a,Cesa-Bianchi2012b}, provide strong theoretical guarantees. However, the bias used there is rather strong, since it assumes the existence of a $2$-clustering of the nodes with a small CC index.

Whereas our focus will be on {\em binary} prediction, researchers have also considered a weighted version of the problem, where edges measure the amount of trust or distrust between two users~(e.g., \cite{guha2004propagation,Tang2013,
%Bachi2012,
Qian2014sn}). Other works have also considered versions of the problem where side information related to the network is available to the learning system. For instance, \cite{EdgeSignsRating15} uses the product purchased on Epinion in conjunction with a neural network, \cite{TrollDetection15} identifies trolls by analysing the textual content of their post, and~\cite{SNTransfer13} uses SVM to perform transfer learning from one network to another. While many of these approaches have interesting performances, they often require extra information which is not always available (or reliable) and, in addition, may face severe scaling issues.
% are not as scalable as ours. 
The recent survey~\cite{Tang2015a} contains pointers to many papers on edge sign prediction for signed networks, especially in the Data Mining area. Additional references, more closely related to our work, will be mentioned at the end of Section~\ref{ss:passive}.
% using the notation we now introduce. 
%for further references to the link classification problem.

### 2017-09-02 troll/notation_and_prel.tex
initial import of troll paper (with citation fix)

\section{Notation and Preliminaries}\label{s:prel}
%
In what follows, we let $G=(V,E)$ be a {\em directed} graph, whose edges $(i,j) \in E$ carry a binary label $y_{i,j}\in \spin$. The edge labeling will sometimes be collectively denoted by the $|V|\times |V|$ matrix $Y = [Y_{i,j}]$, where $Y_{i,j} = y_{i,j}$ if $(i,j) \in E$, and $Y_{i,j} = 0$, otherwise. The corresponding edge-labeled graph will be denoted by $G(Y) = (V,E(Y))$. We use $\Nin(i)$ and $\Nout(i)$ to denote, respectively, the set of edges incoming to and outgoing from node $i \in V$, with $\din(i) = \big|\Nin(i)\big|$ and $\dout(i) = \big|\Nout(i)\big|$ being the in-degree and the out-degree of $i$. Moreover, $\din^+(i)$ is the number of edges $(k,i)\in\Nin(i)$ such that $y_{k,i} = +1$. We define $\din^-(i)$, $\dout^+(i)$, and $\dout^-(i)$ similarly, so that, for instance, $\dout^-(i)/\dout(i)$ is the fraction of outgoing edges from node $i$ whose label in $G(Y)$ is $-1$. We call $tr(i) = \dout^-(i)/\dout(i)$ the {\em trollness} of node $i$, and $un(i) = \din^-(i)/\din(i)$ the {\em untrustworthiness} of node $i$. Finally, we also use the notation $\NNin(i)$ and $\NNout(i)$ to represent, respectively, the in-neighborhood and the out-neighborhood of node $i \in V$.

Given the directed graph $G = (V,E)$, we define two {\em edge-to-node reductions} transforming the original graph $G$ into other graphs. As we see later, these reductions are useful in turning the edge sign prediction problem into a {\em node} sign prediction problem (often called node classification problem), for which many algorithms are indeed available
%---see, e.g., \cite{BC01,LabelPropa03,BDL06,HP07,HLP09,VCGZ12,HPV12,WTA13,HPG15}.
---see, e.g., \cite{BC01,LabelPropa03,HP07,HLP09,WTA13}.
% In fact, when facing a specific edge sign prediction problem, we may find it convenient to operate on the {\em nodes} of the transformed graph, rather than on the {\em edges} of the original graph $G$.
Although any node classification method could in principle be used, the reductions we describe next are essentially aimed at preparing the ground for quadratic energy-minimization approaches computed through a {\em Label Propagation} algorithm~(e.g., \cite{LabelPropa03,BDL06}).

The first reduction, called $G \rightarrow G'$, builds an {\em undirected} graph $G' = (V',E')$ as follows. Each node $i \in V$ has two copies in $V'$, call them $\iin$ and $\iout$. Each directed edge $(i,j)$ in $E$ is associated with one node, call it $e_{i,j}$, in $V'$, along with the two undirected edges $(\iout,e_{i,j})$ and  $(e_{i,j},\jin)$. Hence $|V'| = 2|V|+|E|$ and $|E'| = 2|E|$. Moreover, if $G = G(Y)$ is edge labeled, then this labeling transfers to the subset of nodes $e_{i,j} \in V'$, so that $G'$ is a graph $G'(Y) = (V'(Y),E')$ with partially-labeled nodes. 
%For convenience (see Section~\ref{s:algbatch}), these labels are also mapped to $\{0,1\}$ through the mapping $y \rightarrow \frac{1+y}{2}$. 
The second reduction, called $G \rightarrow G''$, builds an {\em undirected and weighted} graph $G'' = (V'',E'')$.
%whose weights depend on a small value $\epsilon > 0$. 
Specifically, we have $V'' \equiv V'$ and $E'' \supset E'$, where the set $E''$ also includes edges $(\iout,\jin)$ for all $i$ and $j$ such that $(i,j) \in E$.
The edges in $E'$ have weight $2$, whereas the edges in $E''\setminus E'$ have weight $-1$.
% As we explain in \S~\ref{ss:passive}, to run Label Propagation on $G''$, we set the weight of edges in $E'$ to $2$ and those in $E''\setminus E'$ to $-1$.
Finally, as in the $G \rightarrow G'$ reduction, if $G = G(Y)$ is edge labeled, then this labeling transfers to the subset of nodes $e_{i,j} \in V''$. Graph $G'$, which will not be used in this paper, is an intermediate structure between $G$ and $G''$ and provides a conceptual link to the standard cutsize measure in node sign classification.
%in the online setting.
%, again after mapping to $\{0,1\}$. 
%, but in addition each label $y_{i,j}$ gets magnified by $2/\epsilon$. 
%Hence $G''$ is a weighted graph $G''(Y) = (V''(Y),E'')$ with partially-labeled nodes whose labels are in $\{0,2/\epsilon\}$. 
Figure~\ref{f:etnr} illustrates the two reductions.

%Moreover, given the outcome $G'$ of an edge-to-node reduction, and a balancedness parameter $b \in [0,1]$, we say that $G'_b$ is a randomly sparsified version of $G'$, when 
%$G'_b$ is the random subgraph produced from $G'$ as follows. Given any pairing of edges $(\iout,e_{i,j})$ and $(e_{i,j},\jin)$ of $G'$, draw an independent Bernoulli variable $B_{i,j}$ such that $\Pr(B_{i,j} = 1) = b$; if $B_{i,j} = 1$ then drop $(e_{i,j},\jin)$, otherwise drop $(\iout,e_{i,j})$ -- see Figure \ref{f:etnr}, right.

%
%\begin{figure}[t*]
%\begin{picture}(-10,242)(-10,242)
%\begin{tabular}{l}
%\includegraphics[width=0.95\textwidth]{etnr3.pdf}
%\end{tabular}
%\end{picture}

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.28\textwidth}
    \centering \includegraphics[height=.10\textheight]{g_latex-crop} \caption{}
  \end{subfigure}~
  \begin{subfigure}[b]{0.30\textwidth}
    \centering \includegraphics[height=.12\textheight]{gprime_latex-crop} \caption{}
  \end{subfigure}~
  \begin{subfigure}[b]{0.36\textwidth}
    \centering \includegraphics[height=.14\textheight]{gsecond_latex-crop} \caption{}
  \end{subfigure}
\caption{\label{f:etnr}
{\bf (a)} A directed edge-labeled graph $G$. {\bf (b)} Its corresponding graph $G'$ resulting from the $G\rightarrow G'$ reduction. The square nodes in $G'$ correspond to the edges in $G$, and carry the same labels as their corresponding edges. On the other hand, the $2|V|$ circle nodes in $G'$ are unlabeled. Observe that some nodes in $G'$ are isolated (and thus unimportant); these are exactly the nodes in $G'$ corresponding to the nodes having in $G$ no outgoing or no incoming edges ---see, e.g., nodes $3$ and $4$ in $G$. {\bf (c)} The weighted graph resulting from the $G\rightarrow G''$ reduction.
}
\end{figure*}

These reductions are meaningful only if they are able to approximately preserve 
% What is important here is to make sure that such reductions are able to (approximately) preserve
label {\em regularity} when moving from edges to nodes. That is, if the edge sign prediction problem is easy for a given $G(Y) = (V,E(Y))$, then the corresponding node sign prediction problems on $G'(Y) = (V'(Y),E')$ and $G''(Y) = (V''(Y),E)$ are also easy, and vice versa.
% We show this happens in the three learning settings within which we consider the
While we could make this argument more quantitative, here we simply observe that if each node in $G$ tends to be either troll or trustworthy, then few labels from the incoming and outgoing edges of each such node are sufficient to predict the labels on the remaining edges in $G$, and this translates to a small cutsize\footnote
{
Recall that the cutsize of an undirected node-labeled graph $G'(Y)$ is the number of edges in $G'$ connecting nodes having mismatching labels. 
}
of $G'(Y)$ over the nodes corresponding to the edges in $G$ (the colored squares in Figure \ref{f:etnr} (b)).
%
Again, we would like to point out that these reductions serve two purposes: %\textcolor{red}{
First, they allow us to use the many algorithms designed for the better studied problem of node sign prediction. Second, the reduction $G\rightarrow G''$ with the specific choice of edge weights is designed to make the Label Propagation solution approximate the maximum likelihood estimator associated with our generative model (see Section~\ref{ss:passive}).%} 
Note also that efficient Label Propagation implementations exist that can leverage the sparsity of $G''$.

We consider two learning settings associated with the problem of edge sign prediction: a batch setting and an online setting. In the batch setting, we assume that a training set of edges $E_0$ has been drawn uniformly at random {\em without replacement} from $E$, we observe the labels in $E_0$, and we are interested in predicting the sign of the remaining edges $E \setminus E_0$ by making as few prediction mistakes as possible. 
% In the batch {\em active} setting, we are given a budget of edge labels to observe, and are free to select them the way we want within $E(Y)$.
% Again, the goal is to make as few mistakes as possible on the remaining edges.
%
%Unlike the online setting, where the underlying labels are generated by an adaptive adversary, 
%
The specific batch setting 
%(passive and active) 
we study here assumes that labels are produced by a generative model which we describe in the next section, and our label regularity measure is a quadratic function (denoted by $\Psi^2_{G''}(Y)$ ---see Section~\ref{s:exp} for a definition), related to this model. $\Psi^2_{G''}(Y)$ is small just when all nodes in $G$ tend to be either troll or trustworthy.

On the other hand, the {\em online} setting we consider is the standard mistake bound model of online learning~\cite{Winnow88} where all edge labels are assumed to be generated by an adversary and sequentially presented to the learner according to an arbitrary permutation. For an online learning algorithm $A$, we are interested in measuring the total number of mistakes $M_A(Y)$ the algorithm makes over $G(Y)$ when the worst possible presentation order of the edge labels in $Y$ is selected by the adversary.
%---the reader is referred, e.g., to~\cite[Section~3]{Cesa-Bianchi2012b} for definitions specific to the edge sign prediction problem.
% In this setting, learning is split into a sequence of rounds: At each round $t=1,2,\dots$ a learning algorithm $A$ outputs a prediction ${\wh y_t} \in \spin$ for the label $y_t \in \spin$ of an edge arbitrarily selected by the adversary. After the prediction is made, label $y_t$ is revealed to the algorithm, hence allowing it to change its internal state. In the next round a new edge is selected by the adversary, and so on. Notice that
%since the underlying labeling over the edges is decided by the adversary once and for all,\footnote
%{
%For simplicity, we assume the adversary is deterministic.
%} 
%all edges occur exactly once within the sequence to be predicted (so that this game lasts exactly $|E|$ rounds). The adversary decides both the underlying labeling $Y$ over the edges of $G$ and the order of their presentation to the learning algorithm. We say that $A$ has made a mistake at time $t$ if ${\wh y_t} \neq y_t$, and we measure $A$'s prediction performance simply through the total number of mistakes $M_A(Y)$ it makes over $G(Y)$ when the worst possible presentation order of the edge labels in $Y$ is selected by the adversary. As is standard, we contrast $M_A(Y)$ to some kind of {\em regularity} measure $\Psi_G(Y)$ of the labeling $Y$ over $G$, so that we are in fact aimed at bounding the (cumulative) regret
%\(
%M_A(Y) - \Psi_G(Y)~.
%\)
Also in the online setting our label regularity measure, denoted here by $\Psi_G(Y)$, is small when nodes in $G$ tend to be either troll or trustworthy. 
%When this happens, few labels from the incoming and outgoing edges of each node are sufficient to predict the labels on the remaining edges. 
Formally, for fixed $G$ and $Y$, let 
$
    \Psiin(j,Y) = \min\big\{\din^-(j),\din^+(j)\big\}
$ and
%\qquad\text{and}\qquad
$
    \Psiout(i,Y) = \min\big\{\dout^-(i),\dout^+(i)\big\}
$.
Let also
$
    \Psiin(Y) = \sum_{j \in V} \Psiin(j,Y)
$ and
%\qquad\text{and}\qquad
$
\Psiout(Y) = \sum_{i \in V} \Psiout(i,Y)
$.
Then we define $\Psi_G(Y) = \min\big\{\Psiin(Y),\Psiout(Y)\big\}$. The two measures $\Psi^2_{G''}(Y)$ and $\Psi_G(Y)$ are conceptually related. %\textcolor{red}{
Indeed, their value on real data is quite similar%}
% our experiments
(see Table~\ref{tab:all_mcc} in Section~\ref{s:exp}).

%NCB What point are we trying to make here?
% Recall that the cutsize of an undirected node-labeled graph $G'(Y)$ is the number of edges in $G'$ connecting nodes having mismatching labels. Now, because the only nodes in $G'$ we are interested in predicting are those corresponding to the edges in $G$ (the colored squares in Figure \ref{f:etnr} (b)), the online prediction problem on the edges of $G$ translates to a node sign prediction problem on a subset of $V'$. The remaining nodes in $V'$ (the circles in Figure \ref{f:etnr} (b)) we are free to assign dummy labels so as to minimize the corresponding mistake bound over $G'$. It is straightforward to see that if we set the labeling $Y$ on $V'$ in such a way that $y_{\iout} = +1$ if $tr(i) < 0$, and $-1$, otherwise, and $y_{\iin} = +1$ if $un(i) < 0$, and $-1$, otherwise, then the cutsize of $G'(Y)$ equals\footnote
%{
%In fact, for the sake of this specific argument, nothing prevents from retaining of $G'$ either only the edges $(\iout,e_{i,j})$ or only the edges $(e_{i,j},\jin)$, resulting in a cutsize of $\Psiout(Y)$ and $\Psiin(Y)$, respectively.
%} 
%$\Psiin(Y) + \Psiout(Y)$.

### 2017-09-02 troll/supp_body.tex
initial import of troll paper (with citation fix)

\section{Proofs from Section~\ref{s:algbatch}}
\subsection{Proof of Theorem~\ref{t:active}}
The following ancillary results will be useful.
%
\begin{lemma}[Hoeffding's inequality for sampling without replacement]
\label{l:hoeff-wr}
Let $\scX = \{x_1,\dots,x_N\}$ be a finite subset of $[0,1]$ and let
\[
	\mu = \frac{1}{N}\sum_{i=1}^N x_i~.
\]
If $X_1,\dots,X_n$ is a random sample drawn at random from $\scX$ without replacement, then, for every $\ve > 0$,
\[
	\Pr\left( \left|\frac{1}{n}\sum_{t=1}^n X_t - \mu \right| \ge \ve \right) \le 2e^{-2n\ve^2}~.
\]
\end{lemma}
%
\begin{lemma}\label{l:bins}
% Given a function $f\,:\, \Nat \rightarrow \R$ such that $f(n) \rightarrow \infty$ as $n \rightarrow \infty$, 
Let $\NN_1, \ldots, \NN_n$ be subsets of a finite set $E$.
% with $|E| \geq n\,f(n)$
% Let $\theta = \frac{|E|}{n\,f(n)}$, 
% and $k = k(\theta)$ be the number of subsets $\NN_i$ such that $|\NN_i| \geq \theta$. Let 
Let $E_0 \subseteq E$ be sampled uniformly at random {\em without} replacement from $E$, with $|E_0| = m$. Then, for $\delta \in (0,1)$, $Q > 0$, and $\theta \geq 2\times\max\left\{Q,4\ln \frac{n}{\delta}\right\}$, we have
\[
\Pr\Bigl( \exists i \,:\, |\NN_i| \geq \theta, |\NN_i \cap E_0| < Q\Bigl) \leq \delta
\]
provided $|E| \geq m \geq \frac{2|E|}{\theta}\times\max\left\{Q,4\ln \frac{n}{\delta}\right\}$.
\end{lemma}
%
\begin{proof}[Proof of Lemma \ref{l:bins}]
Set for brevity $p_i = |\NN_i|/|E|$. Then, due to the sampling without replacement, each random variable $|\NN_i \cap E_0|$ is the sum of $m$ dependent Bernoulli random variables $X_{i,1}, \ldots, X_{i,m}$ such that $\Pr(X_{i,t} = 1) = p_i$, for $t = 1, \ldots, m$. Let $i$ be such that $|\NN_i| \geq \theta$. Then the condition $m \geq \frac{2|E|Q}{\theta}$ implies
\[
Q \leq \frac{m\theta}{2|E|} \leq \frac{m\,p_i}{2} = \frac{\E\big[|\NN_i \cap E_0|\big]}{2}~.
\]
Since the variables $X_{i,j}$ are negatively associated, we may apply a (multiplicative) Chernoff bound~\cite[Section~3.1]{dpbook}.
This gives
\[
\Pr\big(|\NN_i \cap E_0| < Q\big) \leq e^{-\frac{m\,p_i}{8}} \leq e^{-\frac{m\,\theta}{8|E|}}
\]
so that
$
\Pr\bigl( \exists i\,:\, |\NN_i| \geq \theta, |\NN_i \cap E_0| < Q\bigr) \leq n\,e^{-\frac{m\,\theta}{8|E|}}
$,
which is in turn upper bounded by $\delta$ whenever 
$m \geq \frac{8|E|}{\theta}\ln \frac{n}{\delta}$. % = 8\,n\,f(n)\,\ln \frac{n}{\delta}$.
\end{proof}

Let now $E_{\theta} = \theset{(i,j) \in E}{\din(j) \ge \theta,\, \dout(i) \ge \theta} \setminus E_0$, where $E_0 \ss E$ is the set of edges sampled by the learning algorithm of Section~\ref{s:algbatch}. Then Theorem~\ref{t:active} in the main paper is an immediate consequence of the following lemma. 
%
\begin{lemma}\label{l:active}
Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the model in Section~\ref{s:gen}.
For all $0 < \alpha,\delta < 1$ and $0 < \ve < \tfrac{1}{16}$, if the learning algorithm of Section~\ref{s:algbatch} is run with parameters $Q = \tfrac{1}{2\ve^2}\ln\frac{4|V|}{\delta}$ and $\alpha$, then with probability at least $1-11\delta$ the predictions $\yhat(i,j)$ satisfy $\yhat(i,j) = y^*(i,j)$ for all $(i,j) \in E_{\theta}$ such that $\big|\eta(i,j) - \tfrac{1}{2}\big| > 8\ve$.
\end{lemma}
%
\begin{proof}[Proof of Lemma \ref{l:active}]
We apply Lemma~\ref{l:bins} with $\theta = \tfrac{2Q}{\alpha} \ge 2\times\max\big\{Q,4\ln \frac{2|V|+1}{\delta}\big\}$ to the $2|V|+1$ subsets of $E$ consisting of $E_L$ and $\Nin(i),\Nout(i)$, for $i \in V$. We have that, with probability at least $1-\delta$, at least $Q$ edges of $E_L$ are sampled, at least $Q$ edges of $\Nin(i)$ are sampled for each $i$ such that $\NNin(i) \ge \theta$, and at least $Q$ edges of $\Nout(j)$ are sampled for each $j$ such that $\NNout(j) \ge \theta$.
For all $(i,j) \in E_{\theta}$ let
\[
	\pbar_j = \frac{1}{\din(j)}\sum_{i\in\NNin(j)} p_i \quad\text{and}\quad \qbar_i = \frac{1}{\dout(i)}\sum_{j\in\NNout(i)} q_j
\]
and set for brevity $\hdeltain(j) = 1-\hun(j)$ and $\hdeltaout(i) = 1-\htr(i)$.
We now prove that $\hdeltain(j)$ and $\hdeltaout(i)$ are concentrated around their expectations for all $(i,j) \in E_{\theta}$. Consider $\hdeltaout(i)$ (the same argument works for $\hdeltain(j)$). Let $J_1,\dots,J_Q$ be the first $Q$ draws in $E_0 \cap \NNout(i)$ and define
\[
	\muhat_p(i) = \frac{1}{Q} \sum_{t=1}^Q \frac{p_i+q_{J_t}}{2}~.
\]
Applying Lemma~\ref{l:hoeff-wr} to the set $\theset{\tfrac{p_i+q_j}{2}}{j \in \NNout(i)}$, and using our choice of $Q$, we get that $\big|\muhat_p(i) - \mu_p(i)\big| \le \ve$ holds with probability at least $1-\delta/(2|V|)$, where
\[
	\mu_p(i) = \frac{1}{\dout(i)}\sum_{j \in \NNout(i)} \frac{p_i + q_j}{2} = \frac{p_i + \qbar_i}{2}~.
\]
Now consider the random variables $Z_t = \Ind{y_{i,J_t}=1}$, for $t=1,\dots,Q$. Conditioned on $J_1,\dots,J_Q$, these are independent Bernoulli random variables with $\E[Z_t \mid J_t] = \tfrac{p_i+q_{J_t}}{2}$. Hence, applying a standard Hoeffding bound for independent variables and using our choice of $Q$, we get that
\[
	\left|\frac{1}{Q}\sum_{t=1}^Q Z_t - \muhat_p(i) \right| \le \ve 
\]
with probability at least $1-\delta/(2|V|)$ for every realization of $J_1,\dots,J_Q$. 
%
Since $\hdeltaout(i) = (Z_1+\cdots+Z_Q)/Q$, we get that $\big|\hdeltaout(i) - \mu_p(i)\big| \le 2\ve$ with probability at least $1-2\delta/(2|V|)$. Applying the same argument to $\hdeltain(j)$, and the union bound\footnote
{
The sample spaces for the ingoing and outgoing edges of the vertices occurring as endpoints in $E_{\theta}$ overlap. Hence, in order to prove a uniform concentration result, we need to apply the union bound over the random variables defined over these sample spaces, which motivates the presence of the factor $\ln(2|V|)$ in the definition of $Q$. 
}
on the set $\theset{\hdeltain(j),\hdeltaout(i)}{(i,j) \in E_{\theta}}$, we get that
\begin{equation}
\label{eq:first}
	\left| \hdeltaout(i) + \hdeltain(j) - \frac{p_i + q_j}{2} - \frac{\pbar_j + \qbar_i}{2} \right| \le 4\ve
\end{equation}
simultaneously holds for all $(i,j) \in E_{\theta}$ with probability at least $1 - 4\delta$. Now notice that $\pbar_j$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables drawn from the prior marginal $\int_0^1 \mu\big(\cdot,q\bigr) dq$ with expectation $\mu_p$. Similarly, $\qbar_i$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables independently drawn from the prior marginal $\int_0^1 \mu\big(p,\cdot\big) dp$ with expectation $\mu_q$. By applying Hoeffding bound for independent variables, together with the union bound to the set of pairs of random variables whose sample means are $\pbar_j$ and $\qbar_i$ for each $(i,j) \in E_{\theta}$ (there are at most $2|V|$ of them) we obtain that
\begin{align*}
	\big|\pbar_j - \mu_p\big| \le \ve
\qquad\text{and}\qquad
	\big|\qbar_i - \mu_q\big| \le \ve
\end{align*}
hold simultaneously for all $(i,j) \in E_{\theta}$ with probability at least $1-2\delta$. Combining with~(\ref{eq:first}) we obtain that
\begin{equation}
\label{eq:this}
	\left| \hdeltaout(i) + \hdeltain(j) - \frac{p_i + q_j}{2} - \frac{\mu_p + \mu_q}{2} \right| \le 5\ve
\end{equation}
simultaneously holds for each $(i,j) \in E_{\theta}$ with probability at least $1 - 6\delta$. Next, let $E_L'$ be the set of the first $Q$ edges drawn in $E_L \cap E_0$. Then
\[
	\E\big[\tauhat\big] = \frac{1}{Q}\sum_{(i,j) \in E_L'} \Pr\big(y_{i,j} = 1\big) = \frac{1}{Q}\sum_{(i,j) \in E_L'} \frac{p_i+q_j}{2}~,
\]
where the expectation is w.r.t.\ the independent draws of the labels $y_{i,j}$ for $(i,j) \in E_L'$. Hence, by applying again Hoeffding bound (this time without the union bound) to the $Q = \tfrac{1}{2\ve^2}\ln\tfrac{2}{\delta}$ independent Bernoulli random variables $\Ind{y_{i,j} = 1}$, $(i,j) \in E_L'$, the event
$
	\big| \tauhat - \E\big[\tauhat\big] \big| \le \ve
$
holds with probability at least $1-\delta$. Now, introduce the function
\[
	F(\bp,\bq) = \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(i,j) \in E_L'} \frac{p_i+q_j}{2}~.
\]
For any realization $\bq_0$ of $\bq$, the function $F_1(\bp) = F(\bp,\bq_0)$ is a sample mean of $Q = \tfrac{1}{2\ve^2}\ln\tfrac{4|V|}{\delta}$ i.i.d.\ $[0,1]$-valued random variables $\theset{p_i}{(i,j) \in E_L'}$ (recall that if $i \in V$ is the origin of an edge $(i,j) \in E_L'$, then it is not the origin of any other edge $(i,j') \in E_L'$). Using again the standard Hoeffding bound, we obtain that
\[
	\left| F(\bp,\bq) - E_{\bp}\big[F(\bp,\bq)\big] \right| \le \ve
\]
holds with probability at least $1 -\delta$ for each $\bq\in [0,1]^{|V|}$. With a similar argument, we obtain that
\[
	\left| E_{\bp}\big[F(\bp,\bq)\big] - E_{\bp,\bq}\big[F(\bp,\bq)\big] \right| \le \ve
\]
also holds with probability at least $1 -\delta$. Since
\[
	E_{\bp,\bq}\big[F(\bp,\bq)\big] = \frac{\mu_p+\mu_q}{2}
\]
we obtain that
\begin{equation}
\label{eq:that}
	\Big| \tauhat - \frac{\mu_p+\mu_q}{2} \Big| \le 3\ve
\end{equation}
with probability at least $1-3\delta$. Combining~(\ref{eq:this}) with~(\ref{eq:that}) we obtain
\[
	\left| \hdeltaout(i) + \hdeltain(j) - \tauhat - \frac{p(i) + q(j)}{2} \right| \le 8\ve
\]
simultaneously holds for each $(i,j) \in E_{\theta}$ with probability at least $1 - 10\delta$. Putting together concludes the proof.
\end{proof}
%

\iffalse
****************************************************
\begin{lemma}[Hoeffding's inequality]
\label{l:hoeff}
Let $X_1,\dots,X_n$ be independent random variables such that $X_i \in [0,1]$ and let
\[
	s_n = \frac{1}{n}\sum_{t=1}^n \big(X_t - \E[X_t]\big)~.
\]
Then, for every $\ve > 0$,
$
	\Pr\big(|s_n| \ge \ve\big) \le 2e^{-2n\ve^2}~.
$
\end{lemma}
%
****************************************************
\fi

\subsection{Derivation of the maximum likelihood equations}
%
Recall that the training set
$
	E_0 = \theset{\big(i_t,j_t),y_{i_t,j_t}\big)}{t=1,\dots,m}
$
is drawn uniformly at random from $E$ without replacement. 
%Set for brevity $m = |E_0|$. 
We can write
%
\begin{align*}
	\Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)\\ 
&= 
	\frac{1}{\binom{|E|}{m}\,m!}\prod_{k=1}^m \left(\frac{p_{i_k}+q_{j_k}}{2} \right)^{\Ind{y_{i_k,j_k}=+1}}\,\\
&\quad\times\,\prod_{k=1}^m\left(1-\frac{p_{i_k}+q_{j_k}}{2} \right)^{\Ind{y_{i_k,j_k}=-1}}\\
&=
\frac{1}{\binom{|E|}{m}\,m!}
\prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
\left(\frac{p_{\ell}+q_{j_k}}{2} \right)^{\Ind{i_k = \ell,\,y_{\ell,j_k}=+1}}\,\\
&\quad\times\,
\prod_{k=1}^m\left(1-\frac{p_{\ell}+q_{j_k}}{2} \right)^{\Ind{i_k = \ell,\,y_{\ell,j_k}=-1}}\Biggl)
\end{align*}
%
so that $\log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)$ is proportional to
%
\begin{align*}
&\sum_{\ell=1}^{|V|}\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=+1}\,
\log\left(\frac{p_{\ell}+q_{j_k}}{2} \right)\,\\
&+\,
\sum_{\ell=1}^{|V|}\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=+1}\, 
\log\left(1-\frac{p_{\ell}+q_{j_k}}{2} \right)
\end{align*}
%
and
%
\begin{align*}
\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial p_{\ell}}
&= 
\sum_{k=1}^m 
\frac{\Ind{i_k = \ell,y_{\ell,j_k}=+1}}{p_{\ell}+q_{j_k}}\,\\
& - \,
\sum_{k=1}^m \frac{\Ind{i_k = \ell,y_{\ell,j_k}=-1}}{2-p_{\ell}-q_{j_k}}\,.
\end{align*}
%
By a similar argument,
%
\begin{align*}
\Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) 
\\ &=
\frac{1}{\binom{|E|}{m}\,m!}
\prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
\left(\frac{p_{i_k}+q_{\ell}}{2} \right)^{\Ind{j_k = \ell,\,y_{i_k,\ell}=+1}}\,\\
&\quad\times\,
\prod_{k=1}^m\left(1-\frac{p_{i_k}+q_{\ell}}{2} \right)^{\Ind{j_k = \ell,\,y_{i_k,\ell}=-1}}\Biggl)
\end{align*}
%
so that
%
\begin{align*}
\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial q_{\ell}}
&= 
\sum_{k=1}^m 
\frac{\Ind{j_k = \ell,y_{i_k,\ell}=+1}}{p_{i_k}+q_{\ell}}\,\\ 
&- \,
\sum_{k=1}^m 
\frac{\Ind{j_k = \ell,y_{i_k,\ell}=-1}}{2-p_{i_k}-q_{\ell}}\,.
\end{align*}
%
We then derive the approximation presented in the main paper. Namely, equating to zero the
gradient of the log likelihood w.r.t $p_\ell$ gives
\begin{align*}
\sum_{k=1}^m 
\frac{\Ind{i_k = \ell,y_{\ell,j_k}=+1}}{p_{\ell}+q_{j_k}}\,
 - \,
\frac{\Ind{i_k = \ell,y_{\ell,j_k}=-1}}{2-p_{\ell}-q_{j_k}} &= 0
\end{align*}
To simplify the notation, let $a_k = \Ind{i_k = \ell,y_{\ell,j_k}=+1}$, $b_k=\Ind{i_k =
\ell,y_{\ell,j_k}=+1}$ and $c_k = p_{\ell} + q_{j_k}$, we can rewrite the previous equation as 
\begin{align*}
\sum_{k=1}^m \frac{a_k}{c_k} - \frac{b_k}{2-c_k} &= 0 \\
\sum_{k=1}^m  \frac{a_k(2-c_k) - b_k c_k }{ c_k(2-c_k) } &=0
% \prod_{k=1}^m c_k\prod_{k=1}^m b_k\sum_{k=1}^m \frac{a_k}{b_k} &=\prod_{k=1}^m c_k\prod_{k=1}^m b_k \sum_{k=1}^m\frac{d_k}{c_k}
\end{align*}

The approximation consists in assuming that the denominator $c_k(2-c_k)$ is a
constant for all $k$, and can therefore be disregarded.
Moving $b_kc_k$ to the right hand side and returning to the original variables,
it yields the approximate equation presented in the main paper, namely
\begin{align*}
\sum_{k=1}^m &\Ind{i_k = \ell,y_{\ell,j_k}=+1} \left(2-p_{\ell}-q_{j_k}\right)\\
=
&\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=-1}
(p_{\ell}+q_{j_k})%, \quad \ell = 1, \ldots, |V|
\end{align*}

\iffalse
By applying a similar process to the gradient w.r.t. $q_\ell$, we obtain the
following approximate equation
\begin{align*}
\sum_{k=1}^m &\Ind{j_k = \ell,y_{i_k,\ell}=+1} \left(2-p_{i_k}-q_{\ell}\right)\\
=
&\sum_{k=1}^m  \Ind{j_k = \ell,y_{i_k,\ell}=-1}
\left(p_{i_k}+q_{\ell}\right)
\end{align*}
\fi

\subsection{Label propagation on $G''$}
%
Here we provide more details on the choice of weight for the edges of $G''$, as
well as an explanation on why we temporarily use symmetrized variables lying in
$[-1, 1]$ (which we will denote with primes, so that for instance $p'_i =
2p_i-1$). Since only the ratio between the negative and positive weights
matters, we fix the negative weight of the edges in $E''\setminus E'$ to be
$-1$ and we denote by $\eps$ the weight of edges in $E'$.  With these notations,
Label Propagation on $G''$ seeks the harmonic minimizer of the following
expression 
\begin{equation*}
  \frac{1}{16} \sum_{i,j \in E} \Bigl[ \eps\left(y_{i,j} - p_i'\right)^2 + \eps\left(y_{i,j} - q_j'\right)^2 + (p'_i+q_j')^2 \Bigr] \\
  % \label{eq:lprop_tmp}
\end{equation*}
which can be successively rewritten as
\begin{align*}
  \frac{1}{16} \sum_{i,j \in E} & \Bigl[ \eps\left(y_{i,j}+1 - 2p_i\right)^2 + \eps\left(y_{i,j}+1 - 2q_j\right)^2
\\&
	+ (2p_i +2q_j -2)^2 \Bigr]
\\=
%  \frac{1}{16} \sum_{i,j \in E} &\! 4\eps\!\left(\frac{y_{i,j}+1}{2} - p_i\right)^2\!\! +  4\eps\!\left(\frac{y_{i,j}+1}{2} - q_j\right)^2\!\! + 16\left(\frac{p_i +q_j -1}{2}\right)^2 \\
	\frac{1}{8} \sum_{i,j \in E} & \Biggl[ 2\eps\!\left(\frac{y_{i,j}+1}{2} - p_i\right)^2\!\! +  2\eps\!\left(\frac{y_{i,j}+1}{2} - q_j\right)^2
\\&
	+ 8\left(\frac{p_i +q_j -1}{2}\right)^2 \Biggr]
\\=
  \frac{1}{8}\sum_{i,j \in E} & \Biggl[ 2\eps\left(\left(\frac{y_{i,j}+1}{2}\right)^2 - p_i(1+y_{i,j}) + p_i^2\right) + \\
                              & 2\eps\left(\left(\frac{y_{i,j}+1}{2}\right)^2 - q_j(1+y_{i,j}) + q_j^2\right) +  \\
                              & 8\left(\left(\frac{p_i +q_j}{2}\right)^2 - \frac{p_i+q_j}{2} + \frac{1}{4} \right) \Biggr]
\\=
  \frac{1}{8}\sum_{i,j \in E} & 4\Biggl(\eps\!\left(\frac{y_{i,j}+1}{2}\right)^2\!\! -
  2\eps\!\left(\frac{y_{i,j}+1}{2}\right)\left(\frac{p_i+q_j}{2}\right)
\\&
  + 2\!\left(\frac{p_i +q_j}{2}\right)^2\Biggr)
\\&+
  \sum_{i,j \in E} \Bigl[ \left(2\eps p_i^2 - 4p_i + 1\right) + \left(2\eps q_j^2 - 4q_j + 1\right) \Bigr]
\end{align*}
%
By setting $\eps=2$, we can factor this expression into 
\begin{align*}
  \sum_{i,j \in E} &\left(\frac{y_{i,j}+1}{2} - \frac{p_i+q_j}{2}\right)^2
\\&
  + \frac{1}{2}\sum_{i,j \in E} \left(\left(p_i - \frac{1}{2}\right )^2 + \left(q_j - \frac{1}{2}\right)^2 \right)~.
\end{align*}

\section{Proofs from Section~\ref{s:algonline}}
%

\begin{proof}[Proof of Theorem~\ref{t:online}]
Let each node $i \in V$ host two instances of the randomized Weighted Majority (RWM) algorithm~\cite{LittlestoneWa94} with an online tuning of their learning rate~\cite{cb+97,acg02}: one instance for predicting the sign of outgoing edges $(i,j)$, and one instance for predicting the sign of incoming edges $(j,i)$. Both instances simply compete against the two constant experts, predicting always $+1$ or always $-1$. Denote by $M(i,j)$ the indicator function (zero-one loss) of a mistake on edge $(i,j)$. Then the expected number of mistakes of each RWM instance satisfy~\cite{cb+97,acg02}:
\[
    \sum_{j \in \NNout(i)} \E\,M(i,j) = \Psiout(i,Y) + O\left(\sqrt{\Psiout(i,Y)}+ 1\right)
\]
and
\[
    \sum_{i \in \NNin(j)} \E\,M(i,j) = \Psiin(j,Y) + O\left(\sqrt{\Psiin(j,Y)} + 1\right)~.
\]
We then define two meta-experts: an ingoing expert, which predicts $y_{i,j}$ using the prediction of the ingoing RWM instance for node $j$, and the outgoing expert, which predicts $y_{i,j}$ using the prediction of the outgoing RWM instance for node $i$.
The number of mistakes of these two experts satisfy
\begin{align*}
\sum_{i \in V}\sum_{j \in \NNout(i)} &\E\,M(i,j)\\ 
&= \Psiout(Y) + O\left(\sqrt{|V|\Psiout(Y)} + |V|\right)\\
\sum_{j \in V}\sum_{i \in \NNin(j)} &\E\,M(i,j)\\ 
&= \Psiin(Y)  + O\left(\sqrt{|V|\Psiin(Y)}  + |V|\right)~,
\end{align*}
where we used $\sum_{j \in V} \sqrt{\Psiin(j,Y)} \leq \sqrt{|V|\Psiin(Y)}$, and similarly for $\Psiout(Y)$.
Finally, let the overall prediction of our algorithm be a RWM instance run on top of the ingoing and the outgoing experts. Then the expected number of mistakes of this predictor satisfies
%
\begin{align*}
    \sum_{(i,j) \in E} \E\,M(i,j) 
&= \Psi_G(Y) + O\Biggl(\sqrt{|V|\Psi_G(Y)} + |V|\\ 
&\,\,+ \sqrt{\left(\Psi_G(Y) + |V| + \sqrt{|V|\Psi_G(Y)}\right)} \Biggl)\\
&= \Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V|\right)~,
\end{align*}
%
as claimed.
\end{proof}

\begin{proof}[Proof sketch of Theorem~\ref{t:mistake_bound}]
Let $\mathcal{Y}_K$ be the set of all labelings $Y$ such that the total number of negative and positive edges are $K$ and $|E|-K$, respectively (without loss of generality we will focus on negative edges). Consider the randomized strategy that draws a labeling 
$Y \in \spin^{|E|}$
% for the edges of the input graph 
uniformly at random from $\mathcal{Y}_K$. For each node $i \in V$, we have $\Psiin(i,Y) \le \din^-(i)$, which implies $\Psiin(Y) \le K$. A very similar argument applies to the outgoing edges, leading to $\Psiout(Y) \le K$. The constraint $\Psi_G(Y) \le K$ is therefore always satisfied.

The adversary will force on average $1/2$ mistakes in each one of the first $K$ rounds of the online protocol by repeating $K$ times the following: (i) A label value $\ell\in\{-1,+1\}$ is selected uniformly at random. (ii) An edge $(i,j)$ is sampled uniformly at random from the set of all edges that were not previously revealed and whose labels are equal to $\ell$.

The learner is required to predict $y_{i,j}$ and, in doing so, $1/2$ mistakes will be clearly made on average because of the randomized labeling procedure. Observe that this holds even when $A$ knows the value of $K$ and $\Psi_G(Y)$. Hence, we can conclude that the expected number of mistakes that $A$ can be forced to make is always at least $K/2$, as claimed.

We now show that, as $\frac{K}{|E|} \rightarrow 0$, the lower bound gets arbitrarily close to $K$ for any $G(Y)$ and any constant $K$.
%
Let $\mathcal{E}$ be the following event: There is at least one unrevealed negative label.
The randomized iterative strategy used to achieve this result is identical to the one described above, except for the stopping criterion. Instead of repeating step (i) and (ii) only for the first $K$ rounds, these steps are repeated until $\mathcal{E}$ is true.
%
Let $m_{r,c}$ be defined as follows: For $c=1$ it is equal to the expected number of mistakes forced in round $r$ when $K=1$. For $c > 1$ it is equal to the difference between the expected number of mistakes forced in round $r$ when $K=c$ and $K=c-1$. One can see that $m_{r,c}$ is 
%either 
null when $r<c$. 
%and, for any given $|E|$ and $K$, when $r>c+|E|-K-1$.
%at least one of the following  conditions is met: , . 
%$c>K$, 
%
When $K=1$, the probability that $\mathcal{E}$ is true in round $r$ is clearly equal to $\frac{1}{2^{r-1}}$.
Hence, the expected number of mistakes made by $A$ when $K=1$ in any round $r$ is equal to 
\(
\frac{1}{2}\,\frac{1}{2^{r-1}} = \frac{1}{2^{r}}.
\)
We can therefore conclude that $m_{r,1}=\frac{1}{2^{r}}$ for all $r$.

%When $r> 1$, 
A simple calculation shows that if $r=c$ then $m_{r,c}=\frac{1}{2^r}$. Furthermore, when $r>1$ and $c>1$, 
%for all $r$ and $c$ values such that $1 < c \le K$ and $c < r \le c+|E|-K-1$, 
we have the following recurrence:
%(holding even when $\frac{K}{|E|}$ does not converge to 0)
\[
m_{r,c}=\frac{m_{r-1,c}+m_{r-1,c-1}}{2}~.
\]  
In order to calculate $m_{r,c}$ for all $r$ and $c$, we will rest on the ancillary quantity $s_j(i)$, recursively defined as specified next.

Given any integer variable $i$, we have
%we define the following recurrence relation identity $s_j(i)$:
$
s_0(i)=1
$
and, for any positive integer $j$, 
\[
s_j(i)=\sum_{k=1}^i s_{j-1}(k)~.
\]
%(see Table~\ref{tab:sij}) for examples)

\iffalse
\begin{center}
  \begin{tabular}{|c |M{.45cm}|M{.45cm}|M{.45cm}|M{.45cm}|N}
    \hline
    \diaghead{j\ i}{\ \ \ 	{\footnotesize \textit{\textbf{j}}}\ }{\ \ {\footnotesize \textit{\textbf{i}}}\ \ \ } & \textbf{0}& \textbf{1} & \textbf{2} & \textbf{3}    &\\[13pt] \hline
    \textbf{1} & 1 & 1 & 1 & 1  &\\[13pt] \hline
    \textbf{2} & 1 & 2 & 3 & 4 &\\[13pt] \hline
    \textbf{3} & 1 & 3 & 6 & 10 &\\[13pt] \hline
    \textbf{4} & 1 & 4 & 10 & 20 &\\[13pt] \hline
    \textbf{5} & 1 & 5 & 15 & 35 &\\[13pt] \hline
    \textbf{6} & 1 & 6 & 21 & 56 &\\[13pt] \hline
    \textbf{7} & 1 & 7 & 28 & 84 &\\[13pt] \hline
  \end{tabular}
  \captionof{table}{\ \ Values of $s_j(i)$ for $i \le 7$  and $j \le 3$.} \label{tab:sij} 
\end{center}
\fi

It is not difficult to verify that 
\[
m_{r,c}=\frac{s_{c-1}(r-c+1)}{2^{r}}. 
\]
Since $s_j(i)=\frac{\langle i\rangle_j}{j!}$, where $\langle i\rangle_j$ is the rising factorial $i(i+1)(i+2)\ldots(i+j-1)$, we have 
\[
m_{r,c}=\frac{\langle r-c+1\rangle_{c-1}}{(c-1)!2^{r}}.
\] 
%(see Table~\ref{tab:mrc}) for examples)

\iffalse
\begin{center}
  \begin{tabular}{|c |M{.77cm}|M{.77cm}|M{.77cm}|M{.77cm}|N}
    \hline
    \diaghead{c\ r}{\ \ \ 	{\footnotesize \textit{\textbf{c}}}\ }{\ \
    {\footnotesize \textit{\textbf{r}}}\ \ \ } & \textbf{1}& \textbf{2} &
      \textbf{3} & \textbf{4}                            &\\[13pt] \hline
    \textbf{1} & $1/2^1$ & 0       & 0        & 0        &\\[13pt] \hline
    \textbf{2} & $1/2^2$ & $1/2^2$ & 0        & 0        &\\[13pt] \hline
    \textbf{3} & $1/2^3$ & $2/2^3$ & $1/2^3$  & 0        &\\[13pt] \hline
    \textbf{4} & $1/2^4$ & $3/2^4$ & $3/2^4$  & $1/2^4$  &\\[13pt] \hline
    \textbf{5} & $1/2^5$ & $4/2^5$ & $6/2^5$  & $4/2^5$  &\\[13pt] \hline
    \textbf{6} & $1/2^6$ & $5/2^6$ & $10/2^6$ & $10/2^6$ &\\[13pt] \hline
    \textbf{7} & $1/2^7$ & $6/2^7$ & $15/2^7$ & $20/2^7$ &\\[13pt] \hline
  \end{tabular}
  \captionof{table}{\ \ Values of $m_{r,c}$ for $r \le 7$, $c \le 4$ and $|E| \rightarrow \infty$.} \label{tab:mrc} 
\end{center}
\fi
\bigskip
%Given any $K'>1$, the difference between the expected number of mistakes %forced when $K=K'$ and $K=K'-1$ is equal to 
%\[
%\sum_{r=K'}^{|E|-1} m_{r,K'}=
%\sum_{r=K'}^{|E|-1} \frac{\langle r-K'+1\rangle_{K'-1}}{(K'-1)!2^{r}}.
%\]

When $\frac{K}{|E|} \rightarrow 0$, given any integer $K'>1$, the difference between the expected number of mistakes forced when $K=K'$ and $K=K'-1$ is equal to
%
\begin{align*}
\sum_{r=K'}^{\infty} m_{r,K'}
&=
\frac{1}{(K'-1)!}\sum_{r=K'}^{\infty} \frac{\langle r-K'+1\rangle_{K'-1}}{2^{r}}\\
&=
\frac{1}{(K'-1)!2^{K'-1}}\sum_{r'=1}^{\infty} \frac{\langle r'\rangle_{K'-1}}{2^{r'}}~,
\end{align*}
% 
where we set $r'=r-K'+1$.
%Taking into account the following recurrence relation sequence 
Setting $i'=i-1$ and recalling that
\[
\langle i\rangle_j=j!{{i+j-1}\choose{i-1}}~,
\]
we have
\[
\frac{1}{j!}\sum_{i=1}^{\infty}\frac{\langle i\rangle_j}{2^i}=
\sum_{i=1}^{\infty}\frac{{{i+j-1}\choose{i-1}}}{2^i}=
\sum_{i'=0}^{\infty}\frac{{{i'+j}\choose{i'}}}{2^{i'+1}}~.
\]
Now, using the identity
\[
{i'+j+1 \choose i'} = {i'+j \choose i'} + {i'+j \choose i'-1}~,
\]
we can easily prove by induction on $j$ that
\[
\sum_{i'=0}^{\infty}\frac{{{i'+j}\choose{i'}}}{2^{i'+1}}=2^j~.
\]
Hence, we have
%\[
%\frac{1}{j!}\sum_{i=1}^{\infty}\frac{\langle i\rangle_j}%{2^i}=2^j~,
%\]
\[
\sum_{r=K'}^{\infty} m_{r,K'}=1.
\]
Moreover, as shown earlier, $m_{r,1}=\frac{1}{2^{r}}$ for all $r$. Hence we can conclude that when $\frac{K}{|E|} \rightarrow 0$ 
\[
\E M_A(Y) \ge
\sum_{r=1}^{\infty} \frac{1}{2^{r}} +
\sum_{K'=2}^{K} \sum_{r=K'}^{\infty} m_{r,K'}=K
\]
for any edge-labeled graph $G(Y)$ and any constant $K$, as claimed.
\end{proof}

\iffalse %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the special case when the bias is
\[
    \Psi(Y) = \min\big\{\Psiin(Y),\Psiout(Y)\big\}
\]
where
\[
    \Psiin(Y) = \sum_{j \in V} \min\big\{\din^-(j),\din^+(j)\big\}
\]
and
\[
    \Psiout(Y) = \sum_{i \in V} \min\big\{\dout^-(i),\dout^+(i)\big\}
\]
then using Randomized Weighted Majority on the two experts
\[
    \Yhat^{(1)}_{i,j} = \sgn\left(\frac{1}{2}-\frac{\hdout^-(i)}{\hdout(i)}\right)
\qquad\text{and}\qquad
    \Yhat^{(2)}_{i,j} = \sgn\left(\frac{1}{2}-\frac{\hdin^-(j)}{\hdin(j)}\right)
\]
we get an expected mistake bound of
\[
    2\Psi(Y) + \sqrt{(4\ln 2) \Psi(Y)}~.
\]
\fi %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Further Experimental Results}
%
This section contains more evidence related to the experiments in Section~\ref{s:exp}.
In particular, we experimentally demonstrate the alignment between \usrule{} and \uslogregp{}.

\iffalse
*************************************
\subsection{Accuracy Results}
%
Table~\ref{tab:all_acc} contains the accuracy results on training sets of increasing size. Figures are averaged over 12 random draws of the training set.

\setlength{\tabcolsep}{3pt}
{\scriptsize
\begin{longtable}{lrccc|cccc}
\caption{Like in the main paper, we tabulate Accuracy $\pm$ one standard deviation with increasing training set size, averaged over 12 random draws of the training set. The best accuracy performance is highlighted in \textbf{\textcolor{brown}{bold brown}} and the second one in \textit{\textcolor{red}{italic red}}. If the difference is statistically significant (as defined by a paired Student's $t$-test $p$-value less than $0.005$), the best score is underlined. The "time" rows contain the time taken to train on a $15\%$ training set. \label{tab:all_acc}}\\
\toprule
                                                  & $\frac{|\trainset{}|}{|E|}$ &       \uslogregp{} &          \usrule{} &     \uslpropGsec{} &                \compranknodes{} &                \compbayesian{} &     \complowrank{} &                \comptriads{} \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\aut{}}} & $3\%$  &  $57.14 \pm 0.78$  &  $57.15 \pm 1.31$  &  $60.76 \pm 2.76$  &               $68.99 \pm 0.89$  &    $\vsecond{72.02} \pm 0.26$  &  $70.18 \pm 0.33$  &   $\vfirst{72.44} \pm 0.13$  \\
                                                  %& $5\%$  &  $62.07 \pm 0.79$  &  $61.97 \pm 1.17$  &  $66.80 \pm 1.53$  &               $69.44 \pm 0.69$  &     $\vfirst{71.90} \pm 0.33$  &  $70.49 \pm 0.27$  &  $\vsecond{71.87} \pm 0.65$  \\
                                                  %& $7\%$  &  $65.01 \pm 0.50$  &  $65.98 \pm 0.72$  &  $68.76 \pm 1.63$  &  $\vsecondSig{71.21} \pm 0.69$  &  $\vfirstSig{72.89} \pm 0.54$  &  $71.03 \pm 0.27$  &            $70.69 \pm 0.54$  \\
                                                  & $9\%$  &  $67.07 \pm 0.56$  &  $68.29 \pm 0.79$  &  $70.65 \pm 1.04$  &  $\vsecondSig{71.75} \pm 0.65$  &  $\vfirstSig{73.37} \pm 0.78$  &  $71.41 \pm 0.42$  &            $70.48 \pm 0.77$  \\
                                                  & $15\%$ &  $71.07 \pm 0.36$  &  $72.79 \pm 0.38$  &  $74.10 \pm 1.07$  &  $\vsecondSig{74.47} \pm 0.56$  &  $\vfirstSig{75.29} \pm 0.61$  &  $72.72 \pm 0.34$  &            $70.46 \pm 0.59$  \\
                                                  & $20\%$ &  $73.10 \pm 0.21$  &  $74.63 \pm 0.64$  &  $75.17 \pm 1.09$  &  $\vsecondSig{75.71} \pm 0.49$  &  $\vfirstSig{76.44} \pm 0.43$  &  $73.48 \pm 0.23$  &            $71.30 \pm 0.49$  \\
                                                  & $25\%$ &  $74.79 \pm 0.26$  &  $76.32 \pm 0.28$  &  $76.74 \pm 0.78$  &     $\vsecond{77.06} \pm 0.59$  &     $\vfirst{77.46} \pm 0.40$  &  $74.49 \pm 0.35$  &            $71.66 \pm 0.35$  \\
                                                  & time (ms)   &                           1.9 &                           0.4 &                          16.1 &              128 &                            5398 &               1894 &              4.6 \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\wik{}}} & $3\%$  &  $74.84 \pm 0.38$  &           $75.12 \pm 0.63$  &            $76.45 \pm 1.50$  &           $77.94 \pm 0.24$  &  $\vsecondSig{78.06} \pm 0.35$  &  $77.97 \pm 0.17$  &  $\vfirstSig{78.76} \pm 0.17$  \\
                                                  %& $5\%$  &  $78.32 \pm 0.34$  &           $78.86 \pm 0.37$  &            $79.33 \pm 0.79$  &  $\vfirst{79.67} \pm 0.48$  &     $\vsecond{79.42} \pm 0.60$  &  $78.78 \pm 0.25$  &              $78.83 \pm 0.17$  \\
                                                  %& $7\%$  &  $80.22 \pm 0.28$  &           $80.55 \pm 0.45$  &   $\vfirst{80.86} \pm 0.78$  &           $80.44 \pm 0.90$  &     $\vsecond{80.68} \pm 0.30$  &  $79.60 \pm 0.22$  &              $79.29 \pm 0.31$  \\
                                                  & $9\%$  &  $81.32 \pm 0.18$  &  $\vfirst{81.98} \pm 0.38$  &  $\vsecond{81.62} \pm 0.84$  &           $81.17 \pm 0.31$  &               $81.61 \pm 0.39$  &  $80.15 \pm 0.14$  &              $79.74 \pm 0.22$  \\
                                                  & $15\%$ &  $83.27 \pm 0.14$  &  $\vfirst{83.78} \pm 0.14$  &            $83.30 \pm 0.44$  &           $82.91 \pm 0.41$  &     $\vsecond{83.47} \pm 0.34$  &  $81.09 \pm 0.21$  &              $80.66 \pm 0.30$  \\
                                                  & $20\%$ &  $84.13 \pm 0.19$  &  $\vfirst{84.60} \pm 0.17$  &            $84.36 \pm 0.22$  &           $83.84 \pm 0.38$  &     $\vsecond{84.39} \pm 0.25$  &  $81.73 \pm 0.16$  &              $81.35 \pm 0.18$  \\
                                                  & $25\%$ &  $84.63 \pm 0.13$  &  $\vfirst{85.09} \pm 0.18$  &            $84.63 \pm 0.28$  &           $84.44 \pm 0.28$  &     $\vsecond{84.96} \pm 0.21$  &  $82.08 \pm 0.16$  &              $81.92 \pm 0.13$  \\
                                                  & time (ms)   &                           4.2 &              1.1 &                           35.3 &              210 &              14089 &              4858 &                                                   10.6\\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\sla{}}} & $3\%$  &  $69.06 \pm 0.19$  &  $69.39 \pm 0.37$  &               $74.01 \pm 0.84$  &  $\vfirstSig{80.74} \pm 0.21$  &               $78.33 \pm 0.14$  &  $\vsecondSig{79.90} \pm 0.17$  &  $74.59 \pm 0.58$  \\
                                                  %& $5\%$  &  $72.92 \pm 0.16$  &  $73.34 \pm 0.32$  &               $76.68 \pm 0.76$  &  $\vfirstSig{81.82} \pm 0.29$  &               $79.41 \pm 0.10$  &  $\vsecondSig{80.38} \pm 0.11$  &  $76.57 \pm 0.54$  \\
                                                  %& $7\%$  &  $75.11 \pm 0.10$  &  $75.58 \pm 0.19$  &               $78.95 \pm 0.36$  &  $\vfirstSig{82.34} \pm 0.30$  &               $79.99 \pm 0.10$  &  $\vsecondSig{80.59} \pm 0.07$  &  $78.03 \pm 0.55$  \\
                                                  & $9\%$  &  $76.58 \pm 0.04$  &  $77.06 \pm 0.21$  &               $79.91 \pm 0.40$  &  $\vfirstSig{82.65} \pm 0.20$  &               $80.64 \pm 0.11$  &  $\vsecondSig{80.82} \pm 0.04$  &  $79.58 \pm 0.60$  \\
                                                  & $15\%$ &  $79.16 \pm 0.08$  &  $79.68 \pm 0.16$  &               $81.95 \pm 0.23$  &  $\vfirstSig{83.12} \pm 0.20$  &  $\vsecondSig{82.02} \pm 0.09$  &               $81.16 \pm 0.09$  &  $81.71 \pm 0.40$  \\
                                                  & $20\%$ &  $80.29 \pm 0.12$  &  $80.99 \pm 0.11$  &  $\vsecondSig{82.89} \pm 0.14$  &  $\vfirstSig{83.26} \pm 0.09$  &               $82.82 \pm 0.09$  &               $81.64 \pm 0.14$  &  $82.73 \pm 0.35$  \\
                                                  & $25\%$ &  $81.08 \pm 0.08$  &  $81.86 \pm 0.05$  &  $\vsecondSig{83.45} \pm 0.10$  &  $\vfirstSig{83.66} \pm 0.14$  &               $83.39 \pm 0.08$  &               $81.94 \pm 0.15$  &  $83.42 \pm 0.16$  \\
& time (ms)   &              21.2 &              5.7 &                           655 &                            1918 &              77042 &              56252 &              78 \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\epi{}}} & $3\%$  &  $80.23 \pm 0.19$  &  $80.44 \pm 0.17$  &               $86.56 \pm 0.30$  &  $\vfirstSig{89.44} \pm 0.10$  &               $86.73 \pm 0.13$  &  $\vsecondSig{86.98} \pm 0.08$  &  $86.37 \pm 0.65$  \\
                                                  %& $5\%$  &  $83.60 \pm 0.14$  &  $83.96 \pm 0.10$  &  $\vsecondSig{88.09} \pm 0.25$  &  $\vfirstSig{90.28} \pm 0.08$  &               $87.59 \pm 0.10$  &               $87.35 \pm 0.06$  &  $86.64 \pm 0.66$  \\
                                                  %& $7\%$  &  $85.24 \pm 0.10$  &  $85.69 \pm 0.12$  &  $\vsecondSig{88.96} \pm 0.16$  &  $\vfirstSig{90.70} \pm 0.08$  &               $88.45 \pm 0.07$  &               $87.63 \pm 0.04$  &  $86.55 \pm 0.37$  \\
                                                  & $9\%$  &  $86.43 \pm 0.07$  &  $86.83 \pm 0.06$  &  $\vsecondSig{89.47} \pm 0.10$  &  $\vfirstSig{91.07} \pm 0.09$  &               $89.10 \pm 0.07$  &               $87.92 \pm 0.10$  &  $87.20 \pm 0.65$  \\
                                                  & $15\%$ &  $88.33 \pm 0.06$  &  $88.76 \pm 0.06$  &               $90.49 \pm 0.06$  &  $\vfirstSig{91.56} \pm 0.05$  &  $\vsecondSig{90.53} \pm 0.10$  &               $88.58 \pm 0.15$  &  $88.10 \pm 0.56$  \\
                                                  & $20\%$ &  $89.19 \pm 0.06$  &  $89.65 \pm 0.06$  &               $90.93 \pm 0.05$  &  $\vfirstSig{91.86} \pm 0.04$  &  $\vsecondSig{91.38} \pm 0.05$  &               $88.96 \pm 0.15$  &  $88.80 \pm 0.58$  \\
                                                  & $25\%$ &  $89.76 \pm 0.06$  &  $90.21 \pm 0.05$  &               $91.75 \pm 0.05$  &  $\vfirstSig{92.07} \pm 0.08$  &  $\vsecondSig{91.95} \pm 0.07$  &               $89.42 \pm 0.20$  &  $89.37 \pm 0.63$  \\
                                                  & time (ms)   &              32 &              7.1 &                            1226 &                           2341 &             116838 &             121530 &              129 \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\kiw{}}} & $3\%$  &  $73.91 \pm 0.37$  &  $73.93 \pm 0.58$  &  $80.44 \pm 1.05$  &  $86.95 \pm 0.19$  &  $\vsecondSig{87.72} \pm 0.32$  &               $87.02 \pm 0.05$  &   $\vfirstSig{87.88} \pm 0.01$  \\
                                                  %& $5\%$  &  $77.89 \pm 0.21$  &  $78.22 \pm 0.20$  &  $82.32 \pm 0.46$  &  $87.65 \pm 0.08$  &   $\vfirstSig{88.40} \pm 0.22$  &               $87.24 \pm 0.05$  &  $\vsecondSig{87.90} \pm 0.01$  \\
                                                  %& $7\%$  &  $79.65 \pm 0.20$  &  $80.34 \pm 0.15$  &  $83.43 \pm 0.23$  &  $86.72 \pm 0.73$  &   $\vfirstSig{88.71} \pm 0.13$  &               $87.37 \pm 0.05$  &  $\vsecondSig{87.53} \pm 0.35$  \\
                                                  & $9\%$  &  $80.87 \pm 0.09$  &  $81.61 \pm 0.09$  &  $83.97 \pm 0.27$  &  $86.53 \pm 0.20$  &   $\vfirstSig{89.02} \pm 0.07$  &  $\vsecondSig{87.49} \pm 0.07$  &               $87.23 \pm 0.38$  \\
                                                  & $15\%$ &  $82.57 \pm 0.07$  &  $83.38 \pm 0.06$  &  $85.03 \pm 0.11$  &  $86.97 \pm 0.29$  &   $\vfirstSig{89.67} \pm 0.05$  &  $\vsecondSig{87.76} \pm 0.04$  &               $87.12 \pm 0.33$  \\
                                                  & $20\%$ &  $83.34 \pm 0.06$  &  $84.11 \pm 0.09$  &  $85.48 \pm 0.13$  &  $87.02 \pm 0.24$  &   $\vfirstSig{90.00} \pm 0.07$  &  $\vsecondSig{87.92} \pm 0.07$  &               $87.12 \pm 0.21$  \\
                                                  & $25\%$ &  $83.82 \pm 0.08$  &  $84.56 \pm 0.06$  &  $85.81 \pm 0.11$  &  $87.02 \pm 0.16$  &   $\vfirstSig{90.29} \pm 0.04$  &  $\vsecondSig{88.15} \pm 0.07$  &               $87.24 \pm 0.20$  \\
                                                  & time (ms)   &                           28 &                           6.6 &                           824 &               2938 &                         103264 &             130036 &              104 \\
\bottomrule
\end{longtable}}
********************************************************
\fi

After training on the two features $1-\htr(i)$ and $1-\hun(j)$, \uslogregp{} has learned three weights
$w_0$, $w_1$ and $w_2$, which allow to predict $y_{i,j}$ according to 
\[ 
  \sgn\Big(\big(w_1(1-\htr(i)\big) + w_2 \big(1-\hun(j)\big) + w_0\Big)~.
\]
This can be rewritten as 
\[
\sgn\Big(\big(1-\htr(i)\big) + w_2'\big(1-\hun(j)\big) - \tfrac{1}{2} -\tau'\Big)~,
\]
with $w_2' = \frac{w_2}{w_1}$ and $\tau' = - \left(\frac{1}{2} + \frac{w_0}{w_1}\right)$~.

As shown in Table~\ref{tab:coeff}, and in accordance with the predictor built out of Equation~\eqref{eq:predictor},
% from the main paper,
$w_2'$ is almost $1$ on {\em all datasets}, while $\tau'$ tends to be always close the fraction of positive edges in the dataset.

\begin{table}[bt]
  \centering
  \caption{Normalized logistic regression coefficients averaged over 12 runs (with one standard deviation) \label{tab:coeff}}
  \small
\begin{tabular}{lrccc}
\toprule
                        & $\frac{|\trainset{}|}{|E|}$ &              $w_2'$ & $\tau'$ \\
% \multirow{5}{*}{\aut{}} & $3\%$  &  $0.909 \pm 0.07$  &  $0.592 \pm 0.04$  \\
%                         & $9\%$  &  $0.985 \pm 0.04$  &  $0.707 \pm 0.03$  \\
\multirow{5}{*}{\aut{}} & $5\%$  &  $0.965 \pm 0.04$  &  $0.662 \pm 0.03$  \\
                        & $10\%$ &  $0.983 \pm 0.03$  &  $0.705 \pm 0.02$  \\
                        & $15\%$ &  $1.001 \pm 0.03$  &  $0.729 \pm 0.03$  \\
                        & $20\%$ &  $1.013 \pm 0.02$  &  $0.747 \pm 0.02$  \\
                        & $25\%$ &  $1.011 \pm 0.02$  &  $0.746 \pm 0.01$  \\
\midrule
% \multirow{5}{*}{\wik{}} & $3\%$  &  $0.919 \pm 0.04$  &  $0.666 \pm 0.02$  \\
%                         & $9\%$  &  $0.940 \pm 0.03$  &  $0.726 \pm 0.01$  \\
\multirow{5}{*}{\wik{}} & $5\%$  &  $0.920 \pm 0.02$  &  $0.691 \pm 0.02$  \\
                        & $10\%$ &  $0.940 \pm 0.01$  &  $0.730 \pm 0.01$  \\
                        & $15\%$ &  $0.947 \pm 0.01$  &  $0.741 \pm 0.01$  \\
                        & $20\%$ &  $0.963 \pm 0.01$  &  $0.760 \pm 0.01$  \\
                        & $25\%$ &  $0.962 \pm 0.02$  &  $0.764 \pm 0.01$  \\
\midrule
% \multirow{5}{*}{\sla{}} & $3\%$  &  $1.012 \pm 0.02$  &  $0.669 \pm 0.01$  \\
%                         & $9\%$  &  $1.014 \pm 0.01$  &  $0.703 \pm 0.01$  \\
\multirow{5}{*}{\sla{}} & $5\%$  &  $1.024 \pm 0.02$  &  $0.693 \pm 0.01$  \\
                        & $10\%$ &  $1.017 \pm 0.01$  &  $0.705 \pm 0.01$  \\
                        & $15\%$ &  $1.007 \pm 0.01$  &  $0.707 \pm 0.01$  \\
                        & $20\%$ &  $1.002 \pm 0.00$  &  $0.710 \pm 0.00$  \\
                        & $25\%$ &  $0.995 \pm 0.01$  &  $0.712 \pm 0.00$  \\
\midrule
% \multirow{5}{*}{\epi{}} & $3\%$  &  $1.124 \pm 0.02$  &  $0.786 \pm 0.02$  \\
%                         & $9\%$  &  $1.066 \pm 0.01$  &  $0.783 \pm 0.01$  \\
\multirow{5}{*}{\epi{}} & $5\%$  &  $1.099 \pm 0.02$  &  $0.791 \pm 0.02$  \\
                        & $10\%$ &  $1.059 \pm 0.01$  &  $0.782 \pm 0.01$  \\
                        & $15\%$ &  $1.037 \pm 0.01$  &  $0.774 \pm 0.01$  \\
                        & $20\%$ &  $1.018 \pm 0.01$  &  $0.765 \pm 0.01$  \\
                        & $25\%$ &  $1.010 \pm 0.01$  &  $0.763 \pm 0.01$  \\
\midrule
% \multirow{5}{*}{\kiw{}} & $3\%$  &  $1.075 \pm 0.02$  &  $0.851 \pm 0.02$  \\
%                         & $9\%$  &  $1.041 \pm 0.01$  &  $0.872 \pm 0.01$  \\
\multirow{5}{*}{\kiw{}} & $5\%$  &  $1.047 \pm 0.02$  &  $0.853 \pm 0.01$  \\
                        & $10\%$ &  $1.038 \pm 0.01$  &  $0.872 \pm 0.01$  \\
                        & $15\%$ &  $1.025 \pm 0.01$  &  $0.876 \pm 0.01$  \\
                        & $20\%$ &  $1.012 \pm 0.01$  &  $0.874 \pm 0.01$  \\
                        & $25\%$ &  $1.007 \pm 0.01$  &  $0.874 \pm 0.01$  \\
\bottomrule
  \end{tabular}
\end{table}

### 2017-09-02 troll/introduction.tex
small pass over introduction

@@
 % novel and cool
-%
-Connections in social networks are mostly driven by the {\em homophily} assumption: linked individuals tend to be similar, sharing personality traits, attitudes, or interests.
-However, homophily alone is clearly not sufficient to explain the variety of social links.
-In fact, sociologists have long studied networks, hereafter called \emph{signed} social networks, where also {\em negative} relationships ---like dissimilarity, disapproval or distrust--- are explicitly displayed.
-The presence of negative relationships is also a feature of many technology-mediated social networks.
-Known examples are \textsc{Ebay}, where users trust or distrust agents in the network based on their personal interactions, \textsc{Slashdot}, where each user can tag another user as friend or foe, and \textsc{Epinion}, where users can rate positively or negatively not only products, but also other users.
-Even in social networks where connections solely represent friendships, negative links can still emerge from the analysis of online debates among users.
+
+Connections in social networks are mostly driven by the \emph{homophily} assumption: linked
+individuals tend to be similar, sharing personality traits, attitudes, or interests. However,
+homophily alone is clearly not sufficient to explain the variety of social links\Todo{link with
+chapter 3}. In fact, sociologists have long studied networks, hereafter called \emph{signed} social
+networks\Todo{DSSN?}, where also {\em negative} relationships ---like dissimilarity, disapproval or
+distrust--- are explicitly displayed. The presence of negative relationships is also a feature of
+many technology-mediated social networks. Known examples are \textsc{Ebay}, where users trust or
+distrust agents in the network based on their personal interactions, \textsc{Slashdot}, where each
+user can tag another user as friend or foe, and \textsc{Epinion}, where users can rate positively
+or negatively not only products, but also other users. Even in social networks where connections
+solely represent friendships, negative links can still emerge from the analysis of online debates
+among users.
 
@@
 Or, the new user could try and disrupt the community by engaging into anti-social behavior, and creating conflictual relationships with other members.
-This behavioral dichotomy is supported by decades of research in psychology, starting with the seminal work~\cite{Dissonance57} about cognitive dissonance (when someone is acting in contradiction with his/her personal beliefs, ideas, or values).
-This kind of attitude expressed publicly on social media leads to the definition of \enquote{trolls}~\cite{Hardaker10}: users whose real intentions are to cause disruption and/or to trigger or exacerbate conflict for the purposes of their own amusement.
-The paper~\cite{Shachaf10} elaborates on their motives, adding that boredom, attention seeking, and revenge motivate trolls; they find pleasure from causing damage to other people or to the community as a whole.
+This behavioral dichotomy is supported by decades of research in psychology, starting with the seminal work~\autocite{Dissonance57} about cognitive dissonance (when someone is acting in contradiction with his/her personal beliefs, ideas, or values).
+This kind of attitude expressed publicly on social media leads to the definition of \enquote{trolls}~\autocite{Hardaker10}: users whose real intentions are to cause disruption and/or to trigger or exacerbate conflict for the purposes of their own amusement.
+The paper~\autocite{Shachaf10} elaborates on their motives, adding that boredom, attention seeking, and revenge motivate trolls; they find pleasure from causing damage to other people or to the community as a whole.
 
 When the social network is signed, specific challenges arise in both network analysis and learning.
-On the one hand, novel methods are required to tackle standard tasks (e.g., user clustering, link prediction, targeted advertising/recommendation,
-%prediction of user interests, and 
-analysis of the spreading of diseases in epidemiological models).
-On the other hand, new problems such as edge sign prediction, which we consider here, naturally emerge.
-Edge sign prediction is the problem of classifying the positive or negative nature of the links based on the network topology.
-Prior knowledge of the network topology is often a realistic assumption, for in several situations the discovery of the link sign can be more costly than acquiring the topological information of the network.
-For instance, when two users of an online social network communicate on a public web page, we immediately detect a link.
-Yet, the classification of the link sign as positive or negative may require complex techniques.
+On the one hand, novel methods are required to tackle standard tasks (e.g., user clustering, link
+prediction, targeted advertising/recommendation, prediction of user interests, and analysis of the
+spreading of diseases in epidemiological models). On the other hand, new problems such as edge sign
+prediction, which we consider here, naturally emerge. Edge sign prediction is the problem of
+classifying the positive or negative nature of the links based on the network topology. Prior
+knowledge of the network topology is often a realistic assumption, for in several situations the
+discovery of the link sign can be more costly than acquiring the topological information of the
+network. For instance, when two users of an online social network communicate on a public web page,
+we immediately detect a link. Yet, the classification of the link sign as positive or negative may
+require complex techniques.
 
-From the modeling and algorithmic viewpoints, because of the huge amount of available networked data, a major concern in developing learning methods for edge sign prediction is algorithmic scalability.
-Many successful, yet simple heuristics for edge sign prediction are based on the troll-trust features, i.e., on the fraction of outgoing negative links (trollness) and incoming positive links (trustworthiness) at each node.
-We study such heuristics by defining a probabilistic generative model for the signs on the directed links of a given network, and show that these heuristics can be understood and analyzed as approximators to the Bayes optimal classifier for our generative model.
-We also gather empirical evidence supporting our probabilistic model by observing that a logistic model trained on trollness and trustworthiness features alone is able to learn weights that, on all datasets considered in our experiments, consistently satisfy the properties predicted by our model.
+From the modeling and algorithmic viewpoints, because of the huge amount of available networked
+data, a major concern in developing learning methods for edge sign prediction is algorithmic
+scalability. Many successful, yet simple heuristics for edge sign prediction are based on the
+troll-trust features, i.e., on the fraction of outgoing negative links (trollness) and incoming
+positive links (trustworthiness) at each node. We study such heuristics by defining a probabilistic
+generative model for the signs on the directed links of a given network, and show that these
+heuristics can be understood and analyzed as approximators to the Bayes optimal classifier for our
+generative model. We also gather empirical evidence supporting our probabilistic model by observing
+that a logistic model trained on trollness and trustworthiness features alone is able to learn
+weights that, on all datasets considered in our experiments, consistently satisfy the properties
+predicted by our model.
 
-We then introduce suitable graph transformations defining reductions from edge sign prediction to node sign prediction problems.
-This opens up the possibility of using the arsenal of known algorithmic techniques developed for node classification.
-In particular, we show that a Label Propagation algorithm, combined with our reduction, approximates the maximum likelihood estimator of our probabilistic generative model.
-Experiments on real-world data show the competitiveness of our approach in terms of both prediction performance (especially in the regime when training data are scarce) and scalability.
+We then introduce suitable graph transformations defining reductions from edge sign prediction to
+node sign prediction problems. This opens up the possibility of using the arsenal of known
+algorithmic techniques developed for node classification. In particular, we show that a Label
+Propagation algorithm, combined with our reduction, approximates the maximum likelihood estimator
+of our probabilistic generative model. Experiments on real-world data show the competitiveness of
+our approach in terms of both prediction performance (especially in the regime when training data
+are scarce) and scalability.
 
-Finally, we point out that the notions of trollness and trustworthiness naturally define a measure of complexity, or learning bias, for the signed network that can also be used to design {\em online} (i.e., sequential) learning algorithms for the edge sign prediction problem.
-The learning bias encourages settings where the nodes in the network have polarized features (e.g., trollness/trustworthiness are either very high or very low).
-Our online analysis holds under adversarial conditions, namely, without any stochastic assumption on the assignment of signs to the network links.
+Finally, we point out that the notions of trollness and trustworthiness naturally define a measure
+of complexity, or learning bias, for the signed network that can also be used to design
+\emph{online} (i.e., sequential) learning algorithms for the edge sign prediction problem. The
+learning bias encourages settings where the nodes in the network have polarized features (e.g.,
+trollness/trustworthiness are either very high or very low). Our online analysis holds under
+adversarial conditions, namely, without any stochastic assumption on the assignment of signs to the
+network links.
 
@@
 \subsection{Related work}
-%
-Interest in signed networks can be traced back to the psychological theory of structural balance~\cite{Cartwright56,HeiderBook58} with its weak version~\cite{davis1967clustering}.
-The advent of online signed social networks has enabled a more thorough and quantitative understanding of that phenomenon.
-Among the several approaches related to our work, some extend the spectral properties of a graph to the signed case in order to find good embeddings for classification~\cite{Kunegis2009,SignedEmbedding15}.
-However, the use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which makes those methods hardly scalable to large graphs.
-Another approach is based on mining ego networks with SVM.
-Although this method seems to deliver good results~\cite{Papaoikonomou2014}, the running time makes it often impractical for large real-world datasets.
-An alternative approach, based on local features only and proposed in~\cite{Leskovec2010}, relies on the so-called status theory for directed graphs~\cite{guha2004propagation}.
-Some works in active learning, using a more sophisticated bias based on the correlation clustering (CC) index~\cite{Cesa-Bianchi2012a,Cesa-Bianchi2012b}, provide strong theoretical guarantees.
-However, the bias used there is rather strong, since it assumes the existence of a $2$-clustering of the nodes with a small CC index.
 
-Whereas our focus will be on {\em binary} prediction, researchers have also considered a weighted version of the problem, where edges measure the amount of trust or distrust between two users~(e.g., \cite{guha2004propagation,Tang2013,
-%Bachi2012,
-Qian2014sn}).
-Other works have also considered versions of the problem where side information related to the network is available to the learning system.
-For instance, \cite{EdgeSignsRating15} uses the product purchased on Epinion in conjunction with a neural network, \cite{TrollDetection15} identifies trolls by analysing the textual content of their post, and~\cite{SNTransfer13} uses SVM to perform transfer learning from one network to another.
-While many of these approaches have interesting performances, they often require extra information which is not always available (or reliable) and, in addition, may face severe scaling issues.
-% are not as scalable as ours. 
-The recent survey~\cite{Tang2015a} contains pointers to many papers on edge sign prediction for signed networks, especially in the Data Mining area.
-Additional references, more closely related to our work, will be mentioned at the end of Section~\ref{ss:passive}.
-% using the notation we now introduce. 
-%for further references to the link classification problem.
+Interest in signed networks can be traced back to the psychological theory of structural
+balance~\autocites{Cartwright56}{HeiderBook58} with its weak
+version~\autocite{davis1967clustering}.  The advent of online signed social networks has enabled a
+more thorough and quantitative understanding of that phenomenon. Among the several approaches
+related to our work, some extend the spectral properties of a graph to the signed case in order to
+find good embeddings for classification~\autocites{Kunegis2009}{SignedEmbedding15}. However, the
+use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which
+makes those methods hardly scalable to large graphs. Another approach is based on mining ego
+networks with SVM. Although this method seems to deliver good results~\autocite{Papaoikonomou2014},
+the running time makes it often impractical for large real-world datasets. An alternative approach,
+based on local features only and proposed in~\autocite{Leskovec2010}, relies on the so-called
+status theory for directed graphs~\autocite{guha2004propagation}. Some works in active learning,
+using a more sophisticated bias based on the correlation clustering (CC)
+index~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}, provide strong theoretical guarantees.
+However, the bias used there is rather strong, since it assumes the existence of a $2$-clustering
+of the nodes with a small CC index.\Todo{don't put it that way since we use that bias later}
+
+Whereas our focus will be on \emph{binary} prediction, researchers have also considered a weighted
+version of the problem, where edges measure the amount of trust or distrust between two
+users~\autocites{guha2004propagation}{Tang2013}{Bachi2012}{Qian2014sn}. Other works have also
+considered versions of the problem where side information related to the network is available to
+the learning system. For instance, \autocite{EdgeSignsRating15} uses the product purchased on
+Epinion in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
+analysing the textual content of their post, and~\autocite{SNTransfer13} uses SVM to perform
+transfer learning from one network to another. While many of these approaches have interesting
+performances, they often require extra information which is not always available (or reliable) and,
+in addition, may face severe scaling issues.
+
+The survey~\autocite{Tang2015a} contains pointers to many papers on edge sign prediction for
+signed networks, especially in the Data Mining area. Additional references, more closely related to
+our work, will be mentioned at the end of \autoref{ss:passive}.

### 2017-09-03 troll/notation_and_prel.tex
small pass over notations

@@
 \section{Notation and Preliminaries}\label{s:prel}
-%
-In what follows, we let $G=(V,E)$ be a {\em directed} graph, whose edges $(i,j) \in E$ carry a binary label $y_{i,j}\in \spin$.
-The edge labeling will sometimes be collectively denoted by the $|V|\times |V|$ matrix $Y = [Y_{i,j}]$, where $Y_{i,j} = y_{i,j}$ if $(i,j) \in E$, and $Y_{i,j} = 0$, otherwise.
-The corresponding edge-labeled graph will be denoted by $G(Y) = (V,E(Y))$.
-We use $\Nin(i)$ and $\Nout(i)$ to denote, respectively, the set of edges incoming to and outgoing from node $i \in V$, with $\din(i) = \big|\Nin(i)\big|$ and $\dout(i) = \big|\Nout(i)\big|$ being the in-degree and the out-degree of $i$.
-Moreover, $\din^+(i)$ is the number of edges $(k,i)\in\Nin(i)$ such that $y_{k,i} = +1$.
-We define $\din^-(i)$, $\dout^+(i)$, and $\dout^-(i)$ similarly, so that, for instance, $\dout^-(i)/\dout(i)$ is the fraction of outgoing edges from node $i$ whose label in $G(Y)$ is $-1$.
-We call $tr(i) = \dout^-(i)/\dout(i)$ the {\em trollness} of node $i$, and $un(i) = \din^-(i)/\din(i)$ the {\em untrustworthiness} of node $i$.
-Finally, we also use the notation $\NNin(i)$ and $\NNout(i)$ to represent, respectively, the in-neighborhood and the out-neighborhood of node $i \in V$.
 
-Given the directed graph $G = (V,E)$, we define two {\em edge-to-node reductions} transforming the original graph $G$ into other graphs.
-As we see later, these reductions are useful in turning the edge sign prediction problem into a {\em node} sign prediction problem (often called node classification problem), for which many algorithms are indeed available
-%---see, e.g., \cite{BC01,LabelPropa03,BDL06,HP07,HLP09,VCGZ12,HPV12,WTA13,HPG15}.
----see, e.g., \cite{BC01,LabelPropa03,HP07,HLP09,WTA13}.
-% In fact, when facing a specific edge sign prediction problem, we may find it convenient to operate on the {\em nodes} of the transformed graph, rather than on the {\em edges} of the original graph $G$.
-Although any node classification method could in principle be used, the reductions we describe next are essentially aimed at preparing the ground for quadratic energy-minimization approaches computed through a {\em Label Propagation} algorithm~(e.g., \cite{LabelPropa03,BDL06}).
+In what follows, we let $G=(V,E)$ be a \emph{directed} graph, whose edges $(u,v) \in E$ carry a
+binary label $\yuv\in \spin$. The edge labeling will sometimes be collectively denoted by the
+$|V|\times |V|$ matrix $Y = [\yuv]$, where $\yuv = \yuv$ if $(u,v) \in E$, and $\yuv = 0$,
+otherwise. The corresponding edge-labeled graph will be denoted by $G(Y) = (V,E(Y))$. We use
+$\Nin(u)$ and $\Nout(u)$ to denote, respectively, the set of edges incoming to and outgoing from
+node $u \in V$, with $\din(u) = \big|\Nin(u)\big|$ and $\dout(u) = \big|\Nout(u)\big|$ being the
+in-degree and the out-degree of $u$. Moreover, $\din^+(u)$ is the number of edges $(k,u)\in\Nin(u)$
+such that $y_{k,u} = +1$. We define $\din^-(u)$, $\dout^+(u)$, and $\dout^-(u)$ similarly, so that,
+for instance, $\dout^-(u)/\dout(u)$ is the fraction of outgoing edges from node $u$ whose label in
+$G(Y)$ is $-1$. We call $tr(u) = \dout^-(u)/\dout(u)$ the \emph{trollness} of node $u$, and $un(u)
+= \din^-(u)/\din(u)$ the \emph{untrustworthiness} of node $u$. Finally, we also use the notation
+$\NNin(u)$ and $\NNout(u)$ to represent, respectively, the in-neighborhood and the out-neighborhood
+of node $u \in V$.
 
-The first reduction, called $G \rightarrow G'$, builds an {\em undirected} graph $G' = (V',E')$ as follows.
-Each node $i \in V$ has two copies in $V'$, call them $\iin$ and $\iout$.
-Each directed edge $(i,j)$ in $E$ is associated with one node, call it $e_{i,j}$, in $V'$, along with the two undirected edges $(\iout,e_{i,j})$ and  $(e_{i,j},\jin)$.
-Hence $|V'| = 2|V|+|E|$ and $|E'| = 2|E|$.
-Moreover, if $G = G(Y)$ is edge labeled, then this labeling transfers to the subset of nodes $e_{i,j} \in V'$, so that $G'$ is a graph $G'(Y) = (V'(Y),E')$ with partially-labeled nodes.
-%For convenience (see Section~\ref{s:algbatch}), these labels are also mapped to $\{0,1\}$ through the mapping $y \rightarrow \frac{1+y}{2}$. 
-The second reduction, called $G \rightarrow G''$, builds an {\em undirected and weighted} graph $G'' = (V'',E'')$.
-%whose weights depend on a small value $\epsilon > 0$. 
-Specifically, we have $V'' \equiv V'$ and $E'' \supset E'$, where the set $E''$ also includes edges $(\iout,\jin)$ for all $i$ and $j$ such that $(i,j) \in E$.
-The edges in $E'$ have weight $2$, whereas the edges in $E''\setminus E'$ have weight $-1$.
-% As we explain in \S~\ref{ss:passive}, to run Label Propagation on $G''$, we set the weight of edges in $E'$ to $2$ and those in $E''\setminus E'$ to $-1$.
-Finally, as in the $G \rightarrow G'$ reduction, if $G = G(Y)$ is edge labeled, then this labeling transfers to the subset of nodes $e_{i,j} \in V''$.
-Graph $G'$, which will not be used in this paper, is an intermediate structure between $G$ and $G''$ and provides a conceptual link to the standard cutsize measure in node sign classification.
-%in the online setting.
-%, again after mapping to $\{0,1\}$. 
-%, but in addition each label $y_{i,j}$ gets magnified by $2/\epsilon$. 
-%Hence $G''$ is a weighted graph $G''(Y) = (V''(Y),E'')$ with partially-labeled nodes whose labels are in $\{0,2/\epsilon\}$. 
-Figure~\ref{f:etnr} illustrates the two reductions.
+Given the directed graph $G = (V,E)$, we define two \emph{edge-to-node reductions} transforming the
+original graph $G$ into other graphs. As we see later, these reductions are useful in turning the
+edge sign prediction problem into a \emph{node} sign prediction problem (often called node
+classification problem), for which many algorithms are indeed available~%
+\autocites{BC01}{LabelPropa03}{BDL06}{HP07}{HLP09}{Vitale2012}{HPV12}{WTA13}{HPG15}.  Although any
+node classification method could in principle be used, the reductions we describe next are
+essentially aimed at preparing the ground for quadratic energy-minimization approaches computed
+through a \emph{Label Propagation} algorithm~\autocites{LabelPropa03}{BDL06}.
 
-%Moreover, given the outcome $G'$ of an edge-to-node reduction, and a balancedness parameter $b \in [0,1]$, we say that $G'_b$ is a randomly sparsified version of $G'$, when 
-%$G'_b$ is the random subgraph produced from $G'$ as follows.
-Given any pairing of edges $(\iout,e_{i,j})$ and $(e_{i,j},\jin)$ of $G'$, draw an independent Bernoulli variable $B_{i,j}$ such that $\Pr(B_{i,j} = 1) = b$; if $B_{i,j} = 1$ then drop $(e_{i,j},\jin)$, otherwise drop $(\iout,e_{i,j})$ -- see Figure \ref{f:etnr}, right.
-
-%
-%\begin{figure}[t*]
-%\begin{picture}(-10,242)(-10,242)
-%\begin{tabular}{l}
-%\includegraphics[width=0.95\textwidth]{etnr3.pdf}
-%\end{tabular}
-%\end{picture}
+The first reduction, called $G \rightarrow G'$, builds an \emph{undirected} graph $G' = (V',E')$ as
+follows. Each node $u \in V$ has two copies in $V'$, call them $\iin$ and $\iout$. Each directed
+edge $(u,v)$ in $E$ is associated with one node, call it $e_{u,v}$, in $V'$, along with the two
+undirected edges $(\iout,e_{u,v})$ and $(e_{u,v},\jin)$. Hence $|V'| = 2|V|+|E|$ and $|E'| = 2|E|$.
+Moreover, if $G = G(Y)$ is edge labeled, then this labeling transfers to the subset of nodes
+$e_{u,v} \in V'$, so that $G'$ is a graph $G'(Y) = (V'(Y),E')$ with partially-labeled nodes. The
+second reduction, called $G \rightarrow G''$, builds an \emph{undirected and weighted} graph $G'' =
+(V'',E'')$. Specifically, we have $V'' \equiv V'$ and $E'' \supset E'$, where the set $E''$ also
+includes edges $(\iout,\jin)$ for all $u$ and $v$ such that $(u,v) \in E$. The edges in $E'$ have
+weight $2$, whereas the edges in $E''\setminus E'$ have weight $-1$. We provide in
+\autoref{ssec:troll_lprop_weights} an analytic justification for this weights choice in the context
+of Label Propagation. Finally, as in the $G \rightarrow G'$ reduction, if $G = G(Y)$ is edge
+labeled, then this labeling transfers to the subset of nodes $e_{u,v} \in V''$. Graph $G'$, which
+will not be used in this paper, is an intermediate structure between $G$ and $G''$ and provides a
+conceptual link to the standard cutsize measure in node sign classification. \autoref{f:etnr}
+illustrates these two reductions.
 
@@
   \begin{subfigure}[b]{0.30\textwidth}
-    \centering \includegraphics[height=.12\textheight]{gprime_latex-crop} \caption{}
+    \centering \includegraphics[height=.12\textheight]{gprime_latex-crop}
+    \caption{\label{fig:troll_reduction_gprime}}
   \end{subfigure}~
@@
   \caption{\label{f:etnr}
-{\bf (a)} A directed edge-labeled graph $G$.
-{\bf (b)} Its corresponding graph $G'$ resulting from the $G\rightarrow G'$ reduction.
-The square nodes in $G'$ correspond to the edges in $G$, and carry the same labels as their corresponding edges.
-On the other hand, the $2|V|$ circle nodes in $G'$ are unlabeled.
-Observe that some nodes in $G'$ are isolated (and thus unimportant); these are exactly the nodes in $G'$ corresponding to the nodes having in $G$ no outgoing or no incoming edges ---see, e.g., nodes $3$ and $4$ in $G$.
-{\bf (c)} The weighted graph resulting from the $G\rightarrow G''$ reduction.
+    \textbf{(a)} A directed edge-labeled graph $G$. \textbf{(b)} Its corresponding graph $G'$
+    resulting from the $G\rightarrow G'$ reduction. The square nodes in $G'$ correspond to the
+    edges in $G$, and carry the same labels as their corresponding edges. On the other hand, the
+    $2|V|$ circle nodes in $G'$ are unlabeled. Observe that some nodes in $G'$ are isolated and
+    thus unimportant. These are exactly the nodes in $G'$ corresponding to the nodes having in $G$
+    no outgoing or no incoming edges: for instance nodes $3$ and $4$ in $G$. \textbf{(c)} The
+    weighted graph resulting from the $G\rightarrow G''$ reduction.
   }
@@
 
-These reductions are meaningful only if they are able to approximately preserve 
-% What is important here is to make sure that such reductions are able to (approximately) preserve
-label {\em regularity} when moving from edges to nodes.
-That is, if the edge sign prediction problem is easy for a given $G(Y) = (V,E(Y))$, then the corresponding node sign prediction problems on $G'(Y) = (V'(Y),E')$ and $G''(Y) = (V''(Y),E)$ are also easy, and vice versa.
-% We show this happens in the three learning settings within which we consider the
-While we could make this argument more quantitative, here we simply observe that if each node in $G$ tends to be either troll or trustworthy, then few labels from the incoming and outgoing edges of each such node are sufficient to predict the labels on the remaining edges in $G$, and this translates to a small cutsize\footnote
-{
-Recall that the cutsize of an undirected node-labeled graph $G'(Y)$ is the number of edges in $G'$ connecting nodes having mismatching labels. 
-}
-of $G'(Y)$ over the nodes corresponding to the edges in $G$ (the colored squares in Figure \ref{f:etnr} (b)).
-%
-Again, we would like to point out that these reductions serve two purposes: %\textcolor{red}{
-First, they allow us to use the many algorithms designed for the better studied problem of node sign prediction.
-Second, the reduction $G\rightarrow G''$ with the specific choice of edge weights is designed to make the Label Propagation solution approximate the maximum likelihood estimator associated with our generative model (see Section~\ref{ss:passive}).%}
-Note also that efficient Label Propagation implementations exist that can leverage the sparsity of $G''$.
+These reductions are meaningful only if they are able to approximately preserve label
+\emph{regularity} when moving from edges to nodes. That is, if the edge sign prediction problem is
+easy for a given $G(Y) = (V,E(Y))$, then the corresponding node sign prediction problems on $G'(Y)
+= (V'(Y),E')$ and $G''(Y) = (V''(Y),E)$ are also easy, and vice versa.  While we could make this
+argument more quantitative, here we simply observe that if each node in $G$ tends to be either
+troll or trustworthy, then few labels from the incoming and outgoing edges of each such node are
+sufficient to predict the labels on the remaining edges in $G$, and this translates to a small
+cutsize\footnote{Recall that the cutsize of an undirected node-labeled graph $G'(Y)$ is the number
+of edges in $G'$ connecting nodes having mismatching labels.} of $G'(Y)$ over the nodes
+corresponding to the edges in $G$ (the colored squares in \autoref{fig:troll_reduction_gprime}).
 
-We consider two learning settings associated with the problem of edge sign prediction: a batch setting and an online setting.
-In the batch setting, we assume that a training set of edges $E_0$ has been drawn uniformly at random {\em without replacement} from $E$, we observe the labels in $E_0$, and we are interested in predicting the sign of the remaining edges $E \setminus E_0$ by making as few prediction mistakes as possible.
-% In the batch {\em active} setting, we are given a budget of edge labels to observe, and are free to select them the way we want within $E(Y)$.
-% Again, the goal is to make as few mistakes as possible on the remaining edges.
-%
-%Unlike the online setting, where the underlying labels are generated by an adaptive adversary, 
-%
-The specific batch setting 
-%(passive and active) 
-we study here assumes that labels are produced by a generative model which we describe in the next section, and our label regularity measure is a quadratic function (denoted by $\Psi^2_{G''}(Y)$ ---see Section~\ref{s:exp} for a definition), related to this model.
-$\Psi^2_{G''}(Y)$ is small just when all nodes in $G$ tend to be either troll or trustworthy.
+Again, we would like to point out that these reductions serve two purposes: First, they allow us to
+use the many algorithms designed for the better studied problem of node sign prediction. Second,
+the reduction $G\rightarrow G''$ with the specific choice of edge weights is designed to make the
+Label Propagation solution approximate the maximum likelihood estimator associated with our
+generative model (see Section~\ref{ss:passive}). Note also that efficient Label Propagation
+implementations exist that can leverage the sparsity of $G''$.
 
-On the other hand, the {\em online} setting we consider is the standard mistake bound model of online learning~\cite{Winnow88} where all edge labels are assumed to be generated by an adversary and sequentially presented to the learner according to an arbitrary permutation.
-For an online learning algorithm $A$, we are interested in measuring the total number of mistakes $M_A(Y)$ the algorithm makes over $G(Y)$ when the worst possible presentation order of the edge labels in $Y$ is selected by the adversary.
-%---the reader is referred, e.g., to~\cite[Section~3]{Cesa-Bianchi2012b} for definitions specific to the edge sign prediction problem.
-% In this setting, learning is split into a sequence of rounds: At each round $t=1,2,\dots$ a learning algorithm $A$ outputs a prediction ${\wh y_t} \in \spin$ for the label $y_t \in \spin$ of an edge arbitrarily selected by the adversary.
-After the prediction is made, label $y_t$ is revealed to the algorithm, hence allowing it to change its internal state.
-In the next round a new edge is selected by the adversary, and so on.
-Notice that
-%since the underlying labeling over the edges is decided by the adversary once and for all,\footnote
-%{
-%For simplicity, we assume the adversary is deterministic.
-%} 
-%all edges occur exactly once within the sequence to be predicted (so that this game lasts exactly $|E|$ rounds).
-The adversary decides both the underlying labeling $Y$ over the edges of $G$ and the order of their presentation to the learning algorithm.
-We say that $A$ has made a mistake at time $t$ if ${\wh y_t} \neq y_t$, and we measure $A$'s prediction performance simply through the total number of mistakes $M_A(Y)$ it makes over $G(Y)$ when the worst possible presentation order of the edge labels in $Y$ is selected by the adversary.
-As is standard, we contrast $M_A(Y)$ to some kind of {\em regularity} measure $\Psi_G(Y)$ of the labeling $Y$ over $G$, so that we are in fact aimed at bounding the (cumulative) regret
-%\(
-%M_A(Y) - \Psi_G(Y)~.
-%\)
-Also in the online setting our label regularity measure, denoted here by $\Psi_G(Y)$, is small when nodes in $G$ tend to be either troll or trustworthy. 
-%When this happens, few labels from the incoming and outgoing edges of each node are sufficient to predict the labels on the remaining edges. 
-Formally, for fixed $G$ and $Y$, let 
-$
-    \Psiin(j,Y) = \min\big\{\din^-(j),\din^+(j)\big\}
-$ and
-%\qquad\text{and}\qquad
-$
-    \Psiout(i,Y) = \min\big\{\dout^-(i),\dout^+(i)\big\}
-$.
+We consider two learning settings associated with the problem of edge sign prediction: a batch
+setting and an online setting. In the batch \emph{passive} setting, we assume that a training set
+of edges $E_0$ has been drawn uniformly at random \emph{without replacement} from $E$, we observe
+the labels in $E_0$, and we are interested in predicting the sign of the remaining edges $E
+\setminus E_0$ by making as few prediction mistakes as possible. In the batch \emph{active}
+setting, we are given a budget of edge labels to observe, and are free to select them the way we
+want within $E(Y)$. Again, the goal is to make as few mistakes as possible on the remaining edges.
+The specific batch setting (passive and active) we study here assumes that labels are produced by a
+generative model which we describe in the next section, and our label regularity measure is a
+quadratic function (denoted by $\Psi^2_{G''}(Y)$ and defined in \autoref{s:exp}), related to this
+model. $\Psi^2_{G''}(Y)$ is small just when all nodes in $G$ tend to be either troll or
+trustworthy.
+
+On the other hand, the \emph{online} setting we consider is the standard mistake bound model of
+online learning~\autocite{Winnow88} where all edge labels are assumed to be generated by an
+adversary and sequentially presented to the learner according to an arbitrary permutation. For an
+online learning algorithm $A$, we are interested in measuring the total number of mistakes $M_A(Y)$
+the algorithm makes over $G(Y)$ when the worst possible presentation order of the edge labels in
+$Y$ is selected by the adversary.
+
+\iffalse
+  In this setting, learning is split into a sequence of rounds: At each round $t=1,2,\dots$ a
+  learning algorithm $A$ outputs a prediction ${\wh y_t} \in \spin$ for the label $y_t \in \spin$
+  of an edge arbitrarily selected by the adversary. After the prediction is made, label $y_t$ is
+  revealed to the algorithm, hence allowing it to change its internal state. In the next round a
+  new edge is selected by the adversary, and so on. Notice that since the underlying labeling over
+  the edges is decided by the adversary once and for all,\footnote{For simplicity, we assume the
+  adversary is deterministic.} all edges occur exactly once within the sequence to be predicted (so
+  that this game lasts exactly $|E|$ rounds). The adversary decides both the underlying labeling
+  $Y$ over the edges of $G$ and the order of their presentation to the learning algorithm. We say
+  that $A$ has made a mistake at time $t$ if ${\wh y_t} \neq y_t$, and we measure $A$'s prediction
+  performance simply through the total number of mistakes $M_A(Y)$ it makes over $G(Y)$ when the
+  worst possible presentation order of the edge labels in $Y$ is selected by the adversary. As is
+  standard, we contrast $M_A(Y)$ to some kind of \emph{regularity} measure $\Psi_G(Y)$ of the
+  labeling $Y$ over $G$, so that we are in fact aimed at bounding the (cumulative) regret \( M_A(Y)
+  - \Psi_G(Y)~.\)
+\fi
+
+Also in the online setting our label regularity measure, denoted here by $\Psi_G(Y)$, is small when
+nodes in $G$ tend to be either troll or trustworthy. When this happens, few labels from the
+incoming and outgoing edges of each node are sufficient to predict the labels on the remaining
+edges. Formally, for fixed $G$ and $Y$, let 
+$$
+\Psiin(v,Y) = \min\big\{\din^-(v),\din^+(v)\big\}
+\qquad\text{and}\qquad
+\Psiout(u,Y) = \min\big\{\dout^-(u),\dout^+(u)\big\}
+$$
 Let also
-$
-    \Psiin(Y) = \sum_{j \in V} \Psiin(j,Y)
-$ and
-%\qquad\text{and}\qquad
-$
-\Psiout(Y) = \sum_{i \in V} \Psiout(i,Y)
-$.
-Then we define $\Psi_G(Y) = \min\big\{\Psiin(Y),\Psiout(Y)\big\}$. The two measures $\Psi^2_{G''}(Y)$ and $\Psi_G(Y)$ are conceptually related. %\textcolor{red}{
-Indeed, their value on real data is quite similar%}
-% our experiments
-(see Table~\ref{tab:all_mcc} in Section~\ref{s:exp}).
+$$
+\Psiin(Y) = \sum_{v \in V} \Psiin(v,Y)
+\qquad\text{and}\qquad
+\Psiout(Y) = \sum_{u \in V} \Psiout(u,Y)
+$$
+Then we define $\Psi_G(Y) = \min\big\{\Psiin(Y),\Psiout(Y)\big\}$. The two measures
+$\Psi^2_{G''}(Y)$ and $\Psi_G(Y)$ are conceptually related and their value on real data is quite
+similar (see \autoref{tab:all_mcc} in \autoref{s:exp}).
 
-%NCB What point are we trying to make here?
-% Recall that the cutsize of an undirected node-labeled graph $G'(Y)$ is the number of edges in $G'$ connecting nodes having mismatching labels.
-Now, because the only nodes in $G'$ we are interested in predicting are those corresponding to the edges in $G$ (the colored squares in Figure \ref{f:etnr} (b)), the online prediction problem on the edges of $G$ translates to a node sign prediction problem on a subset of $V'$.
-The remaining nodes in $V'$ (the circles in Figure \ref{f:etnr} (b)) we are free to assign dummy labels so as to minimize the corresponding mistake bound over $G'$.
-It is straightforward to see that if we set the labeling $Y$ on $V'$ in such a way that $y_{\iout} = +1$ if $tr(i) < 0$, and $-1$, otherwise, and $y_{\iin} = +1$ if $un(i) < 0$, and $-1$, otherwise, then the cutsize of $G'(Y)$ equals\footnote
-%{
-%In fact, for the sake of this specific argument, nothing prevents from retaining of $G'$ either only the edges $(\iout,e_{i,j})$ or only the edges $(e_{i,j},\jin)$, resulting in a cutsize of $\Psiout(Y)$ and $\Psiin(Y)$, respectively.
-%} 
-%$\Psiin(Y) + \Psiout(Y)$.
+\iffalse
+  Recall that the cutsize of an undirected node-labeled graph $G'(Y)$ is the number of edges in
+  $G'$ connecting nodes having mismatching labels. Now, because the only nodes in $G'$ we are
+  interested in predicting are those corresponding to the edges in $G$ (the colored squares in
+  Figure \ref{f:etnr} (b)), the online prediction problem on the edges of $G$ translates to a node
+  sign prediction problem on a subset of $V'$. The remaining nodes in $V'$ (the circles in Figure
+  \ref{f:etnr} (b)) we are free to assign dummy labels so as to minimize the corresponding mistake
+  bound over $G'$. It is straightforward to see that if we set the labeling $Y$ on $V'$ in such a
+  way that $y_{\iout} = +1$ if $tr(u) < 0$, and $-1$, otherwise, and $y_{\iin} = +1$ if $un(u) <
+  0$, and $-1$, otherwise, then the cutsize of $G'(Y)$ equals\footnote{In fact, for the sake of
+  this specific argument, nothing prevents from retaining of $G'$ either only the edges
+  $(\iout,e_{u,v})$ or only the edges $(e_{u,v},\jin)$, resulting in a cutsize of $\Psiout(Y)$ and
+  $\Psiin(Y)$, respectively.} $\Psiin(Y) + \Psiout(Y)$.
+\fi

### 2017-09-03 troll/supp_body.tex
small pass over notations

@@
 \subsection{Label propagation on $G''$}
-%
+\label{ssec:troll_lprop_weights}
 Here we provide more details on the choice of weight for the edges of $G''$, as

### 2017-09-04 troll/algs_batch.tex
a pass on the rest of the troll main text

@@
 \section{Algorithms in the Batch Setting}\label{s:algbatch}
-Given $G(Y) =(V,E(Y))$, we have at our disposal a training set $E_0$ of labeled edges from $E(Y)$, our goal being that of building a predictive model for the labels of the remaining edges.
 
-Our first algorithm is an approximation to the Bayes optimal predictor $y^*(i,j)$.
-Let us denote by $\htr(i)$ and $\hun(i)$ the trollness and the untrustworthiness of node $i$ when both are computed on the subgraph induced by the training edges.
-We now design and analyze an edge classifier of the form
-%
+Given $G(Y) =(V,E(Y))$, we have at our disposal a training set \trainset{} of labeled edges from
+$E(Y)$, our goal being that of building a predictive model for the labels of the remaining edges.
+
+Our first algorithm is an approximation to the Bayes optimal predictor $y^*(u,v)$. Let us denote by
+$\htr(u)$ and $\hun(u)$ the trollness and the untrustworthiness of node $u$ when both are computed
+on the subgraph induced by the training edges. We now design and analyze an edge classifier of the
+form
+
 \begin{equation}
   \label{eq:predictor}
-	\sgn\Big(\big(1-\htr(i)\big) + \big(1-\hun(j)\big) - \tfrac{1}{2} -\tau\Big)~,
+  \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) - \tfrac{1}{2} -\tau\Big)~,
 \end{equation}
-%
-where $\tau \ge 0$ is the only parameter to be trained.
-Despite its simplicity, this classifier works reasonably well in practice, as demonstrated by our experiments (see Section~\ref{s:exp}).
-Moreover, unlike previous edge sign prediction methods for directed graphs, our classifier comes with a rigorous theoretical motivation, since it approximates the Bayes optimal classifier $y^*(i,j)$ with respect to the generative model defined in Section~\ref{s:gen}.
-It is important to point out that when we use $1-\htr(i)$ and $1-\hun(j)$ to estimate $p_i$ and $q_j$, an additive bias shows up due to~(\ref{e:pout}).
-This motivates the need of a threshold parameter $\tau$ to cancel this bias.
-Yet, the presence of a prior distribution $\mu(p,q)$ ensures that this bias is the same for all edges $(i,j) \in E$.
 
-%\subsection{Approximation to Bayes via Active Learning}\label{ss:active}
-%
-% We now introduce an active learning algorithm approximating the Bayes optimal predictions $y^*(i,j)$.
+where $\tau \ge 0$ is the only parameter to be trained. Despite its simplicity, this classifier
+works reasonably well in practice, as demonstrated by our experiments (see \autoref{s:exp}).
+Moreover, unlike previous edge sign prediction methods for directed graphs, our classifier comes
+with a rigorous theoretical motivation, since it approximates the Bayes optimal classifier
+$y^*(u,v)$ with respect to the generative model defined in \autoref{s:gen}. It is important to point
+out that when we use $1-\htr(u)$ and $1-\hun(v)$ to estimate $p_u$ and $q_v$, an additive bias shows
+up due to \eqref{e:pout} and \eqref{e:pin}. This motivates the need of a threshold parameter $\tau$
+to cancel this bias. Yet, the presence of a prior distribution $\mu(p,q)$ ensures that this bias is
+the same for all edges $(u,v) \in E$.
 
 \iffalse
-***********************************************************************
-Given a positive integer parameter $Q < \tfrac{|E|}{2|V|+1}$ such that there exists a set\footnote
-{
-This set is needed to find an estimate $\tauhat$ of $\tau$ in~(\ref{eq:predictor}), see Step~3 of the algorithm. It can be any set of directed paths/cycles in $G$ that are pairwise vertex-disjoint.
-}
-$E_L \ss E$ of size $Q$ where each vertex $i \in V$ appearing as an endpoint of some edge in $E_L$ occurs at most once as origin ---i.e., $(i,j)$, and at most once as destination ---i.e., $(j,i)$.
-The algorithm performs the following steps:
+\subsection{Approximation to Bayes via Active Learning}\label{ss:active}
 
-\textbf{1.} For each $i \in V$ such that $\din(i) \ge Q$, $Q$ edges are drawn at random without replacement from $\Nin(i)$, and their are labels queried. Let
-$
-	\hdeltain(i) = \hdin^+(i)/Q
-$,
-where $\hdin^+(i)$ is the number of positive edges sampled from $\Nin(i)$;
+We now introduce a non-adaptive active learning algorithm approximating the Bayes optimal
+predictions $y^*(u,v)$.
 
-\textbf{2.} For each $i \in V$ such that $\dout(i) \ge Q$, $Q$ edges are drawn at random without replacement from $\Nout(i)$ and their labels are queried. Let
-$
-	\hdeltaout(i) = \hdout^+(i)/Q
-$,
-where $\hdout^+(i)$ is the number of positive edges sampled from $\Nout(i)$;
+Given a positive integer parameter $Q < \tfrac{|E|}{2|V|+1}$ such that there exists a
+set\footnote{This set is needed to find an estimate $\tauhat$ of $\tau$ in~(\ref{eq:predictor}), see
+Step~3 of the algorithm. It can be any set of directed paths/cycles in $G$ that are pairwise
+vertex-disjoint.} $E_L \ss E$ of size $Q$ where each vertex $u \in V$ appearing as an endpoint of
+some edge in $E_L$ occurs at most once as origin ---ie{}., $(u,v)$, and at most once as destination
+---ie{}., $(v,u)$. The algorithm performs the following steps:
 
-\textbf{3.} Sample any edge in $E_L$ not yet sampled, and let $\tauhat$ be the fraction of positive edges in $E_L$;
+\textbf{1.} For each $u \in V$ such that $\din(u) \ge Q$, $Q$ edges are drawn at random without
+replacement from $\Nin(u)$, and their are labels queried. Let $\hdeltain(u) = \hdin^+(u)/Q$, where
+$\hdin^+(u)$ is the number of positive edges sampled from $\Nin(u)$;
 
-\textbf{4.} Any remaining non-sampled edge $(i,j)$ is predicted as
-$
-	\yhat(i,j) = \sgn\big(\hdeltaout(i) + \hdeltain(j) - \tfrac{1}{2} - \tauhat \big)
-$.
-***********************************************************************
+\textbf{2.} For each $u \in V$ such that $\dout(u) \ge Q$, $Q$ edges are drawn at random without
+replacement from $\Nout(u)$ and their labels are queried. Let $ \hdeltaout(u) = \hdout^+(u)/Q $,
+where $\hdout^+(u)$ is the number of positive edges sampled from $\Nout(u)$;
+
+\textbf{3.} Sample any edge in $E_L$ not yet sampled, and let $\tauhat$ be the fraction of positive
+edges in $E_L$;
+
+\textbf{4.} Any remaining non-sampled edge $(u,v)$ is predicted as $ \yhat(u,v) =
+\sgn\big(\hdeltaout(u) + \hdeltain(v) - \tfrac{1}{2} - \tauhat \big) $.
 \fi
 
-Our algorithm works under the assumption that for given parameters $Q$ (a positive integer) and $\alpha \in (0,1)$ there exists a set\footnote
-{
-$E_L$ is needed to find an estimate $\tauhat$ of $\tau$ in~(\ref{eq:predictor}) ---see Step~3 of the algorithm.
-Any undirected matching of $G$ of size $\mathcal{O}(\log|V|)$ can be used.
-In practice, however, we never computed $E_L$, and estimated $\tau$ on the entire training set $E_0$.
-}
-$E_L \ss E$ of size $\tfrac{2Q}{\alpha}$ where each vertex $i \in V$ appearing as an endpoint of some edge in $E_L$ occurs at most once as origin ---i.e., $(i,j)$--- and at most once as destination ---i.e., $(j,i)$.
-Moreover, we assume $E_0$ has been drawn from $E$ at random {\em without} replacement, with $m = |E_0| = \alpha\,|E|$.
-The algorithm performs the following steps:
+Our algorithm works under the assumption that for given parameters $Q$ (a positive integer) and
+$\alpha \in (0,1)$ there exists a set\footnote{$E_L$ is needed to find an estimate $\tauhat$ of
+$\tau$ in~(\ref{eq:predictor}) ---see Step~3 of the algorithm. Any undirected matching of $G$ of
+size $\mathcal{O}(\log|V|)$ can be used, for instance using the blossom
+algorithm~\autocite{matching65}. In practice, however, we never computed $E_L$, and estimated $\tau$
+on the entire training set \trainset{}.} $E_L \ss E$ of size $\tfrac{2Q}{\alpha}$ where each vertex
+$u \in V$ appearing as an endpoint of some edge in $E_L$ occurs at most once as origin ---ie{}.,
+$(u,v)$--- and at most once as destination ---ie{}., $(v,u)$. Moreover, we assume \trainset{} has
+been drawn from $E$ at random {\em without} replacement, with $m = |E_0| = \alpha\,|E|$. The
+algorithm performs the following steps:
 
-%\textbf{1.} Draw a subset $E_0$ of $m = \alpha|E|$ edges from $E$ at random {\em without} replacement, and observe the corresponding labels.
+\begin{enumerate}[label=\textbf{\arabic*.}]
+	\item For each $v \in V$, let $\hun(v) = \hdin^-(v)/\hdin(v)$, \ie{} the fraction of
+	  negative edges found in $\Nin(v) \cap \trainset{}$.
 
-\begin{enumerate}[leftmargin=*,nosep,label=\textbf{\arabic*.}]
-	\item For each $j \in V$, let
-		$
-		\hun(j) = \hdin^-(j)/\hdin(j)
-		$,
-		i.e., the fraction of negative edges found in $\Nin(j) \cap E_0$.
+	\item For each $u \in V$, let $\htr(u) = \hdout^-(u)/\hdout(u)$, \ie{} the fraction of
+	  negative edges found in $\Nout(u) \cap \trainset{}$.
 
-	\item For each $i \in V$, let
-		$
-		\htr(i) = \hdout^-(i)/\hdout(i)
-		$,
-		i.e., the fraction of negative edges found in $\Nout(i) \cap E_0$.
+	\item Let $\tauhat$ be the fraction of positive edges in $E_L\cap \trainset{}$.
 
-	\item Let $\tauhat$ be the fraction of positive edges in $E_L\cap E_0$.
-
-	\item Any remaining edge $(i,j) \in E\setminus E_0$ is predicted as
-		$
-		\yhat(i,j) = \sgn\Big(\big(1-\htr(i)\big) + \big(1-\hun(j)\big) - \tfrac{1}{2} -\tauhat\Big)
-		$.
+	\item Any remaining edge $(u,v) \in E\setminus \trainset{}$ is predicted as
+		$$
+		\yhat(u,v) = \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) - \tfrac{1}{2} -\tauhat\Big)
+		$$
 \end{enumerate}
 
-The next result\footnote
-{
-All proofs are in the supplementary material.
-} 
-shows that if the graph is not too sparse, then the above algorithm can approximate the Bayes optimal predictor on nodes whose in-degree and out-degree is not too small.
-%
-% Let $n = |V|$.
-Given a function $f\,:\, \Nat \rightarrow \R$, such that $f(n) \rightarrow \infty$ as $n \rightarrow \infty$ (e.g., $f(n) = \ln n$) and a threshold $\theta = \frac{1}{n\,f(n)}$, let $\kintheta$ be the number of nodes $i \in V$ with $\din(i) \geq \theta$, and $\kouttheta$ be the number of nodes $i \in V$ with $\dout(i) \geq \theta$.
-%
+The next result\footnote{All proofs are in the supplementary material.} shows that if the graph is
+not too sparse, then the above algorithm can approximate the Bayes optimal predictor on nodes whose
+in-degree and out-degree is not too small.
 
@@
   \label{t:active}
-Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the model in Section~\ref{s:gen}.
-%, and let $E_0$ of $m = |E_0| = \alpha|E|$. 
-If the algorithm is run with parameter $Q = \Omega(\ln|V|)$, and $\alpha \in (0,1)$ such that the above assumptions are satisfied, then $\yhat(i,j) = y^*(i,j)$ holds with high probability simultaneously for all test edges $(i,j) \in E$ such that $\dout(i),\din(j) = \Omega(\ln|V|)$, and $\eta(i,j) = \Pr(y_{i,j}=1)$ is bounded away from $\tfrac{1}{2}$.
-%The overall number of training edges is $\Theta\big(|V|\ln|V|\big)$.
+  Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the
+  model in \autoref{s:gen}.
+  If the algorithm is run with parameter $Q = \Omega(\ln|V|)$, and $\alpha \in (0,1)$ such that the
+  above assumptions are satisfied, then $\yhat(u,v) = y^*(u,v)$ holds with high probability
+  simultaneously for all test edges $(u,v) \in E$ such that $\dout(u),\din(v) = \Omega(\ln|V|)$, and
+  $\eta(u,v) = \Pr(\yuv=1)$ is bounded away from $\tfrac{1}{2}$.
+  The overall number of training edges is $\Theta\big(|V|\ln|V|\big)$.
 \end{theorem}
-%
-% While learning the Bayes optimal predictor is alluring, in practice
 
-%NCB The approach leading to Theorem~\ref{t:active} needs the graph to be sufficiently dense and the bias $\tau$ to be the same for all edges.
-In order to address these limitations, we now introduce a second algorithm.
-
-%\textcolor{blue}{
-% The approach leading to Theorem~\ref{t:active} needs the graph to be sufficiently dense and the bias $\tau$ to be the same for all edges.
-In order to address these limitations, we now introduce a second method, which is used to derive the \usrule\ algorithm assessed in our experiments of Section~\ref{s:exp}.%}
-The approach leading to Theorem~\ref{t:active} lets us derive the \usrule{}
-algorithm assessed in our experiments of Section~\ref{s:exp}, but it needs the graph to be
-sufficiently dense and the bias $\tau$ to be the same for all edges. In order to
-address these limitations, we now introduce a second method based on label
-propagation.
-
-% From this algorithm we derive the \textsc{blc} algorithm we use in our
-% experiments of Section~\ref{s:exp}.
-%we show that \textsc{blc} works well even in a passive learning setting.
-
-%{\bf CG: also pointer to logistic regression from here?}
+The approach leading to Theorem~\ref{t:active} lets us derive the \usrule{} algorithm assessed in
+our experiments of \autoref{s:exp}, but it needs the graph to be sufficiently dense and the bias
+$\tau$ to be the same for all edges. In order to address these limitations, we now introduce a
+second method based on label propagation.
 
@@
 
-For simplicity, assume the joint prior distribution $\mu(p,q)$ is uniform over $[0,1]^2$ with independent marginals, and suppose that we draw at random without replacement the training set $E_0 = \big((i_1,j_1),y_{i_1,j_1}), ((i_2,j_2),y_{i_2,j_2}), \ldots, ((i_m,j_m),y_{i_m,j_m}\big)$, with $m = |E_0|$.
-Then a reasonable approach to approximate $y^*(i,j)$ would be to resort to a maximum likelihood estimator of the parameters $\{p_i, q_i\}_{i=1}^{|V|}$ based on $E_0$.
-% Working out the math (details in the supplementary material), one can easily see that
-As showed in the supplementary material,
-the gradient of the log-likelihood function w.r.t.\ $\{p_i, q_i\}_{i=1}^{|V|}$ satisfies
-%
-%
-\begin{small}
+For simplicity, assume the joint prior distribution $\mu(p,q)$ is uniform over $[0,1]^2$ with
+independent marginals, and suppose that we draw at random without replacement the training set $E_0
+= \big((u_1,v_1),y_{u_1,v_1}), ((u_2,v_2),y_{u_2,v_2}), \ldots, ((u_m,v_m),y_{u_m,v_m}\big)$, with
+$m = |E_0|$. Then a reasonable approach to approximate $y^*(u,v)$ would be to resort to a maximum
+likelihood estimator of the parameters $\{p_i, q_i\}_{i=1}^{|V|}$ based on \trainset{}.
+As showed in the supplementary material, the gradient of the log-likelihood function w.r.t.\ $\{p_i,
+q_i\}_{i=1}^{|V|}$ satisfies
+
+\begin{equation}\label{e:mlp}
+\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_i, q_i\}_{u=1}^{|V|}\right)}{\partial p_{\ell}}
+=
+\sum_{k=1}^m
+\frac{\Ind{u_k = \ell,y_{\ell,v_k}=+1}}{p_{\ell}+q_{v_k}}\,
+- \sum_{k=1}^m
+\frac{\Ind{u_k = \ell,y_{\ell,v_k}=-1}}{2-p_{\ell}-q_{v_k}}
+\end{equation}
+
 \begin{align}
-&\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial p_{\ell}}\label{e:mlp}\\
+&\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_i, q_i\}_{u=1}^{|V|}\right)}{\partial q_{\ell}}\label{e:mlq}\\
 &\ \ =
 \sum_{k=1}^m
-\frac{\Ind{i_k = \ell,y_{\ell,j_k}=+1}}{p_{\ell}+q_{j_k}}\, 
+\frac{\Ind{v_k = \ell,y_{u_k,\ell}=+1}}{p_{u_k}+q_{\ell}}\,
 - \sum_{k=1}^m
-\frac{\Ind{i_k = \ell,y_{\ell,j_k}=-1}}{2-p_{\ell}-q_{j_k}}\,,\notag
+\frac{\Ind{v_k = \ell,y_{u_k,\ell}=-1}}{2-p_{u_k}-q_{\ell}}\,,\notag
 \end{align}
-%
-\begin{align}
-&\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial q_{\ell}}\label{e:mlq}\\
-&\ \ = 
-\sum_{k=1}^m
-\frac{\Ind{j_k = \ell,y_{i_k,\ell}=+1}}{p_{i_k}+q_{\ell}}\, 
-- \sum_{k=1}^m
-\frac{\Ind{j_k = \ell,y_{i_k,\ell}=-1}}{2-p_{i_k}-q_{\ell}}\,,\notag
-\end{align}
-\end{small}
-%
+
 where $\Ind{\cdot}$ is the indicator function of the event at argument.
-Unfortunately, equating~(\ref{e:mlp}) and~(\ref{e:mlq}) to zero, and solving for parameters $\{p_i, q_i\}_{i=1}^{|V|}$ gives rise to a hard set of nonlinear equations.
-Moreover, some such parameters may never occur in these equations, namely whenever $\Nout(i)$ or $\Nin(j)$ are not represented in $E_0$ for some $i,j\in V$. 
-%Specifically, if $E_0 \cap \Nout(i) = \emptyset$ then $p_i$ does not occur, and if $E_0 \cap \Nin(j) = \emptyset$ then $q_j$ does not.
-%
-Our \emph{first approximation} is therefore to replace the nonlinear equations resulting from~(\ref{e:mlp}) and~(\ref{e:mlq}) by the following set of linear equations\footnote{Details are provided in the supplementary material.}, one for each $\ell \in V$:
-%
+Unfortunately, equating \eqref{e:mlp} and \eqref{e:mlq} to zero, and solving for parameters
+$\{p_i, q_i\}_{i=1}^{|V|}$ gives rise to a hard set of nonlinear equations.
+Moreover, some such parameters may never occur in these equations, namely whenever $\Nout(u)$ or
+$\Nin(v)$ are not represented in \trainset{} for some $u,v\in V$.
+
+Our \emph{first approximation} is therefore to replace the nonlinear equations resulting
+from \eqref{e:mlp} and \eqref{e:mlq} by the following set of linear equations\footnote{Details
+are provided in the supplementary material.}, one for each $\ell \in V$:
+\begin{equation*}
+  \sum_{k=1}^m \Ind{u_k = \ell,y_{\ell,v_k}=+1} \left(2-p_{\ell}-q_{v_k}\right)
+  = \sum_{k=1}^m \Ind{u_k = \ell,y_{\ell,v_k}=-1}
+  (p_{\ell}+q_{v_k})
+\end{equation*}
+
 \begin{align*}
-\sum_{k=1}^m &\Ind{i_k = \ell,y_{\ell,j_k}=+1} \left(2-p_{\ell}-q_{j_k}\right)\\
+  \sum_{k=1}^m &\Ind{v_k = \ell,y_{u_k,\ell}=+1} \left(2-p_{u_k}-q_{\ell}\right)\\
   =
-&\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=-1}
-(p_{\ell}+q_{j_k})%, \quad \ell = 1, \ldots, |V|
+  &\sum_{k=1}^m  \Ind{v_k = \ell,y_{u_k,\ell}=-1}
+  \left(p_{u_k}+q_{\ell}\right)
 \end{align*}
-%
-%\ncb{Claudio, che significa la doppia somma qui sotto?}
-% CG: un typo...
-%
-\begin{align*}
-\sum_{k=1}^m &\Ind{j_k = \ell,y_{i_k,\ell}=+1} \left(2-p_{i_k}-q_{\ell}\right)\\
-=
-&\sum_{k=1}^m  \Ind{j_k = \ell,y_{i_k,\ell}=-1}
-\left(p_{i_k}+q_{\ell}\right)~.%,\ \ \ \ell = 1, \ldots, |V|\,,
-\end{align*}
-%
-%
-%
-\iffalse
-************************************
-Denoting for brevity ${\wh \dout(i)}$ and  ${\wh \din(i)}$ the out-degre and the in-degree 
-of node $i$ over the training set, these can be rewritten as
-% 
-\begin{align*}
-\sum_{k=1\,:\,i_k = \ell}^m \left(p_{\ell}+q_{j_k}\right) 
-&= 
-{\wh \dout(\ell)} + \sum_{k=1\,:\,i_k = \ell} y_{\ell,j_k}, \qquad \ell = 1, \ldots, |V|~,\\
-\sum_{k=1\,:\,j_k = \ell}^m \left(p_{i_k}+q_{\ell}\right) 
-&= 
-{\wh \din(\ell)} + \sum_{k=1\,:\,j_k = \ell} y_{i_k,\ell}, \qquad \ell = 1, \ldots, |V|\,,
-\end{align*}
-%
-************************************
-\fi
-The solution to these equations are precisely the points where the gradient w.r.t.\ $(\bp,\bq) =\{p_i, q_i\}_{i=1}^{|V|}$ of the quadratic function
+
+The solution to these equations are precisely the points where the gradient w.r.t.\ $(\bp,\bq)
+=\{p_i, q_i\}_{i=1}^{|V|}$ of the quadratic function
 \[
-f_{E_0}(\bp,\bq) = \sum_{(i,j) \in E_0} \left(\frac{1+y_{i,j}}{2} - \frac{p_i+q_j}{2} \right)^2
+f_{E_0}(\bp,\bq) = \sum_{(u,v) \in E_0} \left(\frac{1+\yuv}{2} - \frac{p_u+q_v}{2} \right)^2
 \]
 vanishes.
-We follow a label propagation approach by adding to $f_{E_0}$ the corresponding test set function $f_{E\setminus E_0}$, and treat the sum of the two as the function to be minimized during training w.r.t.\ both $(\bp,\bq)$ and all $y_{i,j} \in [-1,+1]$ for $(i,j)\in E\setminus E_0$, i.e.,
-%
+We follow a label propagation approach by adding to $f_{E_0}$ the corresponding test set function
+$f_{E\setminus \trainset}$, and treat the sum of the two as the function to be minimized during
+training w.r.t.\ both $(\bp,\bq)$ and all $\yuv \in [-1,+1]$ for $(u,v)\in E\setminus \trainset{}$:
+
 \begin{equation}\label{e:quadratic}
-\min_{(\bp,\bq), y_{i,j} \in [-1,+1],\,(i,j)\in E\setminus E_0} 
-\left(f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\right)\,.
+  \min_{(\bp,\bq), \yuv \in [-1,+1],\,(u,v)\in E\setminus E_0}
+  \left(f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\right)
 \end{equation}
-%
-Binary $\pm 1$ predictions on the test set $E\setminus E_0$ are then obtained by thresholding the obtained values $y_{i,j}$ at $0$.
+
+Binary $\pm 1$ predictions on the test set $E\setminus \trainset{}$ are then obtained by
+thresholding the obtained values $\yuv$ at $0$.
 
@@
 ********************************************************
-and then solve for parameters $\{p_i, q_i\}_{i=1}^{|V|}$.
+and then solve for parameters $\{p_u, q_i\}_{u=1}^{|V|}$.
 Yet, it may well be the case that some such parameters never occur in all these equations.\footnote
 {
-On the other hand, recall that no pairing $(i,j) \in E$ can occur more than once here, since $E_0$ is sampled without replacement.
+On the other hand, recall that no pairing $(u,v) \in E$ can occur more than once here, since \trainset{} is sampled without replacement.
 }
-This will happen precisely whenever $\Nout(i)$ or $\Nin(j)$ are not represented in $E_0$.
-Specifically, if $E_0 \cap \Nout(i) = \emptyset$ then $p_i$ does not occur, and if $E_0 \cap \Nin(j) = \emptyset$ then $q_j$ does not occur.
-Hence, for each unsampled edge $(\ell,j)\in E\setminus E_0$, we add to~(\ref{e:mlp2}) the equations
+This will happen precisely whenever $\Nout(u)$ or $\Nin(v)$ are not represented in \trainset{}.
+Specifically, if $E_0 \cap \Nout(u) = \emptyset$ then $p_u$ does not occur, and if $E_0 \cap \Nin(v) = \emptyset$ then $q_v$ does not occur.
+Hence, for each unsampled edge $(\ell,v)\in E\setminus \trainset{}$, we add to~(\ref{e:mlp2}) the equations
 \[
-p_{\ell}+q_j = 1+y_{\ell,j}~,
+p_{\ell}+q_v = 1+y_{\ell,v}~,
 \]
-motivated by the fact that $\E \left[\frac{1+y_{\ell,j}}{2}\,|\, (\ell,j)\right] =  \frac{p_{\ell}+q_j}{2}$. Similarly, we add to~(\ref{e:mlq2}) the equations
+motivated by the fact that $\E \left[\frac{1+y_{\ell,v}}{2}\,|\, (\ell,v)\right] =  \frac{p_{\ell}+q_v}{2}$. Similarly, we add to~(\ref{e:mlq2}) the equations
 \[
-p_i+q_{\ell} = 1+y_{i,\ell}~.
+p_u+q_{\ell} = 1+y_{u,\ell}~.
 \]
@@
 \begin{align}
-p_{i}+ \frac{1}{\dout(i)}\,\sum_{j \in \Nout(i)} q_{j}
+p_{u}+ \frac{1}{\dout(u)}\,\sum_{v \in \Nout(u)} q_{v}
 &=
-\frac{1}{\dout(i)}\,\sum_{j=1}^{|V|} (1+y_{i,j}), \qquad i = 1, \ldots, |V|~,\label{e:mlpa}\\
-q_{j} +  \frac{1}{\din(j)}\,\sum_{i \in \Nin(j)} p_{i}
+\frac{1}{\dout(u)}\,\sum_{v=1}^{|V|} (1+\yuv), \qquad u = 1, \ldots, |V|~,\label{e:mlpa}\\
+q_{v} +  \frac{1}{\din(v)}\,\sum_{u \in \Nin(v)} p_{u}
 &=
-\frac{1}{\din(j)}\,\sum_{i=1}^{|V|} (1+y_{i,j}), \qquad j = 1, \ldots, |V|\,,\label{e:mlqa}\\
-1+y_{i,j} &= p_i+q_{j},\qquad\qquad\qquad\quad\,\, (i,j) \in E\setminus E_0~.\label{e:mlya}
+\frac{1}{\din(v)}\,\sum_{u=1}^{|V|} (1+\yuv), \qquad v = 1, \ldots, |V|\,,\label{e:mlqa}\\
+1+\yuv &= p_u+q_{v},\qquad\qquad\qquad\quad\,\, (u,v) \in E\setminus E_0~.\label{e:mlya}
 \end{align}
@@
 
-%\textcolor{red}{
-We now proceed to solve \eqref{e:quadratic} via label propagation~\cite{LabelPropa03} on the graph $G''$ obtained through the $G \rightarrow G''$ reduction of Section~\ref{s:prel}.%} 
+We now proceed to solve \eqref{e:quadratic} via label propagation~\autocite{LabelPropa03} on the
+graph $G''$ obtained through the $G \rightarrow G''$ reduction of \autoref{s:prel}.
 However, because of the presence of negative edge weights in $G''$, we first have to symmetrize\footnote{%
-While we note here that such linear transformation of the variables does not change the problem, we provide more details in Section~1.3 of the supplementary material.} variables $p_i, q_i, y_{i,j}$ so as they all lie in the interval $[-1,+1]$.
-After this step, one can 
-%easily 
-see that, once we get back to the original variables, label propagation computes the harmonic solution minimizing the function
-%
-%\hspace*{-0.01cm}\vbox{%
+While we note here that such linear transformation of the variables does not change the problem, we
+provide more details in Section~1.3 of the supplementary material.} variables $p_i, q_i$ and $\yuv$
+so as they all lie in the interval $[-1,+1]$.
+After this step, one can see that, once we get back to the original variables, label propagation
+computes the harmonic solution minimizing the function
+
+\begin{equation*}
+{\widehat f}\big(\bp,\bq,{\yuv}_{(u,v) \in E\setminus E_0}\big)
+= f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)
++ \frac{1}{2}\sum_{u\in V}
+\left(\dout(u)\Bigl(p_u-\frac{1}{2}\Bigl)^2+\din(u)\Bigl(q_i-\frac{1}{2}\Bigl)^2\right)
+\end{equation*}
+
+The function ${\widehat f}$ is thus a regularized version of the target function $f_{E_0} +
+f_{E\setminus E_0}$ in (\ref{e:quadratic}), where the regularization term tries to enforce the extra
+constraint that whenever a node $u$ has a high out-degree then the corresponding $p_u$ should be
+close to \shalf. Thus, on any edge $(u,v)$ departing from $u$, the Bayes optimal predictor $y^*(u,v)
+= \sgn(p_u+q_v-1)$ will mainly depend on $q_v$ being larger or smaller than \shalf{} (assuming $v$
+has small in-degree). Similarly, if $u$ has a high in-degree, then the corresponding $q_u$ should be
+close to \shalf{} implying that on any edge $(v,u)$ arriving at $u$ the Bayes optimal predictor
+$y^*(v,u)$ will mainly depend on $p_v$ (assuming $v$ has small out-degree). Put differently, a node
+having a huge out-neighborhood makes each outgoing edge \enquote{count less} than a node having only
+a small number of outgoing edges, and similarly for in-neighborhoods.
+
+The label propagation algorithm operating on $G''$ does so (see again
+\autoref{fig:troll_reduction_gsecond}) by iteratively updating as follows:
+
 \begin{align*}
-{\widehat f}&\big(\bp,\bq,{y_{i,j}}_{(i,j) \in E\setminus E_0}\big)
-%&=
-%\sum_{(i,j) \in E}
-%\Biggl(
-%\frac{1}{2}\left(\frac{1+y_{i,j}}{2}-p_i\right)^2 + \frac{1}{2}\left(\frac{1+y_{i,j}}{2}-q_j\right)^2 \\
-%&\qquad\qquad+ \left(\frac{p_i+q_j-1}{2}\right)^2\Biggl)\\
-= f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\\ 
-& \qquad + \frac{1}{2}\sum_{i\in V} \left(\dout(i)\Bigl(p_i-\frac{1}{2}\Bigl)^2+\din(i)\Bigl(q_i-\frac{1}{2}\Bigl)^2\right)\,.
+  p_{u}  & \leftarrow \frac{-\sum_{v \in \NNout(u)} q_{v} + \sum_{v \in \NNout(u)} (1+\yuv)}{3\,\dout(u)}\,\quad\forall u\in V\\
+  q_{v}  & \leftarrow \frac{-\sum_{u \in  \NNin(v)} p_{u} + \sum_{u \in \NNin(v) } (1+\yuv)}{3\,\din(v)}\qquad\forall v \in V\\
+  \yuv & \leftarrow \frac{p_u + q_v}{2}~ \quad \forall (u,v) \in E\setminus E_0~.
 \end{align*}
-%}
-%
-The function ${\widehat f}$ is thus a regularized version of the target function $f_{E_0} + f_{E\setminus E_0}$ in (\ref{e:quadratic}), where the regularization term 
-%$\sum_{i\in V} \big(\dout(i)r(p_i)+\din(i)r(q_i)\big)$ 
-tries to enforce the extra constraint that whenever a node $i$ has a high out-degree then the corresponding $p_i$ should be close to $1/2$.
-Thus, on any edge $(i,j)$ departing from $i$, the Bayes optimal predictor $y^*(i,j) = \sgn(p_i+q_j-1)$ will mainly depend on $q_j$ being larger or smaller than $\tfrac{1}{2}$ (assuming $j$ has small in-degree).
-Similarly, if $i$ has a high in-degree, then the corresponding $q_i$ should be close to $1/2$ implying that on any edge $(j,i)$ arriving at $i$ the Bayes optimal predictor $y^*(j,i)$ will mainly depend on $p_j$ (assuming $j$ has small out-degree).
-Put differently, a node having a huge out-neighborhood makes each outgoing edge ``count less" than a node having only a small number of outgoing edges, and similarly for in-neighborhoods.
-%
-%
-\iffalse
-trying to establish the relative contribution of $p_i$ and $q_j$ to $\Pr(y_{i,j}=1) = \frac{p_i+q_j}{2}$  as follows: If $(i,j)$ is such that $i$ has high outdegree, then $q_j$ is more relevant than $p_i$; symmetrically, if  $(i,j)$ is such that $j$ has high indegree, then $p_i$ is more relevant than $q_j$.
-%when both are viewed as functions of $\bp$, $\bq$, and $y_{i,j} \in [-1,+1]$, for $(i,j)\in E\setminus E_0$. 
-\fi
-%
-%
-The label propagation algorithm operating on $G''$ does so (see again Figure \ref{f:etnr} (c)) by iteratively updating as follows:
-%
-\begin{align*}
-	p_{i}  & \leftarrow \frac{-\sum_{j \in \NNout(i)} q_{j} + \sum_{j \in \NNout(i)} (1+y_{i,j})}{3\,\dout(i)}\,\quad\forall i\in V\\ 
-	q_{j}  & \leftarrow \frac{-\sum_{i \in  \NNin(j)} p_{i} + \sum_{i \in \NNin(j) } (1+y_{i,j})}{3\,\din(j)}\qquad\forall j \in V\\
-  y_{i,j} & \leftarrow \frac{p_i + q_j}{2}~ \quad \forall (i,j) \in E\setminus E_0~.
-\end{align*}
-%
-The algorithm is guaranteed to converge~\cite{LabelPropa03} to the minimizer of ${\widehat f}$.
-Notice that the presence of negative weights on the edges of $G''$ does not prevent label propagation from converging.
-%Indeed, any node classification algorithm handling both positive and negative weights on the edges of $G''$ could be used instead of label propagation.
-One alternative would be the {\textsc wta} algorithm from \cite{}.
-This is the algorithm we will be championing in our experiments of Section~\ref{s:exp}.
 
-% \footnote
-% {
-% Notice that in the case of negative weights $w_{i,j}$ over the edges, the label propagation update should read as
-% \(
-% y_{i,j} = \frac{\sum_{j \in \Nout(i)} w_{i,j} y_{i,j}}{\sum_{j \in \Nout(i)} |w_{i,j}|}~.
-% \)
-% }
-%
-
-%A computation-friendly alternative is to follow a {\em method of moments}-like approach.
-We start off from the statistics $\{1-\htr(i),1-\hun(i)\}_{i=1}^{|V|}$, and observe that, even if we are sampling $E_0$ without replacement, $1-\htr(i)$ is an unbiased estimator of (\ref{e:pout}), and $1-\hun(i)$ is an unbiased estimator of (\ref{e:pin}).
+The algorithm is guaranteed to converge~\autocite{LabelPropa03} to the minimizer of ${\widehat f}$.
+Notice that the presence of negative weights on the edges of $G''$ does not prevent label
+propagation from converging. Indeed, any node classification algorithm handling both positive and
+negative weights on the edges of $G''$ could be used instead of label propagation. One alternative
+would be the \textsc{wta} algorithm from \autocite{WTA13}.  This is the algorithm we will be
+championing in our experiments of Section~\ref{s:exp}.
 
 \label{ssec:further_related}
-{\bf Further related work.} The vast majority of existing edge sign prediction algorithms for directed graphs are based on the computation of local features of the graph.
-These features are evaluated on the subgraph induced by the training edges, and the resulting values are used to train a supervised classification algorithm (e.g., logistic regression).
-The most basic set of local features used to classify a given edge $(i,j)$ are defined by
-% the neighbourhoods\footnote
-% {
-% We abuse notation here, and define the quantities $\din^+(j),\din^-(j),\dout^+(i),\dout^-(i)$ on the subgraph induced by the training edges.
-% }
-$\din^+(j),\din^-(j),\dout^+(i),\dout^-(i)$ computed over the training set $E_0$, and by the embeddedness coefficient $\big|\Nout(i) \cap \Nin(j)\big|$.
-In turn, these can be used to define more complicated features, such as
+{\bf Further related work.} The vast majority of existing edge sign prediction algorithms for
+directed graphs are based on the computation of local features of the graph. These features are
+evaluated on the subgraph induced by the training edges, and the resulting values are used to train
+a supervised classification algorithm (e.g., logistic regression). The most basic set of local
+features used to classify a given edge $(u,v)$ are defined by
+$\din^+(v),\din^-(v),\dout^+(u),\dout^-(u)$ computed over the training set \trainset{}, and by the
+embeddedness coefficient $\big|\Nout(u) \cap \Nin(v)\big|$. In turn, these can be used to define
+more complicated features, such as
 \(
-	\frac{\din^+(j) + |E^+|\uin(j)}{\din(j) + \uin(j)}
+	\frac{\din^+(v) + |E^+|\uin(v)}{\din(v) + \uin(v)}
 \quad\text{and}\quad
-	\frac{\dout^+(i) + p^+\uout(i)}{\dout(i) + \uout(i)}
+	\frac{\dout^+(u) + p^+\uout(u)}{\dout(u) + \uout(u)}
 \)
-introduced in~\cite{Bayesian15}, together with their negative counterparts, where $|E^+|$ is the overall fraction of positive edges, and $\uin(j),\uout(i)$ are, respectively, the number of test edges outgoing from $i$ and the number of test edges incoming to $j$.
-Other types of features are derived from social status theory~(e.g., \cite{Leskovec2010}), and involve the so-called triads; namely, the triangles formed by $(i,j)$ together with $(i,w)$ and $(w,j)$ for any $w \in \NNout(i) \cap \NNin(j)$.
-A third group of features is based on node ranking scores.
-These scores are computed using a variery of methods, including Prestige~\cite{zolfaghar2010mining}, exponential ranking~\cite{traag2010exponential}, PageTrust~\cite{de2008pagetrust}, Bias and Deserve~\cite{mishra2011finding}, TrollTrust~\cite{wu2016troll}, and generalizations of PageRank and HITS to signed networks~\cite{shahriari2014ranking}.
-Examples of features using such scores are \textsl{reputation} and \textsl{optimism}~\cite{shahriari2014ranking}, defined for a node $i$ by
+introduced in~\autocite{Bayesian15}, together with their negative counterparts, where $|E^+|$ is the
+overall fraction of positive edges, and $\uin(v),\uout(u)$ are, respectively, the number of test
+edges outgoing from $u$ and the number of test edges incoming to $v$.
+Other types of features are derived from social status theory~(e.g., \autocite{Leskovec2010}), and
+involve the so-called triads; namely, the triangles formed by $(u,v)$ together with $(u,w)$ and
+$(w,v)$ for any $w \in \NNout(u) \cap \NNin(v)$. A third group of features is based on node ranking
+scores. These scores are computed using a variery of methods, including
+Prestige~\autocite{zolfaghar2010mining}, exponential ranking~\autocite{traag2010exponential},
+PageTrust~\autocite{de2008pagetrust}, Bias and Deserve~\autocite{mishra2011finding},
+TrollTrust~\autocite{wu2016troll}, and generalizations of PageRank and HITS to signed
+networks~\autocite{shahriari2014ranking}. Examples of features using such scores are \textsl{reputation}
+and \textsl{optimism}~\autocite{shahriari2014ranking}, defined for a node $u$ by
 \(
-	\frac{\sum_{j \in \NNin(i)} y_{j,i}\sigma(j)}{\sum_{j \in \NNin(i)} \sigma(j)}
+	\frac{\sum_{v \in \NNin(u)} y_{v,u}\sigma(v)}{\sum_{v \in \NNin(u)} \sigma(v)}
 \quad\text{and}\quad
-	\frac{\sum_{j \in \NNout(i)} Y_{i,j}\sigma(j)}{\sum_{j \in \NNout(i)} \sigma(j)}\,,
+	\frac{\sum_{v \in \NNout(u)} \yuv\sigma(v)}{\sum_{v \in \NNout(u)} \sigma(v)}\,,
 \)
-where $\sigma(j)$ is the ranking score assigned to node $j$. Some of these algorithms will be used as representative competitors in our experimental study of Section~\ref{s:exp}.
+where $\sigma(v)$ is the ranking score assigned to node $v$. Some of these algorithms will be used
+as representative competitors in our experimental study of \autoref{s:exp}.

### 2017-09-04 troll/algs_online.tex
a pass on the rest of the troll main text

@@
 For the online scenario, we have the following result.
-%
+
 \begin{theorem}\label{t:online}
-There exists a randomized online prediction algorithm $A$ whose expected number of mistakes satisfies
-$
-\E M_A(Y) = \Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V| \right)
-$
+  There exists a randomized online prediction algorithm $A$ whose expected number of mistakes
+  satisfies $\E M_A(Y) = \Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V| \right)$
   on any edge-labeled graph $G(Y) = (V,E(Y))$.
 \end{theorem}
-%
-The algorithm used in Theorem~\ref{t:online} is a combination of randomized Weighted Majority instances.
-Details are reported in the supplementary material.
-We complement the above result by providing a mistake lower bound.
-Like Theorem~\ref{t:online}, the following result holds for all graphs, and for all label irregularity levels $\Psi_G(Y)$.
-%
-\begin{theorem}
-\label{t:mistake_bound}
-Given any edge-labeled graph $G(Y) = (V,E(Y))$ and any integer $K \le \big\lfloor \tfrac{|E|}{2}\big\rfloor$, a randomized labeling $Y\in\spin^{|E|}$ exists such that $\Psi_G(Y) \leq K$, and the expected number of mistakes that any online algorithm $A$ can be forced to make satisfies
-\(
-\E M_A(Y) \ge \frac{K}{2}\,.
-\)
-Moreover, as $\frac{K}{|E|} \rightarrow 0$ then
-\(
-\E M_A(Y) = K
-\).
+
+The algorithm used in \autoref{t:online} is a combination of randomized Weighted Majority instances.
+Details are reported in the supplementary material. We complement the above result by providing a
+mistake lower bound. Like \autoref{t:online}, the following result holds for all graphs, and for all
+label irregularity levels $\Psi_G(Y)$.
+
+\begin{theorem}\label{t:mistake_bound}
+  Given any edge-labeled graph $G(Y) = (V,E(Y))$ and any integer $K \le \big\lfloor
+  \tfrac{|E|}{2}\big\rfloor$, a randomized labeling $Y\in\spin^{|E|}$ exists such that $\Psi_G(Y)
+  \leq K$, and the expected number of mistakes that any online algorithm $A$ can be forced to make
+  satisfies $\E M_A(Y) \ge \frac{K}{2}$.  Moreover, as $\frac{K}{|E|} \rightarrow 0$ then $\E M_A(Y)
+  = K$.
 \end{theorem}
-%
 
-%\footnote
-%{
-%In fact, we can refine the above statement by proving that, as $\frac{K}{|E|} \rightarrow 0$, the lower bound gets arbitrarily close to $K$ for any $G(Y)$, hence asymptotically matching the upper bound of Theorem~\ref{t:online} (\textit{proof in the supplementary material}).
-%}
+In fact, we can refine the above statement by proving that, as $\frac{K}{|E|} \rightarrow 0$, the
+lower bound gets arbitrarily close to $K$ for any $G(Y)$, hence asymptotically matching the upper
+bound of \autoref{t:online} (\textit{proof in the supplementary material}).

### 2017-09-04 troll/experiments.tex
a pass on the rest of the troll main text

@@
 \section{Experimental Analysis}\label{s:exp}
-We now evaluate our edge sign classification methods on representative real-world datasets of varying density and label regularity,
-% within the batch settings,
-showing that our methods compete well against existing approaches in terms of both predictive and computational performance.
-We are especially interested in small training set regimes, and have restricted our comparison to the batch learning scenario since all competing methods we are aware of have been developed in that setting only.
 
-{\bf Datasets.} We considered five real-world classification datasets.
-The first three are directed \ssn{} widely used as benchmarks for this task~(e.g.,\cite{Leskovec2010,shahriari2014ranking,wu2016troll}): In \wik{}, there is an edge from user $i$ to user $j$ if $j$ applies for an admin position and $i$ votes for or against that promotion.
-In \sla{}, a news sharing and commenting website, member $i$ can tag other members $j$ as friends or foes.
-Finally, in \epi{}, an online shopping website, user $j$ reviews products and, based on these reviews, another user $i$ can display whether he considers $j$ to be reliable or not.
-In addition to these three datasets, we considered two other \ssn{} where the signs are inferred automatically.
-%rather than given explicitly by the users.
-In \kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $i$ to user $j$ indicates whether they edited the same article in a constructive manner or not.\footnote
-{
-This is the \href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of
-the \enquote{Wikisigned} dataset}, from which we removed self-loops.
-}
-Finally, in the \aut{}~\cite{kumar2016structure} network, an author $i$ cites another author $j$ by either endorsing or criticizing $j$'s work.
-The edge sign is derived by classifying the citation sentiment with a simple, yet powerful, keyword-based technique using a list of positive and negative words.
-See \cite{kumar2016structure} for more details.\footnote{
-We again removed self-loops and merged multi-edges which are all of the same sign.
-}
+We now evaluate our edge sign classification methods on representative real-world datasets of
+varying density and label regularity. This shows showing that our methods compete well against
+existing approaches in terms of both predictive and computational performance. We are especially
+interested in small training set regimes, and have restricted our comparison to the batch learning
+scenario since all competing methods we are aware of have been developed in that setting only.
+
+\paragraph{Datasets}
+
+We considered five real-world classification datasets. The first three are directed \ssn{} widely
+used as benchmarks for this task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}):
+In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
+votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $u$
+can tag other members $v$ as friends or foes. Finally, in \epi{}, an online shopping website, user
+$v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
+$v$ to be reliable or not. In addition to these three datasets, we considered two other \ssn{} where
+the signs are inferred automatically, rather than given explicitly by the users.  In
+\kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $u$ to user $v$ indicates whether they edited
+the same article in a constructive manner or not.\footnote{This is the
+\href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of the
+\enquote{Wikisigned} dataset}, from which we removed self-loops.} Finally, in the
+\aut{}~\cite{kumar2016structure} network, an author $u$ cites another author $v$ by either endorsing
+or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
+simple, yet powerful, keyword-based technique using a list of positive and negative words. See
+\cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
+multi-edges which are all of the same sign.}
 
@@
   \centering
-  \small
-  \caption{Dataset properties. The 5th column gives the fraction of positive labels. The last two columns provide two different measures of label regularity ---see main text.\label{tab:dataset}}
+  \caption{Dataset properties. The 5th column gives the fraction of positive labels. The last two
+  columns provide two different measures of label regularity ---see main text.\label{tab:dataset}}
   \begin{tabular}{lrrrrrrrr}
@@
     \midrule
-    \aut{} &   \np{4831} &  \np{39452} & 8.1  &              72.33\% &                    .076 &                       .191 \\ %&      5.07\% &            27.10\% \\
-    \wik{} &   \np{7114} & \np{103108} & 14.5 &              78.79\% &                    .063 &                       .142 \\ %&      5.56\% &            10.04\% \\
-    \sla{} &  \np{82140} & \np{549202} & 6.7  &              77.40\% &                    .059 &                       .143 \\ %&     17.74\% &             4.00\% \\
-    \kiw{} & \np{138587} & \np{740106} & 5.3  &              87.89\% &                    .034 &                       .086 \\ %&      6.55\% &             2.09\% \\
-    \epi{} & \np{131580} & \np{840799} & 6.4  &              85.29\% &                    .031 &                       .074 \\ %&     30.83\% &            14.64\% \\
+    \aut{} &   \np{4831} &  \np{39452} & 8.1  &              72.3\% &                    .076 &                       .191 \\ %&      5.07\% &            27.10\% \\
+    \wik{} &   \np{7114} & \np{103108} & 14.5 &              78.8\% &                    .063 &                       .142 \\ %&      5.56\% &            10.04\% \\
+    \sla{} &  \np{82140} & \np{549202} & 6.7  &              77.4\% &                    .059 &                       .143 \\ %&     17.74\% &             4.00\% \\
+    \kiw{} & \np{138587} & \np{740106} & 5.3  &              87.9\% &                    .034 &                       .086 \\ %&      6.55\% &             2.09\% \\
+    \epi{} & \np{131580} & \np{840799} & 6.4  &              85.3\% &                    .031 &                       .074 \\ %&     30.83\% &            14.64\% \\
     \bottomrule
@@
 
-\autoref{tab:dataset} summarizes statistics for these datasets.
-We note that most edge labels are positive.
-Hence, test set accuracy is not an appropriate measure of prediction performance.
-We instead evaluated our performance using the so-called Matthews Correlation Coefficient (MCC)~(e.g., \cite{MCC00}), defined as
-%\footnote
-%{
-%For completeness, we also report accuracy results in the supplementary material.
-%}  
+\autoref{tab:dataset} summarizes statistics for these datasets. We note that most edge labels are
+positive. Hence, test set accuracy is not an appropriate measure of prediction performance. We
+instead evaluated our performance using the so-called Matthews Correlation Coefficient
+(MCC)~\autocite{MCC00}), defined as \Todo{For completeness, we could also report accuracy results in
+the supplementary material.}  
 \[
-	\mathrm{MCC} = \frac{tp\times tn-fp\times fn} {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } }\,.
-			% = \pm \sqrt{\frac{\chi^2}{n}}
+	\mathrm{MCC} = \frac{tp\times tn-fp\times fn}%
+        {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } } = \pm \sqrt{\frac{\chi^2}{n}}
 \]
-MCC combines all the four quantities found in a binary confusion matrix ($t$rue $p$ositive, $t$rue $n$egative, $f$alse $p$ositive and $f$alse $n$egative) into a single metric which ranges from $-1$ (when all predictions are incorrect) to $+1$ (when all predictions are correct).% through $0$ (when predictions are made uniformly at random).
+MCC combines all the four quantities found in a binary confusion matrix ($t$rue $p$ositive, $t$rue
+$n$egative, $f$alse $p$ositive and $f$alse $n$egative) into a single metric which ranges from $-1$
+(when all predictions are incorrect) to $+1$ (when all predictions are correct) through $0$ (when
+predictions are made uniformly at random).
 
-Although the semantics of the edge signs is not the same across these networks, we can see from \autoref{tab:dataset} that our generative model essentially fits all of them.
-Specifically, the last two columns of the table report the rate of label (ir)regularity, as measured by $\Psi^2_{G''}(Y)/|E|$ (second-last column) and $\Psi_{G}(Y)/|E|$ (last column), where
+Although the semantics of the edge signs is not the same across these networks, we can see from
+\autoref{tab:dataset} that our generative model essentially fits all of them. Specifically, the last
+two columns of the table report the rate of label (ir)regularity, as measured by
+$\Psi^2_{G''}(Y)/|E|$ (second-last column) and $\Psi_{G}(Y)/|E|$ (last column), where 
 \[
@@
 \]
-%$\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} f_{E}(\bp,\bq)$, 
-%
-$f_{E_0}$ and $f_{E\setminus E_0}$ being the quadratic criterions of Section~\ref{ss:passive}, viewed as functions of both $(\bp,\bq)$, and $y_{i,j}$,
-% when all labels are known,
-\iffalse
-*****************
-\[
-\Psi_{G''}(Y) = \min_{\bp,\bq\in[0,1]^{|V|}}
-\sum_{(i,j) \in E} \left(\frac{1+y_{i,j}}{2} - \frac{p_i+q_j}{2}\right)^2\,,
-\]
-i.e., the quadratic criterion $f_{E}(\bp,\bq)$ of Section~\ref{ss:passive} when all labels are known,
-*****************
-\fi
-%essentially hinges upon, 
-and $\Psi_{G}(Y)$ is the label regularity measure adopted in the online setting, as defined in Section~\ref{s:prel}. 
-%The two measures have been crafted here so as to be on the same scale, hence they can be contrasted to each other on the same dataset. 
-It is reasonable to expect that higher label irregularity corresponds to lower prediction performance.
-This trend is in fact confirmed by our experimental findings: whereas \epi{} tends to be easy, \aut{} tends to be hard, and this holds for all algorithms we tested, even if they do not explicitly comply with our inductive bias principles.
-Moreover, $\Psi^2_{G''}(Y)/|E|$ tends to be proportional to $\Psi_{G}(Y)/|E|$ across datasets, hence confirming the anticipated connection between the two regularity measures.
-
-%
-%Another notable feature of our datasets is the relative regularity of their
-%labeling with respect to our two complexity measures defined in
-%\autoref{s:analysis}
-%(here  where the minimum
-%is found using the L-BFGS-B algorithm)
-%(especially in \epi{} case, which is confirmed by higher performance).
-%
+$f_{E_0}$ and $f_{E\setminus E_0}$ being the quadratic criterions of \autoref{ss:passive}, viewed as
+functions of both $(\bp,\bq)$, and $y_{i,j}$ when all labels are known, and $\Psi_{G}(Y)$ is the
+label regularity measure adopted in the online setting, as defined in \autoref{s:prel}. It is
+reasonable to expect that higher label irregularity corresponds to lower prediction performance.
+This trend is in fact confirmed by our experimental findings: whereas \epi{} tends to be easy,
+\aut{} tends to be hard, and this holds for all algorithms we tested, even if they do not explicitly
+comply with our inductive bias principles. Moreover, $\Psi^2_{G''}(Y)/|E|$ tends to be proportional
+to $\Psi_{G}(Y)/|E|$ across datasets, hence confirming the anticipated connection between the two
+regularity measures.
 
@@
 
-{\bf Algorithms and parameter tuning.} 
+\paragraph{Algorithms and parameter tuning.} 
 We compared the following algorithms:
-%
-%\begin{itemize}%[nosep,leftmargin=2em]
-%
-%\item
 
-{\bf 1.} The label propagation algorithm of Section~\ref{ss:passive} (referred to as \uslpropGsec{}). The actual binarizing threshold was set by cross-validation on the training set.
-%
-%\item
+\begin{enumerate}[label=\textbf{\arabic*.}]
+  \item The label propagation algorithm of \autoref{ss:passive} (referred to as \uslpropGsec{}).
+    The actual binarizing threshold was set by cross-validation on the training set.
 
-{\bf 2.} The algorithm analyzed at the beginning of Section~\ref{s:algbatch}, which we call \usrule{} (Bayes Learning Classifier based on {\em tr}ollness and {\em un}trustworthiness).
-After computing $\htr(i)$ and $\hun(i)$ on training set $E_0$ for all $i \in V$ (or setting those values to $\frac{1}{2}$ in case there is no outgoing or incoming edges for some node), we use Eq.~\eqref{eq:predictor} and estimate $\tau$ on $E_0$.
-%
-%\item
+  \item The algorithm analyzed at the beginning of \autoref{s:algbatch}, which we call \usrule{}
+    (Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). After
+    computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all $u \in V$ (or setting those
+    values to $\frac{1}{2}$ in case there is no outgoing or incoming edges for some node), we use
+    equation \eqref{eq:predictor} and estimate $\tau$ on $E_0$.
 
-{\bf 3.} A logistic regression model where each edge $(i,j)$ is associated with the features $[1-\htr(i), 1-\hun(j)]$ computed again on $E_0$ (we call this method \uslogregp{}).
-Best binary thresholding is again computed on $E_0$.
-Experimenting with this logistic model serves to support the claim we made in the introduction that our generative model in Section~\ref{s:gen} is a good fit for the data.
+  \item A logistic regression model where each edge $(u,v)$ is associated with the features
+    $[1-\htr(u), 1-\hun(v)]$ computed again on \trainset{} (we call this method \uslogregp{}). Best
+    binary thresholding is again computed on \trainset{}. Experimenting with this logistic model
+    serves to support the claim we made in the introduction that our generative model in
+    \autoref{s:gen} is a good fit for the data.
 
-%
-%\item 
-{\bf 4.} The solution obtained by directly solving the unregularized problem (\ref{e:quadratic}) through a fast constrained minimization algorithm (referred to as \qoptim{}).
-Again, the actual binarizing threshold was set by cross-validation on the training set.\footnote
-{
-We have also tried to minimize (\ref{e:quadratic}) by removing the $[-1,+1]$ constraints, but got similar MCC results
-as the ones we report for \qoptim{}
-}
-%
+  \item  The solution obtained by directly solving the unregularized problem \eqref{e:quadratic}
+    through a fast constrained minimization algorithm (referred to as \qoptim{}). Again, the actual
+    binarizing threshold was set by cross-validation on the training set.\footnote{We have also
+    tried to minimize \eqref{e:quadratic} by removing the $[-1,+1]$ constraints, but got similar MCC
+    results as the ones we report for \qoptim{}}
 
-{\bf 5.} The matrix completion method from~\cite{LowRankCompletion14} based on \complowrank{} matrix factorization.
-Since the authors showed their method to be robust to the choice of the rank parameter $k$, we picked $k=7$ in our experiments.
-%
-%\item
+  \item  The matrix completion method from~\autocite{LowRankCompletion14} based on \complowrank{}
+    matrix factorization. Since the authors showed their method to be robust to the choice of the
+    rank parameter $k$, we picked $k=7$ in our experiments.
 
-{\bf 6.} A logistic regression model built on \comptriads{} features derived from status theory~\cite{Leskovec2010}.
-%
-%\item
+  \item A logistic regression model built on \comptriads{} features derived from status
+    theory~\autocite{Leskovec2010}.
 
-{\bf 7.} The PageRank-inspired algorithm from \cite{wu2016troll}, where a recursive notion of trollness is computed by solving a suitable set of nonlinear equations through an iterative method, and then used to assign ranking scores to nodes, from which (un)trustworthiness features are finally extracted for each edge.
-We call this method \compranknodes{}.
-As for hyperparameter tuning ($\beta$ and $\lambda_1$ in~\cite{wu2016troll}), we closely followed the authors' suggestion of doing cross validation.
-%
-%\item
+  \item The PageRank-inspired algorithm from \autocite{wu2016troll}, where a recursive notion of
+    trollness is computed by solving a suitable set of nonlinear equations through an iterative
+    method, and then used to assign ranking scores to nodes, from which (un)trustworthiness features
+    are finally extracted for each edge. We call this method \compranknodes{}. As for hyperparameter
+    tuning ($\beta$ and $\lambda_1$ in~\autocite{wu2016troll}), we closely followed the authors'
+    suggestion of doing cross validation.
 
-{\bf 8.} The last competitor is the logistic regression model whose features have been build according to \cite{Bayesian15}. We call this method \compbayesian{}.
-%\end{itemize}
+  \item  The last competitor is the logistic regression model whose features have been build
+    according to \autocite{Bayesian15}. We call this method \compbayesian{}.
+\end{enumerate}
 
-The above methods can be roughly divided into {\em local} and {\em global} methods.
-A local method hinges on building local predictive features, based on neighborhoods: \usrule{}, \uslogregp{}, \comptriads{}, and \compbayesian{} essentially fall into this category.
-The remaining methods
-%(\uslpropGsec{}, \complowrank{}, and \compranknodes{}) 
-are global in that their features are designed to depend on global properties of the graph topology.
+The above methods can be roughly divided into \emph{local} and \emph{global} methods. A local method
+hinges on building local predictive features, based on neighborhoods: \usrule{}, \uslogregp{},
+\comptriads{}, and \compbayesian{} essentially fall into this category. The remaining methods
+(\uslpropGsec{}, \complowrank{}, and \compranknodes{}) are global in that their features are
+designed to depend on global properties of the graph topology.
 
@@
 
-Then we again associate each
-edge $\eij$ with $a_{i,j} = \left(1-\htr(i)\right) + \left(1-\hun(j)\right)$
-and find the $\tau$ of \eqref{eq:predictor} that minimizes the empirical risk.
-Another way to exploit those quantities is to
-train a Logistic Regression model where each edge is associated with two
-features: $[1-\htr(i), 1-\hun(j)]$ (we call this method \uslogregp{}).
+Then we again associate each edge $\eij$ with $a_{i,j} = \left(1-\htr(i)\right) +
+\left(1-\hun(j)\right)$ and find the $\tau$ of \eqref{eq:predictor} that minimizes the empirical
+risk.  Another way to exploit those quantities is to train a Logistic Regression model where each
+edge is associated with two features: $[1-\htr(i), 1-\hun(j)]$ (we call this method \uslogregp{}).
 
-We compare these methods with four competitors, each exploiting the sign
-information and network topology in a different way:
+We compare these methods with four competitors, each exploiting the sign information and network
+topology in a different way:
 
-Looking at undirected graphs with a global perspective,
-\citet{LowRankCompletion14} consider the observed adjacency matrix $A$, made of
-the edges in \trainset{}, as a noisy sampling of the adjacency matrix
-$A^\star$ of an underlying complete graph satisfying the weakly balance
-condition (that is with no triangle containing only one negative edge).
+Looking at undirected graphs with a global perspective, \textcite{LowRankCompletion14} consider the
+observed adjacency matrix $A$, made of the edges in \trainset{}, as a noisy sampling of the
+adjacency matrix $A^\star$ of an underlying complete graph satisfying the weakly balance condition
+(that is with no triangle containing only one negative edge).
 %TODO the previous sentence is too long
-This condition implies the existence of a small number $k$ of node clusters
-with positive links within clusters and negative links across clusters, which
-in turn implies $\rank A^\star=k$.
-By recovering a complete matrix $\tilde{A}$ which matches the non-zeros entries
-of $A$, it is possible to predict the sign of $\eij \notin \trainset$
-as $\yhat_{i,j} = \sgn(\tilde{A}_{i,j})$. Although the exact version of this
-problem is NP-hard, authors assume that $k$ is an hyperparameter known
-beforehand\footnote{The method is showed to be robust to the choice of its
-value, thus we pick $k=7$.} and look for two matrices $W,H\in
-\mathbb{R}^{k\times|V|}$ which minimise a sigmoidal reconstruction loss with
-$A$, subject to a regularization term. We refer to this method as
+This condition implies the existence of a small number $k$ of node clusters with positive links
+within clusters and negative links across clusters, which in turn implies $\rank A^\star=k$.  By
+recovering a complete matrix $\tilde{A}$ which matches the non-zeros entries of $A$, it is possible
+to predict the sign of $\eij \notin \trainset$ as $\yhat_{i,j} = \sgn(\tilde{A}_{i,j})$. Although
+the exact version of this problem is NP-hard, authors assume that $k$ is an hyperparameter known
+beforehand\footnote{The method is showed to be robust to the choice of its value, thus we pick
+$k=7$.} and look for two matrices $W,H\in \mathbb{R}^{k\times|V|}$ which minimise a sigmoidal
+reconstruction loss with $A$, subject to a regularization term. We refer to this method as
 \emph{\complowrank{}}.
 
-The status theory heuristic posits that a positive link from $i$ to $j$ denotes
-that user $i$ considers user $j$ as having a higher status or skill than
-himself~\cite{Leskovec2010}. It implies that among the 16 possible triads of
-three nodes with two directed signed edges among them, some must be more
-represented than others. The \emph{\comptriads{}} method exploits this fact by
-counting for each edge in the training set how frequently it is involved in
-each of the 16 triad types. Based on these edge features and 7 degree features
-of each edge endpoints, a Logistic Regression model is trained and used to
-predict the sign of the test edges.
+The status theory heuristic posits that a positive link from $i$ to $j$ denotes that user $i$
+considers user $j$ as having a higher status or skill than himself~\autocite{Leskovec2010}. It
+implies that among the 16 possible triads of three nodes with two directed signed edges among them,
+some must be more represented than others. The \emph{\comptriads{}} method exploits this fact by
+counting for each edge in the training set how frequently it is involved in each of the 16 triad
+types. Based on these edge features and 7 degree features of each edge endpoints, a Logistic
+Regression model is trained and used to predict the sign of the test edges.
 
-Inspired by the PageRank algorithm, \citet{wu2016troll} propose a recursive
-definition of trollness based on the opinion of one node's neighbors weighted
-by their own trollness, which allow to assign a trustworthiness $\pi(i)$ to
-each node $i$ through a set of non linear equations solved by an iterative
-method. This $\pi$ values are used to rank the nodes by computing their
-Reputation and Optimism scores, thus providing four features for each edge,
-which are in turn used to train a Logistic Regression model for the
-classification task. We refer to this method as \emph{\compranknodes{}}.
-Computing $\pi$ requires to choose two hyperparameters of the model, $\beta$
-and $\lambda_1$ and the authors suggest to do it by holding out some training
-data to perform cross validation but in practice we found little difference in
-performance in our implementation while it consumes a lot of time.
+Inspired by the PageRank algorithm, \textcite{wu2016troll} propose a recursive definition of
+trollness based on the opinion of one node's neighbors weighted by their own trollness, which allow
+to assign a trustworthiness $\pi(i)$ to each node $i$ through a set of non linear equations solved
+by an iterative method. This $\pi$ values are used to rank the nodes by computing their Reputation
+and Optimism scores, thus providing four features for each edge, which are in turn used to train a
+Logistic Regression model for the classification task. We refer to this method as
+\emph{\compranknodes{}}.  Computing $\pi$ requires to choose two hyperparameters of the model,
+$\beta$ and $\lambda_1$ and the authors suggest to do it by holding out some training data to
+perform cross validation but in practice we found little difference in performance in our
+implementation while it consumes a lot of time.
 
-The last competitors is also local. \citet{Bayesian15} note that  a node can
-belong to one of the 16 types based of whether is number of its positive (resp.
-negative) outgoing (resp. incoming) edges is zero or not.  The number of
-unobserved incoming and outgoing edges of each node $i$ let us define its
-16-dimensional vector $V_i$ containing the probability of transitioning to any
-other state.  Then for each edge \eij we compute the outer product of $V_i$
-with $V_j$ and also include additional features such as triads count and degree
-information before training a Logistic Regression model. We refer to this
-method as \emph{\compbayesian{}}.
-
-***************************************************
-\fi
-
-\iffalse
-****************************************************
-\begin{table*}[t]
-  \centering
-\setlength{\tabcolsep}{3pt}
-\scriptsize
-% \begin{longtable}{lrccc|cccc}
-\caption{MCC with increasing training set size, with one standard deviation over 12 random sampling of $\trainset$.
-The last four columns refer to the methods we took from the literature.
-For the sake of readability, we multiplied all MCC values by 100.
-The best number in each row is highlighted in \textbf{\textcolor{brown}{bold brown}} and the second one in \textit{\textcolor{red}{italic red}}.
-If the difference is statistically significant ($p$-value of a paired Student's $t$-test less than $0.005$), the best score is underlined.
-The ``time" rows contain the time taken to train on a $15\%$ training set.\label{tab:all_mcc}}
-\begin{tabular}{lrccc|cccc}
-\toprule
-                                                  & $\frac{|\trainset{}|}{|E|}$ &                    \uslogregp{} &                       \usrule{} &                 \uslpropGsec{} &   \compranknodes{} &                 \compbayesian{} &     \complowrank{} &      \comptriads{} \\
-\midrule
-\multirow{7}{*}{\rotatebox[origin=c]{90}{\aut{}}}
-                                                  & $3\%$  &  $\vsecondSig{15.19} \pm 0.93$  &               $15.09 \pm 0.92$  &  $\vfirstSig{19.00} \pm 1.16$  &  $12.28 \pm 1.70$  &               $10.91 \pm 1.31$  &   $8.85 \pm 1.07$  &   $8.62 \pm 0.97$  \\
-                                                  %& $5\%$  &  $\vsecondSig{19.54} \pm 0.48$  &               $19.53 \pm 0.43$  &  $\vfirstSig{24.11} \pm 0.68$  &  $16.65 \pm 0.59$  &               $14.53 \pm 0.85$  &  $12.07 \pm 0.88$  &  $11.47 \pm 1.23$  \\
-                                                  %& $7\%$  &  $\vsecondSig{23.58} \pm 0.55$  &               $23.55 \pm 0.43$  &  $\vfirstSig{27.64} \pm 1.08$  &  $21.38 \pm 0.86$  &               $19.75 \pm 1.02$  &  $15.15 \pm 0.54$  &  $14.58 \pm 1.15$  \\
-                                                  & $9\%$  &  $\vsecondSig{26.46} \pm 0.96$  &               $26.40 \pm 0.96$  &  $\vfirstSig{30.25} \pm 1.16$  &  $24.44 \pm 1.07$  &               $23.75 \pm 1.37$  &  $17.08 \pm 1.15$  &  $16.42 \pm 1.13$  \\
-                                                  & $15\%$ &  $\vsecondSig{32.98} \pm 0.61$  &               $32.98 \pm 0.65$  &  $\vfirstSig{35.73} \pm 0.70$  &  $31.03 \pm 0.97$  &               $32.25 \pm 0.89$  &  $22.57 \pm 0.99$  &  $22.01 \pm 0.62$  \\
-                                                  & $20\%$ &               $36.57 \pm 0.49$  &  $\vsecondSig{36.72} \pm 0.44$  &  $\vfirstSig{38.53} \pm 0.55$  &  $34.57 \pm 0.79$  &               $36.52 \pm 0.78$  &  $25.57 \pm 0.71$  &  $24.77 \pm 0.64$  \\
-                                                  & $25\%$ &               $39.90 \pm 0.65$  &               $40.16 \pm 0.63$  &  $\vfirstSig{41.32} \pm 0.67$  &  $38.26 \pm 0.77$  &  $\vsecondSig{40.32} \pm 0.70$  &  $29.24 \pm 1.10$  &  $27.13 \pm 0.38$  \\
-                                                  & time (ms)   &                           1.9 &                           0.4 &                          16.1 &              128 &                         5398 &               1894 &              4.6 \\
-\midrule
-\multirow{7}{*}{\rotatebox[origin=c]{90}{\wik{}}} 
-                                                  & $3\%$  &  $\vsecondSig{32.32} \pm 0.60$  &  $31.83 \pm 0.63$  &   $\vfirstSig{33.92} \pm 0.54$  &  $26.90 \pm 1.00$  &  $19.94 \pm 1.19$  &  $19.45 \pm 0.64$  &   $4.29 \pm 0.80$  \\
-                                                  %& $5\%$  &  $\vsecondSig{38.62} \pm 0.63$  &  $37.82 \pm 0.66$  &   $\vfirstSig{39.44} \pm 0.44$  &  $33.47 \pm 1.12$  &  $26.68 \pm 0.72$  &  $24.18 \pm 1.04$  &  $11.36 \pm 1.32$  \\
-                                                  %& $7\%$  &     $\vsecond{42.81} \pm 0.80$  &  $42.15 \pm 0.76$  &      $\vfirst{43.28} \pm 0.61$  &  $38.13 \pm 0.80$  &  $33.18 \pm 1.27$  &  $27.89 \pm 0.64$  &  $18.57 \pm 1.38$  \\
-                                                  & $9\%$  &     $\vsecond{45.57} \pm 0.48$  &  $44.74 \pm 0.58$  &      $\vfirst{45.75} \pm 0.46$  &  $41.60 \pm 0.73$  &  $38.25 \pm 1.02$  &  $30.75 \pm 0.68$  &  $24.04 \pm 1.27$  \\
-                                                  & $15\%$ &      $\vfirst{50.70} \pm 0.28$  &  $49.64 \pm 0.57$  &     $\vsecond{50.44} \pm 0.38$  &  $48.02 \pm 0.38$  &  $46.82 \pm 0.40$  &  $35.31 \pm 0.75$  &  $34.42 \pm 1.00$  \\
-                                                  & $20\%$ &   $\vfirstSig{52.98} \pm 0.50$  &  $52.00 \pm 0.47$  &  $\vsecondSig{52.58} \pm 0.40$  &  $51.42 \pm 0.33$  &  $50.45 \pm 0.40$  &  $38.16 \pm 0.53$  &  $38.55 \pm 0.72$  \\
-                                                  & $25\%$ &      $\vfirst{54.49} \pm 0.30$  &  $53.52 \pm 0.41$  &     $\vsecond{54.22} \pm 0.44$  &  $53.42 \pm 0.36$  &  $52.78 \pm 0.61$  &  $39.94 \pm 0.48$  &  $41.51 \pm 0.46$  \\
-                                                  & time (ms)   &                           4.2 &              1.1 &                           35.3 &              210 &              14089 &               4858 &              10.6 \\
-\midrule
-\multirow{7}{*}{\rotatebox[origin=c]{90}{\sla{}}} 
-                                                  & $3\%$  &  $32.34 \pm 0.27$  &  $31.78 \pm 0.81$  &  $\vsecondSig{36.62} \pm 0.42$  &   $\vfirstSig{42.90} \pm 0.54$  &  $25.11 \pm 0.66$  &  $34.32 \pm 0.64$  &  $20.95 \pm 0.86$  \\
-                                                  %& $5\%$  &  $36.93 \pm 0.32$  &  $36.25 \pm 0.63$  &  $\vsecondSig{40.72} \pm 0.24$  &   $\vfirstSig{45.47} \pm 0.37$  &  $29.50 \pm 0.17$  &  $36.95 \pm 0.44$  &  $28.26 \pm 1.06$  \\
-                                                  %& $7\%$  &  $39.96 \pm 0.27$  &  $39.26 \pm 0.39$  &  $\vsecondSig{43.62} \pm 0.21$  &   $\vfirstSig{46.80} \pm 0.64$  &  $33.57 \pm 0.27$  &  $38.25 \pm 0.31$  &  $33.89 \pm 1.41$  \\
-                                                  & $9\%$  &  $42.16 \pm 0.16$  &  $41.19 \pm 0.30$  &  $\vsecondSig{45.70} \pm 0.26$  &   $\vfirstSig{47.46} \pm 0.27$  &  $37.00 \pm 0.17$  &  $39.42 \pm 0.20$  &  $39.14 \pm 1.43$  \\
-                                                  & $15\%$ &  $46.44 \pm 0.11$  &  $45.23 \pm 0.32$  &   $\vfirstSig{49.65} \pm 0.23$  &  $\vsecondSig{48.59} \pm 0.72$  &  $43.28 \pm 0.21$  &  $41.09 \pm 0.29$  &  $46.27 \pm 1.00$  \\
-                                                  & $20\%$ &  $48.71 \pm 0.22$  &  $47.79 \pm 0.29$  &     $\vsecond{51.88} \pm 0.28$  &      $\vfirst{52.09} \pm 0.28$  &  $47.03 \pm 0.24$  &  $43.10 \pm 0.42$  &  $49.44 \pm 0.76$  \\
-                                                  & $25\%$ &  $50.23 \pm 0.16$  &  $49.43 \pm 0.22$  &     $\vsecond{53.30} \pm 0.15$  &      $\vfirst{53.46} \pm 0.23$  &  $49.46 \pm 0.24$  &  $44.37 \pm 0.50$  &  $51.51 \pm 0.48$  \\
-                                                  & time (ms)   &              21.2 &              5.7 &                           655 &                            1918 &              77042 &              56252 &              78 \\
-\midrule
-\multirow{7}{*}{\rotatebox[origin=c]{90}{\epi{}}} 
-                                                  & $3\%$  &  $43.51 \pm 0.35$  &  $41.39 \pm 1.11$  &     $\vsecond{51.47} \pm 0.28$  &     $\vfirst{52.04} \pm 0.52$  &  $31.00 \pm 0.66$  &  $36.84 \pm 0.43$  &  $34.42 \pm 3.59$  \\
-                                                  %& $5\%$  &  $49.16 \pm 0.30$  &  $46.65 \pm 0.83$  &  $\vsecondSig{54.69} \pm 0.33$  &  $\vfirstSig{56.29} \pm 0.42$  &  $38.00 \pm 0.37$  &  $39.73 \pm 0.37$  &  $43.24 \pm 1.81$  \\
-                                                  %& $7\%$  &  $52.35 \pm 0.21$  &  $50.27 \pm 0.55$  &  $\vsecondSig{56.77} \pm 0.33$  &  $\vfirstSig{58.40} \pm 0.46$  &  $43.91 \pm 0.43$  &  $41.93 \pm 0.28$  &  $46.05 \pm 1.17$  \\
-                                                  & $9\%$  &  $54.85 \pm 0.21$  &  $53.23 \pm 0.62$  &  $\vsecondSig{58.43} \pm 0.34$  &  $\vfirstSig{60.21} \pm 0.48$  &  $48.24 \pm 0.38$  &  $43.95 \pm 0.54$  &  $49.94 \pm 1.50$  \\
-                                                  & $15\%$ &  $59.29 \pm 0.15$  &  $57.76 \pm 0.42$  &  $\vsecondSig{61.41} \pm 0.19$  &  $\vfirstSig{62.69} \pm 0.34$  &  $56.88 \pm 0.56$  &  $48.61 \pm 0.88$  &  $54.56 \pm 1.27$  \\
-                                                  & $20\%$ &  $61.45 \pm 0.13$  &  $60.06 \pm 0.31$  &  $\vsecondSig{63.14} \pm 0.18$  &  $\vfirstSig{64.13} \pm 0.19$  &  $61.49 \pm 0.29$  &  $51.43 \pm 0.75$  &  $56.96 \pm 1.49$  \\
-                                                  & $25\%$ &  $62.95 \pm 0.18$  &  $61.93 \pm 0.29$  &  $\vsecondSig{64.47} \pm 0.29$  &  $\vfirstSig{65.22} \pm 0.61$  &  $64.45 \pm 0.39$  &  $54.51 \pm 0.99$  &  $58.73 \pm 1.65$  \\
-                                                  & time (ms)   &              32 &              7.1 &                            1226 &                           2341 &             116838 &             121530 &              129 \\
-\midrule
-\multirow{7}{*}{\rotatebox[origin=c]{90}{\kiw{}}} 
-                                                  & $3\%$  &               $26.02 \pm 0.27$  &  $\vsecondSig{26.23} \pm 0.46$  &   $\vfirstSig{33.92} \pm 0.50$  &  $23.59 \pm 0.43$  &              $20.02 \pm 0.68$  &  $20.13 \pm 0.56$  &   $1.11 \pm 0.26$  \\
-                                                  %& $5\%$  &               $30.76 \pm 0.18$  &  $\vsecondSig{30.95} \pm 0.29$  &   $\vfirstSig{36.46} \pm 0.53$  &  $26.60 \pm 0.53$  &              $26.33 \pm 0.77$  &  $23.09 \pm 0.47$  &   $3.35 \pm 0.58$  \\
-                                                  %& $7\%$  &               $33.44 \pm 0.26$  &  $\vsecondSig{33.45} \pm 0.26$  &   $\vfirstSig{37.57} \pm 0.47$  &  $30.79 \pm 1.43$  &              $30.53 \pm 0.50$  &  $25.00 \pm 0.39$  &   $6.54 \pm 1.42$  \\
-                                                  & $9\%$  &  $\vsecondSig{35.27} \pm 0.19$  &               $35.13 \pm 0.23$  &   $\vfirstSig{38.33} \pm 0.94$  &  $33.38 \pm 0.47$  &              $33.87 \pm 0.28$  &  $26.68 \pm 0.36$  &  $11.07 \pm 0.67$  \\
-                                                  & $15\%$ &               $38.21 \pm 0.10$  &               $37.72 \pm 0.15$  &  $\vsecondSig{38.63} \pm 0.59$  &  $36.81 \pm 0.31$  &  $\vfirstSig{40.14} \pm 0.19$  &  $29.97 \pm 0.37$  &  $18.12 \pm 1.04$  \\
-                                                  & $20\%$ &  $\vsecondSig{39.58} \pm 0.17$  &               $38.74 \pm 0.18$  &               $39.16 \pm 0.80$  &  $38.56 \pm 0.21$  &  $\vfirstSig{43.37} \pm 0.26$  &  $31.89 \pm 0.46$  &  $21.53 \pm 0.65$  \\
-                                                  & $25\%$ &  $\vsecondSig{40.48} \pm 0.11$  &               $39.48 \pm 0.14$  &               $39.14 \pm 0.78$  &  $39.80 \pm 0.26$  &  $\vfirstSig{45.76} \pm 0.15$  &  $34.23 \pm 0.42$  &  $23.89 \pm 0.54$  \\
-                                                  & time (ms)   &                           28 &                           6.6 &                           824 &               2938 &                         103264 &             130036 &              104 \\
-\bottomrule
-\end{tabular}
-\end{table*}
-***********************************************
+The last competitors is also local. \textcite{Bayesian15} note that  a node can belong to one of the
+16 types based of whether is number of its positive (resp.  negative) outgoing (resp. incoming)
+edges is zero or not.  The number of unobserved incoming and outgoing edges of each node $i$ let us
+define its 16-dimensional vector $V_i$ containing the probability of transitioning to any other
+state.  Then for each edge \eij we compute the outer product of $V_i$ with $V_j$ and also include
+additional features such as triads count and degree information before training a Logistic
+Regression model. We refer to this method as \emph{\compbayesian{}}.
 \fi
@@
   \scriptsize
-% \begin{longtable}{lrccc|cccc}
-\caption{MCC with increasing training set size, with one standard deviation over 12 random sampling of $\trainset$.
-The last four columns refer to the methods we took from the literature.
-For the sake of readability, we multiplied all MCC values by 100.
-The best number in each row is highlighted in \textbf{\textcolor{brown}{bold brown}} and the second one in \textit{\textcolor{red}{italic red}}.
-If the difference is statistically significant ($p$-value of a paired Student's $t$-test less than $0.005$), the best score is underlined.
-The ``time" rows contain the time taken to train on a $15\%$ training set.\label{tab:all_mcc}}
+  \caption{MCC with increasing training set size, with one standard deviation over 12 random
+  sampling of $\trainset$. The last four columns refer to the methods we took from the literature.
+For the sake of readability, we multiplied all MCC values by 100. The best number in each row is
+highlighted in \textbf{\textcolor{brown}{bold brown}} and the second one in
+\textit{\textcolor{red}{italic red}}. If the difference is statistically significant ($p$-value of a
+paired Student's $t$-test less than $0.005$), the best score is underlined. The ``time" rows contain
+the time taken to train on a $15\%$ training set.\label{tab:all_mcc}}
   \begin{tabular}{lrcccc|cccc}
@@
     \midrule
-% \multirow{7}{*}{\rotatebox[origin=c]{90}{\aut{}}} & $3\%$  &  $\vfirstSig{19.07} \pm 1.38$  &               $14.97 \pm 0.72$  &  $\vsecondSig{15.05} \pm 0.73$  &  $10.88 \pm 0.83$  &   $8.89 \pm 1.44$  &   $7.54 \pm 1.42$  &  $12.48 \pm 1.18$  &            $11.06 \pm 1.64$  \\
-%                                                   & $9\%$  &  $\vfirstSig{29.64} \pm 0.74$  &  $\vsecondSig{25.96} \pm 0.61$  &               $25.90 \pm 0.67$  &  $23.51 \pm 0.72$  &  $17.05 \pm 0.70$  &  $16.13 \pm 0.64$  &  $23.84 \pm 0.91$  &            $23.44 \pm 0.81$  \\
     \multirow{7}{*}{\rotatebox[origin=c]{90}{\aut{}}} & $5\%$                       & $\vfirstSig{24.54} \pm 0.69$  & $\vsecondSig{20.21} \pm 0.66$ & $20.19 \pm 0.71$              & $15.86 \pm 0.81$ & $12.76 \pm 0.65$ & $11.04 \pm 0.81$ & $17.18 \pm 1.11$             & $15.28 \pm 1.31$ \\
@@
     \midrule
-% \multirow{7}{*}{\rotatebox[origin=c]{90}{\wik{}}} & $3\%$  &   $\vfirstSig{34.08} \pm 1.03$  &  $31.65 \pm 0.99$  &  $\vsecondSig{32.08} \pm 0.94$  &  $28.50 \pm 0.92$  &  $19.73 \pm 0.99$  &   $3.98 \pm 0.84$  &  $27.43 \pm 0.99$  &  $20.11 \pm 1.14$  \\
-%                                                   & $9\%$  &      $\vfirst{45.80} \pm 0.43$  &  $44.92 \pm 0.48$  &     $\vsecond{45.72} \pm 0.42$  &  $43.02 \pm 0.52$  &  $30.24 \pm 0.54$  &  $23.50 \pm 1.66$  &  $41.36 \pm 0.66$  &  $38.05 \pm 0.93$  \\
     \multirow{7}{*}{\rotatebox[origin=c]{90}{\wik{}}} & $5\%$                       & $\vfirstSig{39.46} \pm 0.79$  & $38.03 \pm 0.97$              & $\vsecondSig{38.50} \pm 0.87$ & $35.72 \pm 0.70$ & $24.58 \pm 1.18$ & $9.59 \pm 1.10$  & $33.60 \pm 0.64$             & $26.45 \pm 0.57$ \\
@@
     \midrule
-% \multirow{7}{*}{\rotatebox[origin=c]{90}{\sla{}}} & $3\%$  &  $\vsecondSig{36.58} \pm 0.43$  &  $31.71 \pm 0.59$  &  $32.29 \pm 0.47$  &  $29.14 \pm 0.56$  &  $34.39 \pm 0.56$  &  $20.57 \pm 0.93$  &   $\vfirstSig{42.98} \pm 0.56$  &  $24.96 \pm 0.43$  \\
-%                                                   & $9\%$  &  $\vsecondSig{45.71} \pm 0.28$  &  $41.06 \pm 0.38$  &  $42.19 \pm 0.18$  &  $39.60 \pm 0.25$  &  $39.39 \pm 0.34$  &  $37.92 \pm 1.57$  &   $\vfirstSig{47.40} \pm 0.42$  &  $36.83 \pm 0.36$  \\
     \multirow{7}{*}{\rotatebox[origin=c]{90}{\sla{}}} & $5\%$                       & $\vsecondSig{40.77} \pm 0.20$ & $36.13 \pm 0.57$              & $37.00 \pm 0.29$              & $33.49 \pm 0.32$ & $36.83 \pm 0.47$ & $27.10 \pm 0.75$ & $\vfirstSig{45.16} \pm 0.59$ & $29.25 \pm 0.23$ \\
@@
     \midrule
-% \multirow{7}{*}{\rotatebox[origin=c]{90}{\epi{}}} & $3\%$  &  $\vsecondSig{51.43} \pm 0.29$  &  $41.42 \pm 0.84$  &  $43.41 \pm 0.34$  &  $36.56 \pm 0.49$  &  $37.01 \pm 0.57$  &  $31.71 \pm 6.20$  &  $\vfirstSig{52.33} \pm 0.42$  &            $31.25 \pm 0.80$  \\
-%                                                   & $9\%$  &  $\vsecondSig{58.48} \pm 0.26$  &  $53.20 \pm 0.62$  &  $54.80 \pm 0.12$  &  $51.58 \pm 0.36$  &  $43.81 \pm 0.50$  &  $50.43 \pm 2.07$  &  $\vfirstSig{60.06} \pm 0.46$  &            $48.06 \pm 0.41$  \\
     \multirow{7}{*}{\rotatebox[origin=c]{90}{\epi{}}} & $5\%$                       & $\vsecondSig{54.83} \pm 0.16$ & $46.94 \pm 0.80$              & $49.16 \pm 0.32$              & $42.79 \pm 0.34$ & $39.96 \pm 0.60$ & $42.94 \pm 2.06$ & $\vfirstSig{56.04} \pm 0.76$ & $37.99 \pm 0.49$ \\
@@
     \midrule
-% \multirow{7}{*}{\rotatebox[origin=c]{90}{\kiw{}}} & $3\%$  &   $\vfirstSig{33.88} \pm 0.54$  &  $\vsecondSig{26.02} \pm 0.80$  &               $26.02 \pm 0.42$  &  $15.13 \pm 0.40$  &  $19.93 \pm 0.56$  &   $0.86 \pm 0.11$  &  $23.50 \pm 0.54$  &              $20.10 \pm 0.60$  \\
-%                                                   & $9\%$  &   $\vfirstSig{38.62} \pm 0.52$  &               $35.10 \pm 0.17$  &  $\vsecondSig{35.29} \pm 0.11$  &  $28.23 \pm 0.33$  &  $26.61 \pm 0.37$  &   $9.99 \pm 0.93$  &  $33.22 \pm 0.29$  &              $33.86 \pm 0.24$  \\
     \multirow{7}{*}{\rotatebox[origin=c]{90}{\kiw{}}} & $5\%$                       & $\vfirstSig{36.36} \pm 0.53$  & $\vsecondSig{30.89} \pm 0.28$ & $30.81 \pm 0.20$              & $21.69 \pm 0.25$ & $23.15 \pm 0.26$ & $3.04 \pm 0.46$  & $26.63 \pm 0.44$             & $26.68 \pm 0.34$ \\
@@
 
-{\bf Results.} Our main results are summarized in Table~\ref{tab:all_mcc}, reporting MCC test set performance after training on sets of varying size (from 5\% to 25\%).
-Results have been averaged over 12 repetitions.
-Because scalability is a major concern
-%when training 
-on sizeable datasets, we also give an idea of relative training times (in milliseconds) by reporting the time it took to train a single run of each algorithm on a training set of size\footnote
-{
-Comparison of training time performances is fair since all algorithms have been carefully implemented using the same stack of Python libraries, and run on the same machine (16 Xeon cores and 192Gb Ram).
-} 
-15\% of $|E|$, and then predict on the test set. Though our experiments are not conclusive, some trends can be 
-%readily 
-spotted:
-%
-%\begin{itemize}%[nosep,leftmargin=2em]
-%
+\paragraph{Results}
+Our main results are summarized in \autoref{tab:all_mcc}, reporting MCC test set performance after
+training on sets of varying size (from 5\% to 25\%). Results have been averaged over 12 repetitions.
+Because scalability is a major concern when training on sizeable datasets, we also give an idea of
+relative training times (in milliseconds) by reporting the time it took to train a single run of
+each algorithm on a training set of size\footnote{Comparison of training time performances is fair
+since all algorithms have been carefully implemented using the same stack of Python libraries, and
+run on the same machine (16 Xeon cores and 192Gb Ram).} 15\% of $|E|$, and then predict on the test
+set. Though our experiments are not conclusive, some trends can be readily spotted:
 
-%\item 
-{\bf 1.}
-Global methods tend to outperform local methods in terms of prediction performance, but are also significantly (or even much) slower (running times can differ by as much as three orders of magnitude).
-This is not surprising, and is in line with previous experimental findings (e.g., \cite{shahriari2014ranking,wu2016troll}).
-\compbayesian{} looks like an exception to this rule, but its running time is indeed in the same ballpark as global methods.
-%
+\begin{enumerate}[leftmargin=2em,label=\textbf{\arabic*.}]
 
-%\item
-{\bf 2.} 
-\uslpropGsec{} always ranks first or at least second in this comparison when MCC is considered.
-On top of it, \uslpropGsec{} is fastest among the global methods (one or even two orders of magnitude faster), thereby showing the benefit of our approach to edge sign prediction.
-%\item
+  \item Global methods tend to outperform local methods in terms of prediction performance, but are
+    also significantly (or even much) slower (running times can differ by as much as three orders of
+    magnitude). This is not surprising, and is in line with previous experimental findings (e.g.,
+    \autocites{shahriari2014ranking}{wu2016troll}). \compbayesian{} looks like an exception to this
+    rule, but its running time is indeed in the same ballpark as global methods.
 
-{\bf 3.} The regularized solution computed by \uslpropGsec{} is always better than the unregularized one computed by \qoptim{} in terms of both MCC and running time.
+  \item \uslpropGsec{} always ranks first or at least second in this comparison when MCC is
+    considered. On top of it, \uslpropGsec{} is fastest among the global methods (one or even two
+    orders of magnitude faster), thereby showing the benefit of our approach to edge sign
+    prediction.
 
-{\bf 4.} 
-As claimed in the introduction, our Bayes approximator \usrule{} closely mirrors in performance the more involved \uslogregp{} model.
-In fact, supporting our generative model of Section~\ref{s:gen}, the logistic regression weights for features $1-\htr(i)$ and $1-\hun(j)$ are almost equal (see Table~2 in the supplementary material), thereby suggesting that predictor~\eqref{eq:predictor}, derived from the theoretical results at the beginning of Section~\ref{s:algbatch}, is {\em also} the best logistic model based on trollness and untrustworthiness.
-%This also contributes to giving our generative model of Section\ref{s:gen} a desireable robustness.
-%\end{itemize}
-%
+  \item The regularized solution computed by \uslpropGsec{} is always better than the unregularized
+    one computed by \qoptim{} in terms of both MCC and running time.
 
-\iffalse
-**********************************************
-In our experiments, we sample uniformly at random $3$, $5$, $7$, $9$, $15$,
-$20$, $30$, $40$, $50$, $70$, $80$ and $90$\% of the edges and train each
-method on that sample before predicting the edges not sampled. This procedure
-is repeated 10 times to account for the randomness of the sampling. As showed
-in \autoref{tab:mcc_all}, in the low regime (that
-is when $|\trainset{}|$ is less than $25\%$) our methods are always first or at
-least second. Furthermore, we can compare their runtime performances since all
-algorithms are implemented using the same stack of Python libraries and run
-on the same machine (16 Xeon cores and 192Gb Ram). As showed in
-\autoref{tab:time_per_edge}, our method are faster
-(\uslprop{}), up to several orders of magnitude (\usrule{}).
-**********************************************
-\fi
-
-\iffalse
-**********************************************************
-\begin{table}[b]
-	\centering
-	\caption{The time needed for each method to train on a $15\%$ training set and make its prediction.
-		It is measured in number of test edges predicted per microsecond, to illustrate that our methods scale linearly with the number of edges.
-		Out of fairness, we did not include the time needed to compute triads features needed by \compbayesian{} and \comptriads{} since our implementation is not optimal. 
-		% However, this is a $O(|V|^3)$ operation which would necessarily comes at a cost.
-	\label{tab:time_per_edge}}
-	\input{exp_data/time_per_edge}
-\end{table}
-
-\begin{figure}[tb]
-  \centering
-  \begin{subfigure}[b]{\textwidth}
-    \centering
-    \caption{The time (in millisecond) needed for each method to train on a
-      $15\%$ training set and make its prediction. Out of fairness, we did not
-			include the time needed to compute triads features needed by \compbayesian{}
-      and \comptriads{} since our implementation is not optimal. 
-		% However, this is a $O(|V|^3)$ operation which would necessarily comes at a cost.
-			\label{tab:time_total}}
-      \input{exp_data/time_total}
-  \end{subfigure}
-  \begin{subfigure}[b]{\textwidth}
-    \centering
-    \caption{The same data scaled by the number of edges. More precisely, it
-      measures the speed of each method by the number of test edges predicted
-      per microsecond.  Note that our methods scale linearly with the number of
-      edges. \label{tab:time_per_edge}}
-      \input{exp_data/time_per_edge}
-  \end{subfigure}
-\end{figure}
-**********************************************************
-\fi
+  \item As claimed in the introduction, our Bayes approximator \usrule{} closely mirrors in
+    performance the more involved \uslogregp{} model. In fact, supporting our generative model of
+    \autoref{s:gen}, the logistic regression weights for features $1-\htr(i)$ and $1-\hun(j)$
+    are almost equal (see Table~2 in the supplementary material), thereby suggesting that
+    predictor~\eqref{eq:predictor}, derived from the theoretical results at the beginning of
+    \autoref{s:algbatch}, is \emph{also} the best logistic model based on trollness and
+    untrustworthiness.
+\end{enumerate}

### 2017-09-04 troll/generative.tex
a pass on the rest of the troll main text

@@
 \section{Generative Model for Edge Labels}\label{s:gen}
+
 We now define the stochastic generative model for edge labels we use in the batch learning setting.
-Given the graph $G = (V,E)$, let the label $y_{i,j} \in \spin$ of directed edge $(i,j) \in E$ be generated as follows.
-Each node $i \in V$ is endowed with two latent parameters $p_i, q_i \in [0,1]$, which we assume to be generated, for each node $i$, by an independent draw from a fixed but unknown joint prior distribution $\mu(p,q)$ over $[0,1]^2$.
-Each label $y_{i,j} \in \spin$ is then generated by an independent draw from the mixture of $p_i$ and $q_j$,
-$
-	\Pr\big( y_{i,j} = 1 \big) = \tfrac{p_i + q_j}{2}~.
-$
-The basic intuition is that the nature $y_{i,j}$ of a relationship $i\rightarrow j$ is stochastically determined by a mixture between how much node $i$ tends to like other people ($p_i$) and how much node $j$ tends to be liked by other people ($q_j$).
-In a certain sense, $1-tr(i)$ is the empirical counterpart to $p_i$, and $1-un(j)$ is the empirical counterpart to $q_j$.\footnote
-{
-One might view our model as reminiscent of standard models for link generation in social network analysis, like the classical $p_1$ model from \cite{hl81}.
-Yet, the similarity falls short, for all these models aim at representing the likelihood of the network topology, rather than the probability of edge signs, once the topology is {\em given}.
-} 
-Notice that the Bayes optimal prediction for $y_{i,j}$ is
-$
-	y^*(i,j) = \sgn\big(\eta(i,j) - \tfrac{1}{2}\big)~,
-$
-where $\eta(i,j) = \Pr\big( y_{i,j} = 1 \big)$.
-Moreover, the probability of drawing at random a $+1$-labeled edge from $\Nout(i)$ and the probability of drawing at random a $+1$-labeled edge from $\Nin(j)$ are respectively equal to
-%
+Given the graph $G = (V,E)$, let the label $\yuv \in \spin$ of directed edge $(u,v) \in E$ be
+generated as follows. Each node $u \in V$ is endowed with two latent parameters $p_u, q_u \in
+[0,1]$, which we assume to be generated, for each node $u$, by an independent draw from a fixed but
+unknown joint prior distribution $\mu(p,q)$ over $[0,1]^2$. Each label $\yuv \in \spin$ is then
+generated by an independent draw from the mixture of $p_u$ and $q_v$
+$$\Pr\big( \yuv = 1 \big) = \tfrac{p_u + q_v}{2}$$
 
-\begin{small}
+The basic intuition is that the nature $\yuv$ of a relationship \euv{} is stochastically
+determined by a mixture between how much node $u$ tends to like other people ($p_u$) and how much
+node $v$ tends to be liked by other people ($q_v$). In a certain sense, $1-tr(u)$ is the empirical
+counterpart to $p_u$, and $1-un(v)$ is the empirical counterpart to $q_v$.\footnote{One might view
+our model as reminiscent of standard models for link generation in social network analysis, like the
+classical $p_1$ model from \cite{hl81}. Yet, the similarity falls short, for all these models aim at
+representing the likelihood of the network topology, rather than the probability of edge signs, once
+the topology is \emph{given}.}
+
+Notice that the Bayes optimal prediction for $\yuv$ is $y^*(u,v) = \sgn\big(\eta(u,v) -
+\tfrac{1}{2}\big)$, where $\eta(u,v) = \Pr\big( \yuv = 1 \big)$. Moreover, the probability of
+drawing at random a $+1$-labeled edge from $\Nout(u)$ equals
 \begin{equation}\label{e:pout}
-\frac{1}{2}\,\Biggl(p_i + \frac{1}{\dout(i)}\!\!\sum_{j\in \NNout(i)} \!\!\! q_j \Biggl) \,\,\,\,\text{and}\,\,\,\,
-%\end{equation}
-%and similarly, the probability of drawing at random a $+1$-labeled edge from $\Nin(j)$ %equals
-%
-%\begin{equation}\label{e:pin}
-\frac{1}{2}\,\Biggl(q_j + \frac{1}{\din(j)}\!\!\sum_{i\in \NNin(j)} \!\!\! p_i \Biggl)~.
+	\frac{1}{2}\,\Biggl(p_u + \frac{1}{\dout(u)} \sum_{v\in \NNout(u)} q_v \Biggl)
 \end{equation}
-\end{small}
-%
+Similarly, the probability of drawing at random a $+1$-labeled edge from $\Nin(v)$ equals
+\begin{equation}\label{e:pin}
+	\frac{1}{2}\,\Biggl(q_v + \frac{1}{\din(v)} \sum_{u\in \NNin(v)} p_u \Biggl)
+\end{equation}

### 2017-09-05 troll/supp_body.tex
first pass on troll appendix

--- 2017-09-02 troll/supp_body.tex
@@
-\section{Proofs from Section~\ref{s:algbatch}}
-\subsection{Proof of Theorem~\ref{t:active}}
+\section{Proofs from \autoref{s:algbatch}}
+
+\subsection{Proof of \autoref{t:active}}
+
 The following ancillary results will be useful.
-%
+
 \begin{lemma}[Hoeffding's inequality for sampling without replacement]
@@
 Let $\scX = \{x_1,\dots,x_N\}$ be a finite subset of $[0,1]$ and let
-\[
-	\mu = \frac{1}{N}\sum_{i=1}^N x_i~.
-\]
-If $X_1,\dots,X_n$ is a random sample drawn at random from $\scX$ without replacement, then, for every $\ve > 0$,
+\[ \mu = \frac{1}{N}\sum_{i=1}^N x_i~.  \]
+If $X_1,\dots,X_n$ is a random sample drawn at random from $\scX$ without replacement, then, for
+every $\ve > 0$,
 \[
@@
 \end{lemma}
-%
+
 \begin{lemma}\label{l:bins}
@@
 \end{lemma}
-%
+
 \begin{proof}[Proof of Lemma \ref{l:bins}]
-Set for brevity $p_i = |\NN_i|/|E|$.
-Then, due to the sampling without replacement, each random variable $|\NN_i \cap E_0|$ is the sum of $m$ dependent Bernoulli random variables $X_{i,1}, \ldots, X_{i,m}$ such that $\Pr(X_{i,t} = 1) = p_i$, for $t = 1, \ldots, m$.
-Let $i$ be such that $|\NN_i| \geq \theta$.
-Then the condition $m \geq \frac{2|E|Q}{\theta}$ implies
+Set for brevity $p_i = |\NN_i|/|E|$. Then, due to the sampling without replacement, each random
+variable $|\NN_i \cap E_0|$ is the sum of $m$ dependent Bernoulli random variables $X_{i,1}, \ldots,
+X_{i,m}$ such that $\Pr(X_{i,t} = 1) = p_i$, for $t = 1, \ldots, m$. Let $i$ be such that $|\NN_i|
+\geq \theta$. Then the condition $m \geq \frac{2|E|Q}{\theta}$ implies
 \[
@@
 \]
-Since the variables $X_{i,j}$ are negatively associated, we may apply a (multiplicative) Chernoff bound~\cite[Section~3.1]{dpbook}.
+Since the variables $X_{u,v}$ are negatively associated, we may apply a (multiplicative) Chernoff
+bound~\autocite[Section~3.1]{dpbook}.
 This gives
@@
 
-Let now $E_{\theta} = \theset{(i,j) \in E}{\din(j) \ge \theta,\, \dout(i) \ge \theta} \setminus E_0$, where $E_0 \ss E$ is the set of edges sampled by the learning algorithm of Section~\ref{s:algbatch}.
-Then Theorem~\ref{t:active} in the main paper is an immediate consequence of the following lemma.
-%
+Let now $E_{\theta} = \theset{(u,v) \in E}{\din(v) \ge \theta,\, \dout(u) \ge \theta} \setminus
+E_0$, where $E_0 \ss E$ is the set of edges sampled by the learning algorithm of
+\autoref{s:algbatch}. Then \autoref{t:active} in the main paper is an immediate consequence of the
+following lemma.
+
 \begin{lemma}\label{l:active}
-Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the model in Section~\ref{s:gen}.
-For all $0 < \alpha,\delta < 1$ and $0 < \ve < \tfrac{1}{16}$, if the learning algorithm of Section~\ref{s:algbatch} is run with parameters $Q = \tfrac{1}{2\ve^2}\ln\frac{4|V|}{\delta}$ and $\alpha$, then with probability at least $1-11\delta$ the predictions $\yhat(i,j)$ satisfy $\yhat(i,j) = y^*(i,j)$ for all $(i,j) \in E_{\theta}$ such that $\big|\eta(i,j) - \tfrac{1}{2}\big| > 8\ve$.
+Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the model
+in \autoref{s:gen}.
+For all $0 < \alpha,\delta < 1$ and $0 < \ve < \tfrac{1}{16}$, if the learning algorithm of
+\autoref{s:algbatch} is run with parameters $Q = \tfrac{1}{2\ve^2}\ln\frac{4|V|}{\delta}$ and
+$\alpha$, then with probability at least $1-11\delta$ the predictions $\yhat(u,v)$ satisfy
+$\yhat(u,v) = y^*(u,v)$ for all $(u,v) \in E_{\theta}$ such that $\big|\eta(u,v) - \tfrac{1}{2}\big|
+> 8\ve$.
 \end{lemma}
-%
+
 \begin{proof}[Proof of Lemma \ref{l:active}]
-We apply Lemma~\ref{l:bins} with $\theta = \tfrac{2Q}{\alpha} \ge 2\times\max\big\{Q,4\ln \frac{2|V|+1}{\delta}\big\}$ to the $2|V|+1$ subsets of $E$ consisting of $E_L$ and $\Nin(i),\Nout(i)$, for $i \in V$.
-We have that, with probability at least $1-\delta$, at least $Q$ edges of $E_L$ are sampled, at least $Q$ edges of $\Nin(i)$ are sampled for each $i$ such that $\NNin(i) \ge \theta$, and at least $Q$ edges of $\Nout(j)$ are sampled for each $j$ such that $\NNout(j) \ge \theta$.
-For all $(i,j) \in E_{\theta}$ let
+We apply \autoref{l:bins} with $\theta = \tfrac{2Q}{\alpha} \ge 2\times\max\big\{Q,4\ln
+\frac{2|V|+1}{\delta}\big\}$ to the $2|V|+1$ subsets of $E$ consisting of $E_L$ and
+$\Nin(u),\Nout(u)$, for $i \in V$. We have that, with probability at least $1-\delta$, at least $Q$
+edges of $E_L$ are sampled, at least $Q$ edges of $\Nin(u)$ are sampled for each $i$ such that
+$\NNin(u) \ge \theta$, and at least $Q$ edges of $\Nout(v)$ are sampled for each $j$ such that
+$\NNout(v) \ge \theta$.
+For all $(u,v) \in E_{\theta}$ let
 \[
-	\pbar_j = \frac{1}{\din(j)}\sum_{i\in\NNin(j)} p_i \quad\text{and}\quad \qbar_i = \frac{1}{\dout(i)}\sum_{j\in\NNout(i)} q_j
+  \pbar_j = \frac{1}{\din(v)}\sum_{i\in\NNin(v)} p_i \quad\text{and}\quad \qbar_i = \frac{1}{\dout(u)}\sum_{j\in\NNout(u)} q_j
 \]
-and set for brevity $\hdeltain(j) = 1-\hun(j)$ and $\hdeltaout(i) = 1-\htr(i)$.
-We now prove that $\hdeltain(j)$ and $\hdeltaout(i)$ are concentrated around their expectations for all $(i,j) \in E_{\theta}$.
-Consider $\hdeltaout(i)$ (the same argument works for $\hdeltain(j)$).
-Let $J_1,\dots,J_Q$ be the first $Q$ draws in $E_0 \cap \NNout(i)$ and define
+and set for brevity $\hdeltain(v) = 1-\hun(v)$ and $\hdeltaout(u) = 1-\htr(u)$.
+We now prove that $\hdeltain(v)$ and $\hdeltaout(u)$ are concentrated around their expectations for
+all $(u,v) \in E_{\theta}$. Consider $\hdeltaout(u)$ (the same argument works for $\hdeltain(v)$).
+Let $J_1,\dots,J_Q$ be the first $Q$ draws in $E_0 \cap \NNout(u)$ and define
 \[
-	\muhat_p(i) = \frac{1}{Q} \sum_{t=1}^Q \frac{p_i+q_{J_t}}{2}~.
+  \muhat_p(u) = \frac{1}{Q} \sum_{t=1}^Q \frac{p_i+q_{J_t}}{2}~.
 \]
-Applying Lemma~\ref{l:hoeff-wr} to the set $\theset{\tfrac{p_i+q_j}{2}}{j \in \NNout(i)}$, and using our choice of $Q$, we get that $\big|\muhat_p(i) - \mu_p(i)\big| \le \ve$ holds with probability at least $1-\delta/(2|V|)$, where
+Applying \autoref{l:hoeff-wr} to the set $\theset{\tfrac{p_i+q_j}{2}}{j \in \NNout(u)}$, and using
+our choice of $Q$, we get that $\big|\muhat_p(u) - \mu_p(u)\big| \le \ve$ holds with probability at
+least $1-\delta/(2|V|)$, where
 \[
-	\mu_p(i) = \frac{1}{\dout(i)}\sum_{j \in \NNout(i)} \frac{p_i + q_j}{2} = \frac{p_i + \qbar_i}{2}~.
+  \mu_p(u) = \frac{1}{\dout(u)}\sum_{j \in \NNout(u)} \frac{p_i + q_j}{2} = \frac{p_i + \qbar_i}{2}~.
 \]
-Now consider the random variables $Z_t = \Ind{y_{i,J_t}=1}$, for $t=1,\dots,Q$.
-Conditioned on $J_1,\dots,J_Q$, these are independent Bernoulli random variables with $\E[Z_t \mid J_t] = \tfrac{p_i+q_{J_t}}{2}$.
-Hence, applying a standard Hoeffding bound for independent variables and using our choice of $Q$, we get that
+Now consider the random variables $Z_t = \Ind{y_{u,v_t}=1}$, for $t=1,\dots,Q$. Conditioned on
+$J_1,\dots,J_Q$, these are independent Bernoulli random variables with $\E[Z_t \mid J_t] =
+\tfrac{p_i+q_{J_t}}{2}$. Hence, applying a standard Hoeffding bound for independent variables and
+using our choice of $Q$, we get that
 \[
-	\left|\frac{1}{Q}\sum_{t=1}^Q Z_t - \muhat_p(i) \right| \le \ve 
+  \left|\frac{1}{Q}\sum_{t=1}^Q Z_t - \muhat_p(u) \right| \le \ve 
 \]
 with probability at least $1-\delta/(2|V|)$ for every realization of $J_1,\dots,J_Q$.
-%
-Since $\hdeltaout(i) = (Z_1+\cdots+Z_Q)/Q$, we get that $\big|\hdeltaout(i) - \mu_p(i)\big| \le 2\ve$ with probability at least $1-2\delta/(2|V|)$.
-Applying the same argument to $\hdeltain(j)$, and the union bound\footnote
-{
-The sample spaces for the ingoing and outgoing edges of the vertices occurring as endpoints in $E_{\theta}$ overlap.
-Hence, in order to prove a uniform concentration result, we need to apply the union bound over the random variables defined over these sample spaces, which motivates the presence of the factor $\ln(2|V|)$ in the definition of $Q$.
-}
-on the set $\theset{\hdeltain(j),\hdeltaout(i)}{(i,j) \in E_{\theta}}$, we get that
+
+Since $\hdeltaout(u) = (Z_1+\cdots+Z_Q)/Q$, we get that $\big|\hdeltaout(u) - \mu_p(u)\big| \le
+2\ve$ with probability at least $1-2\delta/(2|V|)$. Applying the same argument to $\hdeltain(v)$,
+and the union bound\footnote{The sample spaces for the ingoing and outgoing edges of the vertices
+occurring as endpoints in $E_{\theta}$ overlap. Hence, in order to prove a uniform concentration
+result, we need to apply the union bound over the random variables defined over these sample spaces,
+which motivates the presence of the factor $\ln(2|V|)$ in the definition of $Q$.}
+on the set $\theset{\hdeltain(v),\hdeltaout(u)}{(u,v) \in E_{\theta}}$, we get that
 \begin{equation}
   \label{eq:first}
-	\left| \hdeltaout(i) + \hdeltain(j) - \frac{p_i + q_j}{2} - \frac{\pbar_j + \qbar_i}{2} \right| \le 4\ve
+  \left| \hdeltaout(u) + \hdeltain(v) - \frac{p_i + q_j}{2} - \frac{\pbar_j + \qbar_i}{2} \right| \le 4\ve
 \end{equation}
-simultaneously holds for all $(i,j) \in E_{\theta}$ with probability at least $1 - 4\delta$.
-Now notice that $\pbar_j$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables drawn from the prior marginal $\int_0^1 \mu\big(\cdot,q\bigr) dq$ with expectation $\mu_p$.
-Similarly, $\qbar_i$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables independently drawn from the prior marginal $\int_0^1 \mu\big(p,\cdot\big) dp$ with expectation $\mu_q$.
-By applying Hoeffding bound for independent variables, together with the union bound to the set of pairs of random variables whose sample means are $\pbar_j$ and $\qbar_i$ for each $(i,j) \in E_{\theta}$ (there are at most $2|V|$ of them) we obtain that
+simultaneously holds for all $(u,v) \in E_{\theta}$ with probability at least $1 - 4\delta$. Now
+notice that $\pbar_j$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables drawn from the
+prior marginal $\int_0^1 \mu\big(\cdot,q\bigr) dq$ with expectation $\mu_p$. Similarly, $\qbar_i$ is
+a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables independently drawn from the prior
+marginal $\int_0^1 \mu\big(p,\cdot\big) dp$ with expectation $\mu_q$. By applying Hoeffding bound
+for independent variables, together with the union bound to the set of pairs of random variables
+whose sample means are $\pbar_j$ and $\qbar_i$ for each $(u,v) \in E_{\theta}$ (there are at most
+$2|V|$ of them) we obtain that
 \begin{align*}
@@
 \end{align*}
-hold simultaneously for all $(i,j) \in E_{\theta}$ with probability at least $1-2\delta$. Combining with~(\ref{eq:first}) we obtain that
+hold simultaneously for all $(u,v) \in E_{\theta}$ with probability at least $1-2\delta$. Combining
+with~\eqref{eq:first} we obtain that
 \begin{equation}
   \label{eq:this}
-	\left| \hdeltaout(i) + \hdeltain(j) - \frac{p_i + q_j}{2} - \frac{\mu_p + \mu_q}{2} \right| \le 5\ve
+  \left| \hdeltaout(u) + \hdeltain(v) - \frac{p_i + q_j}{2} - \frac{\mu_p + \mu_q}{2} \right| \le 5\ve
 \end{equation}
-simultaneously holds for each $(i,j) \in E_{\theta}$ with probability at least $1 - 6\delta$. Next, let $E_L'$ be the set of the first $Q$ edges drawn in $E_L \cap E_0$. Then
+simultaneously holds for each $(u,v) \in E_{\theta}$ with probability at least $1 - 6\delta$. Next,
+let $E_L'$ be the set of the first $Q$ edges drawn in $E_L \cap E_0$. Then
 \[
-	\E\big[\tauhat\big] = \frac{1}{Q}\sum_{(i,j) \in E_L'} \Pr\big(y_{i,j} = 1\big) = \frac{1}{Q}\sum_{(i,j) \in E_L'} \frac{p_i+q_j}{2}~,
+  \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(u,v) \in E_L'} \Pr\big(y_{u,v} = 1\big) = \frac{1}{Q}\sum_{(u,v) \in E_L'} \frac{p_i+q_j}{2}~,
 \]
-where the expectation is w.r.t.\ the independent draws of the labels $y_{i,j}$ for $(i,j) \in E_L'$.
-Hence, by applying again Hoeffding bound (this time without the union bound) to the $Q = \tfrac{1}{2\ve^2}\ln\tfrac{2}{\delta}$ independent Bernoulli random variables $\Ind{y_{i,j} = 1}$, $(i,j) \in E_L'$, the event
+where the expectation is w.r.t.\ the independent draws of the labels $y_{u,v}$ for $(u,v) \in E_L'$.
+Hence, by applying again Hoeffding bound (this time without the union bound) to the $Q =
+\tfrac{1}{2\ve^2}\ln\tfrac{2}{\delta}$ independent Bernoulli random variables $\Ind{y_{u,v} = 1}$,
+$(u,v) \in E_L'$, the event
 $
@@
 \[
-	F(\bp,\bq) = \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(i,j) \in E_L'} \frac{p_i+q_j}{2}~.
+  F(\bp,\bq) = \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(u,v) \in E_L'} \frac{p_i+q_j}{2}~.
 \]
-For any realization $\bq_0$ of $\bq$, the function $F_1(\bp) = F(\bp,\bq_0)$ is a sample mean of $Q = \tfrac{1}{2\ve^2}\ln\tfrac{4|V|}{\delta}$ i.i.d.\ $[0,1]$-valued random variables $\theset{p_i}{(i,j) \in E_L'}$ (recall that if $i \in V$ is the origin of an edge $(i,j) \in E_L'$, then it is not the origin of any other edge $(i,j') \in E_L'$).
-Using again the standard Hoeffding bound, we obtain that
+For any realization $\bq_0$ of $\bq$, the function $F_1(\bp) = F(\bp,\bq_0)$ is a sample mean of $Q
+= \tfrac{1}{2\ve^2}\ln\tfrac{4|V|}{\delta}$ i.i.d.\ $[0,1]$-valued random variables
+$\theset{p_i}{(u,v) \in E_L'}$ (recall that if $i \in V$ is the origin of an edge $(u,v) \in E_L'$,
+then it is not the origin of any other edge $(u,v') \in E_L'$). Using again the standard Hoeffding
+bound, we obtain that
 \[
@@
 \]
-holds with probability at least $1 -\delta$ for each $\bq\in [0,1]^{|V|}$. With a similar argument, we obtain that
+holds with probability at least $1 -\delta$ for each $\bq\in [0,1]^{|V|}$. With a similar argument,
+we obtain that
 \[
@@
 \[
-	\left| \hdeltaout(i) + \hdeltain(j) - \tauhat - \frac{p(i) + q(j)}{2} \right| \le 8\ve
+  \left| \hdeltaout(u) + \hdeltain(v) - \tauhat - \frac{p(u) + q(v)}{2} \right| \le 8\ve
 \]
-simultaneously holds for each $(i,j) \in E_{\theta}$ with probability at least $1 - 10\delta$. Putting together concludes the proof.
+simultaneously holds for each $(u,v) \in E_{\theta}$ with probability at least $1 - 10\delta$.
+Putting together concludes the proof.
 \end{proof}
-%
-
-\iffalse
-****************************************************
-\begin{lemma}[Hoeffding's inequality]
-\label{l:hoeff}
-Let $X_1,\dots,X_n$ be independent random variables such that $X_i \in [0,1]$ and let
-\[
-	s_n = \frac{1}{n}\sum_{t=1}^n \big(X_t - \E[X_t]\big)~.
-\]
-Then, for every $\ve > 0$,
-$
-	\Pr\big(|s_n| \ge \ve\big) \le 2e^{-2n\ve^2}~.
-$
-\end{lemma}
-%
-****************************************************
-\fi
 
 \subsection{Derivation of the maximum likelihood equations}
-%
-Recall that the training set
-$
-	E_0 = \theset{\big(i_t,j_t),y_{i_t,j_t}\big)}{t=1,\dots,m}
-$
+
+Recall that the training set $\trainset = \theset{\big(u_t,v_t),y_{u_t,v_t}\big)}{t=1,\dots,m}$
 is drawn uniformly at random from $E$ without replacement. 
-%Set for brevity $m = |E_0|$. 
 We can write
-%
+
 \begin{align*}
-	\Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)\\ 
-&= 
-	\frac{1}{\binom{|E|}{m}\,m!}\prod_{k=1}^m \left(\frac{p_{i_k}+q_{j_k}}{2} \right)^{\Ind{y_{i_k,j_k}=+1}}\,\\
-&\quad\times\,\prod_{k=1}^m\left(1-\frac{p_{i_k}+q_{j_k}}{2} \right)^{\Ind{y_{i_k,j_k}=-1}}\\
+  \Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) = 
+  \frac{1}{\binom{|E|}{m}\,m!}\prod_{k=1}^m \left(\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=+1}}\,
+  \times\,\prod_{k=1}^m\left(1-\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=-1}}\\
   &=
@@
   \prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
-\left(\frac{p_{\ell}+q_{j_k}}{2} \right)^{\Ind{i_k = \ell,\,y_{\ell,j_k}=+1}}\,\\
-&\quad\times\,
-\prod_{k=1}^m\left(1-\frac{p_{\ell}+q_{j_k}}{2} \right)^{\Ind{i_k = \ell,\,y_{\ell,j_k}=-1}}\Biggl)
+    \left(\frac{p_{\ell}+q_{v_k}}{2} \right)^{\Ind{u_k = \ell,\,y_{\ell,v_k}=+1}}\,\times\,
+    \prod_{k=1}^m\left(1-\frac{p_{\ell}+q_{v_k}}{2} \right)^{\Ind{u_k = \ell,\,y_{\ell,v_k}=-1}}\Biggl)
 \end{align*}
-%
+
 so that $\log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)$ is proportional to
-%
+
+\begin{equation*}
+  \sum_{\ell=1}^{|V|}\left(\sum_{k=1}^m \Ind{u_k = \ell,y_{\ell,v_k}=+1}\,
+  \log\left(\frac{p_{\ell}+q_{v_k}}{2} \right)\, +\,
+  \sum_{\ell=1}^{|V|}\sum_{k=1}^m \Ind{u_k = \ell,y_{\ell,v_k}=+1}\, 
+\log\left(1-\frac{p_{\ell}+q_{v_k}}{2} \right)\right)
+\end{equation*}
+
+and
+
+\begin{equation*}
+  \frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial p_{\ell}}
+  = \sum_{k=1}^m 
+  \frac{\Ind{u_k = \ell,y_{\ell,v_k}=+1}}{p_{\ell}+q_{v_k}}\,
+   - \,
+  \sum_{k=1}^m \frac{\Ind{u_k = \ell,y_{\ell,v_k}=-1}}{2-p_{\ell}-q_{v_k}}\,.
+\end{equation*}
+
+By a similar argument,
+
 \begin{align*}
-&\sum_{\ell=1}^{|V|}\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=+1}\,
-\log\left(\frac{p_{\ell}+q_{j_k}}{2} \right)\,\\
-&+\,
-\sum_{\ell=1}^{|V|}\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=+1}\, 
-\log\left(1-\frac{p_{\ell}+q_{j_k}}{2} \right)
-\end{align*}
-%
-and
-%
-\begin{align*}
-\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial p_{\ell}}
-&= 
-\sum_{k=1}^m 
-\frac{\Ind{i_k = \ell,y_{\ell,j_k}=+1}}{p_{\ell}+q_{j_k}}\,\\
-& - \,
-\sum_{k=1}^m \frac{\Ind{i_k = \ell,y_{\ell,j_k}=-1}}{2-p_{\ell}-q_{j_k}}\,.
-\end{align*}
-%
-By a similar argument,
-%
-\begin{align*}
-\Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) 
-\\ &=
+  \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) 
+=
   \frac{1}{\binom{|E|}{m}\,m!}
   \prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
-\left(\frac{p_{i_k}+q_{\ell}}{2} \right)^{\Ind{j_k = \ell,\,y_{i_k,\ell}=+1}}\,\\
-&\quad\times\,
-\prod_{k=1}^m\left(1-\frac{p_{i_k}+q_{\ell}}{2} \right)^{\Ind{j_k = \ell,\,y_{i_k,\ell}=-1}}\Biggl)
+    \left(\frac{p_{u_k}+q_{\ell}}{2} \right)^{\Ind{v_k = \ell,\,y_{u_k,\ell}=+1}}\,
+\times\,
+    \prod_{k=1}^m\left(1-\frac{p_{u_k}+q_{\ell}}{2} \right)^{\Ind{v_k = \ell,\,y_{u_k,\ell}=-1}}\Biggl)
 \end{align*}
-%
+
 so that
-%
+
 \begin{align*}
 \frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial q_{\ell}}
-&= 
-\sum_{k=1}^m 
-\frac{\Ind{j_k = \ell,y_{i_k,\ell}=+1}}{p_{i_k}+q_{\ell}}\,\\ 
-&- \,
-\sum_{k=1}^m 
-\frac{\Ind{j_k = \ell,y_{i_k,\ell}=-1}}{2-p_{i_k}-q_{\ell}}\,.
+= \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=+1}}{p_{u_k}+q_{\ell}}\,
+- \, \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=-1}}{2-p_{u_k}-q_{\ell}}\,.
 \end{align*}
-%
+
 We then derive the approximation presented in the main paper. Namely, equating to zero the
@@
 \sum_{k=1}^m 
-\frac{\Ind{i_k = \ell,y_{\ell,j_k}=+1}}{p_{\ell}+q_{j_k}}\,
+\frac{\Ind{u_k = \ell,y_{\ell,v_k}=+1}}{p_{\ell}+q_{v_k}}\,
  - \,
-\frac{\Ind{i_k = \ell,y_{\ell,j_k}=-1}}{2-p_{\ell}-q_{j_k}} &= 0
+\frac{\Ind{u_k = \ell,y_{\ell,v_k}=-1}}{2-p_{\ell}-q_{v_k}} &= 0
 \end{align*}
-To simplify the notation, let $a_k = \Ind{i_k = \ell,y_{\ell,j_k}=+1}$, $b_k=\Ind{i_k =
-\ell,y_{\ell,j_k}=+1}$ and $c_k = p_{\ell} + q_{j_k}$, we can rewrite the previous equation as 
+
+To simplify the notation, let $a_k = \Ind{u_k = \ell,y_{\ell,v_k}=+1}$, $b_k=\Ind{u_k =
+\ell,y_{\ell,v_k}=+1}$ and $c_k = p_{\ell} + q_{v_k}$, we can rewrite the previous equation as 
 \begin{align*}
-\sum_{k=1}^m \frac{a_k}{c_k} - \frac{b_k}{2-c_k} &= 0 \\
-\sum_{k=1}^m  \frac{a_k(2-c_k) - b_k c_k }{ c_k(2-c_k) } &=0
-% \prod_{k=1}^m c_k\prod_{k=1}^m b_k\sum_{k=1}^m \frac{a_k}{b_k} &=\prod_{k=1}^m c_k\prod_{k=1}^m b_k \sum_{k=1}^m\frac{d_k}{c_k}
+  \sum_{k=1}^m \frac{a_k}{c_k} - \frac{b_k}{2-c_k} = 0 =
+  \sum_{k=1}^m  \frac{a_k(2-c_k) - b_k c_k }{ c_k(2-c_k) }
 \end{align*}
@@
 \begin{align*}
-\sum_{k=1}^m &\Ind{i_k = \ell,y_{\ell,j_k}=+1} \left(2-p_{\ell}-q_{j_k}\right)\\
-=
-&\sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,j_k}=-1}
-(p_{\ell}+q_{j_k})%, \quad \ell = 1, \ldots, |V|
+  \sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,v_k}=+1} \left(2-p_{\ell}-q_{v_k}\right)
+  = \sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,v_k}=-1}
+  (p_{\ell}+q_{v_k})
 \end{align*}
 
-\iffalse
 By applying a similar process to the gradient w.r.t. $q_\ell$, we obtain the
@@
 \begin{align*}
-\sum_{k=1}^m &\Ind{j_k = \ell,y_{i_k,\ell}=+1} \left(2-p_{i_k}-q_{\ell}\right)\\
-=
-&\sum_{k=1}^m  \Ind{j_k = \ell,y_{i_k,\ell}=-1}
-\left(p_{i_k}+q_{\ell}\right)
+\sum_{k=1}^m \Ind{v_k = \ell,y_{u_k,\ell}=+1} \left(2-p_{u_k}-q_{\ell}\right)
+= \sum_{k=1}^m  \Ind{v_k = \ell,y_{u_k,\ell}=-1}
+\left(p_{u_k}+q_{\ell}\right)
 \end{align*}
-\fi
 
 \subsection{Label propagation on $G''$}
-%
-Here we provide more details on the choice of weight for the edges of $G''$, as
-well as an explanation on why we temporarily use symmetrized variables lying in
-$[-1, 1]$ (which we will denote with primes, so that for instance $p'_i =
-2p_i-1$). Since only the ratio between the negative and positive weights
-matters, we fix the negative weight of the edges in $E''\setminus E'$ to be
-$-1$ and we denote by $\eps$ the weight of edges in $E'$.  With these notations,
-Label Propagation on $G''$ seeks the harmonic minimizer of the following
-expression 
+\label{ssec:troll_lprop_weights}
+
+Here we provide more details on the choice of weight for the edges of $G''$, as well as an
+explanation on why we temporarily use symmetrized variables lying in $[-1, 1]$ (which we will denote
+with primes, so that for instance $p'_u = 2p_u-1$). Since only the ratio between the negative and
+positive weights matters, we fix the negative weight of the edges in $E''\setminus E'$ to be $-1$
+and we denote by $\eps$ the weight of edges in $E'$.  With these notations, Label Propagation on
+$G''$ seeks the harmonic minimizer of the following expression
+
 \begin{equation*}
-  \frac{1}{16} \sum_{i,j \in E} \Bigl[ \eps\left(y_{i,j} - p_i'\right)^2 + \eps\left(y_{i,j} - q_j'\right)^2 + (p'_i+q_j')^2 \Bigr] \\
-  % \label{eq:lprop_tmp}
+  \frac{1}{16} \sum_{u,v \in E} \Bigl[ \eps\left(y_{u,v} - p_u'\right)^2 + \eps\left(y_{u,v} - q_v'\right)^2 + (p'_i+q_v')^2 \Bigr]
 \end{equation*}
+
 which can be successively rewritten as
+
 \begin{align*}
-  \frac{1}{16} \sum_{i,j \in E} & \Bigl[ \eps\left(y_{i,j}+1 - 2p_i\right)^2 + \eps\left(y_{i,j}+1 - 2q_j\right)^2
-\\&
-	+ (2p_i +2q_j -2)^2 \Bigr]
+  \frac{1}{16} \sum_{u,v \in E} & \Bigl[ \eps\left(y_{u,v}+1 - 2p_u\right)^2 + \eps\left(y_{u,v}+1 - 2q_v\right)^2
+	+ (2p_u +2q_v -2)^2 \Bigr]
 \\=
-%  \frac{1}{16} \sum_{i,j \in E} &\! 4\eps\!\left(\frac{y_{i,j}+1}{2} - p_i\right)^2\!\! +  4\eps\!\left(\frac{y_{i,j}+1}{2} - q_j\right)^2\!\! + 16\left(\frac{p_i +q_j -1}{2}\right)^2 \\
-	\frac{1}{8} \sum_{i,j \in E} & \Biggl[ 2\eps\!\left(\frac{y_{i,j}+1}{2} - p_i\right)^2\!\! +  2\eps\!\left(\frac{y_{i,j}+1}{2} - q_j\right)^2
-\\&
-	+ 8\left(\frac{p_i +q_j -1}{2}\right)^2 \Biggr]
+	\frac{1}{8} \sum_{u,v \in E} & \Biggl[ 2\eps\!\left(\frac{y_{u,v}+1}{2} - p_u\right)^2\!\! +  2\eps\!\left(\frac{y_{u,v}+1}{2} - q_v\right)^2
+	+ 8\left(\frac{p_u +q_v -1}{2}\right)^2 \Biggr]
 \\=
-  \frac{1}{8}\sum_{i,j \in E} & \Biggl[ 2\eps\left(\left(\frac{y_{i,j}+1}{2}\right)^2 - p_i(1+y_{i,j}) + p_i^2\right) + \\
-                              & 2\eps\left(\left(\frac{y_{i,j}+1}{2}\right)^2 - q_j(1+y_{i,j}) + q_j^2\right) +  \\
-                              & 8\left(\left(\frac{p_i +q_j}{2}\right)^2 - \frac{p_i+q_j}{2} + \frac{1}{4} \right) \Biggr]
+  \frac{1}{8}\sum_{u,v \in E} & \Biggl[ 2\eps\left(\left(\frac{y_{u,v}+1}{2}\right)^2 -
+    p_u(1+y_{u,v}) + p_u^2\right) + %\\
+                              2\eps\left(\left(\frac{y_{u,v}+1}{2}\right)^2 - q_v(1+y_{u,v}) + q_v^2\right) +  \\
+                              & 8\left(\left(\frac{p_u +q_v}{2}\right)^2 - \frac{p_u+q_v}{2} + \frac{1}{4} \right) \Biggr]
 \\=
-  \frac{1}{8}\sum_{i,j \in E} & 4\Biggl(\eps\!\left(\frac{y_{i,j}+1}{2}\right)^2\!\! -
-  2\eps\!\left(\frac{y_{i,j}+1}{2}\right)\left(\frac{p_i+q_j}{2}\right)
-\\&
-  + 2\!\left(\frac{p_i +q_j}{2}\right)^2\Biggr)
+  \frac{1}{8}\sum_{u,v \in E} & 4\Biggl(\eps\!\left(\frac{y_{u,v}+1}{2}\right)^2\!\! -
+  2\eps\!\left(\frac{y_{u,v}+1}{2}\right)\left(\frac{p_u+q_v}{2}\right)
+% \\&
+  + 2\!\left(\frac{p_u +q_v}{2}\right)^2\Biggr)
 \\&+
-  \sum_{i,j \in E} \Bigl[ \left(2\eps p_i^2 - 4p_i + 1\right) + \left(2\eps q_j^2 - 4q_j + 1\right) \Bigr]
+  \sum_{u,v \in E} \Bigl[ \left(2\eps p_u^2 - 4p_u + 1\right) + \left(2\eps q_v^2 - 4q_v + 1\right) \Bigr]
 \end{align*}
-%
+
 By setting $\eps=2$, we can factor this expression into 
 \begin{align*}
-  \sum_{i,j \in E} &\left(\frac{y_{i,j}+1}{2} - \frac{p_i+q_j}{2}\right)^2
-\\&
-  + \frac{1}{2}\sum_{i,j \in E} \left(\left(p_i - \frac{1}{2}\right )^2 + \left(q_j - \frac{1}{2}\right)^2 \right)~.
+  \sum_{u,v \in E} &\left(\frac{y_{u,v}+1}{2} - \frac{p_u+q_v}{2}\right)^2
+  + \frac{1}{2}\sum_{u,v \in E} \left(\left(p_u - \frac{1}{2}\right )^2 + \left(q_v -
+  \frac{1}{2}\right)^2 \right)~.
 \end{align*}
 
-\section{Proofs from Section~\ref{s:algonline}}
-%
+\section{Proofs from \autoref{s:algonline}}
 
-\begin{proof}[Proof of Theorem~\ref{t:online}]
-Let each node $i \in V$ host two instances of the randomized Weighted Majority (RWM) algorithm~\cite{LittlestoneWa94} with an online tuning of their learning rate~\cite{cb+97,acg02}: one instance for predicting the sign of outgoing edges $(i,j)$, and one instance for predicting the sign of incoming edges $(j,i)$.
-Both instances simply compete against the two constant experts, predicting always $+1$ or always $-1$.
-Denote by $M(i,j)$ the indicator function (zero-one loss) of a mistake on edge $(i,j)$.
-Then the expected number of mistakes of each RWM instance satisfy~\cite{cb+97,acg02}:
+\begin{proof}[Proof of \autoref{t:online}]
+
+Let each node $u \in V$ host two instances of the randomized Weighted Majority (RWM)
+algorithm~\cite{LittlestoneWa94} with an online tuning of their learning rate~\cite{cb+97,acg02}:
+one instance for predicting the sign of outgoing edges $(u,v)$, and one instance for predicting the
+sign of incoming edges $(v,u)$. Both instances simply compete against the two constant experts,
+predicting always $+1$ or always $-1$. Denote by $M(u,v)$ the indicator function (zero-one loss) of
+a mistake on edge $(u,v)$. Then the expected number of mistakes of each RWM instance
+satisfy~\cite{cb+97,acg02}:
 \[
-    \sum_{j \in \NNout(i)} \E\,M(i,j) = \Psiout(i,Y) + O\left(\sqrt{\Psiout(i,Y)}+ 1\right)
+    \sum_{v \in \NNout(u)} \E\,M(u,v) = \Psiout(u,Y) + O\left(\sqrt{\Psiout(u,Y)}+ 1\right)
 \]
@@
 \[
-    \sum_{i \in \NNin(j)} \E\,M(i,j) = \Psiin(j,Y) + O\left(\sqrt{\Psiin(j,Y)} + 1\right)~.
+    \sum_{u \in \NNin(v)} \E\,M(u,v) = \Psiin(v,Y) + O\left(\sqrt{\Psiin(v,Y)} + 1\right)~.
 \]
-We then define two meta-experts: an ingoing expert, which predicts $y_{i,j}$ using the prediction of the ingoing RWM instance for node $j$, and the outgoing expert, which predicts $y_{i,j}$ using the prediction of the outgoing RWM instance for node $i$.
+We then define two meta-experts: an ingoing expert, which predicts $y_{u,v}$ using the prediction of
+the ingoing RWM instance for node $v$, and the outgoing expert, which predicts $y_{u,v}$ using the
+prediction of the outgoing RWM instance for node $u$.
 The number of mistakes of these two experts satisfy
 \begin{align*}
-\sum_{i \in V}\sum_{j \in \NNout(i)} &\E\,M(i,j)\\ 
+  \sum_{u \in V}\sum_{v \in \NNout(u)} \E\,M(u,v)
                                        &= \Psiout(Y) + O\left(\sqrt{|V|\Psiout(Y)} + |V|\right)\\
-\sum_{j \in V}\sum_{i \in \NNin(j)} &\E\,M(i,j)\\ 
+  \sum_{v \in V}\sum_{u \in \NNin(v)} \E\,M(u,v)
                                       &= \Psiin(Y)  + O\left(\sqrt{|V|\Psiin(Y)}  + |V|\right)~,
 \end{align*}
-where we used $\sum_{j \in V} \sqrt{\Psiin(j,Y)} \leq \sqrt{|V|\Psiin(Y)}$, and similarly for $\Psiout(Y)$.
-Finally, let the overall prediction of our algorithm be a RWM instance run on top of the ingoing and the outgoing experts. Then the expected number of mistakes of this predictor satisfies
-%
+where we used $\sum_{v \in V} \sqrt{\Psiin(v,Y)} \leq \sqrt{|V|\Psiin(Y)}$, and similarly for
+$\Psiout(Y)$.
+Finally, let the overall prediction of our algorithm be a RWM instance run on top of the ingoing and
+the outgoing experts. Then the expected number of mistakes of this predictor satisfies~
 \begin{align*}
-    \sum_{(i,j) \in E} \E\,M(i,j) 
-&= \Psi_G(Y) + O\Biggl(\sqrt{|V|\Psi_G(Y)} + |V|\\ 
-&\,\,+ \sqrt{\left(\Psi_G(Y) + |V| + \sqrt{|V|\Psi_G(Y)}\right)} \Biggl)\\
-&= \Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V|\right)~,
+  \sum_{(u,v) \in E} \E\,M(u,v) &= \Psi_G(Y) + O\Biggl(\sqrt{|V|\Psi_G(Y)} + |V|
+    + \sqrt{\left(\Psi_G(Y) + |V| + \sqrt{|V|\Psi_G(Y)}\right)} \Biggl)\\
+    &= \Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V|\right)
 \end{align*}
-%
 as claimed.
@@
 
-\begin{proof}[Proof sketch of Theorem~\ref{t:mistake_bound}]
-Let $\mathcal{Y}_K$ be the set of all labelings $Y$ such that the total number of negative and positive edges are $K$ and $|E|-K$, respectively (without loss of generality we will focus on negative edges).
-Consider the randomized strategy that draws a labeling
-$Y \in \spin^{|E|}$
-% for the edges of the input graph 
-uniformly at random from $\mathcal{Y}_K$.
-For each node $i \in V$, we have $\Psiin(i,Y) \le \din^-(i)$, which implies $\Psiin(Y) \le K$.
-A very similar argument applies to the outgoing edges, leading to $\Psiout(Y) \le K$.
-The constraint $\Psi_G(Y) \le K$ is therefore always satisfied.
+\begin{proof}[Proof sketch of \autoref{t:mistake_bound}]
 
-The adversary will force on average $1/2$ mistakes in each one of the first $K$ rounds of the online protocol by repeating $K$ times the following: (i) A label value $\ell\in\{-1,+1\}$ is selected uniformly at random.
-(ii) An edge $(i,j)$ is sampled uniformly at random from the set of all edges that were not previously revealed and whose labels are equal to $\ell$.
+Let $\mathcal{Y}_K$ be the set of all labelings $Y$ such that the total number of negative and
+positive edges are $K$ and $|E|-K$, respectively (without loss of generality we will focus on
+negative edges). Consider the randomized strategy that draws a labeling $Y \in \spin^{|E|}$ of the
+edges of the input graph \uar{} from $\mathcal{Y}_K$. For each node $u \in V$, we have $\Psiin(u,Y)
+\le \din^-(u)$, which implies $\Psiin(Y) \le K$. A very similar argument applies to the outgoing
+edges, leading to $\Psiout(Y) \le K$. The constraint $\Psi_G(Y) \le K$ is therefore always
+satisfied.
 
-The learner is required to predict $y_{i,j}$ and, in doing so, $1/2$ mistakes will be clearly made on average because of the randomized labeling procedure.
-Observe that this holds even when $A$ knows the value of $K$ and $\Psi_G(Y)$.
-Hence, we can conclude that the expected number of mistakes that $A$ can be forced to make is always at least $K/2$, as claimed.
+The adversary will force on average \shalf{} mistakes in each one of the first $K$ rounds of the
+online protocol by repeating $K$ times the following: (i) A label value $\ell\in\{-1,+1\}$ is
+selected uniformly at random. (ii) An edge $(u,v)$ is sampled uniformly at random from the set of
+all edges that were not previously revealed and whose labels are equal to $\ell$.
 
-We now show that, as $\frac{K}{|E|} \rightarrow 0$, the lower bound gets arbitrarily close to $K$ for any $G(Y)$ and any constant $K$.
+The learner is required to predict $y_{u,v}$ and, in doing so, \shalf{} mistakes will be clearly
+made on average because of the randomized labeling procedure. Observe that this holds even when $A$
+knows the value of $K$ and $\Psi_G(Y)$. Hence, we can conclude that the expected number of mistakes
+that $A$ can be forced to make is always at least $K/2$, as claimed.
+
+\medskip
+
+We now show that, as $\frac{K}{|E|} \rightarrow 0$, the lower bound gets arbitrarily close to $K$
+for any $G(Y)$ and any constant $K$.
+Let $\mathcal{E}$ be the following event: There is at least one unrevealed negative label.
+The randomized iterative strategy used to achieve this result is identical to the one described
+above, except for the stopping criterion. Instead of repeating step (i) and (ii) only for the first
+$K$ rounds, these steps are repeated until $\mathcal{E}$ is true.
 %
-Let $\mathcal{E}$ be the following event: There is at least one unrevealed negative label.
-The randomized iterative strategy used to achieve this result is identical to the one described above, except for the stopping criterion.
-Instead of repeating step (i) and (ii) only for the first $K$ rounds, these steps are repeated until $\mathcal{E}$ is true.
-%
-Let $m_{r,c}$ be defined as follows: For $c=1$ it is equal to the expected number of mistakes forced in round $r$ when $K=1$.
-For $c > 1$ it is equal to the difference between the expected number of mistakes forced in round $r$ when $K=c$ and $K=c-1$.
-One can see that $m_{r,c}$ is
+Let $m_{r,c}$ be defined as follows: For $c=1$ it is equal to the expected number of mistakes forced
+in round $r$ when $K=1$. For $c > 1$ it is equal to the difference between the expected number of
+mistakes forced in round $r$ when $K=c$ and $K=c-1$. One can see that $m_{r,c}$ is 
 %either 
@@
 %
-When $K=1$, the probability that $\mathcal{E}$ is true in round $r$ is clearly equal to $\frac{1}{2^{r-1}}$.
+When $K=1$, the probability that $\mathcal{E}$ is true in round $r$ is clearly equal to
+$\frac{1}{2^{r-1}}$.
 Hence, the expected number of mistakes made by $A$ when $K=1$ in any round $r$ is equal to 
@@
 \]  
-In order to calculate $m_{r,c}$ for all $r$ and $c$, we will rest on the ancillary quantity $s_j(i)$, recursively defined as specified next.
-
-Given any integer variable $i$, we have
-%we define the following recurrence relation identity $s_j(i)$:
-$
-s_0(i)=1
-$
-and, for any positive integer $j$, 
+In order to calculate $m_{r,c}$ for all $r$ and $c$, we will rest on the ancillary quantity
+$s_j(i)$, defined recursively for any positive integer variables $i$ and $j$ by
+$ s_0(i)=1 $
+and
 \[
@@
 \]
-%(see Table~\ref{tab:sij}) for examples)
+(see \autoref{tab:sij}) for examples)
 
-\iffalse
 \begin{center}
@@
   \end{tabular}
-  \captionof{table}{\ \ Values of $s_j(i)$ for $i \le 7$  and $j \le 3$.} \label{tab:sij} 
+  \captionof{table}{Values of $s_j(i)$ for $i \le 7$  and $j \le 3$.} \label{tab:sij} 
 \end{center}
-\fi
 
@@
 \]
-Since $s_j(i)=\frac{\langle i\rangle_j}{j!}$, where $\langle i\rangle_j$ is the rising factorial $i(i+1)(i+2)\ldots(i+j-1)$, we have 
+Since $s_j(i)=\frac{\langle i\rangle_j}{j!}$, where $\langle i\rangle_j$ is the rising factorial
+$i(i+1)(i+2)\ldots(i+j-1)$, we have 
 \[
@@
 \] 
-%(see Table~\ref{tab:mrc}) for examples)
+(see \autoref{tab:mrc}) for examples)
 
-\iffalse
 \begin{center}
@@
     \hline
-    \diaghead{c\ r}{\ \ \ 	{\footnotesize \textit{\textbf{c}}}\ }{\ \
-    {\footnotesize \textit{\textbf{r}}}\ \ \ } & \textbf{1}& \textbf{2} &
-      \textbf{3} & \textbf{4}                            &\\[13pt] \hline
-    \textbf{1} & $1/2^1$ & 0       & 0        & 0        &\\[13pt] \hline
-    \textbf{2} & $1/2^2$ & $1/2^2$ & 0        & 0        &\\[13pt] \hline
-    \textbf{3} & $1/2^3$ & $2/2^3$ & $1/2^3$  & 0        &\\[13pt] \hline
-    \textbf{4} & $1/2^4$ & $3/2^4$ & $3/2^4$  & $1/2^4$  &\\[13pt] \hline
-    \textbf{5} & $1/2^5$ & $4/2^5$ & $6/2^5$  & $4/2^5$  &\\[13pt] \hline
-    \textbf{6} & $1/2^6$ & $5/2^6$ & $10/2^6$ & $10/2^6$ &\\[13pt] \hline
-    \textbf{7} & $1/2^7$ & $6/2^7$ & $15/2^7$ & $20/2^7$ &\\[13pt] \hline
+    \diaghead{\theadfont i}%
+    {{\footnotesize \textit{\textbf{c}}}}{{\footnotesize \textit{\textbf{r}}}}
+               & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} &  \\ [13pt] \hline
+    \textbf{1} & $1/2^1$    & 0          & 0          & 0          &  \\ [13pt] \hline
+    \textbf{2} & $1/2^2$    & $1/2^2$    & 0          & 0          &  \\ [13pt] \hline
+    \textbf{3} & $1/2^3$    & $2/2^3$    & $1/2^3$    & 0          &  \\ [13pt] \hline
+    \textbf{4} & $1/2^4$    & $3/2^4$    & $3/2^4$    & $1/2^4$    &  \\ [13pt] \hline
+    \textbf{5} & $1/2^5$    & $4/2^5$    & $6/2^5$    & $4/2^5$    &  \\ [13pt] \hline
+    \textbf{6} & $1/2^6$    & $5/2^6$    & $10/2^6$   & $10/2^6$   &  \\ [13pt] \hline
+    \textbf{7} & $1/2^7$    & $6/2^7$    & $15/2^7$   & $20/2^7$   &  \\ [13pt] \hline
   \end{tabular}
-  \captionof{table}{\ \ Values of $m_{r,c}$ for $r \le 7$, $c \le 4$ and $|E| \rightarrow \infty$.} \label{tab:mrc} 
+  \captionof{table}{Values of $m_{r,c}$ for $r \le 7$, $c \le 4$ and $|E| \rightarrow \infty$.}
+  \label{tab:mrc} 
 \end{center}
-\fi
+
 \bigskip
@@
 
-When $\frac{K}{|E|} \rightarrow 0$, given any integer $K'>1$, the difference between the expected number of mistakes forced when $K=K'$ and $K=K'-1$ is equal to
+When $\frac{K}{|E|} \rightarrow 0$, given any integer $K'>1$, the difference between the expected
+number of mistakes forced when $K=K'$ and $K=K'-1$ is equal to
 %
@@
 \[
-    \Psiin(Y) = \sum_{j \in V} \min\big\{\din^-(j),\din^+(j)\big\}
+    \Psiin(Y) = \sum_{j \in V} \min\big\{\din^-(v),\din^+(v)\big\}
 \]
@@
 \[
-    \Psiout(Y) = \sum_{i \in V} \min\big\{\dout^-(i),\dout^+(i)\big\}
+    \Psiout(Y) = \sum_{i \in V} \min\big\{\dout^-(u),\dout^+(u)\big\}
 \]
@@
 \[
-    \Yhat^{(1)}_{i,j} = \sgn\left(\frac{1}{2}-\frac{\hdout^-(i)}{\hdout(i)}\right)
+    \Yhat^{(1)}_{u,v} = \sgn\left(\frac{1}{2}-\frac{\hdout^-(u)}{\hdout(u)}\right)
 \qquad\text{and}\qquad
-    \Yhat^{(2)}_{i,j} = \sgn\left(\frac{1}{2}-\frac{\hdin^-(j)}{\hdin(j)}\right)
+    \Yhat^{(2)}_{u,v} = \sgn\left(\frac{1}{2}-\frac{\hdin^-(v)}{\hdin(v)}\right)
 \]
@@
 \section{Further Experimental Results}
-%
-This section contains more evidence related to the experiments in Section~\ref{s:exp}.
+
+This section contains more evidence related to the experiments in \autoref{s:exp}.
 In particular, we experimentally demonstrate the alignment between \usrule{} and \uslogregp{}.
@@
 \iffalse
-*************************************
 \subsection{Accuracy Results}
 %
-Table~\ref{tab:all_acc} contains the accuracy results on training sets of increasing size. Figures are averaged over 12 random draws of the training set.
+\autoref{tab:all_acc} contains the accuracy results on training sets of increasing size. Figures are
+averaged over 12 random draws of the training set.
 
@@
 
-After training on the two features $1-\htr(i)$ and $1-\hun(j)$, \uslogregp{} has learned three weights
-$w_0$, $w_1$ and $w_2$, which allow to predict $y_{i,j}$ according to 
+After training on the two features $1-\htr(u)$ and $1-\hun(v)$, \uslogregp{} has learned three
+weights $w_0$, $w_1$ and $w_2$, which allow to predict $y_{u,v}$ according to 
 \[ 
-  \sgn\Big(\big(w_1(1-\htr(i)\big) + w_2 \big(1-\hun(j)\big) + w_0\Big)~.
+  \sgn\Big(\big(w_1(1-\htr(u)\big) + w_2 \big(1-\hun(v)\big) + w_0\Big)~.
 \]
@@
 \[
-\sgn\Big(\big(1-\htr(i)\big) + w_2'\big(1-\hun(j)\big) - \tfrac{1}{2} -\tau'\Big)~,
+\sgn\Big(\big(1-\htr(u)\big) + w_2'\big(1-\hun(v)\big) - \tfrac{1}{2} -\tau'\Big)~,
 \]
@@
 
-As shown in Table~\ref{tab:coeff}, and in accordance with the predictor built out of Equation~\eqref{eq:predictor},
-% from the main paper,
-$w_2'$ is almost $1$ on {\em all datasets}, while $\tau'$ tends to be always close the fraction of positive edges in the dataset.
+As shown in \autoref{tab:coeff}, and in accordance with the predictor built out of Equation
+\eqref{eq:predictor}, $w_2'$ is almost $1$ on \emph{all datasets}, while $\tau'$ tends to be always
+close the fraction of positive edges in the dataset.
 
@@
   \centering
-  \caption{Normalized logistic regression coefficients averaged over 12 runs (with one standard deviation) \label{tab:coeff}}
+  \caption{Normalized logistic regression coefficients averaged over 12 runs (with one standard
+  deviation) \label{tab:coeff}}
   \small

### 2017-09-05 assets/tikz/troll_notations_tikz.tex
move and add figures in the troll notations part

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
  \begin{tikzpicture}[auto,vertex/.append style={minimum width=10mm,font=\Large,fill=White},
        scale=1.0,]
      \tikzset{>=latex}
      \tikzstyle{edge}=[line width=2pt,color=Black,-{Latex[length=3mm,width=1.0mm,angle'=40]}]
      \tikzstyle{trainEdge}=[line width=10pt,color=Black!15,opacity=.7]
      \tikzstyle{elabel}=[color=Black]
      \tikzstyle{posc}=[scale=1.2,text=Green]
      \tikzstyle{negc}=[scale=1.2,text=red]
      \tikzstyle{testset}=[opacity=1.0]
      \tikzstyle{revealed}=[circle,draw=none,fill=Black!15,inner sep=1pt,]

      \node[vertex] (u) at (0,0) {\LARGE $u$};
      \node[vertex] (v) at (5.5,-1.26) {\LARGE $v$};

      \draw[trainEdge] (u) --                                                       (5, -2.5);
      \draw[edge] (u) -- node [revealed,posc,xshift=-2mm,] {$\mathbf{+}$}           (5, -2.5);

      \draw[testset,edge] (u) edge  node [posc,xshift=-5mm]
        {${\textcolor{Black}{y_{u,v}=}}{\textcolor{Green}{+}}$} (v);
      \draw[testset,edge] (u) edge  node [posc,xshift=2mm,pos=0.7] {$\mathbf{+}$}            (5, 0.66);
      \draw[testset,edge, dashed] (u)   edge   node [negc,xshift=2mm] {$\mathbf{-}$} (5, 2);

      \begin{scope}[on background layer]
        \draw[trainEdge] (-5, -2) --    (u.south west);
        \draw[edge, dashed] (-5, -2)       -- node[revealed,negc,yshift=-3mm,xshift=-5pt]
          {$\mathbf{-}$} (u.south west);
        \draw[trainEdge,]          (-5, 2) -- (u.north west);
        \draw[edge,]               (-5, 2) -- node[revealed,posc,above,pos=0.2,xshift=5pt] {$\mathbf{+}$} (u.north west);
      \end{scope}

      \draw[testset,edge,dashed] (-5, -1)-- node[negc,yshift=-1mm] {$\mathbf{-}$} (u);
      \draw[testset,edge,dashed] (-5, 0) -- node[negc] {$\mathbf{-}$} (u);
      \draw[testset,edge,]       (-5, 1) -- node[posc,xshift=-4mm] {$\mathbf{+}$} (u);

      \node[fill=white!90!DarkOrange,fill opacity=.9,text opacity=1,rounded corners,inner sep=2pt] at (1,-1.5) {$tr(u)=\frac{1}{4}$};
      \node[fill=white!90!Blue,fill opacity=.5,text opacity=1,rounded corners,inner sep=2pt] at (-1.0,1.7) {$un(u)=\frac{3}{5}$};
\end{tikzpicture}
\end{document}

### 2017-09-05 assets/tikz/troll_genmodel_tikz.tex
add one picture for generative model

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
  \begin{tikzpicture}[auto,vertex/.append style={minimum width=10mm,font=\Large},
    edge/.append style={-{Latex[length=4mm,width=1.5mm,angle'=40]}},]
    \node[vertex] (u) at (0,0) {$u$};
    \node[anchor=south] (iparams) at (0,0.5) {$(\outqt{p_u}, q_u) \sim \mu(p, q)$};
    \node[vertex,] (v) at (6,0) {$v$};
    \node[anchor=south] (jparams) at (6,0.5) {$(p_v, \inqt{q_v}) \sim \mu(p, q)$};
    \draw[edge,opacity=1] (u) edge node[below,yshift=-2mm] {$\mathbb{P}(y_{u,v}=+1) = \frac{1}{2}\left(\outqt{p_u}+\inqt{q_v}\right)$} (v);
  \end{tikzpicture}
\end{document}

### 2017-09-05 troll/generative.tex
add one picture for generative model

@@
 $$\Pr\big( \yuv = 1 \big) = \tfrac{p_u + q_v}{2}$$
+This can be seen pictorially in \autoref{fig:troll_genmodel}.
+\begin{marginfigure}
+	\centering
+	\includegraphics[width=0.98\linewidth]{tikz/troll_genmodel_tikz}
+	\caption{The sign \yuv{} of the edge \euv{} is positive with probability $\frac{1}{2}(p_u+q_v)$.
+	\label{fig:troll_genmodel}}
+\end{marginfigure}
 

### 2017-09-05 troll/algs_batch.tex
move all troll related work in one file

@@
 championing in our experiments of Section~\ref{s:exp}.
-
-\label{ssec:further_related}
-{\bf Further related work.} The vast majority of existing edge sign prediction algorithms for
-directed graphs are based on the computation of local features of the graph. These features are
-evaluated on the subgraph induced by the training edges, and the resulting values are used to train
-a supervised classification algorithm (e.g., logistic regression). The most basic set of local
-features used to classify a given edge $(u,v)$ are defined by
-$\din^+(v),\din^-(v),\dout^+(u),\dout^-(u)$ computed over the training set \trainset{}, and by the
-embeddedness coefficient $\big|\Nout(u) \cap \Nin(v)\big|$. In turn, these can be used to define
-more complicated features, such as
-\(
-	\frac{\din^+(v) + |E^+|\uin(v)}{\din(v) + \uin(v)}
-\quad\text{and}\quad
-	\frac{\dout^+(u) + p^+\uout(u)}{\dout(u) + \uout(u)}
-\)
-introduced in~\autocite{Bayesian15}, together with their negative counterparts, where $|E^+|$ is the
-overall fraction of positive edges, and $\uin(v),\uout(u)$ are, respectively, the number of test
-edges outgoing from $u$ and the number of test edges incoming to $v$.
-Other types of features are derived from social status theory~(e.g., \autocite{Leskovec2010}), and
-involve the so-called triads; namely, the triangles formed by $(u,v)$ together with $(u,w)$ and
-$(w,v)$ for any $w \in \NNout(u) \cap \NNin(v)$. A third group of features is based on node ranking
-scores. These scores are computed using a variery of methods, including
-Prestige~\autocite{zolfaghar2010mining}, exponential ranking~\autocite{traag2010exponential},
-PageTrust~\autocite{de2008pagetrust}, Bias and Deserve~\autocite{mishra2011finding},
-TrollTrust~\autocite{wu2016troll}, and generalizations of PageRank and HITS to signed
-networks~\autocite{shahriari2014ranking}. Examples of features using such scores are \textsl{reputation}
-and \textsl{optimism}~\autocite{shahriari2014ranking}, defined for a node $u$ by
-\(
-	\frac{\sum_{v \in \NNin(u)} y_{v,u}\sigma(v)}{\sum_{v \in \NNin(u)} \sigma(v)}
-\quad\text{and}\quad
-	\frac{\sum_{v \in \NNout(u)} \yuv\sigma(v)}{\sum_{v \in \NNout(u)} \sigma(v)}\,,
-\)
-where $\sigma(v)$ is the ranking score assigned to node $v$. Some of these algorithms will be used
-as representative competitors in our experimental study of \autoref{s:exp}.

### 2017-09-05 troll/experiments.tex
move all troll related work in one file

@@
 \iffalse
-***************************************************
 %Given a training set \trainset{}, we exploit the revealed signs in two ways.
@@
 edge is associated with two features: $[1-\htr(i), 1-\hun(j)]$ (we call this method \uslogregp{}).
-
-We compare these methods with four competitors, each exploiting the sign information and network
-topology in a different way:
-
-Looking at undirected graphs with a global perspective, \textcite{LowRankCompletion14} consider the
-observed adjacency matrix $A$, made of the edges in \trainset{}, as a noisy sampling of the
-adjacency matrix $A^\star$ of an underlying complete graph satisfying the weakly balance condition
-(that is with no triangle containing only one negative edge).
-%TODO the previous sentence is too long
-This condition implies the existence of a small number $k$ of node clusters with positive links
-within clusters and negative links across clusters, which in turn implies $\rank A^\star=k$.  By
-recovering a complete matrix $\tilde{A}$ which matches the non-zeros entries of $A$, it is possible
-to predict the sign of $\eij \notin \trainset$ as $\yhat_{i,j} = \sgn(\tilde{A}_{i,j})$. Although
-the exact version of this problem is NP-hard, authors assume that $k$ is an hyperparameter known
-beforehand\footnote{The method is showed to be robust to the choice of its value, thus we pick
-$k=7$.} and look for two matrices $W,H\in \mathbb{R}^{k\times|V|}$ which minimise a sigmoidal
-reconstruction loss with $A$, subject to a regularization term. We refer to this method as
-\emph{\complowrank{}}.
-
-The status theory heuristic posits that a positive link from $i$ to $j$ denotes that user $i$
-considers user $j$ as having a higher status or skill than himself~\autocite{Leskovec2010}. It
-implies that among the 16 possible triads of three nodes with two directed signed edges among them,
-some must be more represented than others. The \emph{\comptriads{}} method exploits this fact by
-counting for each edge in the training set how frequently it is involved in each of the 16 triad
-types. Based on these edge features and 7 degree features of each edge endpoints, a Logistic
-Regression model is trained and used to predict the sign of the test edges.
-
-Inspired by the PageRank algorithm, \textcite{wu2016troll} propose a recursive definition of
-trollness based on the opinion of one node's neighbors weighted by their own trollness, which allow
-to assign a trustworthiness $\pi(i)$ to each node $i$ through a set of non linear equations solved
-by an iterative method. This $\pi$ values are used to rank the nodes by computing their Reputation
-and Optimism scores, thus providing four features for each edge, which are in turn used to train a
-Logistic Regression model for the classification task. We refer to this method as
-\emph{\compranknodes{}}.  Computing $\pi$ requires to choose two hyperparameters of the model,
-$\beta$ and $\lambda_1$ and the authors suggest to do it by holding out some training data to
-perform cross validation but in practice we found little difference in performance in our
-implementation while it consumes a lot of time.
-
-The last competitors is also local. \textcite{Bayesian15} note that  a node can belong to one of the
-16 types based of whether is number of its positive (resp.  negative) outgoing (resp. incoming)
-edges is zero or not.  The number of unobserved incoming and outgoing edges of each node $i$ let us
-define its 16-dimensional vector $V_i$ containing the probability of transitioning to any other
-state.  Then for each edge \eij we compute the outer product of $V_i$ with $V_j$ and also include
-additional features such as triads count and degree information before training a Logistic
-Regression model. We refer to this method as \emph{\compbayesian{}}.
 \fi

### 2017-09-05 troll/introduction.tex
move all troll related work in one file

@@
 %In this paper, we consider the problem of learning link classifiers, and study this problem from different viewpoints. Our contributions are summarized as follows.
-
-\subsection{Related work}
-
-Interest in signed networks can be traced back to the psychological theory of structural
-balance~\autocites{Cartwright56}{HeiderBook58} with its weak
-version~\autocite{davis1967clustering}.  The advent of online signed social networks has enabled a
-more thorough and quantitative understanding of that phenomenon. Among the several approaches
-related to our work, some extend the spectral properties of a graph to the signed case in order to
-find good embeddings for classification~\autocites{Kunegis2009}{SignedEmbedding15}. However, the
-use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which
-makes those methods hardly scalable to large graphs. Another approach is based on mining ego
-networks with SVM. Although this method seems to deliver good results~\autocite{Papaoikonomou2014},
-the running time makes it often impractical for large real-world datasets. An alternative approach,
-based on local features only and proposed in~\autocite{Leskovec2010}, relies on the so-called
-status theory for directed graphs~\autocite{guha2004propagation}. Some works in active learning,
-using a more sophisticated bias based on the correlation clustering (CC)
-index~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}, provide strong theoretical guarantees.
-However, the bias used there is rather strong, since it assumes the existence of a $2$-clustering
-of the nodes with a small CC index.\Todo{don't put it that way since we use that bias later}
-
-Whereas our focus will be on \emph{binary} prediction, researchers have also considered a weighted
-version of the problem, where edges measure the amount of trust or distrust between two
-users~\autocites{guha2004propagation}{Tang2013}{Bachi2012}{Qian2014sn}. Other works have also
-considered versions of the problem where side information related to the network is available to
-the learning system. For instance, \autocite{EdgeSignsRating15} uses the product purchased on
-Epinion in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
-analysing the textual content of their post, and~\autocite{SNTransfer13} uses SVM to perform
-transfer learning from one network to another. While many of these approaches have interesting
-performances, they often require extra information which is not always available (or reliable) and,
-in addition, may face severe scaling issues.
-
-The survey~\autocite{Tang2015a} contains pointers to many papers on edge sign prediction for
-signed networks, especially in the Data Mining area. Additional references, more closely related to
-our work, will be mentioned at the end of \autoref{ss:passive}.

### 2017-09-05 troll/related.tex
move all troll related work in one file

\section{Related work}

Interest in signed networks can be traced back to the psychological theory of structural
balance~\autocites{Cartwright56}{HeiderBook58} with its weak
version~\autocite{davis1967clustering}.  The advent of online signed social networks has enabled a
more thorough and quantitative understanding of that phenomenon. Among the several approaches
related to our work, some extend the spectral properties of a graph to the signed case in order to
find good embeddings for classification~\autocites{Kunegis2009}{SignedEmbedding15}. However, the
use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which
makes those methods hardly scalable to large graphs. Another approach is based on mining ego
networks with SVM. Although this method seems to deliver good results~\autocite{Papaoikonomou2014},
the running time makes it often impractical for large real-world datasets. An alternative approach,
based on local features only and proposed in~\autocite{Leskovec2010}, relies on the so-called
status theory for directed graphs~\autocite{guha2004propagation}. Some works in active learning,
using a more sophisticated bias based on the correlation clustering (CC)
index~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}, provide strong theoretical guarantees.
However, the bias used there is rather strong, since it assumes the existence of a $2$-clustering
of the nodes with a small CC index.\Todo{don't put it that way since we use that bias later}

Whereas our focus will be on \emph{binary} prediction, researchers have also considered a weighted
version of the problem, where edges measure the amount of trust or distrust between two
users~\autocites{guha2004propagation}{Tang2013}{Bachi2012}{Qian2014sn}. Other works have also
considered versions of the problem where side information related to the network is available to
the learning system. For instance, \autocite{EdgeSignsRating15} uses the product purchased on
Epinion in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
analysing the textual content of their post, and~\autocite{SNTransfer13} uses SVM to perform
transfer learning from one network to another. While many of these approaches have interesting
performances, they often require extra information which is not always available (or reliable) and,
in addition, may face severe scaling issues.

The survey~\autocite{Tang2015a} contains pointers to many papers on edge sign prediction for
signed networks, especially in the Data Mining area.

The vast majority of existing edge sign prediction algorithms for
directed graphs are based on the computation of local features of the graph. These features are
evaluated on the subgraph induced by the training edges, and the resulting values are used to train
a supervised classification algorithm (e.g., logistic regression). The most basic set of local
features used to classify a given edge $(u,v)$ are defined by
$\din^+(v),\din^-(v),\dout^+(u),\dout^-(u)$ computed over the training set \trainset{}, and by the
embeddedness coefficient $\big|\Nout(u) \cap \Nin(v)\big|$. In turn, these can be used to define
more complicated features, such as
\(
	\frac{\din^+(v) + |E^+|\uin(v)}{\din(v) + \uin(v)}
\quad\text{and}\quad
	\frac{\dout^+(u) + p^+\uout(u)}{\dout(u) + \uout(u)}
\)
introduced in~\autocite{Bayesian15}, together with their negative counterparts, where $|E^+|$ is the
overall fraction of positive edges, and $\uin(v),\uout(u)$ are, respectively, the number of test
edges outgoing from $u$ and the number of test edges incoming to $v$.
Other types of features are derived from social status theory~(e.g., \autocite{Leskovec2010}), and
involve the so-called triads; namely, the triangles formed by $(u,v)$ together with $(u,w)$ and
$(w,v)$ for any $w \in \NNout(u) \cap \NNin(v)$. A third group of features is based on node ranking
scores. These scores are computed using a variery of methods, including
Prestige~\autocite{zolfaghar2010mining}, exponential ranking~\autocite{traag2010exponential},
PageTrust~\autocite{de2008pagetrust}, Bias and Deserve~\autocite{mishra2011finding},
TrollTrust~\autocite{wu2016troll}, and generalizations of PageRank and HITS to signed
networks~\autocite{shahriari2014ranking}. Examples of features using such scores are \textsl{reputation}
and \textsl{optimism}~\autocite{shahriari2014ranking}, defined for a node $u$ by
\(
	\frac{\sum_{v \in \NNin(u)} y_{v,u}\sigma(v)}{\sum_{v \in \NNin(u)} \sigma(v)}
\quad\text{and}\quad
	\frac{\sum_{v \in \NNout(u)} \yuv\sigma(v)}{\sum_{v \in \NNout(u)} \sigma(v)}\,,
\)
where $\sigma(v)$ is the ranking score assigned to node $v$. Some of these algorithms will be used
as representative competitors in our experimental study of \autoref{s:exp}.

\iffalse
We compare these methods with four competitors, each exploiting the sign information and network
topology in a different way:

Looking at undirected graphs with a global perspective, \textcite{LowRankCompletion14} consider the
observed adjacency matrix $A$, made of the edges in \trainset{}, as a noisy sampling of the
adjacency matrix $A^\star$ of an underlying complete graph satisfying the weakly balance condition
(that is with no triangle containing only one negative edge).
%TODO the previous sentence is too long
This condition implies the existence of a small number $k$ of node clusters with positive links
within clusters and negative links across clusters, which in turn implies $\rank A^\star=k$.  By
recovering a complete matrix $\tilde{A}$ which matches the non-zeros entries of $A$, it is possible
to predict the sign of $\eij \notin \trainset$ as $\yhat_{i,j} = \sgn(\tilde{A}_{i,j})$. Although
the exact version of this problem is NP-hard, authors assume that $k$ is an hyperparameter known
beforehand\footnote{The method is showed to be robust to the choice of its value, thus we pick
$k=7$.} and look for two matrices $W,H\in \mathbb{R}^{k\times|V|}$ which minimise a sigmoidal
reconstruction loss with $A$, subject to a regularization term. We refer to this method as
\emph{\complowrank{}}.

The status theory heuristic posits that a positive link from $i$ to $j$ denotes that user $i$
considers user $j$ as having a higher status or skill than himself~\autocite{Leskovec2010}. It
implies that among the 16 possible triads of three nodes with two directed signed edges among them,
some must be more represented than others. The \emph{\comptriads{}} method exploits this fact by
counting for each edge in the training set how frequently it is involved in each of the 16 triad
types. Based on these edge features and 7 degree features of each edge endpoints, a Logistic
Regression model is trained and used to predict the sign of the test edges.

Inspired by the PageRank algorithm, \textcite{wu2016troll} propose a recursive definition of
trollness based on the opinion of one node's neighbors weighted by their own trollness, which allow
to assign a trustworthiness $\pi(i)$ to each node $i$ through a set of non linear equations solved
by an iterative method. This $\pi$ values are used to rank the nodes by computing their Reputation
and Optimism scores, thus providing four features for each edge, which are in turn used to train a
Logistic Regression model for the classification task. We refer to this method as
\emph{\compranknodes{}}.  Computing $\pi$ requires to choose two hyperparameters of the model,
$\beta$ and $\lambda_1$ and the authors suggest to do it by holding out some training data to
perform cross validation but in practice we found little difference in performance in our
implementation while it consumes a lot of time.

The last competitors is also local. \textcite{Bayesian15} note that  a node can belong to one of the
16 types based of whether is number of its positive (resp.  negative) outgoing (resp. incoming)
edges is zero or not.  The number of unobserved incoming and outgoing edges of each node $i$ let us
define its 16-dimensional vector $V_i$ containing the probability of transitioning to any other
state.  Then for each edge \eij we compute the outer product of $V_i$ with $V_j$ and also include
additional features such as triads count and degree information before training a Logistic
Regression model. We refer to this method as \emph{\compbayesian{}}.
\fi

### 2017-09-10 troll/introduction.tex
add some applications in the troll intro

@@
-\section{Introduction}
-% say that we are facing edge sign prediction from many viewpoints.
-% say about trolls and trustworthy people in a social network
-% say that we are using reductions to node classification (novelty)
-% generative model
-% novel and cool
+As we saw in the introduction, many situations in various fields can be modeled by signed graphs.
+One of the most natural usage of signed graphs, though, is the study of social networks. That is
+indeed what first motivated their inception. Quite naturally, we thus begin our exploration of the
+role of edges in complex networks by a binary classification problem in signed graphs. Namely, we
+want to predict whether an observed interaction between two nodes is positive or negative and we
+call this problem \esp{}.\footnote{This chapter is closely based on a paper written with several
+co-authors~\autocite{trollSign17}.} To motivate this problem from a practical point of view, let us
+first define precisely the type of social networks we consider. We call \emph{Directed Signed Social
+Network} (DSSN) a directed graph whose nodes are either human beings (often called users if they are
+the members of an online community) or artifacts directly created by individuals\footnote{This
+semantic detail allows us to treat citation networks as social networks.}. Furthermore, a directed
+edge from node $u$ to node $v$ denote an interaction initiated by the user $u$ and whose object is
+user $v$. Such an interaction can be positive (to praise, to support, to trust, to befriend) or
+negative (to criticize, to oppose, to distrust, to befoe).
 
-Connections in social networks are mostly driven by the \emph{homophily} assumption: linked
-individuals tend to be similar, sharing personality traits, attitudes, or interests. However,
-homophily alone is clearly not sufficient to explain the variety of social links\Todo{link with
-chapter 3}. In fact, sociologists have long studied networks, hereafter called \emph{signed} social
-networks\Todo{DSSN?}, where also {\em negative} relationships ---like dissimilarity, disapproval or
-distrust--- are explicitly displayed. The presence of negative relationships is also a feature of
-many technology-mediated social networks. Known examples are \textsc{Ebay}, where users trust or
-distrust agents in the network based on their personal interactions, \textsc{Slashdot}, where each
-user can tag another user as friend or foe, and \textsc{Epinion}, where users can rate positively
-or negatively not only products, but also other users. Even in social networks where connections
-solely represent friendships, negative links can still emerge from the analysis of online debates
-among users.
+Known examples are \textsc{Ebay}, where users trust or distrust agents in the network based on their
+personal interactions, \textsc{Slashdot}, where each user can tag another user as friend or foe,
+\textsc{BlaBlaCar}, a carpooling website where users can evaluate drivers and passengers as pleasant
+travelers or not, and \textsc{Epinion}, where users can rate positively or negatively not only
+products, but also other users. In such examples, the sign of the interaction is explicit. However,
+there are other cases where such interactions are only displayed in their raw form and need further
+processing to be given a sign. The typical situation where this happens is when interactions are
+mediated through text, for instance in Twitter or in the comment section of some online content. It
+is then possible to extract signs using sentiment analysis methods~\autocite{signedText12}.
+
+Another distinction is that in most cases, only the owner of the network can observe its graph.
+Think for instance of \textsc{YouTube}, where users can like or dislike the video of another user
+but where only aggregate anonymous statistics are available publicly. The situation is similar on
+the \textsc{Stack Overflow} community, where users can upvote or downvote answers according to their
+perceived quality, but where only the total score are available.
+
+\bigskip
+
+With this definition of social networks, we now look at how the learning problem of \esp{} can be
+used in real world applications. More precisely, we present four possible applications, and order
+them by how severely negative interactions impact the well-being of users of the networks.
+First, it could help improve the quality of link recommendation in solely positive networks. Such
+networks indeed contain implicit negative links~\autocite{Yang2012}. By asking users to label a
+small proportion of existing links as truly positive or actually negative, recommended links could
+be discarded if they are predicted to be negative.
+%
+Second, this could be used to monitor news textual comments in a scalable fashion and thus ensure
+that online debates remain courteous and constructive. For instance, \textcite{betterDebate14} show
+that by visually reminding comment authors of the importance of respecting the plurality of
+opinions, the quality of debates is improved (other references can be found
+in~\autocite{journalism15}).
+%
+Third, like many other fields, human resource management is undergoing a transformation through the
+increasing use of Machine Learning~\autocite{MLinHR16}. Among other tools, assessing personal
+relationships between employees is crucial to maintain a productive work environment. Whereas this
+can be done by invasive wearable sensors~\autocite{WearableBehavior09}, our methods could predict
+which employees would form the tighter teams from a small amount of labeled data.
+%
+Fourth, and most critical, predicting the sign of interactions could be a tool to detect users with
+a high proportion of negative interactions. Indeed, because people tend to be less inhibited in
+their online interactions~\cite{Suler04}, some users may join an online community with the main goal
+to disrupt it, by engaging into anti-social behavior and creating conflictual relationships with
+other members. This kind of attitude expressed publicly on social media leads to the following
+definition of \emph{trolls}: \enquote{users whose real intentions are to cause disruption and/or to
+trigger or exacerbate conflict for the purposes of their own amusement}~\autocite{Hardaker10}.
+\Textcite{Shachaf10} elaborate on their motives, adding that boredom, attention seeking, and
+revenge motivate trolls; they find pleasure from causing damage to other people or to the community
+as a whole. An extreme form of trolling is bullying (\ie{} the behavior of someone intentionally and
+repeatedly harming a victim that is unable to defend himself or herself) and just like anything
+else, it has an online version called cyber bullying~\autocite{cyberbullying13}. Such behavior is
+rather widespread, one study revealing that 72\% of \np{1454} surveyed teenagers reported at least
+one occurrence of cyberbullying, most of them through instant messaging~\autocite{Juvonen08}. While
+we have not conducted experimental study on this topic, we believe that being able to predict such
+harmful interactions would be beneficial for the majority of users.
 

### 2017-09-10 troll/introduction.tex
add pointers to other sections in troll intro

@@
 
-When the social network is signed, specific challenges arise in both network analysis and learning.
-On the one hand, novel methods are required to tackle standard tasks (e.g., user clustering, link
-prediction, targeted advertising/recommendation, prediction of user interests, and analysis of the
-spreading of diseases in epidemiological models). On the other hand, new problems such as edge sign
-prediction, which we consider here, naturally emerge. Edge sign prediction is the problem of
-classifying the positive or negative nature of the links based on the network topology. Prior
-knowledge of the network topology is often a realistic assumption, for in several situations the
-discovery of the link sign can be more costly than acquiring the topological information of the
-network. For instance, when two users of an online social network communicate on a public web page,
-we immediately detect a link. Yet, the classification of the link sign as positive or negative may
-require complex techniques.
+% When the social network is signed, specific challenges arise in both network analysis and
+% learning. On the one hand, novel methods are required to tackle standard tasks (e.g., user
+% clustering, link prediction, targeted advertising/recommendation, prediction of user interests,
+% and analysis of the spreading of diseases in epidemiological models). On the other hand, new
+% problems such as edge sign prediction, which we consider here, naturally emerge.
+
+\bigskip
+
+Such applications motivate studying \esp{}, which is the problem of classifying the positive or
+negative nature of the links based on the
+network topology. Prior knowledge of the network topology is often a realistic assumption, for in
+several situations the discovery of the link sign can be more costly than acquiring the topological
+information of the network. For instance, when two users of an online social network communicate on
+a public web page, we immediately detect a link. Yet, the classification of the link sign as
+positive or negative may require more involved techniques.
 
 From the modeling and algorithmic viewpoints, because of the huge amount of available networked
-data, a major concern in developing learning methods for edge sign prediction is algorithmic
-scalability. Many successful, yet simple heuristics for edge sign prediction are based on the
-troll-trust features, i.e., on the fraction of outgoing negative links (trollness) and incoming
-positive links (trustworthiness) at each node. We study such heuristics by defining a probabilistic
-generative model for the signs on the directed links of a given network, and show that these
+data, a major concern in developing learning methods for \esp{} is algorithmic
+scalability. Many successful, yet simple heuristics for \esp{} are based on the
+troll-trust features, \ie{} on the fraction of outgoing negative links (trollness) and incoming
+positive links (trustworthiness) at each node. We first define these notions and more notations in
+\autoref{s:prel}, along with others tools and precise problem statements. Then we study such
+troll-trust heuristics by defining in \autoref{s:gen} a probabilistic
+generative model for the signs on the directed links of a given network. We also show that these
 heuristics can be understood and analyzed as approximators to the Bayes optimal classifier for our
-generative model. We also gather empirical evidence supporting our probabilistic model by observing
+generative model. In this context, we design our first batch algorithm and show in
+\autoref{s:algbatch} that it provably approximates the Bayes classifier on dense graphs.
+We furthermore gather empirical evidence supporting our probabilistic model by observing
 that a logistic model trained on trollness and trustworthiness features alone is able to learn
@@
 
-We then introduce suitable graph transformations defining reductions from edge sign prediction to
+We then introduce suitable graph transformations defining reductions from \esp{} to
 node sign prediction problems. This opens up the possibility of using the arsenal of known
-algorithmic techniques developed for node classification. In particular, we show that a Label
-Propagation algorithm, combined with our reduction, approximates the maximum likelihood estimator
-of our probabilistic generative model. Experiments on real-world data show the competitiveness of
-our approach in terms of both prediction performance (especially in the regime when training data
-are scarce) and scalability.
+algorithmic techniques developed for node classification. In particular, we introduce in
+\autoref{ss:passive} our second batch algorithm. It takes the form of a Label
+Propagation algorithm that, combined with our problem reduction, approximates the maximum likelihood
+estimator of our probabilistic generative model.
+In order to compare our two algorithms with existing work, we then describe with more details
+previous heuristics and related methods in \autoref{sec:troll_related}.
+We then experimentally evaluate our proposed approach in \autoref{s:exp}. On synthetic data, we
+confirm the quality of our approximation of the Bayes predictor. More importantly, on real-world
+data, we show the competitiveness of our approach in terms of both prediction performance
+(especially in the regime when training data are scarce) and scalability.
 
-Finally, we point out that the notions of trollness and trustworthiness naturally define a measure
+Finally, in \autoref{s:algonline}, we point out that the notions of trollness and trustworthiness
+naturally define a measure
 of complexity, or learning bias, for the signed network that can also be used to design
-\emph{online} (i.e., sequential) learning algorithms for the edge sign prediction problem. The
+\emph{online} (\ie{} sequential) learning algorithms for the \esp{} problem. The
 learning bias encourages settings where the nodes in the network have polarized features (e.g.,
@@
 network links.
-
-%Hence, it is necessary to depart from well-established yet computationally expensive approaches by relying on novel algorithmic techniques.
-
-%In this paper, we consider the problem of learning link classifiers, and study this problem from different viewpoints. Our contributions are summarized as follows.

### 2017-09-10 assets/tikz/troll_notations_tikz.tex
augment the troll notations figure

@@
       \tikzstyle{edge}=[line width=2pt,color=Black,-{Latex[length=3mm,width=1.0mm,angle'=40]}]
-      \tikzstyle{trainEdge}=[line width=10pt,color=Black!15,opacity=.7]
+      \tikzstyle{trainEdge}=[line width=10pt,color=Black!15,opacity=0]
       \tikzstyle{elabel}=[color=Black]
+      \tikzstyle{holder}=[draw,circle,black,inner sep=1pt, minimum width=3mm]
       \tikzstyle{posc}=[scale=1.2,text=Green]
@@
       \tikzstyle{testset}=[opacity=1.0]
-      \tikzstyle{revealed}=[circle,draw=none,fill=Black!15,inner sep=1pt,]
+      \tikzstyle{revealed}=[circle,draw=none,fill=none,inner sep=1pt,]
 
+      \draw[DarkOrange,line width=1.1pt] (5.2, -0.3) ellipse [x radius=1,y radius=2.8] node[yshift=3.05cm] {$\NNout(u)$};
+      \draw[DarkOrange,line width=1.1pt] (2.9, 0) ellipse [x radius=1.1,y radius=1.7] node[yshift=2.25cm] {$\Nout(u)$};
+      \draw[Blue,line width=1.1pt] (-5, 0) ellipse [x radius=0.7,y radius=2.5] node[yshift=2.8cm] {$\NNin(u)$};
+      \draw[Blue,line width=1.1pt] (-2.9, 0.2) ellipse [x radius=1,y radius=1.9] node[yshift=2.25cm] {$\Nin(u)$};
       \node[vertex] (u) at (0,0) {\LARGE $u$};
@@
 
-      \draw[trainEdge] (u) --                                                       (5, -2.5);
-      \draw[edge] (u) -- node [revealed,posc,xshift=-2mm,] {$\mathbf{+}$}           (5, -2.5);
+      \node[holder] (fromu1) at (5, -2.5) {};
+      % \draw[trainEdge] (u) --                                                       (5, -2.5);
+      \draw[edge] (u) -- node [revealed,posc,xshift=-2mm,] {$\mathbf{+}$}           (fromu1);
 
@@
         {${\textcolor{Black}{y_{u,v}=}}{\textcolor{Green}{+}}$} (v);
-      \draw[testset,edge] (u) edge  node [posc,xshift=2mm,pos=0.7] {$\mathbf{+}$}            (5, 0.66);
-      \draw[testset,edge, dashed] (u)   edge   node [negc,xshift=2mm] {$\mathbf{-}$} (5, 2);
+      \node[holder] (fromu2) at (5, .66) {};
+      \node[holder] (fromu3) at (5, 2) {};
+      \draw[testset,edge] (u) edge  node [posc,xshift=2mm,pos=0.7] {$\mathbf{+}$}            (fromu2);
+      \draw[testset,edge, dashed] (u)   edge   node [negc,xshift=2mm] {$\mathbf{-}$} (fromu3);
 
+      \node[holder] (tou1) at (-5, -2) {};
+      \node[holder] (tou2) at (-5, -1) {};
+      \node[holder] (tou3) at (-5, 0) {};
+      \node[holder] (tou4) at (-5, 1) {};
+      \node[holder] (tou5) at (-5, 2) {};
       \begin{scope}[on background layer]
-        \draw[trainEdge] (-5, -2) --    (u.south west);
-        \draw[edge, dashed] (-5, -2)       -- node[revealed,negc,yshift=-3mm,xshift=-5pt]
-          {$\mathbf{-}$} (u.south west);
-        \draw[trainEdge,]          (-5, 2) -- (u.north west);
-        \draw[edge,]               (-5, 2) -- node[revealed,posc,above,pos=0.2,xshift=5pt] {$\mathbf{+}$} (u.north west);
+        % \draw[trainEdge] (-5, -2) --    (u.south west);
+        \draw[edge, dashed] (tou1)       -- node[revealed,negc,yshift=-3mm,xshift=-5pt] {$\mathbf{-}$} (u.south west);
+        % \draw[trainEdge,]          (-5, 2) -- (u.north west);
+        \draw[edge,]               (tou5) -- node[revealed,posc,above,pos=0.35,xshift=5pt] {$\mathbf{+}$} (u.north west);
       \end{scope}
 
-      \draw[testset,edge,dashed] (-5, -1)-- node[negc,yshift=-1mm] {$\mathbf{-}$} (u);
-      \draw[testset,edge,dashed] (-5, 0) -- node[negc] {$\mathbf{-}$} (u);
-      \draw[testset,edge,]       (-5, 1) -- node[posc,xshift=-4mm] {$\mathbf{+}$} (u);
+      \draw[testset,edge,dashed] (tou2)-- node[negc,yshift=-1mm] {$\mathbf{-}$} (u);
+      \draw[testset,edge,dashed] (tou3) -- node[negc] {$\mathbf{-}$} (u);
+      \draw[testset,edge,]       (tou4) -- node[posc,xshift=-4mm] {$\mathbf{+}$} (u);
 

### 2017-09-11 troll/notation_and_prel.tex
a pass on the rest of the troll main text
move and add figures in the troll notations part
small modifs in troll notations and cutsize explanation

@@
 
-In what follows, we let $G=(V,E)$ be a \emph{directed} graph, whose edges $(u,v) \in E$ carry a
+In this chapter and the next one, we let $G=(V,E)$ be a \emph{directed} graph, whose edges $(u,v)
+\in E$ carry a
 binary label $\yuv\in \spin$. The edge labeling will sometimes be collectively denoted by the
-$|V|\times |V|$ matrix $Y = [\yuv]$, where $\yuv = \yuv$ if $(u,v) \in E$, and $\yuv = 0$,
+$|V|\times |V|$ matrix $Y = [\yuv]$, where $Y_{u,v} = \yuv$ if $(u,v) \in E$, and $\yuv = 0$
 otherwise. The corresponding edge-labeled graph will be denoted by $G(Y) = (V,E(Y))$. We use
@@
 $\NNin(u)$ and $\NNout(u)$ to represent, respectively, the in-neighborhood and the out-neighborhood
-of node $u \in V$.
+of node $u \in V$. Most of these notations are illustrated on \autoref{fig:troll_Gexample}.
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.9\linewidth]{tikz/troll_notations_tikz}
+  \caption{Part of a DSSN centered on node $u$. \label{fig:troll_Gexample}}
+\end{figure}
 
@@
 original graph $G$ into other graphs. As we see later, these reductions are useful in turning the
-edge sign prediction problem into a \emph{node} sign prediction problem (often called node
+\esp{} problem into a \emph{node} sign prediction problem (often called node
 classification problem), for which many algorithms are indeed available~%
@@
 labeled, then this labeling transfers to the subset of nodes $e_{u,v} \in V''$. Graph $G'$, which
-will not be used in this paper, is an intermediate structure between $G$ and $G''$ and provides a
+will not be used in the following, is an intermediate structure between $G$ and $G''$ and provides a
 conceptual link to the standard cutsize measure in node sign classification. \autoref{f:etnr}
@@
   \begin{subfigure}[b]{0.28\textwidth}
-    \centering \includegraphics[height=.10\textheight]{g_latex-crop} \caption{}
+    \centering \includegraphics[height=.10\textheight]{raw/g_latex-crop} \caption{}
   \end{subfigure}~
   \begin{subfigure}[b]{0.30\textwidth}
-    \centering \includegraphics[height=.12\textheight]{gprime_latex-crop}
+    \centering \includegraphics[height=.12\textheight]{raw/gprime_latex-crop}
     \caption{\label{fig:troll_reduction_gprime}}
@@
   \begin{subfigure}[b]{0.36\textwidth}
-    \centering \includegraphics[height=.14\textheight]{gsecond_latex-crop} \caption{}
+    \centering \includegraphics[height=.14\textheight]{raw/gsecond_latex-crop}
+    \caption{\label{fig:troll_reduction_gsecond}}
   \end{subfigure}
@@
 
-These reductions are meaningful only if they are able to approximately preserve label
-\emph{regularity} when moving from edges to nodes. That is, if the edge sign prediction problem is
-easy for a given $G(Y) = (V,E(Y))$, then the corresponding node sign prediction problems on $G'(Y)
-= (V'(Y),E')$ and $G''(Y) = (V''(Y),E)$ are also easy, and vice versa.  While we could make this
+\iffalse
+  While we could make this
 argument more quantitative, here we simply observe that if each node in $G$ tends to be either
@@
 corresponding to the edges in $G$ (the colored squares in \autoref{fig:troll_reduction_gprime}).
+\fi
 
-Again, we would like to point out that these reductions serve two purposes: First, they allow us to
+These reductions serve two purposes: First, they allow us to
 use the many algorithms designed for the better studied problem of node sign prediction. Second,
@@
 
-We consider two learning settings associated with the problem of edge sign prediction: a batch
-setting and an online setting. In the batch \emph{passive} setting, we assume that a training set
+\bigskip
+
+We consider two learning settings associated with the problem of \esp{}: a batch
+setting and an online setting. In the batch setting, we assume that a training set
 of edges $E_0$ has been drawn uniformly at random \emph{without replacement} from $E$, we observe
 the labels in $E_0$, and we are interested in predicting the sign of the remaining edges $E
-\setminus E_0$ by making as few prediction mistakes as possible. In the batch \emph{active}
-setting, we are given a budget of edge labels to observe, and are free to select them the way we
-want within $E(Y)$. Again, the goal is to make as few mistakes as possible on the remaining edges.
-The specific batch setting (passive and active) we study here assumes that labels are produced by a
+\setminus E_0$ by making as few prediction mistakes as possible.
+% In the batch \emph{active}
+% setting, we are given a budget of edge labels to observe, and are free to select them the way we
+% want within $E(Y)$. Again, the goal is to make as few mistakes as possible on the remaining edges.
+The specific batch setting
+% (passive and active)
+we study here assumes that labels are produced by a
 generative model which we describe in the next section, and our label regularity measure is a
@@
 online learning~\autocite{Winnow88} where all edge labels are assumed to be generated by an
-adversary and sequentially presented to the learner according to an arbitrary permutation. For an
+adversary and sequentially presented to the learner according to an arbitrary permutation.
+\iffalse
+For an
 online learning algorithm $A$, we are interested in measuring the total number of mistakes $M_A(Y)$
@@
 $Y$ is selected by the adversary.
-
-\iffalse
-  In this setting, learning is split into a sequence of rounds: At each round $t=1,2,\dots$ a
+\fi
+More precisely,
+  in this setting, learning is split into a sequence of rounds: At each round $t=1,2,\ldots$ a
   learning algorithm $A$ outputs a prediction ${\wh y_t} \in \spin$ for the label $y_t \in \spin$
   of an edge arbitrarily selected by the adversary. After the prediction is made, label $y_t$ is
-  revealed to the algorithm, hence allowing it to change its internal state. In the next round a
+  revealed to the algorithm, hence allowing it to change its internal state. In the next round, a
   new edge is selected by the adversary, and so on. Notice that since the underlying labeling over
@@
   - \Psi_G(Y)~.\)
-\fi
 
-Also in the online setting our label regularity measure, denoted here by $\Psi_G(Y)$, is small when
-nodes in $G$ tend to be either troll or trustworthy. When this happens, few labels from the
+We want to design our label regularity measure such that it is small when
+nodes in $G$ tend to be either troll or trustworthy. Indeed, when this happens, few labels from the
 incoming and outgoing edges of each node are sufficient to predict the labels on the remaining
-edges. Formally, for fixed $G$ and $Y$, let 
+edges and the \esp{} problem is to some extent \enquote{easy}. Formally, for fixed $G$ and $Y$, let 
 $$
@@
 $$
+be respectively the number of the least used label in the incoming edges to $v$ and the outgoing
+edges from $u$.
 Let also
@@
 $$
+be the sum of irregularity over all the nodes of the graph.
 Then we define $\Psi_G(Y) = \min\big\{\Psiin(Y),\Psiout(Y)\big\}$. The two measures
 $\Psi^2_{G''}(Y)$ and $\Psi_G(Y)$ are conceptually related and their value on real data is quite
-similar (see \autoref{tab:all_mcc} in \autoref{s:exp}).
+similar (see \autoref{tab:all_mcc} \vpageref{s:exp}).
 
-\iffalse
-  Recall that the cutsize of an undirected node-labeled graph $G'(Y)$ is the number of edges in
-  $G'$ connecting nodes having mismatching labels. Now, because the only nodes in $G'$ we are
+The reductions presented above are meaningful only if they are able to approximately preserve the
+two label regularity measures $\Psi_G(Y)$ and $\Psi^2_{G''}(Y)$ when moving from edges to nodes.
+That is, if the \esp{} prediction problem is
+easy for a given $G(Y) = (V,E(Y))$, then the corresponding node sign prediction problems on $G'(Y)
+= (V'(Y),E')$ and $G''(Y) = (V''(Y),E)$ are also easy, and vice versa.
+Let us first introduce the notion of cutsize of an undirected node-labeled graph $G'(Y)$, which is
+the number of edges in $G'$ connecting nodes having mismatching labels.
+  Now, because the only nodes in $G'$ we are
   interested in predicting are those corresponding to the edges in $G$ (the colored squares in
-  Figure \ref{f:etnr} (b)), the online prediction problem on the edges of $G$ translates to a node
-  sign prediction problem on a subset of $V'$. The remaining nodes in $V'$ (the circles in Figure
-  \ref{f:etnr} (b)) we are free to assign dummy labels so as to minimize the corresponding mistake
+  \autoref{fig:troll_reduction_gprime}), the online prediction problem on the edges of $G$ translates to a node
+  sign prediction problem on a subset of $V'$. As for the remaining nodes in $V'$ (the circles in
+  \autoref{fig:troll_reduction_gprime}), we are free to assign dummy labels so as to minimize the corresponding mistake
   bound over $G'$. It is straightforward to see that if we set the labeling $Y$ on $V'$ in such a
   way that $y_{\iout} = +1$ if $tr(u) < 0$, and $-1$, otherwise, and $y_{\iin} = +1$ if $un(u) <
-  0$, and $-1$, otherwise, then the cutsize of $G'(Y)$ equals\footnote{In fact, for the sake of
+  0$, and $-1$, otherwise, then the cutsize of $G'(Y)$ equals  $\Psiin(Y) + \Psiout(Y)$.%
+  \footnote{In fact, for the sake of
   this specific argument, nothing prevents from retaining of $G'$ either only the edges
   $(\iout,e_{u,v})$ or only the edges $(e_{u,v},\jin)$, resulting in a cutsize of $\Psiout(Y)$ and
-  $\Psiin(Y)$, respectively.} $\Psiin(Y) + \Psiout(Y)$.
-\fi
+  $\Psiin(Y)$, respectively.}
+  A similar reasoning apply in the batch case with the $\Psi^2_{G''}(Y)$ measure.

### 2017-09-11 troll/generative.tex
minor

@@
 $$\Pr\big( \yuv = 1 \big) = \tfrac{p_u + q_v}{2}$$
-This can be seen pictorially in \autoref{fig:troll_genmodel}.
+This process is illustrated in \autoref{fig:troll_genmodel}.
 \begin{marginfigure}
@@
 node $v$ tends to be liked by other people ($q_v$). In a certain sense, $1-tr(u)$ is the empirical
-counterpart to $p_u$, and $1-un(v)$ is the empirical counterpart to $q_v$.\footnote{One might view
+counterpart to $p_u$, and $1-un(v)$ is the empirical counterpart to $q_v$. One might view
 our model as reminiscent of standard models for link generation in social network analysis, like the
-classical $p_1$ model from \cite{hl81}. Yet, the similarity falls short, for all these models aim at
+classical $p_1$ model from \cite{hl81}. However, note that the similarity falls short, for all these
+models aim at
 representing the likelihood of the network topology, rather than the probability of edge signs, once

### 2017-09-11 troll/algs_batch.tex
small clarification to the first batch algo

@@
 
-Given $G(Y) =(V,E(Y))$, we have at our disposal a training set \trainset{} of labeled edges from
-$E(Y)$, our goal being that of building a predictive model for the labels of the remaining edges.
+Given $G(Y) =(V,E(Y))$, in the batch setting we have at our disposal a training set \trainset{} of
+labeled edges from $E(Y)$. We want to build a predictive model for the labels of the remaining
+edges. We present two algorithms to do so, which both compute estimates of the all parameters $p$
+and $q$ of our generative model. They differ in the approximation guarantees they provide, and in
+the class of graphs to which they apply. The first algorithm runs on the graph $G$ and estimates
+locally the parameters by their empirical means in the training set, which under some density
+assumptions are showed to concentrate around their true values. The second algorithm, on the other
+hand, exploits the reduced graph $G''$ and computes a maximum likelihood estimation of the
+parameters through a global label propagation approach, without making any density assumption.
+
+\subsection{Approximation to Bayes via dense sampling}\label{ss:bayes_approx}
 
@@
   \label{eq:predictor}
-  \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) - \tfrac{1}{2} -\tau\Big)~,
+  \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) -\tau - \tfrac{1}{2}\Big)~,
 \end{equation}
@@
 \iffalse
-\subsection{Approximation to Bayes via Active Learning}\label{ss:active}
 
@@
 Our algorithm works under the assumption that for given parameters $Q$ (a positive integer) and
-$\alpha \in (0,1)$ there exists a set\footnote{$E_L$ is needed to find an estimate $\tauhat$ of
-$\tau$ in~(\ref{eq:predictor}) ---see Step~3 of the algorithm. Any undirected matching of $G$ of
-size $\mathcal{O}(\log|V|)$ can be used, for instance using the blossom
-algorithm~\autocite{matching65}. In practice, however, we never computed $E_L$, and estimated $\tau$
+$\alpha \in (0,1)$, there exists a set\footnote{$E_L$ is needed to find an estimate $\tauhat$ of
+$\tau$ in~(\ref{eq:predictor}), as done in Step~3 of the algorithm. Any undirected matching of $G$ of
+size $O(\log|V|)$ can be used, obtained for instance by the blossom algorithm~\autocite{matching65}.
+In practice, however, we never computed $E_L$, and estimated $\tau$
 on the entire training set \trainset{}.} $E_L \ss E$ of size $\tfrac{2Q}{\alpha}$ where each vertex
-$u \in V$ appearing as an endpoint of some edge in $E_L$ occurs at most once as origin ---ie{}.,
-$(u,v)$--- and at most once as destination ---ie{}., $(v,u)$. Moreover, we assume \trainset{} has
+$u \in V$ appearing as an endpoint of some edge in $E_L$ occurs at most once as origin ---\ie{}
+$(u,v)$--- and at most once as destination ---\ie{} $(v,u)$. Moreover, we assume \trainset{} has
 been drawn from $E$ at random {\em without} replacement, with $m = |E_0| = \alpha\,|E|$. The
@@
 		$$
-		\yhat(u,v) = \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) - \tfrac{1}{2} -\tauhat\Big)
+		\yhat(u,v) = \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) -\tauhat - \tfrac{1}{2}\Big)
 		$$
@@
 
-The next result\footnote{All proofs are in the supplementary material.} shows that if the graph is
-not too sparse, then the above algorithm can approximate the Bayes optimal predictor on nodes whose
-in-degree and out-degree is not too small.
+The next result shows that the above algorithm can approximate the Bayes optimal predictor on nodes
+whose in-degree and out-degree is not too small.
 
@@
   model in \autoref{s:gen}.
-  If the algorithm is run with parameter $Q = \Omega(\ln|V|)$, and $\alpha \in (0,1)$ such that the
+  If the algorithm is run with parameter $Q = \Omega(\log|V|)$, and $\alpha \in (0,1)$ such that the
   above assumptions are satisfied, then $\yhat(u,v) = y^*(u,v)$ holds with high probability
-  simultaneously for all test edges $(u,v) \in E$ such that $\dout(u),\din(v) = \Omega(\ln|V|)$, and
+  simultaneously for all test edges $(u,v) \in E$ such that $\dout(u),\din(v) = \Omega(\log|V|)$, and
   $\eta(u,v) = \Pr(\yuv=1)$ is bounded away from $\tfrac{1}{2}$.
-  The overall number of training edges is $\Theta\big(|V|\ln|V|\big)$.
 \end{theorem}
 
-The approach leading to Theorem~\ref{t:active} lets us derive the \usrule{} algorithm assessed in
-our experiments of \autoref{s:exp}, but it needs the graph to be sufficiently dense and the bias
-$\tau$ to be the same for all edges. In order to address these limitations, we now introduce a
-second method based on label propagation.
+While we defer the full proof of \autoref{t:active} to the additional material at the end of this
+chapter (\autoref{ssec:troll_proof_bayes}), here we give a sketch of the method used, focusing on
+parameter $p$. First, by \autoref{l:bins} on negatively associated random variables (stated
+\vpageref{l:bins}), we show that with our choice of $Q$, for $\theta=\frac{2Q}{\alpha}$ and for any
+$u$ having enough out neighbors (\ie{} $\NNout(u) \ge \theta$), at least $Q$ edges of $\Nout(u)$
+are in the training set with high probability.
+We use these $Q$ edges to compute $1-\htr(u)$, the empirical probability of drawing a $+1$-labeled
+edge from $\Nout(i)$. We show that it is close to the real probability, which according to equation
+\eqref{e:pout} is equal to $\frac{1}{2}\,\left(p_u + \qbar_u\right)$. $\qbar_u$ being a sample mean
+of i.i.d.~$[0,1]$-valued random variables independently drawn from the prior marginal $\int_0^1
+\mu\big(p,\cdot\big) dp$, it concentrates around its expectation $\mu_q$.
+
+By following the same reasoning on the parameters $q$ and using an union bound, we then show that
+$\left(1-\htr(u)\right) + \left(1-\hun(v)\right) - \frac{\mu_p+\mu_q}{2}$ is close to
+$\frac{p_u+q_v}{2}$ simultaneously for all edges. Another set of concentration arguments then show
+that our estimate $\tauhat$ is close to $\frac{\mu_p+\mu_q}{2}$. Combining these two facts ends the
+proof.
+
+\medskip
+
+The approach leading to \autoref{t:active} needs the graph to be sufficiently dense. At first sight,
+having $Q = \Omega(\log|V|)$ training edges per nodes does not sound that like such a stringent
+requirement. Consider for instance Facebook ---which is neither signed nor directed though. It has
+two billion
+users~\bottomfootnote{\url{https://investor.fb.com/investor-news/press-release-details/2017/Facebook-Reports-Second-Quarter-2017-Results/}},
+each of them having $155\approx 7.2\log|V|$ friends on average~\autocite{facebookFriend16}. However,
+the constant hidden in the $\Omega$ notation, and needed to guarantee a good behavior, is much
+larger than $7.2$. We will nonetheless see in the experiments that we can still apply the resulting
+algorithm (we call it \usrule{}) with satisfying results. Additionally, and in order to address this
+density limitation, we now introduce a second method based on label propagation.
 
@@
 Binary $\pm 1$ predictions on the test set $E\setminus \trainset{}$ are then obtained by
-thresholding the obtained values $\yuv$ at $0$.
+thresholding the computed values $\yuv$ at $0$.
 

### 2017-09-11 troll/experiments.tex
small clarification to the first batch algo

@@
 
-  \item The algorithm analyzed at the beginning of \autoref{s:algbatch}, which we call \usrule{}
+  \item The algorithm analyzed \autoref{ss:bayes_approx}, which we call \usrule{}
     (Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). After
@@
     are almost equal (see Table~2 in the supplementary material), thereby suggesting that
-    predictor~\eqref{eq:predictor}, derived from the theoretical results at the beginning of
-    \autoref{s:algbatch}, is \emph{also} the best logistic model based on trollness and
+    predictor~\eqref{eq:predictor}, derived from the theoretical results in
+    \autoref{ss:bayes_approx}, is \emph{also} the best logistic model based on trollness and
     untrustworthiness.

### 2017-09-11 troll/generative.tex
small clarification to the first batch algo

@@
 \begin{equation}\label{e:pout}
-	\frac{1}{2}\,\Biggl(p_u + \frac{1}{\dout(u)} \sum_{v\in \NNout(u)} q_v \Biggl)
+	\frac{1}{2}\,\Biggl(p_u + \frac{1}{\dout(u)} \sum_{v\in \nnout(u)} q_v \Biggl) =
+		\frac{1}{2}\,\left(p_u + \qbar_u\right)\,,
 \end{equation}
+where $\qbar_u = \frac{1}{\dout(u)} \sum_{v\in \nnout(u)} q_v$ is the average $q$ value of $u$ out
+neighbors.
 Similarly, the probability of drawing at random a $+1$-labeled edge from $\Nin(v)$ equals
 \begin{equation}\label{e:pin}
-	\frac{1}{2}\,\Biggl(q_v + \frac{1}{\din(v)} \sum_{u\in \NNin(v)} p_u \Biggl)
+	\frac{1}{2}\,\Biggl(q_v + \frac{1}{\din(v)} \sum_{u\in \NNin(v)} p_u \Biggl)=
+		\frac{1}{2}\,\left(q_v + \pbar_v\right)
 \end{equation}

### 2017-09-11 troll/supp_body.tex
small clarification to the first batch algo

@@
 \subsection{Proof of \autoref{t:active}}
+\label{ssec:troll_proof_bayes}
 
@@
 % Given a function $f\,:\, \Nat \rightarrow \R$ such that $f(n) \rightarrow \infty$ as $n \rightarrow \infty$, 
-Let $\NN_1, \ldots, \NN_n$ be subsets of a finite set $E$.
+Let $\NN_1, \ldots, \NN_n$ be $n$ subsets of a finite set $E$.
 % with $|E| \geq n\,f(n)$
@@
 Let $E_0 \subseteq E$ be sampled uniformly at random {\em without} replacement from $E$, with $|E_0| = m$.
-Then, for $\delta \in (0,1)$, $Q > 0$, and $\theta \geq 2\times\max\left\{Q,4\ln \frac{n}{\delta}\right\}$, we have
+Then, for $\delta \in (0,1)$, $Q > 0$, and $\theta \geq 2\times\max\left\{Q,4\log \frac{n}{\delta}\right\}$, we have
 \[
@@
 \]
-provided $|E| \geq m \geq \frac{2|E|}{\theta}\times\max\left\{Q,4\ln \frac{n}{\delta}\right\}$.
+provided $|E| \geq m \geq \frac{2|E|}{\theta}\times\max\left\{Q,4\log \frac{n}{\delta}\right\}$.
 \end{lemma}
 
-\begin{proof}[Proof of Lemma \ref{l:bins}]
-Set for brevity $p_i = |\NN_i|/|E|$. Then, due to the sampling without replacement, each random
+\begin{proof}[Proof of \autoref{l:bins}]
+  Set for brevity $p_i = \nicefrac{|\NN_i|}{|E|}$. Then, due to the sampling without replacement,
+  each random
 variable $|\NN_i \cap E_0|$ is the sum of $m$ dependent Bernoulli random variables $X_{i,1}, \ldots,
 X_{i,m}$ such that $\Pr(X_{i,t} = 1) = p_i$, for $t = 1, \ldots, m$. Let $i$ be such that $|\NN_i|
-\geq \theta$. Then the condition $m \geq \frac{2|E|Q}{\theta}$ implies
+\geq \theta$. Then, since $\theta \geq 2Q$, we may assume that $m\frac{2Q}{\theta} \leq |E|$, which implies
 \[
@@
 \]
-Since the variables $X_{u,v}$ are negatively associated, we may apply a (multiplicative) Chernoff
+Since the variables $X_{i,j}$ are negatively associated, we may apply a (multiplicative) Chernoff
 bound~\autocite[Section~3.1]{dpbook}.
@@
 which is in turn upper bounded by $\delta$ whenever 
-$m \geq \frac{8|E|}{\theta}\ln \frac{n}{\delta}$. % = 8\,n\,f(n)\,\ln \frac{n}{\delta}$.
+$m \geq \frac{8|E|}{\theta}\log \frac{n}{\delta}$ (again a natural assumption since $\theta \geq
+8\log\frac{n}{\delta}$). % = 8\,n\,f(n)\,\log \frac{n}{\delta}$.
 \end{proof}
@@
 Let now $E_{\theta} = \theset{(u,v) \in E}{\din(v) \ge \theta,\, \dout(u) \ge \theta} \setminus
-E_0$, where $E_0 \ss E$ is the set of edges sampled by the learning algorithm of
-\autoref{s:algbatch}. Then \autoref{t:active} in the main paper is an immediate consequence of the
+E_0$, where $E_0 \ss E$ is the set of sampled edges provided to the learning algorithm of
+\autoref{ss:bayes_approx}. Then \autoref{t:active} in the main paper is an immediate consequence of the
 following lemma.
@@
 For all $0 < \alpha,\delta < 1$ and $0 < \ve < \tfrac{1}{16}$, if the learning algorithm of
-\autoref{s:algbatch} is run with parameters $Q = \tfrac{1}{2\ve^2}\ln\frac{4|V|}{\delta}$ and
-$\alpha$, then with probability at least $1-11\delta$ the predictions $\yhat(u,v)$ satisfy
+\autoref{ss:bayes_approx} is run with parameters $Q = \tfrac{1}{2\ve^2}\log\frac{4|V|}{\delta}$ and
+$\alpha$, then with probability at least $1-9\delta$ the predictions $\yhat(u,v)$ satisfy
 $\yhat(u,v) = y^*(u,v)$ for all $(u,v) \in E_{\theta}$ such that $\big|\eta(u,v) - \tfrac{1}{2}\big|
@@
 
-\begin{proof}[Proof of Lemma \ref{l:active}]
-We apply \autoref{l:bins} with $\theta = \tfrac{2Q}{\alpha} \ge 2\times\max\big\{Q,4\ln
-\frac{2|V|+1}{\delta}\big\}$ to the $2|V|+1$ subsets of $E$ consisting of $E_L$ and
-$\Nin(u),\Nout(u)$, for $i \in V$. We have that, with probability at least $1-\delta$, at least $Q$
-edges of $E_L$ are sampled, at least $Q$ edges of $\Nin(u)$ are sampled for each $i$ such that
-$\NNin(u) \ge \theta$, and at least $Q$ edges of $\Nout(v)$ are sampled for each $j$ such that
-$\NNout(v) \ge \theta$.
+\begin{proof}[Proof of \autoref{l:active}]
+We apply \autoref{l:bins} with $\theta = \tfrac{2Q}{\alpha} \ge 2\times\max\big\{Q,4\log
+\frac{2|V|+1}{\delta}\big\}$ to the $2|V|+1$ subsets of $E$ consisting of $E_L$ and, for all $u \in
+V$, of $\Nin(u)$ and $\Nout(u)$. We have that, with probability at least $1-\delta$, at least $Q$
+edges of $E_L$ are sampled, at least $Q$ edges of $\Nin(v)$ are sampled for each $v$ such that
+$\NNin(v) \ge \theta$, and at least $Q$ edges of $\Nout(u)$ are sampled for each $u$ such that
+$\NNout(u) \ge \theta$.
 For all $(u,v) \in E_{\theta}$ let
 \[
-  \pbar_j = \frac{1}{\din(v)}\sum_{i\in\NNin(v)} p_i \quad\text{and}\quad \qbar_i = \frac{1}{\dout(u)}\sum_{j\in\NNout(u)} q_j
+  \pbar_v = \frac{1}{\din(v)}\sum_{u\in\NNin(v)} p_u \quad\text{and}\quad
+  \qbar_u = \frac{1}{\dout(u)}\sum_{v\in\NNout(u)} q_v
 \]
@@
 all $(u,v) \in E_{\theta}$. Consider $\hdeltaout(u)$ (the same argument works for $\hdeltain(v)$).
-Let $J_1,\dots,J_Q$ be the first $Q$ draws in $E_0 \cap \NNout(u)$ and define
+Let $V_1,\dots,V_Q$ be the first $Q$ draws in $E_0 \cap \NNout(u)$ and define
 \[
-  \muhat_p(u) = \frac{1}{Q} \sum_{t=1}^Q \frac{p_i+q_{J_t}}{2}~.
+  \muhat_p(u) = \frac{1}{Q} \sum_{t=1}^Q \frac{p_u+q_{V_t}}{2}~.
 \]
@@
 \[
-  \mu_p(u) = \frac{1}{\dout(u)}\sum_{j \in \NNout(u)} \frac{p_i + q_j}{2} = \frac{p_i + \qbar_i}{2}~.
+  \mu_p(u) = \frac{1}{\dout(u)}\sum_{v \in \NNout(u)} \frac{p_u + q_v}{2} = \frac{p_u + \qbar_u}{2}~.
 \]
 Now consider the random variables $Z_t = \Ind{y_{u,v_t}=1}$, for $t=1,\dots,Q$. Conditioned on
-$J_1,\dots,J_Q$, these are independent Bernoulli random variables with $\E[Z_t \mid J_t] =
-\tfrac{p_i+q_{J_t}}{2}$. Hence, applying a standard Hoeffding bound for independent variables and
+$V_1,\dots,V_Q$, these are independent Bernoulli random variables with $\E[Z_t \mid V_t] =
+\tfrac{p_u+q_{V_t}}{2}$. Hence, applying a standard Hoeffding bound for independent variables and
 using our choice of $Q$, we get that
@@
 \]
-with probability at least $1-\delta/(2|V|)$ for every realization of $J_1,\dots,J_Q$.
+with probability at least $1-\delta/(2|V|)$ for every realization of $V_1,\dots,V_Q$.
 
@@
 occurring as endpoints in $E_{\theta}$ overlap. Hence, in order to prove a uniform concentration
-result, we need to apply the union bound over the random variables defined over these sample spaces,
-which motivates the presence of the factor $\ln(2|V|)$ in the definition of $Q$.}
+result, we need to apply the union bound over the random variables defined over these sample spaces.}
 on the set $\theset{\hdeltain(v),\hdeltaout(u)}{(u,v) \in E_{\theta}}$, we get that
@@
   \label{eq:first}
-  \left| \hdeltaout(u) + \hdeltain(v) - \frac{p_i + q_j}{2} - \frac{\pbar_j + \qbar_i}{2} \right| \le 4\ve
+  \left| \hdeltaout(u) + \hdeltain(v) - \frac{\pbar_v + \qbar_u}{2} - \frac{p_u + q_v}{2} \right| \le 4\ve
 \end{equation}
 simultaneously holds for all $(u,v) \in E_{\theta}$ with probability at least $1 - 4\delta$. Now
-notice that $\pbar_j$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables drawn from the
-prior marginal $\int_0^1 \mu\big(\cdot,q\bigr) dq$ with expectation $\mu_p$. Similarly, $\qbar_i$ is
+notice that $\pbar_v$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables drawn from the
+prior marginal $\int_0^1 \mu\big(\cdot,q\bigr) dq$ with expectation $\mu_p$. Similarly, $\qbar_u$ is
 a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables independently drawn from the prior
@@
 for independent variables, together with the union bound to the set of pairs of random variables
-whose sample means are $\pbar_j$ and $\qbar_i$ for each $(u,v) \in E_{\theta}$ (there are at most
+whose sample means are $\pbar_v$ and $\qbar_u$ for each $(u,v) \in E_{\theta}$ (there are at most
 $2|V|$ of them) we obtain that
 \begin{align*}
-  \big|\pbar_j - \mu_p\big| \le \ve
+  \big|\pbar_v - \mu_p\big| \le \ve
   \qquad\text{and}\qquad
-  \big|\qbar_i - \mu_q\big| \le \ve
+  \big|\qbar_u - \mu_q\big| \le \ve
 \end{align*}
@@
   \label{eq:this}
-  \left| \hdeltaout(u) + \hdeltain(v) - \frac{p_i + q_j}{2} - \frac{\mu_p + \mu_q}{2} \right| \le 5\ve
+  \left| \hdeltaout(u) + \hdeltain(v) - \frac{\mu_p + \mu_q}{2} - \frac{p_u + q_v}{2} \right| \le 5\ve
 \end{equation}
@@
 \[
-  \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(u,v) \in E_L'} \Pr\big(y_{u,v} = 1\big) = \frac{1}{Q}\sum_{(u,v) \in E_L'} \frac{p_i+q_j}{2}~,
+  \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(u,v) \in E_L'} \Pr\big(y_{u,v} = 1\big) = \frac{1}{Q}\sum_{(u,v) \in E_L'} \frac{p_u+q_v}{2}~,
 \]
@@
 Hence, by applying again Hoeffding bound (this time without the union bound) to the $Q =
-\tfrac{1}{2\ve^2}\ln\tfrac{2}{\delta}$ independent Bernoulli random variables $\Ind{y_{u,v} = 1}$,
+\tfrac{1}{2\ve^2}\log\tfrac{4|V|}{\delta}$ independent Bernoulli random variables $\Ind{y_{u,v} = 1}$,
 $(u,v) \in E_L'$, the event
@@
 \[
-  F(\bp,\bq) = \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(u,v) \in E_L'} \frac{p_i+q_j}{2}~.
+  F(\bp,\bq) = \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(u,v) \in E_L'} \frac{p_u+q_v}{2}~.
 \]
 For any realization $\bq_0$ of $\bq$, the function $F_1(\bp) = F(\bp,\bq_0)$ is a sample mean of $Q
-= \tfrac{1}{2\ve^2}\ln\tfrac{4|V|}{\delta}$ i.i.d.\ $[0,1]$-valued random variables
-$\theset{p_i}{(u,v) \in E_L'}$ (recall that if $i \in V$ is the origin of an edge $(u,v) \in E_L'$,
+= \tfrac{1}{2\ve^2}\log\tfrac{4|V|}{\delta}$ i.i.d.\ $[0,1]$-valued random variables
+$\theset{p_u}{(u,v) \in E_L'}$ (recall that if $u \in V$ is the origin of an edge $(u,v) \in E_L'$,
 then it is not the origin of any other edge $(u,v') \in E_L'$). Using again the standard Hoeffding
@@
 \[
-  \left| \hdeltaout(u) + \hdeltain(v) - \tauhat - \frac{p(u) + q(v)}{2} \right| \le 8\ve
+  \left| \hdeltaout(u) + \hdeltain(v) - \tauhat - \frac{p_u+q_v}{2} \right| \le 8\ve
 \]
-simultaneously holds for each $(u,v) \in E_{\theta}$ with probability at least $1 - 10\delta$.
+simultaneously holds for each $(u,v) \in E_{\theta}$ with probability at least $1 - 9\delta$.
 Putting together concludes the proof.
@@
 \[
-    2\Psi(Y) + \sqrt{(4\ln 2) \Psi(Y)}~.
+    2\Psi(Y) + \sqrt{(4\log 2) \Psi(Y)}~.
 \]

### 2017-09-12 troll/generative.tex
typo

@@
 representing the likelihood of the network topology, rather than the probability of edge signs, once
-the topology is \emph{given}.}
+the topology is \emph{given}.
 
@@
 \begin{equation}\label{e:pout}
-	\frac{1}{2}\,\Biggl(p_u + \frac{1}{\dout(u)} \sum_{v\in \nnout(u)} q_v \Biggl) =
+	\frac{1}{2}\,\Biggl(p_u + \frac{1}{\dout(u)} \sum_{v\in \NNout(u)} q_v \Biggl) =
 		\frac{1}{2}\,\left(p_u + \qbar_u\right)\,,
 \end{equation}
-where $\qbar_u = \frac{1}{\dout(u)} \sum_{v\in \nnout(u)} q_v$ is the average $q$ value of $u$ out
+where $\qbar_u = \frac{1}{\dout(u)} \sum_{v\in \NNout(u)} q_v$ is the average $q$ value of $u$ out
 neighbors.

### 2017-09-12 troll/algs_batch.tex
pass over the Max Likelihood algo

@@
 For simplicity, assume the joint prior distribution $\mu(p,q)$ is uniform over $[0,1]^2$ with
-independent marginals, and suppose that we draw at random without replacement the training set $E_0
+independent marginals,\footnote{As we will see, in real data, around $80\%$ if the edges are
+positive, meaning this assumption of uniformity over $[0,1]^2$ is unlikely to fully holds, for
+otherwise the signs would more balanced} and suppose that we draw at random without replacement the
+training set $E_0
 = \big((u_1,v_1),y_{u_1,v_1}), ((u_2,v_2),y_{u_2,v_2}), \ldots, ((u_m,v_m),y_{u_m,v_m}\big)$, with
@@
 likelihood estimator of the parameters $\{p_i, q_i\}_{i=1}^{|V|}$ based on \trainset{}.
-As showed in the supplementary material, the gradient of the log-likelihood function w.r.t.\ $\{p_i,
-q_i\}_{i=1}^{|V|}$ satisfies
+As showed in the supplementary material \vpageref{ssec:troll_mle_deriv}, the gradient of the
+log-likelihood function w.r.t.\ $\{p_i, q_i\}_{i=1}^{|V|}$ satisfies
 
@@
 
-\begin{align}
-&\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_i, q_i\}_{u=1}^{|V|}\right)}{\partial q_{\ell}}\label{e:mlq}\\
-&\ \ =
-\sum_{k=1}^m
+\begin{equation}\label{e:mlq}
+\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_i, q_i\}_{u=1}^{|V|}\right)}{\partial q_{\ell}}
+= \sum_{k=1}^m
 \frac{\Ind{v_k = \ell,y_{u_k,\ell}=+1}}{p_{u_k}+q_{\ell}}\,
 - \sum_{k=1}^m
-\frac{\Ind{v_k = \ell,y_{u_k,\ell}=-1}}{2-p_{u_k}-q_{\ell}}\,,\notag
-\end{align}
+\frac{\Ind{v_k = \ell,y_{u_k,\ell}=-1}}{2-p_{u_k}-q_{\ell}}\,,
+\end{equation}
 
@@
 Our \emph{first approximation} is therefore to replace the nonlinear equations resulting
-from \eqref{e:mlp} and \eqref{e:mlq} by the following set of linear equations\footnote{Details
-are provided in the supplementary material.}, one for each $\ell \in V$:
+from \eqref{e:mlp} and \eqref{e:mlq} by a set of linear equations. In the case of \eqref{e:mlp}, for
+a given $\ell \in V$, we make the assumption that $(p_{\ell}+q_{v_k}) (2-p_{\ell}-q_{v_k})$ is
+constant for every $k \in [1,\ldots,m]$. Multiplying \eqref{e:mlp} by this constant
+quantity and setting the resulting equation to zero, we obtain for each $\ell \in V$:
 \begin{equation*}
@@
 
-\begin{align*}
-  \sum_{k=1}^m &\Ind{v_k = \ell,y_{u_k,\ell}=+1} \left(2-p_{u_k}-q_{\ell}\right)\\
-  =
-  &\sum_{k=1}^m  \Ind{v_k = \ell,y_{u_k,\ell}=-1}
+We apply a similar transformation to \eqref{e:mlq} in order to obtain, for each $\ell \in V$:
+\begin{equation*}
+  \sum_{k=1}^m \Ind{v_k = \ell,y_{u_k,\ell}=+1} \left(2-p_{u_k}-q_{\ell}\right)
+  = \sum_{k=1}^m  \Ind{v_k = \ell,y_{u_k,\ell}=-1}
   \left(p_{u_k}+q_{\ell}\right)
-\end{align*}
+\end{equation*}
 
@@
 Binary $\pm 1$ predictions on the test set $E\setminus \trainset{}$ are then obtained by
-thresholding the computed values $\yuv$ at $0$.
+thresholding the computed values $\yuv$.
 
@@
 While we note here that such linear transformation of the variables does not change the problem, we
-provide more details in Section~1.3 of the supplementary material.} variables $p_i, q_i$ and $\yuv$
-so as they all lie in the interval $[-1,+1]$.
+provide more details in \autoref{ssec:troll_lprop_weights} of the supplementary material.} variables
+$p_i, q_i$ and $\yuv$ so as they all lie in the interval $[-1,+1]$.
 After this step, one can see that, once we get back to the original variables, label propagation
@@
 The function ${\widehat f}$ is thus a regularized version of the target function $f_{E_0} +
-f_{E\setminus E_0}$ in (\ref{e:quadratic}), where the regularization term tries to enforce the extra
+f_{E\setminus E_0}$ in \eqref{e:quadratic}, where the regularization term tries to enforce the extra
 constraint that whenever a node $u$ has a high out-degree then the corresponding $p_u$ should be
@@
 Notice that the presence of negative weights on the edges of $G''$ does not prevent label
-propagation from converging. Indeed, any node classification algorithm handling both positive and
+propagation from converging. In fact, any node classification algorithm handling both positive and
 negative weights on the edges of $G''$ could be used instead of label propagation. One alternative
-would be the \textsc{wta} algorithm from \autocite{WTA13}.  This is the algorithm we will be
-championing in our experiments of Section~\ref{s:exp}.
+would thus be the \textsc{wta} algorithm from \autocite{WTA13}. However, our label propagation
+algorithm is the one we will be championing in our experiments of \autoref{s:exp}.

### 2017-09-12 troll/supp_body.tex
pass over the Max Likelihood algo

@@
 \subsection{Derivation of the maximum likelihood equations}
+\label{ssec:troll_mle_deriv}
 
@@
 - \, \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=-1}}{2-p_{u_k}-q_{\ell}}\,.
-\end{align*}
-
-We then derive the approximation presented in the main paper. Namely, equating to zero the
-gradient of the log likelihood w.r.t $p_\ell$ gives
-\begin{align*}
-\sum_{k=1}^m 
-\frac{\Ind{u_k = \ell,y_{\ell,v_k}=+1}}{p_{\ell}+q_{v_k}}\,
- - \,
-\frac{\Ind{u_k = \ell,y_{\ell,v_k}=-1}}{2-p_{\ell}-q_{v_k}} &= 0
-\end{align*}
-
-To simplify the notation, let $a_k = \Ind{u_k = \ell,y_{\ell,v_k}=+1}$, $b_k=\Ind{u_k =
-\ell,y_{\ell,v_k}=+1}$ and $c_k = p_{\ell} + q_{v_k}$, we can rewrite the previous equation as 
-\begin{align*}
-  \sum_{k=1}^m \frac{a_k}{c_k} - \frac{b_k}{2-c_k} = 0 =
-  \sum_{k=1}^m  \frac{a_k(2-c_k) - b_k c_k }{ c_k(2-c_k) }
-\end{align*}
-
-The approximation consists in assuming that the denominator $c_k(2-c_k)$ is a
-constant for all $k$, and can therefore be disregarded.
-Moving $b_kc_k$ to the right hand side and returning to the original variables,
-it yields the approximate equation presented in the main paper, namely
-\begin{align*}
-  \sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,v_k}=+1} \left(2-p_{\ell}-q_{v_k}\right)
-  = \sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,v_k}=-1}
-  (p_{\ell}+q_{v_k})
-\end{align*}
-
-By applying a similar process to the gradient w.r.t. $q_\ell$, we obtain the
-following approximate equation
-\begin{align*}
-\sum_{k=1}^m \Ind{v_k = \ell,y_{u_k,\ell}=+1} \left(2-p_{u_k}-q_{\ell}\right)
-= \sum_{k=1}^m  \Ind{v_k = \ell,y_{u_k,\ell}=-1}
-\left(p_{u_k}+q_{\ell}\right)
 \end{align*}

### 2017-09-12 intro/use_of_sign_graphs.tex
add 2 refs

@@
       the information contained within the negative links to enhance the visualization of social
-      networks~\autocite{Luca10}. Note however that those three last papers not explicitly use a
-      \pcc{} formulation.
+      networks~\autocites{Luca10}{drawingSignedGraphs11}. Note however that those four last papers
+      not explicitly use a \pcc{} formulation.
 

### 2017-09-13 troll/algs_online.tex
move some proofs back to the online section

@@
 
-For the online scenario, we have the following result.
+In the online scenario, recall we do not assume anymore that the signs are generated by the model of
+\autoref{s:gen}. Instead, they are chosen adversarially, and presented sequentially in order to
+force as many mistakes as possible. This is achieved by deviating from the regular labeling where
+every user is either perfectly troll or trustworthy, as measured by the regularity measure
+$\Psi_G(Y)$ described at the end of \autoref{s:prel}. We start by presenting an algorithm that
+combines randomized Weighted Majority instances and makes little more than $\Psi_G(Y)$ mistakes on
+any edge-labeled graph $G(Y) = (V,E(Y))$.  We then show there is not much room for improvement, for
+as long as $\Psi_G(Y)$ is within a budget of $K$ irregularities, any online algorithm is condemned
+to err at least $\nicefrac{K}{2}$ times.
 
@@
 \end{theorem}
+\begin{proof}
 
-The algorithm used in \autoref{t:online} is a combination of randomized Weighted Majority instances.
-Details are reported in the supplementary material. We complement the above result by providing a
+Let each node $u \in V$ host two instances of the randomized Weighted Majority (RWM)
+algorithm~\cite{LittlestoneWa94} with an online tuning of their learning rate~\cite{cb+97,acg02}:
+one instance for predicting the sign of outgoing edges $(u,v)$, and one instance for predicting the
+sign of incoming edges $(v,u)$. Both instances simply compete against the two constant experts,
+predicting always $+1$ or always $-1$. Denote by $M(u,v)$ the indicator function (zero-one loss) of
+a mistake on edge $(u,v)$. Then the expected number of mistakes of each RWM instance
+satisfy~\cite{cb+97,acg02}:
+\[
+    \sum_{v \in \NNout(u)} \E\,M(u,v) = \Psiout(u,Y) + O\left(\sqrt{\Psiout(u,Y)}+ 1\right)
+\]
+and
+\[
+    \sum_{u \in \NNin(v)} \E\,M(u,v) = \Psiin(v,Y) + O\left(\sqrt{\Psiin(v,Y)} + 1\right)~.
+\]
+We then define two meta-experts: an ingoing expert, which predicts $\yuv$ using the prediction of
+the ingoing RWM instance for node $v$, and the outgoing expert, which predicts $\yuv$ using the
+prediction of the outgoing RWM instance for node $u$.
+The number of mistakes of these two experts satisfy
+\begin{align*}
+  \sum_{u \in V}\sum_{v \in \NNout(u)} \E\,M(u,v)
+                                       &= \Psiout(Y) + O\left(\sqrt{|V|\Psiout(Y)} + |V|\right)\\
+  \sum_{v \in V}\sum_{u \in \NNin(v)} \E\,M(u,v)
+                                      &= \Psiin(Y)  + O\left(\sqrt{|V|\Psiin(Y)}  + |V|\right)~,
+\end{align*}
+where we used $\sum_{v \in V} \sqrt{\Psiin(v,Y)} \leq \sqrt{|V|\Psiin(Y)}$, and similarly for
+$\Psiout(Y)$.
+Finally, let the overall prediction of our algorithm be a RWM instance run on top of the ingoing and
+the outgoing experts. Then the expected number of mistakes of this predictor satisfies~
+\begin{align*}
+  \sum_{(u,v) \in E} \E\,M(u,v) &= \Psi_G(Y) + O\Biggl(\sqrt{|V|\Psi_G(Y)} + |V|
+    + \sqrt{\left(\Psi_G(Y) + |V| + \sqrt{|V|\Psi_G(Y)}\right)} \Biggl)\\
+    &= \Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V|\right)
+\end{align*}
+as claimed.
+\end{proof}
+
+We complement the above result by providing a
 mistake lower bound. Like \autoref{t:online}, the following result holds for all graphs, and for all
@@
   \tfrac{|E|}{2}\big\rfloor$, a randomized labeling $Y\in\spin^{|E|}$ exists such that $\Psi_G(Y)
-  \leq K$, and the expected number of mistakes that any online algorithm $A$ can be forced to make
-  satisfies $\E M_A(Y) \ge \frac{K}{2}$.  Moreover, as $\frac{K}{|E|} \rightarrow 0$ then $\E M_A(Y)
-  = K$.
+  \leq K$, and the expected number of mistakes that any online algorithm $A$ make can be forced to 
+  satisfies $\E M_A(Y) \ge \frac{K}{2}$.
 \end{theorem}
+
+\begin{proof}
+  Let $\mathcal{Y}_K$ be the set of all labelings $Y$ such that the total number of negative and
+  positive edges are $K$ and $|E|-K$, respectively (without loss of generality we will focus on
+  negative edges). Consider the randomized strategy that draws a labeling $Y \in \spin^{|E|}$ of the
+  edges of the input graph \uar{} from $\mathcal{Y}_K$. For each node $u \in V$, we have
+  $\Psiin(u,Y) \le \din^-(u)$, which implies $\Psiin(Y) \le K$. A very similar argument applies to
+  the outgoing edges, leading to $\Psiout(Y) \le K$. The constraint $\Psi_G(Y) \le K$ is therefore
+  always satisfied.
+
+  The adversary will force on average \shalf{} mistakes in each one of the first $K$ rounds of the
+  online protocol by repeating $K$ times the following: (i) A label value $\ell\in\{-1,+1\}$ is
+  selected uniformly at random. (ii) An edge $(u,v)$ is sampled uniformly at random from the set of
+  all edges that were not previously revealed and whose labels are equal to $\ell$.
+
+  The learner is required to predict $y_{u,v}$ and, in doing so, \shalf{} mistakes will be made on
+  average because of the randomized labeling procedure. Observe that this holds even when $A$ knows
+  the value of $K$ and $\Psi_G(Y)$. Hence, we can conclude that the expected number of mistakes that
+  $A$ can be forced to make is always at least $K/2$, as claimed.
+\end{proof}
 
@@
 lower bound gets arbitrarily close to $K$ for any $G(Y)$, hence asymptotically matching the upper
-bound of \autoref{t:online} (\textit{proof in the supplementary material}).
+bound of \autoref{t:online}. A sketch of the proof can be found \vpageref{proof:troll_K_mistakes}.

### 2017-09-13 troll/supp_body.tex
move some proofs back to the online section

@@
 
-\section{Proofs from \autoref{s:algonline}}
+\section{Proof from \autoref{s:algonline}}
 
-\begin{proof}[Proof of \autoref{t:online}]
+\begin{proof}[Proof sketch that, in the context of \autoref{t:mistake_bound}, as $\frac{K}{|E|}
+  \rightarrow 0$ then $\E M_A(Y)=K$  ]\label{proof:troll_K_mistakes}
 
-Let each node $u \in V$ host two instances of the randomized Weighted Majority (RWM)
-algorithm~\cite{LittlestoneWa94} with an online tuning of their learning rate~\cite{cb+97,acg02}:
-one instance for predicting the sign of outgoing edges $(u,v)$, and one instance for predicting the
-sign of incoming edges $(v,u)$. Both instances simply compete against the two constant experts,
-predicting always $+1$ or always $-1$. Denote by $M(u,v)$ the indicator function (zero-one loss) of
-a mistake on edge $(u,v)$. Then the expected number of mistakes of each RWM instance
-satisfy~\cite{cb+97,acg02}:
-\[
-    \sum_{v \in \NNout(u)} \E\,M(u,v) = \Psiout(u,Y) + O\left(\sqrt{\Psiout(u,Y)}+ 1\right)
-\]
-and
-\[
-    \sum_{u \in \NNin(v)} \E\,M(u,v) = \Psiin(v,Y) + O\left(\sqrt{\Psiin(v,Y)} + 1\right)~.
-\]
-We then define two meta-experts: an ingoing expert, which predicts $y_{u,v}$ using the prediction of
-the ingoing RWM instance for node $v$, and the outgoing expert, which predicts $y_{u,v}$ using the
-prediction of the outgoing RWM instance for node $u$.
-The number of mistakes of these two experts satisfy
-\begin{align*}
-  \sum_{u \in V}\sum_{v \in \NNout(u)} \E\,M(u,v)
-                                       &= \Psiout(Y) + O\left(\sqrt{|V|\Psiout(Y)} + |V|\right)\\
-  \sum_{v \in V}\sum_{u \in \NNin(v)} \E\,M(u,v)
-                                      &= \Psiin(Y)  + O\left(\sqrt{|V|\Psiin(Y)}  + |V|\right)~,
-\end{align*}
-where we used $\sum_{v \in V} \sqrt{\Psiin(v,Y)} \leq \sqrt{|V|\Psiin(Y)}$, and similarly for
-$\Psiout(Y)$.
-Finally, let the overall prediction of our algorithm be a RWM instance run on top of the ingoing and
-the outgoing experts. Then the expected number of mistakes of this predictor satisfies~
-\begin{align*}
-  \sum_{(u,v) \in E} \E\,M(u,v) &= \Psi_G(Y) + O\Biggl(\sqrt{|V|\Psi_G(Y)} + |V|
-    + \sqrt{\left(\Psi_G(Y) + |V| + \sqrt{|V|\Psi_G(Y)}\right)} \Biggl)\\
-    &= \Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V|\right)
-\end{align*}
-as claimed.
-\end{proof}
+Let $\mathcal{E}$ be the following event: There is at least one unrevealed negative label.
+The randomized iterative strategy used to force a number of mistakes arbitrarily close to $K$ is identical to the one described
+in the proof of \autoref{t:mistake_bound} \vpageref{t:mistake_bound}, except for the stopping criterion.
+Namely, the adversary draws a labeling $Y$ \uar{} from $\mathcal{Y}_K$, the set of all labelings $Y$
+such that the total number of negative and positive edges are $K$ and $|E|-K$. Then the adversary
+repeats the following two steps: (i) A label value $\ell\in\{-1,+1\}$ is selected \uar{}. (ii) An
+edge $(u,v)$ is sampled \uar{} from the set of all edges that were not previously revealed and whose
+labels are equal to $\ell$. This is done until $\mathcal{E}$ is true.
 
-\begin{proof}[Proof sketch of \autoref{t:mistake_bound}]
-
-Let $\mathcal{Y}_K$ be the set of all labelings $Y$ such that the total number of negative and
-positive edges are $K$ and $|E|-K$, respectively (without loss of generality we will focus on
-negative edges). Consider the randomized strategy that draws a labeling $Y \in \spin^{|E|}$ of the
-edges of the input graph \uar{} from $\mathcal{Y}_K$. For each node $u \in V$, we have $\Psiin(u,Y)
-\le \din^-(u)$, which implies $\Psiin(Y) \le K$. A very similar argument applies to the outgoing
-edges, leading to $\Psiout(Y) \le K$. The constraint $\Psi_G(Y) \le K$ is therefore always
-satisfied.
-
-The adversary will force on average \shalf{} mistakes in each one of the first $K$ rounds of the
-online protocol by repeating $K$ times the following: (i) A label value $\ell\in\{-1,+1\}$ is
-selected uniformly at random. (ii) An edge $(u,v)$ is sampled uniformly at random from the set of
-all edges that were not previously revealed and whose labels are equal to $\ell$.
-
-The learner is required to predict $y_{u,v}$ and, in doing so, \shalf{} mistakes will be clearly
-made on average because of the randomized labeling procedure. Observe that this holds even when $A$
-knows the value of $K$ and $\Psi_G(Y)$. Hence, we can conclude that the expected number of mistakes
-that $A$ can be forced to make is always at least $K/2$, as claimed.
-
-\medskip
-
-We now show that, as $\frac{K}{|E|} \rightarrow 0$, the lower bound gets arbitrarily close to $K$
-for any $G(Y)$ and any constant $K$.
-Let $\mathcal{E}$ be the following event: There is at least one unrevealed negative label.
-The randomized iterative strategy used to achieve this result is identical to the one described
-above, except for the stopping criterion. Instead of repeating step (i) and (ii) only for the first
-$K$ rounds, these steps are repeated until $\mathcal{E}$ is true.
-%
 Let $m_{r,c}$ be defined as follows: For $c=1$ it is equal to the expected number of mistakes forced

### 2017-09-14 troll/diff_mcc_pp.tex
add new dataset, competitor and results

\begin{table}[htpb]
  \centering
  \caption{100 time the difference of MCC between using the twin heuristic and not using it for our
  two main methods. Those values are computed on the same training/testing split as of
\autoref{tab:all_mcc}.\label{tab:twin}}
  \begin{tabular}{lrrrrrrr}
    \toprule
                                & $5\%$                    & $10\%$                   & $15\%$                   & $20\%$                   & $25\%$                   & $50\%$                   & $90\%$                   \\
    \midrule
    \multicolumn{8}{c}{\aut{}} \\
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.7$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Green}{$0.1$} & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.7$}  \\
    \multicolumn{8}{c}{\adv{}} \\
    \uslpropGsecTwin{}          & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.6$}  & \textcolor{Red}{$-1.0$}  & \textcolor{Red}{$-1.6$}  & \textcolor{Red}{$-2.2$}  & \textcolor{Red}{$-5.4$}  & \textcolor{Red}{$-10.8$} \\
    \usruleTwin{}               & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.5$}  & \textcolor{Red}{$-1.0$}  & \textcolor{Red}{$-1.6$}  & \textcolor{Red}{$-2.2$}  & \textcolor{Red}{$-5.7$}  & \textcolor{Red}{$-11.1$} \\
    \multicolumn{8}{c}{\wik{}} \\
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.2$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.2$}  \\
    \multicolumn{8}{c}{\sla{}} \\
    \uslpropGsecTwin{}          & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.5$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$0.9$} & \textcolor{Green}{$1.1$} & \textcolor{Green}{$2.0$} & \textcolor{Green}{$3.3$} \\
    \usruleTwin{}               & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.6$} & \textcolor{Green}{$0.9$} & \textcolor{Green}{$1.1$} & \textcolor{Green}{$1.3$} & \textcolor{Green}{$2.2$} & \textcolor{Green}{$3.5$} \\
    \multicolumn{8}{c}{\epi{}} \\
    \uslpropGsecTwin{}          & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.5$} & \textcolor{Green}{$0.7$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$1.0$} & \textcolor{Green}{$1.9$} & \textcolor{Green}{$3.0$} \\
    \usruleTwin{}               & \textcolor{Green}{$0.4$} & \textcolor{Green}{$0.6$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$1.0$} & \textcolor{Green}{$1.2$} & \textcolor{Green}{$1.8$} & \textcolor{Green}{$2.8$} \\
    \multicolumn{8}{c}{\kiw{}} \\
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.3$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.5$}  \\
    \bottomrule
  \end{tabular}
\end{table}

### 2017-09-14 troll/experiments.tex
add new dataset, competitor and results

@@
 
-We considered five real-world classification datasets. The first three are directed \ssn{} widely
-used as benchmarks for this task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}):
+We considered six real-world classification datasets. The first four are \dssn{} widely
+used as benchmarks for this
+task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
+In \adv{}, a trust-based social network for open source developer, a user $u$ can certify another
+user $v$ with different degrees of trust: \enquote{Observer}, \enquote{Apprentice} (both of which we
+consider negative), \enquote{Journeyer} and \enquote{Master} (both of which we consider
+positive).\footnote{We download the 7\thup{} of July, 2014 version from
+\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric and its
+resistance to attacks is available in the PhD thesis of the website's creator~\autocite[Section
+4]{AdvogatoTrustThesis02}).
 In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
@@
 $v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
-$v$ to be reliable or not. In addition to these three datasets, we considered two other \ssn{} where
+$v$ to be reliable or not. In addition to these four datasets, we considered two other \ssn{} where
 the signs are inferred automatically, rather than given explicitly by the users.  In
@@
   \centering
-  \caption{Dataset properties. The 5th column gives the fraction of positive labels. The last two
-  columns provide two different measures of label regularity ---see main text.\label{tab:dataset}}
+  \small
+  \caption{Dataset properties. The 5\thup{} column gives the fraction of positive labels. The next two
+  columns provide two different measures of label regularity, while the last two columns give the
+  proportion of reciprocal edges, and among them the fraction with different signs.\label{tab:dataset}}
   \begin{tabular}{lrrrrrrrr}
     \toprule
-    Dataset &       $|V|$ &       $|E|$ &$\frac{|E|}{|V|}$ & $\frac{|E^+|}{|E|}$ & $\frac{\Psi^2_{G''}(Y)}{|E|}$ & $\frac{\Psi_G(Y)}{|E|}$ \\ % & twin edges & twin disagreement \\
+    Dataset & $|V|$       & $|E|$       & $\frac{|E|}{|V|}$ & $\frac{|E^+|}{|E|}$ & $\frac{\Psi^2_{G''}(Y)}{|E|}$ & $\frac{\Psi_G(Y)}{|E|}$ & twin edges & twin disagreement \\
     \midrule                                                                                                   
-    \aut{} &   \np{4831} &  \np{39452} & 8.1  &              72.3\% &                    .076 &                       .191 \\ %&      5.07\% &            27.10\% \\
-    \wik{} &   \np{7114} & \np{103108} & 14.5 &              78.8\% &                    .063 &                       .142 \\ %&      5.56\% &            10.04\% \\
-    \sla{} &  \np{82140} & \np{549202} & 6.7  &              77.4\% &                    .059 &                       .143 \\ %&     17.74\% &             4.00\% \\
-    \kiw{} & \np{138587} & \np{740106} & 5.3  &              87.9\% &                    .034 &                       .086 \\ %&      6.55\% &             2.09\% \\
-    \epi{} & \np{131580} & \np{840799} & 6.4  &              85.3\% &                    .031 &                       .074 \\ %&     30.83\% &            14.64\% \\
+    \aut{}  & \np{4831}   & \np{39452}  & 8.2               & 72.3\%              & .076                          & .191                    & 5.1\%      & 27.1\%            \\
+    \adv{}  & \np{5417}   & \np{51312}  & 9.5               & 75.1\%              & .061                          & .132                    & 33.6\%     & 28.6\%            \\
+    \wik{}  & \np{7114}   & \np{103108} & 14.5              & 78.8\%              & .063                          & .142                    & 5.6\%      & 10.0\%            \\
+    \sla{}  & \np{82140}  & \np{549202} & 6.7               & 77.4\%              & .059                          & .143                    & 17.7\%     & 4.0\%             \\
+    \epi{}  & \np{131580} & \np{840799} & 6.4               & 85.3\%              & .031                          & .074                    & 30.8\%     & 2.1\%             \\
+    \kiw{}  & \np{138587} & \np{740106} & 5.3               & 87.9\%              & .034                          & .086                    & 6.5\%      & 14.6\%            \\
     \bottomrule
@@
 
-\iffalse
-***********************************************************************
-Finally, there is a low fraction of reciprocal (or \emph{twin}) edges (i.e.\
-$\eij \in E$ and $\eji \in E$, which is a common mechanism of link formation in
-directed networks~\cite{DirectedReciprocity04}), and in most cases they do not
-disagree, meaning they have the same sign. In practice, we use that fact to
-improve our accuracy at no additional computational cost: when predicting $\eij
-\in E_{\mathrm{test}}$, if the reciprocal edge \eji{} is part of the training set, we
-set $\sgn(\eij) = \sgn(\eji)$.
-
-from http://www.nature.com/articles/srep02729
-
-The study of link reciprocity in binary directed networks1,2, or the tendency
-of vertex pairs to form mutual connections, has received an increasing
-attention in recent years3,4,5,6,7,8,9,10,11,12,13,14. Among other things,
-reciprocity has been shown to be crucial in order to classify3 and model4
-directed networks, understand the effects of network structure on dynamical
-processes (e.g. diffusion or percolation processes5,6,7), explain patterns of
-growth in out-of-equilibrium networks (as in the case of the Wikipedia8 or the
-World Trade Web9,10), and study the onset of higher-order structures such as
-correlations11,12 and triadic motifs13,14,15,16
-************************************************************************
-\fi
+Finally, there is a low fraction of reciprocal (or \emph{twin}) edges (\ie{} both $\euv \in E$ and $\evu
+\in E$), which is a common mechanism of link formation in directed
+networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases such twin edges do not disagree,
+\ie{} they have the same sign. In practice, we can use that fact to improve our accuracy at no
+additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
+\evu{} is part of the training set, we set $\sgn(\euv) = \sgn(\evu)$. For clarity, when comparing
+our methods with existing approaches, we not use that trick. But afterwards, we show in
+\autoref{tab:all_mcc_pp} when it can be beneficial.
+% more references in  http://www.nature.com/articles/srep02729
 
@@
 
-\begin{table*}[t]
-  \centering
-  \setlength{\tabcolsep}{3pt}
-  \scriptsize
-  \caption{MCC with increasing training set size, with one standard deviation over 12 random
-  sampling of $\trainset$. The last four columns refer to the methods we took from the literature.
-For the sake of readability, we multiplied all MCC values by 100. The best number in each row is
-highlighted in \textbf{\textcolor{brown}{bold brown}} and the second one in
-\textit{\textcolor{red}{italic red}}. If the difference is statistically significant ($p$-value of a
-paired Student's $t$-test less than $0.005$), the best score is underlined. The ``time" rows contain
-the time taken to train on a $15\%$ training set.\label{tab:all_mcc}}
-  \begin{tabular}{lrcccc|cccc}
-    \toprule
-    & $\frac{|\trainset{}|}{|E|}$ &                 \uslpropGsec{} &                       \usrule{} &                    \uslogregp{} &         \qoptim{} &     \complowrank{} &      \comptriads{} &   \compranknodes{} &              \compbayesian{} \\
-    \midrule
-    \multirow{7}{*}{\rotatebox[origin=c]{90}{\aut{}}} & $5\%$                       & $\vfirstSig{24.54} \pm 0.69$  & $\vsecondSig{20.21} \pm 0.66$ & $20.19 \pm 0.71$              & $15.86 \pm 0.81$ & $12.76 \pm 0.65$ & $11.04 \pm 0.81$ & $17.18 \pm 1.11$             & $15.28 \pm 1.31$ \\
-                                                      & $10\%$                      & $\vfirstSig{31.20} \pm 0.58$  & $\vsecondSig{27.54} \pm 0.56$ & $27.49 \pm 0.62$              & $25.36 \pm 0.78$ & $17.81 \pm 0.76$ & $16.99 \pm 0.63$ & $25.36 \pm 0.85$             & $24.74 \pm 0.59$ \\
-                                                      & $15\%$ &  $\vfirstSig{35.66} \pm 0.68$  &  $\vsecondSig{32.87} \pm 0.58$  &               $32.79 \pm 0.60$  &  $31.39 \pm 0.75$  &  $22.58 \pm 0.53$  &  $21.55 \pm 0.91$  &  $30.60 \pm 0.87$  &            $31.71 \pm 0.99$  \\
-                                                      & $20\%$ &  $\vfirstSig{38.67} \pm 0.48$  &  $\vsecondSig{36.94} \pm 0.51$  &               $36.86 \pm 0.48$  &  $35.47 \pm 0.41$  &  $25.80 \pm 0.94$  &  $24.27 \pm 0.56$  &  $35.01 \pm 0.83$  &            $36.13 \pm 0.75$  \\
-                                                      & $25\%$ &     $\vfirst{41.05} \pm 0.73$  &               $39.83 \pm 0.58$  &               $39.76 \pm 0.59$  &  $38.48 \pm 0.55$  &  $29.67 \pm 0.78$  &  $26.85 \pm 0.87$  &  $38.06 \pm 0.86$  &  $\vsecond{40.34} \pm 0.94$  \\
-                                                      & time   &                           19.6 &                             0.6 &                             2.6 &               2835 &               3279 &                6.2 &                155 &                         4813 \\
-    \midrule
-    \multirow{7}{*}{\rotatebox[origin=c]{90}{\wik{}}} & $5\%$                       & $\vfirstSig{39.46} \pm 0.79$  & $38.03 \pm 0.97$              & $\vsecondSig{38.50} \pm 0.87$ & $35.72 \pm 0.70$ & $24.58 \pm 1.18$ & $9.59 \pm 1.10$  & $33.60 \pm 0.64$             & $26.45 \pm 0.57$ \\
-                                                      & $10\%$                      & $\vsecond{47.17} \pm 0.35$    & $46.03 \pm 0.49$              & $\vfirst{47.22} \pm 0.40$     & $44.53 \pm 0.48$ & $31.72 \pm 0.61$ & $26.36 \pm 0.83$ & $43.21 \pm 0.81$             & $40.28 \pm 0.69$ \\
-                                                      & $15\%$ &     $\vsecond{50.49} \pm 0.33$  &  $49.89 \pm 0.40$  &      $\vfirst{50.87} \pm 0.36$  &  $49.08 \pm 0.33$  &  $35.77 \pm 0.58$  &  $33.64 \pm 0.83$  &  $48.50 \pm 0.47$  &  $47.07 \pm 0.38$  \\
-                                                      & $20\%$ &  $\vsecondSig{52.74} \pm 0.31$  &  $52.24 \pm 0.49$  &   $\vfirstSig{53.13} \pm 0.27$  &  $51.79 \pm 0.35$  &  $37.90 \pm 0.27$  &  $38.41 \pm 0.53$  &  $51.49 \pm 0.43$  &  $50.54 \pm 0.39$  \\
-                                                      & $25\%$ &     $\vsecond{54.00} \pm 0.63$  &  $53.42 \pm 0.59$  &      $\vfirst{54.26} \pm 0.37$  &  $53.31 \pm 0.37$  &  $40.16 \pm 0.57$  &  $41.34 \pm 1.07$  &  $53.30 \pm 0.37$  &  $52.92 \pm 0.48$  \\
-                                                      & time   &                            41.9 &                1.6 &                             6.0 &              10629 &               8523 &               14.8 &                249 &              12507 \\
-    \midrule
-    \multirow{7}{*}{\rotatebox[origin=c]{90}{\sla{}}} & $5\%$                       & $\vsecondSig{40.77} \pm 0.20$ & $36.13 \pm 0.57$              & $37.00 \pm 0.29$              & $33.49 \pm 0.32$ & $36.83 \pm 0.47$ & $27.10 \pm 0.75$ & $\vfirstSig{45.16} \pm 0.59$ & $29.25 \pm 0.23$ \\
-                                                      & $10\%$                      & $\vsecondSig{46.61} \pm 0.29$ & $41.89 \pm 0.39$              & $43.15 \pm 0.21$              & $40.92 \pm 0.23$ & $39.57 \pm 0.27$ & $40.38 \pm 1.47$ & $\vfirstSig{47.84} \pm 0.50$ & $38.25 \pm 0.21$ \\
-                                                      & $15\%$ &   $\vfirstSig{49.62} \pm 0.22$  &  $45.42 \pm 0.36$  &  $46.42 \pm 0.16$  &  $45.56 \pm 0.19$  &  $41.21 \pm 0.19$  &  $45.88 \pm 1.01$  &  $\vsecondSig{48.75} \pm 0.71$  &  $43.47 \pm 0.16$  \\
-                                                      & $20\%$ &     $\vsecond{51.88} \pm 0.24$  &  $47.78 \pm 0.25$  &  $48.66 \pm 0.10$  &  $48.10 \pm 0.30$  &  $42.74 \pm 0.44$  &  $48.79 \pm 0.57$  &      $\vfirst{52.10} \pm 0.33$  &  $46.89 \pm 0.27$  \\
-                                                      & $25\%$ &     $\vsecond{53.12} \pm 0.20$  &  $49.39 \pm 0.24$  &  $50.22 \pm 0.12$  &  $50.11 \pm 0.20$  &  $44.24 \pm 0.44$  &  $50.62 \pm 0.53$  &      $\vfirst{53.29} \pm 0.22$  &  $49.42 \pm 0.22$  \\
-                                                      & time   &                             677 &                8.3 &               32.8 &              78537 &              69988 &                131 &                            2441 &              68085 \\
-    \midrule
-    \multirow{7}{*}{\rotatebox[origin=c]{90}{\epi{}}} & $5\%$                       & $\vsecondSig{54.83} \pm 0.16$ & $46.94 \pm 0.80$              & $49.16 \pm 0.32$              & $42.79 \pm 0.34$ & $39.96 \pm 0.60$ & $42.94 \pm 2.06$ & $\vfirstSig{56.04} \pm 0.76$ & $37.99 \pm 0.49$ \\
-                                                      & $10\%$                      & $\vsecondSig{58.94} \pm 0.27$ & $54.03 \pm 0.46$              & $55.90 \pm 0.13$              & $53.43 \pm 0.39$ & $44.50 \pm 0.52$ & $50.29 \pm 1.07$ & $\vfirstSig{60.60} \pm 0.32$ & $49.90 \pm 0.36$ \\
-                                                      & $15\%$ &  $\vsecondSig{61.47} \pm 0.21$  &  $57.63 \pm 0.45$  &  $59.25 \pm 0.17$  &  $58.80 \pm 0.32$  &  $48.24 \pm 0.58$  &  $54.64 \pm 1.62$  &  $\vfirstSig{62.69} \pm 0.21$  &            $56.94 \pm 0.65$  \\
-                                                      & $20\%$ &  $\vsecondSig{63.17} \pm 0.13$  &  $60.15 \pm 0.40$  &  $61.45 \pm 0.17$  &  $61.86 \pm 0.13$  &  $52.21 \pm 0.37$  &  $57.27 \pm 1.42$  &  $\vfirstSig{64.10} \pm 0.12$  &            $61.18 \pm 0.45$  \\
-                                                      & $25\%$ &               $64.05 \pm 0.20$  &  $61.88 \pm 0.38$  &  $62.89 \pm 0.12$  &  $63.42 \pm 0.14$  &  $54.68 \pm 0.62$  &  $58.42 \pm 1.59$  &     $\vfirst{65.40} \pm 0.85$  &  $\vsecond{64.59} \pm 0.30$  \\
-                                                      & time   &                            1329 &               10.1 &               54.0 &             143881 &             127654 &                209 &                           3174 &                       104305 \\
-    \midrule
-    \multirow{7}{*}{\rotatebox[origin=c]{90}{\kiw{}}} & $5\%$                       & $\vfirstSig{36.36} \pm 0.53$  & $\vsecondSig{30.89} \pm 0.28$ & $30.81 \pm 0.20$              & $21.69 \pm 0.25$ & $23.15 \pm 0.26$ & $3.04 \pm 0.46$  & $26.63 \pm 0.44$             & $26.68 \pm 0.34$ \\
-                                                      & $10\%$                      & $\vfirstSig{38.58} \pm 0.74$  & $35.68 \pm 0.22$              & $\vsecondSig{35.93} \pm 0.16$ & $29.75 \pm 0.21$ & $27.07 \pm 0.44$ & $12.34 \pm 0.79$ & $33.85 \pm 0.33$             & $35.00 \pm 0.34$ \\
-                                                      & $15\%$ &  $\vsecondSig{39.08} \pm 0.55$  &               $37.77 \pm 0.22$  &               $38.27 \pm 0.19$  &  $33.61 \pm 0.11$  &  $30.05 \pm 0.29$  &  $17.95 \pm 0.92$  &  $36.88 \pm 0.32$  &  $\vfirstSig{40.00} \pm 0.26$  \\
-                                                      & $20\%$ &               $39.04 \pm 0.69$  &               $38.88 \pm 0.36$  &  $\vsecondSig{39.55} \pm 0.11$  &  $35.04 \pm 0.17$  &  $32.17 \pm 0.31$  &  $21.44 \pm 0.67$  &  $38.60 \pm 0.31$  &  $\vfirstSig{43.32} \pm 0.22$  \\
-                                                      & $25\%$ &               $38.90 \pm 0.45$  &               $39.41 \pm 0.16$  &  $\vsecondSig{40.44} \pm 0.14$  &  $36.18 \pm 0.20$  &  $33.94 \pm 0.74$  &  $23.41 \pm 0.41$  &  $39.75 \pm 0.32$  &  $\vfirstSig{45.76} \pm 0.29$  \\
-                                                      & time   &                             927 &                             9.6 &                            46.8 &             219109 &             129460 &                177 &               3890 &                          92719 \\
-    \bottomrule
-  \end{tabular}
-\end{table*}
+\input{all_mcc}
 
@@
 Our main results are summarized in \autoref{tab:all_mcc}, reporting MCC test set performance after
-training on sets of varying size (from 5\% to 25\%). Results have been averaged over 12 repetitions.
+training on sets of varying size (from 5\% to 25\%, plus 50\% and 90\%). Results have been averaged
+over 12 repetitions.
 Because scalability is a major concern when training on sizeable datasets, we also give an idea of
@@
     \autoref{s:gen}, the logistic regression weights for features $1-\htr(i)$ and $1-\hun(j)$
-    are almost equal (see Table~2 in the supplementary material), thereby suggesting that
+    are almost equal (see \autoref{tab:coeff} in the supplementary material), thereby suggesting that
     predictor~\eqref{eq:predictor}, derived from the theoretical results in
@@
 \end{enumerate}
+
+We perform two other experiments. The first one is to assess the effect of predicting reciprocal
+edges by their value if available. As expected, the results in \autoref{tab:twin} demonstrate that
+it improves MCC when the network has enough twin edges with low disagreements, like \sla{} and
+\epi{}. It has no effect when there are few twin edges, like in \aut{}, \wik{} and \kiw{}. And it is
+detrimental if when there are many twin edges that do  not agree enough, like in \adv{}, in which
+case it is better to rely solely on the learned model.
+
+\input{diff_mcc_pp}
+
+In the second, we study the effect of our hypothesis that the labeled edges of the training set are
+chosen \uar{}. In two of our datasets (\wik{} and \epi{}), edges come the timestamp at which they
+were created. A more realistic way of choosing the training set is therefore, for a given training
+size $m$, to let $E_0$ be the $m$ oldest edges, and try to predict the remaining, newest one. This
+is a common experimental setting in link prediction, where the goal is to infer which pair of nodes
+will be connected in the future~\autocite{linkPredSurvey16}. However, as showed in
+\input{ts_mcc}
+\autoref{tab:troll_ts_mcc}, this makes the problem much more challenging, both for our methods and
+the existing ones. It is quite surprising that even when the training set is as large as 90\% of all
+edges, the MCC are so low compared with those reported \autoref{tab:all_mcc}. The explanation of
+why these two sampling strategies produce so different results can be inferred from
+\autoref{tab:troll_early_sampling}. Letting $V_{\mathrm{out}}$ be the subset of node of $V$ with at
+least one outgoing edge, and defining similarly $V_{\mathrm{in}}$, we see that the node coverage is
+naturally larger when sampling at random. Yet when the training set reach 90\% size, with both
+sampling strategies we cover roughly 90\% of the nodes in $V_{\mathrm{out}}$ and in
+$V_{\mathrm{out}}$. The difference in MCC is thus explained by the last three columns. It shows, for
+testing edges \euv{}, the breakdown between three situations, from the least to the most
+informative:
+\begin{enumerate}[(i),nosep]
+  \item no sampled edge outgoing from $u$ nor incoming to $v$;
+  \item some sampled edges, but only either outgoing from $u$ or incoming to $v$; and
+  \item some sampled edges, both outgoing from $u$ and incoming to $v$;
+\end{enumerate}
+When the training set is build \uar{}, the vast majority of testing edges falls into the case (iii).
+This is no longer true with temporal training set, where situations (ii) and even (i) are more
+common, making prediction more difficult. The difference of repartition is particularly strong in
+\wik{}, justifying the better performance in \epi{}.
+
+\input{sampling}

### 2017-09-14 troll/introduction.tex
small clarification to the first batch algo
add two refs

@@
 opinions, the quality of debates is improved (other references can be found
-in~\autocite{journalism15}).
+in~\autocite{journalism15}). In a more extreme case, banning users with aggressive and hateful
+behavior has proved effective in the case of Reddit~\autocite{RedditBan15}.
 %
@@
 generative model. In this context, we design our first batch algorithm and show in
-\autoref{s:algbatch} that it provably approximates the Bayes classifier on dense graphs.
+\autoref{ss:bayes_approx} that it provably approximates the Bayes classifier on dense graphs.
 We furthermore gather empirical evidence supporting our probabilistic model by observing

### 2017-09-17 cc/bias_balance.tex
reorganize CC state of the art
rewrite related troll works

@@
 at weak balance:
-\begin{definition}[Weak balance]
+\begin{definition}[Weak balance]\label{def:weak_balance}
 	A triangle is weakly balanced if it has zero, two or three negative signs. Otherwise it is
@@
 Similar results also hold for weak balance.
-\begin{theorem}
+\begin{theorem}\label{thm:weakbal-consistent}
   A graph $G=(V,E)$ is weakly balanced if and only if $V$ admits a $K$-consistent clustering.

### 2017-09-17 troll/related.tex
add pointers to other sections in troll intro
rewrite related troll works

@@
 \section{Related work}
+\label{sec:troll_related}
 
 Interest in signed networks can be traced back to the psychological theory of structural
-balance~\autocites{Cartwright56}{HeiderBook58} with its weak
-version~\autocite{davis1967clustering}.  The advent of online signed social networks has enabled a
-more thorough and quantitative understanding of that phenomenon. Among the several approaches
-related to our work, some extend the spectral properties of a graph to the signed case in order to
-find good embeddings for classification~\autocites{Kunegis2009}{SignedEmbedding15}. However, the
+balance~\autocites{Cartwright56}{HeiderBook58} and its weak version~\autocite{davis1967clustering},
+that we will describe with more details in \autoref{sub:bias_balance} \vpageref{sub:bias_balance}.
+The advent of online signed social networks has enabled a
+more thorough and quantitative understanding of that phenomenon. In this section, we provide an
+overview of methods tackling the same \esp{} problem as us. Along the way, we give five of them a
+name in small capitals, for they are recent and effective. Therefore, we will compare our approaches
+with those methods in the experiments of the next section. At the end, we mention some closely
+related variants of the original \esp{} problem.
+
+\bigskip
+
+Existing methods can be broadly divided into three strategies, which share some similarities:
+\begin{enumerate}[1),nosep]
+	\item embedding the nodes of the graph in a low dimensional space using spectral or neural
+		techniques, before using node positions as features for a classifier;
+	\item completing the adjacency matrix through a global optimization algorithm; and
+	\item computing local features of the edges with several heuristics and train a classifier such as
+		logistic regression or SVM.
+\end{enumerate}
+
+In the first direction, the spectral embedding is illustrated by works from
+\textcites{Kunegis2009}{SignedEmbedding15}. We shall describe them more thoroughly in
+\autoref{par:cc_spectral}, but here we note that, following the natural orientation of the study of
+graph spectrum, they focus more on clustering than \esp{}. Furthermore, the
 use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which
-makes those methods hardly scalable to large graphs. Another approach is based on mining ego
-networks with SVM. Although this method seems to deliver good results~\autocite{Papaoikonomou2014},
-the running time makes it often impractical for large real-world datasets. An alternative approach,
-based on local features only and proposed in~\autocite{Leskovec2010}, relies on the so-called
-status theory for directed graphs~\autocite{guha2004propagation}. Some works in active learning,
-using a more sophisticated bias based on the correlation clustering (CC)
-index~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}, provide strong theoretical guarantees.
-However, the bias used there is rather strong, since it assumes the existence of a $2$-clustering
-of the nodes with a small CC index.\Todo{don't put it that way since we use that bias later}
+makes those methods hardly scalable to large graphs. More recently, there has been great interest in
+adapting word embedding techniques such as \textsf{word2vec}~\autocite{word2vec13} to a
+\enquote{corpus} of random walks that are considered as documents, while nodes play the role of
+word. This allows the unsupervised extraction of node features that can then used to train
+downstream classifier, see \autocite{nodeEmbeddingsurvey16} for a recent survey. It can also be
+tailored to exploit the specificity of signed graphs.
+The goal is to find for every node $u$ a vector $x_u\in\Rbb^d$ in such a way that in this new space,
+nodes are close to their positive neighbors and far from their negative neighbors. For instance in
+SiNE~\autocite{neuralSigned17}, for a node triplet $u,v,w$ such that the edge $(u,v)$ is positive
+and the edge $(u,w)$ is negative, the objective is to find vectors maximizing $f(x_u, x_w) + \delta
+- f(x_u, x_v)$. The similarity function $f$ is chosen to be a Siamese multilayers neural network,
+whose parameters are learned by back propagation. In SIGNet~\autocite{SIGNet17}, the similarity
+function is a sigmoid whose argument $x_u^Tx_v$ is weighted by the sign of the corresponding edge,
+with an optimization technique closer to \textsf{word2vec}. Finally, SNE~\autocite{SNE17} uses a
+log-bilinear model, and the training objective is to predict, given a path $\mathcal{P}$, which node
+should follow $\mathcal{P}$, weighting differently the vector nodes in $\mathcal{P}$ depending of
+whether they are the source of a positive or negative edge in $\mathcal{P}$.
 
-Whereas our focus will be on \emph{binary} prediction, researchers have also considered a weighted
-version of the problem, where edges measure the amount of trust or distrust between two
-users~\autocites{guha2004propagation}{Tang2013}{Bachi2012}{Qian2014sn}. Other works have also
+\medskip
+
+Next we look at matrix completion approaches, which are global by nature.
+For instance, \Textcite{LowRankCompletion14} tackle the \esp{} problem through this lens,
+restricting themselves to undirected graphs. They consider the
+observed adjacency matrix $A$, made of the edges in the training set \trainset{}, as a noisy
+sampling of the adjacency matrix $A^\star$ of an underlying complete graph satisfying the weakly
+balance condition (that is with no cycle containing only one negative edge, see
+\autoref{def:weak_balance} \vpageref{def:weak_balance}).
+This condition implies the existence of a small number $k$ of node clusters with positive links
+within clusters and negative links across clusters, which in turn implies $\rank A^\star=k$.  By
+recovering a complete matrix $\tilde{A}$ that matches the non-zeros entries of $A$, it is possible
+to predict the sign of $(u,v) \notin \trainset$ as $\yhat_{u,v} = \sgn(\tilde{A}_{u,v})$. Although
+the exact version of this problem is \NPh{}, the authors assume that $k$ is an hyperparameter known
+beforehand and look for two matrices $W,H\in \mathbb{R}^{k\times|V|}$ that minimise a sigmoidal
+reconstruction loss of $A$, subject to a nuclear norm regularization term. The minimization is
+carried out by Stochastic Gradient Descent and we refer to this method as \emph{\complowrank{}}.
+The approach of \textcite{OnlineCompletion17} is similar, but they consider directed graphs, use the
+logistic loss and compute at each iteration a threshold optimizing the $F_1$-score on the observed
+signs. Furthermore, they argue that to better handles class imbalance, it is preferable to minimize
+the maxnorm of the recovered matrix, which is a tighter approximation to the rank function than the
+nuclear norm. We thus refer to this method as \compmaxnorm{}.
+
+\medskip
+
+Finally, the last set of methods
+are based on the computation of local features of the graph. These features are
+evaluated on the subgraph induced by the training edges, and the resulting values are used to train
+a supervised classification algorithm. The most basic set of local
+features used to classify a given edge \euv{} are defined by
+$\din^+(v),\din^-(v),\dout^+(u),\dout^-(u)$ computed over the training set \trainset{}, and by the
+embeddedness coefficient $\big|\Nout(u) \cap \Nin(v)\big|$, which is the number of
+common neighbors of $u$ and $v$. In turn, these degree features can be used to define more
+complicated features, such as a notion of similarity between two nodes based on how they rate and
+are rated by their neighbors~\autocite{Yuan2017}.
+Another way of looking at neighborhoods and degrees is to mine ego networks\footnote{Recall that the
+ego network of a node $u$ is the subgraph induced by the neighbors $\nei(u)$ of $u$.} with a
+SVM~\autocite{Papaoikonomou2014}. \Textcite{Bachi2012} also use an approach based on ego networks,
+in a data mining fashion. Namely, they extract frequent small subgraphs from the collection of all
+ego networks of $G$. Then, they construct rules, which are made of two frequent subgraphs differing
+by a single edge.
+
+A sophisticated take on degree features is presented by \textcite{Bayesian15}, who note that a node
+can belong to one of the 16 types based of whether the number of its positive (respectively
+negative) outgoing (respectively incoming) edges is zero or not. The number of unobserved incoming
+and outgoing edges of each node $u$ let us define a 16-dimensional vector $V_u$ containing the
+probability of transitioning to any other type. Then for each edge \euv, they compute the outer
+product of $V_u$ with $V_v$ and also include additional features such as triads count and degree
+information before training a Logistic Regression model. We refer to this method as
+\emph{\compbayesian{}}.
+
+Other types of features are derived from social status theory, which posits that a positive link
+from $u$ to $v$ denotes that user $u$ considers user $v$ as having a higher status or skill than
+herself~\autocite{Leskovec2010}. This has implications on the distribution of the so called
+\emph{triads} in the network. A triad is a triangle formed by \euv{} together with $u \rightarrow w$
+(or $w \rightarrow u$) and $v \rightarrow w$ (or $w \rightarrow v$) for any $w \in \NNout(u) \cap
+\NNin(v)$. Taking signs and direction into account, there are 16 possible triads but according to
+the status theory some must be more represented than others. The \emph{\comptriads{}}
+method~\autocite{Leskovec2010} exploits this fact by counting for each edge in the training set how
+frequently it is involved in each of the 16 triad types. It also adds 7 degree features before
+training a Logistic Regression model.
+
+A third group of features is based on node ranking scores. These scores (usually one or two per
+node) are computed using a variety of methods, including
+\begin{itemize}
+	\item	PageTrust~\autocite{de2008pagetrust}, which adapts the random walk of
+		PageRank~\autocite{pagerank99}, by making walkers keep in memory a list of nodes they do not
+		like. If they ever reach such a node, their walk stop. The final PageTrust is computed
+		iteratively until convergence. 
+	\item Prestige~\autocite{zolfaghar2010mining}, which can be seen as a compounded degree feature,
+		as it is defined by $P(u) = \frac{\dinp(u)-\dinm(u)}{\din(u)}$.
+	\item	exponential ranking~\autocite{traag2010exponential}, which is the fixed point of the
+		equation $\pi = A^T\frac{\exp{\frac{1}{\mu}\pi}}{||\exp{\frac{1}{\mu}\pi}||_1}$. This follows
+		from the discrete choice theory and by assuming that we observe the reputation with some noise
+		that is double exponentially distributed with parameter $\mu$.
+	\item	Bias and Deserve~\autocite{mishra2011finding}, which are defined in terms of each other in
+		weighted graphs. The bias is the tendency of a node $u$ to trust/mistrust others, that is the
+		difference between its opinion of a neighbor $v$ and what $v$ truly deserve according to the
+		network: $bias(u) = \frac{1}{\dout(u)}\sum_{v\in\NNout(u)}(\wuv - deserve(v))$. The deserve of a
+		node $v$ is the aggregated opinion of its in-neighbors, discounted by their bias: $deserve(v) =
+		\frac{1}{\din(v)}\sum_{u\in\NNin(v)}\left(\wuv(1-\max\{0, \sgn(\wuv)\times bias(u)\})\right)$.
+	\item	Reputation and Optimism~\autocite{shahriari2014ranking}, defined for a node $u$ by
+		$\frac{\sum_{v \in \NNin(u)} y_{v,u}\sigma(v)}{\sum_{v \in \NNin(u)} \sigma(v)}$ and $
+		\frac{\sum_{v \in \NNout(u)} \yuv\sigma(v)}{\sum_{v \in \NNout(u)} \sigma(v)}$, where
+		$\sigma(v)$ is the ranking score assigned to node $v$. The former can be seen as a weighted
+		version of Prestige, while the latter is its outgoing counterpart.
+	\item	TrollTrust~\autocite{wu2016troll}, which builds upon \autocite{shahriari2014ranking} but
+		defined the ranking $\sigma(u)$ as the trustworthiness of $u$. It is follows from a recursive
+		definition of trollness based on the opinion of one node's neighbors weighted by their own
+		trollness, which allow to assign a $\sigma(u)$ to each node $u$ through a set of non linear
+		equations solved by an iterative method. These $\sigma$ values are used to compute Reputation
+		and Optimism scores, thus providing four features for each edge, which are in turn used to train
+		a Logistic Regression model for the classification task. We refer to this method as
+		\emph{\compranknodes{}}.
+% Computing $\pi$ requires to choose two hyperparameters of the model,
+% $\beta$ and $\lambda_1$ and the authors suggest to do it by holding out some training data to
+% perform cross validation but in practice we found little difference in performance in our
+% implementation while it consumes a lot of time.
+\end{itemize}
+
+\bigskip
+
+Other works have also\Todo{maybe that part would be better in the conclusion, especially since side
+information would help with the cold start problem showed in the experiments and trust version is
+one of the future direction.}
 considered versions of the problem where side information related to the network is available to
 the learning system. For instance, \autocite{EdgeSignsRating15} uses the product purchased on
-Epinion in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
+\epi{} in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
 analysing the textual content of their post, and~\autocite{SNTransfer13} uses SVM to perform
-transfer learning from one network to another. While many of these approaches have interesting
+transfer learning from one network to another.
+\autocite{Tang2013} uses a matrix completion approach, approximating $A$ by $PCP^T$, where the
+row $P_u$ of the matrix $P\in\Rbb^{n\times d}$ is the low rank representation of the node $u$ and
+$C\in\Rbb^{d\times d}$ is the correlation matrix between these representations. They also assume
+they are given a symmetric matrix $Z$ of the homophily coefficients $\zeta_{u,v}$ between $u$ and
+$v$ and add the following regularization term to be minimized: $\sum_{u<v} \zeta_{u,v}||P_u -
+P_v||_2^2 = Tr(U^T L_Z U)$, where $L_Z$ is the Laplacian of $Z$.
+While many of these approaches have interesting
 performances, they often require extra information which is not always available (or reliable) and,
@@
 
-The survey~\autocite{Tang2015a} contains pointers to many papers on edge sign prediction for
-signed networks, especially in the Data Mining area.
+\medskip
 
-The vast majority of existing edge sign prediction algorithms for
-directed graphs are based on the computation of local features of the graph. These features are
-evaluated on the subgraph induced by the training edges, and the resulting values are used to train
-a supervised classification algorithm (e.g., logistic regression). The most basic set of local
-features used to classify a given edge $(u,v)$ are defined by
-$\din^+(v),\din^-(v),\dout^+(u),\dout^-(u)$ computed over the training set \trainset{}, and by the
-embeddedness coefficient $\big|\Nout(u) \cap \Nin(v)\big|$. In turn, these can be used to define
-more complicated features, such as
-\(
-	\frac{\din^+(v) + |E^+|\uin(v)}{\din(v) + \uin(v)}
-\quad\text{and}\quad
-	\frac{\dout^+(u) + p^+\uout(u)}{\dout(u) + \uout(u)}
-\)
-introduced in~\autocite{Bayesian15}, together with their negative counterparts, where $|E^+|$ is the
-overall fraction of positive edges, and $\uin(v),\uout(u)$ are, respectively, the number of test
-edges outgoing from $u$ and the number of test edges incoming to $v$.
-Other types of features are derived from social status theory~(e.g., \autocite{Leskovec2010}), and
-involve the so-called triads; namely, the triangles formed by $(u,v)$ together with $(u,w)$ and
-$(w,v)$ for any $w \in \NNout(u) \cap \NNin(v)$. A third group of features is based on node ranking
-scores. These scores are computed using a variery of methods, including
-Prestige~\autocite{zolfaghar2010mining}, exponential ranking~\autocite{traag2010exponential},
-PageTrust~\autocite{de2008pagetrust}, Bias and Deserve~\autocite{mishra2011finding},
-TrollTrust~\autocite{wu2016troll}, and generalizations of PageRank and HITS to signed
-networks~\autocite{shahriari2014ranking}. Examples of features using such scores are \textsl{reputation}
-and \textsl{optimism}~\autocite{shahriari2014ranking}, defined for a node $u$ by
-\(
-	\frac{\sum_{v \in \NNin(u)} y_{v,u}\sigma(v)}{\sum_{v \in \NNin(u)} \sigma(v)}
-\quad\text{and}\quad
-	\frac{\sum_{v \in \NNout(u)} \yuv\sigma(v)}{\sum_{v \in \NNout(u)} \sigma(v)}\,,
-\)
-where $\sigma(v)$ is the ranking score assigned to node $v$. Some of these algorithms will be used
-as representative competitors in our experimental study of \autoref{s:exp}.
+Whereas our focus is on \emph{binary} prediction, researchers have also considered a weighted
+version of the problem, where edges measure the amount of trust or distrust between two users. Note
+that typically, the embedding methods we discussed at the beginning of this section are able to
+handle weighted networks.
+%
+One of the early and influential work on modeling how distrust propagate among users is
+\autocite{guha2004propagation}. They propose to represent atomic conclusions (such as if $u$ trusts
+$v$ and $v$ trusts $w$ then $u$ is likely to trusts $w$) as matrix operators and define four of them
+that are assembled by a weighted linear combination. Starting from the observed adjacency matrix
+$A$, they repeatedly apply this operator (with potentially a discount factor) to obtain a final
+weigh matrix $F$ (that can also be rounded to provide sign prediction).
+%
+\Textcite{Qian2014sn} extend the binary case with categorical relationships (such as strongly
+positive or weakly negative) and describe how every unbalanced triads experience some stress
+depending on the strength of its contradictory relationships. Arguing that the network should
+converge to a balanced state with as little change as possible, they express finding this minimal
+transformation as a Multidimensional Scaling problem, and the resulting graph to characterize
+unlabeled edges.
+%
+Finally, \textcite{edgeWeights16} explicitly consider the weight prediction problem in signed graph.
+They use a procedure similar to the ranking methods discussed above. Namely, they defined two scores
+for each node that are computed iteratively from a uniform initialization. Assuming that weights lie
+in $[-1,1]$, the fairness $f(u)$ of $u$ is a measure of how fair or reliable $u$ is when rating
+others nodes, while the goodness $g(v)$ of $v$ measure how trustworthy is $v$ when evaluated with
+complete fairness.  Formally, $f(u) =  1 - \frac{1}{2\NNout(u)} \sum_{v\in\NNout(u)}
+\frac{|\wuv-g(v)|}{2}$ and $g(v) = \frac{1}{\NNin(v)}\sum_{u\in\NNin(v)}f(u)\wuv$. Note that these
+definitions are very close to bias and deserve from \textcite{mishra2011finding}, but the absolute
+value in the factor $2$ in the expression of fairness allows for better convergence property.
+Finally, the weight of a test edge \euv is predicted as $f(u)\times g(v)$.
 
-\iffalse
-We compare these methods with four competitors, each exploiting the sign information and network
-topology in a different way:
+\medskip
 
-Looking at undirected graphs with a global perspective, \textcite{LowRankCompletion14} consider the
-observed adjacency matrix $A$, made of the edges in \trainset{}, as a noisy sampling of the
-adjacency matrix $A^\star$ of an underlying complete graph satisfying the weakly balance condition
-(that is with no triangle containing only one negative edge).
-%TODO the previous sentence is too long
-This condition implies the existence of a small number $k$ of node clusters with positive links
-within clusters and negative links across clusters, which in turn implies $\rank A^\star=k$.  By
-recovering a complete matrix $\tilde{A}$ which matches the non-zeros entries of $A$, it is possible
-to predict the sign of $\eij \notin \trainset$ as $\yhat_{i,j} = \sgn(\tilde{A}_{i,j})$. Although
-the exact version of this problem is NP-hard, authors assume that $k$ is an hyperparameter known
-beforehand\footnote{The method is showed to be robust to the choice of its value, thus we pick
-$k=7$.} and look for two matrices $W,H\in \mathbb{R}^{k\times|V|}$ which minimise a sigmoidal
-reconstruction loss with $A$, subject to a regularization term. We refer to this method as
-\emph{\complowrank{}}.
+While we presented methods that operate in the batch setting, and will present an online algorithm
+as well in \autoref{s:algsonline}, other works have attacked the \esp{} problem from an active
+learning point of view~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}.  Recall that in the active
+setting, we are given a budget of edge labels to observe, and are free to select them the way we
+want within $E(Y)$. Again, the goal is to make as few mistakes as possible when predicting the sign
+of the remaining edges. These two methods build a spanning subgraph of $G$ and query all its edge,
+but they differ in its the construction. \Textcite{Cesa-Bianchi2012a} partition $G$ into stars and
+connect them in a tree, whereas \textcite{Cesa-Bianchi2012b} cover the graph with cycles, each
+containing one test edge and being queried for the other, while respecting the user specified query
+budget.  This relies on a different bias than our generative model, and this will be the subject of
+our next chapter.  One might also consider an online version of the problem where the topology is
+not known in advance but discovered as prediction are made~\autocite{Gentile2013}. This naturally
+increases and the difficulty of the problem, as reflected by the computational cost of the solution
+proposed, which is quadratic in $|V|$ \emph{at each prediction}.
 
-The status theory heuristic posits that a positive link from $i$ to $j$ denotes that user $i$
-considers user $j$ as having a higher status or skill than himself~\autocite{Leskovec2010}. It
-implies that among the 16 possible triads of three nodes with two directed signed edges among them,
-some must be more represented than others. The \emph{\comptriads{}} method exploits this fact by
-counting for each edge in the training set how frequently it is involved in each of the 16 triad
-types. Based on these edge features and 7 degree features of each edge endpoints, a Logistic
-Regression model is trained and used to predict the sign of the test edges.
-
-Inspired by the PageRank algorithm, \textcite{wu2016troll} propose a recursive definition of
-trollness based on the opinion of one node's neighbors weighted by their own trollness, which allow
-to assign a trustworthiness $\pi(i)$ to each node $i$ through a set of non linear equations solved
-by an iterative method. This $\pi$ values are used to rank the nodes by computing their Reputation
-and Optimism scores, thus providing four features for each edge, which are in turn used to train a
-Logistic Regression model for the classification task. We refer to this method as
-\emph{\compranknodes{}}.  Computing $\pi$ requires to choose two hyperparameters of the model,
-$\beta$ and $\lambda_1$ and the authors suggest to do it by holding out some training data to
-perform cross validation but in practice we found little difference in performance in our
-implementation while it consumes a lot of time.
-
-The last competitors is also local. \textcite{Bayesian15} note that  a node can belong to one of the
-16 types based of whether is number of its positive (resp.  negative) outgoing (resp. incoming)
-edges is zero or not.  The number of unobserved incoming and outgoing edges of each node $i$ let us
-define its 16-dimensional vector $V_i$ containing the probability of transitioning to any other
-state.  Then for each edge \eij we compute the outer product of $V_i$ with $V_j$ and also include
-additional features such as triads count and degree information before training a Logistic
-Regression model. We refer to this method as \emph{\compbayesian{}}.
-\fi
+The survey~\autocite{Tang2015a} contains pointers to many papers on signed networks, in particular
+for the \esp{} problem.

### 2017-09-19 troll/conclusions.tex
write troll conclusion

@@
-\section{Conclusions and Ongoing Research}
-%
-%FV
-We have studied the edge sign prediction problem in directed graphs in both batch and online learning settings.
-In both cases, the underlying modeling assumption hinges on the trollness and (un)trustworthiness predictive features.
-We have introduced a simple generative model for the edge labels
-%FV that serves both as a theoretical justification for many successful heuristic methods to edge sign prediction and 
-%FV as a way 
-to craft this problem as a node sign prediction problem to be efficiently tackled by standard Label Propagation algorithms. 
-%FV We then observed that the two features used
-%FV notion of trollness and trustworthiness
-%FV naturally lend themselves to define a notion of edge sign regularity that can also be analyzed in an (adversarial) online setting. In this setting, 
-%FV 
-Furthermore, we have studied the problem in an (adversarial) online setting providing 
-upper and (almost matching) lower bounds on the expected number of prediction mistakes.
+\section{Conclusion}
 
-%FV Finally, we complemented the above with experiments on five real-world datasets in the small training set regime.
+In this chapter, we started our characterization of edges in complex networks by studying the \esp{}
+problem in \dssn{}. As mentioned in the introduction, given a network topology and a few labeled
+interactions, being able to infer the polarity of the remaining interactions is valuable to improve
+user experience. Our main insight is that two features of users behavior, trollness and
+(un)trustworthiness, are key to understand their interactions. This leads us to design a simple sign
+generative model. Such a model serves both as a theoretical justification for many successful
+heuristics and as the underpinning of our \usrule{} algorithm. This algorithm is fast, trivially
+parallelizable, provably close to the Bayes optimal predictor on dense graphs and experimentally
+accurate even on sparse graphs. We further exploit this model in the batch setting through an
+approximation of the maximum likelihood estimation procedure. While making this approximation, we
+moreover cast the problem as a node prediction problem. This can be efficiently tackled by standard
+Label Propagation algorithms.
 
-%FV 
-Finally, we validated our theoretical results by experimentally assessing our methods on five real-world datasets in the small training set regime. 
-Two interesting conclusions from our experiments are: i.
-Our generative model is robust, for it produces Bayes optimal predictors which tend to be empirically best also within the larger set of models that includes all logistic regressors based on trollness and trustworthiness alone; ii.
-%We showed that 
-our methods are in practice either strictly better than their competitors in terms of prediction quality or, when they are not, they are faster.
-We are currently engaged in 
-%FV gathering more experimental evidence, as well as in 
-extending our approach so as to incorporate further predictive features (e.g., side information, when available).
+We validate our theoretical results by experimentally assessing these two methods in the small
+training set regime, on synthetic and real-world datasets. We draw two main conclusions from our
+experiments. First, our generative model is robust across domains. Indeed, it produces predictors
+that are empirically both close to the Bayes optimal and to general linear models trained on
+trollness and trustworthiness features. This is notable as this result is achieved by simply
+counting negative edges, without resorting to Stochastic Gradient Descent methods typically involved
+in training large scale linear models. Second, our methods are in practice either strictly better
+than their competitors in terms of prediction quality or, when they are not, they are faster.
+
+Finally, we study the online adversarial setting, where trollness and (un)trustworthiness naturally
+lend themselves to define a notion of edge sign regularity. Based on this regularity, we provide an
+upper and an (almost matching) lower bounds on the expected number of prediction mistakes. Before
+exploring in the next chapter another learning bias for \esp{}, which is not based on trollness and
+(un)trustworthiness, we present directions in which this work could be extended.
+
+\bigskip
+
+Given the fundamental role of our generative model, we would like to adapt it in order to handle
+the three problem extensions presented in the related work of \autoref{sec:troll_related}.
+
+\begin{enumerate}[1.,]
+
+  \item When available, using \textbf{side information} about users is an alluring option to
+    alleviate the cold start problem. Recall that in our experiments, we saw that if we learn our
+    model on historical data, we cannot successfully predict the interactions involving new users,
+    for their trollness and trustworthiness are not yet known to us. However, we expect that for
+    existing users, side information is correlated with their value of $p$ and $q$. This knowledge
+    thus provides a parameterised prior for the $p$ and $q$ values of news users. As we observe
+    their behavior in the network, our estimation of $p$ and $q$ would rely less and less on prior
+    information and more and more on observed data, as it is common in a Bayesian approach. Note
+    that this introduces an additional learning stage.
+
+  \item Our generative model, and in particular the prior distribution $\mu(p,q)$, could be
+    exploited to guide label queries in an adaptive \textbf{active learning} setting. Given a budget
+    of queries $B$, and after an initial phase of querying edge labels \uar{}, one might use
+    concentration results on $\mu$\footnote{As done in the proof of \autoref{t:active},
+    \vpageref{ssec:troll_proof_bayes}.} to devise an informativeness criterion, indicating for which
+    node the value of $p$ and $q$ are the most uncertain. Combined with the graph topology, this
+    would suggest which edge is the most important to query next, until the budget is exhausted.
+
+  \item So far we discussed binary classification of edges. In \textbf{weighted graphs} though, a
+    refined problem would be to predict the weight of unlabeled edges. Let us assume for simplicity
+    that the weights are bounded and within $[-1, 1]$. The most immediate solution would be, for an
+    edge \euv{}, to linearly shift our estimation of $\frac{p_u+q_v}{2}=\etauv{}$ from $[0,1]$ to
+    $[-1,1]$ and use this quantity as our prediction. We could also change the semantic of \etauv{}
+    from being the probability of \euv{} to be positive to the mean of a narrow Gaussian from which
+    the weight of \euv{} is drawn. Finally, by letting $p_u$ range from $-1$ to $+1$, we note a
+    similarity with fairness and goodness, suggesting $p_u\times q_v$ could approximate the weights.
+\end{enumerate}

### 2017-09-19 troll/algs_batch.tex
mark content added from the aistats paper

@@
 
+\begin{newcontent}
 Given $G(Y) =(V,E(Y))$, in the batch setting we have at our disposal a training set \trainset{} of
@@
 parameters through a global label propagation approach, without making any density assumption.
+\end{newcontent}
 
@@
 
+\begin{newcontent}
 While we defer the full proof of \autoref{t:active} to the additional material at the end of this
@@
 density limitation, we now introduce a second method based on label propagation.
+\end{newcontent}
 

### 2017-09-19 troll/experiments.tex
mark content added from the aistats paper

@@
 task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
+\begin{newcontent}
 In \adv{}, a trust-based social network for open source developer, a user $u$ can certify another
@@
 4]{AdvogatoTrustThesis02}).
+\end{newcontent}
 In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
@@
 
+\begin{newcontent}
 Finally, there is a low fraction of reciprocal (or \emph{twin}) edges (\ie{} both $\euv \in E$ and $\evu
@@
 our methods with existing approaches, we not use that trick. But afterwards, we show in
-\autoref{tab:all_mcc_pp} when it can be beneficial.
+\autoref{tab:twin} when it can be beneficial.
 % more references in  http://www.nature.com/articles/srep02729
+\end{newcontent}
 
@@
 
-\input{all_mcc}
-
 \paragraph{Results}
@@
 
+\input{all_mcc}
+
+\begin{newcontent}
 We perform two other experiments. The first one is to assess the effect of predicting reciprocal
@@
 will be connected in the future~\autocite{linkPredSurvey16}. However, as showed in
-\input{ts_mcc}
 \autoref{tab:troll_ts_mcc}, this makes the problem much more challenging, both for our methods and
 the existing ones. It is quite surprising that even when the training set is as large as 90\% of all
-edges, the MCC are so low compared with those reported \autoref{tab:all_mcc}. The explanation of
+edges, the MCC are so low compared with those reported \autoref{tab:all_mcc}.
+
+\input{ts_mcc}
+
+The explanation of
 why these two sampling strategies produce so different results can be inferred from
@@
 \input{sampling}
+
+\end{newcontent}

### 2017-09-19 troll/related.tex
mark content added from the aistats paper

@@
 
+\begin{newcontent}
 Interest in signed networks can be traced back to the psychological theory of structural
@@
 While we presented methods that operate in the batch setting, and will present an online algorithm
-as well in \autoref{s:algsonline}, other works have attacked the \esp{} problem from an active
+as well in \autoref{s:algonline}, other works have attacked the \esp{} problem from an active
 learning point of view~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}.  Recall that in the active
@@
 proposed, which is quadratic in $|V|$ \emph{at each prediction}.
+\end{newcontent}
 

### 2017-09-19 troll/experiments.tex
typos

@@
 We now evaluate our edge sign classification methods on representative real-world datasets of
-varying density and label regularity. This shows showing that our methods compete well against
+varying density and label regularity. This shows that our methods compete well against
 existing approaches in terms of both predictive and computational performance. We are especially
@@
 
-We considered six real-world classification datasets. The first four are \dssn{} widely
+We consider six real-world classification datasets. The first four are \dssn{} widely
 used as benchmarks for this
@@
 \begin{newcontent}
-In \adv{}, a trust-based social network for open source developer, a user $u$ can certify another
+In \adv{}, a trust-based social network for open source developers, a user $u$ can certify another
 user $v$ with different degrees of trust: \enquote{Observer}, \enquote{Apprentice} (both of which we
@@
 \url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric and its
-resistance to attacks is available in the PhD thesis of the website's creator~\autocite[Section
+resistance to attacks are available in the PhD thesis of the website's creator~\autocite[Section
 4]{AdvogatoTrustThesis02}).
@@
 instead evaluated our performance using the so-called Matthews Correlation Coefficient
-(MCC)~\autocite{MCC00}), defined as \Todo{For completeness, we could also report accuracy results in
-the supplementary material.}  
+(MCC)~\autocite{MCC00}), defined as
 \[
@@
 \in E$), which is a common mechanism of link formation in directed
-networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases such twin edges do not disagree,
-\ie{} they have the same sign. In practice, we can use that fact to improve our accuracy at no
+networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases, such twin edges do not disagree,
+\ie{} they have the same sign. In practice, we can use this fact to improve our accuracy at no
 additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
-\evu{} is part of the training set, we set $\sgn(\euv) = \sgn(\evu)$. For clarity, when comparing
+\evu{} is part of the training set, we set $\yhat_{u,v} = y_{v,u}$. For clarity, when comparing
 our methods with existing approaches, we not use that trick. But afterwards, we show in
@@
 \begin{newcontent}
-We perform two other experiments. The first one is to assess the effect of predicting reciprocal
+We perform two other sets of experiments. The first one assesses the effect of predicting reciprocal
 edges by their value if available. As expected, the results in \autoref{tab:twin} demonstrate that
-it improves MCC when the network has enough twin edges with low disagreements, like \sla{} and
+it improves MCC when the network has enough twin edges with low disagreements, like in \sla{} and
 \epi{}. It has no effect when there are few twin edges, like in \aut{}, \wik{} and \kiw{}. And it is
-detrimental if when there are many twin edges that do  not agree enough, like in \adv{}, in which
+detrimental when there are many twin edges that do not agree enough, like in \adv{}, in which
 case it is better to rely solely on the learned model.
@@
 
-In the second, we study the effect of our hypothesis that the labeled edges of the training set are
-chosen \uar{}. In two of our datasets (\wik{} and \epi{}), edges come the timestamp at which they
+In the second set of experiments, we study the effect of our hypothesis that the labeled edges of the training set are
+chosen \uar{}. In two of our datasets (\wik{} and \epi{}), edges come with the timestamp at which they
 were created. A more realistic way of choosing the training set is therefore, for a given training
-size $m$, to let $E_0$ be the $m$ oldest edges, and try to predict the remaining, newest one. This
+size $m$, to let $E_0$ be the $m$ oldest edges, and try to predict the remaining, newest ones. This
 is a common experimental setting in link prediction, where the goal is to infer which pair of nodes
@@
 \autoref{tab:troll_ts_mcc}, this makes the problem much more challenging, both for our methods and
-the existing ones. It is quite surprising that even when the training set is as large as 90\% of all
-edges, the MCC are so low compared with those reported \autoref{tab:all_mcc}.
+out competitors. It is quite surprising that even when the training set is as large as 90\% of all
+edges, the MCC are so low compared with those reported in \autoref{tab:all_mcc}.
 
@@
 least one outgoing edge, and defining similarly $V_{\mathrm{in}}$, we see that the node coverage is
-naturally larger when sampling at random. Yet when the training set reach 90\% size, with both
-sampling strategies we cover roughly 90\% of the nodes in $V_{\mathrm{out}}$ and in
+naturally larger when sampling at random. Yet, when the training set reaches 90\% size, both
+sampling strategies cover roughly 90\% of the nodes in $V_{\mathrm{out}}$ and in
 $V_{\mathrm{out}}$. The difference in MCC is thus explained by the last three columns. It shows, for

### 2017-09-19 troll/experiments.tex
add a new set of troll experiments

@@
 We now evaluate our edge sign classification methods on representative real-world datasets of
-varying density and label regularity. This shows that our methods compete well against
+varying density and label regularity.
+After presenting the data and our evaluation criterion, we proceed in two steps. First, we simulate
+our generative model on real networks to give them signs, then we study to which extent we can recover
+the parameters $p$ and $q$ of each node, and how this translates to accurate prediction. Second, we
+select random training sets from the actual signs.
+This shows that our methods compete well against
 existing approaches in terms of both predictive and computational performance. We are especially
@@
 \end{newcontent}
+
+\begin{newcontent}
+\paragraph{Synthetic signs}
+
+Recall that according to our generative model of \autoref{s:gen}, each node $u$ has parameter $p_u$
+governing its sending behavior and each node $v$ has a parameter $q_v$ governing its receiving
+behavior such that the sign of the edge \euv{} is positive with probability $\frac{p_u+q_v}{2} =
+\etauv$. Given a topology $G=(V,E)$, we start by assigning a $p$ and $q$ values to each node. If we
+want \etauv{} to be uniform over $E$, we have to take into account the out-degree of nodes with at
+least one outgoing edges, and likewise the in-degree of nodes with at least one incoming edges. For
+that, we partition the interval $[0,1]$ into $|V|$ segments of size proportional to $\dout(u_1),
+\dout(u_2), \ldots, \dout(u_{|V|})$. We shuffle these segments and draw a number \uar{} from each of
+them, which we set as the $p$ value of the corresponding node. Then, because we want to have more
+positive edges, we apply an exponential transformation \[ p \rightarrow \frac{1}{1-e^{-\lambda}}
+\left( 1-e^{-\lambda p} \right) \,,\] where we choose $\lambda=3$. We do the same for $q$, then for
+every edge \euv{}, we set it positive with probability \etauv{}. An example of the distributions we
+obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}.
+
+Once the signs are generated, we select a training set \trainset{} \uar{}, and we want to predict
+the signs of the remaining edges in the testing set $E\setminus \trainset{}$. The Bayes optimal
+predictor classifies an edge \euv{} as positive if $\etauv \geq \shalf$ and negative otherwise. 
+We compare it with the algorithm analyzed in \autoref{ss:bayes_approx}, which we call \usrule{}
+(Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). In practice,
+\usrule{} proceeds as follow: after computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all
+$u \in V$ (or setting those values to \shalf{} in case there is no outgoing or incoming edges for
+some node), we use the equation \eqref{eq:predictor}, having estimated $\tau$ on $E_0$.
+
+The results of this comparison on the six datasets are showed in \autoref{tab:troll_synth_global}.
+For each network, we generated the signs once. Then for different training size (20\%, 40\% or 80\%
+of $E$), we sampled a training set and predict using the knowledge of the true $\eta$ (Bayes
+predictor) or its estimated value (\usrule{}).
+The MCC of the Bayes predictor is the same on all datasets (around $30$), and so is its accuracy
+(around $75$). The corresponding values for \usrule{} are close on all datasets (although some are
+easier) and the gap naturally decreases as the training size increases. Another interesting
+quantity is how close can \usrule{} estimates $\eta$ using $\wh{\eta} = \left(1-\htr(u)\right) +
+\left(1-\hun(v)\right) -\tauhat$ as defined in \autoref{ss:bayes_approx}. We compute the mean
+absolute error (MAE) between $\eta$ and $\wh{\eta}$ for all testing edges, and specifically for
+the testing edges whose both endpoints have been sampled above a certain threshold in the training
+set. As expected, the estimation is more accurate with increasing training set size, and with
+increasing number of samples for a given edge.
+
+\input{synth_res}
+
+Finally, we can also look in more details at the role of $p$ and $q$ by building a 2D histogram of
+the testing edges \euv{} based on their coordinate $(p_u, q_v)$. For instance on
+\autoref{fig:tr_synth_mae}, we see that the MAE between $\eta$ and $\wh{\eta}$ is not uniform over
+the $p,q$-space. It seems to decrease along the diagonal from $(0,0)$ to $(1,1)$. Note though that
+\autoref{fig:tr_synth_sizemae} suggests this is likely due to the higher number of edges in the top
+right corner (reflecting the imbalance toward positive edges), which allows better estimation.
+
+In this case, we cannot use the MCC because off-diagonal, the Bayes predictor classifies edges
+either all positive or all negative, resulting in a division by zero in the definition. We thus
+fall back on accuracy and display in \autoref{fig:tr_synth_acc} the difference between the accuracy
+of the Bayes predictor and the accuracy of \usrule{}. This time, the anti diagonal where
+$\frac{p+q}{2}=\lhalf$ seems to play a special role. Indeed, as showed more clearly in
+\autoref{fig:tr_synth_cmpacc}, the gap in accuracy between the Bayes predictor and \usrule{}
+increases symmetrically as $\frac{p+q}{2}$ moves away from \shalf{}. The gap also tends to disappear
+when $\frac{p+q}{2}=\lhalf$, most likely because in that area, prediction can only be made by
+chance.
+
+While we show those patterns on \wik{} with a 40\% training size, they are consistent across
+datasets and training size.
+
+\end{newcontent}
+
+\begin{figure*}[hbtp]
+  % \centering
+  \begin{subfigure}[b]{0.504\textwidth} 
+    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_mae.pdf} 
+    \caption{Mean absolute error (MAE) of the \etauv{} estimation.} 
+    \label{fig:tr_synth_mae}
+  \end{subfigure}~
+  \begin{subfigure}[b]{0.504\textwidth} 
+    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_sizemae.pdf} 
+    \caption{MAE in each cell as a function of the number of sampled edges.} 
+    \label{fig:tr_synth_sizemae}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.504\textwidth} 
+    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_acc.pdf} 
+    \caption{100 times the accuracy.} 
+    \label{fig:tr_synth_acc}
+  \end{subfigure}~
+  \begin{subfigure}[b]{0.504\textwidth} 
+    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_cmpacc.pdf} 
+    \caption{Difference of accuracy between \usrule{} and Bayes as $\frac{p+q}{2}$ moves away from $\frac{1}{2}$.} 
+    \label{fig:tr_synth_cmpacc}
+  \end{subfigure}
+  \caption{Results on \wik{} with \usrule{} on a training set of size 40\%}
+\end{figure*}
 

### 2017-09-19 troll/experiments.tex
minor reorganization

@@
 
-\paragraph{Datasets}
+\subsection{Datasets}
 
@@
 \begin{newcontent}
-\paragraph{Synthetic signs}
+\subsection{Synthetic signs}
 
@@
 
-\paragraph{Algorithms and parameter tuning.} 
+\subsection{Real signs} 
 We compared the following algorithms:
@@
 
-  \item The algorithm analyzed \autoref{ss:bayes_approx}, which we call \usrule{}
-    (Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). After
-    computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all $u \in V$ (or setting those
-    values to $\frac{1}{2}$ in case there is no outgoing or incoming edges for some node), we use
-    equation \eqref{eq:predictor} and estimate $\tau$ on $E_0$.
+  \item The \usrule{} algorithm, as described above.
 
   \item A logistic regression model where each edge $(u,v)$ is associated with the features
-    $[1-\htr(u), 1-\hun(v)]$ computed again on \trainset{} (we call this method \uslogregp{}). Best
+    $[1-\htr(u), 1-\hun(v)]$ computed again on \trainset{} (we call this method \uslogregp{}). The best
     binary thresholding is again computed on \trainset{}. Experimenting with this logistic model
@@
 
+  \item The other \compmaxnorm{} matrix completion method from~\autocite{OnlineCompletion17},
+    setting the parameter $\lambda$ to $1.2$ as advised in their paper.
+
   \item A logistic regression model built on \comptriads{} features derived from status
@@
 
-  \item The PageRank-inspired algorithm from \autocite{wu2016troll}, where a recursive notion of
-    trollness is computed by solving a suitable set of nonlinear equations through an iterative
-    method, and then used to assign ranking scores to nodes, from which (un)trustworthiness features
-    are finally extracted for each edge. We call this method \compranknodes{}. As for hyperparameter
-    tuning ($\beta$ and $\lambda_1$ in~\autocite{wu2016troll}), we closely followed the authors'
-    suggestion of doing cross validation.
+  \item The TrollTrust algorithm from~\autocite{wu2016troll}, naming it \compranknodes{}. As for
+    hyperparameter tuning ($\beta$ and $\lambda_1$ in~\autocite{wu2016troll}), we closely followed
+    the authors' suggestion of doing cross validation.
 
@@
 \comptriads{}, and \compbayesian{} essentially fall into this category. The remaining methods
-(\uslpropGsec{}, \complowrank{}, and \compranknodes{}) are global in that their features are
+(\uslpropGsec{}, \complowrank{}, \compmaxnorm{} and \compranknodes{}) are global in that their
+features are
 designed to depend on global properties of the graph topology.
@@
 
-\paragraph{Results}
+\bigskip
+
 Our main results are summarized in \autoref{tab:all_mcc}, reporting MCC test set performance after
@@
 \begin{newcontent}
+\subsection{Additional experiments}
+
 We perform two other sets of experiments. The first one assesses the effect of predicting reciprocal

### 2017-09-20 troll/diff_mcc_pp.tex
mark content added from the aistats paper
minor corrections in troll experiments and add missing figures

@@
-\begin{table}[htpb]
-  \centering
-  \caption{100 time the difference of MCC between using the twin heuristic and not using it for our
+% \begin{table}[htpb]
+%   \centering
+%   \caption{
+\begin{center}
+   \captionsetup{font=small}
+   \captionof{table}{%
+    100 time the difference of MCC between using the twin heuristic and not using it for our
   two main methods. Those values are computed on the same training/testing split as of
@@
     \midrule
-    \multicolumn{8}{c}{\aut{}} \\
+    & \multicolumn{7}{c}{\aut{}} \\ \cmidrule(lr){2-8}
     \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.7$}  \\
     \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Green}{$0.1$} & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.7$}  \\
-    \multicolumn{8}{c}{\adv{}} \\
+    & \multicolumn{7}{c}{\adv{}} \\ \cmidrule(lr){2-8}
     \uslpropGsecTwin{}          & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.6$}  & \textcolor{Red}{$-1.0$}  & \textcolor{Red}{$-1.6$}  & \textcolor{Red}{$-2.2$}  & \textcolor{Red}{$-5.4$}  & \textcolor{Red}{$-10.8$} \\
     \usruleTwin{}               & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.5$}  & \textcolor{Red}{$-1.0$}  & \textcolor{Red}{$-1.6$}  & \textcolor{Red}{$-2.2$}  & \textcolor{Red}{$-5.7$}  & \textcolor{Red}{$-11.1$} \\
-    \multicolumn{8}{c}{\wik{}} \\
+    & \multicolumn{7}{c}{\wik{}} \\ \cmidrule(lr){2-8}
     \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.2$}  \\
     \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.2$}  \\
-    \multicolumn{8}{c}{\sla{}} \\
+    & \multicolumn{7}{c}{\sla{}} \\ \cmidrule(lr){2-8}
     \uslpropGsecTwin{}          & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.5$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$0.9$} & \textcolor{Green}{$1.1$} & \textcolor{Green}{$2.0$} & \textcolor{Green}{$3.3$} \\
     \usruleTwin{}               & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.6$} & \textcolor{Green}{$0.9$} & \textcolor{Green}{$1.1$} & \textcolor{Green}{$1.3$} & \textcolor{Green}{$2.2$} & \textcolor{Green}{$3.5$} \\
-    \multicolumn{8}{c}{\epi{}} \\
+    & \multicolumn{7}{c}{\epi{}} \\ \cmidrule(lr){2-8}
     \uslpropGsecTwin{}          & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.5$} & \textcolor{Green}{$0.7$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$1.0$} & \textcolor{Green}{$1.9$} & \textcolor{Green}{$3.0$} \\
     \usruleTwin{}               & \textcolor{Green}{$0.4$} & \textcolor{Green}{$0.6$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$1.0$} & \textcolor{Green}{$1.2$} & \textcolor{Green}{$1.8$} & \textcolor{Green}{$2.8$} \\
-    \multicolumn{8}{c}{\kiw{}} \\
+    & \multicolumn{7}{c}{\kiw{}} \\ \cmidrule(lr){2-8}
     \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.3$}  \\
@@
   \end{tabular}
-\end{table}
+% \end{table}
+\end{center}

### 2017-09-20 troll/experiments.tex
minor corrections in troll experiments and add missing figures

@@
 
-We now evaluate our edge sign classification methods on representative real-world datasets of
+We now evaluate our \esp{} methods on representative real-world datasets of
 varying density and label regularity.
@@
 our generative model on real networks to give them signs, then we study to which extent we can recover
-the parameters $p$ and $q$ of each node, and how this translates to accurate prediction. Second, we
-select random training sets from the actual signs.
+the parameters $p$ and $q$ of each node, and how predictions based on these estimation compare with
+the Bayes optimal. Second, we select random training sets from the actual signs.
 This shows that our methods compete well against
@@
 positive).\footnote{We download the 7\thup{} of July, 2014 version from
-\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric and its
-resistance to attacks are available in the PhD thesis of the website's creator~\autocite[Section
+\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric, and its
+resistance to attacks, is available in the PhD thesis of the website's creator~\autocite[Section
 4]{AdvogatoTrustThesis02}).
@@
 or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
-simple, yet powerful, keyword-based technique using a list of positive and negative words. See
+technique using a list of positive and negative words; see
 \cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
@@
 
-\begin{table*}[bt]
+\begin{table}[bt]
   \centering
@@
     \toprule
-    Dataset & $|V|$       & $|E|$       & $\frac{|E|}{|V|}$ & $\frac{|E^+|}{|E|}$ & $\frac{\Psi^2_{G''}(Y)}{|E|}$ & $\frac{\Psi_G(Y)}{|E|}$ & twin edges & twin disagreement \\
+    Dataset & \thead{$|V|$}       & \thead{$|E|$}       & \thead{$\frac{|E|}{|V|}$} &
+    \thead{$\frac{|E^+|}{|E|}$} & \thead{$\frac{\Psi^2_{G''}(Y)}{|E|}$} &
+    \thead{$\frac{\Psi_G(Y)}{|E|}$} & \thead{reciprocal\\ edges} &
+    \thead{reciprocal\\ disagreement} \\
     \midrule                                                                                                   
@@
   \end{tabular}
-\end{table*}
+\end{table}
 
@@
 instead evaluated our performance using the so-called Matthews Correlation Coefficient
-(MCC)~\autocite{MCC00}), defined as
+(MCC)~\autocite{MCC00}, defined as
 \[
 	\mathrm{MCC} = \frac{tp\times tn-fp\times fn}%
-        {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } } = \pm \sqrt{\frac{\chi^2}{n}}
+        {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } }% = \pm \sqrt{\frac{\chi^2}{n}}
 \]
@@
 Although the semantics of the edge signs is not the same across these networks, we can see from
-\autoref{tab:dataset} that our generative model essentially fits all of them. Specifically, the last
+\autoref{tab:dataset} that our generative model essentially fits all of them. Specifically,
 two columns of the table report the rate of label (ir)regularity, as measured by
-$\Psi^2_{G''}(Y)/|E|$ (second-last column) and $\Psi_{G}(Y)/|E|$ (last column), where 
+$\Psi^2_{G''}(Y)/|E|$ (6\thup{} column) and $\Psi_{G}(Y)/|E|$ (7\thup{} column), where 
 \[
@@
 $f_{E_0}$ and $f_{E\setminus E_0}$ being the quadratic criterions of \autoref{ss:passive}, viewed as
-functions of both $(\bp,\bq)$, and $y_{i,j}$ when all labels are known, and $\Psi_{G}(Y)$ is the
+functions of both $(\bp,\bq)$, and \yuv{} when all labels are known, and $\Psi_{G}(Y)$ is the
 label regularity measure adopted in the online setting, as defined in \autoref{s:prel}. It is
@@
 \begin{newcontent}
-Finally, there is a low fraction of reciprocal (or \emph{twin}) edges (\ie{} both $\euv \in E$ and $\evu
+Finally, there is a low fraction of reciprocal edges (\ie{} both $\euv \in E$ and $\evu
 \in E$), which is a common mechanism of link formation in directed
-networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases, such twin edges do not disagree,
+networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases, such reciprocal edges do not disagree,
 \ie{} they have the same sign. In practice, we can use this fact to improve our accuracy at no
@@
 
-Recall that according to our generative model of \autoref{s:gen}, each node $u$ has parameter $p_u$
+Recall that according to our generative model of \autoref{s:gen}, each node $u$ has a parameter $p_u$
 governing its sending behavior and each node $v$ has a parameter $q_v$ governing its receiving
-behavior such that the sign of the edge \euv{} is positive with probability $\frac{p_u+q_v}{2} =
+behavior, such that the sign of the edge \euv{} is positive with probability $\frac{p_u+q_v}{2} =
 \etauv$. Given a topology $G=(V,E)$, we start by assigning a $p$ and $q$ values to each node. If we
 want \etauv{} to be uniform over $E$, we have to take into account the out-degree of nodes with at
-least one outgoing edges, and likewise the in-degree of nodes with at least one incoming edges. For
-that, we partition the interval $[0,1]$ into $|V|$ segments of size proportional to $\dout(u_1),
+least one outgoing edge, and likewise the in-degree of nodes with at least one incoming edge. For
+that, in the case of $p$, we partition the interval $[0,1]$ into $|V|$ segments of size proportional
+to $\dout(u_1),
 \dout(u_2), \ldots, \dout(u_{|V|})$. We shuffle these segments and draw a number \uar{} from each of
-them, which we set as the $p$ value of the corresponding node. Then, because we want to have more
+them, which we set as the $p$ value of the corresponding node. Then, because we want to model real
+sign distribution and have more
 positive edges, we apply an exponential transformation \[ p \rightarrow \frac{1}{1-e^{-\lambda}}
@@
 every edge \euv{}, we set it positive with probability \etauv{}. An example of the distributions we
-obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}.
+obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}, giving $71.9\%$ of
+positive edges.
 
-Once the signs are generated, we select a training set \trainset{} \uar{}, and we want to predict
+\vspace{\baselineskip}
+\noindent\begin{minipage}{187mm} 
+  \begin{minipage}{46mm}
+    \centering
+    \includegraphics[width=.95\linewidth]{raw/synth_wik_pdense.pdf}
+    \captionof{subfigure}{Density of $p$.}
+    % \label{fig:figure1}
+  \end{minipage}%
+  \begin{minipage}{46mm}
+    \centering 
+    \includegraphics[width=.95\linewidth]{raw/synth_wik_qdense.pdf} 
+    \captionof{subfigure}{Density of $q$.} 
+    % \label{fig:figure2} 
+  \end{minipage} 
+  \begin{minipage}{46mm}
+    \centering 
+    \includegraphics[width=.95\linewidth]{raw/synth_wik_etadense.pdf} 
+    \captionof{subfigure}{Density of $\eta$.} 
+    % \label{fig:figure2} 
+  \end{minipage} 
+  \begin{minipage}{46mm}
+    \centering 
+    \includegraphics[width=.95\linewidth]{raw/synth_wik_signdense.pdf}
+    \captionof{subfigure}{Density of $\eta$ for both signs.} 
+    % \label{fig:figure2} 
+  \end{minipage} 
+  \captionof{figure}{Synthetic distributions on \wik{}.} 
+  \label{fig:troll_synth_distrib} 
+\end{minipage}
+\vspace{\baselineskip}
+
+Once the signs are generated, we select a training set \trainset{} \uar{}, and predict
 the signs of the remaining edges in the testing set $E\setminus \trainset{}$. The Bayes optimal
@@
 $u \in V$ (or setting those values to \shalf{} in case there is no outgoing or incoming edges for
-some node), we use the equation \eqref{eq:predictor}, having estimated $\tau$ on $E_0$.
+some node), we use the equation \eqref{eq:predictor}\footnote{We reproduce this equation here for
+convenience: $\sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) -\tau - \tfrac{1}{2}\Big)$.},
+having estimated $\tau$ on $E_0$.
 
-The results of this comparison on the six datasets are showed in \autoref{tab:troll_synth_global}.
-For each network, we generated the signs once. Then for different training size (20\%, 40\% or 80\%
+The result of this comparison on the six datasets is showed in \autoref{tab:troll_synth_global}.
+For each network, we generated the signs once. Then, for different training size (20\%, 40\% or 80\%
 of $E$), we sampled a training set and predict using the knowledge of the true $\eta$ (Bayes
 predictor) or its estimated value (\usrule{}).
-The MCC of the Bayes predictor is the same on all datasets (around $30$), and so is its accuracy
-(around $75$). The corresponding values for \usrule{} are close on all datasets (although some are
+The MCC of the Bayes predictor is the same on all datasets (around $30$\footnote{Here and in the
+following, we multiply all MCC value by 100 to improve readability.}), and so is its accuracy
+(around $75\%$). The corresponding values for \usrule{} are close on all datasets (although some are
 easier) and the gap naturally decreases as the training size increases. Another interesting
@@
 \left(1-\hun(v)\right) -\tauhat$ as defined in \autoref{ss:bayes_approx}. We compute the mean
-absolute error (MAE) between $\eta$ and $\wh{\eta}$ for all testing edges, and specifically for
+absolute error (MAE) between $\eta$ and $\wh{\eta}$ for all testing edges, and then specifically for
 the testing edges whose both endpoints have been sampled above a certain threshold in the training
@@
 right corner (reflecting the imbalance toward positive edges), which allows better estimation. 
-
 In this case, we cannot use the MCC because off-diagonal, the Bayes predictor classifies edges
@@
 increases symmetrically as $\frac{p+q}{2}$ moves away from \shalf{}. The gap also tends to disappear
-when $\frac{p+q}{2}=\lhalf$, most likely because in that area, prediction can only be made by
-chance.
+when $\frac{p+q}{2}=\lhalf$, for in that region, both predictors can only rely on random predictions.
 
@@
 datasets and training size.
-
 \end{newcontent}
@@
 
-We perform two other sets of experiments. The first one assesses the effect of predicting reciprocal
+We perform two other sets of experiments. The first one evaluates the effect of predicting reciprocal
 edges by their value if available. As expected, the results in \autoref{tab:twin} demonstrate that
-it improves MCC when the network has enough twin edges with low disagreements, like in \sla{} and
-\epi{}. It has no effect when there are few twin edges, like in \aut{}, \wik{} and \kiw{}. And it is
-detrimental when there are many twin edges that do not agree enough, like in \adv{}, in which
+it improves MCC when the network has enough reciprocal edges with low disagreements, like in \sla{} and
+\epi{}. It has no effect when there are few reciprocal edges, like in \aut{}, \wik{} and \kiw{}. And it is
+detrimental when there are many reciprocal edges that do not agree enough, like in \adv{}, in which
 case it is better to rely solely on the learned model.

### 2017-09-20 cc/gtx_related.tex
add missing labels

@@
+\label{sub:gtx_state_of_the_art}
 Looking for a subgraph $H$ of $G$ that best preserve the distance in $G$ while being sparse is an old

### 2017-09-21 edge/intro.tex
write about a possible connection between signed graphs and vector graphs

\paragraph{Generalization of signed graphs}
\label{par:generalization_of_signed_graphs}

Here we show why this formulation can model signed graphs as a special case. Remember the learning
bias we introduced in the previous chapter \vpageref{text:cc_new_bias}: every node belongs to
exactly one of the $K$ clusters available, is connected positively with the other nodes belonging to
the same cluster, and is connected negatively with the nodes belonging to every other clusters. Now
let the dimension of profiles be $d=K$ and the number of directions $k=2$. Moreover, we define the
two directions vectors as $w_1 = \frac{\onev}{\sqrt{d}} = w^+$ and $w_2 = -w_+ = w_-$, where
$\onev$ is the vector of dimension $d$ with only $1$ coordinates. Finally, for a node $u$ in
cluster $i$, we assume its profile $x_u$ is set as follow: all component are equal to $-b$ except
for the \ith{} one, which is equal to $a$, where $a$ and $b$ are constants we define now.

First, because $x_u$ is a unit vector, we have that $a^2 + (K-1)b^2=1$, from which $a =
\sqrt{1-(K-1)b^2}$ whenever $1-(K-1)b^2 \geq 0$, or equivalently $b \leq \frac{1}{\sqrt{K-1}}$.
Next, we look at $\onev^T\left(x_u \circ x_v\right) = \onev^T m_{uv} = \frac{1}{\sqrt{d}}
\sum_{i=1}^d m_{{uv}_i}$ for any edge $(u,v) \in E$. If $u$ and $v$ are in the same cluster, then
$\sum_{i=1}^d m_{{uv}_i} = a^2 + (K-1)b^2 = 1$, implying that $w_+^T m_{uv} = \frac{1}{d} > w_-^T
m_{uv} = -\frac{1}{d}$. Otherwise, $\sum_{i=1}^d m_{{uv}_i} = (K-2)b^2 - 2ab = (K-2)b^2 -
2b\sqrt{1-(K-1)b^2} = f_K(b)$. By computing the derivative of $f_K$ with respect to $b$, we find
that it reaches a minimum of $-\frac{1}{K-1}$ at $b = \frac{1}{\sqrt{K(K-1)}} \leq
\frac{1}{\sqrt{K-1}}$. In this case, $w_+^T m_{uv} = -\frac{1}{(K-1) \sqrt{d}} < w_-^T m_{uv} =
\frac{1}{(K-1) \sqrt{d}}$. To maximize $\sum_{(u,v)\in E} \max_{1 \leq \ell \leq d} w_\ell^T
m_{uv}$, we therefore assign $w_+$ to edges whose both endpoints belong to the same cluster, and
$w_-$ to edges joining two different clusters.

Moreover, given this assignment, we can show that $w_+$ and $w_-$ are the optimal directions, under
the additional uniformity assumption that there are exactly $r$ edges within every cluster, and $s$
edges between any two clusters. First, we define
\begin{equation*}
  m_{\mathrm{inner}} = \sum_{(u,v) \in E;\, \cluster(u)=\cluster(v)} m_{uv} \qquad \text{and} \qquad
  m_{\mathrm{outer}} = \sum_{(u,v) \in E;\, \cluster(u)\neq \cluster(v)} m_{uv}
\end{equation*}
to be the sum of all edge vectors within and between clusters respectively. By our choice of $a$ and
$b$, we have that $m_{\mathrm{inner}} = r\onev$ and $m_{\mathrm{outer}} = -\frac{s}{2}\onev$. Now
recall that the solution of $\argmax_{w\in \dsphere} w^T c$ is $\frac{c}{||c||}$. Therefore, the
best direction to assign to inner edges is indeed $w_+$, and so is $w_-$ to outer edges.

\begin{aside}
For inner edge vectors, $m_{uv}$ in cluster $i$ has $a^2$ as its \ith{} component, and $b^2$
elsewhere. There are $r$ of them in cluster $i$, so summing over all clusters we get
$m_{\mathrm{inner}} = r\left(a^2 + (K-1)b^2\right)\onev = r\onev$.
For outer edge vectors, there are a few more steps. Consider the $s$ edges between $C_1$ and $C_2$,
whose $m_{uv}$ is
$\begin{pmatrix}
  -ab \\
  -ab \\
  b^2 \\
  b^2 \\
  \vdots \\  
\end{pmatrix}$. Likewise, between $C_1$ and $C_3$ we have
$m_{uv}=\begin{pmatrix}
  -ab \\
  b^2 \\
  -ab \\
  b^2 \\
  \vdots \\  
\end{pmatrix}$. We thus we see that the first component is always $-ab$, that there is another
such term at the other cluster index, while the rest is $b^2$. Summing all the vectors incident to
$C_1$, we get
$s\begin{pmatrix}
  -ab(K-1) \\
  b^2(K-2) -ab \\
  b^2(K-2) -ab \\
  \vdots \\  
\end{pmatrix}$. Similarly, the sum of all vectors incident to $C_2$ is
$s\begin{pmatrix}
  b^2(K-2) -ab \\
  -ab(K-1) \\
  b^2(K-2) -ab \\
  \vdots \\  
\end{pmatrix}$, and so on for every cluster $i$. Summing over all clusters and dividing by $2$ to
avoid double counting, we get $m_{\mathrm{outer}} = \frac{s}{2}\left( (K-1)\left(b^2(K-2) -ab\right)
-ab(K-1) \right)\onev = \frac{s(K-1)}{2}\left(b^2(K-2) -2ab\right)\onev =
\frac{s(K-1)}{2}f_K(b)\onev = -\frac{s(K-1)}{2} \frac{1}{K-1} \onev = -\frac{s}{2}\onev$.
\end{aside}

Granted, the situation where all the nodes of a cluster have exactly the same profile and where the
connections are so regular is highly idealized. Moreover, we did not formally prove that there is
not another assignment (with different direction vectors) that would give a better objective
value\footnote{Although given the symmetry of the problem, that would be surprising.}. Our point is
rather that profiles and directions can be seen as a \enquote{latent} explanation for the signs in
a balanced signed graphs, like our generative model of the first chapter.

### 2017-09-25 edge/model.tex
start writing about the edge label problem and formulation

\section{Settings and learning problem}
\label{sec:edge_model}

As in the previous chapters, we are given an unweighted graph $G=(V,E)$\stodo{undirected?}. This
time though, we also have side information in addition to the graph topology. Namely, each node $u$
is associated with a $d$-dimensional feature vector $x_u \in \Rbb^d$. As an example, in a social
network, $x_u$ would be the profile of user $u$, describing $u$'s demographics and preferences. We
are also given a combination operator between feature vectors $s$ such that $s(x_u, x_v) = s_{uv}
\in \Rbb^d$. Finally, we consider distinguished feature vectors $w_{uv}$\stodo{is it better to
introduce $w$ with a single $\ell$ index, to emphasise that later, it will not be unique per edge?},
along which the combination $s_{uv}$ between two nodes $u$ and $v$ is evaluated. We call such
$w_{uv}$ \emph{directions}, and constrain them to have unit norm. As we will discuss, this
constraint can be seen as a way to couple the different dimensions of the profiles.

Depending on the application, and the semantic we ascribe to links between nodes, we can design and
interpret the combination operator $s$ as a difference, a similarity or a distance. Likewise, we
also have some freedom in defining how $s_{uv}$ and $w_{uv}$ are combined to assign a single real
value $g(s_{uv}, w_{uv})$ to the edge $(u,v)$. Let us call $g(w_{uv}, s_{uv})$ the \emph{goodness}
of $w_{uv}$ with respect to the edge $(u,v)$. We view this value as how adequately the direction
$w_{uv}$ \emph{explains}, or describes, the connection between $u$ and $v$.

\medskip

We assume in the following that both the profiles and the directions are vectors with components in
$[-1, 1]$. A natural choice for the operator $s$ is then the difference between the profile, that is
$s_{uv} = x_u - x_v$. In that spirit, if $u$ and $v$ are close along a direction $w_{uv}$, we expect
${x_u}^T w_{uv}$ to be close to ${x_v}^T w_{uv}$, as measured by $|{x_u}^T w_{uv} - {x_v}^T
w_{uv}|$. Because we prefer to work with smooth functions, we define the goodness to be $g(s_{uv},
w_{uv}) = \left({s_{uv}}^T w_{uv} \right)^2$, and we want to minimize it with respect to $w_{uv}$.
When ${x_u}_{;i} - {x_v}_{;i}$ is close to zero, the minimization and the fact that $w_{uv}$ is
unit-norm impose that ${w_{uv}}_{;i}$ should be large. On the other hand, with this formulation,
there is no way to set ${w_{uv}}_{;i}$ in order to \enquote{reward/favor} the fact that
${x_u}_{;i}$ and ${x_v}_{;i}$ are widely different.
Whereas there might be situations where this is not a problem, as we discussed in the introduction,
here we seek to model heterophilic interactions, at least along a subset of dimensions. Therefore, in
the following, we will adopt a refined approach.

Namely, we define $s$ to be the Hadamard (or component wise) product: $s_{uv} = x_u \circ x_v$, and
$g$ to be the dot product: $g(s_{uv}, w_{uv}) = {s_{uv}}^T w_{uv}$. The goodness of a direction $w$
for a given edge can thus be seen as a weighted sum of the per component similarity of $u$ and $v$
profiles. Therefore, it must be maximized with respect to $w_{uv}$. For a given component $i$, when
both ${x_u}_{;i}$ and ${x_v}_{;i}$ are large and of the same sign (that is close to either $+1$ or
$-1$), we see that $u$ and $v$ agree on the \ith{} dimension, and ${w_{uv}}_{;i}$ must accordingly
be a large positive value. On the other hand, when ${x_u}_{;i}$ and ${x_v}_{;i}$ are both large but
of different signs (\eg ${x_u}_{;i} = +1$ and ${x_v}_{;i} = -1$), $u$ and $v$ strongly disagree on
the \ith{} dimension, requiring in a large negative value for ${w_{uv}}_{;i}$.  Finally, if one of
${x_u}_{;i}$ or ${x_v}_{;i}$ is zero, denoting either an unimportant feature or a missing value,
then the value of ${w_{uv}}_{;i}$ does not matter, and will be set to zero by the maximization
procedure because of the unit norm constraint on $w_{uv}$.

\begin{aside}
When $s$ is a distance, and therefore positive, we impose $w_{uv}\geq 0$. A link $(u,v)$ is then
well explained by $w_{uv}$ when observed differences along this direction are small. More precisely,
when the quantity ${s_{uv}}^T w_{uv}$ is close to zero. Because $w_{uv}$ is unit norm, the \ith{}
component of $w_{uv}$ should be large when the \ith{} feature of $x_u$ and $x_v$ are close. Hence, a
null value for a component $i$ of $s_{uv}$ should not mean anything else than $u$ and $v$ share a
known and similar feature $i$ (\eg{} the unknown value should be different from $0$). When $s$ is
the difference $x_u - x_v$, it can be negative and the same reasoning can be applied, but with
$w_{uv}\in [-1,1]^d$ and a link value of $\left({s_{uv}}^T w_{uv}\right)^2$.

An alternative is to consider that the combination is a similarity. Links are created when the
observed similarity is high. Again, if $s$ is positive, it seems reasonable to impose that
$w_{uv}\geq0$ (and if $s$ can be negative, then $w_{uv}$ should be in $[-1,1]$). One main advantage
of similarities is maybe to avoid (or minimize) problems with missing values setting them to 0.
\end{aside}

\medskip

Based on these specific definitions of $s$ and $g$, we now formally write down the problem of
finding and assigning the most informative directions to every edge of $G$
\begin{problem}[]
  \label{p:edge_full}
  Given a graph $G=(V, E)$, node profiles $X\in [-1, 1]^{n\times d}$ and an integer $k \in \Nbb$,
  find a set of $k$ unit-norm directions $\mathcal{D}_k = \{w_1, w_2, \ldots, w_k\} \in \dsphere$
  and associate to every edge of $E$ the direction with the maximal goodness. Formally, solve
  \begin{equation}
    \label{eq:edge_full}
    \argmax_{w_1,\ldots,w_k \in \dsphere}
    \sum_{u,v \in E} \max_{\ell \in \rangesk} g(s_{uv}, w_{\ell}) =
    \sum_{u,v \in E} \max_{\ell \in \rangesk} \left(x_u \circ x_v\right)^T w_{\ell} = f(k)
  \end{equation}
\end{problem}

If $k=|E|$, the topology of the graph becomes irrelevant, as we can simply exploit the trivial
solution of having a single direction $w_{uv} = \frac{s_{uv}}{||s_{uv}||}$ for every edge $(u,v)$.
More generally, as $k$ increases, we expect the value $\max_{\mathcal{D}_k} f(k)$ to increases as
well, at the cost of interpretability. There could be a principled approach to finding the
\enquote{best} $k$, based on a information theoretical measure of the complexity of $\mathcal{D}_k$
and the minimum description length principle~\autocite{grunwald2005tutorial}. However, in the
interest of simplicity, in the following we focus on the formulation stated in
\autoref{p:edge_full}, where $k$ is given. In practice, we further constrain $k$ to be small (say
less than \np{10}). Within the multilayer framework, $k$ could also be seen as the number of
underlying layers\stodo{elaborate on that connection}.

\begin{aside}
Generalizing the balance theory from signed graphs, we could imagine that in triangles or short
cycles, only certain combinations of directions are valid (or at least desirable). As a nice side
effect, this would likely require more innovative optimization algorithms. However, it sounds very
application-dependant. Furthermore, in our current setting, we do not know in advance in the
semantic of each direction, and we have no labeled edges. Therefore, it is not clear how those
constraints would be specified or whether they can be learned, if applicable. 
\end{aside}

Another way to leverage the graph topology is to define local constraints at the node level. Since
the beginning we considered that the user profiles were given, but from now on we further assume
that they are normalized so that they are unit norm, just like directions. In practice, if nodes are
users of a social network and the profiles measure their activity across several domains, this
allows to compare users with various level of total activity. One way to integrate that fact in
our optimization formulation is to assume that profiles are linear combination of directions, that
is $x_u = b_u + \sum_{v \in \nei(u)} a_{uv} w_{uv}$, where $a_{uv}$ are real coefficients and $b_u
\in \Rbb^d$. Viewed in the other direction, this can be interpreted as follow: each time a node $u$
connect to another node $v$ along a direction $w_{uv}$, it consume a part of its profile
proportional to $w_{uv}$. Therefore, we subtract the term $\sum_{u\in V} \left|\left| x_u - b_u -
\sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2$ to \eqref{eq:edge_full}.
%
We can also make the assumption that each node $u$ is only involved in $k_\mathrm{local} < k$
different directions. The intuition is that when $d$ is large enough, users have to focus their
energy and can only express their interest in a few number of dimensions. In other words, all the
edges incident to $u$ can only be associated with one of $k_\mathrm{local}$ directions. This is
difficult to incorporate directly in the optimization formulation, as for certain value of
$k_\mathrm{local}$, this is a \NPc{} graph colouring problem. Still, it can be seen as a learning
bias and will be used to generate synthetic data.
These two constraints can be combined, so that the profiles are close to a linear combination of a
small number of directions.

% $\theset{(u,v) \in E}{v \in \nei(u)}$

\bigskip

In order to make the optimization more tractable, we relax the formulation of \autoref{p:edge_full}
by replacing the $\max$ inside the sum of \eqref{eq:edge_full}, either by a continuous approximation
or by a low rank matrix approach.

One can check that given a set of numbers $S=\{a_1, a_2, \ldots, a_{|S|}\}$ and real number $\beta >
0$, we have
\begin{equation*}
  \max_{a_i \in S} a_i = \lim_{\beta \rightarrow + \infty} \frac{1}{\beta}
  \log\left( \sum_{i=1}^{|S|} \exp{\beta a_i} \right)
\end{equation*}
Using this fact, and averaging the two components of our objective function, we can rewrite
\eqref{eq:edge_full} as:
\begin{equation}
  \argmax_{w_1,\ldots,w_k \in \dsphere,\, a_{uv}\in\Rbb,\, b_u \in \Rbb^d}
  \frac{1}{|E|\beta} \sum_{u,v\in E}
  \log\Bigl(\sum_{\ell=1}^k\exp\left(\beta {s_{uv}}^T w_\ell \right)\Bigr)
  - \frac{1}{|V|}\sum_{u\in V}
  \left|\left| x_u - b_u - \sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2
\end{equation}

Another way to remove the $\max$ is to look for a single direction per edge, stacked horizontally in
one matrix. Then we can directly minimize \eqref{eq:edge_full} while enforcing the matrix
$W\in\mathbb{R}^{d\times |E|}$ of all directions to be low-rank, so that each $w_{uv}$ (given by the
$(u-1)n+v$-th column of $W$) can be expressed as a linear combination of a small number of basis
vectors. Formally, if $W$ has rank $k$, then there exists $P\in\mathbb{R}^{d\times k}$ and
$Q\in\mathbb{R}^{|E|\times k}$ such that $W=PQ^T$ and $w_{uv} = Pq_{uv}^T$. $P$ defines $k$
representative directions in $\Rbb^d$ while $Q$ gives, for each edge, the weights of each
representative direction. Similar to what is done in spectral clustering, we could cluster the
$q_{uv}$'s to obtain an edge labeling.
%
Denoting by $\mathbb{M}^{d\times |E|}$ the set of $d\times |E|$ matrices with unit $L_2$ norm
columns, a convex formulation of the problem is as follows:
\begin{equation}
  \label{eq:fact_convex}
  \min_{W\in\mathbb{M}^{d\times |E|}} \quad
  \sum_{i,j \in E} {s_{uv}}^T w_{uv} + \lambda \|W\|_*,
\end{equation}
where $w_{uv} \in \Rbb$ denotes the $(u-1)n+v$-th column of $W$, $\|W\|_* =
\sum_{i=1}^{\min(d,|E|)}\sigma_i(W)$ is the sum of the singular values of $W$ (called the nuclear
norm or trace norm), and $\lambda \geq 0$ is a regularization parameter.\footnote{To make the
optimal solution unique, we can use an additional Frobenius norm regularization on $W$.} The nuclear
favors low-rank solutions as it is essentially an $L_1$ norm on the spectrum.
Problem~\eqref{eq:fact_convex} can be solved using proximal gradient descent (requires singular
value decomposition at each iteration) \autocite{Parikh2013a} or Frank-Wolfe (requires only to find
the largest singular vector) \autocite{Jaggi2013a}.

Alternatively, we can consider a nonconvex formulation where we make the low-rank decomposition of
$W$ explicit. Given a rank $k\leq \min(d,|E|)$:
\begin{equation}
  \label{eq:fact_nonconvex}
  \min_{P\in\mathbb{M}^{d\times k}, Q\in\mathbb{M}^{|E|\times k}} \quad
  \sum_{u,v \in E} {s_{uv}}^T w_{uv}.
\end{equation}
Problem~\eqref{eq:fact_nonconvex} can be solved via alternating optimization over $P$ and $Q$ using
standard (projected) gradient descent algorithms.

Note however that this matrix formulation requires learning more parameters, as it each edge has a
different, albeit made up of a small number of basis direction.  Impose sparsity on $Q$?

Max margin, replace the max by a difference of the currently assigned direction with the other
(trade off parameter). At least I think I used it in synthetic data.

### 2017-09-28 intro/draft.tex
start of the draft of the intro for the edspi report

Graphs are useful to represent relationships between entities. In particular, a lot can be learned
from graphs encompassing several types of relations in one integrated structure. However, many
existing works assume that such graphs are given, before studying interesting properties like
community structure and information spreading. In this thesis instead, we develop methods to predict
the type of edges given a graph topology, a few initial label and possibly attributes of the nodes.
\smallskip

\paragraph{Learning on graphs}

Graph theory has been around for almost three centuries~\autocite{biggs1976graph}, and it provides a
conceptually simple yet immensely rich framework to model situations where entities are connected
with each other~\autocite{ComplexNetworksApp11}. Coupled with the increasing availability of large
amount of relational data, it has recently spurred a lot of interest across various and tangible
applications:
\begin{itemize}[nosep,leftmargin=*]
  \item community detection ~\autocite{FortunatoSurvey10} Describe the task in one sentence and give
    some nice examples from the survey, if possible supported by real-world numbers;
  \item semi-supervised learning ~\autocites{SSL06}{graphSSL14} Idem
  \item node classification ~\autocite{nodeClassif11} Idem
  \item node embedding ~\autocite{representationLearning17} Idem
  \item link prediction ~\autocite{linkPredSurvey16} Idem
  \item influence maximization ~\autocites{infmaxKempe15}{infmax13} Idem
  \item network evolution ~\autocite{networkEvolution14} Idem
\end{itemize}

\vspace{-\baselineskip}
\paragraph{Complex networks}

The prevalent setting for most of the works discussed above is that the graph connections are: (i)
homogeneous, in the sense that all edges are of a single type, having the same semantics  between
any pairs of nodes; and (ii) based on the \emph{homophily} principle~\autocite{Homophily01}, roughly
meaning that nodes are linked because they are similar. However, many real world situations require
a more complex model, with several kind of edges and where nodes can be connected if they are
similar only on some dimension or even if they are dissimilar. For instance, a e-commerce website
could allow its users to express whether they find the review of other users trustworthy or not. As
more involved example, consider a social network such as Facebook. Whereas the website offers only a
single \emph{friendship} relation, we may want to distinguish between family, colleague and sport
buddies~\autocite{egoNips12}. Moreover, among one user's \enquote{friend}, there might be some
relations that are actually not positive~\autocite{Yang2012}, whether because the receiving end if
spam or behave in strange way. Note finally that friendship is only kind of relationship between two
users of Facebook, we could also link them whenever they attend the same event or are tagged in the
same post. Yet another possibility would be to study the same set of users across multiple social
networks~\autocite{mergingNetworks16}, in which case the edge type could denote some common
combinations of connections (\eg Twitter+LinkedIn or Facebook+Instagram). In this thesis, we thus
defend the following statement: \textbf{There exists efficient and accurate methods to predict edge
type in complex networks (reyling either only on the graph topology or also on node metadata).}

A first motivation is that by labeling all edges, we can split a complex networks in a superposition
of simple networks with a single edge type, on which we can apply existing methods. Of course,
different kind of relation are intertwined, thus a second stronger motivation is to identify the
different layers in order to study their interactions.

\vspace{-\baselineskip}
\paragraph{Predicting polarity in directed social networks}

The first kind of such complex networks we consider are \emph{signed graphs}~\autocite{Tang2015a},
where edges carry a positive or a negative sign. Among the many applications of signed graphs, a
natural, and historical~\autocite{harary1953}, one is directed social networks. In such systems,
users can explicitly tag others, stating whether they trust them, support them, like them, or not.
Being able to predict edge signs in such networks improve the user experience, for instance by
promoting reasonable online debates~\autocite{journalism15}, assembling efficient teams within
companies~\autocite{MLinHR16} and mitigating cyberbullying~\autocite{cyberbullying13}. Existing
approaches attack the problem along several angles: finding a low dimensional representation of the
nodes based on their connection patterns~\autocites{SIGNet17}{SNE17}, finding low rank approximation
of the adjacency matrix~\autocites{LowRankCompletion14}{OnlineCompletion17} or extracting
topological features and training classifiers~\autocites{Leskovec2010}{Bayesian15}{wu2016troll}.
Guided by the idea that users exhibit various degree of trollness, we propose a simple sign
generative model. It naturally leads to two batch algorithms, one that probably approximate the
Bayes optimal predictor when the training set is dense enough and one that approximate the maximum
likelihood estimates through label propagation. Extensive experiments on synthetic and real datasets
show that in addition to their sound theoretical foundations, our methods are competitive with the
state of the art in predictive performance, while being more scalable. To illustrate this point, a
short extract of our results is showed in \autoref{tab:troll}.

\begin{table}[ptbh]
  \centering
  \caption{Matthews Correlation Coefficient measuring (the closer to 100 the better) in a \aut{}
    network as the proportion of labeled training edges grows. The best result is in red and the
  second one in blue.\label{tab:troll}}
  \vspace{0.5\baselineskip}
  \begin{tabular}{lccccc|r}
    \toprule
    {}               & 3\%                & 9\%                & 15\%               & 20\%               & 25\%               & time (ms)            \\
    \midrule
    \uslogregp{}     & \vsecondSig{15.19} & \vsecondSig{26.46} & 32.98              & 36.57              & 39.90              & 2                    \\
    \rowcolor{gray!20!white}
    \usrule{}        & 15.09              & 26.40              & \vsecondSig{32.98} & \vsecondSig{36.72} & 40.16              & \textbf{\textless 1} \\
    \rowcolor{gray!20!white}
    \uslpropGsec{}   & \vfirstSig{19.00}  & \vfirstSig{30.25}  & \vfirstSig{35.73}  & \vfirstSig{38.53}  & \vfirstSig{41.32}  & \textbf{16}          \\
    \midrule
    \compranknodes{} & 12.28              & 24.44              & 31.03              & 34.57              & 38.26              & 128         \\
    \complowrank{}   & 8.85               & 17.08              & 22.57              & 25.57              & 29.24              & 1894                 \\
    \compbayesian{}  & 10.91              & 23.75              & 32.25              & 36.52              & \vsecondSig{40.32} & 5398                 \\
    \comptriads{}    & 8.62               & 16.42              & 22.01              & 24.77              & 27.13              & \textbf{5}           \\
    \bottomrule
  \end{tabular}
\end{table}

\vspace{-\baselineskip}
\paragraph{Active learning in undirected signed graphs}

There are many other domains where signed graphs have proved useful. For instance in computer
vision, it is used to segment images in 2D~\autocites{Kim2011}{Bagon2011}{CellSeg14} or
3D~\autocites{VolumeSegmentation12}{Beier2015}, to simplify 3D shapes~\autocite{Shape3D17} and to
track target across sequential video frames~\autocite{multiTracking15}. When the nodes are words,
signed graphs are employed to identify coreference~\autocites{graphicalCoreference04}{Elsner2009} or
cluster synonym words~\autocite{SignedWordRatings}. They plays a large role in biology, to clusters
genes~\autocite{Ben-Dor99}, identify mutation regions in chromosomes~\autocite{Das2015} or finding
stable subsystems~\autocite{monotoneBiology07}. Finally, given large databases with duplicated
records, they naturally help solving entity
resolution~\autocites{Crosslingual07}{DeDup09}{LargeScaleDeDup09}{WeightedER14}.

In all these cases, graphs are not directed and nodes are not human beings, meaning we can hardly
assign them trollness.  Therefore, we make a different assumption, positing that nodes are
partitioned in $K$ groups in such a way that links within groups are positive and links across
groups are negative. Recovering the optimal partition after observing all the signs of a graph is
called \pcc{}~\autocite{Bansal2002} in the Machine Learning field. Whereas it has been proved to be
\NPc{} to even approximate in the worst case~\autocites{Charikar2003}{Demaine2006}, our thorough
literature review paints a less gloomy landscape. Not only does there exist
exact~\autocite{Berg2015} and fixed parameters algorithms~\autocites{GoldenCE12}{Fomin2014},
efficient approximation~\autocites{CCPivotConf05}{ChawlaArxiv14} and large scale
heuristics~\autocites{Levorato2015}{Facchetti2011isingmodel}{Kappes2016}, but beyond the worst case,
we are interested in instances that exhibit stability under
perturbation~\autocites{clusteringFeasibility15}{StableCC09}{StableLP09} or are obtained through
random moves from an ideal case~\autocites{plantedAilon09}{Makarychev2014}. This leads us to develop
a spanning tree tailored to predicting edge signs in such a setting, which we call \gtx{}. It does
so by aiming to approximately preserve the distances in the
graph~\autocites{LowerBound95}{Abraham2012}. We present preliminary results showing that on certain
kind of graphs, this indeed results in better predictive performance than a competitive baseline.
Namely, we create synthetic graphs of growing size with the \lpa{} model~\autocite{Barabasi1999},
partition the nodes in two clusters, set the edge signs accordingly and finally flip a fraction of
them \uar{}. As showed on \autoref{fig:gtx_xp_pasynthmcc}, \gtx{} always predict signs better than a
\bfs{} tree rooted at the highest degree node, especially when the noise level is low, and is less
sensible to the graph size.

\begin{figure}[phtb]
  \centering
  \includegraphics[width=.6\textwidth]{gtx_exp/pasynthmcc}
  \caption{Synthetic \lpa{} }\label{fig:gtx_xp_pasynthmcc}
\end{figure}

\vspace{-\baselineskip}
\paragraph{Predicting more than two kinds of relations}

Finally, we consider the case where there are more than two possible type of relations, which is
often called \emph{multilayer graphs}~\autocites{Kivela2014}{multiSurvey14}. In addition, we assume
that the graph is attributed, that is every node is described by a vector of numeric features called
the \emph{profile} of the node. Our assumption is then that links cannot be fully described by a
global homophily. Thus we take a more nuanced view, combining partial homophily (that is, nodes are
connected when they are similar on a subset of the attributes) and heterophily (that is, nodes are
connected when they are dissimilar on a subset of the attributes). As examples of the latter, think
of dating websites ---where most users are linked with users of the opposite
gender~\autocites{homophilyMyspace09}{Tinder16}; spreading innovation ---where meeting people with
different background and point of view is crucial to favor diversity and
creativity~\autocite{rogers2003diffusion}; and online news consumption ---where connecting people
from different sides of the political spectrum helps avoiding echo chambers and fueling a democratic
debate~\autocite{balancedNews17}. More generally, multilayer graphs have applications in various
fields. In social networks, where players of the Pardus massively multilayer online game can have
six types of relations (friendship, enmity, private message, trading, aggression and
bounty)~\autocite{Szell2010} or where photo sharers of the Flickr website can interact in eleven
ways (either directly or through comments, tags, groups and so on)~\autocite{RecoFlickrMulti11}. In
citation networks, where authors of the DBLP dataset are connected if they have co-authored a paper
in one the \np{1000} conferences that form the edge type~\autocite{communityDBLPbyConf05}, or where
\np{5000} SIAM papers can be connected for five reasons (abstract similarity, title similarity,
keywords similarity, author similarity or citation)~\autocite{articlesMultiSim11} in economic
networks, where 951 ports are connected by 3 kind of ships (bulk dry carriers, container ships and
oil tankers)~\autocite{ports3kindofships10} or in the International Trade Network, where 162
countries are connected by 96 kind of commodities they can exchange~\autocite{worldTradeNetwork10}.
And in biology, where genes co-expression network under 130 different experimental
conditions~\autocite{bioLayerExp11}

Existing works ~\autocite{egoNips12}

The social circles model defines a generative for connection inside ego
networks~\autocites{LeskovecEgo12}{LeskovecEgo14}. Basically, each node makes friends along $K$
categories and we learn the parameters that maximize the probability for each edge to belong to one
of the $K$ categories\footnote{But those categories can change from node to node.}.

Yixiang Fang, Reynold Cheng, Siqiang Luo, and Jiafeng Hu. 2016. Effective community search for large
attributed graphs. Proc. VLDB Endow. 9, 12 (August 2016), 1233-1244. DOI:
\url{http://dx.doi.org/10.14778/2994509.2994538}

Look for attributed community,  which are subgraphs satisfying both structure cohesiveness (i.e.,
its vertices are tightly connected) and keyword cohesiveness (i.e., its vertices share common
keywords),  enabling a better understanding of how and why a community is formed
To enable efficient AC search, we develop the CL-tree index structure and three algorithms based on
it. We evaluate our solutions on four large graphs, namely Flickr, DBLP, Tencent, and DBpedia. Our
results show that ACs are more effective and efficient than existing community retrieval approaches.

Find partitions by maximizing the densite of intra community edges, but weight the edge between $i$
and $j$ by a function that depends both of the $i$ and $j$ profile similarities and a learned weight
vector per community~\autocite{ZhangModelFree16}. Solved by alternating optimizing over node
assignment and over community weight learning.
vector

each node can be seen as a point in $R^d$ and because the dimensions are assumed to be not
independent, each point belong to a subspace. It's a well studied problem~\autocite{SCSurvey11}

Recovering $d$ different metric spaces whose intersection form the observed social multiplex in
almost linear time with bounded distortion~\autocite{Abraham2012a}. Assume there are $K$ social
categories modelled by Euclidean spaces $\mathcal{D}_i$. Users in there are near uniformly
distributed and categories have small local correlation:
\[ \forall i\neq i', \; \forall r, r' = O(polylog(n)),\,
  \forall u, u',\, |\mathcal{B}_i(u, r) \cap \mathcal{B}_{i'}(u', r')| \leq O(\log n) \]
These spaces give rise to small world networks $\mathcal{G}_i$ with edge probability $\propto
\mathcal{D}_i(u, v)^{-d}$ and we observe the real network $\mathcal{G} =\bigcup_i \mathcal{G}_i$.
From $\mathcal{G}$, the proposed algorithm recovers in $n \mathrm{polylog} n$ time a bounded
approximation $\mathcal{D}'_i$ of all $\mathcal{D}_i$
\[ \sigma \mathcal{D}_i(u, v) \leq \mathcal{D}'_i(u, v) \leq \delta \mathcal{D}_i(u, v) + \Delta \]

As hinted by the title, \autocite{Ahmed2016}\footnote{published version in
\autocite{ahmed2017roles}?} is a specific instantiation of a more general framework
about Role Discovery in Networks by the same authors. Originally, two nodes were said to have the
same role if they were structurally equivalent (that is, connected to the rest of the network in
identical ways). Hereafter, structural equivalence is relaxed into mere similarity. Furthermore, the
entities under consideration are now edges instead of nodes.
The framework has two main steps:
1) Automatically extracting a feature matrix X from an attributed graph
2) Applying clustering or factorization methods on X to assign roles to edges

The first step is mostly based on counting the number of k-motifs in which edges is involved,
although the authors stresses that their method could easily be extended to handle features derived
from node attributes. Letting m be the number of edges, the output of this step is a $m\times f$
matrix X of edge features.

In the second step, the authors seek a low rank approximation of X. Namely, they
want to minimize the Bregman divergence between X and UVT, where U and V are
two non-negative matrix of rank at most r. The optimization is carried out by a
parallel coordinate descent method called PCMF-BASIC and described in a separate
journal article: Parallel collective factorization for modeling large heterogeneous
networks.
There is also a regularization term, which is not given explicitly (although it
might have something to do with ensuring UVT sparsity). The rank r is chosen
dynamically through a Minimum Description Length approach.

The article ends with an extension to time evolving graph and some experimental
illustrations on the Enron mail dataset (and a citation one I assume). They also
give an example of application: detecting malicious user in Social Network. While its
node and most of its edges are normal, some edges might not be (harassment,
scam, …)

Stattner, Erick, and Martine Collard. "Link Clustering for Extracting Collaborative Patterns in a Scientific Co-Authored Network.", FAB workshop, ANOSAM 2017
they use their method of conceptual links
Groups of nodes are first created when they have common attributes. The set of links between two
groups is called a conceptual link in the sense that it corresponds to a relationship between two
sets of attributes, namely two concepts in the field of formal concept analysis [7]. The number of
links between two groups is then evaluated and when the frequency is greater than a given support
threshold $\beta$, we talk about frequent conceptual link. Note that we only keep maximal frequent conceptual links (mfcl), i.e. those who are not included in others.
Finally, the set of maximal frequent conceptual links is used to create a new network structure called a conceptual view which summarizes all the knowledge extracted from the initial network. In a conceptual view, each node corresponds to a set of attributes (in this context we call them Meta-nodes) and a link corresponds to maximal frequent conceptual links.
A formal presentation of frequent conceptual links can be found in [18].

Our approach is to seek a small number of vectors that, once assigned to every edge, best explain
the graph (\ie maximize a score function between the edge vector and the profiles of its endpoints).
From this initial formulation, we derive two optimization problems, and initially solve them on
synthetic data, generated with a few natural topological constraints. We compared these solutions
with the result of a baseline approach that simply clusters the pairwise profile vectors using
$k$-means.

% Surveys are also available about applications of unsigned networks
% such as data classification [Sen et al. 2008], recommendation [Tang et al. 2013], and
% information propagation [Chen et al. 2013a]

While our thesis is therefore primarily concerned with learning on graphs, instead of focusing at
nodes (to predict their class, their cluster or to find their representation), we look at edges (\ie
pairs of nodes), and instead of having one kind of relationship, we have two, or more than two when
the nodes are attributed.

### 2017-09-29 intro/draft.tex
first finished version of the edspi report

@@
 \begin{itemize}[nosep,leftmargin=*]
-  \item community detection ~\autocite{FortunatoSurvey10} Describe the task in one sentence and give
-    some nice examples from the survey, if possible supported by real-world numbers;
-  \item semi-supervised learning ~\autocites{SSL06}{graphSSL14} Idem
-  \item node classification ~\autocite{nodeClassif11} Idem
-  \item node embedding ~\autocite{representationLearning17} Idem
-  \item link prediction ~\autocite{linkPredSurvey16} Idem
-  \item influence maximization ~\autocites{infmaxKempe15}{infmax13} Idem
-  \item network evolution ~\autocite{networkEvolution14} Idem
+  \item community detection~\autocite{FortunatoSurvey10}, where the goal is to cluster nodes in
+    tightly connected group that are loosely connected with the rest of the graph, in order to
+    better understand their relationships and the graph organization. Is has been used for instance
+    to identify proteins functional groups~\autocite{clusterBio03} or see how different scientific
+    fields relate based on publication data~\autocite{clusterScience08}.
+  \item semi-supervised learning~\autocites{SSL06}{graphSSL14}, where during training, we are
+    provided with unlabeled data at the same time as labeled data. Connecting similar instances
+    allow propagating information along the graph, with applications in classifying text
+    documents~\autocite{sslText09} or aligning categories and relations across multiple knowledge
+    base~\autocite{sslKB13}.
+  % \item node classification~\autocite{nodeClassif11} Idem
+    % could add visualisation (Roberto Tamassia. 2013. Handbook of Graph Drawing and Visualization.
+    % CRC Press.) but it's not exactly learning
+  \item node embedding~\autocite{representationLearning17}, where we seek a low dimensional
+    representation of nodes based on their structural patterns, usually in an unsupervised way,
+    although it is also possible to include problem supervision when available. Such representation can
+    then be used in downstream tasks, for instance visualisation~\autocite{LINE15} or the
+    aforementioned node classification, even when new nodes join the graph after
+    training~\autocite{inductiveRepresentation17}.
+  \item link prediction~\autocite{linkPredSurvey16}, where the goal is, given a snapshot of the
+    graph at time $t$, to return a set of non-existing links that will be created by time
+    $t+\Delta$. Most methods are based on the assumption that link creation is driven by node
+    similarity and it has been successfully applied to inferring potential interaction between
+    proteins without expensive experiments~\autocite{linkPredBio06} and uncovering hidden
+    association in criminal networks~\autocite{linkPredCrime08}.
+  \item information and influence propagation~\autocite{infmax13}, which is the study of processes
+    by which content is spread across networks, and how such processes can be influenced to speed
+    them up or slow them down. Two main applications are selecting the best seed in a social
+    network to promote a viral marketing campaign~\autocite{infmaxKempe15} and better contain the
+    diffusion of actual biological viruses~\autocite{influenceBio13}.
+  \item network evolution~\autocite{networkEvolution14}, which focuses on the mechanisms and
+    consequences of the growth of most networks, as well as how to keep the results on some data
+    mining algorithms \enquote{fresh} and relevant. Monitoring the changes in the interactions of
+    proteins can indeed be used as an early indicator for some kind of
+    diseases~\autocite{evolBio10}, while sudden changes in networks of computers are usually worth
+    investigating for they might signal some external attack~\autocite{evolSecurity04}.
 \end{itemize}
@@
 from different sides of the political spectrum helps avoiding echo chambers and fueling a democratic
-debate~\autocite{balancedNews17}. More generally, multilayer graphs have applications in various
-fields. In social networks, where players of the Pardus massively multilayer online game can have
-six types of relations (friendship, enmity, private message, trading, aggression and
-bounty)~\autocite{Szell2010} or where photo sharers of the Flickr website can interact in eleven
-ways (either directly or through comments, tags, groups and so on)~\autocite{RecoFlickrMulti11}. In
-citation networks, where authors of the DBLP dataset are connected if they have co-authored a paper
-in one the \np{1000} conferences that form the edge type~\autocite{communityDBLPbyConf05}, or where
-\np{5000} SIAM papers can be connected for five reasons (abstract similarity, title similarity,
-keywords similarity, author similarity or citation)~\autocite{articlesMultiSim11} in economic
-networks, where 951 ports are connected by 3 kind of ships (bulk dry carriers, container ships and
-oil tankers)~\autocite{ports3kindofships10} or in the International Trade Network, where 162
-countries are connected by 96 kind of commodities they can exchange~\autocite{worldTradeNetwork10}.
-And in biology, where genes co-expression network under 130 different experimental
-conditions~\autocite{bioLayerExp11}
+debate~\autocite{balancedNews17}. Our goal is to predict to which category a given belong. This is
+widely applicable, for multilayer graphs are present  in various fields. In social networks, where
+players of the Pardus massively multilayer online game can have six types of relations (friendship,
+enmity, private message, trading, aggression and bounty)~\autocite{Szell2010} or where photo sharers
+of the Flickr website can interact in eleven ways (either directly or through comments, tags, groups
+and so on)~\autocite{RecoFlickrMulti11}. In citation networks, where authors of the DBLP dataset are
+connected if they have co-authored a paper in one the \np{1000} conferences that form the edge
+type~\autocite{communityDBLPbyConf05}, or where \np{5000} SIAM papers can be connected for five
+reasons (abstract similarity, title similarity, keywords similarity, author similarity or
+citation)~\autocite{articlesMultiSim11} in economic networks, where 951 ports are connected by 3
+kind of ships (bulk dry carriers, container ships and oil tankers)~\autocite{ports3kindofships10} or
+in the International Trade Network, where 162 countries are connected by 96 kind of commodities they
+can exchange~\autocite{worldTradeNetwork10}. And in biology, where genes are connected in a
+co-expression network under 130 different experimental conditions~\autocite{bioLayerExp11}.
 
-Existing works ~\autocite{egoNips12}
-
-The social circles model defines a generative for connection inside ego
-networks~\autocites{LeskovecEgo12}{LeskovecEgo14}. Basically, each node makes friends along $K$
-categories and we learn the parameters that maximize the probability for each edge to belong to one
-of the $K$ categories\footnote{But those categories can change from node to node.}.
-
-Yixiang Fang, Reynold Cheng, Siqiang Luo, and Jiafeng Hu. 2016. Effective community search for large
-attributed graphs. Proc. VLDB Endow. 9, 12 (August 2016), 1233-1244. DOI:
-\url{http://dx.doi.org/10.14778/2994509.2994538}
-
-Look for attributed community,  which are subgraphs satisfying both structure cohesiveness (i.e.,
-its vertices are tightly connected) and keyword cohesiveness (i.e., its vertices share common
-keywords),  enabling a better understanding of how and why a community is formed
-To enable efficient AC search, we develop the CL-tree index structure and three algorithms based on
-it. We evaluate our solutions on four large graphs, namely Flickr, DBLP, Tencent, and DBpedia. Our
-results show that ACs are more effective and efficient than existing community retrieval approaches.
-
-Find partitions by maximizing the densite of intra community edges, but weight the edge between $i$
-and $j$ by a function that depends both of the $i$ and $j$ profile similarities and a learned weight
-vector per community~\autocite{ZhangModelFree16}. Solved by alternating optimizing over node
-assignment and over community weight learning.
-vector
-
-each node can be seen as a point in $R^d$ and because the dimensions are assumed to be not
-independent, each point belong to a subspace. It's a well studied problem~\autocite{SCSurvey11}
-
-Recovering $d$ different metric spaces whose intersection form the observed social multiplex in
-almost linear time with bounded distortion~\autocite{Abraham2012a}. Assume there are $K$ social
-categories modelled by Euclidean spaces $\mathcal{D}_i$. Users in there are near uniformly
-distributed and categories have small local correlation:
-\[ \forall i\neq i', \; \forall r, r' = O(polylog(n)),\,
-  \forall u, u',\, |\mathcal{B}_i(u, r) \cap \mathcal{B}_{i'}(u', r')| \leq O(\log n) \]
-These spaces give rise to small world networks $\mathcal{G}_i$ with edge probability $\propto
-\mathcal{D}_i(u, v)^{-d}$ and we observe the real network $\mathcal{G} =\bigcup_i \mathcal{G}_i$.
-From $\mathcal{G}$, the proposed algorithm recovers in $n \mathrm{polylog} n$ time a bounded
-approximation $\mathcal{D}'_i$ of all $\mathcal{D}_i$
-\[ \sigma \mathcal{D}_i(u, v) \leq \mathcal{D}'_i(u, v) \leq \delta \mathcal{D}_i(u, v) + \Delta \]
-
-As hinted by the title, \autocite{Ahmed2016}\footnote{published version in
-\autocite{ahmed2017roles}?} is a specific instantiation of a more general framework
-about Role Discovery in Networks by the same authors. Originally, two nodes were said to have the
-same role if they were structurally equivalent (that is, connected to the rest of the network in
-identical ways). Hereafter, structural equivalence is relaxed into mere similarity. Furthermore, the
-entities under consideration are now edges instead of nodes.
-The framework has two main steps:
-1) Automatically extracting a feature matrix X from an attributed graph
-2) Applying clustering or factorization methods on X to assign roles to edges
-
-The first step is mostly based on counting the number of k-motifs in which edges is involved,
-although the authors stresses that their method could easily be extended to handle features derived
-from node attributes. Letting m be the number of edges, the output of this step is a $m\times f$
-matrix X of edge features.
-
-In the second step, the authors seek a low rank approximation of X. Namely, they
-want to minimize the Bregman divergence between X and UVT, where U and V are
-two non-negative matrix of rank at most r. The optimization is carried out by a
-parallel coordinate descent method called PCMF-BASIC and described in a separate
-journal article: Parallel collective factorization for modeling large heterogeneous
-networks.
-There is also a regularization term, which is not given explicitly (although it
-might have something to do with ensuring UVT sparsity). The rank r is chosen
-dynamically through a Minimum Description Length approach.
-
-The article ends with an extension to time evolving graph and some experimental
-illustrations on the Enron mail dataset (and a citation one I assume). They also
-give an example of application: detecting malicious user in Social Network. While its
-node and most of its edges are normal, some edges might not be (harassment,
-scam, …)
-
-Stattner, Erick, and Martine Collard. "Link Clustering for Extracting Collaborative Patterns in a Scientific Co-Authored Network.", FAB workshop, ANOSAM 2017
-they use their method of conceptual links
-Groups of nodes are first created when they have common attributes. The set of links between two
-groups is called a conceptual link in the sense that it corresponds to a relationship between two
-sets of attributes, namely two concepts in the field of formal concept analysis [7]. The number of
-links between two groups is then evaluated and when the frequency is greater than a given support
-threshold $\beta$, we talk about frequent conceptual link. Note that we only keep maximal frequent conceptual links (mfcl), i.e. those who are not included in others.
-Finally, the set of maximal frequent conceptual links is used to create a new network structure called a conceptual view which summarizes all the knowledge extracted from the initial network.
-In a conceptual view, each node corresponds to a set of attributes (in this context we call them Meta-nodes) and a link corresponds to maximal frequent conceptual links.
-A formal presentation of frequent conceptual links can be found in [18].
+There exists previous work that leans towards analyzing and characterising edges in such multilayer
+graphs, although they do not share exact our goal of predicting category. For instance, one can
+maximize the probability of a generative model in ego networks with $K$
+categories~\autocites{LeskovecEgo12}{LeskovecEgo14}, look for communities that are both tightly
+connected and share common keywords~\autocite{AttributedCommunity16}, find a partition maximizing
+the density of intra community edges ---weighted by the endpoints profile similarities and a learned
+community weight vector~\autocite{ZhangModelFree16}, recover with low distortion $K$ metric spaces
+whose small-world superposition yield the observed graph~\autocite{Abraham2012a}, discover edge role
+by computing high order topological features~\autocite{ahmed2017roles}, group nodes with common
+attributes and extract frequent links between such group as edge type~\autocite{conceptualLinks12},
+or cluster the profiles into a set of affine subspace of $\mathbb{R}^d$~\autocite{SCSurvey11}.
 
@@
 with the result of a baseline approach that simply clusters the pairwise profile vectors using
-$k$-means.
+$k$-means. As showed in \autoref{fig:edge}, on a graph with \np{500} nodes, \np{1300} edges and 7
+types of edges (where each node is involved in at most 3 types of relation), we recover the hidden
+edge assignment more accurately than the $k$-means baseline.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.7\linewidth]{assets/raw/Gfixed_over0_step2_7dirs3_ambi.pdf}
+  \caption{Distribution of the Adjusted Mutual Information (the closer to one the better) over
+  \np{600} generation of 7 random directions, comparing our \textsf{optim} method with the $k$-means
+baseline. \label{fig:edge}}
+\end{figure}
 

### 2017-09-30 intro/draft.tex
proof reading of the edspi report

@@
-Graphs are useful to represent relationships between entities. In particular, a lot can be learned
+Graphs are well-suited to represent relationships between entities. In particular, a lot can be learned
 from graphs encompassing several types of relations in one integrated structure. However, many
-existing works assume that such graphs are given, before studying interesting properties like
-community structure and information spreading. In this thesis instead, we develop methods to predict
-the type of edges given a graph topology, a few initial label and possibly attributes of the nodes.
+existing works that study interesting properties of such graphs, like community structure and
+information spreading, assume that the graph if fully given. In this thesis instead, we develop
+methods to predict
+the type of edges given a graph topology, a few initial labels or alternatively attributes of the nodes.
 \smallskip
@@
 with each other~\autocite{ComplexNetworksApp11}. Coupled with the increasing availability of large
-amount of relational data, it has recently spurred a lot of interest across various and tangible
-applications:
+amount of relational data, learning on graphs has recently spurred a lot of interest across various
+lines of research, with tangible benefits:
 \begin{itemize}[nosep,leftmargin=*]
   \item community detection~\autocite{FortunatoSurvey10}, where the goal is to cluster nodes in
-    tightly connected group that are loosely connected with the rest of the graph, in order to
-    better understand their relationships and the graph organization. Is has been used for instance
+    tightly connected groups that are loosely connected with the rest of the graph, in order to
+    better understand the graph organization the nodes relationships. It has been used for instance
     to identify proteins functional groups~\autocite{clusterBio03} or see how different scientific
     fields relate based on publication data~\autocite{clusterScience08}.
-  \item semi-supervised learning~\autocites{SSL06}{graphSSL14}, where during training, we are
-    provided with unlabeled data at the same time as labeled data. Connecting similar instances
+  \item semi-supervised learning~\autocites{SSL06}{graphSSL14}, where in addition to labeled data,
+    the learner is also provided with unlabeled data at training time, and its goal is to classify
+    nodes. Connecting similar instances
     allow propagating information along the graph, with applications in classifying text
     documents~\autocite{sslText09} or aligning categories and relations across multiple knowledge
-    base~\autocite{sslKB13}.
+    bases~\autocite{sslKB13}.
   % \item node classification~\autocite{nodeClassif11} Idem
@@
   \item link prediction~\autocite{linkPredSurvey16}, where the goal is, given a snapshot of the
-    graph at time $t$, to return a set of non-existing links that will be created by time
-    $t+\Delta$. Most methods are based on the assumption that link creation is driven by node
-    similarity and it has been successfully applied to inferring potential interaction between
+    graph at time $t$, to return a set of links that do not exist at time $t$ but will be created by time
+    $t+\Delta_t$. Most methods are based on the assumption that link creation is driven by node
+    similarity, and it has been successfully applied to inferring potential interactions between
     proteins without expensive experiments~\autocite{linkPredBio06} and uncovering hidden
-    association in criminal networks~\autocite{linkPredCrime08}.
+    associations in criminal networks~\autocite{linkPredCrime08}.
   \item information and influence propagation~\autocite{infmax13}, which is the study of processes
     by which content is spread across networks, and how such processes can be influenced to speed
-    them up or slow them down. Two main applications are selecting the best seed in a social
-    network to promote a viral marketing campaign~\autocite{infmaxKempe15} and better contain the
+    them up or slow them down. Two representative applications are selecting the best seeds in a social
+    network to promote a viral marketing campaign~\autocite{infmaxKempe15} and containing more effectively the
     diffusion of actual biological viruses~\autocite{influenceBio13}.
   \item network evolution~\autocite{networkEvolution14}, which focuses on the mechanisms and
-    consequences of the growth of most networks, as well as how to keep the results on some data
+    consequences of the growth of networks, as well as how to keep the results on some data
     mining algorithms \enquote{fresh} and relevant. Monitoring the changes in the interactions of
     proteins can indeed be used as an early indicator for some kind of
-    diseases~\autocite{evolBio10}, while sudden changes in networks of computers are usually worth
-    investigating for they might signal some external attack~\autocite{evolSecurity04}.
+    diabetes~\autocite{evolBio10}, while sudden changes in a network of computers are usually worth
+    investigating, for they might signal external attacks~\autocite{evolSecurity04}.
 \end{itemize}
@@
 homogeneous, in the sense that all edges are of a single type, having the same semantics  between
-any pairs of nodes; and (ii) based on the \emph{homophily} principle~\autocite{Homophily01}, roughly
+any pair of nodes; and (ii) based on the \emph{homophily} principle~\autocite{Homophily01}, roughly
 meaning that nodes are linked because they are similar. However, many real world situations require
 a more complex model, with several kind of edges and where nodes can be connected if they are
-similar only on some dimension or even if they are dissimilar. For instance, a e-commerce website
-could allow its users to express whether they find the review of other users trustworthy or not. As
+similar only on some dimensions, or even if they are dissimilar. For instance, an e-commerce website
+could let its users express whether they find the reviews written by other users helpful or not. As a
 more involved example, consider a social network such as Facebook. Whereas the website offers only a
-single \emph{friendship} relation, we may want to distinguish between family, colleague and sport
-buddies~\autocite{egoNips12}. Moreover, among one user's \enquote{friend}, there might be some
-relations that are actually not positive~\autocite{Yang2012}, whether because the receiving end if
-spam or behave in strange way. Note finally that friendship is only kind of relationship between two
+single \emph{friendship} relation, we may want to distinguish between family ties, work ties and
+hobbies ties~\autocite{egoNips12}. Moreover, among one user's \enquote{friends}, there might be some
+relations that are actually not positive~\autocite{Yang2012}, whether because of spam, noise or
+friendship drift.. Note finally that friendship is only one possible kind of relationship between two
 users of Facebook, we could also link them whenever they attend the same event or are tagged in the
@@
 defend the following statement: \textbf{There exists efficient and accurate methods to predict edge
-type in complex networks (reyling either only on the graph topology or also on node metadata).}
+type in complex networks, reyling either only on the graph topology or also on node metadata.}
 
-A first motivation is that by labeling all edges, we can split a complex networks in a superposition
+A first motivation for such methods is that by labeling all edges, we can split a complex networks
+in a superposition
 of simple networks with a single edge type, on which we can apply existing methods. Of course,
@@
 
-The first kind of such complex networks we consider are \emph{signed graphs}~\autocite{Tang2015a},
+The first kind of complex networks we consider are \emph{signed graphs}~\autocite{Tang2015a},
 where edges carry a positive or a negative sign. Among the many applications of signed graphs, a
-natural, and historical~\autocite{harary1953}, one is directed social networks. In such systems,
+natural (and historical~\autocite{harary1953}) one is directed social networks. In such systems,
 users can explicitly tag others, stating whether they trust them, support them, like them, or not.
 Being able to predict edge signs in such networks improve the user experience, for instance by
-promoting reasonable online debates~\autocite{journalism15}, assembling efficient teams within
+promoting courteous online debates~\autocite{journalism15}, assembling efficient teams within
 companies~\autocite{MLinHR16} and mitigating cyberbullying~\autocite{cyberbullying13}. Existing
 approaches attack the problem along several angles: finding a low dimensional representation of the
-nodes based on their connection patterns~\autocites{SIGNet17}{SNE17}, finding low rank approximation
+nodes based on their connection patterns~\autocites{SIGNet17}{SNE17}, finding a low rank approximation
 of the adjacency matrix~\autocites{LowRankCompletion14}{OnlineCompletion17} or extracting
 topological features and training classifiers~\autocites{Leskovec2010}{Bayesian15}{wu2016troll}.
-Guided by the idea that users exhibit various degree of trollness, we propose a simple sign
-generative model. It naturally leads to two batch algorithms, one that probably approximate the
-Bayes optimal predictor when the training set is dense enough and one that approximate the maximum
+Guided by the idea that every users exhibit trollness to various extent, we propose a simple sign
+generative model. This model naturally leads to two batch algorithms, one that probably approximate the
+Bayes optimal predictor when the training set is dense enough, and one that approximate the maximum
 likelihood estimates through label propagation. Extensive experiments on synthetic and real datasets
@@
 state of the art in predictive performance, while being more scalable. To illustrate this point, a
-short extract of our results is showed in \autoref{tab:troll}.
+short extract of our results is presented in \autoref{tab:troll}.
 
@@
   \centering
-  \caption{Matthews Correlation Coefficient measuring (the closer to 100 the better) in a \aut{}
-    network as the proportion of labeled training edges grows. The best result is in red and the
-  second one in blue.\label{tab:troll}}
+  \caption{Matthews Correlation Coefficient measuring binary prediction performance (the closer to
+    100 the better) in a \aut{}
+    network as the proportion of labeled training edges grows. The best result is in red, the
+  second one in blue, and our two methods are highlighted in gray.\label{tab:troll}}
   \vspace{0.5\baselineskip}
@@
     \toprule
-    {}               & 3\%                & 9\%                & 15\%               & 20\%               & 25\%               & time (ms)            \\
+    method           & 5\%               & 10\%              & 15\%              & 20\%              & 25\%              & time (ms)            \\
     \midrule
-    \uslogregp{}     & \vsecondSig{15.19} & \vsecondSig{26.46} & 32.98              & 36.57              & 39.90              & 2                    \\
+    \uslogregp{}     & 19.8              & 27.9              & \vsecondSig{33.2} & 36.9              & 39.7              & \np{4}               \\
     \rowcolor{gray!20!white}
-    \usrule{}        & 15.09              & 26.40              & \vsecondSig{32.98} & \vsecondSig{36.72} & 40.16              & \textbf{\textless 1} \\
+    \usrule{}        & \vsecondSig{19.8} & \vsecondSig{28.0} & 33.1              & \vsecondSig{37.1} & 39.7              & \textbf{\textless 1} \\
     \rowcolor{gray!20!white}
-    \uslpropGsec{}   & \vfirstSig{19.00}  & \vfirstSig{30.25}  & \vfirstSig{35.73}  & \vfirstSig{38.53}  & \vfirstSig{41.32}  & \textbf{16}          \\
+    \uslpropGsec{}   & \vfirstSig{24.2}  & \vfirstSig{31.7}  & \vfirstSig{36.1}  & \vfirstSig{38.9}  & \vfirstSig{41.1}  & \textbf{\np{19}}     \\
     \midrule
-    \compranknodes{} & 12.28              & 24.44              & 31.03              & 34.57              & 38.26              & 128         \\
-    \complowrank{}   & 8.85               & 17.08              & 22.57              & 25.57              & 29.24              & 1894                 \\
-    \compbayesian{}  & 10.91              & 23.75              & 32.25              & 36.52              & \vsecondSig{40.32} & 5398                 \\
-    \comptriads{}    & 8.62               & 16.42              & 22.01              & 24.77              & 27.13              & \textbf{5}           \\
+    \autocite{LowRankCompletion14}   & 12.4              & 17.9              & 22.0              & 25.7              & 29.0              & \np{3222}            \\
+    \autocite{OnlineCompletion17}   & 1.2               & 12.6              & 22.2              & 30.3              & 36.5              & \np{23229}           \\
+    \autocite{Leskovec2010}    & 11.4              & 17.2              & 21.0              & 24.3              & 27.0              & \textbf{\np{7}}      \\
+    \autocite{wu2016troll} & 17.5              & 25.1              & 31.2              & 35.2              & 37.8              & \np{157}             \\
+    \autocite{Bayesian15}  & 15.2              & 25.5              & 32.0              & 36.7              & \vsecondSig{39.8} & \np{4787}            \\
     \bottomrule
@@
 stable subsystems~\autocite{monotoneBiology07}. Finally, given large databases with duplicated
-records, they naturally help solving entity
-resolution~\autocites{Crosslingual07}{DeDup09}{LargeScaleDeDup09}{WeightedER14}.
+records, they naturally help de-deduplicating
+entities~\autocites{Crosslingual07}{DeDup09}{LargeScaleDeDup09}{WeightedER14}.
 
@@
 partitioned in $K$ groups in such a way that links within groups are positive and links across
-groups are negative. Recovering the optimal partition after observing all the signs of a graph is
+groups are negative. Recovering the optimal partition from the observation of all the edge signs is
 called \pcc{}~\autocite{Bansal2002} in the Machine Learning field. Whereas it has been proved to be
@@
 literature review paints a less gloomy landscape. Not only does there exist
-exact~\autocite{Berg2015} and fixed parameters algorithms~\autocites{GoldenCE12}{Fomin2014},
+exact~\autocite{Berg2015} and fixed-parameters tractable algorithms~\autocites{GoldenCE12}{Fomin2014},
 efficient approximation~\autocites{CCPivotConf05}{ChawlaArxiv14} and large scale
@@
 random moves from an ideal case~\autocites{plantedAilon09}{Makarychev2014}. This leads us to develop
-a spanning tree tailored to predicting edge signs in such a setting, which we call \gtx{}. It does
-so by aiming to approximately preserve the distances in the
-graph~\autocites{LowerBound95}{Abraham2012}. We present preliminary results showing that on certain
+a spanning tree tailored to predicting edge signs in an active setting~\autocite{Cesa-Bianchi2012b},
+where the learner is allowed to select its training labels in the most informative way. The idea is
+to ensure that for every edge we will predict, there will be a path in the tree as short as
+possible~\autocites{LowerBound95}{Abraham2012} from which we have queried all the signs. We call
+this structure \gtx{}, as it recursively build stars of stars. 
+We present preliminary results showing that on certain
 kind of graphs, this indeed results in better predictive performance than a competitive baseline.
@@
 partition the nodes in two clusters, set the edge signs accordingly and finally flip a fraction of
-them \uar{}. As showed on \autoref{fig:gtx_xp_pasynthmcc}, \gtx{} always predict signs better than a
-\bfs{} tree rooted at the highest degree node, especially when the noise level is low, and is less
-sensible to the graph size.
+the signs \uar{}. As showed on \autoref{fig:gtx_xp_pasynthmcc}, \gtx{} always predict signs better than a
+\bfs{} tree rooted at the highest degree node, especially when the noise level is low. Furthermore,
+\gtx{} is less sensible to the graph size.
 
@@
   \centering
-  \includegraphics[width=.6\textwidth]{gtx_exp/pasynthmcc}
-  \caption{Synthetic \lpa{} }\label{fig:gtx_xp_pasynthmcc}
+  \includegraphics[width=.8\textwidth]{gtx_exp/pasynthmcc}
+  \caption{The Matthews Correlation Coefficient of \bfs{} and \gtx{} on synthetic \lpa{} graphs as a
+    function of the number of nodes and of the fraction of signs who have been flipped from the
+    hidden two-clustering.  \label{fig:gtx_xp_pasynthmcc}}
 \end{figure}
@@
 Finally, we consider the case where there are more than two possible type of relations, which is
-often called \emph{multilayer graphs}~\autocites{Kivela2014}{multiSurvey14}. In addition, we assume
+often referred to as \emph{multilayer graphs}~\autocites{Kivela2014}{multiSurvey14}. In addition, we assume
 that the graph is attributed, that is every node is described by a vector of numeric features called
-the \emph{profile} of the node. Our assumption is then that links cannot be fully described by a
+the \emph{profile} of the node. Our assumption is then that links cannot be fully described by
 global homophily. Thus we take a more nuanced view, combining partial homophily (that is, nodes are
@@
 of dating websites ---where most users are linked with users of the opposite
-gender~\autocites{homophilyMyspace09}{Tinder16}; spreading innovation ---where meeting people with
-different background and point of view is crucial to favor diversity and
+gender~\autocites{homophilyMyspace09}{Tinder16}; diffusing innovations ---where meeting people with
+different backgrounds and point of views is crucial to favor diversity and
 creativity~\autocite{rogers2003diffusion}; and online news consumption ---where connecting people
 from different sides of the political spectrum helps avoiding echo chambers and fueling a democratic
-debate~\autocite{balancedNews17}. Our goal is to predict to which category a given belong. This is
+debate~\autocite{balancedNews17}. Our goal in this setting remains the same as before: predict the
+type of a given edge. This is
 widely applicable, for multilayer graphs are present  in various fields. In social networks, where
 players of the Pardus massively multilayer online game can have six types of relations (friendship,
-enmity, private message, trading, aggression and bounty)~\autocite{Szell2010} or where photo sharers
+enmity, private message, trading, aggression and bounty)~\autocite{Szell2010}, or where photo sharers
 of the Flickr website can interact in eleven ways (either directly or through comments, tags, groups
@@
 connected if they have co-authored a paper in one the \np{1000} conferences that form the edge
-type~\autocite{communityDBLPbyConf05}, or where \np{5000} SIAM papers can be connected for five
+types~\autocite{communityDBLPbyConf05}, or where \np{5000} SIAM papers can be connected for five
 reasons (abstract similarity, title similarity, keywords similarity, author similarity or
-citation)~\autocite{articlesMultiSim11} in economic networks, where 951 ports are connected by 3
+citation)~\autocite{articlesMultiSim11}. In economic networks, where 951 ports are connected by 3
 kind of ships (bulk dry carriers, container ships and oil tankers)~\autocite{ports3kindofships10} or
-in the International Trade Network, where 162 countries are connected by 96 kind of commodities they
+where 162 countries of the International Trade Network are connected by 96 kind of commodities they
 can exchange~\autocite{worldTradeNetwork10}. And in biology, where genes are connected in a
@@
 
-There exists previous work that leans towards analyzing and characterising edges in such multilayer
-graphs, although they do not share exact our goal of predicting category. For instance, one can
+There exists previous works that lean towards analyzing and characterising edges in such multilayer
+graphs, although they do not share exactly our goal of predicting edge type. For instance, one can
 maximize the probability of a generative model in ego networks with $K$
@@
 whose small-world superposition yield the observed graph~\autocite{Abraham2012a}, discover edge role
-by computing high order topological features~\autocite{ahmed2017roles}, group nodes with common
-attributes and extract frequent links between such group as edge type~\autocite{conceptualLinks12},
+by computing high-order topological features~\autocite{ahmed2017roles}, group nodes with common
+attributes and extract frequent links between such groups as edge types~\autocite{conceptualLinks12},
 or cluster the profiles into a set of affine subspace of $\mathbb{R}^d$~\autocite{SCSurvey11}.
@@
 Our approach is to seek a small number of vectors that, once assigned to every edge, best explain
-the graph (\ie maximize a score function between the edge vector and the profiles of its endpoints).
+the graph (\ie maximize a score function between the vector assigned to an edge  and the profiles of
+this edge's endpoints).
 From this initial formulation, we derive two optimization problems, and initially solve them on
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{assets/raw/Gfixed_over0_step2_7dirs3_ambi.pdf}
-  \caption{Distribution of the Adjusted Mutual Information (the closer to one the better) over
-  \np{600} generation of 7 random directions, comparing our \textsf{optim} method with the $k$-means
+  \includegraphics[width=0.8\linewidth]{assets/raw/Gfixed_over0_step2_7dirs3_ambi.pdf}
+  \caption{Distribution of the Adjusted Mutual Information (the closer to $1$ the better) over
+  \np{600} generations of 7 random directions, comparing our \textsf{optim} method with the $k$-means
 baseline. \label{fig:edge}}
@@
 
-% Surveys are also available about applications of unsigned networks
-% such as data classification [Sen et al. 2008], recommendation [Tang et al. 2013], and
-% information propagation [Chen et al. 2013a]
-
-While our thesis is therefore primarily concerned with learning on graphs, instead of focusing at
+While our thesis is therefore primarily concerned with learning on graphs, instead of focusing on
 nodes (to predict their class, their cluster or to find their representation), we look at edges (\ie

### 2017-09-30 intro/draft.tex
add some refs

@@
 single \emph{friendship} relation, we may want to distinguish between family ties, work ties and
-hobbies ties~\autocite{egoNips12}. Moreover, among one user's \enquote{friends}, there might be some
+hobbies ties~\autocite{LeskovecEgo12}. Moreover, among one user's \enquote{friends}, there might be some
 relations that are actually not positive~\autocite{Yang2012}, whether because of spam, noise or

### 2017-09-30 troll/related.tex
add some refs

@@
 word. This allows the unsupervised extraction of node features that can then used to train
-downstream classifier, see \autocite{nodeEmbeddingsurvey16} for a recent survey. It can also be
+downstream classifier, see
+\autocites{nodeEmbeddingsurvey16}{representationLearning17}{graphEmbeddingSurvey17} for three recent
+surveys. It can also be
 tailored to exploit the specificity of signed graphs.

### 2017-09-30 intro/draft.tex
LT typos

@@
     although it is also possible to include problem supervision when available. Such representation can
-    then be used in downstream tasks, for instance visualisation~\autocite{LINE15} or the
+    then be used in downstream tasks, for instance visualization~\autocite{LINE15} or the
     aforementioned node classification, even when new nodes join the graph after
@@
 meaning that nodes are linked because they are similar. However, many real world situations require
-a more complex model, with several kind of edges and where nodes can be connected if they are
+a more complex model, with several kinds of edge and where nodes can be connected if they are
 similar only on some dimensions, or even if they are dissimilar. For instance, an e-commerce website
@@
 relations that are actually not positive~\autocite{Yang2012}, whether because of spam, noise or
-friendship drift.. Note finally that friendship is only one possible kind of relationship between two
+friendship drift. Note finally that friendship is only one possible kind of relationship between two
 users of Facebook, we could also link them whenever they attend the same event or are tagged in the
@@
 topological features and training classifiers~\autocites{Leskovec2010}{Bayesian15}{wu2016troll}.
-Guided by the idea that every users exhibit trollness to various extent, we propose a simple sign
+Guided by the idea that every user exhibit trollness to various extent, we propose a simple sign
 generative model. This model naturally leads to two batch algorithms, one that probably approximate the
@@
 signed graphs are employed to identify coreference~\autocites{graphicalCoreference04}{Elsner2009} or
-cluster synonym words~\autocite{SignedWordRatings}. They plays a large role in biology, to clusters
+cluster synonym words~\autocite{SignedWordRatings}. They play a large role in biology, to clusters
 genes~\autocite{Ben-Dor99}, identify mutation regions in chromosomes~\autocite{Das2015} or finding
@@
 creativity~\autocite{rogers2003diffusion}; and online news consumption ---where connecting people
-from different sides of the political spectrum helps avoiding echo chambers and fueling a democratic
+from different sides of the political spectrum helps to avoid echo chambers and fuel a democratic
 debate~\autocite{balancedNews17}. Our goal in this setting remains the same as before: predict the
@@
 reasons (abstract similarity, title similarity, keywords similarity, author similarity or
-citation)~\autocite{articlesMultiSim11}. In economic networks, where 951 ports are connected by 3
-kind of ships (bulk dry carriers, container ships and oil tankers)~\autocite{ports3kindofships10} or
-where 162 countries of the International Trade Network are connected by 96 kind of commodities they
+citation)~\autocite{articlesMultiSim11}. In economic networks, where 951 ports are connected by three
+kinds of ships (bulk dry carriers, container ships and oil tankers)~\autocite{ports3kindofships10} or
+where 162 countries of the International Trade Network are connected by 96 kinds of commodities they
 can exchange~\autocite{worldTradeNetwork10}. And in biology, where genes are connected in a
@@
 
-There exists previous works that lean towards analyzing and characterising edges in such multilayer
+There exists previous works that lean towards analyzing and characterizing edges in such multilayer
 graphs, although they do not share exactly our goal of predicting edge type. For instance, one can
@@
 with the result of a baseline approach that simply clusters the pairwise profile vectors using
-$k$-means. As showed in \autoref{fig:edge}, on a graph with \np{500} nodes, \np{1300} edges and 7
+$k$-means. As showed on \autoref{fig:edge}, on a graph with \np{500} nodes, \np{1300} edges and 7
 types of edges (where each node is involved in at most 3 types of relation), we recover the hidden

### 2017-10-02 assets/raw/troll_labeled_gprime.tex
explain better the cutsize with a figure

\documentclass[svgnames,11pt]{standalone}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{multicol}
\usepackage{tikz}
\usepackage{times}
\usepackage{xcolor}
\usetikzlibrary{arrows,automata}
\newcommand{\poslabel}{$1$}
\newcommand{\neglabel}{$0$}
\newcommand{\epslabel}{}
\newcommand{\minuslabel}{$-1$}
\begin{document}

\begin{tikzpicture}[auto, every label/.style={font=\Large}]
\tikzset{>=latex}
\tikzstyle{peers}=[draw,circle,black,minimum width=11mm,inner sep=0]
\tikzstyle{nedge}=[draw,rectangle,black,minimum width=8.5mm,minimum height=8.5mm, inner sep=0]
\tikzstyle{edge}=[line width=0.5pt,color=Black,->]
\tikzstyle{uedge}=[line width=0.5pt,color=Black]
\tikzstyle{phi}=[line width=2.5pt,line cap=round,shorten <= 1pt, shorten >= 1pt]
\tikzstyle{pnode}=[fill=White]
\tikzstyle{elabel}=[color=Black]
\tikzstyle{qnode}=[fill=White]
% \tikzstyle{posc}=[color=blue]
% \tikzstyle{negc}=[color=red]

\tikzstyle{posc}=[color=blue, fill=blue!30!white]
\tikzstyle{negc}=[color=red, fill=red!30!white]
\node[peers,pnode,posc,label={$+1$}] (p1) at (1.3, -1.0) {\LARGE$\mathbf{1_{\mathrm{out}}}$}; 
\node[peers,pnode,negc,label={$-1$}] (p2) at (6.4, -2.0) {\LARGE$\mathbf{2_{\mathrm{out}}}$}; 
\node[peers,pnode,posc,label=190:{$+1$}] (p3) at (1.3, -6.0) {\LARGE$\mathbf{3_{\mathrm{out}}}$}; 
\node[peers,pnode,posc,label={$+1$}] (p4) at (6.4, -5.0) {\LARGE$\mathbf{4_{\mathrm{out}}}$}; 
\node[peers,qnode,posc,label={$+1$}] (q1) at (0.0, -2.0) {\LARGE$\mathbf{1_{\mathrm{in}}}$}; 
\node[peers,qnode,posc,label={$+1$}] (q2) at (5.1, -1.0) {\LARGE$\mathbf{2_{\mathrm{in}}}$}; 
\node[peers,qnode,posc,label={$+1$}] (q3) at (0.0, -5.0) {\LARGE$\mathbf{3_{\mathrm{in}}}$}; 
\node[peers,qnode,posc,label=110:$+1$] (q4) at (5.1, -6.0) {\LARGE$\mathbf{4_{\mathrm{in}}}$}; 
 \node[nedge,posc] (e1) at (3.2, -1.0) {\LARGE$\mathbf{1}$}; 
 \node[nedge,negc] (e2) at (3.2, -2.0) {\LARGE$\mathbf{-1}$}; 
 \node[nedge,negc] (e3) at (5.2, -3.7) {\LARGE$\mathbf{-1}$}; 
 \node[nedge,posc] (e4) at (1.2, -3.7) {\LARGE$\mathbf{1}$}; 
 \node[nedge,negc] (e5) at (3.2, -3.7) {\LARGE$\mathbf{-1}$}; 
 \node[nedge,posc] (e6) at (3.2, -6.0) {\LARGE$\mathbf{1}$};

\draw[uedge] (p1) edge [bend left=0] node [elabel,below] {\epslabel} (e1);
\draw[uedge] (e1) edge [bend left=0] node [elabel,below] {\epslabel} (q2);
\draw[uedge] (p2) edge [bend left=0] node [elabel,above,yshift=-1mm,pos=.25] {\epslabel} (e2);
\draw[uedge,phi] (e2) edge [bend left=0] node [elabel,above,yshift=-1mm] {\epslabel} (q1);
\draw[uedge] (p2) edge [bend left=0] node [elabel] {\epslabel} (e3);
\draw[uedge,phi] (e3) edge [bend left=0] node [elabel] {\epslabel} (q4);
\draw[uedge] (p3) edge [bend left=0] node [elabel] {\epslabel} (e4);
\draw[uedge] (e4) edge [bend left=0] node [elabel,above,xshift=1mm,pos=.35] {\epslabel} (q1);
\draw[uedge,phi] (p3) edge [bend left=0] node [elabel,above,xshift=-1mm] {\epslabel} (e5);
\draw[uedge,phi] (e5) edge [bend left=0] node [elabel,below,xshift=1mm,pos=0.3] {\epslabel} (q2);
\draw[uedge] (p3) edge [bend left=0] node [elabel] {\epslabel} (e6);
\draw[uedge] (e6) edge [bend left=0] node [elabel] {\epslabel} (q4);

\end{tikzpicture}
\end{document}

### 2017-10-02 troll/introduction.tex
mark content added from the aistats paper
explain better the cutsize with a figure

@@
 user $v$. Such an interaction can be positive (to praise, to support, to trust, to befriend) or
-negative (to criticize, to oppose, to distrust, to befoe).
+negative (to criticize, to oppose, to distrust, to make enemy with).
 
@@
 data, we show the competitiveness of our approach in terms of both prediction performance
-(especially in the regime when training data are scarce) and scalability.
+(especially in the regime where training data are scarce) and scalability.
 

### 2017-10-02 troll/notation_and_prel.tex
mark content added from the aistats paper
explain better the cutsize with a figure

@@
 
-The first reduction, called $G \rightarrow G'$, builds an \emph{undirected} graph $G' = (V',E')$ as
+The first reduction builds an \emph{undirected} graph $G' = (V',E')$ as
 follows. Each node $u \in V$ has two copies in $V'$, call them $\iin$ and $\iout$. Each directed
@@
 $e_{u,v} \in V'$, so that $G'$ is a graph $G'(Y) = (V'(Y),E')$ with partially-labeled nodes. The
-second reduction, called $G \rightarrow G''$, builds an \emph{undirected and weighted} graph $G'' =
+second reduction builds an \emph{undirected and weighted} graph $G'' =
 (V'',E'')$. Specifically, we have $V'' \equiv V'$ and $E'' \supset E'$, where the set $E''$ also
@@
 \autoref{ssec:troll_lprop_weights} an analytic justification for this weights choice in the context
-of Label Propagation. Finally, as in the $G \rightarrow G'$ reduction, if $G = G(Y)$ is edge
+of Label Propagation. Finally, as in the first reduction, if $G = G(Y)$ is edge
 labeled, then this labeling transfers to the subset of nodes $e_{u,v} \in V''$. Graph $G'$, which
 will not be used in the following, is an intermediate structure between $G$ and $G''$ and provides a
-conceptual link to the standard cutsize measure in node sign classification. \autoref{f:etnr}
-illustrates these two reductions.
+conceptual link to the standard cutsize measure in node sign classification, as we will describe
+shortly. \autoref{f:etnr} illustrates these two reductions. Note that because $|V''|=2|V|$ and
+$|E''|=3|E|$, the reduction yielding $G''$ can be computed in linear time, and does not require the
+knowledge of the edge label, meaning it can be done in parallel with potential label querying.
+Furthermore, it can be updated incrementally if new nodes or edges are added to the original graph
+$G$.
 
@@
   \begin{subfigure}[b]{0.28\textwidth}
-    \centering \includegraphics[height=.10\textheight]{raw/g_latex-crop} \caption{}
+    \centering \includegraphics[height=.10\textheight]{raw/g_latex-crop}
+    \caption{\label{fig:troll_reduction_g}}
   \end{subfigure}~
@@
     \textbf{(a)} A directed edge-labeled graph $G$. \textbf{(b)} Its corresponding graph $G'$
-    resulting from the $G\rightarrow G'$ reduction. The square nodes in $G'$ correspond to the
+    resulting from the first reduction we describe. The square nodes in $G'$ correspond to the
     edges in $G$, and carry the same labels as their corresponding edges. On the other hand, the
@@
     no outgoing or no incoming edges: for instance nodes $3$ and $4$ in $G$. \textbf{(c)} The
-    weighted graph resulting from the $G\rightarrow G''$ reduction.
+    weighted graph resulting from the second reduction we describe.
   }
@@
 use the many algorithms designed for the better studied problem of node sign prediction. Second,
-the reduction $G\rightarrow G''$ with the specific choice of edge weights is designed to make the
+the reduction yielding $G''$, along its specific choice of edge weights, is designed to make the
 Label Propagation solution approximate the maximum likelihood estimator associated with our
@@
 generative model which we describe in the next section, and our label regularity measure is a
-quadratic function (denoted by $\Psi^2_{G''}(Y)$ and defined in \autoref{s:exp}), related to this
-model. $\Psi^2_{G''}(Y)$ is small just when all nodes in $G$ tend to be either troll or
-trustworthy.
+quadratic function (denoted by $\Psi^2_{G''}(Y)$ and defined in \autoref{s:exp} as a regularized
+energy function of $G''$\footnote{If we denote the value of the square node $e_{u,v}$ of $G''$ as
+$\frac{1+\yuv}{2}$, the value of the circle node $u_{\mathrm{out}}$ as $p_u$ and the value of the
+circle node $v_{\mathrm{in}}$ as $q_v$, we have that
+$\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \sum_{(u,v) \in E} \left(\frac{1+\yuv}{2} - \frac{p_u+q_v}{2}
+\right)^2$, where $(\bp,\bq) =\{p_u, q_u\}_{u=1}^{|V|}$.})
+related to this model. $\Psi^2_{G''}(Y)$ is small just when all nodes in $G$ tend to be either troll
+or trustworthy.
 
@@
   performance simply through the total number of mistakes $M_A(Y)$ it makes over $G(Y)$ when the
-  worst possible presentation order of the edge labels in $Y$ is selected by the adversary. As is
-  standard, we contrast $M_A(Y)$ to some kind of \emph{regularity} measure $\Psi_G(Y)$ of the
+  worst possible presentation order of the edge labels in $Y$ is selected by the adversary. As it is
+  standard practice, we contrast $M_A(Y)$ to some kind of \emph{regularity} measure $\Psi_G(Y)$ of the
   labeling $Y$ over $G$, so that we are in fact aimed at bounding the (cumulative) regret \( M_A(Y)
@@
 Then we define $\Psi_G(Y) = \min\big\{\Psiin(Y),\Psiout(Y)\big\}$. The two measures
-$\Psi^2_{G''}(Y)$ and $\Psi_G(Y)$ are conceptually related and their value on real data is quite
+$\Psi_G(Y)$ and $\Psi^2_{G''}(Y)$ are conceptually related and their value on real data is quite
 similar (see \autoref{tab:all_mcc} \vpageref{s:exp}).
@@
 Let us first introduce the notion of cutsize of an undirected node-labeled graph $G'(Y)$, which is
-the number of edges in $G'$ connecting nodes having mismatching labels.
+the number of edges in $G'$ connecting nodes that have mismatching labels.
   Now, because the only nodes in $G'$ we are
@@
   sign prediction problem on a subset of $V'$. As for the remaining nodes in $V'$ (the circles in
-  \autoref{fig:troll_reduction_gprime}), we are free to assign dummy labels so as to minimize the corresponding mistake
-  bound over $G'$. It is straightforward to see that if we set the labeling $Y$ on $V'$ in such a
-  way that $y_{\iout} = +1$ if $tr(u) < 0$, and $-1$, otherwise, and $y_{\iin} = +1$ if $un(u) <
-  0$, and $-1$, otherwise, then the cutsize of $G'(Y)$ equals  $\Psiin(Y) + \Psiout(Y)$.%
+  \autoref{fig:troll_reduction_gprime}), we are free to assign arbitrary labels so as to minimize the corresponding mistake
+  bound over $G'$. Such an assignment is constructed as follow: we set the labeling $Y$ on $V'$ in such a
+  way that $y_{\iout} = +1$ if $tr(u) \leq \shalf$ and $-1$ otherwise. Similarly, we have $y_{\iin} = +1$ if
+  $un(u) \leq \shalf$,
+  and $-1$ otherwise. For instance, given the trollness and untrustworthiness values for the graph $G$ of
+  \autoref{fig:troll_reduction_g} reported in \autoref{tab:troll_psi}, following this construction
+  would result in a label version of $G'$ showed on \autoref{fig:troll_psi}.
+\begin{marginfigure}
+	\centering
+	\includegraphics[width=0.98\linewidth]{assets/raw/troll_labeled_gprime.pdf}
+	\caption{The node-labeling of $G'$ illustrating the relation between the edge label
+	  complexity and the cutsize. Here the four edges part of the cut are in heavy black.
+	\label{fig:troll_psi}}
+\end{marginfigure}
+  We see that by labeling circle nodes by the majority label of the square nodes there are connected
+  to, we have that the cutsize of $G'(Y)$ equals  $\Psiin(Y) + \Psiout(Y)$.%
   \footnote{In fact, for the sake of
@@
   $\Psiin(Y)$, respectively.}
-  A similar reasoning apply in the batch case with the $\Psi^2_{G''}(Y)$ measure.
+  A similar reasoning apply in the batch case with the $\Psi^2_{G''}(Y)$ measure, which can be seen
+  as a soft quadratic version of the cutsize.
+
+\begin{table}[htb]
+  \centering
+  \caption{The label regularity values for the nodes of the example $G$ graph of
+  \autoref{fig:troll_reduction_g}.  \label{tab:troll_psi}}
+  \begin{tabular}{lccccr}
+    \toprule
+    node $u$       & $1$    & $2$    & $3$               & $4$    & total \\
+    \midrule
+    $\Psiout(u,Y)$ & $0$    & $0$    & $1$               & $0$    & $1$   \\
+    $\Psiin(u,Y)$  & $1$    & $1$    & $0$               & $1$    & $3$   \\
+    $tr(u)$        & $0$    & $1$    & $\nicefrac{1}{3}$ & $1$    & ---   \\
+    $un(u)$        & \shalf & \shalf & \shalf            & \shalf & ---   \\
+    \bottomrule
+  \end{tabular}
+\end{table}

### 2017-10-02 troll/supp_body.tex
add new dataset, competitor and results
only one section for troll additional material
explain better the cutsize with a figure

@@
-\section{Proofs from \autoref{s:algbatch}}
+% \section{Proofs from \autoref{s:algbatch}}
+\section{Additional material}
+\subsection{Proofs from \autoref{s:algbatch}}
 
-\subsection{Proof of \autoref{t:active}}
+\subsubsection{Proof of \autoref{t:active}}
 \label{ssec:troll_proof_bayes}
@@
 \]
-Applying \autoref{l:hoeff-wr} to the set $\theset{\tfrac{p_i+q_j}{2}}{j \in \NNout(u)}$, and using
+Applying \autoref{l:hoeff-wr} to the set $\theset{\tfrac{p_u+q_v}{2}}{v \in \NNout(u)}$, and using
 our choice of $Q$, we get that $\big|\muhat_p(u) - \mu_p(u)\big| \le \ve$ holds with probability at
@@
 
-\subsection{Derivation of the maximum likelihood equations}
+\subsubsection{Derivation of the maximum likelihood equations}
 \label{ssec:troll_mle_deriv}
 
-Recall that the training set $\trainset = \theset{\big(u_t,v_t),y_{u_t,v_t}\big)}{t=1,\dots,m}$
+Recall that the training set $\trainset = \theset{\big(u_k,v_k),y_{u_k,v_k}\big)}{k=1,\dots,m}$
 is drawn uniformly at random from $E$ without replacement. 
@@
 \begin{align*}
-  \Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) = 
+  \Pr&\left(E_0\,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right) = 
   \frac{1}{\binom{|E|}{m}\,m!}\prod_{k=1}^m \left(\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=+1}}\,
@@
 
-so that $\log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)$ is proportional to
+so that $\log \Pr\left(E_0\,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right)$ is proportional to
 
@@
 \begin{equation*}
-  \frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial p_{\ell}}
+  \frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right)}{\partial p_{\ell}}
   = \sum_{k=1}^m 
@@
 \begin{align*}
-  \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) 
+  \Pr\left(E_0\,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right) 
 =
@@
 \begin{align*}
-\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial q_{\ell}}
+\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right)}{\partial q_{\ell}}
 = \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=+1}}{p_{u_k}+q_{\ell}}\,
@@
 
-\subsection{Label propagation on $G''$}
+\subsubsection{Label propagation on $G''$}
 \label{ssec:troll_lprop_weights}
@@
 \begin{equation*}
-  \frac{1}{16} \sum_{u,v \in E} \Bigl[ \eps\left(y_{u,v} - p_u'\right)^2 + \eps\left(y_{u,v} - q_v'\right)^2 + (p'_i+q_v')^2 \Bigr]
+  \frac{1}{16} \sum_{u,v \in E} \Bigl[ \eps\left(y_{u,v} - p_u'\right)^2 + \eps\left(y_{u,v} - q_v'\right)^2 + (p'_u+q_v')^2 \Bigr]
 \end{equation*}
@@
 
-\section{Proof from \autoref{s:algonline}}
+\subsection{Proof from \autoref{s:algonline}}
 
@@
 
-\section{Further Experimental Results}
+\subsection{Further Experimental Results}
 
@@
 
-\begin{table}[bt]
-  \centering
-  \caption{Normalized logistic regression coefficients averaged over 12 runs (with one standard
-  deviation) \label{tab:coeff}}
-  \small
-  \begin{tabular}{lrccc}
-    \toprule
-    & $\frac{|\trainset{}|}{|E|}$ &              $w_2'$ & $\tau'$ \\
-    % \multirow{5}{*}{\aut{}} & $3\%$  &  $0.909 \pm 0.07$  &  $0.592 \pm 0.04$  \\
-    %                         & $9\%$  &  $0.985 \pm 0.04$  &  $0.707 \pm 0.03$  \\
-    \multirow{5}{*}{\aut{}} & $5\%$  &  $0.965 \pm 0.04$  &  $0.662 \pm 0.03$  \\
-                            & $10\%$ &  $0.983 \pm 0.03$  &  $0.705 \pm 0.02$  \\
-                            & $15\%$ &  $1.001 \pm 0.03$  &  $0.729 \pm 0.03$  \\
-                            & $20\%$ &  $1.013 \pm 0.02$  &  $0.747 \pm 0.02$  \\
-                            & $25\%$ &  $1.011 \pm 0.02$  &  $0.746 \pm 0.01$  \\
-    \midrule
-    % \multirow{5}{*}{\wik{}} & $3\%$  &  $0.919 \pm 0.04$  &  $0.666 \pm 0.02$  \\
-    %                         & $9\%$  &  $0.940 \pm 0.03$  &  $0.726 \pm 0.01$  \\
-    \multirow{5}{*}{\wik{}} & $5\%$  &  $0.920 \pm 0.02$  &  $0.691 \pm 0.02$  \\
-                            & $10\%$ &  $0.940 \pm 0.01$  &  $0.730 \pm 0.01$  \\
-                            & $15\%$ &  $0.947 \pm 0.01$  &  $0.741 \pm 0.01$  \\
-                            & $20\%$ &  $0.963 \pm 0.01$  &  $0.760 \pm 0.01$  \\
-                            & $25\%$ &  $0.962 \pm 0.02$  &  $0.764 \pm 0.01$  \\
-    \midrule
-    % \multirow{5}{*}{\sla{}} & $3\%$  &  $1.012 \pm 0.02$  &  $0.669 \pm 0.01$  \\
-    %                         & $9\%$  &  $1.014 \pm 0.01$  &  $0.703 \pm 0.01$  \\
-    \multirow{5}{*}{\sla{}} & $5\%$  &  $1.024 \pm 0.02$  &  $0.693 \pm 0.01$  \\
-                            & $10\%$ &  $1.017 \pm 0.01$  &  $0.705 \pm 0.01$  \\
-                            & $15\%$ &  $1.007 \pm 0.01$  &  $0.707 \pm 0.01$  \\
-                            & $20\%$ &  $1.002 \pm 0.00$  &  $0.710 \pm 0.00$  \\
-                            & $25\%$ &  $0.995 \pm 0.01$  &  $0.712 \pm 0.00$  \\
-    \midrule
-    % \multirow{5}{*}{\epi{}} & $3\%$  &  $1.124 \pm 0.02$  &  $0.786 \pm 0.02$  \\
-    %                         & $9\%$  &  $1.066 \pm 0.01$  &  $0.783 \pm 0.01$  \\
-    \multirow{5}{*}{\epi{}} & $5\%$  &  $1.099 \pm 0.02$  &  $0.791 \pm 0.02$  \\
-                            & $10\%$ &  $1.059 \pm 0.01$  &  $0.782 \pm 0.01$  \\
-                            & $15\%$ &  $1.037 \pm 0.01$  &  $0.774 \pm 0.01$  \\
-                            & $20\%$ &  $1.018 \pm 0.01$  &  $0.765 \pm 0.01$  \\
-                            & $25\%$ &  $1.010 \pm 0.01$  &  $0.763 \pm 0.01$  \\
-    \midrule
-    % \multirow{5}{*}{\kiw{}} & $3\%$  &  $1.075 \pm 0.02$  &  $0.851 \pm 0.02$  \\
-    %                         & $9\%$  &  $1.041 \pm 0.01$  &  $0.872 \pm 0.01$  \\
-    \multirow{5}{*}{\kiw{}} & $5\%$  &  $1.047 \pm 0.02$  &  $0.853 \pm 0.01$  \\
-                            & $10\%$ &  $1.038 \pm 0.01$  &  $0.872 \pm 0.01$  \\
-                            & $15\%$ &  $1.025 \pm 0.01$  &  $0.876 \pm 0.01$  \\
-                            & $20\%$ &  $1.012 \pm 0.01$  &  $0.874 \pm 0.01$  \\
-                            & $25\%$ &  $1.007 \pm 0.01$  &  $0.874 \pm 0.01$  \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+\input{all_coeff}

### 2017-10-02 troll/algs_batch.tex
explain better the first troll batch algo

--- 2017-09-12 troll/algs_batch.tex
@@
 Given $G(Y) =(V,E(Y))$, in the batch setting we have at our disposal a training set \trainset{} of
-labeled edges from $E(Y)$. We want to build a predictive model for the labels of the remaining
+labeled edges from $E(Y)$. Formally, we have a training set $\trainset
+= \big((u_1,v_1),y_{u_1,v_1}), ((u_2,v_2),y_{u_2,v_2}), \ldots, ((u_m,v_m),y_{u_m,v_m}\big)$, that
+has been drawn from $E\times Y$ \uar{} with replacement, with
+$m = |E_0|$. Then a reasonable approach to approximate $y^*(u,v)$ would be to resort to a maximum
+ We want to build a predictive model for the labels of the remaining
 edges. We present two algorithms to do so, which both compute estimates of the all parameters $p$
@@
 $\htr(u)$ and $\hun(u)$ the trollness and the untrustworthiness of node $u$ when both are computed
-on the subgraph induced by the training edges. We now design and analyze an edge classifier of the
-form
-
+on the subgraph induced by the training edges.
+Recall that the Bayes optimal predictor classifies an edge \euv{} using the following rule:
+$y^*(u,v) = \sgn\left( \frac{p_u+q_v}{2} - \frac{1}{2} \right)$. Our approximation thus consists in
+using the quantities $\htr(u)$ and $\hun(u)$ to estimate $\frac{p_u+q_v}{2}$. This results in the
+following rule:
 \begin{equation}
   \label{eq:predictor}
-  \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) -\tau - \tfrac{1}{2}\Big)~,
+  \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) - \tau - \tfrac{1}{2}\Big)~,
 \end{equation}
+where $\tau \ge 0$ is the only parameter to be trained. We now give an intuition to justify this
+equation, and defer the technical arguments to end of the chapter,
+\vpageref{ssec:troll_proof_bayes}, where we will formalize what it means for a quantity to be
+\enquote{close} to another. Note first that $1-\htr(u) = \frac{\hdout^+(u)}{\hdout(u)}$ is the
+empirical mean\footnote{The hat symbol denote quantities computed solely on the training set.} of
+the probability of a random edge outgoing from $u$ to be positive, and is therefore \enquote{close}
+to $\frac{1}{2}\,\left(p_u + \qbar_u\right)$ according to \eqref{e:pout}. By the same reasoning,
+$1-\hun(v) \approx \frac{1}{2}\,\left(q_v + \pbar_v\right)$ in accordance with \eqref{e:pin}. At
+this stage, $\left( 1 - \htr(u) \right) + \left( 1 - \hun(v) \right) - \frac{p_u+q_v}{2} \approx
+\lhalf \left( \pbar_v + \qbar_u \right)$. Since $\pbar_v$ is a sample mean of i.i.d.~$[0,1]$-valued
+random variables independently drawn from the prior marginal $\int_0^1 \mu\big(\cdot, q\big) dq$, it
+concentrates around its expectation $\mu_p$. Likewise, $\qbar_u \approx \mu_q$. Now we seek an
+estimation of $ \lhalf \left( \pbar_v + \qbar_u \right) \approx \lhalf \left( \mu_p + \mu_q
+\right)$. For that, we compute the expected number of positive edges in the graph, which, as in the
+previous section, is the expected value of the random variable $Z = \sum_{\euv \in E} \Ind{\yuv =
++1}$. Let $V_{\mathrm{out}}$ be the subset of node of $V$ with at least one outgoing edge, and
+define similarly $V_{\mathrm{in}}$ as $\theset{v\in V}{\din(v) > 0}$. According to our generative
+model, $\E\left(Z \right) = \sum_{\euv \in E} \frac{p_u+q_v}{2} = \sum_{u \in V_{\mathrm{out}}}
+\left( \sum_{v \in \NNout(u)} \frac{p_u+q_v}{2} \right)$. Observe that
+\begin{align*}
+  \sum_{u \in V_{\mathrm{out}}} \left( \sum_{v \in \NNout(u)} p_u \right) &=
+  \sum_{v \in V_{\mathrm{in}}} \left( \sum_{u \in \NNin(v)} p_u \right) =
+  \sum_{v \in V_{\mathrm{in}}} \din(v) \pbar_v \approx
+  \sum_{v \in V_{\mathrm{in}}} \din(v) \mu_p = |E| \mu_p \label{eq:troll_mup}\quad \text{and} \\
+  \sum_{u \in V_{\mathrm{out}}} \left( \sum_{v \in \NNout(u)} q_v \right) &=
+  \sum_{u \in V_{\mathrm{out}}} \dout(u) \qbar_u \approx
+  \sum_{u \in V_{\mathrm{out}}} \dout(u) \mu_q = |E| \mu_q \,.% \label{eq:troll_muq} \,.
+\end{align*}
+Therefore, letting $\tau$ be the fraction of positive edge in the graph, we have that $\tau =
+\frac{\E\left(Z \right)}{|E|} \approx \lhalf \left( \mu_p + \mu_q \right)$. This means that if
+$\tauhat$ is the quantity $\tau$ computed on the training, then $\tauhat$ is close to $\left(
+\pbar_v + \qbar_u \right)$. Putting together all this approximations thus provide an informal
+justification of the rule \eqref{eq:predictor}.
 
-where $\tau \ge 0$ is the only parameter to be trained. Despite its simplicity, this classifier
+We are now ready to describe our algorithm, which follows naturally from the formula derived above.
+It takes as input a training set \trainset{} drawn at random with replacement and outputs a sign
+prediction for all the edges in $E\setminus \trainset$. We call it \usrule{}, which stands for Bayes
+Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness.
+\iffalse
+Despite its apparent simplicity, \usrule{}
 works reasonably well in practice, as demonstrated by our experiments (see \autoref{s:exp}).
@@
 the same for all edges $(u,v) \in E$.
-
-\iffalse
 
@@
 \sgn\big(\hdeltaout(u) + \hdeltain(v) - \tfrac{1}{2} - \tauhat \big) $.
-\fi
 
@@
 algorithm performs the following steps:
+\fi
 
 \begin{enumerate}[label=\textbf{\arabic*.}]
+	\item For each $u \in V$, let $\htr(u) = \hdout^-(u)/\hdout(u)$, \ie{} the fraction of
+	  negative edges found in $\Nout(u) \cap \trainset{}$.
+
 	\item For each $v \in V$, let $\hun(v) = \hdin^-(v)/\hdin(v)$, \ie{} the fraction of
@@
 
-	\item For each $u \in V$, let $\htr(u) = \hdout^-(u)/\hdout(u)$, \ie{} the fraction of
-	  negative edges found in $\Nout(u) \cap \trainset{}$.
+	\item Let $\tauhat$ be the fraction of positive edges in $E_L\cap \trainset{}$ (this set
+	  $E_L$ is a technical requirement that we shall define shortly).
 
-	\item Let $\tauhat$ be the fraction of positive edges in $E_L\cap \trainset{}$.
-
-	\item Any remaining edge $(u,v) \in E\setminus \trainset{}$ is predicted as
+	\item Any edge $(u,v) \in E\setminus \trainset{}$ is predicted as
 		$$
@@
 
-The next result shows that the above algorithm can approximate the Bayes optimal predictor on nodes
-whose in-degree and out-degree is not too small.
+Despite its apparent simplicity, \usrule{}
+works reasonably well in practice, as demonstrated by our experiments (see \autoref{s:exp}).
+Moreover, unlike previous edge sign prediction methods for directed graphs, our classifier comes
+with a rigorous theoretical motivation, since it approximates the Bayes optimal classifier
+$y^*(u,v)$ with respect to the generative model defined in \autoref{s:gen}. Indeed, we quantify this
+approximation on nodes whose in-degree and out-degree are not too small in the next result. However,
+this first requires a few extra assumption on the training set. Namely, given parameters $Q$ (a
+positive integer) and $\alpha = \nicefrac{|\trainset|}{|E|} \in (0,1)$, we assume there exists a set
+$E_L \ss E$ of size $\tfrac{2Q}{\alpha}$ where each vertex $u \in V$ appearing as an endpoint of
+some edge in $E_L$ occurs at most once as origin ---\ie{} \euv{}--- and at most once as destination
+---\ie{} \evu{}. While the definition of $E_L$ is not immediately intuitive, this set is needed to
+find an estimate $\tauhat$ of $\tau$ in~\eqref{eq:predictor} during Step~3 of \usrule{}, and its
+definition allows us to applying an Hoeffding bound on independent variables. Any undirected
+matching of $G$ of size $O(\log|V|)$ can be used, obtained for instance by the blossom
+algorithm~\autocite{matching65}. In practice, however, we never computed $E_L$, and estimated $\tau$
+on the entire training set \trainset{} (instead of $E_L \cap \trainset$).
 
@@
   If the algorithm is run with parameter $Q = \Omega(\log|V|)$, and $\alpha \in (0,1)$ such that the
-  above assumptions are satisfied, then $\yhat(u,v) = y^*(u,v)$ holds with high probability
+  above assumption about $E_L$ is satisfied, then $\yhat(u,v) = y^*(u,v)$ holds with high probability
   simultaneously for all test edges $(u,v) \in E$ such that $\dout(u),\din(v) = \Omega(\log|V|)$, and
@@
 While we defer the full proof of \autoref{t:active} to the additional material at the end of this
-chapter (\autoref{ssec:troll_proof_bayes}), here we give a sketch of the method used, focusing on
-parameter $p$. First, by \autoref{l:bins} on negatively associated random variables (stated
+chapter (\autoref{ssec:troll_proof_bayes}), here we give a sketch of the method used.
+First, by \autoref{l:bins} on negatively associated random variables (stated
 \vpageref{l:bins}), we show that with our choice of $Q$, for $\theta=\frac{2Q}{\alpha}$ and for any
-$u$ having enough out neighbors (\ie{} $\NNout(u) \ge \theta$), at least $Q$ edges of $\Nout(u)$
-are in the training set with high probability.
-We use these $Q$ edges to compute $1-\htr(u)$, the empirical probability of drawing a $+1$-labeled
-edge from $\Nout(i)$. We show that it is close to the real probability, which according to equation
-\eqref{e:pout} is equal to $\frac{1}{2}\,\left(p_u + \qbar_u\right)$. $\qbar_u$ being a sample mean
-of i.i.d.~$[0,1]$-valued random variables independently drawn from the prior marginal $\int_0^1
-\mu\big(p,\cdot\big) dp$, it concentrates around its expectation $\mu_q$.
-
-By following the same reasoning on the parameters $q$ and using an union bound, we then show that
-$\left(1-\htr(u)\right) + \left(1-\hun(v)\right) - \frac{\mu_p+\mu_q}{2}$ is close to
-$\frac{p_u+q_v}{2}$ simultaneously for all edges. Another set of concentration arguments then show
-that our estimate $\tauhat$ is close to $\frac{\mu_p+\mu_q}{2}$. Combining these two facts ends the
-proof.
+$u$ having enough out neighbors (\ie{} $\dout(u) \ge \theta$), at least $Q$ edges of $\Nout(u)$
+are in the training set with high probability. Likewise, for any $v$ with enough in neighbors (\ie
+$\din(v) \ge \theta$), at least $Q$ edges of $\Nin(v)$ are in \trainset with high probability.
+With this number of samples, we then show a chain of concentration results roughly following the
+outline given earlier to justify the equation~\eqref{eq:predictor}, culminating in proving that 
+for any $0 < \ve < \tfrac{1}{16}$,
+\[ \left| \left( 1 - \htr(u) \right) + \left( 1 - \hun(v) \right) - \tauhat
+  - \frac{p_u+q_v}{2} \right| \le 8\ve \]
+simultaneously holds with high probability for each $(u,v) \in \theset{(u,v) \in E}{\din(v) \ge
+  \theta,\, \dout(u) \ge \theta} \setminus E_0$.
 
@@
 
-The approach leading to \autoref{t:active} needs the graph to be sufficiently dense. At first sight,
-having $Q = \Omega(\log|V|)$ training edges per nodes does not sound that like such a stringent
-requirement. Consider for instance Facebook ---which is neither signed nor directed though. It has
-two billion
-users~\bottomfootnote{\url{https://investor.fb.com/investor-news/press-release-details/2017/Facebook-Reports-Second-Quarter-2017-Results/}},
+The approach leading to \autoref{t:active} requires the graph to be sufficiently dense. At first sight,
+having $Q = \Omega(\log|V|)$ training edges per nodes appears to be a reasonable assumption.
+Consider for instance Facebook ---which is neither signed nor directed though. It has two billion
+users as of 2017~\bottomfootnote{%
+\url{https://investor.fb.com/investor-news/press-release-details/2017/Facebook-Reports-Second-Quarter-2017-Results/}},
 each of them having $155\approx 7.2\log|V|$ friends on average~\autocite{facebookFriend16}. However,
-the constant hidden in the $\Omega$ notation, and needed to guarantee a good behavior, is much
-larger than $7.2$. We will nonetheless see in the experiments that we can still apply the resulting
-algorithm (we call it \usrule{}) with satisfying results. Additionally, and in order to address this
-density limitation, we now introduce a second method based on label propagation.
+the constant in the $\Omega$ notation is a trade off between the number of edges to sample per
+node, and the quality guarantee of the $\frac{p_u+q_v}{2}$ estimation. In the Facebook example,
+having a good guarantee that holds simultaneously for all the test edges might require more than
+$7.2\log|V|$ samples per node. We will nonetheless see in the experiments that we can still apply
+\usrule{} with satisfying results, especially since its simplicity makes it very scalable.
+Additionally, and in order to sidestep this density limitation, we now introduce a second method
+based on label propagation.
 
@@
 We now proceed to solve \eqref{e:quadratic} via label propagation~\autocite{LabelPropa03} on the
-graph $G''$ obtained through the $G \rightarrow G''$ reduction of \autoref{s:prel}.
+graph $G''$ obtained through the second reduction of \autoref{s:prel}.
 However, because of the presence of negative edge weights in $G''$, we first have to symmetrize\footnote{%

### 2017-10-02 troll/generative.tex
explain better the first troll batch algo

@@
 Notice that the Bayes optimal prediction for $\yuv$ is $y^*(u,v) = \sgn\big(\eta(u,v) -
-\tfrac{1}{2}\big)$, where $\eta(u,v) = \Pr\big( \yuv = 1 \big)$. Moreover, the probability of
+\tfrac{1}{2}\big)$, where $\eta(u,v) = \Pr\big( \yuv = 1 \big)$. Moreover, once all the signs have
+been generated, we can compute the expected number of positive edges outgoing from $u$ as the
+expected value of the random variable $P_u = \sum_{v\in \NNout(u)} \Ind{\yuv = +1}$. Using this, we
+have that the probability of
 drawing at random a $+1$-labeled edge from $\Nout(u)$ equals
 \begin{equation}\label{e:pout}
+	\frac{\E\left(P_u\right)}{\dout(u)} =
+	\frac{1}{\dout(u)} \sum_{v\in \NNout(u)} \frac{p_u + q_v}{2} =
 	\frac{1}{2}\,\Biggl(p_u + \frac{1}{\dout(u)} \sum_{v\in \NNout(u)} q_v \Biggl) =
@@
 \end{equation}
-where $\qbar_u = \frac{1}{\dout(u)} \sum_{v\in \NNout(u)} q_v$ is the average $q$ value of $u$ out
-neighbors.
+where $\qbar_u = \frac{1}{\dout(u)} \left( \sum_{v\in \NNout(u)} q_v \right)$ is the average $q$
+value of $u$ out neighbors.
 Similarly, the probability of drawing at random a $+1$-labeled edge from $\Nin(v)$ equals

### 2017-10-02 troll/algs_batch.tex
a tad more explaination on the label prop

@@
 
-For simplicity, assume the joint prior distribution $\mu(p,q)$ is uniform over $[0,1]^2$ with
+Remember we suppose that the training set \trainset{} has been drawn \uar{} with replacement, with
+$m = |E_0|$. Then a reasonable approach to approximate $y^*(u,v)$ would be to resort to a maximum
+likelihood estimator of the parameters $\{p_u, q_u\}_{u=1}^{|V|}$ based on \trainset{}.
+If we further assume, in order to make the computation more tractable, that
+the joint prior distribution $\mu(p,q)$ is uniform over $[0,1]^2$ with
 independent marginals,\footnote{As we will see, in real data, around $80\%$ if the edges are
 positive, meaning this assumption of uniformity over $[0,1]^2$ is unlikely to fully holds, for
-otherwise the signs would more balanced} and suppose that we draw at random without replacement the
-training set $E_0
-= \big((u_1,v_1),y_{u_1,v_1}), ((u_2,v_2),y_{u_2,v_2}), \ldots, ((u_m,v_m),y_{u_m,v_m}\big)$, with
-$m = |E_0|$. Then a reasonable approach to approximate $y^*(u,v)$ would be to resort to a maximum
-likelihood estimator of the parameters $\{p_i, q_i\}_{i=1}^{|V|}$ based on \trainset{}.
-As showed in the supplementary material \vpageref{ssec:troll_mle_deriv}, the gradient of the
-log-likelihood function w.r.t.\ $\{p_i, q_i\}_{i=1}^{|V|}$ satisfies
+otherwise the signs would more balanced}
+we show in the supplementary material \vpageref{ssec:troll_mle_deriv} that for $\ell \in \{1,
+\ldots, |V|\}$ the gradient of the log-likelihood function with respect to a given $p_\ell$ and
+$q_\ell$ satisfies
 
 \begin{equation}\label{e:mlp}
-\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_i, q_i\}_{u=1}^{|V|}\right)}{\partial p_{\ell}}
+\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right)}{\partial p_{\ell}}
 =
@@
 \begin{equation}\label{e:mlq}
-\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_i, q_i\}_{u=1}^{|V|}\right)}{\partial q_{\ell}}
+\frac{\partial \log \Pr\left(E_0 \,\Big|\, \{p_u, q_u\}_{u=1}^{|V|}\right)}{\partial q_{\ell}}
 = \sum_{k=1}^m
@@
 Unfortunately, equating \eqref{e:mlp} and \eqref{e:mlq} to zero, and solving for parameters
-$\{p_i, q_i\}_{i=1}^{|V|}$ gives rise to a hard set of nonlinear equations.
+$\{p_u, q_u\}_{u=1}^{|V|}$ gives rise to a hard set of nonlinear equations.
 Moreover, some such parameters may never occur in these equations, namely whenever $\Nout(u)$ or
@@
 quantity and setting the resulting equation to zero, we obtain for each $\ell \in V$:
-\begin{equation*}
+\begin{equation}
+  \label{eq:troll_appp}
   \sum_{k=1}^m \Ind{u_k = \ell,y_{\ell,v_k}=+1} \left(2-p_{\ell}-q_{v_k}\right)
@@
   (p_{\ell}+q_{v_k})
-\end{equation*}
+\end{equation}
 
 We apply a similar transformation to \eqref{e:mlq} in order to obtain, for each $\ell \in V$:
-\begin{equation*}
+\begin{equation}
+  \label{eq:troll_appq}
   \sum_{k=1}^m \Ind{v_k = \ell,y_{u_k,\ell}=+1} \left(2-p_{u_k}-q_{\ell}\right)
@@
   \left(p_{u_k}+q_{\ell}\right)
+\end{equation}
+
+At this point, we find convenient to take a step back and define one of the label
+regularity measure we introduced in \autoref{s:prel}. Namely, recall we said $\Psi^2_{G''}(Y)$ could
+be seen as a soft quadratic version of the cutsize. With the notation of our generative model, we
+can write it as
+\begin{equation*}
+  \Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \sum_{(u,v) \in E}
+  \left(\frac{1+\yuv}{2} - \frac{p_u+q_v}{2} \right)^2 =  
+  \min_{(\bp,\bq)} f_E(\bp, \bq)\, ,
 \end{equation*}
+where $(\bp,\bq) =\{p_u, q_u\}_{u=1}^{|V|}$ is the set of the model parameters. Intuitively,
+minimizing $f_E$ with respect to $(\bp,\bq)$ is similar to the maximum likelihood approach, as we
+seek the parameters $(\bp, \bq)$ that \enquote{best agree} with the observed signs. It also turns
+out that if we restrict the minimization problem to the observed edges (\ie $\min_{(\bp,\bq)}
+f_{\trainset}(\bp,\bq)$) and set the derivative of $f_{\trainset}$ with respect to $p_\ell$ and
+$q_\ell$ to zero, we recover the equations \eqref{eq:troll_appp} and \eqref{eq:troll_appq}
+respectively.
 
-The solution to these equations are precisely the points where the gradient w.r.t.\ $(\bp,\bq)
-=\{p_i, q_i\}_{i=1}^{|V|}$ of the quadratic function
-\[
-f_{E_0}(\bp,\bq) = \sum_{(u,v) \in E_0} \left(\frac{1+\yuv}{2} - \frac{p_u+q_v}{2} \right)^2
-\]
-vanishes.
-We follow a label propagation approach by adding to $f_{E_0}$ the corresponding test set function
+To include the full topology of the graph and not restrain ourselves to the observed edges,
+we follow a label propagation approach by adding to $f_{E_0}$ the corresponding test set function
 $f_{E\setminus \trainset}$, and treat the sum of the two as the function to be minimized during
@@
 graph $G''$ obtained through the second reduction of \autoref{s:prel}.
+Indeed, one can show\footnote{as we do in \autoref{ssec:troll_lprop_weights}.} that this objective is
+equal ---up to a regularization term--- to the quadratic energy objective minimized by label
+propagation methods.
 However, because of the presence of negative edge weights in $G''$, we first have to symmetrize\footnote{%

### 2017-10-03 troll/experiments.tex
corrections in troll related and experiments.

@@
 task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
-\begin{newcontent}
 In \adv{}, a trust-based social network for open source developers, a user $u$ can certify another
@@
 4]{AdvogatoTrustThesis02}).
-\end{newcontent}
 In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
@@
 \[
-\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \left(f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\right)\,,
+  \Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \sum_{(u,v) \in E}
+  \left(\frac{1+\yuv}{2} - \frac{p_u+q_v}{2} \right)^2
 \]
-$f_{E_0}$ and $f_{E\setminus E_0}$ being the quadratic criterions of \autoref{ss:passive}, viewed as
-functions of both $(\bp,\bq)$, and \yuv{} when all labels are known, and $\Psi_{G}(Y)$ is the
+as first described in \autoref{ss:passive}
+and $\Psi_{G}(Y)$ is the
 label regularity measure adopted in the online setting, as defined in \autoref{s:prel}. It is
@@
 
-\begin{newcontent}
 Finally, there is a low fraction of reciprocal edges (\ie{} both $\euv \in E$ and $\evu
@@
 % more references in  http://www.nature.com/articles/srep02729
-\end{newcontent}
 
-\begin{newcontent}
 \subsection{Synthetic signs}
@@
 \dout(u_2), \ldots, \dout(u_{|V|})$. We shuffle these segments and draw a number \uar{} from each of
-them, which we set as the $p$ value of the corresponding node. Then, because we want to model real
+them, which we set as the initial $p'$ value of the corresponding node. Then, because we want to model real
 sign distribution and have more
-positive edges, we apply an exponential transformation \[ p \rightarrow \frac{1}{1-e^{-\lambda}}
-\left( 1-e^{-\lambda p} \right) \,,\] where we choose $\lambda=3$. We do the same for $q$, then for
+positive than negative edges, we apply an exponential transformation \[ p = \frac{1}{1-e^{-\lambda}}
+\left( 1-e^{-\lambda p'} \right) \,,\] where we choose $\lambda=3$. We do the same for $q$, then for
 every edge \euv{}, we set it positive with probability \etauv{}. An example of the distributions we
-obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}, giving $71.9\%$ of
-positive edges.
+obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}\footnote{More precisely,
+these figures are normalized histogram, meaning the counts are scale down so that the bars total
+area sums up to one.}, giving $71.9\%$ of positive edges.
 
+% TODO: do a proper subfigure instead of minipage?
 \vspace{\baselineskip}
@@
 predictor classifies an edge \euv{} as positive if $\etauv \geq \shalf$ and negative otherwise. 
-We compare it with the algorithm analyzed in \autoref{ss:bayes_approx}, which we call \usrule{}
-(Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). In practice,
+We compare it with the \usrule{} algorithm analyzed in \autoref{ss:bayes_approx}. Recall that
 \usrule{} proceeds as follow: after computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all
@@
 right corner (reflecting the imbalance toward positive edges), which allows better estimation. 
-In this case, we cannot use the MCC because off-diagonal, the Bayes predictor classifies edges
-either all positive or all negative, resulting in a division by zero in the definition. We thus
+In this case, we cannot use the MCC because outside the diagonal, the Bayes predictor classifies edges
+either all positive or all negative, resulting in a division by zero in the definition of MCC. We thus
 fall back on accuracy and display in \autoref{fig:tr_synth_acc} the difference between the accuracy
-of the Bayes predictor and the accuracy of \usrule{}. This time, the anti diagonal where
+of of \usrule{} and the accuracy of the Bayes predictor. This time, the anti diagonal where
 $\frac{p+q}{2}=\lhalf$ seems to play a special role. Indeed, as showed more clearly in
-\autoref{fig:tr_synth_cmpacc}, the gap in accuracy between the Bayes predictor and \usrule{}
+\autoref{fig:tr_synth_cmpacc}, the gap in accuracy between \usrule{} and the Bayes predictor
 increases symmetrically as $\frac{p+q}{2}$ moves away from \shalf{}. The gap also tends to disappear
@@
 datasets and training size.
-\end{newcontent}
 
@@
     \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_acc.pdf} 
-    \caption{100 times the accuracy.} 
+    \caption{100 time the difference of accuracy between \usrule{} and Bayes.} 
     \label{fig:tr_synth_acc}
@@
 
-  \item The \usrule{} algorithm, as described above.
+  \item The \usrule{} algorithm described in \autoref{ss:passive}.
 
@@
 Because scalability is a major concern when training on sizeable datasets, we also give an idea of
-relative training times (in milliseconds) by reporting the time it took to train a single run of
+relative training times (in milliseconds) by reporting separately in \autoref{tab:all_time} the time it took to train a single run of
 each algorithm on a training set of size\footnote{Comparison of training time performances is fair
@@
 run on the same machine (16 Xeon cores and 192Gb Ram).} 15\% of $|E|$, and then predict on the test
-set. Though our experiments are not conclusive, some trends can be readily spotted:
+set. Some conclusions emerge from those experiments:
 
@@
 
-  \item \uslpropGsec{} always ranks first or at least second in this comparison when MCC is
-    considered. On top of it, \uslpropGsec{} is fastest among the global methods (one or even two
-    orders of magnitude faster), thereby showing the benefit of our approach to edge sign
-    prediction.
+  \item \uslpropGsec{} almost always ranks first or at least second in this comparison when MCC is
+    considered, at least in the small training set regime (that is when having access to at most
+    $25\%$ of the labels, which might be more realistic in many real world situations).
+    On top of it, \uslpropGsec{} is fastest among the global methods (one or even two
+    orders of magnitude faster), thereby showing the benefit of our approach to \esp{}.
+
+  \item Two methods are competitive with us, \compbayesian{} and \compranknodes{}. \compbayesian{}
+    achieves its best results on the larger dataset (\sla{}, \epi{} and \kiw{}) when the training
+    set size is large. Indeed, it learns a much more complex model than our methods, with $256$
+    parameters\footnote{corresponding to $16\times 16$ feature per edge}, which also requires a lot
+    of time to be trained. \compranknodes{} also shines in those large datasets, even at small
+    training size. Note however that the difference with \uslpropGsec{} is usually less than one point,
+    at the expense of a five times longer learning time.
 
@@
 
-\begin{newcontent}
+\input{all_time}
+
 \subsection{Additional experiments}
@@
 \autoref{tab:troll_ts_mcc}, this makes the problem much more challenging, both for our methods and
-out competitors. It is quite surprising that even when the training set is as large as 90\% of all
+our competitors. It is quite surprising that even when the training set is as large as 90\% of all
 edges, the MCC are so low compared with those reported in \autoref{tab:all_mcc}.
@@
 
+This conclusion is not in contradiction with our previous experiments, since our algorithms were
+developed under the assumption that both the revealed labels and the ones to predict were
+distributed uniformly within the graph. Recall for instance that \autoref{t:active} requires to have
+$\Omega(\log |V|)$ labels outgoing from $u$ and incoming to $v$ to guarantee an accurate prediction
+of the sign of \euv{}. Therefore in the next section, we present an online algorithm, whose goal is
+to guarantee good performance when the labeling is regular enough, despite the adversarial order in
+which the predictions are asked.
+
 \input{sampling}
-
-\end{newcontent}

### 2017-10-03 troll/related.tex
corrections in troll related and experiments.

@@
 
-\begin{newcontent}
 Interest in signed networks can be traced back to the psychological theory of structural
@@
 A sophisticated take on degree features is presented by \textcite{Bayesian15}, who note that a node
-can belong to one of the 16 types based of whether the number of its positive (respectively
+can belong to one of the 16 node-types based of whether the number of its positive (respectively
 negative) outgoing (respectively incoming) edges is zero or not. The number of unobserved incoming
 and outgoing edges of each node $u$ let us define a 16-dimensional vector $V_u$ containing the
-probability of transitioning to any other type. Then for each edge \euv, they compute the outer
+probability of transitioning to any other type once the unobserved signs are revealed. Then each
+edge \euv is associated with a feature vector consisting of the outer
 product of $V_u$ with $V_v$ and also include additional features such as triads count and degree
@@
 
-Other works have also\Todo{maybe that part would be better in the conclusion, especially since side
+Other works have also\done\Todo{maybe that part would be better in the conclusion, especially since side
 information would help with the cold start problem showed in the experiments and trust version is
@@
 proposed, which is quadratic in $|V|$ \emph{at each prediction}.
-\end{newcontent}
 

### 2017-10-03 troll/conclusions.tex
mark content added from the aistats paper
split the troll conclusion in two parts

@@
-\section{Conclusion}
-
-In this chapter, we started our characterization of edges in complex networks by studying the \esp{}
-problem in \dssn{}. As mentioned in the introduction, given a network topology and a few labeled
-interactions, being able to infer the polarity of the remaining interactions is valuable to improve
-user experience. Our main insight is that two features of users behavior, trollness and
-(un)trustworthiness, are key to understand their interactions. This leads us to design a simple sign
-generative model. Such a model serves both as a theoretical justification for many successful
-heuristics and as the underpinning of our \usrule{} algorithm. This algorithm is fast, trivially
-parallelizable, provably close to the Bayes optimal predictor on dense graphs and experimentally
-accurate even on sparse graphs. We further exploit this model in the batch setting through an
-approximation of the maximum likelihood estimation procedure. While making this approximation, we
-moreover cast the problem as a node prediction problem. This can be efficiently tackled by standard
-Label Propagation algorithms.
-
-We validate our theoretical results by experimentally assessing these two methods in the small
-training set regime, on synthetic and real-world datasets. We draw two main conclusions from our
-experiments. First, our generative model is robust across domains. Indeed, it produces predictors
-that are empirically both close to the Bayes optimal and to general linear models trained on
-trollness and trustworthiness features. This is notable as this result is achieved by simply
-counting negative edges, without resorting to Stochastic Gradient Descent methods typically involved
-in training large scale linear models. Second, our methods are in practice either strictly better
-than their competitors in terms of prediction quality or, when they are not, they are faster.
-
-Finally, we study the online adversarial setting, where trollness and (un)trustworthiness naturally
-lend themselves to define a notion of edge sign regularity. Based on this regularity, we provide an
-upper and an (almost matching) lower bounds on the expected number of prediction mistakes. Before
-exploring in the next chapter another learning bias for \esp{}, which is not based on trollness and
-(un)trustworthiness, we present directions in which this work could be extended.
-
-\bigskip
+\section{Open questions}
 
 Given the fundamental role of our generative model, we would like to adapt it in order to handle
-the three problem extensions presented in the related work of \autoref{sec:troll_related}.
+the three problem extensions presented in the related works of \autoref{sec:troll_related}.
 
@@
 \end{enumerate}
+
+\section{Summary}
+
+In this chapter, we started our characterization of edges in complex networks by studying the \esp{}
+problem in \dssn{}. As mentioned in the introduction, given a network topology and a few labeled
+interactions, being able to infer the polarity of the remaining interactions is valuable to improve
+user experience. Our main insight is that two features of users behavior, trollness and
+(un)trustworthiness, are key to understand their interactions. This leads us to design a simple sign
+generative model. Such a model serves both as a theoretical justification for many successful
+heuristics and as the underpinning of our \usrule{} algorithm. This algorithm is fast, trivially
+parallelizable, provably close to the Bayes optimal predictor on dense graphs and experimentally
+accurate even on sparse graphs. We further exploit this model in the batch setting through an
+approximation of the maximum likelihood estimation procedure. While making this approximation, we
+moreover cast the problem as a node prediction problem. This can be efficiently tackled by standard
+Label Propagation algorithms.
+
+We validate our theoretical results by experimentally assessing these two methods in the small
+training set regime, on synthetic and real-world datasets. We draw two main conclusions from our
+experiments. First, our generative model is robust across domains. Indeed, it produces predictors
+that are empirically both close to the Bayes optimal and to general linear models trained on
+trollness and trustworthiness features. This is notable as this result is achieved by simply
+counting negative edges, without resorting to Stochastic Gradient Descent methods typically involved
+in training large scale linear models. Second, our methods are in practice either strictly better
+than their competitors in terms of prediction quality or, when they are not, they are faster.
+
+Finally, we study the online adversarial setting, where trollness and (un)trustworthiness naturally
+lend themselves to define a notion of edge sign regularity. Based on this regularity, we provide an
+upper and an (almost matching) lower bounds on the expected number of prediction mistakes. We
+conclude by presenting three directions in which our generative model could be extended: namely
+use side information as priors, guide budgeted queries in active learning and generate weights in
+addition to signs.
+
+In the next chapter, we will go even further than this generative model and explore another learning
+bias for \esp{}, which is not based on trollness and (un)trustworthiness, but rather on the social
+balance theory.

### 2017-10-03 troll/related.tex
one more ref

@@
 \epi{} in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
-analysing the textual content of their post, and~\autocite{SNTransfer13} uses SVM to perform
+analysing the textual content of their post, \autocite{attributedSNE17} improve
+the embedding approach of \textcite{neuralSigned17} by considering the words
+written in reviews as attributes of the users,  and~\autocite{SNTransfer13} uses SVM to perform
 transfer learning from one network to another.

### 2017-10-04 edge/conclusions.tex
add some draft of other edge section

The framework we proposed is agnostic to the question of the interplay between connection and
linking attributes, although this is an interesting and active research question.

Homophily and Contagion Are Generically Confounded in Observational Social Network Studies
Cosma Rohilla Shalizi, Andrew C. Thomas
Sociological Methods \& Research
Vol 40, Issue 2, pp. 211-239
DOI: 10.1177/0049124111404820

Digital Footprints: Opportunities and Challenges for Online Social Research
Annual Review of Sociology
Vol. 40:129-152 
First published online as a Review in Advance on June 16, 2014
\url{https://doi.org/10.1146/annurev-soc-071913-043145}

Here is what the “Digital Footprints” writes about the previous paper
An important limitation is the difficulty distinguishing between homophily and contagion. Homophily
refers to a variety of selection mechanisms by which a social tie is more likely between individuals
with similar attributes and environmental exposures (McPherson et al.  2001). Contagion refers to
influence mechanisms (e.g., imitation or peer pressure) by which traits diffuse along network edges.
Homophily and contagion offer competing explanations for network autocorrelation, which refers to
the greater similarity in the attributes of closely connected nodes. Based on simulated networks,
Shalizi \& Thomas (2011, p. 216) conclude that “there is just no way to separate selection from
influence observationally” (see also Manski 1993). This does not mean that observational studies
using online networks are useless, but researchers need to refrain from assuming that the observed
network autocorrelation reflects contagion effects and to acknowledge that the similarity between
adjacent nodes may reflect the mutually reinforcing effects of influence and selection whose
separate contributions may be impossible to tease apart.  For example, although Ugander et al.
(2012) controlled for demographic similarity (sex, age, and nationality), there are countless other
ways in which shared environments, affiliations, interests, and personality traits might cause two
friends to join Facebook independently but not on the same day, making it look like the “early
adopter” influenced the friend they invited who would have joined anyway.

Furthermore, we could also consider the role of time with dynamic networks. For instance, while the
photo similarity between two Flickr users rises before they connect, it later decreases if they have
the same level of popularity, as they try to differentiate themselves
Zeng and Wei: Social Ties and User Content Generation: Evidence from Flickr, Information Systems
Research 24(1), pp. 71–87

### 2017-10-04 edge/related.tex
add some draft of other edge section

Stattner, Erick, and Martine Collard. "Link Clustering for Extracting Collaborative Patterns in a Scientific Co-Authored Network.", FAB workshop, ANOSAM 2017
they use their method of conceptual links

Groups of nodes are first created when they have common attributes. The set of links between two
groups is called a conceptual link in the sense that it corresponds to a relationship between two
sets of attributes, namely two concepts in the field of formal concept analysis [7]. The number of
links between two groups is then evaluated and when the frequency is greater than a given support
threshold $\beta$, we talk about frequent conceptual link. Note that we only keep maximal frequent conceptual links (mfcl), i.e. those who are not included in others.
Finally, the set of maximal frequent conceptual links is used to create a new network structure called a conceptual view which summarizes all the knowledge extracted from the initial network. In a conceptual view, each node corresponds to a set of attributes (in this context we call them Meta-nodes) and a link corresponds to maximal frequent conceptual links.
A formal presentation of frequent conceptual links can be found in [18].

\Textcite[section 7.2 edge labelling]{nodeClassif11} mention a cool paper from Darja Krushevskaja
but I can't find it online.

\fullcite{Zhuang2012} and 
section 2.3 of Semantic Mining of Social Networks (same thing I believe.
Actually they also talk about a WSDM paper on infering ties using transfer
learning~\autocite{transferTiesPred16})

\Textcite[Section 4.2.2]{miningSocialTheoriesSurvey14} is about \emph{Social Tie
Prediction}, but they only mention \autocite{transferTiesPred16} and
\autocite{Yang2012}.

H. Liang, K. Wang and F. Zhu, "Mining social ties beyond homophily," 2016 IEEE 32nd International Conference on Data Engineering (ICDE), Helsinki, 2016, pp. 421-432.
\url{http://doi.org/10.1109/ICDE.2016.7498259} very datamining approach, they
look for link pattern with support and confidence above some threshold, that
also diverge from homophily by excludingi attributes taking the same value (so
it's not exactly like us once again)

The kind of graphs with talked in this paper is a special case of what is
called \emph{heterogenous information network} in \autocite{HINSurvey17}.

### 2017-10-04 edge/model.tex
move link with signed graph from intro to model part

@@
 \end{problem}
+
+\paragraph{Generalization of signed graphs}
+\label{par:generalization_of_signed_graphs}
+
+Here we show why \autoref{p:edge_full} can model signed graphs as a special case, namely with $k=2$.
+Remember the learning
+bias we introduced in the previous chapter \vpageref{text:cc_new_bias}: every node belongs to
+exactly one of the $K$ clusters available, is connected positively with the other nodes belonging to
+the same cluster, and is connected negatively with the nodes belonging to every other clusters. Now
+let the dimension of profiles be $d=K$ and the number of directions $k=2$. Moreover, we define the
+two directions vectors as $w_1 = \frac{\onev}{\sqrt{d}} = w^+$ and $w_2 = -w_+ = w_-$, where
+$\onev$ is the vector of dimension $d$ with only $1$ coordinates. Finally, for a node $u$ in
+cluster $i$, we assume its profile $x_u$ is set as follow: all component are equal to $-b$ except
+for the \ith{} one, which is equal to $a$, where $a$ and $b$ are constants we define now.
+
+First, because $x_u$ is a unit vector, we have that $a^2 + (K-1)b^2=1$, from which $a =
+\sqrt{1-(K-1)b^2}$ whenever $1-(K-1)b^2 \geq 0$, or equivalently $b \leq \frac{1}{\sqrt{K-1}}$.
+Next, we look at $\onev^T\left(x_u \circ x_v\right) = \onev^T m_{uv} = \frac{1}{\sqrt{d}}
+\sum_{i=1}^d m_{{uv}_i}$ for any edge $(u,v) \in E$. If $u$ and $v$ are in the same cluster, then
+$\sum_{i=1}^d m_{{uv}_i} = a^2 + (K-1)b^2 = 1$, implying that $w_+^T m_{uv} = \frac{1}{d} > w_-^T
+m_{uv} = -\frac{1}{d}$. Otherwise, $\sum_{i=1}^d m_{{uv}_i} = (K-2)b^2 - 2ab = (K-2)b^2 -
+2b\sqrt{1-(K-1)b^2} = f_K(b)$. By computing the derivative of $f_K$ with respect to $b$, we find
+that it reaches a minimum of $-\frac{1}{K-1}$ at $b = \frac{1}{\sqrt{K(K-1)}} \leq
+\frac{1}{\sqrt{K-1}}$. In this case, $w_+^T m_{uv} = -\frac{1}{(K-1) \sqrt{d}} < w_-^T m_{uv} =
+\frac{1}{(K-1) \sqrt{d}}$. To maximize $\sum_{(u,v)\in E} \max_{1 \leq \ell \leq d} w_\ell^T
+m_{uv}$, we therefore assign $w_+$ to edges whose both endpoints belong to the same cluster, and
+$w_-$ to edges joining two different clusters.
+
+Moreover, given this assignment, we can show that $w_+$ and $w_-$ are the optimal directions, under
+the additional uniformity assumption that there are exactly $r$ edges within every cluster, and $s$
+edges between any two clusters. First, we define
+\begin{equation*}
+  m_{\mathrm{inner}} = \sum_{(u,v) \in E;\, \cluster(u)=\cluster(v)} m_{uv} \qquad \text{and} \qquad
+  m_{\mathrm{outer}} = \sum_{(u,v) \in E;\, \cluster(u)\neq \cluster(v)} m_{uv}
+\end{equation*}
+to be the sum of all edge vectors within and between clusters respectively. By our choice of $a$ and
+$b$, we have that $m_{\mathrm{inner}} = r\onev$ and $m_{\mathrm{outer}} = -\frac{s}{2}\onev$. Now
+recall that the solution of $\argmax_{w\in \dsphere} w^T c$ is $\frac{c}{||c||}$. Therefore, the
+best direction to assign to inner edges is indeed $w_+$, and so is $w_-$ to outer edges.
+
+\begin{aside}
+For inner edge vectors, $m_{uv}$ in cluster $i$ has $a^2$ as its \ith{} component, and $b^2$
+elsewhere. There are $r$ of them in cluster $i$, so summing over all clusters we get
+$m_{\mathrm{inner}} = r\left(a^2 + (K-1)b^2\right)\onev = r\onev$.
+For outer edge vectors, there are a few more steps. Consider the $s$ edges between $C_1$ and $C_2$,
+whose $m_{uv}$ is
+$\begin{pmatrix}
+  -ab \\
+  -ab \\
+  b^2 \\
+  b^2 \\
+  \vdots \\  
+\end{pmatrix}$. Likewise, between $C_1$ and $C_3$ we have
+$m_{uv}=\begin{pmatrix}
+  -ab \\
+  b^2 \\
+  -ab \\
+  b^2 \\
+  \vdots \\  
+\end{pmatrix}$. We thus we see that the first component is always $-ab$, that there is another
+such term at the other cluster index, while the rest is $b^2$. Summing all the vectors incident to
+$C_1$, we get
+$s\begin{pmatrix}
+  -ab(K-1) \\
+  b^2(K-2) -ab \\
+  b^2(K-2) -ab \\
+  \vdots \\  
+\end{pmatrix}$. Similarly, the sum of all vectors incident to $C_2$ is
+$s\begin{pmatrix}
+  b^2(K-2) -ab \\
+  -ab(K-1) \\
+  b^2(K-2) -ab \\
+  \vdots \\  
+\end{pmatrix}$, and so on for every cluster $i$. Summing over all clusters and dividing by $2$ to
+avoid double counting, we get $m_{\mathrm{outer}} = \frac{s}{2}\left( (K-1)\left(b^2(K-2) -ab\right)
+-ab(K-1) \right)\onev = \frac{s(K-1)}{2}\left(b^2(K-2) -2ab\right)\onev =
+\frac{s(K-1)}{2}f_K(b)\onev = -\frac{s(K-1)}{2} \frac{1}{K-1} \onev = -\frac{s}{2}\onev$.
+\end{aside}
+
+Granted, the situation where all the nodes of a cluster have exactly the same profile and where the
+connections are so regular is highly idealized. Moreover, we did not formally prove that there is
+not another assignment (with different direction vectors) that would give a better objective
+value\footnote{Although given the symmetry of the problem, that would be surprising.}. Our point is
+rather that profiles and directions can be seen as a \enquote{latent} explanation for the signs in
+a balanced signed graphs, like our generative model of the first chapter.
+
+\bigskip
 
@@
 standard (projected) gradient descent algorithms.
+See \autocite[Section 6]{attributedSNE17} for a write up example of matrix optimization.
 

### 2017-10-13 edge/experiments.tex
first writeup of the synthetic experiments

\paragraph{Synthetic data}
\label{par:synthetic_data}

While there exists several methods to generate attributed graph with community
structure~\autocites{Yang2013}{XuBayesian14}{Kataoka2016}, here we present the one we adopted for
our initial experiments, as it is specifically tailored to the model introduced in the previous
section. Given a number of directions $k$, a maximum number $k_\mathrm{local}$ of directions
incident to each node, the dimension of profiles $d$ and a number of node $n$, it return a graph
$G=(V,E)$, a profile matrix $X\in \Rbb^{n\times d}$, a set of $k$ directions $\mathcal{D}_k$ and an
assignment from every edge $(u,v)$ to a direction $w_{uv} \in \mathcal{D}_k$ such that in most
cases, $\forall w_\ell \in \mathcal{D}_k \setminus \{w_{uv}\}\,, {s_{uv}}^T w_{uv} \geq  {s_{uv}}^T
w_\ell$.

We first generate the $k$ unit-norm directions $\mathcal{D}_k=\{w_1, \ldots, w_k\}$, independently
of the graph. Furthermore, we control the number of coordinates $n_o$ that overlap with each other.
That is, when $n_o = 0$, each direction has exactly \nicefrac{k}{d}\footnote{We choose $d$ to be a
multiple of $k$.} non-zero dimension that are disjoint. On the other hand, if for instance $n_o=5$,
then $5$ of the $d$ dimensions will have two directions with non-zero components on it. Those
non-zero components are drawn \uar{} from $[-1, 1]$ and the each directions is then normalized. For
instance, with $d=6$, $k=3$ and $n_o=2$, we could generate the following directions, where the dots
represents non-zero numbers and the red ones are overlapping:
\vspace{-.5\baselineskip}
\begin{center}
	\begin{tabular}{ccc}
		$w_1$ & $w_2$ & $w_3$ \\
		$\begin{pmatrix}
			\bullet \\
			0 \\
			\bullet \\
			0 \\
			0 \\
			\textcolor{Red}{\bullet}
		\end{pmatrix}$ & 
		$\begin{pmatrix}
			0 \\
			\bullet \\
			0 \\
			\textcolor{Red}{\bullet} \\
			\bullet \\
			0
		\end{pmatrix}$ &
		$\begin{pmatrix}
			0 \\
			0 \\
			0 \\
			\bullet \\
			0 \\
			\bullet
		\end{pmatrix}$
	\end{tabular}
\end{center}

The next step is to create the graph topology, which we do by assembling small Erdős-Rényi
subgraphs~\autocites{erdos1959random}{gilbertRG59} that we call \emph{blocks}. Furthermore, driven
by a topological constraint mentioned earlier, we assign to each node a $k_\mathrm{local}$-tuple of
directions such that two adjacent node have \emph{at least} one common direction that can later be
assigned to the edge between them. For convenience, let us first identify directions with colors and
thus call such a $k_\mathrm{local}$-tuple of directions a \emph{palette}. For a given palette $p$,
we call adjacent palettes, denoted $\adj(p)$, all the palettes different from $p$\footnote{We ruled
out $p$ being adjacent to itself to avoid the trivial solution of all nodes being assigned the same
palette.} but sharing one
color with $p$. Finally, we say that an edge $u,v$ is \emph{colorless} if the palette of $u$ and $v$ have
no color in common. Because in the general case, it is not always possible to assign palette to
every nodes such that there is no colorless edge,\footnote{For instance, consider a $4$-clique with $k=4$
and $k_\mathrm{local}=2$. Without loss of generality, say we assign the palette $(1,2)$ to node
$1$. The adjacent palettes are then $(1,3)$, $(1,4)$, $(2,3)$ and $(2,4)$. If we assign those
starting by $1$ (respectively $2$) to the second and third node, then the fourth palette has to
contain a $1$ (respectively $2$), which is not possible (because two connected nodes cannot have the
same palette). On the other hand, if we assign $(1,3)$ to the second, the third node can only have
$(2,3)$, meaning the fourth node has again no palette available. The same situation with $(1,4)$ and
$(2,4)$.} we now describe a simple heuristic. It performs a breath first visit of a subgraph,
starting from a random node. Upon visiting an uncolored node $u$, it builds the intersection $P$ of
all the adjacent palettes of the palette of its colored neighbors (\ie{} $P=\bigcap_{v\in \nei(u)}
\adj(p_v)$). Then it selects \uar{} a palette from $P$, or an arbitrary palette if $P$ is empty (see
\begin{marginfigure}
  \centering
	\includegraphics[width=\textwidth]{tikz/edge_palette_tikz.pdf}
	\caption{A small example of the node palette assignment, with $k=4$ colors (blue, green, red and
		orange) and palette of size $k_\mathrm{local}=3$. We assume nodes $1$ and $2$ have already been
		visited, and got assigned the palette $(\text{blue, green, red})$ and $(\text{blue, green,
		orange})$ respectively. Moreover, we are currently at $3$ while $4$ is yet uncolored. In that
		case, there are two possible palettes for $3$. If we select $(\text{green, red, orange})$, a
		possible color assignment for the edges $(1,2)$, $(1, 3)$ and $(2, 3)$ is respectively blue,
		red, and green.
	\label{fig:edge_palette}}
\end{marginfigure}
\autoref{fig:edge_palette} for a small illustration). Once all nodes have been colored, we count the
number of edges that are colorless. We repeat this procedure, keeping track of the assignment
minimizing the number of colorless edges. Once we colored the nodes of every block, we look at pair
of blocks, build a list of all edges between blocks that are not colorless and sample from it to
connect blocks. The last step is to a assign a color to each edge from the shared color of its
endpoint (or an arbitrary color for colorless edges). Note this is reminiscent of the stochastic
block model, although the probabilities of edges between blocks are not uniform anymore.

The directions $\mathcal{D}_k$, and the graph $G$ along its edge assignment are the two sources of
randomness we will consider in the following.

\bigskip

Now that we have a direction $w_{uv} \in \mathcal{D}_k$ assigned to every edge $(u,v)$, we want to
find a set of user profiles $X=\{x_u\}_{u\in V}$ that maximize the edge score and minimize the node
loss, defined respectively by equations \eqref{eq:edge_full}\stodo{not really} and \eqref{} in the
previous section as: 
\begin{align*}
  \mathcal{L}_{\mathrm{edge}} &=
  \sum_{u,v \in E} \left(x_u \circ x_v\right)^T w_{uv} \quad \text{and} \\
  \mathcal{L}_{\mathrm{node}} &=
  \sum_{u\in V} \left|\left| x_u - b_u - \sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2\,,
\end{align*}
where for simplicity, we fix $b_u = \zerov$ and $a_{uv} = 1$. Furthermore, to ensure that the
assigned direction of any edge $(u,v)$ achieve a better score than the other directions (that is
$\forall w_\ell \in \mathcal{D}_k \setminus \{w_{uv}\}\,, {s_{uv}}^T w_{uv} \geq  {s_{uv}}^T
w_\ell$) we also minimize cross-entropy loss commonly used in non-binary classification
problem~\autocite[Section 4.3.4]{PRML06}.  Specifically, recall that the goodness of direction
$w_\ell$ for the edge $(u,v)$ is $g(s_{uv}, w_{uv})$. If we denote by $p_{uv,\ell} = \frac{\exp
\left( g(s_{uv}, w_{uv}) \right)}{\sum_{\ell=1}^k \exp \left( g(s_{uv}, w_\ell) \right)}$ the
softmax \enquote{probability} of $s_{uv}$ being explained by direction $\ell$, the cross-entropy
loss is defined by
\begin{equation*}
  \mathcal{L}_{\mathrm{cross-entropy}} = -\sum_{u,v \in E} \sum_{l=1}^k
  \Ind{w_{uv} = w_\ell} p_{uv,\ell}
\end{equation*}

In practice, we first minimize $\lambda \mathcal{L}_{\mathrm{cross-entropy}} -
\mathcal{L}_{\mathrm{edge}}$ with respect to $X$ using the Adam algorithm~\autocite{Adam15} and
automatic differentiation~\autocite{autograd15} as implemented by the Pytorch
package\footnote{\url{http://pytorch.org/}} and projecting the current iterate back to the set of
matrix with unit $L_2$ norm columns $\mathbb{M}^{n\times d}$ at each step.

Finally, to minimize $\mathcal{L}_{\mathrm{node}}$, we iterate over the nodes and take some gradient
steps to minimize $\left|\left| x_u - \sum_{v\in\nei{u}} w_{uv}\right|\right|^2$, but only to the
extent that $w_{uv}$ remains the direction with the largest score on the edge $u,v \in E$.

\subparagraph{Results}
\label{par:edge_synth_results}

In the following we generate one graph topology, along with palette assignment to nodes and
direction assignment to edges. Then we generate a set of unit-norm directions,
parametrized by four numbers (and their default value): $k=7$, $d=35$, $n_o=0$ and
$k_{\mathrm{local}}=3$. To account for the randomness of this generation process, we repeat our
measurement over 200 such $\mathcal{D}_k$. On the other hand, we verify experimentally that
generating various graphs while keeping $\mathcal{D}_k$ fixed gives the same results, hence we will
not report them in the main text.
We also try other configurations to study the parameter sensibility of our methods. Specifically, we
experiment with less directions ($k=5$) or more directions ($k=9, d=36$)\footnote{We add one
dimension to ensure that $d$ remains a multiple of $k$}, with two higher levels of overlap between
the directions ($n_o=6$ and $n_o=12$), with more directions per node ($k_{\mathrm{local}}=4$) and in
larger dimension ($d=77$).

After having assigned a single direction to every edge and generated the user profiles, we can
measure how those two quantities *agree* with the edge scores. Namely, recall we predict the
direction of an edge $(u,v)$ as $\argmax_{\ell \in \rangesk} g(s_{uv}, w_\ell) = {s_{uv}}^T w_\ell$.
However, in the following, we do not evaluate such predictions using accuracy. Indeed, while the
methods we evaluate all assign one among $k$ directions to every edge, there is no guarantee those
directions are recovered in the same order as they were generated. In case of strong disagreement
between the predictions and ground truth labels, it might additionally be difficult to find a
permutation to conciliate them.  Therefore, we see the problem as a clustering one with known ground
truth, and turn to the Adjusted Mutual Information (AMI)~\autocite{AMI09}. It is an information
theoretic measure that enjoys several properties: it is invariant to permutations of the labels,
invariant to the shape of clusters and it is bounded between $0$ and $1$, where an AMI of $1$
indicates perfect correlation with the ground truth labels and $0$ indicates perfect independence
with the ground truth labels.

First, as we can see in the table below, in the \default{} configuration, we do not always
generate user profiles leading to a perfect assignment of edge according to their score/goodness. In
general this is not always possible. Consider a $k$-regular subgraph, where the $k$ edges incident
to every node are each assigned a different directions. By symmetry, it is impossible to find node
profiles that would achieve maximum goodness for all the edges. This is partly what motivated the
$k_{\mathrm{local}}$ local constraint. Even so, and more pragmatically, whereas those AMI scores
could be made higher by increasing the coefficient in front of the cross entropy loss, this would
imply a lower edge goodness. Even with our choice, only a few percent of the edges have a mismatch
between the directions we assigned them and the one with the highest goodness. Note also that
depending of the parameters of the generation, the AMI is not the same. Therefore, in the following,
we divide all the AMI scores by this natural score, in order to obtain a normalized measure.
\newline\vspace{.5\baselineskip}\noindent%
\begin{tabular}{lccccccc}
  \toprule
  {}  & {\smallk{}}    & {\default{}}   & {\largek{}}    & {\smallo{}}    & {\largeo{}}    & {\fdirs{}}     & {\larged{}}    \\
  \midrule
  AMI & $.925 \pm .03$ & $.922 \pm .03$ & $.905 \pm .02$ & $.880 \pm .04$ & $.874 \pm .05$ & $.893 \pm .03$ & $.973 \pm .01$ \\
  \bottomrule
\end{tabular}
\vspace{.5\baselineskip}

Besides this normalized AMI, we also measure how far the recovered directions $\mathcal{\wh{D}}_k$
are from the actual ones, generated initially. More precisely, we associate each vector $w_\ell \in
\mathcal{D}_k$ with the closest one in $\mathcal{\wh{D}}_k$ and report the average $\ell_2$ distance
between the two elements of these $k$ pairs, that is 
\begin{equation*}
  d(\mathcal{D}_k, \mathcal{\wh{D}}_k) =
  \frac{1}{k} \sum_{\ell \in \rangesk} \min_{w'_i \in \mathcal{\wh{D}}_k} \| w_\ell - w'_i \|_2
\end{equation*}

In total we compare five of the methods introduced earlier\stodo{make sure it's clear how we obtain
directions and prediction from each method}:
\begin{enumerate}[1.]
  \item \textsc{$k$-means} clustering of the normalized $s_{uv}$ edge similarity vectors;
  \item \textsc{Lloyd}-like iterative refinement of the $k$-means solution;
  \item \textsc{Combined} optimization of $\mathcal{L}_{\mathrm{edge}}$ and
    $\mathcal{L}_{\mathrm{node}}$;
  \item \textsc{Frank--Wolfe} maximization of $\mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(W)$ over
    the $\delta$-ball of the nuclear norm, with a regularization of the $\ell_2$ norm of the columns
    of $W$. In other words, denoting by $\langle A, B \rangle_\mathrm{F} = \tr\left(A^T B\right)$
    the Frobenius inner product of two real matrices, we solve
    \begin{equation*}
      \min_{\|W\|_* \leq \delta} -\frop{S^T}{W} + \mu \frop{W}{W}\,\text{; and}
    \end{equation*}
  \item \textsc{Explicit} low-rank $PQ^T$ factorization, initialized with the solution of
    \textsc{Combined}\stodo{describe formally} and minimizing the same objective in matric form.
\end{enumerate}

As showed in \autoref{tab:edge_Wvaries_nami}, \combined{} is always the best method and, excluding
\pqt{} for now, the \lloyd{} heuristic is second except in one case, while \kmeans{} is third. Note
that although the differences are not large in absolute value, they are generally statistically
significant. Coming back to \pqt{}, not only does it have the same performance as \combined{}, but
we verified that it actually returns the same predictions, up to a few edges that are classified
differently. This can be explained because they optimize the same objective and \pqt{} is
initialized with the solution from \combined{}. However, we also verified that both edge score and
node loss are respectively higher and lower at the end of the optimization. Finally, the results of
\fwa{} are much worse than all the others, despite the optimization leading to edge scores
comparable to the other methods. This suggests that the subsequent clustering is not
adequate\stodo{Maybe it would be better to save the list of atoms or do a SVD at the end and cluster
from there.}. Regarding the generative parameters, and compared with the default configuration, the
problem is easier when the $k$ directions are spread in a larger number of dimension, and more
difficult when each nodes is involved in four directions instead of three. On the other hand,
whereas we expect the performance to decrease with increase in the overall number of directions or
their overlap, there is not consistent evidence of that.

\begin{table}[htb]
  \centering
  % \setlength{\tabcolsep}{3pt}
  \small
  \caption{Normalized AMI of 5 methods, when generating direction with 7 configurations. For each
    configuration, we generate directions 200 times, and report the mean normalized AMI along with
    the standard deviation. Among the first three methods (that do not use matrix formulation), we
    highlight the best one in \textbf{\textcolor{brown}{bold brown}} and the second best one in
    \textit{\textcolor{red}{italic red}}. When the difference between a score and the next best one
    is statistically significant (that is we can confidently reject the hypothesis that the two
    distribution have the same mean), we furthermore underline that score and show in parenthesis the
    $p$-value of a paired Student's $t$-test.
  \label{tab:edge_Wvaries_nami}}
  \begin{tabular}{llll|ll}
    \toprule
    {} &         \thead{\kmeans{}} &                                  \thead{\lloyd{}} &                              \thead{\combined{}} &  \thead{\fwa{}} &  \thead{\pqt{}} \\
    \midrule
    {\smallk{}}  &            $.836 \pm .08$ &                          $\vsecond{.838 \pm .07}$ &  $\vfirstSig{.875 \pm .06}\,\,\spval{2.17}{-58}$ &  $.213 \pm .11$ &  $.875 \pm .06$ \\
    {\default{}} &            $.818 \pm .06$ &  $\vsecondSig{.873 \pm .05}\,\,\spval{1.25}{-63}$ &  $\vfirstSig{.893 \pm .04}\,\,\spval{5.68}{-33}$ &  $.381 \pm .05$ &  $.893 \pm .04$ \\
    {\largek{}}  &            $.803 \pm .04$ &  $\vsecondSig{.881 \pm .04}\,\,\spval{2.66}{-94}$ &  $\vfirstSig{.894 \pm .04}\,\,\spval{8.98}{-17}$ &  $.421 \pm .04$ &  $.894 \pm .04$ \\
    {\smallo{}}  &            $.813 \pm .07$ &   $\vsecondSig{.824 \pm .06}\,\,\spval{7.57}{-6}$ &  $\vfirstSig{.856 \pm .06}\,\,\spval{2.99}{-57}$ &  $.378 \pm .05$ &  $.855 \pm .06$ \\
    {\largeo{}}  &  $\vsecond{.827 \pm .07}$ &                                    $.823 \pm .06$ &  $\vfirstSig{.852 \pm .06}\,\,\spval{1.90}{-25}$ &  $.370 \pm .06$ &  $.851 \pm .06$ \\
    {\fdirs{}}   &            $.772 \pm .07$ &  $\vsecondSig{.814 \pm .07}\,\,\spval{6.02}{-42}$ &  $\vfirstSig{.853 \pm .06}\,\,\spval{2.13}{-47}$ &  $.320 \pm .06$ &  $.853 \pm .06$ \\
    {\larged{}}  &            $.905 \pm .05$ &  $\vsecondSig{.933 \pm .04}\,\,\spval{1.32}{-31}$ &  $\vfirstSig{.941 \pm .03}\,\,\spval{1.77}{-22}$ &  $.222 \pm .10$ &  $.931 \pm .04$ \\
    \bottomrule
  \end{tabular}
\end{table}

Similar conclusions carry out when we evaluate methods according to their ability of recovering the
original directions. Keeping in mind that in the case of perfect recovery, $d(\mathcal{D}_k,
\mathcal{\wh{D}}_k)$ would be equal to zero, while the distance between two $d$-dimensional unit
vectors is $\sqrt{2}$ in expectation\footnote{One can derive this from the fact after normalization,
vectors whose coordinates are drawn from a standard normal distribution are uniformly distributed on
the unit sphere~\autocite{gaussianSphere59}.}, we see in \autoref{tab:edge_Wvaries_dst} that no
method gets very close to the original directions. \combined{} is again the closer overall, but now
\kmeans{} and \lloyd{} are alternating at the second place. The fact $\mathcal{D}_k$ cannot be
completely recover is not surprising, for in absence of prior information the problem is
under-constrained. Indeed, while \combined{} and \pqt{} delivers almost the exact same prediction,
their directions are clearly different.

\begin{table}[hbt]
  \centering
  % \setlength{\tabcolsep}{3pt}
  \small
  \caption{Same as \autoref{tab:edge_Wvaries_nami}, but reporting $d(\mathcal{D}_k,
  \mathcal{\wh{D}}_k)$, which should be as close as possible to $0$.
  \label{tab:edge_Wvaries_dst}}
    \begin{tabular}{llll|ll}
      \toprule
      {} &                                 \thead{\kmeans{}} &                                  \thead{\lloyd{}} &                              \thead{\combined{}} &   \thead{\fwa{}} &  \thead{\pqt{}} \\
      \midrule
      {\smallk{}}  &  $\vsecondSig{.581 \pm .08}\,\,\spval{4.78}{-30}$ &                                    $.606 \pm .09$ &  $\vfirstSig{.560 \pm .05}\,\,\spval{8.07}{-24}$ &  $1.007 \pm .05$ &  $.633 \pm .06$ \\
      {\default{}} &                                    $.533 \pm .08$ &   $\vsecondSig{.528 \pm .09}\,\,\spval{9.61}{-4}$ &   $\vfirstSig{.516 \pm .06}\,\,\spval{4.95}{-5}$ &   $.877 \pm .05$ &  $.564 \pm .06$ \\
      {\largek{}}  &                                    $.541 \pm .07$ &  $\vsecondSig{.521 \pm .07}\,\,\spval{5.08}{-23}$ &                          $\vfirst{.520 \pm .05}$ &   $.811 \pm .06$ &  $.551 \pm .06$ \\
      {\smallo{}}  &  $\vsecondSig{.554 \pm .08}\,\,\spval{2.74}{-52}$ &                                    $.589 \pm .08$ &  $\vfirstSig{.538 \pm .05}\,\,\spval{2.38}{-12}$ &   $.908 \pm .05$ &  $.596 \pm .07$ \\
      {\largeo{}}  &  $\vsecondSig{.554 \pm .09}\,\,\spval{1.50}{-58}$ &                                    $.595 \pm .08$ &   $\vfirstSig{.545 \pm .06}\,\,\spval{4.08}{-5}$ &   $.935 \pm .05$ &  $.604 \pm .08$ \\
      {\fdirs{}}   &                                    $.565 \pm .08$ &                          $\vsecond{.564 \pm .09}$ &   $\vfirstSig{.546 \pm .05}\,\,\spval{6.30}{-8}$ &   $.928 \pm .06$ &  $.588 \pm .06$ \\
      {\larged{}}  &  $\vsecondSig{.571 \pm .07}\,\,\spval{2.02}{-89}$ &                                    $.602 \pm .07$ &                          $\vfirst{.567 \pm .05}$ &  $1.033 \pm .06$ &  $.641 \pm .05$ \\
      \bottomrule
    \end{tabular}
\end{table}

\iffalse
\begin{table}[hbt]
  \centering
  % \setlength{\tabcolsep}{3pt}
  \small
  \caption{Same as \autoref{tab:edge_Wvaries_nami}, but when $\mathcal{D}_k$ is fixed and generating
    200 graphs.  \label{tab:edge_Gvaries_nami}}
    \begin{tabular}{llll|ll}
      \toprule
      {} &                                 \thead{\kmeans{}} &                                  \thead{\lloyd{}} &                              \thead{\combined{}} &  \thead{\fwa{}} &  \thead{\pqt{}} \\
      \midrule
      {\smallk{}}  &  $\vsecondSig{.678 \pm .07}\,\,\spval{4.16}{-22}$ &                                    $.657 \pm .07$ &  $\vfirstSig{.729 \pm .08}\,\,\spval{1.13}{-62}$ &  $.330 \pm .08$ &  $.729 \pm .08$ \\
      {\default{}} &                                    $.745 \pm .04$ &  $\vsecondSig{.799 \pm .04}\,\,\spval{2.88}{-65}$ &  $\vfirstSig{.839 \pm .03}\,\,\spval{1.75}{-68}$ &  $.379 \pm .04$ &  $.839 \pm .03$ \\
      {\largek{}}  &                                    $.769 \pm .03$ &  $\vsecondSig{.843 \pm .04}\,\,\spval{5.06}{-84}$ &  $\vfirstSig{.870 \pm .04}\,\,\spval{8.54}{-40}$ &  $.431 \pm .04$ &  $.870 \pm .04$ \\
      {\smallo{}}  &                                    $.813 \pm .06$ &  $\vsecondSig{.845 \pm .04}\,\,\spval{1.44}{-27}$ &  $\vfirstSig{.869 \pm .04}\,\,\spval{3.98}{-55}$ &  $.371 \pm .05$ &  $.868 \pm .04$ \\
      {\largeo{}}  &                                    $.788 \pm .06$ &   $\vsecondSig{.799 \pm .05}\,\,\spval{1.54}{-6}$ &  $\vfirstSig{.830 \pm .05}\,\,\spval{2.36}{-57}$ &  $.408 \pm .05$ &  $.830 \pm .05$ \\
      {\fdirs{}}   &                                    $.678 \pm .05$ &  $\vsecondSig{.700 \pm .06}\,\,\spval{1.65}{-24}$ &  $\vfirstSig{.761 \pm .06}\,\,\spval{2.73}{-80}$ &  $.336 \pm .05$ &  $.761 \pm .06$ \\
      {\larged{}}  &                                    $.863 \pm .06$ &  $\vsecondSig{.893 \pm .05}\,\,\spval{8.68}{-44}$ &  $\vfirstSig{.907 \pm .05}\,\,\spval{5.58}{-31}$ &  $.276 \pm .08$ &  $.895 \pm .05$ \\
      \bottomrule
    \end{tabular}
\end{table}

\begin{table}[hbt]
  \centering
  % \setlength{\tabcolsep}{3pt}
  \small
  \caption{Same as \autoref{tab:edge_Wvaries_dst}, but when $\mathcal{D}_k$ is fixed and generating
    200 graphs.  \label{tab:edge_Gvaries_dst}}
    \begin{tabular}{llll|ll}
      \toprule
      {} &                                 \thead{\kmeans{}} &                                  \thead{\lloyd{}} &                              \thead{\combined{}} &   \thead{\fwa{}} &  \thead{\pqt{}} \\
      \midrule
      {\smallk{}}  &  $\vsecondSig{.702 \pm .08}\,\,\spval{2.27}{-69}$ &                                    $.754 \pm .06$ &  $\vfirstSig{.647 \pm .06}\,\,\spval{1.59}{-81}$ &   $.975 \pm .06$ &  $.734 \pm .07$ \\
      {\default{}} &                          $\vsecond{.607 \pm .06}$ &                                    $.608 \pm .07$ &  $\vfirstSig{.567 \pm .04}\,\,\spval{2.55}{-69}$ &   $.897 \pm .04$ &  $.622 \pm .05$ \\
      {\largek{}}  &                                    $.584 \pm .06$ &  $\vsecondSig{.563 \pm .07}\,\,\spval{2.79}{-24}$ &                          $\vfirst{.561 \pm .05}$ &   $.804 \pm .05$ &  $.582 \pm .06$ \\
      {\smallo{}}  &  $\vsecondSig{.530 \pm .08}\,\,\spval{1.20}{-67}$ &                                    $.571 \pm .07$ &   $\vfirstSig{.518 \pm .05}\,\,\spval{5.10}{-8}$ &   $.894 \pm .05$ &  $.572 \pm .06$ \\
      {\largeo{}}  &  $\vsecondSig{.607 \pm .08}\,\,\spval{1.24}{-70}$ &                                    $.647 \pm .07$ &  $\vfirstSig{.578 \pm .05}\,\,\spval{6.06}{-35}$ &   $.896 \pm .06$ &  $.636 \pm .07$ \\
      {\fdirs{}}   &  $\vsecondSig{.655 \pm .08}\,\,\spval{2.30}{-32}$ &                                    $.677 \pm .08$ &  $\vfirstSig{.608 \pm .06}\,\,\spval{6.88}{-71}$ &   $.938 \pm .04$ &  $.671 \pm .07$ \\
      {\larged{}}  &  $\vsecondSig{.603 \pm .07}\,\,\spval{9.25}{-95}$ &                                    $.636 \pm .07$ &  $\vfirstSig{.589 \pm .05}\,\,\spval{6.97}{-12}$ &  $1.003 \pm .07$ &  $.662 \pm .06$ \\
      \bottomrule
    \end{tabular}
\end{table}
\fi

### 2017-10-14 edge/model.tex
describe edge matrix formulation in more detail
explain Frank Wolfe better, as well as the kmean baseline and its
iterative improvement

@@
   \end{equation}
+  where $s_{uv} = x_u \circ x_v$ and $g(s_{uv}, w_{\ell}) = {s_{uv}}^T w_{\ell}$.
 \end{problem}
@@
 
-If $k=|E|$, the topology of the graph becomes irrelevant, as we can simply exploit the trivial
+At the other extreme of $k=2$, when $k=|E|$, the topology of the graph becomes irrelevant,
+as we can simply exploit the trivial
 solution of having a single direction $w_{uv} = \frac{s_{uv}}{||s_{uv}||}$ for every edge $(u,v)$.
@@
 
+\stodo{move that in the conclusion}
 \begin{aside}
@@
 \end{aside}
+\stodo{Specification of constraints sounds related to network schema in \autocite[Definition
+3]{HINSurvey17}, except we have a single node type and are looking at higher order patter.}
 
@@
 \sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2$ to \eqref{eq:edge_full}.
-%
+
 We can also make the assumption that each node $u$ is only involved in $k_\mathrm{local} < k$
@@
 edges incident to $u$ can only be associated with one of $k_\mathrm{local}$ directions. This is
-difficult to incorporate directly in the optimization formulation, as for certain value of
+difficult to incorporate directly in the optimization formulation\stodo{actually it's not that
+difficult}, as for certain value of
 $k_\mathrm{local}$, this is a \NPc{} graph colouring problem. Still, it can be seen as a learning
@@
 \begin{equation}
-  \argmax_{w_1,\ldots,w_k \in \dsphere,\, a_{uv}\in\Rbb,\, b_u \in \Rbb^d}
+  \label{eq:edge_soft}
+  \argmax_{{\substack{w_1,\ldots,w_k \in \dsphere,\\
+  a_{uv}\in\Rbb,\, b_u \in \Rbb^d}}}
   \frac{1}{|E|\beta} \sum_{u,v\in E}
@@
 
-Another way to remove the $\max$ is to look for a single direction per edge, stacked horizontally in
+Another way to remove the $\max$ is to first seek one direction for every edge, that is to maximize
+$\sum_{u,v\in E} {s_{uv}}^T w_{uv}$. This objective can be written in matric form as $\tr(SW) =
+\frop{S^T}{W}$, where $S \in \Rbb^{|E| \times d}$ is the $s_{uv}$ vectors for all edges stacked
+vertically and $W \in \Rbb^{d \times |E|}$ is the matrix of all directions stacked horizontally.
+More specifically, because of the unit norm constraint on directions, we look for $W$ in the set
+$\mathbb{M}^{d\times |E|}$ of $d\times |E|$ matrices with unit $\ell_2$ norm columns. To avoid the
+trivial solution of letting each direction $w_{uv}$ merely be $\frac{s_{uv}}{\|s_{uv}\|}$, we
+further constraint $W$ to be of low rank. That is, what we really want to optimize is \[\min_{W\in
+\mathbb{M}^{d\times |E|}} -\frop{S^T}{W} + \rank(W)\,.\] However, while the Frobenius inner product is
+linear in $W$, the rank operator is highly irregular and optimizing it directly is \NPc{}. Therefore
+we relax the rank term by the nuclear norm of $W$, defined as $\|W\|_* = \sum_{i=1}^{\min(d,|E|)}
+\sigma_i(W)$, where $\sigma_i(W)$ is the \ith{} largest singular value of $W$. The intuition is
+that the rank of $W$ is
+equal to the number of non-zero singular value while the nuclear norm is essentially an $\ell_1$ norm on
+the spectrum of $W$. Therefore, as minimizing the $\ell_1$ enforce sparsity, minimizing the nuclear
+norm tends to make the rank smaller by ensuring the presence of zeros in the singular values of $W$.
+\footnote{As mentioned in \autoref{sec:troll_related} about \autocite{OnlineCompletion17}, a tighter
+relaxation of the low-rank constraint is the max norm, but is it convex?} This
+results in the following optimization problem: \[\min_{W\in \mathbb{M}^{d\times |E|}} -\frop{S^T}{W}
++ \lambda \|W\|_*\,,\] where $\lambda$ is a regularization parameter. Since $\|\cdot\|_*$ is a norm, this
+objective function is convex (although the domain over which we optimize is not), but it does not
+have a gradient, forcing us to rely on potentially costly proximal gradient descent
+methods~\autocite{Parikh2013a}. Instead, we consider the equivalent problem obtained by replacing
+the regularization parameter with an upper bound $\delta$ on the nuclear norm of $W$, yielding:
+\begin{equation*}
+  \min_{\substack{W\in \mathbb{M}^{d\times |E|} \\ \|W\|_* \leq \delta}} -\frop{S^T}{W}\,.
+\end{equation*}
+
+At this stage, we could also add the node loss term, express in matric form as $\|X - B -
+C\left(A\circ W^T \right)\|_F^2$, where $B\in \Rbb^{|V| \times d}$ is the node bias vectors $b_u$
+stacked vertically, $C \in \Rbb^{|V| \times |E|}$ is the incidence matrix of $G$\footnote{The
+  incidence matrix of a graph $G$, usually denoted by $B$, is the $|V| \times |E|$ matrix such that
+$B_{i,j} = 1$ if the node $v_i$ and edge $e_j$ are incident and 0 otherwise}, $A \in \Rbb^{|E|
+\times d}$ is the matrix whose each row is the corresponding $a_{uv}\onev$ vector and $\|\cdot\|_F$
+denotes the Frobenius norm. Note that this term is jointly convex in $A$, $B$ and $W$.\footnote{As
+one can check that the inequality defining convexity holds for this function.} However, for the
+clarity of exposition, we refrain to do so for now.
+
+The remaining issue is that $W\in \mathbb{M}^{d\times |E|}$ is not a convex constraint. As in the
+previous formulation, we thus relax it by simply imposing that the $\ell_2$ norm of the columns of
+$W$ are not too large. In fact, we already have one bound on these column norms, because of the
+nuclear norm constraint. One can indeed show that for $W\in \Rbb^{d\times |E|}$ such that $\|W\|_*
+\leq \delta$, the \ith{} column $W_{:,i}$ of $W$ satisfies $\|W_{:,i}\|_2 \leq \delta\sqrt{d}$.
+However, setting $\delta = \nicefrac{1}{\sqrt{d}}$ is too stringent in high dimension, and because
+this inequality is rather loose, it will likely results in very small columns norm. Instead of
+relying solely on the nuclear norm constraint, we therefore add an additional regularization term
+equal to $\mu \sum_{i=1}^{|E|} \|W_{:,i}\|_2 = \mu \frop{W}{W}$. Our final convex formulation is
+then:
+\begin{equation}
+  \label{eq:edge_FW}
+  \min_{\substack{W\in \Rbb^{d\times |E|} \\ \|W\|_* \leq \delta}} \frop{\mu W - S^T}{W} \,.
+\end{equation}
+
+\begin{aside}
+  Let $W$ be a matrix in $\Rbb^{d\times m}$ with $d < m$ such that $\|W\|_* \leq \delta$.
+  Furthermore, let its singular value decomposition be such that we can write $W = \sum_{l=1}^d
+  \sigma_l u_l v_l^T$, where $u_l$ and $v_l$ are $d$ unit-norm vectors, of dimension $d$ and $m$
+  respectively. We can thus express the general term $W_{k,i}$ as $\sum_{l=1}^d \sigma_l v_{l,i}
+  u_{l,k}$. Because $v_l$ is unit norm, $v_{l,i} \leq 1$ and $W_{k,i}^2 \leq \left( \sum_{l=1}^d
+  \sigma_l u_{l,k} \right)^2 \leq \sum_{l=1}^d \sigma_l^2 \sum_{l=1}^d u_{l,k}^2$, using
+  Cauchy--Schwartz inequality. The norm of the \ith{} column of $W$ then satisfies:
+  \begin{equation*}
+    \|W_{:,i}\|_2 = \left( \sum_{k=1}^m W_{k,i}^2 \right)^\lhalf
+    \leq \left( \sum_{k=1}^m \left( \sum_{l=1}^d \sigma_l^2 \sum_{l=1}^d u_{l,k}^2 \right) \right)^\lhalf
+    = \left( \left( \sum_{l=1}^d \sigma_l^2 \right) \left( \sum_{l=1}^d \sum_{k=1}^m u_{l,k}^2 \right) \right)^\lhalf
+  \end{equation*}
+  From $\sum_{l=1}^d |\sigma_l | \leq \delta$, we have that $\sum_{l=1}^d \sigma_l^2 \leq \delta^2$.
+  Combined with $u_l$ being unit norm, we conclude that $\|W_{:,i}\|_2 \leq \delta\sqrt{d}$.
+\end{aside}
+
+Because \eqref{eq:edge_FW} is of the form $\min_{x\in \mathcal{X}} f(x)$ where $f$ and $\mathcal{X}$
+are convex, we can solve it using the Frank--Wolfe algorithm~\autocites{FrankWolfe56}{Jaggi2013a},
+which enjoy properties like a $O(\frac{1}{t})$ convergence rate, low computational cost by avoiding
+projection step and sparse solution. It is an iterative procedure that maintains an estimate
+$x^{(t)}$ of the solution and works succinctly as follow: it linearizes the objective $f$ at the
+current position $x^{(t)}$ by computing $\nabla f(x^{(t)})$, it finds a minimum $s$ of that
+linearization within the domain $\mathcal{X}$ by solving $\argmin_{s\in \mathcal{X}}
+\innerp{s}{\nabla f(x^{(t)})}$ and it moves toward that minimizer by a step $\gamma$, setting
+$x^{(t+1)} = (1-\gamma) x^{(t)} + \gamma s$.\footnote{$\gamma$ is either found by a line search of
+set to $\nicefrac{t}{2+t}$.} In the case of \eqref{eq:edge_FW}, the linear minimization step amounts
+to $\argmin_{\|W\|_* \leq \delta} 2\mu W - S^T = \argmax_{\|W\|_* \leq \delta} S^T - 2\mu W$.
+Moreover, the $\delta$-ball of the nuclear norm is the convex hull of the matrix of the form $\delta
+u v^T$ with $\|u\|_2 = 1 = \|v\|_2$~\autocite{Jaggi2013a}. By letting $u_1$ and $v_1$ be
+respectively the left and right vectors associated with the largest singular value of $S^T - 2\mu
+W$, the minimizer \enquote{$s$} we are looking for is therefore $\delta u_1 v_1^T$.
+
+\iffalse
+stacked horizontally in
 one matrix. Then we can directly minimize \eqref{eq:edge_full} while enforcing the matrix
@@
 representative directions in $\Rbb^d$ while $Q$ gives, for each edge, the weights of each
-representative direction. Similar to what is done in spectral clustering, we could cluster the
+representative direction. Similar to what is done in spectral clustering, we then cluster the
 $q_{uv}$'s to obtain an edge labeling.
@@
 Denoting by $\mathbb{M}^{d\times |E|}$ the set of $d\times |E|$ matrices with unit $L_2$ norm
-columns, a convex formulation of the problem is as follows:
-\begin{equation}
+columns, a convex formulation\stodo{justify the convexity later: $L_edge$ is linear and both
+$L_node$ and $\|W\|_*$ are norm and thus convex} of the problem is
+as follows:
+\begin{align}
   \label{eq:fact_convex}
-  \min_{W\in\mathbb{M}^{d\times |E|}} \quad
-  \sum_{i,j \in E} {s_{uv}}^T w_{uv} + \lambda \|W\|_*,
-\end{equation}
-where $w_{uv} \in \Rbb$ denotes the $(u-1)n+v$-th column of $W$, $\|W\|_* =
+  \max_{W\in\mathbb{M}^{d\times |E|}} h(W) &=
+  \mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(W) -
+  \mu \mathcal{L}_{\mathrm{node}}^{\mathrm{mat}}(W) - \lambda \|W\|_* \\
+  &= \frac{1}{|E|} \tr\left(SW\right) -
+  \frac{\mu}{|V|} \|X - B - C\left(A\circ W^T \right)\|_2^2 - \lambda \|W\|_*
+  % \sum_{i,j \in E} {s_{uv}}^T w_{uv} + \lambda \|W\|_*,
+\end{align}
+where $S \in \Rbb^{|E| \times d}$ is the $s_{uv}$ vectors for all edges stacked vertically, $\mu
+\geq 0$ is a trade-off parameter, $B\in \Rbb^{|V| \times d}$ is the node bias vectors $b_u$ stacked
+vertically, $C \in \Rbb^{|V| \times |E|}$ is the incidence matrix of $G$\footnote{The incidence
+matrix of a graph $G$, usually denoted by $B$, is the $|V| \times |E|$ matrix such that $B_{i,j} =
+1$ if the node $v_i$ and edge $e_j$ are incident and 0 otherwise}, $A \in \Rbb^{|E| \times d}$ is
+the matrix whose each row is the corresponding $a_{uv}\onev$ vector, $\|\cdot\|_2$ denotes the
+Frobenius norm, $\|W\|_* =
 \sum_{i=1}^{\min(d,|E|)}\sigma_i(W)$ is the sum of the singular values of $W$ (called the nuclear
 norm or trace norm), and $\lambda \geq 0$ is a regularization parameter.\footnote{To make the
-optimal solution unique, we can use an additional Frobenius norm regularization on $W$.} The nuclear
-favors low-rank solutions as it is essentially an $L_1$ norm on the spectrum.
+optimal solution unique, we can use an additional Frobenius norm regularization on $W$.} The nuclear norm
+favors low-rank solutions as it is essentially an $L_1$ norm on the spectrum.\footnote{As mentioned
+in \autoref{sec:troll_related} about \autocite{OnlineCompletion17}, a tighter relaxation of the
+low-rank constraint is the max norm.}
+Both $\mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(W)$ and
+$\mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(W)$ are convex in $W$, with respective gradient
+\begin{align*}
+  \frac{\partial \mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(W)}{\partial W} &=
+  \frac{1}{|E|} S^T \\
+  \frac{\partial \mathcal{L}_{\mathrm{node}}^{\mathrm{mat}}(W)}{\partial W} &= \frac{-2}{|V|}
+  A^T \circ \left(\Big(\left(X-B\right)^T -A^T \circ \left(W C^T\right) \Big) C\right)
+\end{align*}
 Problem~\eqref{eq:fact_convex} can be solved using proximal gradient descent (requires singular
 value decomposition at each iteration) \autocite{Parikh2013a} or Frank-Wolfe (requires only to find
-the largest singular vector) \autocite{Jaggi2013a}.
+the largest singular vector) \autocite{Jaggi2013a}\stodo{look at the rest of the tutorial and
+comment on the difficulty of the LMO step}.
+\fi
 
-Alternatively, we can consider a nonconvex formulation where we make the low-rank decomposition of
-$W$ explicit. Given a rank $k\leq \min(d,|E|)$:
+We can avoid the rank regularization term (and thus simplify these projection issues) at the cost of
+convexity by making low-rank decomposition of $W$ explicit. Given a rank $k\leq \min(d,|E|)$:
 \begin{equation}
   \label{eq:fact_nonconvex}
-  \min_{P\in\mathbb{M}^{d\times k}, Q\in\mathbb{M}^{|E|\times k}} \quad
-  \sum_{u,v \in E} {s_{uv}}^T w_{uv}.
+  \max_{P\in\mathbb{M}^{d\times k}, Q\in\mathbb{M}^{|E|\times k}} \quad \tilde{h}(P, Q) =
+  \mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(PQ^T) -
+  \mu \mathcal{L}_{\mathrm{node}}^{\mathrm{mat}}(PQ^T)
 \end{equation}
 Problem~\eqref{eq:fact_nonconvex} can be solved via alternating optimization over $P$ and $Q$ using
-standard (projected) gradient descent algorithms.
+standard (projected) gradient descent algorithms, with the following partial derivatives:
+\begin{align*}
+  \frac{\partial \tilde{h}}{\partial P} &=
+  \frac{1}{|E|} S^T Q + \frac{2\mu}{|V|}
+  A^T \circ (((X-B)^T -A^T \circ (PQ^T )C^T )C)Q 
+  \\
+  \frac{\partial \tilde{h}}{\partial Q} &=
+  \frac{1}{|E|} S P + \frac{2\mu}{|V|}
+  A\circ (C^T (X-B-C(A\circ (QP^T ))))P.
+\end{align*}
 See \autocite[Section 6]{attributedSNE17} for a write up example of matrix optimization.
 
-Note however that this matrix formulation requires learning more parameters, as it each edge has a
-different, albeit made up of a small number of basis direction.  Impose sparsity on $Q$?
+Compared with \refeq{eq:edge_soft}, this matrix formulation requires learning more parameters, as it
+each edge has a different direction, albeit made up of a small number of basis direction.  Impose
+sparsity on $Q$?
 
-Max margin, replace the max by a difference of the currently assigned direction with the other
-(trade off parameter). At least I think I used it in synthetic data.
+A simple and natural baseline is to cluster the set of all similarity edge vectors $
+\theset{s_{uv}}{(u,v) \in E}$ using the $k$-means algorithm, where $k$ is the number of directions
+set in \autoref{p:edge_full}. Formally, given the graph $G=(V,E)$ and the node profiles $X$ as
+input, we order the edges from $e_1$ to $e_m$ and build the matrix $S \in \Rbb^{m\times d}$ whose
+\ith{} row is the similarity between the profiles of the endpoints of $e_i=(u,v)$, that is $s_{uv} =
+x_u \circ x_v$. Running $k$-means will thus partition $E$ into $k$ sets of edges, call them $E_1,
+\ldots, E_k$. The first drawback of that simplicity is that it does not solve \autoref{p:edge_full}.
+Indeed, the partition of $E$ does not provide the set of directions $\mathcal{D}_k = \{w_1, w_2,
+\ldots, w_k\} \in \dsphere$ we are looking for. A natural way of obtaining directions from that
+partition is to set $w_\ell = \frac{s_\ell}{\left\|s_\ell\right\|}$, where $s_\ell = \sum_{u,v \in
+E_\ell} s_{uv}$ is the sum of all the similarity vectors of the edge in $E_\ell$. Note however that
+this does not guarantee that we have assigned the best direction to every edge. We therefore propose
+the following heuristic, inspired by the Lloyd algorithm for $k$-means, but where the directions in
+$\mathcal{D}_k$ play the role of centroids. Namely, we alternate between two steps
+\begin{enumerate}[(i), nosep]
+  \item create a new partition $\{E'_1, \ldots, E'_k\}$ by assigning every edge $(u,v)$ to the
+    direction $w$ maximizing its score ${s_{uv}}^T w$, that is $E'_\ell = \theset{(u,v) \in E}{
+    w_\ell = \argmax_{w_i \in \mathcal{D}_k} {s_{uv}}^T w_i}$, and
+  \item update the directions to make them optimal with respect to the edges currently assigned to
+    them, that is as previously, set $\mathcal{D}_k = \left\{ w_\ell = \frac{s_\ell}{\left\| s_\ell
+    \right\|}\right\}_{\ell=1}^k $, where $s_\ell = \sum_{u,v \in E'_\ell} s_{uv}$,
+\end{enumerate}
+until the edge assignment is stable or we reach a maximum number of iterations. By analogy with the
+$k$-means algorithm, obtaining convergence guarantees in the general case would be surprising, but in
+practice we expect that only a small number of iterations would be needed.

### 2017-10-15 edge/model.tex
small precision on PQ, soft and kmeans

@@
 \end{equation}
+Note that in second term, because we do not know the edge assignment, we have to decide which of the
+$w_1, \ldots, w_k$ is $w_{uv}$ equal to. According to our definition of the goodness function, we
+want to let $w_{uv} = \argmax_{w_\ell \in \mathcal{D}_k} {s_{uv}}^T w_\ell$. However, the $\argmax$
+function is not convex and we thus use the same relaxation as before. Namely, by stacking all the
+vectors of $\mathcal{D}_k$ into a single vector $W$ of size $k\times d$, the relaxation is:
+$$ \begin{array}[t]{lrcl}
+  \mathcal{A} : & \Rbb^m & \longrightarrow & \Rbb^d \\
+      & (w_1,\ldots,w_k) & \longmapsto &
+  \sum_{i=1}^k \frac{\exp(\beta{s_{uv}}^T w_i)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_i)} w_i
+\end{array} $$
+This function has an explicit derivative, but it is rather costly to compute and in practice we
+found automatic differentiation~\autocite{autograd15} to be more efficient. While this modified
+objective function is now jointly convex in all its variables, the set $\dsphere$ is not. To ease
+the optimization procedure, we thus only impose the directions $w_\ell$ to lie within the unit ball
+and expect them to have norm close to one in order to maximize the dot product in the ${s_{uv}}^T
+w_\ell$ term.
 
@@
 
-We can avoid the rank regularization term (and thus simplify these projection issues) at the cost of
-convexity by making low-rank decomposition of $W$ explicit. Given a rank $k\leq \min(d,|E|)$:
+We can avoid the rank regularization term altogether by making low-rank decomposition of $W$
+explicit, that by writing $W=PQ^T$ and optimizing over both $P$ and $Q$. However, this comes at cost
+of convexity. Indeed, given a rank $k\leq \min(d,|E|)$, the problem becomes, assuming $A$ and $B$
+are fixed:
 \begin{equation}
-  \label{eq:fact_nonconvex}
-  \max_{P\in\mathbb{M}^{d\times k}, Q\in\mathbb{M}^{|E|\times k}} \quad \tilde{h}(P, Q) =
-  \mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(PQ^T) -
-  \mu \mathcal{L}_{\mathrm{node}}^{\mathrm{mat}}(PQ^T)
+  \label{eq:edge_PQ}
+  \min_{P\in\mathbb{M}^{d\times k}, Q\in\mathbb{M}^{|E|\times k}} \quad \tilde{h}(P, Q) =
+  - \frop{S}{PQ^T} + \mu \frop{} \|X - B - C\left(A\circ QP^T \right)\|_F^2
 \end{equation}
-Problem~\eqref{eq:fact_nonconvex} can be solved via alternating optimization over $P$ and $Q$ using
-standard (projected) gradient descent algorithms, with the following partial derivatives:
+While \eqref{eq:fact_nonconvex} is not jointly convex in both $P$ and $Q$, it is convex in one of
+them while the other is fixed. Therefore it can be solved via alternating optimization over $P$ and
+$Q$ using standard (projected) gradient descent algorithms, with the following partial derivatives:
 \begin{align*}
@@
 \end{align*}
-See \autocite[Section 6]{attributedSNE17} for a write up example of matrix optimization.
+Because the problem is not jointly convex, we are less concerned with the domains of the variables
+not being convex, as we anyway have no guarantee to find a global minimum. We also do not add a
+regularization term of the norm of the columns of $PQ^T$. Indeed, since $P\in\mathbb{M}^{d\times k}$
+and $Q\in\mathbb{M}^{|E|\times k}$, each column of the resulting $PQ^T$ is a linear combination of
+$k$ unit norm vectors of $P$ with weights whose square sum to $1$ and thus their norm is bounded by
+$\sqrt{k}$.
 
-Compared with \refeq{eq:edge_soft}, this matrix formulation requires learning more parameters, as it
-each edge has a different direction, albeit made up of a small number of basis direction.  Impose
-sparsity on $Q$?
+Compared with \refeq{eq:edge_soft}, these two matrix formulations require learning more parameters,
+as it each edge has a different direction, albeit made up of a small number of basis directions. In
+the second case, we could furthermore impose a sparsity constraint on $Q$.
 
@@
 \ldots, w_k\} \in \dsphere$ we are looking for. A natural way of obtaining directions from that
-partition is to set $w_\ell = \frac{s_\ell}{\left\|s_\ell\right\|}$, where $s_\ell = \sum_{u,v \in
-E_\ell} s_{uv}$ is the sum of all the similarity vectors of the edge in $E_\ell$. Note however that
+partition is to set $w_\ell$ to be the normalized cluster center of $E_\ell$. Note however that
 this does not guarantee that we have assigned the best direction to every edge. We therefore propose
@@
     them, that is as previously, set $\mathcal{D}_k = \left\{ w_\ell = \frac{s_\ell}{\left\| s_\ell
-    \right\|}\right\}_{\ell=1}^k $, where $s_\ell = \sum_{u,v \in E'_\ell} s_{uv}$,
+    \right\|}\right\}_{\ell=1}^k $, where $s_\ell = \sum_{u,v \in E'_\ell} s_{uv}$ is the sum of all
+    the similarity vectors of the edge in $E'_\ell$,
 \end{enumerate}

### 2017-10-16 edge/model.tex
explain how we could model k_local constraint

@@
 energy and can only express their interest in a few number of dimensions. In other words, all the
-edges incident to $u$ can only be associated with one of $k_\mathrm{local}$ directions. This is
-difficult to incorporate directly in the optimization formulation\stodo{actually it's not that
-difficult}, as for certain value of
-$k_\mathrm{local}$, this is a \NPc{} graph colouring problem. Still, it can be seen as a learning
-bias and will be used to generate synthetic data.
+edges incident to $u$ can only be associated with one of $k_\mathrm{local}$ directions. For
+certain value of
+$k_\mathrm{local}$, this is a \NPc{} graph colouring problem and thus cannot necessarily be enforced
+exactly. Still, it can be seen as a learning bias and will be used to generate synthetic data.
+Furthermore, we can also express it as an optimization constraint. Specifically, suppose that for
+an edge $(u,v)$ associated with the direction $w_i \in \mathcal{D}_k$, we have a $k$-dimensional
+vector $y_{u,v}$ equal to $e_i$ (that is all zero except for the \ith{} component set to one). Let
+us stack all such vectors into a $|E| \times k$ matrix $Y$, and let $C \in \Rbb^{|V| \times |E|}$ be
+the incidence matrix of $G$\footnote{The incidence matrix of a graph $G$, usually denoted by $B$, is
+the $|V| \times |E|$ matrix such that $B_{i,j} = 1$ if the node $v_i$ and edge $e_j$ are incident
+and 0 otherwise.} and $c_u$ be its $u^{th}$ row. Then $Yc_u$ is a $k$-dimensional vector that counts
+how many times each direction is incident to node $u$. We thus want the $\ell_0$ norm of all $Yc_u$
+(\ie its number of non-zero component) to be upper bounded by $k_\mathrm{local}$. As previously,
+this is not convex but we can relax this constraint, as such sparsity inducing problem have been
+well studied~\autocite{sparseOptim12}. Namely, we let $y_{uv}$ be a softmax membership vector, that
+is we set $y_{uv,i} = \frac{\exp(\beta{s_{uv}}^T w_i)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_i)}$,
+and we replace the $\ell_0$ norm by a $\ell_1$ or even a $k$-support norm~\autocite{KsupportNorm12}.
 These two constraints can be combined, so that the profiles are close to a linear combination of a
@@
 C\left(A\circ W^T \right)\|_F^2$, where $B\in \Rbb^{|V| \times d}$ is the node bias vectors $b_u$
-stacked vertically, $C \in \Rbb^{|V| \times |E|}$ is the incidence matrix of $G$\footnote{The
-  incidence matrix of a graph $G$, usually denoted by $B$, is the $|V| \times |E|$ matrix such that
-$B_{i,j} = 1$ if the node $v_i$ and edge $e_j$ are incident and 0 otherwise}, $A \in \Rbb^{|E|
+stacked vertically, $C \in \Rbb^{|V| \times |E|}$ is the incidence matrix of $G$, $A \in \Rbb^{|E|
 \times d}$ is the matrix whose each row is the corresponding $a_{uv}\onev$ vector and $\|\cdot\|_F$

### 2017-10-16 edge/experiments.tex
ove methods list, fix indentation and colored result macros

@@
-\paragraph{Synthetic data}
-\label{par:synthetic_data}
+\section{Experimental analysis}
+\label{sec:edge_exp}
+
+We start by extracting from the conceptual discussion of the previous section the five methods we
+compare in our experiments. They all take as input an attributed graph, \ie{} $G=(V, E)$ and $X$.
+For evaluation purposes, we describe how to extract from their output a set of $k$ directions and an
+assignment $\yhat$ of every edge to one of these directions (that is $\yhat$ is a function from $E$
+to $\rangesk$ such that $\yhat_{uv}$ is the index of the direction assigned to the edge $(u,v)$).
+
+\begin{enumerate}[1.]
+
+  \item Our baseline is a \kmeans{} clustering of the normalized $s_{uv}$ edge similarity
+    vectors. This naturally provides a \yhat{} assignment, and the directions are the normalized
+    cluster centers.
+
+  \item Our first method is a \lloyd{}-like iterative refinement of the $k$-means solution, where we
+    alternatively reassign edges based on their score and recompute the directions to maximize those
+    scores.
+
+  \item Then we consider the \textsc{Combined} optimization of the two edge and node terms of
+    \eqref{eq:edge_full}, which directly provides a set of directions. The assignment is then simply
+    the direction that gives the highest goodness.
+
+  \item As for matrix methods, we first do a \textsc{Frank--Wolfe} maximization of
+    \eqref{eq:edge_FW} over the $\delta$-ball of the nuclear norm. Afterwards, we perform a
+    \kmeans{} clustering of its columns and use it as assignment, while the directions are again the
+    normalized cluster centers.
+
+  \item \textsc{Explicit} low-rank $PQ^T$ factorization, initialized with the solution of
+    \combined{} (that is we set $P_0 \in \mathbb{M}^{d\times k}$ to be the final directions found by
+    \combined{} and the rows of $Q_0 \in \mathbb{M}^{|E|\times k}$ to be equal to the corresponding
+    $e_{y_{uv}}$) and minimizing the same objective of \eqref{eq:edge_PQ}. We take the directions to
+    be the columns of the resulting $P$, and the assignment to be a \kmeans{} clustering of the rows
+    of $Q$.
+\end{enumerate}
+
+\subsection{Synthetic data}
+\label{ssec:edge_synth}
+
+We try those five methods on data generated synthetically to fit our model describe previously and
+study how they perform under various conditions.
+
+\paragraph{Generation}
+\label{par:edge_generation}
 
@@
 
-\subparagraph{Results}
+\paragraph{Results}
 \label{par:edge_synth_results}
@@
 
-In total we compare five of the methods introduced earlier\stodo{make sure it's clear how we obtain
-directions and prediction from each method}:
-\begin{enumerate}[1.]
-  \item \textsc{$k$-means} clustering of the normalized $s_{uv}$ edge similarity vectors;
-  \item \textsc{Lloyd}-like iterative refinement of the $k$-means solution;
-  \item \textsc{Combined} optimization of $\mathcal{L}_{\mathrm{edge}}$ and
-    $\mathcal{L}_{\mathrm{node}}$;
-  \item \textsc{Frank--Wolfe} maximization of $\mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(W)$ over
-    the $\delta$-ball of the nuclear norm, with a regularization of the $\ell_2$ norm of the columns
-    of $W$. In other words, denoting by $\langle A, B \rangle_\mathrm{F} = \tr\left(A^T B\right)$
-    the Frobenius inner product of two real matrices, we solve
-    \begin{equation*}
-      \min_{\|W\|_* \leq \delta} -\frop{S^T}{W} + \mu \frop{W}{W}\,\text{; and}
-    \end{equation*}
-  \item \textsc{Explicit} low-rank $PQ^T$ factorization, initialized with the solution of
-    \textsc{Combined}\stodo{describe formally} and minimizing the same objective in matric form.
-\end{enumerate}
-
 As showed in \autoref{tab:edge_Wvaries_nami}, \combined{} is always the best method and, excluding
@@
     is statistically significant (that is we can confidently reject the hypothesis that the two
-    distribution have the same mean), we furthermore underline that score and show in parenthesis the
+    distribution have the same mean), we furthermore show in parenthesis the supporting
     $p$-value of a paired Student's $t$-test.
@@
     \midrule
-    {\smallk{}}  &            $.836 \pm .08$ &                          $\vsecond{.838 \pm .07}$ &  $\vfirstSig{.875 \pm .06}\,\,\spval{2.17}{-58}$ &  $.213 \pm .11$ &  $.875 \pm .06$ \\
-    {\default{}} &            $.818 \pm .06$ &  $\vsecondSig{.873 \pm .05}\,\,\spval{1.25}{-63}$ &  $\vfirstSig{.893 \pm .04}\,\,\spval{5.68}{-33}$ &  $.381 \pm .05$ &  $.893 \pm .04$ \\
-    {\largek{}}  &            $.803 \pm .04$ &  $\vsecondSig{.881 \pm .04}\,\,\spval{2.66}{-94}$ &  $\vfirstSig{.894 \pm .04}\,\,\spval{8.98}{-17}$ &  $.421 \pm .04$ &  $.894 \pm .04$ \\
-    {\smallo{}}  &            $.813 \pm .07$ &   $\vsecondSig{.824 \pm .06}\,\,\spval{7.57}{-6}$ &  $\vfirstSig{.856 \pm .06}\,\,\spval{2.99}{-57}$ &  $.378 \pm .05$ &  $.855 \pm .06$ \\
-    {\largeo{}}  &  $\vsecond{.827 \pm .07}$ &                                    $.823 \pm .06$ &  $\vfirstSig{.852 \pm .06}\,\,\spval{1.90}{-25}$ &  $.370 \pm .06$ &  $.851 \pm .06$ \\
-    {\fdirs{}}   &            $.772 \pm .07$ &  $\vsecondSig{.814 \pm .07}\,\,\spval{6.02}{-42}$ &  $\vfirstSig{.853 \pm .06}\,\,\spval{2.13}{-47}$ &  $.320 \pm .06$ &  $.853 \pm .06$ \\
-    {\larged{}}  &            $.905 \pm .05$ &  $\vsecondSig{.933 \pm .04}\,\,\spval{1.32}{-31}$ &  $\vfirstSig{.941 \pm .03}\,\,\spval{1.77}{-22}$ &  $.222 \pm .10$ &  $.931 \pm .04$ \\
+    {\smallk{}}  &            $.836 \pm .08$ &                          $\esecond{.838 \pm .07}$ &  $\efirstSig{.875 \pm .06}\,\,\spval{2.17}{-58}$ &  $.213 \pm .11$ &  $.875 \pm .06$ \\
+    {\default{}} &            $.818 \pm .06$ &  $\esecondSig{.873 \pm .05}\,\,\spval{1.25}{-63}$ &  $\efirstSig{.893 \pm .04}\,\,\spval{5.68}{-33}$ &  $.381 \pm .05$ &  $.893 \pm .04$ \\
+    {\largek{}}  &            $.803 \pm .04$ &  $\esecondSig{.881 \pm .04}\,\,\spval{2.66}{-94}$ &  $\efirstSig{.894 \pm .04}\,\,\spval{8.98}{-17}$ &  $.421 \pm .04$ &  $.894 \pm .04$ \\
+    {\smallo{}}  &            $.813 \pm .07$ &   $\esecondSig{.824 \pm .06}\,\,\spval{7.57}{-6}$ &  $\efirstSig{.856 \pm .06}\,\,\spval{2.99}{-57}$ &  $.378 \pm .05$ &  $.855 \pm .06$ \\
+    {\largeo{}}  &  $\esecond{.827 \pm .07}$ &                                    $.823 \pm .06$ &  $\efirstSig{.852 \pm .06}\,\,\spval{1.90}{-25}$ &  $.370 \pm .06$ &  $.851 \pm .06$ \\
+    {\fdirs{}}   &            $.772 \pm .07$ &  $\esecondSig{.814 \pm .07}\,\,\spval{6.02}{-42}$ &  $\efirstSig{.853 \pm .06}\,\,\spval{2.13}{-47}$ &  $.320 \pm .06$ &  $.853 \pm .06$ \\
+    {\larged{}}  &            $.905 \pm .05$ &  $\esecondSig{.933 \pm .04}\,\,\spval{1.32}{-31}$ &  $\efirstSig{.941 \pm .03}\,\,\spval{1.77}{-22}$ &  $.222 \pm .10$ &  $.931 \pm .04$ \\
     \bottomrule
@@
       \midrule
-      {\smallk{}}  &  $\vsecondSig{.581 \pm .08}\,\,\spval{4.78}{-30}$ &                                    $.606 \pm .09$ &  $\vfirstSig{.560 \pm .05}\,\,\spval{8.07}{-24}$ &  $1.007 \pm .05$ &  $.633 \pm .06$ \\
-      {\default{}} &                                    $.533 \pm .08$ &   $\vsecondSig{.528 \pm .09}\,\,\spval{9.61}{-4}$ &   $\vfirstSig{.516 \pm .06}\,\,\spval{4.95}{-5}$ &   $.877 \pm .05$ &  $.564 \pm .06$ \\
-      {\largek{}}  &                                    $.541 \pm .07$ &  $\vsecondSig{.521 \pm .07}\,\,\spval{5.08}{-23}$ &                          $\vfirst{.520 \pm .05}$ &   $.811 \pm .06$ &  $.551 \pm .06$ \\
-      {\smallo{}}  &  $\vsecondSig{.554 \pm .08}\,\,\spval{2.74}{-52}$ &                                    $.589 \pm .08$ &  $\vfirstSig{.538 \pm .05}\,\,\spval{2.38}{-12}$ &   $.908 \pm .05$ &  $.596 \pm .07$ \\
-      {\largeo{}}  &  $\vsecondSig{.554 \pm .09}\,\,\spval{1.50}{-58}$ &                                    $.595 \pm .08$ &   $\vfirstSig{.545 \pm .06}\,\,\spval{4.08}{-5}$ &   $.935 \pm .05$ &  $.604 \pm .08$ \\
-      {\fdirs{}}   &                                    $.565 \pm .08$ &                          $\vsecond{.564 \pm .09}$ &   $\vfirstSig{.546 \pm .05}\,\,\spval{6.30}{-8}$ &   $.928 \pm .06$ &  $.588 \pm .06$ \\
-      {\larged{}}  &  $\vsecondSig{.571 \pm .07}\,\,\spval{2.02}{-89}$ &                                    $.602 \pm .07$ &                          $\vfirst{.567 \pm .05}$ &  $1.033 \pm .06$ &  $.641 \pm .05$ \\
+      {\smallk{}}  &  $\esecondSig{.581 \pm .08}\,\,\spval{4.78}{-30}$ &                                    $.606 \pm .09$ &  $\efirstSig{.560 \pm .05}\,\,\spval{8.07}{-24}$ &  $1.007 \pm .05$ &  $.633 \pm .06$ \\
+      {\default{}} &                                    $.533 \pm .08$ &   $\esecondSig{.528 \pm .09}\,\,\spval{9.61}{-4}$ &   $\efirstSig{.516 \pm .06}\,\,\spval{4.95}{-5}$ &   $.877 \pm .05$ &  $.564 \pm .06$ \\
+      {\largek{}}  &                                    $.541 \pm .07$ &  $\esecondSig{.521 \pm .07}\,\,\spval{5.08}{-23}$ &                          $\efirst{.520 \pm .05}$ &   $.811 \pm .06$ &  $.551 \pm .06$ \\
+      {\smallo{}}  &  $\esecondSig{.554 \pm .08}\,\,\spval{2.74}{-52}$ &                                    $.589 \pm .08$ &  $\efirstSig{.538 \pm .05}\,\,\spval{2.38}{-12}$ &   $.908 \pm .05$ &  $.596 \pm .07$ \\
+      {\largeo{}}  &  $\esecondSig{.554 \pm .09}\,\,\spval{1.50}{-58}$ &                                    $.595 \pm .08$ &   $\efirstSig{.545 \pm .06}\,\,\spval{4.08}{-5}$ &   $.935 \pm .05$ &  $.604 \pm .08$ \\
+      {\fdirs{}}   &                                    $.565 \pm .08$ &                          $\esecond{.564 \pm .09}$ &   $\efirstSig{.546 \pm .05}\,\,\spval{6.30}{-8}$ &   $.928 \pm .06$ &  $.588 \pm .06$ \\
+      {\larged{}}  &  $\esecondSig{.571 \pm .07}\,\,\spval{2.02}{-89}$ &                                    $.602 \pm .07$ &                          $\efirst{.567 \pm .05}$ &  $1.033 \pm .06$ &  $.641 \pm .05$ \\
       \bottomrule
@@
 
-\iffalse
-\begin{table}[hbt]
-  \centering
+\begin{aside}
+
+% \begin{table}[hbt]
+%   \centering
   % \setlength{\tabcolsep}{3pt}
+  % \caption{%
+\begin{center}
   \small
-  \caption{Same as \autoref{tab:edge_Wvaries_nami}, but when $\mathcal{D}_k$ is fixed and generating
+   \captionsetup{font=small}
+   \captionof{table}{%
+    Same as \autoref{tab:edge_Wvaries_nami}, but when $\mathcal{D}_k$ is fixed and generating
     200 graphs.  \label{tab:edge_Gvaries_nami}}
@@
       \midrule
-      {\smallk{}}  &  $\vsecondSig{.678 \pm .07}\,\,\spval{4.16}{-22}$ &                                    $.657 \pm .07$ &  $\vfirstSig{.729 \pm .08}\,\,\spval{1.13}{-62}$ &  $.330 \pm .08$ &  $.729 \pm .08$ \\
-      {\default{}} &                                    $.745 \pm .04$ &  $\vsecondSig{.799 \pm .04}\,\,\spval{2.88}{-65}$ &  $\vfirstSig{.839 \pm .03}\,\,\spval{1.75}{-68}$ &  $.379 \pm .04$ &  $.839 \pm .03$ \\
-      {\largek{}}  &                                    $.769 \pm .03$ &  $\vsecondSig{.843 \pm .04}\,\,\spval{5.06}{-84}$ &  $\vfirstSig{.870 \pm .04}\,\,\spval{8.54}{-40}$ &  $.431 \pm .04$ &  $.870 \pm .04$ \\
-      {\smallo{}}  &                                    $.813 \pm .06$ &  $\vsecondSig{.845 \pm .04}\,\,\spval{1.44}{-27}$ &  $\vfirstSig{.869 \pm .04}\,\,\spval{3.98}{-55}$ &  $.371 \pm .05$ &  $.868 \pm .04$ \\
-      {\largeo{}}  &                                    $.788 \pm .06$ &   $\vsecondSig{.799 \pm .05}\,\,\spval{1.54}{-6}$ &  $\vfirstSig{.830 \pm .05}\,\,\spval{2.36}{-57}$ &  $.408 \pm .05$ &  $.830 \pm .05$ \\
-      {\fdirs{}}   &                                    $.678 \pm .05$ &  $\vsecondSig{.700 \pm .06}\,\,\spval{1.65}{-24}$ &  $\vfirstSig{.761 \pm .06}\,\,\spval{2.73}{-80}$ &  $.336 \pm .05$ &  $.761 \pm .06$ \\
-      {\larged{}}  &                                    $.863 \pm .06$ &  $\vsecondSig{.893 \pm .05}\,\,\spval{8.68}{-44}$ &  $\vfirstSig{.907 \pm .05}\,\,\spval{5.58}{-31}$ &  $.276 \pm .08$ &  $.895 \pm .05$ \\
+      {\smallk{}}  &  $\esecondSig{.678 \pm .07}\,\,\spval{4.16}{-22}$ &                                    $.657 \pm .07$ &  $\efirstSig{.729 \pm .08}\,\,\spval{1.13}{-62}$ &  $.330 \pm .08$ &  $.729 \pm .08$ \\
+      {\default{}} &                                    $.745 \pm .04$ &  $\esecondSig{.799 \pm .04}\,\,\spval{2.88}{-65}$ &  $\efirstSig{.839 \pm .03}\,\,\spval{1.75}{-68}$ &  $.379 \pm .04$ &  $.839 \pm .03$ \\
+      {\largek{}}  &                                    $.769 \pm .03$ &  $\esecondSig{.843 \pm .04}\,\,\spval{5.06}{-84}$ &  $\efirstSig{.870 \pm .04}\,\,\spval{8.54}{-40}$ &  $.431 \pm .04$ &  $.870 \pm .04$ \\
+      {\smallo{}}  &                                    $.813 \pm .06$ &  $\esecondSig{.845 \pm .04}\,\,\spval{1.44}{-27}$ &  $\efirstSig{.869 \pm .04}\,\,\spval{3.98}{-55}$ &  $.371 \pm .05$ &  $.868 \pm .04$ \\
+      {\largeo{}}  &                                    $.788 \pm .06$ &   $\esecondSig{.799 \pm .05}\,\,\spval{1.54}{-6}$ &  $\efirstSig{.830 \pm .05}\,\,\spval{2.36}{-57}$ &  $.408 \pm .05$ &  $.830 \pm .05$ \\
+      {\fdirs{}}   &                                    $.678 \pm .05$ &  $\esecondSig{.700 \pm .06}\,\,\spval{1.65}{-24}$ &  $\efirstSig{.761 \pm .06}\,\,\spval{2.73}{-80}$ &  $.336 \pm .05$ &  $.761 \pm .06$ \\
+      {\larged{}}  &                                    $.863 \pm .06$ &  $\esecondSig{.893 \pm .05}\,\,\spval{8.68}{-44}$ &  $\efirstSig{.907 \pm .05}\,\,\spval{5.58}{-31}$ &  $.276 \pm .08$ &  $.895 \pm .05$ \\
       \bottomrule
     \end{tabular}
-\end{table}
+% \end{table}
+\end{center}
 
-\begin{table}[hbt]
-  \centering
+% \begin{table}[hbt]
+%   \centering
   % \setlength{\tabcolsep}{3pt}
+\begin{center}
   \small
-  \caption{Same as \autoref{tab:edge_Wvaries_dst}, but when $\mathcal{D}_k$ is fixed and generating
+   \captionsetup{font=small}
+   \captionof{table}{%
+  % \caption{%
+    Same as \autoref{tab:edge_Wvaries_dst}, but when $\mathcal{D}_k$ is fixed and generating
     200 graphs.  \label{tab:edge_Gvaries_dst}}
@@
       \midrule
-      {\smallk{}}  &  $\vsecondSig{.702 \pm .08}\,\,\spval{2.27}{-69}$ &                                    $.754 \pm .06$ &  $\vfirstSig{.647 \pm .06}\,\,\spval{1.59}{-81}$ &   $.975 \pm .06$ &  $.734 \pm .07$ \\
-      {\default{}} &                          $\vsecond{.607 \pm .06}$ &                                    $.608 \pm .07$ &  $\vfirstSig{.567 \pm .04}\,\,\spval{2.55}{-69}$ &   $.897 \pm .04$ &  $.622 \pm .05$ \\
-      {\largek{}}  &                                    $.584 \pm .06$ &  $\vsecondSig{.563 \pm .07}\,\,\spval{2.79}{-24}$ &                          $\vfirst{.561 \pm .05}$ &   $.804 \pm .05$ &  $.582 \pm .06$ \\
-      {\smallo{}}  &  $\vsecondSig{.530 \pm .08}\,\,\spval{1.20}{-67}$ &                                    $.571 \pm .07$ &   $\vfirstSig{.518 \pm .05}\,\,\spval{5.10}{-8}$ &   $.894 \pm .05$ &  $.572 \pm .06$ \\
-      {\largeo{}}  &  $\vsecondSig{.607 \pm .08}\,\,\spval{1.24}{-70}$ &                                    $.647 \pm .07$ &  $\vfirstSig{.578 \pm .05}\,\,\spval{6.06}{-35}$ &   $.896 \pm .06$ &  $.636 \pm .07$ \\
-      {\fdirs{}}   &  $\vsecondSig{.655 \pm .08}\,\,\spval{2.30}{-32}$ &                                    $.677 \pm .08$ &  $\vfirstSig{.608 \pm .06}\,\,\spval{6.88}{-71}$ &   $.938 \pm .04$ &  $.671 \pm .07$ \\
-      {\larged{}}  &  $\vsecondSig{.603 \pm .07}\,\,\spval{9.25}{-95}$ &                                    $.636 \pm .07$ &  $\vfirstSig{.589 \pm .05}\,\,\spval{6.97}{-12}$ &  $1.003 \pm .07$ &  $.662 \pm .06$ \\
+      {\smallk{}}  &  $\esecondSig{.702 \pm .08}\,\,\spval{2.27}{-69}$ &                                    $.754 \pm .06$ &  $\efirstSig{.647 \pm .06}\,\,\spval{1.59}{-81}$ &   $.975 \pm .06$ &  $.734 \pm .07$ \\
+      {\default{}} &                          $\esecond{.607 \pm .06}$ &                                    $.608 \pm .07$ &  $\efirstSig{.567 \pm .04}\,\,\spval{2.55}{-69}$ &   $.897 \pm .04$ &  $.622 \pm .05$ \\
+      {\largek{}}  &                                    $.584 \pm .06$ &  $\esecondSig{.563 \pm .07}\,\,\spval{2.79}{-24}$ &                          $\efirst{.561 \pm .05}$ &   $.804 \pm .05$ &  $.582 \pm .06$ \\
+      {\smallo{}}  &  $\esecondSig{.530 \pm .08}\,\,\spval{1.20}{-67}$ &                                    $.571 \pm .07$ &   $\efirstSig{.518 \pm .05}\,\,\spval{5.10}{-8}$ &   $.894 \pm .05$ &  $.572 \pm .06$ \\
+      {\largeo{}}  &  $\esecondSig{.607 \pm .08}\,\,\spval{1.24}{-70}$ &                                    $.647 \pm .07$ &  $\efirstSig{.578 \pm .05}\,\,\spval{6.06}{-35}$ &   $.896 \pm .06$ &  $.636 \pm .07$ \\
+      {\fdirs{}}   &  $\esecondSig{.655 \pm .08}\,\,\spval{2.30}{-32}$ &                                    $.677 \pm .08$ &  $\efirstSig{.608 \pm .06}\,\,\spval{6.88}{-71}$ &   $.938 \pm .04$ &  $.671 \pm .07$ \\
+      {\larged{}}  &  $\esecondSig{.603 \pm .07}\,\,\spval{9.25}{-95}$ &                                    $.636 \pm .07$ &  $\efirstSig{.589 \pm .05}\,\,\spval{6.97}{-12}$ &  $1.003 \pm .07$ &  $.662 \pm .06$ \\
       \bottomrule
     \end{tabular}
-\end{table}
-\fi
+% \end{table}
+\end{center}
+
+\end{aside}

### 2017-10-16 edge/experiments.tex
reorganize the model material

@@
 find a set of user profiles $X=\{x_u\}_{u\in V}$ that maximize the edge score and minimize the node
-loss, defined respectively by equations \eqref{eq:edge_full}\stodo{not really} and \eqref{} in the
-previous section as:
+loss, defined respectively by equations \eqref{eq:edge_full} and
+\eqref{eq:edge_node_loss} in the previous section as:
 \begin{align*}

### 2017-10-16 edge/model.tex
reorganize the model material

@@
-\section{Settings and learning problem}
-\label{sec:edge_model}
+\section{Learning problem}
+\label{sec:edge_problem}
 
-As in the previous chapters, we are given an unweighted graph $G=(V,E)$\stodo{undirected?}. This
+As in the previous chapter, we are given an undirected, unweighted graph $G=(V,E)$. This
 time though, we also have side information in addition to the graph topology. Namely, each node $u$
@@
 network, $x_u$ would be the profile of user $u$, describing $u$'s demographics and preferences. We
-are also given a combination operator between feature vectors $s$ such that $s(x_u, x_v) = s_{uv}
-\in \Rbb^d$. Finally, we consider distinguished feature vectors $w_{uv}$\stodo{is it better to
-introduce $w$ with a single $\ell$ index, to emphasise that later, it will not be unique per edge?},
+are also given a combination operator $s$ between feature vectors, such that $s(x_u, x_v) = s_{uv}
+\in \Rbb^d$. Finally, we consider distinguished feature vectors $w_{uv}$,
 along which the combination $s_{uv}$ between two nodes $u$ and $v$ is evaluated. We call such
-$w_{uv}$ \emph{directions}, and constrain them to have unit norm. As we will discuss, this
-constraint can be seen as a way to couple the different dimensions of the profiles.
+$w_{uv}$ \emph{directions}, and constrain them to have unit norm\stodo{or bounded?}. As we will discuss, this
+constraint can be seen as a way to couple the different dimensions of the profiles%
+\stodo{don't remember what I meant here}.
 
@@
 We assume in the following that both the profiles and the directions are vectors with components in
-$[-1, 1]$. A natural choice for the operator $s$ is then the difference between the profile, that is
+$[-1, 1]$. A natural choice for the operator $s$ is then the difference between the profiles, that is
 $s_{uv} = x_u - x_v$. In that spirit, if $u$ and $v$ are close along a direction $w_{uv}$, we expect
@@
 w_{uv}) = \left({s_{uv}}^T w_{uv} \right)^2$, and we want to minimize it with respect to $w_{uv}$.
-When ${x_u}_{;i} - {x_v}_{;i}$ is close to zero, the minimization and the fact that $w_{uv}$ is
+When on the \ith{} dimension ${x_u}_{;i} - {x_v}_{;i}$ is close to zero, the minimization objective
+and the fact that $w_{uv}$ is
 unit-norm impose that ${w_{uv}}_{;i}$ should be large. On the other hand, with this formulation,
-there is no way to set ${w_{uv}}_{;i}$ in order to \enquote{reward/favor} the fact that
+there is no way to set ${w_{uv}}_{;i}$ in order to \enquote{highlight} the fact that
 ${x_u}_{;i}$ and ${x_v}_{;i}$ are widely different.
@@
 
+\iffalse
 \begin{aside}
@@
 \end{aside}
+\fi
 
@@
 
-Here we show why \autoref{p:edge_full} can model signed graphs as a special case, namely with $k=2$.
+Here we show why \autoref{p:edge_full} can model signed graphs as a special case, namely with $k=2$
+and specifically crafted profiles.
 Remember the learning
@@
 let the dimension of profiles be $d=K$ and the number of directions $k=2$. Moreover, we define the
-two directions vectors as $w_1 = \frac{\onev}{\sqrt{d}} = w^+$ and $w_2 = -w_+ = w_-$, where
+two direction vectors as $w_1 = \frac{\onev}{\sqrt{d}} = w^+$ and $w_2 = -w_+ = w_-$, where
 $\onev$ is the vector of dimension $d$ with only $1$ coordinates. Finally, for a node $u$ in
 cluster $i$, we assume its profile $x_u$ is set as follow: all component are equal to $-b$ except
-for the \ith{} one, which is equal to $a$, where $a$ and $b$ are constants we define now.
+for the \ith{} one, which is equal to $a$, where $a$ and $b$ are real constants we define now.
 
@@
 \sqrt{1-(K-1)b^2}$ whenever $1-(K-1)b^2 \geq 0$, or equivalently $b \leq \frac{1}{\sqrt{K-1}}$.
-Next, we look at $\onev^T\left(x_u \circ x_v\right) = \onev^T m_{uv} = \frac{1}{\sqrt{d}}
-\sum_{i=1}^d m_{{uv}_i}$ for any edge $(u,v) \in E$. If $u$ and $v$ are in the same cluster, then
-$\sum_{i=1}^d m_{{uv}_i} = a^2 + (K-1)b^2 = 1$, implying that $w_+^T m_{uv} = \frac{1}{d} > w_-^T
-m_{uv} = -\frac{1}{d}$. Otherwise, $\sum_{i=1}^d m_{{uv}_i} = (K-2)b^2 - 2ab = (K-2)b^2 -
+Next, we look at $\left(x_u \circ x_v\right)^T \onev = {s_{uv}}^T \onev = \frac{1}{\sqrt{d}}
+\sum_{i=1}^d  {s_{uv}}_{;i}$ for any edge $(u,v) \in E$. If $u$ and $v$ are in the same cluster, then
+$\sum_{i=1}^d  {s_{uv}}_{;i} = a^2 + (K-1)b^2 = 1$, implying that ${s_{uv}}^T w_+ = \frac{1}{d} >
+{s_{uv}}^T w_- = -\frac{1}{d}$. Otherwise, $\sum_{i=1}^d  {s_{uv}}_{;i} = (K-2)b^2 - 2ab = (K-2)b^2 -
 2b\sqrt{1-(K-1)b^2} = f_K(b)$. By computing the derivative of $f_K$ with respect to $b$, we find
 that it reaches a minimum of $-\frac{1}{K-1}$ at $b = \frac{1}{\sqrt{K(K-1)}} \leq
-\frac{1}{\sqrt{K-1}}$. In this case, $w_+^T m_{uv} = -\frac{1}{(K-1) \sqrt{d}} < w_-^T m_{uv} =
+\frac{1}{\sqrt{K-1}}$. In this case, ${s_{uv}}^T w_+ = -\frac{1}{(K-1) \sqrt{d}} < {s_{uv}}^T w_- =
 \frac{1}{(K-1) \sqrt{d}}$. To maximize $\sum_{(u,v)\in E} \max_{1 \leq \ell \leq d} w_\ell^T
-m_{uv}$, we therefore assign $w_+$ to edges whose both endpoints belong to the same cluster, and
+s_{uv}$, we therefore assign $w_+$ to edges whose both endpoints belong to the same cluster, and
 $w_-$ to edges joining two different clusters.
@@
 Moreover, given this assignment, we can show that $w_+$ and $w_-$ are the optimal directions, under
-the additional uniformity assumption that there are exactly $r$ edges within every cluster, and $s$
+the additional uniformity assumption that there are exactly $p$ edges within every cluster, and $q$
 edges between any two clusters. First, we define
 \begin{equation*}
-  m_{\mathrm{inner}} = \sum_{(u,v) \in E;\, \cluster(u)=\cluster(v)} m_{uv} \qquad \text{and} \qquad
-  m_{\mathrm{outer}} = \sum_{(u,v) \in E;\, \cluster(u)\neq \cluster(v)} m_{uv}
+  s_{\mathrm{inner}} = \sum_{(u,v) \in E;\, \cluster(u)=\cluster(v)} s_{uv} \qquad \text{and} \qquad
+  s_{\mathrm{outer}} = \sum_{(u,v) \in E;\, \cluster(u)\neq \cluster(v)} s_{uv}
 \end{equation*}
 to be the sum of all edge vectors within and between clusters respectively. By our choice of $a$ and
-$b$, we have that $m_{\mathrm{inner}} = r\onev$ and $m_{\mathrm{outer}} = -\frac{s}{2}\onev$. Now
+$b$, we have that $s_{\mathrm{inner}} = p\onev$ and $s_{\mathrm{outer}} = -\frac{q}{2}\onev$. Now
 recall that the solution of $\argmax_{w\in \dsphere} w^T c$ is $\frac{c}{||c||}$. Therefore, the
@@
 \begin{aside}
-For inner edge vectors, $m_{uv}$ in cluster $i$ has $a^2$ as its \ith{} component, and $b^2$
-elsewhere. There are $r$ of them in cluster $i$, so summing over all clusters we get
-$m_{\mathrm{inner}} = r\left(a^2 + (K-1)b^2\right)\onev = r\onev$.
-For outer edge vectors, there are a few more steps. Consider the $s$ edges between $C_1$ and $C_2$,
-whose $m_{uv}$ is
+For inner edge vectors, $s_{uv}$ in cluster $i$ has $a^2$ as its \ith{} component, and $b^2$
+elsewhere. There are $p$ of them in cluster $i$, so summing over all clusters we get
+$s_{\mathrm{inner}} = p\left(a^2 + (K-1)b^2\right)\onev = p\onev$.
+For outer edge vectors, we first consider the $q$ edges between $C_1$ and $C_2$,
+whose $s_{uv}$ is
 $\begin{pmatrix}
@@
 \end{pmatrix}$. Likewise, between $C_1$ and $C_3$ we have
-$m_{uv}=\begin{pmatrix}
+$s_{uv}=\begin{pmatrix}
   -ab \\
@@
 $C_1$, we get
-$s\begin{pmatrix}
+$q\begin{pmatrix}
   -ab(K-1) \\
@@
 \end{pmatrix}$. Similarly, the sum of all vectors incident to $C_2$ is
-$s\begin{pmatrix}
+$q\begin{pmatrix}
   b^2(K-2) -ab \\
@@
 \end{pmatrix}$, and so on for every cluster $i$. Summing over all clusters and dividing by $2$ to
-avoid double counting, we get $m_{\mathrm{outer}} = \frac{s}{2}\left( (K-1)\left(b^2(K-2) -ab\right)
--ab(K-1) \right)\onev = \frac{s(K-1)}{2}\left(b^2(K-2) -2ab\right)\onev =
-\frac{s(K-1)}{2}f_K(b)\onev = -\frac{s(K-1)}{2} \frac{1}{K-1} \onev = -\frac{s}{2}\onev$.
+avoid double counting, we get $s_{\mathrm{outer}} = \frac{q}{2}\left( (K-1)\left(b^2(K-2) -ab\right)
+-ab(K-1) \right)\onev = \frac{q(K-1)}{2}\left(b^2(K-2) -2ab\right)\onev =
+\frac{q(K-1)}{2}f_K(b)\onev = -\frac{q(K-1)}{2} \frac{1}{K-1} \onev = -\frac{q}{2}\onev$.
 \end{aside}
@@
 
-At the other extreme of $k=2$, when $k=|E|$, the topology of the graph becomes irrelevant,
-as we can simply exploit the trivial
+At the other extreme of having only $k=2$ directions, when $k=|E|$, the topology of the graph
+becomes irrelevant, as we can simply exploit the trivial
 solution of having a single direction $w_{uv} = \frac{s_{uv}}{||s_{uv}||}$ for every edge $(u,v)$.
-More generally, as $k$ increases, we expect the value $\max_{\mathcal{D}_k} f(k)$ to increases as
+More generally, as $k$ increases, we expect the value $\max_{\mathcal{D}_k} f(k)$ to increase as
 well, at the cost of interpretability. There could be a principled approach to finding the
@@
 less than \np{10}). Within the multilayer framework, $k$ could also be seen as the number of
-underlying layers\stodo{elaborate on that connection}.
+underlying layers. To retain a fair level of interpretability, we deem appropriate to have $k$ not
+too large.
 
-\stodo{move that in the conclusion}
-\begin{aside}
-Generalizing the balance theory from signed graphs, we could imagine that in triangles or short
-cycles, only certain combinations of directions are valid (or at least desirable). As a nice side
-effect, this would likely require more innovative optimization algorithms. However, it sounds very
-application-dependant. Furthermore, in our current setting, we do not know in advance in the
-semantic of each direction, and we have no labeled edges. Therefore, it is not clear how those
-constraints would be specified or whether they can be learned, if applicable. 
-\end{aside}
-\stodo{Specification of constraints sounds related to network schema in \autocite[Definition
-3]{HINSurvey17}, except we have a single node type and are looking at higher order patter.}
-
-Another way to leverage the graph topology is to define local constraints at the node level. Since
-the beginning we considered that the user profiles were given, but from now on we further assume
-that they are normalized so that they are unit norm, just like directions. In practice, if nodes are
+In addition to assume that the input graph is the superposition of a small number of layers,
+another way to leverage the graph topology is to define local constraints at the node level. 
+Recall that we consider node profiles to be normalized and having unit norm. A practical
+justification is that if nodes are
 users of a social network and the profiles measure their activity across several domains, this
@@
 connect to another node $v$ along a direction $w_{uv}$, it consume a part of its profile
-proportional to $w_{uv}$. Therefore, we subtract the term $\sum_{u\in V} \left|\left| x_u - b_u -
-\sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2$ to \eqref{eq:edge_full}.
+proportional to $w_{uv}$. Therefore, we subtract the term 
+\begin{equation}
+  \label{eq:edge_node_loss}
+  \mathcal{L}_{\mathrm{node}} = 
+  \sum_{u\in V} \left|\left| x_u - b_u - \sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2
+\end{equation}
+from \eqref{eq:edge_full}.
 
@@
 
-% $\theset{(u,v) \in E}{v \in \nei(u)}$
+\section{Proposed approaches}
+\label{sec:edge_methods}
 
-\bigskip
+After having introduced \autoref{p:edge_full} and some additional constraints, we now present
+methods to solve it. We start with a naive baseline, and propose a post-processing to improve it.
+Then we describe a convex relaxation. Finally, we depart slightly from finding $k$ directions by
+considering a low-rank matrix formulation, where each edge is assigned a linear combination of a
+small number of base directions, paving the way to overlapping clustering.
 
-In order to make the optimization more tractable, we relax the formulation of \autoref{p:edge_full}
-by replacing the $\max$ inside the sum of \eqref{eq:edge_full}, either by a continuous approximation
-or by a low rank matrix approach.
+\subsection{\kmeans{} baseline and improvement}
+\label{sub:edge_baseline}
 
-One can check that given a set of numbers $S=\{a_1, a_2, \ldots, a_{|S|}\}$ and real number $\beta >
-0$, we have
+A simple and natural baseline is to cluster the set of all similarity edge vectors $
+\theset{s_{uv}}{(u,v) \in E}$ using the $k$-means algorithm, where $k$ is the number of directions
+set in \autoref{p:edge_full}. Formally, given the graph $G=(V,E)$ and the node profiles $X$ as
+input, we order the edges from $e_1$ to $e_m$ and build the matrix $S \in \Rbb^{m\times d}$ whose
+\ith{} row is the similarity between the profiles of the endpoints of $e_i=(u,v)$, that is $s_{uv} =
+x_u \circ x_v$. Running $k$-means will thus partition $E$ into $k$ sets of edges, call them $E_1,
+\ldots, E_k$. The first drawback of that simplicity is that it does not solve \autoref{p:edge_full}.
+Indeed, the partition of $E$ does not provide the set of directions $\mathcal{D}_k = \{w_1, w_2,
+\ldots, w_k\} \in \dsphere$ we are looking for. A natural way of obtaining directions from that
+partition is to set $w_\ell$ to be the normalized cluster center of $E_\ell$. Note however that
+this does not guarantee that we have assigned the best direction to every edge. We therefore propose
+the following heuristic, inspired by the Lloyd algorithm for $k$-means, but where the directions in
+$\mathcal{D}_k$ play the role of centroids. Namely, we alternate between two steps
+\begin{enumerate}[(i), nosep]
+  \item create a new partition $\{E'_1, \ldots, E'_k\}$ by assigning every edge $(u,v)$ to the
+    direction $w$ maximizing its score ${s_{uv}}^T w$, that is $E'_\ell = \theset{(u,v) \in E}{
+    w_\ell = \argmax_{w_i \in \mathcal{D}_k} {s_{uv}}^T w_i}$, and
+  \item update the directions to make them optimal with respect to the edges currently assigned to
+    them, that is as previously, set $\mathcal{D}_k = \left\{ w_\ell = \frac{s_\ell}{\left\| s_\ell
+    \right\|}\right\}_{\ell=1}^k $, where $s_\ell = \sum_{u,v \in E'_\ell} s_{uv}$ is the sum of all
+    the similarity vectors of the edge in $E'_\ell$,
+\end{enumerate}
+until the edge assignment is stable or we reach a maximum number of iterations. By analogy with the
+$k$-means algorithm, obtaining convergence guarantees in the general case would be surprising, but in
+practice we expect that only a small number of iterations would be needed.
+
+\subsection{Vector optimization}
+\label{sub:edge_vector}
+
+The difficulty in optimizing directly \eqref{eq:edge_full} stem from the $\max$ operator, which we
+replace here by a convex relaxation. Indeed, one can check that given a set of numbers $S=\{a_1,
+a_2, \ldots, a_{|S|}\}$ and real number $\beta > 0$, we have
 \begin{equation*}
@@
 
+\subsection{Matrix optimization}
+\label{sub:edge_matrix}
+
 Another way to remove the $\max$ is to first seek one direction for every edge, that is to maximize
@@
 \frop{S^T}{W}$, where $S \in \Rbb^{|E| \times d}$ is the $s_{uv}$ vectors for all edges stacked
-vertically and $W \in \Rbb^{d \times |E|}$ is the matrix of all directions stacked horizontally.
+vertically and $W \in \Rbb^{d \times |E|}$ is the matrix of all directions stacked horizontally
+and $\frop{A}{B} = \tr\left(A^T B\right)$ is the Frobenius inner product of two real matrices.
 More specifically, because of the unit norm constraint on directions, we look for $W$ in the set
@@
 \footnote{As mentioned in \autoref{sec:troll_related} about \autocite{OnlineCompletion17}, a tighter
-relaxation of the low-rank constraint is the max norm, but is it convex?} This
+convex relaxation of the low-rank constraint is the max norm, but we do not pursue that venue
+further.} This
 results in the following optimization problem: \[\min_{W\in \mathbb{M}^{d\times |E|}} -\frop{S^T}{W}
@@
 
-\iffalse
-stacked horizontally in
-one matrix. Then we can directly minimize \eqref{eq:edge_full} while enforcing the matrix
-$W\in\mathbb{R}^{d\times |E|}$ of all directions to be low-rank, so that each $w_{uv}$ (given by the
-$(u-1)n+v$-th column of $W$) can be expressed as a linear combination of a small number of basis
-vectors. Formally, if $W$ has rank $k$, then there exists $P\in\mathbb{R}^{d\times k}$ and
-$Q\in\mathbb{R}^{|E|\times k}$ such that $W=PQ^T$ and $w_{uv} = Pq_{uv}^T$. $P$ defines $k$
-representative directions in $\Rbb^d$ while $Q$ gives, for each edge, the weights of each
-representative direction. Similar to what is done in spectral clustering, we then cluster the
-$q_{uv}$'s to obtain an edge labeling.
-%
-Denoting by $\mathbb{M}^{d\times |E|}$ the set of $d\times |E|$ matrices with unit $L_2$ norm
-columns, a convex formulation\stodo{justify the convexity later: $L_edge$ is linear and both
-$L_node$ and $\|W\|_*$ are norm and thus convex} of the problem is
-as follows:
-\begin{align}
-  \label{eq:fact_convex}
-  \max_{W\in\mathbb{M}^{d\times |E|}} h(W) &=
-  \mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(W) -
-  \mu \mathcal{L}_{\mathrm{node}}^{\mathrm{mat}}(W) - \lambda \|W\|_* \\
-  &= \frac{1}{|E|} \tr\left(SW\right) -
-  \frac{\mu}{|V|} \|X - B - C\left(A\circ W^T \right)\|_2^2 - \lambda \|W\|_*
-  % \sum_{i,j \in E} {s_{uv}}^T w_{uv} + \lambda \|W\|_*,
-\end{align}
-where $S \in \Rbb^{|E| \times d}$ is the $s_{uv}$ vectors for all edges stacked vertically, $\mu
-\geq 0$ is a trade-off parameter, $B\in \Rbb^{|V| \times d}$ is the node bias vectors $b_u$ stacked
-vertically, $C \in \Rbb^{|V| \times |E|}$ is the incidence matrix of $G$\footnote{The incidence
-matrix of a graph $G$, usually denoted by $B$, is the $|V| \times |E|$ matrix such that $B_{i,j} =
-1$ if the node $v_i$ and edge $e_j$ are incident and 0 otherwise}, $A \in \Rbb^{|E| \times d}$ is
-the matrix whose each row is the corresponding $a_{uv}\onev$ vector, $\|\cdot\|_2$ denotes the
-Frobenius norm, $\|W\|_* =
-\sum_{i=1}^{\min(d,|E|)}\sigma_i(W)$ is the sum of the singular values of $W$ (called the nuclear
-norm or trace norm), and $\lambda \geq 0$ is a regularization parameter.\footnote{To make the
-optimal solution unique, we can use an additional Frobenius norm regularization on $W$.} The nuclear norm
-favors low-rank solutions as it is essentially an $L_1$ norm on the spectrum.\footnote{As mentioned
-in \autoref{sec:troll_related} about \autocite{OnlineCompletion17}, a tighter relaxation of the
-low-rank constraint is the max norm.}
-Both $\mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(W)$ and
-$\mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(W)$ are convex in $W$, with respective gradient
-\begin{align*}
-  \frac{\partial \mathcal{L}_{\mathrm{edge}}^{\mathrm{mat}}(W)}{\partial W} &=
-  \frac{1}{|E|} S^T \\
-  \frac{\partial \mathcal{L}_{\mathrm{node}}^{\mathrm{mat}}(W)}{\partial W} &= \frac{-2}{|V|}
-  A^T \circ \left(\Big(\left(X-B\right)^T -A^T \circ \left(W C^T\right) \Big) C\right)
-\end{align*}
-Problem~\eqref{eq:fact_convex} can be solved using proximal gradient descent (requires singular
-value decomposition at each iteration) \autocite{Parikh2013a} or Frank-Wolfe (requires only to find
-the largest singular vector) \autocite{Jaggi2013a}\stodo{look at the rest of the tutorial and
-comment on the difficulty of the LMO step}.
-\fi
+\medskip
 
@@
   \min_{P\in\mathbb{M}^{d\times k}, Q\in\mathbb{M}^{|E|\times k}} \quad \tilde{h}(P, Q) =
-  - \frop{S}{PQ^T} + \mu \frop{} \|X - B - C\left(A\circ QP^T \right)\|_F^2
+  - \frop{S}{PQ^T} + \mu \|X - B - C\left(A\circ QP^T \right)\|_F^2
 \end{equation}
-While \eqref{eq:fact_nonconvex} is not jointly convex in both $P$ and $Q$, it is convex in one of
+While \eqref{eq:edge_PQ} is not jointly convex in both $P$ and $Q$, it is convex in one of
 them while the other is fixed. Therefore it can be solved via alternating optimization over $P$ and
@@
 the second case, we could furthermore impose a sparsity constraint on $Q$.
-
-A simple and natural baseline is to cluster the set of all similarity edge vectors $
-\theset{s_{uv}}{(u,v) \in E}$ using the $k$-means algorithm, where $k$ is the number of directions
-set in \autoref{p:edge_full}. Formally, given the graph $G=(V,E)$ and the node profiles $X$ as
-input, we order the edges from $e_1$ to $e_m$ and build the matrix $S \in \Rbb^{m\times d}$ whose
-\ith{} row is the similarity between the profiles of the endpoints of $e_i=(u,v)$, that is $s_{uv} =
-x_u \circ x_v$. Running $k$-means will thus partition $E$ into $k$ sets of edges, call them $E_1,
-\ldots, E_k$. The first drawback of that simplicity is that it does not solve \autoref{p:edge_full}.
-Indeed, the partition of $E$ does not provide the set of directions $\mathcal{D}_k = \{w_1, w_2,
-\ldots, w_k\} \in \dsphere$ we are looking for. A natural way of obtaining directions from that
-partition is to set $w_\ell$ to be the normalized cluster center of $E_\ell$. Note however that
-this does not guarantee that we have assigned the best direction to every edge. We therefore propose
-the following heuristic, inspired by the Lloyd algorithm for $k$-means, but where the directions in
-$\mathcal{D}_k$ play the role of centroids. Namely, we alternate between two steps
-\begin{enumerate}[(i), nosep]
-  \item create a new partition $\{E'_1, \ldots, E'_k\}$ by assigning every edge $(u,v)$ to the
-    direction $w$ maximizing its score ${s_{uv}}^T w$, that is $E'_\ell = \theset{(u,v) \in E}{
-    w_\ell = \argmax_{w_i \in \mathcal{D}_k} {s_{uv}}^T w_i}$, and
-  \item update the directions to make them optimal with respect to the edges currently assigned to
-    them, that is as previously, set $\mathcal{D}_k = \left\{ w_\ell = \frac{s_\ell}{\left\| s_\ell
-    \right\|}\right\}_{\ell=1}^k $, where $s_\ell = \sum_{u,v \in E'_\ell} s_{uv}$ is the sum of all
-    the similarity vectors of the edge in $E'_\ell$,
-\end{enumerate}
-until the edge assignment is stable or we reach a maximum number of iterations. By analogy with the
-$k$-means algorithm, obtaining convergence guarantees in the general case would be surprising, but in
-practice we expect that only a small number of iterations would be needed.

### 2017-10-17 edge/conclusions.tex
edge: future directions

@@
-The framework we proposed is agnostic to the question of the interplay between connection and
-linking attributes, although this is an interesting and active research question.
+\section{Conclusions}
+\label{sec:edge_conc}
 
-Homophily and Contagion Are Generically Confounded in Observational Social Network Studies
-Cosma Rohilla Shalizi, Andrew C. Thomas
-Sociological Methods \& Research
-Vol 40, Issue 2, pp. 211-239
-DOI: 10.1177/0049124111404820
+We could extend our approach to more general types of graphs and tasks:
 
-Digital Footprints: Opportunities and Challenges for Online Social Research
-Annual Review of Sociology
-Vol. 40:129-152 
-First published online as a Review in Advance on June 16, 2014
-\url{https://doi.org/10.1146/annurev-soc-071913-043145}
+\begin{itemize}[leftmargin=*]
 
-Here is what the “Digital Footprints” writes about the previous paper
-An important limitation is the difficulty distinguishing between homophily and contagion. Homophily
-refers to a variety of selection mechanisms by which a social tie is more likely between individuals
-with similar attributes and environmental exposures (McPherson et al.  2001). Contagion refers to
-influence mechanisms (e.g., imitation or peer pressure) by which traits diffuse along network edges.
-Homophily and contagion offer competing explanations for network autocorrelation, which refers to
-the greater similarity in the attributes of closely connected nodes. Based on simulated networks,
-Shalizi \& Thomas (2011, p. 216) conclude that “there is just no way to separate selection from
-influence observationally” (see also Manski 1993). This does not mean that observational studies
-using online networks are useless, but researchers need to refrain from assuming that the observed
-network autocorrelation reflects contagion effects and to acknowledge that the similarity between
-adjacent nodes may reflect the mutually reinforcing effects of influence and selection whose
-separate contributions may be impossible to tease apart.  For example, although Ugander et al.
-(2012) controlled for demographic similarity (sex, age, and nationality), there are countless other
-ways in which shared environments, affiliations, interests, and personality traits might cause two
-friends to join Facebook independently but not on the same day, making it look like the “early
-adopter” influenced the friend they invited who would have joined anyway.
+  \item First, as we mentioned earlier, \textbf{networks are dynamic} and evolve over time, with
+    nodes and edges constantly appearing and disappearing~\autocite{networkEvolution14}. Slicing
+    time into successive periods is one way to create edge types, one per period. For instance,
+    \textcite{TimedUSSenate10} study the \np{110} congresses of the US senate between the years 1789
+    and 2008, where \np{1884} individual Senators are linked by their voting similarity, and show
+    how it improves the detection of relevant communities. Beyond predicting the relative order of
+    links, it would also be interesting to include time into nodes similarity itself.
+    %
+    For instance, while the photo similarity between two Flickr users rises before they connect, it
+    later decreases if they have the same level of popularity, as both try to differentiate
+    themselves~\autocite{dynamicFlickr13}.
+    %
+    More generally, by defining our problem on a static snapshot of a graph, our model is agnostic
+    to the question of whether edges were created because of the similarity/dissimilarity of
+    the linked nodes' attributes or because attributes started to change after the edge creation.
+    This is traditionally framed as dichotomy between homophily and contagion, and is an active topic
+    of research in social networks analysis~\autocites{InfluenceVsHomo08}{ConfoundedHomophily11} as
+    noted by \textcite{OSNreview14}: \textquote{Homophily refers to a variety of selection
+    mechanisms by which a social tie is more likely between individuals with similar attributes and
+    environmental exposures~\autocite{Homophily01}. Contagion refers to influence mechanisms (\eg{}
+    imitation or peer pressure) by which traits diffuse along network edges. Homophily and contagion
+    offer competing explanations for network autocorrelation, which refers to the greater similarity
+    in the attributes of closely connected nodes.}
 
-Furthermore, we could also consider the role of time with dynamic networks. For instance, while the
-photo similarity between two Flickr users rises before they connect, it later decreases if they have
-the same level of popularity, as they try to differentiate themselves
-Zeng and Wei: Social Ties and User Content Generation: Evidence from Flickr, Information Systems
-Research 24(1), pp. 71–87
+  \item Another rich class of graphs were predicting edge types would be a challenging and rewarding
+    task is \textbf{multigraph}, which allow several edges between two nodes. It can also be seen as
+    the flattening of several graphs sharing the same nodeset, but where the provenance of each edge
+    would have been lost and need to be recover. With our original formulation of goodness as
+    ${s_{uv}}^T w_{uv}$, only one direction can achieve the higher score (discarding the rare case
+    of tie). One alternative we proposed was to let each $w_{uv}$ be a linear combination of a small
+    number of base directions, although experimentally it did not fit well our synthetic data.
+    Symmetrically, another idea could be to have more than one way to compute the similarity
+    $s_{uv}$ between two nodes.
+
+  \item Generalizing the balance theory from signed graphs, we could imagine that in triangles or
+    short cycles, only certain combinations of directions are valid (or at least desirable).
+    Interestingly, this would likely require different optimization algorithms, probably non convex.
+    However, while this is a general concept, its exact implementation might be
+    application-dependant. Furthermore, in our current setting, we do not know in advance in the
+    semantic of each direction, and we have no labeled edges. Therefore, it is not clear how those
+    constraints would be specified or whether they can be learned, if applicable. A limited solution
+    exists in the context of heterogeneous information networks, where in addition to having several
+    types of edges, there are also several types of nodes. Since not all type of node can be
+    connected with any type of edges, the possible connections are themselves represented as a meta
+    graph called the network schema~\autocite[Definition 3]{HINSurvey17}. This approach would need
+    to be extended to deal with higher order patterns.
+
+  \item Among tasks that could benefit from labeling the edge of complex networks, we mention two.
+    The first is \textbf{graph summarization} where the idea is to transmit a fraction of a large
+    graph plus additional information that can be used to reconstruct the original graph. Namely,
+    our goal here is to send only a few node profiles along with the edge directions, enabling the
+    missing profiles to be reconstructed from the known goodness values. Note that we spare
+    information regarding the profiles but still transmit the full topology. Therefore, this differs
+    from lossless schemes~\autocites{compressSN09}{compressSN12} that leverage typical edge locality
+    patterns to encode edges using as few bits as possible, or from lossy graph sketching
+    methods~\autocite{graphSketch12} discussed in \autoref{sub:gtx_state_of_the_art}, which preserve
+    spectral and topological properties of the original graph.
+    %
+    The second is \textbf{node embedding}, where every node $u$ is represented by a vector of size
+    $k$ containing the relative proportion of each direction with the edges incident to $u$.
+\end{itemize}

### 2017-10-18 edge/conclusions.tex
add one ref

@@
   \item Another rich class of graphs were predicting edge types would be a challenging and rewarding
-    task is \textbf{multigraph}, which allow several edges between two nodes. It can also be seen as
+    task is \textbf{multigraph}, which allow several edges between two
+    nodes~\autocite{typedMultigraph11}. It can also be seen as
     the flattening of several graphs sharing the same nodeset, but where the provenance of each edge

### 2017-10-18 edge/intro.tex
first version of edge intro

@@
-\paragraph{Generalization of signed graphs}
-\label{par:generalization_of_signed_graphs}
+In this chapter, we extend the problem of predicting edge types beyond the binary case.
+Furthermore, we assume that we are provided with node attributes in addition to the mere graph
+topology we used in the previous chapters. Our motivation is twofold: intrinsic and extrinsic.
+First, we want an interpretable model that, given two connected nodes, can explain why they are
+linked. This will lead to the use of linear models, for they are both easily amenable to
+interpretation and computationally efficient. Second, edge-labeled graphs are susceptible to further
+studies by existing networks analysis methods, as we will describe shortly. Thus, our method is a
+useful way of enriching data to extract new insights from it.
 
-Here we show why this formulation can model signed graphs as a special case. Remember the learning
-bias we introduced in the previous chapter \vpageref{text:cc_new_bias}: every node belongs to
-exactly one of the $K$ clusters available, is connected positively with the other nodes belonging to
-the same cluster, and is connected negatively with the nodes belonging to every other clusters. Now
-let the dimension of profiles be $d=K$ and the number of directions $k=2$. Moreover, we define the
-two directions vectors as $w_1 = \frac{\onev}{\sqrt{d}} = w^+$ and $w_2 = -w_+ = w_-$, where
-$\onev$ is the vector of dimension $d$ with only $1$ coordinates. Finally, for a node $u$ in
-cluster $i$, we assume its profile $x_u$ is set as follow: all component are equal to $-b$ except
-for the \ith{} one, which is equal to $a$, where $a$ and $b$ are constants we define now.
+As demonstrated by our prior extensive coverage, predicting edge signs is an interesting and
+challenging problem in its own right. Yet it is natural to extend and generalize it.  First, there
+are other edge binary classification problems that arise in graphs. For instance, predicting whether
+a link in a social network is strong or weak~\autocites{communityWeakTies14}{communityWeakTies17};
+predicting whether two authors in a citation network are advisor-advisee or regular
+coauthors~\autocite{Advisor10}; predicting whether two persons in a company network have a
+relationship of manager-subordinate or not~\autocite{Manager07}; predicting whether two proteins
+interact physically or genetically; or whether two neurons have a chemical or ionic link between
+them~\autocite{BioMultiplex15}.
 
-First, because $x_u$ is a unit vector, we have that $a^2 + (K-1)b^2=1$, from which $a =
-\sqrt{1-(K-1)b^2}$ whenever $1-(K-1)b^2 \geq 0$, or equivalently $b \leq \frac{1}{\sqrt{K-1}}$.
-Next, we look at $\onev^T\left(x_u \circ x_v\right) = \onev^T m_{uv} = \frac{1}{\sqrt{d}}
-\sum_{i=1}^d m_{{uv}_i}$ for any edge $(u,v) \in E$. If $u$ and $v$ are in the same cluster, then
-$\sum_{i=1}^d m_{{uv}_i} = a^2 + (K-1)b^2 = 1$, implying that $w_+^T m_{uv} = \frac{1}{d} > w_-^T
-m_{uv} = -\frac{1}{d}$. Otherwise, $\sum_{i=1}^d m_{{uv}_i} = (K-2)b^2 - 2ab = (K-2)b^2 -
-2b\sqrt{1-(K-1)b^2} = f_K(b)$. By computing the derivative of $f_K$ with respect to $b$, we find
-that it reaches a minimum of $-\frac{1}{K-1}$ at $b = \frac{1}{\sqrt{K(K-1)}} \leq
-\frac{1}{\sqrt{K-1}}$. In this case, $w_+^T m_{uv} = -\frac{1}{(K-1) \sqrt{d}} < w_-^T m_{uv} =
-\frac{1}{(K-1) \sqrt{d}}$. To maximize $\sum_{(u,v)\in E} \max_{1 \leq \ell \leq d} w_\ell^T
-m_{uv}$, we therefore assign $w_+$ to edges whose both endpoints belong to the same cluster, and
-$w_-$ to edges joining two different clusters.
+Second, and more generally, we are also interested in multiclass problems, within the context of
+multilayer graphs~\autocites{Kivela2014}{multiSurvey14}. While \textcite{Kivela2014} define an
+elaborate and highly expressive notion of multilayer graph, in this chapter we shall restrict
+ourselves to the following\footnote{Which is roughly corresponding to \emph{single aspect}
+\emph{node-aligned} multilayer graphs of \autocite{Kivela2014}.}:
+\begin{definition}[$k$-multilayer graph]
+  A $k$-multilayer graph $G$ is a pair $(V, \{E_1, \ldots, E_k \})$, where $V$ is a set of $n$ nodes
+  and each $E_i \subset V\times V$, $i \in \rangesk$ is a set of edges. We call the subgraph $G_i =
+  (V, E_i)$ the \ith{} layer of $G$ and it will be convenient to think of $G$ as the superposition
+  of its $k$ layers, so that by a slight abuse of notation we can write $ G = \bigcup_{i=1}^k G_i $.
+\end{definition}%\vspace{-0.25\baselineskip}\noindent%
+We naturally identify each layer with a type of relation, so that our classification problem is to
+decide to which layer a given edge belongs. Furthermore, we make the simplifying assumption that two
+nodes can be connected in at most one layer (\ie $\forall u,v \in V,\, (u,v)\in E_i \Rightarrow
+\forall j \neq i,\, (u,v) \notin E_j$), meaning that the resulting superposition of all layers does
+not turn into a multigraph. As mentioned in the introduction, such multigraphs can be used to model
+different relation types in networks related to social
+interactions~\autocites{Szell2010}{RecoFlickrMulti11}, bibliographical
+citation~\autocites{communityDBLPbyConf05}{articlesMultiSim11}, economic
+exchange~\autocites{ports3kindofships10}{worldTradeNetwork10} and biology~\autocite{bioLayerExp11}.
+This additional information on the nature of the relationships allow traditional graph tasks to be
+performed more precisely and with higher granularity, such as node ranking~\autocite{HRRanking13};
+node clustering \autocites[Section 2]{surveyAttributedClustering15}[Section 4.5.1]{Kivela2014} and
+the study of information spreading \autocite{SpreadingMultiSurvey15}. However, to produce such finer
+results, the methods we just referenced requires labeled that in many cases are not available,
+making our work rather useful. In the following table we give some examples of networks whose edges
+are currently unlabeled and why it would be beneficial to have them.
 
-Moreover, given this assignment, we can show that $w_+$ and $w_-$ are the optimal directions, under
-the additional uniformity assumption that there are exactly $r$ edges within every cluster, and $s$
-edges between any two clusters. First, we define
-\begin{equation*}
-  m_{\mathrm{inner}} = \sum_{(u,v) \in E;\, \cluster(u)=\cluster(v)} m_{uv} \qquad \text{and} \qquad
-  m_{\mathrm{outer}} = \sum_{(u,v) \in E;\, \cluster(u)\neq \cluster(v)} m_{uv}
-\end{equation*}
-to be the sum of all edge vectors within and between clusters respectively. By our choice of $a$ and
-$b$, we have that $m_{\mathrm{inner}} = r\onev$ and $m_{\mathrm{outer}} = -\frac{s}{2}\onev$. Now
-recall that the solution of $\argmax_{w\in \dsphere} w^T c$ is $\frac{c}{||c||}$. Therefore, the
-best direction to assign to inner edges is indeed $w_+$, and so is $w_-$ to outer edges.
+\noindent\begin{tabulary}{\textwidth}{lLLLL}
+  \toprule
+  Source & Node & Edge & Node profiles & Possible edge meanings \\
+  \midrule
+  DBLP & author & citation & keyword count of their papers & \\
+  Github & developer & following & number of lines written in various programming languages & \\
+  Amazon & video game products & \enquote{frequently bought together} & price, genre, time of release, popularity and evaluation & \\
+  Wikipedia & article & internal link & category, length, edit history, bag of words & \\
+  Flickr & user & following & activity statistics, groups membership, count of photo tags & \\
+  Foursquare & venue & \enquote{frequently visited together} & location, popularity, category, the time distribution at which they are visited & \\
+  IMDB & movie & sharing actors & genre, release date and popularity & \\
+  Blablacar & driver or passenger & shared ride & experience, location, preferences, rating & \\
+  \bottomrule
+\end{tabulary}
 
-\begin{aside}
-For inner edge vectors, $m_{uv}$ in cluster $i$ has $a^2$ as its \ith{} component, and $b^2$
-elsewhere. There are $r$ of them in cluster $i$, so summing over all clusters we get
-$m_{\mathrm{inner}} = r\left(a^2 + (K-1)b^2\right)\onev = r\onev$.
-For outer edge vectors, there are a few more steps. Consider the $s$ edges between $C_1$ and $C_2$,
-whose $m_{uv}$ is
-$\begin{pmatrix}
-  -ab \\
-  -ab \\
-  b^2 \\
-  b^2 \\
-  \vdots \\  
-\end{pmatrix}$. Likewise, between $C_1$ and $C_3$ we have
-$m_{uv}=\begin{pmatrix}
-  -ab \\
-  b^2 \\
-  -ab \\
-  b^2 \\
-  \vdots \\  
-\end{pmatrix}$. We thus we see that the first component is always $-ab$, that there is another
-such term at the other cluster index, while the rest is $b^2$. Summing all the vectors incident to
-$C_1$, we get
-$s\begin{pmatrix}
-  -ab(K-1) \\
-  b^2(K-2) -ab \\
-  b^2(K-2) -ab \\
-  \vdots \\  
-\end{pmatrix}$. Similarly, the sum of all vectors incident to $C_2$ is
-$s\begin{pmatrix}
-  b^2(K-2) -ab \\
-  -ab(K-1) \\
-  b^2(K-2) -ab \\
-  \vdots \\  
-\end{pmatrix}$, and so on for every cluster $i$. Summing over all clusters and dividing by $2$ to
-avoid double counting, we get $m_{\mathrm{outer}} = \frac{s}{2}\left( (K-1)\left(b^2(K-2) -ab\right)
--ab(K-1) \right)\onev = \frac{s(K-1)}{2}\left(b^2(K-2) -2ab\right)\onev =
-\frac{s(K-1)}{2}f_K(b)\onev = -\frac{s(K-1)}{2} \frac{1}{K-1} \onev = -\frac{s}{2}\onev$.
-\end{aside}
+Besides predicting potentially more than two classes, the other difference with the setting of the
+previous chapters is that we now assume that nodes are associated with attributes, which we call
+\emph{profiles}. One could argue this makes our framework less general, for attributes are not
+always available, or may come with missing values. Another recent finding is that, despite a natural
+assumption, the nodes profiles in attributed graphs are not necessarily correlated with known ground
+truth communities~\autocite{FortunatoGroundTruth14}, casting further doubt on the usefulness of that
+extra information.
+This raises the question of why we cannot simply extend the troll/trust method of counting the type
+of edges incident to every node and learn a model from such features.
+A similar tension between relying solely on always available yet sparse topological information or
+leveraging possibly richer attributes information also exists in link
+prediction~\autocite{linkPredSurvey16}\stodo{What they actually write is \enquote{Topology-based
+techniques are more versatile than attribute-based methods since they are not domain specific} which
+sounds a bit unfair: only the attributes are domain specific, not necessarily the methods
+themselves}.
+% examples of using node attributes for link prediction, taken from that survey:
+% https://doi.org/10.1007/s11432-014-5237-y
+% - http://dl.acm.org/citation.cfm?id=2124378
+% - using graph kernel https://doi.org/10.1016/j.dss.2012.09.019
+% - matrix factorization “these latent features (learned from topology) may be combined with
+%   optional explicit features for nodes or edges, which yields better performance than using either
+%   type of feature exclusively.” https://doi.org/10.1007/978-3-642-23783-6_28
+However, the presence of profiles is motivated by the last fundamental difference with the previous
+problems. In this chapter, we do not have access to any direct supervision in the form of labels.
+Rather than a strict classification problem, we are thus faced with unsupervised clustering, guided
+by the information contained in node profiles. Note that contrary to the relation between \pcc{} and
+\esp{}, this time the objects we seek to cluster are edges and not nodes. Furthermore, we stress
+that the profiles crucially increases the interpretability of our models.
 
-Granted, the situation where all the nodes of a cluster have exactly the same profile and where the
-connections are so regular is highly idealized. Moreover, we did not formally prove that there is
-not another assignment (with different direction vectors) that would give a better objective
-value\footnote{Although given the symmetry of the problem, that would be surprising.}. Our point is
-rather that profiles and directions can be seen as a \enquote{latent} explanation for the signs in
-a balanced signed graphs, like our generative model of the first chapter.
+A common learning bias in attributed graphs\stodo{better transition from attributes to homophily} is
+that nodes are connected together because they are similar, or as often said proverbially:
+\enquote{birds of a feather flock together}. Formally this is called the homophily
+principle~\autocite{Homophily01} and has been consistently verified, both in
+offline~\autocites{homoAttitude78}{homoEdu85} and online social
+networks~\autocites{homoSN09}{homoSN12}. While our every day life experience makes homophily not
+surprising in social networks, the same assortative patterns~\autocite{AssortativeMixing02} are
+present in other kinds of networks~\autocite{AssortativeMixing09}. Here though, we assume that links
+cannot only be explained by a global homophily along all attributes but rather by partial homophily,
+partial heterophily or even both. An example of this combination is balanced news
+diet~\autocite{balancedNews17}, where we try to connect users of a social network to prevent them
+from being locked in echo chambers. In that situation, a user $A$ might be more likely to listen to
+the opposite point of view of user $B$ about gun control if they both share common demographics and
+interests.
+
+After having motivated our problem of predicting multiple edge types in attributed graphs, we carry
+on by stating more formally the \ecp{} problem in \autoref{sec:edge_problem}. Namely, we define a
+similarity between nodes profiles and seek a small number of explanations (corresponding to edge
+types) achieving good scores as prescribed by a linear model. After explaining how this problem can
+model a situation closely resembling the case of signed graphs
+\vpageref{par:generalization_of_signed_graphs}, we introduce in \autoref{sec:edge_methods} several
+methods to solve it in practice. This ranges from tailoring the \kmeans{} algorithm to our goodness
+measure (\autoref{sub:edge_baseline}), to a convex relaxation of our objective function
+(\autoref{sub:edge_vector}) all the way to a richer and higher capacity matrix formulation allowing
+overlapping clustering of edges (\autoref{sub:edge_matrix}). Because finding publicly available real
+data with both node attributes and edge types proved difficult, we describe a synthetic model in
+\autoref{ssec:edge_synth} on which we perform extensive experiments. We then survey in
+\autoref{sec:edge_related} works that are related to the \ecp{}, while noting that very few address
+it directly. Finally, we suggest in \autoref{sec:edge_conc} several directions in which our model
+could be extended to handle more general classes of graphs.

### 2017-10-23 intro/draft.tex
tiny fixes and ref all over the places

@@
 defend the following statement: \textbf{There exists efficient and accurate methods to predict edge
-type in complex networks, reyling either only on the graph topology or also on node metadata.}
+type in complex networks, relying either only on the graph topology or also on node metadata.}
 
@@
 show that in addition to their sound theoretical foundations, our methods are competitive with the
-state of the art in predictive performance, while being more scalable. To illustrate this point, a
+state of the art in predictive performance, while being more scalable.
+\iffalse
+To illustrate this point, a
 short extract of our results is presented in \autoref{tab:troll}.
@@
 \end{table}
+\fi
 
@@
 this structure \gtx{}, as it recursively build stars of stars. 
+\iffalse
 We present preliminary results showing that on certain
@@
 this edge's endpoints).
+\iffalse
 From this initial formulation, we derive two optimization problems, and initially solve them on

### 2017-10-23 packages.tex
add new dataset, competitor and results
mark content added from the aistats paper
minor corrections in troll experiments and add missing figures
write about a possible connection between signed graphs and vector graphs
add some refs
tiny fixes and ref all over the places

@@
 \addbibresource{gtx.bib}
+\addbibresource{rapport.bib}
+\addbibresource{/home/orphee/data/projects/magnet/work/edge-vector-label/gsc.bib}
 \usepackage{xcolor}
@@
 \usepackage{amsthm}
+\usepackage{amssymb}
+% \usepackage{bbm}
+% \usepackage{bbold}
+\usepackage{bm}
+% \usepackage{dsfont}
 \usepackage{import}
@@
 \let\bottomfootnote\footnote
-\renewcommand{\footnote}[1]{\sidenote{\footnotesize #1}}
+% \renewcommand{\footnote}[1]{\sidenote{\footnotesize #1}}
 \usepackage[shortlabels]{enumitem}
 \usepackage{tabulary}
-\usepackage{bm}
 \usepackage{multirow}
@@
 \usepackage{mdframed}
+\usepackage[breakable]{tcolorbox}
 \newenvironment{aside}%
@@
 {\end{mdframed}}
+% \newenvironment{newcontent}%
+% {\begin{mdframed}[hidealllines=true, leftline=true, rightline=true, innertopmargin=0.3ex,
+%       innerbottommargin=0.3ex,
+%       innermargin=-.75em, leftmargin=-.75em,
+%       outermargin=-.75em, rightmargin=-.75em,
+%       innerleftmargin=0.75em, innerrightmargin=0.75em,
+%       linewidth=1.3pt, skipabove=0,
+%       skipbelow=0, linecolor=Black!20!White]}%
+% {\end{mdframed}}
 \usepackage{stmaryrd}
@@
 \usepackage{caption}
-\usepackage[margin=0pt,font+=small,labelformat=parens,labelsep=space,
+\usepackage[margin=0pt,labelformat=parens,labelsep=space,%font+=small,
 skip=6pt,list=false,hypcap=false]{subcaption}
 \captionsetup{compatibility=false}
+\captionsetup[subfigure]{labelformat=parens,labelsep=space}
 \usepackage{graphicx}
 \usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{makecell}
+\renewcommand\theadfont{\small}
 \graphicspath{{./assets/}}
@@
 \usepackage[]{extdash}
+\usepackage[]{calc}
+\newlength{\fullpage}

### 2017-10-24 assets/tikz/edge_example_tikz.tex
take Marc feedback on the edge model into account
- split into more subsection
- add an example
- describe and name the methods earlier
- add a paragraph comparing all the methods
- make it more clear there will be only synthetic data

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
  \begin{tikzpicture}[auto,vertex/.append style={inner sep=1pt},
    ]
    \tikzstyle{w1}=[red]
    \tikzstyle{w2}=[Lime]
    \tikzstyle{young}=[minimum size=6mm]
    \tikzstyle{old}=[minimum size=12mm]
    \tikzstyle{opneg}=[shape=rectangle ]
    \tikzstyle{opneu}=[shape=circle]
    \tikzstyle{oppos}=[shape=diamond]
    \tikzstyle{compG}=[fill=Orange]
    \tikzstyle{compF}=[fill=DodgerBlue!50!White]
    \tikzstyle{compboth}=[left color=Orange, right color=DodgerBlue!50!White,fading speed=100]
    \node[vertex,label={above:$u$},young, opneu,compG] (u) at (0, 0) {\male};
    \node[vertex,label={above:$v$},young, oppos,compboth] (v) at (2, 0) {\male};
    \node[vertex,label={below:$w$},old, oppos,compG] (w) at (0, -2) {\female};
    \node[vertex,label={below:$x$},old, opneg,compboth] (x) at (2, -2) {\female};
    \node[vertex,label={above:$y$},old, opneg,compF] (y) at (4, 0) {\male};
    \node[vertex,label={below:$z$},young, oppos,compF] (z) at (4, -2) {\female};

    \draw[edge,w1] (u) edge node[above] {$.083$} (v);
    \draw[edge,w1] (u) edge node[right] {$.104$} (w);
    \draw[edge,w2] (v) edge node[left] {$.077$} (x);
    \draw[edge,w2] (v) edge node[above] {$.094$} (y);
    \draw[edge,w1] (w) edge node[above] {$.061$} (x);
    \draw[edge,w1] (x) edge node[left] {$.027$} (y);
    \draw[edge,w2] (x) edge node[above] {$.092$} (z);
    \draw[edge,w2] (y) edge node[left] {$.126$} (z);
  \end{tikzpicture}
\end{document}

### 2017-10-24 edge/experiments.tex
take Marc feedback on the edge model into account
- split into more subsection
- add an example
- describe and name the methods earlier
- add a paragraph comparing all the methods
- make it more clear there will be only synthetic data

@@
-\section{Experimental analysis}
+\section{Synthetic experiments}
 \label{sec:edge_exp}
 
-We start by extracting from the conceptual discussion of the previous section the five methods we
-compare in our experiments. They all take as input an attributed graph, \ie{} $G=(V, E)$ and $X$.
-For evaluation purposes, we describe how to extract from their output a set of $k$ directions and an
-assignment $\yhat$ of every edge to one of these directions (that is $\yhat$ is a function from $E$
-to $\rangesk$ such that $\yhat_{uv}$ is the index of the direction assigned to the edge $(u,v)$).
+In the previous section, we introduced five methods to address \autoref{p:edge_full}, which we now
+compare in our experiments. Specifically, this includes:
+\begin{enumerate}[1.,nosep]
+\item the \kmeans{} baseline,
+\item our \lloyd{}-like iterative refinement,
+\item the \combined{} convex optimization of \eqref{eq:edge_soft},
+\item the \fwa{} convex optimization of \eqref{eq:edge_FW} and
+\item the \pqt{} explicit low-rank factorization, optimizing \eqref{eq:edge_PQ}.
+\end{enumerate}
+They all take as input a node-attributed graph, that is a pair $G=(V, E)$ and $X$.
+For evaluation purposes, we extract from their output a set of $k$ directions and an
+assignment of every edge to one of these directions.
+We try those five methods on data generated synthetically to fit our model described previously in
+\autoref{sub:edge_setting}, and study how they perform under various conditions.
 
-\begin{enumerate}[1.]
+% (that is $\yhat$ is a function from $E$
+% to $\rangesk$ such that $\yhat_{uv}$ is the index of the direction assigned to the edge $(u,v)$).
 
-  \item Our baseline is a \kmeans{} clustering of the normalized $s_{uv}$ edge similarity
-    vectors. This naturally provides a \yhat{} assignment, and the directions are the normalized
-    cluster centers.
-
-  \item Our first method is a \lloyd{}-like iterative refinement of the $k$-means solution, where we
-    alternatively reassign edges based on their score and recompute the directions to maximize those
-    scores.
-
-  \item Then we consider the \textsc{Combined} optimization of the two edge and node terms of
-    \eqref{eq:edge_full}, which directly provides a set of directions. The assignment is then simply
-    the direction that gives the highest goodness.
-
-  \item As for matrix methods, we first do a \textsc{Frank--Wolfe} maximization of
-    \eqref{eq:edge_FW} over the $\delta$-ball of the nuclear norm. Afterwards, we perform a
-    \kmeans{} clustering of its columns and use it as assignment, while the directions are again the
-    normalized cluster centers.
-
-  \item \textsc{Explicit} low-rank $PQ^T$ factorization, initialized with the solution of
-    \combined{} (that is we set $P_0 \in \mathbb{M}^{d\times k}$ to be the final directions found by
-    \combined{} and the rows of $Q_0 \in \mathbb{M}^{|E|\times k}$ to be equal to the corresponding
-    $e_{y_{uv}}$) and minimizing the same objective of \eqref{eq:edge_PQ}. We take the directions to
-    be the columns of the resulting $P$, and the assignment to be a \kmeans{} clustering of the rows
-    of $Q$.
-\end{enumerate}
-
-\subsection{Synthetic data}
-\label{ssec:edge_synth}
-
-We try those five methods on data generated synthetically to fit our model describe previously and
-study how they perform under various conditions.
-
-\paragraph{Generation}
-\label{par:edge_generation}
+\subsection{Generation}
+\label{sub:edge_generation}
 
@@
 instance, with $d=6$, $k=3$ and $n_o=2$, we could generate the following directions, where the dots
-represents non-zero numbers and the red ones are overlapping:
+represents non-zero numbers and the blue ones are overlapping:
 \vspace{-.5\baselineskip}
@@
       0 \\
-      \textcolor{Red}{\bullet}
+      \textcolor{DodgerBlue}{\bullet}
     \end{pmatrix}$ &
@@
       0 \\
-      \textcolor{Red}{\bullet} \\
+      \textcolor{DodgerBlue}{\bullet} \\
       \bullet \\
@@
 
-\paragraph{Results}
-\label{par:edge_synth_results}
+\subsection{Results}
+\label{sub:edge_synth_results}
 

### 2017-10-24 edge/model.tex
take Marc feedback on the edge model into account
- split into more subsection
- add an example
- describe and name the methods earlier
- add a paragraph comparing all the methods
- make it more clear there will be only synthetic data

@@
-\section{Learning problem}
+\section{Attributed graphs and problem definition}
 \label{sec:edge_problem}
+
+We start by introducing some notations and terminology about edge types in node-attributed graphs.
+Then we present two options to model the interactions between nodes and the role of attributes in
+\enquote{explaining} edges (\autoref{sub:edge_setting}). Choosing of these two options leads us to
+formulate the \ecp{} problem in \autoref{sub:edge_constraints}. We then describe additional
+constraints to take into account the topology of the graph. Finally we elaborate on the relation
+between \ecp{} and \esp{} in signed graphs.
+
+\subsection{Setting and modelling}
+\label{sub:edge_setting}
 
@@
 along which the combination $s_{uv}$ between two nodes $u$ and $v$ is evaluated. We call such
-$w_{uv}$ \emph{directions}, and constrain them to have unit norm\stodo{or bounded?}. As we will discuss, this
-constraint can be seen as a way to couple the different dimensions of the profiles%
-\stodo{don't remember what I meant here}.
+$w_{uv}$ \emph{directions}, and constrain them to have unit norm.
+While being exactly of norm one is not primordial, the role of bounding the norm of the directions
+(\ie{} having $\sum_{i=1}^d {w_{uv}}_{;i}^2 \leq B^2$) is to tie the $d$ dimensions together.
+Without such a constraint, we could simply study each dimension separately, whereas this bound
+adds a coupling across different attributes, since increasing the weight of an attribute mechanically
+affects the others.
 
@@
 
-We assume in the following that both the profiles and the directions are vectors with components in
-$[-1, 1]$. A natural choice for the operator $s$ is then the difference between the profiles, that is
+As written in the introduction, because we are interested in (partial) heterophilic links, we assume
+in the following that both the profiles and the directions are vectors with components in $[-1, 1]$.
+Let us take the example of encoding the gender of a user in a social network, assuming there is only
+two values possible: male and female. Traditionally, this attribute would be set to $1$ if the user
+is female and $0$ otherwise. Instead, we use $+1$ and $-1$, to clearly emphasize the difference
+between the two values. The same is valid for continuous variables. For instance, an attribute could
+range from $-1$ to $+1$ to represent the ideological stance of a user on the conservative/liberal
+scale~\autocite{newsBias15}.\footnote{Note of course that in both cases, negative values do not
+represent a judgement but are merely a reversible mathematical convenience.} This applies even to
+non diverging positive quantities such as age, which can be projected to $[-1, 1]$ using
+pre-processing such as standardization or $z$-score.
+
+A natural choice for the operator $s$ is then the difference between the profiles, that is
 $s_{uv} = x_u - x_v$. In that spirit, if $u$ and $v$ are close along a direction $w_{uv}$, we expect
@@
 w_{uv}) = \left({s_{uv}}^T w_{uv} \right)^2$, and we want to minimize it with respect to $w_{uv}$.
-When on the \ith{} dimension ${x_u}_{;i} - {x_v}_{;i}$ is close to zero, the minimization objective
+When, on the \ith{} dimension, ${x_u}_{;i} - {x_v}_{;i}$ is close to zero, the minimization objective
 and the fact that $w_{uv}$ is
@@
 ${x_u}_{;i}$ and ${x_v}_{;i}$ are widely different.
-Whereas there might be situations where this is not a problem, as we discussed in the introduction,
-here we seek to model heterophilic interactions, at least along a subset of dimensions. Therefore, in
-the following, we will adopt a refined approach.
+There might be situations where this is not a problem. But again,
+here we seek to model heterophilic interactions, at least along a subset of dimensions. Therefore,
+we now introduce a more flexible approach.
 
@@
 
-Based on these specific definitions of $s$ and $g$, we now formally write down the problem of
+This second choice of $s$ and $g$ is indeed more flexible, for it encompasses all cases of
+homophily, indifference and heterophily. Furthermore, since the profiles are given, the resulting
+$s_{uv}$ can be seen as constants, meaning that $g$ is linear and therefore interpretable. Finally,
+it also supports categorical data. We shall demonstrate all these features on a small scale example,
+but we first define formally our problem.
+
+\subsection{Learning problem and additional constraints}
+\label{sub:edge_constraints}
+
+\paragraph{Problem definition}
+
+With working definitions of $s$ and $g$ at hand, we now formally write down the problem of
 finding and assigning the most informative directions to every edge of $G$
-\begin{problem}[]
+\begin{problem}[\ecp{}]
   \label{p:edge_full}
@@
     \argmax_{w_1,\ldots,w_k \in \dsphere}
-    \sum_{u,v \in E} \max_{\ell \in \rangesk} g(s_{uv}, w_{\ell}) =
-    \sum_{u,v \in E} \max_{\ell \in \rangesk} \left(x_u \circ x_v\right)^T w_{\ell} = f(k)
+    \sum_{u,v \in E} \max_{\ell \in \rangesk} g(s_{uv}, w_{\ell})
   \end{equation}
@@
 \end{problem}
+
+\paragraph{Example}
+
+In \autoref{fig:edge_exe}, we present a small instance of the \ecp{} problem, where we are given the
+graph shown in \autoref{fig:edge_exe_graph}, whose node attributes are listed
+\autoref{tab:edge_exe}. The first two attributes are categorical, denoting the company employing
+each user, either \textsf{G} or \textsf{F}. Instead of a classic binary one-hot encoding, we
+additionally use $-1$ to denote that a user had been employed by a given company. \textsf{Gender} is
+encoded as presented earlier, and we also apply a pre processing to transform \textsf{age} between
+$-1$ and $1$. Finally, \textsf{opinion} represents the view of former \textsf{F} employees about a
+societal issue, with $0$ denoting a neutral position. In \autoref{tab:edge_exe}, we also show two
+directions that explain the connections in the graph $G$. The first covers being of the same age and
+same companies, especially at \textsf{G}, while the second directions covers former employees of
+\textsf{F} having different \textsf{age} and \textsf{opinion}. Not only are the results
+interpretable as claimed before, but we can also assess the strength of each explanation, using
+their goodness. For instance, we see that the edge between $x$ and $y$ has a low score, because
+those two users have indeed the same \textsf{age} and \textsf{opinion}.
+
+\begin{figure*}[htpb]
+  % \centering
+  \begin{subfigure}[t]{0.55\textwidth}
+    \centering
+    \includegraphics[width=.9\textwidth]{tikz/edge_example_tikz.pdf}
+    \caption{The graph $G$. Larger nodes represent old users while smaller ones are younger. Their
+    gender is inscribed in the node and their shape represent their opinion: rectangle for $-1$,
+    circle for $0$ and diamond for $+1$. Finally, orange stands for being employed at company G and
+    faded blue for having been employed at company F. As for the edges, the red ones are labeled by
+    the first direction and the green ones by the second direction, and we label them with the
+    associated (normalized) goodness. } \label{fig:edge_exe_graph}
+  \end{subfigure}~
+  \begin{subfigure}[t]{0.42\textwidth}
+    \centering
+    \small
+    \begin{tabular}{lrrrrr}
+      \toprule
+      {}    & G   & F    & gender & age    & opinion \\
+      \midrule
+      $u$   &  $+1$ &   $0$ &   $-1$ &  $-0.8$ &     $0$ \\
+      $v$   &  $+1$ &  $-1$ &   $-1$ &  $-0.7$ &    $+1$ \\
+      $w$   &  $+1$ &   $0$ &   $+1$ &   $0.6$ &    $+1$ \\
+      $x$   &  $+1$ &  $-1$ &   $+1$ &   $0.9$ &    $-1$ \\
+      $y$   &   $0$ &  $-1$ &   $-1$ &   $0.7$ &    $-1$ \\
+      $z$   &   $0$ &  $-1$ &   $+1$ &  $-0.6$ &    $+1$ \\
+      \midrule
+      $w_1$ &  $+3$ &  $+1$ &    $0$ &    $+1$ &     $0$ \\
+      $w_2$ &  $0$  &  $+4$ &    $0$ &    $-2$ &    $-4$ \\
+      \bottomrule
+    \end{tabular}
+    \caption{The numerical attributes of the six nodes. They are followed by the (non normalized)
+    two directions $w_1$ and $w_2$.} \label{tab:edge_exe}
+  \end{subfigure}
+  \caption{A small instance of \ecp{} and an handcrafted solution, albeit non-optimal.}
+  \label{fig:edge_exe}
+\end{figure*}
+
+\paragraph{Choice of $k$}
+\label{par:edge_choose_k}
+
+For clarity, the previous example of \autoref{fig:edge_exe} has only a handful of nodes and
+dimensions, and can therefore reasonably be explained by two directions. But in larger graphs, the
+choice of $k$ is a legitimate question.
+At one extreme, picking $k=|E|$ renders the topology of the graph
+irrelevant, as we can simply exploit the trivial
+solution of having a single direction $w_{uv} = \frac{s_{uv}}{||s_{uv}||}$ for every edge $(u,v)$.
+More generally, as $k$ increases, we expect the value $\max_{\mathcal{D}_k} \sum_{u,v \in E}
+\max_{\ell \in \rangesk} g(s_{uv}, w_{\ell})$ to increase as
+well, at the cost of interpretability. There could be a principled approach to finding the
+\enquote{best} $k$, based on a information theoretical measure of the complexity of $\mathcal{D}_k$
+and the minimum description length principle~\autocite{grunwald2005tutorial}. However, in the
+interest of simplicity, in the following we focus on the formulation stated in
+\autoref{p:edge_full}, where $k$ is given. In practice, we further constrain $k$ to be small (say
+less than \np{10}). Within the multilayer framework, $k$ could also be seen as the number of
+underlying layers. To retain a fair level of interpretability, we deem appropriate to have $k$ not
+too large.
+
+\paragraph{Topological constraints}
+
+In addition to assume that the input graph is the superposition of a small number of layers,
+another way to leverage the graph topology is to define local constraints at the node level.
+For that we will consider node profiles to be normalized and having unit norm. A practical
+justification is that if nodes are
+users of a social network and the profiles measure their activity across several domains, this
+allows to compare users with various level of total activity. Furthermore, we assume that profiles
+are linear combination of directions, that
+is $x_u = b_u + \sum_{v \in \nei(u)} a_{uv} w_{uv}$, where $a_{uv}$ are real weights and $b_u
+\in \Rbb^d$ is the bias vector of $u$, its inherent profile take in isolation.
+Viewed in the other direction, this can be interpreted as follow: each time a node $u$
+connect to another node $v$ along a direction $w_{uv}$, it \enquote{consumes} a part of its profile
+proportional to $w_{uv}$. Therefore, we subtract the term
+\begin{equation}
+  \label{eq:edge_node_loss}
+  \mathcal{L}_{\mathrm{node}} =
+  \sum_{u\in V} \left|\left| x_u - b_u - \sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2_2
+\end{equation}
+from \eqref{eq:edge_full}.
+
+\medskip
+
+We also make the assumption that each node $u$ is only involved in $k_\mathrm{local} < k$
+different directions. The intuition is that when $d$ is large enough, users have to focus their
+energy and can only express their interest in a few number of dimensions. In other words, all the
+edges incident to $u$ can only be associated with one of $k_\mathrm{local}$ directions. For
+some values of
+$k_\mathrm{local}$, this is a \NPc{} graph colouring problem and thus cannot necessarily be enforced
+exactly. Still, it can be seen as a learning bias and will be used to generate synthetic data.
+Furthermore, we can also express it as an optimization constraint. Specifically, suppose that for
+an edge $(u,v)$ associated with the direction $w_i \in \mathcal{D}_k$, we have a $k$-dimensional
+vector $y_{uv}$ equal to $e_i$ (that is all zero except for the \ith{} component set to one). Let
+us stack all such vectors into a $|E| \times k$ matrix $Y$, and let $C \in \Rbb^{|V| \times |E|}$ be
+the incidence matrix of $G$\footnote{The incidence matrix of a graph $G$, usually denoted by $B$, is
+the $|V| \times |E|$ matrix such that $B_{i,j} = 1$ if the node $v_i$ and edge $e_j$ are incident
+and 0 otherwise.}, with $c_u$ being its $u^{th}$ row. Then $c_u^TY$ is a $k$-dimensional vector that counts
+how many times each direction is incident to node $u$. We thus want the $\ell_0$ norm of $c_u^TY$
+(\ie the number of its non-zero component) to be upper bounded by $k_\mathrm{local}$.
+This is not a convex constraint but we can relax it, for such sparsity inducing problems have been
+well studied~\autocite{sparseOptim12}. Namely, we let $y_{uv}$ be a softmax membership
+$k$-dimensional vector, that is we set $\forall i \in \rangesk\, y_{uv,i} =
+\frac{\exp(\beta{s_{uv}}^T w_i)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_j)} $, with $\beta>0$ a large
+positive constant.  And we replace the $\ell_0$ norm by a $\ell_1$ norm (or even a $k$-support
+norm~\autocite{KsupportNorm12}), yielding the following $\mathcal{L}_{\mathrm{local}}$ to be
+minimized:
+\begin{equation}
+  \label{eq:edge_local_loss}
+  \mathcal{L}_{\mathrm{local}} =
+  \sum_{u \in V} \left\| \sum_{v \in \nei(u)} y_{uv} \right\|_1 =
+  \sum_{u \in V} \sum_{\ell = 1}^k \sum_{v \in \nei(u)}
+  \frac{\exp(\beta{s_{uv}}^T w_\ell)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_j)}
+\end{equation}
+
+These two constraints can be combined, so that the profiles are close to a linear combination of a
+small number of directions.
 
@@
 let the dimension of profiles be $d=K$ and the number of directions $k=2$. Moreover, we define the
-two direction vectors as $w_1 = \frac{\onev}{\sqrt{d}} = w^+$ and $w_2 = -w_+ = w_-$, where
+two direction vectors as $w_1 = \frac{\onev}{\sqrt{d}} = w_+$ and $w_2 = -w_+ = w_-$, where
 $\onev$ is the vector of dimension $d$ with only $1$ coordinates. Finally, for a node $u$ in
@@
 \frac{1}{\sqrt{K-1}}$. In this case, ${s_{uv}}^T w_+ = -\frac{1}{(K-1) \sqrt{d}} < {s_{uv}}^T w_- =
-\frac{1}{(K-1) \sqrt{d}}$. To maximize $\sum_{(u,v)\in E} \max_{1 \leq \ell \leq d} w_\ell^T
-s_{uv}$, we therefore assign $w_+$ to edges whose both endpoints belong to the same cluster, and
+\frac{1}{(K-1) \sqrt{d}}$. To maximize $\sum_{(u,v)\in E} \max_{\ell \in \{1, 2\}} {s_{uv}}^T
+w_\ell$, we therefore assign $w_+$ to edges whose both endpoints belong to the same cluster, and
 $w_-$ to edges joining two different clusters.
@@
 
-\bigskip
-
-At the other extreme of having only $k=2$ directions, when $k=|E|$, the topology of the graph
-becomes irrelevant, as we can simply exploit the trivial
-solution of having a single direction $w_{uv} = \frac{s_{uv}}{||s_{uv}||}$ for every edge $(u,v)$.
-More generally, as $k$ increases, we expect the value $\max_{\mathcal{D}_k} f(k)$ to increase as
-well, at the cost of interpretability. There could be a principled approach to finding the
-\enquote{best} $k$, based on a information theoretical measure of the complexity of $\mathcal{D}_k$
-and the minimum description length principle~\autocite{grunwald2005tutorial}. However, in the
-interest of simplicity, in the following we focus on the formulation stated in
-\autoref{p:edge_full}, where $k$ is given. In practice, we further constrain $k$ to be small (say
-less than \np{10}). Within the multilayer framework, $k$ could also be seen as the number of
-underlying layers. To retain a fair level of interpretability, we deem appropriate to have $k$ not
-too large.
-
-In addition to assume that the input graph is the superposition of a small number of layers,
-another way to leverage the graph topology is to define local constraints at the node level. 
-Recall that we consider node profiles to be normalized and having unit norm. A practical
-justification is that if nodes are
-users of a social network and the profiles measure their activity across several domains, this
-allows to compare users with various level of total activity. One way to integrate that fact in
-our optimization formulation is to assume that profiles are linear combination of directions, that
-is $x_u = b_u + \sum_{v \in \nei(u)} a_{uv} w_{uv}$, where $a_{uv}$ are real coefficients and $b_u
-\in \Rbb^d$. Viewed in the other direction, this can be interpreted as follow: each time a node $u$
-connect to another node $v$ along a direction $w_{uv}$, it consume a part of its profile
-proportional to $w_{uv}$. Therefore, we subtract the term 
-\begin{equation}
-  \label{eq:edge_node_loss}
-  \mathcal{L}_{\mathrm{node}} = 
-  \sum_{u\in V} \left|\left| x_u - b_u - \sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2
-\end{equation}
-from \eqref{eq:edge_full}.
-
-We can also make the assumption that each node $u$ is only involved in $k_\mathrm{local} < k$
-different directions. The intuition is that when $d$ is large enough, users have to focus their
-energy and can only express their interest in a few number of dimensions. In other words, all the
-edges incident to $u$ can only be associated with one of $k_\mathrm{local}$ directions. For
-certain value of
-$k_\mathrm{local}$, this is a \NPc{} graph colouring problem and thus cannot necessarily be enforced
-exactly. Still, it can be seen as a learning bias and will be used to generate synthetic data.
-Furthermore, we can also express it as an optimization constraint. Specifically, suppose that for
-an edge $(u,v)$ associated with the direction $w_i \in \mathcal{D}_k$, we have a $k$-dimensional
-vector $y_{u,v}$ equal to $e_i$ (that is all zero except for the \ith{} component set to one). Let
-us stack all such vectors into a $|E| \times k$ matrix $Y$, and let $C \in \Rbb^{|V| \times |E|}$ be
-the incidence matrix of $G$\footnote{The incidence matrix of a graph $G$, usually denoted by $B$, is
-the $|V| \times |E|$ matrix such that $B_{i,j} = 1$ if the node $v_i$ and edge $e_j$ are incident
-and 0 otherwise.} and $c_u$ be its $u^{th}$ row. Then $Yc_u$ is a $k$-dimensional vector that counts
-how many times each direction is incident to node $u$. We thus want the $\ell_0$ norm of all $Yc_u$
-(\ie its number of non-zero component) to be upper bounded by $k_\mathrm{local}$. As previously,
-this is not convex but we can relax this constraint, as such sparsity inducing problem have been
-well studied~\autocite{sparseOptim12}. Namely, we let $y_{uv}$ be a softmax membership vector, that
-is we set $y_{uv,i} = \frac{\exp(\beta{s_{uv}}^T w_i)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_i)}$,
-and we replace the $\ell_0$ norm by a $\ell_1$ or even a $k$-support norm~\autocite{KsupportNorm12}.
-These two constraints can be combined, so that the profiles are close to a linear combination of a
-small number of directions.
-
 \section{Proposed approaches}
@@
 set in \autoref{p:edge_full}. Formally, given the graph $G=(V,E)$ and the node profiles $X$ as
-input, we order the edges from $e_1$ to $e_m$ and build the matrix $S \in \Rbb^{m\times d}$ whose
-\ith{} row is the similarity between the profiles of the endpoints of $e_i=(u,v)$, that is $s_{uv} =
-x_u \circ x_v$. Running $k$-means will thus partition $E$ into $k$ sets of edges, call them $E_1,
-\ldots, E_k$. The first drawback of that simplicity is that it does not solve \autoref{p:edge_full}.
+input, we order the edges from $e_1$ to $e_m$. Then we build the matrix $S \in \Rbb^{m\times d}$, whose
+\ith{} row is the similarity $s_{uv} = x_u \circ x_v$ between the profiles of the endpoints of $e_i=(u,v)$.
+Running $k$-means on the rows of $S$ will thus partition $E$ into $k$ sets of edges, call them $E_1,
+\ldots, E_k$. A drawback of that simplicity is that it does not solve \autoref{p:edge_full}.
 Indeed, the partition of $E$ does not provide the set of directions $\mathcal{D}_k = \{w_1, w_2,
 \ldots, w_k\} \in \dsphere$ we are looking for. A natural way of obtaining directions from that
-partition is to set $w_\ell$ to be the normalized cluster center of $E_\ell$. Note however that
+partition is to set $w_\ell$ to be the normalized cluster center of $E_\ell$. We refer to this
+method as \kmeans{}, and assuming it requires $T$ iterations to converge, it has a linear complexity
+of $O(T|E|)$.
+\iffalse
+\begin{center}
+  \rule{\textwidth}{.3pt}
+  \begin{algorithmic}[1]
+    \Function{\kmeans{}}{graph $G=(V,E=\{e_1, \ldots, e_m \})$, profiles $X$}
+    \State build the matrix $S$ such that $S_{i,j} = {s_{e_i}}_{;j}$
+    \State let $\yhat{}$ and $\{c_1, \ldots, c_k\}$ be the assignment and centroids resulting of
+    running $k$-means on the rows of $S$
+    \State \textbf{return} \yhat{}, $\mathcal{D}_k = \left\{\frac{c_\ell}{\left\| c_\ell
+    \right\|}\right\}_{\ell=1}^k $
+    \EndFunction
+  \end{algorithmic}
+  \rule{\textwidth}{.3pt}
+\end{center}
+\fi
+
+However,
 this does not guarantee that we have assigned the best direction to every edge. We therefore propose
 the following heuristic, inspired by the Lloyd algorithm for $k$-means, but where the directions in
-$\mathcal{D}_k$ play the role of centroids. Namely, we alternate between two steps
+$\mathcal{D}_k$ play the role of centroids. Namely, we alternate between two steps:
 \begin{enumerate}[(i), nosep]
   \item create a new partition $\{E'_1, \ldots, E'_k\}$ by assigning every edge $(u,v)$ to the
-    direction $w$ maximizing its score ${s_{uv}}^T w$, that is $E'_\ell = \theset{(u,v) \in E}{
+    direction $w$ maximizing its goodness ${s_{uv}}^T w$, that is $E'_\ell = \theset{(u,v) \in E}{
     w_\ell = \argmax_{w_i \in \mathcal{D}_k} {s_{uv}}^T w_i}$, and
   \item update the directions to make them optimal with respect to the edges currently assigned to
-    them, that is as previously, set $\mathcal{D}_k = \left\{ w_\ell = \frac{s_\ell}{\left\| s_\ell
+    them. Specifically, set $\mathcal{D}_k = \left\{ w_\ell = \frac{s_\ell}{\left\| s_\ell
     \right\|}\right\}_{\ell=1}^k $, where $s_\ell = \sum_{u,v \in E'_\ell} s_{uv}$ is the sum of all
@@
 $k$-means algorithm, obtaining convergence guarantees in the general case would be surprising, but in
-practice we expect that only a small number of iterations would be needed.
+practice we expect that only a small number $T$ of iterations would be needed. We call this method
+\lloyd{} and note it also has a linear complexity of $O(T|E|)$.
 
-\subsection{Vector optimization}
+\subsection{Convex relaxation}
 \label{sub:edge_vector}
 
-The difficulty in optimizing directly \eqref{eq:edge_full} stem from the $\max$ operator, which we
+The difficulty in optimizing directly \eqref{eq:edge_full} stems from the inner $\max$ operator, which we
 replace here by a convex relaxation. Indeed, one can check that given a set of numbers $S=\{a_1,
@@
   \max_{a_i \in S} a_i = \lim_{\beta \rightarrow + \infty} \frac{1}{\beta}
-  \log\left( \sum_{i=1}^{|S|} \exp{\beta a_i} \right)
+  \log\left( \sum_{i=1}^{|S|} \exp{(\beta a_i)} \right)
 \end{equation*}
@@
 \begin{equation}
-  \label{eq:edge_soft}
+  \label{eq:edge_soft_early}
   \argmax_{{\substack{w_1,\ldots,w_k \in \dsphere,\\
@@
   \log\Bigl(\sum_{\ell=1}^k\exp\left(\beta {s_{uv}}^T w_\ell \right)\Bigr)
-  - \frac{1}{|V|}\sum_{u\in V}
-  \left|\left| x_u - b_u - \sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2
+  - \frac{\mu}{|V|}\sum_{u\in V}
+  \left|\left| x_u - b_u - \sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2,
 \end{equation}
-Note that in second term, because we do not know the edge assignment, we have to decide which of the
+where $\mu > 0$ is a trade-off parameter.
+Note that in the second term, because we do not know the edge assignment, we have to decide which of the
 $w_1, \ldots, w_k$ is $w_{uv}$ equal to. According to our definition of the goodness function, we
 want to let $w_{uv} = \argmax_{w_\ell \in \mathcal{D}_k} {s_{uv}}^T w_\ell$. However, the $\argmax$
-function is not convex and we thus use the same relaxation as before. Namely, by stacking all the
-vectors of $\mathcal{D}_k$ into a single vector $W$ of size $k\times d$, the relaxation is:
+function is again not convex and we thus use the same relaxation as before. Namely, this relaxation
+$\mathcal{A}$ takes as input a $k$-tuple of directions and return their sum weighted by the softmax
+function as follow:
 $$ \begin{array}[t]{lrcl}
-  \mathcal{A} : & \Rbb^m & \longrightarrow & \Rbb^d \\
+  \mathcal{A} : & \Rbb^d \times \ldots \times \Rbb^d & \longrightarrow & \Rbb^d \\
       & (w_1,\ldots,w_k) & \longmapsto &
-  \sum_{i=1}^k \frac{\exp(\beta{s_{uv}}^T w_i)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_i)} w_i
+  \sum_{i=1}^k \frac{\exp(\beta{s_{uv}}^T w_i)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_j)} w_i
 \end{array} $$
-This function has an explicit derivative, but it is rather costly to compute and in practice we
-found automatic differentiation~\autocite{autograd15} to be more efficient. While this modified
-objective function is now jointly convex in all its variables, the set $\dsphere$ is not. To ease
+This function has an explicit derivative with respect to any $w_i$, but it is rather costly to
+compute and in practice we
+found automatic differentiation~\autocite{autograd15} to be more efficient. Whereas the modified
+objective function \eqref{eq:edge_soft_early} is now jointly convex in all its variables, the set
+$\dsphere$ is not. To ease
 the optimization procedure, we thus only impose the directions $w_\ell$ to lie within the unit ball
 and expect them to have norm close to one in order to maximize the dot product in the ${s_{uv}}^T
-w_\ell$ term.
+w_\ell$ term. A further simplification is that while \eqref{eq:edge_soft_early} requires to optimize
+$kd + |E| + |V|d$ parameters, we fix $a_{uv} = 1$ and $b_u = \onev$ for all edges and nodes in order
+to be independent of the size of the graph. The final problem solved by our \combined{} method using
+(full) projected gradient descent is therefore:
+\begin{equation}
+  \label{eq:edge_soft}
+\!  \argmax_{w_1,\ldots,w_k \in \dball}
+  \frac{1}{|E|\beta} \sum_{u,v\in E}\!
+  \log\Bigl(\sum_{\ell=1}^k\exp\left(\beta {s_{uv}}^T w_\ell \right)\Bigr)
+  - \frac{\mu}{|V|}\sum_{u\in V}
+  \left|\left| x_u  - \sum_{v \in \nei(u)}\!\! \mathcal{A}(w_1,\ldots,w_k) \right|\right|^2
+\end{equation}
 
@@
 
-Another way to remove the $\max$ is to first seek one direction for every edge, that is to maximize
-$\sum_{u,v\in E} {s_{uv}}^T w_{uv}$. This objective can be written in matric form as $\tr(SW) =
+Besides relaxing the $\max$ operator from \eqref{eq:edge_full}, another solution to remove it would
+be to increase $k$ to be equal to $E$. As we already mentioned, this makes the problem trivial. To
+avoid this, we add the constraint that all those $|E|$ directions are linear combination of a small
+number $k$ of basis directions. In order to formulate this more clearly, we rewrite the previous
+objectives with matrices. We first show this can express in a form that is amenable to the
+Frank--Wolfe algorithm. Then we take the alternative road of explicit low-rank matrix factorization,
+and we conclude by commenting on these approach.
+
+\subsubsection{\fwa{} method}
+\label{ssub:edge_frank_wolfe}
+
+Seeking one direction for every edge, our goal is thus to maximize
+$\sum_{u,v\in E} {s_{uv}}^T w_{uv}$. This objective can be written in matricial form as $\tr(SW) =
 \frop{S^T}{W}$, where $S \in \Rbb^{|E| \times d}$ is the $s_{uv}$ vectors for all edges stacked
@@
 trivial solution of letting each direction $w_{uv}$ merely be $\frac{s_{uv}}{\|s_{uv}\|}$, we
-further constraint $W$ to be of low rank. That is, what we really want to optimize is \[\min_{W\in
+further constraint $W$ to be of low rank. That is, what we want to optimize is \[\min_{W\in
 \mathbb{M}^{d\times |E|}} -\frop{S^T}{W} + \rank(W)\,.\] However, while the Frobenius inner product is
-linear in $W$, the rank operator is highly irregular and optimizing it directly is \NPc{}. Therefore
+linear in $W$, the rank operator is highly irregular and optimizing it directly is \NPc{}.  Therefore,
 we relax the rank term by the nuclear norm of $W$, defined as $\|W\|_* = \sum_{i=1}^{\min(d,|E|)}
@@
 that the rank of $W$ is
-equal to the number of non-zero singular value while the nuclear norm is essentially an $\ell_1$ norm on
-the spectrum of $W$. Therefore, as minimizing the $\ell_1$ enforce sparsity, minimizing the nuclear
+equal to the number of non-zero singular value, while the nuclear norm is essentially an $\ell_1$ norm on
+the spectrum of $W$. Therefore, as minimizing the $\ell_1$ enforces sparsity, minimizing the nuclear
 norm tends to make the rank smaller by ensuring the presence of zeros in the singular values of $W$.
@@
 
-At this stage, we could also add the node loss term, express in matric form as $\|X - B -
+At this stage, we could also add the node loss term \eqref{eq:edge_node_loss}, express in matricial
+form as $\|X - B -
 C\left(A\circ W^T \right)\|_F^2$, where $B\in \Rbb^{|V| \times d}$ is the node bias vectors $b_u$
@@
 previous formulation, we thus relax it by simply imposing that the $\ell_2$ norm of the columns of
-$W$ are not too large. In fact, we already have one bound on these column norms, because of the
-nuclear norm constraint. One can indeed show that for $W\in \Rbb^{d\times |E|}$ such that $\|W\|_*
-\leq \delta$, the \ith{} column $W_{:,i}$ of $W$ satisfies $\|W_{:,i}\|_2 \leq \delta\sqrt{d}$.
+$W$ are not too large. In fact, we already have a bound on these column norms, because of the
+nuclear norm constraint.
+\begin{prop}
+For any $W\in \Rbb^{d\times |E|}$ such that $\|W\|_* \leq \delta$, the \ith{} column $W_{:,i}$ of
+$W$ satisfies $\|W_{:,i}\|_2 \leq \delta\sqrt{d}$.
+\end{prop}
+\begin{proof}
+  Let $W$ be a matrix in $\Rbb^{d\times m}$ with $d < m$ such that $\|W\|_* \leq \delta$.
+  Furthermore, let its singular value decomposition be such that we can write $W = \sum_{l=1}^d
+  \sigma_l u_l v_l^T$, where $u_l$ and $v_l$ are $d$ unit-norm vectors, of dimension $d$ and $m$
+  respectively. We can thus express the general term $W_{k,i}$ as $\sum_{l=1}^d \sigma_l v_{l,i}
+  u_{l,k}$. Because $v_l$ is unit norm, $v_{l,i} \leq 1$ and $$W_{k,i}^2 \leq \left( \sum_{l=1}^d
+  \sigma_l u_{l,k} \right)^2 \leq \sum_{l=1}^d \sigma_l^2 \sum_{l=1}^d u_{l,k}^2$$ using the
+  Cauchy--Schwartz inequality. The norm of the \ith{} column of $W$ then satisfies:
+  \begin{equation*}
+    \|W_{:,i}\|_2 = \left( \sum_{k=1}^m W_{k,i}^2 \right)^\lhalf
+    \leq \left( \sum_{k=1}^m \left( \sum_{l=1}^d \sigma_l^2 \sum_{l=1}^d u_{l,k}^2 \right) \right)^\lhalf
+    = \left( \left( \sum_{l=1}^d \sigma_l^2 \right) \left( \sum_{l=1}^d \sum_{k=1}^m u_{l,k}^2 \right) \right)^\lhalf
+  \end{equation*}
+  From $\sum_{l=1}^d |\sigma_l | \leq \delta$, we have that $\sum_{l=1}^d \sigma_l^2 \leq \delta^2$.
+  Combined with $u_l$ being unit norm, we conclude that $\|W_{:,i}\|_2 \leq \delta\sqrt{d}$.
+\end{proof}
 However, setting $\delta = \nicefrac{1}{\sqrt{d}}$ is too stringent in high dimension, and because
@@
 
-\begin{aside}
-  Let $W$ be a matrix in $\Rbb^{d\times m}$ with $d < m$ such that $\|W\|_* \leq \delta$.
-  Furthermore, let its singular value decomposition be such that we can write $W = \sum_{l=1}^d
-  \sigma_l u_l v_l^T$, where $u_l$ and $v_l$ are $d$ unit-norm vectors, of dimension $d$ and $m$
-  respectively. We can thus express the general term $W_{k,i}$ as $\sum_{l=1}^d \sigma_l v_{l,i}
-  u_{l,k}$. Because $v_l$ is unit norm, $v_{l,i} \leq 1$ and $W_{k,i}^2 \leq \left( \sum_{l=1}^d
-  \sigma_l u_{l,k} \right)^2 \leq \sum_{l=1}^d \sigma_l^2 \sum_{l=1}^d u_{l,k}^2$, using
-  Cauchy--Schwartz inequality. The norm of the \ith{} column of $W$ then satisfies:
-  \begin{equation*}
-    \|W_{:,i}\|_2 = \left( \sum_{k=1}^m W_{k,i}^2 \right)^\lhalf
-    \leq \left( \sum_{k=1}^m \left( \sum_{l=1}^d \sigma_l^2 \sum_{l=1}^d u_{l,k}^2 \right) \right)^\lhalf
-    = \left( \left( \sum_{l=1}^d \sigma_l^2 \right) \left( \sum_{l=1}^d \sum_{k=1}^m u_{l,k}^2 \right) \right)^\lhalf
-  \end{equation*}
-  From $\sum_{l=1}^d |\sigma_l | \leq \delta$, we have that $\sum_{l=1}^d \sigma_l^2 \leq \delta^2$.
-  Combined with $u_l$ being unit norm, we conclude that $\|W_{:,i}\|_2 \leq \delta\sqrt{d}$.
-\end{aside}
-
 Because \eqref{eq:edge_FW} is of the form $\min_{x\in \mathcal{X}} f(x)$ where $f$ and $\mathcal{X}$
 are convex, we can solve it using the Frank--Wolfe algorithm~\autocites{FrankWolfe56}{Jaggi2013a},
-which enjoy properties like a $O(\frac{1}{t})$ convergence rate, low computational cost by avoiding
-projection step and sparse solution. It is an iterative procedure that maintains an estimate
+which enjoy properties like a $O(\frac{1}{t})$ convergence rate, a low computational cost by avoiding
+projection step and the sparsity of its solution. It is an iterative procedure that maintains an estimate
 $x^{(t)}$ of the solution and works succinctly as follow: it linearizes the objective $f$ at the
-current position $x^{(t)}$ by computing $\nabla f(x^{(t)})$, it finds a minimum $s$ of that
+current position $x^{(t)}$ by computing $\nabla f(x^{(t)})$, it finds a minimum $s^{(t)}$ of that
 linearization within the domain $\mathcal{X}$ by solving $\argmin_{s\in \mathcal{X}}
 \innerp{s}{\nabla f(x^{(t)})}$ and it moves toward that minimizer by a step $\gamma$, setting
-$x^{(t+1)} = (1-\gamma) x^{(t)} + \gamma s$.\footnote{$\gamma$ is either found by a line search of
-set to $\nicefrac{t}{2+t}$.} In the case of \eqref{eq:edge_FW}, the linear minimization step amounts
+$x^{(t+1)} = (1-\gamma) x^{(t)} + \gamma s^{(t)}$.\footnote{$\gamma$ is either found by a line search or,
+in our case, set to $\nicefrac{t}{2+t}$.} In the case of \eqref{eq:edge_FW}, the linear minimization step amounts
 to $\argmin_{\|W\|_* \leq \delta} 2\mu W - S^T = \argmax_{\|W\|_* \leq \delta} S^T - 2\mu W$.
 Moreover, the $\delta$-ball of the nuclear norm is the convex hull of the matrix of the form $\delta
-u v^T$ with $\|u\|_2 = 1 = \|v\|_2$~\autocite{Jaggi2013a}. By letting $u_1$ and $v_1$ be
+u v^T$ with $\|u\|_2 = 1 = \|v\|_2$~\autocite{Jaggi2013a}. By letting $u_1^{(t)}$ and $v_1^{(t)}$ be
 respectively the left and right vectors associated with the largest singular value of $S^T - 2\mu
-W$, the minimizer \enquote{$s$} we are looking for is therefore $\delta u_1 v_1^T$.
+W^{(t)}$, the minimizer $s^{(t)}$ we are looking for is therefore $\delta u_1^{(t)} {v_1^{(t)}}^T$,
+and we have $W^{(t+1)} = (1-\gamma) W^{(t)} + \frac{t}{2+t} \delta u_1^{(t)} {v_1^{(t)}}^T$. We
+perform a small number $T$ of such iterations, and these two eigenvectors are computed in a
+time linear in the number of non-zero entries of $S^T - 2\mu W^{(t)}$~\autocite{topEigenvalue92}.
+Thus the overall complexity of this method, which we call \fwa{}, is $O(Td|E|)$.
 
-\medskip
+As said at the beginning of this section, the matrix formulation is not exactly solving
+\autoref{p:edge_full}. However, for evaluation purposes, here we describe two ways to extract $k$
+directions and an edge assignment from the matrix $W$ obtained through this \fwa{} method. The first
+is rather straightforward. It performs a $k$-means clustering of the columns of $W$ and use it as
+assignment, while the directions are the normalized cluster centers. The second involves an extra
+optimization step. Specifically, letting $r$ be the rank of $W$, we first compute a reduced SVD such
+that $W = U \Sigma V^T$ and let $P = U \in \mathbb{M}^{d\times r}$ and $Q^T = \Sigma V^T \in \Rbb^{r
+\times |E|}$. We then note that for any invertible matrix $R \in \Rbb^{r \times r}$, we have $W =
+PR^{-1} RQ^T$. Thus we look for a matrix $R$ that makes the columns of $RQ^T$ as close as possible
+of having unit $\ell_2$-norm, that is $R = \min_{R \in \Rbb^{r \times r}} \sum_{i=1}^E \left| \left\|
+{RQ^T}_{:,i} \right\|_2^2 - 1 \right|$. Finally, as before,  we cluster the columns of $RQ^T$ using
+$k$-means, and use the columns of $PR^{-1}$ as basis directions, noting there are not
+necessarily unit norm anymore.
 
-We can avoid the rank regularization term altogether by making low-rank decomposition of $W$
-explicit, that by writing $W=PQ^T$ and optimizing over both $P$ and $Q$. However, this comes at cost
+\subsubsection{\pqt{} low rank factorization}
+
+We can avoid the rank regularization term altogether by making the low-rank decomposition of $W$
+explicit, that by writing $W=PQ^T$ and optimizing over both $P$ and $Q$. We thus call this method
+\pqt{}. However, this comes at cost
 of convexity. Indeed, given a rank $k\leq \min(d,|E|)$, the problem becomes, assuming $A$ and $B$
@@
 While \eqref{eq:edge_PQ} is not jointly convex in both $P$ and $Q$, it is convex in one of
-them while the other is fixed. Therefore it can be solved via alternating optimization over $P$ and
+them when the other is fixed. Therefore it can be solved via alternating optimization over $P$ and
 $Q$ using standard (projected) gradient descent algorithms, with the following partial derivatives:
@@
 $k$ unit norm vectors of $P$ with weights whose square sum to $1$ and thus their norm is bounded by
-$\sqrt{k}$.
+$\sqrt{k}$. As opposed to all the previous methods, computing the full gradient is not linear in the
+number of edges. Indeed, computing the product $S^T Q$ requires $O(|E|^\omega)$ arithmetic
+operations, where $\omega \geq 2$\footnote{$\omega$ is the optimal exponent of matrix multiplication
+algorithm.}. This could be alleviate by using stochastic gradient descent. Finally, to obtain a
+solution of the \ecp{} \autoref{p:edge_full}, we take the directions to be the columns of the
+resulting $P$, and the assignment to be a $k$-means clustering of the rows of $Q$.
 
-Compared with \refeq{eq:edge_soft}, these two matrix formulations require learning more parameters,
-as it each edge has a different direction, albeit made up of a small number of basis directions. In
-the second case, we could furthermore impose a sparsity constraint on $Q$.
+\paragraph{Discussion}
+
+While the \pqt{} formulation is not convex, in practice we initialize it with the solution from
+\combined{} and it thus converge after few iterations. Namely, we set $P_0 \in \mathbb{M}^{d\times
+k}$ to be the final directions found by \combined{} and the rows of $Q_0 \in \mathbb{M}^{|E|\times
+k}$ to be equal to the corresponding $e_{yhat_{uv}}$\stodo{notation}. Furthermore, in the projection
+step of the $Q$ variable, we actually ensure that all the entries of $Q$ are positive, which makes
+them easier to  interpret as weights for the basis directions of $P$. This is something that we
+cannot guarantee on the decomposition of the matrix $W$ obtained by the \fwa{} method. On the other
+hand, if we run the \fwa{} algorithm for more than $k$ iterations, the rank of $W$ can be higher
+than $k$, which might allow a finer clustering of the edges.
+
+Compared with the \combined{} approach of the previous \autoref{sub:edge_vector}, these two matrix
+formulations require learning more parameters. Indeed, it increases from $dk$ to $d|E|$, as it each
+edge has a different direction, albeit made up of a small number of basis directions. In the case of
+\pqt{}, we could furthermore impose a sparsity constraint on the columns of $Q$ to reduces the
+number of coefficient to learn. Yet this additional complexity has the advantage of allowing
+overlapping clustering of the edges. This is useful in cases where two nodes have a relationship
+that is best described by a combination of the $k$ base directions. Moreover, in the case of \fwa{}
+this additional flexibility comes with no increase in computational cost.

### 2017-10-26 edge/model.tex
move relation between edge and signs to its own file

@@
 
-\paragraph{Generalization of signed graphs}
-\label{par:generalization_of_signed_graphs}
-
-Here we show why \autoref{p:edge_full} can model signed graphs as a special case, namely with $k=2$
-and specifically crafted profiles.
-Remember the learning
-bias we introduced in the previous chapter \vpageref{text:cc_new_bias}: every node belongs to
-exactly one of the $K$ clusters available, is connected positively with the other nodes belonging to
-the same cluster, and is connected negatively with the nodes belonging to every other clusters. Now
-let the dimension of profiles be $d=K$ and the number of directions $k=2$. Moreover, we define the
-two direction vectors as $w_1 = \frac{\onev}{\sqrt{d}} = w_+$ and $w_2 = -w_+ = w_-$, where
-$\onev$ is the vector of dimension $d$ with only $1$ coordinates. Finally, for a node $u$ in
-cluster $i$, we assume its profile $x_u$ is set as follow: all component are equal to $-b$ except
-for the \ith{} one, which is equal to $a$, where $a$ and $b$ are real constants we define now.
-
-First, because $x_u$ is a unit vector, we have that $a^2 + (K-1)b^2=1$, from which $a =
-\sqrt{1-(K-1)b^2}$ whenever $1-(K-1)b^2 \geq 0$, or equivalently $b \leq \frac{1}{\sqrt{K-1}}$.
-Next, we look at $\left(x_u \circ x_v\right)^T \onev = {s_{uv}}^T \onev = \frac{1}{\sqrt{d}}
-\sum_{i=1}^d  {s_{uv}}_{;i}$ for any edge $(u,v) \in E$. If $u$ and $v$ are in the same cluster, then
-$\sum_{i=1}^d  {s_{uv}}_{;i} = a^2 + (K-1)b^2 = 1$, implying that ${s_{uv}}^T w_+ = \frac{1}{d} >
-{s_{uv}}^T w_- = -\frac{1}{d}$. Otherwise, $\sum_{i=1}^d  {s_{uv}}_{;i} = (K-2)b^2 - 2ab = (K-2)b^2 -
-2b\sqrt{1-(K-1)b^2} = f_K(b)$. By computing the derivative of $f_K$ with respect to $b$, we find
-that it reaches a minimum of $-\frac{1}{K-1}$ at $b = \frac{1}{\sqrt{K(K-1)}} \leq
-\frac{1}{\sqrt{K-1}}$. In this case, ${s_{uv}}^T w_+ = -\frac{1}{(K-1) \sqrt{d}} < {s_{uv}}^T w_- =
-\frac{1}{(K-1) \sqrt{d}}$. To maximize $\sum_{(u,v)\in E} \max_{\ell \in \{1, 2\}} {s_{uv}}^T
-w_\ell$, we therefore assign $w_+$ to edges whose both endpoints belong to the same cluster, and
-$w_-$ to edges joining two different clusters.
-
-Moreover, given this assignment, we can show that $w_+$ and $w_-$ are the optimal directions, under
-the additional uniformity assumption that there are exactly $p$ edges within every cluster, and $q$
-edges between any two clusters. First, we define
-\begin{equation*}
-  s_{\mathrm{inner}} = \sum_{(u,v) \in E;\, \cluster(u)=\cluster(v)} s_{uv} \qquad \text{and} \qquad
-  s_{\mathrm{outer}} = \sum_{(u,v) \in E;\, \cluster(u)\neq \cluster(v)} s_{uv}
-\end{equation*}
-to be the sum of all edge vectors within and between clusters respectively. By our choice of $a$ and
-$b$, we have that $s_{\mathrm{inner}} = p\onev$ and $s_{\mathrm{outer}} = -\frac{q}{2}\onev$. Now
-recall that the solution of $\argmax_{w\in \dsphere} w^T c$ is $\frac{c}{||c||}$. Therefore, the
-best direction to assign to inner edges is indeed $w_+$, and so is $w_-$ to outer edges.
-
-\begin{aside}
-For inner edge vectors, $s_{uv}$ in cluster $i$ has $a^2$ as its \ith{} component, and $b^2$
-elsewhere. There are $p$ of them in cluster $i$, so summing over all clusters we get
-$s_{\mathrm{inner}} = p\left(a^2 + (K-1)b^2\right)\onev = p\onev$.
-For outer edge vectors, we first consider the $q$ edges between $C_1$ and $C_2$,
-whose $s_{uv}$ is
-$\begin{pmatrix}
-  -ab \\
-  -ab \\
-  b^2 \\
-  b^2 \\
-  \vdots \\  
-\end{pmatrix}$. Likewise, between $C_1$ and $C_3$ we have
-$s_{uv}=\begin{pmatrix}
-  -ab \\
-  b^2 \\
-  -ab \\
-  b^2 \\
-  \vdots \\  
-\end{pmatrix}$. We thus we see that the first component is always $-ab$, that there is another
-such term at the other cluster index, while the rest is $b^2$. Summing all the vectors incident to
-$C_1$, we get
-$q\begin{pmatrix}
-  -ab(K-1) \\
-  b^2(K-2) -ab \\
-  b^2(K-2) -ab \\
-  \vdots \\  
-\end{pmatrix}$. Similarly, the sum of all vectors incident to $C_2$ is
-$q\begin{pmatrix}
-  b^2(K-2) -ab \\
-  -ab(K-1) \\
-  b^2(K-2) -ab \\
-  \vdots \\  
-\end{pmatrix}$, and so on for every cluster $i$. Summing over all clusters and dividing by $2$ to
-avoid double counting, we get $s_{\mathrm{outer}} = \frac{q}{2}\left( (K-1)\left(b^2(K-2) -ab\right)
--ab(K-1) \right)\onev = \frac{q(K-1)}{2}\left(b^2(K-2) -2ab\right)\onev =
-\frac{q(K-1)}{2}f_K(b)\onev = -\frac{q(K-1)}{2} \frac{1}{K-1} \onev = -\frac{q}{2}\onev$.
-\end{aside}
-
-Granted, the situation where all the nodes of a cluster have exactly the same profile and where the
-connections are so regular is highly idealized. Moreover, we did not formally prove that there is
-not another assignment (with different direction vectors) that would give a better objective
-value\footnote{Although given the symmetry of the problem, that would be surprising.}. Our point is
-rather that profiles and directions can be seen as a \enquote{latent} explanation for the signs in
-a balanced signed graphs, like our generative model of the first chapter.
+\input{signed}
 

### 2017-10-26 edge/signed.tex
move relation between edge and signs to its own file

\paragraph{Generalization of signed graphs}
\label{par:generalization_of_signed_graphs}

Here we show why \autoref{p:edge_full} can model signed graphs as a special case, namely with $k=2$
and specifically crafted profiles.
Remember the learning
bias we introduced in the previous chapter \vpageref{text:cc_new_bias}: every node belongs to
exactly one of the $K$ clusters available, is connected positively with the other nodes belonging to
the same cluster, and is connected negatively with the nodes belonging to every other clusters. Now
let the dimension of profiles be $d=K$ and the number of directions $k=2$. Moreover, we define the
two direction vectors as $w_1 = \frac{\onev}{\sqrt{d}} = w_+$ and $w_2 = -w_+ = w_-$, where
$\onev$ is the vector of dimension $d$ with only $1$ coordinates. Finally, for a node $u$ in
cluster $i$, we assume its profile $x_u$ is set as follow: all component are equal to $-b$ except
for the \ith{} one, which is equal to $a$, where $a$ and $b$ are real constants we define now.

First, because $x_u$ is a unit vector, we have that $a^2 + (K-1)b^2=1$, from which $a =
\sqrt{1-(K-1)b^2}$ whenever $1-(K-1)b^2 \geq 0$, or equivalently $b \leq \frac{1}{\sqrt{K-1}}$.
Next, we look at $\left(x_u \circ x_v\right)^T \onev = {s_{uv}}^T \onev = \frac{1}{\sqrt{d}}
\sum_{i=1}^d  {s_{uv}}_{;i}$ for any edge $(u,v) \in E$. If $u$ and $v$ are in the same cluster, then
$\sum_{i=1}^d  {s_{uv}}_{;i} = a^2 + (K-1)b^2 = 1$, implying that ${s_{uv}}^T w_+ = \frac{1}{d} >
{s_{uv}}^T w_- = -\frac{1}{d}$. Otherwise, $\sum_{i=1}^d  {s_{uv}}_{;i} = (K-2)b^2 - 2ab = (K-2)b^2 -
2b\sqrt{1-(K-1)b^2} = f_K(b)$. By computing the derivative of $f_K$ with respect to $b$, we find
that it reaches a minimum of $-\frac{1}{K-1}$ at $b = \frac{1}{\sqrt{K(K-1)}} \leq
\frac{1}{\sqrt{K-1}}$. In this case, ${s_{uv}}^T w_+ = -\frac{1}{(K-1) \sqrt{d}} < {s_{uv}}^T w_- =
\frac{1}{(K-1) \sqrt{d}}$. To maximize $\sum_{(u,v)\in E} \max_{\ell \in \{1, 2\}} {s_{uv}}^T
w_\ell$, we therefore assign $w_+$ to edges whose both endpoints belong to the same cluster, and
$w_-$ to edges joining two different clusters.

Moreover, given this assignment, we can show that $w_+$ and $w_-$ are the optimal directions, under
the additional uniformity assumption that there are exactly $p$ edges within every cluster, and $q$
edges between any two clusters. First, we define
\begin{equation*}
  s_{\mathrm{inner}} = \sum_{(u,v) \in E;\, \cluster(u)=\cluster(v)} s_{uv} \qquad \text{and} \qquad
  s_{\mathrm{outer}} = \sum_{(u,v) \in E;\, \cluster(u)\neq \cluster(v)} s_{uv}
\end{equation*}
to be the sum of all edge vectors within and between clusters respectively. By our choice of $a$ and
$b$, we have that $s_{\mathrm{inner}} = p\onev$ and $s_{\mathrm{outer}} = -\frac{q}{2}\onev$. Now
recall that the solution of $\argmax_{w\in \dsphere} w^T c$ is $\frac{c}{||c||}$. Therefore, the
best direction to assign to inner edges is indeed $w_+$, and so is $w_-$ to outer edges.

\begin{aside}
For inner edge vectors, $s_{uv}$ in cluster $i$ has $a^2$ as its \ith{} component, and $b^2$
elsewhere. There are $p$ of them in cluster $i$, so summing over all clusters we get
$s_{\mathrm{inner}} = p\left(a^2 + (K-1)b^2\right)\onev = p\onev$.
For outer edge vectors, we first consider the $q$ edges between $C_1$ and $C_2$,
whose $s_{uv}$ is
$\begin{pmatrix}
  -ab \\
  -ab \\
  b^2 \\
  b^2 \\
  \vdots \\  
\end{pmatrix}$. Likewise, between $C_1$ and $C_3$ we have
$s_{uv}=\begin{pmatrix}
  -ab \\
  b^2 \\
  -ab \\
  b^2 \\
  \vdots \\  
\end{pmatrix}$. We thus we see that the first component is always $-ab$, that there is another
such term at the other cluster index, while the rest is $b^2$. Summing all the vectors incident to
$C_1$, we get
$q\begin{pmatrix}
  -ab(K-1) \\
  b^2(K-2) -ab \\
  b^2(K-2) -ab \\
  \vdots \\  
\end{pmatrix}$. Similarly, the sum of all vectors incident to $C_2$ is
$q\begin{pmatrix}
  b^2(K-2) -ab \\
  -ab(K-1) \\
  b^2(K-2) -ab \\
  \vdots \\  
\end{pmatrix}$, and so on for every cluster $i$. Summing over all clusters and dividing by $2$ to
avoid double counting, we get $s_{\mathrm{outer}} = \frac{q}{2}\left( (K-1)\left(b^2(K-2) -ab\right)
-ab(K-1) \right)\onev = \frac{q(K-1)}{2}\left(b^2(K-2) -2ab\right)\onev =
\frac{q(K-1)}{2}f_K(b)\onev = -\frac{q(K-1)}{2} \frac{1}{K-1} \onev = -\frac{q}{2}\onev$.
\end{aside}

Granted, the situation where all the nodes of a cluster have exactly the same profile and where the
connections are so regular is highly idealized. Moreover, we did not formally prove that there is
not another assignment (with different direction vectors) that would give a better objective
value\footnote{Although given the symmetry of the problem, that would be surprising.}. Our point is
rather that profiles and directions can be seen as a \enquote{latent} explanation for the signs in
a balanced signed graphs, like our generative model of the first chapter.

### 2017-10-27 edge/signed.tex
2-CC version of the connection between edge and signed
The previous approaches was indeed not correct, for reasons explained
in the aside paragraph :(

@@
 
-Here we show why \autoref{p:edge_full} can model signed graphs as a special case, namely with $k=2$
-and specifically crafted profiles.
-Remember the learning
-bias we introduced in the previous chapter \vpageref{text:cc_new_bias}: every node belongs to
-exactly one of the $K$ clusters available, is connected positively with the other nodes belonging to
-the same cluster, and is connected negatively with the nodes belonging to every other clusters. Now
-let the dimension of profiles be $d=K$ and the number of directions $k=2$. Moreover, we define the
-two direction vectors as $w_1 = \frac{\onev}{\sqrt{d}} = w_+$ and $w_2 = -w_+ = w_-$, where
-$\onev$ is the vector of dimension $d$ with only $1$ coordinates. Finally, for a node $u$ in
-cluster $i$, we assume its profile $x_u$ is set as follow: all component are equal to $-b$ except
-for the \ith{} one, which is equal to $a$, where $a$ and $b$ are real constants we define now.
+The learning bias of \autoref{chap:cc} implicitly defined an idealized \esp{} problem as:
+\begin{problem}[idealized \esp{}]
+  \label{p:edge_esp}
+  Given $n$ nodes belonging latently to $K$ groups, predict whether two nodes $u$ and $v$ belong to
+  the same group (that is $(u,v)$ is positive) or not (that is $(u,v)$ is negative).
+\end{problem}
+In \autoref{sec:gtx} we study in particular the important special case where $K=2$. Now consider an
+instance of \autoref{p:edge_full} where $k=2$, and where the profiles and edges are based on group
+membership. We show that the optimal solution of such an instance is exactly the solution of this
+idealized \esp{} problem on the same graph.
 
-First, because $x_u$ is a unit vector, we have that $a^2 + (K-1)b^2=1$, from which $a =
-\sqrt{1-(K-1)b^2}$ whenever $1-(K-1)b^2 \geq 0$, or equivalently $b \leq \frac{1}{\sqrt{K-1}}$.
-Next, we look at $\left(x_u \circ x_v\right)^T \onev = {s_{uv}}^T \onev = \frac{1}{\sqrt{d}}
-\sum_{i=1}^d  {s_{uv}}_{;i}$ for any edge $(u,v) \in E$. If $u$ and $v$ are in the same cluster, then
-$\sum_{i=1}^d  {s_{uv}}_{;i} = a^2 + (K-1)b^2 = 1$, implying that ${s_{uv}}^T w_+ = \frac{1}{d} >
-{s_{uv}}^T w_- = -\frac{1}{d}$. Otherwise, $\sum_{i=1}^d  {s_{uv}}_{;i} = (K-2)b^2 - 2ab = (K-2)b^2 -
-2b\sqrt{1-(K-1)b^2} = f_K(b)$. By computing the derivative of $f_K$ with respect to $b$, we find
-that it reaches a minimum of $-\frac{1}{K-1}$ at $b = \frac{1}{\sqrt{K(K-1)}} \leq
-\frac{1}{\sqrt{K-1}}$. In this case, ${s_{uv}}^T w_+ = -\frac{1}{(K-1) \sqrt{d}} < {s_{uv}}^T w_- =
-\frac{1}{(K-1) \sqrt{d}}$. To maximize $\sum_{(u,v)\in E} \max_{\ell \in \{1, 2\}} {s_{uv}}^T
-w_\ell$, we therefore assign $w_+$ to edges whose both endpoints belong to the same cluster, and
-$w_-$ to edges joining two different clusters.
+Let us first define this instance formally. Given a graph $G=(V,E)$, we partition the nodes in two
+groups $L$ and $R$. We then define the nodes profiles $X$ as follow. We first draw \uar{} a vector $p$
+from $\{-1, 1\}^d$, then let the profiles in $L$ be $\frac{p}{\sqrt{d}}$ and the
+profiles in $R$ be $-\frac{p}{\sqrt{d}}$. We call this instance $\mathcal{I} = (G, X, k=2)$. With
+this choice, we have
+\begin{equation*}
+  s_{uv} =
+  \begin{cases}
+    \frac{\onev}{d} = s_{\mathrm{inner}} & \quad \text{if $u$ and $v$ are in the same group,} \\
+    -\frac{\onev}{d} = s_{\mathrm{outer}} & \quad \text{if $u$ and $v$ are in different groups.}
+  \end{cases}
+\end{equation*}
+Let us denote by $m_i$ the number of edges that are internal to $L$ and $R$, and by $m_o = |E| -
+m_i$ the number of edges between $L$ and $R$.
 
-Moreover, given this assignment, we can show that $w_+$ and $w_-$ are the optimal directions, under
-the additional uniformity assumption that there are exactly $p$ edges within every cluster, and $q$
-edges between any two clusters. First, we define
+\begin{prop}
+  Let $\mathcal{D}_2 = \{w_1^\star, w_2^\star\}$ be an optimal solution of \autoref{p:edge_full} on the
+  instance $\mathcal{I}$. Then $\mathcal{D}_2$ provides the following optimal solution to
+  \autoref{p:edge_esp}: each every edge explained by $w_1^\star$ is predicted positive, and every
+  edge explained by $w_2^\star$ is predicted negative.
+\end{prop}
+\begin{proof}
+  Take any $w_1, w_2 \in \dsphere \times \dsphere$ such that $w_1 \neq w_2$. Assume without loss of
+  generality that $w_1 = \argmax_{w\in \{w_1, w_2\}} {s_{\mathrm{inner}}}^T w$. This implies that:
   \begin{equation*}
-  s_{\mathrm{inner}} = \sum_{(u,v) \in E;\, \cluster(u)=\cluster(v)} s_{uv} \qquad \text{and} \qquad
-  s_{\mathrm{outer}} = \sum_{(u,v) \in E;\, \cluster(u)\neq \cluster(v)} s_{uv}
+    {s_{\mathrm{inner}}}^T w_1 = \frac{1}{d}\sum_{i=1}^d {w_1}_{;i} >
+    \frac{1}{d}\sum_{i=1}^d {w_2}_{;i} = {s_{\mathrm{inner}}}^T w_2 \,.
   \end{equation*}
-to be the sum of all edge vectors within and between clusters respectively. By our choice of $a$ and
-$b$, we have that $s_{\mathrm{inner}} = p\onev$ and $s_{\mathrm{outer}} = -\frac{q}{2}\onev$. Now
-recall that the solution of $\argmax_{w\in \dsphere} w^T c$ is $\frac{c}{||c||}$. Therefore, the
-best direction to assign to inner edges is indeed $w_+$, and so is $w_-$ to outer edges.
+  In turn, we have that:
+  \begin{equation*}
+    {s_{\mathrm{outer}}}^T w_2 = -\frac{1}{d}\sum_{i=1}^d {w_2}_{;i} >
+    -\frac{1}{d}\sum_{i=1}^d {w_1}_{;i} = {s_{\mathrm{outer}}}^T w_1 \,.
+  \end{equation*}
+  We can thus rewrite the maximization objective \eqref{eq:edge_full} of \autoref{p:edge_full} as:
+  \begin{align}
+    \label{eq:edge_special}
+    \argmax_{w_1,w_2 \in \dsphere} & \quad \sum_{u,v \in E} \max_{w \in \{w_1, w_2\}} {s_{uv}}^T w = \\
+    \argmax_{w_1,w_2 \in \dsphere} & \quad
+    \sum_{\substack{u,v \in E\\ s_{uv} = s_{\mathrm{inner}}}} {s_{\mathrm{inner}}}^T w_1 +
+    \sum_{\substack{u,v \in E\\ s_{uv} = s_{\mathrm{outer}}}} {s_{\mathrm{outer}}}^T w_2 = \notag \\
+    \argmax_{w_1,w_2 \in \dsphere} & \quad m_i \frac{1}{d} \onev + m_o \frac{-1}{d} \onev \notag
+  \end{align}
+  One can check using Lagrange multipliers that for any $c \in \Rbb^n$, $\max_{w \in \dsphere} c^T w
+  = \frac{c}{\|c\|}$. The solution of \eqref{eq:edge_special} is therefore:
+  \begin{equation*}
+    w_1^\star = \frac{\onev}{\sqrt{d}} \qquad \text{and} \qquad w_2^\star = \frac{-\onev}{\sqrt{d}}
+  \end{equation*}
+  As claimed, the edges associated with $w_1^\star$ (respectively $w_2^\star$) are the edges that
+  are to be predicted positive (respectively negative) in \autoref{p:edge_esp}.
+\end{proof}
 
 \begin{aside}
-For inner edge vectors, $s_{uv}$ in cluster $i$ has $a^2$ as its \ith{} component, and $b^2$
-elsewhere. There are $p$ of them in cluster $i$, so summing over all clusters we get
-$s_{\mathrm{inner}} = p\left(a^2 + (K-1)b^2\right)\onev = p\onev$.
-For outer edge vectors, we first consider the $q$ edges between $C_1$ and $C_2$,
-whose $s_{uv}$ is
-$\begin{pmatrix}
-  -ab \\
-  -ab \\
-  b^2 \\
-  b^2 \\
-  \vdots \\  
-\end{pmatrix}$. Likewise, between $C_1$ and $C_3$ we have
-$s_{uv}=\begin{pmatrix}
-  -ab \\
-  b^2 \\
-  -ab \\
-  b^2 \\
-  \vdots \\  
-\end{pmatrix}$. We thus we see that the first component is always $-ab$, that there is another
-such term at the other cluster index, while the rest is $b^2$. Summing all the vectors incident to
-$C_1$, we get
-$q\begin{pmatrix}
-  -ab(K-1) \\
-  b^2(K-2) -ab \\
-  b^2(K-2) -ab \\
-  \vdots \\  
-\end{pmatrix}$. Similarly, the sum of all vectors incident to $C_2$ is
-$q\begin{pmatrix}
-  b^2(K-2) -ab \\
-  -ab(K-1) \\
-  b^2(K-2) -ab \\
-  \vdots \\  
-\end{pmatrix}$, and so on for every cluster $i$. Summing over all clusters and dividing by $2$ to
-avoid double counting, we get $s_{\mathrm{outer}} = \frac{q}{2}\left( (K-1)\left(b^2(K-2) -ab\right)
--ab(K-1) \right)\onev = \frac{q(K-1)}{2}\left(b^2(K-2) -2ab\right)\onev =
-\frac{q(K-1)}{2}f_K(b)\onev = -\frac{q(K-1)}{2} \frac{1}{K-1} \onev = -\frac{q}{2}\onev$.
+  It is natural to ask how this result can be extended to more than $K=2$ groups. The most natural
+  idea is to use one-hot encoding of the group membership as profiles, with additional negative
+  values as in the example of \autoref{fig:edge_exe}. Namely, set the profiles of the \ith{} group
+  as follow: all component are equal to $-b$ except for the \ith{} one, which is equal to $a$, where
+  $a$ and $b$ are real constants such that $a^2 + (K-1)b^2=1$. However, in that case, we check
+  numerically that $w_+ = \frac{\onev}{\sqrt{d}}$ and $w_- = \frac{-\onev}{\sqrt{d}}$ is the not the
+  optimal solution, and that a solution achieving a larger objective function value does not cluster
+  the edges between inner and outer.
+
+  The next thing I tried was to use the approach described later in \autoref{sub:edge_generation} in
+  order to optimize the profiles of a complete graphs with $K$ groups, given $w_+$, $w_-$ and the
+  inner/outer edges assignment. As one could expect from the symmetry of the problem, all the nodes
+  in the same group have the same profile. However, there are no clear pattern in those profile, and
+  once again there exists better solutions than $\{w_+, w_-\}$, which partition the edges in two
+  roughly balanced sets.
 \end{aside}
-
-Granted, the situation where all the nodes of a cluster have exactly the same profile and where the
-connections are so regular is highly idealized. Moreover, we did not formally prove that there is
-not another assignment (with different direction vectors) that would give a better objective
-value\footnote{Although given the symmetry of the problem, that would be surprising.}. Our point is
-rather that profiles and directions can be seen as a \enquote{latent} explanation for the signs in
-a balanced signed graphs, like our generative model of the first chapter.

### 2017-10-31 cc/bias_directed.tex
add missing experiments at the beginning of CC chapter

@@
 with the \complowrank{} approach ran directly on the undirected graphs. This is a fair comparison,
-for \complowrank{} is designed to works solely on undirected graphs.\Todo{run the undirected
-experiments.}
+for \complowrank{} is designed to works solely on undirected graphs.
+
+As we can see in \autoref{tab:bias_exp_undir}, our methods perform worse than when running on the
+original directed graphs. Looking at \uslpropGsec{} on a 15\% training set as an example, we observe
+that the MCC decreases by one (\kiw{}) to almost ten (\epi{}) points. On the other hand, the
+\complowrank{} seems to perform better. Recall however that because it is an undirected method in
+the first place and because we roughly double the number of signed edges, a 15\% training size means
+the algorithm gets to observe more signs and therefore perform better. This small experiment thereby
+shows that although the performance remains decent, not having direction information hurt our bias.
+
+\input{bias_exp_undir}

### 2017-10-31 cc/bias_model.tex
add some ref
add missing experiments at the beginning of CC chapter

@@
 reactions accelerated by other nodes. Intuitively, this is quite far-fetched. To demonstrate this,
-we borrow several gene regulatory networks from \autocite[Table 1]{BioSigned09} and perform the same
-experiments as in the previous chapter.\Todo{get the data and run the experiments.}
+we borrow the following three gene regulatory networks from \autocite[Table 1]{BioSigned09}, and
+display their statistics in \autoref{tab:bias_bio_dataset}:
+\begin{enumerate}[1.]
+  \item The signaling pathways of the \hip{} CA1 neuron~\autocite{Hippocamp05}.
+  \item The interactions between genes and their products that are involved in \can{} development in
+    humans and mouses~\autocite{Cancer07}.
+  \item The \reg{} database\autocite{RegulonDB16}, summarizing the connection between transcription
+    factors and their targets in the Escherichia coli K-12 bacteria. We download four
+    files\footnote{\url{http://regulondb.ccg.unam.mx/menu/download/datasets/index.jsp}} containing
+    experimental evidence and keep only the positive and negative edges with \textsf{Strong}
+    support.
+\end{enumerate}
+%\url{http://causalbionet.com/}
+
+\begin{table}[bt]
+  \centering
+  \small
+  \caption{Biological dataset properties. The columns have the same meaning as in
+    \autoref{tab:dataset} \vpageref{tab:dataset}\label{tab:bias_bio_dataset}}
+  \begin{tabular}{lrrrrrrrr}
+    \toprule
+    Dataset & \thead{$|V|$}       & \thead{$|E|$}       & \thead{$\frac{|E|}{|V|}$} &
+    \thead{$\frac{|E^+|}{|E|}$} & \thead{$\frac{\Psi^2_{G''}(Y)}{|E|}$} &
+    \thead{$\frac{\Psi_G(Y)}{|E|}$} & \thead{reciprocal\\ edges} &
+    \thead{reciprocal\\ disagreement} \\
+    \midrule                                                                                                   
+    \hip{} &  \np{501} & \np{1046} &               2.1 &               69.5\% &                       .056 &                    .124 &       0.2\% &              0.0\% \\
+    \can{} & \np{1240} & \np{3065} &               2.5 &               78.4\% &                       .047 &                    .108 &       5.5\% &             27.1\% \\
+    \reg{} & \np{1700} & \np{2570} &               1.5 &               50.3\% &                       .060 &                    .148 &       0.0\% &              0.0\% \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\input{bias_exp_bio.tex}
+
+Compared with the \dssn{} of the previous chapter, those biological are smaller, have lower average
+degree and little to no reciprocal edges. Trying to solve the \esp{} problem using the same
+procedure as in \autoref{s:exp}, we read in \autoref{tab:cc_bio_exp} that our methods have lower
+absolute MCC performance (for instance, \uslpropGsec{} with a 15\% training size achieves $20$, $24$
+and $45$ compared with values between $36$ and $61$ on \dssn{}).  To further illustrate the mismatch
+between the troll bias and biological networks, notice that the standard deviation figures are
+several times larger than those reported in \autoref{tab:all_mcc}.

### 2017-10-31 edge/intro.tex
small pass over the edge intro

@@
 First, we want an interpretable model that, given two connected nodes, can explain why they are
-linked. This will lead to the use of linear models, for they are both easily amenable to
+linked. This lead us to use a linear model, for it is both easily amenable to
 interpretation and computationally efficient. Second, edge-labeled graphs are susceptible to further
@@
 elaborate and highly expressive notion of multilayer graph, in this chapter we shall restrict
-ourselves to the following\footnote{Which is roughly corresponding to \emph{single aspect}
-\emph{node-aligned} multilayer graphs of \autocite{Kivela2014}.}:
+ourselves to the following\footnote{Which roughly corresponds to the \emph{single aspect},
+\emph{node-aligned} multilayer graph of \autocite{Kivela2014}.}:
 \begin{definition}[$k$-multilayer graph]
   A $k$-multilayer graph $G$ is a pair $(V, \{E_1, \ldots, E_k \})$, where $V$ is a set of $n$ nodes
-  and each $E_i \subset V\times V$, $i \in \rangesk$ is a set of edges. We call the subgraph $G_i =
+  and each $E_i \subset V\times V$ is a set of edges, for $i \in \rangesk$. We call the subgraph $G_i =
   (V, E_i)$ the \ith{} layer of $G$ and it will be convenient to think of $G$ as the superposition
@@
 This additional information on the nature of the relationships allow traditional graph tasks to be
-performed more precisely and with higher granularity, such as node ranking~\autocite{HRRanking13};
+performed more precisely and with higher granularity. Consider for instance node ranking~\autocite{HRRanking13};
 node clustering \autocites[Section 2]{surveyAttributedClustering15}[Section 4.5.1]{Kivela2014} and
 the study of information spreading \autocite{SpreadingMultiSurvey15}. However, to produce such finer
-results, the methods we just referenced requires labeled that in many cases are not available,
-making our work rather useful. In the following table we give some examples of networks whose edges
-are currently unlabeled and why it would be beneficial to have them.
+results, the methods we just referenced require labeled edges that in many cases are not available,
+thus justifying the usefulness of our work. Indeed, in \autoref{tab:edge_apps}, we give some examples of
+networks whose edges are currently unlabeled and what type of relations they could possibly denotes.
+In other words, given a small list of possible reasons in terms of attributes and two nodes, why are
+they connected?
 
-\noindent\begin{tabulary}{\textwidth}{lLLLL}
+\setlength{\fullpage}{\textwidth+\marginparsep+\marginparwidth}
+\begin{table}[h]
+  \small
+  \setlength{\tabcolsep}{4pt}
+  \centering
+  \caption{Real attributed graphs that we have collected or are easily accessible. Unfortunately,
+    none of them comes with known edge types, which implies the last column is merely speculation.
+  \label{tab:edge_apps}}
+  {\renewcommand{\arraystretch}{0.95}%
+  \begin{tabulary}{\fullpage}{lLLLL}
     \toprule
-  Source & Node & Edge & Node profiles & Possible edge meanings \\
+    {Source} & {Node} & {Edge} & {Node profiles} & {Possible edge meanings} \\
     \midrule
-  DBLP & author & citation & keyword count of their papers & \\
-  Github & developer & following & number of lines written in various programming languages & \\
-  Amazon & video game products & \enquote{frequently bought together} & price, genre, time of release, popularity and evaluation & \\
-  Wikipedia & article & internal link & category, length, edit history, bag of words & \\
-  Flickr & user & following & activity statistics, groups membership, count of photo tags & \\
-  Foursquare & venue & \enquote{frequently visited together} & location, popularity, category, the time distribution at which they are visited & \\
-  IMDB & movie & sharing actors & genre, release date and popularity & \\
-  Blablacar & driver or passenger & shared ride & experience, location, preferences, rating & \\
+    DBLP & author & citation & keyword count of their papers & scientific topic \\
+    Github & developer & following & popularity, number of lines written in various programming languages &
+    common hobby, involvement in the same open source project, same
+    workplace, famous role model \\
+    Amazon & video game products & \enquote{frequently bought together} & price, genre, time of release, popularity and evaluation &
+    part of a series, cheap bundle, same genre and level of satisfaction \\
+    Wikipedia & article & internal link & category, length, edit history, bag of words &
+    generalization, specialization, organisation \\
+    Flickr & user & following & activity statistics, groups membership, count of photo tags &
+    geography, same tags distribution, famous role model,  \\ 
+    Foursquare & venue & \enquote{frequently visited together} & location, popularity, category, the time distribution at which they are visited &
+    closeness, part of a \enquote{series} (\eg{} movie theater $\rightarrow$ restaurant $\rightarrow$
+    bar $\rightarrow$ club) \\
+    IMDB & movie & sharing actors & genre, release date and popularity &
+    part of a series, common theme, same time period and genre \\
+    BlaBlaCar & driver or passenger & shared ride & experience, location, preferences, rating &
+    geography, professional, leisure \\
     \bottomrule
-\end{tabulary}
+  \end{tabulary}}
+\end{table}
 
+\medskip
 Besides predicting potentially more than two classes, the other difference with the setting of the
@@
 
-A common learning bias in attributed graphs\stodo{better transition from attributes to homophily} is
+It would be tempting to assume that nodes are connected because they share the same attributes.
+Indeed, a common learning bias in graphs is
 that nodes are connected together because they are similar, or as often said proverbially:
@@
 cannot only be explained by a global homophily along all attributes but rather by partial homophily,
-partial heterophily or even both. An example of this combination is balanced news
-diet~\autocite{balancedNews17}, where we try to connect users of a social network to prevent them
-from being locked in echo chambers. In that situation, a user $A$ might be more likely to listen to
+partial heterophily or even both. An example of this combination is the balanced news
+diet problem~\autocite{balancedNews17}, where we try to connect users of a social network to prevent them
+from being locked into topical echo chambers. In that situation, a user $A$ might be more likely to listen to
 the opposite point of view of user $B$ about gun control if they both share common demographics and
-interests.
+interests. This suggests, as we make it formal in the next section, that node attributes can include
+polarized opinions, \ie{} both positive and negative numerical values.
+
+\medskip
 
@@
 similarity between nodes profiles and seek a small number of explanations (corresponding to edge
-types) achieving good scores as prescribed by a linear model. After explaining how this problem can
-model a situation closely resembling the case of signed graphs
+types) achieving good scores as prescribed by a linear model. After explaining how this problem relates
+to the \esp{} problem on 2-clusters signed graphs
 \vpageref{par:generalization_of_signed_graphs}, we introduce in \autoref{sec:edge_methods} several
@@
 measure (\autoref{sub:edge_baseline}), to a convex relaxation of our objective function
-(\autoref{sub:edge_vector}) all the way to a richer and higher capacity matrix formulation allowing
+(\autoref{sub:edge_vector}) all the way through a richer and higher capacity matrix formulation allowing
 overlapping clustering of edges (\autoref{sub:edge_matrix}). Because finding publicly available real
 data with both node attributes and edge types proved difficult, we describe a synthetic model in
-\autoref{ssec:edge_synth} on which we perform extensive experiments. We then survey in
+\autoref{ssec:edge_exp}, on which we perform extensive experiments. We then survey in
 \autoref{sec:edge_related} works that are related to the \ecp{}, while noting that very few address

### 2017-10-31 edge/model.tex
pass over edge model and methods

@@
 Then we present two options to model the interactions between nodes and the role of attributes in
-\enquote{explaining} edges (\autoref{sub:edge_setting}). Choosing of these two options leads us to
+\enquote{explaining} edges (\autoref{sub:edge_setting}). Choosing one of these two options leads us to
 formulate the \ecp{} problem in \autoref{sub:edge_constraints}. We then describe additional
-constraints to take into account the topology of the graph. Finally we elaborate on the relation
+constraints allowing us to take into account the topology of the graph. Finally, we elaborate on the relation
 between \ecp{} and \esp{} in signed graphs.
@@
 adds a coupling across different attributes, since increasing the weight of an attribute mechanically
-affects the others.
+affects the weights of the other attributes.
 
@@
 $s_{uv}$ can be seen as constants, meaning that $g$ is linear and therefore interpretable. Finally,
-it also supports categorical data. We shall demonstrate all these features on a small scale example,
+it also supports categorical data. We shall demonstrate all these features on a small example,
 but we first define formally our problem.
@@
 graph shown in \autoref{fig:edge_exe_graph}, whose node attributes are listed
-\autoref{tab:edge_exe}. The first two attributes are categorical, denoting the company employing
+in Table~\ref{tab:edge_exe}. The first two attributes are categorical, denoting the company employing
 each user, either \textsf{G} or \textsf{F}. Instead of a classic binary one-hot encoding, we
@@
 $-1$ and $1$. Finally, \textsf{opinion} represents the view of former \textsf{F} employees about a
-societal issue, with $0$ denoting a neutral position. In \autoref{tab:edge_exe}, we also show two
+societal issue, with $0$ denoting a neutral position. In Table~\ref{tab:edge_exe}, we also show two
 directions that explain the connections in the graph $G$. The first covers being of the same age and
@@
 their goodness. For instance, we see that the edge between $x$ and $y$ has a low score, because
-those two users have indeed the same \textsf{age} and \textsf{opinion}.
+although those two users have the same \textsf{age} and used to work at \textsf{F}, they do not
+share a position at \textsf{G}.
 
@@
 is $x_u = b_u + \sum_{v \in \nei(u)} a_{uv} w_{uv}$, where $a_{uv}$ are real weights and $b_u
-\in \Rbb^d$ is the bias vector of $u$, its inherent profile take in isolation.
+\in \Rbb^d$ is the bias vector of $u$, \ie{} its inherent profile taken in isolation.
 Viewed in the other direction, this can be interpreted as follow: each time a node $u$
-connect to another node $v$ along a direction $w_{uv}$, it \enquote{consumes} a part of its profile
+connects to another node $v$ along a direction $w_{uv}$, it \enquote{consumes} a part of its profile
 proportional to $w_{uv}$. Therefore, we subtract the term
@@
 
-\medskip
+\bigskip
 
@@
 well studied~\autocite{sparseOptim12}. Namely, we let $y_{uv}$ be a softmax membership
-$k$-dimensional vector, that is we set $\forall i \in \rangesk\, y_{uv,i} =
+$k$-dimensional vector, that is we set $\forall i \in \rangesk\,, y_{uv,i} =
 \frac{\exp(\beta{s_{uv}}^T w_i)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_j)} $, with $\beta>0$ a large
 positive constant.  And we replace the $\ell_0$ norm by a $\ell_1$ norm (or even a $k$-support
-norm~\autocite{KsupportNorm12}), yielding the following $\mathcal{L}_{\mathrm{local}}$ to be
+norm~\autocite{KsupportNorm12}), yielding the following $\mathcal{L}_{\mathrm{local}}$ term to be
 minimized:
@@
   \sum_{u \in V} \left\| \sum_{v \in \nei(u)} y_{uv} \right\|_1 =
-  \sum_{u \in V} \sum_{\ell = 1}^k \sum_{v \in \nei(u)}
-  \frac{\exp(\beta{s_{uv}}^T w_\ell)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_j)}
+  \sum_{u \in V} \left( \sum_{i = 1}^k \sum_{v \in \nei(u)}
+  \frac{\exp(\beta{s_{uv}}^T w_i)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_j)} \right)
 \end{equation}
@@
 method as \kmeans{}, and assuming it requires $T$ iterations to converge, it has a linear complexity
-of $O(T|E|)$.
+of $O(Td|E|)$.
 \iffalse
@@
 practice we expect that only a small number $T$ of iterations would be needed. We call this method
-\lloyd{} and note it also has a linear complexity of $O(T|E|)$.
+\lloyd{} and note it also has a linear complexity of $O(Td|E|)$.
 
@@
 and expect them to have norm close to one in order to maximize the dot product in the ${s_{uv}}^T
-w_\ell$ term. A further simplification is that while \eqref{eq:edge_soft_early} requires to optimize
+w_\ell$ term\footnote{Indeed, since the first term is convex, its maximum can only be reach at the
+boundary of the feasible domain~\autocite[Theorem 32.1]{convexAnalysis70}. However, we are also
+subtracting another convex term and the resulting function is therefore more complicated.}.
+A further simplification is that while \eqref{eq:edge_soft_early} requires to optimize
 $kd + |E| + |V|d$ parameters, we fix $a_{uv} = 1$ and $b_u = \onev$ for all edges and nodes in order
@@
 be to increase $k$ to be equal to $E$. As we already mentioned, this makes the problem trivial. To
-avoid this, we add the constraint that all those $|E|$ directions are linear combination of a small
+avoid this, we add the constraint that all those $|E|$ directions are linear combinations of a small
 number $k$ of basis directions. In order to formulate this more clearly, we rewrite the previous
-objectives with matrices. We first show this can express in a form that is amenable to the
+objectives with matrices. We first show this can be expressed in a form that is amenable to the
 Frank--Wolfe algorithm. Then we take the alternative road of explicit low-rank matrix factorization,
-and we conclude by commenting on these approach.
+and we conclude by commenting on these approaches.
 
@@
 
-At this stage, we could also add the node loss term \eqref{eq:edge_node_loss}, express in matricial
+At this stage, we could also add the node loss term \eqref{eq:edge_node_loss}, expressed in matricial
 form as $\|X - B -
@@
 However, setting $\delta = \nicefrac{1}{\sqrt{d}}$ is too stringent in high dimension, and because
-this inequality is rather loose, it will likely results in very small columns norm. Instead of
+this inequality is rather loose, it will likely results in very small column norms. Instead of
 relying solely on the nuclear norm constraint, we therefore add an additional regularization term
@@
 {RQ^T}_{:,i} \right\|_2^2 - 1 \right|$. Finally, as before,  we cluster the columns of $RQ^T$ using
-$k$-means, and use the columns of $PR^{-1}$ as basis directions, noting there are not
+$k$-means, and use the columns of $PR^{-1}$ as basis directions, noting they are not
 necessarily unit norm anymore.
@@
 operations, where $\omega \geq 2$\footnote{$\omega$ is the optimal exponent of matrix multiplication
-algorithm.}. This could be alleviate by using stochastic gradient descent. Finally, to obtain a
+algorithm.}. This could be mitigated by using stochastic gradient descent. Finally, to obtain a
 solution of the \ecp{} \autoref{p:edge_full}, we take the directions to be the columns of the
@@
 k}$ to be the final directions found by \combined{} and the rows of $Q_0 \in \mathbb{M}^{|E|\times
-k}$ to be equal to the corresponding $e_{yhat_{uv}}$\stodo{notation}. Furthermore, in the projection
+k}$ to be equal to the corresponding $e_{\yhat_{uv}}$\stodo{notation}. Furthermore, in the projection
 step of the $Q$ variable, we actually ensure that all the entries of $Q$ are positive, which makes
@@
 overlapping clustering of the edges. This is useful in cases where two nodes have a relationship
-that is best described by a combination of the $k$ base directions. Moreover, in the case of \fwa{}
+that is best described by a combination of the $k$ base directions. Moreover, in the case of \fwa{},
 this additional flexibility comes with no increase in computational cost.

### 2017-10-31 edge/signed.tex
pass over edge model and methods

@@
     \sum_{\substack{u,v \in E\\ s_{uv} = s_{\mathrm{outer}}}} {s_{\mathrm{outer}}}^T w_2 = \notag \\
-    \argmax_{w_1,w_2 \in \dsphere} & \quad m_i \frac{1}{d} \onev + m_o \frac{-1}{d} \onev \notag
+    \argmax_{w_1,w_2 \in \dsphere} & \quad m_i {\frac{\onev}{d}}^T w_1  + m_o {\frac{-\onev}{d}} ^T w_2 \notag
   \end{align}
@@
   optimal solution, and that a solution achieving a larger objective function value does not cluster
-  the edges between inner and outer.
+  the edges between inner and outer groups.
 

### 2017-11-01 edge/conclusions.tex
pass over edge experiments and conclusions

@@
     \textcite{TimedUSSenate10} study the \np{110} congresses of the US senate between the years 1789
-    and 2008, where \np{1884} individual Senators are linked by their voting similarity, and show
+    and 2008, where \np{1884} individual Senators are linked by their voting similarity. They show
     how it improves the detection of relevant communities. Beyond predicting the relative order of
@@
     graph called the network schema~\autocite[Definition 3]{HINSurvey17}. This approach would need
-    to be extended to deal with higher order patterns.
+    to be extended to deal with higher order patterns, for instance in a manner reminiscent of
+    composition of paths in multilayer graphs~\autocites{metapath11}[Section
+    4.2]{KnowledgeGraphSurvey17}.
 

### 2017-11-01 edge/experiments.tex
pass over edge experiments and conclusions

@@
 structure~\autocites{Yang2013}{XuBayesian14}{Kataoka2016}, here we present the one we adopted for
-our initial experiments, as it is specifically tailored to the model introduced in the previous
-section. Given a number of directions $k$, a maximum number $k_\mathrm{local}$ of directions
+our initial experiments, as it is specifically tailored to the model introduced in \autoref{sub:edge_setting}.
+Given a number of directions $k$, a maximum number $k_\mathrm{local}$ of directions
 incident to each node, the dimension of profiles $d$ and a number of node $n$, it return a graph
@@
 of the graph. Furthermore, we control the number of coordinates $n_o$ that overlap with each other.
-That is, when $n_o = 0$, each direction has exactly \nicefrac{k}{d}\footnote{We choose $d$ to be a
-multiple of $k$.} non-zero dimension that are disjoint. On the other hand, if for instance $n_o=5$,
+That is, when $n_o = 0$, each direction has exactly \nicefrac{k}{d} non-zero dimension that are
+disjoint.\footnote{We choose $d$ to be a multiple of $k$.} On the other hand, if for instance $n_o=5$,
 then $5$ of the $d$ dimensions will have two directions with non-zero components on it. Those
@@
 subgraphs~\autocites{erdos1959random}{gilbertRG59} that we call \emph{blocks}. Furthermore, driven
-by a topological constraint mentioned earlier, we assign to each node a $k_\mathrm{local}$-tuple of
-directions such that two adjacent node have \emph{at least} one common direction that can later be
+by the topological constraint approximated by \eqref{eq:edge_local_loss} mentioned earlier, we
+assign to each node a $k_\mathrm{local}$-tuple of
+directions such that two adjacent nodes have \emph{at least} one common direction that can later be
 assigned to the edge between them. For convenience, let us first identify directions with colors and
@@
 color with $p$. Finally, we say that an edge $u,v$ is \emph{colorless} if the palette of $u$ and $v$ have
-no color in common. Because in the general case, it is not always possible to assign palette to
-every nodes such that there is no colorless edge,\footnote{For instance, consider a $4$-clique with $k=4$
+no color in common. Because in the general case, it is not always possible to assign a palette to
+every node such that there is no colorless edge,\footnote{For instance, consider a $4$-clique with $k=4$
 and $k_\mathrm{local}=2$. Without loss of generality, say we assign the palette $(1,2)$ to node
@@
 same palette). On the other hand, if we assign $(1,3)$ to the second, the third node can only have
-$(2,3)$, meaning the fourth node has again no palette available. The same situation with $(1,4)$ and
-$(2,4)$.} we now describe a simple heuristic. It performs a breath first visit of a subgraph,
+$(2,3)$, meaning the fourth node has again no palette available. The same situation arises with $(1,4)$ and
+$(2,4)$ respectively.} we now describe a simple heuristic. It performs a breath first visit of a subgraph,
 starting from a random node. Upon visiting an uncolored node $u$, it builds the intersection $P$ of
@@
 \autoref{fig:edge_palette} for a small illustration). Once all nodes have been colored, we count the
-number of edges that are colorless. We repeat this procedure, keeping track of the assignment
+number of edges that are colorless. We repeat this procedure, keeping track of the palettes assignment
 minimizing the number of colorless edges. Once we colored the nodes of every block, we look at pair
@@
 loss, defined respectively by equations \eqref{eq:edge_full} and
-\eqref{eq:edge_node_loss} in the previous section as:
+\eqref{eq:edge_node_loss} in \autoref{sec:edge_problem} as:
 \begin{align*}
@@
 $\forall w_\ell \in \mathcal{D}_k \setminus \{w_{uv}\}\,, {s_{uv}}^T w_{uv} \geq  {s_{uv}}^T
-w_\ell$) we also minimize cross-entropy loss commonly used in non-binary classification
+w_\ell$) we also minimize a cross-entropy loss commonly used in non-binary classification
 problem~\autocite[Section 4.3.4]{PRML06}.  Specifically, recall that the goodness of direction
@@
 loss is defined by
-\begin{equation*}
+\begin{equation}
+  \label{eq:edge_cross}
   \mathcal{L}_{\mathrm{cross-entropy}} = -\sum_{u,v \in E} \sum_{l=1}^k
   \Ind{w_{uv} = w_\ell} p_{uv,\ell}
-\end{equation*}
+\end{equation}
 
@@
 \mathcal{L}_{\mathrm{edge}}$ with respect to $X$ using the Adam algorithm~\autocite{Adam15} and
-automatic differentiation~\autocite{autograd15} as implemented by the Pytorch
-package\footnote{\url{http://pytorch.org/}} and projecting the current iterate back to the set of
+automatic differentiation~\autocite{autograd15}\footnote{Both being implemented by the Pytorch
+package: \url{http://pytorch.org/}.}, projecting the current iterate back to the set of
 matrix with unit $L_2$ norm columns $\mathbb{M}^{n\times d}$ at each step.
@@
 experiment with less directions ($k=5$) or more directions ($k=9, d=36$)\footnote{We add one
-dimension to ensure that $d$ remains a multiple of $k$}, with two higher levels of overlap between
+dimension to ensure that $d$ remains a multiple of $k$.}, with two higher levels of overlap between
 the directions ($n_o=6$ and $n_o=12$), with more directions per node ($k_{\mathrm{local}}=4$) and in
@@
 After having assigned a single direction to every edge and generated the user profiles, we can
-measure how those two quantities *agree* with the edge scores. Namely, recall we predict the
+measure how the assignment and the profiles match, in the sense that the assigned direction of every
+edge is the one with the largest goodness. Indeed, recall we predict the
 direction of an edge $(u,v)$ as $\argmax_{\ell \in \rangesk} g(s_{uv}, w_\ell) = {s_{uv}}^T w_\ell$.
@@
 
-First, as we can see in the table below, in the \default{} configuration, we do not always
-generate user profiles leading to a perfect assignment of edge according to their score/goodness. In
+First, as we can see in the table at the end of this paragraph, in the \default{} configuration, we do not always
+generate user profiles leading to a perfect assignment of edge according to their goodness. In
 general this is not always possible. Consider a $k$-regular subgraph, where the $k$ edges incident
@@
 $k_{\mathrm{local}}$ local constraint. Even so, and more pragmatically, whereas those AMI scores
-could be made higher by increasing the coefficient in front of the cross entropy loss, this would
+could be made higher by increasing the coefficient $\lambda$ in~\eqref{eq:edge_cross}, this would
 imply a lower edge goodness. Even with our choice, only a few percent of the edges have a mismatch
-between the directions we assigned them and the one with the highest goodness. Note also that
+between the directions we assigned them and the one with the largest goodness. Note also that
 depending of the parameters of the generation, the AMI is not the same. Therefore, in the following,
 we divide all the AMI scores by this natural score, in order to obtain a normalized measure.
-\newline\vspace{.5\baselineskip}\noindent%
+\newline\vspace{.8\baselineskip}\noindent%
 \begin{tabular}{lccccccc}
@@
 comparable to the other methods. This suggests that the subsequent clustering is not
-adequate\stodo{Maybe it would be better to save the list of atoms or do a SVD at the end and cluster
-from there.}. Regarding the generative parameters, and compared with the default configuration, the
+adequate. Regarding the generative parameters, and compared with the default configuration, the
 problem is easier when the $k$ directions are spread in a larger number of dimension, and more
@@
   \small
-  \caption{Normalized AMI of 5 methods, when generating direction with 7 configurations. For each
+  \caption{Normalized AMI of 5 methods, when generating directions with 7 configurations. For each
     configuration, we generate directions 200 times, and report the mean normalized AMI along with
@@
     \textit{\textcolor{red}{italic red}}. When the difference between a score and the next best one
-    is statistically significant (that is we can confidently reject the hypothesis that the two
-    distribution have the same mean), we furthermore show in parenthesis the supporting
+    is statistically significant (\ie{} when we can confidently reject the hypothesis that the two
+    distributions have the same mean), we furthermore show in parenthesis the supporting
     $p$-value of a paired Student's $t$-test.
@@
 completely recover is not surprising, for in absence of prior information the problem is
-under-constrained. Indeed, while \combined{} and \pqt{} delivers almost the exact same prediction,
+under-constrained. Indeed, while \combined{} and \pqt{} deliver almost the exact same prediction,
 their directions are clearly different.

### 2017-11-02 edge/model.tex
tiny changes and bib data for edges

@@
 them when the other is fixed. Therefore it can be solved via alternating optimization over $P$ and
-$Q$ using standard (projected) gradient descent algorithms, with the following partial derivatives:
+$Q$~\autocite{AlternatingOptim02} using standard (projected) gradient descent algorithms, with the
+following partial derivatives:
 \begin{align*}

### 2017-11-02 edge/signed.tex
tiny changes and bib data for edges

@@
 \end{aside}
+
+Unpublished paper from CMU, give a link between edge types and k-CC: the inner and outer edge
+vectors are given, and assumed to be drawn from two different distributions. The goal is to recover
+the planted node partition using maximum likelihood estimation, which in this case can be formulated
+as \mind{} \pcc{} problem~\autocite{Barnes2016a}

### 2017-11-06 edge/related.tex
write edge related work part

@@
-Stattner, Erick, and Martine Collard. "Link Clustering for Extracting Collaborative Patterns in a Scientific Co-Authored Network.", FAB workshop, ANOSAM 2017
-they use their method of conceptual links
+\section{Related work}
+\label{sec:edge_related}
+
+To the best of our knowledge, there are not many works that directly attempt to classify edges in
+attributed networks. A natural idea is thus to rely on the numerous existing node clustering
+algorithms, by considering the line graph $L(G)$ of $G$. $L(G)$ is a graph whose each of the $|E|$
+nodes corresponds to an edge of $G$, and where two nodes of $L(G)$ are connected whenever their
+corresponding edges share a common node in $G$. \Textcite{LineGraph09} apply a modularity algorithm
+to $L(G)$ is order to partition the edges of $G$, but contrary to us, their final goal is to
+discover an overlapping clustering of the nodes of $G$, and they rely solely on the graph topology.
+In their review of mining social networks, \textcite[Section 2.3]{SemanticMining15} present several
+methods to solve what they call \emph{inferring social ties}. The first one is based on a graphical
+model where each node is an edge. The label of an edge is influenced by the attributes of its
+endpoints, the correlation between the type of neighboring edges and some global constraints. These
+three influences are modeled by pairwise factors. The parameters of the model are learned by
+maximizing the likelihood of the observed label through gradient descent. Those parameters are then
+use to predict the remaining label that further maximize the total
+likelihood~\autocite{graphicalModelTies11}. An extension to this graphical model is to actively
+query the most informative edges, in the case where it is costly to ask social network user to label
+their relationships~\autocite{Zhuang2012}. Another way to leverage additional supervision is to use
+transfer learning~\autocite{transferTiesPred16}. Namely, given two distinct and partially labeled
+graphs, where the fraction of label is much higher for the first graph than the second one, the goal
+is to fully label the edges of the second graph, exploiting the rich information of the first one.
+This requires the attributes to be comparable between the two graphs, and such attributes are
+therefore computed based on topological features. Finally, a last direct approach to labelling edges
+is presented by \textcite{Aggarwal2016a}. They consider that the input graph has three type of
+edges: labeled by $-1$, labeled by $+1$ or unlabeled. In each of the three induced subgraphs, they
+define a similarity between two nodes $u$ and $v$ as the Jaccard coefficient of the two sets of
+edges incident to $u$ and $v$. A weighted combination of these three coefficients is the final
+similarity between $u$ and $v$. Then $S_k(u)$ are the top-$k$ most similar nodes to $u$, and the
+label of the edge $(u,v)$ is chosen as the majority label in $S_k(u)\times S_k(v) \bigcap E$.
+All these works assign a label to the edges of the input graph, but either they need supervision or
+they do not use attributes, and in both cases they provide little interpretability on their results.
+
+Besides social network, the task of predicting the semantic of an edge has also been addressed in
+the context of knowledge graphs. In such graphs, nodes are abstract concepts and concrete entities
+from the real world, and edges are directed predicates representing facts connecting two
+entities~\autocite{KnowledgeGraphSurvey16}. A typical example is an edge $(h, \ell, t) =
+($\textsf{Donald J. Trump}, \textsf{President of}, \textsf{USA}$)$\footnote{At least at the time of
+writing} where the $(h, \ell, t)$ notation stands respectively for head, type of relation and tail.
+This illustrates some differences with our setting. First, there might be a very large number of
+type of edges in knowledge graphs, since they cover a domain as large as possible. Second, this
+coverage makes it difficult to describe directly entities by a consistent set of $d$ attributes, and
+one usually has to rely on low dimensional embeddings. \autocite{transE13} is primarily concerned
+with learning such embeddings in $\Rbb^k$ for both nodes and relations, that are denoted by the
+boldface letters $\bm{h}$, $\bm{\ell}$, $\bm{t}$. The intuition, similar to word vector, is that
+relations can be modeled by translations, so the existence of an $(h,l,t)$ edge offers evidence for
+$\bm{h} + \bm{\ell} \approx \bm{l}$. Given a training set of existing triplet $S$ and a
+dissimilarity measure $D$, the authors learn embeddings in a fashion reminiscent of metric
+learning~\autocite{MetricLearning15}. Namely, they create negative examples by corrupting at random
+training examples and then minimize with a margin the positive part of the difference between
+dissimilarity of positive and negative examples. More generally, we refer the interested reader to a
+recent survey on embedding knowledge graphs~\autocite{KnowledgeGraphSurvey17}, which among other
+ideas present the interesting notion of composing relation along paths between two entities. We note
+that, in contrast with us, those methods require some form of supervision when it comes to predict
+edge types.  One way to avoid supervision is to rely on a large amount of unstructured text, extract
+entities from this corpus, map entities and possible relations to a knowledge base and finally embed
+them in a low dimensional space using text features such as POS tags~\autocite{Ren2017}.
+
+% One can also graph convolutional network to learn such embeddings~\autocite{Schlichtkrull2017}
+
+A related idea is embedding edges in general graphs, not necessarily in the context of knowledge
+graphs. This is the counterpart of node embeddings techniques we mentioned in
+\autoref{sec:troll_related}. The most straightforward way is to keep track of edges instead of nodes
+when performing random walks, learn an embedding from this corpus using word2vec-like methods and
+then cluster these representations using $k$-means~\autocite{edgerep17}. We cannot directly compare
+this with our work since there is no attribute and thus the clustering is not interpretable (indeed
+the goal of \autocite{edgerep17} is eventually to find an overlapping clustering of the nodes, not
+the edges). More closely related to our work, \textcite{ahmed2017roles} seek to assign a role to the
+edges of a graph. First they learn edge features as follow: starting from topological features, they
+use combination operators (like max, min, mean, sum, product and so on) to iteratively learn higher
+order features while pruning those that are correlated in order to avoid a combinatorial explosion.
+Once they obtain such a $d$-dimensional feature vector for each edge, they stack them in a matrix $S
+\in \Rbb^{|E|\times d}$ and look for $U^{|E| \times r}$ and $B^{r \times d}$ such that the distance
+between $S$ and $UV^T$ is minimized. $V$ is a description of $r$ roles in terms of features while
+$U$ is the mixed membership of each edge. Finally the authors choose the number of roles $r$ using
+the Minimum Description Length principle. While their approach scales linearly with the number of
+edges and is partially parallelizable, the process in which edge feature are learned does not lend
+itself easily to interpretation.
+% Learn a function that take as input the embedding of two nodes and output the predicted strength of
+% the relationship~\autocite{EdgeAwareEmbedding17}, could it be modified to output a cluster index?
+% (I think this is too far actually, sure they learn this edge function, but their motivations are
+% edge asymmetry and model size, working on non attributed, non-typed graph. so leave it out)
+%\autocites[section 2.5]{representationLearning17}{ahmed2017roles}{Ahmed2016}
+
+Another direction is to see the problem as clustering the similarity vectors $s_{uv}$. Because of
+our bias that nodes are connected through partial homophily and heterophily, this can more precisely
+be casted as a subspace clustering problem~\autocite{SCSurvey11}. Namely, we are given a set of $m$
+points in $\Rbb^d$, with the assumption they belong to the union of affine subspaces of unknown
+dimension. The goal is to recover the number $K$ of such subspaces, their dimension, their
+parameters and the assignment of the $m$ point to those subspace. If $K$ is known, one can use
+iterative, $k$-mean like methods (reminiscent of our \lloyd{} method). Otherwise it also possible to
+factorize the matrix of all the data points and interpret as a similarity matrix, use iterative
+statistical approaches such as random sample consensus or build a similarity graph and apply
+spectral clustering algorithms to it. Finally, it is possible to leverage sparsity assumptions and
+express each point as a sparse linear combination of other points, and use these coefficients as
+weights of a similarity graph, which is then clustered into subspaces by a spectral
+method~\autocite{SparseSC13}.
+While subspace clustering comes with information theoretic guarantees of its ability to retrieve
+optimal subspaces, when formulating directly over the $\{s_{uv}\}_{(u,v) \in E}$, it makes no use of
+graph topology, as we consider each edge independently of all others. As an example of applying
+related ideas to graph data, \textcite{Huang2015} build a $d$-dimensional grid of the attributes
+space and use subspace clustering to find cells that have low entropy and high connectivity, before
+merging such cells into node clusters.
+% \textcite{subgraphEdgeLabels13} propose a method to extract subgraphs that are dense across
+% several edge types. (was updated in \autocite{Boden2017}
+
+\Textcite{Taskar04} represent a node-attributed graphs with partially labelled edges as a
+probabilistic graphical model (called Relational Markov Model) and learns its parameters from the
+data using gradient descent. It is a flexible modeling approach that let the designer express
+complex graph patterns. Using a slightly less principled approach, \textcite{conceptualLinks12}
+first cluster nodes based solely on their attributes and then count the links between such groups to
+keep only the maximally frequent ones, arguing this reveals the set of attributes that support the
+connection between two nodes. In the same data mining vein, and related to our departure from global
+homophily, \textcite{beyondHome16} look for link patterns whose support and confidence is not only
+above some given thresholds, but that also diverge from homophily. In the author explicitly exclude
+attributes taking the same value within such link patterns.
 
@@
 links between two groups is then evaluated and when the frequency is greater than a given support
-threshold $\beta$, we talk about frequent conceptual link. Note that we only keep maximal frequent conceptual links (mfcl), i.e. those who are not included in others.
-Finally, the set of maximal frequent conceptual links is used to create a new network structure called a conceptual view which summarizes all the knowledge extracted from the initial network.
-In a conceptual view, each node corresponds to a set of attributes (in this context we call them Meta-nodes) and a link corresponds to maximal frequent conceptual links.
-A formal presentation of frequent conceptual links can be found in [18].
+threshold $\beta$, we talk about frequent conceptual link. Note that we only keep maximal frequent
+conceptual links (mfcl), i.e. those who are not included in others.  Finally, the set of maximal
+frequent conceptual links is used to create a new network structure called a conceptual view which
+summarizes all the knowledge extracted from the initial network. In a conceptual view, each node
+corresponds to a set of attributes (in this context we call them Meta-nodes) and a link corresponds
+to maximal frequent conceptual links~\autocite{conceptualLinks12}.
+The objective of \textcite{ZhangModelFree16} is eventually to hard cluster the nodes into
+communities but their work similarities with ours. Namely, they tweak the traditional modularity
+objective (maximizing the density of intra community edges) by weighting the edges with the
+similarity of their endpoints, and simultaneously learning a weight vector $w_\ell$ for a community
+$C_\ell$. Formally they defined the $R$ criterion as :
+\begin{equation*}
+  R(\mathcal{C}, \beta;\alpha, \mu) = \sum_{\ell=1}^k \frac{1}{|\mathcal{C}_\ell|^\alpha}
+  \sum_{u,v \in \mathcal{C}_\ell} A_{uv} \mu -e^{-g(s_{uv}, w_\ell)}
+\end{equation*}
+where $A$ is the adjacency matrix, $\alpha$ penalizes unbalanced communities, $\mu$ is a trade off
+parameter between information from edges and node features, and $g$ is exactly our goodness
+function.  They optimize it by alternately optimizing over the labels with fixed parameters and over
+the parameters with fixed labels, using block coordinate descent.
+Finally we mention a more quantitative work. \Textcite{Abraham2012a} assume that there are $K$
+social categories modelled by $K$ Euclidean spaces $\mathcal{D}_i$. Nodes of the graph have an
+associated point in each of these spaces. The key assumption is that space have small local
+correlation: informally, the intersection of any two small balls from two distinct spaces is small.
+These $K$ spaces give rise to small world networks $\mathcal{G}_i$, where the edge probability
+$\propto \mathcal{D}_i(u, v)^{-d}$ and we observe the real network $\mathcal{G} =\bigcup_i
+\mathcal{G}_i$. From $\mathcal{G}$, the proposed algorithm recovers in $n \mathrm{polylog} n$ time a
+bounded approximation $\mathcal{D}'_i$ of all $\mathcal{D}_i$, that is
+\begin{equation*}
+  \sigma \mathcal{D}_i(u, v) \leq \mathcal{D}'_i(u, v) \leq \delta \mathcal{D}_i(u, v) + \Delta
+\end{equation*}
+
+\iffalse
+adhoc~\autocites{conceptualLinks12}{Abraham2012a}{Taskar04}.
+A supervised probabilistic model
+find a partition maximizing the density of intra community edges ---weighted by the endpoints
+profile similarities and a learned community weight vector~\autocite{ZhangModelFree16}
+
+Let $A$ be the unweighted adjacency matrix and  $\mathcal{E} = \bigcup_{i=1}^K
+\mathcal{E}_k$ be a partition of the node, which alternatively is represented by
+the community memberships vector $e$.  They start by presenting a simple
+community goodness criterion, which like modularity basically score a partition
+based on the number (or density) of edges within communities
+$$ R(e;\alpha) = \sum_{k=1}^K \frac{1}{|\mathcal{E}_k|^\alpha} \sum_{i,j\in \mathcal{E}_k}A_{ij} $$
+where $\alpha$ is a tuning parameters which penalizes unbalanced communities as
+it gets larger.
+The idea would be to add weight on the edges based on the node features
+$f_i\in \mathbb{R}^p$ so as to increase that score. Because features and
+communities don't correlate perfectly, they introduce a more flexible criterion
+$$ R(e,\beta;\alpha,w_n) = \sum_{k=1}^K \frac{1}{|\mathcal{E}_k|^\alpha}
+\sum_{i,j\in \mathcal{E}_k}A_{ij}W(f_i, f_j,\beta_k;w_n) $$
+where
+$\beta_k\in \mathbb{R}^p$ is a weight vector for the $k$th community and $w_n$
+is trade off parameters between information from edges and node features.
+Basically, $W(f_i, f_j,\beta_k;w_n)$ is a function of the similarity
+$\phi_{ij}=\phi(f_i, f_j)$ between two nodes. This reminds of our approach, by
+assigning one vector $w$ to each community. As a practical example, they use
+$$
+w_{ijk} = W(f_i, f_j,\beta_k;w_n) = w_n -e^{-\langle \phi_{ij},\beta_k\rangle}
+$$
+
+This \enquote{criterion needs to be optimized over both the community
+assignments $e$ and the feature parameters $\beta$. Using block coordinate
+descent, we optimize it by alternately optimizing over the labels with fixed
+parameters and over the parameters with fixed labels, and iterating until
+convergence}. Again, this is similar to what we tried at first.
+
+\fi
+
+\bigskip
+
+Although there has not been so many work on clustering edges in node attributed networks, there is a
+wealth of papers on clustering nodes in such networks. Since this is not exactly our topic, we only
+mention a selection of the most relevant to our objective (especially the overlapping ones), and
+refer the interested reader to the survey of
+\textcite{surveyAttributedClustering15}\footnote{Especially the section 2 dealing with edge labeled
+graphs.}. We also reiterate the warning given in the introduction that in large graphs, node
+attributes are not necessarily aligned with annotated communities. Such attributes thus need to be
+used \emph{in addition} to topological information. This is further covered in~\autocite[Section
+3.4]{Fortunato2016}.
+
+An early attempt is the \textsf{SA-Cluster} algorithm of \textcite{Zhou2009} that partitions the
+nodes of a graph based on a distance combining nodes structural and attributes similarities, along
+with its faster incremental version~\autocite{Zhou2010}. Nodes in the same cluster are well
+connected and have a set of similar attributes. The balance between these two objectives was further
+studied in a setting where it is tunable by the user of the algorithm~\textcite{Baroni2017}. In a
+more \enquote{data mining fashion}, the same problem can also be worded as finding subgraphs induced
+by a set of attributes that are more dense than what one would expect in a null
+model~\autocite{Silva2012}. Instead of detecting all such subgraphs or communities, one can adopt a
+query-based approach. Namely, given a query node $u$ and a set $S$ of attributes, find all the
+connected subgraphs containing $u$ that are both tightly connected and share enough common
+attributes~\autocite{AttributedCommunity16}. In that spirit of focusing on a given node,
+\textcite{LeskovecEgo12} introduce a generative model for ego networks in social networks where the
+neighbors of a node $u$ can belong to $k$ categories (such as family, colleagues, school friends).
+Using our notations, those $k$ categories (or circle) $\{C_i\}_{i=1}^k$ of a node $u$ are defined by
+a vector $w_i$ and a weight $\alpha_i$. The probability $p(e)$ of an edge $e=(u,v)$ favors the
+presence of edges with high goodness within circles:
+\begin{equation}
+  \label{eq:edge_ego}
+  p((u, v)=e \in E) \propto \exp \left(
+    \sum_{C_i  \supseteq {u,v}} g(s_{uv}, w_i) -
+    \sum_{C_i \nsupseteq {u,v}} \alpha_i g(s_{uv}, w_i)
+  \right)% = \exp\left(\Phi(e)\right)
+\end{equation}
+The unsupervised problem of maximizing the likelihood of observing the input graph according to this
+probability distribution is solved by alternating two steps: assigning nodes to circles and
+optimizing the parameters $\{w_i, \alpha_i\}_{i=1}^k$ given a circle assignment. Similar to us, the
+vector $w_i$ explain why some nodes belong to the \ith{} circle. However, it is not obvious how to
+transfer this knowledge to the edges themselves.
+
+In fact, this generative model approach has proved very fruitful when it comes to community
+detection in node attributed graphs~\autocites{Xu2014}{Zhao2017}{Yang2013}{Kataoka2016}{Weng2016}{Newman2016}.
+The general idea is the following. First, design a model to generate some of the following aspects:
+the nodes attributes, the topology of the graph and the community membership of the nodes. Then,
+infer the parameters that maximize the likelihood of observing the current graph. Finally, extract
+from these parameters community membership for every nodes. We give a very succinct description of
+the generative models of some representative recent works in \autoref{tab:edge_genmodel}. Seen at
+the light of our \autoref{p:edge_full}, given the overlapping membership obtained by inference of a
+generative model, and an edge $(u,v)$, what could look at the highest share community coefficient
+between $u$ and $v$ and use that to explain the edge (among $\frac{K(K-1)}{2}$ edge type). However,
+this is clearly an ad-hod post processing, as indeed these works are concerned with node and not
+edge clustering.
+
+\setlength{\fullpage}{\textwidth+\marginparsep+\marginparwidth}
+\begin{table}[htpb]
+  \centering
+  \small
+  \caption{We summarize how each model generates: 1) the membership of a node $u$ to a community $c
+  \in \rangesk$, 2) the attributes of $u$ knowing its community membership and 3) the edges between
+nodes.}
+  \label{tab:edge_genmodel}
+  \bgroup
+  \def\arraystretch{1.5}
+  \begin{tabulary}{\fullpage}{LLLL}
+    \toprule
+    ref & community & attributes & links \\
+    \midrule
+
+    \autocite{Xu2014} &
+    single community drawn from a multinomial &
+    drawn from a distribution parametrized by the node community &
+    drawn from a distribution parametrized by the pair of endpoints communities \\
+
+    \autocite{Yang2013} &
+    intensity of membership $F_{uc} \in [0, \infty)$ are given &
+    logistic function of $F_{u1},\ldots,F_{uC}$ &
+    sharing more communities makes link more likely:
+    $P_{uv} = 1-\exp(-\sum_c F_{uc}F_{vc})$ \\
+
+    \autocite{Kataoka2016} &
+    multinomial of dimension $k$ &
+    drawn from one normal distribution per community &
+    Stochastic Block Model (SBM), where blocks are identified with communities \\
+
+    \midrule
+
+    \autocite{Zhao2017} &
+    gamma distribution, whose parameters depends of the nodes attribute &
+    given binary attributes &
+    SBM, with the block matrix drawn from a hierarchical relational gamma process \\
+
+    \autocites{LeskovecEgo12}{LeskovecEgo14} &
+    overlapping circle defined by $\theta_k$, but not generated &
+    given &
+    higher probability of appearing within common circle \eqref{eq:edge_ego} \\
+
+    \autocite{Weng2016} &
+    logistic function of the attributes &
+    given &
+    SBM \\
+
+    \autocite{Newman2016} &
+    one multinomial for each discrete value of the single attribute &
+    given &
+    degree corrected SBM \\
+    \bottomrule
+  \end{tabulary}
+  \egroup
+\end{table}
+
+We also point to the work of~\textcite{DeBacco2017}, which despite not considering node attributes,
+present a model with interesting applications. Specifically, the nodes have mixed membership to $K$
+overlapping groups and each $L$ layer is generated by a specific $K\times K$ block matrix, taking
+into account the group membership, like in a degree corrected SBM. Unlike us, this allows for
+multigraphs, although the authors assume for simplicity that this does not happen. Once its
+parameters are found, this model can then be used to predict the existence of extra edges in each
+layer, which can be seen as edge type prediction. Furthermore, measuring the extent to which one
+layer helps predict links in another layer provides a way to measure the relationships between the
+layers, from redundancy to complete independence, allowing information compression.
+
+\iffalse
+\begin{itemize}[leftmargin=*]
+\item \href{http://dl.acm.org/citation.cfm?id=2629616}{A General Bayesian
+  Framework for Attributed Graph Clustering}~\autocite{Xu2014}
+
+They present a generative model of clustered attributed graph. That is, given a
+vertex set $V$, \enquote{a joint probability distribution over the space of all
+possible attributed graphs and all possible partitions over $V$. The
+clustering with the highest probability is this clustering best explains the
+attribute values and edge patterns of the given graph.}
+
+Let $X$ be the $n\times n$ adjacency matrix, $Y$ be the $n\times T$ attribute
+matrix and $Z$ be the $n$-long cluster indicator vector (taking values in
+${1,\ldots,k}$). The generative process is as follow, for each node $v_i\in V$:
+
+\begin{itemize}
+\item
+  Choose the cluster according to a multinomial distribution:
+  $Z_i \sim \mathtt{Multinomial}(\alpha)$. That is
+  $P(Z_i = k|\alpha) = \alpha_k$ with $\alpha_k \in [0,1]$ and
+  $\sum_k \alpha_k = 1$.
+\item
+  Then sample the attribute of the node, from a distribution
+  parametrized according to the chosen cluster
+  $Y_i \sim P(Y_i |\theta_{Z_i}) = \prod_{t=1}^{T} P(Y_{it}|\theta_{Z_{i},t})$.
+  The exact form of $P$ depends of the kind of attribute being
+  modelled (binary, categorical, continuous and so on).
+\item
+  Finally, we potentially connect the vertex $v_i$ to the previous
+  ones ${v_j, j<i}$ according to
+  $X_{ij} \sim P(X_{ij} |\Phi_{Z_iZ_j})$. The intuition is that two
+  vertices from the same cluster will have similar connectivity pattern
+  with the rest of the graph.
+\end{itemize}
+
+In Section 3.2.2, the authors also described the prior placed on $\alpha$,
+$\theta$ and $\Phi$. Note that contrary to the previous model (and because the
+focus is on clustering) the edge probabilities only depend indirectly of the
+attributes (through the $Z$ variable).
+
+They use a variational approach to perform the parameters inference.
+
+TODO: add this one \fullcite{Zhao2017}
+
+\item \href{https://cs.stanford.edu/people/jure/pubs/cesna-icdm13.pdf}{Community
+  Detection in Networks with Node Attributes}~\autocite{Yang2013}
+
+CESNA is based on the following properties:
+
+\begin{quote}
+\begin{itemize}[leftmargin=*,nosep]
+\item
+  ``Nodes that belong to the same communities are likely to be connected
+  to each other.
+\item
+  Communities can overlap, as individual nodes may belong to multiple
+  communities.
+\item
+  If two nodes belong to multiple common communities, they are more
+  likely to be connected than if they share only a single common
+  community (i.e., overlapping communities are denser).
+\item
+  Nodes in the same community are likely to share common attributes ---
+  for example, a community might consist of friends attending a same
+  school.''
+\end{itemize}
+\end{quote}
+
+The graph $G$ has $C$ communities and $n$ nodes, each $u$ of
+them having $K$ binary attributes in its profile vector $X_u$ and
+belonging to a community $c$ with intensity
+$F_{uc} \in [0, \infty)$.
+
+Link generation is done by considering that the probability of linking
+two nodes $u$ and $v$ belonging to the same community $c$ is
+$P_{uv}(c) = 1-\exp(F_{uc}F_{vc})$ (which is $0$ is one node does
+not belong to $c$). Then, assuming that each community connects $u$
+and $v$ independently, we have that
+$1-P_{uv} = \prod_c (1-P_{uv}(c)) = \exp(-\sum_c F_{uc}F_{vc})$ from
+which $P_{uv} = 1-\exp(-\sum_c F_{uc}F_{vc})$.
+
+\begin{minipage}[t]{0.6\linewidth}
+Likewise community memberships generate attributes. The intuition is
+that given its community memberships, it should be possible to predict a
+node's attributes. That is, considering $F_{u1},\ldots,F_{uC}$ as
+input feature of a logistic model, and given the logistic weight
+$W_{kc}$ (for attribute $k$ and community $c$), we have
+\[Q_{uk} = \frac{1}{1+\exp(-\sum_c W_{kc}F_{uc})}\] and
+$X_{uk} \sim \mathrm{Bernoulli}(Q_{uk})$.
+\end{minipage}
+\begin{minipage}[t]{0.37\linewidth}
+  \centering
+  \vspace{-\ht\strutbox}
+  \includegraphics[width=1\linewidth]{raw/cesna.png}
+\end{minipage}
+
+\item add this one (not a generative model) \fullcite{Chesnokov2017}
+
+\item \href{https://arxiv.org/abs/1608.00920}{Community Detection Algorithm
+  Combining Stochastic Block Model and Attribute Data Clustering}~\autocite{Kataoka2016}
+
+That one is familiar. First they also made the assumption that the adjacency
+matrix is conditionally independent of attribute data given the community
+assignment. Then for a given node $i$, first they pick its community $x_i$ from
+a multinomial parametrized by $\gamma$.  Then its (one dimensional continuous)
+attribute is drawn from the normal distribution of the community $\ell$:
+$\mathcal{N}(\mu_\ell, \sigma_\ell)$. Finally the edges are created by a
+Stochastic Block Model.
+
+\item \href{https://arxiv.org/abs/1610.09735}{Community Detection with Nodal
+  Information}~\autocite{Weng2016}
+
+Here $A$ is the adjacency matrix, $X$ the attribute matrix and $c$ the community
+assignment. Their first assumption is that $A \bot X | c$, \enquote{a strong
+constraint stating that given the community membership, what nodes are like
+(described by covariates $X$) does not affect how they are connected (encoded
+in $A$)}.
+
+This motivates them to consider the following model \emph{Node-coupled
+Stochastic Block Model (NSBM)}:
+\begin{align*}
+  P(c|X) &= \prod_i \frac{\exp( \beta^T_{c_i} x_i)}{\sum^K_{k =1}\exp( \beta^T_{c_i} x_i)} \\
+  P(A|c) &= \prod_{ i<j} B^{A_{ij}}_{c_i c_j}(1 - B_{c_i c_j})^{(1-A_{ij})}
+\end{align*}
+where $B = ( B_{ab}) \in [0, 1]^{K\times K}$ is symmetric and
+$\beta = ( \beta_1 , \ldots, \beta_K) \in \mathbb{R}^{Kp}$.
+
+\begin{quote}
+\enquote{
+The distribution $P(A|c)$ follows the stochastic block model (SBM), which, as
+the fundamental model, has been extensively studied in the literature. The SBM
+implies that the distribution of an edge between node $i$ and $j$ only depends
+on their community membership $c_i$ and $c_j$.
+
+The nodes from the same community are stochastically equivalent. The element
+$B_{ab}$ in the matrix $B$ represents the probability of edge connection between
+a node in community $a$ and a node in community $b$.
+
+The $P ( c | X)$ simply takes a multi-logistic regression form, where we will
+assume $\beta_K = 0$ for identifiability.}
+\end{quote}
+
+\item\href{https://arxiv.org/abs/1507.04001}{Structure and Inference in
+  Annotated Networks}~\autocite{Newman2016}
+
+Consider $n$ nodes labeled $u=1,\ldots,n$, divided into $k$ communities ($u$
+belongs to community $s_u$), and associated with one metadata $x_u \in [1, K]$
+(that is a simplified version for introductory purpose, later metadata can also
+be continuous). Given $\{x_u\}_{u=1}^n$ and the degree sequence
+$\{d_u\}_{u=1}^n$, the network is generated as follow. First each node $u$ is
+assigned to community $s$ with a probability $\gamma_{s,x_u}$ (the $k\times K$
+matrix $\Gamma$ holds these parameters). One each node is assigned a community,
+we use a degree corrected block model to create the network.  Edge between $u$
+and $v$ is created with probability $p_{uv} = d_ud_v\theta_{s_u,s_v}$. This
+time, $\theta$ parameters are stored in a $k\times k$ symmetric matrix $\Theta$.
+
+\item \href{http://doi.org/10.1103/PhysRevE.95.042317}{\citefield{DeBacco2017}{title}}
+  \autocite{DeBacco2017}
+
+\end{itemize}
+
+It has been implemented in many ways (make a table?)
+Replace distance measure by a Bayesian model~\autocites{Xu2012}{Xu2014}{Zhao2017}.
+CESNA (see maingraphSC)~\autocite{Yang2013}, overlapping, compare with CESNA so put them together
+\autocite{Chesnokov2017}
+generative models and then node membership inference~\autocites{Kataoka2016}{Weng2016}{Newman2016}
+\fi
+
+% keep checking in mendeley and community from maingraphSC
+%
+% \subsection{Mendeley}
+
+% \begin{description}
+% \item[\citefield{Zhou2009}{title}] -- \fullcite{Zhou2009}\\ \citefield{Zhou2009}{abstract}
+% \item[\citefield{Zhou2010}{title}] -- \fullcite{Zhou2010}\\ \citefield{Zhou2010}{abstract}
+
+% \item[\citefield{Xu2012}{title}] -- \fullcite{Xu2012}\\ \citefield{Xu2012}{abstract}
+% \item[\citefield{Xu2014}{title}] -- \fullcite{Xu2014}\\ \citefield{Xu2014}{abstract}
+
+% \item[\citefield{Silva2012}{title}] -- \fullcite{Silva2012}\\ \citefield{Silva2012}{abstract}
+
+  % too far
+% \item[\citefield{Ruan2013}{title}] -- \fullcite{Ruan2013}\\ \citefield{Ruan2013}{abstract}
+
+% CESNA
+% \item[\citefield{Yang2013}{title}] -- \fullcite{Yang2013}\\ \citefield{Yang2013}{abstract}
+
+% subspace
+% \item[\citefield{Huang2015}{title}] -- \fullcite{Huang2015}\\ \citefield{Huang2015}{abstract}
+
+% Too far an not super clear, I think they do node classification
+% \item[\citefield{Bayer2015}{title}] -- \fullcite{Bayer2015}\\ \citefield{Bayer2015}{abstract}
+
+% \item[\citefield{Kataoka2016}{title}] -- \fullcite{Kataoka2016}\\ \citefield{Kataoka2016}{abstract}
+
+% \item[\citefield{Weng2016}{title}] -- \fullcite{Weng2016}\\ \citefield{Weng2016}{abstract}
+
+% link with k-CC
+% \item[\citefield{Barnes2016a}{title}] -- \fullcite{Barnes2016a}\\ \citefield{Barnes2016a}{abstract}
+
+% \item[\citefield{Newman2016}{title}] -- \fullcite{Newman2016}\\ \citefield{Newman2016}{abstract}
+
+  % Unpublished, benchmarks? (Mason Porter though)
+% \item[\citefield{Bazzi2016}{title}] -- \fullcite{Bazzi2016}\\ \citefield{Bazzi2016}{abstract}
+
+  % KG
+% \item[\citefield{Ren2017}{title}] -- \fullcite{Ren2017}\\ \citefield{Ren2017}{abstract}
+
+  % fun extension, each nodes is itself a single-attribute graph…
+% \item[\citefield{Guo2017}{title}] -- \fullcite{Guo2017}\\ \citefield{Guo2017}{abstract}
+
+% MiMAG mining dense subgraphs featuring a single edge type, so it's more of downstream application
+  % “we aim at finding clusters of vertices that are densely connected by edges with similar labels
+  % in a subset of the graph layers” paper version of
+% \item[\citefield{Boden2017}{title}] -- \fullcite{Boden2017}\\ \citefield{Boden2017}{abstract}
+
+  % Unpublished, quite far anyway
+% \item[\citefield{Wu2017}{title}] -- \fullcite{Wu2017}\\ \citefield{Wu2017}{abstract}
+
+  % share the idea of explaining edges: “Given a collection of attributed subgraphs from different
+  % classes, how can we discover the attributes that characterize their differences?” thing is, it's
+  % about differences so… and there Focus attribute KDD'14 paper is only about node
+% \item[\citefield{Rezaei2017}{title}] -- \fullcite{Rezaei2017}\\ \citefield{Rezaei2017}{abstract}
+
+  % Anomaly detection is a nice application in edge-labeled graphs: Novel distributed algorithms
+  % that identify problematic nodes whose influence can only be detected on their neighbors,
+  % validated through the analysis of data breaches in bank transactions. (except there's no label
+  % here) cite it rather for the tensor stuff
+% \item[\citefield{Araujo2017}{title}] -- \fullcite{Araujo2017}\\ \citefield{Araujo2017}{abstract}
+
+% \item[\citefield{Baroni2017}{title}] -- \fullcite{Baroni2017}\\ \citefield{Baroni2017}{abstract}
+
+  % nice keywords in the abstract: overlapping, disassortative, layer merging for compression,
+  % generative model: the model assigns probabilities to each pair of nodes that they have an
+  % unobserved link of each type. However, it needs supervision
+  % node have mixed membership to K groups and each layer is generated by a K×K matrix given the
+  % density of edges between groups in that specific layer, taking into account the group
+  % membership, like in a degree corrected SBM. Unlike us, this allows for multigraphs, but the EM
+  % procedure presented to recover the model parameters require observing all the layers, that is
+  % full supervision. Once its parameters are found, this model can then be used to predict the
+  % existence of extra edges in each layer, which can be seen as edge type prediction. Furthermore,
+  % “measuring the extent to which one layer helps predict links in another layer provide a
+  % way to measure the relationships between the layers, from redundancy to independence, allowing
+  % information compression.
+% \item[\citefield{DeBacco2017}{title}] -- \fullcite{DeBacco2017}\\ \citefield{DeBacco2017}{abstract}
+
+  % SNE embedding, move it to conclusion
+% \item[\citefield{Liao2017}{title}] -- \fullcite{Liao2017}\\ \citefield{Liao2017}{abstract}
+
+  % Too far, detecting localized attributes on a graph
+% \item[\citefield{Chen2017}{title}] -- \fullcite{Chen2017}\\ \citefield{Chen2017}{abstract}
+
+  %ICML, another generative model with fast, fully Bayesian inference but no edge type.
+% \item[\citefield{Zhao2017}{title}] -- \fullcite{Zhao2017}\\ \citefield{Zhao2017}{abstract}
+
+  % overlapping, compare with CESNA so put them together
+% \item[\citefield{Chesnokov2017}{title}] -- \fullcite{Chesnokov2017}\\ OVERLAPPING \citefield{Chesnokov2017}{abstract}
+
+% \end{description}
+
+% \cite{Zhou2009,Zhou2010,Xu2012,Silva2012,Ruan2013,Yang2013,Xu2014,Huang2015,Bayer2015,BOTHOREL2015,Kataoka2016,Weng2016,Barnes2016a,Newman2016,Bazzi2016,Ren2017,Guo2017,Boden2017,Wu2017,Rezaei2017,Araujo2017,Baroni2017,DeBacco2017,Liao2017,Chen2017,Zhao2017}
+
+\iffalse
 
@@
 look for link pattern with support and confidence above some threshold, that
-also diverge from homophily by excludingi attributes taking the same value (so
+also diverge from homophily by excluding attributes taking the same value (so
 it's not exactly like us once again)
@@
 called \emph{heterogenous information network} in \autocite{HINSurvey17}.
+
+learn a function that take as input the embedding of two nodes and output the predicted strength of
+the relationship~\autocite{EdgeAwareEmbedding17}, could it be modified to output a cluster index?
+\fi

### 2017-11-07 edge/related.tex
remove draft text on edge related

@@
 attributes taking the same value within such link patterns.
-
-Groups of nodes are first created when they have common attributes. The set of links between two
-groups is called a conceptual link in the sense that it corresponds to a relationship between two
-sets of attributes, namely two concepts in the field of formal concept analysis [7]. The number of
-links between two groups is then evaluated and when the frequency is greater than a given support
-threshold $\beta$, we talk about frequent conceptual link. Note that we only keep maximal frequent
-conceptual links (mfcl), i.e. those who are not included in others.  Finally, the set of maximal
-frequent conceptual links is used to create a new network structure called a conceptual view which
-summarizes all the knowledge extracted from the initial network. In a conceptual view, each node
-corresponds to a set of attributes (in this context we call them Meta-nodes) and a link corresponds
-to maximal frequent conceptual links~\autocite{conceptualLinks12}.
 The objective of \textcite{ZhangModelFree16} is eventually to hard cluster the nodes into
@@
 \end{equation*}
-
-\iffalse
-adhoc~\autocites{conceptualLinks12}{Abraham2012a}{Taskar04}.
-A supervised probabilistic model
-find a partition maximizing the density of intra community edges ---weighted by the endpoints
-profile similarities and a learned community weight vector~\autocite{ZhangModelFree16}
-
-Let $A$ be the unweighted adjacency matrix and  $\mathcal{E} = \bigcup_{i=1}^K
-\mathcal{E}_k$ be a partition of the node, which alternatively is represented by
-the community memberships vector $e$.  They start by presenting a simple
-community goodness criterion, which like modularity basically score a partition
-based on the number (or density) of edges within communities
-$$ R(e;\alpha) = \sum_{k=1}^K \frac{1}{|\mathcal{E}_k|^\alpha} \sum_{i,j\in \mathcal{E}_k}A_{ij} $$
-where $\alpha$ is a tuning parameters which penalizes unbalanced communities as
-it gets larger.
-The idea would be to add weight on the edges based on the node features
-$f_i\in \mathbb{R}^p$ so as to increase that score. Because features and
-communities don't correlate perfectly, they introduce a more flexible criterion
-$$ R(e,\beta;\alpha,w_n) = \sum_{k=1}^K \frac{1}{|\mathcal{E}_k|^\alpha}
-\sum_{i,j\in \mathcal{E}_k}A_{ij}W(f_i, f_j,\beta_k;w_n) $$
-where
-$\beta_k\in \mathbb{R}^p$ is a weight vector for the $k$th community and $w_n$
-is trade off parameters between information from edges and node features.
-Basically, $W(f_i, f_j,\beta_k;w_n)$ is a function of the similarity
-$\phi_{ij}=\phi(f_i, f_j)$ between two nodes. This reminds of our approach, by
-assigning one vector $w$ to each community. As a practical example, they use
-$$
-w_{ijk} = W(f_i, f_j,\beta_k;w_n) = w_n -e^{-\langle \phi_{ij},\beta_k\rangle}
-$$
-
-This \enquote{criterion needs to be optimized over both the community
-assignments $e$ and the feature parameters $\beta$. Using block coordinate
-descent, we optimize it by alternately optimizing over the labels with fixed
-parameters and over the parameters with fixed labels, and iterating until
-convergence}. Again, this is similar to what we tried at first.
-
-\fi
 
@@
 layers, from redundancy to complete independence, allowing information compression.
-
-\iffalse
-\begin{itemize}[leftmargin=*]
-\item \href{http://dl.acm.org/citation.cfm?id=2629616}{A General Bayesian
-  Framework for Attributed Graph Clustering}~\autocite{Xu2014}
-
-They present a generative model of clustered attributed graph. That is, given a
-vertex set $V$, \enquote{a joint probability distribution over the space of all
-possible attributed graphs and all possible partitions over $V$. The
-clustering with the highest probability is this clustering best explains the
-attribute values and edge patterns of the given graph.}
-
-Let $X$ be the $n\times n$ adjacency matrix, $Y$ be the $n\times T$ attribute
-matrix and $Z$ be the $n$-long cluster indicator vector (taking values in
-${1,\ldots,k}$). The generative process is as follow, for each node $v_i\in V$:
-
-\begin{itemize}
-\item
-  Choose the cluster according to a multinomial distribution:
-  $Z_i \sim \mathtt{Multinomial}(\alpha)$. That is
-  $P(Z_i = k|\alpha) = \alpha_k$ with $\alpha_k \in [0,1]$ and
-  $\sum_k \alpha_k = 1$.
-\item
-  Then sample the attribute of the node, from a distribution
-  parametrized according to the chosen cluster
-  $Y_i \sim P(Y_i |\theta_{Z_i}) = \prod_{t=1}^{T} P(Y_{it}|\theta_{Z_{i},t})$.
-  The exact form of $P$ depends of the kind of attribute being
-  modelled (binary, categorical, continuous and so on).
-\item
-  Finally, we potentially connect the vertex $v_i$ to the previous
-  ones ${v_j, j<i}$ according to
-  $X_{ij} \sim P(X_{ij} |\Phi_{Z_iZ_j})$. The intuition is that two
-  vertices from the same cluster will have similar connectivity pattern
-  with the rest of the graph.
-\end{itemize}
-
-In Section 3.2.2, the authors also described the prior placed on $\alpha$,
-$\theta$ and $\Phi$. Note that contrary to the previous model (and because the
-focus is on clustering) the edge probabilities only depend indirectly of the
-attributes (through the $Z$ variable).
-
-They use a variational approach to perform the parameters inference.
-
-TODO: add this one \fullcite{Zhao2017}
-
-\item \href{https://cs.stanford.edu/people/jure/pubs/cesna-icdm13.pdf}{Community
-  Detection in Networks with Node Attributes}~\autocite{Yang2013}
-
-CESNA is based on the following properties:
-
-\begin{quote}
-\begin{itemize}[leftmargin=*,nosep]
-\item
-  ``Nodes that belong to the same communities are likely to be connected
-  to each other.
-\item
-  Communities can overlap, as individual nodes may belong to multiple
-  communities.
-\item
-  If two nodes belong to multiple common communities, they are more
-  likely to be connected than if they share only a single common
-  community (i.e., overlapping communities are denser).
-\item
-  Nodes in the same community are likely to share common attributes ---
-  for example, a community might consist of friends attending a same
-  school.''
-\end{itemize}
-\end{quote}
-
-The graph $G$ has $C$ communities and $n$ nodes, each $u$ of
-them having $K$ binary attributes in its profile vector $X_u$ and
-belonging to a community $c$ with intensity
-$F_{uc} \in [0, \infty)$.
-
-Link generation is done by considering that the probability of linking
-two nodes $u$ and $v$ belonging to the same community $c$ is
-$P_{uv}(c) = 1-\exp(F_{uc}F_{vc})$ (which is $0$ is one node does
-not belong to $c$). Then, assuming that each community connects $u$
-and $v$ independently, we have that
-$1-P_{uv} = \prod_c (1-P_{uv}(c)) = \exp(-\sum_c F_{uc}F_{vc})$ from
-which $P_{uv} = 1-\exp(-\sum_c F_{uc}F_{vc})$.
-
-\begin{minipage}[t]{0.6\linewidth}
-Likewise community memberships generate attributes. The intuition is
-that given its community memberships, it should be possible to predict a
-node's attributes. That is, considering $F_{u1},\ldots,F_{uC}$ as
-input feature of a logistic model, and given the logistic weight
-$W_{kc}$ (for attribute $k$ and community $c$), we have
-\[Q_{uk} = \frac{1}{1+\exp(-\sum_c W_{kc}F_{uc})}\] and
-$X_{uk} \sim \mathrm{Bernoulli}(Q_{uk})$.
-\end{minipage}
-\begin{minipage}[t]{0.37\linewidth}
-  \centering
-  \vspace{-\ht\strutbox}
-  \includegraphics[width=1\linewidth]{raw/cesna.png}
-\end{minipage}
-
-\item add this one (not a generative model) \fullcite{Chesnokov2017}
-
-\item \href{https://arxiv.org/abs/1608.00920}{Community Detection Algorithm
-  Combining Stochastic Block Model and Attribute Data Clustering}~\autocite{Kataoka2016}
-
-That one is familiar. First they also made the assumption that the adjacency
-matrix is conditionally independent of attribute data given the community
-assignment. Then for a given node $i$, first they pick its community $x_i$ from
-a multinomial parametrized by $\gamma$.  Then its (one dimensional continuous)
-attribute is drawn from the normal distribution of the community $\ell$:
-$\mathcal{N}(\mu_\ell, \sigma_\ell)$. Finally the edges are created by a
-Stochastic Block Model.
-
-\item \href{https://arxiv.org/abs/1610.09735}{Community Detection with Nodal
-  Information}~\autocite{Weng2016}
-
-Here $A$ is the adjacency matrix, $X$ the attribute matrix and $c$ the community
-assignment. Their first assumption is that $A \bot X | c$, \enquote{a strong
-constraint stating that given the community membership, what nodes are like
-(described by covariates $X$) does not affect how they are connected (encoded
-in $A$)}.
-
-This motivates them to consider the following model \emph{Node-coupled
-Stochastic Block Model (NSBM)}:
-\begin{align*}
-  P(c|X) &= \prod_i \frac{\exp( \beta^T_{c_i} x_i)}{\sum^K_{k =1}\exp( \beta^T_{c_i} x_i)} \\
-  P(A|c) &= \prod_{ i<j} B^{A_{ij}}_{c_i c_j}(1 - B_{c_i c_j})^{(1-A_{ij})}
-\end{align*}
-where $B = ( B_{ab}) \in [0, 1]^{K\times K}$ is symmetric and
-$\beta = ( \beta_1 , \ldots, \beta_K) \in \mathbb{R}^{Kp}$.
-
-\begin{quote}
-\enquote{
-The distribution $P(A|c)$ follows the stochastic block model (SBM), which, as
-the fundamental model, has been extensively studied in the literature. The SBM
-implies that the distribution of an edge between node $i$ and $j$ only depends
-on their community membership $c_i$ and $c_j$.
-
-The nodes from the same community are stochastically equivalent. The element
-$B_{ab}$ in the matrix $B$ represents the probability of edge connection between
-a node in community $a$ and a node in community $b$.
-
-The $P ( c | X)$ simply takes a multi-logistic regression form, where we will
-assume $\beta_K = 0$ for identifiability.}
-\end{quote}
-
-\item\href{https://arxiv.org/abs/1507.04001}{Structure and Inference in
-  Annotated Networks}~\autocite{Newman2016}
-
-Consider $n$ nodes labeled $u=1,\ldots,n$, divided into $k$ communities ($u$
-belongs to community $s_u$), and associated with one metadata $x_u \in [1, K]$
-(that is a simplified version for introductory purpose, later metadata can also
-be continuous). Given $\{x_u\}_{u=1}^n$ and the degree sequence
-$\{d_u\}_{u=1}^n$, the network is generated as follow. First each node $u$ is
-assigned to community $s$ with a probability $\gamma_{s,x_u}$ (the $k\times K$
-matrix $\Gamma$ holds these parameters). One each node is assigned a community,
-we use a degree corrected block model to create the network.  Edge between $u$
-and $v$ is created with probability $p_{uv} = d_ud_v\theta_{s_u,s_v}$. This
-time, $\theta$ parameters are stored in a $k\times k$ symmetric matrix $\Theta$.
-
-\item \href{http://doi.org/10.1103/PhysRevE.95.042317}{\citefield{DeBacco2017}{title}}
-  \autocite{DeBacco2017}
-
-\end{itemize}
-
-It has been implemented in many ways (make a table?)
-Replace distance measure by a Bayesian model~\autocites{Xu2012}{Xu2014}{Zhao2017}.
-CESNA (see maingraphSC)~\autocite{Yang2013}, overlapping, compare with CESNA so put them together
-\autocite{Chesnokov2017}
-generative models and then node membership inference~\autocites{Kataoka2016}{Weng2016}{Newman2016}
-\fi
-
-% keep checking in mendeley and community from maingraphSC
-%
-% \subsection{Mendeley}
-
-% \begin{description}
-% \item[\citefield{Zhou2009}{title}] -- \fullcite{Zhou2009}\\ \citefield{Zhou2009}{abstract}
-% \item[\citefield{Zhou2010}{title}] -- \fullcite{Zhou2010}\\ \citefield{Zhou2010}{abstract}
-
-% \item[\citefield{Xu2012}{title}] -- \fullcite{Xu2012}\\ \citefield{Xu2012}{abstract}
-% \item[\citefield{Xu2014}{title}] -- \fullcite{Xu2014}\\ \citefield{Xu2014}{abstract}
-
-% \item[\citefield{Silva2012}{title}] -- \fullcite{Silva2012}\\ \citefield{Silva2012}{abstract}
-
-  % too far
-% \item[\citefield{Ruan2013}{title}] -- \fullcite{Ruan2013}\\ \citefield{Ruan2013}{abstract}
-
-% CESNA
-% \item[\citefield{Yang2013}{title}] -- \fullcite{Yang2013}\\ \citefield{Yang2013}{abstract}
-
-% subspace
-% \item[\citefield{Huang2015}{title}] -- \fullcite{Huang2015}\\ \citefield{Huang2015}{abstract}
-
-% Too far an not super clear, I think they do node classification
-% \item[\citefield{Bayer2015}{title}] -- \fullcite{Bayer2015}\\ \citefield{Bayer2015}{abstract}
-
-% \item[\citefield{Kataoka2016}{title}] -- \fullcite{Kataoka2016}\\ \citefield{Kataoka2016}{abstract}
-
-% \item[\citefield{Weng2016}{title}] -- \fullcite{Weng2016}\\ \citefield{Weng2016}{abstract}
-
-% link with k-CC
-% \item[\citefield{Barnes2016a}{title}] -- \fullcite{Barnes2016a}\\ \citefield{Barnes2016a}{abstract}
-
-% \item[\citefield{Newman2016}{title}] -- \fullcite{Newman2016}\\ \citefield{Newman2016}{abstract}
-
-  % Unpublished, benchmarks? (Mason Porter though)
-% \item[\citefield{Bazzi2016}{title}] -- \fullcite{Bazzi2016}\\ \citefield{Bazzi2016}{abstract}
-
-  % KG
-% \item[\citefield{Ren2017}{title}] -- \fullcite{Ren2017}\\ \citefield{Ren2017}{abstract}
-
-  % fun extension, each nodes is itself a single-attribute graph…
-% \item[\citefield{Guo2017}{title}] -- \fullcite{Guo2017}\\ \citefield{Guo2017}{abstract}
-
-% MiMAG mining dense subgraphs featuring a single edge type, so it's more of downstream application
-  % “we aim at finding clusters of vertices that are densely connected by edges with similar labels
-  % in a subset of the graph layers” paper version of
-% \item[\citefield{Boden2017}{title}] -- \fullcite{Boden2017}\\ \citefield{Boden2017}{abstract}
-
-  % Unpublished, quite far anyway
-% \item[\citefield{Wu2017}{title}] -- \fullcite{Wu2017}\\ \citefield{Wu2017}{abstract}
-
-  % share the idea of explaining edges: “Given a collection of attributed subgraphs from different
-  % classes, how can we discover the attributes that characterize their differences?” thing is, it's
-  % about differences so… and there Focus attribute KDD'14 paper is only about node
-% \item[\citefield{Rezaei2017}{title}] -- \fullcite{Rezaei2017}\\ \citefield{Rezaei2017}{abstract}
-
-  % Anomaly detection is a nice application in edge-labeled graphs: Novel distributed algorithms
-  % that identify problematic nodes whose influence can only be detected on their neighbors,
-  % validated through the analysis of data breaches in bank transactions. (except there's no label
-  % here) cite it rather for the tensor stuff
-% \item[\citefield{Araujo2017}{title}] -- \fullcite{Araujo2017}\\ \citefield{Araujo2017}{abstract}
-
-% \item[\citefield{Baroni2017}{title}] -- \fullcite{Baroni2017}\\ \citefield{Baroni2017}{abstract}
-
-  % nice keywords in the abstract: overlapping, disassortative, layer merging for compression,
-  % generative model: the model assigns probabilities to each pair of nodes that they have an
-  % unobserved link of each type. However, it needs supervision
-  % node have mixed membership to K groups and each layer is generated by a K×K matrix given the
-  % density of edges between groups in that specific layer, taking into account the group
-  % membership, like in a degree corrected SBM. Unlike us, this allows for multigraphs, but the EM
-  % procedure presented to recover the model parameters require observing all the layers, that is
-  % full supervision. Once its parameters are found, this model can then be used to predict the
-  % existence of extra edges in each layer, which can be seen as edge type prediction. Furthermore,
-  % “measuring the extent to which one layer helps predict links in another layer provide a
-  % way to measure the relationships between the layers, from redundancy to independence, allowing
-  % information compression.
-% \item[\citefield{DeBacco2017}{title}] -- \fullcite{DeBacco2017}\\ \citefield{DeBacco2017}{abstract}
-
-  % SNE embedding, move it to conclusion
-% \item[\citefield{Liao2017}{title}] -- \fullcite{Liao2017}\\ \citefield{Liao2017}{abstract}
-
-  % Too far, detecting localized attributes on a graph
-% \item[\citefield{Chen2017}{title}] -- \fullcite{Chen2017}\\ \citefield{Chen2017}{abstract}
-
-  %ICML, another generative model with fast, fully Bayesian inference but no edge type.
-% \item[\citefield{Zhao2017}{title}] -- \fullcite{Zhao2017}\\ \citefield{Zhao2017}{abstract}
-
-  % overlapping, compare with CESNA so put them together
-% \item[\citefield{Chesnokov2017}{title}] -- \fullcite{Chesnokov2017}\\ OVERLAPPING \citefield{Chesnokov2017}{abstract}
-
-% \end{description}
-
-% \cite{Zhou2009,Zhou2010,Xu2012,Silva2012,Ruan2013,Yang2013,Xu2014,Huang2015,Bayer2015,BOTHOREL2015,Kataoka2016,Weng2016,Barnes2016a,Newman2016,Bazzi2016,Ren2017,Guo2017,Boden2017,Wu2017,Rezaei2017,Araujo2017,Baroni2017,DeBacco2017,Liao2017,Chen2017,Zhao2017}
-
-\iffalse
-
-\Textcite[section 7.2 edge labelling]{nodeClassif11} mention a cool paper from Darja Krushevskaja
-but I can't find it online.
-
-\fullcite{Zhuang2012} and
-section 2.3 of Semantic Mining of Social Networks (same thing I believe.
-Actually they also talk about a WSDM paper on infering ties using transfer
-learning~\autocite{transferTiesPred16})
-
-\Textcite[Section 4.2.2]{miningSocialTheoriesSurvey14} is about \emph{Social Tie
-Prediction}, but they only mention \autocite{transferTiesPred16} and
-\autocite{Yang2012}.
-
-H. Liang, K. Wang and F. Zhu, "Mining social ties beyond homophily," 2016 IEEE 32nd International Conference on Data Engineering (ICDE), Helsinki, 2016, pp. 421-432.
-\url{http://doi.org/10.1109/ICDE.2016.7498259} very datamining approach, they
-look for link pattern with support and confidence above some threshold, that
-also diverge from homophily by excluding attributes taking the same value (so
-it's not exactly like us once again)
-
-The kind of graphs with talked in this paper is a special case of what is
-called \emph{heterogenous information network} in \autocite{HINSurvey17}.
-
-learn a function that take as input the embedding of two nodes and output the predicted strength of
-the relationship~\autocite{EdgeAwareEmbedding17}, could it be modified to output a cluster index?
-\fi

### 2017-11-13 edge/related.tex
pass over edge related work

@@
 
+Whereas we believe we are the first to study \autoref{p:edge_full}, the prevalence of multilayer
+graphs in real world applications, as well as the wealth of information provided by nodes attributes
+has attracted a lot of attention. We start by presenting methods to classify edges, noting they
+either require supervision or do not consider node attributes. We then discuss works predicting
+relations in knowledge graphs, and more generally embedding edges in low dimensional spaces. Another
+approach we highlight includes combining topology and subspace clustering. We also note how ideas
+such as heterophily, learning $k$ ways to weight edge's endpoint similarity and distinguishing a
+superposition of layers have been used in different contexts. Finally, we conclude by the more
+distant but very rich topic of generative model for attributed graphs and how it enables overlapping
+node clustering.
+
+\paragraph{Edge classification}
+\label{par:erw_direct}
+
 To the best of our knowledge, there are not many works that directly attempt to classify edges in
-attributed networks. A natural idea is thus to rely on the numerous existing node clustering
+attributed networks. A first natural idea is actually to rely on the numerous existing node clustering
 algorithms, by considering the line graph $L(G)$ of $G$. $L(G)$ is a graph whose each of the $|E|$
@@
 corresponding edges share a common node in $G$. \Textcite{LineGraph09} apply a modularity algorithm
-to $L(G)$ is order to partition the edges of $G$, but contrary to us, their final goal is to
+to $L(G)$ is order to partition the edges of $G$ but, contrary to us, their final goal is to
 discover an overlapping clustering of the nodes of $G$, and they rely solely on the graph topology.
@@
 three influences are modeled by pairwise factors. The parameters of the model are learned by
-maximizing the likelihood of the observed label through gradient descent. Those parameters are then
+maximizing the likelihood of the observed labels through gradient descent. Those parameters are then
 use to predict the remaining label that further maximize the total
-likelihood~\autocite{graphicalModelTies11}. An extension to this graphical model is to actively
-query the most informative edges, in the case where it is costly to ask social network user to label
+likelihood~\autocite{graphicalModelTies11}. An extension of this graphical model is to actively
+query the most informative edges, in the case where it is costly to ask social network users to label
 their relationships~\autocite{Zhuang2012}. Another way to leverage additional supervision is to use
 transfer learning~\autocite{transferTiesPred16}. Namely, given two distinct and partially labeled
-graphs, where the fraction of label is much higher for the first graph than the second one, the goal
+graphs, where the fraction of labelled edges is much higher for the first graph than the second one, the goal
 is to fully label the edges of the second graph, exploiting the rich information of the first one.
@@
 edges incident to $u$ and $v$. A weighted combination of these three coefficients is the final
-similarity between $u$ and $v$. Then $S_k(u)$ are the top-$k$ most similar nodes to $u$, and the
-label of the edge $(u,v)$ is chosen as the majority label in $S_k(u)\times S_k(v) \bigcap E$.
-All these works assign a label to the edges of the input graph, but either they need supervision or
-they do not use attributes, and in both cases they provide little interpretability on their results.
+similarity between $u$ and $v$. Then, $S_k(u)$ are the top-$k$ most similar nodes to $u$, and the
+label of the edge $(u,v)$ is chosen as the majority label in $\left( S_k(u)\times S_k(v) \right) \bigcap E$.
+All these works assign a label to the edges of the input graph, but they either require supervision or
+they do not use attributes, and in both cases they provide little interpretability over their results.
+
+\paragraph{Prediction in knowledge graphs}
+\label{par:erw_pred_kg}
 
@@
 This illustrates some differences with our setting. First, there might be a very large number of
-type of edges in knowledge graphs, since they cover a domain as large as possible. Second, this
+types of edge in knowledge graphs, since they cover a domain as large as possible. Second, this
 coverage makes it difficult to describe directly entities by a consistent set of $d$ attributes, and
@@
 with learning such embeddings in $\Rbb^k$ for both nodes and relations, that are denoted by the
-boldface letters $\bm{h}$, $\bm{\ell}$, $\bm{t}$. The intuition, similar to word vector, is that
-relations can be modeled by translations, so the existence of an $(h,l,t)$ edge offers evidence for
-$\bm{h} + \bm{\ell} \approx \bm{l}$. Given a training set of existing triplet $S$ and a
-dissimilarity measure $D$, the authors learn embeddings in a fashion reminiscent of metric
+boldface letters $\bm{h}$, $\bm{\ell}$, $\bm{t}$. Similarly to word embedding, the intuition is that
+relations can be modeled by translations. Thus, the existence of an $(h,l,t)$ edge offers evidence that
+$\bm{h} + \bm{\ell} \approx \bm{l}$. Given a training set of existing triplets $S$ and a
+dissimilarity measure $D$, the authors learn an embedding in a fashion reminiscent of metric
 learning~\autocite{MetricLearning15}. Namely, they create negative examples by corrupting at random
@@
 
+\paragraph{Edge embedding}
+\label{par:erw_embed}
+
 A related idea is embedding edges in general graphs, not necessarily in the context of knowledge
-graphs. This is the counterpart of node embeddings techniques we mentioned in
+graphs. This is the counterpart of node embedding techniques we mentioned in
 \autoref{sec:troll_related}. The most straightforward way is to keep track of edges instead of nodes
@@
 $U$ is the mixed membership of each edge. Finally the authors choose the number of roles $r$ using
-the Minimum Description Length principle. While their approach scales linearly with the number of
-edges and is partially parallelizable, the process in which edge feature are learned does not lend
+the Minimum Description Length principle. Their approach bears some similarities with our \pqt{}
+method, but have the additional properties of scaling linearly with the number of
+edges and being partially parallelizable. However, the process in which edge feature are learned does not lend
 itself easily to interpretation.
@@
 
-Another direction is to see the problem as clustering the similarity vectors $s_{uv}$. Because of
+\paragraph{Subspace clustering}
+\label{par:erw_subs}
+
+Another direction is to see the problem as clustering the similarity vectors $s_{uv}$, like our
+approaches in \autoref{sub:edge_baseline}. Because of
 our bias that nodes are connected through partial homophily and heterophily, this can more precisely
@@
 dimension. The goal is to recover the number $K$ of such subspaces, their dimension, their
-parameters and the assignment of the $m$ point to those subspace. If $K$ is known, one can use
-iterative, $k$-mean like methods (reminiscent of our \lloyd{} method). Otherwise it also possible to
-factorize the matrix of all the data points and interpret as a similarity matrix, use iterative
-statistical approaches such as random sample consensus or build a similarity graph and apply
-spectral clustering algorithms to it. Finally, it is possible to leverage sparsity assumptions and
+parameters and the assignment of the $m$ point to those subspaces. If $K$ is known, one can use
+iterative, $k$-mean like methods (reminiscent of our \lloyd{} method). Otherwise, it also possible to
+factorize the matrix of all the data points and interpret it as a similarity matrix, to use iterative
+statistical approaches such as random sample consensus or to build a similarity graph and apply
+spectral clustering algorithms to it. Finally, it is also possible to leverage sparsity assumptions and
 express each point as a sparse linear combination of other points, and use these coefficients as
@@
 While subspace clustering comes with information theoretic guarantees of its ability to retrieve
-optimal subspaces, when formulating directly over the $\{s_{uv}\}_{(u,v) \in E}$, it makes no use of
+optimal subspaces, when formulated directly over the $\{s_{uv}\}_{(u,v) \in E}$, it makes no use of
 graph topology, as we consider each edge independently of all others. As an example of applying
@@
 
-\Textcite{Taskar04} represent a node-attributed graphs with partially labelled edges as a
+\paragraph{Further similar ideas}
+\label{par:erw_misc}
+
+\Textcite{Taskar04} represent a node-attributed graph with partially labelled edges as a
 probabilistic graphical model (called Relational Markov Model) and learns its parameters from the
@@
 first cluster nodes based solely on their attributes and then count the links between such groups to
-keep only the maximally frequent ones, arguing this reveals the set of attributes that support the
+keep only the maximally frequent ones, arguing this reveals the sets of attributes that support the
 connection between two nodes. In the same data mining vein, and related to our departure from global
-homophily, \textcite{beyondHome16} look for link patterns whose support and confidence is not only
-above some given thresholds, but that also diverge from homophily. In the author explicitly exclude
+homophily, \textcite{beyondHome16} look for link patterns whose support and confidence are not only
+above some given thresholds, but that also diverge from homophily. Indeed, the author explicitly exclude
 attributes taking the same value within such link patterns.
 The objective of \textcite{ZhangModelFree16} is eventually to hard cluster the nodes into
-communities but their work similarities with ours. Namely, they tweak the traditional modularity
+communities but their work shares similarities with ours. Namely, they tweak the traditional modularity
 objective (maximizing the density of intra community edges) by weighting the edges with the
-similarity of their endpoints, and simultaneously learning a weight vector $w_\ell$ for a community
+similarity of their endpoints, and simultaneously learn a weight vector $w_\ell$ for a community
 $C_\ell$. Formally they defined the $R$ criterion as :
@@
   R(\mathcal{C}, \beta;\alpha, \mu) = \sum_{\ell=1}^k \frac{1}{|\mathcal{C}_\ell|^\alpha}
-  \sum_{u,v \in \mathcal{C}_\ell} A_{uv} \mu -e^{-g(s_{uv}, w_\ell)}
+  \sum_{u,v \in \mathcal{C}_\ell} A_{uv} \left( \mu -e^{-g(s(u,v), w_\ell)} \right)
 \end{equation*}
 where $A$ is the adjacency matrix, $\alpha$ penalizes unbalanced communities, $\mu$ is a trade off
-parameter between information from edges and node features, and $g$ is exactly our goodness
-function.  They optimize it by alternately optimizing over the labels with fixed parameters and over
+parameter between information from edges and node features, and $g$ corresponds to our goodness
+function while $s$ is an arbitrary similarity function between nodes.
+They optimize it by alternately optimizing over the labels with fixed parameters and over
 the parameters with fixed labels, using block coordinate descent.
-Finally we mention a more quantitative work. \Textcite{Abraham2012a} assume that there are $K$
+Finally, we mention a more quantitative work. \Textcite{Abraham2012a} assume that there are $K$
 social categories modelled by $K$ Euclidean spaces $\mathcal{D}_i$. Nodes of the graph have an
@@
 These $K$ spaces give rise to small world networks $\mathcal{G}_i$, where the edge probability
-$\propto \mathcal{D}_i(u, v)^{-d}$ and we observe the real network $\mathcal{G} =\bigcup_i
+is proportional to $\mathcal{D}_i(u, v)^{-d}$ and we observe the real network $\mathcal{G} =\bigcup_i
 \mathcal{G}_i$. From $\mathcal{G}$, the proposed algorithm recovers in $n \mathrm{polylog} n$ time a
-bounded approximation $\mathcal{D}'_i$ of all $\mathcal{D}_i$, that is
-\begin{equation*}
-  \sigma \mathcal{D}_i(u, v) \leq \mathcal{D}'_i(u, v) \leq \delta \mathcal{D}_i(u, v) + \Delta
-\end{equation*}
+bounded approximation $\mathcal{D}'_i$ of all $\mathcal{D}_i$, that is there exists positive
+constants $\sigma$, $\delta$ and $\Delta$ such that $
+\sigma \mathcal{D}_i(u, v) \leq \mathcal{D}'_i(u, v) \leq \delta \mathcal{D}_i(u, v) + \Delta$.
 
-\bigskip
+\paragraph{Generative models and node clustering}
+\label{par:erw_genmod}
 
-Although there has not been so many work on clustering edges in node attributed networks, there is a
+Although there has not been so many works on clustering edges in node attributed networks, there is a
 wealth of papers on clustering nodes in such networks. Since this is not exactly our topic, we only
@@
 connected and have a set of similar attributes. The balance between these two objectives was further
-studied in a setting where it is tunable by the user of the algorithm~\textcite{Baroni2017}. In a
+studied in a setting where it is tunable by the user of the algorithm~\autocite{Baroni2017}. In a
 more \enquote{data mining fashion}, the same problem can also be worded as finding subgraphs induced
@@
 query-based approach. Namely, given a query node $u$ and a set $S$ of attributes, find all the
-connected subgraphs containing $u$ that are both tightly connected and share enough common
+subgraphs containing $u$ that are both tightly connected and share enough common
 attributes~\autocite{AttributedCommunity16}. In that spirit of focusing on a given node,
@@
 a vector $w_i$ and a weight $\alpha_i$. The probability $p(e)$ of an edge $e=(u,v)$ favors the
-presence of edges with high goodness within circles:
+presence of edges with high goodness within circles, as expressed by
 \begin{equation}
@@
   p((u, v)=e \in E) \propto \exp \left(
-    \sum_{C_i  \supseteq {u,v}} g(s_{uv}, w_i) -
-    \sum_{C_i \nsupseteq {u,v}} \alpha_i g(s_{uv}, w_i)
-  \right)% = \exp\left(\Phi(e)\right)
+    \sum_{C_i  \supseteq {u,v}} g(s(u,v), w_i) -
+    \sum_{C_i \nsupseteq {u,v}} \alpha_i g(s(u,v), w_i)
+  \right) \,,% = \exp\left(\Phi(e)\right)
 \end{equation}
+where $s$ is an arbitrary similarity function between two node profiles.
 The unsupervised problem of maximizing the likelihood of observing the input graph according to this
@@
 the light of our \autoref{p:edge_full}, given the overlapping membership obtained by inference of a
-generative model, and an edge $(u,v)$, what could look at the highest share community coefficient
-between $u$ and $v$ and use that to explain the edge (among $\frac{K(K-1)}{2}$ edge type). However,
-this is clearly an ad-hod post processing, as indeed these works are concerned with node and not
-edge clustering.
+generative model, and an edge $(u,v)$, one could look at the highest shared community coefficient
+between $u$ and $v$ and use that to explain the edge (among the $\frac{K(K+1)}{2}$ possible edge
+types induced by the $K$ communities and their $\frac{K(K-1)}{2}$ pairs). However,
+this is clearly an ad-hod post processing, as indeed these works are concerned with nodes and not
+edges clustering.
 
@@
     given &
-    higher probability of appearing within common circle \eqref{eq:edge_ego} \\
+    higher probability of appearing within common circle, as given in \eqref{eq:edge_ego} \\
 

### 2017-11-14 edge/experiments.tex
take feedback on edge chapter into account
- rewrite the setting part, giving the definition of s and g right away,
and adding a “prediction” operator
- try to motivate better the | x - b - \sum a⋅w | term in (1.2)
- acknowledge the signed case is only a sanity check (I couldn't find a
closed form solution for the noisy case)
- give a list of methods at the beginning of 1.2
- fix typos and math
- split the generation process in 4 steps and explicitly state its input
and output
- reorganize the results writeup in 4 parts: generation, measures,
initial AMI, commented tables

@@
 \item our \lloyd{}-like iterative refinement,
-\item the \combined{} convex optimization of \eqref{eq:edge_soft},
+\item the \combined{} optimization of the two terms in \eqref{eq:edge_soft},
 \item the \fwa{} convex optimization of \eqref{eq:edge_FW} and
@@
 
-% (that is $\yhat$ is a function from $E$
-% to $\rangesk$ such that $\yhat_{uv}$ is the index of the direction assigned to the edge $(u,v)$).
-
-\subsection{Generation}
+\subsection{Data generation}
 \label{sub:edge_generation}
@@
 our initial experiments, as it is specifically tailored to the model introduced in \autoref{sub:edge_setting}.
-Given a number of directions $k$, a maximum number $k_\mathrm{local}$ of directions
-incident to each node, the dimension of profiles $d$ and a number of node $n$, it return a graph
-$G=(V,E)$, a profile matrix $X\in \Rbb^{n\times d}$, a set of $k$ directions $\mathcal{D}_k$ and an
-assignment from every edge $(u,v)$ to a direction $w_{uv} \in \mathcal{D}_k$ such that in most
-cases, $\forall w_\ell \in \mathcal{D}_k \setminus \{w_{uv}\}\,, {s_{uv}}^T w_{uv} \geq  {s_{uv}}^T
-w_\ell$.
+It takes the following parameters as input:
+\begin{itemize}[nosep]
+  \item the number of nodes $n$,
+  \item the dimension of their profiles $d$,
+  \item the total number of directions $k$,
+  \item the maximum number $k_\mathrm{local}$ of directions incident to each node, and
+  \item an integer $n_o$ controlling to which extent two distinct directions explain common dimensions.
+\end{itemize}
+It then returns:
+\begin{itemize}[nosep]
+  \item a graph topology $G=(V,E)$,
+  \item a profile matrix $X\in \Rbb^{n\times d}$,
+  \item a set of $k$ directions $\mathcal{D}_k \subset \dball$, and
+  \item an assignment from every edge $(u,v)$ to a direction index $y_{uv} \in \rangesk$ such that in
+    most cases, $\mathcal{E}(u,v) = y_{uv}$. In other words, the direction assigned to $(u,v)$ is
+    indeed the one achieving maximum goodness. We will explain later why this is not always the
+    case.
+\end{itemize}
+Our methods essentially proceeds as follow:
+\begin{enumerate}[1),nosep]
+  \item We create an appropriate number of small Erdős-Rényi subgraphs and assign to each node a set of
+    $k_\mathrm{local}$ directions such that most pairs of adjacent nodes have at least one
+    directions in common (as we motivated after having introduced equation \eqref{eq:edge_local_loss}
+    \vpageref{eq:edge_local_loss}).
+  \item We connect the blocks by several pairs of nodes having one direction in common, and pick for
+    each edge $(u,v)$ a direction index $y_{uv} \in \rangesk$ among the ones shared by its endpoint.
+  \item We draw the $k$ directions at random, ensuring they respect the parameter $n_o$.
+  \item Finally, we optimize the profiles so as to simultaneously maximize the edges goodness,
+    minimize the term $\mathcal{L}_{\mathrm{node}}$ of equation \eqref{eq:edge_node_loss} and
+    enforces as much as possible that for every edge $(u,v) \in E$, $\mathcal{E}(u,v) = y_{uv}$.
+\end{enumerate}
+Steps 1 and 2 generate a graph topology $(V, E)$, while steps 3 and 4 generates directions and
+profiles $(\mathcal{D_k}, X)$. Note however that step 3 is independent of the steps 1 and 2, while
+step 4 is dependent of all the previous steps.
+We now describe each of these steps in more details.
 
-We first generate the $k$ unit-norm directions $\mathcal{D}_k=\{w_1, \ldots, w_k\}$, independently
-of the graph. Furthermore, we control the number of coordinates $n_o$ that overlap with each other.
-That is, when $n_o = 0$, each direction has exactly \nicefrac{k}{d} non-zero dimension that are
-disjoint.\footnote{We choose $d$ to be a multiple of $k$.} On the other hand, if for instance $n_o=5$,
-then $5$ of the $d$ dimensions will have two directions with non-zero components on it. Those
-non-zero components are drawn \uar{} from $[-1, 1]$ and the each directions is then normalized. For
-instance, with $d=6$, $k=3$ and $n_o=2$, we could generate the following directions, where the dots
-represents non-zero numbers and the blue ones are overlapping:
-\vspace{-.5\baselineskip}
-\begin{center}
-  \begin{tabular}{ccc}
-    $w_1$ & $w_2$ & $w_3$ \\
-    $\begin{pmatrix}
-      \bullet \\
-      0 \\
-      \bullet \\
-      0 \\
-      0 \\
-      \textcolor{DodgerBlue}{\bullet}
-    \end{pmatrix}$ &
-    $\begin{pmatrix}
-      0 \\
-      \bullet \\
-      0 \\
-      \textcolor{DodgerBlue}{\bullet} \\
-      \bullet \\
-      0
-    \end{pmatrix}$ &
-    $\begin{pmatrix}
-      0 \\
-      0 \\
-      0 \\
-      \bullet \\
-      0 \\
-      \bullet
-    \end{pmatrix}$
-  \end{tabular}
-\end{center}
-
-The next step is to create the graph topology, which we do by assembling small Erdős-Rényi
-subgraphs~\autocites{erdos1959random}{gilbertRG59} that we call \emph{blocks}. Furthermore, driven
-by the topological constraint approximated by \eqref{eq:edge_local_loss} mentioned earlier, we
+\begin{enumerate}[1),leftmargin=*]
+\item
+We first create small Erdős-Rényi
+subgraphs~\autocites{erdos1959random}{gilbertRG59} that we call \emph{blocks}. Then we
 assign to each node a $k_\mathrm{local}$-tuple of
-directions such that two adjacent nodes have \emph{at least} one common direction that can later be
-assigned to the edge between them. For convenience, let us first identify directions with colors and
+directions such that two adjacent nodes have \emph{at least} one direction in common.
+For convenience, let us first identify directions with colors and
 thus call such a $k_\mathrm{local}$-tuple of directions a \emph{palette}. For a given palette $p$,
@@
 $(2,4)$ respectively.} we now describe a simple heuristic. It performs a breath first visit of a subgraph,
-starting from a random node. Upon visiting an uncolored node $u$, it builds the intersection $P$ of
-all the adjacent palettes of the palette of its colored neighbors (\ie{} $P=\bigcap_{v\in \nei(u)}
-\adj(p_v)$). Then it selects \uar{} a palette from $P$, or an arbitrary palette if $P$ is empty (see
-\begin{marginfigure}
+starting from a random node. Upon visiting an uncolored node $u$, it builds a set $P$ of palette
+respecting the colouring constraint. Namely, for every colored neighbor $v$ of $u$, it retrieves the
+adjacen palettes of $v$'s palette and builds their intersection:
+\begin{equation*}
+  P=\bigcap_{v\in \nei(u)} \adj(p_v)\,.
+\end{equation*}
+Then it selects \uar{} a palette from $P$, or an arbitrary palette if $P$ is empty: see
+\autoref{fig:edge_palette} for a small illustration. Once all nodes have been colored, we count the
+number of edges that are colorless. We repeat this procedure, keeping track of the palettes assignment
+minimizing the number of colorless edges.
+
+\item
+Once we colored the nodes of every block, we look at pair
+of blocks, build a list of all edges between blocks that are not colorless and sample from it to
+connect blocks. The last step is to a assign a color $y_{uv}$ to each edge $(u,v)$ from the shared color of its
+endpoint (or an arbitrary color for colorless edges). We note that this is reminiscent of the stochastic
+block model, although the probabilities of edges between blocks are not uniform, for they depend of
+the palette assigned to every node.
+\begin{figure}[bht]
   \centering
-  \includegraphics[width=\textwidth]{tikz/edge_palette_tikz.pdf}
+  \includegraphics[width=0.4\textwidth]{tikz/edge_palette_tikz.pdf}
   \caption{A small example of the node palette assignment, with $k=4$ colors (blue, green, red and
-    orange) and palette of size $k_\mathrm{local}=3$. We assume nodes $1$ and $2$ have already been
+    orange) and palettes of size $k_\mathrm{local}=3$. We assume nodes $1$ and $2$ have already been
     visited, and got assigned the palette $(\text{blue, green, red})$ and $(\text{blue, green,
-    orange})$ respectively. Moreover, we are currently at $3$ while $4$ is yet uncolored. In that
-    case, there are two possible palettes for $3$. If we select $(\text{green, red, orange})$, a
+    orange})$ respectively. Moreover, we are currently visiting node $3$ while node $4$ is yet uncolored. In that
+    case, there are two possible palettes for node $3$. If we select $(\text{green, red, orange})$, a
     possible color assignment for the edges $(1,2)$, $(1, 3)$ and $(2, 3)$ is respectively blue,
@@
   \label{fig:edge_palette}}
-\end{marginfigure}
-\autoref{fig:edge_palette} for a small illustration). Once all nodes have been colored, we count the
-number of edges that are colorless. We repeat this procedure, keeping track of the palettes assignment
-minimizing the number of colorless edges. Once we colored the nodes of every block, we look at pair
-of blocks, build a list of all edges between blocks that are not colorless and sample from it to
-connect blocks. The last step is to a assign a color to each edge from the shared color of its
-endpoint (or an arbitrary color for colorless edges). Note this is reminiscent of the stochastic
-block model, although the probabilities of edges between blocks are not uniform anymore.
+\end{figure}
 
-The directions $\mathcal{D}_k$, and the graph $G$ along its edge assignment are the two sources of
-randomness we will consider in the following.
+  \item 
+We then generate $k$ sparse, unit-norm directions $\mathcal{D}_k=\{w_1, \ldots, w_k\}$,
+independently of the graph. An underlying assumption of our model is that each direction should
+provide an explanation of why two nodes are connected that is markedly different from the
+explanation of another direction. A way to achieve that is to create directions with only a few
+non-zeros components, and ensure that the non-zero components of one direction do not appear at the
+same position as the non-zero components of another direction. Therefore, we control through $n_o$
+the number of dimensions where more than one non-zero component exists across the $k$ directions.
+That is, when $n_o = 0$, each direction has exactly \nicefrac{k}{d} non-zero components that are
+disjoint.\footnote{We choose $d$ to be a multiple of $k$.} On the other hand, if for instance $n_o=5$,
+then $5$ of the $d$ dimensions will have two directions with non-zero components on it. Those
+non-zero components are drawn \uar{} from $[-1, 1]$ and each direction is then normalized. For
+instance, with $d=6$, $k=3$ and $n_o=2$, we could generate the following directions, where the blue
+non-zero components are \enquote{overlapping} with the black ones:
+% \vspace{-.5\baselineskip}
 
-\bigskip
+\begin{center}
+  \begin{tabular}{ccc}
+    $w_1$ & $w_2$ & $w_3$ \\
+    $\begin{pmatrix}
+      \phantom{-}0.729 \\
+      0    \\
+      0    \\
+      0    \\
+      \phantom{-}\textcolor{DodgerBlue}{0.450} \\
+      -0.516 \\
+    \end{pmatrix}$ &
+    $\begin{pmatrix}
+      0    \\
+      0    \\
+      \textcolor{DodgerBlue}{-0.483} \\
+      -0.533 \\
+      \phantom{-}0.694 \\
+      0    \\
+    \end{pmatrix}$ &
+    $\begin{pmatrix}
+      0    \\
+      -0.639 \\
+      -0.769 \\
+      0    \\
+      0    \\
+      0    \\
+    \end{pmatrix}$
+  \end{tabular}
+\end{center}
 
-Now that we have a direction $w_{uv} \in \mathcal{D}_k$ assigned to every edge $(u,v)$, we want to
-find a set of user profiles $X=\{x_u\}_{u\in V}$ that maximize the edge score and minimize the node
-loss, defined respectively by equations \eqref{eq:edge_full} and
-\eqref{eq:edge_node_loss} in \autoref{sec:edge_problem} as:
+\item
+Returning to directions instead of colors, we have a direction $w_{y_{uv}} \in \mathcal{D}_k$
+assigned to every edge $(u,v)$. For simplicity, in that paragraph, we will let $w_{uv} = w_{y_{uv}}$.
+The final step in generating our node-attributed graph is to
+find a set of user profiles $X=\{x_u\}_{u\in V}$ that maximize the edge goodness and minimize the node
+loss term, defined respectively as:
 \begin{align*}
   \mathcal{L}_{\mathrm{edge}} &=
-  \sum_{u,v \in E} \left(x_u \circ x_v\right)^T w_{uv} \quad \text{and} \\
+  \sum_{u,v \in E} {s_{uv}}^T w_{uv} \quad \text{and} \\
   \mathcal{L}_{\mathrm{node}} &=
@@
 \end{align*}
-where for simplicity, we fix $b_u = \zerov$ and $a_{uv} = 1$. Furthermore, to ensure that the
-assigned direction of any edge $(u,v)$ achieve a better score than the other directions (that is
-$\forall w_\ell \in \mathcal{D}_k \setminus \{w_{uv}\}\,, {s_{uv}}^T w_{uv} \geq  {s_{uv}}^T
-w_\ell$) we also minimize a cross-entropy loss commonly used in non-binary classification
-problem~\autocite[Section 4.3.4]{PRML06}.  Specifically, recall that the goodness of direction
-$w_\ell$ for the edge $(u,v)$ is $g(s_{uv}, w_{uv})$. If we denote by $p_{uv,\ell} = \frac{\exp
+where for simplicity, we fix as before $b_u = \zerov$ and $a_{uv} = \frac{1}{\deg(u)}$. Furthermore, to ensure that the
+direction assigned to any edge $(u,v)$ achieves a higher goodness than the other directions (that is
+$\forall w_\ell \in \mathcal{D}_k \setminus \{w_{uv}\}\,, g(s_{uv}, w_{uv}) \geq  g(s_{uv},
+w_\ell)$), we also minimize a cross-entropy loss commonly used in non-binary classification
+problem~\autocite[Section 4.3.4]{PRML06}.  Specifically, denoting by $p_{uv,\ell} = \frac{\exp
 \left( g(s_{uv}, w_{uv}) \right)}{\sum_{\ell=1}^k \exp \left( g(s_{uv}, w_\ell) \right)}$ the
 softmax \enquote{probability} of $s_{uv}$ being explained by direction $\ell$, the cross-entropy
-loss is defined by
+loss is defined by:
 \begin{equation}
@@
   \mathcal{L}_{\mathrm{cross-entropy}} = -\sum_{u,v \in E} \sum_{l=1}^k
-  \Ind{w_{uv} = w_\ell} p_{uv,\ell}
+  \Ind{w_{uv} = w_\ell} p_{uv,\ell}\,.
 \end{equation}
 
-In practice, we first minimize $\lambda \mathcal{L}_{\mathrm{cross-entropy}} -
-\mathcal{L}_{\mathrm{edge}}$ with respect to $X$ using the Adam algorithm~\autocite{Adam15} and
+In practice, we first minimize
+\begin{equation}
+  \label{eq:edge_cross_edge}
+  \lambda \mathcal{L}_{\mathrm{cross-entropy}} - \mathcal{L}_{\mathrm{edge}}
+\end{equation}
+with respect to $X$ using the Adam algorithm~\autocite{Adam15} and
 automatic differentiation~\autocite{autograd15}\footnote{Both being implemented by the Pytorch
@@
 Finally, to minimize $\mathcal{L}_{\mathrm{node}}$, we iterate over the nodes and take some gradient
-steps to minimize $\left|\left| x_u - \sum_{v\in\nei{u}} w_{uv}\right|\right|^2$, but only to the
-extent that $w_{uv}$ remains the direction with the largest score on the edge $u,v \in E$.
+steps to minimize $\left|\left| x_u - \frac{1}{\deg(u)} \sum_{v\in\nei{u}} w_{uv}\right|\right|^2$, but only to the
+extent that $w_{uv}$ remains the direction with the highest goodness on the edge $(u,v) \in E$.
+\end{enumerate}
 
@@
 
-In the following we generate one graph topology, along with palette assignment to nodes and
-direction assignment to edges. Then we generate a set of unit-norm directions,
-parametrized by four numbers (and their default value): $k=7$, $d=35$, $n_o=0$ and
-$k_{\mathrm{local}}=3$. To account for the randomness of this generation process, we repeat our
-measurement over 200 such $\mathcal{D}_k$. On the other hand, we verify experimentally that
+In this section, we start by describing how we generate test instances of \autoref{p:edge_full} for
+the five methods we want to compare.  Then we present two evaluation measures: how well those
+methods classify edges, and how well they recover known directions. We use the first measure to
+comment on the quality of the generated instances. Finally we present the results of the five
+methods in \autoref{tab:edge_Wvaries_nami} and \autoref{tab:edge_Wvaries_dst}.
+% - describe how we generate instances
+% - present our two evaluation measures how well we classify edges and how well we recover known direction
+% - comment on their quality (but this requires AMI…)
+% - comparison of the 5 methods
+
+Our test instances are generated as follow. We first choose $n=500$, and generate once a graph
+topology with directions assigned to all edges according to steps 1 and 2. This results in a graph
+$G$ with roughly $\np{1350}$ edges. We then generate the directions $\mathcal{D}_k$ according to
+step 3 and the following choice of parameters: $d=35$, $k=7$, $k_{\mathrm{local}}=3$ and $n_o=0$.
+Finally, we create the profiles $X$ as described in step 4.
+To account for the randomness of this generation process, we repeat our
+measurement over \np{200} such generation of $\mathcal{D}_k$ and $X$. On the other hand, we verify experimentally that
 generating various graphs while keeping $\mathcal{D}_k$ fixed gives the same results, hence we will
 not report them in the main text.
+
+In the rest of this section, we will refer to the set of parameters $d=35$, $k=7$,
+$k_{\mathrm{local}}=3$ and $n_o=0$ as the \emph{default} configuration.
 We also try other configurations to study the parameter sensibility of our methods. Specifically, we
@@
 
-After having assigned a single direction to every edge and generated the user profiles, we can
-measure how the assignment and the profiles match, in the sense that the assigned direction of every
-edge is the one with the largest goodness. Indeed, recall we predict the
-direction of an edge $(u,v)$ as $\argmax_{\ell \in \rangesk} g(s_{uv}, w_\ell) = {s_{uv}}^T w_\ell$.
-However, in the following, we do not evaluate such predictions using accuracy. Indeed, while the
-methods we evaluate all assign one among $k$ directions to every edge, there is no guarantee those
+\medskip
+% After having assigned a single direction to every edge and generated the user profiles, we can
+% measure how the assignment and the profiles match, in the sense that the assigned direction of every
+% edge is the one with the largest goodness. Indeed, recall we predict the
+% direction of an edge $(u,v)$ as $\mathcal{E}(u,v) = \argmax_{\ell \in \rangesk} g(s_{uv}, w_\ell)$.
+% To do so, we however do not rely on the accuracy measure.
+
+Let us now describe how we evaluate the results of our methods on such instances. As said at the
+beginning of this section, we extract from the output of each method a set $\mathcal{\wh{D}}_k =
+\{\wh{w}_1, \ldots, \wh{w}_k\}$ of $k$ directions, and assign a 
+direction to an edge $(u,v)$ as $\mathcal{E}(u,v) = \argmax_{\ell \in \rangesk} g(s_{uv}, \wh{w}_\ell)$.
+Recall that during the second of the step of the instance generation, we assign a direction index
+$y_{uv}$ to every edge.
+However, we cannot simply evaluate our predictions by simply comparing $\mathcal{E}(u,v)$ and
+$y_{uv}$. Indeed, here is no guarantee that the $\mathcal{\wh{D}}_k$
 directions are recovered in the same order as they were generated. In case of strong disagreement
@@
 
-First, as we can see in the table at the end of this paragraph, in the \default{} configuration, we do not always
-generate user profiles leading to a perfect assignment of edge according to their goodness. In
-general this is not always possible. Consider a $k$-regular subgraph, where the $k$ edges incident
-to every node are each assigned a different directions. By symmetry, it is impossible to find node
-profiles that would achieve maximum goodness for all the edges. This is partly what motivated the
-$k_{\mathrm{local}}$ local constraint. Even so, and more pragmatically, whereas those AMI scores
-could be made higher by increasing the coefficient $\lambda$ in~\eqref{eq:edge_cross}, this would
-imply a lower edge goodness. Even with our choice, only a few percent of the edges have a mismatch
-between the directions we assigned them and the one with the largest goodness. Note also that
-depending of the parameters of the generation, the AMI is not the same. Therefore, in the following,
-we divide all the AMI scores by this natural score, in order to obtain a normalized measure.
-\newline\vspace{.8\baselineskip}\noindent%
-\begin{tabular}{lccccccc}
-  \toprule
-  {}  & {\smallk{}}    & {\default{}}   & {\largek{}}    & {\smallo{}}    & {\largeo{}}    & {\fdirs{}}     & {\larged{}}    \\
-  \midrule
-  AMI & $.925 \pm .03$ & $.922 \pm .03$ & $.905 \pm .02$ & $.880 \pm .04$ & $.874 \pm .05$ & $.893 \pm .03$ & $.973 \pm .01$ \\
-  \bottomrule
-\end{tabular}
-\vspace{.5\baselineskip}
-
-Besides this normalized AMI, we also measure how far the recovered directions $\mathcal{\wh{D}}_k$
+Besides the AMI score, we also measure how far the recovered directions $\mathcal{\wh{D}}_k$
 are from the actual ones, generated initially. More precisely, we associate each vector $w_\ell \in
@@
 \end{equation*}
+
+\medskip
+
+Before moving to the results, let us first note that,
+in the \default{} configuration, we do not always
+generate user profiles leading to a perfect assignment of edge according to their goodness. This can
+be seen in \autoref{tab:edge_badami}. In
+general, this is not always possible. Consider a $k$-regular subgraph, where the $k$ edges incident
+to every node are each assigned a different directions. By symmetry, it is impossible to find node
+profiles that would achieve maximum goodness for all the edges. This is partly what motivated the
+$k_{\mathrm{local}}$ local constraint. Even so, and more pragmatically, whereas those AMI scores
+could be made higher by increasing the coefficient $\lambda$ in~\eqref{eq:edge_cross_edge}, this would
+imply a lower edge goodness. With our choice of $\lambda$, only a few percent of the edges have a mismatch
+between the directions we assigned them and the one with the largest goodness. Note also that
+depending of the parameters of the generation, the AMI is not the same. Therefore, in the following,
+we divide all the AMI scores by this natural score, in order to obtain a \enquote{standardized} measure.
+% \newline\vspace{1\baselineskip}\noindent%
+
+\begin{table}[htb]
+  \centering
+  \small
+  \caption{The degree to which the edge assignment of step 2 agree with the directions and profiles
+    of steps 3 and 4. This is measured by the AMI between the assignment $\{y_{uv}\}_{uv \in E}$ and
+    what maximal goodness would assign, \ie{} $\{\mathcal{E}(u,v)\}_{uv \in E}$. We report the
+    average AMI and standard deviation over 200 generations of $\mathcal{D}_k$, for different
+    configurations.  \label{tab:edge_badami}}
+\begin{tabular}{lccccccc}
+  \toprule
+  {$\mathcal{D}_k$ parameters}  & {\default{}}   & {\smallk{}}    & {\largek{}}    & {\smallo{}}    & {\largeo{}}    & {\fdirs{}}     & {\larged{}}    \\
+  \midrule
+  AMI & $.922 \pm .03$ & $.925 \pm .03$ & $.905 \pm .02$ & $.880 \pm .04$ & $.874 \pm .05$ & $.893 \pm .03$ & $.973 \pm .01$ \\
+  \bottomrule
+\end{tabular}
+\end{table}
 
@@
   \small
-  \caption{Normalized AMI of 5 methods, when generating directions with 7 configurations. For each
-    configuration, we generate directions 200 times, and report the mean normalized AMI along with
+  \caption{Standardized AMI of 5 methods, when generating directions with 7 configurations. For each
+    configuration, we generate directions 200 times, and report the mean standardized AMI along with
     the standard deviation. Among the first three methods (that do not use matrix formulation), we
@@
     \toprule
-    {} &         \thead{\kmeans{}} &                                  \thead{\lloyd{}} &                              \thead{\combined{}} &  \thead{\fwa{}} &  \thead{\pqt{}} \\
+    \thead{$\mathcal{D}_k$ parameters} &         \thead{\kmeans{}} &                                  \thead{\lloyd{}} &                              \thead{\combined{}} &  \thead{\fwa{}} &  \thead{\pqt{}} \\
     \midrule
+    {\default{}} &            $.818 \pm .06$ &  $\esecondSig{.873 \pm .05}\,\,\spval{1.25}{-63}$ &  $\efirstSig{.893 \pm .04}\,\,\spval{5.68}{-33}$ &  $.381 \pm .05$ &  $.893 \pm .04$ \\
     {\smallk{}}  &            $.836 \pm .08$ &                          $\esecond{.838 \pm .07}$ &  $\efirstSig{.875 \pm .06}\,\,\spval{2.17}{-58}$ &  $.213 \pm .11$ &  $.875 \pm .06$ \\
-    {\default{}} &            $.818 \pm .06$ &  $\esecondSig{.873 \pm .05}\,\,\spval{1.25}{-63}$ &  $\efirstSig{.893 \pm .04}\,\,\spval{5.68}{-33}$ &  $.381 \pm .05$ &  $.893 \pm .04$ \\
     {\largek{}}  &            $.803 \pm .04$ &  $\esecondSig{.881 \pm .04}\,\,\spval{2.66}{-94}$ &  $\efirstSig{.894 \pm .04}\,\,\spval{8.98}{-17}$ &  $.421 \pm .04$ &  $.894 \pm .04$ \\
@@
       \toprule
-      {} &                                 \thead{\kmeans{}} &                                  \thead{\lloyd{}} &                              \thead{\combined{}} &   \thead{\fwa{}} &  \thead{\pqt{}} \\
+      \thead{$\mathcal{D}_k$ parameters} &                                 \thead{\kmeans{}} &                                  \thead{\lloyd{}} &                              \thead{\combined{}} &   \thead{\fwa{}} &  \thead{\pqt{}} \\
       \midrule
+      {\default{}} &                                    $.533 \pm .08$ &   $\esecondSig{.528 \pm .09}\,\,\spval{9.61}{-4}$ &   $\efirstSig{.516 \pm .06}\,\,\spval{4.95}{-5}$ &   $.877 \pm .05$ &  $.564 \pm .06$ \\
       {\smallk{}}  &  $\esecondSig{.581 \pm .08}\,\,\spval{4.78}{-30}$ &                                    $.606 \pm .09$ &  $\efirstSig{.560 \pm .05}\,\,\spval{8.07}{-24}$ &  $1.007 \pm .05$ &  $.633 \pm .06$ \\
-      {\default{}} &                                    $.533 \pm .08$ &   $\esecondSig{.528 \pm .09}\,\,\spval{9.61}{-4}$ &   $\efirstSig{.516 \pm .06}\,\,\spval{4.95}{-5}$ &   $.877 \pm .05$ &  $.564 \pm .06$ \\
       {\largek{}}  &                                    $.541 \pm .07$ &  $\esecondSig{.521 \pm .07}\,\,\spval{5.08}{-23}$ &                          $\efirst{.520 \pm .05}$ &   $.811 \pm .06$ &  $.551 \pm .06$ \\
@@
 \end{aside}
+\fi

### 2017-11-14 edge/model.tex
take feedback on edge chapter into account
- rewrite the setting part, giving the definition of s and g right away,
and adding a “prediction” operator
- try to motivate better the | x - b - \sum a⋅w | term in (1.2)
- acknowledge the signed case is only a sanity check (I couldn't find a
closed form solution for the noisy case)
- give a list of methods at the beginning of 1.2
- fix typos and math
- split the generation process in 4 steps and explicitly state its input
and output
- reorganize the results writeup in 4 parts: generation, measures,
initial AMI, commented tables

@@
 
-As in the previous chapter, we are given an undirected, unweighted graph $G=(V,E)$. This
-time though, we also have side information in addition to the graph topology. Namely, each node $u$
-is associated with a $d$-dimensional feature vector $x_u \in \Rbb^d$. As an example, in a social
-network, $x_u$ would be the profile of user $u$, describing $u$'s demographics and preferences. We
-are also given a combination operator $s$ between feature vectors, such that $s(x_u, x_v) = s_{uv}
-\in \Rbb^d$. Finally, we consider distinguished feature vectors $w_{uv}$,
-along which the combination $s_{uv}$ between two nodes $u$ and $v$ is evaluated. We call such
-$w_{uv}$ \emph{directions}, and constrain them to have unit norm.
-While being exactly of norm one is not primordial, the role of bounding the norm of the directions
-(\ie{} having $\sum_{i=1}^d {w_{uv}}_{;i}^2 \leq B^2$) is to tie the $d$ dimensions together.
-Without such a constraint, we could simply study each dimension separately, whereas this bound
-adds a coupling across different attributes, since increasing the weight of an attribute mechanically
+Like we wrote in the introduction, to classify edges, we make use of both the topology of the graph
+and the attributes of its nodes. Namely, we are given an undirected, unweighted graph $G=(V,E)$.  We
+also have side information in addition to the graph topology. It takes the form, for every node $u$,
+of a $d$-dimensional feature vector $x_u \in \Rbb^d$ that we call the \emph{profile} of $u$.
+Because we want to model heterophily, we more explicitly have that $x_u \in [-1, 1]^d$. To justify
+this statement, let us take the example of encoding the gender of a user in a social network,
+assuming there is only two values possible: male and female. Traditionally, this attribute would be
+set to $1$ if the user is female and $0$ otherwise. Instead, we use $+1$ and $-1$, to clearly
+emphasize the difference between the two values. The same is valid for continuous variables. For
+instance, an attribute could range from $-1$ to $+1$ to represent the ideological stance of a user
+on the conservative/liberal scale~\autocite{newsBias15}.\footnote{Note of course that in both cases,
+negative values do not represent a judgement but are merely a reversible mathematical convenience.}
+This applies even to non diverging positive quantities such as age, which can be projected to $[-1,
+1]$ using pre-processing like standardization or $z$-score.
+
+Stacking all these node profiles in a $|V| \times d$ matrix $X$, our input so far is $(G, X)$. As it
+is standard in classification problem, we are also given the number of edge types
+$k$.\footnote{Because our problem is essentially unsupervised, assuming that we are given $k$ is a
+simplification, on which we elaborate \vpageref{par:edge_choose_k}.} Moreover, recall that we do not
+merely want to classify edges into type, but also provide an explanation of why two nodes are
+connected. This requires the introduction of three further concepts:
+\begin{inparaenum}[1)]
+\item how to represent a connection (\ie{} an edge),
+\item how to represent an edge's explanation, and
+\item how to evaluate which explanation is the best.
+\end{inparaenum}
+We now present our modelling choice:
+
+\begin{enumerate}[1),nosep,leftmargin=*]
+  \item We naturally describe an edge $(u,v)$ through the profiles of its endpoints. More precisely,
+    we combine $x_u$ and $x_v$ by an operator $s$ defined as:
+    \begin{equation*}
+      s(u, v) = x_u \circ x_v = s_{uv} \in \Rbb^d \, ,
+    \end{equation*}
+    where $\circ$ stands for the Hadamard (or component wise) product. As showed in
+    \begin{marginfigure}
+      % \small
+      \captionsetup{font=small,type=table}
+      \captionof{table}{We show the effect of the Hadamard product in combining information from two
+        users $u$ and $v$. Instead of actual numbers, in this table we use $+$ to denote a large
+        positive value, $-$ to denote a large negative value, $0$ to denote a small absolute value
+        and $\star$ to denote an arbitrary value.
+      \label{tab:edge_sim}}
+      \begin{tabular}{ccc}
+        \toprule
+        ${x_u}_{;i}$ & ${x_v}_{;i}$ & ${s_{uv}}_{;i}$ \\
+        \midrule
+        $+$          & $+$          & $+$             \\
+        $-$          & $-$          & $+$             \\
+        $+$          & $-$          & $-$             \\
+        $0$          & $\star$      & $0$             \\
+        \bottomrule
+      \end{tabular}
+    \end{marginfigure}
+    \autoref{tab:edge_sim}, for each dimension $i \in \rangesd$, the $s_{uv}$ holds a value
+    indicating whether the profiles of $u$ and $v$ agree, disagree or are indifferent along that
+    dimension.
+  \item The $k$ explanations take the form of bounded norm vectors $\mathcal{D}_k = \{w_1, w_2,
+    \ldots, w_k\} \subset \dball$, where $\dball$ is the $\ell_2$-unit ball of $\Rbb^d$. We call such a
+    vector $w_\ell$, for $\ell \in \rangesk$, a \emph{direction}. As we shall explain more formally
+    later on, those directions are exactly the parameters learned while solving our problem.
+  \item Given an edge $(u,v)$, all the $k$ directions $\{w_1, w_2, \ldots, w_k\}$ provide a possible
+    explanation of why $u$ and $v$ are connected. We score these explanations by assigning them a
+    \emph{goodness} of explanation\footnote{Whereas the name is inspired by the concept of
+    \emph{goodness of fit}, there is no formal relation.} defined as:
+    \begin{equation*}
+      g(s_{uv}, w_\ell) = {s_{uv}}^T w_\ell \, .
+    \end{equation*}
+    As the $w$ notation hints, a direction can be interpreted as a set of weights. These weights
+    indicates to which extent each attribute contributes to the connection, according to that
+    direction. For a given set of directions $\mathcal{D}_k$, the classification of edges is
+    naturally performed by the following operator $\mathcal{E}$, returning the index of the
+    direction with the maximal goodness:
+    \begin{equation*}
+      \begin{array}[t]{lrcl}
+        \mathcal{E} : & E \times \left( \Rbb^d \times \ldots \times \Rbb^d \right)
+                      & \longrightarrow & \rangesk \\
+                      & (u,v), \mathcal{D}_k & \longmapsto
+                      & \argmax_{\ell \in \rangesk} g(s_{uv}, w_\ell)
+      \end{array}
+    \end{equation*}
+    Because it will be clear from context, we lighten the notation and simply write $\mathcal{E}(u,
+    v)$ instead of $\mathcal{E}\left((u, v), \mathcal{D}_k \right)$.
+\end{enumerate}
+
+This choice of $s$ and $g$ enjoy two interesting properties. First, it encompasses all cases of
+homophily, indifference and heterophily. For instance, given $x_u = \begin{pmatrix} 1 & 0 & -1
+\end{pmatrix}^T$ and $x_v = \begin{pmatrix} -1 & 1 & -1 \end{pmatrix}^T$, the direction $w =
+\begin{pmatrix} -\nicefrac{1}{\sqrt{2}} & 0 & \nicefrac{1}{\sqrt{2}} \end{pmatrix}^T \in \dball$
+achieves the maximal goodness. In doing so, it highlights that $u$ and $v$ disagree on the first
+component (\ie{} heterophily), agree on the last component (\ie{} homophily) and that the middle
+component is irrelevant (\ie{} indifference). This also shows that, second, given the linear nature
+of the goodness $g$ once $s_{uv}$ has been computed, it provides interpretable explanation of why
+two nodes are connected.
+
+Those two properties of $s$ and $g$ should not be taken for granted. Indeed, 
+a natural choice for the operator $s$ is the difference between the profiles, that is
+$s_{uv} = x_u - x_v$. In that spirit, if $u$ and $v$ are close along a direction $w_\ell$, we expect
+${x_u}^T w_\ell$ to be close to ${x_v}^T w_\ell$, as measured by $|{x_u}^T w_\ell - {x_v}^T
+w_\ell|$. And because it is easier to deal with smooth functions, we define the goodness to be $g(s_{uv},
+w_\ell) = -\left({s_{uv}}^T w_\ell \right)^2$, where the minus preserve the semantic than the
+\enquote{better} the direction $w_\ell$, the larger the goodness. However, with this formulation,
+there is no way to set ${w_\ell}_{;i}$ in order to \enquote{highlight} the fact that ${x_u}_{;i}$
+and ${x_v}_{;i}$ are widely different. Furthermore, the square in the goodness expression loses the
+linearity.
+
+We conclude with a last comment on the norm of the directions. While being exactly of norm one is
+not primordial, the role of bounding the norm of the directions (\ie{} having $\sum_{i=1}^d
+{w_\ell}_{;i}^2 \leq B^2$) is to tie the $d$ dimensions together. Without such a constraint, and we
+our choice of $s$ and $g$, we could simply study each dimension separately, whereas this bound adds
+a coupling across different attributes, since increasing the weight of an attribute mechanically
 affects the weights of the other attributes.
-
-Depending on the application, and the semantic we ascribe to links between nodes, we can design and
-interpret the combination operator $s$ as a difference, a similarity or a distance. Likewise, we
-also have some freedom in defining how $s_{uv}$ and $w_{uv}$ are combined to assign a single real
-value $g(s_{uv}, w_{uv})$ to the edge $(u,v)$. Let us call $g(w_{uv}, s_{uv})$ the \emph{goodness}
-of $w_{uv}$ with respect to the edge $(u,v)$. We view this value as how adequately the direction
-$w_{uv}$ \emph{explains}, or describes, the connection between $u$ and $v$.
-
-\medskip
-
-As written in the introduction, because we are interested in (partial) heterophilic links, we assume
-in the following that both the profiles and the directions are vectors with components in $[-1, 1]$.
-Let us take the example of encoding the gender of a user in a social network, assuming there is only
-two values possible: male and female. Traditionally, this attribute would be set to $1$ if the user
-is female and $0$ otherwise. Instead, we use $+1$ and $-1$, to clearly emphasize the difference
-between the two values. The same is valid for continuous variables. For instance, an attribute could
-range from $-1$ to $+1$ to represent the ideological stance of a user on the conservative/liberal
-scale~\autocite{newsBias15}.\footnote{Note of course that in both cases, negative values do not
-represent a judgement but are merely a reversible mathematical convenience.} This applies even to
-non diverging positive quantities such as age, which can be projected to $[-1, 1]$ using
-pre-processing such as standardization or $z$-score.
-
-A natural choice for the operator $s$ is then the difference between the profiles, that is
-$s_{uv} = x_u - x_v$. In that spirit, if $u$ and $v$ are close along a direction $w_{uv}$, we expect
-${x_u}^T w_{uv}$ to be close to ${x_v}^T w_{uv}$, as measured by $|{x_u}^T w_{uv} - {x_v}^T
-w_{uv}|$. Because we prefer to work with smooth functions, we define the goodness to be $g(s_{uv},
-w_{uv}) = \left({s_{uv}}^T w_{uv} \right)^2$, and we want to minimize it with respect to $w_{uv}$.
-When, on the \ith{} dimension, ${x_u}_{;i} - {x_v}_{;i}$ is close to zero, the minimization objective
-and the fact that $w_{uv}$ is
-unit-norm impose that ${w_{uv}}_{;i}$ should be large. On the other hand, with this formulation,
-there is no way to set ${w_{uv}}_{;i}$ in order to \enquote{highlight} the fact that
-${x_u}_{;i}$ and ${x_v}_{;i}$ are widely different.
-There might be situations where this is not a problem. But again,
-here we seek to model heterophilic interactions, at least along a subset of dimensions. Therefore,
-we now introduce a more flexible approach.
-
-Namely, we define $s$ to be the Hadamard (or component wise) product: $s_{uv} = x_u \circ x_v$, and
-$g$ to be the dot product: $g(s_{uv}, w_{uv}) = {s_{uv}}^T w_{uv}$. The goodness of a direction $w$
-for a given edge can thus be seen as a weighted sum of the per component similarity of $u$ and $v$
-profiles. Therefore, it must be maximized with respect to $w_{uv}$. For a given component $i$, when
-both ${x_u}_{;i}$ and ${x_v}_{;i}$ are large and of the same sign (that is close to either $+1$ or
-$-1$), we see that $u$ and $v$ agree on the \ith{} dimension, and ${w_{uv}}_{;i}$ must accordingly
-be a large positive value. On the other hand, when ${x_u}_{;i}$ and ${x_v}_{;i}$ are both large but
-of different signs (\eg ${x_u}_{;i} = +1$ and ${x_v}_{;i} = -1$), $u$ and $v$ strongly disagree on
-the \ith{} dimension, requiring in a large negative value for ${w_{uv}}_{;i}$.  Finally, if one of
-${x_u}_{;i}$ or ${x_v}_{;i}$ is zero, denoting either an unimportant feature or a missing value,
-then the value of ${w_{uv}}_{;i}$ does not matter, and will be set to zero by the maximization
-procedure because of the unit norm constraint on $w_{uv}$.
-
-\iffalse
-\begin{aside}
-When $s$ is a distance, and therefore positive, we impose $w_{uv}\geq 0$. A link $(u,v)$ is then
-well explained by $w_{uv}$ when observed differences along this direction are small. More precisely,
-when the quantity ${s_{uv}}^T w_{uv}$ is close to zero. Because $w_{uv}$ is unit norm, the \ith{}
-component of $w_{uv}$ should be large when the \ith{} feature of $x_u$ and $x_v$ are close. Hence, a
-null value for a component $i$ of $s_{uv}$ should not mean anything else than $u$ and $v$ share a
-known and similar feature $i$ (\eg{} the unknown value should be different from $0$). When $s$ is
-the difference $x_u - x_v$, it can be negative and the same reasoning can be applied, but with
-$w_{uv}\in [-1,1]^d$ and a link value of $\left({s_{uv}}^T w_{uv}\right)^2$.
-
-An alternative is to consider that the combination is a similarity. Links are created when the
-observed similarity is high. Again, if $s$ is positive, it seems reasonable to impose that
-$w_{uv}\geq0$ (and if $s$ can be negative, then $w_{uv}$ should be in $[-1,1]$). One main advantage
-of similarities is maybe to avoid (or minimize) problems with missing values setting them to 0.
-\end{aside}
-\fi
-
-\medskip
-
-This second choice of $s$ and $g$ is indeed more flexible, for it encompasses all cases of
-homophily, indifference and heterophily. Furthermore, since the profiles are given, the resulting
-$s_{uv}$ can be seen as constants, meaning that $g$ is linear and therefore interpretable. Finally,
-it also supports categorical data. We shall demonstrate all these features on a small example,
-but we first define formally our problem.
+For consistency, we also consider node profiles to be normalized and having unit norm. A further
+practical justification is that if nodes are users of a social network and the profiles measure
+their activity across several domains, this allows to compare users with various level of total
+activity.
 
@@
 With working definitions of $s$ and $g$ at hand, we now formally write down the problem of
-finding and assigning the most informative directions to every edge of $G$
+finding the directions maximizing the goodness of the graph.
 \begin{problem}[\ecp{}]
@@
   Given a graph $G=(V, E)$, node profiles $X\in [-1, 1]^{n\times d}$ and an integer $k \in \Nbb$,
-  find a set of $k$ unit-norm directions $\mathcal{D}_k = \{w_1, w_2, \ldots, w_k\} \in \dsphere$
+  find a set of $k$ norm-bounded directions $\mathcal{D}_k = \{w_1, w_2, \ldots, w_k\} \in \dball$
   and associate to every edge of $E$ the direction with the maximal goodness. Formally, solve
@@
     \label{eq:edge_full}
-    \argmax_{w_1,\ldots,w_k \in \dsphere}
-    \sum_{u,v \in E} \max_{\ell \in \rangesk} g(s_{uv}, w_{\ell})
+    \argmax_{\mathcal{D}_k = \{w_1,\ldots,w_k\} \subset \dball}
+    \sum_{u,v \in E}  g(s_{uv}, w_{\mathcal{E}(u,v)})
   \end{equation}
-  where $s_{uv} = x_u \circ x_v$ and $g(s_{uv}, w_{\ell}) = {s_{uv}}^T w_{\ell}$.
+  where $s_{uv} = x_u \circ x_v$, $g(s_{uv}, w_{\ell}) = {s_{uv}}^T w_{\ell}$ and $\mathcal{E}(u,v)
+  = \argmax_{\ell \in \rangesk} g(s_{uv}, w_\ell)$.
 \end{problem}
@@
 their goodness. For instance, we see that the edge between $x$ and $y$ has a low score, because
-although those two users have the same \textsf{age} and used to work at \textsf{F}, they do not
-share a position at \textsf{G}.
+although those two users have the same \textsf{age} and used to work at \textsf{F}, not both of them
+works at \textsf{G}.
 
@@
       \toprule
-      {}    & G   & F    & gender & age    & opinion \\
+      {} & \textsf{G} & \textsf{F} & \textsf{gender} & \textsf{age} & \textsf{opinion} \\
       \midrule
@@
     \end{tabular}
-    \caption{The numerical attributes of the six nodes. They are followed by the (non normalized)
-    two directions $w_1$ and $w_2$.} \label{tab:edge_exe}
+    \caption{The numerical attributes of the six nodes. All are categorical encoded with a ternary
+      scheme, except for \textsf{age}, which has been transformed to fit in $[-1, 1]$.
+      They are followed by the (non normalized) two directions $w_1$ and $w_2$.}
+      \label{tab:edge_exe}
   \end{subfigure}
@@
 another way to leverage the graph topology is to define local constraints at the node level.
-For that we will consider node profiles to be normalized and having unit norm. A practical
-justification is that if nodes are
-users of a social network and the profiles measure their activity across several domains, this
-allows to compare users with various level of total activity. Furthermore, we assume that profiles
-are linear combination of directions, that
-is $x_u = b_u + \sum_{v \in \nei(u)} a_{uv} w_{uv}$, where $a_{uv}$ are real weights and $b_u
-\in \Rbb^d$ is the bias vector of $u$, \ie{} its inherent profile taken in isolation.
-Viewed in the other direction, this can be interpreted as follow: each time a node $u$
-connects to another node $v$ along a direction $w_{uv}$, it \enquote{consumes} a part of its profile
-proportional to $w_{uv}$. Therefore, we subtract the term
+Indeed, one of our motivation is that it is not enough to look at edges in isolation, but rather we
+expect the connections of one node to influence the connections of its neighbors. We now present two
+of these local, node-level constraints.
+
+First, if a graph was to perfectly follow our model, every edge $(u,v)$ would be maximally explained
+by one of the $k$ directions. This would happen if the profiles of $u$ and $v$ were collinear with
+this direction, up to sign reversal. When considering several edges, we could imagine that the nodes
+profiles result from the following process. Two nodes $u$ and $v$ establish a connection by first
+picking a direction $w_\ell \in \mathcal{D}_k$ and add to their base vector $b_u$ and $b_v \in
+\Rbb^d$ a fraction $a_{uv} > 0$ of $w_\ell$. We therefore subtract the following term
 \begin{equation}
@@
   \mathcal{L}_{\mathrm{node}} =
-  \sum_{u\in V} \left|\left| x_u - b_u - \sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2_2
+  \sum_{u\in V} \left|\left| x_u - b_u -
+  \sum_{v \in \nei(u)} a_{uv} w_{\mathcal{E}(u,v)} \right|\right|^2_2
 \end{equation}
-from \eqref{eq:edge_full}.
+from \eqref{eq:edge_full}. The cost we pay is introducing the extra parameters $\{a_{uv}\}_{uv \in
+E}$ and $\{b_u\}_{u \in V}$ to be learned, but we will see in the next section how to handle that in
+practice. On the other hand, it has two advantages. First, this is another way, besides goodness, in
+which the (given) profiles provide prior information about the (unknown) directions, and thus
+restrain the search space. Second, this makes the directions dependent of the topology of the graph,
+by coupling them with the neighborhood of each node.
 
-\bigskip
-
-We also make the assumption that each node $u$ is only involved in $k_\mathrm{local} < k$
+Second, we also make the assumption that each node $u$ is only involved in $k_\mathrm{local} < k$
 different directions. The intuition is that when $d$ is large enough, users have to focus their
@@
 well studied~\autocite{sparseOptim12}. Namely, we let $y_{uv}$ be a softmax membership
-$k$-dimensional vector, that is we set $\forall i \in \rangesk\,, y_{uv,i} =
-\frac{\exp(\beta{s_{uv}}^T w_i)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_j)} $, with $\beta>0$ a large
+$k$-dimensional vector, that is we set, $\forall i \in \rangesk$,
+\begin{equation*}
+  y_{uv,i} = \frac{\exp(\beta g(s_{uv}, w_i))}{\sum_{j=1}^k \exp(\beta g(s_{uv}, w_j))}\,,
+\end{equation*}
+with $\beta>0$ a large
 positive constant.  And we replace the $\ell_0$ norm by a $\ell_1$ norm (or even a $k$-support
@@
   \sum_{u \in V} \left( \sum_{i = 1}^k \sum_{v \in \nei(u)}
-  \frac{\exp(\beta{s_{uv}}^T w_i)}{\sum_{j=1}^k \exp(\beta {s_{uv}}^T w_j)} \right)
+  \frac{\exp(\beta g(s_{uv}, w_i))}{\sum_{j=1}^k \exp(\beta g(s_{uv}, w_j))} \right)
 \end{equation}
-
-These two constraints can be combined, so that the profiles are close to a linear combination of a
-small number of directions.
+It is reasonable to further expect that two neighboring nodes share one common direction among their
+$k_\mathrm{local}$ ones. One way to ensure this condition is to imagine that the set of allowed
+$k_\mathrm{local}$ directions changes smoothly over the graph. This property depends of the
+topology, as we can see by considering two extreme cases: a line and a complete graph. On the line,
+once a node has picked two allowed directions from its two neighbors, the remaining ones can be
+chosen arbitrarily. On the complete graph on the other hand, this choice is must more constrained.
 
@@
 
-After having introduced \autoref{p:edge_full} and some additional constraints, we now present
-methods to solve it. We start with a naive baseline, and propose a post-processing to improve it.
-Then we describe a convex relaxation. Finally, we depart slightly from finding $k$ directions by
+We now introduce the five methods solving \autoref{p:edge_full} that we will experimentally compare
+in the next section. Recall that given a graph topology $(V, E)$ and node profiles $X$, we first
+build the edges representation $\mathcal{S} = \{s_{uv}\}_{(u,v) \in E}$ and then look for $k$
+bounded directions that maximize the total goodness of the graph, possibly under topological
+constraints. We start with a straightforward baseline, which simply performs a \newline
+\hspace*{14pt} 1) \kmeans{} clustering of $\mathcal{S}$, \newline
+and propose a post-processing to improve it, plugging our custom goodness metric into the existing
+\newline \hspace*{14pt} 2) \lloyd{} algorithm for clustering. \newline
+Because these two methods consider each edge in isolation with no regard for the topology, we then
+describe a convex relaxation of the objective in equation \eqref{eq:edge_full}, which is further
+\newline \hspace*{14pt} 3) \combined{} with the node constraint of equation
+\eqref{eq:edge_node_loss}. \newline
+Finally, we depart slightly from finding $k$ directions by
 considering a low-rank matrix formulation, where each edge is assigned a linear combination of a
-small number of base directions, paving the way to overlapping clustering.
+small number of base directions, paving the way to overlapping clustering. More precisely, we
+give a convex formulation of that objective that can be solved by a \newline
+\hspace*{14pt} 4) \fwa{} algorithm, \newline
+as well an alternating optimization method, where we make the low-rank factorization \newline
+\hspace*{14pt} 5) \pqt{}.
 
@@
 set in \autoref{p:edge_full}. Formally, given the graph $G=(V,E)$ and the node profiles $X$ as
-input, we order the edges from $e_1$ to $e_m$. Then we build the matrix $S \in \Rbb^{m\times d}$, whose
+input, we order the edges from $e_1$ to $e_m$. Then we build the matrix $S \in \Rbb^{|E|\times d}$, whose
 \ith{} row is the similarity $s_{uv} = x_u \circ x_v$ between the profiles of the endpoints of $e_i=(u,v)$.
@@
 
+In both cases, those methods do not take the topology of the graph into account, motivating the
+following approaches.
+
 \subsection{Convex relaxation}
@@
 
-The difficulty in optimizing directly \eqref{eq:edge_full} stems from the inner $\max$ operator, which we
+By unrolling the operator $\mathcal{E}$ in \eqref{eq:edge_full}, we can rewrite our optimization
+objective as
+\begin{equation}
+  \label{eq:edge_unroll}
+  \argmax_{\mathcal{D}_k = \{w_1,\ldots,w_k\} \subset \dball}
+  \sum_{u,v \in E}  \max_{\ell \in \rangesk} g(s_{uv}, w_\ell)
+\end{equation}
+The difficulty in optimizing directly \eqref{eq:edge_unroll} stems from the inner $\max$ operator, which we
 replace here by a convex relaxation. Indeed, one can check that given a set of numbers $S=\{a_1,
@@
 \end{equation*}
-Using this fact, and averaging the two components of our objective function, we can rewrite
-\eqref{eq:edge_full} as:
+  \label{}
+Using this fact, and dividing the goodness term and the $\mathcal{L}_{\mathrm{node}}$ term of
+\eqref{eq:edge_node_loss} respectively by the number of edges and nodes so that they have comparable
+magnitude, we now have the following objective:
 \begin{equation}
@@
   - \frac{\mu}{|V|}\sum_{u\in V}
-  \left|\left| x_u - b_u - \sum_{v \in \nei(u)} a_{uv} w_{uv} \right|\right|^2,
+  \left|\left| x_u - b_u - \sum_{v \in \nei(u)} a_{uv} w_{\mathcal{E}(u,v)} \right|\right|^2,
 \end{equation}
 where $\mu > 0$ is a trade-off parameter.
-Note that in the second term, because we do not know the edge assignment, we have to decide which of the
-$w_1, \ldots, w_k$ is $w_{uv}$ equal to. According to our definition of the goodness function, we
-want to let $w_{uv} = \argmax_{w_\ell \in \mathcal{D}_k} {s_{uv}}^T w_\ell$. However, the $\argmax$
+
+Note that in the second term, $w_{\mathcal{E}(u,v)} = \argmax_{w_\ell \in \mathcal{D}_k} {s_{uv}}^T
+w_\ell$. However, the $\argmax$
 function is again not convex and we thus use the same relaxation as before. Namely, this relaxation
@@
 compute and in practice we
-found automatic differentiation~\autocite{autograd15} to be more efficient. Whereas the modified
-objective function \eqref{eq:edge_soft_early} is now jointly convex in all its variables, the set
-$\dsphere$ is not. To ease
+found automatic differentiation~\autocite{autograd15} to be more efficient.
+
+We make two further simplifications. First, notice that the set $\dsphere \times \ldots \times
+\dsphere$ is not convex. To ease
 the optimization procedure, we thus only impose the directions $w_\ell$ to lie within the unit ball
@@
 subtracting another convex term and the resulting function is therefore more complicated.}.
-A further simplification is that while \eqref{eq:edge_soft_early} requires to optimize
-$kd + |E| + |V|d$ parameters, we fix $a_{uv} = 1$ and $b_u = \onev$ for all edges and nodes in order
+Second, while \eqref{eq:edge_soft_early} requires to optimize
+$kd + |E| + |V|d$ parameters, we fix $a_{uv} = \deg(u)$ and $b_u = \zerov$ for all edges and nodes in order
 to be independent of the size of the graph. The final problem solved by our \combined{} method using
@@
   \log\Bigl(\sum_{\ell=1}^k\exp\left(\beta {s_{uv}}^T w_\ell \right)\Bigr)
-  - \frac{\mu}{|V|}\sum_{u\in V}
-  \left|\left| x_u  - \sum_{v \in \nei(u)}\!\! \mathcal{A}(w_1,\ldots,w_k) \right|\right|^2
+  - \frac{\mu}{|V|}\sum_{u\in V} \left|\left| x_u  - \frac{1}{\deg(u)}
+  \sum_{v \in \nei(u)}\!\! \mathcal{A}(w_1,\ldots,w_k) \right|\right|^2
 \end{equation}
@@
 Besides relaxing the $\max$ operator from \eqref{eq:edge_full}, another solution to remove it would
-be to increase $k$ to be equal to $E$. As we already mentioned, this makes the problem trivial. To
+be to increase $k$ to be equal to $|E|$. As we already mentioned, this makes the problem trivial. To
 avoid this, we add the constraint that all those $|E|$ directions are linear combinations of a small
@@
 
-At this stage, we could also add the node loss term \eqref{eq:edge_node_loss}, expressed in matricial
-form as $\|X - B -
-C\left(A\circ W^T \right)\|_F^2$, where $B\in \Rbb^{|V| \times d}$ is the node bias vectors $b_u$
-stacked vertically, $C \in \Rbb^{|V| \times |E|}$ is the incidence matrix of $G$, $A \in \Rbb^{|E|
-\times d}$ is the matrix whose each row is the corresponding $a_{uv}\onev$ vector and $\|\cdot\|_F$
-denotes the Frobenius norm. Note that this term is jointly convex in $A$, $B$ and $W$.\footnote{As
-one can check that the inequality defining convexity holds for this function.} However, for the
-clarity of exposition, we refrain to do so for now.
-
 The remaining issue is that $W\in \mathbb{M}^{d\times |E|}$ is not a convex constraint. As in the
@@
 relying solely on the nuclear norm constraint, we therefore add an additional regularization term
-equal to $\mu \sum_{i=1}^{|E|} \|W_{:,i}\|_2 = \mu \frop{W}{W}$. Our final convex formulation is
+equal to $\mu \sum_{i=1}^{|E|} \|W_{:,i}\|_2^2 = \mu \frop{W}{W}$. Our final convex formulation is
 then:
@@
   \min_{P\in\mathbb{M}^{d\times k}, Q\in\mathbb{M}^{|E|\times k}} \quad \tilde{h}(P, Q) =
-  - \frop{S}{PQ^T} + \mu \|X - B - C\left(A\circ QP^T \right)\|_F^2
+  - \frop{S^T}{PQ^T} + \mu \|X - B - C\left(A\circ QP^T \right)\|_F^2
 \end{equation}
+
+The second term corresponds to the node loss term of equation \eqref{eq:edge_node_loss}, expressed
+in matricial form as $\|X - B -
+C\left(A\circ W^T \right)\|_F^2$, where $B\in \Rbb^{|V| \times d}$ is the node bias vectors $b_u$
+stacked vertically, $C \in \Rbb^{|V| \times |E|}$ is the incidence matrix of $G$, $A \in \Rbb^{|E|
+\times d}$ is the matrix whose each row is the corresponding $a_{uv}\onev$ vector and $\|\cdot\|_F$
+denotes the Frobenius norm. As previously, we set $B=0$ and the rows of $A$ to be one over the
+degree of the node at the origin of the corresponding edge.
+
 While \eqref{eq:edge_PQ} is not jointly convex in both $P$ and $Q$, it is convex in one of
@@
   \frac{\partial \tilde{h}}{\partial P} &=
-  \frac{1}{|E|} S^T Q + \frac{2\mu}{|V|}
+  S^T Q + 2\mu
   A^T \circ (((X-B)^T -A^T \circ (PQ^T )C^T )C)Q 
@@
   \frac{\partial \tilde{h}}{\partial Q} &=
-  \frac{1}{|E|} S P + \frac{2\mu}{|V|}
+  S P + 2\mu
   A\circ (C^T (X-B-C(A\circ (QP^T ))))P.
@@
 number of edges. Indeed, computing the product $S^T Q$ requires $O(|E|^\omega)$ arithmetic
-operations, where $\omega \geq 2$\footnote{$\omega$ is the optimal exponent of matrix multiplication
-algorithm.}. This could be mitigated by using stochastic gradient descent. Finally, to obtain a
+operations, where $\omega \geq 2$ is the optimal exponent of matrix multiplication
+algorithm. This could be mitigated by using stochastic gradient descent. Finally, to obtain a
 solution of the \ecp{} \autoref{p:edge_full}, we take the directions to be the columns of the
@@
 k}$ to be the final directions found by \combined{} and the rows of $Q_0 \in \mathbb{M}^{|E|\times
-k}$ to be equal to the corresponding $e_{\yhat_{uv}}$\stodo{notation}. Furthermore, in the projection
+k}$ to be equal to the corresponding $e_{\mathcal{E}(u,v)}$. Furthermore, in the projection
 step of the $Q$ variable, we actually ensure that all the entries of $Q$ are positive, which makes

### 2017-11-14 edge/signed.tex
take feedback on edge chapter into account
- rewrite the setting part, giving the definition of s and g right away,
and adding a “prediction” operator
- try to motivate better the | x - b - \sum a⋅w | term in (1.2)
- acknowledge the signed case is only a sanity check (I couldn't find a
closed form solution for the noisy case)
- give a list of methods at the beginning of 1.2
- fix typos and math
- split the generation process in 4 steps and explicitly state its input
and output
- reorganize the results writeup in 4 parts: generation, measures,
initial AMI, commented tables

--- 2017-10-31 edge/signed.tex
@@
-\paragraph{Generalization of signed graphs}
+\paragraph{Connection with signed graphs}
 \label{par:generalization_of_signed_graphs}
 
-The learning bias of \autoref{chap:cc} implicitly defined an idealized \esp{} problem as:
+The learning bias of \autoref{sub:bias_balance} implicitly defined an idealized \esp{} problem as:
 \begin{problem}[idealized \esp{}]
@@
 instance of \autoref{p:edge_full} where $k=2$, and where the profiles and edges are based on group
-membership. We show that the optimal solution of such an instance is exactly the solution of this
-idealized \esp{} problem on the same graph.
+membership. We now check that the optimal solution of such an instance is consistent with the
+solution of this idealized \esp{} problem on the same graph.
 
-Let us first define this instance formally. Given a graph $G=(V,E)$, we partition the nodes in two
-groups $L$ and $R$. We then define the nodes profiles $X$ as follow. We first draw \uar{} a vector $p$
+Let us first define this instance formally. Given a graph $G=(V,E)$, call the $K=2$ groups of
+\autoref{p:edge_esp} $L$ and $R$. We then define the nodes profiles $X$ as follow. We first draw \uar{} a vector $p$
 from $\{-1, 1\}^d$, then let the profiles in $L$ be $\frac{p}{\sqrt{d}}$ and the
@@
   Let $\mathcal{D}_2 = \{w_1^\star, w_2^\star\}$ be an optimal solution of \autoref{p:edge_full} on the
-  instance $\mathcal{I}$. Then $\mathcal{D}_2$ provides the following optimal solution to
+  instance $\mathcal{I}$. Then $\mathcal{D}_2$ provides the following solution to
   \autoref{p:edge_esp}: each every edge explained by $w_1^\star$ is predicted positive, and every
@@
 
+This is nothing more than a sanity check, as the profiles explicitly encode the membership of each
+node to one of the two groups. Indeed, the proof makes no use of the dimension $d$ of the profiles,
+so that $d=1$ would be enough. Slightly more interesting is the case were those profiles are set up
+as before, but perturbed \uar{}. Namely, the sign of each coordinate is changed with probability $p
+< \shalf$. One can then check that
+\begin{equation*}
+  \E \left[ s_{\mathrm{inner}} \right] = \frac{(1-2p)^2}{d} \onev =
+  -\E \left[ s_{\mathrm{inner}} \right] 
+\end{equation*}
+and that furthermore, the variance of each component is $\frac{8(p-p^2) - 16(p-p^2)^2}{d^2}$. While
+the previous $\{w_1^\star, w_2^\star\}$ solution is now only optimal in expectation, we see that the
+variance decrease quadratically with the dimension of the profiles. In practice we thus expect this
+solution to be very close to the optimal. Intuitively this can be seen as taking the majority of $d$
+votes on whether each node belongs to one group or the other.
+
 \begin{aside}
@@
   the edges between inner and outer groups.
-
-  The next thing I tried was to use the approach described later in \autoref{sub:edge_generation} in
-  order to optimize the profiles of a complete graphs with $K$ groups, given $w_+$, $w_-$ and the
-  inner/outer edges assignment. As one could expect from the symmetry of the problem, all the nodes
-  in the same group have the same profile. However, there are no clear pattern in those profile, and
-  once again there exists better solutions than $\{w_+, w_-\}$, which partition the edges in two
-  roughly balanced sets.
 \end{aside}

### 2017-11-19 intro/learning.tex
first draft of the intro's intro plus first section

The birth of graph theory in 1736 is credited to \textcite{euler1741} for his elegant solution of
the Seven Bridges of Königsberg
problem\footnote{\url{https://en.wikipedia.org/wiki/Seven_Bridges_of_Königsberg}}. Since then,
it has been a very active branch of mathematics ~\autocite{biggs1976graph}. Indeed, it provides a
conceptually simple yet immensely rich framework to model phenomena where entities are connected
with each other~\autocite{ComplexNetworksApp11}. Coupled with the increasing availability of large
amount of relational data, learning on graphs has recently spurred a lot of interest across various
lines of research, with tangible benefits:

\begin{description}[leftmargin=*]

  \item[community detection~\autocite{FortunatoSurvey10}] The goal is to cluster nodes in tightly
    connected groups that are loosely connected with the rest of the graph. This allows for a better
    understanding of the graph organization, and present a higher level view than looking at
    individual nodes relationships. For instance, it has been used to identify proteins functional
    groups~\autocite{clusterBio03} or see how different scientific fields relate based on
    publication data~\autocite{clusterScience08}.

  \item[semi-supervised learning~\autocites{SSL06}{graphSSL14}] In addition to labeled data,
    the learner is also provided with unlabeled data at training time, and its goal is to classify
    nodes. Connecting similar instances allow propagating information along the graph. This has
    found applications in classifying text documents~\autocite{sslText09} or aligning categories and
    relations across multiple knowledge bases~\autocite{sslKB13}.

  % \item node classification~\autocite{nodeClassif11} Idem
    % could add visualisation (Roberto Tamassia. 2013. Handbook of Graph Drawing and Visualization.
    % CRC Press.) but it's not exactly learning

  \item[node embedding~\autocite{representationLearning17}] The goal is to find a low dimensional
    representation of the nodes, based on their structural patterns. This usually performed in an
    unsupervised way, although it is also possible to include problem supervision when available.
    Such representation can then be used in downstream tasks, for instance
    visualization~\autocite{LINE15} or the aforementioned node classification, even in the inductive
    setting where new nodes can join the graph after training~\autocite{inductiveRepresentation17}.

  \item[link prediction~\autocite{linkPredSurvey16}] Given a snapshot of the graph at time $t$, the
    goal of link prediction is to return a set of links that do not exist at time $t$ but will be
    created by time $t+\Delta_t$. Most methods are based on the assumption that link creation is
    driven by node similarity. It has been successfully applied to inferring potential interactions
    between proteins without expensive experiments~\autocite{linkPredBio06} and uncovering hidden
    associations in criminal networks~\autocite{linkPredCrime08}.

  \item[information and influence propagation~\autocite{infmax13}] The study of processes by which
    content is spread across networks, and how such processes can be influenced to speed them up or
    slow them down. Two representative applications are selecting the best seeds in a social network
    to promote a viral marketing campaign~\autocite{infmaxKempe15} and containing more effectively
    the diffusion of actual biological viruses~\autocite{influenceBio13}.

  \item[network evolution~\autocite{networkEvolution14}] These methods focus on the mechanisms and
    consequences of the growth of networks, as well as how to keep the results of some data mining
    algorithms up to date and relevant. Monitoring the changes in the interactions of proteins
    can indeed be used as an early indicator for some kind of diabetes~\autocite{evolBio10}.
    Furthermore, sudden changes in a network of computers are usually worth investigating, for they
    might signal external attacks~\autocite{evolSecurity04}.

\end{description}

This list of graph learning problems and their applications to real world scenario is incomplete.
Yet it already demonstrates the impact of inferring patterns in relational data over many aspects of
our lives. However, we argue that more can be done. Indeed, all the methods presented above only
consider graphs with a single type of edge and where nodes are connected because based on some
domain-dependant notion of similarity.

### 2017-11-19 intro/main.tex
first draft of the intro's intro plus first section

Graphs are a natural way to represent the relationships between a set of entities. Because of the
simplicity and flexibility of this formalism, graphs are ubiquitous and have been used in countless
fields. To make the rest of our explanations more concrete, we now give three examples of graphs
that we consider later on.
\begin{enumerate}[nosep]%,leftmargin=*]
  \item In a social network\footnote{In the rest of this thesis, we use the terms \emph{graph} and
    \emph{network} interchangeably.}, the nodes of the graph are human users, and relationships
    between nodes denote interactions between the corresponding users. One social network we study
    in \autoref{chap:troll} summarizes the connections between the editors of \wik{}. These editors
    can be promoted to administrators after a vote from their peers. An edge \euv{} in this network
    means that user $u$ has voted on the possible promotion of $v$.
  \item In computer vision, we can represent an image as a graph. Each pixel is a node, and those
    nodes are connected to four neighbors, namely the adjacent pixel from the top, bottom, left and
    right sides.
  \item In e-commerce, we can consider a co-purchase network. Nodes are products being sold on a
    website, and two nodes are connected if the corresponding products are frequently bought
    together by customers, for instance a phone and a memory card.
\end{enumerate}

Not only are there graphs in many domains, but the progress of technology in the last few decades
has made it easier to collect many graphs in every single domain, sometimes with up to billion of
nodes and hundred of billions of edges. The availability of such large amount of structured data has
prompted the development of automated methods to extract insight from it. For instance, it is
possible to cluster nodes into coherent groups, predict the category of the nodes or study how to
best propagate information within a graph. We describe in more details these possibilities and
others in \autoref{sec:learning_in_graphs}. By doing so, our goal is to illustrate the wide
potential of \emph{learning in graphs}.

At this point of the discussion though, we have only considered the most common and simple kind of
graph, one representing a single type of relation between nodes. However, in many situations, nodes
have more than one type of relation among each other. As a case in point, let us look again at our
three examples. In the \wik{} network, a vote can support the promotion or oppose it. This
additional information enrich our understanding of the relationships between the nodes in the graph.
Likewise in an image, an edge between two pixels is positive if the two pixels belong to the same
object (say a car or a building) and negative otherwise. Such information could be used to segment
the image. In the co-purchase network, there is even more than two types of relation. Assuming the
products are movies, two movies can be frequently bought together because they are part of a series
(like Star Wars), because they have the same director, because they receive the same prestigious
awards in different years, and so on. In \autoref{sec:edge_semantics}, we give additional examples
of so-called \emph{complex networks}, where edges have different \emph{semantics}, or types. As
showed in our examples, we distinguish between two cases. The first is when there are two types of
edge having opposite semantics. Such graphs are called \emph{signed graphs} and have been
extensively studied since the fifties, for we shall see they have many applications. The second case
is when there are more than two types of edge. Such graphs have also been studied for a long time
under different names and we refer to them as \emph{multilayer graphs}.

It is natural to expect that we can extract finer insights from graphs with edge semantics. However,
in many cases, the information about edge types is not available. We therefore present in
\autoref{sec:predicting_edge_type} the problem of characterizing edges in complex networks.
Informally, given an input graph and possibly some extra information, we want to predict the type of
every edge. This problem can take several forms depending of what information is available as input.
First, the graph may be directed or undirected. Second, there might be two edge types (in the case of
signed graphs) or more (for general multilayer graphs). Third, the problem can be supervised or not.
In the supervised setting, we are provided with labels for some of the edges. In the unsupervised
setting, there is no label. We thus make the additional assumption that we observe the
\emph{attributes} of the nodes. Once again, we return to our three examples to illustrate what are
those attributes. In the \wik{} graph, we could have for each user data about age, experience and
area of expertise. In an image, each pixel is associated with a color, along with higher order
visual features. Each product of a co-purchase network comes with information about \eg{}, its price,
popularity, category and availability. In the absence of label, our intuition is that these
attributes can inform us about the type of edge between nodes.

We list in \autoref{sec:outline} three concrete instantiations of this general problem that we
consider in this thesis. Our solutions to these three problems offer evidence in the defense of our
thesis statement:
\begin{quote}
  \bf
  There exists efficient and accurate methods to predict edge type in complex networks, relying
  only on the graph topology or also on node attributes.
\end{quote}

\section{Learning in graphs}
\label{sec:learning_in_graphs}
\input{learning}

\section{Graph with several edge semantics}
\label{sec:edge_semantics}
\input{semantics}

\section{Predicting edge type}
\label{sec:predicting_edge_type}
\input{predicting}

\section{Outline}
\label{sec:outline}
\input{outline}

### 2017-11-21 intro/main.tex
say complex networks have 2 features: k edge types and homo/hetero mix

@@
 
+\medskip
+
 At this point of the discussion though, we have only considered the most common and simple kind of
-graph, one representing a single type of relation between nodes. However, in many situations, nodes
-have more than one type of relation among each other. As a case in point, let us look again at our
+graph, one representing a single type of relation between nodes. However, in many situations, there
+are two dimensions along which graphs exhibit more complexity. First, nodes have more than one type
+of relation among each other. Second, two nodes are not only connected because of their global
+similarity but also for more nuanced reasons. As a case in point, let us look again at our
 three examples. In the \wik{} network, a vote can support the promotion or oppose it. This
 additional information enrich our understanding of the relationships between the nodes in the graph.
+It also points out that two editors can be connected because they share a common topic of interest,
+but come from different socio-economic backgrounds and therefore cannot agree on this topic.
 Likewise in an image, an edge between two pixels is positive if the two pixels belong to the same
-object (say a car or a building) and negative otherwise. Such information could be used to segment
-the image. In the co-purchase network, there is even more than two types of relation. Assuming the
+object (say a car or a building) and negative otherwise. This information could be used to segment
+the image. In the co-purchase network, there are even more than two types of relation. Assuming the
 products are movies, two movies can be frequently bought together because they are part of a series
-(like Star Wars), because they have the same director, because they receive the same prestigious
-awards in different years, and so on. In \autoref{sec:edge_semantics}, we give additional examples
-of so-called \emph{complex networks}, where edges have different \emph{semantics}, or types. As
-showed in our examples, we distinguish between two cases. The first is when there are two types of
+(like Star Wars), because they have the same director but different gender, because they receive the
+same prestigious awards in different years, and so on. The second type of relation (\enquote{same
+director} and \enquote{different gender}) is actually an example of a mixed relation, for it mixes
+similarity and dissimilarity.
+
+We call \emph{complex networks} graphs where edges have different \emph{semantics}, or types, and
+where connections are explained by more than simple global similarity. In
+\autoref{sec:edge_semantics}, we give additional examples of such complex networks. As
+showed in our three previous examples, we distinguish between two cases. The first is when there are two types of
 edge having opposite semantics. Such graphs are called \emph{signed graphs} and have been
@@
 under different names and we refer to them as \emph{multilayer graphs}.
+
+\medskip
 
@@
 In the supervised setting, we are provided with labels for some of the edges. In the unsupervised
-setting, there is no label. We thus make the additional assumption that we observe the
-\emph{attributes} of the nodes. Once again, we return to our three examples to illustrate what are
-those attributes. In the \wik{} graph, we could have for each user data about age, experience and
+setting, there is no label at all. We thus make the additional assumption that we observe some
+\emph{attributes} of the nodes. Once again, we return to our three examples to illustrate what
+those attributes can be. In the \wik{} graph, we could have for each user data about age, experience and
 area of expertise. In an image, each pixel is associated with a color, along with higher order
-visual features. Each product of a co-purchase network comes with information about \eg{}, its price,
+visual features. Each product of a co-purchase network comes with information about, \eg{}, its price,
 popularity, category and availability. In the absence of label, our intuition is that these
 attributes can inform us about the type of edge between nodes.
+
+\medskip
 

### 2017-11-23 intro/semantics.tex
examples of signed graphs in intro

The three graphs we described in the introduction are examples of what we call complex networks.
Such graphs have more than one type of edge, and edges are not explained by mere global node
similarity. The point of this section is to illustrate that this notion is not simply a mathematical
variation of a \enquote{simple} graph, where the adjacency matrix would take values in $\{-1, 0, 1\}$ or
$\{0, 1, \ldots, k\}$ instead of $\{0, 1\}$. Rather, we reviews many applications in various domains
where complex networks are the right model to represent rich networked data. First, we showcase many
uses of signed networks. In that case, we do not insist on the fact that similarity is not the only
driver of node connections, for it is implicit that negative edges actually denote dissimilarity.
Second, we present multilayer graphs, that is graphs with more than two types of edges. When
applicable, we highlight how the connections in such graphs indeed rely on a nuanced similarity or
dissimilarity across some of the nodes attributes.

To elaborate on this last point, our hypothesis is that connections in complex networks are the
results of both partial homophily (that is, nodes are connected when they are similar on a subset of
the attributes) and partial heterophily (that is, nodes are connected when they are dissimilar on a
subset of the attributes). As examples of the latter, think of dating websites ---where most users
are linked with users of the opposite gender~\autocites{homophilyMyspace09}{Tinder16}; diffusing
innovations ---where meeting people with different backgrounds and point of views is crucial to
favor diversity and creativity~\autocite{rogers2003diffusion}; and online news consumption ---where
connecting people from different sides of the political spectrum helps to avoid echo chambers and
fuel a democratic debate~\autocite{balancedNews17}.

\subsection{Signed graphs}
\label{sub:intro_signed_graphs}

In this section, we present a list of signed graph usages, sorted by domains. Many of those signed
graphs are the input of some clustering algorithms. In the context of signed graph, the clustering
task can be captured by the \pcc{} problem. We provide a thorough overview of this problem in
\autoref{chap:cc}. Here we simply give a broad, informal definition. The objective in \pcc{} is to
cluster the nodes of a signed graph in a way that minimizes the number of positive edges across
clusters and the number of negative edges within clusters. Such edges are called disagreements.

\input{use_of_sign_graphs.tex}

### 2017-11-23 intro/use_of_sign_graphs.tex
examples of signed graphs in intro

@@
 %TODO turn those items into paragraphs?
-\begin{description}
-   \item[Computer Vision]
+% \begin{description}[leftmargin=*]
+   \paragraph{Computer Vision}
       The ubiquitous task of segmenting an image into homogeneous regions is a prerequisite for many
-      further processing, and can be tackled with \pcc{}. For instance, to segment cell in
+      further processing. As we mentioned earlier in one example, building a signed graph can help,
+      although it might be costly to do it at the pixel level. For instance, to segment cell in
       microscopy imagery, \textcite{CellSeg14} first use generic image features to classify pixel in
@@
 
-   \item[Natural Language Processing] 
+      \paragraph{Natural Language Processing} 
       Coreference resolution is the task of finding all expressions that refer to the same entity in
       a text. Like image segmentation, it is a preprocessing step that can later be used in document
-      summarization, question answering, and information extraction. \Textcite[Section
-      2.3]{graphicalCoreference04} tackle coreference resolution by using an undirected graphical
+      summarization, question answering, and information extraction. Furthermore, in that case, it
+      is also natural to build a graph of words. One then add negative edges between words that
+      cannot refer to the same entity (for instance because they are of different gender) and
+      positive edges between words with linguistic cues indicating they might refer to the same
+      entity. 
+      \Textcite[Section
+      2.3]{graphicalCoreference04} instead tackle coreference resolution by using an undirected graphical
       model on which performing inference is equivalent to \pcc{}. On small scale instances,
-      \textcite{Elsner2009} first obtain an upper bound of the optimal solution by solving a SDP
+      \textcite{Elsner2009} use the signed graph procedure outlined above. They first obtain an
+      upper bound of the optimal solution by solving a SDP
       relaxation of the problem. They then compare various heuristics and show that best performing
@@
       step, such as the Best One Element Move~\autocite{Gionis2007}. Further NLP tasks amenable to
-      \pcc{} formulation are referenced in their paper. Another possible task would be 
-      to cluster words based on distributional embedding vector while adding antonym
-      constraints~\autocite{SignedWordRatings}, although the authors of that paper choose to use
-      signed spectral clustering.
+      a signed graph representation are referenced in their paper. Another task is
+      to cluster words based on distributional embedding vectors while adding antonym
+      constraints~\autocite{SignedWordRatings}.
 
-   \item[Biology]
-      Clustering tasks also abounds in Biology.
+      \paragraph{Biology}
+      Signed graphs are also abundant in Biology.
       A typical input is a similarity matrix between genes expression level in various experimental conditions
@@
       number of edges to remove in order to decompose a dynamics system into a collection of
-      monotone system. This corresponds to the \mind{} problem and allows to study
-      the complete system more easily. Many more applications of \mind{} on weighted complete graphs
+      monotone system. This allows to study
+      the complete system more easily. More applications of weighted complete signed graphs in biology
       are presented in~\autocite[Section 6]{clusterEditSurvey13}.
@@
 
-   \item[Network science]
-      As mentioned in the first Chapter, one early use of signed graphs was to model social
-      interactions and given the large body of work on community
-      detection~\autocite{FortunatoSurvey10}, it is therefore not surprising that \pcc{} has been
-      used to study such networks. For instance, one can extract all the votes of the members of a
+      \paragraph{Network science}
+      One early use of signed graphs was to model social
+      interactions~\autocites{harary1953}{HeiderBook58}. Here we present
+      some recent references. For instance, one can extract all the votes of the members of a
       political parliament and form a graph whose nodes are politicians and edge weights quantify how
@@
       the information contained within the negative links to enhance the visualization of social
-      networks~\autocites{Luca10}{drawingSignedGraphs11}. Note however that those four last papers
-      not explicitly use a \pcc{} formulation.
+      networks~\autocites{Luca10}{drawingSignedGraphs11}.
 
-   \item[Others]
+      \paragraph{Others}
       \begin{itemize}[leftmargin=*]
@@
 	    On a high level, a solution to this problem is to build a graph of all the available
-	    objects, define a similarity between them and run a \pcc{} algorithm. This was indeed
-	    this kind of problems at Whizbang! Labs that partly motivated one of the early \pcc{}
-	    paper~\autocite{Bansal2002}. The main challenge
+	    objects, define a similarity between them and run a \pcc{} algorithm.
+	    % This was indeed this kind of problems at Whizbang! Labs that partly motivated one of the early \pcc{} paper~\autocite{Bansal2002}.
+	    The main challenge
 	    thus lies in devising an appropriate similarity measure, given that object can have very
@@
 	    for a two-clustering of the nodes that will minimize the number of disagreements.
+	 \item In finance, one can represent an investment portfolio as a signed
+	    graph~\autocite{portfolio02}. Each node is a security, and the edge between two
+	    securities is weighted by their correlation, which can be negative. For instance, a
+	    graph with only positive edges is speculative, as all the securities move in the same
+	    direction, either up or down. On the other hand, if the securities can be partitioned in
+	    groups without disagreement, the risk is limited, for two clusters will move in opposite
+	    directions, providing the investors with a hedging guarantee.
+	 \item In wireless networks, signed graphs can be used to solve optimization problems
+	    involved in determining energy-saving routes~\autocite{signedRouting12} or to exchange
+	    cryptographic keys in a secure and efficient manner~\autocite{signedKey17}.
       \end{itemize}
-\end{description}
+% \end{description}

### 2017-11-25 edge/intro.tex
take feedback on edge chapter into account
- rewrite the setting part, giving the definition of s and g right away,
and adding a “prediction” operator
- try to motivate better the | x - b - \sum a⋅w | term in (1.2)
- acknowledge the signed case is only a sanity check (I couldn't find a
closed form solution for the noisy case)
- give a list of methods at the beginning of 1.2
- fix typos and math
- split the generation process in 4 steps and explicitly state its input
and output
- reorganize the results writeup in 4 parts: generation, measures,
initial AMI, commented tables
example of multilayer

@@
 citation~\autocites{communityDBLPbyConf05}{articlesMultiSim11}, economic
-exchange~\autocites{ports3kindofships10}{worldTradeNetwork10} and biology~\autocite{bioLayerExp11}.
+exchange~\autocites{ports3kindofships10}{worldTradeNetwork10}{KantPeace15} and
+biology~\autocites{bioLayerExp11}{EcoChile15}{Neuroscience16}.
 This additional information on the nature of the relationships allow traditional graph tasks to be
@@
 data with both node attributes and edge types proved difficult, we describe a synthetic model in
-\autoref{ssec:edge_exp}, on which we perform extensive experiments. We then survey in
+\autoref{sec:edge_exp}, on which we perform extensive experiments. We then survey in
 \autoref{sec:edge_related} works that are related to the \ecp{}, while noting that very few address

### 2017-11-25 intro/semantics.tex
example of multilayer graphs in intro

@@
 \input{use_of_sign_graphs.tex}
+
+\subsection{Multilayer graphs}
+\label{sub:intro_multilayer_graphs}
+
+Besides signed graphs, in this thesis we also consider at multilayer
+graphs~\autocites{Kivela2014}{multiSurvey14}. Those are graphs with $k$ edge types, and the name
+refer to the fact we can see them as the superposition of the $k$ subgraphs induced by each edge
+type. Even when $k=2$, we make a distinction between signed graphs and multilayer graphs. In signed
+graphs, the two types of edge have reverse semantic, whereas in a 2-layers graph, they simply
+denotes two possible interaction, for instance advisor-advisee or regular coauthors relation in a
+citation network~\autocite{Advisor10}. In general though, we focus on cases where $k$ is larger than
+$2$, and not larger than a few dozen in order to preserve interpretability.
+Like signed graphs, these multilayer graphs are versatile enough to be used in many fields.
+
+\paragraph{Social networks} \Textcite{Szell2010} model the interactions of the players of
+\emph{Pardus}, a massively multilayer online game. These players can be friend or enemy, send
+private messages, trade resources, attack each other and set a bounty on the head of another
+players. These six types of interactions are either positive or negative, but their nuances cannot
+simply by explained by global similarity and dissimilarity. Another examples are the users of the
+photo sharing website Flickr.  They can interact in eleven ways, either directly or through
+comments, shared tags, groups membership and so on~\autocite{RecoFlickrMulti11}. Again, while being
+part of a common group denote shared interests between two users, overall they must also differs in
+some other attributes (for instance location) in order to bring diversity to this groups.
+
+\paragraph{Citation networks} In these networks, nodes are research papers or authors, and edges
+typically connect two nodes whenever one cites the other. For instance, using the DBLP dataset,
+\textcite{communityDBLPbyConf05} connect two authors if they have co-authored a paper in one the
+\np{1000} conferences appearing in the data. One can then consider that the resulting graph was
+obtained as the superposition of these \np{1000} subgraphs. Besides direct citations,
+\textcite{articlesMultiSim11} consider four others reasons to connect \np{5000} SIAM papers, based
+on their similarity in terms of abstract words, title words, keywords and authors.
+
+\paragraph{Economic networks} In our ever increasingly globalized economy, entities around the world
+are getting more and more tightly connected. However, it would be overly simplistic to consider
+there is only kind of link between entities. For something seemingly as simple as the connections
+between the largest \np{951} ports, one must already noticed that these connections can be
+implemented by any combinations of three kinds of ships: bulk dry carriers, container ships and oil
+tankers~\autocite{ports3kindofships10}. Likewise, the 162 countries of the International Trade
+Network are connected by 96 kinds of commodities they can exchange~\autocite{worldTradeNetwork10}.
+Finally, \textcite{KantPeace15} studies international relations through the lens of Kant three
+folded program for peace,  based on democracy, economic dependence and supra national governance.
+They build the graph of all countries and connect them in three layers. All democracies form a
+clique in the first layer, countries are connected with weights proportional to the amount of yearly
+trade in the second layer and with weights proportional to the number of international organisations
+they belong to together in the third layer. While the essence of commerce is to exploit differential
+between the partners involved (\ie{} heterophily), trade (and in the last example participating in
+common institution) intensity also involve geographical, historical and cultural ties (\ie{}
+homophily).
+
+\paragraph{Biological networks} The interactions in biology also take several forms and studying
+them as a whole has proved fruitful. One example is a genes co-expression network, where each
+connection was tested under 130 different experimental conditions, providing as many
+layers~\autocite{bioLayerExp11}.
+Multilayer graphs are also a relevant way to represent ecological
+networks~\autocite{EcologyMultiReview17}. For instance, \textcite{EcoChile15} build the graph of
+more than 100 species living on the Chilean
+coast\footnote{\url{http://app.mappr.io/play/chile-marine-intertidal-network}} and divide their
+interactions in three categories: trophic (that is one specie eats another one), positive
+non-trophic (\eg{} refuge providing) and negative non-trophic (\eg{} competition for shared
+resources).
+Finally, in neuroscience, multilayer graphs have recently emerged as a useful tool to better
+understand the human brain~\autocite{Neuroscience16}. The nodes of such graphs are neurons, and the
+edges can be labeled in various ways: some correspond to actual physical links while others are
+functional ---\ie{} neurons responding in the same way to external stimuli, some are present in
+healthy subject and others in patient cohort, some are acquired through MRI and others by EEG.

### 2017-11-25 intro/learning.tex
typo

@@
-The birth of graph theory in 1736 is credited to \textcite{euler1741} for his elegant solution of
+The birth of graph theory is credited to \textcite{euler1741} in 1736 for his elegant solution of
 the Seven Bridges of Königsberg
@@
 amount of relational data, learning on graphs has recently spurred a lot of interest across various
-lines of research, with tangible benefits:
+lines of research, with tangible benefits.
 
@@
 our lives. However, we argue that more can be done. Indeed, all the methods presented above only
-consider graphs with a single type of edge and where nodes are connected because based on some
+consider graphs with a single type of edge and where nodes are connected based on some
 domain-dependant notion of similarity.

### 2017-11-26 intro/outline.tex
first draft of the last two sections of the intro

The contributions of this thesis are organized as follow:

\begin{itemize}

  \item We start in \autoref{chap:troll} by addressing the \esp{} problem in \dssn{}. Our goal is to
    design a method that is scalable, principled and accurate. For that, we start by introducing a
    generative model for signs, and derive approximations of the optimal Bayes predictor or the
    maximum likelihood predictor. We confirm the theoretical soundness of these approaches by
    performing extensive synthetic and real world experiments. Finally, we are also the first to our
    knowledge to give an online algorithm for the \esp{} problem on directed graphs. This chapter is
    based on an existing publication~\autocite{trollSign17}.

  \item We then move from directed to undirected signed graphs, and explain in \autoref{chap:cc} why
    this requires another learning bias. Namely, we assume that the nodes belong to $k$ groups, and
    that the sign between two nodes is positive when both nodes belong to the same group, negative
    otherwise. We describe how this new bias is related to the \pcc{} problem, and present a
    thorough overview of existing approaches. Finally, we provide the first implementation of an
    existing spanner construction~\autocite{gtxFabio}, and give a preview of its performance on
    synthetic and real signed graphs.

 \item In \autoref{chap:vector}, we consider the \ecp{} problem on multilayer graphs with node
   attributes. Our approach is to seek a small number of vectors that, once assigned to every edge,
   best explain the graph (\ie{} maximize a score function between the vector assigned to an edge
   and the profiles of this edge's endpoints). From this initial formulation, we derive two
   optimization problems, and initially solve them on synthetic data, generated with a few natural
   topological constraints. We compared these solutions with the result of a baseline approach that
   simply clusters the pairwise profile vectors using $k$-means.

  \item Finally, reflecting on our treatment of the three previous problems, we discuss in
    \autoref{chap:conclusion} other settings and methods that might extend the problem of
    characterizing edges in complex networks beyond the frame of this thesis.

\end{itemize}

### 2017-11-26 intro/predicting.tex
first draft of the last two sections of the intro

Let us summarize in one sentence the two previous sections. Learning in graphs provides many
insights, and many graphs are complex, in the sense of having edges expressing different semantics
and created for mixed reasons. The logical conclusion is that we want to learn in complex graphs.
However, this requires the edges of such graphs to be labeled with one of $k$ types of relationship.
In the examples presented above, this was already done. Yet we argue that efficiently labeling the
edges is an interesting problem, especially in the following three situations:

\begin{enumerate}

  \item There exists a function that can perfectly label any edge because it is tailored to this
    specific graph, but calling it is expensive. For instance, say we want to know whether two
    connected users of a social network are tied to each other through family, work, school or
    hobby. The labeling function in this case simply asks users to label their relationships,
    assuming the answers do not contain any noise. At the scale of Facebook, this would require
    asking hundred of questions to each user on average, which is quite time consuming. Furthermore,
    it would also cost marketing resource to convince users this is beneficial for them, and
    engineering time to ensure this information remain confidential. Likewise in a biological
    network, determining whether two proteins interact positively or negatively with each other is
    achieves by a lab experiment, which requires time and material.

  \item All edges are already labeled, perfectly and without any cost, but the number $k$ of edge
    types is very large. Indeed, we have seen examples where they are hundreds or even thousands of
    edge types. Similarly to what happen in dimensionality
    reduction~\autocite{DimensionReduction10}, to make intuitive sense of such data, we want to
    reduce the number of edge types to less than ten. A natural way is to ask a domain expert to
    cluster edge types together. However, this is again time consuming, and does not necessarily
    make use of the topology of the graph.

  \item The input graph is actually unlabeled. That is, we observe interactions between the nodes,
    and we assume from domain knowledge that these interactions fall into $k$ categories. However,
    there is no reasonable way to come up with a specific labeling function. In
    \autoref{tab:edge_apps} \vpageref{tab:edge_apps}, we present several such examples, but for now
    we simply recall our earlier co-purchase network. As we mentioned, there are several reasons why
    two products could be bought together. Yet it is unlikely that customers will provide this type
    of feedback, for it does not bring them immediate advantage.
\end{enumerate}

In this thesis, we consider three versions of the problem of predicting edge types.
\begin{enumerate}

  \item The \esp{} problem, which takes as input the topology of a directed signed graph, and the
    label of some of the edges. The output is the label of the remaining edges. Therefore, it can be
    seen as a supervised binary classification problem.

  \item The same \esp{} problem, where the input graph is undirected. In this case, we consider an
    active variant where instead of being given a random training set of labels, we have a budget of
    queries we can use to request arbitrary labels.

  \item The \ecp{} problem, which takes as input an unlabeled, undirected multilayer graph, the
    attributes of all its nodes, and a number of edge types $k$. The output is a $k$ clustering of
    the edges, and a set of $k$ vectors describing the clusters in terms of node attributes.
    Therefore, it can be seen as an unsupervised clustering problem with side information.
\end{enumerate}

We now review existing approaches addressing these problems, in order to highlight our contribution
in the next section.

The modern formulation of the \esp{} on directed graphs can be attributed to
\textcite{Leskovec2010}. Their idea is to compute local features of the nodes based on the training
labels. Such features include variation on the node degree, for instance the number of positive
outgoing edges or the number of negative incoming edges. These node features are combined into edge
features, and a supervised classification algorithm is trained. Several works have devised
additional features~\autocites{Bayesian15}{Yuan2017}, most notably based on scoring the nodes in a
Page Rank fashion~\autocites{traag2010exponential}{shahriari2014ranking}{wu2016troll}. These methods
are accurate, fast ---for they are local, and interpretable ---for the features are hand crafted.
The drawback is that this feature engineering is done mostly in an ad hoc way. On the other hand, it
is also possible to look at the problem from a global point of view. Namely, by completing the
signed adjacency matrix through low rank
factorization~\autocites{LowRankCompletion14}{OnlineCompletion17}. These methods are also accurate
and bypass feature engineering, but by nature, they require careful algorithms to scale with larger
networks. Finally, during the writing of this manuscript, several papers have been published, which
use embedding of the nodes in a low dimensional space, based on the training
label~\autocites{SIGNet17}{SNE17}. This also avoids feature engineering but is global in nature and
not so interpretable.

One way to solve the problem on undirected graphs would be to apply the previous methods, having
first replaced every edge $(u,v)$ by two directed edges \euv{} and \evu{}. This is not quite
satisfactory, both from a theoretical but also practical point of view (see
\autoref{sub:need_for_a_directed_graph}). Instead, \textcite{Cesa-Bianchi2012b} draw a connection
between the \esp{} problem and the \pcc{} problem. Recall that given a fully labeled signed graph,
the solution to \pcc{} is a partition of the nodes that minimize the number of disagreement
edges~\autocite{Bansal2002}. \Textcite{Cesa-Bianchi2012b} assume that the signs are originally
consistent with an underlying, hidden $2$-clustering, but we only observe the signs after they have
been flipped \uar{}. In this case, they show that the optimal number of disagreements is a lower
bound of the number of mistakes of any active \esp{} algorithm. One natural approach would then be
to solve \pcc{} based on the observed signs, and predict the remaining signs consistently with the
inferred clusters. At first it seems hopeless, as \pcc{} is difficult to approximate on general
graphs~\autocite{Charikar2003}, even when there are only two clusters~\autocite{Giotis2006}.
However, this worst case analysis does not forbid more positive results on signed graphs that
exhibit stability under perturbation~\autocites{clusteringFeasibility15}{StableCC09}{StableLP09} or
are obtained through perturbations from an ideal case~\autocites{plantedAilon09}{Makarychev2014}.
Furthermore, in the active setting, the learner gets to choose which signs are observed. The general
idea of \textcite{Cesa-Bianchi2012b} is thus to use the query budget to build fully labeled paths
between each connected nodes. Those paths must be as short as possible, since the predicted sign of
$(u,v)$ is the product of the signs along the path from $u$ to $v$. Therefore, the shorter the path
and the less influence of the random perturbations. At the same times, those paths must span the
whole graph. This is the topic of an active research area~\autocites{Abraham2012}{Spanner17}.

As for the \ecp{} problem, to the best of our knowledge, it has not been studied under our
assumptions. A more common problem in attributed graphs is to cluster nodes into
communities~\autocites{LeskovecEgo12}{Yang2013}{Xu2014}{ZhangModelFree16}. However, it is not
immediate how such methods, generally based on generative models, could be adapted to our problem.
Direct approaches to classify edges have been proposed, based on graphical
models~\autocite{graphicalModelTies11}, nearest neighbors with a customized
distance~\autocite{Aggarwal2016a} and edge embedding in knowledge graphs~\autocite{transE13}. Yet
they all rely on having training label as supervision. A simple unsupervised method is to cluster
the line graph of the input graph~\autocite{LineGraph09}, but this does not take advantage of
attributes. Another method uses topological features~\autocite{ahmed2017roles} but is rather complex
and thus not very interpretable. If we further assume that each edge type is associated with an
Euclidean space, and that the position of nodes in one space are not correlated with their position
in another space, then it is possible to recover an approximation of these Euclidean metrics in
polynomial time without any supervision~\autocite{Abraham2012a}.

### 2017-11-27 intro/outline.tex
typos

@@
     maximum likelihood predictor. We confirm the theoretical soundness of these approaches by
-    performing extensive synthetic and real world experiments. Finally, we are also the first to our
-    knowledge to give an online algorithm for the \esp{} problem on directed graphs. This chapter is
-    based on an existing publication~\autocite{trollSign17}.
+    performing extensive synthetic and real world experiments. Finally, to the best of our
+    knowledge, we are the first to give an online algorithm for the \esp{} problem on directed
+    graphs. This chapter is based on an existing publication~\autocite{trollSign17}.
 
@@
    and the profiles of this edge's endpoints). From this initial formulation, we derive two
-   optimization problems, and initially solve them on synthetic data, generated with a few natural
-   topological constraints. We compared these solutions with the result of a baseline approach that
-   simply clusters the pairwise profile vectors using $k$-means.
+   optimization problems, and show how to solve them on synthetic data.
 

### 2017-11-27 intro/predicting.tex
typos

@@
     it would also cost marketing resource to convince users this is beneficial for them, and
-    engineering time to ensure this information remain confidential. Likewise in a biological
+    engineering time to ensure this information remain confidential. Likewise, in a biological
     network, determining whether two proteins interact positively or negatively with each other is
-    achieves by a lab experiment, which requires time and material.
+    achieved by a lab experiment, which requires time and material.
 
@@
 \textcite{Leskovec2010}. Their idea is to compute local features of the nodes based on the training
-labels. Such features include variation on the node degree, for instance the number of positive
+labels. Such features include variations on the node degree, such as the number of positive
 outgoing edges or the number of negative incoming edges. These node features are combined into edge
@@
 The drawback is that this feature engineering is done mostly in an ad hoc way. On the other hand, it
-is also possible to look at the problem from a global point of view. Namely, by completing the
+is also possible to look at the problem from a global point of view, by completing the
 signed adjacency matrix through low rank
@@
 use embedding of the nodes in a low dimensional space, based on the training
-label~\autocites{SIGNet17}{SNE17}. This also avoids feature engineering but is global in nature and
+labels~\autocites{SIGNet17}{SNE17}. This also avoids feature engineering but is global in nature and
 not so interpretable.
@@
 distance~\autocite{Aggarwal2016a} and edge embedding in knowledge graphs~\autocite{transE13}. Yet
-they all rely on having training label as supervision. A simple unsupervised method is to cluster
+they all rely on having training labels as supervision. A simple unsupervised method is to cluster
 the line graph of the input graph~\autocite{LineGraph09}, but this does not take advantage of

### 2017-11-27 intro/semantics.tex
typos

@@
 
-In this section, we present a list of signed graph usages, sorted by domains. Many of those signed
+In this section, we present a list of signed graph usages, sorted by domains~\autocite{Tang2015a}. Many of those signed
 graphs are the input of some clustering algorithms. In the context of signed graph, the clustering
@@
 graphs, the two types of edge have reverse semantic, whereas in a 2-layers graph, they simply
-denotes two possible interaction, for instance advisor-advisee or regular coauthors relation in a
+denotes two possible interactions, for instance advisor-advisee or regular coauthors relation in a
 citation network~\autocite{Advisor10}. In general though, we focus on cases where $k$ is larger than
@@
 comments, shared tags, groups membership and so on~\autocite{RecoFlickrMulti11}. Again, while being
-part of a common group denote shared interests between two users, overall they must also differs in
-some other attributes (for instance location) in order to bring diversity to this groups.
+part of a common group denote shared interests between two users, overall they must also differ in
+some other attributes (for instance location) in order to bring diversity to this group.
 
@@
 They build the graph of all countries and connect them in three layers. All democracies form a
-clique in the first layer, countries are connected with weights proportional to the amount of yearly
-trade in the second layer and with weights proportional to the number of international organisations
+clique in the first layer, countries are connected with weight proportional to the amount of yearly
+trade in the second layer and with weight proportional to the number of international organisations
 they belong to together in the third layer. While the essence of commerce is to exploit differential

### 2017-11-28 cc/bias_balance.tex
add missing labels
typos and stability in cc

@@
 are possible. By forbidding triangles with a single negative edge, weak balance gives rise to a
-$K$-\emph{consistent} clusters structure on complete graphs. After making these notions more precise
+$K$-\emph{consistent} clusters structure on complete graphs. After making these notions more detailed
 in the rest of this section, we conclude by precisely exposing our learning bias.
@@
 
-We just saw that if a complete graphs is strongly balanced, then it admits a $2$-consistent
-clustering (\autoref{thm:structural} proves the converse is true). It is natural to ask under which
+We just saw that if a complete graph is strongly balanced, then it admits a $2$-consistent
+clustering (\autoref{thm:structural} proves that the converse is true). It is natural to ask under which
 conditions a complete signed graph admits a $K>2$ consistent clustering. This is where the notion of
@@
 % will incur some disagreements. 
-\marginpars{The proof makes essential use of the strong balance (which is ok since next section also
-deal with 2-CC) and the completeness of the graph (which is a far less realistic assumption) yet
-it would be interesting to see if it can be extended.}
+%\marginpars{The proof makes essential use of the strong balance (which is ok since next section also
+%deal with 2-CC) and the completeness of the graph (which is a far less realistic assumption) yet
+%it would be interesting to see if it can be extended.}
 \Textcite[Section 5.5]{BookKleinberg2010}
@@
 
+\label{text:cc_new_bias}
 Using these results on weak balance, we can now make our new bias explicit. Drawing parallel with
@@
 consistent clustering. This can still be seen as a generative process because of the initial integer
-drawn \uar{},\Todo{Although I'm not sure how I would write this likelihood to be fair} but one that
+drawn \uar{}, but one that
+%\Todo{Although I'm not sure how I would write this likelihood to be fair}
 avoids the two pitfalls discussed before. Indeed, there is no decision to be made by nodes, since

### 2017-11-28 cc/cc_non_worst_case.tex
typos and stability in cc

@@
 on true negative edges is smaller than $\mu^-<0$. They give a finite-sample bound on the number of
-node misclustered w.r.t the planted partition as a function on the probability distribution
+nodes misclustered w.r.t the planted partition as a function on the probability distribution
 parameters. Indeed, as pointed out by \textcite{plantedAilon09}, independent and uniform noise is
@@
 is often the result of a preprocessing, which may present strong correlations. Therefore, instead of
-measuring the quality of a solution against in the input (\ie{} the similarity information), they
+measuring the quality of a solution against the input (\ie{} the similarity information), they
 argue it is more sensible to measure it against the (unknown) true optimal clustering that gave rise
-to the input, and show that \ccpivot{} allows that thanks to a new analysis. \Textcite{Mathieu2010}
+to the input, and show that \ccpivot{} allows this, thanks to a new analysis. \Textcite{Mathieu2010}
 also consider an adversarial model, where all edges are flipped with probability $p$ but the
-adversary then decide whether to show us the true sign or the flipped sign. On complete unweighted
+adversary then decides whether to reveal the true sign or the flipped sign. On complete unweighted
 graphs, they find a solution of \mind{} at most $1 + O(n^{-\nicefrac{1}{6}})$ times the optimal
-whenever $p \leq \shalf - n^{-\nicefrac{1}{3}}$ by rounding the usual SDP solution. If in addition
-$p\leq\nicefrac{1}{3}$, there is no adversary and each planted cluster has at least
+whenever $p \leq \shalf - n^{-\nicefrac{1}{3}}$ by rounding the usual SDP solution. If, in addition,
+there is no adversary, $p\leq\nicefrac{1}{3}$ and each planted cluster has at least
 $\Theta(\sqrt{n})$ nodes, then the planted partition can be recovered exactly.
@@
 present another algorithm that finds the ground truth clustering with an arbitrarily small
-classification error.\footnote{They also mention a potentially interesting related
-work~\autocite{unobservedCC14}.} % http://www.jmlr.org/papers/v15/chen14a.html
+classification error.
+% \footnote{They also mention a potentially interesting related work~\autocite{unobservedCC14}.}
+% http://www.jmlr.org/papers/v15/chen14a.html
 % \textcite{Noisy2CCLocality16} study the recovery of an arbitrary planted 2-partition under noisy
@@
 
-Whereas clustering objective functions are \NPh{} to optimize, we expect that meaningful instances
-in which we are interested have additional structure which allows for guaranteed polynomial time
-algorithms, see for instance \autocite{clusteringFeasibility15} for a critical overview of some
-proposed such notions of structure. Informally, a general idea is that the clustering should not
+Whereas clustering objective functions are \NPh{} to optimize, we expect meaningful instances
+in which we are interested to have additional structure which allows for guaranteed polynomial time
+algorithms. We refer the reader to a critical overview~\autocite{clusteringFeasibility15} of some
+such notions of structure proposed recently. Informally, a general idea is that the clustering should not
 change (or at least very little) if the data are slightly perturbed. For instance, a weighted graph
@@
 $\alpha$.
-Finally, we say that an algorithm is \emph{robust} if, given an instance $\mathcal{I}$ and
-in polynomial time, it: returns the optimal solution of $\mathcal{I}$ if $\mathcal{I}$ is
-$\alpha$-stable; and if $\mathcal{I}$ is not $\alpha$-stable, either returns the optimal solution of
-$\mathcal{I}$ or reports that $\mathcal{I}$ is not $\alpha$-stable. This is a handy property, as in
-general, we cannot check the stability of an instance in practice as we do not know the optimal
-solution hence we cannot tell whether it changes or not under perturbations.
+Another notion is the $(c, \epsilon)$-approximation stability. Formally, a dataset $X$ is $(c,
+\epsilon)$-stable if, for an objective function $\Phi$, any clustering whose cost is within a factor
+$c$ of the optimal cost $OPT_{\Phi}$ is $\epsilon$-close to the optimal clustering (as measured by
+the fraction of points in which the two clustering disagree). When the objective is \mind{} on
+complete graphs, \textcite{StableCC09} show that for $(1+\alpha, \epsilon)$-instances, the
+approximation algorithms we describe in \autoref{ssub:cc_harness_approx} find a solution that is
+$(\nicefrac{49}{\alpha} + 1)\epsilon$ close to the optimal clustering.
 
-See \autocite{StableCC09} for application to \mind{} on complete graphs and \autocite{StableLP09}
-for more general study of the solution stability of the canonical \mind{} LP in any kind of graph.
+It is also possible to study the stability of a \pcc{} instance over a general graph w.r.t edge
+weights via its canonical \mind{} LP. \Textcite{StableLP09} present a method to determine to which
+extent the weights of a \pcc{} instance can be perturbed before the optimal solution changes.
 

### 2017-11-28 cc/cc_problem_and_esp.tex
typos and stability in cc

@@
 $u$ and $v$ is associated with two positive numbers:
-$w_{u,v}^+$ denotes the strength of the similarity between $u$ and $v$;
+$w_{u,v}^+$ denotes the strength of the similarity between $u$ and $v$, while
 $w_{u,v}^-$ denotes the strength of the dissimilarity between $u$ and $v$.
@@
 that best approximates the observed signs, but each weakly unbalanced cycle is gonna cost us at
-least one unavoidable mistakes. Despite this fundamental difference in their nature, \pcc{} is
+least one unavoidable mistake. Despite this fundamental difference in their nature, \pcc{} is
 connected with \esp{} at two levels, theoretical and practical. On the theoretical side, computing
@@
 n}$~\autocite[Theorem 6]{Cesa-Bianchi2012b}. These theoretical considerations naturally lead to
-practical batch algorithm, which first finds a clustering minimizing as much as possible the number
+practical a batch algorithm, which first finds a clustering minimizing as much as possible the number
 of disagreements on the training set and then predicts signs according to this clustering. Our
@@
 
-Furthermore, as mentioned in the introduction of this thesis, \pcc{} has also been used in several
+Furthermore, as mentioned in the introduction of this thesis \vpageref{sub:intro_signed_graphs},
+\pcc{} have been used in many
 domains. For instance in computer vision, it is used to segment images in

### 2017-11-28 cc/cc_state_of_art.tex
add some ref
add two refs
tiny fixes and ref all over the places
typos and stability in cc

@@
 case it is reasonable to expect finding the optimal clustering. Furthermore, this can also be useful
-to evaluate in practice the quality of heuristic methods, albeit in non asymptotic setting. The
+to evaluate in practice the quality of heuristic methods, albeit in non asymptotic settings. The
 methods used to solve \pcc{} exactly depends on the type of graph considered. More precisely, the
@@
 find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
-union of cliques. It is possible to navigate between an instance $H$ of cluster editing and an
+union of cliques. It is possible to alternate between an instance $H$ of cluster editing and an
 instance $G$ of \pcc{} on complete graphs in polynomial time. To see why, let $G=(V,(E^+,E^-))$
@@
 unweighted signed graphs, they show that both \mind{} and \maxa{} are \NPc{}. Along the way, they
-give a $17429$-approximation of \mind{} and a PTAS%
+give a $17429$-approximation of \mind{} and a PTAS
 that, for any $\epsilon \in [0,1]$, runs in
@@
 construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
-4.4]{Demaine2006})\Todo{give the intuition of the proof} that it takes a linear time to construct a
+4.4]{Demaine2006}) that it takes a linear time to construct a
 multicut of weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.
+%\Todo{give the intuition of the proof}
 
@@
    \includegraphics[width=0.8\linewidth]{assets/raw/cc_to_mmc.png}
-   \caption{The transformation from \pcc{} on $G$ to \mmc{} on $H$ (temporarily reproduced from
+   \caption{The transformation from \pcc{} on $G$ to \mmc{} on $H$ (reproduced from
    \autocite{Demaine2006}) \label{fig:cc_mmc}}
@@
 Conjecture of~\textcite{UGC02}.}, then for every constant $c>0$ it is \NPh{} to approximate \mmc{}
-within a factor $c$~\autocite{}. On general graphs, the best known approximation factor in $O(\log
+within a factor $c$~\autocite{MulticutHardness06}. On general graphs, the best known approximation factor in $O(\log
 k)$~\autocite{RegionGrowing93}, which according to the presented reduction translates to
@@
 leaving the cluster is less than $c\log(n+1)$ times the fractional weights of the positive edges
-inside the cluster.\Todo{The full proof can be found in pages 7--9 of
-\href{http://erikdemaine.org/papers/Clustering_TCS/paper.pdf}{this online pdf} but in my opinion, it
-doesn't have much to do with machine learning.}
+inside the cluster.
+%\Todo{The full proof can be found in pages 7--9 of
+%\href{http://erikdemaine.org/papers/Clustering_TCS/paper.pdf}{this online pdf} but in my opinion, it
+%doesn't have much to do with machine learning.}
 
@@
 solution of the canonical LP to obtain different approximation factor depending on the
-constraints imposed on the weight.\Todo{Comment on those constraints, especially the triangular ones
-in the context of consensus clustering.} Recall that in the general formulation of the problem, each
+% \Todo{Comment on those constraints, especially the triangular ones in the context of consensus clustering.}
+constraints imposed on the weight. Recall that in the general formulation of the problem, each
 edge carries two positive numbers: $w_{u,v}^+$ and $w_{u,v}^-$. If the weights respect the
@@
 one first solve the canonical LP, and use the \regionGrow{} method to round the fractional solution,
-resulting in an $O(log n)$ approximation. On complete graphs, the simplest solution is the
+resulting in an $O(\log n)$ approximation. On complete graphs, the simplest solution is the
 \ccpivot{} algorithm, that grows clusters out of randomly selected pivots, in a way reminiscent of
@@
 application at hand.
+% \url{http://ieeexplore.ieee.org/document/7321476}
 
@@
 Editing problem that are then fed to an integer program.
-This local search followed by random perturbation is reminiscent of genetic algorithms, and such an
-approach is sketched in~\autocite{GeneticCC08}.\footnote{but that's a terrible paper!} Instead of
+% This local search followed by random perturbation is reminiscent of genetic algorithms, and such an
+% approach is sketched in~\autocite{GeneticCC08}.\footnote{but that's a terrible paper!}
+Instead of
 moving nodes, \textcite{restoreNeighborhood13} starts from the observation that in a balanced
@@
 probability to each cut of a signed graph proportional to the exponential of the number of
-disagreements of that cut. They also develop efficient cut sampling methods.
+disagreements of that cut. They also develop efficient cut sampling methods. Several of these
+methods have recently been evaluated empirically by \textcite{energyCCBench17}.
 
@@
 between all pairs of nodes.
+% this was improved in https://arxiv.org/abs/1709.07308
 
@@
 requires solving linear programs with $m$ variables and up to $O(n^3)$ constraints, which results in
-a high complexity of $O(m^{4.5})$~\autocite[Section 7.2]{LPBook07}. Fortunately we can obtain
+a high complexity of $O(m^{4.5})$~\autocite[Section 7.2]{LPBook07}. Fortunately, we can obtain
 reasonable solutions in practice, thanks to efficient heuristics (using the output of \ccpivot{} as

### 2017-11-28 cc/cc_variants.tex
add some ref
add 2 refs
rewrite related troll works
add some refs
tiny fixes and ref all over the places
typos and stability in cc

@@
 %   \url{https://doi.org/10.1007/s41060-016-0005-2}
+% - CC with triplet constraint instead of signs \autocite{relativeCC17}
 
@@
 graph is bipartite and then move to variants. We will consider more general objectives, when the
-edges are labeled categorically instead of binary, when nodes can belong simultaneously to several
+edge labels are categorical instead of binary, when nodes can belong simultaneously to several
 clusters or when we optimize local objectives per nodes instead of global ones. Finally we will also look at
@@
 deterministic $3$-approximation using the \regionGrow{} method. Later \textcite{Bonizzoni2008} show that
-the minimization version is \APXh{}, even when the input is made of three clusterings and give a
+the minimization version is \APXh{}, even when the input is made of only three clusterings and give a
 combinatorial $\nicefrac{4}{5}$-approximation for the maximization problem. Experimental evaluations are
@@
 beforehand, starting with an initial partition and then alternating between finding the majority
-color of clusters and finding better clusters. Unfortunately, the maximum degree of a graph can be
+color of clusters and finding better clusters. An improved heuristic algorithm is given
+in~\autocite{CCCheuristic16}. Unfortunately, the maximum degree of a graph can be
 as large as $n$. However, \textcite{Anava2015} present constant factor approximations. Namely, they
@@
 vector $x$ indexed by $V$ such that $x_{uv} \in [0, 1]$ for all $uv \in \binom{V}{2}$ and such that
-$x_{vz} \leq x_{vw} + x_{wz}$ for all distinct $v, w, z \in V$}. They also define \enquote{The
+$x_{vz} \leq x_{vw} + x_{wz}$ for all distinct $v, w, z \in V$}. They also define \enquote{the
 error vector $err(x)$ of $x$, as a real vector indexed by $V$ whose coordinates are}
@@
 \paragraph{Spectral Clustering}
+\label{par:cc_spectral}
 
@@
 (corresponding to assortative clusters in $G^+$) and of $Q^-$ (corresponding to disassortative
-clusters in $G^-$), while this is note the case for the arithmetic mean of Laplacians.
+clusters in $G^-$), while this is not the case for the arithmetic mean of Laplacians.
 
@@
 In a similar spirit to \maxa{}$[k]$, \textcite{SignedGang} focus on finding $k$ subgraphs dense in
-positive edges and densely connected by negative edges to each other. Thee dub such subgraph
+positive edges and densely connected by negative edges to each other. They dub such subgraph
 \emph{Oppositive Cohesive Groups}, or more vividly \emph{Gangs in War}, and after formulating the
 problem as a constrained quadratic optimization, they propose a faster iterative local search
-heuristic.
+heuristic. When $k=2$, these subgraphs are called antagonistic communities and a specific data
+mining approach was proposed by \textcite{quasiGang16}.
 % relaxed structural balance \autocite{Doreian2009} but used the same approach as \textcite{Early96}

### 2017-11-28 cc/gtx_related.tex
typos and stability in cc

@@
 \url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
-for instance the Elkin paper \enquote{Our centralized randomized algorithm computes (with
+for instance the Elkin paper~\autocite{Spanner17} \enquote{Our centralized randomized algorithm computes (with
 probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in

### 2017-12-02 cc/cc_main.tex
reorganize non worst case CC and add two more refs
typos and stability in cc
review gtx description and experiments
in particular, avoid mentioning strech by name, and say we only have a
conjecture regarding the runtime
also make the transition between cc and gtx smoother

@@
 
+\subsection{Beyond worst case instances}
+\label{sub:cc_non_worst_case}
+
+\input{cc_non_worst_case}
+
+\begin{aside}
 \subsection{Variants and extensions}
@@
 \input{cc_variants}
+\end{aside}
 
-\subsubsection{\pcc{} under stability assumption}
-\label{ssub:cc_under_stability_assumption}
-\input{cc_multimincut_stable}
+Let us review this material about \pcc{} in the light of our thesis objective: efficiently and
+accurately characterize edges in complex networks, or rather signed graphs in this chapter. The
+results on the hardness of \pcc{}, and the fact that the best approximations rely on a linear
+program with a large number of constraints seems to run counter to such efficiency and accuracy
+requirements. To avoid this pitfall, we presented many existing heuristics. More importantly, recall
+that our learning bias is that nodes are assigned to $K$ groups, and that signs are consistent
+with those groups. Informally, the closer an input signed graph is to this ideal situation, the more
+regular it is with respect to this bias, and thus the easier the \esp{} problem is. Accordingly, the
+works described in \autoref{sub:cc_non_worst_case} point out that such non worst case instances are
+indeed where we expect clustering algorithms to be able to identify those $K$ groups with little to
+no error. In the next section, we present an algorithm relying on a similar intuition, in an active
+setting and where $K=2$.

### 2017-12-02 cc/gtx_algorithm.tex
tiny fixes and ref all over the places
review gtx description and experiments
in particular, avoid mentioning strech by name, and say we only have a
conjecture regarding the runtime
also make the transition between cc and gtx smoother

@@
 connected pair of nodes in the original graph, we rely on a particular subgraph structure, namely
-the star.  We therefore introduce two algorithmic primitives: \extractStar{}, which partition a
-graph $G$ into a set of disjoint stars; and \collapseStar{}, which selects edges from $E$ to
+the star.  We therefore introduce two algorithmic primitives. \extractStar{}, which partition a
+graph $G$ into a set of disjoint stars. And \collapseStar{}, which selects edges from $E$ to
 assemble these stars into a new, smaller graph. Given a graph topology $G_0=(V_0, E_0)$ and assuming
@@
 complexity. Then we state formally the complete \gtx{} algorithm, prove its termination and
-correctness, and show a detailed example of its execution.  Finally, we study its properties, such
-as the number of iterations needed to finish and the stretch of the resulting tree.
+correctness, and show a detailed example of its execution.  Finally, we provide a conjecture on
+the number of iterations needed to finish.
 
@@
 We also assume that $G$ is the adjacency list of the graph, so that $G[u]$ is the set of neighbors of
-$u$, \ie{} $G[u] \equiv \mathcal{N}(u)$. Finally $membership$ is a map and we
-define the \textsc{Star} function, which creates a star given a center $c$, a list $periphery$ of
+$u$, \ie{} $G[u] \equiv \mathcal{N}(u)$. Finally $membership$ is the map storing for each node the
+index of the star it belongs to. This map is updated by the
+\textsc{Star} function, which creates a star given a center $c$, a list $periphery$ of
 peripheral nodes, and a star index $i$. After creating the \ith{} star, for every node $u$ belonging
@@
 maximum of the queue (line 8) and testing if $c$ is still part of the graph, which happens $|V|$
-time. Then we build the corresponding star (line 11--14). In total, we test the membership of $|V|$
+times in total. Then we build the corresponding star (line 11--14). In total, we test the membership
+of $|V|$
 nodes in line 11, $\textsc{Star}$ updates the $membership$ map $|V|$ times in line 12,
@@
 with the $membership$ result of \extractStar{} and an $eccentricity$ array. It builds a new graph
-$G_{t+1}$ where each star from \extractStar{} becomes a node and where there is a link between two
+$G_{t+1}$ where each star $s_u$ from \extractStar{} becomes a node and where there is a link between two
 nodes $s_u$ and $s_v$ if the nodes in $V_t$ making up $s_u$ and $s_v$ are connected in $E_t$. The
@@
   stars $s_1$ and $s_2$, and all are part of $cross\_edges[(s_1, s_2)]$, here represented in light
-red. Those edges are labeled with the sum of eccentricity of their underlying endpoints, and the
-minimal one is linking a peripheral node of $s_1$ directly to the center of $s_2$.}
+red. Those edges are labeled with the sum of eccentricity of their underlying endpoints. White node
+have an eccentricity of $0$, and gray node an eccentricity of $1$. The edge with the
+minimal eccentricity is linking a peripheral node of $s_1$ directly to the center of $s_2$.}
   \label{fig:gtx_cross_edge}
@@
       \State Let $cross\_edges$ be the map described above
-      \ForAll{edge $(u, v)$ in $E_t$}
+      \For{\textbf{every} edge $(u, v)$ in $E_t$}
       \Let{$s_u$}{$membership[u]$}
@@
       \EndFor
-      \ForAll{pair of will-be connected stars $(s_u,s_v)$ in $cross\_edges$}
+      \For{\textbf{every} pair of will-be connected stars $(s_u,s_v)$ in $cross\_edges$}
         \Let{$possible\_underlying\_edges$}{$cross\_edges[(s_u, s_v)]$}
-        \marginpars{\small Line 11 is actually a simplification since the $eccentricity$ array is only
-        defined for nodes in $V_0$ so when \collapseStar{} is called on $G_3$ for instance, $u$ and $v$
-        are nodes of $V_3$ and we have to retrieve the actual corresponding edge in $E_0$. Maybe I could
-        simply add a map from $E_t$ to $E_0$ like in the real python code…}
+        %\marginpars{\small Line 11 is actually a simplification since the $eccentricity$ array is
+        %only defined for nodes in $V_0$ so when \collapseStar{} is called on \eg{} $G_3$, $u$
+        %and $v$ are nodes of $V_3$ and we have to retrieve the actual corresponding edge in $E_0$.
+        %Maybe I could simply add a map from $E_t$ to $E_0$ like in the real python code…}
         \State $$(u_0, v_0) \gets \argmin_{(u,v)\in possible\_underlying\_edges} eccentricity[u]+eccentricity[v]$$
@@
 our earlier promise, this entails showing some bookkeeping code, because it has an influence on the
-runtime of \gtx{}. \Todo{Those maps could probably use better names}
+runtime of \gtx{}. %\Todo{Those maps could probably use better names}
 Namely in the listing of \autoref{alg:gtx} \vpageref{alg:gtx}, we use the three following maps:
@@
 \vspace{0.5\baselineskip}
-\noindent\begin{tabulary}{187mm}{lllL}
+\setlength{\fullpage}{\textwidth+\marginparsep+\marginparwidth}
+\noindent\begin{tabulary}{\fullpage}{lllL}
   \toprule
@@
 graph (line 9). We perform these operations until there are no edge connecting stars anymore. At
-this point, we revisit every outer edges to build the spanning tree (line 13).
+this point, we revisit every outer edge to build the spanning tree (line 13).
 
@@
   |V_t|$. Note also that for all $t$, $|V_t| > 0$ and that when $|V_t|=1$, \extractStar{} creates a
-  singleton and \collapseStar{} does not return any outer edges so \gtx{} finishes, proving that the
-  number of iterations $K$ satisfies $K \leq |V_0|$.
+  singleton and \collapseStar{} does not return any outer edges. At this point, \gtx{} finishes,
+  proving that the number of iterations $K$ satisfies $K \leq |V_0|$.
 
@@
  star, although it's difficult to tell how it would affect the quality of the resulting tree.}
- \textsc{Update-Nodes-Mapping} take $O(|V_0|)$ time since they go through every node of the original
+ \textsc{Update-Nodes-Mapping} take $O(|V_0|)$ time, since they go through every node of the original
  graph. Because $G_0$ is connected, $|V_0| \leq |E_0|$, and since for all $t$, $|E_t| \leq |E_0|$,
@@
 The bound of \autoref{prop:gtx_correct} amounts to $O(|V_0||E_0|)$, which we believe is overly
-pessimistic. First because we expect the number of nodes to decrease by more than one at each step,
-maybe leading to only $O(\log |V_0|)$ iterations. More importantly, instead of bounding
-$\sum_{t=0}^K |E_t|$ by $K|E_0|$, we hope to show that $E_{t+1} = rE_t$ with $r<1$ so that this sum
-is $O(|E_0|)$.
+pessimistic. Indeed, we expect the number of nodes to decrease by more than one at each step,
+potentially leading to only $O(\log |V_0|)$ iterations.
 
@@
 We now run \extractStar{} on $G_1$. Because all nodes have degree $2$, they could all be chosen to
-be the center of a star yet we again they are picked according to their index and therefore we
+be the center of a star. We again assume they are picked according to their index and therefore we
 choose $\starone{1}$ to be the center of the star $\startwo{1}$ with peripheral nodes $\starone{2}$
@@
 \vpageref{fig:gtx_spanning_tree}. For completeness, we can also look at the edges which are not part
-of the spanning tree and therefore contribute to the stretch of the tree. In that case the average
-stretch is $7$, as showed in \autoref{tab:gtx_example_stretch}.
+of the spanning tree, that is the test edges, and compute how long are the paths joining their
+endpoints. Indeed, this is the second quantity appearing in the mistakes bound of
+equation~\eqref{eq:stretch_mistakes}. As shown in \autoref{tab:gtx_example_stretch}, this average
+length is $7$.
 
@@
   \centering
-  \caption{Stretch of the example tree}
+  \caption{Length of the paths not in the resulting spanning tree.}
   \label{tab:gtx_example_stretch}
@@
 
-\paragraph{Number of iterations needed}\label{par:number_of_iteration}%
-
-\textcolor{red}{\LARGE Draft}
+\paragraph{Discussion on the number of iterations needed}\label{par:number_of_iteration}%
 
@@
   \includegraphics[width=0.9\linewidth]{assets/tikz/gtx_line_tikz.pdf}
-  \caption{A line graph with stars in blue}
+  \caption{A line graph with stars in blue.}
   \label{fig:gtx_line_graph}
 \end{marginfigure}
-This will results in $\nicefrac{2n}{5}$ stars, which is less than half of $n$ and because in that
-case $n=m+1$, \gtx{} will finish after $O(\log m)$ iterations. Note that a barbell graph (two
+This will results in $\nicefrac{2n}{5}$ stars organized in a line, which is less than half of $n$.
+Thus we can conclude that in this case, \gtx{} will finish after $O(\log n)$ iterations. Note that a barbell graph (two
 cliques connected by a line) would requires a number of iterations logarithmic in the length of that
-central line, despite having many more edges. This suggests unsurprisingly that the diameter of the graph
+central line, despite having many more nodes and edges.
+
+In fact, it is helpful to look at this question from the other direction. Namely, we can wonder what
+is the minimum number $n(K)$ of nodes necessary for a graph to requires at least $K$ collapses
+before being reduced to a single node. Checking by hand, we have that $n(1)=2$, $n(2)=4$ and
+$n(3)=8$. Furthermore, this number is achieved by a line or circle consisting of exactly $2^K$
+nodes. By induction, we can even show that on any line or circle graphs with $2^n$ nodes, the \gtx{}
+algorithm will terminate after $n$ collapses. Assuming we could prove that $n(K)=2^K$, given a graph
+$G$ with $n=2^{\log_2 n}$ nodes, by definition of $n(K)=2^K$, it would require less than
+$\ceil{\log_2 n} = O(\log n)$ collapses before being reduced to a single node. However, we were not
+able to rule out the possibility that there is a graph more parsimonious than a line still able to
+survive the same number of collapses. Therefore, we cannot conclude that $n(K)=2^K$ and we leave
+this statement as a conjecture.
+
+\iffalse
+This suggests unsurprisingly that the diameter of the graph
 could be a good parameter to quantify the number of iterations needed. Indeed, take a tree and consider
@@
 \end{figure*}
+\fi
 
@@
 The execution of \extractStar{} is mainly deterministic, except for the fact the ties between nodes
-with the same highest degree are broken arbitrarily. While this allows for an efficient
-implementation, and simplify the analysis of the resulting sequence of stars and therefore the
-induced spanning tree, in can be detrimental in an adversarial context, where we could end up with a
+with the same highest degree are broken arbitrarily. This allows for an efficient
+implementation, and simplifies the analysis of the resulting sequence of stars and therefore the
+induced spanning tree. However, it can be detrimental in an adversarial context, where we could end up with a
 tree forcing a lot of mistakes. We add an element of randomization to \extractStar{} by letting it
@@
 A variant of the \gtx{} algorithm as a whole we did not explore so much in practice is its ability
-to produce spanner by stopping early. Basically if we stop at iteration $t$, we output the graph
+to produce spanners by stopping early. Basically if we stop at iteration $t$, we output the graph
 $G_t$ (which in general is not a tree) with its edges unfolded to lie in $E_0$. This corresponds to
-a trade off between having more edges than $|V_0|-1$ but potentially making shorter connection and
-thus having a lower stretch. This would also be interesting in the sign prediction setting. Assuming
-the treewidth of $G_t$ is decreasing with $t$, maybe we could show there is \enquote{reasonable}
+a trade off between having more edges than $|V_0|-1$ but potentially making shorter connection.
+This would also be interesting \esp{} problem. Assuming
+the treewidth of $G_t$ is decreasing with $t$, we could try to show there is \enquote{reasonable}
 number of paths between $u$ and $v$, compute their parity and take a majority vote on the results.

### 2017-12-02 cc/gtx_experiments.tex
reorganize files and write some intro
review gtx description and experiments
in particular, avoid mentioning strech by name, and say we only have a
conjecture regarding the runtime
also make the transition between cc and gtx smoother

--- 2017-06-14 gtx_experiments.tex
@@
-In this \namecref{sub:gtx_empirical_evaluation}, we provide empirical evidences of the properties of
-\gtx{} over several classes of graph, and compare it with a \bfs{} baseline.\marginpars{If time
-allows, it would be interesting to implement some methods of \vref{sub:gtx_state_of_the_art} and add
-them to the comparison}\todo*{implement more low stretch methods} Namely, we consider three kinds of
+In this section, we provide empirical evidences of the properties of
+\gtx{} over several classes of graph, and compare it with a \bfs{} baseline.
+%\marginpars{If time allows, it would be interesting to implement some methods of \vref{sub:gtx_state_of_the_art} and add them to the comparison}\todo*{implement more low stretch methods}
+Namely, we consider three kinds of
 graph topology (with both synthetic and real world instances that carry signs on their edges) and
@@
 The three kinds of topology we consider are:
-\begin{description}
+\begin{description}[leftmargin=*]
 	\item[\grid{}] which are 2D lattices, where each node has four neighbors except on the boundary.
@@
 		crude but reasonable approximation of online social networks. Therefore, the real world
-		instances of the \lpa{} model are \wik{}, \sla{} and \epi{}\marginpars{as used in the first
-		chapter}\todo*{add a ref to first chapter} along with \gplus{}. The last one is constructed from
+		instances of the \lpa{} model are \wik{}, \sla{} and \epi{} networks from \autoref{chap:troll}
+		 along with \gplus{}. The last one is constructed from
 		ego networks of \gplus{}\footnote{Available at
@@
 		presented in (\autoref{tab:gtx_xp_dataset}). 
-	\item[\triangle{}] which consists of a Delaunay triangulation of random 2D points\footnote{As
-		implemented by the \textsf{graph-tool} library (\url{https://graph-tool.skewed.de})}.
+	\item[\triangle{}] which consists of a Delaunay triangulation of a set of points randomly
+		located in a 2D space.\footnote{As
+		implemented by the \textsf{graph-tool} library (\url{https://graph-tool.skewed.de})}
 \end{description}
@@
 The first property of Galaxy trees we wish to evaluate is their stretch, which depends only of graph
-topology. Namely, let $G$ be a graph over vertex set $V$ with $|V|=n$ and edge set
-$E$.\Todo[MOVE]{Stretch definition is likely to happen somewhere earlier} Furthermore, let $T$ be a
-spanning tree of $G$ and $\etest{}$ the edges of $G$ not in $T$. Then we define the \emph{average
-test edge stretch} as $\frac{1}{|\etest{}|} \sum_{(u,v) \in \etest{}} |\mathrm{path}^T_{u,v}|$,
+topology. Recall that following equation~\autoref{eq:test_stretch_def}, we define the average test
+edge stretch as $\frac{1}{|\etest{}|} \sum_{(u,v) \in \etest{}} |\mathrm{path}^T_{u,v}|$,
 where $|\mathrm{path}^T_{u,v}|$ is the unique path between $u$ and $v$ in $T$.
@@
 rooted at the highest degree node and obtained through a breadth first visit of the graph. This
-involves randomness in order in which nodes are visited. Likewise in \gtx{}, the choice of the edge
+involves randomness in the order in which nodes are visited. Likewise in \gtx{}, the choice of the edge
 linking two stars is not always unique, meaning that we have to break ties at random.  Therefore,
-for each graph, we repeat the tree construction 12 times and present the average result, noting that
+for each graph, we repeat the tree construction \np{12} times and present the average result, noting that
 the variance (showed as error bar in \autoref{fig:gtx_xp_st}) is small.
 
-On \lpa{} and \triangle{}, we see that both trees exhibits logarithmic stretch, although with a
+On \lpa{} and \triangle{}, we see in \autoref{fig:gtx_xp_st} that both trees exhibits logarithmic stretch, although with a
 larger constant for \gtx{}. Note that this is also the case for others low stretch tree methods
-\autocite[\S 5.3.1]{papplow}. On \grid{} however, \gtx{} preserves this logarithmic stretch growth
+\autocite[Section 5.3.1]{papplow}. On \grid{} however, \gtx{} preserves this logarithmic stretch growth
 while this is visually no longer the case for \bfs{}.
+In that case, we cannot expect a better stretch than $\frac{\log n}{2048}$ according to
+\autocite[Theorem 6.6]{LowerBound95}.
 
@@
 Except for the three real datasets that already include signs\footnote{We nonetheless perform some
-preprocessing in order to make them undirected to remove the small proportion of conflicting edges
+preprocessing in order to make them undirected and to remove the small proportion of conflicting edges
 (e.g. positive from $u$ to $v$ but negative from $v$ to $u$).}, all the other are constructed,
@@
 be $+$ if they are in the same cluster and $-$ otherwise.  Predicting using path parity will thus
-gives perfect result. To test performance in real or adversarial situation, we then add noise, that
+gives perfect result. To test performance in real situations, we then add noise, that
 is we select a fraction of edges uniformly at random and flip their sign.
 
-We evaluate the performance of our prediction using the Matthews Correlation\Todo{merge this
-definition of MCC with the one in first chapter} Coefficient (MCC)~\autocite{MCC00} \[
-	\mathrm{MCC} = \frac{ TP \times TN - FP \times FN } {\sqrt{ (TP + FP) ( TP + FN ) (
-			TN + FP ) ( TN + FN ) } } = \pm \sqrt{\frac{\chi^2}{n}}
-\]
-Since we do not have confidence score, we cannot use AUC. Yet we have to account for the large sign
-unbalance and thus cannot rely on accuracy or $F_1$ measure.  Therefore we choose MCC, which
-combines all the four numbers of the confusion matrix in a single metric. It ranges from $+1$
-(perfect prediction) to $-1$ (inverse prediction) through $0$ (random prediction). As a
-demonstration of MCC usefulness, predicting all edges but one to be positive on Slashdot gives
-$.764$ accuracy, $.886$ $F_1$ score\marginpars{Actually the $F_1$ score is $.866$ for positive
-edges, $0$ for negative ones and $.661$ if we can take an average weighted by class size.} but
-$-0.0007$ MCC.
-
+Like in \autoref{s:exp}, we evaluate the performance of our prediction using the Matthews
+Correlation Coefficient (MCC), defined in equation~\autoref{eq:troll_mcc} \vpageref{eq:troll_mcc}.
 As showed in \autoref{fig:gtx_xp_mcc}, when the noise level is low, \gtx{} performs better than
@@
 
-To further assess the quality of our trees, we plug them in them into a successful heuristic method
-to predict edge sign: \asym{}~\autocite{Kunegis2009}. \Todo{It might also be interesting to see if
-that would be a good training set for our troll method, although it has to be checked it makes sense
-from a running time point of view.} It computes the exponential of the adjacency matrix after it has
+To further assess the quality of our trees, we plug them in them into an existing heuristic method
+to predict edge sign: \asym{}~\autocite{Kunegis2009}.
+%\Todo{It might also be interesting to see if that would be a good training set for our troll method, although it has to be checked it makes sense from a running time point of view.}
+It computes the exponential of the adjacency matrix after it has
 been reduce to $z$ dimension. This allows to count the sign of all paths between two pairs of nodes
@@
 forming a \gtx{} $iii)$ $|V|-1$ edges chosen uniformly at random.
-
 We set the parameter $z$ equal to $15$ because $i)$ it is one of the best in \cite[Fig.
-11]{Kunegis2009}, $ii)$ it performs well on real dataset in \cite[Fig.3]{Cesa-Bianchi2012a}, and
-$iii)$ it was good in our initial testing (\texttt{20150401\_wed\_spectral.ipynb}).
+11]{Kunegis2009} and $ii)$ it performs well on real datasets in \cite[Fig.3]{Cesa-Bianchi2012a}.
+% , and $iii)$ it was good in our initial testing (\texttt{20150401\_wed\_spectral.ipynb}).
 
@@
 graphs used previously are not all included. The conclusion of \autoref{fig:gtx_xp_asym} is that
-except on social network, it is better to use spanning trees than random edges. Specifically, \gtx{}
+except on social networks, it is better to use spanning trees than random edges. Specifically, \gtx{}
 on \grid{} and \bfs{} elsewhere.
@@
 
+\iffalse
 Finally\marginpars{Actually I never did it because \shz{} wasn't implemented at the time, so now is
@@
 the task of nodes prediction using \shz{} algorithm~\autocite{Vitale2012}.
+\fi

### 2017-12-02 cc/gtx_main.tex
review gtx description and experiments
in particular, avoid mentioning strech by name, and say we only have a
conjecture regarding the runtime
also make the transition between cc and gtx smoother

@@
-Given a graph $G$, and following the social balance theory roughly summarized as \enquote{the enemy
-of my enemy is my friend}, assume that the labelling of $E$ is consistent with a two-clustering of
-$V$. Namely, assume that $V$ can be partitioned in two clusters such that edges within each cluster
-are positive and edges across clusters are negative. In that case, the following
+We now show how to apply the learning bias of \autoref{sec:new_bias} on undirected graphs.
+Furthermore, we assume that the strong balance holds, meaning that there are only two groups
+according to \autoref{thm:structural}. In other words, the labelling of $E$ is consistent with a
+two-clustering of $V$. Namely, $V$ can be partitioned in two clusters such that edges within each
+cluster are positive and edges across clusters are negative.
+In that case, the following
 \emph{multiplicative rule} holds: for any nodes $u$, $v$ in $V$, and any path $p$ between $u$ and
-$v$ in $G$, the sign $\yuv$ is equal to the product of the signs along $p$ (we call this product the
-parity of $p$ and denote it $\pi(p)$). While it is a simple and convenient model, this is too strong
-of a requirement to be satisfied in practice. Therefore, we relax it by assuming that we start with
-a consistent labeling $Y$ but can only observe a randomly perturbed version $Y'$ of $Y$.
-Specifically, given a constant $q\in [0, \nicefrac{1}{2})$, each sign of $Y$ is flipped with a
-probability smaller than $q$.
+$v$ in $G$, the sign $\yuv$ is equal to the product of the signs along $p$. Hereafter, we call this
+product the parity of $p$, and denote it by $\pi(p)$.
+While it is a simple and convenient hypothesis, this is too strong
+of a requirement to be satisfied in practice. Therefore, we relax it by assuming that, starting from
+a consistent labeling $Y$, we can only observe a randomly perturbed version $Y'$ of $Y$.
+Specifically, given a constant $q\in [0, \nicefrac{1}{2})$, every sign of $Y$ is flipped with a
+probability smaller than $q$. We denote by $E_{\mathrm{flip}} \subset E$ the set of edges whose sign
+has been flipped.
 
-In this section, we are interested in active learning algorithms, that first query a subset
+In this section, we are interested in active learning algorithms that first query a subset
 $\etrain$ of the edges, observe the signs in $\etrain$ and use them to predict the remaining signs.
-More precisely, we focus on an algorithm that queries spanning tree $T$ of $G$ and predicts the sign
-of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. While this
-can\Todo{state it formally then ;)} be stated more formally~\autocite[Section
-4.1]{Cesa-Bianchi2012b}, it makes sense intuitively that since each sign has been potentially
-flipped, the longer the path in $T$, the more likely its parity will be not be equal to the true
-sign $\yuv{}$. Therefore we would like each such path to be as short as possible. In the following,
-we describe a way to build spanning trees tailored for this situation.
+More precisely, we focus on an algorithm that queries a spanning tree $T$ of $G$ and predicts the
+sign of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. Intuitively, since
+each sign has been potentially flipped, the longer the path in $T$, the more likely its parity will
+be not be equal to the true sign $\yuv{}$. Therefore we would like each such path to be as short as
+possible.
+Formally, the number of mistakes of such an algorithm is upper bounded by~\autocite[Equation
+(3)]{Cesa-Bianchi2012b}
+\begin{equation*}
+  |E_{\mathrm{flip}}| + \sum_{(u,v) \in \etest}
+  \sum_{e\in E} \Ind{e \in \pathtuv} \Ind{e \in E_{\mathrm{flip}}}
+\end{equation*}
+which in expectation is equal to:
+\begin{equation}
+  \label{eq:stretch_mistakes}
+  q\left(|E| + \sum_{(u,v) \in \etest} |\pathtuv| \right)
+\end{equation}
+In the following, we describe a way to build spanning trees tailored for this situation. More
+precisely, we implement and analyze a suggestion made to us by~\autocite{gtxFabio}.
 

### 2017-12-02 troll/experiments.tex
review gtx description and experiments
in particular, avoid mentioning strech by name, and say we only have a
conjecture regarding the runtime
also make the transition between cc and gtx smoother

@@
 (MCC)~\autocite{MCC00}, defined as
-\[
+\begin{equation}
+  \label{eq:troll_mcc}
 	\mathrm{MCC} = \frac{tp\times tn-fp\times fn}%
         {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } }% = \pm \sqrt{\frac{\chi^2}{n}}
-\]
+\end{equation}
 MCC combines all the four quantities found in a binary confusion matrix ($t$rue $p$ositive, $t$rue

### 2017-12-03 cc/gtx_related.tex
rewrite gtx related work

@@
 \label{sub:gtx_state_of_the_art}
+
 Looking for a subgraph $H$ of $G$ that best preserve the distance in $G$ while being sparse is an old
-problem, especially driven by network design in fields such as transportation~\autocite{RoadNetworks60}
-and electrical circuits~\autocite{electricalNetworks60}. For instance, \textcite{Johnson1978} define
-the following problem\footnote{We adapt their notations to match ours}:  
-\begin{problem}[Network Design Problem]
-  \label{prob:gtx_ndp}
-  Given an undirected integer-weighted graph $G=(V, E, w)$, a budget $B\in\Nbb$ and a criterion
-  threshold $C\in \Nbb$, does there exist a spanning subgraph $G'=(V, E')$ of $G$ with weight
-  $w(E') \leq B$ and criterion value $F(G') \leq C$, where the criterion function $F(G')$ denotes
-  the sum of the weights of the shortest paths in $G'$ between all vertex pairs?
-\end{problem}
-Note that their definition of stretch includes all possible edges in $G$ and not only those in $E$.
-They prove that this problem is \NPc{} by exhibiting a reduction from the \textsc{Knapsack}
-problem. Finally they prove that the simpler problem of finding a spanning tree on an unweighted
-graph, that is
-\vspace{-.5\baselineskip}
-\begin{problem}[Simple Network Design Problem]
-  \autoref{prob:gtx_ndp} with $w$ being the equal to $1$ for all edges in $E$ and $B=|V|-1$.
-\end{problem}%
-\vspace{-.5\baselineskip}
-\noindent is also \NPc{} by reduction from \textsc{Exact 3-Cover}.
-In this setting of a spanning tree minimizing the stretch over all pairs of nodes, \enquote{
-\textcite{AllPairStrech08} obtained a $(c log n)^{O(\log \Delta)}$ factor, where $c$ is the optimal
-maximum stretch and $\Delta$ is the diameter.  They also showed $O(1)$ approximation for the case
-where the graph is unweighted. The constant was later reduced to $6$ by \textcite{AllPairStrech10}}
+problem, driven originally by network design in fields such as transportation~\autocite{RoadNetworks60}
+and electrical circuits~\autocite{electricalNetworks60}. The way we define \enquote{preserving the
+distance}, and the exact form of $H$ give rise to several problems, which we summarize later in
+\autoref{tab:gtx_related_stretch} \vpageref{tab:gtx_related_stretch}. We first give some
+definitions, then cover the most relevant problems in details, and finally give some pointers for
+the others problems.
 
-\begin{problem}[Optimal Communication Spanning Tree~\autocite{Requirements74}]
+Let the distance between $u$ and $v$ in $G$ be
+\begin{equation*}
+  d_G(u,v) = \sum_{e \in \pathguv} \ell(e)\,,
+\end{equation*}
+where $\ell(e)$ is the \emph{length} of the edge $e$ and \pathguv{} is the shortest path between $u$
+and $v$ in $G$. In the following, we consider only the uniform case, in which the length of an edge
+is equal to its weight. The stretch of an edge $(u,v)$ in $H$ is defined as
+\begin{equation*}
+  \estr(u,v) = \frac{d_H(u,v)}{d_G(u,v)}.
+\end{equation*}
+
+We may then want to minimize the stretch of:
+\begin{enumerate}[1),nosep]%,leftmargin=*]
+  \item some pairs of nodes. That is, given $L$ and $R$ in $V$, minimize $\sum_{u \in L, v\in R}
+    \estr(u,v)$
+  \item all pairs of nodes corresponding to edges of $G$, \ie{} minimize $\sum_{(u,v) \in E}
+    \estr(u,v)$
+  \item all pairs of nodes, \ie{}  minimize $\sum_{(u,v) \in V^2} \estr(u,v)$
+\end{enumerate}
+Note that for an unweighted graph, the second problem reduces to minimizing $\sum_{(u,v) \in E}
+|\pathhuv|$. If furthermore $H$ is tree, this is equivalent to minimize the second term of equation
+\eqref{eq:stretch_mistakes}. Therefore we focus mainly of that definition of stretch, and consider
+the other two only briefly.
+
+The second point affecting the problem is the structure of $H$. The only requirements are that it
+must be spanning all the nodes involved in the computation of the chosen stretch, and that $\forall
+(u,v) \in E,\, d_H(u,v) \geq d_G(u,v)$. Beside that, $H$ can be a tree of $G$, a general subgraph of
+$G$ or even a subset of $V^2$ (\ie{} containing edges not in $E$). We focus mainly on the first two
+cases, since they are covered by the \gtx{} algorithm.
+
+% Namely, let $G$ be a graph over vertex set $V$ with $|V|=n$ and edge set $E$. Furthermore, let $T$
+% be a spanning tree of $G$ and $\etest{}$ the edges of $G$ not in $T$. Then we define the
+% \emph{average test edge stretch} as $\frac{1}{|\etest{}|} \sum_{(u,v) \in \etest{}}
+% |\mathrm{path}^T_{u,v}|$, where $|\mathrm{path}^T_{u,v}|$ is the unique path between $u$ and $v$ in
+% $T$.
+
+% However, \textcite[Section 3, page 453]{lognMetricBoundConf03} claim a $O(\log n)$ approximation so maybe I'm
+% wrong. Turns out, they refer to a distribution over trees and this $O(\log n)$ is the expected
+% stretch of a tree sampled from this distribution
+
+% This defines two kind of structures, spanning trees and spanners (which are still sparse subgraphs
+% yet containing more than $|V|-1$ edges).
+
+\paragraph{Trees}
+\label{par:trees}
+
+One early mention of seeking a low-stretch spanning tree is given by \textcite{Requirements74},
+albeit in more general form:
+\begin{problem}[Optimal Communication Spanning Tree]
 Given a set of nodes $V=\{v_1, \ldots, v_n\}$, a set of distances $d_{ij}$ and a set of requirements
@@
 \end{problem}
+For a weighted graph $G=(V,E,w)$, by letting $d_{ij} = w_{ij}$ and $r_{i,j}= \Ind{(i,j) \in E}$,
+finding an Optimal Communication Spanning Tree thus amounts to finding a low-stretch spanning tree.
+\autoref{tab:gtx_related} present a list of works where the stretch was improved.
 
-Given a weighted graph $G=(V,E,w)$, if we let $d_{ij}$ be $w((i,j))$ and $r_{ij}$ be $1$ if
-$(i,j)\in E$ and $0$ otherwise, it seems to me this correspond to the weighted low stretch problem.
-% However, \textcite[Section 3, page 453]{lognMetricBoundConf03} claim a $O(\log n)$ approximation so maybe I'm
-% wrong. Turns out, they refer to a distribution over trees and this $O(\log n)$ is the expected
-% stretch of a tree sampled from this distribution
-
-This defines two kind of structures, spanning trees and spanners (which are still sparse subgraphs
-yet containing more than $|V|-1$ edges).
-
-\paragraph{Trees}
-\label{par:trees}
-
-Earlier but slightly less related, \textcite{OptimalNetwork69} studies the following problem:
-\begin{problem}[Optimal Network Problem]
-  \label{prob:gtx_scott}
-  Given a set $V$ of $n$ vertices, find a set of spanning edges $E\subset V^2$ that minimizes
-  the sum of the length of the shortest paths  between all vertex pairs while the
-  total length of the resulting network does not exceed some upper bound $B\in\Nbb$.
-\end{problem}
-This can be seen as a special case of \autoref{prob:gtx_ndp} with $G$ being the unweighted
-\marginpars{according to the experiments, it's not clear whether there are weights or
-not…} $n$-complete graph. He proposes a backtracking solution and two local search approximate
-algorithms. Some early branch and bound heuristic solutions to \autoref{prob:gtx_scott} are surveyed
-in~\autocite[Section 2.3.2]{networkDesignSurvey89} although they do not come with asymptotic
-guarantee on the stretch. Furthermore, \textcite{optimApproxNP80} proves that for any $\epsilon \in
-(0,1)$, finding a $|V|^{1-\epsilon}$ approximation is \NPc{}.
-
-Next is the seminal paper of \textcite{LowerBound95}. It touches on many topics, and frame the
+We start with the seminal paper of \textcite{LowerBound95}. It touches on many topics, and frame the
 problem in a game theoretic way but here we only focus on two of their results: a lower bound of
@@
   \centering
-  \caption{Reproduction of Table 1 from~\autocite{Abraham2012} showing the evolution of the best
+  \caption{Reproduction of Table 1 from~\autocite{Abraham2012}, showing the evolution of the best
   asymptotic average stretch over time.}\label{tab:gtx_related}
@@
 
-\autocite{LowerStretch05} is the star decomposition described in section 3 of \autocite{papplow},
-which was modified by \autocite{nearlyTight08} to improve the stretch. Then \textcite{TighterSDD11}
-improved the runtime by rounding the edge weights to the closest power of $2$ and using a modified
+The idea of recursively partitioning the graph and construction a low-stretch spanning tree in each
+part is common to all the papers of \autoref{tab:gtx_related}. \Textcite{LowerStretch05} devise a
+$(\delta, \epsilon)$-star decomposition such that all the stars have comparably low radius.
+It was modified in~\autocite{nearlyTight08} to improve the stretch. Then \textcite{TighterSDD11}
+improve the runtime by rounding the edge weights to the closest power of $2$ and using a modified
 implementation of the Dijkstra's algorithm in the case of at most $k$ distinct edge
@@
 complex but tighter petal decomposition.
+\iffalse
 \begin{marginfigure}
@@
 \end{marginfigure}
-
-If the tree is non spanning (\ie{} it can include edges not from the original graph as long as the
-distances in tree remain larger the distances in the graph), the average stretch can be reduced to
-$\Theta(\log n)$~\autocite{lognMetricBoundConf03}. Furthermore, when the objective if to minimize
-the average stretch over all possible pairs of node, then \textcite{constantDistortion07} achieve a
-universal constant bound.
 
@@
 dually chordal graphs, and others) admit tree $t$-spanners for small values of $t$}
-
-A tangential problem is \enquote{Light Approximate Shortest-path Trees (LASTs).  Given a set of
-“sources” and a sink, a LAST is a tree of weight close to the minimum spanning tree (MST) on the
-sources and the sink, such that the tree distance from any source to the sink is close to the
-shortest-path distance. Khuller, Raghavachari, and Young (KRY95) defined and studied LASTs in the
-offline setting and showed that one can get constant stretch with cost constant times the
-MST.}\footnote{FOCS'17 \href{https://arxiv.org/abs/1611.00052}{arxiv.org/abs/1611.00052}}
+\fi
 
@@
 
+As we mentioned, by stopping the \gtx{} algorithm before it finishes, we obtain a set of edges
+spanning the graphs that is not a tree. Such structure are called \emph{spanner}. More precisely,
+the subgraph $H$ is said to be an $t$-spanner of $G$ if, for a parameter $t \geq 1$, and for every
+pair $u, v \in V$ of vertices, it holds that $d_H(u, v) \leq t \cdot d_G(u, v)$. The problem was
+introduced by \textcites{SpannerFirst89}{SpannerSecond89} and has been extensively studied since
+then, for it has many applications in network design. It was also showed to be \NPh{} to
+approximate~\autocite{SpannerNPHard07}. The most simple construction is a greedy
+algorithm~\autocite{greedySpanner93} that works similarly to the minimum spanning tree construction.
+Starting from an empty subgraph $H$, it goes through every edge $(u, v)$ of $G$ sorted by weight and
+check if there is a path between $u$ and $v$ in $H$ of length at most $t$. If it is case the edge
+$(u,v)$ is dropped, otherwise it is inserted in $H$. This results in a $(2t - 1)$-spanner with
+$O(n^{1+1/t})$ edges, which is an optimal trade-off between those two quantifies.  Furthermore on
+weighted graphs, the greedy spanner total weight is essentially optimal~\autocite{GreedyOpt16}.
+However, the best implementation of it, using a dynamic data structure~\autocite{fastGreedy04} is
+not scalable for it runs in $O(t n^{2+\nicefrac{1}{t}})$ and cannot easily be parallelized.
+Parallelization therefore requires other kind of approaches~\autocites{parSpanner08}{parSpanner15}.
+Recently, \textcite{Spanner17} showed how to obtain, for any $\epsilon > 0$, a $(2t - 1)$-spanner
+with $O(n^{1+1/k}/\epsilon)$ edges in $t$ rounds, with probability at least $1 - \epsilon$.
+
+\iffalse
 \url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
@@
 n}{n}\right)\right)$ edges in $O(|E|)$ time.
-
-the subgraph $H$ is said to be an $(\alpha, \beta)$-spanner of $G$ if, for a pair of parameters
-$\alpha \geq 1$, $\beta \geq 0$, and for every pair $u, v \in V$ of vertices, it holds that $d_H(u,
-v) \leq \alpha\cdot d_G(u, v)+\beta$.
 
@@
 problems, Computational Geometry, Volume 46, Issue 7, 2013) for more on the 2D case.
+
 In undirected graph, finding a spanner with less than $k$ edges is \NPc{} (theorem 2.2)
@@
 	mostly with the Laplacian of the tree ]
+\fi
+
+\paragraph{Other problems}
+
+Finding low stretch trees and spanners with respect to the existing edges is the most relevant
+problem when addressing the \esp{} problem. For the sake of completeness, we nonetheless give an
+overview of some related problems.
+
+For instance, \textcite{Johnson1978} define the following problem, where the stretch is defined over
+all possible pairs of nodes\footnote{We adapt their notations
+to match ours}:  
+\begin{problem}[Network Design Problem]
+  \label{prob:gtx_ndp}
+  Given an undirected integer-weighted graph $G=(V, E, w)$, a budget $B\in\Nbb$ and a criterion
+  threshold $C\in \Nbb$, does there exist a spanning subgraph $G'=(V, E')$ of $G$ with weight
+  $w(E') \leq B$ and criterion value $F(G') \leq C$, where the criterion function $F(G')$ denotes
+  the sum of the weights of the shortest paths in $G'$ between all vertex pairs?
+\end{problem}
+They prove that finding such a subgraph is \NPc{}, by exhibiting a reduction from the
+\textsc{Knapsack} problem. They also prove that the less general problem of finding a spanning tree
+on an unweighted graph, that is
+\vspace{-.5\baselineskip}
+\begin{problem}[Simple Network Design Problem]
+  \autoref{prob:gtx_ndp} with $w$ being the equal to $1$ for all edges in $E$ and $B=|V|-1$.
+\end{problem}%
+\vspace{-.5\baselineskip}
+\noindent is also \NPc{} by reduction from \textsc{Exact 3-Cover}.
+However, it has recently been show that this Simple Network Design problem can be approximated to a
+constant factor $6$~\autocite{AllPairStrech10}. Moreover, even when the graph is weighted,
+\textcite{constantDistortion07} achieve a universal constant bound for any weighted graph.
+
+Another problem appear when the low-stretch structure $H$ can include edges not in $G$ (as long as
+the distances in $H$ remain larger than the distances in $G$). This is captured by the following
+problem~\autocite{OptimalNetwork69}:
+\begin{problem}[Optimal Network Problem]
+  \label{prob:gtx_scott}
+  Given a set $V$ of $n$ vertices, find a set of spanning edges $E\subset V^2$ that minimizes
+  the sum of the length of the shortest paths  between all vertex pairs while the
+  total length of the resulting network does not exceed some upper bound $B\in\Nbb$.
+\end{problem}
+This can be seen as a special case of \autoref{prob:gtx_ndp} with $G$ being the unweighted
+$n$-complete graph. \Textcite{OptimalNetwork69} proposes a backtracking solution and two local search approximate
+algorithms. Some early branch and bound heuristic solutions to \autoref{prob:gtx_scott} are surveyed
+in~\autocite[Section 2.3.2]{networkDesignSurvey89} although they do not come with asymptotic
+guarantee on the stretch. Furthermore, \textcite{optimApproxNP80} proves that for any $\epsilon \in
+(0,1)$, finding a $|V|^{1-\epsilon}$ approximation is \NPc{}.
+However, if we consider the average stretches over a distribution of trees, then this approximation
+factor can be reduced to $\Theta(\log n)$~\autocite{lognMetricBoundConf03}.
+
+Finally, the stretch can also be computed for a subset of the edges. This is useful in cases where
+we have prior information on the importance of individual nodes or edges.  For instance,
+\textcite{RamseyTree17} show that for every $t$, any $n$-nodes graph $G=(V,E)$ has a subset $S$ of
+size at least $n^{1 - \nicefrac{1}{k}}$, and a spanning tree that has stretch $O ( k \log \log n)$
+between any node in $S$ and any node in $V$. Likewise, \textcite{mLAST17} describe how to maintain a
+light subgraph $H$ that minimizes the distance between pairs of source and sink that are given in an
+online fashion.
+
+As shown by \autoref{tab:gtx_related_stretch}, those problems defined in the seventies are still
+being discussed nowadays in top tier conferences, proving their relevance and impact beyond the
+\esp{} problem.
+
+\setlength{\fullpage}{\textwidth+\marginparsep+\marginparwidth}
+\begin{table}[htbp]
+  \centering
+  \caption{A summary of the lowest stretches achievable for various problems.
+  \label{tab:gtx_related_stretch}}
+  \begin{tabulary}{\fullpage}{LCCL}
+    \toprule
+    kind of stretch    & \multicolumn{2}{c}{only existing edges}  & extra edges allowed   \\
+    \midrule
+                       & tree                                     & not tree             &\\
+    \cmidrule(r){2-3}
+    some pairs         & $O(k\log\log n)$~\autocite{RamseyTree17} & \autocite[Section 4]{mLAST17} & --- \\
+    all existing pairs & $O\left(\log n (\log\log n)\right)$~\autocite{Abraham2012}
+		       & $(2t - 1)$-spanner, $O(n^{1+1/t})$ edges~\autocite{greedySpanner93}
+		       & $\Theta(\log n)$ in expectation~\autocite{lognMetricBoundConf03} \\
+    all possible pairs & $6$ for unweighted graphs \autocite{AllPairStrech10} and $O(1)$
+                         in general \autocite{constantDistortion07}
+		       & ---
+		       & no need for extra edges in that case                              \\
+    \bottomrule
+  \end{tabulary}
+\end{table}

### 2017-12-06 cc/conclusions.tex
write a conclusion for cc chapter

@@
-\begin{itemize}
-	\item improve stable \pcc{}
-	\item give guarantees for \gtx{}
-	\item handle weights in \gtx{}
-\end{itemize}
+\subsection{Summary}
+
+In this chapter, we expanded our efforts on characterizing edges in complex networks by addressing
+the \esp{} problem in undirected signed graphs. We first demonstrated, conceptually and practically,
+that the methods we designed in \autoref{chap:troll} are not suitable to this new setting. Drawing
+on the well established theory of social balance, we therefore came up with a different learning
+bias. More precisely, we posit that nodes latently belong to $K$ groups. Then that the edge signs
+are initially set to be positive if both endpoints belong to the same group and negative otherwise.
+And that finally we observe such assignment after it had been perturbed \uar{}.
+
+We connected twice this bias for the \esp{} problem with the \pcc{} problem. First, solving \pcc{}
+on a training set would provide us with the most reasonable partition of the nodes into $K$ groups,
+allowing us to predict the test signs accordingly. Second, the optimal value of the \pcc{} objective
+is a measure of the difficulty of the \esp{} problem~\autocite{Cesa-Bianchi2012b}. We thus presented
+a detailed overview of the \pcc{} problem and its proposed solution. Our goal was to point out that
+despite its hardness in the worst case, there exist efficient heuristics in practice, as well as
+methods with formal guarantees in more favorable cases. We furthermore noted that such cases are
+those when the sign assignment is close to follow our bias.
+
+Such general \pcc{} approaches would typically be used in a batch setting. In the last part of this
+chapter, we instead shifted our attention to the active setting, in the special case where $K=2$.
+This allowed us to exploit our bias in a more specific way. Namely, we described an implementation
+of the \gtx{} spanning data structure~\autocite{gtxFabio}. The idea is to query the sign of edges
+that connect the endpoints of every other edge by the shortest possible path.  Our experiments
+showed that this \gtx{} construction used in active \esp{} is competitive for some classes of
+graphs.
+
+\subsection{Future work}
+
+Beyond those preliminary experiments, we could push further our methods along several directions:
+
+\begin{enumerate}[1)]
+
+	\item The basic analysis \autoref{prop:gtx_correct} shows that on any graph with $n$ nodes, the
+		\gtx{} will perform at most $n$ collapses before terminating. However, as we mentioned
+		\vpageref{par:number_of_iteration} at the end of \autoref{sub:gtx_algo}, we believe that this
+		maximum number of collapses might be as low as $O(\log n)$. It turns out we were not able to
+		prove this statement. This is unfortunate, for besides giving us guarantees on the algorithm
+		runtime, this would also inform us about the stretch of the resulting tree. Indeed, let us
+		consider the reverse direction in which the \gtx{} algorithm operates. That is, instead of a
+		sequence of collapses transforming the original graph into a single node, we look at a sequence
+		of expansions turning this single node into the original graph. Informally, the length of the
+		longest path can only increase by a factor of five as each pair of nodes in $G_{t+1}$ expands
+		into two connected stars in $G_{t}$.\footnote{More precisely, consider the edge $(u,v)$ in
+		$G_{t+1}$. The two nodes are expanded into the two stars $s_u$ and $s_v$ in $G_t$, respectively
+		with peripheral nodes $\{u_1, \ldots, u_k\}$ and $\{v_1, \ldots, v_\ell\}$ and center $c_u$ and
+		$c_v$. In the worst case, there is now a path of length $5$ in $G_t$, \eg{} $(u_1, c_u, u_k,
+		v_\ell, c_v, c_1)$.} Whereas the behavior of the average stretch is less straightforward, it is
+		reasonable to expect that the smaller the number of expansions, the lower the stretch.
+
+	\item As mentioned in \autoref{sec:troll_related}, \esp{} could be extended to weighted signed
+		graphs. This is also relevant in undirected signed graphs. For instance, the signs might have
+		been generated by thresholding an expensive symmetric similarity function. In that case, we
+		might want to avoid extra evaluation of that function by predicting the signed weight of some
+		new edges. A natural way to modify the \gtx{} algorithm is the following. When choosing the
+		center of stars, pick the node with the current highest weighted degree. When choosing an edge
+		to connect two center, pick among the edges with the lowest endpoints eccentricity the one with
+		the lowest weight. What is less clear is to which extent such changes would affect the analysis.
+
+	\item We wish we had more time to further explore \pcc{} under stability assumptions, especially
+		as a way to handle the case where $K>2$. Recall that a clustering instance, made of a weighted
+		graph and a partition objective, is \emph{$\alpha$-stable} (with $\alpha>1$) if its optimal
+		partition remains the same whenever every weight $w_i$ is multiplied by a factor $c_i$ between
+		$1$ and $\alpha$. Furthermore, we say that an algorithm is \emph{robust} if, given a clustering
+		instance $\mathcal{I}$, it yields of the two following outcomes in polynomial time:
+		\begin{enumerate}[(i),nosep]
+			\item if $\mathcal{I}$ is $\alpha$-stable, returns the optimal solution of $\mathcal{I}$;
+			\item if $\mathcal{I}$ is not $\alpha$-stable, either returns the optimal solution of
+				$\mathcal{I}$ or reports that $\mathcal{I}$ is not $\alpha$-stable.
+		\end{enumerate}
+		This is a handy property because, in general, we cannot practically check the stability of an
+		instance. Indeed, since we do not know the optimal solution, we cannot tell whether it changes
+		or not under perturbations.
+
+		As an example, let us consider the \minmc{} problem. Given a weighted undirected graph $G$ and a
+		set of $k$ terminal nodes, the goal is to partition $G$ into $k$ clusters, each containing
+		exactly one terminal, such that the total weight of edges across clusters is minimized. This
+		bears some superficial similarity with \mmc{}. Recall from \autoref{ssub:cc_harness_approx} that
+		\mmc{} requires us the find the lightest set of edges to cut in order to separate $k$ pairs of
+		nodes.  Recall further that \mmc{} is equivalent to \pcc{}. As for \minmc{}, we can cast it as a
+		\pcc{} problem~\autocite[Theorem 24]{Bansal2004}. Interestingly, there exists a robust
+		algorithm for $2-\nicefrac{2}{k}$-stable instances of \minmc{}~\autocite{StableCC17}.  Under the
+		assumption that graphs obeying our learning bias are $\alpha$-stable, for some $\alpha$
+		depending of the graph irregularity, having such a stable algorithm for \mmc{} would be very
+		valuable.  However, this might prove challenging. Indeed, \mmc{} is harder than \minmc{}, as
+		illustrated by their respective approximation factor, $O(\log n)$ and less than
+		$1.3$~\autocite{minmcApprox17}.
+
+\end{enumerate}

### 2017-12-06 intro/main.tex
typos
take some intro feedback into account

@@
-Graphs are a natural way to represent the relationships between a set of entities. Because of the
+Graphs are a natural way to represent the relationships over a set of entities. Because of the
 simplicity and flexibility of this formalism, graphs are ubiquitous and have been used in countless
@@
 nodes are connected to four neighbors, namely the adjacent pixel from the top, bottom, left and
-right sides.
+right sides. The relationship in this graph is therefore adjacency in an image.
 \item In e-commerce, we can consider a co-purchase network. Nodes are products being sold on a
@@
 nodes and hundred of billions of edges. The availability of such large amount of structured data has
-prompted the development of automated methods to extract insight from it. For instance, it is
+prompted the development of automated methods to extract insights from them. For instance, it is
 possible to cluster nodes into coherent groups, predict the category of the nodes or study how to
@@
 At this point of the discussion though, we have only considered the most common and simple kind of
-graph, one representing a single type of relation between nodes. However, in many situations, there
+graph, one representing a single type of relation between nodes.\footnote{Note that by \enquote{type
+of relation}, we do not refer to some graph-theoretical characteristic of an edge such as being directed or weighted.
+Indeed, all the graphs we consider have homogeneous edges with that respect. Instead we mean domain
+specific semantic, as we will make clear in examples.} However, in many situations, there
 are two dimensions along which graphs exhibit more complexity. First, nodes have more than one type
@@
 products are movies, two movies can be frequently bought together because they are part of a series
-(like Star Wars), because they have the same director but different gender, because they receive the
+(like Star Wars), because they have the same director but different genre, because they receive the
 same prestigious awards in different years, and so on. The second type of relation (\enquote{same
-director} and \enquote{different gender}) is actually an example of a mixed relation, for it mixes
+director} and \enquote{different genre}) is actually an example of a mixed relation, for it mixes
 similarity and dissimilarity.
@@
 edge having opposite semantics. Such graphs are called \emph{signed graphs} and have been
-extensively studied since the fifties, for we shall see they have many applications. The second case
+extensively studied since the fifties, for we shall see they have many applications. Their name
+comes from the fact that edges are typically labeled $+1$ and $-1$. The second case
 is when there are more than two types of edge. Such graphs have also been studied for a long time
@@
 It is natural to expect that we can extract finer insights from graphs with edge semantics. However,
-in many cases, the information about edge types is not available. We therefore present in
+in many cases, the information about edge types is not available, at least not in a convenient form.
+We therefore present in
 \autoref{sec:predicting_edge_type} the problem of characterizing edges in complex networks.

### 2017-12-06 intro/semantics.tex
take some intro feedback into account

@@
 private messages, trade resources, attack each other and set a bounty on the head of another
-players. These six types of interactions are either positive or negative, but their nuances cannot
+player. These six types of interactions are either positive or negative, but their nuances cannot
 simply by explained by global similarity and dissimilarity. Another examples are the users of the
@@
 \paragraph{Economic networks} In our ever increasingly globalized economy, entities around the world
-are getting more and more tightly connected. However, it would be overly simplistic to consider
-there is only kind of link between entities. For something seemingly as simple as the connections
-between the largest \np{951} ports, one must already noticed that these connections can be
+are getting more and more tightly connected. However, those connections take on many different
+forms simultaneously. For something seemingly as simple as the connections
+between the largest \np{951} ports in the world, one must already notice that these connections can be
 implemented by any combinations of three kinds of ships: bulk dry carriers, container ships and oil

### 2017-12-06 intro/use_of_sign_graphs.tex
typos
take some intro feedback into account

@@
       \paragraph{Biology}
-      Signed graphs are also abundant in Biology.
+      Signed graphs are also abundant in biology.
       A typical input is a similarity matrix between genes expression level in various experimental conditions
@@
 	    solutions to that problem by \textcite{DeDup09}, who note that their non optimized
-	    implementation does not perform the best.
+	    implementation does not perform best.
 	 \item 

### 2017-12-10 conclusion/main.tex
write the conclusion

\label{chap:conclusion}

\section{Contributions}

The objective of this thesis was to provide \enquote{efficient and accurate methods to predict edge
type in complex networks}. By complex networks, we mean graphs where edges can have one among $k$
semantics (or types) and where nodes are connected by both partial homophily and heterophily.

\begin{itemize}

  \item In \autoref{chap:troll}~\autocite{trollSign17}, we first focused on directed signed graphs,
    as they model social interactions in several domains, such as e-commerce and collaborations.
    Here edges have two possible semantics, positive or negative. Our goal was to solve a supervised
    binary classification problem by predicting the sign of test edges. For that, we first
    introduced a simple sign generative model. It endows nodes with two parameters --trollness and
    unpleasantness--, governing their incoming and outgoing behavior. This might seem like a
    simplistic way of expressing the notion of partial homophily and heterophily. However, we
    presented two methods to approximate this Bayes predictor, suggesting a trade off between
    theoretical guarantees and input requirement. We showed experimentally that those methods are
    competitive with the state of the art, while being more scalable. Motivated by the regularity
    measure offered by trollness and unpleasantness, and the inability of existing methods to cope
    with new nodes joining the graph, we furthermore develop the first online algorithm for the
    \esp{} problem.

  \item We started \autoref{chap:cc} by showing that the methods of the previous chapter are not
    straightforward to extend to undirected signed graphs. Therefore, we changed our learning bias
    from a generative model to a clustering assumption inspired by social balance theories. Namely,
    we posit that the nodes are assigned to one of $K$ groups, and that the sign of the edge $(u,v)$
    is positive if the nodes $u$ and $v$ belong to the same group and negative otherwise. We then
    presented two approaches to leverage this bias. First, we drew a connection between the \esp{}
    problem and the \pcc{} problem. While \pcc{} is hard to approximate in general, we suggested,
    based on existing works, that the more an instance complies with our bias, the better \pcc{}
    methods could recover the $K$ underlying groups and thus provide accurate prediction. Second, we
    addressed the problem in an active setting, where $K=2$. For that, we implemented a promising
    spanning tree algorithm~\autocite{gtxFabio}, that queries some edges in order to connect the
    remaining edges by paths as short as possible. Experimentally, we showed that on both synthetic
    and real graphs, our method is competitive in terms of sign prediction with a strong natural
    baseline.

  \item Finally, in \autoref{chap:edge}, we considered the case where there are more than two types
    of edge. This means that instead of signed graphs, our input shifted to multilayer graphs, which
    are also ubiquitous to describe real world phenomena. Moreover, we modified the learning
    setting, so that no edge label are available, but instead nodes carry a profile vector. However,
    our goal remained to predict edge type, in what could then be described as a clustering problem
    with side information and interpretability requirements. In order to do so, we first defined a
    similarity between nodes, which accommodate both heterophily and homophily. We then formulate
    the problem as simultaneously finding $k$ special vectors called directions, and assigning one
    of them to every edge in order to maximize a scalar \emph{goodness} function. After introducing
    baselines oblivious of the graph topology, we added constraints resulting in one vectorial and
    one matricial optimization objectives. Finally, we showed on synthetic data how those methods
    can recover the planted ground truth.

\end{itemize}

\section{Future work}

In this section, we revisit the problems summarized above by showing how they could be applied in a
new context or enriched by existing node embedding methods.

\subsection{Reciprocal recommendation}

In this paragraph, we present the reciprocal recommendation problem~\autocite{Pizzato2013}. As we
shall see, it incorporates aspects from all the problems we discussed previously. Broadly speaking,
given two sets of users $L$ and $R$, we want to recommend users in $R$ to users in $L$, and vice
versa. These recommendations are based on mutual interest, but in most cases there is an asymmetry
between the two groups. To better illustrate this general definition, we now give some concrete
examples.

\begin{itemize}
  \item One recent application that quickly rose to prominence is online
    dating~\autocites{Krzywicki2015}{Xia2016a}{Alanazi}. In the most common case, users are divided
    between men and women, and we want to recommend users of the opposite gender in order to create
    couples. A well known example of such a system is Tinder~\autocite{TinderDesc16}.
  \item In another domain, we can also match job seekers and
    employers~\autocites{Siting2012}{Hong2013}{Kille2015}.
  \item As a specific case in point, \textcite{Zhang2016a} propose a method to match prospective PhD
    students and their future supervisors.
  \item In large organizations or associations, having a mentor is usually beneficial for
    both employees and employers~\autocite{mentorBenef04} and finding mentee/mentor pairs can be
    cast as a reciprocal recommendation problem~\autocite{OSSMentors12}.
    %\url{http://ieeexplore.ieee.org/abstract/document/7207130/}
  \item After moving to their new jobs, people also need a place to live. Again, we can match
    tenants and landlords, or people to share a flat.
  \item Finally, in a more critical setting, finding relevant matches between patients and organ
    donors can improve the success of transplants~\autocite{patientDonor17}.
\end{itemize}

Reciprocal recommendations present specific challenges~\autocite{Andrews2015}. Among them are
asymmetry and volume.  Asymmetry refers to the imbalance between the two groups, in terms of size
and bargaining power. For instance, depending of the economic situation and the field, there might
be a lot of job seekers for a few open positions, in which case employers can be very selective.
Likewise, it is generally observed that dating websites have a higher proportion of men, meaning
that women can afford to choose among numerous suitors~\autocite{TinderDesc16}. By volume, we mean
that users on both side can only handle a few recommendations at the same time. This is sharp
contrast with traditional user-item recommendation systems, where recommending simultaneously the
same item (say a movie) to thousands of users is not an issue. On the other hand, we have to keep
user engaged and prevent them from being idle. This is done by ensuring they are regularly the
subject or object of some recommendations.

As the name implies, reciprocal recommendation has often be seen as a special case of recommender
system~\autocite{Andrews2015}. Besides this paradigm, we could also cast it as an adaptive, active,
cost sensitive \esp{} problem on bipartite directed signed graphs with side information. Let us
decompose this last sentence, starting with the input graph. It is naturally bipartite, because of
the presence of two groups in which there is no internal recommendation. Furthermore, every time one
user expresses interest or disinterest in another user, this creates a directed signed edge. Such a
graph is thus a special case of the \dssn{} studied in \autoref{chap:troll}. Each nodes carries two
kind of information. The first is explicitly provided by the user, in order to arise interest in
users from the other group. The second is implicitly inferred from the user activity. In particular,
the interactions with other nodes allow us to build an implicit preference model for every user.
Combining these two sources of information yield a profile of each user, like in
\autoref{chap:vector}. Making recommendations in such a graph therefore boils down to finding non
existing positive edges. However, to cope with the two challenges presented in the previous
paragraph, we follow the setting of \autoref{chap:cc} and use an active approach. More precisely,
respecting the volume constraints would require an adaptive algorithm, which acquires the most
useful information without over- or under-loading any nodes. As for the group asymmetry, this could
be addressed by using a cost sensitive classification error, so that the error rate at a given node
is inversely proportional to its importance in the network.

Unfortunately, we did not study this problem further in this thesis. Indeed, getting data and
testing algorithms usually require an industrial partner.\footnote{See for
instance~\autocites{Andrews2015}{Kille2015}{Krzywicki2015}{Xia2015}{Alanazi}}

\subsection{Representation learning}
\label{par:representation_learning}

As we mentioned already in \autoref{sec:troll_related} and \autoref{sec:edge_related}, a recent and
promising trend regarding learning in graph is node embedding. In addition to the references cited
earlier, this also witnessed by recent works on embedding attributed nodes, either in signed
graphs~\autocites{SignedFS17}{attributedSNE17}{SHINE18} or in general
graphs~\autocites{Liao2017}{AANE17}{DynamicEmbedding17}. At first sight, those methods seem
radically different from traditional graph methods. However, just like word embedding methods have
been linked to matrix
factorization~\autocites{WordEmbeddingAsFactorization14}{Hellinger14}{AroraRandWalk16}, so have been
popular and successful node embedding methods~\autocite{embeddingAsFactorization18}. These methods
are the simple ones that do not consider node attributes and multiple types of edges, such as
DeepWalk~\autocite{DeepWalk14}, LINE~\autocite{LINE15} and node2vec~\autocite{node2vec16}. For
instance, \textcite{embeddingAsFactorization18} show that the node embedding of DeepWalk on the
weighted graph $G$ are approximately the left singular vectors of the following matrix:
\begin{equation}
  \label{eq:deepwalk}
  \log \left( vol(G) \left( \frac{1}{T} \sum_{r=1}^T (D^{-1}A)^r\right) D^{-1} \right) - \log b
\end{equation}
where $vol(G)$ is the sum of all weighted degrees, $T$ is the context window size, $D$ is the
diagonal matrix of degrees, $A$ is the adjacency matrix of the graph and $r$ is the number of
negative samples in skip-gram. There are several ways such findings could be used in our context:
\begin{itemize}

  \item If the edges of the input graph are fully labeled into $k$ types, then we have $k$ matrices
    $D_k$ and $A_k$, by considering the subgraphs induced by each edge type. If we simply decompose
    the $k$ matrices constructed according to equation \eqref{eq:deepwalk}, we would have $k$
    embedding for each node. Instead, it seems natural to stack those matrices together to form two
    third order tensors $\mathcal{D} \in \Rbb^{n \times n \times k}$ and $\mathcal{A} \in \Rbb^{n
    \times n \times k}$. One could then use one tensor decomposition
    methods~\autocites{tensorReview09}{tensorReview17} to extract node embedding taking into account
    the multiplicity of edge types.

  \item If the edges of the input graph are partially labeled, then predicting edge type can be seen
    as a tensor completion problem. Like in the matrix case, we seek a low rank tensor that
    coincides with the observed entries of the input data. Furthermore, when we have access to node
    profiles, we can use methods tailored to leverage such side information~\autocite[Section
    4]{tensorCompletion17}.

  \item If the input graph has no edge labels at all but we are given node profiles, we can use
    existing attributed embedding approaches~\autocites{Liao2017}{AANE17}{DynamicEmbedding17} to
    generate extra information about each node, before applying the methods presented in
    \autoref{chap:vector} on these extended profiles. We note however that
    doing so might reduce the interpretability of the final result.

\end{itemize}

### 2017-12-11 abstract.tex
first version of abstract

In this thesis, we develop methods to efficiently and accurately characterize edges in complex
networks. In simple graphs, nodes are connected by a single semantic. For instance, two users are
friends in a social networks, or there is a hypertext link from one webpage to another. Furthermore,
those connections are typically driven by node similarity, in what is known as the homophily
mechanism. In the previous examples, users become friends because of common features, and webpages
link to each other based on common topics. By contrast, complex networks are graphs where every
connection has one semantic among $k$ possible ones. Those connections are moreover based on both
partial homophily and heterophily of their endpoints. This additional information enable finer
analysis of real world graphs. However, it can be expensive to acquire, or is sometimes not known
beforehand. We address the problems of inferring edge semantics in various settings. First, we
consider graphs where edges have two opposite semantics, and where we observe the label of some
edges. These so-called \emph{signed graphs} are a convenient way to represent polarized
interactions. We propose two learning biases suited for directed and undirected signed graphs
respectively. This leads us to design several algorithms leveraging the graph topology to solve a
binary classification problem that we call \esp{}. Second, we consider graphs with $k \geq 2$
available semantics for edge. In that case of \emph{multilayer graphs}, we are not provided with any
edge label, but instead are given one feature vector for each node. Faced with such an unsupervised
\ecp{} problem, we devise a quality criterion expressing how well an edge $k$-partition and $k$
semantical vectors explains the observed connections. We optimize this \emph{goodness of
explanation} criterion in vectorial and matricial forms, and show how those two methods perform on
synthetic data.

### 2017-12-12 cc/gtx_related.tex
fix missing ref

@@
 \end{enumerate}
-Note that for an unweighted graph, the second problem reduces to minimizing $\sum_{(u,v) \in E}
-|\pathhuv|$. If furthermore $H$ is tree, this is equivalent to minimize the second term of equation
+Note that for unweighted graphs, the second problem reduces to minimizing
+\begin{equation}
+  \label{eq:test_stretch_def}
+  \estr(H) = \sum_{(u,v) \in E} |\pathhuv|.
+\end{equation}
+If furthermore $H$ is tree, this is equivalent to minimize the second term of equation
 \eqref{eq:stretch_mistakes}. Therefore we focus mainly of that definition of stretch, and consider

### 2017-12-12 conclusion/main.tex
fix missing ref

@@
 
-  \item Finally, in \autoref{chap:edge}, we considered the case where there are more than two types
+  \item Finally, in \autoref{chap:vector}, we considered the case where there are more than two types
     of edge. This means that instead of signed graphs, our input shifted to multilayer graphs, which
@@
 
-In this paragraph, we present the reciprocal recommendation problem~\autocite{Pizzato2013}. As we
-shall see, it incorporates aspects from all the problems we discussed previously. Broadly speaking,
+We first present a task tightly related to our edge characterization problem, namely the reciprocal
+recommendation problem~\autocite{Pizzato2013}. As we shall see, it indeed incorporates aspects from
+all the problems we discussed previously, along with new constraints. Broadly speaking,
 given two sets of users $L$ and $R$, we want to recommend users in $R$ to users in $L$, and vice
@@
 
-\begin{itemize}
+\begin{itemize}[leftmargin=*]
   \item One recent application that quickly rose to prominence is online

### 2017-12-12 conclusion/main.tex
be smoother in the future work conclusion

@@
 
-Unfortunately, we did not study this problem further in this thesis. Indeed, getting data and
-testing algorithms usually require an industrial partner.\footnote{See for
-instance~\autocites{Andrews2015}{Kille2015}{Krzywicki2015}{Xia2015}{Alanazi}}
+As in any worthy machine learning endeavor, reliably evaluating and comparing approaches require
+large amount of real data. This is particularly true in such an applied setting, but the most
+realistic evaluation is to perform A/B testing on a live system with enough traffic to give some
+statistical significance. This explains why most previous studies rely on the cooperation of
+industrial partners~\autocites{Andrews2015}{Kille2015}{Krzywicki2015}{Xia2015}{Alanazi}. Thus, this
+is also something we would like to do in the future.
 
@@
 
-As we mentioned already in \autoref{sec:troll_related} and \autoref{sec:edge_related}, a recent and
+A fundamental question underlying many machine learning approaches is the representation of data.
+When extracting information from the real world, we are usually given some freedom in how to present
+it to learning algorithms. Ideally, we would like to use the representation providing the best
+performance in the learning task. One solution is to ask expert domains to craft relevant features
+based on their prior knowledge. While this has achieved successes, it is not scalable to the present
+day situation, where the deluge of available data has prompted the use of machine learning in every
+corner of the society. An alluring alternative to expert knowledge is to learn the data
+representation itself, guided by the task.
+
+As an illustration in the context of this thesis, we assume that data was organized as a graph,
+where nodes were possibly associated with a feature vector. However, there might an different
+representation of those nodes that would better suit our needs. Indeed,
+we mentioned already in \autoref{sec:troll_related} and \autoref{sec:edge_related} that a recent and
 promising trend regarding learning in graph is node embedding. In addition to the references cited

### 2017-12-13 final.tex
start changing the layout

\documentclass[a4paper,svgnames,twoside,onecolumn,final,titlepage,11pt]{report}
\input{pmin}
\input{abbreviations}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}

\title{Characterizing edges in signed and vector-valued graphs}
\author{Géraud Le Falher}

\begin{document}
\newgeometry{vmargin=1.5cm,inner=3.8cm,bindingoffset=0.0cm,outer=3.3cm,marginparsep=0pt,marginparwidth=0pt}

\maketitle

% \section*{Abstract}
% \input{abstract}
%
% \tableofcontents
% \newpage
%
% \input{notations}
%
% \chapter{Introduction}
% \subimport*{intro/}{main}

\chapter{On the Troll-Trust Model for edge sign prediction in Social Networks}
\subimport*{troll/}{main}
\end{document}

\chapter{Edge sign prediction in general graphs and Correlation Clustering}
\subimport*{cc/}{main}

\chapter{Edge clustering in node attributed graphs}
\subimport*{edge/}{main}

\chapter{Conclusion}
\subimport*{conclusion/}{main}

\vspace{\baselineskip}
\begingroup
\newgeometry{hmargin=1.5cm,vmargin=1.3cm}
\setstretch{0.87}
\setlength\bibitemsep{1pt}
\todos
\printbibliography
\restoregeometry
\endgroup
\end{document}

### 2017-12-13 troll/algs_batch.tex
finish troll layout

@@
   \sum_{v \in V_{\mathrm{in}}} \din(v) \pbar_v \approx
-  \sum_{v \in V_{\mathrm{in}}} \din(v) \mu_p = |E| \mu_p \label{eq:troll_mup}\quad \text{and} \\
-  \sum_{u \in V_{\mathrm{out}}} \left( \sum_{v \in \NNout(u)} q_v \right) &=
+  \sum_{v \in V_{\mathrm{in}}} \din(v) \mu_p = |E| \mu_p \label{eq:troll_mup} \\
+	&\text{and} \\ \sum_{u \in V_{\mathrm{out}}} \left( \sum_{v \in \NNout(u)} q_v \right) &=
   \sum_{u \in V_{\mathrm{out}}} \dout(u) \qbar_u \approx
@@
   - \frac{p_u+q_v}{2} \right| \le 8\ve \]
-simultaneously holds with high probability for each $(u,v) \in \theset{(u,v) \in E}{\din(v) \ge
-  \theta,\, \dout(u) \ge \theta} \setminus E_0$.
+simultaneously holds with high probability for each \[(u,v) \in \theset{(u,v) \in E}{\din(v) \ge
+\theta,\, \dout(u) \ge \theta} \setminus E_0.\]
 
@@
 
-\begin{equation*}
+\begin{align*}
 {\widehat f}\big(\bp,\bq,{\yuv}_{(u,v) \in E\setminus E_0}\big)
-= f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)
-+ \frac{1}{2}\sum_{u\in V}
+= &f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq) +\\
+  &\frac{1}{2}\sum_{u\in V}
 \left(\dout(u)\Bigl(p_u-\frac{1}{2}\Bigl)^2+\din(u)\Bigl(q_i-\frac{1}{2}\Bigl)^2\right)
-\end{equation*}
+\end{align*}
 

### 2017-12-13 troll/generative.tex
finish troll layout

@@
 
+\begin{wrapfigure}[8]{o}[10mm]{50mm}
+	\centering
+	\includegraphics[width=0.99\linewidth]{tikz/troll_genmodel_tikz}
+	\caption{\small The sign \yuv{} of the edge \euv{} is positive with probability $\frac{1}{2}(p_u+q_v)$.
+	\label{fig:troll_genmodel}}
+\end{wrapfigure}
 We now define the stochastic generative model for edge labels we use in the batch learning setting.
-Given the graph $G = (V,E)$, let the label $\yuv \in \spin$ of directed edge $(u,v) \in E$ be
+Given the graph $G = (V,E)$, let the label $\yuv \in \spin$ of directed edge $(u,v) \in E$ be%
 generated as follows. Each node $u \in V$ is endowed with two latent parameters $p_u, q_u \in
@@
 This process is illustrated in \autoref{fig:troll_genmodel}.
-\begin{marginfigure}
-	\centering
-	\includegraphics[width=0.98\linewidth]{tikz/troll_genmodel_tikz}
-	\caption{The sign \yuv{} of the edge \euv{} is positive with probability $\frac{1}{2}(p_u+q_v)$.
-	\label{fig:troll_genmodel}}
-\end{marginfigure}
 

### 2017-12-13 troll/notation_and_prel.tex
start changing the layout
finish troll layout

@@
 of node $u \in V$. Most of these notations are illustrated on \autoref{fig:troll_Gexample}.
-\begin{figure}[htpb]
+\begin{figure}[hbpt]
   \centering
-  \includegraphics[width=0.9\linewidth]{tikz/troll_notations_tikz}
+  \includegraphics[width=0.8\linewidth]{tikz/troll_notations_tikz}
   \caption{Part of a DSSN centered on node $u$. \label{fig:troll_Gexample}}
@@
 
-\begin{figure*}[t]
+\begin{figure*}[tbp]
   \centering
@@
 Let us first introduce the notion of cutsize of an undirected node-labeled graph $G'(Y)$, which is
+\begin{wrapfigure}{o}[10mm]{45mm}
+	\centering
+	\includegraphics[width=0.85\linewidth]{assets/raw/troll_labeled_gprime.pdf}
+	\caption{\small The node-labeling of $G'$ illustrating the relation between the edge label
+	  complexity and the cutsize. Here the four edges part of the cut are in heavy black.
+	\label{fig:troll_psi}}
+\end{wrapfigure}
 the number of edges in $G'$ connecting nodes that have mismatching labels.
@@
   would result in a label version of $G'$ showed on \autoref{fig:troll_psi}.
-\begin{marginfigure}
-	\centering
-	\includegraphics[width=0.98\linewidth]{assets/raw/troll_labeled_gprime.pdf}
-	\caption{The node-labeling of $G'$ illustrating the relation between the edge label
-	  complexity and the cutsize. Here the four edges part of the cut are in heavy black.
-	\label{fig:troll_psi}}
-\end{marginfigure}
   We see that by labeling circle nodes by the majority label of the square nodes there are connected

### 2017-12-13 troll/related.tex
finish troll layout

@@
 		difference between its opinion of a neighbor $v$ and what $v$ truly deserve according to the
-		network: $bias(u) = \frac{1}{\dout(u)}\sum_{v\in\NNout(u)}(\wuv - deserve(v))$. The deserve of a
-		node $v$ is the aggregated opinion of its in-neighbors, discounted by their bias: $deserve(v) =
-		\frac{1}{\din(v)}\sum_{u\in\NNin(v)}\left(\wuv(1-\max\{0, \sgn(\wuv)\times bias(u)\})\right)$.
+		network: \[bias(u) = \frac{1}{\dout(u)}\sum_{v\in\NNout(u)}(\wuv - deserve(v)).\] The deserve of a
+		node $v$ is the aggregated opinion of its in-neighbors, discounted by their bias: \[deserve(v) =
+		\frac{1}{\din(v)}\sum_{u\in\NNin(v)}\left(\wuv(1-\max\{0, \sgn(\wuv)\times bias(u)\})\right).\]
 	\item	Reputation and Optimism~\autocite{shahriari2014ranking}, defined for a node $u$ by

### 2017-12-13 edge/intro.tex
fix layout of edge|vector chapter

@@
 
-\setlength{\fullpage}{\textwidth+\marginparsep+\marginparwidth}
+\setlength{\fullpage}{179mm}
 \begin{table}[h]
+\begin{adjustwidth}{-2cm}{}
   \small
@@
   \end{tabulary}}
+\end{adjustwidth}
 \end{table}
@@
 leveraging possibly richer attributes information also exists in link
-prediction~\autocite{linkPredSurvey16}\stodo{What they actually write is \enquote{Topology-based
-techniques are more versatile than attribute-based methods since they are not domain specific} which
-sounds a bit unfair: only the attributes are domain specific, not necessarily the methods
-themselves}.
+prediction~\autocite{linkPredSurvey16}. %\stodo{What they actually write is \enquote{Topology-based
+%techniques are more versatile than attribute-based methods since they are not domain specific} which
+%sounds a bit unfair: only the attributes are domain specific, not necessarily the methods
+%themselves}.
 % examples of using node attributes for link prediction, taken from that survey:

### 2017-12-13 edge/model.tex
typos
fix layout of edge|vector chapter

@@
 connected. This requires the introduction of three further concepts:
-\begin{inparaenum}[1)]
+\begin{enumerate*}[1)]
 \item how to represent a connection (\ie{} an edge),
@@
 \item how to evaluate which explanation is the best.
-\end{inparaenum}
+\end{enumerate*}
 We now present our modelling choice:
@@
     where $\circ$ stands for the Hadamard (or component wise) product. As showed in
-    \begin{marginfigure}
+    \autoref{tab:edge_sim}, for each dimension $i \in \rangesd$, the $s_{uv}$ holds a value
+    indicating whether the profiles of $u$ and $v$ agree, disagree or are indifferent along that
+    dimension.
+    \begin{table}[hb]
+      \centering
       % \small
-      \captionsetup{font=small,type=table}
-      \captionof{table}{We show the effect of the Hadamard product in combining information from two
+      % \captionsetup{font=small,type=table}
+      \caption{\small We show the effect of the Hadamard product in combining information from two
         users $u$ and $v$. Instead of actual numbers, in this table we use $+$ to denote a large
@@
       \end{tabular}
-    \end{marginfigure}
-    \autoref{tab:edge_sim}, for each dimension $i \in \rangesd$, the $s_{uv}$ holds a value
-    indicating whether the profiles of $u$ and $v$ agree, disagree or are indifferent along that
-    dimension.
+    \end{table}
   \item The $k$ explanations take the form of bounded norm vectors $\mathcal{D}_k = \{w_1, w_2,
@@
 solution of having a single direction $w_{uv} = \frac{s_{uv}}{||s_{uv}||}$ for every edge $(u,v)$.
-More generally, as $k$ increases, we expect the value $\max_{\mathcal{D}_k} \sum_{u,v \in E}
-\max_{\ell \in \rangesk} g(s_{uv}, w_{\ell})$ to increase as
+More generally, as $k$ increases, we expect the value \[\max_{\mathcal{D}_k} \sum_{u,v \in E}
+\max_{\ell \in \rangesk} g(s_{uv}, w_{\ell})\] to increase as
 well, at the cost of interpretability. There could be a principled approach to finding the
@@
   \item create a new partition $\{E'_1, \ldots, E'_k\}$ by assigning every edge $(u,v)$ to the
-    direction $w$ maximizing its goodness ${s_{uv}}^T w$, that is $E'_\ell = \theset{(u,v) \in E}{
-    w_\ell = \argmax_{w_i \in \mathcal{D}_k} {s_{uv}}^T w_i}$, and
+    direction $w$ maximizing its goodness ${s_{uv}}^T w$, that is \[E'_\ell = \theset{(u,v) \in E}{
+    w_\ell = \argmax_{w_i \in \mathcal{D}_k} {s_{uv}}^T w_i}\]
   \item update the directions to make them optimal with respect to the edges currently assigned to

### 2017-12-13 cc/bias_balance.tex
fix cc chapter layout

@@
 
+\begin{wrapfigure}{o}[10mm]{45mm}
+  \centering
+  \includegraphics[width=.9\linewidth]{assets/raw/2cc.pdf}
+  \caption{\small A two-clustering of a complete strongly balanced graph}
+  \label{fig:cc_twostrongcc}
+\end{wrapfigure}
 It is interesting to note that this local property of triangles gives raise to a global structure.
@@
 a negative edge.
-\begin{marginfigure}
-  \centering
-  \includegraphics[width=28mm]{assets/raw/2cc.pdf}
-  \caption{A two-clustering of a complete strongly balanced graph}
-  \label{fig:cc_twostrongcc}
-\end{marginfigure}
 We say that $L$ and $R$ are $2$-consistent clusters and more generally, we define

### 2017-12-13 cc/gtx_algorithm.tex
fix cc chapter layout

@@
 \paragraph{\extractStar{}}\label{par:extractstar}%
+\begin{wrapfigure}{o}[10mm]{45mm}
+  \centering
+  \includegraphics[width=.9\linewidth]{assets/tikz/gtx_star_tikz.pdf}
+  \caption[A sample star]{\small A sample star created during the \tth{} collapse level. The black node
+    is the center $c_i$ of the star $S_i^t$, which is also made of the four light gray peripheral nodes
+  as well as the solid edges. The 2-hops neighbors of $c_i$ are the white nodes
+  $h_1$ to $h_3$, whose degree will decrease once $S_i^t$ is removed from $G_t$.}
+  \label{fig:gtx_star_simple}
+\end{wrapfigure}
 \extractStar{} takes as input a graph $G_t=(V_t, E_t)$.
@@
 well as the lookup of a value associated with a particular key.
-\begin{marginfigure}
-  \centering
-  \includegraphics[height=0.15\textheight]{assets/tikz/gtx_star_tikz.pdf}
-  \caption[A sample star]{A sample star created during the \tth{} collapse level. The black node
-    % \tikz{\node[vertex,rare] {$c_i$};}
-    is the center $c_i$ of the star $S_i^t$, which is also made of the four light gray peripheral nodes
-  % \tikz{\node[vertex,medium] {$p_1$};} to \tikz{\node[vertex,medium] {$p_4$};}
-  as well as the solid edges. The 2-hops neighbors of $c_i$ are the white nodes
-  % \tikz{\node[vertex] {$h_1$};} to \tikz{\node[vertex] {$h_3$};}
-  $h_1$ to $h_3$, whose degree will decrease once $S_i^t$ is removed from $G_t$.}
-  \label{fig:gtx_star_simple}
-\end{marginfigure}
 
@@
   \end{algorithmic}
-  \rule{\textwidth}{.3pt}
+  % \rule{\textwidth}{.3pt}
 \end{center}
 
+\vspace*{-2\baselineskip}
 \begin{prop}
@@
 
-\begin{marginfigure}
+\collapseStar{} not only returns the new graph $G_{t+1}$ but also a map $cross\_edges$. This map is
+used to keep track of which edge in $E_t$ connects any pair of node in $V_{t+1}$ (as illustrated in
+\autoref{fig:gtx_cross_edge}). It associates to any edge in $E_{t+1}$ a set of edges from $E_t$.
+This set can have an arbitrary size during the execution of \collapseStar{}, yet it is guaranteed to
+contain a single edge upon termination. Although this does not appear in the following pseudo code,
+in practice we ensure that $(s_u, s_v)$ and $(s_v, s_u)$ refer to the same edge in $cross\_edges$.
+
+\begin{figure}[bpt]
   \centering
-  \includegraphics[width=0.92\linewidth]{assets/tikz/gtx_cross_edges_tikz.pdf}
+  \includegraphics[width=0.4\linewidth]{assets/tikz/gtx_cross_edges_tikz.pdf}
   \caption[Cross edge representation]{In this graph, there are four possible edges between the two
@@
   \label{fig:gtx_cross_edge}
-\end{marginfigure}%
-\collapseStar{} not only returns the new graph $G_{t+1}$ but also a map $cross\_edges$. This map is
-used to keep track of which edge in $E_t$ connects any pair of node in $V_{t+1}$ (as illustrated in
-\autoref{fig:gtx_cross_edge}). It associates to any edge in $E_{t+1}$ a set of edges from $E_t$.
-This set can have an arbitrary size during the execution of \collapseStar{}, yet it is guaranteed to
-contain a single edge upon termination. Although this does not appear in the following pseudo code,
-in practice we ensure that $(s_u, s_v)$ and $(s_v, s_u)$ refer to the same edge in $cross\_edges$.
+\end{figure}
 
@@
   \end{algorithmic}
-  \rule{\textwidth}{.3pt}
+  \vspace*{-.5\baselineskip}
+  % \rule{\textwidth}{.3pt}
 \end{center}
@@
 \vspace{0.5\baselineskip}
-\setlength{\fullpage}{\textwidth+\marginparsep+\marginparwidth}
+\setlength{\fullpage}{179mm}
+\begin{center}
+\begin{adjustwidth}{-2cm}{}
 \noindent\begin{tabulary}{\fullpage}{lllL}
@@
 \end{tabulary}
+\end{adjustwidth}
+\end{center}
 \vspace{0.5\baselineskip}
@@
 		\caption{Spanning tree of $G_1$}\label{fig:gtx_spanning_tree_1}
-	\end{subfigure}~
+	\end{subfigure}
+
 	\begin{subfigure}[b]{53mm}
@@
 and two nodes stars, as in \autoref{fig:gtx_line_graph}.%
-\begin{marginfigure}
+This will results in $\nicefrac{2n}{5}$ stars organized in a line, which is less than half of $n$.
+Thus we can conclude that in this case, \gtx{} will finish after $O(\log n)$ iterations. Note that a barbell graph (two
+cliques connected by a line) would requires a number of iterations logarithmic in the length of that
+central line, despite having many more nodes and edges.
+\begin{figure}[bth]
   \centering
@@
   \label{fig:gtx_line_graph}
-\end{marginfigure}
-This will results in $\nicefrac{2n}{5}$ stars organized in a line, which is less than half of $n$.
-Thus we can conclude that in this case, \gtx{} will finish after $O(\log n)$ iterations. Note that a barbell graph (two
-cliques connected by a line) would requires a number of iterations logarithmic in the length of that
-central line, despite having many more nodes and edges.
+\end{figure}
 

### 2017-12-13 cc/gtx_experiments.tex
fix cc chapter layout

@@
 		The synthetic ones are square, while the \enquote{real world} ones represents the four neighbors
-		pixel connectivity of the pictures showed in \autoref{fig:gtx_xp_bwpics}.
+		pixel connectivity of the pictures showed in \autoref{fig:gtx_xp_bwpics}
+		\vpageref{fig:gtx_xp_bwpics}.
 	\item[\lpa{}] which are built synthetically according to the model of \textcite{Barabasi1999}.
@@
 
-On \lpa{} and \triangle{}, we see in \autoref{fig:gtx_xp_st} that both trees exhibits logarithmic stretch, although with a
+On \lpa{} and \triangle{}, we see in \autoref{fig:gtx_xp_st} \vpageref{fig:gtx_xp_st} that both trees exhibits logarithmic stretch, although with a
 larger constant for \gtx{}. Note that this is also the case for others low stretch tree methods
@@
 Correlation Coefficient (MCC), defined in equation~\autoref{eq:troll_mcc} \vpageref{eq:troll_mcc}.
-As showed in \autoref{fig:gtx_xp_mcc}, when the noise level is low, \gtx{} performs better than
+As showed in \autoref{fig:gtx_xp_mcc} \vpageref{fig:gtx_xp_mcc}, when the noise level is low, \gtx{} performs better than
 \bfs{}. As the noise level gets higher, they have similar performance. Note also than in
@@
 As the \asym{} has a $O(n^3)$ complexity and uses quite some memory at prediction time, the larger
-graphs used previously are not all included. The conclusion of \autoref{fig:gtx_xp_asym} is that
+graphs used previously are not all included. The conclusion of \autoref{fig:gtx_xp_asym} \vpageref{fig:gtx_xp_asym} is that
 except on social networks, it is better to use spanning trees than random edges. Specifically, \gtx{}
@@
 	\centering
-	\begin{subfigure}[b]{0.47\textwidth}
+	\begin{subfigure}[b]{0.475\textwidth}
 		\includegraphics[width=\textwidth]{gtx_exp/grsynthasym}
@@
 	\end{subfigure}~
-	\begin{subfigure}[b]{0.47\textwidth}
+	\begin{subfigure}[b]{0.475\textwidth}
 		\includegraphics[width=\textwidth]{gtx_exp/grrwasym}
@@
 	\end{subfigure}
-	\begin{subfigure}[b]{0.47\textwidth}
+	\begin{subfigure}[b]{0.475\textwidth}
 		\includegraphics[width=\textwidth]{gtx_exp/pasynthasym}
@@
 	\end{subfigure}~
-	\begin{subfigure}[b]{0.47\textwidth}
+	\begin{subfigure}[b]{0.475\textwidth}
 		\includegraphics[width=\textwidth]{gtx_exp/trasym}
@@
 	\end{subfigure}
-	\begin{subfigure}[b]{0.47\textwidth}
+	\begin{subfigure}[b]{0.475\textwidth}
 		\includegraphics[width=\textwidth]{gtx_exp/parwasym}

### 2017-12-16 intro/learning.tex
pass over the intro

@@
     connected groups that are loosely connected with the rest of the graph. This allows for a better
-    understanding of the graph organization, and present a higher level view than looking at
-    individual nodes relationships. For instance, it has been used to identify proteins functional
-    groups~\autocite{clusterBio03} or see how different scientific fields relate based on
+    understanding of the graph organization, and present a higher level view than looking at the
+    individual node relationships. For instance, it has been used to identify proteins functional
+    groups~\autocite{clusterBio03}, or to see how different scientific fields relate based on
     publication data~\autocite{clusterScience08}.
@@
   \item[network evolution~\autocite{networkEvolution14}] These methods focus on the mechanisms and
-    consequences of the growth of networks, as well as how to keep the results of some data mining
+    consequences of the growth of networks. They also seek ways to keep the results of some data mining
     algorithms up to date and relevant. Monitoring the changes in the interactions of proteins

### 2017-12-16 intro/main.tex
pass over the intro

@@
 \item In computer vision, we can represent an image as a graph. Each pixel is a node, and those
-nodes are connected to four neighbors, namely the adjacent pixel from the top, bottom, left and
+nodes are connected to four neighbors, namely the adjacent pixels from the top, bottom, left and
 right sides. The relationship in this graph is therefore adjacency in an image.
@@
 Not only are there graphs in many domains, but the progress of technology in the last few decades
-has made it easier to collect many graphs in every single domain, sometimes with up to billion of
-nodes and hundred of billions of edges. The availability of such large amount of structured data has
+has made it easier to collect many graphs in every single domain, sometimes with up to billions of
+nodes and hundreds of billions of edges. The availability of such large amounts of structured data has
 prompted the development of automated methods to extract insights from them. For instance, it is
@@
 graph, one representing a single type of relation between nodes.\footnote{Note that by \enquote{type
-of relation}, we do not refer to some graph-theoretical characteristic of an edge such as being directed or weighted.
-Indeed, all the graphs we consider have homogeneous edges with that respect. Instead we mean domain
-specific semantic, as we will make clear in examples.} However, in many situations, there
+of relation}, we do not refer to some graph-theoretical characteristic of an edge, such as being directed or weighted.
+Indeed, all the graphs we consider have homogeneous edges with that respect. Instead, we mean \emph{domain
+specific semantic}, as we will make clear in examples.} However, in many situations, there
 are two dimensions along which graphs exhibit more complexity. First, nodes have more than one type
@@
 three examples. In the \wik{} network, a vote can support the promotion or oppose it. This
-additional information enrich our understanding of the relationships between the nodes in the graph.
+additional information enrich our understanding of the relationships among the nodes in the graph.
 It also points out that two editors can be connected because they share a common topic of interest,
@@
 (like Star Wars), because they have the same director but different genre, because they receive the
-same prestigious awards in different years, and so on. The second type of relation (\enquote{same
-director} and \enquote{different genre}) is actually an example of a mixed relation, for it mixes
-similarity and dissimilarity.
+same prestigious award in different years, and so on. The second type of relation (\enquote{same
+director} and \enquote{different genre}) is actually an example of a mixed relation, for it combines
+similarity and dissimilarity over several features of the nodes.
 
@@
 Informally, given an input graph and possibly some extra information, we want to predict the type of
-every edge. This problem can take several forms depending of what information is available as input.
+every edge. This problem can take several forms depending on what information is available as input.
 First, the graph may be directed or undirected. Second, there might be two edge types (in the case of
@@
 \emph{attributes} of the nodes. Once again, we return to our three examples to illustrate what
-those attributes can be. In the \wik{} graph, we could have for each user data about age, experience and
+those attributes can be. In the \wik{} graph, we could have for each user data about her age, experience and
 area of expertise. In an image, each pixel is associated with a color, along with higher order
@@
 popularity, category and availability. In the absence of label, our intuition is that these
-attributes can inform us about the type of edge between nodes.
+attributes can inform us about the types of edge among nodes.
 

### 2017-12-16 intro/outline.tex
pass over the intro

@@
     design a method that is scalable, principled and accurate. For that, we start by introducing a
-    generative model for signs, and derive approximations of the optimal Bayes predictor or the
+    generative model for signs, and derive approximations of the optimal Bayes predictor and of the
     maximum likelihood predictor. We confirm the theoretical soundness of these approaches by

### 2017-12-16 intro/predicting.tex
pass over the intro

@@
     assuming the answers do not contain any noise. At the scale of Facebook, this would require
-    asking hundred of questions to each user on average, which is quite time consuming. Furthermore,
+    asking hundred of questions to each user on average, which is quite time-consuming. Furthermore,
     it would also cost marketing resource to convince users this is beneficial for them, and
@@
   \item The same \esp{} problem, where the input graph is undirected. In this case, we consider an
-    active variant where instead of being given a random training set of labels, we have a budget of
+    active variant where, instead of being given a random training set of labels, we have a budget of
     queries we can use to request arbitrary labels.
@@
 
-We now review existing approaches addressing these problems, in order to highlight our contribution
+We now briefly review existing approaches addressing these problems, in order to highlight our contribution
 in the next section.
@@
 signed adjacency matrix through low rank
-factorization~\autocites{LowRankCompletion14}{OnlineCompletion17}. These methods are also accurate
+factorization~\autocites{LowRankCompletion14}{OnlineCompletion17}. These methods are also accurate,
 and bypass feature engineering, but by nature, they require careful algorithms to scale with larger
@@
 edges~\autocite{Bansal2002}. \Textcite{Cesa-Bianchi2012b} assume that the signs are originally
-consistent with an underlying, hidden $2$-clustering, but we only observe the signs after they have
+consistent with an underlying, hidden $2$-clustering, but that we only observe the signs after they have
 been flipped \uar{}. In this case, they show that the optimal number of disagreements is a lower
-bound of the number of mistakes of any active \esp{} algorithm. One natural approach would then be
+bound of the number of mistakes made by any active \esp{} algorithm. One natural approach would then be
 to solve \pcc{} based on the observed signs, and predict the remaining signs consistently with the
-inferred clusters. At first it seems hopeless, as \pcc{} is difficult to approximate on general
+inferred clusters. At first, it seems hopeless, as \pcc{} is difficult to approximate on general
 graphs~\autocite{Charikar2003}, even when there are only two clusters~\autocite{Giotis2006}.
@@
 $(u,v)$ is the product of the signs along the path from $u$ to $v$. Therefore, the shorter the path
-and the less influence of the random perturbations. At the same times, those paths must span the
+and the less influence of the random perturbations. At the same time, those paths must span the
 whole graph. This is the topic of an active research area~\autocites{Abraham2012}{Spanner17}.
@@
 attributes. Another method uses topological features~\autocite{ahmed2017roles} but is rather complex
-and thus not very interpretable. If we further assume that each edge type is associated with an
+and thus not very interpretable. If we further assume that each edge type is associated with a
 Euclidean space, and that the position of nodes in one space are not correlated with their position

### 2017-12-16 intro/semantics.tex
pass over the intro

@@
 variation of a \enquote{simple} graph, where the adjacency matrix would take values in $\{-1, 0, 1\}$ or
-$\{0, 1, \ldots, k\}$ instead of $\{0, 1\}$. Rather, we reviews many applications in various domains
+$\{0, 1, \ldots, k\}$ instead of $\{0, 1\}$. Rather, we review many applications in various domains
 where complex networks are the right model to represent rich networked data. First, we showcase many
@@
 subset of the attributes). As examples of the latter, think of dating websites ---where most users
-are linked with users of the opposite gender~\autocites{homophilyMyspace09}{Tinder16}; diffusing
+are linked with users of the opposite gender~\autocites{homophilyMyspace09}{TinderDesc16}; diffusing
 innovations ---where meeting people with different backgrounds and point of views is crucial to
@@
 connecting people from different sides of the political spectrum helps to avoid echo chambers and
-fuel a democratic debate~\autocite{balancedNews17}.
+instead fuel a democratic debate~\autocite{balancedNews17}.
 
@@
 
-Besides signed graphs, in this thesis we also consider at multilayer
+Besides signed graphs, in this thesis we also consider multilayer
 graphs~\autocites{Kivela2014}{multiSurvey14}. Those are graphs with $k$ edge types, and the name
@@
 graphs, the two types of edge have reverse semantic, whereas in a 2-layers graph, they simply
-denotes two possible interactions, for instance advisor-advisee or regular coauthors relation in a
+denote two possible interactions, for instance advisor-advisee or regular coauthors relation in a
 citation network~\autocite{Advisor10}. In general though, we focus on cases where $k$ is larger than
@@
 player. These six types of interactions are either positive or negative, but their nuances cannot
-simply by explained by global similarity and dissimilarity. Another examples are the users of the
-photo sharing website Flickr.  They can interact in eleven ways, either directly or through
+simply by explained by global similarity and dissimilarity. Another example is photo sharing website
+Flickr and its users. They can interact in eleven ways, either directly or through
 comments, shared tags, groups membership and so on~\autocite{RecoFlickrMulti11}. Again, while being
-part of a common group denote shared interests between two users, overall they must also differ in
+part of a common group denotes shared interests between two users, overall they must also differ in
 some other attributes (for instance location) in order to bring diversity to this group.
@@
 Network are connected by 96 kinds of commodities they can exchange~\autocite{worldTradeNetwork10}.
-Finally, \textcite{KantPeace15} studies international relations through the lens of Kant three
+Finally, \textcite{KantPeace15} studies international relations through the lens of Kant's three
 folded program for peace,  based on democracy, economic dependence and supra national governance.
@@
 clique in the first layer, countries are connected with weight proportional to the amount of yearly
-trade in the second layer and with weight proportional to the number of international organisations
+trade in the second layer and with weight proportional to the number of international organizations
 they belong to together in the third layer. While the essence of commerce is to exploit differential
-between the partners involved (\ie{} heterophily), trade (and in the last example participating in
-common institution) intensity also involve geographical, historical and cultural ties (\ie{}
+between the partners involved (\ie{} heterophily), trade intensity and participating in
+common institutions also involve geographical, historical and cultural ties (\ie{}
 homophily).
@@
 coast\footnote{\url{http://app.mappr.io/play/chile-marine-intertidal-network}} and divide their
-interactions in three categories: trophic (that is one specie eats another one), positive
+interactions in three categories: trophic (\ie{} one specie eating another one), positive
 non-trophic (\eg{} refuge providing) and negative non-trophic (\eg{} competition for shared
@@
 edges can be labeled in various ways: some correspond to actual physical links while others are
-functional ---\ie{} neurons responding in the same way to external stimuli, some are present in
-healthy subject and others in patient cohort, some are acquired through MRI and others by EEG.
+functional (\ie{} neurons responding in the same way to external stimuli), some are present in
+healthy subjects and others in treated patients, some are acquired through MRI and others by EEG.

### 2017-12-16 intro/use_of_sign_graphs.tex
pass over the intro

@@
       further processing. As we mentioned earlier in one example, building a signed graph can help,
-      although it might be costly to do it at the pixel level. For instance, to segment cell in
+      although it might be costly to do it at the pixel level. For instance, to segment cells in
       microscopy imagery, \textcite{CellSeg14} first use generic image features to classify pixel in
       belonging to region boundaries or not. Then, they extract small scale regions called
-      superpixels. After building the adjacency graph of these superpixels, they assigned edge
+      superpixels. After building the adjacency graph of these superpixels, they assign edge
       weights by averaging the boundary probabilities of the pixels separating adjacent
@@
       upper bound of the optimal solution by solving a SDP
-      relaxation of the problem. They then compare various heuristics and show that best performing
+      relaxation of the problem. They then compare various heuristics and show that the best performing
       ones are within few percents of the optimum, provided they are followed by a local search
@@
       Signed graphs are also abundant in biology.
-      A typical input is a similarity matrix between genes expression level in various experimental conditions
-      and the goal is to cluster those genes into groups which react similarly. \Textcite[Section
+      A typical input is a similarity matrix between genes expression level in various experimental conditions,
+      and the goal is to cluster those genes into groups that react similarly. \Textcite[Section
       4]{Ben-Dor99} gives three examples: 112 genes involved in the rat central nervous system, 1246
-      genes of the roundworm \emph{C. elegans} and 2000 human genes obtained from 40 tumor and 22
+      genes of the roundworm \emph{C. elegans} and 2000 human genes obtained from 40 tumorous and 22
       normal colon samples. \Textcite{Mason2009} analyze a signed co-expression network of genes
@@
       single nucleotide polymorphisms (SNPs), where a single base differs between the two DNA
-      sequences, leading to different alleles of the corresponding gene. An haplotype is the list of
+      sequences, leading to different alleles of the corresponding gene. A haplotype is the list of
       all alleles at a contiguous region of a single chromosome, and this information is
@@
       interactions~\autocites{harary1953}{HeiderBook58}. Here we present
-      some recent references. For instance, one can extract all the votes of the members of a
+      some recent references within this line of research. For instance, one can extract all the
+      votes of the members of a
       political parliament and form a graph whose nodes are politicians and edge weights quantify how
@@
 	    The main challenge
-	    thus lies in devising an appropriate similarity measure, given that object can have very
+	    thus lies in devising an appropriate similarity measure, given that objects can have very
 	    different features from one database to another. \Textcite{LargeScaleDeDup09} propose a

### 2017-12-17 troll/algs_batch.tex
pass over troll

@@
 = \big((u_1,v_1),y_{u_1,v_1}), ((u_2,v_2),y_{u_2,v_2}), \ldots, ((u_m,v_m),y_{u_m,v_m}\big)$, that
-has been drawn from $E\times Y$ \uar{} with replacement, with
-$m = |E_0|$. Then a reasonable approach to approximate $y^*(u,v)$ would be to resort to a maximum
- We want to build a predictive model for the labels of the remaining
+has been drawn from $E\times Y$ \uar{} without replacement, with
+$m = |E_0|$. We want to build a predictive model for the labels of the remaining
 edges. We present two algorithms to do so, which both compute estimates of the all parameters $p$
@@
 We are now ready to describe our algorithm, which follows naturally from the formula derived above.
-It takes as input a training set \trainset{} drawn at random with replacement and outputs a sign
+It takes as input a training set \trainset{} drawn at random without replacement and outputs a sign
 prediction for all the edges in $E\setminus \trainset$. We call it \usrule{}, which stands for Bayes
@@
 
-Remember we suppose that the training set \trainset{} has been drawn \uar{} with replacement, with
+Remember we suppose that the training set \trainset{} has been drawn \uar{} without replacement, with
 $m = |E_0|$. Then a reasonable approach to approximate $y^*(u,v)$ would be to resort to a maximum
@@
 positive, meaning this assumption of uniformity over $[0,1]^2$ is unlikely to fully holds, for
-otherwise the signs would more balanced}
+otherwise the signs would more balanced.}
 we show in the supplementary material \vpageref{ssec:troll_mle_deriv} that for $\ell \in \{1,
@@
 has small in-degree). Similarly, if $u$ has a high in-degree, then the corresponding $q_u$ should be
-close to \shalf{} implying that on any edge $(v,u)$ arriving at $u$ the Bayes optimal predictor
+close to \shalf{}, implying that on any edge $(v,u)$ arriving at $u$, the Bayes optimal predictor
 $y^*(v,u)$ will mainly depend on $p_v$ (assuming $v$ has small out-degree). Put differently, a node

### 2017-12-17 troll/experiments.tex
finish troll layout
pass over troll

@@
 used as benchmarks for this
-task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
+task~\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}.
 In \adv{}, a trust-based social network for open source developers, a user $u$ can certify another
@@
   \small
-  \caption{Dataset properties. The 5\thup{} column gives the fraction of positive labels. The next two
+  \caption[\dssn{} dataset properties]{Dataset properties. The 5\thup{} column gives the fraction of positive labels. The next two
   columns provide two different measures of label regularity, while the last two columns give the
@@
 networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases, such reciprocal edges do not disagree,
-\ie{} they have the same sign. In practice, we can use this fact to improve our accuracy at no
+\ie{} they have the same sign. In practice, we can use this fact to improve our prediction at no
 additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
 \evu{} is part of the training set, we set $\yhat_{u,v} = y_{v,u}$. For clarity, when comparing
-our methods with existing approaches, we not use that trick. But afterwards, we show in
+our methods with existing approaches, we dot not use that heuristic. But afterwards, we show in
 \autoref{tab:twin} when it can be beneficial.
@@
 obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}\footnote{More precisely,
-these figures are normalized histogram, meaning the counts are scale down so that the bars total
+these figures are normalized histograms, meaning the counts are scaled down so that the bars total
 area sums up to one.}, giving $71.9\%$ of positive edges.
 
-% TODO: do a proper subfigure instead of minipage?
-\vspace{\baselineskip}
-\noindent\begin{minipage}{187mm} 
-  \begin{minipage}{46mm}
+\begin{figure*}[hbtp]
+  \centering
+  \begin{subfigure}[b]{0.401\textwidth} 
     \centering
     \includegraphics[width=.95\linewidth]{raw/synth_wik_pdense.pdf}
-    \captionof{subfigure}{Density of $p$.}
-    % \label{fig:figure1}
-  \end{minipage}%
-  \begin{minipage}{46mm}
+    \caption{Density of $p$.}
+  \end{subfigure}\quad
+  \begin{subfigure}[b]{0.401\textwidth} 
     \centering
     \includegraphics[width=.95\linewidth]{raw/synth_wik_qdense.pdf}
-    \captionof{subfigure}{Density of $q$.} 
-    % \label{fig:figure2} 
-  \end{minipage} 
-  \begin{minipage}{46mm}
+    \caption{Density of $q$.}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.401\textwidth} 
     \centering
     \includegraphics[width=.95\linewidth]{raw/synth_wik_etadense.pdf}
-    \captionof{subfigure}{Density of $\eta$.} 
-    % \label{fig:figure2} 
-  \end{minipage} 
-  \begin{minipage}{46mm}
+    \caption{Density of $\eta$.}
+  \end{subfigure}\quad
+  \begin{subfigure}[b]{0.401\textwidth} 
     \centering
     \includegraphics[width=.95\linewidth]{raw/synth_wik_signdense.pdf}
-    \captionof{subfigure}{Density of $\eta$ for both signs.} 
-    % \label{fig:figure2} 
-  \end{minipage} 
-  \captionof{figure}{Synthetic distributions on \wik{}.} 
+    \caption{Density of $\eta$ for both signs.}
+  \end{subfigure}
+  \caption{Synthetic distributions on \wik{}.} 
   \label{fig:troll_synth_distrib} 
-\end{minipage}
-\vspace{\baselineskip}
+\end{figure*}
 
@@
 
+\input{synth_res}
 The result of this comparison on the six datasets is showed in \autoref{tab:troll_synth_global}.
@@
 increasing number of samples for a given edge.
-
-\input{synth_res}
 
@@
 
-  \item The \usrule{} algorithm described in \autoref{ss:passive}.
+  \item The \usrule{} algorithm described in \autoref{ss:bayes_approx}.
 
@@
   \item  The solution obtained by directly solving the unregularized problem \eqref{e:quadratic}
-    through a fast constrained minimization algorithm (referred to as \qoptim{}). Again, the actual
+    through a constrained minimization algorithm (referred to as \qoptim{}). Again, the actual
     binarizing threshold was set by cross-validation on the training set.\footnote{We have also
@@
 Because scalability is a major concern when training on sizeable datasets, we also give an idea of
-relative training times (in milliseconds) by reporting separately in \autoref{tab:all_time} the time it took to train a single run of
+relative training times by reporting separately in \autoref{tab:all_time} the time (in milliseconds) it took to train a single run of
 each algorithm on a training set of size\footnote{Comparison of training time performances is fair
 since all algorithms have been carefully implemented using the same stack of Python libraries, and
-run on the same machine (16 Xeon cores and 192Gb Ram).} 15\% of $|E|$, and then predict on the test
+run on the same machine (16 Xeon cores and 192GB Ram).} 15\% of $|E|$, and then predict on the test
 set. Some conclusions emerge from those experiments:
@@
 sampling strategies cover roughly 90\% of the nodes in $V_{\mathrm{out}}$ and in
-$V_{\mathrm{out}}$. The difference in MCC is thus explained by the last three columns. It shows, for
+$V_{\mathrm{in}}$. The difference in MCC is thus explained by the last three columns. It shows, for
 testing edges \euv{}, the breakdown between three situations, from the least to the most

### 2017-12-17 troll/introduction.tex
tiny fixes and ref all over the places
pass over troll

@@
 As we saw in the introduction, many situations in various fields can be modeled by signed graphs.
-One of the most natural usage of signed graphs, though, is the study of social networks. That is
+Nonetheless, one of the most natural usage of signed graphs is the study of social networks. It is
 indeed what first motivated their inception. Quite naturally, we thus begin our exploration of the
 role of edges in complex networks by a binary classification problem in signed graphs. Namely, we
-want to predict whether an observed interaction between two nodes is positive or negative and we
+want to predict whether an observed interaction between two nodes is positive or negative, and we
 call this problem \esp{}.\footnote{This chapter is closely based on a paper written with several
@@
 
-Another distinction is that in most cases, only the owner of the network can observe its graph.
+Another distinction is that in most cases, only the owner of the network can observe the detail of
+all those interactions.
 Think for instance of \textsc{YouTube}, where users can like or dislike the video of another user
@@
 
-With this definition of social networks, we now look at how the learning problem of \esp{} can be
+With this definition of social networks at hand, we now look at how the learning problem of \esp{} can be
 used in real world applications. More precisely, we present four possible applications, and order
-them by how severely negative interactions impact the well-being of users of the networks.
+them by how severely negative interactions impact the well-being of users of the network.
 First, it could help improve the quality of link recommendation in solely positive networks. Such
@@
 small proportion of existing links as truly positive or actually negative, recommended links could
-be discarded if they are predicted to be negative.
+later be discarded if they are predicted to be negative.
 %
@@
 increasing use of Machine Learning~\autocite{MLinHR16}. Among other tools, assessing personal
-relationships between employees is crucial to maintain a productive work environment. Whereas this
+relationships between employees is crucial to maintain a productive work
+environment~\autocite{friendshipTeam02}. Whereas this
 can be done by invasive wearable sensors~\autocite{WearableBehavior09}, our methods could predict
@@
 positive links (trustworthiness) at each node. We first define these notions and more notations in
-\autoref{s:prel}, along with others tools and precise problem statements. Then we study such
+\autoref{s:prel}, along with others tools and precise problem statements.
+We also introduce suitable graph transformations defining reductions from \esp{} to
+node sign prediction problems.
+Then we study such
 troll-trust heuristics by defining in \autoref{s:gen} a probabilistic
@@
 
-We then introduce suitable graph transformations defining reductions from \esp{} to
-node sign prediction problems. This opens up the possibility of using the arsenal of known
+Our graph transformations opens up the possibility of using the arsenal of known
 algorithmic techniques developed for node classification. In particular, we introduce in

### 2017-12-17 troll/notation_and_prel.tex
pass over troll

@@
 
-In this chapter and the next one, we let $G=(V,E)$ be a \emph{directed} graph, whose edges $(u,v)
-\in E$ carry a
+In this chapter\footnote{And in the next chapter when applicable.}, we let $G=(V,E)$ be a
+\emph{directed} graph, whose edges $(u,v) \in E$ carry a
 binary label $\yuv\in \spin$. The edge labeling will sometimes be collectively denoted by the
@@
 node $u \in V$, with $\din(u) = \big|\Nin(u)\big|$ and $\dout(u) = \big|\Nout(u)\big|$ being the
-in-degree and the out-degree of $u$. Moreover, $\din^+(u)$ is the number of edges $(k,u)\in\Nin(u)$
-such that $y_{k,u} = +1$. We define $\din^-(u)$, $\dout^+(u)$, and $\dout^-(u)$ similarly, so that,
+in-degree and the out-degree of $u$. Moreover, $\din^+(u)$ is the number of edges $(w,u)\in\Nin(u)$
+such that $y_{w,u} = +1$. We define $\din^-(u)$, $\dout^+(u)$, and $\dout^-(u)$ similarly, so that,
 for instance, $\dout^-(u)/\dout(u)$ is the fraction of outgoing edges from node $u$ whose label in
@@
   \end{subfigure}
-  \caption{\label{f:etnr}
+  \caption[Illustration of the graph transformations turning \esp{} into node
+  classification]{\label{f:etnr}
     \textbf{(a)} A directed edge-labeled graph $G$. \textbf{(b)} Its corresponding graph $G'$
@@
 two label regularity measures $\Psi_G(Y)$ and $\Psi^2_{G''}(Y)$ when moving from edges to nodes.
-That is, if the \esp{} prediction problem is
+That is, if the \esp{} problem is
 easy for a given $G(Y) = (V,E(Y))$, then the corresponding node sign prediction problems on $G'(Y)
 = (V'(Y),E')$ and $G''(Y) = (V''(Y),E)$ are also easy, and vice versa.
-Let us first introduce the notion of cutsize of an undirected node-labeled graph $G'(Y)$, which is
+
 \begin{wrapfigure}{o}[10mm]{45mm}
@@
 \end{wrapfigure}
+Let us first introduce the notion of cutsize of an undirected node-labeled graph $G'(Y)$, which is
 the number of edges in $G'$ connecting nodes that have mismatching labels.
@@
   \autoref{fig:troll_reduction_g} reported in \autoref{tab:troll_psi}, following this construction
-  would result in a label version of $G'$ showed on \autoref{fig:troll_psi}.
+  would result in a labeled version of $G'$ shown in \autoref{fig:troll_psi}.
   We see that by labeling circle nodes by the majority label of the square nodes there are connected

### 2017-12-17 troll/related.tex
pass over troll

@@
 probability of transitioning to any other type once the unobserved signs are revealed. Then each
-edge \euv is associated with a feature vector consisting of the outer
-product of $V_u$ with $V_v$ and also include additional features such as triads count and degree
+edge \euv{} is associated with a feature vector consisting of the outer
+product of $V_u$ with $V_v$ and also including additional features such as triads count and degree
 information before training a Logistic Regression model. We refer to this method as
@@
 (or $w \rightarrow u$) and $v \rightarrow w$ (or $w \rightarrow v$) for any $w \in \NNout(u) \cap
-\NNin(v)$. Taking signs and direction into account, there are 16 possible triads but according to
+\NNin(v)$. Taking signs and directions into account, there are 16 possible triads but according to
 the status theory some must be more represented than others. The \emph{\comptriads{}}
@@
 		weighted graphs. The bias is the tendency of a node $u$ to trust/mistrust others, that is the
-		difference between its opinion of a neighbor $v$ and what $v$ truly deserve according to the
+		difference between its opinion of a neighbor $v$ and what $v$ truly deserves according to the
 		network: \[bias(u) = \frac{1}{\dout(u)}\sum_{v\in\NNout(u)}(\wuv - deserve(v)).\] The deserve of a
@@
 	\item	TrollTrust~\autocite{wu2016troll}, which builds upon \autocite{shahriari2014ranking} but
-		defined the ranking $\sigma(u)$ as the trustworthiness of $u$. It is follows from a recursive
+		defines the ranking $\sigma(u)$ as the trustworthiness of $u$. It follows from a recursive
 		definition of trollness based on the opinion of one node's neighbors weighted by their own
@@
 \epi{} in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
-analysing the textual content of their post, \autocite{attributedSNE17} improve
+analysing the textual content of their post, \autocite{attributedSNE17} improves
 the embedding approach of \textcite{neuralSigned17} by considering the words
@@
 %
-Finally, \textcite{edgeWeights16} explicitly consider the weight prediction problem in signed graph.
+Finally, \textcite{edgeWeights16} explicitly consider the weight prediction problem in signed graphs.
 They use a procedure similar to the ranking methods discussed above. Namely, they defined two scores
@@
 value in the factor $2$ in the expression of fairness allows for better convergence property.
-Finally, the weight of a test edge \euv is predicted as $f(u)\times g(v)$.
+Finally, the weight of a test edge \euv{} is predicted as $f(u)\times g(v)$.
 
@@
 While we presented methods that operate in the batch setting, and will present an online algorithm
-as well in \autoref{s:algonline}, other works have attacked the \esp{} problem from an active
+as well in \autoref{s:algonline}, other works have addressed the \esp{} problem from an active
 learning point of view~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}.  Recall that in the active

### 2017-12-18 cc/bias_directed.tex
typos and stability in cc
pass over CC

@@
-Having two parameters per node, one for outgoing edges and another one for incoming edges clearly
+Having two parameters per node, one for outgoing edges and another one for incoming edges, clearly
 targets directed graphs. Many online interactions are inherently directed, for instance friendship,
@@
 freedom from the model. To illustrate this point, we conduct the following experiment. We use our
-previous \dssn{} datasets and remove the edge direction. As mentioned earlier, for few pairs of
-nodes there are reciprocal edges of different signs, in which case we pick a sign arbitrarily. Given
+previous \dssn{} datasets and remove the edge direction. As mentioned earlier, for a few pair of
+nodes, there are reciprocal edges of different signs, in which case we pick a sign arbitrarily. Given
 those undirected graphs, we orient them using the approach described above and compare our method
@@
 the algorithm gets to observe more signs and therefore perform better. This small experiment thereby
-shows that although the performance remains decent, not having direction information hurt our bias.
+shows that although the performance remains decent, not having direction information hurts our bias.
 

### 2017-12-18 cc/bias_main.tex
move social balance to first section
typos and stability in cc
pass over CC

@@
 chosen to be positive with probability $\frac{p_u+q_v}{2}$. This suggests that nodes in the
-graph have a form of agency, and this imposes that edges are directed. We show experimentally that
-failing to meet these two assumptions harms the performance of our previous method. This is not
+graph have a form of agency, and further imposes that edges are directed. We show experimentally that
+failing to meet these two assumptions harms the performance of our previous methods. This is not
 surprising, for our bias is no more justified in that case. Therefore, we suggest a different bias,
@@
 \input{bias_directed}
+
+\subsection{Social balance as a learning bias}
+\label{sub:bias_balance}
+
+\input{bias_balance}

### 2017-12-18 cc/bias_model.tex
pass over CC

@@
     humans and mouses~\autocite{Cancer07}.
-  \item The \reg{} database\autocite{RegulonDB16}, summarizing the connection between transcription
+  \item The \reg{} database~\autocite{RegulonDB16}, summarizing the connection between transcription
     factors and their targets in the Escherichia coli K-12 bacteria. We download four
@@
 
-Compared with the \dssn{} of the previous chapter, those biological are smaller, have lower average
+Compared with the \dssn{} of the previous chapter, those biological networks are smaller, have lower average
 degree and little to no reciprocal edges. Trying to solve the \esp{} problem using the same

### 2017-12-18 cc/cc_non_worst_case.tex
pass over CC

@@
 We now look at two settings where our bias is likely to be present.  In the first one, we assume the
-input graph is randomly perturbed version of a consistent clustering. In the second one, we do not
+input signs are randomly perturbed of an initial assignment derived from a consistent clustering. In the second one, we do not
 assume anything about such a consistent clustering but expect instead that the optimal solution is
@@
 recovering with high probability the planted partition of an unweighted graph in $O(n^2(\log n)^c)$,
-where $c$ depends on the size of the smallest cluster and the noise probability.
+where $c$ depends on the size of the smallest cluster and on the noise probability.
 
@@
 nodes misclustered w.r.t the planted partition as a function on the probability distribution
-parameters. Indeed, as pointed out by \textcite{plantedAilon09}, independent and uniform noise is
+parameters. Indeed, as pointed out by \textcite{plantedAilon09}, uniform and independent noise is
 not a good model of real situations, where the input of \pcc{}, \ie{} the similarity between nodes,
@@
 $c$ of the optimal cost $OPT_{\Phi}$ is $\epsilon$-close to the optimal clustering (as measured by
-the fraction of points in which the two clustering disagree). When the objective is \mind{} on
+the fraction of points on which the two clusterings disagree). When the objective is \mind{} on
 complete graphs, \textcite{StableCC09} show that for $(1+\alpha, \epsilon)$-instances, the

### 2017-12-18 cc/cc_problem_and_esp.tex
pass over CC

@@
 fixed, as it consists of all possible partitions of the nodeset. We are trying to find the partition
-that best approximates the observed signs, but each weakly unbalanced cycle is gonna cost us at
+that best approximates the observed signs, but each weakly unbalanced cycle will cost us at
 least one unavoidable mistake. Despite this fundamental difference in their nature, \pcc{} is
@@
 n}$~\autocite[Theorem 6]{Cesa-Bianchi2012b}. These theoretical considerations naturally lead to
-practical a batch algorithm, which first finds a clustering minimizing as much as possible the number
+a practical batch algorithm, which first finds a clustering minimizing as much as possible the number
 of disagreements on the training set and then predicts signs according to this clustering. Our
@@
 2D~\autocites{Kim2011}{Bagon2011}{CellSeg14} or 3D~\autocites{VolumeSegmentation12}{Beier2015}, to
-simplify 3D shapes~\autocite{Shape3D17} and to track target across sequential video
+simplify 3D shapes~\autocite{Shape3D17} and to track targets across sequential video
 frames~\autocite{multiTracking15}. When the nodes are words, \pcc{} is employed to identify

### 2017-12-18 cc/cc_state_of_art.tex
fix cc chapter layout
pass over CC

@@
 we next present hardness results about \pcc{}, namely that it is \NPh{} to approximate with some
-constant factor greater than one, and then describe some proposed approximations. Whereas this
-sounds like bad news for us, we show that heuristics can solve large instances with satisfactory
+constant factor greater than one, and then describe some proposed approximations. Although this
+appears rather negative at first, we show that heuristics can solve large instances with satisfactory
 results. We conclude by further highlighting the learning aspect of \pcc{} through approaches in
@@
 of the problem. However, we can imagine offline signed social networks with only few nodes, in which
-case it is reasonable to expect finding the optimal clustering. Furthermore, this can also be useful
+case it is reasonable to expect to find the optimal clustering. Furthermore, this can also be useful
 to evaluate in practice the quality of heuristic methods, albeit in non asymptotic settings. The
-methods used to solve \pcc{} exactly depends on the type of graph considered. More precisely, the
+methods used to solve \pcc{} exactly depend on the type of graph considered. More precisely, the
 fixed parameter algorithms we describe now only handle complete graphs, while the next set of
@@
 instance $G$ of \pcc{} on complete graphs in polynomial time. To see why, let $G=(V,(E^+,E^-))$
-be a complete signed graph with nodeset $V$, $E^+$ be the set $E$ of edges of $H$ and $E^-$ be all
-the edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
+be a complete signed graph where $V$ is the nodeset of $h$, $E^+$ is the set $E$ of edges of $H$ and
+$E^-$ is the set of all
+edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
 disagreements within clusters are edges added to $H$ and the positive disagreements between clusters
@@
 
-For that we follow~\autocite{Demaine2006} but omit the full proofs to avoid redundancy. Recall that
+For that we follow~\autocite{Demaine2006} but omit the full proofs to avoid redundancy. In \mmc{},
 given a weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and targets
-$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, in \mmc{} we want to find the lightest set of edges $T\in E_H$
+$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, we want to find the lightest set of edges $T\in E_H$
 whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
@@
 
-\begin{table}[b]
+\begin{table}[tb]
+\begin{adjustwidth}{-2cm}{}
    \centering
@@
    \caption{Hardness results of \pcc{}} \label{tab:cc_cpx}
-   \begin{tabulary}{187mm}{lCCCC}
+   \begin{tabulary}{179mm}{lCCCC}
       \toprule
@@
    \end{tabulary}
+\end{adjustwidth}
 \end{table}
@@
 We next present the polynomial transformation from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
-$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v)$ of
+$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v) \in E^-$ of
 weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
@@
 If a certain conjecture in computational complexity is true\footnote{Namely, the Unique Game
-Conjecture of~\textcite{UGC02}.}, then for every constant $c>0$ it is \NPh{} to approximate \mmc{}
+Conjecture of~\textcite{UGC02}.}, then for every constant $c>0$, it is \NPh{} to approximate \mmc{}
 within a factor $c$~\autocite{MulticutHardness06}. On general graphs, the best known approximation factor in $O(\log
@@
 After solving the SDP, a clustering can be obtained by a general rounding technique $H_t$: pick $t$
-random hyperplanes and divides the nodes in $2^t$ clusters. \Textcite[Theorem 3]{Charikar2003} prove
+random hyperplanes and divide the nodes in $2^t$ clusters. \Textcite[Theorem 3]{Charikar2003} prove
 that taking the best results of $H_2$ and $H_3$ gives in a $0.7664$ approximation on general graph.
@@
 Combining \mind{} and \maxa{}, \textcite[Section 4]{Charikar2004} give a $\Omega(\frac{1}{\log n})$
-approximation of the \textsc{MaxCorr} problem, which is maximizing \eqref{eq:maxa} - \eqref{eq:mind}
+approximation of the \textsc{MaxCorr} problem, which is maximizing \eqref{eq:maxa} $-$ \eqref{eq:mind}
 and can be formulated as a quadratic programming problem solved in polynomial time.
@@
 \begin{table}
-   \caption{Best current results on \pcc{} problems. The \enquote{easiest} setting is \maxa{} on
+  \caption[Best current results on \pcc{} problems]{Best current results on \pcc{} problems. The \enquote{easiest} setting is \maxa{} on
       complete graphs, for it admits PTASs. All others cases are \APXh{}. However, we see that on
@@
 equivalence with \mmc{}.
-Currently there are three main approaches to approximate \mind{}. On a general graph,
-one first solve the canonical LP, and use the \regionGrow{} method to round the fractional solution,
-resulting in an $O(\log n)$ approximation. On complete graphs, the simplest solution is the
+Currently there are three main approaches to approximate \mind{}. First, on a general graph,
+we can solve the canonical LP, and use the \regionGrow{} method to round the fractional solution,
+resulting in an $O(\log n)$ approximation. Second, on complete graphs, the simplest solution is the
 \ccpivot{} algorithm, that grows clusters out of randomly selected pivots, in a way reminiscent of
-the proof of \autoref{thm:weakbal-consistent}. The approximation factor of this general idea can be
+the proof of \autoref{thm:weakbal-consistent}. Third, the approximation factor of this general idea can be
 improved by interpreting the fractional solution of the linear program as probability to be included
@@
 $O(\frac{n}{\epsilon})$ time and queries~\autocite[Theorem 3.3]{Bonchi2013}.
-Roughly stated, the second one relies on an existing low-rank approximation of the adjacency matrix,
+Roughly stated, the second technique relies on an existing low-rank approximation of the adjacency matrix,
 that partition the graph into similar sized classes such that edges between those classes behave as
@@
    % 187mm - 3em = https://tex.stackexchange.com/a/8263
-   \begin{tabulary}{177mm}{lLL}
+   \begin{tabulary}{150mm}{lLL}
       \toprule
@@
 presented in~\autocites{Levorato2015}{Levorato2017}. Each iteration of this algorithm starts by
-greedily build a clustering in fashion similar to \textsc{Vote}, albeit with more randomness in the
+greedily building a clustering in a fashion similar to \textsc{Vote}, albeit with more randomness in the
 node ordering. This clustering is locally improved by moving blocks of $r\in\{1,2\}$ nodes from one
@@
 centroid, and query the edges connecting the centroid as ordered by an entropy-based criterion.
-In the case the oracle answering the queries is consistent, then there exists an information
+In the case the oracle answering the queries is consistent, there exists an information
 theoretic bound on the number of queries needed to recover the clusters, and an algorithm matching
@@
 $v$. Whenever
-$p < \shalf$, they show that $O(n^{1+\frac{1}{\log\log n}}\log n)$ queries are enough are to recover
+$p < \shalf$, they show that $O(n^{1+\frac{1}{\log\log n}}\log n)$ queries are enough to recover
 the planted partition in polynomial time and with high probability. Those queries are actually
@@
 hand, it can be solved with approximation guarantees, at least in theory. Indeed, the best methods
-requires solving linear programs with $m$ variables and up to $O(n^3)$ constraints, which results in
+require solving linear programs with $m$ variables and up to $O(n^3)$ constraints, which results in
 a high complexity of $O(m^{4.5})$~\autocite[Section 7.2]{LPBook07}. Fortunately, we can obtain
@@
 almost exactly. We next present settings where instances are indeed close to our bias.
-Before that, we remark that whereas \ccpivot{} is often used on general graphs in practice by
-assuming the missing edges are negative, doing so does not preserve its approximation guarantee.
-\begin{marginfigure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/raw/pivotstar.pdf}
-  \caption{A positive star with few negative edges}
-  \label{fig:cc_star_pivot}
-\end{marginfigure}
+Before that, we remark that while \ccpivot{} is often used on general graphs in practice (by
+assuming the missing edges are negative), doing so does not preserve its approximation guarantee.
 
 \begin{aside}
+\begin{wrapfigure}{i}[10mm]{45mm}
+  \centering
+  \includegraphics[width=.9\linewidth]{assets/raw/pivotstar.pdf}
+  \caption{\small A positive star with few negative edges}
+  \label{fig:cc_star_pivot}
+\end{wrapfigure}
    Consider the graph $G$ showed in \autoref{fig:cc_star_pivot}. Assume there are $n$ nodes

### 2017-12-18 cc/cc_variants.tex
pass over CC

@@
 appropriate choice of $k$, it is possible to use this PTAS to solve the general \maxa{} problem.
+Beyond those results on \pcc{}, we further discuss bipartite signed graphs in \autoref{ssec:conc_rr}.
 
@@
 color. Namely, a disagreement is now a negative edge between clusters or a within-cluster edge
-whose color differs from the majority color of that cluster. This is motivated by edge-labeled graph
-in social networks, biology and citation networks and will discuss such applications in
+whose color differs from the majority color of that cluster. This is motivated by edge-labeled graphs
+in social networks, biology and citation networks and we will further discuss such applications in
 \autoref{chap:vector}. As a generalisation of \pcc{}, it is \NPc{} but \textcite{Bonchi2012a}
@@
 as large as $n$. However, \textcite{Anava2015} present constant factor approximations. Namely, they
-show the problem can be reduced to classical \pcc{} by setting all edges incident to a node $u$ to
+show that the problem can be reduced to classical \pcc{} by setting all edges incident to a node $u$ to
 negative if they are not of the majority color of $u$. They then apply the regular \ccpivot{} and
@@
 \cluster(v)) - w_{uv}\right|\] where $H$ is a similarity function between two sets of labels, chosen
-here to be the Jaccard similarity or a $0/1$ indicator of nonempty intersection. These problems are
-showed \NPc{} and approximated by a local search algorithm, iteratively optimizing the assignment of
-one node while all others are fixed. As one of the demonstration on their theoretical work,
+in their paper to be the Jaccard similarity or a $0/1$ indicator of nonempty intersection. These problems are
+showed to be \NPc{}, and approximated by a local search algorithm, iteratively optimizing the assignment of
+one node while all others are fixed. As one of the demonstration of their theoretical work,
 \textcite{WeightedTheta15} show a faster solution based on a weighted extension of the Lovász's
@@
 In classical \pcc{}, all nodes have an identical role, in the sense that they contribute equally to
-the final objective in terms of (dis)agreements. Here we instead look at approaches where we either
-add a local penalty to each in order to better control their behavior or where we altogether modify
+the final objective in terms of (dis)agreements. Here, we instead look at approaches where we either
+add a local penalty to each in order to better control their behavior, or where we altogether modify
 the objective to focus on (dis)agreements at specific nodes.
@@
 number node and create a cluster from its center with a radius of $\nicefrac{3}{7}$. They also show
-that on general weighted graphs, the LP has a large integrality gap of $\nicefrac{n}{2}$ yet they
+that on general weighted graphs, the LP has a large integrality gap of $\nicefrac{n}{2}$. Yet they
 combine it with a combinatorial approach to reach a $O(\sqrt{n})$ approximation. Finally, they
-consider the complementary problem of maximizing the minimum number of agreements reach at a single
+consider the complementary problem of maximizing the minimum number of agreements at a single
 node, and provide a $\frac{1}{2+\epsilon}$ approximation.
@@
 Laplacian $Q^-$ of the negative subgraph $G^-=(V, E^-)$, where $Q^- = D^- + A^-$. They suggest
-instead to use a geometric mean defined for two positive matrices $A$ and $B$ as $A\#B =
+instead to use a geometric mean, defined for two positive matrices $A$ and $B$ as $A\#B =
 A^{\shalf{}} \left(A^{-\shalf{}} BA^{-\shalf{}} \right)^{\shalf{}} A^{\shalf{}}$. This suggestion is
@@
 then find a threshold  based on the number of disagreements.
-The one node moves local heuristic that we described in the Physics-inspired paragraph
+The one node move local heuristic that we described in the Physics-inspired paragraph
 \vpageref{par:cc_physics} can also be formulated as genetic algorithms that simultaneously try to
@@
 modularity~\autocites{Li2013}{Amelio2013}. \Textcite{Anchuri2012} also consider these two objectives
-by seeing them as eigenvalue problems and devise a iterative splitting procedure.
+by seeing them as eigenvalue problems and devise an iterative splitting procedure.
 The overlapping community detection variant is considered by \textcite{Chen14}, who used a signed

### 2017-12-18 cc/conclusions.tex
pass over CC

@@
 
-	\item The basic analysis \autoref{prop:gtx_correct} shows that on any graph with $n$ nodes, the
+	\item The basic analysis of \autoref{prop:gtx_correct} shows that on any graph with $n$ nodes, the
 		\gtx{} will perform at most $n$ collapses before terminating. However, as we mentioned
@@
 		center of stars, pick the node with the current highest weighted degree. When choosing an edge
-		to connect two center, pick among the edges with the lowest endpoints eccentricity the one with
+		to connect two centers, pick among the edges with the lowest endpoints eccentricity the one with
 		the lowest weight. What is less clear is to which extent such changes would affect the analysis.
@@
 		$1$ and $\alpha$. Furthermore, we say that an algorithm is \emph{robust} if, given a clustering
-		instance $\mathcal{I}$, it yields of the two following outcomes in polynomial time:
+		instance $\mathcal{I}$, it yields one of the two following outcomes in polynomial time:
 		\begin{enumerate}[(i),nosep]
@@
 		bears some superficial similarity with \mmc{}. Recall from \autoref{ssub:cc_harness_approx} that
-		\mmc{} requires us the find the lightest set of edges to cut in order to separate $k$ pairs of
+		\mmc{} requires us to find the lightest set of edges to cut in order to separate $k$ pairs of
 		nodes.  Recall further that \mmc{} is equivalent to \pcc{}. As for \minmc{}, we can cast it as a

### 2017-12-18 cc/gtx_algorithm.tex
pass over CC

@@
   \end{algorithmic}
-  % \rule{\textwidth}{.3pt}
+  \rule{\textwidth}{.3pt}
 \end{center}
 
-\vspace*{-2\baselineskip}
+% \vspace*{-2\baselineskip}
 \begin{prop}
@@
 called at most $|E|$ times. Since all queue operations require constant time when using a Strict
-Fibonacci Heap~\autocite{FibonacciHeaps12}, the complexity is $O(|E|+|V|)$, which is also $O(|E|)$
-since $G$ is connected.
+Fibonacci Heap~\autocite{FibonacciHeaps12}, the complexity is $O(|E|+|V|)$, which is also $O(|E|)$.
 \end{proof}
@@
   \includegraphics[width=0.4\linewidth]{assets/tikz/gtx_cross_edges_tikz.pdf}
-  \caption[Cross edge representation]{In this graph, there are four possible edges between the two
+  \caption[Cross edges between two stars]{In this graph, there are four possible edges between the two
   stars $s_1$ and $s_2$, and all are part of $cross\_edges[(s_1, s_2)]$, here represented in light
@@
 
- Then we analyze the time complexity. We already now that during the \tth{} iteration,
+ Then we analyze the time complexity. We already know that during the \tth{} iteration,
  \extractStar{} and \collapseStar{} take time $O(|E_t|)$. We only provide the actual python code
@@
  rid of this $O(|V_0|)$ at each iteration by ignoring eccentricity and taking a random edge between
- star, although it's difficult to tell how it would affect the quality of the resulting tree.}
+ star, although it is not clear at the moment how it would affect the quality of the resulting tree.}
  \textsc{Update-Nodes-Mapping} take $O(|V_0|)$ time, since they go through every node of the original
@@
 
-  Finally we prove that \gtx{} indeed returns a spanning tree of $G_0$. Specifically, we show that
+  Finally, we prove that \gtx{} indeed returns a spanning tree of $G_0$. Specifically, we show that
   by starting from $G_K$ and going in reverse order, we build a spanning tree $T_t$ of every $G_t$,
-  including eventually $G_0$. This process is illustrated on \autoref{fig:gtx_spanning_tree} and
-  works as follow: when moving from $G_{t+1}$ to $G_t$, we expand the nodes of $G_{t+1}$ into star
-  of nodes of $V_t$ and update all the outer edges we met so far to have their endpoints in $V_t$.
+  including eventually $G_0$. This process is illustrated in \autoref{fig:gtx_spanning_tree} and
+  works as follow: when moving from $G_{t+1}$ to $G_t$, we expand each node of $G_{t+1}$ into a star
+  of nodes of $V_t$, and update all the outer edges we met so far to have their endpoints in $V_t$.
   These outer edges are exactly $T_{t+1}$ and by translating them into $E_t$ and adding the inner
@@
 \end{proof}
-\vspace{-2\baselineskip}
+% \vspace{-2\baselineskip}
 \begin{figure*}[h]
@@
 
-We illustrate the operation of the \gtx{} algorithm on a small (and somewhat contrived) example.
+We illustrate the operation of the \gtx{} algorithm on a small example.
 Let us start with the initial graph $G_0$ depicted in \autoref{fig:gtx_eccentricity}
@@
 
-We then call \collapseStar{}. This will connect all possible pairs of star. For instance, the edge
+We then call \collapseStar{}. This will connect all possible pairs of stars. For instance, the edge
 between nodes $19$ and $29$ leads to the edge between $\starone{4}$ and $\starone{6}$. This is
@@
 consecutive nodes. In the worst case, centers will be chosen such that we have a succession of three
-and two nodes stars, as in \autoref{fig:gtx_line_graph}.%
+and two nodes stars, as in \autoref{fig:gtx_line_graph}.
 This will results in $\nicefrac{2n}{5}$ stars organized in a line, which is less than half of $n$.
@@
 tree forcing a lot of mistakes. We add an element of randomization to \extractStar{} by letting it
-use of two optional arguments, a \emph{threshold function} $\tau$ or a \emph{degree function}
+use two optional arguments, a \emph{threshold function} $\tau$ or a \emph{degree function}
 $\widetilde{d}$. Such functions modify the center sampling process in the following way:
@@
 a trade off between having more edges than $|V_0|-1$ but potentially making shorter connection.
-This would also be interesting \esp{} problem. Assuming
+This would also be interesting for the \esp{} problem. Assuming
 the treewidth of $G_t$ is decreasing with $t$, we could try to show there is \enquote{reasonable}

### 2017-12-18 cc/gtx_related.tex
fix cc chapter layout
pass over CC

@@
 
-Looking for a subgraph $H$ of $G$ that best preserve the distance in $G$ while being sparse is an old
+Looking for a subgraph $H$ of $G$ that best preserves the distance in $G$ while being sparse is an old
 problem, driven originally by network design in fields such as transportation~\autocite{RoadNetworks60}
@@
 distance}, and the exact form of $H$ give rise to several problems, which we summarize later in
-\autoref{tab:gtx_related_stretch} \vpageref{tab:gtx_related_stretch}. We first give some
+\autoref{tab:gtx_related_stretch}. We first give some
 definitions, then cover the most relevant problems in details, and finally give some pointers for
@@
 Another interesting idea from this paper is to consider a distribution over trees instead of a
-single instance, especially when one is concern about the maximum stretch instead of the average
+single instance, especially when one is concerned about the maximum stretch instead of the average
 one. For instance, on a cycle with $n$ nodes, a tree is obtained by removing one edge, and that edge
@@
 Starting from an empty subgraph $H$, it goes through every edge $(u, v)$ of $G$ sorted by weight and
-check if there is a path between $u$ and $v$ in $H$ of length at most $t$. If it is case the edge
+check if there is a path between $u$ and $v$ in $H$ of length at most $t$. If this is the case, the edge
 $(u,v)$ is dropped, otherwise it is inserted in $H$. This results in a $(2t - 1)$-spanner with
@@
 
-\setlength{\fullpage}{\textwidth+\marginparsep+\marginparwidth}
+\setlength{\fullpage}{179mm}
 \begin{table}[htbp]
+\begin{adjustwidth}{-2cm}{}
   \centering
@@
   \end{tabulary}
+\end{adjustwidth}
 \end{table}

### 2017-12-19 edge/conclusions.tex
tiny changes and bib data for edges
pass over edge

@@
-\section{Conclusions}
+\section{Open directions}
 \label{sec:edge_conc}
@@
     how it improves the detection of relevant communities. Beyond predicting the relative order of
-    links, it would also be interesting to include time into nodes similarity itself.
+    links, it would also be interesting to include time into node similarity itself.
     %
@@
 
-  \item Another rich class of graphs were predicting edge types would be a challenging and rewarding
+  \item Another rich class of graphs where predicting edge types would be a challenging and rewarding
     task is \textbf{multigraph}, which allow several edges between two
@@
     the flattening of several graphs sharing the same nodeset, but where the provenance of each edge
-    would have been lost and need to be recover. With our original formulation of goodness as
+    would have been lost and need to be recovered. With our original formulation of goodness as
     ${s_{uv}}^T w_{uv}$, only one direction can achieve the higher score (discarding the rare case
     of tie). One alternative we proposed was to let each $w_{uv}$ be a linear combination of a small
-    number of base directions, although experimentally it did not fit well our synthetic data.
+    number of base directions.
     Symmetrically, another idea could be to have more than one way to compute the similarity
@@
     However, while this is a general concept, its exact implementation might be
-    application-dependant. Furthermore, in our current setting, we do not know in advance in the
-    semantic of each direction, and we have no labeled edges. Therefore, it is not clear how those
-    constraints would be specified or whether they can be learned, if applicable. A limited solution
+    application-dependant. Furthermore, in our current setting, we do not know in advance the
+    semantic of each direction, nor do we have labeled edges. Therefore, it is not clear how those
+    constraints would be specified, or whether they can be learned, if applicable. A limited solution
     exists in the context of heterogeneous information networks, where in addition to having several
-    types of edges, there are also several types of nodes. Since not all type of node can be
-    connected with any type of edges, the possible connections are themselves represented as a meta
+    types of edges, there are also several types of nodes. Since not all types of node can be
+    connected with any types of edge, the possible connections are themselves represented as a meta
     graph called the network schema~\autocite[Definition 3]{HINSurvey17}. This approach would need

### 2017-12-19 edge/experiments.tex
fix layout of edge|vector chapter
pass over edge

@@
 While there exists several methods to generate attributed graph with community
-structure~\autocites{Yang2013}{XuBayesian14}{Kataoka2016}, here we present the one we adopted for
+structure~\autocites{Yang2013}{XuBayesian14}{Kataoka2016}, here we present the one we devised for
 our initial experiments, as it is specifically tailored to the model introduced in \autoref{sub:edge_setting}.
@@
 minimizing the number of colorless edges.
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.6\textwidth]{tikz/edge_palette_tikz.pdf}
+  \caption[Finding the $k_\mathrm{local}$ directions of node when generating synthetic graph]{A
+    small example of the node palette assignment, with $k=4$ colors (blue, green, red and
+    orange) and palettes of size $k_\mathrm{local}=3$. We assume nodes $1$ and $2$ have already been
+    visited, and got assigned the palette $(\text{blue, green, red})$ and $(\text{blue, green,
+    orange})$ respectively. Moreover, we are currently visiting node $3$ while node $4$ is yet uncolored. In that
+    case, there are two possible palettes for node $3$. If we select $(\text{green, red, orange})$, a
+    possible color assignment for the edges $(1,2)$, $(1, 3)$ and $(2, 3)$ is respectively blue,
+    red, and green.
+  \label{fig:edge_palette}}
+\end{figure}
 
@@
 the palette assigned to every node.
-\begin{figure}[bht]
-  \centering
-  \includegraphics[width=0.4\textwidth]{tikz/edge_palette_tikz.pdf}
-  \caption{A small example of the node palette assignment, with $k=4$ colors (blue, green, red and
-    orange) and palettes of size $k_\mathrm{local}=3$. We assume nodes $1$ and $2$ have already been
-    visited, and got assigned the palette $(\text{blue, green, red})$ and $(\text{blue, green,
-    orange})$ respectively. Moreover, we are currently visiting node $3$ while node $4$ is yet uncolored. In that
-    case, there are two possible palettes for node $3$. If we select $(\text{green, red, orange})$, a
-    possible color assignment for the edges $(1,2)$, $(1, 3)$ and $(2, 3)$ is respectively blue,
-    red, and green.
-  \label{fig:edge_palette}}
-\end{figure}
 
@@
 Returning to directions instead of colors, we have a direction $w_{y_{uv}} \in \mathcal{D}_k$
-assigned to every edge $(u,v)$. For simplicity, in that paragraph, we will let $w_{uv} = w_{y_{uv}}$.
+assigned to every edge $(u,v)$. For easy of notation, in that paragraph, we will let $w_{uv} = w_{y_{uv}}$.
 The final step in generating our node-attributed graph is to
-find a set of user profiles $X=\{x_u\}_{u\in V}$ that maximize the edge goodness and minimize the node
+find a set of user profiles $X=\{x_u\}_{u\in V}$ that maximizes the edge goodness and minimizes the node
 loss term, defined respectively as:
@@
 To account for the randomness of this generation process, we repeat our
-measurement over \np{200} such generation of $\mathcal{D}_k$ and $X$. On the other hand, we verify experimentally that
+measurement over \np{200} such generations of $\mathcal{D}_k$ and $X$. On the other hand, we verify experimentally that
 generating various graphs while keeping $\mathcal{D}_k$ fixed gives the same results, hence we will
@@
 direction to an edge $(u,v)$ as $\mathcal{E}(u,v) = \argmax_{\ell \in \rangesk} g(s_{uv}, \wh{w}_\ell)$.
-Recall that during the second of the step of the instance generation, we assign a direction index
+Recall that during the second of the step of the instance generation, we assigned a direction index
 $y_{uv}$ to every edge.
@@
 \begin{table}[htb]
+\begin{adjustwidth}{-1.5cm}{}
   \centering
   \small
-  \caption{The degree to which the edge assignment of step 2 agree with the directions and profiles
+  \caption[Qualite of our synthetic instances according to generation parameters]{The degree to
+    which the edge assignment of step 2 agree with the directions and profiles
     of steps 3 and 4. This is measured by the AMI between the assignment $\{y_{uv}\}_{uv \in E}$ and
@@
 \end{tabular}
+\end{adjustwidth}
 \end{table}
@@
 adequate. Regarding the generative parameters, and compared with the default configuration, the
-problem is easier when the $k$ directions are spread in a larger number of dimension, and more
+problem is easier when the $k$ directions are spread in a larger number of dimensions, and more
 difficult when each nodes is involved in four directions instead of three. On the other hand,
@@
 
-\begin{table}[htb]
+\begin{table*}[htb]
+\begin{adjustwidth}{-2cm}{}
   \centering
@@
   \small
-  \caption{Standardized AMI of 5 methods, when generating directions with 7 configurations. For each
+  \caption[Performance of our proposed methods on synthetic instances]{Standardized AMI of 5
+    methods, when generating directions with 7 configurations. For each
     configuration, we generate directions 200 times, and report the mean standardized AMI along with
@@
   \label{tab:edge_Wvaries_nami}}
-  \begin{tabular}{llll|ll}
+  \begin{tabulary}{179mm}{llll|ll}
     \toprule
@@
     \bottomrule
-  \end{tabular}
-\end{table}
+  \end{tabulary}
+\end{adjustwidth}
+\end{table*}
 
@@
 
-\begin{table}[hbt]
+\begin{table*}[hbt]
+\begin{adjustwidth}{-2cm}{}
   \centering
@@
   \label{tab:edge_Wvaries_dst}}
-    \begin{tabular}{llll|ll}
+  \begin{tabulary}{179mm}{llll|ll}
       \toprule
@@
       \bottomrule
-    \end{tabular}
-\end{table}
+    \end{tabulary}
+\end{adjustwidth}
+\end{table*}
 

### 2017-12-19 edge/intro.tex
pass over edge

@@
 First, we want an interpretable model that, given two connected nodes, can explain why they are
-linked. This lead us to use a linear model, for it is both easily amenable to
+linked. This leads us to use a linear model, for it is both easily amenable to
 interpretation and computationally efficient. Second, edge-labeled graphs are susceptible to further
@@
 
-As demonstrated by our prior extensive coverage, predicting edge signs is an interesting and
+As demonstrated by our prior extensive coverage, predicting edge sign is an interesting and
 challenging problem in its own right. Yet it is natural to extend and generalize it.  First, there
@@
 of edges incident to every node and learn a model from such features.
-A similar tension between relying solely on always available yet sparse topological information or
+A similar tension between relying solely on always available albeit sparse topological information or
 leveraging possibly richer attributes information also exists in link
@@
 \esp{}, this time the objects we seek to cluster are edges and not nodes. Furthermore, we stress
-that the profiles crucially increases the interpretability of our models.
+that the profiles crucially increase the interpretability of our models.
 

### 2017-12-19 edge/model.tex
pass over edge

@@
     where $\circ$ stands for the Hadamard (or component wise) product. As showed in
-    \autoref{tab:edge_sim}, for each dimension $i \in \rangesd$, the $s_{uv}$ holds a value
+    \autoref{tab:edge_sim}, for each dimension $i \in \rangesd$, the vector $s_{uv}$ holds a value
     indicating whether the profiles of $u$ and $v$ agree, disagree or are indifferent along that
@@
 not primordial, the role of bounding the norm of the directions (\ie{} having $\sum_{i=1}^d
-{w_\ell}_{;i}^2 \leq B^2$) is to tie the $d$ dimensions together. Without such a constraint, and we
+{w_\ell}_{;i}^2 \leq B^2$) is to tie the $d$ dimensions together. Without such a constraint, and with
 our choice of $s$ and $g$, we could simply study each dimension separately, whereas this bound adds
@@
 $-1$ and $1$. Finally, \textsf{opinion} represents the view of former \textsf{F} employees about a
-societal issue, with $0$ denoting a neutral position. In Table~\ref{tab:edge_exe}, we also show two
+societal issue, with $0$ denoting a neutral or unknown position. In Table~\ref{tab:edge_exe}, we also show two
 directions that explain the connections in the graph $G$. The first covers being of the same age and
@@
 
-In addition to assume that the input graph is the superposition of a small number of layers,
+In addition to assuming that the input graph is the superposition of a small number of layers,
 another way to leverage the graph topology is to define local constraints at the node level.
@@
 \end{equation}
-from \eqref{eq:edge_full}. The cost we pay is introducing the extra parameters $\{a_{uv}\}_{uv \in
+from \eqref{eq:edge_full}. The cost we pay is introducing the extra parameters $\{a_{uv}\}_{\euv{} \in
 E}$ and $\{b_u\}_{u \in V}$ to be learned, but we will see in the next section how to handle that in
@@
 \begin{equation*}
-  y_{uv,i} = \frac{\exp(\beta g(s_{uv}, w_i))}{\sum_{j=1}^k \exp(\beta g(s_{uv}, w_j))}\,,
+  {y_{uv}}_{;i} = \frac{\exp(\beta g(s_{uv}, w_i))}{\sum_{j=1}^k \exp(\beta g(s_{uv}, w_j))}\,,
 \end{equation*}
@@
 considering a low-rank matrix formulation, where each edge is assigned a linear combination of a
-small number of base directions, paving the way to overlapping clustering. More precisely, we
+small number of base directions, paving the way to overlapping edge clustering. More precisely, we
 give a convex formulation of that objective that can be solved by a \newline
 \hspace*{14pt} 4) \fwa{} algorithm, \newline
-as well an alternating optimization method, where we make the low-rank factorization \newline
+as well as by an alternating optimization method, where we make the low-rank factorization \newline
 \hspace*{14pt} 5) \pqt{}.
@@
 replace here by a convex relaxation. Indeed, one can check that given a set of numbers $S=\{a_1,
-a_2, \ldots, a_{|S|}\}$ and real number $\beta > 0$, we have
+a_2, \ldots, a_{|S|}\}$ and a real number $\beta > 0$, we have
 \begin{equation*}
@@
 \end{equation*}
-  \label{}
 Using this fact, and dividing the goodness term and the $\mathcal{L}_{\mathrm{node}}$ term of
@@
 Frank--Wolfe algorithm. Then we take the alternative road of explicit low-rank matrix factorization,
-and we conclude by commenting on these approaches.
+and we conclude by commenting on the respective merits and disadvantage of all these approaches.
 
@@
 are convex, we can solve it using the Frank--Wolfe algorithm~\autocites{FrankWolfe56}{Jaggi2013a},
-which enjoy properties like a $O(\frac{1}{t})$ convergence rate, a low computational cost by avoiding
+which enjoys properties like a $O(\frac{1}{t})$ convergence rate, a low computational cost by avoiding
 projection step and the sparsity of its solution. It is an iterative procedure that maintains an estimate
@@
 As said at the beginning of this section, the matrix formulation is not exactly solving
-\autoref{p:edge_full}. However, for evaluation purposes, here we describe two ways to extract $k$
+\autoref{p:edge_full}. Therefore, for evaluation purposes, here we describe two ways to extract $k$
 directions and an edge assignment from the matrix $W$ obtained through this \fwa{} method. The first
@@
 optimization step. Specifically, letting $r$ be the rank of $W$, we first compute a reduced SVD such
-that $W = U \Sigma V^T$ and let $P = U \in \mathbb{M}^{d\times r}$ and $Q^T = \Sigma V^T \in \Rbb^{r
+that $W = U \Sigma V^T$, and then let $P = U \in \mathbb{M}^{d\times r}$ and $Q^T = \Sigma V^T \in \Rbb^{r
 \times |E|}$. We then note that for any invertible matrix $R \in \Rbb^{r \times r}$, we have $W =
@@
 
-Compared with the \combined{} approach of the previous \autoref{sub:edge_vector}, these two matrix
+Compared with the \combined{} approach of \autoref{sub:edge_vector}, these two matrix
 formulations require learning more parameters. Indeed, it increases from $dk$ to $d|E|$, as it each

### 2017-12-19 edge/related.tex
fix layout of edge|vector chapter
pass over edge

@@
 maximizing the likelihood of the observed labels through gradient descent. Those parameters are then
-use to predict the remaining label that further maximize the total
+use to predict the remaining labels that further maximize the total
 likelihood~\autocite{graphicalModelTies11}. An extension of this graphical model is to actively
@@
 therefore computed based on topological features. Finally, a last direct approach to labelling edges
-is presented by \textcite{Aggarwal2016a}. They consider that the input graph has three type of
-edges: labeled by $-1$, labeled by $+1$ or unlabeled. In each of the three induced subgraphs, they
+is presented by \textcite{Aggarwal2016a}. They consider that the input graph has three types of
+edges labeled by $-1$, labeled by $+1$ or unlabeled. In each of the three induced subgraphs, they
 define a similarity between two nodes $u$ and $v$ as the Jaccard coefficient of the two sets of
@@
 
-Besides social network, the task of predicting the semantic of an edge has also been addressed in
+Besides social networks, the task of predicting the semantic of an edge has also been addressed in
 the context of knowledge graphs. In such graphs, nodes are abstract concepts and concrete entities
@@
 relations can be modeled by translations. Thus, the existence of an $(h,l,t)$ edge offers evidence that
-$\bm{h} + \bm{\ell} \approx \bm{l}$. Given a training set of existing triplets $S$ and a
+$\bm{h} + \bm{\ell} \approx \bm{t}$. Given a training set of existing triplets $S$ and a
 dissimilarity measure $D$, the authors learn an embedding in a fashion reminiscent of metric
@@
 recent survey on embedding knowledge graphs~\autocite{KnowledgeGraphSurvey17}, which among other
-ideas present the interesting notion of composing relation along paths between two entities. We note
+ideas present the interesting notion of composing relations along paths between two entities. We note
 that, in contrast with us, those methods require some form of supervision when it comes to predict
@@
 the goal of \autocite{edgerep17} is eventually to find an overlapping clustering of the nodes, not
-the edges). More closely related to our work, \textcite{ahmed2017roles} seek to assign a role to the
+of the edges). More closely related to our work, \textcite{ahmed2017roles} seek to assign a role to the
 edges of a graph. First they learn edge features as follow: starting from topological features, they
@@
 Once they obtain such a $d$-dimensional feature vector for each edge, they stack them in a matrix $S
-\in \Rbb^{|E|\times d}$ and look for $U^{|E| \times r}$ and $B^{r \times d}$ such that the distance
+\in \Rbb^{|E|\times d}$ and look for $U\in \Rbb^{|E| \times r}$ and $V\in \Rbb^{r \times d}$ such that the distance
 between $S$ and $UV^T$ is minimized. $V$ is a description of $r$ roles in terms of features while
@@
 space and use subspace clustering to find cells that have low entropy and high connectivity, before
-merging such cells into node clusters.
+merging such cells into clusters of nodes.
 % \textcite{subgraphEdgeLabels13} propose a method to extract subgraphs that are dense across
@@
 homophily, \textcite{beyondHome16} look for link patterns whose support and confidence are not only
-above some given thresholds, but that also diverge from homophily. Indeed, the author explicitly exclude
+above some given thresholds, but that also diverge from homophily. Indeed, the authors explicitly exclude
 attributes taking the same value within such link patterns.
@@
 is proportional to $\mathcal{D}_i(u, v)^{-d}$ and we observe the real network $\mathcal{G} =\bigcup_i
-\mathcal{G}_i$. From $\mathcal{G}$, the proposed algorithm recovers in $n \mathrm{polylog} n$ time a
+\mathcal{G}_i$. From $\mathcal{G}$, the proposed algorithm recovers in $O(n \polylog n)$ time a
 bounded approximation $\mathcal{D}'_i$ of all $\mathcal{D}_i$, that is there exists positive
@@
 \textcite{surveyAttributedClustering15}\footnote{Especially the section 2 dealing with edge labeled
-graphs.}. We also reiterate the warning given in the introduction that in large graphs, node
+graphs.}. We also reiterate the warning given in the introduction of this chapter. Namely, in large graphs, node
 attributes are not necessarily aligned with annotated communities. Such attributes thus need to be
@@
 
-\setlength{\fullpage}{\textwidth+\marginparsep+\marginparwidth}
-\begin{table}[htpb]
+\setlength{\fullpage}{179mm}
+\begin{table*}[tbh]
+\begin{adjustwidth}{-2cm}{}
   \centering
   \small
-  \caption{We summarize how each model generates: 1) the membership of a node $u$ to a community $c
+  \caption[Generative model of node attributed graphs with community structure]{We summarize how
+    each model generates: 1) the membership of a node $u$ to a community $c
   \in \rangesk$, 2) the attributes of $u$ knowing its community membership and 3) the edges between
@@
   \egroup
-\end{table}
+\end{adjustwidth}
+\end{table*}
 
@@
 present a model with interesting applications. Specifically, the nodes have mixed membership to $K$
-overlapping groups and each $L$ layer is generated by a specific $K\times K$ block matrix, taking
+overlapping groups and each of the $L$ layers is generated by a specific $K\times K$ block matrix, taking
 into account the group membership, like in a degree corrected SBM. Unlike us, this allows for

### 2017-12-19 notations.tex
add missing labels
start changing the layout
pass over edge

@@
 
+\autoref{tab:notations} presents a tentative list of notations, but is still subject to change.
+
 \begin{table*}[thpb]
+\begin{adjustwidth}{-2cm}{}
   \centering
@@
   \vspace{-.5\baselineskip}
-  \begin{tabulary}{187mm}{LL}
+  \begin{tabulary}{179mm}{LL}
     \toprule
@@
     \rangek{} & The natural integers from $1$ to $K$, \ie{} $\{1, 2, \ldots, K \}$ \\
-    $G$ & An arbitrary graph. It should be clear from the context whether it is weighted or not, and directed or not \\
+    $\Ind{p}$ & An indicator function evaluating to $1$ if the predicate $p$ is true and $0$ otherwise \\
+    $x_{;i}$ & For a vector $x\in \Rbb^d$, $x_{;i}$ denotes the value of its \ith{} component. \\
+    $\dsphere{}$ & The sphere of $\ell_2$ unit vectors in dimension $d$, \ie{} $\theset{x \in
+    \Rbb^d}{\|x\|_2 = 1}$ \\
+    $G$ & An unweighted graph. It should be clear from the context whether it is directed or not \\
     $V$ & The set of all the nodes of a graph, with $|V|=n$. Unless noted otherwise, nodes are indexed from $1$ to $n$ \\
@@
     that $|\nei(u)| = \degr(u)$ \\
+    $\neip(u)$ & The set of all positive neighbors of $u$, regardless of edge direction. That is,
+    $\forall v \in \neip(u),\, \yuv{} = +1$. \\
+    $\nein(u)$ & The set of all negative neighbors of $u$, regardless of edge direction. That is,
+    $\forall v \in \neip(u),\, \yuv{} = -1$. \\
     $T$ & An unweighted and undirected tree \\
@@
     $|\pathtuv|$ & The length of the path between $u$ and $v$ in the tree $T$, that is its number of edges \\
+    $\diam(G)$ & The diameter of $G$, that is the longest shortest path between any two nodes. \\
     \bottomrule
   \end{tabulary}
+\end{adjustwidth}
 \end{table*}
+
+% LP
+% NLP
+% SDP

### 2017-12-20 conclusion/main.tex
pass over conclusion

@@
     as they model social interactions in several domains, such as e-commerce and collaborations.
-    Here edges have two possible semantics, positive or negative. Our goal was to solve a supervised
+    In such graphs, edges have two possible semantics, positive or negative. Our goal was to solve a supervised
     binary classification problem by predicting the sign of test edges. For that, we first
@@
     measure offered by trollness and unpleasantness, and the inability of existing methods to cope
-    with new nodes joining the graph, we furthermore develop the first online algorithm for the
+    with new nodes joining the graph, we furthermore developed the first online algorithm for the
     \esp{} problem.
@@
     of edge. This means that instead of signed graphs, our input shifted to multilayer graphs, which
-    are also ubiquitous to describe real world phenomena. Moreover, we modified the learning
+    are also ubiquitous in describing real world phenomena. Moreover, we modified the learning
     setting, so that no edge label are available, but instead nodes carry a profile vector. However,
@@
     with side information and interpretability requirements. In order to do so, we first defined a
-    similarity between nodes, which accommodate both heterophily and homophily. We then formulate
+    similarity between nodes, which accommodates both heterophily and homophily. We then formulated
     the problem as simultaneously finding $k$ special vectors called directions, and assigning one
@@
 \subsection{Reciprocal recommendation}
+\label{ssec:conc_rr}
 
@@
 same item (say a movie) to thousands of users is not an issue. On the other hand, we have to keep
-user engaged and prevent them from being idle. This is done by ensuring they are regularly the
+users engaged and prevent them from being idle. This is done by ensuring they are regularly the
 subject or object of some recommendations.
@@
 graph is thus a special case of the \dssn{} studied in \autoref{chap:troll}. Each nodes carries two
-kind of information. The first is explicitly provided by the user, in order to arise interest in
+kind of (side) information. The first is explicitly provided by the user, in order to arise interest in
 users from the other group. The second is implicitly inferred from the user activity. In particular,
@@
 As in any worthy machine learning endeavor, reliably evaluating and comparing approaches require
-large amount of real data. This is particularly true in such an applied setting, but the most
-realistic evaluation is to perform A/B testing on a live system with enough traffic to give some
+large amount of real data. This is particularly true in such an applied setting, where the most
+realistic evaluation is to perform A/B testing on a live system with enough traffic to deliver
 statistical significance. This explains why most previous studies rely on the cooperation of
@@
 
-As an illustration in the context of this thesis, we assume that data was organized as a graph,
+As an illustration in the context of this thesis, we assumed that data was organized as a graph,
 where nodes were possibly associated with a feature vector. However, there might an different

### 2017-12-21 abstract.fr.tex
add french abstract

Dans cette thèse, nous proposons des méthodes pour caractériser efficacement et précisément les
arêtes au sein de réseaux complexes. Dans les graphes simples, les nœuds sont liés au travers d'une
sémantique unique. Par exemple, deux utilisateurs sont amis dans un réseau social, ou une page web
contient un lien hypertexte pointant vers un autre page. De plus, ces connexions sont généralement
guidées par la similarité entre les nœuds, au travers d'un mécanisme appelé homophilie. Dans les
exemples précédents, les utilisateurs deviennent amis à cause de caractéristiques communes, et les
pages web sont reliées les unes aux autres sur la base de sujets communs. En revanche, les réseaux
complexes sont des graphes où chaque connexion possède une sémantique parmi $k$ possibles. Ces
connexions sont en outre basées à la fois sur une homophilie et une hétérophilie partielle des nœuds
à leurs extrémité. Cette information supplémentaire permet une analyse plus fine des graphes issus
d'applications réelles. Cependant, elle peut être coûteuse à acquérir, ou n'est pas toujours
disponible a priori. Nous abordons donc le problème d'inférer la sémantique des arêtes dans
plusieurs contextes. Tout d'abord, nous considérons les graphes où les arêtes ont deux sémantiques
opposées, et où nous observons l'étiquette de certaines arêtes. Ces \enquote{graphes signés} sont
une façon élégante de représenter des interactions polarisées. Nous proposons deux biais
d'apprentissage, adaptés respectivement aux graphes signés dirigés et non dirigés. Ceci nous amène à
concevoir plusieurs algorithmes utilisant la topologie du graphe pour résoudre un problème de
classification binaire que nous appelons \esp{}. Deuxièmement, nous considérons les graphes avec $k
\geq 2$ sémantiques possibles pour les arêtes. Dans ce cas, nous ne recevons pas d'étiquette
d'arêtes, mais plutôt un vecteur de caractéristiques pour chaque nœud. Face à ce problème non
supervisé d'\ecp{}, nous concevons un critère de qualité exprimant dans quelle mesure une
$k$-partition des arêtes et $k$ vecteurs sémantiques expliquent les connexions observées. Nous
optimisons ce critère \enquote{qualité explicative} sous une forme vectorielle et matricielle et
illustrons le comportement de ces deux méthodes sur des données synthétiques.

### 2017-12-21 assets/tikz/cc_2cc_tikz.tex
redraw 3 figures

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
\begin{tikzpicture}[auto,vertex/.append style={minimum width=5mm},
  edge/.append style={line width=1.2pt,color=Black}]

  \node[vertex,medium,] (w) at (0.000, -0.000) {$w$};
  \node[vertex,medium,] (v) at (-.60, -.8315) {$v$};
  \node[vertex,medium,] (u) at (1.53, -1.03) {$u$};

  \draw[edge,Green] (w) -- node [] {$+$} (u) ;
  \draw[edge,Green] (v) -- node [below] {$+$} (u) ;
  \draw[edge,dashed] (v) -- node [] {?} (w) ;

  \draw[color=DodgerBlue] (.5, -.5) circle [x radius=1.8, y radius=1.1]  node[xshift=30, yshift=15,text=Black] {$\mathbf{L}$};

  \node[vertex,medium,] (y) at (-0.200, -4.000) {$y$};
  \node[vertex,medium,] (x) at (1.400, -4.300) {$x$};
  \draw[edge,dashed] (x) -- node [] {?} (y) ;

  \draw[edge,dashed] (v) -- node [] {?} (y) ;

  \draw[edge,Red] (y) -- node [] {$-$} (u) ;
  \draw[edge,Red] (x) -- node [] {$-$} (u) ;

  \draw[color=DodgerBlue] (.4, -4.5) circle [x radius=1.5, y radius=1]  node[xshift=22, yshift=-10,text=Black] {$\mathbf{R}$};

\end{tikzpicture}
\end{document}

### 2017-12-21 assets/tikz/cc_4triads_tikz.tex
redraw 3 figures

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
\begin{tikzpicture}[auto,vertex/.append style={minimum width=5mm},
  edge/.append style={line width=1.2pt,color=Black}]

  \node[vertex,medium,] (t11) at (0,0) {};
  \node[vertex,medium,] (t12) at (2,0) {};
  \node[vertex,medium,] (t13) at (1, 1.732) {};
  \draw[edge,Green] (t11) -- node[below] {$+$} (t12);
  \draw[edge,Green] (t11) -- node[left] {$+$} (t13);
  \draw[edge,Green] (t12) -- node[right] {$+$} (t13);
  \node[anchor=north,font=\small,scale=.8] (t1) at (1, -.5) {$T_3$ Mutual friends};

  \node[vertex,medium,] (t21) at (3.5,0) {};
  \node[vertex,medium,] (t22) at (5.5,0) {};
  \node[vertex,medium,] (t23) at (4.5, 1.732) {};
  \draw[edge,Green] (t21) -- node[below] {$+$} (t22);
  \draw[edge,Red] (t21) -- node[left] {$-$} (t23);
  \draw[edge,Red] (t22) -- node[right] {$-$} (t23);
  \node[anchor=north,font=\small,scale=.8] (t1) at (4.5, -.5) {$T_1$ Common enemy};

  \draw[edge,black,very thick] (-.20, 2.15) -- node[] {Balanced} (6.5, 2.15);

  \node[vertex,medium,] (t31) at (7,0) {};
  \node[vertex,medium,] (t32) at (9,0) {};
  \node[vertex,medium,] (t33) at (8, 1.732) {};
  \draw[edge,Red] (t31) -- node[below] {$-$} (t32);
  \draw[edge,Red] (t31) -- node[left] {$-$} (t33);
  \draw[edge,Red] (t32) -- node[right] {$-$} (t33);
  \node[anchor=north,font=\small,scale=.8] (t1) at (8, -.5) {$T_0$ Mutual enemies};

  \node[vertex,medium,] (t41) at (10.5,0) {};
  \node[vertex,medium,] (t42) at (12.5,0) {};
  \node[vertex,medium,] (t43) at (11.5, 1.732) {};
  \draw[edge,Red] (t41) -- node[below] {$-$} (t42);
  \draw[edge,Green] (t41) -- node[left] {$+$} (t43);
  \draw[edge,Green] (t42) -- node[right] {$+$} (t43);
  \node[anchor=north,font=\small,scale=.8] (t1) at (11.5, -.5) {$T_2$ Antagonist friends};
  \draw[edge,black,very thick,below] (-.20, -1.2) -- node[] {Weakly Balanced} (10, -1.2);
\end{tikzpicture}
\end{document}

### 2017-12-21 cc/bias_balance.tex
pass over CC
redraw 3 figures

@@
 
-\begin{wrapfigure}{o}[10mm]{45mm}
+\begin{wrapfigure}{o}[10mm]{40mm}
   \centering
-  \includegraphics[width=.9\linewidth]{assets/raw/2cc.pdf}
+  \includegraphics[width=.8\linewidth]{assets/tikz/cc_2cc_tikz.pdf}
   \caption{\small A two-clustering of a complete strongly balanced graph}
@@
 such that all edges within $L$ and $R$ are positive, while all edges between $L$ and $R$ are
-negative. Indeed, let us pick an arbitrary node $u$ and let $L=\neip(u)\setminus \{u\}$ and $R=V\setminus L$. Now
+negative. Indeed, let us pick an arbitrary node $u$ and let $L=\neip(u)\setminus \{u\}$ and
+$R=V\setminus L$ (where $\neip(u)$ denotes the set of all positive neighbors of $u$). Now
 let $v,w\in L\setminus \{u\}$ and $x,y\in R$. From \autoref{fig:cc_twostrongcc} and the fact that
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{assets/raw/triangle_smaller.pdf}
+	\includegraphics[width=0.8\linewidth]{assets/tikz/cc_4triads_tikz.pdf}
 	\caption{The four possible undirected triads, as classified by the two structural balance theories

### 2017-12-21 intro/main.tex
typos

@@
 \bf
-There exists efficient and accurate methods to predict edge type in complex networks, relying
+There exist efficient and accurate methods to predict edge type in complex networks, relying
 only on the graph topology or also on node attributes.

### 2017-12-21 troll/related.tex
typos

@@
 
-Other works have also\done\Todo{maybe that part would be better in the conclusion, especially since side
-information would help with the cold start problem showed in the experiments and trust version is
-one of the future direction.}
+Other works have also
 considered versions of the problem where side information related to the network is available to

### 2018-02-25 ack.tex
add title page and acknowlegments

First and foremost, I would like to express my deepest gratitude to my
supervisor, Pr. Marc Tommasi. From the day we first got in touch in that spring
four years ago and all the way towards my defense, he has provided me with
thorough guidance. In terms of research of course, but on many other topics as
well, from professional development to folkloric music. Moreover, this has
always been done in the kindest way possible. Therefore I'm doubly indebted to
him, for not only I would haven't complete my PhD without him, but he made it a
very pleasant experience. Second, such feelings carried over to include my
advisor, Dr. Fabio Vitale. To be honest, there had been time where his sense of
rigor has challenged me. But at the end, it was an important learning
experience, and his ability to crack a joke at the least expected moment of a
long research meeting has proved very useful: the fact we went to bars in at
least three different countries speaks for it. I would also like to thanks Pr.
Claudio Gentile for hosting me in Varese, and giving me sharp advice along
those three years.

Then I want to thank Alessandro Provetti and Liva Ralaivola who kindly agreed
to review my manuscript, as well as Elisa Fromont and Sophie Tison for having
accepted to be part of my committee.
%
To conclude this academic paragraph, I'd like to sincerely acknowledge how
lucky I was to collaborate with all my talented co-authors: Nicolò
Cesa-Bianchi, Emre Çelikten, Claudio Gentile, Michael Mathioudakis and Fabio
Vitale.

Although doing a PhD is a rather personal enterprise, I was fortunate to be
part of the awesome Magnet
team\footnote{\url{https://www.inria.fr/en/teams/magnet}}, whose past and
current members I met include, in seating order: Thomas, Pauline, David,
Mathieu (Thursday French Fries forever), Thanh, Carlos, Pierre, Paul, Juhi,
Quentin, William, Mikaela, Pascal (special thanks for great accommodation and
sportive spirit), Rémi, Aurélien and Hippolyte. Thanks for stimulating
discussions on many topics, constant readiness to help and answer my questions,
life long lessons of table football, and quite importantly, thanks for
patiently bearing with me while I was finishing lunches (and/or fiddling with
my phone)! Of course, this wouldn't be proper acknowledgement without
mentioning the best office in the world, namely B224, and the best possible
office mates: Nathalie (for all the cookies) and Thibault (because there were
some genuinely good jokes among all of them). Whereas I always been slightly
worried about logistics, I never had to think about it thanks to the fantastic
work of Julie. Finally, one last Magnet member, Antonino, left before I joined
but nonetheless deserve my gratitude for he helped me find my next job. Same
credit goes to Michal, in addition to being a great company in all occasions.
At this stage, I might as well thanks the rest of the Sequel team, and
especially Julien, Alexandre, Émilie, Frédéric, Romaric, Matteo, Daniele,
Julien, Bilal, Marta, Tomáš,
Daniele, Florian, Jean-Bastien, Olivier and Ronan.

Doing a PhD is also supposed to be a rather labor intensive enterprise, but
again, I was fortunate to live in the student city of Lille. There I met many
wonderful people, each of them who contributed in a way to this manuscript.  It
all started when a young version of myself attended a student event organized
by Tilda and later had to pleasure to collaborate with Guillaume, Antoine,
Fabien, Alexandre, and by extension with Carmelo, Jason, Cindy, Antoine
(Dujardin), Clément, Ilkay, Benjamin and Émilie. On my way to organize an
elusive café lingua, I also enjoyed the company of many members of a meetup
group, including Samir, Nathalie, Marie-Pierre, Anne Charlotte, Léa, Géraldine,
Fabienne, George, Norida and Thomas, as well as Rashida, Mairead and Yash, who
beared with me during circumstances better not written here. Another great
event in Lille is undeniably the \enquote{apéro culture}, where I was
privileged to meet Cyrille, Élise, Audrey, Wasilla and Olmo.  This picture
wouldn't be complete without Stéfana, Arthur, Esmeralda, Daniel, and my
flatmate Édouard for some nice video games nights.

I also received support from outside Lille, thanks to Tristan, Olivier (may the
for be with you guys), Aloïs, Juliette, Florent, Isabelle, Camille and Claire.
In these cold days of February, I also have a warm thoughts for people I met
during my visits in Aalto: Luiza, Annika and Klaudia as well as Sanja and Kiran
for their long standing support and comments on some part of this thesis.
Special mention to Vera, who relentlessly corrected my typos despite having to
endure countless pictures of food at any time of day.

At this point dear reader, you might become suspicious of whether a single man
can really be so lucky. Yet this is not over and I was again fortunate to met
just before starting my PhD a wonderful friend. Over those years, Nataša has
given me support that words probably can't describe. But it is safe to say that
without her, this thesis and even myself would both be very different.

\begin{otherlanguage}{french}
Enfin, j'aimerais remercier ma mère et mon frère, pour leur soutien
inconditionel sans lequel toute cette aventure n'aurais pas été possible.
\end{otherlanguage}

\vspace{5\baselineskip}
\begin{flushright}
  {\itshape
    Berlin, February 26, 2018
  }
\end{flushright}

### 2018-02-25 final.tex
compile all the content
add french abstract
add nice headers
add title page and acknowlegments

@@
-\documentclass[a4paper,svgnames,twoside,onecolumn,final,titlepage,11pt]{report}
+\documentclass[a4paper,svgnames,twoside,onecolumn,final,titlepage,11pt]{book}
 \input{pmin}
@@
 
+\begin{titlepage}
+%\setcounter{page}{0}  % to make the backref work properly on arabic numbers!!
+%\thispagestyle{empty}
+\newgeometry{top=.7in,bottom=0in,left=1.5cm,right=2.1cm}
+\noindent\rule{\textwidth}{0.4pt}\vspace{0.3cm}
+%\vrule height 1pt width.5\hsize
+\begin{center}
+{\LARGE \textbf{\includegraphics[scale=.14]{./assets/raw/univlille.png} \\ \vspace{0.15cm} École Doctorale Sciences Pour l'Ingénieur }}\\ \vspace{0.5cm}
+{\huge \textbf{\textsc{  Th\`ese de Doctorat}}}\\ \vspace{.8cm}
+{\large préparée au sein de  \raisebox{-1.2ex}{\includegraphics[scale=.088]{./assets/raw//logoCristal.jpg}}\\ \vspace{-.1cm} et du centre de recherche~~\raisebox{-.1ex}{\includegraphics[scale=.148]{./assets/raw//logo-inria.pdf}}~ Lille - Nord Europe\\ financée par \quad \raisebox{-.1ex}{\includegraphics[scale=.148]{./assets/raw//logo-inria.pdf}} \quad
+% et \raisebox{-3.0ex}{\includegraphics[scale=.1]{./assets/raw//hautsdefrance.jpg}}
+\\}
+\end{center}
+
+\vspace{.5cm}
+\begin{center}
+
+%présentée pour obtenir le grade de\\
+
+%\textbf{DOCTEUR EN SCIENCES DE L'USTL}\\
+
+{\LARGE Spécialité : \textbf{Informatique}}
+\end{center}
+
+\vspace{0.6cm}
+
+\begin{center}
+{\large présentée par \\
+\LARGE \textbf{Géraud \textsc{Le Falher}} \\}
+\end{center}
+
+\vspace{1cm}
+\addtolength{\oddsidemargin}{.275in}
+	\addtolength{\evensidemargin}{.225in}
+	\addtolength{\textwidth}{1.25in}
+\begin{center}
+\begin{minipage}{0.8\textwidth}
+\begin{center}
+\rule{\textwidth}{0.4pt}%\vspace{.1cm}
+\begin{Huge}\textsc{ \textbf{Characterizing edges in signed \\[.2cm] and vector-valued graphs  }}\end{Huge}
+\vspace{-.31cm}
+\rule{\textwidth}{0.4pt}
+%\begin{large}Subtitle\end{large}
+\end{center}
+
+\end{minipage}
+
+\end{center}
+\vspace{1.4cm}
+\begin{center}
+{\large sous la direction de \textbf{Dr. Marc \textsc{Tommasi}} \\}
+{\large et l'encadrement de \textbf{Fabio VITALE} \\}
+\end{center}
+\vspace{2.8cm}
+\begin{center}
+\vrule height .4pt width.5\hsize
+
+\vspace{0.3cm}
+{Soutenue publiquement à \textbf{Villeneuve d'Ascq}, le \textbf{16 avril 2018} devant le jury composé de :}
+
+\vspace{0.3cm}
+\begin{large}
+  \begin{tabular}{l@{\hskip 0.6in}l@{\hskip 0.5in}l}
+    M. Liva \textbf{RALAIVOLA}      & Univ. Marseille           & Rapporteur  \\
+    M. Alessandro \textbf{PROVETTI} & Birbeck College in London & Rapporteur  \\
+    Mme Elisa \textbf{FROMONT}      & Univ. Rennes              & Examinateur \\
+    Mme Sophie \textbf{TISON}       & Univ. Lille               & Examinateur \\
+    M. Claudio \textbf{GENTILE}     & Inria                     & Invité      \\
+    M. Fabio \textbf{VITALE}        & Univ. Lille               & Encadrant   \\
+    M. Marc \textbf{TOMMASI}        & Univ. Lille               & Directeur   \\
+  \end{tabular}
+\end{large}
+
+\end{center}\restoregeometry
+
+\end{titlepage}
+
 \maketitle
+\frontmatter
 
-% \section*{Abstract}
-% \input{abstract}
-%
-% \tableofcontents
-% \newpage
-%
-% \input{notations}
-%
-% \chapter{Introduction}
-% \subimport*{intro/}{main}
+\section*{Abstract}
+\input{abstract}
+
+\begin{otherlanguage}{french}
+  \section*{Résumé}
+  \input{abstract.fr.tex}
+\end{otherlanguage}
+
+\section*{Acknowledgements}
+\label{sec:acknowledgements}
+\input{ack}
+
+\tableofcontents
+\listoffigures
+\listoftables
+\newpage
+
+\input{notations}
+
+\mainmatter
+\pagestyle{fancy}
+
+\chapter{Introduction}
+\subimport*{intro/}{main}
 
@@
 \subimport*{troll/}{main}
-\end{document}
 
@@
 
+\backmatter
 \vspace{\baselineskip}
@@
 \setlength\bibitemsep{1pt}
-\todos
 \printbibliography

### 2018-02-26 ack.tex
pass of acknowledgments

@@
 supervisor, Pr. Marc Tommasi. From the day we first got in touch in that spring
-four years ago and all the way towards my defense, he has provided me with
+four years ago, and all the way towards my defense, he has provided me with
 thorough guidance. In terms of research of course, but on many other topics as
 well, from professional development to folkloric music. Moreover, this has
-always been done in the kindest way possible. Therefore I'm doubly indebted to
-him, for not only I would haven't complete my PhD without him, but he made it a
-very pleasant experience. Second, such feelings carried over to include my
-advisor, Dr. Fabio Vitale. To be honest, there had been time where his sense of
+always been done in the kindest way possible. Therefore, I'm doubly indebted to
+him, for not only I wouldn't have complete my PhD without him, but he made it a
+very pleasant experience. Second, such feelings carries over to include my
+advisor, Dr. Fabio Vitale. There has been time when his sense of
 rigor has challenged me. But at the end, it was an important learning
@@
 long research meeting has proved very useful: the fact we went to bars in at
-least three different countries speaks for it. I would also like to thanks Pr.
+least three different countries speaks for it. I would also like to thank Pr.
 Claudio Gentile for hosting me in Varese, and giving me sharp advice along
@@
 lucky I was to collaborate with all my talented co-authors: Nicolò
-Cesa-Bianchi, Emre Çelikten, Claudio Gentile, Michael Mathioudakis and Fabio
+Cesa-Bianchi, Emre Çelikten, Claudio Gentile, Aristides Gionis, Michael Mathioudakis and Fabio
 Vitale.
+
+\medskip
 
@@
 patiently bearing with me while I was finishing lunches (and/or fiddling with
-my phone)! Of course, this wouldn't be proper acknowledgement without
+my phone)! Of course, this wouldn't be proper acknowledgment without
 mentioning the best office in the world, namely B224, and the best possible
 office mates: Nathalie (for all the cookies) and Thibault (because there were
-some genuinely good jokes among all of them). Whereas I always been slightly
+some genuinely good jokes among all of them). Whereas I've always been slightly
 worried about logistics, I never had to think about it thanks to the fantastic
@@
 Julien, Bilal, Marta, Tomáš,
-Daniele, Florian, Jean-Bastien, Olivier and Ronan.
+Daniele, Florian, Jean-Bastien, Olivier and Ronan (too bad my defense is not on Friday, for I know
+you wouldn't be long to hit the dance floor). Finally I want to thank all the
+support crews from Inria, as they made it a great place for doing research, as
+well as the University Lille 3 for making my short teaching experience so
+pleasant.
+
+\medskip
 
@@
 Fabienne, George, Norida and Thomas, as well as Rashida, Mairead and Yash, who
-beared with me during circumstances better not written here. Another great
+bore with me during circumstances better not written here. Another great
 event in Lille is undeniably the \enquote{apéro culture}, where I was
@@
 for be with you guys), Aloïs, Juliette, Florent, Isabelle, Camille and Claire.
-In these cold days of February, I also have a warm thoughts for people I met
-during my visits in Aalto: Luiza, Annika and Klaudia as well as Sanja and Kiran
-for their long standing support and comments on some part of this thesis.
-Special mention to Vera, who relentlessly corrected my typos despite having to
-endure countless pictures of food at any time of day.
+In these cold days of February, I also have a warm thought for people I met
+during my visits in Aalto: Luiza, Annika, Eric, Polina and Klaudia as well as
+Kiran, Vera and Sanja
+for their long-standing support and comments on some part of this thesis.
 
-At this point dear reader, you might become suspicious of whether a single man
+You astute reader may at this point grow suspicious of whether a single man
 can really be so lucky. Yet this is not over and I was again fortunate to met
 just before starting my PhD a wonderful friend. Over those years, Nataša has
-given me support that words probably can't describe. But it is safe to say that
-without her, this thesis and even myself would both be very different.
+given me support that words probably can't describe. But it's safe to say that
+without her, this thesis and even myself would both be lacking something
+essential.
+
+\medskip
 
@@
   {\itshape
-    Berlin, February 26, 2018
+    Berlin, Germany, February 27, 2018 \\
+    Géraud Le Falher
   }

### 2018-03-31 edge/model.tex
shorten table captions in edge chapter
typo

@@
       % \captionsetup{font=small,type=table}
-      \caption{\small We show the effect of the Hadamard product in combining information from two
+      \caption[Combining node profiles through the Hadamard product]{%
+        \small We show the effect of the Hadamard product in combining information from two
         users $u$ and $v$. Instead of actual numbers, in this table we use $+$ to denote a large
@@
 once a node has picked two allowed directions from its two neighbors, the remaining ones can be
-chosen arbitrarily. On the complete graph on the other hand, this choice is must more constrained.
+chosen arbitrarily. On the complete graph on the other hand, this choice is much more constrained.
 

### 2018-04-09 assets/tikz/defense_original_tikz.tex
material for the talk

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
       \begin{tikzpicture}[auto, every node/.style={transform shape},baseline=(b.base)]
         \tikzset{>=latex}
         \tikzstyle{peers}=[draw,circle,black,minimum width=9mm,inner sep=2pt]
         \tikzstyle{nedge}=[draw,rectangle,black,minimum width=8.5mm,minimum height=8.5mm, inner sep=0]
         \tikzstyle{edge}=[line width=2pt,color=Black,->]
         \tikzstyle{uedge}=[line width=2pt,color=Black]
         \tikzstyle{pnode}=[fill=White]
         \tikzstyle{elabel}=[color=Black]
         \tikzstyle{qnode}=[fill=White]
         \tikzstyle{posc}=[color=Green]
         \tikzstyle{negc}=[color=red]
         \foreach \place/\name in { {(0,0)/1}, {(3,0)/2}, {(0,-3)/3}, {(3,-3)/4}}
         \node[peers,pnode] (\name) at \place {\large $\name$};
         \draw[edge] (3) -- node [posc]                            {$+$}  (1);
         \draw[edge] (3) -- node [posc]                           {$+$}  (4);
         \draw[edge, dashed] (3) -- node [negc]                    {$-$}  (2);
         \draw[edge, dashed] (2) -- node [negc]                    {$-$}  (4);
         \draw[edge] (1) edge [bend left =10] node [posc]         {$+$} (2);
         \draw[edge, dashed] (2) edge [bend left =10] node [negc] {$-$} (1);
         \node[] (caption) at (0.5, 0.5) {$G$};
         \node[] (b) at (1.5, -1.5) {};
       \end{tikzpicture}
\end{document}

### 2018-04-09 assets/tikz/defense_signed_tikz.tex
material for the talk

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
\begin{tikzpicture}[auto,vertex/.append style={minimum width=10mm}]
  \tikzset{>=latex}
  \tikzstyle{peers}=[draw,circle,black,inner sep=2pt,fill opacity=.0, text opacity=1]
  \tikzstyle{edge}=[line width=1.5pt,color=Black,-{Latex[length=3mm,width=1.0mm,angle'=40]}]
  \tikzstyle{trainEdge}=[line width=6pt,color=Black!15,opacity=.7]
  \tikzstyle{elabel}=[color=Black]
  \tikzstyle{posc}=[scale=1.0,text=Green]
  \tikzstyle{negc}=[scale=1.0,text=red]
  \tikzstyle{testset}=[opacity=1.0]
  \tikzstyle{revealed}=[circle,draw=none,fill=Black!15,inner sep=1pt,]

  \node[peers] (i) at (0,0) {$u$};
  \node[peers] (j) at (3.5,-0.66) {$v$};

  \draw[trainEdge] (i) --                                                           (3, -2);
  \draw[edge] (i) -- node [revealed,posc,xshift=-2mm,] {$\mathbf{+}$}                      (3, -2);

  \draw[testset,edge] (i) edge  node [posc,xshift=-2mm] {$\textcolor{Black}{y_{uv}=}+$} (j);
  \draw[testset,edge] (i) edge  node [posc,xshift=2mm] {$\mathbf{+}$}                    (3, 0.66);
  \draw[testset,edge, dashed] (i)   edge   node [negc,xshift=2mm] {$\mathbf{-}$} (3, 2);

  \draw[trainEdge] (-3, -2) --    (i);
  \draw[edge, dashed] (-3, -2) --   node [revealed,negc,yshift=-3mm,xshift=-5pt] {$\mathbf{-}$} (i);

  \draw[testset,edge, dashed] (-3, -1) --   node [negc,yshift=-1mm] {$\mathbf{-}$} (i);
  \draw[testset,edge, dashed] (-3, 0)  --  node [negc] {$\mathbf{-}$} (i);
  \draw[testset,edge,]                (-3, 1) --   node [posc,xshift=-4mm] {$\mathbf{+}$} (i);

  \draw[trainEdge,]           (-3, 2) --   (i);
  \draw[edge,]                (-3, 2) --   node [revealed,posc,above,pos=0.2,xshift=5pt] {$\mathbf{+}$} (i);

  % \node[fill=white!90!DarkOrange,fill opacity=.9,text opacity=1,rounded corners,inner sep=2pt] at (1,-1.5) {$tr(i)=\frac{1}{4}$};
  % \node[fill=white!90!Blue,fill opacity=.5,text opacity=1,rounded corners,inner sep=2pt] at (-1.0,1.7) {$un(i)=\frac{3}{5}$};
\end{tikzpicture}
\end{document}

### 2018-04-09 assets/tikz/defense_weighted_tikz.tex
material for the talk

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\newcommand{\epslabel}{{$2$}}
\usepackage{fixltx2e,graphicx}
\newcommand{\unaryminus}{\scalebox{0.5}[1.0]{\(-\)}}
\newcommand{\minuslabel}{{$\unaryminus1$}}

\begin{document}
       \begin{tikzpicture}[auto, every node/.style={transform shape}]
         \tikzset{>=latex}
         \tikzstyle{peers}=[font=\small,draw,circle,black,minimum width=11mm,inner sep=0,fill opacity=.0, text opacity=1]
         \tikzstyle{nedge}=[font=\small,draw,rectangle,black,minimum width=9mm,minimum height=9mm, inner sep=0]
         \tikzstyle{uedge}=[line width=1.5pt,color=Black]
         \tikzstyle{edge}=[uedge,-{Latex[length=4mm,width=1.5mm,angle'=40]}]
         \tikzstyle{pnode}=[fill=White]
         \tikzstyle{elabel}=[color=Black,font=\footnotesize]
         \tikzstyle{qnode}=[fill=White]
         \tikzstyle{posc}=[color=Green, fill=Green!30!white]
         \tikzstyle{negc}=[color=red, fill=red!30!white]
         \tikzstyle{neuc}=[color=black!80!white, fill=black!30!white]
         \tikzstyle{eps}=[color=DarkViolet]
         \tikzstyle{minus}=[color=Goldenrod]

         \node[peers,pnode] (p1) at (1.3, -1.0) {$1_{\mathrm{out}}$}; 
         \node[peers,pnode] (p2) at (6.4, -2.0) {$2_{\mathrm{out}}$}; 
         \node[peers,pnode] (p3) at (1.3, -6.0) {$3_{\mathrm{out}}$}; 
         \node[peers,pnode] (p4) at (6.4, -5.0) {$4_{\mathrm{out}}$}; 
         \node[peers,qnode] (q1) at (0.0, -2.0) {$1_{\mathrm{in}}$}; 
         \node[peers,qnode] (q2) at (5.1, -1.0) {$2_{\mathrm{in}}$}; 
         \node[peers,qnode] (q3) at (0.0, -5.0) {$3_{\mathrm{in}}$}; 
         \node[peers,qnode] (q4) at (5.1, -6.0) {$4_{\mathrm{in}}$}; 
         \node[nedge,posc]  (e1) at (3.2, -1.0) {$1$}; 
         \node[nedge,negc]  (e2) at (3.2, -2.0) {$\unaryminus1$}; 
         \node[nedge,negc]  (e3) at (5.2, -3.7) {$\unaryminus1$}; 
         \node[nedge,posc]  (e4) at (1.2, -3.7) {$1$}; 
         \node[nedge,posc]  (e5) at (3.2, -3.7) {$1$}; 
         \node[nedge,negc]  (e6) at (3.2, -6.0) {$\unaryminus1$};
         \node[] (caption) at (0.2, -.5) {\small $G''$};

         \draw[uedge,eps] (p1) edge [bend left=0] node [elabel,]               {\epslabel} (e1);
         \draw[uedge,eps] (p2) edge [bend left=0] node [elabel,pos=.25]       {\epslabel} (e2);
         \draw[uedge,eps] (e2) edge [bend left=0] node [elabel,]               {\epslabel} (q1);
         \draw[uedge,eps] (p2) edge [bend left=0] node [elabel,]               {\epslabel} (e3);
         \draw[uedge,eps] (e3) edge [bend left=0] node [elabel,]               {\epslabel} (q4);
         \draw[uedge,eps] (e4) edge [bend left=0] node [elabel,pos=.35]       {\epslabel} (q1);
         \draw[uedge,eps] (e6) edge [bend left=0] node [elabel,]               {\epslabel} (q4);
         \draw[uedge,eps] (p3) edge [bend left=0] node [elabel,]                {\epslabel} (e4);
         \draw[uedge,eps] (p3) edge [bend left=0] node [elabel,]               {\epslabel} (e5);
         \draw[uedge,eps] (p3) edge [bend left=0] node [elabel,]               {\epslabel} (e6);
         \draw[uedge,eps] (e1) edge [bend left=0] node [elabel,]               {\epslabel} (q2);
         \draw[uedge,eps] (e5) edge [bend left=0] node [elabel,pos=0.3]       {\epslabel} (q2);

         \draw[uedge,minus] (p1) edge [bend left=28] node  [elabel,]          {\minuslabel} (q2);
         \draw[uedge,minus] (p2) edge [bend left=18] node  [elabel,pos=.55]   {\minuslabel} (q1);
         \draw[uedge,minus] (p2) edge [bend left=10] node  [elabel,pos=.4]    {\minuslabel} (q4);
         \draw[uedge,minus] (p3) edge [bend left=10] node  [elabel,pos=.6]     {\minuslabel} (q1);
         \draw[uedge,minus] (p3) edge [bend left=-25] node [elabel,pos=0.45,below=5pt] {\minuslabel} (q2);
         \draw[uedge,minus] (p3) edge [bend right=28] node [elabel,pos=.8]           {\minuslabel} (q4);
       \end{tikzpicture}
\end{document}

### 2018-04-09 soutenance/talk.tex
material for the talk

\documentclass[10pt,svgnames,ignorenonframetext,final]{beamer}
% \usepackage{polyglossia}
% \setmainlanguage{english}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{tabulary}
\usepackage{bm}
\usepackage{changepage}
\usepackage[np]{numprint}
\input{common}
\usepackage{colortbl}
\usepackage{xcolor}
\usepackage{appendixnumberbeamer}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
% titleformat plain=allcaps,
%titleformat frame=smallcaps,
\usetheme[numbering=fraction,progressbar=frametitle,block=fill,]{metropolis}
\definecolor{lightyellow}{RGB}{255,240,210}
\setbeamercolor{frametitle}{bg=lightyellow,fg=DarkOrange}
\setbeamerfont{frametitle}{size=\Large,shape=\scshape}
\setbeamertemplate{frametitle}[default][center]
% \usepackage{lmodern}
\usepackage{amssymb,amsmath}
% \iffalse
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
\usepackage[]{hyperref}
\hypersetup{
            pdftitle={Characterizing edges in signed and vector-valued graphs},
            pdfauthor={Géraud Le Falher},
            pdfborder={0 0 0},
            unicode=true,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
% \fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
% \IfFileExists{microtype.sty}{%
% \usepackage{microtype}
% \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
% }{}
\newif\ifbibliography
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight0.8\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

% Prevent slide breaks in the middle of a paragraph:
\widowpenalties 1 10000
\raggedbottom

\usepackage[citestyle=numeric-comp,bibstyle=ieee,isbn=false,maxnames=1,minnames=1,sorting=none,backend=biber,defernumbers=true]{biblatex}
\AtEveryBibitem{
   \clearfield{arxivId}
   % \clearfield{booktitle}
   \clearfield{doi}
   \clearfield{eprint}
   \clearfield{eventdate}
   \clearfield{isbn}
   \clearfield{issn}
   % \clearfield{journaltitle}
   \clearfield{month}
   % \clearfield{number}
   % \clearfield{pages}
   \clearfield{series}
   % \clearfield{url}
   \clearfield{urldate}
   \clearfield{venue}
   % \clearfield{volume}
   \clearlist{location} % alias to field 'address'
   \clearlist{publisher}
   \clearname{editor}
}
\addbibresource{../short_biber.bib}

% Comment these out if you don't want a slide with just the
% part/section/subsection/subsubsection title:
% \AtBeginPart{
%   \let\insertpartnumber\relax
%   \let\partname\relax
%   \frame{\partpage}
% }
\AtBeginSection{
  % \ifbibliography
  % \else
    \let\insertsectionnumber\relax
    \let\sectionname\relax
    \frame{\sectionpage}
  % \fi
}
\iffalse
\AtBeginSubsection{
  \let\insertsubsectionnumber\relax
  \let\subsectionname\relax
  \frame{\subsectionpage}
}
\fi

\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\providecommand{\largelist}{%
  \setlength{\itemsep}{8pt}\setlength{\parskip}{3pt}}
\setcounter{secnumdepth}{0}

\tikzset{
  every overlay node/.style={
    % draw=black,fill=white,rounded corners,
    fill=white!98!black, anchor=north west,
  },
}
% Usage:
% \tikzoverlay at (-1cm,-5cm) {content};
% or
% \tikzoverlay[text width=5cm] at (-1cm,-5cm) {content};
\def\tikzoverlay{%
   \tikz[baseline,overlay]\node[every overlay node]
}%

% \AtBeginSection{}
\title{Characterizing edges in signed and vector-valued graphs}
\author{Géraud Le Falher}
\date{April 16, 2018}

\begin{document}

\frame{\titlepage}

\section{Introduction (8 minutes)}\label{introduction-8-minutes}

\begin{frame}{General context}
\protect\hypertarget{general-context}{}

\begin{itemize}
\item
  In Machine Learning, we seek to automatically extract patterns from
  data and exploit them on future data.
\item
  In this work, our data take the form of graphs, basically a set of
  entities (nodes) and their relationship (edges)
% \end{itemize}
\begin{center}
\includegraphics{../assets/tikz/defense_graph_exe_tikz.pdf}
\end{center}

% \begin{itemize}
% \tightlist
\item
  Because of their simplicity and recent availability, graphs are
  ubiquitous, supporting tasks such as community detection,
  semi-supervised learning, link prediction and influence maximization.
\end{itemize}

\end{frame}

\begin{frame}{More complex graphs -- \wik{}}
  \begin{itemize}
    \tightlist
    \item Some graphs are more \alert{complex}: multiple types of edge and nodes with attributes.
    \item Votes on Wikipedia: nodes are editors and edges represent for or against vote to get
      promoted to administrators
\end{itemize}

  \begin{center}
  \includegraphics[height=6cm]{wiki.pdf}
  \end{center}
\end{frame}
\begin{frame}{More complex graphs -- Bipartite purchase}
  Bipartite purchase network: nodes are customers and products, edges
  are reviews

  \begin{center}
  \includegraphics[height=7cm]{bipartite.pdf}
  \end{center}
\end{frame}
\begin{frame}{More complex graphs -- Co-purchase}
  Attributed co-purchase network: nodes are products and their
  characteristic, edges denotes “frequently purchased together”

  \begin{center}
  \includegraphics[height=8.5cm]{movies2.png}
  \end{center}
\end{frame}

\begin{frame}{Thesis statement}
\protect\hypertarget{thesis-statement}{}

\begin{itemize}
\item
  Such graphs are richer, but edge label might be costly to obtain,
  too numerous or missing.
\item
  Fortunately, I will show that \textbf{there exist \alert{efficient and accurate} methods to
  \alert{predict edge type} in \alert{complex networks}, relying only on the \alert{graph topology}
or also on \alert{node attributes}.}
\end{itemize}

\end{frame}

  \begin{frame}
    \frametitle{Outline}
    \tableofcontents%[currentsection]
  \end{frame}

\section{I - Directed signed graphs (8 mins)}
\label{i---directed-signed-graphs-8-minutes}

\begin{frame}{Motivations}

    More Directed signed social networks (DSSN) besides \wik{} editors:
\begin{description}
    \largelist
\item[\aut{}]
\(i\) cites the work of \(j\) to praise it or criticise it.

\item[\sla{}]
\(i\) considers \(j\) as a friend or foe.

\item[\epi{}]
\(i\) trusts or not the reviews made by \(j\).

\item[\kiw{}]
\(i\) reacted to a Wikipedia edit made by \(j\), to enhance it or revert it.
\end{description}

\end{frame}

\begin{frame}{Problem statement}

  Given the topology of a directed graph and the signs of some edges:
  predict the remaining signs \(\rightarrow\) batch binary
  classification

\begin{columns}[c]
\begin{column}{.55\textwidth}
    \includegraphics[height=5cm]{../assets/tikz/defense_signed_tikz.pdf}
\end{column}%
\hfill%
\begin{column}{.44\textwidth}
  Being able to predict edge signs let us solve \textbf{practical, real world problems}:
\end{column}%
\end{columns}
  % \begin{center}
  % \end{center}
  \begin{itemize}
    \tightlist
  \item
    ``Frenemy'' detection%;~\autocite{frenemy12};
  \item
    Automatic moderation of large scale online interactions
  \item
    Cyber bullying prevention, at school or in online games%~\autocite{CyberbullyingCHI15}.
\end{itemize}

\end{frame}

\begin{frame}{Solution Part 1: Generative model}
\protect\hypertarget{solution-part-1-generative-model}{}

The first ingredient is a sign generative model, made of 2 parameters $p$ and $q$
governing emitting and receiving behavior and drawn from a prior distribution $\mu$.

\begin{center}
\includegraphics{../assets/tikz/troll_genmodel_tikz.pdf}
\end{center}

\end{frame}

\begin{frame}{Solution Part 2: Graph transformation}
\protect\hypertarget{solution-part-2-graph-transformation}{}

The second one is a linear transformation of the graph, turning the
problem into node classification. Doing label propagation on that new
graph is an approximation of computing the maximum likelihood estimator
of \(p\) and \(q\).

\begin{columns}[c]
\begin{column}{.49\textwidth}
\includegraphics[height=5cm]{../assets/tikz/defense_original_tikz.pdf}
\end{column}%
\hfill%
\begin{column}{.49\textwidth}
\includegraphics[height=5cm]{../assets/tikz/defense_weighted_tikz.pdf}
\end{column}%
\end{columns}

\end{frame}

\begin{frame}{Solution Part 3: Theoretical guarantees}
\protect\hypertarget{solution-part-3-theoretical-guarantees}{}

Finally we show that when the graph is dense and the sampling uniform,
directly estimating \(p\) and \(q\) on training data has good
theoretical property, and performs well in practice anyway.

Namely, when we have $Q = \frac{1}{2\ve^2}\ln\frac{4|V|}{\delta}$ samples of outgoing and incoming
edges for every nodes, then, letting $\outqt{\htr}$, $\inqt{\hun}$ and $\biasqt{\tauhat}$ being
empirical estimates, we have that:

\begin{equation*}
\left|\Big[\outqt{\big(1-\htr_u\big)} + \inqt{\big(1-\hun_v\big)} -\biasqt{\tauhat} \Big] -
\Big[\frac{\outqt{p_u}+\inqt{q_v}}{2}\Big]\right| \leq 8\ve
\end{equation*}

\alert{holds with probability at least $1-10\delta$} simultaneously for all
non-queried edges $(u,v) \in E$ such that $\dout(u),\din(v) \ge Q$.

\end{frame}

\begin{frame}{Experiments: \epi{} dataset}
\protect\hypertarget{experiments-epinion-dataset}{}

\begin{table}[p]
  % \begin{adjustwidth}{-0.8cm}{}
    \small
  \centering
  \caption{$100\times$ MCC results on \epi{} as $|\Etrain|$ grows}
  \begin{tabular}{lcccccc|r}
    \toprule
    {}               & Global     & 3\%             & 9\%                & 15\%               & 20\%               & 25\%               & time (ms)     \\
    \midrule
    \uslogregp{}     &            & 43.51           & 54.85              & 59.29              & 61.45              & 62.95              & 32            \\
    \rowcolor{lightyellow}
    \usrule{}        &            & 41.39           & 53.23              & 57.76              & 60.06              & 61.93              & \textbf{7}    \\
    \rowcolor{lightyellow}
    \uslpropGsec{}   & \checkmark & \vsecond{51.47} & \vsecondSig{58.43} & \vsecondSig{61.41} & \vsecondSig{63.14} & \vsecondSig{64.47} & \textbf{1226} \\
    \midrule
    \compranknodes{} & \checkmark & \vfirst{52.04}  & \vfirstSig{60.21}  & \vfirstSig{62.69}  & \vfirstSig{64.13}  & \vfirstSig{65.22}  & \textbf{2341} \\
    \complowrank{}   & \checkmark & 36.84           & 43.95              & 48.61              & 51.43              & 54.51              & 121530        \\
    \compbayesian{}  &            & 31.00           & 48.24              & 56.88              & 61.49              & 64.45              & 116838        \\
    \comptriads{}    &            & 34.42           & 49.94              & 54.56              & 56.96              & 58.73              & 129           \\
    \bottomrule
  \end{tabular}
  % \end{adjustwidth}
\end{table}

\end{frame}

\begin{frame}{Experiments: \aut{} dataset}
\protect\hypertarget{experiments-citation-dataset}{}

\begin{table}[p]
  % \begin{adjustwidth}{-0.8cm}{}
  \centering
  \small
  \caption{$100\times$ MCC results on \aut{} as $|\Etrain|$ grows}
    \begin{tabular}{lcccccc|r}
    \toprule
    {}               & Global     & 3\%                & 9\%                & 15\%               & 20\%               & 25\%               & time (ms)            \\
    \midrule
    \uslogregp{}     &            & \vsecondSig{15.19} & \vsecondSig{26.46} & 32.98              & 36.57              & 39.90              & 2                    \\
    \rowcolor{lightyellow}
    \usrule{}        &            & 15.09              & 26.40              & \vsecondSig{32.98} & \vsecondSig{36.72} & 40.16              & \textbf{\textless 1} \\
    \rowcolor{lightyellow}
    \uslpropGsec{}   & \checkmark & \vfirstSig{19.00}  & \vfirstSig{30.25}  & \vfirstSig{35.73}  & \vfirstSig{38.53}  & \vfirstSig{41.32}  & \textbf{16}          \\
    \midrule
    \compranknodes{} & \checkmark & 12.28              & 24.44              & 31.03              & 34.57              & 38.26              & \textbf{128}         \\
    \complowrank{}   & \checkmark & 8.85               & 17.08              & 22.57              & 25.57              & 29.24              & 1894                 \\
    \compbayesian{}  &            & 10.91              & 23.75              & 32.25              & 36.52              & \vsecondSig{40.32} & 5398                 \\
    \comptriads{}    &            & 8.62               & 16.42              & 22.01              & 24.77              & 27.13              & 5                    \\
    \bottomrule
    \end{tabular}
  % \end{adjustwidth}
\end{table}

\end{frame}

\begin{frame}{Conclusion}
\protect\hypertarget{conclusion}{}

Because we only use the topology, our solution is both:

\begin{itemize}
\tightlist
\item
  domain agnostic
\item
  fast
\end{itemize}

\end{frame}

\section{II - Undirected signed graphs (8 minutes)}
\label{ii---undirected-signed-graphs-8-minutes}

\begin{frame}{Problem}

\begin{itemize}
\tightlist
\item
  Our sign generative model is not suitable for undirected graphs.\\
  Yet such graphs are important, for instance for recommendation in
  bipartite graphs.
\end{itemize}

\begin{columns}[T] % align columns
\begin{column}{.49\textwidth}
\includegraphics[height=5cm]{suggestions.png}
\end{column}%
\hfill%
\begin{column}{.49\textwidth}
\includegraphics[height=5cm]{sugbipart.pdf}
\end{column}%
\end{columns}

\begin{itemize}
\tightlist
\item
  We consider the active setting, where we first select a training set,
  query its signs and predict the remaining edges.
\end{itemize}

\end{frame}

\begin{frame}{New bias}

\begin{itemize}
\item
  Assume that nodes belongs to \(K\) latent groups, and that signs are
  governed by those groups, modulo some irregularities.
  \begin{center}
  \includegraphics[height=3cm]{../assets/tikz/defense_kgroups_tikz.pdf}
  \end{center}
\item This is motivated in social networks by the balance theory and in other graphs by assortative/dissortative patterns.
\item
  Recovering those \(K\) groups in the presence of noise is the well
  studied Correlation Clustering problem (CC). The minimal objective
  value of CC when \(K=2\) is a bound on the number of mistakes for any
  active algorithms \autocite{Cesa-Bianchi2012b}.
\end{itemize}

\end{frame}

\begin{frame}{Solution: spanning structure}

\begin{itemize}
\item
  An interesting case is \(K\) = 2 (strong balance), if there was no
  noise, the sign of any edge would the parity of any path between its
  endpoint.
\item Since the noise is uniform (with probability $q$), the longer such path, the more likely we make an error by
  predicting using parity of observed edges. Overall, the expected number of errors is:
  \begin{equation*}
    q \left(|E| + \sum_{(u,v) \in E_{\mathrm{test}}} |\mathrm{path}^T(u, v)| \right)
  \end{equation*}
\item
  Thus we look for a spanning structure with few edges that aims to minimize the stretch.
\end{itemize}
\end{frame}

\begin{frame}{Construction}
  \begin{center}
    \includegraphics{../assets/tikz/gtx_eccentricity_tikz.pdf}
  \end{center}
\end{frame}

\begin{frame}{Experiments - Stretch}
\includegraphics{../assets/gtx_exp/gridst.pdf}
\end{frame}

\begin{frame}{Experiments - Stretch}
\includegraphics{../assets/gtx_exp/pasynthmcc.pdf}
\end{frame}

\section{III - Node attributed graphs (15 minutes)}
\label{iii---node-attributed-graphs-15-minutes}

\begin{frame}{Problem}
\protect\hypertarget{problem}{}

\begin{itemize}
\item
  Still characterizing edges, two big differences:

  \begin{itemize}
  \tightlist
  \item
    more than 2 types of edges \(\rightarrow\) \(k\)-multilayer graphs
  \item
    no label are provided at any point \(\rightarrow\) unsupervised
    problem
  \end{itemize}
\item
  Additional information: node \(u\) has a profile \(x_u \in [-1, 1]^d\)
\item
  Find \(k\) bounded “directions” and associate to every edge the “most
  explanatory” direction among those \(k\)
\end{itemize}

\end{frame}

\begin{frame}{Motivations}
\protect\hypertarget{motivations}{}

\begin{itemize}
\item
  BlaBlaCar: nodes are users described by experience, location,
  preferences (music, smoke, talk, pet), reviews, while edges are shared
  rides. Those rides could be to go to work, cultural events, vacations
  and so on.
\item
  Foursquare: nodes area venues described by location, reviews,
  category, time of visits, while edges are “frequently visited
  together” relation. Those venue groups can be part of party night,
  Sunday morning or sport routine.
\item
  We model nuanced relations driven both by partial homophily and
  heterophily.
\end{itemize}

\end{frame}

\begin{frame}{Formalization}
\protect\hypertarget{formalization}{}

\begin{itemize}
\item
  Combine the profiles of the
  \textcolor{DodgerBlue}{edge $u,v$ by entrywise product} and
  \textcolor{Orange}{score it} with
  \textcolor{Green}{direction $w_\ell$}:
\begin{equation*}
  \mathcolor{Orange}{\big(}\mathcolor{DodgerBlue}{x_u \circ x_v}\mathcolor{Orange}{\big)^T}
  \mathcolor{Green}{w_\ell}
\end{equation*}
\end{itemize}
\begin{columns}[T] % align columns
\begin{column}{.59\textwidth}
\begin{itemize}
\item
  Associate the \textcolor{HotPink}{"most explanatory" direction} to
  this edge:
  \begin{equation*}
    \mathcolor{HotPink}{\max_{\ell \in \{1, \ldots, k\}}} 
    \mathcolor{Orange}{\big(}\mathcolor{DodgerBlue}{x_u \circ x_v}\mathcolor{Orange}{\big)^T}
    \mathcolor{Green}{w_{\mathcolor{HotPink}{\ell}}}\quad\text{Let}\,
    \mathcolor{Green}{w_{\mathcolor{HotPink}{\ell}}} =
    \mathcolor{Green}{w(}\mathcolor{HotPink}{u,v}\mathcolor{Green}{)}
  \end{equation*}
\item
  Find the \textcolor{Teal}{overall best $k$ directions}:
  \begin{equation*}
    \mathcolor{Teal}{\argmax_{\{w_1,\ldots,w_k\} \subset \dball}
    \sum_{u,v \in E}}
    \mathcolor{HotPink}{\max_{\ell \in \{1, \ldots, k\}}} 
    \mathcolor{Orange}{\big(}\mathcolor{DodgerBlue}{x_u \circ x_v}\mathcolor{Orange}{\big)^T}
    \mathcolor{Green}{w_{\mathcolor{HotPink}{\ell}}}
  \end{equation*}
\end{itemize}
\end{column}%
\hfill%
\begin{column}{.42\textwidth}

{\small
  \setlength{\tabcolsep}{4pt}
  \begin{tabular}{rrrrrr}
    \toprule
    ${x_u}$ & ${x_v}$ & ${\mathcolor{DodgerBlue}{x_u \circ x_v}}$
            & $\mathcolor{Green}{w_1}$ & $\mathcolor{Green}{w_2}$ \\
    \midrule
    $0.9$  & $0.8$  & $0.72$  & $0.6$  & $0.1$ \\
    $-0.8$ & $-0.9$ & $0.72$  & $0.1$  & $0.4$ \\
    $0.9$  & $-0.8$ & $-0.72$ & $-0.7$ & $0.6$ \\
    $0.0$  & $0.5$  & $0.0$   & $0.1$  & $0.4$ \\
    \bottomrule
  \end{tabular}
}

$\mathcolor{Orange}{\big(}\mathcolor{DodgerBlue}{x_u \circ x_v}\mathcolor{Orange}{\big)^T} \mathcolor{Green}{w_1} = 1.008$
$\mathcolor{Orange}{\big(}\mathcolor{DodgerBlue}{x_u \circ x_v}\mathcolor{Orange}{\big)^T} \mathcolor{Green}{w_2} = -0.072$
\end{column}%
\end{columns}

\end{frame}

\begin{frame}{Two Topological Constraints}
\protect\hypertarget{two-topological-constraints}{}

\begin{enumerate}
  \item The profile of each node is (close to) a linear combination of its incident directions: this
    introduces additional parameters ($\{a_{uv}\}_{u,v \in E}$ and $\{b_u\}_{u \in V}$) but provides more guidance.

\begin{equation*}
  \mathcal{L}_{\mathrm{node}} =
  \sum_{u\in V} \left|\left| x_u - b_u -
  \sum_{v \in \mathcal{N}(u)} a_{uv} 
    \mathcolor{Green}{w(}\mathcolor{HotPink}{u,v}\mathcolor{Green}{)}
 \right|\right|^2_2
\end{equation*}

  \item Each node can only be involved in \(k_\mathrm{local} < k\) directions.

\begin{equation*}
  \label{eq:edge_local_loss}
  \mathcal{L}_{\mathrm{local}} =
  \sum_{u \in V} \left\| \sum_{v \in \nei(u)} y_{uv} \right\|_1
  % = \sum_{u \in V} \left( \sum_{i = 1}^k \sum_{v \in \nei(u)}
  % \frac{\exp(\beta g(s_{uv}, w_i))}{\sum_{j=1}^k \exp(\beta g(s_{uv}, w_j))} \right)
\end{equation*}
Think of $y_{uv}$ as an indicator vector with $k$ components. Its $i^{th}$ component can be relaxed
to

\begin{equation*}
  {y_{uv}}_{;i} = \frac{\mathcolor{Sienna}{\exp(\beta} 
    \mathcolor{Orange}{\big(}\mathcolor{DodgerBlue}{x_u \circ x_v}\mathcolor{Orange}{\big)^T}\mathcolor{Green}{w_i}
  \mathcolor{Sienna}{)}}%
  {\mathcolor{Sienna}{\sum_{j=1}^k \exp(\beta} 
    \mathcolor{Orange}{\big(}\mathcolor{DodgerBlue}{x_u \circ x_v}\mathcolor{Orange}{\big)^T}\mathcolor{Green}{w_j}
  \mathcolor{Sienna}{)}}
\end{equation*}
\end{enumerate}

\end{frame}

\begin{frame}{First Solutions}
\protect\hypertarget{first-solutions}{}

\begin{itemize}
\item
  A baseline is simply to cluster the \(|E| \, \mathcolor{DodgerBlue}{x_u \circ x_v}\) in \(k\)
  clusters.
\item
  It can be improved by plugging our scoring function in the Lloyd
  algorithm.
\item
  Another approach is to directly optimize our objective function. We first make it convex by
  \textcolor{Sienna}{relaxing \(\mathcolor{HotPink}{\max}\) using softmax}.

  \begin{equation*}
    \mathcolor{Teal}{\argmax_{\{w_1,\ldots,w_k\} \subset \dball}
    \sum_{u,v \in E}}
    \mathcolor{Sienna}{\frac{1}{\beta}}
    \mathcolor{Sienna}{\sum_{\ell=1}^k\log\Big(\exp\big(\beta}
        \mathcolor{Orange}{\big(}\mathcolor{DodgerBlue}{x_u \circ x_v}\mathcolor{Orange}{\big)^T}
        \mathcolor{Green}{w_{\mathcolor{Sienna}{\ell}}}
    \mathcolor{Sienna}{\big)\Big)}
  \end{equation*}
Then we take topology into account, by relaxing as well the two previous
  regularization terms, and optimize the difference of convex terms by gradient
  descent\footnote{Disciplined Convex-Concave Programming}
\end{itemize}

\end{frame}

\begin{frame}{Matrix Solution -- Convex formulation}
\protect\hypertarget{matrix-solutions}{}

\begin{itemize}
\item
  Find a direction for each edge, but make sure those directions are
  linear combination of \(k\) “base” directions \(\rightarrow\) \textcolor{Peru}{low rank
  matrix}.

  \begin{equation*}
    \sum_{u,v \in E}
    \mathcolor{Orange}{\big(}\mathcolor{DodgerBlue}{x_u \circ x_v}\mathcolor{Orange}{\big)^T}
    \mathcolor{Green}{w_{uv}}
    = \mathcolor{Orange}{\big\langle} \mathcolor{DodgerBlue}{S^T} \mathcolor{Orange}{,}
    \mathcolor{Green}{W} \mathcolor{Orange}{\big\rangle_\mathrm{F}}
  \end{equation*}

  \begin{equation*}
    \mathcolor{Teal}{\min_{W\in \mathbb{M}^{d\times |E|}}}
    -\mathcolor{Orange}{\big\langle} \mathcolor{DodgerBlue}{S^T} \mathcolor{Orange}{,}
    \mathcolor{Green}{W} \mathcolor{Orange}{\big\rangle_\mathrm{F}} + \mathcolor{Peru}{\rank(W)}
  \end{equation*}

\item
  Relax low rank by \textcolor{Peru}{nuclear norm} and ensure \textcolor{Peru}{the norm of \(W\)
  columns are not too large}. Optimized by the Frank Wolfe method, only requiring top singular
  vector at each iteration.

  \begin{equation*}
    \mathcolor{Teal}{\min_{\substack{W\in \Rbb^{d\times |E|} \\ \mathcolor{Peru}{\|W\|_* \leq \delta}}}}
    \mathcolor{Orange}{\big\langle} \mathcolor{Peru}{\mu W}
      -\mathcolor{DodgerBlue}{S^T} \mathcolor{Orange}{,}
    \mathcolor{Green}{W} \mathcolor{Orange}{\big\rangle_\mathrm{F}}
  \end{equation*}
\end{itemize}
\end{frame}

\begin{frame}{Matrix Solution -- Alternating formulation}
\begin{itemize}
\item
  Write \(W\) explicitly as \(PQ^T\) and use alternating optimization.
\item
  Those formulations have more parameters, but allow mixed membership.
\end{itemize}

\end{frame}

\begin{frame}{Synthetic experiments}
\protect\hypertarget{synthetic-experiments}{}

% \begin{itemize}
% \tightlist
% \item
% \end{itemize}
  we generate data in order to have ground truth: (500 nodes, 1300 edges, 200 times $k$ directions
  and corresponding profiles)

% \begin{adjustwidth}{-.5cm}{}
  % \small
  \begin{tabulary}{\textwidth}{llll|ll}
    \toprule
    \thead{$\mathcal{D}_k$} & \thead{\kmeans{}} & \thead{\lloyd{}}                      &
    \thead{\combined{}}                  & \thead{\textsc{Frank Wolfe}} & \thead{\pqt{}} \\
    \midrule
    {\default{}}            & $.818 $           & $\esecondSig{.873 }\spval{1.25}{-63}$ & $\efirstSig{.893 }\spval{5.68}{-33}$ & $.381 $                      & $.893 $        \\
    {\smallk{}}             & $.836 $           & $\esecond{.838 }$                     & $\efirstSig{.875 }\spval{2.17}{-58}$ & $.213 $                      & $.875 $        \\
    {\largek{}}             & $.803 $           & $\esecondSig{.881 }\spval{2.66}{-94}$ & $\efirstSig{.894 }\spval{8.98}{-17}$ & $.421 $                      & $.894 $        \\
    {\smallo{}}             & $.813 $           & $\esecondSig{.824 }\spval{7.57}{-6}$  & $\efirstSig{.856 }\spval{2.99}{-57}$ & $.378 $                      & $.855 $        \\
    {\largeo{}}             & $\esecond{.827 }$ & $.823 $                               & $\efirstSig{.852 }\spval{1.90}{-25}$ & $.370 $                      & $.851 $        \\
    {\fdirs{}}              & $.772 $           & $\esecondSig{.814 }\spval{6.02}{-42}$ & $\efirstSig{.853 }\spval{2.13}{-47}$ & $.320 $                      & $.853 $        \\
    {\larged{}}             & $.905 $           & $\esecondSig{.933 }\spval{1.32}{-31}$ & $\efirstSig{.941 }\spval{1.77}{-22}$ & $.222 $                      & $.931 $        \\
    \bottomrule
  \end{tabulary}
% \end{adjustwidth}

\end{frame}

\begin{frame}{Conclusions}

\begin{itemize}
    \item
      Given a graph with node profiles, find $k$ directions explaining the existing connections
    \item
      linear model for performance and interpretability, two formulations:
    \begin{itemize}
      \item difference of convex scalar functions
      \item Matrix convex expression suitable for Frank--Wolfe algorithm
    \end{itemize}
    \item
      Future work:
    \begin{itemize}
      \item take time into account (homophily vs contagion)
      \item multigraph, more than one type of edge between two nodes
    \end{itemize}
\end{itemize}

\end{frame}

\section{IV - Conclusion (6 minutes)}
\label{iv---conclusion-6-minutes}

\begin{frame}{Contribution}
\protect\hypertarget{contribution}{}

\textbf{efficient and accurate methods to predict edge type in complex
networks, relying only on the graph topology or also on node
attributes.}

\begin{adjustwidth}{-.31cm}{}
  \small
  \setlength{\tabcolsep}{4pt}
  \begin{tabulary}{0.9\paperwidth}{lLLL}
    \toprule
    & I & II & III \\
    \midrule
    graph            & directed, 2 types      & undirected, 2 types           & attributed, $k$ types \\
    learning setting & batch label            & active label                  & no label \\
    approach         & estimate parameters    & find spanning structure       & matrix optimization \\
    efficient        & $O(|E|)$, parallel & $O(|E|\log |E|)$, conjectured & $O(dT|E|)$ \\
    accurate &
    close to Bayes predictor, efficient in practice &
    Close to Correlation Clustering bound &
    convex problem: global max (but no ground truth) \\
    \bottomrule
  \end{tabulary}
\end{adjustwidth}

\end{frame}

\begin{frame}{Future work}
\protect\hypertarget{future-work}{}

\begin{itemize}
\tightlist
\item
  Two ways relation with representation learning in graphs:

  \begin{itemize}
  \tightlist
  \item
    can give more accurate input to RL methods
  \item
    can exploit RL features as node attributes
  \end{itemize}
\item
  \emph{link prediction as Liva suggested?}
\end{itemize}

\end{frame}

\begin{frame}{References}
  \printbibliography
\end{frame}

\begin{frame}[plain,c]

\begin{center}
\Huge Thank you! \\ Questions?
\end{center}

\end{frame}

\appendix

\begin{frame}{Setting}

The signs are \textbf{adversarial} rather than generated by our model.

At each round, the learner is asked to predict one label, which is then revealed to him and the
procedure repeats.

\end{frame}

\begin{frame}{Labeling regularity}

% Inspired by trollness and trustworthiness, we define the regularity of a labeling \(Y\).

Letting $Y$ be the vector of all labels, $\Psiout(i, Y)$ is the number of least used label outgoing
from $i$, and $\Psiout(Y) = \sum_{i \in V} \Psiout(i,Y)$.

Likewise for incoming edges, \(\Psiin(Y) = \sum_{j \in V} \Psiin(j,Y)\) and finally
\(\Psi_G(Y) = \min\big\{\Psiin(Y), \Psiout(Y)\big\}\).

% This can be read from the cutsize of \(G'\)
\begin{columns}[T]
  \begin{column}{0.6\textwidth}
    % \vspace{-2cm}
    \begin{tabular}{lcccc|l}
      \toprule
      node $i$        & $1$ & $2$ & $3$ & $4$ &                  \\
      \midrule
      $\Psiout(i, Y)$ & 0   & 0   & 1   & 0   & $\Psiout(Y) = 1$  \\
      $\Psiin(i, Y)$  & 1   & 0   & 0   & 1   & $\Psiin(Y) = 2$ \\
      \bottomrule
    \end{tabular}
  \end{column}%
  \hfill%
  \begin{column}{0.39\textwidth}
    \centering \includegraphics[height=.35\textheight]{../assets/raw/g_latex-crop}
  \end{column}%
\end{columns}

\end{frame}

\begin{frame}{Online algorithm and bounds}
  \begin{itemize}[<+->]
    \largelist
  \item 
    Our algorithm is a combination of Randomized Weighted Majority instances built on top of each other.

  \item 
    On average, it makes $\Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V|\right)$ mistakes.

  \item 
    On the lower side,
    % Up to this $O\left(\sqrt{|V|\Psi_G(Y)} + |V|\right)$ factor, it is optimal as we can show that 
    for any directed graph $G$ and any integer $K$,
    there exists a labeling $Y$ forcing at least $\frac{K}{2}$ mistakes to any online algorithms,
    while $\Psi_G(Y) \leq K$.
\end{itemize}

\end{frame}

\begin{frame}{Online algorithm, 1. RWM node instances}

For each node \(i\), we predict the sign of edge outgoing from \(i\) by
relying on two constant experts, always predicting \(-1\) or always
predicting \(+1\). The best one will make \(\Psiout(i, Y)\) mistakes. We
combine them in a Randomized Weighted Majority algorithm (RWM) instance
associated with \(i\), call it \(RWM_{out}(i)\). The instance expected
number of mistakes is therefore~\autocite{acg02},
denoting by \(M(i,j)\) the indicator function of a mistake on edge
\((i,j)\)
\[\sum_{j \in \Nout(i)} \E\,M(i,j) = \Psiout(i,Y) + O\left(\sqrt{\Psiout(i,Y)}+ 1\right)\]

We use the same technique to predict incoming edges of each node \(j\),
the instance \(RWM_{in}(j)\) having the following average number of
mistakes
\[\sum_{i \in \Nin(j)} \E\,M(i,j) = \Psiin(j,Y) + O\left(\sqrt{\Psiin(j,Y)} + 1\right)\]

\end{frame}

\begin{frame}{Online algorithm, 2. combining instances}

We then define two meta experts: \(RWM_{out}\), which predicts
\(y_{i,j}\) as \(RWM_{out}(i)\), and \(RWM_{in}\), which predicts
\(y_{i,j}\) as \(RWM_{in}(j)\). Summing over all nodes, the number of
mistakes of these two experts satisfy

\begin{align*}
    \sum_{i \in V}\sum_{j \in \Nout(i)} \E\,M(i,j) &= \Psiout(Y) + O\left(\sqrt{|V|\Psiout(Y)} + |V|\right) \\
    \sum_{j \in V}\sum_{i \in \Nin(j)} \E\,M(i,j)  &= \Psiin(Y)  + O\left(\sqrt{|V|\Psiin(Y)}  + |V|\right)
\end{align*}
\end{frame}

\begin{frame}{Online algorithm, 3. final prediction}

Our final predictor is a RWM combination of \(RWM_{out}\) and
\(RWM_{out}\), whose expected number of mistakes is
\begin{alignat*}{3}
    \sum_{(i,j) \in E} \E\,M(i,j) 
    &= \Psi_G(Y) + O\Biggl(&&\sqrt{|V|\Psi_G(Y)} + |V| \\
  & &&+ \sqrt{\Bigl(\Psi_G(Y) + |V| + \sqrt{|V|\Psi_G(Y)}\Bigr)} \Biggr)\\
    &= \Psi_G(Y) + O\Bigl( &&\sqrt{|V|\Psi_G(Y)} + |V|\Bigr)
\end{alignat*}

\end{frame}

\begin{frame}{Datasets properties}

\begin{table}
  \centering
  % \small
  \caption{Dataset properties. \label{tab:dataset}}
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Dataset & $|V|$       & $|E|$       & $\frac{|E|}{|V|}$ & $\frac{|E^+|}{|E|}$ & $\frac{\Psi_{G''}(Y)}{|E|}$ & $\frac{\Psi_G(Y)}{|E|}$ \\
    \midrule
    \aut{}  & \np{4831}   & \np{39452}  & 8.1               & 72.33\%             & .076                        & .191                    \\
    \wik{}  & \np{7114}   & \np{103108} & 14.5              & 78.79\%             & .063                        & .142                    \\
    \sla{}  & \np{82140}  & \np{549202} & 6.7               & 77.40\%             & .059                        & .143                    \\
    \epi{}  & \np{131580} & \np{840799} & 6.4               & 85.29\%             & .031                        & .074                    \\
    \kiw{}  & \np{138587} & \np{740106} & 5.3               & 87.89\%             & .034                        & .086                    \\
    \bottomrule
  \end{tabular}
\end{table}

$$\Psi_{G''}(Y) = 
\min_{\bp,\bq\in[0,1]^{|V|}}
\sum_{(i,j) \in E} \left(\frac{1+y_{i,j}}{2} - \frac{p_i+q_j}{2}\right)^2$$

\end{frame}

\begin{frame}{Data generation}
\protect\hypertarget{data-generation}{}

\begin{itemize}
\item
  create a random graph and pick for each a set of \(k_\mathrm{local}\)
  directions.
\item
  pick for each edge \((u,v)\) a direction index \(y_{uv} \in \rangesk\)
  among the ones shared by its endpoint.
\item
  draw the \(k\) directions at random
\item
  optimize the profiles so as to maximize the edges goodness, minimize
  the term \(\mathcal{L}_{\mathrm{node}}\) and enforces as much as
  possible that for every edge \((u,v) \in E\),
  \(\mathcal{E}(u,v) = y_{uv}\).
\end{itemize}

\emph{(take a look a Liva comment about Lagrangian relaxation of integer
program\ldots{})}

\end{frame}
\end{document}
